### 2019-11-18 pre-reg/pre-reg_merir-scale.Rmd
update1118

---
title: "Pre-reg, scale perception meritocracy"
author:
- Luis Maldonado
- Juan Carlos Castillo
- Julio Iturra
- Jorge Atria
- Francisco Meneses
date: "13-11-2019"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float: true
    toc_depth: 2
    number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Research questions

* How valid is a measurement model for the perception of 4-factor meritocracy, which is grouped into: a first-order   factor related to meritocracy and a second-order factor referred to inequalities of opportunity?

# General description and hypothesis

To measure the ideals of meritocracy, it is essential to differentiate preferences (i.e. what is considered desirable) and perceptions (i.e. how reality is believed to be) (duru-bellat, 2012). This work focuses on the perceptions of meritocracy.

Although a group of authors has measured the perception of meritocracy as if the perception of inequality of opportunities were the  side opposite of the continuum (Kunovich, 2007; Newman, 2015), others have shown that both concepts are rather different dimensions(Reynolds , 2014). Following this evidence, we chose to model both constructs as distinct latent variables. Different types of inequalities such as origin, contact or discrimination, cannot be considered in a single factor, since they refer to different realities and perceptions, therefore, it is consistent that they are independent latent variables. We also consider the existence of a latent second order variable, referring to the belief of inequality of opportunities in general.

Consistently, we consider that questions regarding meritocratic preferences will be grouped into two factors, meritocratic preferences and non-meritocratic preferences.

Additionally, it is important to mention that we will exploratoryly test the relationship between the “get a head” questions and questions that refer to whether “In Chile, people are rewarded or manage to succeed according to different qualities (i.e., hardworking, talented, with contacts)”.

* H1. The perception of meritocracy is a factor in the drift of the importance attributed to talent and the effort to get ahead in life.
* H2. The factors of family origin, contacts and discrimination are grouped into a second order factor that relates to non-meritocratic perceptions.
* H3. Preferences are also grouped into meritocratic and non-meritocratic preferences.
* H4 (exploratory) the way in which the perception of inequality of opportunities is asked influences the perception declared by the interviewees.

# Type of study

The present study is of an observational type focused on the measurement of the meritocracy preception variable. This study is part of an experimental study, applied online, which will be longitudinal panel type, that is, the same subjects will be surveyed 3 different occasions. The subjects will be randomly selected to be assigned to a placebo or to either of the two experimental procedures (change of poverty in Chile or international comparison of inequality)
To access the cases we turn to the Netquest survey company, which has coverage in several Latin American countries.

# Sample

## Detailed data and explanations of its use.

The Sample that will be necessary for this study has not yet been produced, nor has it begun its field work. However, in order to test the online questionnaire and the functionality of the quota offered by the Qualtrics platform, two previous studies have been carried out. The subjects participated in these two opportunities are excluded to the central sample, which corresponds to the 3-wave panel study.
These data from the two pilot studies were used for methodological decisions, descriptive stadistics, correlations and confirmation tests.
Due to the above, although the definitive sample has not yet been produced and therefore has not been explored, it is responsible to say, in addition to the review of previous research and international databases, it has some notions of the relationship among the variables product of the analysis of the previous samples

## Magnitude of the sample and explanation of the magnitude.

A representative sample of the large cities of Chile (Gran Santiago, Valparaíso-Viña del Mar, Concepción and Antofagasta) is expected, which will have an initial number for the first wave of 2100 cases, which are expected to be at least 1800 in the second wave. From the second to the third wave it has been considered as “free fall” that is, we are not sure of the retention rate of the interviewees.
The sample size was calculated from power analysis that used information from the pilot study that was applied in June 2019. This pilot study consisted of an online experiment applied to 949 subjects, randomly separating participants to two stimuli and a placebo, leaving a third of the sample in each condition. From these data, the low power of a two-tailed test was calculated, considering an average treatment effect of 0.26, with a standard deviation of the result of the treatment group of 1.23 and a significance level of 0.05. Based on multiple comparison correlations (Bonferroni), we adjusted alpha for six tests (three results * two comparisons between conditions and placebo). The results indicated that the number of subjects exposed to each stimulus should be 580, so we will work with three samples of 600 cases, which gives us a total of 1800 cases for the second wave in which the experiment will be applied. To guarantee the above and in consideration of a high non-response rate, we will start with a sample of 2100 subjects.

### Data collection procedure:
To access this target population, a company has been hired dedicated to the online surveys that will be in charge of getting the respondents. These will be invited to participate in the survey under the incentive provided by the external company. Respondents have 4 business days to answer the survey, with a maximum of 20 minutes per survey. In order to achieve greater representativeness of the sample, the quota method will be used, that is, the program will only allow respondents to respond while more people with their demographic characteristics are required. The quotas used were age, sex and education. It should be noted that the time between wave and wave is between 7 and 9 days.

### Stop Rule:
The rule of detention is to reach the number of respondents indicated, adjusted to the corresponding quotas. However, there may be modifications depending on the contingencies of the field work, depending on the response rates.

# Variables to use:
The variables used for the perception factor of meritocracy are:

* Para surgir en la vida, ¿cuán importante cree
usted que es?
    + Tener ambición
    + El trabajo duro

On the other hand, the questions related to inequality of opportunities are:

* Para surgir en la vida, ¿cuán importante cree
usted que es?
    + provenir de una familia rica o con muchos recursos
    + Tener padres con altos niveles de educación
    + Conocer a las personas adecuadas
    + Tener contactos políticos
    + La raza u origen étnico de las personas
    + El sexo de las personas, ser hombre o mujer

The items of preferences are:

* ¿En qué medida se encuentra usted de acuerdo o en desacuerdo con cada una de las siguientes afirmaciones?
    + Quienes más se esfuerzan deberían obtener mayores recompensas que quienes se esfuerzan menos
    + Quienes poseen más talento deberían obtener mayores recompensas que quienes tienen menos talento
    + Está bien que quienes tienen padres ricos salgan adelante
    + Está bien que quienes tienen buenos contactos salgan adelante

The questions that will be used for the exploratory analysis indicated are:

* ¿En qué medida se encuentra usted de acuerdo o en
desacuerdo con cada una de las siguientes
afirmaciones? En Chile,
    + Quienes más se esfuerzan logran obtener mayores recompensas.
    + Quienes tienen más talento logran obtener mayores recompensas.
    + Las personas con padres ricos logran salir adelante.
    + Los que tienen buenos contactos logran salir adelante.

### Variable manipulation

The variables will be worked on the same scale in which they were asked, without resorting to manipulations of the variables to assess the validity of the measurement model.

### Indexes

The investigation of the validity of this scale does not resort to indices intended to represent the perception of meritocracy in an observed way, but rather confirmatory factor analysis will be used to estimate the value of the latent variables underlying the indicators.

### Blinding of information

No special blinding is used for this study.

### Randomization
To avoid the existence of biases in the order of the questions, they have been randomized within a thematic block of questions, so, randomly, people answer questions in different orders regarding their perception of meritocracy.

# Analysis plan
To evaluate these hypotheses, confirmatory factor analysis will be used, since we have strong theory regarding the ordering of the factors. In addition, exploratory factor analysis will be used to evaluate the exploratory hypothesis indicated. The Lavaan R package will be used.

### Inference criteria,
The values that will be used as evaluation criteria for the goodness of the adjustment of the model were taken from the proposal of Ruiz, Partdo and San Martin (2010) and are the following:
*Chi-square:> 0.05
*Chi-square ratio:> 3
*Comparative adjustment goodness index (CFI):> 0.95
*Tucker-Lewis Index (TLI):> 0.95
*Root of the average squared residual approximation <0.08.

### Data Exclusion
All cases will be used as long as they do not show missing valuesin any of the items on the scale. No imputation criteria will be used.

### Ethics. JI
The experiment and survey instruments are approved by the IRB of the University of Chile.

# bibliography

Ruiz, M., Pardo, A. & San Martín, R. (2010). Modelos de Ecuaciones Estructurales. -Papeles del psicólogo, 31-(1), pp. 34-45.

Duru-Bellat, M., & Tenret, E. (2012). Who’s for Meritocracy? Individual and Contextual Variations in the Faith. Chicago Journal, 56(2), 223-247.

Kunovich, S., & Slomczynski, K. M. (2007). Systems of Distribution and a Sense of Equity: A Multilevel Analysis of Meritocratic Attitudes in
Post-industrial Societies. European Sociological Review, 23(5), 649–663. doi:10.1093/esr/jcm026

Newman, B. J., Johnston, C. D., & Lown, P. L. (2014). False Consciousness or Class Awareness? Local Income Inequality, Personal Economic Position, and Belief in American Meritocracy. American Journal of Political Science, 59(2), 326-340. doi:10.1111/ajps.12153

Reynolds, J., & Xian, H. (2014). Perceptions of meritocracy in the land of opportunity. Research in Social Stratification and Mobility, 36, 121–137. doi:10.1016/j.rssm.2014.03.001

### 2019-11-18 index.Rmd
update1119

--- 2019-11-18 pre-reg/pre-reg_merir-scale.Rmd
@@
 ---
-title: "Pre-reg, scale perception meritocracy"
-author:
-- Luis Maldonado
-- Juan Carlos Castillo
-- Julio Iturra
-- Jorge Atria
-- Francisco Meneses
-date: "13-11-2019"
+title: "Meritorcacy Scale: Preferences and Perceptions"
+csl: apa6.csl
+bibliography: zlibrary.bib
+css: style.css
 output:
+  pdf_document:
+    keep_tex: false
+    template: null
+    number_sections: true
+    toc: true
+    toc_depth: 2
   html_document:
     theme: flatly
-    highlight: tango
+    number_sections: true
     toc: true
     toc_float: true
-    toc_depth: 2
-    number_sections: false
+geometry: margin=0.78in
+header-includes:
+  # - \usepackage[spanish,es-tabla,es-nodecimaldot]{babel}
+  - \usepackage{caption}
+  - \captionsetup[figure]{labelfont={bf},labelformat={default},labelsep=period}
+  - \usepackage{graphicx}
+  - \usepackage{float}
+linestretch: '1.5'
+link-citations: yes
+linkcolor: blue
 ---
+
+```{r eval=FALSE, echo=FALSE}
+# Para obtener doble output (html/pdf) correr directamente en consola/terminal:
+rmarkdown::render('pre-reg_merir-scale.rmd', output_format = 'all')
+```
 

### 2019-11-21 index.Rmd
update211119

--- 2019-11-18 pre-reg/pre-reg_merir-scale.Rmd
@@
 ---
-title: "Pre-reg, scale perception meritocracy"
-author:
-- Luis Maldonado
-- Juan Carlos Castillo
-- Julio Iturra
-- Jorge Atria
-- Francisco Meneses
-date: "13-11-2019"
+title: "Meritocracy Scale: Preferences and Perceptions"
+csl: apa6.csl
+bibliography: zlibrary.bib
+css: style.css
 output:
+  pdf_document:
+    keep_tex: false
+    template: null
+    number_sections: true
+    toc: true
   html_document:
     theme: flatly
-    highlight: tango
+    number_sections: true
     toc: true
     toc_float: true
-    toc_depth: 2
-    number_sections: false
+geometry: margin=0.78in
+header-includes:
+  # - \usepackage[spanish,es-tabla,es-nodecimaldot]{babel}
+  - \usepackage{caption}
+  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
+  - \usepackage{graphicx}
+  - \usepackage{float}
+  - \usepackage{booktabs}
+  - \usepackage{longtable}
+  - \usepackage{array}
+  - \usepackage{multirow}
+  - \usepackage{wrapfig}
+  - \usepackage{float}
+  - \usepackage{colortbl}
+  - \usepackage{pdflscape}
+  - \usepackage{tabu}
+  - \usepackage{threeparttable}
+linestretch: '1.5'
+link-citations: yes
+linkcolor: blue
 ---
+
+```{r eval=FALSE, echo=FALSE}
+# Para obtener doble output (html/pdf) correr directamente en consola/terminal:
+rmarkdown::render('index.rmd', output_format = 'all')
+```
 
@@
 knitr::opts_chunk$set(echo = TRUE)
+```
+
+```{r include=FALSE}
+#Determine the output format of the document
+outputFormat   = knitr::opts_knit$get("rmarkdown.pandoc.to")
+
+#Figure and Table Caption Numbering, for HTML do it manually
+capTabNo = 1; capFigNo = 1;
+
+#Function to add the Table Number
+capTab = function(x){
+  if(outputFormat == 'html'){
+    x = paste0("Table ",capTabNo,". ",x)
+    capTabNo <<- capTabNo + 1
+  }; x
+}
+
+#Function to add the Figure Number
+capFig = function(x){
+  if(outputFormat == 'html'){
+    x = paste0("**","Figura ",capFigNo,": ","**",x)
+    capFigNo <<- capFigNo + 1
+  }; x
+}
 ```
@@
 what is considered desirable) and perceptions (i.e.
-how reality is believed to be) (duru-bellat, 2012).
+how reality is believed to be) [@Duru-bellat2012].
 This work focuses on the perceptions of meritocracy.
 
-Although a group of authors has measured the perception of meritocracy as if the perception of inequality of opportunities were the  side opposite of the continuum (Kunovich, 2007; Newman, 2015), others have shown that both concepts are rather different dimensions(Reynolds , 2014).
+Although a group of authors has measured the perception of meritocracy as if the perception of inequality of opportunities were the  side opposite of the continuum [@Kunovich2007; @Newman2015], others have shown that both concepts are rather different dimensions [@Reynolds2014].
 Following this evidence, we chose to model both constructs as distinct latent variables.
@@
 
-Additionally, it is important to mention that we will exploratoryly test the relationship between the “get a head” questions and questions that refer to whether “In Chile, people are rewarded or manage to succeed according to different qualities (i.e., hardworking, talented, with contacts)”.
+Additionally, it is important to mention that we will exploratory analysis the relationship between the “get a head” questions and questions that refer to whether “In Chile, people are rewarded or manage to succeed according to different qualities (i.e., hardworking, talented, with contacts)”.
 
-* H1. The perception of meritocracy is a factor in the drift of the importance attributed to talent and the effort to get ahead in life.
-* H2. The factors of family origin, contacts and discrimination are grouped into a second order factor that relates to non-meritocratic perceptions.
-* H3. Preferences are also grouped into meritocratic and non-meritocratic preferences.
-* H4 (exploratory) the way in which the perception of inequality of opportunities is asked influences the perception declared by the interviewees.
+* H1. The perception of meritocracy is a latent variable in the drift of the importance attributed to talent and the effort to get ahead in life.
+* H2. Non-meritocratic perception is a latent variable that derives from two elements related to the agreement with the statement that people with contacts and rich parents, manage to get ahead
+* H3. Meritocratic preferences behave as a latent variable.
+* H4. Non-meritocratic preferences behave as a latent variable.
 
@@
 
-The present study is of an observational type focused on the measurement of the meritocracy preception variable.
+The present study is of an observational type focused on the measurement of the meritocracy perception variable.
 This study is part of an experimental study, applied online, which will be longitudinal panel type, that is, the same subjects will be surveyed 3 different occasions.
@@
 The subjects participated in these two opportunities are excluded to the central sample, which corresponds to the 3-wave panel study.
-These data from the two pilot studies were used for methodological decisions, descriptive stadistics, correlations and confirmation tests.
+These data from the two pilot studies were used for methodological decisions, descriptive statistics, correlations and confirmation tests.
 Due to the above, although the definitive sample has not yet been produced and therefore has not been explored, it is responsible to say, in addition to the review of previous research and international databases, it has some notions of the relationship among the variables product of the analysis of the previous samples
@@
 # Variables to use:
-The variables used for the perception factor of meritocracy are:
 
-* Para surgir en la vida, ¿cuán importante cree
-usted que es?
-    + Tener ambición
-    + El trabajo duro
+All the variables used in the scale use the following question and answer options
 
-On the other hand, the questions related to inequality of opportunities are:
+* To what extent do you agree or in
+disagree with each of the following
+affirmations?
+    + Strongly disagree
+    + Disagree
+    + Neither agree nor disagree
+    + Agree
+    + Totally agree
 
-* Para surgir en la vida, ¿cuán importante cree
-usted que es?
-    + provenir de una familia rica o con muchos recursos
-    + Tener padres con altos niveles de educación
-    + Conocer a las personas adecuadas
-    + Tener contactos políticos
-    + La raza u origen étnico de las personas
-    + El sexo de las personas, ser hombre o mujer
+```{r include=FALSE}
+# Get output document type and use as the format argument in kable function
+library(knitr)
+library(kableExtra)
 
-The items of preferences are:
+table_format = if(is_html_output()) {
+  "html"
+} else if(is_latex_output()) {
+  "latex"
+}
+```
 
-* ¿En qué medida se encuentra usted de acuerdo o en desacuerdo con cada una de las siguientes afirmaciones?
-    + Quienes más se esfuerzan deberían obtener mayores recompensas que quienes se esfuerzan menos
-    + Quienes poseen más talento deberían obtener mayores recompensas que quienes tienen menos talento
-    + Está bien que quienes tienen padres ricos salgan adelante
-    + Está bien que quienes tienen buenos contactos salgan adelante
+```{r echo=FALSE}
 
-The questions that will be used for the exploratory analysis indicated are:
+tab01 <- read.csv(file = "materials/table01.csv",header = 1,sep = ";")
 
-* ¿En qué medida se encuentra usted de acuerdo o en
-desacuerdo con cada una de las siguientes
-afirmaciones? En Chile,
-    + Quienes más se esfuerzan logran obtener mayores recompensas.
-    + Quienes tienen más talento logran obtener mayores recompensas.
-    + Las personas con padres ricos logran salir adelante.
-    + Los que tienen buenos contactos logran salir adelante.
+cnames <- c("Dimension", "Factor","Statement (english)" , "Statement (spanish)")
+
+caption01 <- capTab("Meritocracy perception and preference")
+kable(tab01,format = table_format,booktabs = T, linesep = "",col.names = cnames, caption = caption01) %>%
+  kable_styling(full_width = T,latex_options = c("hold_position","scale_down"),position = "center",font_size = 9) %>%
+  column_spec(column = 1,width = "1.5cm",) %>%
+  column_spec(column = 2,width = "2 cm") %>%
+  collapse_rows(columns = 1:2,valign = "middle")
+
+```
 
@@
 # Analysis plan
-To evaluate these hypotheses, confirmatory factor analysis will be used, since we have strong theory regarding the ordering of the factors.
+
+To evaluate these hypotheses, confirmatory factor analysis (CFA) will be used, since we have strong theory regarding the ordering of the factors.
 In addition, exploratory factor analysis will be used to evaluate the exploratory hypothesis indicated.
-The Lavaan R package will be used.
+The lavaan [@Rosseel2012] R package will be used.
 
-### Inference criteria,
-The values that will be used as evaluation criteria for the goodness of the adjustment of the model were taken from the proposal of Ruiz, Partdo and San Martin (2010) and are the following:
-*Chi-square:> 0.05
-*Chi-square ratio:> 3
-*Comparative adjustment goodness index (CFI):> 0.95
-*Tucker-Lewis Index (TLI):> 0.95
-*Root of the average squared residual approximation <0.08.
+### Inference criteria
+
+The values that will be used as evaluation criteria for the goodness of the adjustment of the model were taken from the proposal of @Brown2008 and are the following:
+
+* Chi-square:> 0.05
+* Chi-square ratio:> 3
+* Comparative adjustment goodness index (CFI):> 0.95
+* Tucker-Lewis Index (TLI):> 0.95
+* Root of the average squared residual approximation <0.08.
+
+### Secondary analysis
+
+To evaluate the metric stability of the measurement model [@Davidovetal2014] proposed above we will conduct a longitudinal invariance test using the first and third wave of our study.
+Following @liuTestingMeasurementInvariance2017, we will test a series of four hierarchical models: Configural, Weak, Strong and Strict invariance models for ordered-categorical indicators based on the assumtion that a five category likert scale cannot be treated as a continous variable because can lead to biased parameter estimates.
 
@@
 
-### Ethics. JI
+### Ethics.
 The experiment and survey instruments are approved by the IRB of the University of Chile.
 
-# bibliography
-
-Ruiz, M., Pardo, A. & San Martín, R. (2010). Modelos de Ecuaciones Estructurales. -Papeles del psicólogo, 31-(1), pp. 34-45.
-
-Duru-Bellat, M., & Tenret, E. (2012). Who’s for Meritocracy? Individual and Contextual Variations in the Faith. Chicago Journal, 56(2), 223-247.
-
-Kunovich, S., & Slomczynski, K. M. (2007). Systems of Distribution and a Sense of Equity: A Multilevel Analysis of Meritocratic Attitudes in
-Post-industrial Societies. European Sociological Review, 23(5), 649–663. doi:10.1093/esr/jcm026
-
-Newman, B.
-J., Johnston, C.
-D., & Lown, P.
-L.
-(2014).
-False Consciousness or Class Awareness?
-Local Income Inequality, Personal Economic Position, and Belief in American Meritocracy.
-American Journal of Political Science, 59(2), 326-340.
-doi:10.1111/ajps.12153
-
-Reynolds, J., & Xian, H. (2014). Perceptions of meritocracy in the land of opportunity. Research in Social Stratification and Mobility, 36, 121–137. doi:10.1016/j.rssm.2014.03.001
+# Bibliography

### 2019-11-21 code/20190524_escala_merito_non.R

@@ -89,6 +89,8 @@ no_pref =~ pref_famrica  + pref_contact
 no_perc~~no_pref
 '
 
+
+library(lavaan)
 fit1 <- cfa(model = model1,data = data, ordered = c("per_esfuerzo","per_talento",
                                                     "pref_esfuerzo","pref_talento",
                                                     "per_famrica","per_contact" ,
@@ -107,7 +109,7 @@ modelp1 <- semPlot::semPaths(fit1,
                              intercepts = FALSE,
                              thresholds = FALSE)
 
-
+modelp1
 
 
 

### 2019-11-29 analysis-cfa.Rmd
update2911

---
title: "Analysis"
date: "29-11-2019"
output: 
  html_document:
    theme: flatly
    number_sections: true
    toc: true
    toc_float: true
---

```{r eval=FALSE, include=FALSE}
rmarkdown::render(input = "analysis-cfa.Rmd",output_format = "html_document")
```

```{r}
library(sjPlot)
library(dplyr)
library(lavaan)

data01 <- sjlabelled::read_spss(path = "data/dat1.sav",verbose = FALSE)

dat01 <- data01 %>% filter(Intro==1) %>% select(starts_with("meritv01")) %>% na.omit()
dat02 <- data01 %>% filter(Intro==1) %>% select(starts_with("meritv02")) %>% na.omit()
dat03 <- data01 %>% filter(Intro==1) %>% select(starts_with("meritv03_p")) %>% na.omit()

```

# Version 01:

1. Percepcion esfuerzo
2. Percepcion talento
3. Percepcion familia rica
4. Percepcion redes
5. Preferencia esfuerzo
6. Preferencia talento
7. Preferencia familia rica
8. Preferencia redes

```{r,results='hold'}
model01 <- 'perc_merit=~meritv01_perc_effort+meritv01_perc_talent 
            perc_nmerit=~meritv01_perc_wpart+meritv01_perc_netw  
            pref_merit=~meritv01_pref_effort+meritv01_pref_talent 
            pref_nmerit=~meritv01_pref_wpart+meritv01_pref_netw'

fit1 <- cfa(model = model01,data = dat01)

summary(fit1,standardized=TRUE, fit.measures=TRUE)

semPlot::semPaths(object = fit1,what = "std")
```

# Version 02:

1. Percepcion esfuerzo
2. Preferencia esfuerzo
3. Percepcion talento
4. Preferencia talento
5. Percepcion familia rica
6. Preferencia familia rica
7. Percepcion redes
8. Preferencia redes

```{r,results='hold'}
model02 <- 'perc_merit=~meritv02_perc_effort+meritv02_perc_talent 
            perc_nmerit=~meritv02_perc_wpart+meritv02_perc_netw  
            pref_merit=~meritv02_pref_effort+meritv02_pref_talent 
            pref_nmerit=~meritv02_pref_wpart+meritv02_pref_netw'

fit2 <- cfa(model = model02,data = dat02)

summary(fit2,standardized=TRUE, fit.measures=TRUE)

semPlot::semPaths(object = fit2,what = "std")
```

# Version 03: batería orden aleatorio

```{r,results='hold'}
model03 <- 'perc_merit=~meritv03_perc_effort+meritv03_perc_talent 
            perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw  
            pref_merit=~meritv03_pref_effort+meritv03_pref_talent 
            pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw'

fit3 <- cfa(model = model03,data = dat03)

summary(fit3,standardized=TRUE, fit.measures=TRUE)

semPlot::semPaths(object = fit3,what = "std")
```

# Version 04:  Modelo muestra completa

```{r,results='hold'}
dat04 <- data01 %>% filter(Intro==1) %>% select(starts_with("meritv01"),starts_with("meritv02"),starts_with("meritv03_p"))

dat04$perc_effort <- rowSums(dat04[,c(matches(match = "perc_effort",vars = names(dat04)))],na.rm = TRUE)
dat04$perc_talent <- rowSums(dat04[,c(matches(match = "perc_talent",vars = names(dat04)))],na.rm = TRUE)
dat04$perc_wpart  <- rowSums(dat04[,c(matches(match = "perc_wpart" ,vars = names(dat04)))],na.rm = TRUE)
dat04$perc_netw   <- rowSums(dat04[,c(matches(match = "perc_netw"  ,vars = names(dat04)))],na.rm = TRUE)

dat04$pref_effort <- rowSums(dat04[,c(matches(match = "pref_effort",vars = names(dat04)))],na.rm = TRUE)
dat04$pref_talent <- rowSums(dat04[,c(matches(match = "pref_talent",vars = names(dat04)))],na.rm = TRUE)
dat04$pref_wpart  <- rowSums(dat04[,c(matches(match = "pref_wpart" ,vars = names(dat04)))],na.rm = TRUE)
dat04$pref_netw   <- rowSums(dat04[,c(matches(match = "pref_netw"  ,vars = names(dat04)))],na.rm = TRUE)

model04 <- 'perc_merit=~perc_effort+perc_talent 
            perc_nmerit=~perc_wpart+perc_netw  
            pref_merit=~pref_effort+pref_talent 
            pref_nmerit=~pref_wpart+pref_netw'

fit4 <- cfa(model = model04,data = dat04)

summary(fit4,standardized=TRUE, fit.measures=TRUE)

semPlot::semPaths(object = fit4,what = "std")
```

### 2019-11-29 index.Rmd
update2911

@@
 ---
-title: "Meritocracy Scale: Preferences and Perceptions"
+title: "Perceptions and Preferences for Meritocracy Scale (PPM-S)"
 csl: apa6.csl
@@
 
-* How valid is a measurement model for the perception of 4-factor meritocracy, which is grouped into: a first-order factor related to meritocracy and a second-order factor referred to inequalities of opportunity?
+Is it possible to identify and measure perceptions and preferences for meritocracy as two distinct but related factors?
 
-# General description and hypothesis
+# General description and hypotheses
 
-To measure the ideals of meritocracy, it is essential to differentiate preferences (i.e.
-what is considered desirable) and perceptions (i.e.
-how reality is believed to be) [@Duru-bellat2012].
-This work focuses on the perceptions of meritocracy.
+This research builds upon previous studies attempting to develop a comprehensive approach to measuring meritocratic beliefs [@Duru-bellat2012, @Kunovich2007, @Newman2015, @Reynolds2014].
+The proposed measurement model is based on two axis of analysis, as depicted in Table 1.
+The first one distinguishes between different types of "beliefs", using instead the terms perceptions and preferences for meritocracy.
+Perceptions refers to the extent to which people see meritocracy working in their society, which in terms of measurement relates to the "reasons to get ahead" battery, whereas preferences refer to normative expectations that are usually linked to a "should" expression (e.g.
+whether hard work should be related to payment).
+The second axis consider the distinction between meritocratic and non-meritocratic dimensions.
+This aspect has been usually treated as different ends of a same continuum in part of the previous research, an assumption that requieres empirical scrutiny.
+These non-meritocratic elements usually refer to the use of personal contacts or family advantages to get ahead in life.
 
-Although a group of authors has measured the perception of meritocracy as if the perception of inequality of opportunities were the  side opposite of the continuum [@Kunovich2007; @Newman2015], others have shown that both concepts are rather different dimensions [@Reynolds2014].
-Following this evidence, we chose to model both constructs as distinct latent variables.
-Different types of inequalities such as origin, contact or discrimination, cannot be considered in a single factor, since they refer to different realities and perceptions, therefore, it is consistent that they are independent latent variables.
-We also consider the existence of a latent second order variable, referring to the belief of inequality of opportunities in general.
+Table 1: Model of perceptions and preferences for meritocracy and non-meritocracy
 
-Consistently, we consider that questions regarding meritocratic preferences will be grouped into two factors, meritocratic preferences and non-meritocratic preferences.
+|                  	  | Perceptions 	| Preferences 	|
+|:-----------------	  |-------------	|-------------	|
+| _Meritocracy_     	|             	|             	|
+| _Non-meritocracy_ 	|             	|             	|
 
-Additionally, it is important to mention that we will exploratory analysis the relationship between the “get a head” questions and questions that refer to whether “In Chile, people are rewarded or manage to succeed according to different qualities (i.e., hardworking, talented, with contacts)”.
+**Hypotheses**
 
-* H1. The perception of meritocracy is a latent variable in the drift of the importance attributed to talent and the effort to get ahead in life.
-* H2. Non-meritocratic perception is a latent variable that derives from two elements related to the agreement with the statement that people with contacts and rich parents, manage to get ahead
-* H3. Meritocratic preferences behave as a latent variable.
-* H4. Non-meritocratic preferences behave as a latent variable.
+* H1. The perception of meritocracy is a latent variable based on indicators of the importance attributed to talent and the effort to get ahead in life.
+* H2. The non-meritocratic perception is a latent variable that derives from two indicators related to the agreement with the statement that people with contacts and rich parents manage to get ahead.
+* H3. Meritocratic preferences behave as the latent variable based on a normative value of effort and talent.
+* H4. Non-meritocratic preferences behave as a latent variable based on the normative value of the use of personal contacts and having rich parents.
 
@@
 
-The present study is of an observational type focused on the measurement of the meritocracy perception variable.
-This study is part of an experimental study, applied online, which will be longitudinal panel type, that is, the same subjects will be surveyed 3 different occasions.
-The subjects will be randomly selected to be assigned to a placebo or to either of the two experimental procedures (change of poverty in Chile or international comparison of inequality)
-To access the cases we turn to the Netquest survey company, which has coverage in several Latin American countries.
+The scale will be included in an online three-wave panel survey study.
+This study covers among other things an experiment of the effect of information about poverty and inequality on opportunity beliefs.
+The scale will be presented in wave 1 and wave 3.
 
@@
 
-The Sample that will be necessary for this study has not yet been produced, nor has it begun its field work.
-However, in order to test the online questionnaire and the functionality of the quota offered by the Qualtrics platform, two previous studies have been carried out.
-The subjects participated in these two opportunities are excluded to the central sample, which corresponds to the 3-wave panel study.
-These data from the two pilot studies were used for methodological decisions, descriptive statistics, correlations and confirmation tests.
-Due to the above, although the definitive sample has not yet been produced and therefore has not been explored, it is responsible to say, in addition to the review of previous research and international databases, it has some notions of the relationship among the variables product of the analysis of the previous samples
+The data will come from a representative sample of the large cities of Chile (Gran Santiago, Valparaíso-Viña del Mar, Concepción and Antofagasta).
+In order to ensure representativeness, quotas for sex, education and age will be provided to the agency in charge of the online fieldwork.
 
@@
 
-A representative sample of the large cities of Chile (Gran Santiago, Valparaíso-Viña del Mar, Concepción and Antofagasta) is expected, which will have an initial number for the first wave of 2100 cases, which are expected to be at least 1800 in the second wave.
-From the second to the third wave it has been considered as “free fall” that is, we are not sure of the retention rate of the interviewees.
-The sample size was calculated from power analysis that used information from the pilot study that was applied in June 2019.
+The sample will have an initial number for the first wave of 2100 cases, which are expected to be at least 1800 in the second wave.
+From the second to the third wave it has been considered as “free fall”, from which we expect to retain about 1500 cases.
+
+The sample size was calculated from power analysis that used information from a pilot study that was applied in June 2019.
 This pilot study consisted of an online experiment applied to 949 subjects, randomly separating participants to two stimuli and a placebo, leaving a third of the sample in each condition.
@@
 ### Data collection procedure:
-To access this target population, a company has been hired dedicated to the online surveys that will be in charge of getting the respondents.
-These will be invited to participate in the survey under the incentive provided by the external company.
+
+Respondents will be invited to participate in the online survey under the incentive provided by the external company.
 Respondents have 4 business days to answer the survey, with a maximum of 20 minutes per survey.
-In order to achieve greater representativeness of the sample, the quota method will be used, that is, the program will only allow respondents to respond while more people with their demographic characteristics are required.
+In order to achieve greater representativeness of the sample, the quota method will be used, that is, the program will only allow respondents with the required demographic characteristics.
 The quotas used were age, sex and education.
-It should be noted that the time between wave and wave is between 7 and 9 days.
+It should be noted that the time between wave and wave is around 7 and 9 days.
 
 ### Stop Rule:
-The rule of detention is to reach the number of respondents indicated, adjusted to the corresponding quotas.
-However, there may be modifications depending on the contingencies of the field work, depending on the response rates.
+
+The rule of detention is to reach the number of respondents indicated in the quotas. However, there may be modifications depending on the contingencies of the field work and the response rates.
 
@@
 
-All the variables used in the scale use the following question and answer options
+All the variables used in the scale use the following question and answer options:
 
-* To what extent do you agree or in
-disagree with each of the following
-affirmations?
+* To what extent do you agree or disagree with each of the following affirmations?
+  - Response scale:
     + Strongly disagree
@@
 
-The variables will be worked on the same scale in which they were asked, without resorting to manipulations of the variables to assess the validity of the measurement model.
+There will not be variable manipulation.
 
@@
 
-The investigation of the validity of this scale does not resort to indices intended to represent the perception of meritocracy in an observed way, but rather confirmatory factor analysis will be used to estimate the value of the latent variables underlying the indicators.
+The investigation of the validity of this scale does not rely on indices but rather confirmatory factor analysis will be used to estimate the value of the latent variables underlying the indicators.
 
@@
 ### Randomization
-To avoid the existence of biases in the order of the questions, they have been randomized within a thematic block of questions, so, randomly, people answer questions in different orders regarding their perception of meritocracy.
+
+To test the items' order effects we present the items in three different order conditions which will be assigned randomly. The conditions are the following:
+
+- Order 1: the order in which items are presented in Table 1
+- Order 2: the items are ordered based on topics (work harder, talent, rich parents, connections), and for each of them they have to respond first about perceptions and then about preferences.
+For instance: a) those who work harder get greater rewards, b) those who work arder should get greater rewards ...
+- Order 3: the order is randomized within this condition
 
@@
 
-To evaluate these hypotheses, confirmatory factor analysis (CFA) will be used, since we have strong theory regarding the ordering of the factors.
-In addition, exploratory factor analysis will be used to evaluate the exploratory hypothesis indicated.
-The lavaan [@Rosseel2012] R package will be used.
+To evaluate the hypotheses, confirmatory factor analysis (CFA) will be used as we rely on a theory regarding the underlying four factors.
+The method of estimation will be Weighted Least Squares Mean Variance Approximation (WLMSV) for categorical ordered indicators.
+For the analysis we will use the R package Lavaan [@Rosseel2012].
 
@@
 
-To evaluate the metric stability of the measurement model [@Davidovetal2014] proposed above we will conduct a longitudinal invariance test using the first and third wave of our study.
-Following @liuTestingMeasurementInvariance2017, we will test a series of four hierarchical models: Configural, Weak, Strong and Strict invariance models for ordered-categorical indicators based on the assumtion that a five category likert scale cannot be treated as a continous variable because can lead to biased parameter estimates.
+To evaluate the metric stability of the measurement model [@Davidovetal2014] we will conduct a longitudinal invariance test using the first and third wave of our study.
+Following @liuTestingMeasurementInvariance2017, we will test a series of four hierarchical models: Configural, Weak, Strong and Strict invariance models for ordered-categorical indicators based on the assumption that a five category likert scale cannot be treated as a continous variable because can lead to biased parameter estimates.
 
 ### Data Exclusion
+
 All cases will be used as long as they do not show missing values in any of the items on the scale. No imputation criteria will be used.
@@
 ### Ethics.
+
 The experiment and survey instruments are approved by the IRB of the University of Chile.

### 2019-11-29 code/20190524_escala_merito_non.R

@@ -1,117 +1,0 @@
-# Merit scale -------------------------------------------------------------
-
-# install.packages("haven","dplyr","sjPlot","psych","lavaan","here")
-
-library(haven)
-library(dplyr)
-library(sjPlot)
-library(psych)
-library(lavaan)
-library(here)
-rm(list=ls())
-setwd(here())
-
-source("code/recode_tratamientos.R") # recode original dataset 
-load(file = "data/pob_04.RData")
-
-#Nota: Quedarme solamente con los datos completos para los ítems de la Escala
-
-# -99 y -999 == NA --------------------------------------------------------
-
-
-for (i in 1:ncol(data)) {
-data[,i][data[,i] == c(-99)]  <- NA #Missing 
-data[,i][data[,i] == c(-999)] <- NA #Missing  
-}
-
-summary(data)
-names(data)
-
-sjPlot::view_df(data)
-
-# Nombres sustantivos variables -------------------------------------------
-data <- rename(data,
-               per_esfuerzo =merit_perc_pref_1,
-               per_talento  =merit_perc_pref_2,
-               per_famrica  =merit_perc_pref_3,
-               per_contact  =merit_perc_pref_4,
-               pref_esfuerzo=merit_perc_pref_5,
-               pref_talento =merit_perc_pref_13,
-               pref_famrica =merit_perc_pref_14,
-               pref_contact =merit_perc_pref_15,
-               reduc=pref_redis_taxB_1,
-               rsalu=pref_redis_taxB_2,
-               rpens=pref_redis_taxB_3,
-               rdesi=pref_redis_taxB_4)
-
-data$per_esfuerzo <- as.numeric(data$per_esfuerzo )
-data$per_talento  <- as.numeric(data$per_talento  )
-data$per_famrica  <- as.numeric(data$per_famrica  )
-data$per_contact  <- as.numeric(data$per_contact  )
-data$pref_esfuerzo<- as.numeric(data$pref_esfuerzo)
-data$pref_talento <- as.numeric(data$pref_talento )
-data$pref_famrica <- as.numeric(data$pref_famrica )
-data$pref_contact <- as.numeric(data$pref_contact )
-
-# Brechas -----------------------------------------------------------------
-data$gap_per <- as.numeric(data$salary_perc_1/data$salary_perc_2)
-data$gap_jus <- as.numeric(data$salary_just_1/data$salary_just_2)
-data$lngap_per <- as.numeric(log(data$salary_perc_1/data$salary_perc_2))
[…]

### 2019-11-29 analysis-cfa.Rmd
update2911

@@
 title: "Analysis"
-date: "29-11-2019"
+date: "`r format(Sys.time(), '%d %B, %Y')`"
 output: 
@@
 ---
+
+# Análisis PPM-S
 
@@
 
-# Version 01:
+## Version 01:
 
@@
 
-fit1 <- cfa(model = model01,data = dat01)
+fit1 <- cfa(model = model01,data = dat01,ordered = c("meritv01_perc_effort","meritv01_perc_talent",
+                                                     "meritv01_perc_wpart","meritv01_perc_netw",
+                                                     "meritv01_pref_effort","meritv01_pref_talent",
+                                                     "meritv01_pref_wpart","meritv01_pref_netw"))
 
 summary(fit1,standardized=TRUE, fit.measures=TRUE)
-
-semPlot::semPaths(object = fit1,what = "std")
+semPlot::semPaths(object = fit1,what = "std",thresholds = FALSE, intercepts = FALSE)
 ```
 
-# Version 02:
+## Version 02:
 
@@
 
-fit2 <- cfa(model = model02,data = dat02)
+fit2 <- cfa(model = model02,data = dat02,ordered = c("meritv02_perc_effort","meritv02_perc_talent",
+                                                     "meritv02_perc_wpart","meritv02_perc_netw",
+                                                     "meritv02_pref_effort","meritv02_pref_talent",
+                                                     "meritv02_pref_wpart","meritv02_pref_netw"))
 
 summary(fit2,standardized=TRUE, fit.measures=TRUE)
-
 semPlot::semPaths(object = fit2,what = "std")
@@
 
-# Version 03: batería orden aleatorio
+## Version 03: orden aleatorio
 
@@
 
-fit3 <- cfa(model = model03,data = dat03)
+fit3 <- cfa(model = model03,data = dat03,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
+                                                     "meritv03_perc_wpart","meritv03_perc_netw",
+                                                     "meritv03_pref_effort","meritv03_pref_talent",
+                                                     "meritv03_pref_wpart","meritv03_pref_netw"))
 
 summary(fit3,standardized=TRUE, fit.measures=TRUE)
-
-semPlot::semPaths(object = fit3,what = "std")
+semPlot::semPaths(object = fit3,what = "std",thresholds = FALSE, intercepts = FALSE)
 ```
 
-# Version 04:  Modelo muestra completa
+## Version 04: muestra completa
 
@@
 fit4 <- cfa(model = model04,data = dat04)
-
 summary(fit4,standardized=TRUE, fit.measures=TRUE)
-
-semPlot::semPaths(object = fit4,what = "std")
+semPlot::semPaths(object = fit4,what = "std",thresholds = FALSE, intercepts = FALSE)
 ```

### 2019-12-12 code/20190524_escala_merito_non.R
update1312

@@ -108,8 +108,3 @@ modelp1 <- semPlot::semPaths(fit1,
                              thresholds = FALSE)
 
 
-
-
-
-
-

### 2019-12-12 about.Rmd

---
title: "Otro"
author: "Julio César Iturra Sanhueza - jciturra@uc.cl"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2020-01-09 index.Rmd
update0901-2020

@@
 
-This research builds upon previous studies attempting to develop a comprehensive approach to measuring meritocratic beliefs [@Duru-bellat2012, @Kunovich2007, @Newman2015, @Reynolds2014].
+This research builds upon previous studies attempting to develop a comprehensive approach to measuring meritocratic beliefs [@Duru-bellat2012; @Kunovich2007; @Newman2015; @Reynolds2014].
 The proposed measurement model is based on two axis of analysis, as depicted in Table 1.

### 2020-01-10 analysis-cfa.Rmd
update0901-2020
Análisis Factorial de Escala actualizado con base de datos 09 enero 2020
commentarios primera versión

@@
 
-data01 <- sjlabelled::read_spss(path = "data/dat1.sav",verbose = FALSE)
+data01 <- sjlabelled::read_spss(path = "data/Estudio_3_ola1_January_9_2020.sav",verbose = FALSE)
 
@@
 
-## Version 01:
+## Version 01: Fixed order by perception / preferences (N=  )
 
@@
 
-## Version 02:
+*summary*: good fit, heywood case item 8.
+
+## Version 02: Fixed order, by topic (N=  )
 
@@
 ```
+*summary*: bad fit, heywood
 
-## Version 03: orden aleatorio
+## Version 03: Random order
 
@@
 
-## Version 04: muestra completa
+*summary*: fit not very goood (RMSEA .7), no heywood. Try tau-equivalence
+
+## Version 04: muestra completa (N=  )
 

### 2020-01-10 index.Rmd
update1001-2020
IPO version of merit-scale

@@
 title: "Perceptions and Preferences for Meritocracy Scale (PPM-S)"
-csl: apa6.csl
-bibliography: zlibrary.bib
-css: style.css
+csl: "input/bib/apa6.csl"
+bibliography: "input/bib/zlibrary.bib"
+css: tyle.css
 output:
@@
 
-tab01 <- read.csv(file = "materials/table01.csv",header = 1,sep = ";")
+tab01 <- read.csv(file = "input/data/original/materials/table01.csv",header = 1,sep = ";")
 

### 2020-01-10 production/analysis-cfa.Rmd
update1001-2020
IPO version of merit-scale

--- 2019-11-29 analysis-cfa.Rmd
@@
     toc_float: true
+editor_options: 
+  chunk_output_type: console
 ---
@@
 ```{r eval=FALSE, include=FALSE}
-rmarkdown::render(input = "analysis-cfa.Rmd",output_format = "html_document")
+rmarkdown::render(input = "production/analysis-cfa.Rmd",output_format = "html_document",output_dir = ".")
 ```
@@
 
-data01 <- sjlabelled::read_spss(path = "data/dat1.sav",verbose = FALSE)
+data01 <- sjlabelled::read_spss(path = "../input/data/original/Estudio_3_ola1_January_9_2020.sav",verbose = FALSE)
 
-dat01 <- data01 %>% filter(Intro==1) %>% select(starts_with("meritv01")) %>% na.omit()
-dat02 <- data01 %>% filter(Intro==1) %>% select(starts_with("meritv02")) %>% na.omit()
-dat03 <- data01 %>% filter(Intro==1) %>% select(starts_with("meritv03_p")) %>% na.omit()
+dat01 <- data01 %>% filter(Intro==1,Finished==1) %>% select(starts_with("meritv01")) %>% na.omit()
+dat02 <- data01 %>% filter(Intro==1,Finished==1) %>% select(starts_with("meritv02")) %>% na.omit()
+dat03 <- data01 %>% filter(Intro==1,Finished==1) %>% select(starts_with("meritv03_p")) %>% na.omit()
+
+# - Nos quedamos solamente con los casos que ACEPTARON participar y terminaron el cuestionario.
+# - Para dejar una base que contenga solamente los casos que tienen información completa en los ítems usamos na.omit()
+
+dat04 <- data01 %>% filter(Intro==1) %>% select(starts_with("meritv01"),starts_with("meritv02"),starts_with("meritv03_p"))
+
+dat04$perc_effort <- rowSums(dat04[,c(matches(match = "perc_effort",vars = names(dat04)))],na.rm = TRUE)
+dat04$perc_talent <- rowSums(dat04[,c(matches(match = "perc_talent",vars = names(dat04)))],na.rm = TRUE)
+dat04$perc_wpart  <- rowSums(dat04[,c(matches(match = "perc_wpart" ,vars = names(dat04)))],na.rm = TRUE)
+dat04$perc_netw   <- rowSums(dat04[,c(matches(match = "perc_netw"  ,vars = names(dat04)))],na.rm = TRUE)
+
+dat04$pref_effort <- rowSums(dat04[,c(matches(match = "pref_effort",vars = names(dat04)))],na.rm = TRUE)
+dat04$pref_talent <- rowSums(dat04[,c(matches(match = "pref_talent",vars = names(dat04)))],na.rm = TRUE)
+dat04$pref_wpart  <- rowSums(dat04[,c(matches(match = "pref_wpart" ,vars = names(dat04)))],na.rm = TRUE)
+dat04$pref_netw   <- rowSums(dat04[,c(matches(match = "pref_netw"  ,vars = names(dat04)))],na.rm = TRUE)
 
@@
 
-## Version 01:
+## Version 01: Fixed order by perception / preferences (N=`r dim(dat01)[1]`)
 
@@
 
-fit1 <- cfa(model = model01,data = dat01,ordered = c("meritv01_perc_effort","meritv01_perc_talent",
+fit1 <- cfa(model = model01,data = dat01,
+            ordered = c("meritv01_perc_effort","meritv01_perc_talent",
                         "meritv01_perc_wpart","meritv01_perc_netw",
@@
 
-## Version 02:
+*summary*: good fit, heywood case item 8.
+
+## Version 02: Fixed order, by topic (N=`r dim(dat02)[1]`)
 
@@
 summary(fit2,standardized=TRUE, fit.measures=TRUE)
-semPlot::semPaths(object = fit2,what = "std")
+semPlot::semPaths(object = fit2,what = "std",thresholds = FALSE, intercepts = FALSE)
 ```
+*summary*: bad fit, heywood
 
-## Version 03: orden aleatorio
+## Version 03: Random order (N=`r dim(dat03)[1]`)
 
@@
 
-## Version 04: muestra completa
+*summary*: fit not very goood (RMSEA .7), no heywood. Try tau-equivalence
+
+## Version 04: muestra completa (N=`r dim(dat04)[1]`)
+
+### Asumiendo items continuos
 
 ```{r,results='hold'}
-dat04 <- data01 %>% filter(Intro==1) %>% select(starts_with("meritv01"),starts_with("meritv02"),starts_with("meritv03_p"))
+model04 <- 'perc_merit=~ perc_effort+perc_talent 
+            perc_nmerit=~perc_wpart +perc_netw  
+            pref_merit=~ pref_effort+pref_talent 
+            pref_nmerit=~pref_wpart +pref_netw'
 
-dat04$perc_effort <- rowSums(dat04[,c(matches(match = "perc_effort",vars = names(dat04)))],na.rm = TRUE)
-dat04$perc_talent <- rowSums(dat04[,c(matches(match = "perc_talent",vars = names(dat04)))],na.rm = TRUE)
-dat04$perc_wpart  <- rowSums(dat04[,c(matches(match = "perc_wpart" ,vars = names(dat04)))],na.rm = TRUE)
-dat04$perc_netw   <- rowSums(dat04[,c(matches(match = "perc_netw"  ,vars = names(dat04)))],na.rm = TRUE)
+fit4 <- cfa(model = model04,
+            data = dat04, estimator="MLR")
 
-dat04$pref_effort <- rowSums(dat04[,c(matches(match = "pref_effort",vars = names(dat04)))],na.rm = TRUE)
-dat04$pref_talent <- rowSums(dat04[,c(matches(match = "pref_talent",vars = names(dat04)))],na.rm = TRUE)
-dat04$pref_wpart  <- rowSums(dat04[,c(matches(match = "pref_wpart" ,vars = names(dat04)))],na.rm = TRUE)
-dat04$pref_netw   <- rowSums(dat04[,c(matches(match = "pref_netw"  ,vars = names(dat04)))],na.rm = TRUE)
-
-model04 <- 'perc_merit=~perc_effort+perc_talent 
-            perc_nmerit=~perc_wpart+perc_netw  
-            pref_merit=~pref_effort+pref_talent 
-            pref_nmerit=~pref_wpart+pref_netw'
-
-fit4 <- cfa(model = model04,data = dat04)
 summary(fit4,standardized=TRUE, fit.measures=TRUE)
+fitmeasures(fit4)[c("cfi","rmsea.robust")]
 semPlot::semPaths(object = fit4,what = "std",thresholds = FALSE, intercepts = FALSE)
 ```
+
+### Asumiendo items ordenados
+
+```{r}
+model04or <- 'perc_merit=~ perc_effort+perc_talent 
+              perc_nmerit=~perc_wpart +perc_netw  
+              pref_merit=~ pref_effort+pref_talent 
+              pref_nmerit=~pref_wpart +pref_netw'
+
+fit4or <- cfa(model = model04or,
+            data = dat04,
+            ordered = c("perc_effort","perc_talent","perc_wpart","perc_netw",
+                        "pref_effort","pref_talent","pref_wpart","pref_netw"))
+
+summary(fit4or,standardized=TRUE, fit.measures=TRUE)
+fitmeasures(fit4or)[c("cfi","rmsea.robust","rmsea.scaled")]
+semPlot::semPaths(object = fit4or,what = "std",thresholds = FALSE, intercepts = FALSE)
+```
+
+### Test V04 Tau equivalente (ordered)
+
+```{r,results='hold'}
+model04.tau01 <- 'perc_merit=~ r1*perc_effort+r1*perc_talent 
+                  perc_nmerit=~perc_wpart +perc_netw  
+                  pref_merit=~ pref_effort+pref_talent 
+                  pref_nmerit=~pref_wpart +pref_netw'
+
+fit4.t01 <- cfa(model = model04.tau01,
+                data = dat04,
+                ordered = c("perc_effort","perc_talent","perc_wpart","perc_netw",
+                            "pref_effort","pref_talent","pref_wpart","pref_netw"))
+summary(fit4.t01,standardized=TRUE, fit.measures=TRUE)
+semPlot::semPaths(object = fit4.t01,what = "std",thresholds = FALSE, intercepts = FALSE)
+fitmeasures(fit4.t01)[c("cfi","rmsea.robust","rmsea.scaled")]
+```
+
+```{r}
+anova(fit4or,fit4.t01)
+```

### 2020-01-10 code/20190524_escala_merito_non.R

@@ -1,110 +1,0 @@
-# Merit scale -------------------------------------------------------------
-
-# install.packages("haven","dplyr","sjPlot","psych","lavaan","here")
-
-library(haven)
-library(dplyr)
-library(sjPlot)
-library(psych)
-library(lavaan)
-library(here)
-rm(list=ls())
-setwd(here())
-
-source("code/recode_tratamientos.R") # recode original dataset 
-load(file = "data/pob_04.RData")
-
-#Nota: Quedarme solamente con los datos completos para los ítems de la Escala
-
-# -99 y -999 == NA --------------------------------------------------------
-
-
-for (i in 1:ncol(data)) {
-data[,i][data[,i] == c(-99)]  <- NA #Missing 
-data[,i][data[,i] == c(-999)] <- NA #Missing  
-}
-
-summary(data)
-names(data)
-
-sjPlot::view_df(data)
-
-# Nombres sustantivos variables -------------------------------------------
-data <- rename(data,
-               per_esfuerzo =merit_perc_pref_1,
-               per_talento  =merit_perc_pref_2,
-               per_famrica  =merit_perc_pref_3,
-               per_contact  =merit_perc_pref_4,
-               pref_esfuerzo=merit_perc_pref_5,
-               pref_talento =merit_perc_pref_13,
-               pref_famrica =merit_perc_pref_14,
-               pref_contact =merit_perc_pref_15,
-               reduc=pref_redis_taxB_1,
-               rsalu=pref_redis_taxB_2,
-               rpens=pref_redis_taxB_3,
-               rdesi=pref_redis_taxB_4)
-
-data$per_esfuerzo <- as.numeric(data$per_esfuerzo )
-data$per_talento  <- as.numeric(data$per_talento  )
-data$per_famrica  <- as.numeric(data$per_famrica  )
-data$per_contact  <- as.numeric(data$per_contact  )
-data$pref_esfuerzo<- as.numeric(data$pref_esfuerzo)
-data$pref_talento <- as.numeric(data$pref_talento )
-data$pref_famrica <- as.numeric(data$pref_famrica )
-data$pref_contact <- as.numeric(data$pref_contact )
-
-# Brechas -----------------------------------------------------------------
-data$gap_per <- as.numeric(data$salary_perc_1/data$salary_perc_2)
-data$gap_jus <- as.numeric(data$salary_just_1/data$salary_just_2)
-data$lngap_per <- as.numeric(log(data$salary_perc_1/data$salary_perc_2))
[…]

### 2020-01-10 index.Rmd
update1001-2020

@@
 # Para obtener doble output (html/pdf) correr directamente en consola/terminal:
-rmarkdown::render('index.rmd', output_format = 'all')
+rmarkdown::render('production/index.rmd', output_format = 'all',output_dir = ".")
 ```

### 2020-01-13 index.Rmd
update1301-2010

@@
 bibliography: "input/bib/zlibrary.bib"
-css: tyle.css
+css: style.css
 output:

### 2020-01-13 production/analysis-cfa.Rmd

@@
 date: "`r format(Sys.time(), '%d %B, %Y')`"
+css: ../style.css
 output: 
@@
   chunk_output_type: console
+
 ---
-
 # Análisis PPM-S
@@
 library(lavaan)
+library(stargazer)
 
@@
 
-dat04 <- data01 %>% filter(Intro==1) %>% select(starts_with("meritv01"),starts_with("meritv02"),starts_with("meritv03_p"))
-
+dat04 <- data01 %>% filter(Intro==1) %>%
+  select(starts_with("meritv01"),
+         starts_with("meritv02"),
+         starts_with("meritv03_p")) 
 dat04$perc_effort <- rowSums(dat04[,c(matches(match = "perc_effort",vars = names(dat04)))],na.rm = TRUE)
@@
 8. Preferencia redes
+
+```{r echo=FALSE,results='asis'}
+stargazer(dat01,type = "html")
+```
 
@@
 
+```{r echo=FALSE,results='asis'}
+stargazer(dat02,type = "html")
+```
+
 ```{r,results='hold'}
@@
 ## Version 03: Random order (N=`r dim(dat03)[1]`)
+
+```{r echo=FALSE,results='asis'}
+stargazer(dat03,type = "html")
+```
 
@@
 ## Version 04: muestra completa (N=`r dim(dat04)[1]`)
+
+```{r echo=FALSE,results='asis'}
+stargazer(dat04,type = "html")
+```
 
@@
 
-```{r,results='hold'}
+```{r, results='hold'}
 model04.tau01 <- 'perc_merit=~ r1*perc_effort+r1*perc_talent 

### 2020-03-19 production/analysis-cfa.Rmd

--- 2020-01-10 analysis-cfa.Rmd
@@
 date: "`r format(Sys.time(), '%d %B, %Y')`"
+css: ../style.css
 output: 
   html_document:
-    theme: flatly
     number_sections: true
@@
     toc_float: true
+    toc_depth: 4
+    code_folding: "hide"
+editor_options: 
+  chunk_output_type: console
 ---
@@
 
+# Wave 01
+
 ```{r eval=FALSE, include=FALSE}
-rmarkdown::render(input = "analysis-cfa.Rmd",output_format = "html_document")
+rmarkdown::render(input = "production/analysis-cfa.Rmd",output_format = "html_document")
+```
+
+```{r include=FALSE}
+knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = TRUE)
+knitr::opts_knit$set(root.dir = "../")
+options(scipen=999)
+rm(list=ls())
+options(knitr.kable.NA = '')
 ```
@@
 library(lavaan)
-
-data01 <- sjlabelled::read_spss(path = "data/Estudio_3_ola1_January_9_2020.sav",verbose = FALSE)
-
-dat01 <- data01 %>% filter(Intro==1) %>% select(starts_with("meritv01")) %>% na.omit()
-dat02 <- data01 %>% filter(Intro==1) %>% select(starts_with("meritv02")) %>% na.omit()
-dat03 <- data01 %>% filter(Intro==1) %>% select(starts_with("meritv03_p")) %>% na.omit()
-
+library(stargazer)
+library(corrplot)
+library(psych)
+library(knitr)
+library(kableExtra)
 ```
 
-## Version 01: Fixed order by perception / preferences (N=  )
+```{r}
+data01 <- sjlabelled::read_spss(path = "input/data/original/Estudio_3_ola1.sav",verbose = FALSE)
+
+dat01 <- data01 %>% filter(Intro==1,Finished==1) %>% select(starts_with("meritv01")) %>% na.omit()
+dat02 <- data01 %>% filter(Intro==1,Finished==1) %>% select(starts_with("meritv02")) %>% na.omit()
+dat03 <- data01 %>% filter(Intro==1,Finished==1) %>% select(starts_with("meritv03_p")) %>% na.omit()
+
+# - Nos quedamos solamente con los casos que ACEPTARON participar y terminaron el cuestionario.
+# - Para dejar una base que contenga solamente los casos que tienen información completa en los ítems usamos na.omit()
+
+dat04 <- data01 %>% filter(Intro==1) %>%
+  select(starts_with("meritv01"),
+         starts_with("meritv02"),
+         starts_with("meritv03_p"),
+         starts_with("FL_21_DO")) 
+dat04$grupo <- NA
+dat04$grupo[dat04$FL_21_DO_merit_perc_pref_julio19v01==1] <- 1
+dat04$grupo[dat04$FL_21_DO_merit_perc_pref_julio19v02==1] <- 2
+dat04$grupo[dat04$FL_21_DO_merit_perc_pref_julio19v03==1] <- 3
+dat04$perc_effort <- rowSums(dat04[,c(matches(match = "perc_effort",vars = names(dat04)))],na.rm = TRUE)
+dat04$perc_talent <- rowSums(dat04[,c(matches(match = "perc_talent",vars = names(dat04)))],na.rm = TRUE)
+dat04$perc_wpart  <- rowSums(dat04[,c(matches(match = "perc_wpart" ,vars = names(dat04)))],na.rm = TRUE)
+dat04$perc_netw   <- rowSums(dat04[,c(matches(match = "perc_netw"  ,vars = names(dat04)))],na.rm = TRUE)
+
+dat04$pref_effort <- rowSums(dat04[,c(matches(match = "pref_effort",vars = names(dat04)))],na.rm = TRUE)
+dat04$pref_talent <- rowSums(dat04[,c(matches(match = "pref_talent",vars = names(dat04)))],na.rm = TRUE)
+dat04$pref_wpart  <- rowSums(dat04[,c(matches(match = "pref_wpart" ,vars = names(dat04)))],na.rm = TRUE)
+dat04$pref_netw   <- rowSums(dat04[,c(matches(match = "pref_netw"  ,vars = names(dat04)))],na.rm = TRUE)
+dat04[dat04==0] <- NA
+```
+
+## Version 01: Fixed order by perception / preferences (N=`r dim(dat01)[1]`)
 
@@
 
-```{r,results='hold'}
-model01 <- 'perc_merit=~meritv01_perc_effort+meritv01_perc_talent 
+```{r echo=FALSE,results='asis'}
+stargazer(dat01,type = "html")
+```
+
+```{r correlacion v01}
+cor01<-polychoric(dat01)
+corrplot(cor01$rho,method = "color" ,type = "upper", order = "hclust", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
+```
+
+```{r}
+model01 <- '
+perc_merit=~meritv01_perc_effort+meritv01_perc_talent 
 perc_nmerit=~meritv01_perc_wpart+meritv01_perc_netw  
@@
 
-fit1 <- cfa(model = model01,data = dat01,ordered = c("meritv01_perc_effort","meritv01_perc_talent",
+fit1_c <- cfa(model = model01,data = dat01,estimator="MLR") # Continuo/ estimador ML Robust
+
+fit1_o <- cfa(model = model01,data = dat01,
+            ordered = c("meritv01_perc_effort","meritv01_perc_talent",
                         "meritv01_perc_wpart","meritv01_perc_netw",
@@
 
-summary(fit1,standardized=TRUE, fit.measures=TRUE)
-semPlot::semPaths(object = fit1,what = "std",thresholds = FALSE, intercepts = FALSE)
+#---------------------------------------------------#
+# #summary(fit1_c,standardized=TRUE, fit.measures=TRUE)}
+fitmeasures(fit1_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+semPlot::semPaths(object = fit1_c,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE, curve = 1.5)
+
+# #summary(fit1_o,standardized=TRUE, fit.measures=TRUE)
+fitmeasures(fit1_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+semPlot::semPaths(object = fit1_o,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE, curve = 1.5)
 ```
 
-*summary*: good fit, heywood case item 8.
+> *#summary*: good fit, heywood case item 8.
 
-## Version 02: Fixed order, by topic (N=  )
+## Version 02: Fixed order, by topic (N=`r dim(dat02)[1]`)
 
@@
 
+```{r echo=FALSE,results='asis'}
+stargazer(dat02,type = "html")
+```
+
+```{r}
+cor02<-polychoric(dat02)
+corrplot(cor02$rho,method = "color", type = "upper", order = "hclust", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
+```
+
 ```{r,results='hold'}
-model02 <- 'perc_merit=~meritv02_perc_effort+meritv02_perc_talent 
+model02 <- '
+perc_merit=~meritv02_perc_effort+meritv02_perc_talent 
 perc_nmerit=~meritv02_perc_wpart+meritv02_perc_netw  
@@
 
-fit2 <- cfa(model = model02,data = dat02,ordered = c("meritv02_perc_effort","meritv02_perc_talent",
+fit2_c <- cfa(model = model02,data = dat02,estimator="MLR") # Continuo/ estimador ML Robust
+
+fit2_o <- cfa(model = model02,data = dat02,ordered = c("meritv02_perc_effort","meritv02_perc_talent",
                                                      "meritv02_perc_wpart","meritv02_perc_netw",
@@
 
-summary(fit2,standardized=TRUE, fit.measures=TRUE)
-semPlot::semPaths(object = fit2,what = "std")
+#---------------------------------------------------#
+#summary(fit2_c,standardized=TRUE, fit.measures=TRUE)
+fitmeasures(fit2_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+semPlot::semPaths(object = fit2_c,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE, curve = 1.5)
+
+#summary(fit2_o,standardized=TRUE, fit.measures=TRUE)
+fitmeasures(fit2_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+semPlot::semPaths(object = fit2_o,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE, curve = 1.5)
 ```
-*summary*: bad fit, heywood
+> **NOTA**: bad fit, heywood en estimación DWLS
 
-## Version 03: Random order
+## Version 03: Random order (N=`r dim(dat03)[1]`)
+
+```{r echo=FALSE,results='asis'}
+stargazer(dat03,type = "html")
+```
+
+```{r}
+cor03<-polychoric(dat03)
+corrplot(cor03$rho,method = "color", type = "upper", order = "hclust", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
+```
 
 ```{r,results='hold'}
-model03 <- 'perc_merit=~meritv03_perc_effort+meritv03_perc_talent 
+model03 <- '
+perc_merit=~meritv03_perc_effort+meritv03_perc_talent 
 perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw  
@@
 
-fit3 <- cfa(model = model03,data = dat03,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
+fit3_c <- cfa(model = model03,data = dat03,estimator="MLR",std.lv=FALSE) # Continuo/ estimador ML Robust
+
+fit3_o <- cfa(model = model03,data = dat03,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
+                                                       "meritv03_perc_wpart","meritv03_perc_netw",
+                                                       "meritv03_pref_effort","meritv03_pref_talent",
+                                                       "meritv03_pref_wpart","meritv03_pref_netw"),std.lv=FALSE)
+#---------------------------------------------------#
+#summary(fit3_c,standardized=TRUE, fit.measures=TRUE)
+fitmeasures(fit3_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+semPlot::semPaths(object = fit3_c,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE, curve = 1.5)
+
+#summary(fit3_o,standardized=TRUE, fit.measures=TRUE)
+fitmeasures(fit3_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+semPlot::semPaths(object = fit3_o,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE, curve = 1.5)
+```
+
+> **NOTA**: fit not very goood (RMSEA .7), no heywood. Try tau-equivalence
+
+```{r,results='hold'}
+model03.tau <- '
+perc_merit=~ r1*meritv03_perc_effort+r1*meritv03_perc_talent 
+perc_nmerit=~r2*meritv03_perc_wpart+ r2*meritv03_perc_netw  
+pref_merit=~ r3*meritv03_pref_effort+r3*meritv03_pref_talent 
+pref_nmerit=~r4*meritv03_pref_wpart+ r4*meritv03_pref_netw'
+
+fit3_c.tau <- cfa(model = model03.tau,data = dat03,estimator="MLR",std.lv=FALSE) # Continuo/ estimador ML Robust
+
+fit3_o.tau <- cfa(model = model03.tau,data = dat03,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
+                                                               "meritv03_perc_wpart","meritv03_perc_netw",
+                                                               "meritv03_pref_effort","meritv03_pref_talent",
+                                                               "meritv03_pref_wpart","meritv03_pref_netw"), std.lv=FALSE)
+#---------------------------------------------------#
+#summary(fit3_c.tau,standardized=TRUE, fit.measures=TRUE)
+fitmeasures(fit3_c.tau)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+semPlot::semPaths(object = fit3_c.tau,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE, curve = 1.5)
+
+#summary(fit3_o,standardized=TRUE, fit.measures=TRUE)
+fitmeasures(fit3_o.tau)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+semPlot::semPaths(object = fit3_o.tau,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE, curve = 1.5)
+```
+
+```{r}
+rbind(fitmeasures(fit3_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+      fitmeasures(fit3_o.tau)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")])
+anova(fit3_o,fit3_o.tau)
+```
+
+> **NOTA:** El modelo con igualdad de cargas factoriales no deteriora significativamente el ajuste del modelo.
+
+## Version 04: muestra completa (N=`r dim(dat04)[1]`)
+
+```{r echo=FALSE,results='asis'}
+stargazer(select(dat04,perc_effort,perc_talent,perc_wpart,perc_netw,pref_effort,pref_talent,pref_wpart,pref_netw),
+          type = "html")
+```
+
+```{r}
+cor04<-polychoric(select(dat04,perc_effort,perc_talent,perc_wpart,perc_netw,pref_effort,pref_talent,pref_wpart,pref_netw))
+corrplot(cor04$rho, method="color",type = "upper", order = "hclust", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
+```
+
+```{r,results='hold'}
+model04 <- '
+perc_merit =~ perc_effort+perc_talent 
+perc_nmerit=~perc_wpart +perc_netw  
+pref_merit =~ pref_effort+pref_talent 
+pref_nmerit=~pref_wpart +pref_netw'
+
+fit4_c <- cfa(model = model04,data = dat04,estimator="MLR",std.lv=FALSE) # Continuo/ estimador ML Robust
+
+fit4_o <- cfa(model = model04,data = dat04,ordered = c("perc_effort","perc_talent",
+                                                       "perc_wpart","perc_netw",
+                                                       "pref_effort","pref_talent",
+                                                       "pref_wpart","pref_netw"),std.lv=FALSE)
+#---------------------------------------------------#
+#summary(fit4_c,standardized=TRUE, fit.measures=TRUE)
+fitmeasures(fit4_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+semPlot::semPaths(object = fit3_c,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE,curve = 2)
+
+#summary(fit4_o,standardized=TRUE, fit.measures=TRUE)
+fitmeasures(fit4_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+semPlot::semPaths(object = fit3_o,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE,curve = 2)
+```
+
+### Resumen ajuste ola 01
+
+```{r}
+sum_fit<- bind_rows(fitmeasures(fit1_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+                    fitmeasures(fit1_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+                    fitmeasures(fit2_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+                    fitmeasures(fit2_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+                    fitmeasures(fit3_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+                    fitmeasures(fit3_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+                    fitmeasures(fit4_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+                    fitmeasures(fit4_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")])
+sum_fit$mod <- c("Modelo 1","Modelo 1","Modelo 2","Modelo 2","Modelo 3","Modelo 3","Modelo 4","Modelo 4")
+sum_fit$est <- c("Continuo","Ordinal","Continuo","Ordinal","Continuo","Ordinal","Continuo","Ordinal")
+sum_fit <- select(sum_fit,mod,est,everything())
+
+kable(sum_fit,digits = 3,format = "html",row.names = F,booktabs=T, caption = "Resumen estadísticos de ajuste ola 1") %>% kable_styling(full_width = F) %>% 
+  collapse_rows(columns = 1,valign = "middle")  %>% 
+  footnote(number = c("Modelo 1: fixed order by percepction/preference", 
+                      "Modelo 2: fixed order by topic (i.e: effort)", 
+                      "Modelo 3: Randomized order",
+                      "Modelo 4: Complete sample"))
+```
+
+## Análisis por Grupos múltiples
+
+```{r}
+model04 <- '
+perc_merit =~ perc_effort+perc_talent 
+perc_nmerit=~perc_wpart +perc_netw  
+pref_merit =~ pref_effort+pref_talent 
+pref_nmerit=~pref_wpart +pref_netw'
+
+fit4_c_mg <- cfa(model = model04,data = dat04,estimator="MLR",std.lv=FALSE,group = "grupo") # Continuo/ estimador ML Robust
+
+fit4_o_mg <- cfa(model = model04,data = dat04,ordered = c("perc_effort","perc_talent",
+                                                       "perc_wpart","perc_netw",
+                                                       "pref_effort","pref_talent",
+                                                       "pref_wpart","pref_netw"),std.lv=FALSE, group = "grupo")
+#---------------------------------------------------#
+#summary(fit4_c_mg,standardized=TRUE, fit.measures=TRUE)
+fitmeasures(fit4_c_mg)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+semPlot::semPaths(object = fit4_c_mg,what = "std",thresholds = FALSE, intercepts = FALSE, fade=FALSE, curve=2)
+
+#summary(fit4_o_mg,standardized=TRUE, fit.measures=TRUE)
+fitmeasures(fit4_o_mg)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+semPlot::semPaths(object = fit4_o_mg,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE, curve=2)
+```
+
+Cosas por probar:
+
+* V01+V02: cargas y fit
+* V01+V03: cargas y fit
+* V02+V03: cargas y fit
+
+# Modelo Factorial de Segundo Orden:
+
+## a. Factor de segundo orden por Percepción y Preferencia
+
+```{r}
+model04b1or <- '
+perc_merit=~ perc_effort+perc_talent 
+perc_nmerit=~perc_wpart +perc_netw  
+pref_merit=~ pref_effort+pref_talent 
+pref_nmerit=~pref_wpart +pref_netw
+percep=~ perc_merit + perc_nmerit
+prefer=~ pref_merit + pref_nmerit'
+
+fit4a_c <- cfa(model = model04b1or,data = dat04, estimator="MLR")
+
+fit4b_c <- cfa(model = model04b1or,data = dat04,ordered = c("perc_effort","perc_talent","perc_wpart","perc_netw",
+                                                            "pref_effort","pref_talent","pref_wpart","pref_netw"))
+
+#---------------------------------------------------#
+#summary(fit4a_c,standardized=TRUE, fit.measures=TRUE)
+fitmeasures(fit4a_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+semPlot::semPaths(object = fit4a_c,what = "std",thresholds = FALSE, intercepts = FALSE, fade=FALSE,curve=2)
+
+#summary(fit4b_c,standardized=TRUE, fit.measures=TRUE)
+fitmeasures(fit4b_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+semPlot::semPaths(object = fit4b_c,what = "std",thresholds = FALSE, intercepts = FALSE, fade=FALSE,curve=2)
+```
+
+> **NOTA:** Cargas factoriales de segundo orden son bajas. El ajuste del modelo no es bueno.
+
+## b. Factor de segundo orden por Meritocracia y No Meritocracia
+
+```{r}
+model04b2or <-'perc_merit=~ perc_effort+perc_talent
+               perc_nmerit=~perc_wpart +perc_netw
+               pref_merit=~ pref_effort+pref_talent
+               pref_nmerit=~pref_wpart +pref_netw
+               merit =~ perc_merit  + pref_merit
+               nmerit=~ perc_nmerit + pref_nmerit'
+
+# fit4b_c <- cfa(model = model04b2or,data = dat04,estimator="MLR") # No converge
+
+fit4b_o <- cfa(model = model04b2or,data = dat04, 
+               ordered = c("perc_effort","perc_talent","perc_wpart","perc_netw",
+                           "pref_effort","pref_talent","pref_wpart","pref_netw"))
+
+#---------------------------------------------------#
+# #summary(fit4b_c,standardized=TRUE, fit.measures=TRUE)
+# fitmeasures(fit4b_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+# semPlot::semPaths(object = fit4b_c,what = "std",thresholds = FALSE, intercepts = FALSE, fade=FALSE, curve=2)
+
+#summary(fit4b_o,standardized=TRUE, fit.measures=TRUE)
+fitmeasures(fit4b_o)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+semPlot::semPaths(object = fit4b_o,what = "std",thresholds = FALSE, intercepts = FALSE, fade=FALSE, curve=2)
+```
+
+> **NOTA:** El modelo estimado con MLR con factores de segundo orden "meritocracia/ no meritocracia" no converge.
+EL modelo estimado a través de DWLS converge, posee cargas muy bajas y una muy alta (merit~pref_merit).
+Fit es malo (RMSEA >.10)
+
+# Wave 02
+
+```{r}
+data02 <- sjlabelled::read_spss(path = "input/data/original/Estudio_3_ola2.sav",verbose = FALSE)
+load(file = "input/data/proc/datawide.RData")
+dat01 <- data02 %>% select(starts_with("meritv03_p")) 
+```
+
+## ATE indicadores meritocracia
+
+```{r, out.width="75%", results='asis'}
+m01<- lm(perc_effort_w2~treat,data = wide01)
+m02<- lm(perc_talent_w2~treat,data = wide01)
+m03<- lm(perc_wpart_w2~treat, data = wide01)
+m04<- lm(perc_netw_w2~treat,  data = wide01)
+#--------------------------------------------#
+m05<- lm(pref_effort_w2~treat,data = wide01)
+m06<- lm(pref_talent_w2~treat,data = wide01)
+m07<- lm(pref_wpart_w2 ~treat, data = wide01)
+m08<- lm(pref_netw_w2  ~treat,  data = wide01)
+#--------------------------------------------#
+stargazer::stargazer(m01,m02,m03,m04,type = "html")
+stargazer::stargazer(m05,m06,m07,m08,type = "html")
+```
+
+> **NOTA**: Los tratamientos de información no se asocian de manera estadísticamente significativa con ninguno de los indicadores de la escala de meritocracia.
+
+## CFA Muestra completa (Control, Desigualdad y Pobreza)
+
+```{r echo=FALSE,results='asis'}
+stargazer(dat01,type = "html")
+```
+
+```{r}
+dat01cor<- select(dat01,meritv03_perc_effort,meritv03_perc_talent,
+                        meritv03_perc_wpart,meritv03_perc_netw,
+                        meritv03_pref_effort,meritv03_pref_talent,
+                        meritv03_pref_wpart,meritv03_pref_netw)
+cor01<-polychoric(dat01cor)
+corrplot(cor01$rho, type = "upper", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
+```
+
+```{r,results='hold'}
+model03 <- '
+perc_merit=~meritv03_perc_effort+meritv03_perc_talent 
+perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw  
+pref_merit=~meritv03_pref_effort+meritv03_pref_talent 
+pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw'
+
+fitw02_c <- cfa(model = model03,data = dat01,estimator="MLR",std.lv=FALSE) # Continuo/ estimador ML Robust
+
+fitw02_o <- cfa(model = model03,data = dat01,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
+                                                          "meritv03_perc_wpart","meritv03_perc_netw",
+                                                          "meritv03_pref_effort","meritv03_pref_talent",
+                                                          "meritv03_pref_wpart","meritv03_pref_netw"))
+#---------------------------------------------------#
+#summary(fitw02_c,standardized=TRUE, fit.measures=TRUE)
+fitmeasures(fitw02_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+semPlot::semPaths(object = fitw02_c,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE,curve=2)
+
+#summary(fitw02_o,standardized=TRUE, fit.measures=TRUE)
+fitmeasures(fitw02_o)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+semPlot::semPaths(object = fitw02_o,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE,curve=2)
+```
+
+> **NOTA**: El ajuste de muestra completa en wave 02 con estimador MLR (robusto) es bastante bueno (CFI, RMSEA). El ajuste con DWLS es bastante bueno, CFI mejora, RMSEA aumenta un poco (0.59).
+
+* **Recomendación**: es probable que el estimador MLR sea mejor que el ordinal.
+
+### Factor de segundo orden
+
+```{r}
+modw2 <- '
+perc_merit=~meritv03_perc_effort+meritv03_perc_talent 
+perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw  
+pref_merit=~meritv03_pref_effort+meritv03_pref_talent 
+pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw 
+merit =~perc_merit+pref_merit  
+nmerit=~perc_nmerit+pref_nmerit'
+
+# mod2_merit_c <- cfa(model = modw2,data = dat01) # Continuo/ estimador ML Robust
+
+mod2_merit_o <- cfa(model = modw2,data = dat01,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
                                                            "meritv03_perc_wpart","meritv03_perc_netw",
@@
 
-summary(fit3,standardized=TRUE, fit.measures=TRUE)
-semPlot::semPaths(object = fit3,what = "std",thresholds = FALSE, intercepts = FALSE)
+#---------------------------------------------------#
+# #summary(mod2_merit_c,standardized=TRUE, fit.measures=TRUE)
+# fitmeasures(mod2_merit_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+# semPlot::semPaths(object = mod2_merit_c,what = "std",thresholds = FALSE, intercepts = FALSE)
+
+#summary(mod2_merit_o,standardized=TRUE, fit.measures=TRUE)
+fitmeasures(mod2_merit_o)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+semPlot::semPaths(object = mod2_merit_o,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE, curve=2, rotation=2)
 ```
 
-*summary*: fit not very goood (RMSEA .7), no heywood. Try tau-equivalence
+> * Factor de segundo orden usando MLR no converge.
+> * Factor de segundo orden usando DWLS converge, pero las cargas no son buenas y el ajuste tampoco.
 
-## Version 04: muestra completa (N=  )
+## CFA Muestra Control
+
+```{r}
+w02con <- data02 %>% filter(FL_6_DO_Control==1) %>% select(starts_with("meritv03_p")) 
+```
+
+```{r echo=FALSE,results='asis'}
+stargazer(w02con,type = "html")
+```
+
+```{r}
+corw02_1<-polychoric(w02con)
+corrplot(corw02_1$rho, type = "upper", order = "hclust", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
+```
 
 ```{r,results='hold'}
-dat04 <- data01 %>% filter(Intro==1) %>% select(starts_with("meritv01"),starts_with("meritv02"),starts_with("meritv03_p"))
+model03 <- '
+perc_merit=~meritv03_perc_effort+meritv03_perc_talent 
+perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw  
+pref_merit=~meritv03_pref_effort+meritv03_pref_talent 
+pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw'
 
-dat04$perc_effort <- rowSums(dat04[,c(matches(match = "perc_effort",vars = names(dat04)))],na.rm = TRUE)
-dat04$perc_talent <- rowSums(dat04[,c(matches(match = "perc_talent",vars = names(dat04)))],na.rm = TRUE)
-dat04$perc_wpart  <- rowSums(dat04[,c(matches(match = "perc_wpart" ,vars = names(dat04)))],na.rm = TRUE)
-dat04$perc_netw   <- rowSums(dat04[,c(matches(match = "perc_netw"  ,vars = names(dat04)))],na.rm = TRUE)
+fitw02a_c <- cfa(model = model03,data = w02con,estimator="MLR") # Continuo/ estimador ML Robust
 
-dat04$pref_effort <- rowSums(dat04[,c(matches(match = "pref_effort",vars = names(dat04)))],na.rm = TRUE)
-dat04$pref_talent <- rowSums(dat04[,c(matches(match = "pref_talent",vars = names(dat04)))],na.rm = TRUE)
-dat04$pref_wpart  <- rowSums(dat04[,c(matches(match = "pref_wpart" ,vars = names(dat04)))],na.rm = TRUE)
-dat04$pref_netw   <- rowSums(dat04[,c(matches(match = "pref_netw"  ,vars = names(dat04)))],na.rm = TRUE)
+fitw02a_o <- cfa(model = model03,data = w02con,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
+                                                           "meritv03_perc_wpart","meritv03_perc_netw",
+                                                           "meritv03_pref_effort","meritv03_pref_talent",
+                                                           "meritv03_pref_wpart","meritv03_pref_netw"))
 
-model04 <- 'perc_merit=~perc_effort+perc_talent 
-            perc_nmerit=~perc_wpart+perc_netw  
-            pref_merit=~pref_effort+pref_talent 
-            pref_nmerit=~pref_wpart+pref_netw'
+#---------------------------------------------------#
+#summary(fitw02a_c,standardized=TRUE, fit.measures=TRUE)
+fitmeasures(fitw02a_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+semPlot::semPaths(object = fitw02a_c,what = "std",thresholds = FALSE, intercepts = FALSE, curve=2)
 
-fit4 <- cfa(model = model04,data = dat04)
-summary(fit4,standardized=TRUE, fit.measures=TRUE)
-semPlot::semPaths(object = fit4,what = "std",thresholds = FALSE, intercepts = FALSE)
+#summary(fitw02a_o,standardized=TRUE, fit.measures=TRUE)
+fitmeasures(fitw02a_o)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+semPlot::semPaths(object = fitw02a_o,what = "std",thresholds = FALSE, intercepts = FALSE, curve=2)
 ```
+
+## CFA Muestra Pobreza
+
+```{r}
+w02pob <- data02 %>% filter(FL_6_DO_Tratamiento==1) %>% select(starts_with("meritv03_p")) 
+```
+
+```{r echo=FALSE,results='asis'}
+stargazer(w02pob,type = "html")
+```
+
+```{r}
+corw02_2<-polychoric(w02pob)
+corrplot(corw02_2$rho, type = "upper", order = "hclust", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
+```
+
+```{r,results='hold'}
+model03 <- '
+perc_merit=~meritv03_perc_effort+meritv03_perc_talent 
+perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw  
+pref_merit=~meritv03_pref_effort+meritv03_pref_talent 
+pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw'
+
+fitw02b_c <- cfa(model = model03,data = w02pob,estimator="MLR") # Continuo/ estimador ML Robust
+
+fitw02b_o <- cfa(model = model03,data = w02pob,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
+                                                     "meritv03_perc_wpart","meritv03_perc_netw",
+                                                     "meritv03_pref_effort","meritv03_pref_talent",
+                                                     "meritv03_pref_wpart","meritv03_pref_netw"))
+
+#---------------------------------------------------#
+#summary(fitw02b_c,standardized=TRUE, fit.measures=TRUE)
+fitmeasures(fitw02b_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+semPlot::semPaths(object = fitw02b_c,what = "std",thresholds = FALSE, intercepts = FALSE, curve=2)
+
+#summary(fitw02b_o,standardized=TRUE, fit.measures=TRUE)
+fitmeasures(fitw02b_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+semPlot::semPaths(object = fitw02b_o,what = "std",thresholds = FALSE, intercepts = FALSE, curve=2)
+```
+
+## CFA Muestra Desigualdad
+
+```{r}
+w02des <- data02 %>% filter(FL_6_DO_Tratamiento_desigualdad==1) %>% select(starts_with("meritv03_p")) 
+```
+
+```{r echo=FALSE,results='asis'}
+stargazer(w02des,type = "html")
+```
+
+```{r}
+corw02_3<-polychoric(w02des)
+corrplot(corw02_3$rho, type = "upper", order = "hclust", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
+```
+
+```{r,results='hold'}
+model03 <- '
+perc_merit=~meritv03_perc_effort+meritv03_perc_talent
+perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw  
+pref_merit=~meritv03_pref_effort+meritv03_pref_talent 
+pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw'
+
+fitw02c_c <- cfa(model = model03,data = w02des,estimator="MLR") # Continuo/ estimador ML Robust
+
+fitw02c_o <- cfa(model = model03,data = w02des,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
+                                                           "meritv03_perc_wpart","meritv03_perc_netw",
+                                                           "meritv03_pref_effort","meritv03_pref_talent",
+                                                           "meritv03_pref_wpart","meritv03_pref_netw"))
+#---------------------------------------------------#
+#summary(fitw02c_c,standardized=TRUE, fit.measures=TRUE)
+fitmeasures(fitw02c_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+semPlot::semPaths(object = fitw02c_c,what = "std",thresholds = FALSE, intercepts = FALSE)
+
+#summary(fitw02b_o,standardized=TRUE, fit.measures=TRUE)
+fitmeasures(fitw02c_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+semPlot::semPaths(object = fitw02c_o,what = "std",thresholds = FALSE, intercepts = FALSE)
+```
+
+# Resumen ajuste ola 2
+```{r}
+sum_fitw02<- bind_rows(fitmeasures(fitw02_c)[ c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+                       fitmeasures(fitw02_o)[ c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+                       fitmeasures(fitw02a_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+                       fitmeasures(fitw02a_o)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+                       fitmeasures(fitw02b_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+                       fitmeasures(fitw02b_o)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+                       fitmeasures(fitw02c_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+                       fitmeasures(fitw02c_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")])
+
+sum_fitw02$mod <- c("Modelo 1","Modelo 1","Modelo 2","Modelo 2","Modelo 3","Modelo 3","Modelo 4","Modelo 4")
+sum_fitw02$est <- c("Continuo","Ordinal","Continuo","Ordinal","Continuo","Ordinal","Continuo","Ordinal")
+sum_fitw02 <- select(sum_fitw02,mod,est,everything())
+
+kable(sum_fitw02,digits = 3,format = "html",row.names = F,booktabs=T, caption = "Resumen ajuste ola 2") %>% 
+  kable_styling(full_width = F) %>% 
+  collapse_rows(columns = 1,valign = "middle") %>% 
+  footnote(number = c("Modelo 1: Muestra completa", 
+                      "Modelo 2: Muestra grupo Control", 
+                      "Modelo 3: Muestra grupo tratamiento pobreza",
+                      "Modelo 4: Muestra grupo tratamiento desigualdad"))
+```
+
+# Wave 03
+
+```{r}
+data03 <- sjlabelled::read_spss(path = "input/data/original/Estudio_3_ola3.sav",verbose = FALSE)
+dat01 <- data03 %>% select(starts_with("meritv03_p")) 
+```
+
+```{r}
+sjt.fa(data = dat01,show.comm = TRUE)
+```
+
+```{r}
+model03 <- '
+perc_merit=~meritv03_perc_effort+meritv03_perc_talent 
+perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw  
+pref_merit=~meritv03_pref_effort+meritv03_pref_talent 
+pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw'
+
+fitw03c_c <- cfa(model = model03,data = dat01,estimator="MLR") # Continuo/ estimador ML Robust
+
+fitw03c_o <- cfa(model = model03,data = dat01,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
+                                                          "meritv03_perc_wpart","meritv03_perc_netw",
+                                                          "meritv03_pref_effort","meritv03_pref_talent",
+                                                          "meritv03_pref_wpart","meritv03_pref_netw"))
+#---------------------------------------------------#
+#summary(fitw03c_c,standardized=TRUE, fit.measures=TRUE)
+fitmeasures(fitw03c_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+
+#summary(fitw03c_o,standardized=TRUE, fit.measures=TRUE)
+fitmeasures(fitw03c_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+```
+**summary**
+> * Ajuste usando estimador MLR  es bueno (CFI = 0.971, RMSEA = 0.052). El chi2 (robusto) es 62.421  
+> * Ajuste usando estimador DWLS es bueno (CFI = 0.974, RMSEA = 0.067). El chi2 (robusto) es 106.111
+
+# Resumen ajuste ola 3
+
+```{r}
+sum_fitw03<- bind_rows(fitmeasures(fitw03c_c)[ c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+                       fitmeasures(fitw03c_o)[ c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")])
+
+sum_fitw03$mod <- c("Modelo 1","Modelo 1")
+sum_fitw03$est <- c("Continuo","Ordinal")
+sum_fitw03 <- select(sum_fitw03,mod,est,everything())
+
+kable(sum_fitw03,digits = 3,format = "html",row.names = F,booktabs=T,caption = "Resumen ajuste ola 3") %>% 
+  kable_styling(full_width = F) %>% 
+  collapse_rows(columns = 1,valign = "middle") %>% 
+  footnote(number = c("Modelo 1: Muestra completa"))
+```
+
+# Análisis longitudinal
+
+```{r}
+library(sjPlot)
+library(dplyr)
+library(lavaan)
+library(stargazer)
+library(corrplot)
+library(psych)
+library(knitr)
+library(kableExtra)
+load(file = "input/data/proc/datawide_123.RData")
+```
+
+## Correlación indicadores w01 y w02
+
+```{r}
+g01<- wide02 %>% filter(grupo==1) %>% select(starts_with("perc_"),starts_with("pref_")) # Fixed by perception/preference
+g02<- wide02 %>% filter(grupo==2) %>% select(starts_with("perc_"),starts_with("pref_")) # Fixed by topic (i.e: effort)
+g03<- wide02 %>% filter(grupo==3) %>% select(starts_with("perc_"),starts_with("pref_")) # Randomized
+```
+
+### Grupo Versión 01
+
+```{r, out.width="75%", results='asis'}
+stargazer(g01,type = "html", median = T,digits = 2)
+```
+
+```{r, out.width=c("75%","75%"), fig.cap=c("Percepción","Preferencia"), fig.align='center'}
+corwide01a<- polychoric(x = select(g01 ,starts_with(match = "perc_")),na.rm = T)
+corrplot(corwide01a$rho, method = "color" ,type = "upper", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
+
+corwide01b<- polychoric(x = select(g01 ,starts_with(match = "pref_")),na.rm = T)
+corrplot(corwide01b$rho, method = "color" ,type = "upper", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
+```
+
+### Grupo Versión 02
+
+```{r, out.width="75%", results='asis'}
+stargazer(g02,type = "html", median = T,digits = 2)
+```
+
+```{r, out.width=c("75%","75%"), fig.cap=c("Percepción","Preferencia"), fig.align='center'}
+corwide02a<- polychoric(x = select(g02 ,starts_with(match = "perc_")),na.rm = T)
+corrplot(corwide02a$rho, method = "color" ,type = "upper", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
+
+corwide02b<- polychoric(x = select(g02 ,starts_with(match = "pref_")),na.rm = T)
+corrplot(corwide02b$rho, method = "color" ,type = "upper", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
+```
+
+### Grupo Versión 03
+
+```{r, results='asis'}
+stargazer(g03,type = "html", median = T,digits = 2)
+```
+
+```{r, out.width=c("75%","75%"), fig.cap=c("Percepción","Preferencia"), fig.align='center'}
+corwide03a<- polychoric(x = select(g03 ,starts_with(match = "perc_")),na.rm = T)
+corrplot(corwide03a$rho, type = "upper", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE,tl.cex = 0.75)
+
+corwide03b<- polychoric(x = select(g03 ,starts_with(match = "pref_")),na.rm = T)
+corrplot(corwide03b$rho, method = "color" ,type = "upper", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
+```
+
+### Muestra completa
+
+```{r, results='asis'}
+stargazer(select(wide02 ,starts_with(match = "perc_"), starts_with("pref_")),type = "html", median = T,digits = 2)
+```
+
+```{r, out.width=c("75%","75%"), fig.cap=c("Percepción","Preferencia"), fig.align='center'}
+corwideA<- polychoric(x = select(wide02 ,starts_with(match = "perc_")),na.rm = T)
+corrplot(corwideA$rho, method = "color" ,type = "upper", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
+
+corwideB<- polychoric(x = select(wide02 ,starts_with(match = "pref_")),na.rm = T)
+corrplot(corwideB$rho, method = "color" ,type = "upper", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
+```
+
+## CFA wave01, wave02 y wave03
+
+```{r}
+model_long <- '
+# wave 01
+perc_merit1 =~perc_effort_w1+perc_talent_w1 
+perc_nmerit1=~perc_wpart_w1 +perc_netw_w1  
+pref_merit1 =~pref_effort_w1+pref_talent_w1 
+pref_nmerit1=~pref_wpart_w1 +pref_netw_w1
+# wave 02
+perc_merit2 =~perc_effort_w2+perc_talent_w2 
+perc_nmerit2=~perc_wpart_w2 +perc_netw_w2  
+pref_merit2 =~pref_effort_w2+pref_talent_w2 
+pref_nmerit2=~pref_wpart_w2 +pref_netw_w2
+# wave 03
+perc_merit3 =~perc_effort_w3+perc_talent_w3 
+perc_nmerit3=~perc_wpart_w3 +perc_netw_w3  
+pref_merit3 =~pref_effort_w3+pref_talent_w3 
+pref_nmerit3=~pref_wpart_w3 +pref_netw_w3'
+
+fitlong_c <- cfa(model = model_long,data = wide02,estimator="MLR") # Continuo/ estimador ML Robust
+
+indicadores <-c("perc_effort_w1","perc_talent_w1","perc_wpart_w1","perc_netw_w1",
+                "pref_effort_w1","pref_talent_w1","pref_wpart_w1","pref_netw_w1",
+                "perc_effort_w2","perc_talent_w2","perc_wpart_w2","perc_netw_w2",
+                "pref_effort_w2","pref_talent_w2","pref_wpart_w2","pref_netw_w2")
+
+fitlong_o <- cfa(model = model_long,data = wide02,ordered = indicadores)
+
+#---------------------------------------------------#
+#summary(fitlong_c,standardized=TRUE, fit.measures=TRUE)
+fitmeasures(fitlong_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+semPlot::semPaths(object = fitlong_c,what = "std",thresholds = FALSE, intercepts = FALSE)
+
+#summary(fitlong_o,standardized=TRUE, fit.measures=TRUE)
+fitmeasures(fitlong_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+semPlot::semPaths(object = fitlong_o,what = "std",thresholds = FALSE, intercepts = FALSE,style = "mx")
+```
+
+```{r,fig.cap="Correlación entre variables latentes w01, w02 y w03"}
+cormat<- as.matrix(inspect(fitlong_o, "cor.lv"))
+corrplot(cormat, method = "color" ,type = "upper", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
+```
+
+# Invarianza
+
+1. Invarianza entre modalidades en ola 1,2 y 3 (between groups invariance)
+2. Invarianza entre tratamientos en ola 2(between groups invariance)
+
+>**NOTA**: En vista de que no existe una librería que permita estimar invarianza longitudinal asumiendo indicadores ordinales, las primeras estimaciones las realizaré asumiendo indicadores continuos (estimator=ML).
+
+```{r}
+library(semTools)
+```
+
+```{r}
+load(file = "input/data/proc/datalong.RData")     # Datos formato long
+load(file = "input/data/proc/datawide_123.RData") # Datos formato long
+```
+
+```{r}
+dim(dat04)
+m_inv_modalidad <- 'perc_merit=~ perc_effort+perc_talent 
+                    perc_nmerit=~perc_wpart +perc_netw  
+                    pref_merit=~ pref_effort+pref_talent 
+                    pref_nmerit=~pref_wpart +pref_netw'
+
+indicators  <-c("perc_effort","perc_talent","perc_wpart","perc_netw",
+                "pref_effort","pref_talent","pref_wpart","pref_netw")
+
+fit.conf=cfa(m_inv_modalidad,data=dat04,group="grupo")
+fit.deb =cfa(m_inv_modalidad,data=dat04,group="grupo",group.equal=c("loadings"))
+fit.fuer=cfa(m_inv_modalidad,data=dat04,group="grupo",group.equal=c("loadings","intercepts"))
+fit.str =cfa(m_inv_modalidad,data=dat04,group="grupo",group.equal=c("loadings","intercepts","residuals"))
+
+fitmeasures(fit.conf)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+fitmeasures(fit.deb )[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+fitmeasures(fit.fuer)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+fitmeasures(fit.str )[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+anova(fit.conf, fit.deb, fit.fuer, fit.str)
+```
+
+```{r}
+configural <- '
+# wave 01
+perc_merit1 =~NA*perc_effort_w1+lambda1*perc_effort_w1+perc_talent_w1 
+perc_nmerit1=~NA*perc_wpart_w1 +lambda1*perc_wpart_w1 +perc_netw_w1  
+pref_merit1 =~NA*pref_effort_w1+lambda1*pref_effort_w1+pref_talent_w1 
+pref_nmerit1=~NA*pref_wpart_w1 +lambda1*pref_wpart_w1 +pref_netw_w1
+# wave 02
+perc_merit2 =~NA*perc_effort_w2+lambda1*perc_effort_w2+perc_talent_w2 
+perc_nmerit2=~NA*perc_wpart_w2 +lambda1*perc_wpart_w2 +perc_netw_w2  
+pref_merit2 =~NA*pref_effort_w2+lambda1*pref_effort_w2+pref_talent_w2 
+pref_nmerit2=~NA*pref_wpart_w2 +lambda1*pref_wpart_w2 +pref_netw_w2
+# wave 03
+perc_merit3 =~NA*perc_effort_w3+lambda1*perc_effort_w3+perc_talent_w3 
+perc_nmerit3=~NA*perc_wpart_w3 +lambda1*perc_wpart_w3 +perc_netw_w3  
+pref_merit3 =~NA*pref_effort_w3+lambda1*pref_effort_w3+pref_talent_w3 
+pref_nmerit3=~NA*pref_wpart_w3 +lambda1*pref_wpart_w3 +pref_netw_w3
+
+# INTERCEPTS: Por cada variable latente, debo fijar el intercepto del primer indicador.
+# wave 01--------------------------
+perc_effort_w1 ~ i1*1   # Percepcion merito        
+perc_talent_w1 ~ 1      #      
+perc_wpart_w1 ~ i1*1    # Percepcion no-merito   
+perc_netw_w1 ~ 1        #
+pref_effort_w1 ~ i1*1   # Preferencia merito    
+pref_talent_w1 ~ 1      #  
+pref_wpart_w1 ~ i1*1    # Preferencia no-merito   
+pref_netw_w1 ~ 1        #
+
+# wave 02--------------------------
+perc_effort_w2 ~ i1*1   # Percepcion merito                    
+perc_talent_w2 ~ 1      #               
+perc_wpart_w2 ~ i1*1    # Percepcion no-merito              
+perc_netw_w2 ~ 1        #       
+pref_effort_w2 ~ i1*1   # Preferencia merito                
+pref_talent_w2 ~ 1      #           
+pref_wpart_w2 ~ i1*1    # Preferencia no-merito              
+pref_netw_w2 ~ 1        #
+
+# wave 03--------------------------
+perc_effort_w3 ~ i1*1   # Percepcion merito                     
+perc_talent_w3 ~ 1      #                            
+perc_wpart_w3 ~ i1*1    # Percepcion no-merito                
+perc_netw_w3 ~ 1        #                  
+pref_effort_w3 ~ i1*1   # Preferencia merito                   
+pref_talent_w3 ~ 1      #                        
+pref_wpart_w3 ~ i1*1    # Preferencia no-merito               
+pref_netw_w3 ~ 1        #
+
+# UNIQUE VARIANCES and COVARIANCES.
+
+# LATENT VARIABLES MEANS: Para cada variable latente fijar la media
+# Fix T1 common factor mean to zero
+perc_merit1 ~ 0*1
+# Freely estimate T2-T3 common factor means
+perc_merit2 ~ 1
+perc_merit3 ~ 1
+
+perc_nmerit1 ~ 0*1
+perc_nmerit2 ~ 1
+perc_nmerit3 ~ 1
+
+pref_merit1 ~ 0*1
+pref_merit2 ~ 1
+pref_merit3 ~ 1
+
+pref_nmerit1 ~ 0*1
+pref_nmerit2 ~ 1
+pref_nmerit3 ~ 1
+
+# LATENT VARIABLE VARIANCES AND COVARIANCES: correlaciones entre variables latentes
+
+'
+
+fit_conf <- cfa(model = configural,data =wide02)
+summary(fit_conf, standardized=TRUE)
+```
+
+* Invarianza Longitudinal en lavaan
+https://quantdev.ssri.psu.edu/tutorials/intro-basics-longitudinal-measurement-invariance
+
+* https://users.ugent.be/~yrosseel/lavaan/multiplegroup6Dec2012.pdf
+
+* Invarianza con indicadores categóricos
+https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5121102/pdf/nihms757737.pdf
+
+* Parametrización tetha: 
+https://groups.google.com/forum/#!topic/lavaan/nfdatPgLLhc
+
+* Colapsing categories:
+https://www.tandfonline.com/doi/abs/10.1080/10705511.2018.1547640?journalCode=hsem20

### 2020-03-19 production/prep-data-A.Rmd

---
title: "Longitudinal data"
date: "`r format(Sys.time(), '%d %B, %Y')`"
css: ../style.css
output: 
  html_document:
    number_sections: true
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: "hide"
editor_options: 
  chunk_output_type: console
---

# Longitudinal data

```{r eval=FALSE, include=FALSE}
rmarkdown::render(input = "production/prep-data-A.Rmd",output_format = "html_document",output_dir = ".")
```

```{r include=FALSE}
knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = TRUE,results = "hold")
knitr::opts_knit$set(root.dir = "../")
options(scipen=999)
rm(list=ls())
options(knitr.kable.NA = '')
```

```{r}
library(sjPlot)
library(dplyr)
library(tidyr)
library(lavaan)
library(stargazer)
library(corrplot)
library(psych)
library(knitr)
library(kableExtra)
```

```{r}
w01 <- sjlabelled::read_spss("input/data/original/Estudio_3_ola1.sav",verbose = F)
w02 <- sjlabelled::read_spss("input/data/original/Estudio_3_ola2.sav",verbose = F)
w03 <- sjlabelled::read_spss("input/data/original/Estudio_3_ola3.sav",verbose = F)
```

```{r}
table(w01$Intro,exclude=NULL) # (2) 221 no aceptan, (1) 2236 aceptan

w01a <- filter(w01,Intro==1) # Elimina los que no aceptan
table(w01a$Intro,exclude=NULL) # 2236

# Variables para ver si finalizo encuesta
table(w01a$Finished,exclude=NULL) # (1) 2100 finalizan, (0) 136 no terminan

w01b <- filter(w01a,Finished==1) # Elimina los que no aceptan
table(w01b$Intro,exclude=NULL) # 2100
```

```{r}
table(w02$Intro,exclude=NULL) # (2)  no aceptan, (1)  aceptan

w02a <- filter(w02,Intro==1) # Elimina los que no aceptan
table(w02a$Intro,exclude=NULL) #

# Variables para ver si finalizo encuesta
table(w02a$Finished,exclude=NULL) # (1)  finalizan, (0)  no terminan

w02b <- filter(w02a,Finished==1) # Elimina los que no aceptan
table(w02b$Intro,exclude=NULL) # 
```

```{r}
table(w03$Finished==1)
w03b <- filter(w03,Finished==1) # Elimina los que no aceptan
dim(w03b)
```

```{r}
w01b$ID<- stringr::str_split_fixed(w01b$ticket,"_", 4)[,1] # En la variable ticket el primer código es el ID, todo lo demás se borra
w02b$ID <- stringr::str_split_fixed(w02b$ticket, "_", 4)[,1]

w03b$ID <- stringr::str_split_fixed(w03b$ticket, "_", 4)[,1] 
```

## prepare wave01

```{r}
w01c <- w01b 
w01c$grupo <- NA
w01c$grupo[w01c$FL_21_DO_merit_perc_pref_julio19v01==1] <- 1
w01c$grupo[w01c$FL_21_DO_merit_perc_pref_julio19v02==1] <- 2
w01c$grupo[w01c$FL_21_DO_merit_perc_pref_julio19v03==1] <- 3
w01c$perc_effort_w1 <- rowSums(w01c[,c(matches(match = "perc_effort",vars = names(w01c)))],na.rm = TRUE)
w01c$perc_talent_w1 <- rowSums(w01c[,c(matches(match = "perc_talent",vars = names(w01c)))],na.rm = TRUE)
w01c$perc_wpart_w1  <- rowSums(w01c[,c(matches(match = "perc_wpart" ,vars = names(w01c)))],na.rm = TRUE)
w01c$perc_netw_w1   <- rowSums(w01c[,c(matches(match = "perc_netw"  ,vars = names(w01c)))],na.rm = TRUE)

w01c$pref_effort_w1 <- rowSums(w01c[,c(matches(match = "pref_effort",vars = names(w01c)))],na.rm = TRUE)
w01c$pref_talent_w1 <- rowSums(w01c[,c(matches(match = "pref_talent",vars = names(w01c)))],na.rm = TRUE)
w01c$pref_wpart_w1  <- rowSums(w01c[,c(matches(match = "pref_wpart" ,vars = names(w01c)))],na.rm = TRUE)
w01c$pref_netw_w1   <- rowSums(w01c[,c(matches(match = "pref_netw"  ,vars = names(w01c)))],na.rm = TRUE)
w01c[w01c==0] <- NA
w01c <- w01c %>% select(ID,grupo,
                        perc_effort_w1,perc_talent_w1,perc_wpart_w1,perc_netw_w1,
                        pref_effort_w1,pref_talent_w1,pref_wpart_w1,pref_netw_w1,
                        sexo_w1=sexo,edad_w1=edad,edcep_w1=edcep)
#-----------------------------------------------------------------------#
skimr::skim(w01c)
```

## prepare wave02

```{r}
w02c <- w02b

w02c$treat <- NA
w02c$treat[w02c$FL_6_DO_Control==1] <- 1 
w02c$treat[w02c$FL_6_DO_Tratamiento==1] <- 2 
w02c$treat[w02c$FL_6_DO_Tratamiento_desigualdad==1] <- 3

w02c$treat <- factor(w02c$treat,levels = c(1,2,3),labels = c("control","pobreza","desiguald"))

w02c <- w02c %>% rename(perc_effort_w2=meritv03_perc_effort,
                        perc_talent_w2=meritv03_perc_talent,
                        perc_wpart_w2 =meritv03_perc_wpart,
                        perc_netw_w2  =meritv03_perc_netw,
                        pref_effort_w2=meritv03_pref_effort,
                        pref_talent_w2=meritv03_pref_talent,
                        pref_wpart_w2 =meritv03_pref_wpart,
                        pref_netw_w2  =meritv03_pref_netw)

w02c <- w02c %>% select(ID,
                        perc_effort_w2,perc_talent_w2,perc_wpart_w2,perc_netw_w2,
                        pref_effort_w2,pref_talent_w2,pref_wpart_w2,pref_netw_w2,
                        sexo_w2=sexo,edad_w2=edad,edcep_w2=edcep, treat)

#-----------------------------------------------------------------------#
skimr::skim(w02c)
```

## prepare wave03

```{r}
w03c <- w03b

w03c <- w03c %>% rename(perc_effort_w3=meritv03_perc_effort,
                        perc_talent_w3=meritv03_perc_talent,
                        perc_wpart_w3 =meritv03_perc_wpart,
                        perc_netw_w3  =meritv03_perc_netw,
                        pref_effort_w3=meritv03_pref_effort,
                        pref_talent_w3=meritv03_pref_talent,
                        pref_wpart_w3 =meritv03_pref_wpart,
                        pref_netw_w3  =meritv03_pref_netw)
w03c <- w03c %>% select(ID,
                        perc_effort_w3,perc_talent_w3,perc_wpart_w3,perc_netw_w3,
                        pref_effort_w3,pref_talent_w3,pref_wpart_w3,pref_netw_w3,
                        sexo_w3=sexo,edad_w3=edad,edcep_w3=edcep)

#-----------------------------------------------------------------------#
skimr::skim(w03c)
```

## Merge wide data

```{r}
wide01 <- inner_join(x = w01c,y = w02c, by = "ID") # Solamente nos quedamos con casos en ambas olas 
dim(wide01) #en w01 y w02
```

```{r}
wide02 <- inner_join(x = wide01,y = w03c, by = "ID") # Solamente nos quedamos con casos en ambas olas 
dim(wide02) #en w01, w02 y w03
```

```{r}
skimr::skim(wide02)
```

```{r}
save(wide01,file = "input/data/proc/datawide.RData")
save(wide02,file = "input/data/proc/datawide_123.RData")
```

# Wide to long data

```{r}
library(datasets)
library(data.table)

wide <- setDT(wide02)

long01<- melt(wide,variable.name = "wave",
              measure = patterns("^perc_effort_","^perc_talent_","^perc_wpart_","^perc_netw_",
                                       "^pref_effort_","^pref_talent_","^pref_wpart_","^pref_netw_",
                                       "^sexo_","^edad_","^edcep_"),
              value.name = c("perc_effort","perc_talent","perc_wpart","perc_netw",
                             "pref_effort","pref_talent","pref_wpart","pref_netw",
                             "sexo","edad","edcep"))

```

```{r}
save(long01,file = "input/data/proc/datalong.RData")
```

### 2020-03-19 production/analysis-cfa.Rmd
update02

@@
 date: "`r format(Sys.time(), '%d %B, %Y')`"
-css: ../style.css
+css: "../input/bib/style.css"
 output: 
@@
 ```{r include=FALSE}
-knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = TRUE)
+knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = FALSE)
 knitr::opts_knit$set(root.dir = "../")

### 2020-03-19 production/prep-data-A.Rmd
update02

@@
 ```{r include=FALSE}
-knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = TRUE,results = "hold")
+knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = FALSE,results = "hold")
 knitr::opts_knit$set(root.dir = "../")

### 2020-03-19 render-site.R

@@ -1,0 +1,20 @@
+
+# Render site ------------------------------------------------------------
+
+site<- list.files(path = "site")
+file.copy(file.path("site",site), "docs", overwrite = TRUE) # copy main text of the site
+
+production<- list.files(path = "production",pattern = ".Rmd") # copy data proc and analysis code
+file.copy(file.path("production",production), "docs",overwrite = TRUE)
+
+rmarkdown::render_site("docs") # Render site
+
+unlink(paste0("docs/",dir(path = "docs",pattern = "_files")), recursive = TRUE) # add folder name to be deleted
+unlink(paste0("docs/",dir(path = "docs",pattern = "_cache")), recursive = TRUE) # add folder name to be deleted
+
+ext <- c(".rmd",".Rmd",".yml",".tex",".log") #add file extension to be deleted
+for (i in 1:length(ext)) {
+  file.remove(paste0("docs/",dir(path="docs", pattern=ext[i])))
+}
+
+

### 2020-03-19 site/preanalisis.Rmd

--- 2020-01-10 index.Rmd
@@
 title: "Perceptions and Preferences for Meritocracy Scale (PPM-S)"
-csl: "input/bib/apa6.csl"
-bibliography: "input/bib/zlibrary.bib"
-css: tyle.css
+csl: "../input/bib/apa6.csl"
+bibliography: "../input/bib/zlibrary.bib"
+css: "../input/bib/style.css"
 output:
@@
   html_document:
-    theme: flatly
     number_sections: true
@@
 
-tab01 <- read.csv(file = "input/data/original/materials/table01.csv",header = 1,sep = ";")
+tab01 <- read.csv(file = "../input/data/original/materials/table01.csv",header = 1,sep = ";")
 

### 2020-03-20 docs/paper.rmd
update20marzo

---
title: "Perceptions and Preferences for Meritocracy Scale (PPM-S)"
linestretch: '1.5'
link-citations: yes
output:
  html_document:
    number_sections: true
    toc: true
    toc_float: true
linkcolor: blue
bibliography: "../input/bib/EconomiaMoral.bib"
css: "../input/bib/style.css"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = TRUE)
```

## Introducción

Tanto las organizaciones internacionales, como parte del mundo académico, presentan una creciente preocupación por el aumento de la desigualdad social y los efectos negativos que esta posee en torno a la legitimidad de nuestras sociedades [ONU,2016; @schlozmanUnheavenlyChorusUnequal2012; @CastilloMeritocraciadesigualdadeconomica2019]. Más relevante aún se vuelve la legitimidad de la desigualdad en un contexto mundial de crecientes protestas sociales en contra de la misma. En el caso Chileno, estas movilizaciones han alcanzado una gran envergadura, relevando la importancia de evaluar los niveles de legitimidad de la desigualdad existente, a la luz de distintas ideologías que podrían justificarla. Dentro de estas ideologías que legitiman la desigualdad, la literatura sobre la meritocracia, presenta un notorio crecimiento [@MadeiraPrimesConsequencesSystematic2019]. Tradicionalmente la meritocracia es entendida como una forma de distribución social de recursos donde el mérito individual (i.e. el esfuerzo y el talento) determinan la distribucion de los privilegios [@breenClassInequalityMeritocracy1999;@youngRiseMeritocracy1994; @OhsfeldtBeliefsInequalityAmericans1987; @Jostrolestereotypingsystemjustification1994], sin la interferencia de factores estructurales como la herencia o los contactos [@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995; @yairMeritocracy2007; @landWeSatTable2006;@youngRiseMeritocracy1994].

Los estudios relativos a la meritocracia se pueden clasificar entre estudios de caracter objetivo y estudios decaracter subjetivo. Los estudios de caracter objetivo, suelen dedicarse a la observar de la movilidad social, evaluando críticamente hasta qué punto las sociedades son efectivamente meritocráticas y su distribución no está influida por contactos o ventajas externas [@alonDiversityOpportunityShifting2007; @dubetRepensarJusticiaSocial2017; @eriksonIntergenerationalClassMobility2010; @rosanvallonSocieteEgaux2011; @barozetValorHistoricoPituto2006; @mcnameeMeritocracyMyth2004; @goldthorpeMythEducationbasedMeritocracy2003] Los segundos de caracter subjetivo, ponen el enfacis no en la observacion de la distribucion material de los recursos sino en las ideas y discursos en torno a la distribucion social, mirando la meritocracia desde el enfoque de la economia moral. Desde esta perspectiva de la economía moral, estudian las creencias meritocráticas y su relación con distintos factores, como el ingreso, el contexto social o los prejuicios a grupos vulnerables [@KunovichSystemsDistributionSense2007;@NewmanFalseConsciousnessClass2015; @ReynoldsPerceptionsmeritocracyland2014; @Duru-BellatWhoMeritocracyIndividual2012]. Pese a que el concepto de creencias meritocráticas es fundamental para esta segunda línea de trabajo, las distintas escalas utilizadas en esta literatura no poseen suficientes evidencias de validez a su favor, lo que releva la importancia de aportar con propuestas para medir este constructo.

Para estudiar las creencias meritocráticas se han utilizado múltiples escalas que suponen una estructura unifactorial de las creencias meritocráticas, sin distinguir entre elementos individuales y estructurales, ni diferenciar las preferencias de las percepciones [@KunovichSystemsDistributionSense2007;@NewmanFalseConsciousnessClass2015; @ReynoldsPerceptionsmeritocracyland2014]. Más aún, se ha optado por índices unidimensionales incluso cuando se ha tenido evidencia en contra de ellos [@KunovichSystemsDistributionSense2007;@ReynoldsPerceptionsmeritocracyland2014]. La mala especificación de las dimensiones del constructo puede llevar a conclusiones erradas sobre las creencias meritocráticas y su rol en la legitimidad de las desigualdades, por lo que es necesario hacer evaluaciones tanto sobre la dimensionalidad del constructo como de los ítems que permiten aproximarse adecuadamente al mismo.

El objetivo de este proyecto fue desarrollar una escala de creencias en la meritocracia, que presentará una adecuada estructura de medición, que además buscamos que fuese estable en el tiempo. Por ello, nuestras preguntas de investigación refieren a la estructura del constructo y a la mantención de su validez a nivel temporal.

## Antecedentes generales, la economía moral y la meritocracia
Las ciencias sociales, desde la perspectiva de la economía moral, han enriquecido la comprensión de las relaciones económicas, aportando a la lógica racional imperante en las teorias economicas, aspectos subjetivos como las normas, los procesos históricos y culturales que estructuran las prácticas tanto las económicas [@mauMoralEconomyWelfare2004; @sachwehMoralEconomyInequality2012; @polanyiGranTransformacionCritica1997; @zelizerEconomicLivesHow2010 ;@fourcadeMoralViewsMarket2007] como la produccion y reproduccion de la desigualdad [@lamontWhatMissingCultural2014].En este marco teorico, se ha  planteado que las creencias meritocráticas pueden legitimar la desigualdad existente [@hadjarMeritokratieAlsLegitimationsprinzip2008]. En la misma linea, se ha planteado que la creencia en la meritocracia, que adjudica causas individuales a la pobreza, se relaciona con la atribucion de prejuicios hacia grupos de bajo estatus [@MadeiraPrimesConsequencesSystematic2019] y con ideas conservadoras de autoritarismo social [@AzevedoNeoliberalIdeologyJustification2019].

## Evidencia de la dimensionalidad de las creencias meritocráticas. 
La mayoría de las investigaciones que han trabajado con las creencias meritocráticas, la han abordado como un concepto unidimensional. No obstante, algunos autores han destacado la diferencia existente entre las preferencias y las percepciones meritocráticas [@Duru-BellatWhoMeritocracyIndividual2012; @CastilloMeritocraciadesigualdadeconomica2019]. Por percepciones meritocráticas refiere al modo en que el entrevistado describe la realidad. Las preferencias refieren a un aspecto normativo, es decir, que tan de acuerdo se está con que la meritocracia es un buen sistema de distribucion. Esta distinción es fundamental, puesto que ambas variables poseen bastante relación entre sí, por lo cual es de suponer que trabajar con estas variables sin controlarlas entre sí nos puede llevar a tener ideas causales erróneas por espuriedad. Así, por ejemplo, puede ser que las conclusiones que indican una mayor percepción de meritocracia en base a los ingresos, posean coeficientes sobreestimados y errores no normales, por la falta de variables relevantes.

Además, de la distinción entre percepciones y preferencias, es necesario hacer la distinción entre elementos meritocráticos y elementos no meritocráticos. Al respecto algunos autores, que trabajan solamente con las percepciones meritocráticas, suponen que creer en el peso de factores externos (e.g. la ventaja de tener padres ricos) es contrario a creer en el rol del esfuerzo. De este modo, las creencias no meritocraticas son cosnideradas es el extremo negativo del continuo, mientras creer que la trayectoria personal depende del esfuerzo y el talento del individuo, corresponde al lado positivo del continuo. Dicho planteamiento deriva teoría que señala que una sociedad meritocrática ideal se define por la distribución de riquezas en base a esfuerzo y no a fuerzas estructurales [@yairMeritocracy2007; @landWeSatTable2006]. Aunque dicho planteamiento es correcto para evaluar que tan meritocratica es una sociedad, no necesariamente este pensamiento se refleja en el modo en como se escructuran los discursos en los imaginarios de la poblacion. De hecho, otros investigadores han evaluado la dimensionalidad de las creencias meritocráticas, generando evidencia sobre la multidimensionalidad de este constructo, dando cuenta de que aspectos meritocraticos y no meritocraticos no forman parte de la misma dimension dentro de la subjetividad de los encuestados. Entre de las investigaciones que han evaluado dicha dimensionalidad se encuentran los trabajos de @KunovichSystemsDistributionSense2007 y @ReynoldsPerceptionsmeritocracyland2014.

La investigación multinivel internacional de @KunovichSystemsDistributionSense2007, que busca explicar las preferencias meritocráticas en el pago utilizando la encuesta ISSP, realizó análisis factoriales para probar la unidimensionalidad sin encontrar resultados satisfactorios, llegando a la conclusión de que las creencias en la meritocracia son más bien un conjunto diverso de ideas agregadas respecto al esfuerzo y la recompensa, razón por la cual decide utilizar un índice sumatorio pese a tener correlaciones insatisfactorias, las cuales estaban más cargadas a los ítems relacionados con aspectos no meritocráticos. En vista de la valioza evidencia producida por esta investigacion, podemos decir que, pese a que se ha trabajado como tal, las creencias meritocraticas parecen no ajustar a un modelo unidimensional.

Por su parte, el artículo de @ReynoldsPerceptionsmeritocracyland2014, utilizando la batería del “get ahead” de la ISSP, plantea que las creencias meritocráticas poseen 3 dimensiones, las cuales refieren a la “percepción de meritocracia” (i.e. importancia atribuida al éxito y al talento para seguir adelante) a la “percepción de la importancia de las redes para salir adelante (i.e tener padres ricos, tener contactos) y la percepción de la discriminación como obstáculo para salir adelante (i.e la importancia del sexo y de la raza de la persona). Pese a esta evidencia, el autor decide hacer un índice unidimensional, donde se incorpora la primera dimensión como el lado positivo del continuo y un agregado de las segundas como el lado negativo del continuo, posteriormente el autor divide este continuo en 4 grupos.

En suma, considerando la evidencia de estos autores, se deduce la multidimensionalidad del constructo, destacándose la diferencia entre los factores de percepciones meritocráticas y las percepciones no meritocráticas.

En consideración de lo discutido, se hipotetiza que las creencias en la meritocracia poseen una estructura multidimensional de 4 factores, los cuales son el resultado del cruce entre las diferencias de percepciones o preferencias y aspectos meritocráticos o no meritocráticos, como se indica en la tabla 1. Así, probaremos un modelo en que las dimensiones sean percepciones meritocráticas, percepciones no meritocráticas, preferencias meritocráticas y preferencias no meritocráticas. Este modelo de medida posee dos ventajas, en primer lugar, no parte del supuesto de la dicotomía entre las creencias meritocráticas y no meritocráticas permitiendo que los sujetos puntúan alto o bajo en las dos variables. Igualmente, es positivo que exista la distinción entre preferencias y percepciones puesto que esto nos permitirá predicciones más robustas y más precisas respecto a las creencias de la meritocracia.

|                  	  | Perceptions 	| Preferences 	|
|:-----------------	  |-------------	|-------------	|
| _Meritocracy_     	|             	|             	|
| _Non-meritocracy_ 	|             	|             	|

# Sección de indicadores 
Si bien no existe un consenso respecto a cuáles son los indicadores más adecuados para medir creencias en la meritocracia, los indicadores más comunes son preguntas del tipo, “qué tan importante es … para salir adelante”, las cuales incluyen características como tener padres ricos, trabajar duro, ser talentoso, el sexo, la raza, entre otras. El problema con este indicador presente en la International Social Survey Program, es que habla de un criterio de importancia y no de suficiencia. Así, por ejemplo, alguien puede creer que el esfuerzo es fundamental para salir adelante, y creer al mismo tiempo que no todas las personas que se esfuerzan más logran obtener más recursos que quienes se esfuerzan menos. Al respecto, hemos encontrado evidencia sistemática de que si bien una gran mayoría de las personas de chile consideran que el esfuerzo es fundamental para salir adelante, no creen que las personas no son recompensadas por su esfuerzo en su país.

Otro indicador de esa misma encuesta, pregunta respecto a si cree que es importante a considerar en el pago de una persona distintas características, incluyendo características relativas al mérito, como educación o responsabilidad, y algunas relativas a las necesidades, como el tener hijos. No obstante, este modo de evaluar las creencias respecto a la meritocracia, si bien posee un aspecto no meritocrático, este no refiere a las desigualdades heredables que son lo propiamente contrario al ideal meritocrático (Yair, 2007; Land, 2006; newman; Reynolds and Xian, 2014).  Además, este indicador no incluye aspectos de percepción, lo cual, como dijimos, puede generar problemas de espuriedad.

En vista de la inexistencia de un consenso respecto a la manera correcta de medir las creencias meritocráticas proponemos nuevos indicadores que buscan ser más enfáticos en el fraseo respecto a la diferencia de preferencias y percepciones.  Así, las preguntas que utilizamos son las presentes en la tabla 2. Estos indicadores, para referirse a las percepciones sitúan al entrevistado en la realidad de su país.

![](input/images/tabla2.png)

# Muestra del Estudio

# Resultados: Estructura factorial, efecto framing e invarianza longitudinal
Este estudio cuenta con una encuesta panel de 3 olas. En la primera de los olas, se incorporo la escala en distintos ordenes. El primer orden, pregunta primero por todas las percepciones y despues por todas las preferencias, el segundo orden, pregunta seguidamente por percepciones y preferencias en torno a un mismo tema y, el tercer orden incorpora los items de manera aleatoria. En la segunda y tercera ola, se utilizo el orden

Para poder evaluar la estructura factorial de la escala propuesta, se realizará un análisis factorial confirmatorio, evaluando el ajuste del modelo propuesto a partir de los estadísticos CFA, TLI, RMSEA y χ2 ratio. Los criterios de ajuste utilizados fueron anteriormente reportados en un preregistro de la investigación en el sitio Open Science FrameWork. Basado en los criterios de Kline (2016)  y de Brown (2008)  esperamos tanto para la prueba CFA como TLI, valores por sobre .95, mientras que para la prueba RMSEA, se consideran aceptables valores bajo .08 e idóneos bajo .05. Por su parte, respecto al χ2, se esperan valores p no significativos como parámetro ideal y se considera aceptable un χ2 ratio menor a 3.

En una primera instancia se probó el ajuste del modelo con tres distintos órdenes de los ítems, los cuales fueron aplicados a un tercio de la muestra de la primera ola cada uno. El primer orden de los ítems corresponde al orden de de la tabla 2, en el que primero se pregunta a los encuestados por todas las percepciones y después por todas las preferencias, en el segundo orden se les pregunta seguidamente por percepciones y preferencias respecto a un mismo tema, por ejemplo, al papel del esfuerzo, y en el tercero, los ítems aparecen de manera aleatoria.

Table 3: **Fit indicators**

|           | Model order 1 | Model order 2 | Model order 3 |
|:---------:|:-------------:|:-------------:|:-------------:|
| **n**     |      697      |      702      |      697      |
| **CFI**   |     0.997     |     0.984     |     0.991     |
| **TLI**   |     0.992     |     0.968     |     0.984     |
| **RMSEA** |     0.037     |     0.071     |     0.051     |
| **χ2**    |     28.03     |     64.156    |     39.090    |
| **p**     |      .000     |      .000     |      .000     |
| **χ2/gl** | 2,00          | 4,58          | 2,78          |

El primero de los órdenes fue el que mejor ajuste obtuvo (CFI=0.998, TLI= 0.995,RMSEA=0.037, χ2(df=14)=28,03, p = 0.014), seguido por el orden aleatorio de los ítems (CFI=0.992, TLI=0.984,RMSEA=0.051, χ2(df=14)=39.09, p < 0.001).  Por su parte la escala ordenada por temáticas genera un efecto framing en el cual la relación entre las percepciones y las preferencias parecen sobreestimadas, lo cual afecta el ajuste, siendo este deficitario, con un χ2 ratio mayor a tres y un RMSEA que se puede encontrar por sobre .8 según el intervalo de confianza correspondiente(CFI=0.984, TLI=0.968,RMSEA=.071, χ2(df=14)=64.156, p < 0.001).

|           | Contrast Model 1 Factor | theoretical model 4 Factors | Model with M.I. |
|:---------:|:-----------------------:|:---------------------------:|:---------------:|
| **n**     |           1769          |             1769            |       1769      |
| **CFI**   |          0.595          |            0.988            |      0.994      |
| **TLI**   |          0.433          |            0.976            |      0.985      |
| **RMSEA** |          0.226          |            0.047            |      0.036      |
| **χ2**    |         1830.839        |            68.661           |      40.250     |
| **p**     |           .000          |             .000            |       .000      |
| **χ2/gl** |          65.38          |             4.90            |       3.35      |

En la segunda ola se utilizó el primer orden de los ítems ya que este fue el que presentó un mejor resultado. EL modelo teorico propuesto de cuatro factores ajustó, como se observa en la tabla 4, mejor que el modelo de contraste de 1 factor. El modelo teórico ajustó de manera relativamente adecuada, pues muestra indicadores óptimos para CFI= 0.987, TLI = 0.975 y RMSEA=.041, aunque posee indicadores deficientes para la prueba χ2(df=14)=67.6, p-value=.000. Para evaluar posibles mejoras de la escala, se analizaron las relaciones propuestas por los índices de modificación. Estos indican la existencia de dos cargas cruzadas no especificadas. Cuando se generó un modelo siguiendo estas recomendaciones, hubo una mejora considerable del modelo, aunque el nuevo modelo tampoco obtuvo un X2 ratio menor a 3 y obtuvo cargas factoriales muy bajas ʎ < 0.15, por lo tanto, siguiendo las recomendación de Brown (2006) de solo aceptar las propuesta de los índices de modificación cuando se posee teoria y evidencia sólida, se ha decidido no incorporar estos parámetros al modelo.

* agregar cuales fueron las cargas sujeridas.

# Validez del constructo

## Validez convergente  
indicar la construcción de los índices -- hacer un anexo.

Esta escala cuenta con validez de constructo, ya que los distintos índices formados según factor, correlacionan coherentemente la escala "Get a head", utilizada por el programa internacional de encuestas sociales (ISSP), y con la escala de valores igualitarios utilizada por la encuesta internacional de valores. **(indicar validez del constructo y bibliografía que  lo use).**

|                                  | Igualitarismo | Get a head |
|----------------------------------|:-------------:|:----------:|
| Percepción <br>Meritocrática     |   -0.143***   | 0.228***   |
| Percepción <br>No Meritocrática  |    0.130***   | -0.210***  |
| Preferencia <br>Meritocrática    |   -0.055**    | 0.001      |
| Preferencia <br>No Meritocratica |   -0.191***   | 0.083***   |

**dar vuelta la tabla, get primero**

Existe coherencia entre el índice "Get a Head" y el índice de preferencias y percepciones meritocráticas. Como era esperable, la relación es positiva entre este índice y la percepción de meritocracia de la escala propuesta (*r*=0.228). En la misma línea, el índice Get a head, correlaciona de manera negativa con la percepción de no meritocracia (c=-0.210;p=0.000). Además, las relaciones con las preferencias son considerablemente menores, e incluso no es significativa respecto a las preferencias meritocráticas. En suma, el "Get a head" que es comúnmente utilizado para medir percepciones, posee relación solo con las variables de percepción y no así con las de preferencias.

Por su parte, el índice de Igualitarismo que refleja cuán problemática consideran la desigualdad los entrevistados, posee también relaciones coherentes. En primer lugar, es necesario destacar que solo posee una relación muy pequeña con la variable de preferencias meritocráticas, lo cual es positivo pues dicha dimensión de la escala no busca reflejar igualitarismo sino de preferencias meritocráticas. En segundo lugar, el índice se relaciona de manera negativa con la preferencia no meritocrática, lo cual es coherente en tanto alguien que valore la igualdad debería tender a no estar de acuerdo con la desigualdad de oportunidades. En tercer lugar, considerar la desigualdad como un problema, desde el índice de igualitarismo, posee como requisito creer que existe desigualdad, al respecto, es coherente que quienes consideran que la desigualdad existente es problemática, perciban igualmente que el éxito no depende sólo del esfuerzo y el talento (c=-0.143;p=0.000) sino que está influido por aspectos no meritocráticos (c=0.130;p=).

**contradicción a considerar: la validez por criterio, en la cual se evalúa la relación con una variable validada, depende de la calidad de la medida. nosotros lo relacionamos con lo que dice que esta mal.get-

# Invarianza longitudinal

# Discusión

 - el texto incorpora una discusión que parte de las cargas cruzadas, dándoles sentido (que es un hallazgo propio del texto) y dando cuenta de cómo esta estructura es equivalente en varias encuestas, posteriormente justifica la existencia de cierto desajuste en base a una hipótesis plausible (diferencia entre autoaplicado y por teléfono)

-  habla de la validez predictiva, habla de bajas correlaciones y plantea una posibilidad de un nuevo estudio con un cambio

- plantea una limitación, su estudio no es experimental y podría estar influido por más por el tipo de encuesta que se contestó.

# Conclusión

En un contexto de crecientes movilizaciones sociales y creciente deslegitimidad de los gobiernos, es necesario contar con instrumentos válidos para la medición de las creencias meritocráticas, las cuales, suelen funcionar como legitimadoras de la desigualdad.  Por ello, es necesario contar con instrumentos breves y validados, para incorporar la medición de esta variable en otros estudios relacionados con temas de legitimidad. Utilizando los datos recolectados en una encuesta panel representativa de Chile, este artículo expone la elaboración y la validación de una escala de creencias meritocráticas de 4 factores: percepciones meritocráticas, percepciones no-meritocráticas, preferencias meritocráticas y preferencias no meritocráticas. La escala de percepciones y preferencias meritocráticas es un instrumento breve, el cual permite medir de manera relativamente sencilla las creencias meritocráticas, las cuales son fundamentales para abordar temas de legitimación de la desigualdad, preferencias redistributivas o prejuicios hacia grupos vulnerables.

# Bibliografía

### 2020-03-20 production/prod_analysis-cfa.rmd
update20marzo

--- 2020-03-19 production/analysis-cfa.Rmd
@@
 ```{r include=FALSE}
-knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = FALSE)
+knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = TRUE)
 knitr::opts_knit$set(root.dir = "../")

### 2020-03-20 render-site.R

@@ -1,20 +1,25 @@
 
 # Render site ------------------------------------------------------------
 
-site<- list.files(path = "site")
-file.copy(file.path("site",site), "docs", overwrite = TRUE) # copy main text of the site
+# INSTRUCTIONS-------------------------:
+# 1. set your root dir at "merit-scale"
+# 2. Run this code.
 
-production<- list.files(path = "production",pattern = ".Rmd") # copy data proc and analysis code
-file.copy(file.path("production",production), "docs",overwrite = TRUE)
+production<- list.files(path = "production",pattern = "prod_")        # a list with the names of the files to copy
+file.copy(file.path("production",production), "docs",overwrite = TRUE)# copy data proc and analysis files 
 
 rmarkdown::render_site("docs") # Render site
 
-unlink(paste0("docs/",dir(path = "docs",pattern = "_files")), recursive = TRUE) # add folder name to be deleted
-unlink(paste0("docs/",dir(path = "docs",pattern = "_cache")), recursive = TRUE) # add folder name to be deleted
+# before you run this line, check if your R Markdown files have a .rmd or .Rmd extension
+# on this case we use both
+ext <- c(grep("^prod_.*\\.rmd$",x = dir(path = "docs"),value = T), # for .rmd 
+         grep("^prod_.*\\.Rmd$",x = dir(path = "docs"),value = T), # for .Rmd
+         ".tex",".log")                                            # for .tex and .log 
 
-ext <- c(".rmd",".Rmd",".yml",".tex",".log") #add file extension to be deleted
 for (i in 1:length(ext)) {
-  file.remove(paste0("docs/",dir(path="docs", pattern=ext[i])))
+  file.remove(paste0("docs/",dir(path="docs", pattern=ext[i]))) # delete files 
 }
 
-
+# Keep only the original .rmd files (site), configuration (_site.yml and _config.yml) 
+# and prod_ folders (cache and files)
+ 

### 2020-03-21 docs/index.Rmd
update - still bugs to fix website

---
title: "PPM-S"
csl: "../input/bib/apa6.csl"
bibliography: "../input/bib/zlibrary.bib"
author: " "
output:
    html_document:
        smart: false
---

## This is the project website for the development and testing of the Perceptions and Preferences for Meritocracy Scale.

last website update: `r format(Sys.time(), "%A %d %B %Y %X")`

### Authors:

- Juan Carlos Castillo, Departamento de Sociología, Universidad de Chile. [juancastillov@uchile.cl](juancastillov@uchile.cl), [jc-castillo.com](http://jc-castillo.com/)
- Julio Iturra, Centro de Estudios de Conflicto y Cohesión Social - COES
- Francisco Meneses, Departamento de Sociología, Universidad de Chile.
- Luis Maldonado, Instituto de Sociología, Pontificia Universidad Católica de Chile

### In brief:

Based on a previous paper that proposed a distinction between perceptions and preferences for meritocracy and tested it with secondary survet data [[@Castilloetal2019]](http://revintsociologia.revistas.csic.es/index.php/revintsociologia/article/view/1017), we designed and tested a scale in a longitudinal three-wave online panel in Chile (N=1.422, final wave) through confirmatory factor analysis and measurement invariance procedures.

### Structure of the project website:

- Pre-analysis: here is the description of the pre-registration of the analysis plan in the Open Science Framework, you can see it directly [here:](https://osf.io/z45y2). To cite: Castillo, J. C., Iturra, J. C., Rivas, F. J. M., & Maldonado, L. (2019, November 21). Perceptions and Preferences for Meritocracy Scale  (PPM-S). Retrieved from osf.io/z45y2

- The other three tabs follow the logic of the [IPO protocol](https://github.com/juancarloscastillo/ipo) for academic workflows, based on three main components = **I**nput -> **P**roduction -> **O**utput. Here we show only one the main parts of each component: Data -> Analysis -> Paper. The complete reproducible files are in the [Github repo: https://juancarloscastillo.github.io/merit-scale/](https://juancarloscastillo.github.io/merit-scale/)

  - _Data_: pre-processing and general descriptives
  - _Analysis_: CFA (Confirmatory Factor Analysis) and measurement invariance
  - _Paper_: final output

## References

### 2020-03-21 docs/paper.rmd
update - still bugs to fix website

@@
 bibliography: "../input/bib/EconomiaMoral.bib"
-css: "../input/bib/style.css"
+# css: "../input/bib/style.css"
 

### 2020-03-21 docs/preanalisis.Rmd
update - still bugs to fix website

--- 2020-01-10 index.Rmd
@@
 ---
-title: "Perceptions and Preferences for Meritocracy Scale (PPM-S)"
-csl: "input/bib/apa6.csl"
-bibliography: "input/bib/zlibrary.bib"
-css: tyle.css
+title: "Pre-registration"
+csl: "../input/bib/apa6.csl"
+bibliography: "../input/bib/zlibrary.bib"
+# css: "../input/bib/style.css"
 output:
@@
     toc_float: true
+    smart: false
 geometry: margin=0.78in
@@
 ```
+# Perceptions and Preferences for Meritocracy Scale (PPM-S)
+
+(link to [PDF file](docs/preanalisis.pdf))
 
@@
 
-tab01 <- read.csv(file = "input/data/original/materials/table01.csv",header = 1,sep = ";")
-
-cnames <- c("Dimension", "Factor","Statement (english)" , "Statement (spanish)")
-
-caption01 <- capTab("Meritocracy perception and preference")
-kable(tab01,format = table_format,booktabs = T, linesep = "",col.names = cnames, caption = caption01) %>%
-  kable_styling(full_width = T,latex_options = c("hold_position","scale_down"),position = "center",font_size = 9) %>%
-  column_spec(column = 1,width = "1.5cm",) %>%
-  column_spec(column = 2,width = "2 cm") %>%
-  collapse_rows(columns = 1:2,valign = "middle")
+# tab01 <- read.csv(file = "input/data/original/materials/table01.csv",header = 1,sep = ";")
+#
+# cnames <- c("Dimension", "Factor","Statement (english)" , "Statement (spanish)")
+#
+# caption01 <- capTab("Meritocracy perception and preference")
+# kable(tab01,format = table_format,booktabs = T, linesep = "",col.names = cnames, caption = caption01) %>%
+#   kable_styling(full_width = T,latex_options = c("hold_position","scale_down"),position = "center",font_size = 9) %>%
+#   column_spec(column = 1,width = "1.5cm",) %>%
+#   column_spec(column = 2,width = "2 cm") %>%
+#   collapse_rows(columns = 1:2,valign = "middle")
 

### 2020-03-21 render-site.R

@@ -1,25 +1,33 @@
 
 # Render site ------------------------------------------------------------
 
+###### logic
+# there are some limitations for rendering websites in Rmarkdown/Githubpages
+# as all files should be in a same directory, which goes against the logic
+# of a clean and organized folder structure. Therefore, this script:
+# 1. bring all necessary files to the "docs" folder, where the site files are hosted
+# 2. renders the site from this location (docs folder)
+# 3. delete the Rmd files that where brought in step 1, leaving only the html files
+
+
 # INSTRUCTIONS-------------------------:
 # 1. set your root dir at "merit-scale"
 # 2. Run this code.
 
 production<- list.files(path = "production",pattern = "prod_")        # a list with the names of the files to copy
-file.copy(file.path("production",production), "docs",overwrite = TRUE)# copy data proc and analysis files 
+file.copy(file.path("production",production), "docs",overwrite = TRUE)# copy data proc and analysis files
 
 rmarkdown::render_site("docs") # Render site
 
 # before you run this line, check if your R Markdown files have a .rmd or .Rmd extension
 # on this case we use both
-ext <- c(grep("^prod_.*\\.rmd$",x = dir(path = "docs"),value = T), # for .rmd 
+ext <- c(grep("^prod_.*\\.rmd$",x = dir(path = "docs"),value = T), # for .rmd
          grep("^prod_.*\\.Rmd$",x = dir(path = "docs"),value = T), # for .Rmd
-         ".tex",".log")                                            # for .tex and .log 
+         ".tex",".log")                                            # for .tex and .log
 
 for (i in 1:length(ext)) {
-  file.remove(paste0("docs/",dir(path="docs", pattern=ext[i]))) # delete files 
+  file.remove(paste0("docs/",dir(path="docs", pattern=ext[i]))) # delete files
 }
 
-# Keep only the original .rmd files (site), configuration (_site.yml and _config.yml) 
+# Keep only the original .rmd files (site), configuration (_site.yml and _config.yml)
 # and prod_ folders (cache and files)
- 

### 2020-03-21 docs/preanalisis.Rmd
update21marzo

--- 2020-03-19 site/preanalisis.Rmd
@@
 knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = TRUE)
+knitr::opts_knit$set(root.dir = "../")
+options(scipen=999)
+rm(list=ls())
+options(knitr.kable.NA = '')
 ```
@@
 
-tab01 <- read.csv(file = "../input/data/original/materials/table01.csv",header = 1,sep = ";")
+tab01 <- read.csv(file = "input/data/original/materials/table01.csv",header = 1,sep = ";")
 

### 2020-04-01 production/prod_analysis-cfa.rmd
update-0103-2020
* analisis de invarianza completo

@@
 ```{r eval=FALSE, include=FALSE}
-rmarkdown::render(input = "production/analysis-cfa.Rmd",output_format = "html_document")
+rmarkdown::render(input = "production/prod_analysis-cfa.Rmd",output_format = "html_document")
 ```
@@
 #---------------------------------------------------#
-# #summary(fit1_c,standardized=TRUE, fit.measures=TRUE)}
-fitmeasures(fit1_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fit1_c,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE, curve = 1.5)
-
-# #summary(fit1_o,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fit1_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fit1_o,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE, curve = 1.5)
+cnames <- c("Factor","Indicador","Loading (MLR)","Loading (DWLS)")
+kable(left_join(x = standardizedsolution(fit1_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),y = standardizedsolution(fit1_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),c("lhs","rhs")),
+      format = "markdown",digits = 2,col.names = cnames)
 ```
@@
 fit2_o <- cfa(model = model02,data = dat02,ordered = c("meritv02_perc_effort","meritv02_perc_talent",
-                                                     "meritv02_perc_wpart","meritv02_perc_netw",
-                                                     "meritv02_pref_effort","meritv02_pref_talent",
-                                                     "meritv02_pref_wpart","meritv02_pref_netw"))
+                                                      "meritv02_perc_wpart"  ,"meritv02_perc_netw",
+                                                      "meritv02_pref_effort" ,"meritv02_pref_talent",
+                                                      "meritv02_pref_wpart"  ,"meritv02_pref_netw"))
 
 #---------------------------------------------------#
-#summary(fit2_c,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fit2_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fit2_c,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE, curve = 1.5)
-
-#summary(fit2_o,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fit2_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fit2_o,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE, curve = 1.5)
+cnames <- c("Factor","Indicador","Loading (MLR)","Loading (DWLS)")
+kable(left_join(x = standardizedsolution(fit2_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),y = standardizedsolution(fit2_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),c("lhs","rhs")),
+      format = "markdown",digits = 2,col.names = cnames)
 ```
@@
 #---------------------------------------------------#
-#summary(fit3_c,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fit3_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fit3_c,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE, curve = 1.5)
-
-#summary(fit3_o,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fit3_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fit3_o,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE, curve = 1.5)
+cnames <- c("Factor","Indicador","Loading (MLR)","Loading (DWLS)")
+kable(left_join(x = standardizedsolution(fit3_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),y = standardizedsolution(fit3_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),c("lhs","rhs")),
+      format = "markdown",digits = 2,col.names = cnames)
 ```
@@
 #---------------------------------------------------#
-#summary(fit3_c.tau,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fit3_c.tau)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fit3_c.tau,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE, curve = 1.5)
-
-#summary(fit3_o,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fit3_o.tau)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fit3_o.tau,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE, curve = 1.5)
-```
-
-```{r}
-rbind(fitmeasures(fit3_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-      fitmeasures(fit3_o.tau)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")])
-anova(fit3_o,fit3_o.tau)
+cnames <- c("Factor","Indicador","Loading (MLR)","Loading (DWLS)")
+kable(left_join(x = standardizedsolution(fit3_c.tau ) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),y = standardizedsolution(fit3_o.tau) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),c("lhs","rhs")),
+      format = "markdown",digits = 2,col.names = cnames)
+```
+
+```{r}
+kable(rbind(ordinal    =fitmeasures(fit3_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            ordinal.tau=fitmeasures(fit3_o.tau)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),format = "markdown",digits = 3)
+
+kable(anova(fit3_o,fit3_o.tau),format = "markdown",digits = 3)
 ```
@@
 ```{r echo=FALSE,results='asis'}
-stargazer(select(dat04,perc_effort,perc_talent,perc_wpart,perc_netw,pref_effort,pref_talent,pref_wpart,pref_netw),
-          type = "html")
+stargazer(select(dat04,perc_effort,perc_talent,perc_wpart,perc_netw,pref_effort,pref_talent,pref_wpart,pref_netw),type = "html")
 ```
@@
 #---------------------------------------------------#
-#summary(fit4_c,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fit4_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fit3_c,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE,curve = 2)
-
-#summary(fit4_o,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fit4_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fit3_o,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE,curve = 2)
+cnames <- c("Factor","Indicador","Loading (MLR)","Loading (DWLS)")
+kable(left_join(x = standardizedsolution(fit4_c ) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),y = standardizedsolution(fit4_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),c("lhs","rhs")),
+      format = "markdown",digits = 2,col.names = cnames)
 ```
@@
 fit4_c_mg <- cfa(model = model04,data = dat04,estimator="MLR",std.lv=FALSE,group = "grupo") # Continuo/ estimador ML Robust
-
 fit4_o_mg <- cfa(model = model04,data = dat04,ordered = c("perc_effort","perc_talent",
@@
 #---------------------------------------------------#
-#summary(fit4_c_mg,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fit4_c_mg)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fit4_c_mg,what = "std",thresholds = FALSE, intercepts = FALSE, fade=FALSE, curve=2)
-
-#summary(fit4_o_mg,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fit4_o_mg)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fit4_o_mg,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE, curve=2)
-```
-
-Cosas por probar:
-
-* V01+V02: cargas y fit
-* V01+V03: cargas y fit
-* V02+V03: cargas y fit
+kable(rbind(continuo=fitmeasures(fit4_c_mg)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            ordinal=  fitmeasures(fit4_o_mg)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
+      format = "markdown",digits = 3)
+
+cnames <- c("Grupo","Factor","Indicador","Loading (MLR)","Loading (DWLS)")
+kable(left_join(x = standardizedsolution(fit4_c_mg) %>% filter(op=="=~") %>% select(group,lhs,rhs,est.std),
+                y = standardizedsolution(fit4_o_mg) %>% filter(op=="=~") %>% select(group,lhs,rhs,est.std),
+                c("group","lhs","rhs")),
+      format = "markdown",digits = 2,col.names = cnames)
+```
 
@@
 #---------------------------------------------------#
-#summary(fit4a_c,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fit4a_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fit4a_c,what = "std",thresholds = FALSE, intercepts = FALSE, fade=FALSE,curve=2)
-
-#summary(fit4b_c,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fit4b_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fit4b_c,what = "std",thresholds = FALSE, intercepts = FALSE, fade=FALSE,curve=2)
+kable(rbind(continuo=fitmeasures(fit4a_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            ordinal= fitmeasures(fit4b_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
+      format = "markdown",digits = 3)
+
+cnames <- c("Factor","Indicador","Loading (MLR)","Loading (DWLS)")
+kable(left_join(x = standardizedsolution(fit4a_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
+                y = standardizedsolution(fit4b_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
+                c("lhs","rhs")),
+      format = "markdown",digits = 2,col.names = cnames)
 ```
@@
 #---------------------------------------------------#
-# #summary(fit4b_c,standardized=TRUE, fit.measures=TRUE)
-# fitmeasures(fit4b_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-# semPlot::semPaths(object = fit4b_c,what = "std",thresholds = FALSE, intercepts = FALSE, fade=FALSE, curve=2)
-
-#summary(fit4b_o,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fit4b_o)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fit4b_o,what = "std",thresholds = FALSE, intercepts = FALSE, fade=FALSE, curve=2)
+kable(rbind(ordinal=fitmeasures(fit4b_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
+      format = "markdown",digits = 3)
+
+cnames <- c("Factor","Indicador","Loading (DWLS)")
+kable(standardizedsolution(fit4b_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
+      format = "markdown",digits = 2,col.names = cnames)
 ```
@@
 
-```{r, out.width="75%", results='asis'}
+```{r, out.width="75%"}
 m01<- lm(perc_effort_w2~treat,data = wide01)
@@
 #--------------------------------------------#
-stargazer::stargazer(m01,m02,m03,m04,type = "html")
-stargazer::stargazer(m05,m06,m07,m08,type = "html")
+stargazer::stargazer(m01,m02,m03,m04,type = "text")
+stargazer::stargazer(m05,m06,m07,m08,type = "text")
 ```
@@
 #---------------------------------------------------#
-#summary(fitw02_c,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fitw02_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fitw02_c,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE,curve=2)
-
-#summary(fitw02_o,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fitw02_o)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fitw02_o,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE,curve=2)
+kable(rbind(continuo=fitmeasures(fitw02_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            ordinal= fitmeasures(fitw02_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
+      format = "markdown",digits = 3)
+
+cnames <- c("Factor","Indicador","Loading (MLR)","Loading (DWLS)")
+kable(left_join(x = standardizedsolution(fitw02_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
+                y = standardizedsolution(fitw02_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
+                c("lhs","rhs")),
+      format = "markdown",digits = 2,col.names = cnames)
 ```
@@
 #---------------------------------------------------#
-# #summary(mod2_merit_c,standardized=TRUE, fit.measures=TRUE)
-# fitmeasures(mod2_merit_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-# semPlot::semPaths(object = mod2_merit_c,what = "std",thresholds = FALSE, intercepts = FALSE)
-
-#summary(mod2_merit_o,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(mod2_merit_o)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = mod2_merit_o,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE, curve=2, rotation=2)
+kable(rbind(ordinal=fitmeasures(mod2_merit_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
+      format = "markdown",digits = 3)
+
+cnames <- c("Factor","Indicador","Loading (DWLS)")
+kable(standardizedsolution(mod2_merit_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
+      format = "markdown",digits = 2,col.names = cnames)
 ```
@@
                                                            "meritv03_pref_wpart","meritv03_pref_netw"))
-
 #---------------------------------------------------#
-#summary(fitw02a_c,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fitw02a_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fitw02a_c,what = "std",thresholds = FALSE, intercepts = FALSE, curve=2)
-
-#summary(fitw02a_o,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fitw02a_o)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fitw02a_o,what = "std",thresholds = FALSE, intercepts = FALSE, curve=2)
+kable(rbind(continuo=fitmeasures(fitw02a_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            ordinal= fitmeasures(fitw02a_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
+      format = "markdown",digits = 3)
+
+cnames <- c("Factor","Indicador","Loading (MLR)","Loading (DWLS)")
+kable(left_join(x = standardizedsolution(fitw02a_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
+                y = standardizedsolution(fitw02a_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
+                c("lhs","rhs")),
+      format = "markdown",digits = 2,col.names = cnames)
 ```
@@
 #---------------------------------------------------#
-#summary(fitw02b_c,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fitw02b_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fitw02b_c,what = "std",thresholds = FALSE, intercepts = FALSE, curve=2)
-
-#summary(fitw02b_o,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fitw02b_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fitw02b_o,what = "std",thresholds = FALSE, intercepts = FALSE, curve=2)
+kable(rbind(continuo=fitmeasures(fitw02b_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            ordinal= fitmeasures(fitw02b_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
+      format = "markdown",digits = 3)
+
+cnames <- c("Factor","Indicador","Loading (MLR)","Loading (DWLS)")
+kable(left_join(x = standardizedsolution(fitw02b_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
+                y = standardizedsolution(fitw02b_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
+                c("lhs","rhs")),
+      format = "markdown",digits = 2,col.names = cnames)
 ```
@@
 #---------------------------------------------------#
-#summary(fitw02c_c,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fitw02c_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fitw02c_c,what = "std",thresholds = FALSE, intercepts = FALSE)
-
-#summary(fitw02b_o,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fitw02c_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fitw02c_o,what = "std",thresholds = FALSE, intercepts = FALSE)
+kable(rbind(continuo=fitmeasures(fitw02c_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            ordinal= fitmeasures(fitw02c_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
+      format = "markdown",digits = 3)
+
+cnames <- c("Factor","Indicador","Loading (MLR)","Loading (DWLS)")
+kable(left_join(x = standardizedsolution(fitw02c_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
+                y = standardizedsolution(fitw02c_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
+                c("lhs","rhs")),
+      format = "markdown",digits = 2,col.names = cnames)
 ```
@@
 # Resumen ajuste ola 2
+
 ```{r}
@@
 ```{r}
-sjt.fa(data = dat01,show.comm = TRUE)
+# sjt.fa(data = sjlabelled::remove_label(dat01),show.comm = TRUE)
 ```
@@
 #---------------------------------------------------#
-#summary(fitw03c_c,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fitw03c_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-
-#summary(fitw03c_o,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fitw03c_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+kable(rbind(continuo=fitmeasures(fitw03c_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            ordinal= fitmeasures(fitw03c_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
+      format = "markdown",digits = 3)
+
+cnames <- c("Factor","Indicador","Loading (MLR)","Loading (DWLS)")
+kable(left_join(x = standardizedsolution(fitw03c_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
+                y = standardizedsolution(fitw03c_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
+                c("lhs","rhs")),
+      format = "markdown",digits = 2,col.names = cnames)
 ```
 **summary**
+
 > * Ajuste usando estimador MLR  es bueno (CFI = 0.971, RMSEA = 0.052). El chi2 (robusto) es 62.421  
@@
 ```
-
-## Correlación indicadores w01 y w02
 
@@
 pref_merit3 =~pref_effort_w3+pref_talent_w3 
-pref_nmerit3=~pref_wpart_w3 +pref_netw_w3'
+pref_nmerit3=~pref_wpart_w3 +pref_netw_w3
+'
 
@@
 #---------------------------------------------------#
-#summary(fitlong_c,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fitlong_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fitlong_c,what = "std",thresholds = FALSE, intercepts = FALSE)
-
-#summary(fitlong_o,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fitlong_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fitlong_o,what = "std",thresholds = FALSE, intercepts = FALSE,style = "mx")
+kable(rbind(continuo=fitmeasures(fitlong_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            ordinal= fitmeasures(fitlong_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
+      format = "markdown",digits = 3)
+
+cnames <- c("Factor","Indicador","Loading (MLR)","Loading (DWLS)")
+kable(left_join(x = standardizedsolution(fitlong_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
+                y = standardizedsolution(fitlong_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
+                c("lhs","rhs")),
+      format = "markdown",digits = 2,col.names = cnames)
 ```
@@
 
-1. Invarianza entre modalidades en ola 1,2 y 3 (between groups invariance)
-2. Invarianza entre tratamientos en ola 2(between groups invariance)
-
->**NOTA**: En vista de que no existe una librería que permita estimar invarianza longitudinal asumiendo indicadores ordinales, las primeras estimaciones las realizaré asumiendo indicadores continuos (estimator=ML).
+## Invarianza entre modalidades en ola 1,2 y 3 (between groups invariance)
+
+**NOTA**: En vista de que no existe una librería que permita estimar invarianza longitudinal asumiendo indicadores ordinales, las primeras estimaciones las realizaré asumiendo indicadores continuos (estimator=MLR).
 
@@
 load(file = "input/data/proc/datalong.RData")     # Datos formato long
-load(file = "input/data/proc/datawide_123.RData") # Datos formato long
-```
-
-```{r}
-dim(dat04)
+load(file = "input/data/proc/datawide_123.RData") # Datos formato wide 3 olas
+
+wide_m1 <- subset(wide02,grupo==1) # subset for mode 01
+wide_m2 <- subset(wide02,grupo==2) # subset for mode 02
+wide_m3 <- subset(wide02,grupo==3) # subset for mode 03
+```
+
+```{r}
 m_inv_modalidad <- 'perc_merit=~ perc_effort+perc_talent 
@@
 
-fit.conf=cfa(m_inv_modalidad,data=dat04,group="grupo")
-fit.deb =cfa(m_inv_modalidad,data=dat04,group="grupo",group.equal=c("loadings"))
-fit.fuer=cfa(m_inv_modalidad,data=dat04,group="grupo",group.equal=c("loadings","intercepts"))
-fit.str =cfa(m_inv_modalidad,data=dat04,group="grupo",group.equal=c("loadings","intercepts","residuals"))
-
-fitmeasures(fit.conf)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-fitmeasures(fit.deb )[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-fitmeasures(fit.fuer)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-fitmeasures(fit.str )[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-anova(fit.conf, fit.deb, fit.fuer, fit.str)
-```
-
-```{r}
-configural <- '
-# wave 01
-perc_merit1 =~NA*perc_effort_w1+lambda1*perc_effort_w1+perc_talent_w1 
-perc_nmerit1=~NA*perc_wpart_w1 +lambda1*perc_wpart_w1 +perc_netw_w1  
-pref_merit1 =~NA*pref_effort_w1+lambda1*pref_effort_w1+pref_talent_w1 
-pref_nmerit1=~NA*pref_wpart_w1 +lambda1*pref_wpart_w1 +pref_netw_w1
-# wave 02
-perc_merit2 =~NA*perc_effort_w2+lambda1*perc_effort_w2+perc_talent_w2 
-perc_nmerit2=~NA*perc_wpart_w2 +lambda1*perc_wpart_w2 +perc_netw_w2  
-pref_merit2 =~NA*pref_effort_w2+lambda1*pref_effort_w2+pref_talent_w2 
-pref_nmerit2=~NA*pref_wpart_w2 +lambda1*pref_wpart_w2 +pref_netw_w2
-# wave 03
-perc_merit3 =~NA*perc_effort_w3+lambda1*perc_effort_w3+perc_talent_w3 
-perc_nmerit3=~NA*perc_wpart_w3 +lambda1*perc_wpart_w3 +perc_netw_w3  
-pref_merit3 =~NA*pref_effort_w3+lambda1*pref_effort_w3+pref_talent_w3 
-pref_nmerit3=~NA*pref_wpart_w3 +lambda1*pref_wpart_w3 +pref_netw_w3
-
-# INTERCEPTS: Por cada variable latente, debo fijar el intercepto del primer indicador.
-# wave 01--------------------------
-perc_effort_w1 ~ i1*1   # Percepcion merito        
-perc_talent_w1 ~ 1      #      
-perc_wpart_w1 ~ i1*1    # Percepcion no-merito   
-perc_netw_w1 ~ 1        #
-pref_effort_w1 ~ i1*1   # Preferencia merito    
-pref_talent_w1 ~ 1      #  
-pref_wpart_w1 ~ i1*1    # Preferencia no-merito   
-pref_netw_w1 ~ 1        #
-
-# wave 02--------------------------
-perc_effort_w2 ~ i1*1   # Percepcion merito                    
-perc_talent_w2 ~ 1      #               
-perc_wpart_w2 ~ i1*1    # Percepcion no-merito              
-perc_netw_w2 ~ 1        #       
-pref_effort_w2 ~ i1*1   # Preferencia merito                
-pref_talent_w2 ~ 1      #           
-pref_wpart_w2 ~ i1*1    # Preferencia no-merito              
-pref_netw_w2 ~ 1        #
-
-# wave 03--------------------------
-perc_effort_w3 ~ i1*1   # Percepcion merito                     
-perc_talent_w3 ~ 1      #                            
-perc_wpart_w3 ~ i1*1    # Percepcion no-merito                
-perc_netw_w3 ~ 1        #                  
-pref_effort_w3 ~ i1*1   # Preferencia merito                   
-pref_talent_w3 ~ 1      #                        
-pref_wpart_w3 ~ i1*1    # Preferencia no-merito               
-pref_netw_w3 ~ 1        #
-
-# UNIQUE VARIANCES and COVARIANCES.
-
-# LATENT VARIABLES MEANS: Para cada variable latente fijar la media
-# Fix T1 common factor mean to zero
-perc_merit1 ~ 0*1
-# Freely estimate T2-T3 common factor means
+fit.conf=cfa(m_inv_modalidad,data=dat04,group="grupo",estimator="MLR")
+fit.deb =cfa(m_inv_modalidad,data=dat04,group="grupo",group.equal=c("loadings"),estimator="MLR")
+fit.fuer=cfa(m_inv_modalidad,data=dat04,group="grupo",group.equal=c("loadings","intercepts"),estimator="MLR")
+fit.str =cfa(m_inv_modalidad,data=dat04,group="grupo",group.equal=c("loadings","intercepts","residuals"),estimator="MLR")
+```
+
+```{r}
+kable(rbind(fit.conf=fitmeasures(fit.conf)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            fit.deb =fitmeasures(fit.deb )[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            fit.fuer=fitmeasures(fit.fuer)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            fit.str =fitmeasures(fit.str )[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),digits = 3,format = "markdown")
+
+kable(anova(fit.conf, fit.deb, fit.fuer, fit.str),format = "markdown",digits = 3)
+```
+
+## Invarianza Longitudinal
+
+> Adapted from Templin (2018), consulted at 24th of march of 2020 from: [here](https://jonathantemplin.com/wp-content/uploads/2017/08/EPSY906_Example07b_CFA_Longitudinal_Invariance.nb_.html)
+
+### 1. Configural Longitudinal Invariance Model (everything separate across time)
+
+```{r configural}
+configuralSyntax = "
+#===================================================================================================
+#Factor loadings all freely estimated in each group with labels
+perc_merit1  =~ L1T1*perc_effort_w1 + L2T1*perc_talent_w1 
+perc_merit2  =~ L1T2*perc_effort_w2 + L2T2*perc_talent_w2
+perc_merit3  =~ L1T3*perc_effort_w3 + L2T3*perc_talent_w3
+
+perc_nmerit1 =~ L3T1*perc_wpart_w1  + L4T1*perc_netw_w1  
+perc_nmerit2 =~ L3T2*perc_wpart_w2  + L4T2*perc_netw_w2  
+perc_nmerit3 =~ L3T3*perc_wpart_w3  + L4T3*perc_netw_w3
+
+pref_merit1  =~ L5T1*pref_effort_w1 + L6T1*pref_talent_w1
+pref_merit2  =~ L5T2*pref_effort_w2 + L6T2*pref_talent_w2
+pref_merit3  =~ L5T3*pref_effort_w3 + L6T3*pref_talent_w3
+
+pref_nmerit1 =~ L7T1*pref_wpart_w1 + L8T1*pref_netw_w1
+pref_nmerit2 =~ L7T2*pref_wpart_w2 + L8T2*pref_netw_w2
+pref_nmerit3 =~ L7T3*pref_wpart_w3 + L8T3*pref_netw_w3
+
+#===================================================================================================
+#Item intercepts all freely estimated in each group with labels (8 items for each wave 8x3 = 16 intercepts)
+perc_effort_w1 ~ I1T1*1; perc_talent_w1 ~ I2T1*1; perc_wpart_w1 ~ I3T1*1; perc_netw_w1 ~ I4T1*1; pref_effort_w1 ~ I5T1*1; pref_talent_w1 ~ I6T1*1; pref_wpart_w1 ~ I7T1*1; pref_netw_w1 ~ I8T1*1; 
+perc_effort_w2 ~ I1T2*1; perc_talent_w2 ~ I2T2*1; perc_wpart_w2 ~ I3T2*1; perc_netw_w2 ~ I4T2*1; pref_effort_w2 ~ I5T2*1; pref_talent_w2 ~ I6T2*1; pref_wpart_w2 ~ I7T2*1; pref_netw_w2 ~ I8T2*1; 
+perc_effort_w3 ~ I1T3*1; perc_talent_w3 ~ I2T3*1; perc_wpart_w3 ~ I3T3*1; perc_netw_w3 ~ I4T3*1; pref_effort_w3 ~ I5T3*1; pref_talent_w3 ~ I6T3*1; pref_wpart_w3 ~ I7T3*1; pref_netw_w3 ~ I8T3*1;
+
+#===================================================================================================
+#Residual variances all freely estimated in each group with labels 
+perc_effort_w1 ~~ E1T1*perc_effort_w1; perc_talent_w1 ~~ E2T1*perc_talent_w1; perc_wpart_w1 ~~ E3T1*perc_wpart_w1; perc_netw_w1 ~~ E4T1*perc_netw_w1; pref_effort_w1 ~~ E5T1*pref_effort_w1; pref_talent_w1 ~~ E6T1*pref_talent_w1; pref_wpart_w1 ~~ E7T1*pref_wpart_w1; pref_netw_w1 ~~ E8T1*pref_netw_w1;
+perc_effort_w2 ~~ E1T2*perc_effort_w2; perc_talent_w2 ~~ E2T2*perc_talent_w2; perc_wpart_w2 ~~ E3T2*perc_wpart_w2; perc_netw_w2 ~~ E4T2*perc_netw_w2; pref_effort_w2 ~~ E5T2*pref_effort_w2; pref_talent_w2 ~~ E6T2*pref_talent_w2; pref_wpart_w2 ~~ E7T2*pref_wpart_w2; pref_netw_w2 ~~ E8T2*pref_netw_w2;
+perc_effort_w3 ~~ E1T3*perc_effort_w3; perc_talent_w3 ~~ E2T3*perc_talent_w3; perc_wpart_w3 ~~ E3T3*perc_wpart_w3; perc_netw_w3 ~~ E4T3*perc_netw_w3; pref_effort_w3 ~~ E5T3*pref_effort_w3; pref_talent_w3 ~~ E6T3*pref_talent_w3; pref_wpart_w3 ~~ E7T3*pref_wpart_w3; pref_netw_w3 ~~ E8T2*pref_netw_w3;
+#===================================================================================================
+#Residual covariances freely estimated in each group with labels (C1T12= covariance 1 between time 1&2)
+
+perc_effort_w1 ~~ C1T12*perc_effort_w2 + C1T13*perc_effort_w3; perc_effort_w2 ~~ C1T23*perc_effort_w3;  
+perc_talent_w1 ~~ C2T12*perc_talent_w2 + C2T13*perc_talent_w3; perc_talent_w2 ~~ C2T23*perc_talent_w3;  
+perc_wpart_w1  ~~ C3T12*perc_wpart_w2  + C3T13*perc_wpart_w3 ; perc_wpart_w2  ~~ C3T23*perc_wpart_w3 ; 
+perc_netw_w1   ~~ C4T12*perc_netw_w2   + C4T13*perc_netw_w3  ; perc_netw_w2   ~~ C4T23*perc_netw_w3  ;
+pref_effort_w1 ~~ C5T12*pref_effort_w2 + C5T13*pref_effort_w3; pref_effort_w2 ~~ C5T23*pref_effort_w3;  
+pref_talent_w1 ~~ C6T12*pref_talent_w2 + C6T13*pref_talent_w3; pref_talent_w2 ~~ C6T23*pref_talent_w3;  
+pref_wpart_w1  ~~ C7T12*pref_wpart_w2  + C7T13*pref_wpart_w3 ; pref_wpart_w2  ~~ C7T23*pref_wpart_w3 ; 
+pref_netw_w1   ~~ C8T12*pref_netw_w2   + C8T13*pref_netw_w3  ; pref_netw_w2   ~~ C8T23*pref_netw_w3  ;
+
+#===================================================================================================
+#Factor variance fixed to 1 for identification in each group
+perc_merit1  ~~ 1*perc_merit1
+perc_merit2  ~~ 1*perc_merit2
+perc_merit3  ~~ 1*perc_merit3
+
+perc_nmerit1 ~~ 1*perc_nmerit1
+perc_nmerit2 ~~ 1*perc_nmerit2
+perc_nmerit3 ~~ 1*perc_nmerit3
+
+pref_merit1  ~~ 1*pref_merit1
+pref_merit2  ~~ 1*pref_merit2
+pref_merit3  ~~ 1*pref_merit3
+
+pref_nmerit1 ~~ 1*pref_nmerit1
+pref_nmerit2 ~~ 1*pref_nmerit2
+pref_nmerit3 ~~ 1*pref_nmerit3
+
+#===================================================================================================
+#Factor mean fixed to zero for identification in each group
+perc_merit1  ~ 0
+perc_merit2  ~ 0
+perc_merit3  ~ 0
+
+perc_nmerit1 ~ 0
+perc_nmerit2 ~ 0
+perc_nmerit3 ~ 0
+
+pref_merit1  ~ 0
+pref_merit2  ~ 0
+pref_merit3  ~ 0
+
+pref_nmerit1 ~ 0
+pref_nmerit2 ~ 0
+pref_nmerit3 ~ 0
+
+#===================================================================================================
+#Factor covariances all freely estimated
+perc_merit1  ~~ perc_merit2 + perc_merit3
+perc_merit2  ~~ perc_merit3
+
+perc_nmerit1 ~~ perc_nmerit2 + perc_nmerit3
+perc_nmerit2 ~~ perc_nmerit3
+
+pref_merit1  ~~ pref_merit2 + pref_merit3
+pref_merit2  ~~ pref_merit3
+
+pref_nmerit1 ~~ pref_nmerit2 + pref_nmerit3
+pref_nmerit2 ~~ pref_nmerit3
+#===================================================================================================
+"
+```
+
+```{r}
+#configural invariance for mode 01
+conf_m1 = lavaan(model = configuralSyntax, data = wide_m1, estimator = "MLR")
+```
+
+```{r}
+#configural invariance for mode 02
+conf_m2 = lavaan(model = configuralSyntax, data = wide_m2, estimator = "MLR")
+```
+
+```{r}
+#configural invariance for mode 03
+conf_m3 = lavaan(model = configuralSyntax, data = wide_m3, estimator = "MLR")
+```
+
+```{r}
+#configural invariance for complete sample (mode 1, 2 & 3)
+conf = lavaan(model = configuralSyntax, data = wide02, estimator = "MLR")
+```
+
+```{r}
+kable(rbind(fit.conf=fitmeasures(conf)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),format = "markdown",digits = 3)
+```
+
+### 2. Metric Invariance Models (ALL loadings held equal across time - identified model using Time1 Factor Variance = 1)
+
+```{r metric}
+metricSyntax1 = "
+#===================================================================================================
+#Factor loadings all constrained across groups with labels *****
+perc_merit1  =~ L1*perc_effort_w1 + L2*perc_talent_w1 
+perc_merit2  =~ L1*perc_effort_w2 + L2*perc_talent_w2
+perc_merit3  =~ L1*perc_effort_w3 + L2*perc_talent_w3
+
+perc_nmerit1 =~ L3*perc_wpart_w1  + L4*perc_netw_w1  
+perc_nmerit2 =~ L3*perc_wpart_w2  + L4*perc_netw_w2  
+perc_nmerit3 =~ L3*perc_wpart_w3  + L4*perc_netw_w3
+
+pref_merit1  =~ L5*pref_effort_w1 + L6*pref_talent_w1
+pref_merit2  =~ L5*pref_effort_w2 + L6*pref_talent_w2
+pref_merit3  =~ L5*pref_effort_w3 + L6*pref_talent_w3
+
+pref_nmerit1 =~ L7*pref_wpart_w1 + L8*pref_netw_w1
+pref_nmerit2 =~ L7*pref_wpart_w2 + L8*pref_netw_w2
+pref_nmerit3 =~ L7*pref_wpart_w3 + L8*pref_netw_w3
+
+#===================================================================================================
+#Item intercepts all freely estimated in each group with labels (8 items for each wave 8x3 = 16 intercepts)
+perc_effort_w1 ~ I1T1*1; perc_talent_w1 ~ I2T1*1; perc_wpart_w1 ~ I3T1*1; perc_netw_w1 ~ I4T1*1; pref_effort_w1 ~ I5T1*1; pref_talent_w1 ~ I6T1*1; pref_wpart_w1 ~ I7T1*1; pref_netw_w1 ~ I8T1*1; 
+perc_effort_w2 ~ I1T2*1; perc_talent_w2 ~ I2T2*1; perc_wpart_w2 ~ I3T2*1; perc_netw_w2 ~ I4T2*1; pref_effort_w2 ~ I5T2*1; pref_talent_w2 ~ I6T2*1; pref_wpart_w2 ~ I7T2*1; pref_netw_w2 ~ I8T2*1; 
+perc_effort_w3 ~ I1T3*1; perc_talent_w3 ~ I2T3*1; perc_wpart_w3 ~ I3T3*1; perc_netw_w3 ~ I4T3*1; pref_effort_w3 ~ I5T3*1; pref_talent_w3 ~ I6T3*1; pref_wpart_w3 ~ I7T3*1; pref_netw_w3 ~ I8T3*1;
+
+#===================================================================================================
+#Residual variances all freely estimated in each group with labels 
+perc_effort_w1 ~~ E1T1*perc_effort_w1; perc_talent_w1 ~~ E2T1*perc_talent_w1; perc_wpart_w1 ~~ E3T1*perc_wpart_w1; perc_netw_w1 ~~ E4T1*perc_netw_w1; pref_effort_w1 ~~ E5T1*pref_effort_w1; pref_talent_w1 ~~ E6T1*pref_talent_w1; pref_wpart_w1 ~~ E7T1*pref_wpart_w1; pref_netw_w1 ~~ E8T1*pref_netw_w1;
+perc_effort_w2 ~~ E1T2*perc_effort_w2; perc_talent_w2 ~~ E2T2*perc_talent_w2; perc_wpart_w2 ~~ E3T2*perc_wpart_w2; perc_netw_w2 ~~ E4T2*perc_netw_w2; pref_effort_w2 ~~ E5T2*pref_effort_w2; pref_talent_w2 ~~ E6T2*pref_talent_w2; pref_wpart_w2 ~~ E7T2*pref_wpart_w2; pref_netw_w2 ~~ E8T2*pref_netw_w2;
+perc_effort_w3 ~~ E1T3*perc_effort_w3; perc_talent_w3 ~~ E2T3*perc_talent_w3; perc_wpart_w3 ~~ E3T3*perc_wpart_w3; perc_netw_w3 ~~ E4T3*perc_netw_w3; pref_effort_w3 ~~ E5T3*pref_effort_w3; pref_talent_w3 ~~ E6T3*pref_talent_w3; pref_wpart_w3 ~~ E7T3*pref_wpart_w3; pref_netw_w3 ~~ E8T2*pref_netw_w3;
+#===================================================================================================
+#Residual covariances freely estimated in each group with labels (C1T12= covariance 1 between time 1&2)
+
+perc_effort_w1 ~~ C1T12*perc_effort_w2 + C1T13*perc_effort_w3; perc_effort_w2 ~~ C1T23*perc_effort_w3;  
+perc_talent_w1 ~~ C2T12*perc_talent_w2 + C2T13*perc_talent_w3; perc_talent_w2 ~~ C2T23*perc_talent_w3;  
+perc_wpart_w1  ~~ C3T12*perc_wpart_w2  + C3T13*perc_wpart_w3 ; perc_wpart_w2  ~~ C3T23*perc_wpart_w3 ; 
+perc_netw_w1   ~~ C4T12*perc_netw_w2   + C4T13*perc_netw_w3  ; perc_netw_w2   ~~ C4T23*perc_netw_w3  ;
+pref_effort_w1 ~~ C5T12*pref_effort_w2 + C5T13*pref_effort_w3; pref_effort_w2 ~~ C5T23*pref_effort_w3;  
+pref_talent_w1 ~~ C6T12*pref_talent_w2 + C6T13*pref_talent_w3; pref_talent_w2 ~~ C6T23*pref_talent_w3;  
+pref_wpart_w1  ~~ C7T12*pref_wpart_w2  + C7T13*pref_wpart_w3 ; pref_wpart_w2  ~~ C7T23*pref_wpart_w3 ; 
+pref_netw_w1   ~~ C8T12*pref_netw_w2   + C8T13*pref_netw_w3  ; pref_netw_w2   ~~ C8T23*pref_netw_w3  ;
+
+#===================================================================================================
+# Factor variance fixed to 1 for identification in first group; estimated in others *****
+perc_merit1  ~~ 1*perc_merit1
+perc_merit2  ~~ perc_merit2  # deje fijado la varianza de los factores en 1 (esto lo cambie yo respecto al script original)
+perc_merit3  ~~ perc_merit3  # deje fijado la varianza de los factores en 1 (esto lo cambie yo respecto al script original)
+
+perc_nmerit1 ~~ 1*perc_nmerit1
+perc_nmerit2 ~~ perc_nmerit2 # deje fijado la varianza de los factores en 1 (esto lo cambie yo respecto al script original)
+perc_nmerit3 ~~ perc_nmerit3 # deje fijado la varianza de los factores en 1 (esto lo cambie yo respecto al script original)
+
+pref_merit1  ~~ 1*pref_merit1
+pref_merit2  ~~ pref_merit2  # deje fijado la varianza de los factores en 1 (esto lo cambie yo respecto al script original)
+pref_merit3  ~~ pref_merit3  # deje fijado la varianza de los factores en 1 (esto lo cambie yo respecto al script original)
+
+pref_nmerit1 ~~ 1*pref_nmerit1
+pref_nmerit2 ~~ pref_nmerit2 # deje fijado la varianza de los factores en 1 (esto lo cambie yo respecto al script original)
+pref_nmerit3 ~~ pref_nmerit3 # deje fijado la varianza de los factores en 1 (esto lo cambie yo respecto al script original)
+
+#===================================================================================================
+#Factor mean fixed to zero for identification in each group
+perc_merit1  ~ 0
+perc_merit2  ~ 0
+perc_merit3  ~ 0
+
+perc_nmerit1 ~ 0
+perc_nmerit2 ~ 0
+perc_nmerit3 ~ 0
+
+pref_merit1  ~ 0
+pref_merit2  ~ 0
+pref_merit3  ~ 0
+
+pref_nmerit1 ~ 0
+pref_nmerit2 ~ 0
+pref_nmerit3 ~ 0
+
+#===================================================================================================
+#Factor covariances all freely estimated
+perc_merit1  ~~ perc_merit2 + perc_merit3
+perc_merit2  ~~ perc_merit3
+
+perc_nmerit1 ~~ perc_nmerit2 + perc_nmerit3
+perc_nmerit2 ~~ perc_nmerit3
+
+pref_merit1  ~~ pref_merit2 + pref_merit3
+pref_merit2  ~~ pref_merit3
+
+pref_nmerit1 ~~ pref_nmerit2 + pref_nmerit3
+pref_nmerit2 ~~ pref_nmerit3
+#===================================================================================================
+"
+```
+
+```{r}
+#metric invariance for mode 01
+metr_m1A = lavaan(model = metricSyntax1, data = wide_m1, estimator = "MLR")
+```
+
+```{r}
+#metric invariance for mode 02
+metr_m2A = lavaan(model = metricSyntax1, data = wide_m2, estimator = "MLR")
+```
+
+```{r}
+#metric invariance for mode 03
+metr_m3A = lavaan(model = metricSyntax1, data = wide_m3, estimator = "MLR")
+```
+
+```{r}
+#metric invariance for complete sample (mode 1, 2 & 3)
+metrA = lavaan(model = metricSyntax1, data = wide02, estimator = "MLR")
+```
+
+```{r}
+#lavTestLRT(metr_m1A, conf_m1,scaled.shifted = TRUE, method = "satorra.bentler.2001")#conf_m1 was not estimated using scaled (robust) fit measures  
+# lavTestLRT(metr_m2A, conf_m2,scaled.shifted = TRUE, method = "satorra.bentler.2001") # invariantes
+# lavTestLRT(metr_m3A, conf_m3,scaled.shifted = TRUE, method = "satorra.bentler.2001") # invariantes
+
+kable(rbind(fit.conf=fitmeasures(conf)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            fit.metrA =fitmeasures(metrA)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),digits = 3,format = "markdown")
+
+kable(lavTestLRT(metrA,conf,scaled.shifted = TRUE, method = "satorra.bentler.2001"),digits = 3,format = "markdown") # no es invariante, ver modind
+```
+
+**Mod Indices**:
+
+```{r, modind metric1}
+#get modification indices for all parameters:
+metricMI1 = modificationindices(metrA, free.remove = FALSE, maximum.number = 10000)
+#restrict output to only factor loading parameters
+metricMI1a = metricMI1[which(metricMI1$op == "=~"),]
+#restrict output to only factors and items with same time point
+metricMI1a$factorTime = metricMI1a$lhs
+#extract time characters from items
+metricMI1a$itemTime = ifelse(test =endsWith(x = metricMI1a$rhs,suffix = "w1"),yes = "wave01",
+                              no = ifelse(test = endsWith(x = metricMI1a$rhs,suffix = "w2"),yes = "wave02",
+                                          no = "wave03"))
+
+metricMI1a$factorTime = ifelse(test =endsWith(x = metricMI1a$lhs,suffix = "1"),yes = "wave01",
+                               no = ifelse(test = endsWith(x = metricMI1a$lhs,suffix = "2"),yes = "wave02",
+                                           no = "wave03"))
+metricMI1a$factor     = ifelse(test =startsWith(x = metricMI1a$lhs,prefix = "perc_merit"),yes = "perc_merit",
+                               no = ifelse(test = startsWith(x = metricMI1a$lhs,prefix = "perc_nmerit"),yes = "perc_nmerit",
+                                           no = ifelse(test = startsWith(x = metricMI1a$lhs,prefix = "pref_merit"),yes = "pref_merit",
+                                                       no = "pref_nmerit" )))
+
+metricMI1a$item     = ifelse(test =startsWith(x = metricMI1a$rhs,prefix = "perc_effort") | startsWith(x = metricMI1a$rhs,prefix = "perc_talent"),yes = "perc_merit",
+                               no = ifelse(test = startsWith(x = metricMI1a$rhs,prefix = "perc_wpart")  | startsWith(x = metricMI1a$rhs,prefix = "perc_netw"),yes = "perc_nmerit",
+                                           no = ifelse(test = startsWith(x = metricMI1a$rhs,prefix = "pref_effort") | startsWith(x = metricMI1a$rhs,prefix = "pref_talent"),yes = "pref_merit",
+                                                       no = "pref_nmerit" )))
+
+metricMI1b<- metricMI1a %>% filter(factorTime==itemTime & factor==item) # quedarme con los que son de la misma ola y mismo factor
+
+#show only MIs for same time
+metricMI1b <- metricMI1b[order(-metricMI1b$mi),]
+```
+
+```{r metric2}
+metricSyntax2 = "
+#===================================================================================================
+#Factor loadings all constrained across groups with labels *****
+perc_merit1  =~ L1*perc_effort_w1 + L2*perc_talent_w1 
+perc_merit2  =~ L1*perc_effort_w2 + L2*perc_talent_w2
+perc_merit3  =~ L1*perc_effort_w3 + L2*perc_talent_w3
+perc_nmerit1 =~ L3*perc_wpart_w1  + L4T1*perc_netw_w1 #FREE PARAMETER    
+perc_nmerit2 =~ L3*perc_wpart_w2  + L4*perc_netw_w2  
+perc_nmerit3 =~ L3*perc_wpart_w3  + L4*perc_netw_w3    
+pref_merit1  =~ L5*pref_effort_w1 + L6*pref_talent_w1
+pref_merit2  =~ L5*pref_effort_w2 + L6*pref_talent_w2
+pref_merit3  =~ L5*pref_effort_w3 + L6*pref_talent_w3
+pref_nmerit1 =~ L7*pref_wpart_w1  + L8*pref_netw_w1    
+pref_nmerit2 =~ L7*pref_wpart_w2  + L8*pref_netw_w2  
+pref_nmerit3 =~ L7*pref_wpart_w3  + L8*pref_netw_w3
+
+#===================================================================================================
+#Item intercepts all freely estimated in each group with labels (8 items for each wave 8x3 = 16 intercepts)
+perc_effort_w1 ~ I1T1*1; perc_talent_w1 ~ I2T1*1; perc_wpart_w1 ~ I3T1*1; perc_netw_w1 ~ I4T1*1; pref_effort_w1 ~ I5T1*1; pref_talent_w1 ~ I6T1*1; pref_wpart_w1 ~ I7T1*1; pref_netw_w1 ~ I8T1*1; 
+perc_effort_w2 ~ I1T2*1; perc_talent_w2 ~ I2T2*1; perc_wpart_w2 ~ I3T2*1; perc_netw_w2 ~ I4T2*1; pref_effort_w2 ~ I5T2*1; pref_talent_w2 ~ I6T2*1; pref_wpart_w2 ~ I7T2*1; pref_netw_w2 ~ I8T2*1; 
+perc_effort_w3 ~ I1T3*1; perc_talent_w3 ~ I2T3*1; perc_wpart_w3 ~ I3T3*1; perc_netw_w3 ~ I4T3*1; pref_effort_w3 ~ I5T3*1; pref_talent_w3 ~ I6T3*1; pref_wpart_w3 ~ I7T3*1; pref_netw_w3 ~ I8T3*1;
+
+#===================================================================================================
+#Residual variances all freely estimated in each group with labels 
+perc_effort_w1 ~~ E1T1*perc_effort_w1; perc_talent_w1 ~~ E2T1*perc_talent_w1; perc_wpart_w1 ~~ E3T1*perc_wpart_w1; perc_netw_w1 ~~ E4T1*perc_netw_w1; pref_effort_w1 ~~ E5T1*pref_effort_w1; pref_talent_w1 ~~ E6T1*pref_talent_w1; pref_wpart_w1 ~~ E7T1*pref_wpart_w1; pref_netw_w1 ~~ E8T1*pref_netw_w1;
+perc_effort_w2 ~~ E1T2*perc_effort_w2; perc_talent_w2 ~~ E2T2*perc_talent_w2; perc_wpart_w2 ~~ E3T2*perc_wpart_w2; perc_netw_w2 ~~ E4T2*perc_netw_w2; pref_effort_w2 ~~ E5T2*pref_effort_w2; pref_talent_w2 ~~ E6T2*pref_talent_w2; pref_wpart_w2 ~~ E7T2*pref_wpart_w2; pref_netw_w2 ~~ E8T2*pref_netw_w2;
+perc_effort_w3 ~~ E1T3*perc_effort_w3; perc_talent_w3 ~~ E2T3*perc_talent_w3; perc_wpart_w3 ~~ E3T3*perc_wpart_w3; perc_netw_w3 ~~ E4T3*perc_netw_w3; pref_effort_w3 ~~ E5T3*pref_effort_w3; pref_talent_w3 ~~ E6T3*pref_talent_w3; pref_wpart_w3 ~~ E7T3*pref_wpart_w3; pref_netw_w3 ~~ E8T2*pref_netw_w3;
+#===================================================================================================
+#Residual covariances freely estimated in each group with labels (C1T12= covariance 1 between time 1&2)
+perc_effort_w1 ~~ C1T12*perc_effort_w2 + C1T13*perc_effort_w3; perc_effort_w2 ~~ C1T23*perc_effort_w3;  
+perc_talent_w1 ~~ C2T12*perc_talent_w2 + C2T13*perc_talent_w3; perc_talent_w2 ~~ C2T23*perc_talent_w3;  
+perc_wpart_w1  ~~ C3T12*perc_wpart_w2  + C3T13*perc_wpart_w3 ; perc_wpart_w2  ~~ C3T23*perc_wpart_w3 ; 
+perc_netw_w1   ~~ C4T12*perc_netw_w2   + C4T13*perc_netw_w3  ; perc_netw_w2   ~~ C4T23*perc_netw_w3  ;
+pref_effort_w1 ~~ C5T12*pref_effort_w2 + C5T13*pref_effort_w3; pref_effort_w2 ~~ C5T23*pref_effort_w3;  
+pref_talent_w1 ~~ C6T12*pref_talent_w2 + C6T13*pref_talent_w3; pref_talent_w2 ~~ C6T23*pref_talent_w3;  
+pref_wpart_w1  ~~ C7T12*pref_wpart_w2  + C7T13*pref_wpart_w3 ; pref_wpart_w2  ~~ C7T23*pref_wpart_w3 ; 
+pref_netw_w1   ~~ C8T12*pref_netw_w2   + C8T13*pref_netw_w3  ; pref_netw_w2   ~~ C8T23*pref_netw_w3  ;
+
+#===================================================================================================
+# Factor variance fixed to 1 for identification in first group; estimated in others *****
+perc_merit1  ~~ 1*perc_merit1
+perc_merit2  ~~ perc_merit2
+perc_merit3  ~~ perc_merit3
+perc_nmerit1 ~~ 1*perc_nmerit1
+perc_nmerit2 ~~ perc_nmerit2
+perc_nmerit3 ~~ perc_nmerit3
+pref_merit1  ~~ 1*pref_merit1
+pref_merit2  ~~ pref_merit2
+pref_merit3  ~~ pref_merit3
+pref_nmerit1 ~~ 1*pref_nmerit1 
+pref_nmerit2 ~~ pref_nmerit2
+pref_nmerit3 ~~ pref_nmerit3
+
+#===================================================================================================
+#Factor mean fixed to zero for identification in each group
+perc_merit1  ~ 0
+perc_merit2  ~ 0
+perc_merit3  ~ 0
+perc_nmerit1 ~ 0
+perc_nmerit2 ~ 0
+perc_nmerit3 ~ 0
+pref_merit1  ~ 0
+pref_merit2  ~ 0
+pref_merit3  ~ 0
+pref_nmerit1 ~ 0
+pref_nmerit2 ~ 0
+pref_nmerit3 ~ 0
+
+#===================================================================================================
+#Factor covariances all freely estimated
+perc_merit1  ~~ perc_merit2 + perc_merit3
+perc_merit2  ~~ perc_merit3
+perc_nmerit1 ~~ perc_nmerit2 + perc_nmerit3
+perc_nmerit2 ~~ perc_nmerit3
+pref_merit1  ~~ pref_merit2 + pref_merit3
+pref_merit2  ~~ pref_merit3
+pref_nmerit1 ~~ pref_nmerit2 + pref_nmerit3
+pref_nmerit2 ~~ pref_nmerit3
+"
+```
+
+```{r}
+#metric invariance (v02) for complete sample (mode 1, 2 & 3) +  free parameter
+metrA2 = lavaan(model = metricSyntax2, data = wide02, estimator = "MLR")
+
+kable(rbind(fit.conf=fitmeasures(conf)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            fit.metrA =fitmeasures(metrA)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            fit.metrA2 =fitmeasures(metrA2)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),digits = 3,format = "markdown")
+
+#fit.metrA  = all factor loading restricted
+#fit.metrA2 = one free estimated factor loading
+kable(lavTestLRT(metrA2,conf,scaled.shifted = TRUE,method = "satorra.bentler.2001"),format = "markdown",digits = 3) #non invariant, see modind metric2 below
+```
+
+**NOTA**: El modelo que deja libre 1 parámetro (metrA2) mejora el ajuste, pero la diff del chi2 sigue siendo significativa al 0.05.
+
+**Mod Indices**:
+
+```{r, modind metric2}
+#get modification indices for all parameters:
+metricMI1 = modificationindices(metrA2, free.remove = FALSE, maximum.number = 10000)
+#restrict output to only factor loading parameters
+metricMI1a = metricMI1[which(metricMI1$op == "=~"),]
+#restrict output to only factors and items with same time point
+metricMI1a$factorTime = metricMI1a$lhs
+#extract time characters from items
+metricMI1a$itemTime   = ifelse(test =endsWith(x = metricMI1a$rhs,suffix = "w1"),yes = "wave01",
+                                no = ifelse(test = endsWith(x = metricMI1a$rhs,suffix = "w2"),yes = "wave02",
+                                            no = "wave03"))
+
+metricMI1a$factorTime = ifelse(test =endsWith(x = metricMI1a$lhs,suffix = "1"),yes = "wave01",
+                               no = ifelse(test = endsWith(x = metricMI1a$lhs,suffix = "2"),yes = "wave02",
+                                           no = "wave03"))
+metricMI1a$factor     = ifelse(test =startsWith(x = metricMI1a$lhs,prefix = "perc_merit"),yes = "perc_merit",
+                               no = ifelse(test = startsWith(x = metricMI1a$lhs,prefix = "perc_nmerit"),yes = "perc_nmerit",
+                                           no = ifelse(test = startsWith(x = metricMI1a$lhs,prefix = "pref_merit"),yes = "pref_merit",
+                                                       no = "pref_nmerit" )))
+
+metricMI1a$item      = ifelse(test =startsWith(x = metricMI1a$rhs,prefix = "perc_effort") | startsWith(x = metricMI1a$rhs,prefix = "perc_talent"),yes = "perc_merit",
+                              no = ifelse(test = startsWith(x = metricMI1a$rhs,prefix = "perc_wpart")  | startsWith(x = metricMI1a$rhs,prefix = "perc_netw"),yes = "perc_nmerit",
+                                          no = ifelse(test = startsWith(x = metricMI1a$rhs,prefix = "pref_effort") | startsWith(x = metricMI1a$rhs,prefix = "pref_talent"),yes = "pref_merit",
+                                                      no = "pref_nmerit" )))
+
+metricMI1b<- metricMI1a %>% filter(factorTime==itemTime & factor==item) # quedarme con los que son de la misma ola y mismo factor
+
+#show only MIs for same time
+metricMI1b <- metricMI1b[order(-metricMI1b$mi),]
+```
+
+```{r metric3}
+metricSyntax3 = "
+#===================================================================================================
+#Factor loadings all constrained across groups with labels *****
+perc_merit1  =~ L1*perc_effort_w1 + L2*perc_talent_w1 
+perc_merit2  =~ L1*perc_effort_w2 + L2*perc_talent_w2
+perc_merit3  =~ L1*perc_effort_w3 + L2*perc_talent_w3
+perc_nmerit1 =~ L3*perc_wpart_w1  + L4T1*perc_netw_w1  #FREE PARAMETER    
+perc_nmerit2 =~ L3*perc_wpart_w2  + L4*perc_netw_w2  
+perc_nmerit3 =~ L3*perc_wpart_w3  + L4*perc_netw_w3    
+pref_merit1  =~ L5*pref_effort_w1 + L6T1*pref_talent_w1 #FREE PARAMETER
+pref_merit2  =~ L5*pref_effort_w2 + L6*pref_talent_w2
+pref_merit3  =~ L5*pref_effort_w3 + L6*pref_talent_w3
+pref_nmerit1 =~ L7*pref_wpart_w1  + L8*pref_netw_w1    
+pref_nmerit2 =~ L7*pref_wpart_w2  + L8*pref_netw_w2  
+pref_nmerit3 =~ L7*pref_wpart_w3  + L8*pref_netw_w3
+
+#===================================================================================================
+#Item intercepts all freely estimated in each group with labels (8 items for each wave 8x3 = 16 intercepts)
+perc_effort_w1 ~ I1T1*1; perc_talent_w1 ~ I2T1*1; perc_wpart_w1 ~ I3T1*1; perc_netw_w1 ~ I4T1*1; pref_effort_w1 ~ I5T1*1; pref_talent_w1 ~ I6T1*1; pref_wpart_w1 ~ I7T1*1; pref_netw_w1 ~ I8T1*1; 
+perc_effort_w2 ~ I1T2*1; perc_talent_w2 ~ I2T2*1; perc_wpart_w2 ~ I3T2*1; perc_netw_w2 ~ I4T2*1; pref_effort_w2 ~ I5T2*1; pref_talent_w2 ~ I6T2*1; pref_wpart_w2 ~ I7T2*1; pref_netw_w2 ~ I8T2*1; 
+perc_effort_w3 ~ I1T3*1; perc_talent_w3 ~ I2T3*1; perc_wpart_w3 ~ I3T3*1; perc_netw_w3 ~ I4T3*1; pref_effort_w3 ~ I5T3*1; pref_talent_w3 ~ I6T3*1; pref_wpart_w3 ~ I7T3*1; pref_netw_w3 ~ I8T3*1;
+
+#===================================================================================================
+#Residual variances all freely estimated in each group with labels 
+perc_effort_w1 ~~ E1T1*perc_effort_w1; perc_talent_w1 ~~ E2T1*perc_talent_w1; perc_wpart_w1 ~~ E3T1*perc_wpart_w1; perc_netw_w1 ~~ E4T1*perc_netw_w1; pref_effort_w1 ~~ E5T1*pref_effort_w1; pref_talent_w1 ~~ E6T1*pref_talent_w1; pref_wpart_w1 ~~ E7T1*pref_wpart_w1; pref_netw_w1 ~~ E8T1*pref_netw_w1;
+perc_effort_w2 ~~ E1T2*perc_effort_w2; perc_talent_w2 ~~ E2T2*perc_talent_w2; perc_wpart_w2 ~~ E3T2*perc_wpart_w2; perc_netw_w2 ~~ E4T2*perc_netw_w2; pref_effort_w2 ~~ E5T2*pref_effort_w2; pref_talent_w2 ~~ E6T2*pref_talent_w2; pref_wpart_w2 ~~ E7T2*pref_wpart_w2; pref_netw_w2 ~~ E8T2*pref_netw_w2;
+perc_effort_w3 ~~ E1T3*perc_effort_w3; perc_talent_w3 ~~ E2T3*perc_talent_w3; perc_wpart_w3 ~~ E3T3*perc_wpart_w3; perc_netw_w3 ~~ E4T3*perc_netw_w3; pref_effort_w3 ~~ E5T3*pref_effort_w3; pref_talent_w3 ~~ E6T3*pref_talent_w3; pref_wpart_w3 ~~ E7T3*pref_wpart_w3; pref_netw_w3 ~~ E8T2*pref_netw_w3;
+#===================================================================================================
+#Residual covariances freely estimated in each group with labels (C1T12= covariance 1 between time 1&2)
+perc_effort_w1 ~~ C1T12*perc_effort_w2 + C1T13*perc_effort_w3; perc_effort_w2 ~~ C1T23*perc_effort_w3;  
+perc_talent_w1 ~~ C2T12*perc_talent_w2 + C2T13*perc_talent_w3; perc_talent_w2 ~~ C2T23*perc_talent_w3;  
+perc_wpart_w1  ~~ C3T12*perc_wpart_w2  + C3T13*perc_wpart_w3 ; perc_wpart_w2  ~~ C3T23*perc_wpart_w3 ; 
+perc_netw_w1   ~~ C4T12*perc_netw_w2   + C4T13*perc_netw_w3  ; perc_netw_w2   ~~ C4T23*perc_netw_w3  ;
+pref_effort_w1 ~~ C5T12*pref_effort_w2 + C5T13*pref_effort_w3; pref_effort_w2 ~~ C5T23*pref_effort_w3;  
+pref_talent_w1 ~~ C6T12*pref_talent_w2 + C6T13*pref_talent_w3; pref_talent_w2 ~~ C6T23*pref_talent_w3;  
+pref_wpart_w1  ~~ C7T12*pref_wpart_w2  + C7T13*pref_wpart_w3 ; pref_wpart_w2  ~~ C7T23*pref_wpart_w3 ; 
+pref_netw_w1   ~~ C8T12*pref_netw_w2   + C8T13*pref_netw_w3  ; pref_netw_w2   ~~ C8T23*pref_netw_w3  ;
+
+#===================================================================================================
+# Factor variance fixed to 1 for identification in first group; estimated in others *****
+perc_merit1  ~~ 1*perc_merit1
+perc_merit2  ~~ perc_merit2
+perc_merit3  ~~ perc_merit3
+perc_nmerit1 ~~ 1*perc_nmerit1
+perc_nmerit2 ~~ perc_nmerit2
+perc_nmerit3 ~~ perc_nmerit3
+pref_merit1  ~~ 1*pref_merit1
+pref_merit2  ~~ pref_merit2
+pref_merit3  ~~ pref_merit3
+pref_nmerit1 ~~ 1*pref_nmerit1 
+pref_nmerit2 ~~ pref_nmerit2
+pref_nmerit3 ~~ pref_nmerit3
+
+#===================================================================================================
+#Factor mean fixed to zero for identification in each group
+perc_merit1  ~ 0
+perc_merit2  ~ 0
+perc_merit3  ~ 0
+perc_nmerit1 ~ 0
+perc_nmerit2 ~ 0
+perc_nmerit3 ~ 0
+pref_merit1  ~ 0
+pref_merit2  ~ 0
+pref_merit3  ~ 0
+pref_nmerit1 ~ 0
+pref_nmerit2 ~ 0
+pref_nmerit3 ~ 0
+
+#===================================================================================================
+#Factor covariances all freely estimated
+perc_merit1  ~~ perc_merit2 + perc_merit3
+perc_merit2  ~~ perc_merit3
+perc_nmerit1 ~~ perc_nmerit2 + perc_nmerit3
+perc_nmerit2 ~~ perc_nmerit3
+pref_merit1  ~~ pref_merit2 + pref_merit3
+pref_merit2  ~~ pref_merit3
+pref_nmerit1 ~~ pref_nmerit2 + pref_nmerit3
+pref_nmerit2 ~~ pref_nmerit3
+"
+```
+
+```{r}
+#metric invariance (v03) for complete sample (mode 1, 2 & 3) +  free parameter
+metrA3 = lavaan(model = metricSyntax3, data = wide02, estimator = "MLR")
+
+#metrA  = all factor loading restricted
+#metrA2 = one free estimated factor loading
+#metrA3 = two free estimated factor loading
+
+kable(rbind(fit.conf   =fitmeasures(conf)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            fit.metrA  =fitmeasures(metrA)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            fit.metrA2 =fitmeasures(metrA2)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            fit.metrA3 =fitmeasures(metrA3)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),digits = 3,format = "markdown")
+
+kable(lavTestLRT(metrA3,conf,scaled.shifted = TRUE,method = "satorra.bentler.2001"),format = "markdown",digits = 3)
+```
+
+**NOTE**: the model with two free parameters (metric v03) at factor loadings works fine. Chi2 diff test is not significant
+
+### 3. Scalar Invariance Models (ALL loadings held equal across time - identified model using Time1 Factor Mean = 1)
+
+```{r, scalar1}
+scalarSyntax1 = "
+#===================================================================================================
+#Factor loadings constrained across groups except non-invariant ones
+perc_merit1  =~ L1*perc_effort_w1 + L2*perc_talent_w1 
+perc_merit2  =~ L1*perc_effort_w2 + L2*perc_talent_w2
+perc_merit3  =~ L1*perc_effort_w3 + L2*perc_talent_w3
+perc_nmerit1 =~ L3*perc_wpart_w1  + L4T1*perc_netw_w1   #FREE PARAMETER    
+perc_nmerit2 =~ L3*perc_wpart_w2  + L4*perc_netw_w2  
+perc_nmerit3 =~ L3*perc_wpart_w3  + L4*perc_netw_w3    
+pref_merit1  =~ L5*pref_effort_w1 + L6T1*pref_talent_w1 #FREE PARAMETER
+pref_merit2  =~ L5*pref_effort_w2 + L6*pref_talent_w2
+pref_merit3  =~ L5*pref_effort_w3 + L6*pref_talent_w3
+pref_nmerit1 =~ L7*pref_wpart_w1  + L8*pref_netw_w1    
+pref_nmerit2 =~ L7*pref_wpart_w2  + L8*pref_netw_w2  
+pref_nmerit3 =~ L7*pref_wpart_w3  + L8*pref_netw_w3
+
+#===================================================================================================
+#Item intercepts all constrained across groups except non-invariant ones ****
+perc_effort_w1 ~ I1*1; perc_talent_w1 ~ I2*1; perc_wpart_w1 ~ I3*1; perc_netw_w1 ~ I4T1*1; pref_effort_w1 ~ I5*1; pref_talent_w1 ~ I6T1*1; pref_wpart_w1 ~ I7*1; pref_netw_w1 ~ I8*1; # TWO FREE INTERCEPTS
+perc_effort_w2 ~ I1*1; perc_talent_w2 ~ I2*1; perc_wpart_w2 ~ I3*1; perc_netw_w2 ~ I4*1; pref_effort_w2   ~ I5*1; pref_talent_w2 ~ I6*1; pref_wpart_w2   ~ I7*1; pref_netw_w2 ~ I8*1; 
+perc_effort_w3 ~ I1*1; perc_talent_w3 ~ I2*1; perc_wpart_w3 ~ I3*1; perc_netw_w3 ~ I4*1; pref_effort_w3   ~ I5*1; pref_talent_w3 ~ I6*1; pref_wpart_w3   ~ I7*1; pref_netw_w3 ~ I8*1;
+
+#===================================================================================================
+#Residual variances all freely estimated in each group with labels 
+perc_effort_w1 ~~ E1T1*perc_effort_w1; perc_talent_w1 ~~ E2T1*perc_talent_w1; perc_wpart_w1 ~~ E3T1*perc_wpart_w1; perc_netw_w1 ~~ E4T1*perc_netw_w1; pref_effort_w1 ~~ E5T1*pref_effort_w1; pref_talent_w1 ~~ E6T1*pref_talent_w1; pref_wpart_w1 ~~ E7T1*pref_wpart_w1; pref_netw_w1 ~~ E8T1*pref_netw_w1;
+perc_effort_w2 ~~ E1T2*perc_effort_w2; perc_talent_w2 ~~ E2T2*perc_talent_w2; perc_wpart_w2 ~~ E3T2*perc_wpart_w2; perc_netw_w2 ~~ E4T2*perc_netw_w2; pref_effort_w2 ~~ E5T2*pref_effort_w2; pref_talent_w2 ~~ E6T2*pref_talent_w2; pref_wpart_w2 ~~ E7T2*pref_wpart_w2; pref_netw_w2 ~~ E8T2*pref_netw_w2;
+perc_effort_w3 ~~ E1T3*perc_effort_w3; perc_talent_w3 ~~ E2T3*perc_talent_w3; perc_wpart_w3 ~~ E3T3*perc_wpart_w3; perc_netw_w3 ~~ E4T3*perc_netw_w3; pref_effort_w3 ~~ E5T3*pref_effort_w3; pref_talent_w3 ~~ E6T3*pref_talent_w3; pref_wpart_w3 ~~ E7T3*pref_wpart_w3; pref_netw_w3 ~~ E8T2*pref_netw_w3;
+#===================================================================================================
+#Residual covariances freely estimated in each group with labels (C1T12= covariance 1 between time 1&2)
+
+perc_effort_w1 ~~ C1T12*perc_effort_w2 + C1T13*perc_effort_w3; perc_effort_w2 ~~ C1T23*perc_effort_w3;  
+perc_talent_w1 ~~ C2T12*perc_talent_w2 + C2T13*perc_talent_w3; perc_talent_w2 ~~ C2T23*perc_talent_w3;  
+perc_wpart_w1  ~~ C3T12*perc_wpart_w2  + C3T13*perc_wpart_w3 ; perc_wpart_w2  ~~ C3T23*perc_wpart_w3 ; 
+perc_netw_w1   ~~ C4T12*perc_netw_w2   + C4T13*perc_netw_w3  ; perc_netw_w2   ~~ C4T23*perc_netw_w3  ;
+pref_effort_w1 ~~ C5T12*pref_effort_w2 + C5T13*pref_effort_w3; pref_effort_w2 ~~ C5T23*pref_effort_w3;  
+pref_talent_w1 ~~ C6T12*pref_talent_w2 + C6T13*pref_talent_w3; pref_talent_w2 ~~ C6T23*pref_talent_w3;  
+pref_wpart_w1  ~~ C7T12*pref_wpart_w2  + C7T13*pref_wpart_w3 ; pref_wpart_w2  ~~ C7T23*pref_wpart_w3 ; 
+pref_netw_w1   ~~ C8T12*pref_netw_w2   + C8T13*pref_netw_w3  ; pref_netw_w2   ~~ C8T23*pref_netw_w3  ;
+
+#===================================================================================================
+#Factor variance fixed to 1 for identification in first group; estimated in others
+perc_merit1  ~~ 1*perc_merit1
+perc_merit2  ~~ perc_merit2
+perc_merit3  ~~ perc_merit3
+
+perc_nmerit1 ~~ 1*perc_nmerit1
+perc_nmerit2 ~~ perc_nmerit2
+perc_nmerit3 ~~ perc_nmerit3
+
+pref_merit1  ~~ 1*pref_merit1
+pref_merit2  ~~ pref_merit2
+pref_merit3  ~~ pref_merit3
+
+pref_nmerit1 ~~ 1*pref_nmerit1
+pref_nmerit2 ~~ pref_nmerit2
+pref_nmerit3 ~~ pref_nmerit3
+
+#===================================================================================================
+#Factor mean fixed to zero for identification first group, estimated in others ****
+perc_merit1  ~ 0
 perc_merit2  ~ 1
@@
 
-perc_nmerit1 ~ 0*1
+perc_nmerit1 ~ 0
 perc_nmerit2 ~ 1
@@
 
-pref_merit1 ~ 0*1
+pref_merit1  ~ 0
 pref_merit2  ~ 1
@@
 
-pref_nmerit1 ~ 0*1
+pref_nmerit1 ~ 0
 pref_nmerit2 ~ 1
@@
 
-# LATENT VARIABLE VARIANCES AND COVARIANCES: correlaciones entre variables latentes
-
-'
-
-fit_conf <- cfa(model = configural,data =wide02)
-summary(fit_conf, standardized=TRUE)
-```
+#===================================================================================================
+#Factor covariances all freely estimated
+perc_merit1  ~~ perc_merit2 + perc_merit3
+perc_merit2  ~~ perc_merit3
+
+perc_nmerit1 ~~ perc_nmerit2 + perc_nmerit3
+perc_nmerit2 ~~ perc_nmerit3
+
+pref_merit1  ~~ pref_merit2 + pref_merit3
+pref_merit2  ~~ pref_merit3
+
+pref_nmerit1 ~~ pref_nmerit2 + pref_nmerit3
+pref_nmerit2 ~~ pref_nmerit3
+#===================================================================================================
+"
+```
+
+```{r}
+#scalar invariance for mode 01
+sca_m1A = lavaan(model = scalarSyntax1, data = wide_m1, estimator = "MLR")
+```
+
+```{r}
+#scalar invariance for mode 02
+sca_m2A = lavaan(model = scalarSyntax1, data = wide_m2, estimator = "MLR")
+```
+
+```{r}
+#scalar invariance for mode 03
+sca_m3A = lavaan(model = scalarSyntax1, data = wide_m3, estimator = "MLR")
+```
+
+```{r}
+#scalar invariance for complete sample (mode 1, 2 & 3)
+scaA = lavaan(model = scalarSyntax1, data = wide02, estimator = "MLR")
+```
+
+```{r}
+kable(rbind(fit.conf   =fitmeasures(conf)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            fit.metrA  =fitmeasures(metrA)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            fit.metrA2 =fitmeasures(metrA2)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            fit.metrA3 =fitmeasures(metrA3)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            fit.scaA   =fitmeasures(scaA)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),digits = 3,format = "markdown")
+```
+
+```{r}
+kable(lavTestLRT(scaA,metrA3,conf,scaled.shifted = TRUE,method = "satorra.bentler.2001"),format = "markdown",digits = 3)
+```
+
+**NOTA**: the chi2 test shows that scalar vs metric models are not significant.
+
+### 4. Residual Variance Invariance Model (error variances held equal for all except non-invariant items)
+
+```{r residual1}
+residVarSyntax1 = "
+#===================================================================================================
+#Factor loadings constrained across groups except non-invariant ones
+perc_merit1  =~ L1*perc_effort_w1 + L2*perc_talent_w1 
+perc_merit2  =~ L1*perc_effort_w2 + L2*perc_talent_w2
+perc_merit3  =~ L1*perc_effort_w3 + L2*perc_talent_w3
+perc_nmerit1 =~ L3*perc_wpart_w1  + L4T1*perc_netw_w1   #FREE PARAMETER    
+perc_nmerit2 =~ L3*perc_wpart_w2  + L4*perc_netw_w2  
+perc_nmerit3 =~ L3*perc_wpart_w3  + L4*perc_netw_w3    
+pref_merit1  =~ L5*pref_effort_w1 + L6T1*pref_talent_w1 #FREE PARAMETER
+pref_merit2  =~ L5*pref_effort_w2 + L6*pref_talent_w2
+pref_merit3  =~ L5*pref_effort_w3 + L6*pref_talent_w3
+pref_nmerit1 =~ L7*pref_wpart_w1  + L8*pref_netw_w1    
+pref_nmerit2 =~ L7*pref_wpart_w2  + L8*pref_netw_w2  
+pref_nmerit3 =~ L7*pref_wpart_w3  + L8*pref_netw_w3
+
+#===================================================================================================
+#Item intercepts all constrained across groups except non-invariant ones 
+perc_effort_w1 ~ I1*1; perc_talent_w1 ~ I2*1; perc_wpart_w1 ~ I3*1; perc_netw_w1 ~ I4T1*1; pref_effort_w1 ~ I5*1; pref_talent_w1 ~ I6T1*1; pref_wpart_w1 ~ I7*1; pref_netw_w1 ~ I8*1; # TWO FREE INTERCEPTS
+perc_effort_w2 ~ I1*1; perc_talent_w2 ~ I2*1; perc_wpart_w2 ~ I3*1; perc_netw_w2 ~ I4*1; pref_effort_w2   ~ I5*1; pref_talent_w2 ~ I6*1; pref_wpart_w2   ~ I7*1; pref_netw_w2 ~ I8*1; 
+perc_effort_w3 ~ I1*1; perc_talent_w3 ~ I2*1; perc_wpart_w3 ~ I3*1; perc_netw_w3 ~ I4*1; pref_effort_w3   ~ I5*1; pref_talent_w3 ~ I6*1; pref_wpart_w3   ~ I7*1; pref_netw_w3 ~ I8*1;
+
+#===================================================================================================
+#Redidual variances all constrained in each group except non-invariant ones ****
+perc_effort_w1 ~~ E1*perc_effort_w1; perc_talent_w1 ~~ E2*perc_talent_w1; perc_wpart_w1 ~~ E3*perc_wpart_w1; perc_netw_w1 ~~ E4T1*perc_netw_w1; pref_effort_w1 ~~ E5*pref_effort_w1; pref_talent_w1 ~~ E6T1*pref_talent_w1; pref_wpart_w1 ~~ E7*pref_wpart_w1; pref_netw_w1 ~~ E8*pref_netw_w1;
+perc_effort_w2 ~~ E1*perc_effort_w2; perc_talent_w2 ~~ E2*perc_talent_w2; perc_wpart_w2 ~~ E3*perc_wpart_w2; perc_netw_w2 ~~ E4*perc_netw_w2; pref_effort_w2   ~~ E5*pref_effort_w2; pref_talent_w2 ~~ E6*pref_talent_w2; pref_wpart_w2   ~~ E7*pref_wpart_w2; pref_netw_w2 ~~ E8*pref_netw_w2;
+perc_effort_w3 ~~ E1*perc_effort_w3; perc_talent_w3 ~~ E2*perc_talent_w3; perc_wpart_w3 ~~ E3*perc_wpart_w3; perc_netw_w3 ~~ E4*perc_netw_w3; pref_effort_w3   ~~ E5*pref_effort_w3; pref_talent_w3 ~~ E6*pref_talent_w3; pref_wpart_w3   ~~ E7*pref_wpart_w3; pref_netw_w3 ~~ E8*pref_netw_w3;
+#===================================================================================================
+#Residual covariances freely estimated in each group with labels (C1T12= covariance 1 between time 1&2)
+
+perc_effort_w1 ~~ C1T12*perc_effort_w2 + C1T13*perc_effort_w3; perc_effort_w2 ~~ C1T23*perc_effort_w3;  
+perc_talent_w1 ~~ C2T12*perc_talent_w2 + C2T13*perc_talent_w3; perc_talent_w2 ~~ C2T23*perc_talent_w3;  
+perc_wpart_w1  ~~ C3T12*perc_wpart_w2  + C3T13*perc_wpart_w3 ; perc_wpart_w2  ~~ C3T23*perc_wpart_w3 ; 
+perc_netw_w1   ~~ C4T12*perc_netw_w2   + C4T13*perc_netw_w3  ; perc_netw_w2   ~~ C4T23*perc_netw_w3  ;
+pref_effort_w1 ~~ C5T12*pref_effort_w2 + C5T13*pref_effort_w3; pref_effort_w2 ~~ C5T23*pref_effort_w3;  
+pref_talent_w1 ~~ C6T12*pref_talent_w2 + C6T13*pref_talent_w3; pref_talent_w2 ~~ C6T23*pref_talent_w3;  
+pref_wpart_w1  ~~ C7T12*pref_wpart_w2  + C7T13*pref_wpart_w3 ; pref_wpart_w2  ~~ C7T23*pref_wpart_w3 ; 
+pref_netw_w1   ~~ C8T12*pref_netw_w2   + C8T13*pref_netw_w3  ; pref_netw_w2   ~~ C8T23*pref_netw_w3  ;
+
+#===================================================================================================
+#Factor variance fixed to 1 for identification in first group; estimated in others
+perc_merit1  ~~ 1*perc_merit1
+perc_merit2  ~~ perc_merit2
+perc_merit3  ~~ perc_merit3
+
+perc_nmerit1 ~~ 1*perc_nmerit1
+perc_nmerit2 ~~ perc_nmerit2
+perc_nmerit3 ~~ perc_nmerit3
+
+pref_merit1  ~~ 1*pref_merit1
+pref_merit2  ~~ pref_merit2
+pref_merit3  ~~ pref_merit3
+
+pref_nmerit1 ~~ 1*pref_nmerit1
+pref_nmerit2 ~~ pref_nmerit2
+pref_nmerit3 ~~ pref_nmerit3
+
+#===================================================================================================
+#Factor mean fixed to zero for identification first group, estimated in others ****
+perc_merit1  ~ 0
+perc_merit2  ~ 1
+perc_merit3  ~ 1
+
+perc_nmerit1 ~ 0
+perc_nmerit2 ~ 1
+perc_nmerit3 ~ 1
+
+pref_merit1  ~ 0
+pref_merit2  ~ 1
+pref_merit3  ~ 1
+
+pref_nmerit1 ~ 0
+pref_nmerit2 ~ 1
+pref_nmerit3 ~ 1
+
+#===================================================================================================
+#Factor covariances all freely estimated
+perc_merit1  ~~ perc_merit2 + perc_merit3
+perc_merit2  ~~ perc_merit3
+
+perc_nmerit1 ~~ perc_nmerit2 + perc_nmerit3
+perc_nmerit2 ~~ perc_nmerit3
+
+pref_merit1  ~~ pref_merit2 + pref_merit3
+pref_merit2  ~~ pref_merit3
+
+pref_nmerit1 ~~ pref_nmerit2 + pref_nmerit3
+pref_nmerit2 ~~ pref_nmerit3
+#===================================================================================================
+"
+```
+
+```{r}
+#residual invariance for complete sample (mode 1, 2 & 3)
+resA = lavaan(model =residVarSyntax1 , data = wide02, estimator = "MLR")
+```
+
+```{r}
+kable(rbind(fit.conf   =fitmeasures(conf)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            fit.metrA  =fitmeasures(metrA)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            fit.metrA2 =fitmeasures(metrA2)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            fit.metrA3 =fitmeasures(metrA3)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            fit.scaA   =fitmeasures(scaA)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            fit.resA   =fitmeasures(resA)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),digits = 3,format = "markdown")
+```
+
+```{r}
+kable(lavTestLRT(resA,scaA,metrA3,conf,scaled.shifted = TRUE,method = "satorra.bentler.2001"),format = "markdown",digits = 3)
+```
+
+**Mod Indices**:
+
+```{r modind residual1}
+#get modification indices for all parameters:
+residVarMI1 = modificationindices(resA, free.remove = FALSE, maximum.number = 10000)
+#restrict output to only residual variance parameters:
+residVarMI1a = residVarMI1[which(residVarMI1$op == "~~" & residVarMI1$lhs == residVarMI1$rhs),]
+residVarMI1a = residVarMI1a[order(-residVarMI1a$mi),] # change: pref_wpart_w2 ~~  pref_wpart_w2
+```
+
+```{r residual2}
+residVarSyntax2 = "
+#===================================================================================================
+#Factor loadings constrained across groups except non-invariant ones
+perc_merit1  =~ L1*perc_effort_w1 + L2*perc_talent_w1 
+perc_merit2  =~ L1*perc_effort_w2 + L2*perc_talent_w2
+perc_merit3  =~ L1*perc_effort_w3 + L2*perc_talent_w3
+perc_nmerit1 =~ L3*perc_wpart_w1  + L4T1*perc_netw_w1   #FREE PARAMETER    
+perc_nmerit2 =~ L3*perc_wpart_w2  + L4*perc_netw_w2  
+perc_nmerit3 =~ L3*perc_wpart_w3  + L4*perc_netw_w3    
+pref_merit1  =~ L5*pref_effort_w1 + L6T1*pref_talent_w1 #FREE PARAMETER
+pref_merit2  =~ L5*pref_effort_w2 + L6*pref_talent_w2
+pref_merit3  =~ L5*pref_effort_w3 + L6*pref_talent_w3
+pref_nmerit1 =~ L7*pref_wpart_w1  + L8*pref_netw_w1    
+pref_nmerit2 =~ L7*pref_wpart_w2  + L8*pref_netw_w2  
+pref_nmerit3 =~ L7*pref_wpart_w3  + L8*pref_netw_w3
+
+#===================================================================================================
+#Item intercepts all constrained across groups except non-invariant ones 
+perc_effort_w1 ~ I1*1; perc_talent_w1 ~ I2*1; perc_wpart_w1 ~ I3*1; perc_netw_w1 ~ I4T1*1; pref_effort_w1 ~ I5*1; pref_talent_w1 ~ I6T1*1; pref_wpart_w1 ~ I7*1; pref_netw_w1 ~ I8*1; # TWO FREE INTERCEPTS
+perc_effort_w2 ~ I1*1; perc_talent_w2 ~ I2*1; perc_wpart_w2 ~ I3*1; perc_netw_w2 ~ I4*1; pref_effort_w2   ~ I5*1; pref_talent_w2 ~ I6*1; pref_wpart_w2   ~ I7*1; pref_netw_w2 ~ I8*1; 
+perc_effort_w3 ~ I1*1; perc_talent_w3 ~ I2*1; perc_wpart_w3 ~ I3*1; perc_netw_w3 ~ I4*1; pref_effort_w3   ~ I5*1; pref_talent_w3 ~ I6*1; pref_wpart_w3   ~ I7*1; pref_netw_w3 ~ I8*1;
+
+#===================================================================================================
+#Redidual variances all constrained in each group except non-invariant ones ****
+perc_effort_w1 ~~ E1*perc_effort_w1; perc_talent_w1 ~~ E2*perc_talent_w1; perc_wpart_w1 ~~ E3*perc_wpart_w1; perc_netw_w1 ~~ E4T1*perc_netw_w1; pref_effort_w1 ~~ E5*pref_effort_w1; pref_talent_w1 ~~ E6T1*pref_talent_w1; pref_wpart_w1 ~~ E7*pref_wpart_w1; pref_netw_w1 ~~ E8*pref_netw_w1;
+perc_effort_w2 ~~ E1*perc_effort_w2; perc_talent_w2 ~~ E2*perc_talent_w2; perc_wpart_w2 ~~ E3*perc_wpart_w2; perc_netw_w2 ~~ E4*perc_netw_w2; pref_effort_w2   ~~ E5*pref_effort_w2; pref_talent_w2 ~~ E6*pref_talent_w2; pref_wpart_w2   ~~ E7T2*pref_wpart_w2; pref_netw_w2 ~~ E8*pref_netw_w2; # F-RES
+perc_effort_w3 ~~ E1*perc_effort_w3; perc_talent_w3 ~~ E2*perc_talent_w3; perc_wpart_w3 ~~ E3*perc_wpart_w3; perc_netw_w3 ~~ E4*perc_netw_w3; pref_effort_w3   ~~ E5*pref_effort_w3; pref_talent_w3 ~~ E6*pref_talent_w3; pref_wpart_w3   ~~ E7*pref_wpart_w3; pref_netw_w3 ~~ E8*pref_netw_w3;
+#===================================================================================================
+#Residual covariances freely estimated in each group with labels (C1T12= covariance 1 between time 1&2)
+
+perc_effort_w1 ~~ C1T12*perc_effort_w2 + C1T13*perc_effort_w3; perc_effort_w2 ~~ C1T23*perc_effort_w3;  
+perc_talent_w1 ~~ C2T12*perc_talent_w2 + C2T13*perc_talent_w3; perc_talent_w2 ~~ C2T23*perc_talent_w3;  
+perc_wpart_w1  ~~ C3T12*perc_wpart_w2  + C3T13*perc_wpart_w3 ; perc_wpart_w2  ~~ C3T23*perc_wpart_w3 ; 
+perc_netw_w1   ~~ C4T12*perc_netw_w2   + C4T13*perc_netw_w3  ; perc_netw_w2   ~~ C4T23*perc_netw_w3  ;
+pref_effort_w1 ~~ C5T12*pref_effort_w2 + C5T13*pref_effort_w3; pref_effort_w2 ~~ C5T23*pref_effort_w3;  
+pref_talent_w1 ~~ C6T12*pref_talent_w2 + C6T13*pref_talent_w3; pref_talent_w2 ~~ C6T23*pref_talent_w3;  
+pref_wpart_w1  ~~ C7T12*pref_wpart_w2  + C7T13*pref_wpart_w3 ; pref_wpart_w2  ~~ C7T23*pref_wpart_w3 ; 
+pref_netw_w1   ~~ C8T12*pref_netw_w2   + C8T13*pref_netw_w3  ; pref_netw_w2   ~~ C8T23*pref_netw_w3  ;
+
+#===================================================================================================
+#Factor variance fixed to 1 for identification in first group; estimated in others
+perc_merit1  ~~ 1*perc_merit1
+perc_merit2  ~~ perc_merit2
+perc_merit3  ~~ perc_merit3
+
+perc_nmerit1 ~~ 1*perc_nmerit1
+perc_nmerit2 ~~ perc_nmerit2
+perc_nmerit3 ~~ perc_nmerit3
+
+pref_merit1  ~~ 1*pref_merit1
+pref_merit2  ~~ pref_merit2
+pref_merit3  ~~ pref_merit3
+
+pref_nmerit1 ~~ 1*pref_nmerit1
+pref_nmerit2 ~~ pref_nmerit2
+pref_nmerit3 ~~ pref_nmerit3
+
+#===================================================================================================
+#Factor mean fixed to zero for identification first group, estimated in others ****
+perc_merit1  ~ 0
+perc_merit2  ~ 1
+perc_merit3  ~ 1
+
+perc_nmerit1 ~ 0
+perc_nmerit2 ~ 1
+perc_nmerit3 ~ 1
+
+pref_merit1  ~ 0
+pref_merit2  ~ 1
+pref_merit3  ~ 1
+
+pref_nmerit1 ~ 0
+pref_nmerit2 ~ 1
+pref_nmerit3 ~ 1
+
+#===================================================================================================
+#Factor covariances all freely estimated
+perc_merit1  ~~ perc_merit2 + perc_merit3
+perc_merit2  ~~ perc_merit3
+
+perc_nmerit1 ~~ perc_nmerit2 + perc_nmerit3
+perc_nmerit2 ~~ perc_nmerit3
+
+pref_merit1  ~~ pref_merit2 + pref_merit3
+pref_merit2  ~~ pref_merit3
+
+pref_nmerit1 ~~ pref_nmerit2 + pref_nmerit3
+pref_nmerit2 ~~ pref_nmerit3
+#===================================================================================================
+"
+```
+
+```{r}
+#residual invariance (v02) for complete sample (mode 1, 2 & 3)
+resA2 = lavaan(model =residVarSyntax2 , data = wide02, estimator = "MLR")
+```
+
+```{r}
+kable(rbind(fit.conf   =fitmeasures(conf)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            fit.metrA  =fitmeasures(metrA)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            fit.metrA2 =fitmeasures(metrA2)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            fit.metrA3 =fitmeasures(metrA3)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            fit.scaA   =fitmeasures(scaA)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            fit.resA   =fitmeasures(resA)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            fit.resA2  =fitmeasures(resA2)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),digits = 3,format = "markdown")
+
+```
+
+```{r}
+kable(lavTestLRT(resA2,scaA,metrA3,conf,scaled.shifted = TRUE,method = "satorra.bentler.2001"),format = "markdown",digits = 3)
+```
+
+**Mod Indices**:
+
+```{r modind residual2}
+#get modification indices for all parameters:
+residVarMI1 = modificationindices(resA2, free.remove = FALSE, maximum.number = 10000)
+#restrict output to only residual variance parameters:
+residVarMI1a = residVarMI1[which(residVarMI1$op == "~~" & residVarMI1$lhs == residVarMI1$rhs),]
+residVarMI1a <- residVarMI1a[order(-residVarMI1a$mi),] # changes: pref_wpart_w3 ~~  pref_wpart_w3
+```
+
+```{r residual3}
+residVarSyntax3 = "
+#===================================================================================================
+#Factor loadings constrained across groups except non-invariant ones
+perc_merit1  =~ L1*perc_effort_w1 + L2*perc_talent_w1 
+perc_merit2  =~ L1*perc_effort_w2 + L2*perc_talent_w2
+perc_merit3  =~ L1*perc_effort_w3 + L2*perc_talent_w3
+perc_nmerit1 =~ L3*perc_wpart_w1  + L4T1*perc_netw_w1   #FREE PARAMETER    
+perc_nmerit2 =~ L3*perc_wpart_w2  + L4*perc_netw_w2  
+perc_nmerit3 =~ L3*perc_wpart_w3  + L4*perc_netw_w3    
+pref_merit1  =~ L5*pref_effort_w1 + L6T1*pref_talent_w1 #FREE PARAMETER
+pref_merit2  =~ L5*pref_effort_w2 + L6*pref_talent_w2
+pref_merit3  =~ L5*pref_effort_w3 + L6*pref_talent_w3
+pref_nmerit1 =~ L7*pref_wpart_w1  + L8*pref_netw_w1    
+pref_nmerit2 =~ L7*pref_wpart_w2  + L8*pref_netw_w2  
+pref_nmerit3 =~ L7*pref_wpart_w3  + L8*pref_netw_w3
+
+#===================================================================================================
+#Item intercepts all constrained across groups except non-invariant ones 
+perc_effort_w1 ~ I1*1; perc_talent_w1 ~ I2*1; perc_wpart_w1 ~ I3*1; perc_netw_w1 ~ I4T1*1; pref_effort_w1 ~ I5*1; pref_talent_w1 ~ I6T1*1; pref_wpart_w1 ~ I7*1; pref_netw_w1 ~ I8*1; # TWO FREE INTERCEPTS
+perc_effort_w2 ~ I1*1; perc_talent_w2 ~ I2*1; perc_wpart_w2 ~ I3*1; perc_netw_w2 ~ I4*1; pref_effort_w2   ~ I5*1; pref_talent_w2 ~ I6*1; pref_wpart_w2   ~ I7*1; pref_netw_w2 ~ I8*1; 
+perc_effort_w3 ~ I1*1; perc_talent_w3 ~ I2*1; perc_wpart_w3 ~ I3*1; perc_netw_w3 ~ I4*1; pref_effort_w3   ~ I5*1; pref_talent_w3 ~ I6*1; pref_wpart_w3   ~ I7*1; pref_netw_w3 ~ I8*1;
+
+#===================================================================================================
+#Redidual variances all constrained in each group except non-invariant ones ****
+perc_effort_w1 ~~ E1*perc_effort_w1; perc_talent_w1 ~~ E2*perc_talent_w1; perc_wpart_w1 ~~ E3*perc_wpart_w1; perc_netw_w1 ~~ E4T1*perc_netw_w1; pref_effort_w1 ~~ E5*pref_effort_w1; pref_talent_w1 ~~ E6T1*pref_talent_w1; pref_wpart_w1 ~~ E7*pref_wpart_w1; pref_netw_w1 ~~ E8*pref_netw_w1;
+perc_effort_w2 ~~ E1*perc_effort_w2; perc_talent_w2 ~~ E2*perc_talent_w2; perc_wpart_w2 ~~ E3*perc_wpart_w2; perc_netw_w2 ~~ E4*perc_netw_w2; pref_effort_w2   ~~ E5*pref_effort_w2; pref_talent_w2 ~~ E6*pref_talent_w2; pref_wpart_w2   ~~ E7T2*pref_wpart_w2; pref_netw_w2 ~~ E8*pref_netw_w2; # F-RES2
+perc_effort_w3 ~~ E1*perc_effort_w3; perc_talent_w3 ~~ E2*perc_talent_w3; perc_wpart_w3 ~~ E3*perc_wpart_w3; perc_netw_w3 ~~ E4*perc_netw_w3; pref_effort_w3   ~~ E5*pref_effort_w3; pref_talent_w3 ~~ E6*pref_talent_w3; pref_wpart_w3   ~~ E7T3*pref_wpart_w3; pref_netw_w3 ~~ E8*pref_netw_w3; # F-RES3
+#===================================================================================================
+#Residual covariances freely estimated in each group with labels (C1T12= covariance 1 between time 1&2)
+
+perc_effort_w1 ~~ C1T12*perc_effort_w2 + C1T13*perc_effort_w3; perc_effort_w2 ~~ C1T23*perc_effort_w3;  
+perc_talent_w1 ~~ C2T12*perc_talent_w2 + C2T13*perc_talent_w3; perc_talent_w2 ~~ C2T23*perc_talent_w3;  
+perc_wpart_w1  ~~ C3T12*perc_wpart_w2  + C3T13*perc_wpart_w3 ; perc_wpart_w2  ~~ C3T23*perc_wpart_w3 ; 
+perc_netw_w1   ~~ C4T12*perc_netw_w2   + C4T13*perc_netw_w3  ; perc_netw_w2   ~~ C4T23*perc_netw_w3  ;
+pref_effort_w1 ~~ C5T12*pref_effort_w2 + C5T13*pref_effort_w3; pref_effort_w2 ~~ C5T23*pref_effort_w3;  
+pref_talent_w1 ~~ C6T12*pref_talent_w2 + C6T13*pref_talent_w3; pref_talent_w2 ~~ C6T23*pref_talent_w3;  
+pref_wpart_w1  ~~ C7T12*pref_wpart_w2  + C7T13*pref_wpart_w3 ; pref_wpart_w2  ~~ C7T23*pref_wpart_w3 ; 
+pref_netw_w1   ~~ C8T12*pref_netw_w2   + C8T13*pref_netw_w3  ; pref_netw_w2   ~~ C8T23*pref_netw_w3  ;
+
+#===================================================================================================
+#Factor variance fixed to 1 for identification in first group; estimated in others
+perc_merit1  ~~ 1*perc_merit1
+perc_merit2  ~~ perc_merit2
+perc_merit3  ~~ perc_merit3
+
+perc_nmerit1 ~~ 1*perc_nmerit1
+perc_nmerit2 ~~ perc_nmerit2
+perc_nmerit3 ~~ perc_nmerit3
+
+pref_merit1  ~~ 1*pref_merit1
+pref_merit2  ~~ pref_merit2
+pref_merit3  ~~ pref_merit3
+
+pref_nmerit1 ~~ 1*pref_nmerit1
+pref_nmerit2 ~~ pref_nmerit2
+pref_nmerit3 ~~ pref_nmerit3
+
+#===================================================================================================
+#Factor mean fixed to zero for identification first group, estimated in others ****
+perc_merit1  ~ 0
+perc_merit2  ~ 1
+perc_merit3  ~ 1
+
+perc_nmerit1 ~ 0
+perc_nmerit2 ~ 1
+perc_nmerit3 ~ 1
+
+pref_merit1  ~ 0
+pref_merit2  ~ 1
+pref_merit3  ~ 1
+
+pref_nmerit1 ~ 0
+pref_nmerit2 ~ 1
+pref_nmerit3 ~ 1
+
+#===================================================================================================
+#Factor covariances all freely estimated
+perc_merit1  ~~ perc_merit2 + perc_merit3
+perc_merit2  ~~ perc_merit3
+
+perc_nmerit1 ~~ perc_nmerit2 + perc_nmerit3
+perc_nmerit2 ~~ perc_nmerit3
+
+pref_merit1  ~~ pref_merit2 + pref_merit3
+pref_merit2  ~~ pref_merit3
+
+pref_nmerit1 ~~ pref_nmerit2 + pref_nmerit3
+pref_nmerit2 ~~ pref_nmerit3
+#===================================================================================================
+"
+```
+
+```{r}
+#residual invariance (v03) for complete sample (mode 1, 2 & 3)
+resA3 = lavaan(model =residVarSyntax3 , data = wide02, estimator = "MLR")
+```
+
+```{r}
+kable(rbind(fit.conf   =fitmeasures(conf)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            fit.metrA  =fitmeasures(metrA)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            fit.metrA2 =fitmeasures(metrA2)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            fit.metrA3 =fitmeasures(metrA3)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            fit.scaA   =fitmeasures(scaA)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            fit.resA   =fitmeasures(resA)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            fit.resA2  =fitmeasures(resA2)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            fit.resA3  =fitmeasures(resA3)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),digits = 3,format = "markdown")
+```
+
+```{r}
+kable(lavTestLRT(resA3,scaA,metrA3,conf,scaled.shifted = TRUE,method = "satorra.bentler.2001"),format = "markdown",digits = 3)
+```
+
+**Mod Indices**:
+
+```{r modind residual3}
+#get modification indices for all parameters:
+residVarMI1 = modificationindices(resA3, free.remove = FALSE, maximum.number = 10000)
+#restrict output to only residual variance parameters:
+residVarMI1a = residVarMI1[which(residVarMI1$op == "~~" & residVarMI1$lhs == residVarMI1$rhs),]
+residVarMI1a <- residVarMI1a[order(-residVarMI1a$mi),] # change: pref_wpart_w3 ~~  pref_wpart_w3 (?)
+```
+
+**NOTE**: The residual is too big. Mod indices suggest to freely estimate the pref_wpart_w3 covariance, but this one is already free.
 
 * Invarianza Longitudinal en lavaan
-https://quantdev.ssri.psu.edu/tutorials/intro-basics-longitudinal-measurement-invariance
+  * https://quantdev.ssri.psu.edu/tutorials/intro-basics-longitudinal-measurement-invariance 
+  * https://jonathantemplin.com/wp-content/uploads/2017/08/EPSY906_Example07b_CFA_Longitudinal_Invariance.nb_.html
 

### 2020-04-05 production/prod_prep-data-A.rmd
data-prerp
add labels

dfSummary

---
title: "Data preparation"
date: "`r format(Sys.time(), '%d %B, %Y')`"
css: "../input/bib/style.css"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_float: false
    toc_depth: 2
    code_folding: "hide"
editor_options: 
  chunk_output_type: console
---

```{r eval=FALSE, include=FALSE}
rmarkdown::render(input = "production/prep-data-A.Rmd",output_format = "html_document",output_dir = ".")
```

```{r include=FALSE}
knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = FALSE,results = "hold")
knitr::opts_knit$set(root.dir = "../")
options(scipen=999)
rm(list=ls())
options(knitr.kable.NA = '')
```

load packages
```{r}
library(sjPlot)
library(dplyr)
library(tidyr)
library(lavaan)
library(stargazer)
library(corrplot)
library(psych)
library(knitr)
library(kableExtra)
library(summarytools)
library(sjlabelled)
library(Hmisc)
```

# Data preparation

```{r}
w01 <- sjlabelled::read_spss("input/data/original/Estudio_3_ola1.sav",verbose = F)
```

##  homogenize names of the items on the scale.

The items on the scale had different orders according to the order in which they were exposed to the respondents. There are three order. The first one the orders keeps together the items of the same factor. The second order then asks for perceptions and preferences regarding the same topic (i.e Distribution of resources according to efforts). In the third order, the items are presented randomly to each respondent.

```{r}
dat04 <- w01 %>% filter(Intro==1) %>%
  select(starts_with("meritv01"),
         starts_with("meritv02"),
         starts_with("meritv03_p"),
         starts_with("FL_21_DO")) 
dat04$grupo <- NA
dat04$grupo[dat04$FL_21_DO_merit_perc_pref_julio19v01==1] <- 1
dat04$grupo[dat04$FL_21_DO_merit_perc_pref_julio19v02==1] <- 2
dat04$grupo[dat04$FL_21_DO_merit_perc_pref_julio19v03==1] <- 3
dat04$perc_effort <- rowSums(dat04[,c(matches(match = "perc_effort",vars = names(dat04)))],na.rm = TRUE)
dat04$perc_talent <- rowSums(dat04[,c(matches(match = "perc_talent",vars = names(dat04)))],na.rm = TRUE)
dat04$perc_wpart  <- rowSums(dat04[,c(matches(match = "perc_wpart" ,vars = names(dat04)))],na.rm = TRUE)
dat04$perc_netw   <- rowSums(dat04[,c(matches(match = "perc_netw"  ,vars = names(dat04)))],na.rm = TRUE)

dat04$pref_effort <- rowSums(dat04[,c(matches(match = "pref_effort",vars = names(dat04)))],na.rm = TRUE)
dat04$pref_talent <- rowSums(dat04[,c(matches(match = "pref_talent",vars = names(dat04)))],na.rm = TRUE)
dat04$pref_wpart  <- rowSums(dat04[,c(matches(match = "pref_wpart" ,vars = names(dat04)))],na.rm = TRUE)
dat04$pref_netw   <- rowSums(dat04[,c(matches(match = "pref_netw"  ,vars = names(dat04)))],na.rm = TRUE)
dat04[dat04==0] <- NA
```

## Translation of variable labels and value labels.  
```{r}
label(dat04$grupo) <- "This variable classifies the interviewees according to the order in which the indicators are shown within the scale. There are three orders  of items on the scale. In the first order (group 1), questions of the same factor are then presented to them, in the second order (group 2) they were separated and in the third order (group 3) at random."

dat04$perc_effort <- set_labels(dat04$perc_effort,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
label(dat04$perc_effort) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - Who
the more they try they manage to get bigger
rewards that those who strive
less."

dat04$perc_talent <- set_labels(dat04$perc_talent,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
label(dat04$perc_talent) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - Who
possess more talent they manage to obtain
greater rewards than those who possess
less talent."

dat04$perc_wpart <- set_labels(dat04$perc_wpart,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
label(dat04$perc_wpart) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - Who
they have rich parents manage to get out
ahead."

dat04$perc_netw <- set_labels(dat04$perc_netw,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
label(dat04$perc_netw) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - Who
they have good contacts they manage to get out
ahead."

dat04$pref_effort <- set_labels(dat04$pref_effort,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
label(dat04$pref_effort) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - Who
the more they try they should get
greater rewards than those who
they try less."

dat04$pref_talent <- set_labels(dat04$pref_talent,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
label(dat04$pref_talent) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - Who
possess more talent they should get
greater rewards than those who possess
less talent."

dat04$pref_wpart <- set_labels(dat04$pref_wpart,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
label(dat04$pref_wpart) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - Is
well that those who have good
contacts get ahead."

dat04$pref_netw <- set_labels(dat04$pref_netw,  labels = c("strongly disagree", "disagreement", "Neither agree nor disagree","Agree", "totally agree"))
label(dat04$pref_netw) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - Is
well that those who have good
contacts get ahead."
```

```{r}
dat04 <-  dat04 %>% select(grupo, perc_effort, perc_talent, perc_wpart, perc_netw, pref_effort, pref_talent, pref_wpart , pref_netw) %>% as.data.frame()
```

```{r} 
print(dfSummary(dat04, valid.col = FALSE, graph.magnif = 0.75), 
      max.tbl.height = 300, method = "render")
```

### Preparation of group 1: order according to the factors.

```{r}
dat01 <- dat04 %>% filter(grupo==1) %>%  na.omit()
```

```{r} 
print(dfSummary(dat01, valid.col = FALSE, graph.magnif = 0.75), 
      max.tbl.height = 300, method = "render")
```

### Preparation of group 2: order according to topics (i.e Effort).

```{r}
dat02 <- dat04 %>% filter(grupo==2) %>% select(perc_effort, pref_effort, perc_talent,pref_talent, perc_wpart,pref_wpart,  perc_netw, pref_netw) %>% na.omit()
```

```{r} 
print(dfSummary(dat02, valid.col = FALSE, graph.magnif = 0.75), 
      max.tbl.height = 300, method = "render")
```

### Preparation of group 3: randomized order.

```{r}
dat03 <- dat04 %>% filter(grupo==3) %>%  na.omit()
```

```{r} 
print(dfSummary(dat03, valid.col = FALSE, graph.magnif = 0.75), 
      max.tbl.height = 300, method = "render")
```

```{r}
save(dat01,file = "input/data/proc/dat01.RData")
save(dat02,file = "input/data/proc/dat02.RData")
save(dat03,file = "input/data/proc/dat03.RData")
save(dat04,file = "input/data/proc/dat04.RData")
```

# Longitudinal data

```{r}
w01 <- sjlabelled::read_spss("input/data/original/Estudio_3_ola1.sav",verbose = F)
w02 <- sjlabelled::read_spss("input/data/original/Estudio_3_ola2.sav",verbose = F)
w03 <- sjlabelled::read_spss("input/data/original/Estudio_3_ola3.sav",verbose = F)
```

## Preparation wave 1

Add tags wave 1
```{r }
label(w01$Intro) <- "If you wish to participate, select the alternative I ACCEPT take part in this investigation. If you do not want to participate in this study, press the alternative I NOT ACCEPT take part in this investigation."

w01$Intro <- set_labels(w01$Intro,  labels = c("I ACCEPT take part in this investigation","I NOT ACCEPT take part in this investigation."))

```

Frequency of respondents who Accept and Finish wave 1.
```{r fig.width= 10}
w01_Intro <- w01 %>% select(Intro,Finished) %>% as.data.frame() 
sjPlot::view_df(w01_Intro, show.type = T,show.frq = T, show.string.values = T, verbose = F, show.id = F)
```

Delete those who do not accept and Delete those who do not accept
```{r}
w01b <- w01 %>% filter(Intro==1, Finished==1) %>% as.data.frame() 
w01b$ID<- stringr::str_split_fixed(w01b$ticket,"_", 4)[,1] # En la variable ticket el primer código es el ID, todo lo demás se borra
```

Frequency of respondents who Accept and Finish wave 1.
```{r fig.width= 20 }
w01b$Intro <- set_labels(w01b$Intro,  labels = c("I ACCEPT take part in this investigation", "I NOT ACCEPT take part in this investigation."))
w01_Introb <- w01b %>% select(Intro,Finished) %>% as.data.frame() 
sjPlot::view_df(w01_Introb, show.type = T,show.frq = T, show.string.values = T, verbose = F, show.id = F)
```

```{r}
w01c <- w01b 
w01c$grupo <- NA
w01c$grupo[w01c$FL_21_DO_merit_perc_pref_julio19v01==1] <- 1
w01c$grupo[w01c$FL_21_DO_merit_perc_pref_julio19v02==1] <- 2
w01c$grupo[w01c$FL_21_DO_merit_perc_pref_julio19v03==1] <- 3
w01c$perc_effort_w1 <- rowSums(w01c[,c(matches(match = "perc_effort",vars = names(w01c)))],na.rm = TRUE)
w01c$perc_talent_w1 <- rowSums(w01c[,c(matches(match = "perc_talent",vars = names(w01c)))],na.rm = TRUE)
w01c$perc_wpart_w1  <- rowSums(w01c[,c(matches(match = "perc_wpart" ,vars = names(w01c)))],na.rm = TRUE)
w01c$perc_netw_w1   <- rowSums(w01c[,c(matches(match = "perc_netw"  ,vars = names(w01c)))],na.rm = TRUE)

w01c$pref_effort_w1 <- rowSums(w01c[,c(matches(match = "pref_effort",vars = names(w01c)))],na.rm = TRUE)
w01c$pref_talent_w1 <- rowSums(w01c[,c(matches(match = "pref_talent",vars = names(w01c)))],na.rm = TRUE)
w01c$pref_wpart_w1  <- rowSums(w01c[,c(matches(match = "pref_wpart" ,vars = names(w01c)))],na.rm = TRUE)
w01c$pref_netw_w1   <- rowSums(w01c[,c(matches(match = "pref_netw"  ,vars = names(w01c)))],na.rm = TRUE)
w01c[w01c==0] <- NA
w01c <- w01c %>% select(grupo,
                        perc_effort_w1,perc_talent_w1,perc_wpart_w1,perc_netw_w1,
                        pref_effort_w1,pref_talent_w1,pref_wpart_w1,pref_netw_w1,
                        sexo_w1=sexo,edad_w1=edad,edcep_w1=edcep, ID)
#-----------------------------------------------------------------------#
print(dfSummary(w01c, valid.col = FALSE, graph.magnif = 0.75), 
      max.tbl.height = 300, method = "render")
```

## Preparation wave 2

Add tags wave 2
```{r }
label(w02$Intro) <- "If you wish to participate, select the alternative I ACCEPT take part in this investigation. If you do not want to participate in this study, press the alternative I NOT ACCEPT take part in this investigation."

w02$Intro <- set_labels(w02$Intro,  labels = c("I ACCEPT take part in this investigation","I NOT ACCEPT take part in this investigation."))

```

Frequency of respondents who Accept and Finish wave 1.
```{r}
w02_Intro <- w02 %>% select(Intro,Finished) %>% as.data.frame() 
sjPlot::view_df(w02_Intro, show.type = T,show.frq = T, show.string.values = T, verbose = F, show.id = F)
```

Delete those who do not accept and Delete those who do not accept
```{r}
w02b <- w02 %>% filter(Intro==1, Finished==1) %>% as.data.frame() 
w02b$ID <- stringr::str_split_fixed(w02b$ticket, "_", 4)[,1]

```

Frequency of respondents who Accept and Finish wave 1.
```{r }
w02b$Intro <- set_labels(w02b$Intro,  labels = c("I ACCEPT take part in this investigation", "I NOT ACCEPT take part in this investigation."))
w02_Introb <- w02b %>% select(Intro,Finished) %>% as.data.frame() 
sjPlot::view_df(w02_Introb, show.type = T,show.frq = T, show.string.values = T, verbose = F, show.id = F)
```

## 
```{r}
w02c <- w02b

w02c$treat <- NA
w02c$treat[w02c$FL_6_DO_Control==1] <- 1 
w02c$treat[w02c$FL_6_DO_Tratamiento==1] <- 2 
w02c$treat[w02c$FL_6_DO_Tratamiento_desigualdad==1] <- 3

w02c$treat <- factor(w02c$treat,levels = c(1,2,3),labels = c("control","pobreza","desiguald"))

w02c <- w02c %>% rename(perc_effort_w2=meritv03_perc_effort,
                        perc_talent_w2=meritv03_perc_talent,
                        perc_wpart_w2 =meritv03_perc_wpart,
                        perc_netw_w2  =meritv03_perc_netw,
                        pref_effort_w2=meritv03_pref_effort,
                        pref_talent_w2=meritv03_pref_talent,
                        pref_wpart_w2 =meritv03_pref_wpart,
                        pref_netw_w2  =meritv03_pref_netw)

w02c <- w02c %>% select(perc_effort_w2,perc_talent_w2,perc_wpart_w2,perc_netw_w2,
                        pref_effort_w2,pref_talent_w2,pref_wpart_w2,pref_netw_w2,
                        sexo_w2=sexo,edad_w2=edad,edcep_w2=edcep, treat, ID)
#-----------------------------------------------------------------------#
print(dfSummary(w02c, valid.col = FALSE, graph.magnif = 0.75), 
      max.tbl.height = 300, method = "render")
```

## Preparation wave 3

```{r}
table(w03$Finished==1)
w03b <- filter(w03,Finished==1) # Elimina los que no aceptan
dim(w03b)
w03b$ID <- stringr::str_split_fixed(w03b$ticket, "_", 4)[,1] 
```

```{r}
w03c <- w03b

w03c <- w03c %>% rename(perc_effort_w3=meritv03_perc_effort,
                        perc_talent_w3=meritv03_perc_talent,
                        perc_wpart_w3 =meritv03_perc_wpart,
                        perc_netw_w3  =meritv03_perc_netw,
                        pref_effort_w3=meritv03_pref_effort,
                        pref_talent_w3=meritv03_pref_talent,
                        pref_wpart_w3 =meritv03_pref_wpart,
                        pref_netw_w3  =meritv03_pref_netw)
w03c <- w03c %>% select(perc_effort_w3,perc_talent_w3,perc_wpart_w3,perc_netw_w3,
                        pref_effort_w3,pref_talent_w3,pref_wpart_w3,pref_netw_w3,
                        sexo_w3=sexo,edad_w3=edad,edcep_w3=edcep, ID)

#-----------------------------------------------------------------------#
print(dfSummary(w03c, valid.col = FALSE, graph.magnif = 0.75), 
      max.tbl.height = 300, method = "render")
```

## Merge wide data

```{r}
wide01 <- inner_join(x = w01c,y = w02c, by = "ID") # Solamente nos quedamos con casos en ambas olas 
dim(wide01) #en w01 y w02
```

```{r}
wide02 <- inner_join(x = wide01,y = w03c, by = "ID") # Solamente nos quedamos con casos en ambas olas 
dim(wide02) #en w01, w02 y w03
```

```{r}
print(dfSummary(w02, valid.col = FALSE, graph.magnif = 0.75), 
      max.tbl.height = 300, method = "render")
```

```{r}
save(wide01,file = "input/data/proc/datawide.RData")
save(wide02,file = "input/data/proc/datawide_123.RData")
```

# Wide to long data

```{r}
library(datasets)
library(data.table)

wide <- setDT(wide02)

long01<- melt(wide,variable.name = "wave",
              measure = patterns("^perc_effort_","^perc_talent_","^perc_wpart_","^perc_netw_",
                                       "^pref_effort_","^pref_talent_","^pref_wpart_","^pref_netw_",
                                       "^sexo_","^edad_","^edcep_"),
              value.name = c("perc_effort","perc_talent","perc_wpart","perc_netw",
                             "pref_effort","pref_talent","pref_wpart","pref_netw",
                             "sexo","edad","edcep"))

```

```{r}
save(long01,file = "input/data/proc/datalong.RData")
```

### 2020-04-06 production/prod_analysis-long_invariance.rmd
update-0604-2020
* Actualización de análisis, con foco en ola 1.
* El documento ahora está completamente en inglés.

* Análisis longitudinal se movió a un archivo independiente (prod_analysis-long-invariance.rmd

--- 2020-04-01 production/prod_analysis-cfa.rmd
@@
 ---
-title: "Analysis"
+title: "Longitudinal analysis "
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
     toc_float: true
-    toc_depth: 4
+    toc_depth: 2
     code_folding: "hide"
@@
 ---
-
-# Análisis PPM-S
-
-# Wave 01
-
-```{r eval=FALSE, include=FALSE}
-rmarkdown::render(input = "production/prod_analysis-cfa.Rmd",output_format = "html_document")
-```
-
-```{r include=FALSE}
-knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = TRUE)
-knitr::opts_knit$set(root.dir = "../")
-options(scipen=999)
-rm(list=ls())
-options(knitr.kable.NA = '')
-```
-
-```{r}
-library(sjPlot)
-library(dplyr)
-library(lavaan)
-library(stargazer)
-library(corrplot)
-library(psych)
-library(knitr)
-library(kableExtra)
-```
-
-```{r}
-data01 <- sjlabelled::read_spss(path = "input/data/original/Estudio_3_ola1.sav",verbose = FALSE)
-
-dat01 <- data01 %>% filter(Intro==1,Finished==1) %>% select(starts_with("meritv01")) %>% na.omit()
-dat02 <- data01 %>% filter(Intro==1,Finished==1) %>% select(starts_with("meritv02")) %>% na.omit()
-dat03 <- data01 %>% filter(Intro==1,Finished==1) %>% select(starts_with("meritv03_p")) %>% na.omit()
-
-# - Nos quedamos solamente con los casos que ACEPTARON participar y terminaron el cuestionario.
-# - Para dejar una base que contenga solamente los casos que tienen información completa en los ítems usamos na.omit()
-
-dat04 <- data01 %>% filter(Intro==1) %>%
-  select(starts_with("meritv01"),
-         starts_with("meritv02"),
-         starts_with("meritv03_p"),
-         starts_with("FL_21_DO")) 
-dat04$grupo <- NA
-dat04$grupo[dat04$FL_21_DO_merit_perc_pref_julio19v01==1] <- 1
-dat04$grupo[dat04$FL_21_DO_merit_perc_pref_julio19v02==1] <- 2
-dat04$grupo[dat04$FL_21_DO_merit_perc_pref_julio19v03==1] <- 3
-dat04$perc_effort <- rowSums(dat04[,c(matches(match = "perc_effort",vars = names(dat04)))],na.rm = TRUE)
-dat04$perc_talent <- rowSums(dat04[,c(matches(match = "perc_talent",vars = names(dat04)))],na.rm = TRUE)
-dat04$perc_wpart  <- rowSums(dat04[,c(matches(match = "perc_wpart" ,vars = names(dat04)))],na.rm = TRUE)
-dat04$perc_netw   <- rowSums(dat04[,c(matches(match = "perc_netw"  ,vars = names(dat04)))],na.rm = TRUE)
-
-dat04$pref_effort <- rowSums(dat04[,c(matches(match = "pref_effort",vars = names(dat04)))],na.rm = TRUE)
-dat04$pref_talent <- rowSums(dat04[,c(matches(match = "pref_talent",vars = names(dat04)))],na.rm = TRUE)
-dat04$pref_wpart  <- rowSums(dat04[,c(matches(match = "pref_wpart" ,vars = names(dat04)))],na.rm = TRUE)
-dat04$pref_netw   <- rowSums(dat04[,c(matches(match = "pref_netw"  ,vars = names(dat04)))],na.rm = TRUE)
-dat04[dat04==0] <- NA
-```
-
-## Version 01: Fixed order by perception / preferences (N=`r dim(dat01)[1]`)
-
-1. Percepcion esfuerzo
-2. Percepcion talento
-3. Percepcion familia rica
-4. Percepcion redes
-5. Preferencia esfuerzo
-6. Preferencia talento
-7. Preferencia familia rica
-8. Preferencia redes
-
-```{r echo=FALSE,results='asis'}
-stargazer(dat01,type = "html")
-```
-
-```{r correlacion v01}
-cor01<-polychoric(dat01)
-corrplot(cor01$rho,method = "color" ,type = "upper", order = "hclust", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
-```
-
-```{r}
-model01 <- '
-perc_merit=~meritv01_perc_effort+meritv01_perc_talent 
-perc_nmerit=~meritv01_perc_wpart+meritv01_perc_netw  
-pref_merit=~meritv01_pref_effort+meritv01_pref_talent 
-pref_nmerit=~meritv01_pref_wpart+meritv01_pref_netw'
-
-fit1_c <- cfa(model = model01,data = dat01,estimator="MLR") # Continuo/ estimador ML Robust
-
-fit1_o <- cfa(model = model01,data = dat01,
-            ordered = c("meritv01_perc_effort","meritv01_perc_talent",
-                        "meritv01_perc_wpart","meritv01_perc_netw",
-                        "meritv01_pref_effort","meritv01_pref_talent",
-                        "meritv01_pref_wpart","meritv01_pref_netw"))
-
-#---------------------------------------------------#
-cnames <- c("Factor","Indicador","Loading (MLR)","Loading (DWLS)")
-kable(left_join(x = standardizedsolution(fit1_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),y = standardizedsolution(fit1_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),c("lhs","rhs")),
-      format = "markdown",digits = 2,col.names = cnames)
-```
-
-> *#summary*: good fit, heywood case item 8.
-
-## Version 02: Fixed order, by topic (N=`r dim(dat02)[1]`)
-
-1. Percepcion esfuerzo
-2. Preferencia esfuerzo
-3. Percepcion talento
-4. Preferencia talento
-5. Percepcion familia rica
-6. Preferencia familia rica
-7. Percepcion redes
-8. Preferencia redes
-
-```{r echo=FALSE,results='asis'}
-stargazer(dat02,type = "html")
-```
-
-```{r}
-cor02<-polychoric(dat02)
-corrplot(cor02$rho,method = "color", type = "upper", order = "hclust", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
-```
-
-```{r,results='hold'}
-model02 <- '
-perc_merit=~meritv02_perc_effort+meritv02_perc_talent 
-perc_nmerit=~meritv02_perc_wpart+meritv02_perc_netw  
-pref_merit=~meritv02_pref_effort+meritv02_pref_talent 
-pref_nmerit=~meritv02_pref_wpart+meritv02_pref_netw'
-
-fit2_c <- cfa(model = model02,data = dat02,estimator="MLR") # Continuo/ estimador ML Robust
-
-fit2_o <- cfa(model = model02,data = dat02,ordered = c("meritv02_perc_effort","meritv02_perc_talent",
-                                                      "meritv02_perc_wpart"  ,"meritv02_perc_netw",
-                                                      "meritv02_pref_effort" ,"meritv02_pref_talent",
-                                                      "meritv02_pref_wpart"  ,"meritv02_pref_netw"))
-
-#---------------------------------------------------#
-cnames <- c("Factor","Indicador","Loading (MLR)","Loading (DWLS)")
-kable(left_join(x = standardizedsolution(fit2_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),y = standardizedsolution(fit2_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),c("lhs","rhs")),
-      format = "markdown",digits = 2,col.names = cnames)
-```
-> **NOTA**: bad fit, heywood en estimación DWLS
-
-## Version 03: Random order (N=`r dim(dat03)[1]`)
-
-```{r echo=FALSE,results='asis'}
-stargazer(dat03,type = "html")
-```
-
-```{r}
-cor03<-polychoric(dat03)
-corrplot(cor03$rho,method = "color", type = "upper", order = "hclust", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
-```
-
-```{r,results='hold'}
-model03 <- '
-perc_merit=~meritv03_perc_effort+meritv03_perc_talent 
-perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw  
-pref_merit=~meritv03_pref_effort+meritv03_pref_talent 
-pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw'
-
-fit3_c <- cfa(model = model03,data = dat03,estimator="MLR",std.lv=FALSE) # Continuo/ estimador ML Robust
-
-fit3_o <- cfa(model = model03,data = dat03,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
-                                                       "meritv03_perc_wpart","meritv03_perc_netw",
-                                                       "meritv03_pref_effort","meritv03_pref_talent",
-                                                       "meritv03_pref_wpart","meritv03_pref_netw"),std.lv=FALSE)
-#---------------------------------------------------#
-cnames <- c("Factor","Indicador","Loading (MLR)","Loading (DWLS)")
-kable(left_join(x = standardizedsolution(fit3_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),y = standardizedsolution(fit3_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),c("lhs","rhs")),
-      format = "markdown",digits = 2,col.names = cnames)
-```
-
-> **NOTA**: fit not very goood (RMSEA .7), no heywood. Try tau-equivalence
-
-```{r,results='hold'}
-model03.tau <- '
-perc_merit=~ r1*meritv03_perc_effort+r1*meritv03_perc_talent 
-perc_nmerit=~r2*meritv03_perc_wpart+ r2*meritv03_perc_netw  
-pref_merit=~ r3*meritv03_pref_effort+r3*meritv03_pref_talent 
-pref_nmerit=~r4*meritv03_pref_wpart+ r4*meritv03_pref_netw'
-
-fit3_c.tau <- cfa(model = model03.tau,data = dat03,estimator="MLR",std.lv=FALSE) # Continuo/ estimador ML Robust
-
-fit3_o.tau <- cfa(model = model03.tau,data = dat03,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
-                                                               "meritv03_perc_wpart","meritv03_perc_netw",
-                                                               "meritv03_pref_effort","meritv03_pref_talent",
-                                                               "meritv03_pref_wpart","meritv03_pref_netw"), std.lv=FALSE)
-#---------------------------------------------------#
-cnames <- c("Factor","Indicador","Loading (MLR)","Loading (DWLS)")
-kable(left_join(x = standardizedsolution(fit3_c.tau ) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),y = standardizedsolution(fit3_o.tau) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),c("lhs","rhs")),
-      format = "markdown",digits = 2,col.names = cnames)
-```
-
-```{r}
-kable(rbind(ordinal    =fitmeasures(fit3_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-            ordinal.tau=fitmeasures(fit3_o.tau)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),format = "markdown",digits = 3)
-
-kable(anova(fit3_o,fit3_o.tau),format = "markdown",digits = 3)
-```
-
-> **NOTA:** El modelo con igualdad de cargas factoriales no deteriora significativamente el ajuste del modelo.
-
-## Version 04: muestra completa (N=`r dim(dat04)[1]`)
-
-```{r echo=FALSE,results='asis'}
-stargazer(select(dat04,perc_effort,perc_talent,perc_wpart,perc_netw,pref_effort,pref_talent,pref_wpart,pref_netw),type = "html")
-```
-
-```{r}
-cor04<-polychoric(select(dat04,perc_effort,perc_talent,perc_wpart,perc_netw,pref_effort,pref_talent,pref_wpart,pref_netw))
-corrplot(cor04$rho, method="color",type = "upper", order = "hclust", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
-```
-
-```{r,results='hold'}
-model04 <- '
-perc_merit =~ perc_effort+perc_talent 
-perc_nmerit=~perc_wpart +perc_netw  
-pref_merit =~ pref_effort+pref_talent 
-pref_nmerit=~pref_wpart +pref_netw'
-
-fit4_c <- cfa(model = model04,data = dat04,estimator="MLR",std.lv=FALSE) # Continuo/ estimador ML Robust
-
-fit4_o <- cfa(model = model04,data = dat04,ordered = c("perc_effort","perc_talent",
-                                                       "perc_wpart","perc_netw",
-                                                       "pref_effort","pref_talent",
-                                                       "pref_wpart","pref_netw"),std.lv=FALSE)
-#---------------------------------------------------#
-cnames <- c("Factor","Indicador","Loading (MLR)","Loading (DWLS)")
-kable(left_join(x = standardizedsolution(fit4_c ) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),y = standardizedsolution(fit4_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),c("lhs","rhs")),
-      format = "markdown",digits = 2,col.names = cnames)
-```
-
-### Resumen ajuste ola 01
-
-```{r}
-sum_fit<- bind_rows(fitmeasures(fit1_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                    fitmeasures(fit1_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                    fitmeasures(fit2_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                    fitmeasures(fit2_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                    fitmeasures(fit3_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                    fitmeasures(fit3_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                    fitmeasures(fit4_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                    fitmeasures(fit4_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")])
-sum_fit$mod <- c("Modelo 1","Modelo 1","Modelo 2","Modelo 2","Modelo 3","Modelo 3","Modelo 4","Modelo 4")
-sum_fit$est <- c("Continuo","Ordinal","Continuo","Ordinal","Continuo","Ordinal","Continuo","Ordinal")
-sum_fit <- select(sum_fit,mod,est,everything())
-
-kable(sum_fit,digits = 3,format = "html",row.names = F,booktabs=T, caption = "Resumen estadísticos de ajuste ola 1") %>% kable_styling(full_width = F) %>% 
-  collapse_rows(columns = 1,valign = "middle")  %>% 
-  footnote(number = c("Modelo 1: fixed order by percepction/preference", 
-                      "Modelo 2: fixed order by topic (i.e: effort)", 
-                      "Modelo 3: Randomized order",
-                      "Modelo 4: Complete sample"))
-```
-
-## Análisis por Grupos múltiples
-
-```{r}
-model04 <- '
-perc_merit =~ perc_effort+perc_talent 
-perc_nmerit=~perc_wpart +perc_netw  
-pref_merit =~ pref_effort+pref_talent 
-pref_nmerit=~pref_wpart +pref_netw'
-
-fit4_c_mg <- cfa(model = model04,data = dat04,estimator="MLR",std.lv=FALSE,group = "grupo") # Continuo/ estimador ML Robust
-fit4_o_mg <- cfa(model = model04,data = dat04,ordered = c("perc_effort","perc_talent",
-                                                          "perc_wpart","perc_netw",
-                                                          "pref_effort","pref_talent",
-                                                          "pref_wpart","pref_netw"),std.lv=FALSE, group = "grupo")
-#---------------------------------------------------#
-kable(rbind(continuo=fitmeasures(fit4_c_mg)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-            ordinal=  fitmeasures(fit4_o_mg)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
-      format = "markdown",digits = 3)
-
-cnames <- c("Grupo","Factor","Indicador","Loading (MLR)","Loading (DWLS)")
-kable(left_join(x = standardizedsolution(fit4_c_mg) %>% filter(op=="=~") %>% select(group,lhs,rhs,est.std),
-                y = standardizedsolution(fit4_o_mg) %>% filter(op=="=~") %>% select(group,lhs,rhs,est.std),
-                c("group","lhs","rhs")),
-      format = "markdown",digits = 2,col.names = cnames)
-```
-
-# Modelo Factorial de Segundo Orden:
-
-## a. Factor de segundo orden por Percepción y Preferencia
-
-```{r}
-model04b1or <- '
-perc_merit=~ perc_effort+perc_talent 
-perc_nmerit=~perc_wpart +perc_netw  
-pref_merit=~ pref_effort+pref_talent 
-pref_nmerit=~pref_wpart +pref_netw
-percep=~ perc_merit + perc_nmerit
-prefer=~ pref_merit + pref_nmerit'
-
-fit4a_c <- cfa(model = model04b1or,data = dat04, estimator="MLR")
-
-fit4b_c <- cfa(model = model04b1or,data = dat04,ordered = c("perc_effort","perc_talent","perc_wpart","perc_netw",
-                                                            "pref_effort","pref_talent","pref_wpart","pref_netw"))
-
-#---------------------------------------------------#
-kable(rbind(continuo=fitmeasures(fit4a_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-            ordinal= fitmeasures(fit4b_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
-      format = "markdown",digits = 3)
-
-cnames <- c("Factor","Indicador","Loading (MLR)","Loading (DWLS)")
-kable(left_join(x = standardizedsolution(fit4a_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
-                y = standardizedsolution(fit4b_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
-                c("lhs","rhs")),
-      format = "markdown",digits = 2,col.names = cnames)
-```
-
-> **NOTA:** Cargas factoriales de segundo orden son bajas. El ajuste del modelo no es bueno.
-
-## b. Factor de segundo orden por Meritocracia y No Meritocracia
-
-```{r}
-model04b2or <-'perc_merit=~ perc_effort+perc_talent
-               perc_nmerit=~perc_wpart +perc_netw
-               pref_merit=~ pref_effort+pref_talent
-               pref_nmerit=~pref_wpart +pref_netw
-               merit =~ perc_merit  + pref_merit
-               nmerit=~ perc_nmerit + pref_nmerit'
-
-# fit4b_c <- cfa(model = model04b2or,data = dat04,estimator="MLR") # No converge
-
-fit4b_o <- cfa(model = model04b2or,data = dat04, 
-               ordered = c("perc_effort","perc_talent","perc_wpart","perc_netw",
-                           "pref_effort","pref_talent","pref_wpart","pref_netw"))
-
-#---------------------------------------------------#
-kable(rbind(ordinal=fitmeasures(fit4b_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
-      format = "markdown",digits = 3)
-
-cnames <- c("Factor","Indicador","Loading (DWLS)")
-kable(standardizedsolution(fit4b_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
-      format = "markdown",digits = 2,col.names = cnames)
-```
-
-> **NOTA:** El modelo estimado con MLR con factores de segundo orden "meritocracia/ no meritocracia" no converge.
-EL modelo estimado a través de DWLS converge, posee cargas muy bajas y una muy alta (merit~pref_merit).
-Fit es malo (RMSEA >.10)
-
-# Wave 02
-
-```{r}
-data02 <- sjlabelled::read_spss(path = "input/data/original/Estudio_3_ola2.sav",verbose = FALSE)
-load(file = "input/data/proc/datawide.RData")
-dat01 <- data02 %>% select(starts_with("meritv03_p")) 
-```
-
-## ATE indicadores meritocracia
-
-```{r, out.width="75%"}
-m01<- lm(perc_effort_w2~treat,data = wide01)
-m02<- lm(perc_talent_w2~treat,data = wide01)
-m03<- lm(perc_wpart_w2~treat, data = wide01)
-m04<- lm(perc_netw_w2~treat,  data = wide01)
-#--------------------------------------------#
-m05<- lm(pref_effort_w2~treat,data = wide01)
-m06<- lm(pref_talent_w2~treat,data = wide01)
-m07<- lm(pref_wpart_w2 ~treat, data = wide01)
-m08<- lm(pref_netw_w2  ~treat,  data = wide01)
-#--------------------------------------------#
-stargazer::stargazer(m01,m02,m03,m04,type = "text")
-stargazer::stargazer(m05,m06,m07,m08,type = "text")
-```
-
-> **NOTA**: Los tratamientos de información no se asocian de manera estadísticamente significativa con ninguno de los indicadores de la escala de meritocracia.
-
-## CFA Muestra completa (Control, Desigualdad y Pobreza)
-
-```{r echo=FALSE,results='asis'}
-stargazer(dat01,type = "html")
-```
-
-```{r}
-dat01cor<- select(dat01,meritv03_perc_effort,meritv03_perc_talent,
-                        meritv03_perc_wpart,meritv03_perc_netw,
-                        meritv03_pref_effort,meritv03_pref_talent,
-                        meritv03_pref_wpart,meritv03_pref_netw)
-cor01<-polychoric(dat01cor)
-corrplot(cor01$rho, type = "upper", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
-```
-
-```{r,results='hold'}
-model03 <- '
-perc_merit=~meritv03_perc_effort+meritv03_perc_talent 
-perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw  
-pref_merit=~meritv03_pref_effort+meritv03_pref_talent 
-pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw'
-
-fitw02_c <- cfa(model = model03,data = dat01,estimator="MLR",std.lv=FALSE) # Continuo/ estimador ML Robust
-
-fitw02_o <- cfa(model = model03,data = dat01,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
-                                                         "meritv03_perc_wpart","meritv03_perc_netw",
-                                                         "meritv03_pref_effort","meritv03_pref_talent",
-                                                         "meritv03_pref_wpart","meritv03_pref_netw"))
-#---------------------------------------------------#
-kable(rbind(continuo=fitmeasures(fitw02_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-            ordinal= fitmeasures(fitw02_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
-      format = "markdown",digits = 3)
-
-cnames <- c("Factor","Indicador","Loading (MLR)","Loading (DWLS)")
-kable(left_join(x = standardizedsolution(fitw02_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
-                y = standardizedsolution(fitw02_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
-                c("lhs","rhs")),
-      format = "markdown",digits = 2,col.names = cnames)
-```
-
-> **NOTA**: El ajuste de muestra completa en wave 02 con estimador MLR (robusto) es bastante bueno (CFI, RMSEA). El ajuste con DWLS es bastante bueno, CFI mejora, RMSEA aumenta un poco (0.59).
-
-* **Recomendación**: es probable que el estimador MLR sea mejor que el ordinal.
-
-### Factor de segundo orden
-
-```{r}
-modw2 <- '
-perc_merit=~meritv03_perc_effort+meritv03_perc_talent 
-perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw  
-pref_merit=~meritv03_pref_effort+meritv03_pref_talent 
-pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw 
-merit =~perc_merit+pref_merit  
-nmerit=~perc_nmerit+pref_nmerit'
-
-# mod2_merit_c <- cfa(model = modw2,data = dat01) # Continuo/ estimador ML Robust
-
-mod2_merit_o <- cfa(model = modw2,data = dat01,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
-                                                           "meritv03_perc_wpart","meritv03_perc_netw",
-                                                           "meritv03_pref_effort","meritv03_pref_talent",
-                                                           "meritv03_pref_wpart","meritv03_pref_netw"))
-
-#---------------------------------------------------#
-kable(rbind(ordinal=fitmeasures(mod2_merit_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
-      format = "markdown",digits = 3)
-
-cnames <- c("Factor","Indicador","Loading (DWLS)")
-kable(standardizedsolution(mod2_merit_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
-      format = "markdown",digits = 2,col.names = cnames)
-```
-
-> * Factor de segundo orden usando MLR no converge.
-> * Factor de segundo orden usando DWLS converge, pero las cargas no son buenas y el ajuste tampoco.
-
-## CFA Muestra Control
-
-```{r}
-w02con <- data02 %>% filter(FL_6_DO_Control==1) %>% select(starts_with("meritv03_p")) 
-```
-
-```{r echo=FALSE,results='asis'}
-stargazer(w02con,type = "html")
-```
-
-```{r}
-corw02_1<-polychoric(w02con)
-corrplot(corw02_1$rho, type = "upper", order = "hclust", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
-```
-
-```{r,results='hold'}
-model03 <- '
-perc_merit=~meritv03_perc_effort+meritv03_perc_talent 
-perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw  
-pref_merit=~meritv03_pref_effort+meritv03_pref_talent 
-pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw'
-
-fitw02a_c <- cfa(model = model03,data = w02con,estimator="MLR") # Continuo/ estimador ML Robust
-
-fitw02a_o <- cfa(model = model03,data = w02con,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
-                                                           "meritv03_perc_wpart","meritv03_perc_netw",
-                                                           "meritv03_pref_effort","meritv03_pref_talent",
-                                                           "meritv03_pref_wpart","meritv03_pref_netw"))
-#---------------------------------------------------#
-kable(rbind(continuo=fitmeasures(fitw02a_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-            ordinal= fitmeasures(fitw02a_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
-      format = "markdown",digits = 3)
-
-cnames <- c("Factor","Indicador","Loading (MLR)","Loading (DWLS)")
-kable(left_join(x = standardizedsolution(fitw02a_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
-                y = standardizedsolution(fitw02a_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
-                c("lhs","rhs")),
-      format = "markdown",digits = 2,col.names = cnames)
-```
-
-## CFA Muestra Pobreza
-
-```{r}
-w02pob <- data02 %>% filter(FL_6_DO_Tratamiento==1) %>% select(starts_with("meritv03_p")) 
-```
-
-```{r echo=FALSE,results='asis'}
-stargazer(w02pob,type = "html")
-```
-
-```{r}
-corw02_2<-polychoric(w02pob)
-corrplot(corw02_2$rho, type = "upper", order = "hclust", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
-```
-
-```{r,results='hold'}
-model03 <- '
-perc_merit=~meritv03_perc_effort+meritv03_perc_talent 
-perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw  
-pref_merit=~meritv03_pref_effort+meritv03_pref_talent 
-pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw'
-
-fitw02b_c <- cfa(model = model03,data = w02pob,estimator="MLR") # Continuo/ estimador ML Robust
-
-fitw02b_o <- cfa(model = model03,data = w02pob,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
-                                                           "meritv03_perc_wpart","meritv03_perc_netw",
-                                                           "meritv03_pref_effort","meritv03_pref_talent",
-                                                           "meritv03_pref_wpart","meritv03_pref_netw"))
-
-#---------------------------------------------------#
-kable(rbind(continuo=fitmeasures(fitw02b_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-            ordinal= fitmeasures(fitw02b_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
-      format = "markdown",digits = 3)
-
-cnames <- c("Factor","Indicador","Loading (MLR)","Loading (DWLS)")
-kable(left_join(x = standardizedsolution(fitw02b_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
-                y = standardizedsolution(fitw02b_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
-                c("lhs","rhs")),
-      format = "markdown",digits = 2,col.names = cnames)
-```
-
-## CFA Muestra Desigualdad
-
-```{r}
-w02des <- data02 %>% filter(FL_6_DO_Tratamiento_desigualdad==1) %>% select(starts_with("meritv03_p")) 
-```
-
-```{r echo=FALSE,results='asis'}
-stargazer(w02des,type = "html")
-```
-
-```{r}
-corw02_3<-polychoric(w02des)
-corrplot(corw02_3$rho, type = "upper", order = "hclust", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
-```
-
-```{r,results='hold'}
-model03 <- '
-perc_merit=~meritv03_perc_effort+meritv03_perc_talent
-perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw  
-pref_merit=~meritv03_pref_effort+meritv03_pref_talent 
-pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw'
-
-fitw02c_c <- cfa(model = model03,data = w02des,estimator="MLR") # Continuo/ estimador ML Robust
-
-fitw02c_o <- cfa(model = model03,data = w02des,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
-                                                           "meritv03_perc_wpart","meritv03_perc_netw",
-                                                           "meritv03_pref_effort","meritv03_pref_talent",
-                                                           "meritv03_pref_wpart","meritv03_pref_netw"))
-#---------------------------------------------------#
-kable(rbind(continuo=fitmeasures(fitw02c_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-            ordinal= fitmeasures(fitw02c_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
-      format = "markdown",digits = 3)
-
-cnames <- c("Factor","Indicador","Loading (MLR)","Loading (DWLS)")
-kable(left_join(x = standardizedsolution(fitw02c_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
-                y = standardizedsolution(fitw02c_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
-                c("lhs","rhs")),
-      format = "markdown",digits = 2,col.names = cnames)
-```
-
-# Resumen ajuste ola 2
-
-```{r}
-sum_fitw02<- bind_rows(fitmeasures(fitw02_c)[ c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                       fitmeasures(fitw02_o)[ c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                       fitmeasures(fitw02a_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                       fitmeasures(fitw02a_o)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                       fitmeasures(fitw02b_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                       fitmeasures(fitw02b_o)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                       fitmeasures(fitw02c_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                       fitmeasures(fitw02c_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")])
-
-sum_fitw02$mod <- c("Modelo 1","Modelo 1","Modelo 2","Modelo 2","Modelo 3","Modelo 3","Modelo 4","Modelo 4")
-sum_fitw02$est <- c("Continuo","Ordinal","Continuo","Ordinal","Continuo","Ordinal","Continuo","Ordinal")
-sum_fitw02 <- select(sum_fitw02,mod,est,everything())
-
-kable(sum_fitw02,digits = 3,format = "html",row.names = F,booktabs=T, caption = "Resumen ajuste ola 2") %>% 
-  kable_styling(full_width = F) %>% 
-  collapse_rows(columns = 1,valign = "middle") %>% 
-  footnote(number = c("Modelo 1: Muestra completa", 
-                      "Modelo 2: Muestra grupo Control", 
-                      "Modelo 3: Muestra grupo tratamiento pobreza",
-                      "Modelo 4: Muestra grupo tratamiento desigualdad"))
-```
-
-# Wave 03
-
-```{r}
-data03 <- sjlabelled::read_spss(path = "input/data/original/Estudio_3_ola3.sav",verbose = FALSE)
-dat01 <- data03 %>% select(starts_with("meritv03_p")) 
-```
-
-```{r}
-# sjt.fa(data = sjlabelled::remove_label(dat01),show.comm = TRUE)
-```
-
-```{r}
-model03 <- '
-perc_merit=~meritv03_perc_effort+meritv03_perc_talent 
-perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw  
-pref_merit=~meritv03_pref_effort+meritv03_pref_talent 
-pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw'
-
-fitw03c_c <- cfa(model = model03,data = dat01,estimator="MLR") # Continuo/ estimador ML Robust
-
-fitw03c_o <- cfa(model = model03,data = dat01,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
-                                                          "meritv03_perc_wpart","meritv03_perc_netw",
-                                                          "meritv03_pref_effort","meritv03_pref_talent",
-                                                          "meritv03_pref_wpart","meritv03_pref_netw"))
-#---------------------------------------------------#
-kable(rbind(continuo=fitmeasures(fitw03c_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-            ordinal= fitmeasures(fitw03c_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
-      format = "markdown",digits = 3)
-
-cnames <- c("Factor","Indicador","Loading (MLR)","Loading (DWLS)")
-kable(left_join(x = standardizedsolution(fitw03c_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
-                y = standardizedsolution(fitw03c_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
-                c("lhs","rhs")),
-      format = "markdown",digits = 2,col.names = cnames)
-```
-**summary**
-
-> * Ajuste usando estimador MLR  es bueno (CFI = 0.971, RMSEA = 0.052). El chi2 (robusto) es 62.421  
-> * Ajuste usando estimador DWLS es bueno (CFI = 0.974, RMSEA = 0.067). El chi2 (robusto) es 106.111
-
-# Resumen ajuste ola 3
-
-```{r}
-sum_fitw03<- bind_rows(fitmeasures(fitw03c_c)[ c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                       fitmeasures(fitw03c_o)[ c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")])
-
-sum_fitw03$mod <- c("Modelo 1","Modelo 1")
-sum_fitw03$est <- c("Continuo","Ordinal")
-sum_fitw03 <- select(sum_fitw03,mod,est,everything())
-
-kable(sum_fitw03,digits = 3,format = "html",row.names = F,booktabs=T,caption = "Resumen ajuste ola 3") %>% 
-  kable_styling(full_width = F) %>% 
-  collapse_rows(columns = 1,valign = "middle") %>% 
-  footnote(number = c("Modelo 1: Muestra completa"))
-```
-
-# Análisis longitudinal
+# Longitudinal analysis
 
@@
 corwide01a<- polychoric(x = select(g01 ,starts_with(match = "perc_")),na.rm = T)
-corrplot(corwide01a$rho, method = "color" ,type = "upper", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
+rownames(corwide01a$rho) <-c("1.
+pc_effortw1","2.
+pc_talentw1","3.
+pc_wpartw1 ","4.
+pc_netww1","5.
+pc_effortw2","6.
+pc_talentw2","7.
+pc_wpartw2","8.
+pc_netww2","9.
+pc_effortw3","10.
+pc_talentw3","11.
+pc_wpartw3 ","12.
+pc_netww3")
+colnames(corwide01a$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)","(9)","(10)","(11)","(12)")
+corrplot(corwide01a$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
 
 corwide01b<- polychoric(x = select(g01 ,starts_with(match = "pref_")),na.rm = T)
-corrplot(corwide01b$rho, method = "color" ,type = "upper", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
+rownames(corwide01b$rho) <-c("1.
+pf_effortw1","2.
+pf_talentw1","3.
+pf_wpartw1 ","4.
+pf_netww1","5.
+pf_effortw2","6.
+pf_talentw2","7.
+pf_wpartw2","8.
+pf_netww2","9.
+pf_effortw3","10.
+pf_talentw3","11.
+pf_wpartw3 ","12.
+pf_netww3")
+colnames(corwide01b$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)","(9)","(10)","(11)","(12)")
+corrplot(corwide01b$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
+
 ```
@@
 corwide02a<- polychoric(x = select(g02 ,starts_with(match = "perc_")),na.rm = T)
-corrplot(corwide02a$rho, method = "color" ,type = "upper", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
+rownames(corwide02a$rho) <-c("1.
+pc_effortw1","2.
+pc_talentw1","3.
+pc_wpartw1 ","4.
+pc_netww1","5.
+pc_effortw2","6.
+pc_talentw2","7.
+pc_wpartw2","8.
+pc_netww2","9.
+pc_effortw3","10.
+pc_talentw3","11.
+pc_wpartw3 ","12.
+pc_netww3")
+colnames(corwide02a$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)","(9)","(10)","(11)","(12)")
+corrplot(corwide02a$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
 
 corwide02b<- polychoric(x = select(g02 ,starts_with(match = "pref_")),na.rm = T)
-corrplot(corwide02b$rho, method = "color" ,type = "upper", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
+rownames(corwide02b$rho) <-c("1.
+pf_effortw1","2.
+pf_talentw1","3.
+pf_wpartw1 ","4.
+pf_netww1","5.
+pf_effortw2","6.
+pf_talentw2","7.
+pf_wpartw2","8.
+pf_netww2","9.
+pf_effortw3","10.
+pf_talentw3","11.
+pf_wpartw3 ","12.
+pf_netww3")
+colnames(corwide02b$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)","(9)","(10)","(11)","(12)")
+corrplot(corwide02b$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
 ```
@@
 corwide03a<- polychoric(x = select(g03 ,starts_with(match = "perc_")),na.rm = T)
-corrplot(corwide03a$rho, type = "upper", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE,tl.cex = 0.75)
+rownames(corwide03a$rho) <-c("1.
+pc_effortw1","2.
+pc_talentw1","3.
+pc_wpartw1 ","4.
+pc_netww1","5.
+pc_effortw2","6.
+pc_talentw2","7.
+pc_wpartw2","8.
+pc_netww2","9.
+pc_effortw3","10.
+pc_talentw3","11.
+pc_wpartw3 ","12.
+pc_netww3")
+colnames(corwide03a$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)","(9)","(10)","(11)","(12)")
+corrplot(corwide03a$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
 
 corwide03b<- polychoric(x = select(g03 ,starts_with(match = "pref_")),na.rm = T)
-corrplot(corwide03b$rho, method = "color" ,type = "upper", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
+rownames(corwide03b$rho) <-c("1.
+pf_effortw1","2.
+pf_talentw1","3.
+pf_wpartw1 ","4.
+pf_netww1","5.
+pf_effortw2","6.
+pf_talentw2","7.
+pf_wpartw2","8.
+pf_netww2","9.
+pf_effortw3","10.
+pf_talentw3","11.
+pf_wpartw3 ","12.
+pf_netww3")
+colnames(corwide03b$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)","(9)","(10)","(11)","(12)")
+corrplot(corwide03b$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
 ```
@@
 corwideA<- polychoric(x = select(wide02 ,starts_with(match = "perc_")),na.rm = T)
-corrplot(corwideA$rho, method = "color" ,type = "upper", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
+rownames(corwideA$rho) <-c("1.
+pc_effortw1","2.
+pc_talentw1","3.
+pc_wpartw1 ","4.
+pc_netww1","5.
+pc_effortw2","6.
+pc_talentw2","7.
+pc_wpartw2","8.
+pc_netww2","9.
+pc_effortw3","10.
+pc_talentw3","11.
+pc_wpartw3 ","12.
+pc_netww3")
+colnames(corwideA$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)","(9)","(10)","(11)","(12)")
+corrplot(corwideA$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
 
 corwideB<- polychoric(x = select(wide02 ,starts_with(match = "pref_")),na.rm = T)
-corrplot(corwideB$rho, method = "color" ,type = "upper", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
+rownames(corwideB$rho) <-c("1.
+pf_effortw1","2.
+pf_talentw1","3.
+pf_wpartw1 ","4.
+pf_netww1","5.
+pf_effortw2","6.
+pf_talentw2","7.
+pf_wpartw2","8.
+pf_netww2","9.
+pf_effortw3","10.
+pf_talentw3","11.
+pf_wpartw3 ","12.
+pf_netww3")
+colnames(corwideB$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)","(9)","(10)","(11)","(12)")
+corrplot(corwideB$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
 ```
@@
 
-fitlong_c <- cfa(model = model_long,data = wide02,estimator="MLR") # Continuo/ estimador ML Robust
-
-indicadores <-c("perc_effort_w1","perc_talent_w1","perc_wpart_w1","perc_netw_w1",
+fitlong_c <- cfa(model = model_long,data = wide02,estimator="MLR") # Continuous/ estimador ML Robust
+
+Indicatores <-c("perc_effort_w1","perc_talent_w1","perc_wpart_w1","perc_netw_w1",
                 "pref_effort_w1","pref_talent_w1","pref_wpart_w1","pref_netw_w1",
@@
 
-fitlong_o <- cfa(model = model_long,data = wide02,ordered = indicadores)
+fitlong_o <- cfa(model = model_long,data = wide02,ordered = Indicatores)
 
 #---------------------------------------------------#
-kable(rbind(continuo=fitmeasures(fitlong_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+kable(rbind(Continuous=fitmeasures(fitlong_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
             ordinal= fitmeasures(fitlong_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
@@
 
-cnames <- c("Factor","Indicador","Loading (MLR)","Loading (DWLS)")
+cnames <- c("Factor","Indicator","Loading (MLR)","Loading (DWLS)")
 kable(left_join(x = standardizedsolution(fitlong_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
@@
 
-**NOTA**: En vista de que no existe una librería que permita estimar invarianza longitudinal asumiendo indicadores ordinales, las primeras estimaciones las realizaré asumiendo indicadores continuos (estimator=MLR).
+**NOTE**: En vista de que no existe una librería que permita estimar invarianza longitudinal asumiendo Indicatores ordinales, las primeras estimaciones las realizaré asumiendo Indicatores Continuouss (estimator=MLR).
 
@@
 
-**NOTA**: El modelo que deja libre 1 parámetro (metrA2) mejora el ajuste, pero la diff del chi2 sigue siendo significativa al 0.05.
+**NOTE**: El modelo que deja libre 1 parámetro (metrA2) mejora el ajuste, pero la diff del chi2 sigue siendo significativa al 0.05.
 
@@
 
-**NOTA**: the chi2 test shows that scalar vs metric models are not significant.
+**NOTE**: the chi2 test shows that scalar vs metric models are not significant.
 
@@
 
-* Invarianza con indicadores categóricos
+* Invarianza con Indicatores categóricos
 https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5121102/pdf/nihms757737.pdf

### 2020-04-06 render-site.R

@@ -29,5 +29,7 @@ for (i in 1:length(ext)) {
   file.remove(paste0("docs/",dir(path="docs", pattern=ext[i]))) # delete files
 }
 
+browseURL(url = "docs/index.html")
+
 # Keep only the original .rmd files (site), configuration (_site.yml and _config.yml)
 # and prod_ folders (cache and files)

### 2020-04-06 production/prod_analysis-cfa.rmd

--- 2020-03-20 production/prod_analysis-cfa.rmd
@@
 
-# Análisis PPM-S
+# Analysis PPM-S
 
-# Wave 01
+This section, documents all the procedures related to the data analysis and measurement models.
+
+# Wave 01 {.tabset .tabset-fade .tabset-pills}
 
 ```{r eval=FALSE, include=FALSE}
-rmarkdown::render(input = "production/analysis-cfa.Rmd",output_format = "html_document")
+rmarkdown::render(input = "production/prod_analysis-cfa.Rmd",output_format = "html_document")
 ```
@@
 
+**Libraries**
 ```{r}
@@
 
+**Load data**
 ```{r}
@@
 
-## Version 01: Fixed order by perception / preferences (N=`r dim(dat01)[1]`)
+---
 
-1. Percepcion esfuerzo
-2. Percepcion talento
-3. Percepcion familia rica
-4. Percepcion redes
-5. Preferencia esfuerzo
-6. Preferencia talento
-7. Preferencia familia rica
-8. Preferencia redes
+**Modality**
+
+Each group correspond to the application modality of the scale.
+
+1. **Version 01:** Fixed order by perception / preferences (N=`r dim(dat01)[1]`)
+
+2. **Version 02:** Fixed order, by topic (N=`r dim(dat02)[1]`)
+
+3.  **Version 03:** Random order (N=`r dim(dat03)[1]`)
+
+4. **Version 04:** Complete sample (N=`r dim(dat04)[1]`)
+
+## Version 01 (N=`r dim(dat01)[1]`){.tabset}
+
+| Spanish (original)          | English (translated)   |
+|-----------------------------|------------------------|
+| 1. Percepción esfuerzo      | Perception effort      |
+| 2. Percepción talento       | Perception talent      |
+| 3. Percepción familia rica  | Perception rich family |
+| 4. Percepción redes         | Perception networks    |
+| 5. Preferencia esfuerzo     | Preference effort      |
+| 6. Preferencia talento      | Preference talent      |
+| 7. Preferencia familia rica | Preference rich family |
+| 8. Preferencia redes        | Preference networks    |
+Table: Mode 01: Fixed order by perception
 
 ```{r echo=FALSE,results='asis'}
-stargazer(dat01,type = "html")
+stargazer(dat01,type = "html",
+          title = "Descriptive statistics",
+          font.size = "small",
+          median = TRUE,
+          header=FALSE,
+          digit.separator = "",
+          digits = 2,
+          covariate.labels = c("1. pc.effort", "2. pc.talent","3. pc.wpart","4. pc.netw","5. pf.effort","6. pf.talent","7. pf.wpart", "8. pf.netw"))
+
 ```
@@
 cor01<-polychoric(dat01)
-corrplot(cor01$rho,method = "color" ,type = "upper", order = "hclust", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
+rownames(cor01$rho) <- c("1. pc_effort", "2. pc_talent","3. pc_wpart","4. pc_netw","5. pf_effort","6. pf_talent","7. pf_wpart", "8. pf_netw")
+colnames(cor01$rho) <- c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)")
+corrplot(cor01$rho,method = "color" ,type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
 ```
@@
 ```{r}
-model01 <- '
-perc_merit=~meritv01_perc_effort+meritv01_perc_talent 
+model01 <- 'perc_merit=~meritv01_perc_effort+meritv01_perc_talent
             perc_nmerit=~meritv01_perc_wpart+meritv01_perc_netw
@@
 
-fit1_c <- cfa(model = model01,data = dat01,estimator="MLR") # Continuo/ estimador ML Robust
+fit1_c <- cfa(model = model01,data = dat01,estimator="MLR") # Continuous/ estimator ML Robust
 
@@
 #---------------------------------------------------#
-# #summary(fit1_c,standardized=TRUE, fit.measures=TRUE)}
-fitmeasures(fit1_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fit1_c,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE, curve = 1.5)
-
-# #summary(fit1_o,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fit1_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fit1_o,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE, curve = 1.5)
+cnames <- c("Factor","Indicator","Loading (MLR)","Loading (DWLS)")
+kable(left_join(x = standardizedsolution(fit1_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),y = standardizedsolution(fit1_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),c("lhs","rhs")),
+      format = "markdown",digits = 2,col.names = cnames, caption = "Factor loadings")
 ```
 
-> *#summary*: good fit, heywood case item 8.
+> **NOTE**: good fit, heywood case item 8.
 
-## Version 02: Fixed order, by topic (N=`r dim(dat02)[1]`)
+## Version 02  (N=`r dim(dat02)[1]`)
 
-1. Percepcion esfuerzo
-2. Preferencia esfuerzo
-3. Percepcion talento
-4. Preferencia talento
-5. Percepcion familia rica
-6. Preferencia familia rica
-7. Percepcion redes
-8. Preferencia redes
+| Spanish (original)          | English (translated)   |
+|-----------------------------|------------------------|
+| 1. Percepción esfuerzo      | Perception effort      |
+| 2. Preferencia esfuerzo     | Preference effort      |
+| 3. Percepción talento       | Perception talent      |
+| 4. Preferencia talento      | Preference talent      |
+| 5. Percepción familia rica  | Perception rich family |
+| 6. Preferencia familia rica | Preference rich family |
+| 7. Percepción redes         | Perception networks    |
+| 8. Preferencia redes        | Preference networks    |
+Table: Mode 02: Fixed order, by topic.
 
 ```{r echo=FALSE,results='asis'}
-stargazer(dat02,type = "html")
+stargazer(dat02,type = "html",
+          title = "Descriptive statistics",
+          font.size = "small",
+          median = TRUE,
+          header=FALSE,
+          digit.separator = "",
+          digits = 2,
+          covariate.labels = c("1. pc.effort","2. pf.effort","3. pc.talent","4. pf.talent","5. pc.wpart","6. pf.wpart","7. pc.netw","8 pc.netw"))
 ```
@@
 cor02<-polychoric(dat02)
-corrplot(cor02$rho,method = "color", type = "upper", order = "hclust", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
+rownames(cor02$rho) <-c("1. pc_effort", "2. pc_talent","3. pc_wpart","4. pc_netw","5. pf_effort","6. pf_talent","7. pf_wpart", "8. pf_netw")
+colnames(cor02$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)")
+corrplot(cor02$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
 ```
@@
 ```{r,results='hold'}
-model02 <- '
-perc_merit=~meritv02_perc_effort+meritv02_perc_talent 
+model02 <- 'perc_merit=~meritv02_perc_effort+meritv02_perc_talent
             perc_nmerit=~meritv02_perc_wpart+meritv02_perc_netw
@@
 
-fit2_c <- cfa(model = model02,data = dat02,estimator="MLR") # Continuo/ estimador ML Robust
+fit2_c <- cfa(model = model02,data = dat02,estimator="MLR") # Continuous/ estimator ML Robust
 
 fit2_o <- cfa(model = model02,data = dat02,ordered = c("meritv02_perc_effort","meritv02_perc_talent",
-                                                     "meritv02_perc_wpart","meritv02_perc_netw",
-                                                     "meritv02_pref_effort","meritv02_pref_talent",
-                                                     "meritv02_pref_wpart","meritv02_pref_netw"))
+                                                       "meritv02_perc_wpart"  ,"meritv02_perc_netw",
+                                                       "meritv02_pref_effort" ,"meritv02_pref_talent",
+                                                       "meritv02_pref_wpart"  ,"meritv02_pref_netw"))
 
 #---------------------------------------------------#
-#summary(fit2_c,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fit2_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fit2_c,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE, curve = 1.5)
+cnames <- c("Factor","Indicator","Loading (MLR)","Loading (DWLS)")
+kable(left_join(x = standardizedsolution(fit2_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),y = standardizedsolution(fit2_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),c("lhs","rhs")),
+      format = "markdown",digits = 2,col.names = cnames, caption = "Factor loadings")
+```
+> **NOTE**: bad fit, heywood case using DWLS
 
-#summary(fit2_o,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fit2_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fit2_o,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE, curve = 1.5)
-```
-> **NOTA**: bad fit, heywood en estimación DWLS
+## Version 03 (N=`r dim(dat03)[1]`)
 
-## Version 03: Random order (N=`r dim(dat03)[1]`)
+| Spanish (original)          | English (translated)   |
+|-----------------------------|------------------------|
+| Orden aleatorio             | Randomized order       |
+
+Table: Mode 03: Randomized order.
 
 ```{r echo=FALSE,results='asis'}
-stargazer(dat03,type = "html")
+stargazer(dat03,type = "html",
+          title = "Descriptive statistics",
+          font.size = "small",
+          median = TRUE,
+          header=FALSE,
+          digit.separator = "",
+          digits = 2,
+          covariate.labels = c("1. pc.effort", "2. pc.talent","3. pc.wpart","4. pc.netw","5. pf.effort","6. pf.talent","7. pf.wpart", "8. pf.netw"))
 ```
@@
 cor03<-polychoric(dat03)
-corrplot(cor03$rho,method = "color", type = "upper", order = "hclust", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
+rownames(cor03$rho) <-c("1. pc_effort", "2. pc_talent","3. pc_wpart","4. pc_netw","5. pf_effort","6. pf_talent","7. pf_wpart", "8. pf_netw")
+colnames(cor03$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)")
+corrplot(cor03$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
 ```
@@
 
-fit3_c <- cfa(model = model03,data = dat03,estimator="MLR",std.lv=FALSE) # Continuo/ estimador ML Robust
+fit3_c <- cfa(model = model03,data = dat03,estimator="MLR",std.lv=FALSE) # Continuous/ estimator ML Robust
 
@@
 #---------------------------------------------------#
-#summary(fit3_c,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fit3_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fit3_c,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE, curve = 1.5)
-
-#summary(fit3_o,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fit3_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fit3_o,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE, curve = 1.5)
+cnames <- c("Factor","Indicator","Loading (MLR)","Loading (DWLS)")
+kable(left_join(x = standardizedsolution(fit3_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),y = standardizedsolution(fit3_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),c("lhs","rhs")),
+      format = "markdown",digits = 2,col.names = cnames, caption = "Factor loadings")
 ```
 
-> **NOTA**: fit not very goood (RMSEA .7), no heywood. Try tau-equivalence
+> **NOTE**: fit not very goood (RMSEA .7), no heywood. Try tau-equivalence
+
+### Tau-equivalence
 
@@
 
-fit3_c.tau <- cfa(model = model03.tau,data = dat03,estimator="MLR",std.lv=FALSE) # Continuo/ estimador ML Robust
+fit3_c.tau <- cfa(model = model03.tau,data = dat03,estimator="MLR",std.lv=FALSE) # Continuous/ estimator ML Robust
 
@@
 #---------------------------------------------------#
-#summary(fit3_c.tau,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fit3_c.tau)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fit3_c.tau,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE, curve = 1.5)
-
-#summary(fit3_o,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fit3_o.tau)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fit3_o.tau,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE, curve = 1.5)
+cnames <- c("Factor","Indicator","Loading (MLR)","Loading (DWLS)")
+kable(left_join(x = standardizedsolution(fit3_c.tau ) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),y = standardizedsolution(fit3_o.tau) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),c("lhs","rhs")),
+      format = "markdown",digits = 2,col.names = cnames)
 ```
@@
 ```{r}
-rbind(fitmeasures(fit3_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-      fitmeasures(fit3_o.tau)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")])
-anova(fit3_o,fit3_o.tau)
+kable(rbind(Ordinal    =fitmeasures(fit3_o    )[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            Ordinal.tau=fitmeasures(fit3_o.tau)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
+      format = "markdown",
+      digits = 3, caption = "Fit tau-equivalence")
+kable(anova(fit3_o,fit3_o.tau),
+      format = "markdown",
+      digits = 3)
 ```
 
-> **NOTA:** El modelo con igualdad de cargas factoriales no deteriora significativamente el ajuste del modelo.
+<!-- > **NOTE:** El modelo con igualdad de cargas factoriales (en Percepción meritocracia) no deteriora significativamente el ajuste del modelo. -->
 
-## Version 04: muestra completa (N=`r dim(dat04)[1]`)
+> **NOTE:** The model with equal factor loadings (in Meritocracy Perception) does not significantly deteriorate the model fit.
+
+## Version 04 (N=`r dim(dat04)[1]`)
+
+| Spanish (original)                           | English (translated)                 |
+|----------------------------------------------|--------------------------------------|
+| Orden fijo, según percepción/preferencia     | Fixed order, by perception/preference|
+| Orden fijo, por tópico                       | Fixed order, by topic.               |
+| Orden aleatorizado                           | Randomized order                     |
+Table: Mode 04: Complete sample
 
@@
 stargazer(select(dat04,perc_effort,perc_talent,perc_wpart,perc_netw,pref_effort,pref_talent,pref_wpart,pref_netw),
-          type = "html")
+          type = "html",
+          title = "Descriptive statistics",
+          font.size = "small",
+          median = TRUE,
+          header=FALSE,
+          digit.separator = "",
+          digits = 2,
+          covariate.labels = c("1. pc.effort", "2. pc.talent","3. pc.wpart","4. pc.netw","5. pf.effort","6. pf.talent","7. pf.wpart", "8. pf.netw"))
 ```
@@
 cor04<-polychoric(select(dat04,perc_effort,perc_talent,perc_wpart,perc_netw,pref_effort,pref_talent,pref_wpart,pref_netw))
-corrplot(cor04$rho, method="color",type = "upper", order = "hclust", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
+rownames(cor04$rho) <-c("1. pc_effort", "2. pc_talent","3. pc_wpart","4. pc_netw","5. pf_effort","6. pf_talent","7. pf_wpart", "8. pf_netw")
+colnames(cor04$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)")
+corrplot(cor04$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
 ```
@@
 
-fit4_c <- cfa(model = model04,data = dat04,estimator="MLR",std.lv=FALSE) # Continuo/ estimador ML Robust
+fit4_c <- cfa(model = model04,data = dat04,estimator="MLR",std.lv=FALSE) # Continuous/ estimator ML Robust
 
@@
 #---------------------------------------------------#
-#summary(fit4_c,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fit4_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fit3_c,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE,curve = 2)
-
-#summary(fit4_o,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fit4_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fit3_o,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE,curve = 2)
+cnames <- c("Factor","Indicator","Loading (MLR)","Loading (DWLS)")
+kable(left_join(x = standardizedsolution(fit4_c ) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),y = standardizedsolution(fit4_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),c("lhs","rhs")),
+      format = "markdown",digits = 2,col.names = cnames, caption = "Factor loadings")
 ```
 
-### Resumen ajuste ola 01
+# Summary fit indices wave 01
 
@@
                     fitmeasures(fit4_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")])
-sum_fit$mod <- c("Modelo 1","Modelo 1","Modelo 2","Modelo 2","Modelo 3","Modelo 3","Modelo 4","Modelo 4")
-sum_fit$est <- c("Continuo","Ordinal","Continuo","Ordinal","Continuo","Ordinal","Continuo","Ordinal")
+sum_fit$mod <- c("Model 1","Model 1","Model 2","Model 2","Model 3","Model 3","Model 4","Model 4")
+sum_fit$est <- c("Continous","Ordinal","Continous","Ordinal","Continous","Ordinal","Continous","Ordinal")
 sum_fit <- select(sum_fit,mod,est,everything())
 
-kable(sum_fit,digits = 3,format = "html",row.names = F,booktabs=T, caption = "Resumen estadísticos de ajuste ola 1") %>% kable_styling(full_width = F) %>% 
+sumtable01<- kable(sum_fit,digits = 3,format = "html",row.names = F,booktabs=T, caption = "Summary fit indices wave 01") %>%     kable_styling(full_width = F)  %>% 
       collapse_rows(columns = 1,valign = "middle")  %>%  
-  footnote(number = c("Modelo 1: fixed order by percepction/preference", 
-                      "Modelo 2: fixed order by topic (i.e: effort)", 
-                      "Modelo 3: Randomized order",
-                      "Modelo 4: Complete sample"))
+      footnote(number = c("Model 1: fixed order by percepction/preference", 
+                          "Model 2: fixed order by topic (i.e: effort)", 
+                          "Model 3: Randomized order",
+                          "Model 4: Complete sample"));sumtable01
+
+# Save sumamry table wave 01------------------------------#
+# as_image(sumtable01,file = "output/sumtablew01.png")
+save_kable(sumtable01,file = "output/tables/sumtablew01.html",self_contained = FALSE)
 ```
 
-## Análisis por Grupos múltiples
+## Other analyses
+
+### Multigroup CFA
+
+* The group specification correspond to the application modality of the scale.
 
@@
 
-fit4_c_mg <- cfa(model = model04,data = dat04,estimator="MLR",std.lv=FALSE,group = "grupo") # Continuo/ estimador ML Robust
-
+fit4_c_mg <- cfa(model = model04,data = dat04,estimator="MLR",std.lv=FALSE,group = "grupo") # Continuous/ estimator ML Robust
 fit4_o_mg <- cfa(model = model04,data = dat04,ordered = c("perc_effort","perc_talent",
@@
 #---------------------------------------------------#
-#summary(fit4_c_mg,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fit4_c_mg)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fit4_c_mg,what = "std",thresholds = FALSE, intercepts = FALSE, fade=FALSE, curve=2)
+kable(rbind(Continuous=fitmeasures(fit4_c_mg)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            Ordinal=  fitmeasures(fit4_o_mg)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
+      format = "markdown",digits = 3,caption = "Fit indices")
 
-#summary(fit4_o_mg,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fit4_o_mg)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fit4_o_mg,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE, curve=2)
+cnames <- c("Grupo","Factor","Indicator","Loading (MLR)","Loading (DWLS)")
+kable(left_join(x = standardizedsolution(fit4_c_mg) %>% filter(op=="=~") %>% select(group,lhs,rhs,est.std),
+                y = standardizedsolution(fit4_o_mg) %>% filter(op=="=~") %>% select(group,lhs,rhs,est.std),
+                c("group","lhs","rhs")),
+      format = "markdown",digits = 2,col.names = cnames,caption = "Factor loadings by group")
 ```
 
-Cosas por probar:
+### Second order CFA
 
-* V01+V02: cargas y fit
-* V01+V03: cargas y fit
-* V02+V03: cargas y fit
-
-# Modelo Factorial de Segundo Orden:
-
-## a. Factor de segundo orden por Percepción y Preferencia
+#### SO-CFA Perception and Preference
 
@@
 #---------------------------------------------------#
-#summary(fit4a_c,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fit4a_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fit4a_c,what = "std",thresholds = FALSE, intercepts = FALSE, fade=FALSE,curve=2)
+kable(rbind(Continuous=fitmeasures(fit4a_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            Ordinal= fitmeasures(fit4b_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
+      format = "markdown",digits = 3)
 
-#summary(fit4b_c,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fit4b_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fit4b_c,what = "std",thresholds = FALSE, intercepts = FALSE, fade=FALSE,curve=2)
+cnames <- c("Factor","Indicator","Loading (MLR)","Loading (DWLS)")
+kable(left_join(x = standardizedsolution(fit4a_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
+                y = standardizedsolution(fit4b_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
+                c("lhs","rhs")),
+      format = "markdown",digits = 2,col.names = cnames,caption = "Factor loadings")
 ```
 
-> **NOTA:** Cargas factoriales de segundo orden son bajas. El ajuste del modelo no es bueno.
+<!-- > **NOTE:** Cargas factoriales de segundo orden son bajas. El ajuste del modelo no es bueno. -->
 
-## b. Factor de segundo orden por Meritocracia y No Meritocracia
+> **NOTE:** Second order factor loads are low. The fit of the model is not good.
+
+#### SO-CFA Meritocratic y Non-meritocratic
 
@@
 #---------------------------------------------------#
-# #summary(fit4b_c,standardized=TRUE, fit.measures=TRUE)
-# fitmeasures(fit4b_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-# semPlot::semPaths(object = fit4b_c,what = "std",thresholds = FALSE, intercepts = FALSE, fade=FALSE, curve=2)
+kable(rbind(Ordinal=fitmeasures(fit4b_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
+      format = "markdown",digits = 3)
 
-#summary(fit4b_o,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fit4b_o)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fit4b_o,what = "std",thresholds = FALSE, intercepts = FALSE, fade=FALSE, curve=2)
+cnames <- c("Factor","Indicator","Loading (DWLS)")
+kable(standardizedsolution(fit4b_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
+      format = "markdown",digits = 2,col.names = cnames,caption = "Factor loadings")
 ```
 
-> **NOTA:** El modelo estimado con MLR con factores de segundo orden "meritocracia/ no meritocracia" no converge.
+<!-- > **NOTE:** El modelo estimado con MLR con factores de segundo orden "meritocracia/ no meritocracia" no converge.
 EL modelo estimado a través de DWLS converge, posee cargas muy bajas y una muy alta (merit~pref_merit).
-Fit es malo (RMSEA >.10)
+Fit es malo (RMSEA >.10) -->
 
-# Wave 02
+> **NOTE:**  The model estimated with MLR with second order factors "meritocracy / non meritocracy" does not converge.
+The model estimated through DWLS converges, has very low loads and a very high one (merit ~ pref_merit).
+Fit is bad (RMSEA> .10)
+
+# Wave 02 {.tabset .tabset-fade .tabset-pills}
+
+Each sample correspond to one of the three conditions generated by the randomized assignment.
+
+1. **Control group:** the group that has been assigned to the control group (placebo condition)
+
+2. **Poverty condition sample** the group that has been assigned to the Poverty treatment condition.
+
+3. **Inequality condition sample:** the group that has been assigned to the Inequality treatment condition.
+
+4. **Complete sample:** Control, Poverty and Inequality treatment.
 
@@
 
-## ATE indicadores meritocracia
+**ATE on observed indicators**
 
-```{r, out.width="75%", results='asis'}
+```{r, out.width="75%",results='asis'}
 m01<- lm(perc_effort_w2~treat,data = wide01)
@@
 
-> **NOTA**: Los tratamientos de información no se asocian de manera estadísticamente significativa con ninguno de los indicadores de la escala de meritocracia.
+<!-- > **NOTE**: Los tratamientos de información no se asocian de manera estadísticamente significativa con ninguno de los Indicatores de la escala de meritocracia.   -->
 
-## CFA Muestra completa (Control, Desigualdad y Pobreza)
+> **NOTE**: There are no statistically significant effects of the treatment conditions on the observed indicators of the meritocracy scale.
+
+## Control group
+
+```{r}
+w02con <- data02 %>% filter(FL_6_DO_Control==1) %>% select(starts_with("meritv03_p")) 
+```
 
 ```{r echo=FALSE,results='asis'}
-stargazer(dat01,type = "html")
+stargazer(w02con,type = "html",          
+          title = "Descriptive statistics",
+          font.size = "small",
+          median = TRUE,
+          header=FALSE,
+          digit.separator = "",
+          digits = 2)
 ```
@@
 ```{r}
-dat01cor<- select(dat01,meritv03_perc_effort,meritv03_perc_talent,
-                        meritv03_perc_wpart,meritv03_perc_netw,
-                        meritv03_pref_effort,meritv03_pref_talent,
-                        meritv03_pref_wpart,meritv03_pref_netw)
-cor01<-polychoric(dat01cor)
-corrplot(cor01$rho, type = "upper", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
+corw02_1<-polychoric(w02con)
+rownames(corw02_1$rho) <-c("1. pc_effort", "2. pc_talent","3. pc_wpart","4. pc_netw","5. pf_effort","6. pf_talent","7. pf_wpart", "8. pf_netw")
+colnames(corw02_1$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)")
+corrplot(corw02_1$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
 ```
@@
 
-fitw02_c <- cfa(model = model03,data = dat01,estimator="MLR",std.lv=FALSE) # Continuo/ estimador ML Robust
+fitw02a_c <- cfa(model = model03,data = w02con,estimator="MLR") # Continuous/ estimator ML Robust
+
+fitw02a_o <- cfa(model = model03,data = w02con,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
+                                                           "meritv03_perc_wpart","meritv03_perc_netw",
+                                                           "meritv03_pref_effort","meritv03_pref_talent",
+                                                           "meritv03_pref_wpart","meritv03_pref_netw"))
+#---------------------------------------------------#
+kable(rbind(Continuous=fitmeasures(fitw02a_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            Ordinal= fitmeasures(fitw02a_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
+      format = "markdown",digits = 3)
+
+cnames <- c("Factor","Indicator","Loading (MLR)","Loading (DWLS)")
+kable(left_join(x = standardizedsolution(fitw02a_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
+                y = standardizedsolution(fitw02a_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
+                c("lhs","rhs")),
+      format = "markdown",digits = 2,col.names = cnames)
+```
+
+## Poverty condition
+
+```{r}
+w02pob <- data02 %>% filter(FL_6_DO_Tratamiento==1) %>% select(starts_with("meritv03_p")) 
+```
+
+```{r echo=FALSE,results='asis'}
+stargazer(w02pob,type = "html",
+          title = "Descriptive statistics",
+          font.size = "small",
+          median = TRUE,
+          header=FALSE,
+          digit.separator = "",
+          digits = 2)
+```
+
+```{r}
+corw02_2<-polychoric(w02pob)
+rownames(corw02_2$rho) <-c("1. pc_effort", "2. pc_talent","3. pc_wpart","4. pc_netw","5. pf_effort","6. pf_talent","7. pf_wpart", "8. pf_netw")
+colnames(corw02_2$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)")
+corrplot(corw02_2$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
+```
+
+```{r,results='hold'}
+model03 <- '
+perc_merit=~meritv03_perc_effort+meritv03_perc_talent 
+perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw  
+pref_merit=~meritv03_pref_effort+meritv03_pref_talent 
+pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw'
+
+fitw02b_c <- cfa(model = model03,data = w02pob,estimator="MLR") # Continuous/ estimator ML Robust
+
+fitw02b_o <- cfa(model = model03,data = w02pob,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
+                                                           "meritv03_perc_wpart","meritv03_perc_netw",
+                                                           "meritv03_pref_effort","meritv03_pref_talent",
+                                                           "meritv03_pref_wpart","meritv03_pref_netw"))
+
+#---------------------------------------------------#
+kable(rbind(Continuous=fitmeasures(fitw02b_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            Ordinal= fitmeasures(fitw02b_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
+      format = "markdown",digits = 3)
+
+cnames <- c("Factor","Indicator","Loading (MLR)","Loading (DWLS)")
+kable(left_join(x = standardizedsolution(fitw02b_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
+                y = standardizedsolution(fitw02b_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
+                c("lhs","rhs")),
+      format = "markdown",digits = 2,col.names = cnames)
+```
+
+## Inequality condition
+
+```{r}
+w02des <- data02 %>% filter(FL_6_DO_Tratamiento_desigualdad==1) %>% select(starts_with("meritv03_p")) 
+```
+
+```{r echo=FALSE,results='asis'}
+stargazer(w02des,type = "html",
+          title = "Descriptive statistics",
+          font.size = "small",
+          median = TRUE,
+          header=FALSE,
+          digit.separator = "",
+          digits = 2)
+```
+
+```{r}
+corw02_3<-polychoric(w02des)
+rownames(corw02_3$rho) <-c("1. pc_effort", "2. pc_talent","3. pc_wpart","4. pc_netw","5. pf_effort","6. pf_talent","7. pf_wpart", "8. pf_netw")
+colnames(corw02_3$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)")
+corrplot(corw02_3$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
+```
+
+```{r,results='hold'}
+model03 <- '
+perc_merit=~meritv03_perc_effort+meritv03_perc_talent
+perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw  
+pref_merit=~meritv03_pref_effort+meritv03_pref_talent 
+pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw'
+
+fitw02c_c <- cfa(model = model03,data = w02des,estimator="MLR") # Continuous/ estimator ML Robust
+
+fitw02c_o <- cfa(model = model03,data = w02des,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
+                                                           "meritv03_perc_wpart","meritv03_perc_netw",
+                                                           "meritv03_pref_effort","meritv03_pref_talent",
+                                                           "meritv03_pref_wpart","meritv03_pref_netw"))
+#---------------------------------------------------#
+kable(rbind(Continuous=fitmeasures(fitw02c_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            Ordinal= fitmeasures(fitw02c_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
+      format = "markdown",digits = 3)
+
+cnames <- c("Factor","Indicator","Loading (MLR)","Loading (DWLS)")
+kable(left_join(x = standardizedsolution(fitw02c_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
+                y = standardizedsolution(fitw02c_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
+                c("lhs","rhs")),
+      format = "markdown",digits = 2,col.names = cnames)
+```
+
+## Complete sample
+
+```{r echo=TRUE,results='asis'}
+stargazer(dat01,type = "html",
+          title = "Descriptive statistics",
+          font.size = "small",
+          median = TRUE,
+          header=FALSE,
+          digit.separator = "",
+          digits = 2)
+```
+
+```{r}
+dat01cor<- select(dat01,meritv03_perc_effort,meritv03_perc_talent,
+                        meritv03_perc_wpart,meritv03_perc_netw,
+                        meritv03_pref_effort,meritv03_pref_talent,
+                        meritv03_pref_wpart,meritv03_pref_netw)
+cor01<-polychoric(dat01cor)
+rownames(cor01$rho) <-c("1. pc_effort", "2. pc_talent","3. pc_wpart","4. pc_netw","5. pf_effort","6. pf_talent","7. pf_wpart", "8. pf_netw")
+colnames(cor01$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)")
+corrplot(cor01$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
+```
+
+```{r,results='hold'}
+model03 <- '
+perc_merit=~meritv03_perc_effort+meritv03_perc_talent 
+perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw  
+pref_merit=~meritv03_pref_effort+meritv03_pref_talent 
+pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw'
+
+fitw02_c <- cfa(model = model03,data = dat01,estimator="MLR",std.lv=FALSE) # Continuous/ estimator ML Robust
 
@@
 #---------------------------------------------------#
-#summary(fitw02_c,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fitw02_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fitw02_c,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE,curve=2)
+kable(rbind(Continuous=fitmeasures(fitw02_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            Ordinal= fitmeasures(fitw02_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
+      format = "markdown",digits = 3)
 
-#summary(fitw02_o,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fitw02_o)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fitw02_o,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE,curve=2)
+cnames <- c("Factor","Indicator","Loading (MLR)","Loading (DWLS)")
+kable(left_join(x = standardizedsolution(fitw02_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
+                y = standardizedsolution(fitw02_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
+                c("lhs","rhs")),
+      format = "markdown",digits = 2,col.names = cnames)
 ```
 
-> **NOTA**: El ajuste de muestra completa en wave 02 con estimador MLR (robusto) es bastante bueno (CFI, RMSEA). El ajuste con DWLS es bastante bueno, CFI mejora, RMSEA aumenta un poco (0.59).
+> **NOTE**: El ajuste de muestra completa en wave 02 con estimator MLR (robusto) es bastante bueno (CFI, RMSEA). El ajuste con DWLS es bastante bueno, CFI mejora, RMSEA aumenta un poco (0.59).
 
-* **Recomendación**: es probable que el estimador MLR sea mejor que el ordinal.
-
-### Factor de segundo orden
+### Second order CFA
 
@@
 
-# mod2_merit_c <- cfa(model = modw2,data = dat01) # Continuo/ estimador ML Robust
+# mod2_merit_c <- cfa(model = modw2,data = dat01) # Continuous/ estimator ML Robust
 
@@
 #---------------------------------------------------#
-# #summary(mod2_merit_c,standardized=TRUE, fit.measures=TRUE)
-# fitmeasures(mod2_merit_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-# semPlot::semPaths(object = mod2_merit_c,what = "std",thresholds = FALSE, intercepts = FALSE)
+kable(rbind(Ordinal=fitmeasures(mod2_merit_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
+      format = "markdown",digits = 3)
 
-#summary(mod2_merit_o,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(mod2_merit_o)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = mod2_merit_o,what = "std",thresholds = FALSE, intercepts = FALSE,fade=FALSE, curve=2, rotation=2)
+cnames <- c("Factor","Indicator","Loading (DWLS)")
+kable(standardizedsolution(mod2_merit_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
+      format = "markdown",digits = 2,col.names = cnames)
 ```
 
-> * Factor de segundo orden usando MLR no converge.
-> * Factor de segundo orden usando DWLS converge, pero las cargas no son buenas y el ajuste tampoco.
+<!-- > * Factor de segundo orden usando MLR no converge. -->
+<!-- > * Factor de segundo orden usando DWLS converge, pero las cargas no son buenas y el ajuste tampoco. -->
 
-## CFA Muestra Control
+> * Second order factor using MLR does not converge.
+> * Second order factor using DWLS converges, but loads are not good and the fit is not.
+
+# Summary fit indices wave 02
 
 ```{r}
-w02con <- data02 %>% filter(FL_6_DO_Control==1) %>% select(starts_with("meritv03_p")) 
-```
-
-```{r echo=FALSE,results='asis'}
-stargazer(w02con,type = "html")
-```
-
-```{r}
-corw02_1<-polychoric(w02con)
-corrplot(corw02_1$rho, type = "upper", order = "hclust", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
-```
-
-```{r,results='hold'}
-model03 <- '
-perc_merit=~meritv03_perc_effort+meritv03_perc_talent 
-perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw  
-pref_merit=~meritv03_pref_effort+meritv03_pref_talent 
-pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw'
-
-fitw02a_c <- cfa(model = model03,data = w02con,estimator="MLR") # Continuo/ estimador ML Robust
-
-fitw02a_o <- cfa(model = model03,data = w02con,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
-                                                           "meritv03_perc_wpart","meritv03_perc_netw",
-                                                           "meritv03_pref_effort","meritv03_pref_talent",
-                                                           "meritv03_pref_wpart","meritv03_pref_netw"))
-
-#---------------------------------------------------#
-#summary(fitw02a_c,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fitw02a_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fitw02a_c,what = "std",thresholds = FALSE, intercepts = FALSE, curve=2)
-
-#summary(fitw02a_o,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fitw02a_o)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fitw02a_o,what = "std",thresholds = FALSE, intercepts = FALSE, curve=2)
-```
-
-## CFA Muestra Pobreza
-
-```{r}
-w02pob <- data02 %>% filter(FL_6_DO_Tratamiento==1) %>% select(starts_with("meritv03_p")) 
-```
-
-```{r echo=FALSE,results='asis'}
-stargazer(w02pob,type = "html")
-```
-
-```{r}
-corw02_2<-polychoric(w02pob)
-corrplot(corw02_2$rho, type = "upper", order = "hclust", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
-```
-
-```{r,results='hold'}
-model03 <- '
-perc_merit=~meritv03_perc_effort+meritv03_perc_talent 
-perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw  
-pref_merit=~meritv03_pref_effort+meritv03_pref_talent 
-pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw'
-
-fitw02b_c <- cfa(model = model03,data = w02pob,estimator="MLR") # Continuo/ estimador ML Robust
-
-fitw02b_o <- cfa(model = model03,data = w02pob,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
-                                                     "meritv03_perc_wpart","meritv03_perc_netw",
-                                                     "meritv03_pref_effort","meritv03_pref_talent",
-                                                     "meritv03_pref_wpart","meritv03_pref_netw"))
-
-#---------------------------------------------------#
-#summary(fitw02b_c,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fitw02b_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fitw02b_c,what = "std",thresholds = FALSE, intercepts = FALSE, curve=2)
-
-#summary(fitw02b_o,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fitw02b_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fitw02b_o,what = "std",thresholds = FALSE, intercepts = FALSE, curve=2)
-```
-
-## CFA Muestra Desigualdad
-
-```{r}
-w02des <- data02 %>% filter(FL_6_DO_Tratamiento_desigualdad==1) %>% select(starts_with("meritv03_p")) 
-```
-
-```{r echo=FALSE,results='asis'}
-stargazer(w02des,type = "html")
-```
-
-```{r}
-corw02_3<-polychoric(w02des)
-corrplot(corw02_3$rho, type = "upper", order = "hclust", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
-```
-
-```{r,results='hold'}
-model03 <- '
-perc_merit=~meritv03_perc_effort+meritv03_perc_talent
-perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw  
-pref_merit=~meritv03_pref_effort+meritv03_pref_talent 
-pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw'
-
-fitw02c_c <- cfa(model = model03,data = w02des,estimator="MLR") # Continuo/ estimador ML Robust
-
-fitw02c_o <- cfa(model = model03,data = w02des,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
-                                                           "meritv03_perc_wpart","meritv03_perc_netw",
-                                                           "meritv03_pref_effort","meritv03_pref_talent",
-                                                           "meritv03_pref_wpart","meritv03_pref_netw"))
-#---------------------------------------------------#
-#summary(fitw02c_c,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fitw02c_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fitw02c_c,what = "std",thresholds = FALSE, intercepts = FALSE)
-
-#summary(fitw02b_o,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fitw02c_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fitw02c_o,what = "std",thresholds = FALSE, intercepts = FALSE)
-```
-
-# Resumen ajuste ola 2
-```{r}
-sum_fitw02<- bind_rows(fitmeasures(fitw02_c)[ c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                       fitmeasures(fitw02_o)[ c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                       fitmeasures(fitw02a_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+sum_fitw02<- bind_rows(fitmeasures(fitw02a_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
                        fitmeasures(fitw02a_o)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
@@
                        fitmeasures(fitw02c_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                       fitmeasures(fitw02c_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")])
+                       fitmeasures(fitw02c_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+                       fitmeasures(fitw02_c)[ c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+                       fitmeasures(fitw02_o)[ c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],)
 
-sum_fitw02$mod <- c("Modelo 1","Modelo 1","Modelo 2","Modelo 2","Modelo 3","Modelo 3","Modelo 4","Modelo 4")
-sum_fitw02$est <- c("Continuo","Ordinal","Continuo","Ordinal","Continuo","Ordinal","Continuo","Ordinal")
+sum_fitw02$mod <- c("Model 1","Model 1","Model 2","Model 2","Model 3","Model 3","Model 4","Model 4")
+sum_fitw02$est <- c("Continuous","Ordinal","Continuous","Ordinal","Continuous","Ordinal","Continuous","Ordinal")
 sum_fitw02 <- select(sum_fitw02,mod,est,everything())
 
-kable(sum_fitw02,digits = 3,format = "html",row.names = F,booktabs=T, caption = "Resumen ajuste ola 2") %>% 
+kable(sum_fitw02,digits = 3,format = "html",row.names = F,booktabs=T, caption = "Summary fit indices wave 02") %>% 
   kable_styling(full_width = F) %>% 
   collapse_rows(columns = 1,valign = "middle") %>% 
-  footnote(number = c("Modelo 1: Muestra completa", 
-                      "Modelo 2: Muestra grupo Control", 
-                      "Modelo 3: Muestra grupo tratamiento pobreza",
-                      "Modelo 4: Muestra grupo tratamiento desigualdad"))
+  footnote(number = c("Model 1: Control group sample", 
+                      "Model 2: Poverty treatment condition sample",
+                      "Model 3: Inequality treatment condition sample",
+                      "Model 4: Complete sample"))
 ```
@@
 
-```{r}
-sjt.fa(data = dat01,show.comm = TRUE)
+```{r echo=TRUE,results='asis'}
+stargazer(dat01,type = "html",
+          title = "Descriptive statistics",
+          font.size = "small",
+          median = TRUE,
+          header=FALSE,
+          digit.separator = "",
+          digits = 2)
 ```
@@
 
-fitw03c_c <- cfa(model = model03,data = dat01,estimator="MLR") # Continuo/ estimador ML Robust
+fitw03c_c <- cfa(model = model03,data = dat01,estimator="MLR") # Continuous/ estimator ML Robust
 
@@
 #---------------------------------------------------#
-#summary(fitw03c_c,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fitw03c_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+kable(rbind(Continuous=fitmeasures(fitw03c_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+            Ordinal= fitmeasures(fitw03c_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
+      format = "markdown",digits = 3)
 
-#summary(fitw03c_o,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fitw03c_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
+cnames <- c("Factor","Indicator","Loading (MLR)","Loading (DWLS)")
+kable(left_join(x = standardizedsolution(fitw03c_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
+                y = standardizedsolution(fitw03c_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
+                c("lhs","rhs")),
+      format = "markdown",digits = 2,col.names = cnames)
 ```
 **summary**
-> * Ajuste usando estimador MLR  es bueno (CFI = 0.971, RMSEA = 0.052). El chi2 (robusto) es 62.421  
-> * Ajuste usando estimador DWLS es bueno (CFI = 0.974, RMSEA = 0.067). El chi2 (robusto) es 106.111
 
-# Resumen ajuste ola 3
+<!-- > * Ajuste usando estimator MLR  es bueno (CFI = 0.971, RMSEA = 0.052). El chi2 (robusto) es 62.421   -->
+<!-- > * Ajuste usando estimator DWLS es bueno (CFI = 0.974, RMSEA = 0.067). El chi2 (robusto) es 106.111 -->
+
+> * Adjustment using MLR estimator is good (CFI = 0.971, RMSEA = 0.052). The chi2 (robust) is 62,421
+> * Fit using DWLS estimator is good (CFI = 0.974, RMSEA = 0.067). The chi2 (robust) is 106,111
+
+# Summary fit indices wave 03
 
@@
 
-sum_fitw03$mod <- c("Modelo 1","Modelo 1")
-sum_fitw03$est <- c("Continuo","Ordinal")
+sum_fitw03$mod <- c("Model 1","Model 1")
+sum_fitw03$est <- c("Continuous","Ordinal")
 sum_fitw03 <- select(sum_fitw03,mod,est,everything())
 
-kable(sum_fitw03,digits = 3,format = "html",row.names = F,booktabs=T,caption = "Resumen ajuste ola 3") %>% 
+kable(sum_fitw03,digits = 3,format = "html",row.names = F,booktabs=T,caption = "Summary fit indices wave 03") %>% 
   kable_styling(full_width = F) %>% 
   collapse_rows(columns = 1,valign = "middle") %>% 
-  footnote(number = c("Modelo 1: Muestra completa"))
+  footnote(number = c("Model 1: Complete sample"))
 ```
-
-# Análisis longitudinal
-
-```{r}
-library(sjPlot)
-library(dplyr)
-library(lavaan)
-library(stargazer)
-library(corrplot)
-library(psych)
-library(knitr)
-library(kableExtra)
-load(file = "input/data/proc/datawide_123.RData")
-```
-
-## Correlación indicadores w01 y w02
-
-```{r}
-g01<- wide02 %>% filter(grupo==1) %>% select(starts_with("perc_"),starts_with("pref_")) # Fixed by perception/preference
-g02<- wide02 %>% filter(grupo==2) %>% select(starts_with("perc_"),starts_with("pref_")) # Fixed by topic (i.e: effort)
-g03<- wide02 %>% filter(grupo==3) %>% select(starts_with("perc_"),starts_with("pref_")) # Randomized
-```
-
-### Grupo Versión 01
-
-```{r, out.width="75%", results='asis'}
-stargazer(g01,type = "html", median = T,digits = 2)
-```
-
-```{r, out.width=c("75%","75%"), fig.cap=c("Percepción","Preferencia"), fig.align='center'}
-corwide01a<- polychoric(x = select(g01 ,starts_with(match = "perc_")),na.rm = T)
-corrplot(corwide01a$rho, method = "color" ,type = "upper", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
-
-corwide01b<- polychoric(x = select(g01 ,starts_with(match = "pref_")),na.rm = T)
-corrplot(corwide01b$rho, method = "color" ,type = "upper", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
-```
-
-### Grupo Versión 02
-
-```{r, out.width="75%", results='asis'}
-stargazer(g02,type = "html", median = T,digits = 2)
-```
-
-```{r, out.width=c("75%","75%"), fig.cap=c("Percepción","Preferencia"), fig.align='center'}
-corwide02a<- polychoric(x = select(g02 ,starts_with(match = "perc_")),na.rm = T)
-corrplot(corwide02a$rho, method = "color" ,type = "upper", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
-
-corwide02b<- polychoric(x = select(g02 ,starts_with(match = "pref_")),na.rm = T)
-corrplot(corwide02b$rho, method = "color" ,type = "upper", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
-```
-
-### Grupo Versión 03
-
-```{r, results='asis'}
-stargazer(g03,type = "html", median = T,digits = 2)
-```
-
-```{r, out.width=c("75%","75%"), fig.cap=c("Percepción","Preferencia"), fig.align='center'}
-corwide03a<- polychoric(x = select(g03 ,starts_with(match = "perc_")),na.rm = T)
-corrplot(corwide03a$rho, type = "upper", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE,tl.cex = 0.75)
-
-corwide03b<- polychoric(x = select(g03 ,starts_with(match = "pref_")),na.rm = T)
-corrplot(corwide03b$rho, method = "color" ,type = "upper", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
-```
-
-### Muestra completa
-
-```{r, results='asis'}
-stargazer(select(wide02 ,starts_with(match = "perc_"), starts_with("pref_")),type = "html", median = T,digits = 2)
-```
-
-```{r, out.width=c("75%","75%"), fig.cap=c("Percepción","Preferencia"), fig.align='center'}
-corwideA<- polychoric(x = select(wide02 ,starts_with(match = "perc_")),na.rm = T)
-corrplot(corwideA$rho, method = "color" ,type = "upper", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
-
-corwideB<- polychoric(x = select(wide02 ,starts_with(match = "pref_")),na.rm = T)
-corrplot(corwideB$rho, method = "color" ,type = "upper", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
-```
-
-## CFA wave01, wave02 y wave03
-
-```{r}
-model_long <- '
-# wave 01
-perc_merit1 =~perc_effort_w1+perc_talent_w1 
-perc_nmerit1=~perc_wpart_w1 +perc_netw_w1  
-pref_merit1 =~pref_effort_w1+pref_talent_w1 
-pref_nmerit1=~pref_wpart_w1 +pref_netw_w1
-# wave 02
-perc_merit2 =~perc_effort_w2+perc_talent_w2 
-perc_nmerit2=~perc_wpart_w2 +perc_netw_w2  
-pref_merit2 =~pref_effort_w2+pref_talent_w2 
-pref_nmerit2=~pref_wpart_w2 +pref_netw_w2
-# wave 03
-perc_merit3 =~perc_effort_w3+perc_talent_w3 
-perc_nmerit3=~perc_wpart_w3 +perc_netw_w3  
-pref_merit3 =~pref_effort_w3+pref_talent_w3 
-pref_nmerit3=~pref_wpart_w3 +pref_netw_w3'
-
-fitlong_c <- cfa(model = model_long,data = wide02,estimator="MLR") # Continuo/ estimador ML Robust
-
-indicadores <-c("perc_effort_w1","perc_talent_w1","perc_wpart_w1","perc_netw_w1",
-                "pref_effort_w1","pref_talent_w1","pref_wpart_w1","pref_netw_w1",
-                "perc_effort_w2","perc_talent_w2","perc_wpart_w2","perc_netw_w2",
-                "pref_effort_w2","pref_talent_w2","pref_wpart_w2","pref_netw_w2")
-
-fitlong_o <- cfa(model = model_long,data = wide02,ordered = indicadores)
-
-#---------------------------------------------------#
-#summary(fitlong_c,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fitlong_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fitlong_c,what = "std",thresholds = FALSE, intercepts = FALSE)
-
-#summary(fitlong_o,standardized=TRUE, fit.measures=TRUE)
-fitmeasures(fitlong_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-semPlot::semPaths(object = fitlong_o,what = "std",thresholds = FALSE, intercepts = FALSE,style = "mx")
-```
-
-```{r,fig.cap="Correlación entre variables latentes w01, w02 y w03"}
-cormat<- as.matrix(inspect(fitlong_o, "cor.lv"))
-corrplot(cormat, method = "color" ,type = "upper", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
-```
-
-# Invarianza
-
-1. Invarianza entre modalidades en ola 1,2 y 3 (between groups invariance)
-2. Invarianza entre tratamientos en ola 2(between groups invariance)
-
->**NOTA**: En vista de que no existe una librería que permita estimar invarianza longitudinal asumiendo indicadores ordinales, las primeras estimaciones las realizaré asumiendo indicadores continuos (estimator=ML).
-
-```{r}
-library(semTools)
-```
-
-```{r}
-load(file = "input/data/proc/datalong.RData")     # Datos formato long
-load(file = "input/data/proc/datawide_123.RData") # Datos formato long
-```
-
-```{r}
-dim(dat04)
-m_inv_modalidad <- 'perc_merit=~ perc_effort+perc_talent 
-                    perc_nmerit=~perc_wpart +perc_netw  
-                    pref_merit=~ pref_effort+pref_talent 
-                    pref_nmerit=~pref_wpart +pref_netw'
-
-indicators  <-c("perc_effort","perc_talent","perc_wpart","perc_netw",
-                "pref_effort","pref_talent","pref_wpart","pref_netw")
-
-fit.conf=cfa(m_inv_modalidad,data=dat04,group="grupo")
-fit.deb =cfa(m_inv_modalidad,data=dat04,group="grupo",group.equal=c("loadings"))
-fit.fuer=cfa(m_inv_modalidad,data=dat04,group="grupo",group.equal=c("loadings","intercepts"))
-fit.str =cfa(m_inv_modalidad,data=dat04,group="grupo",group.equal=c("loadings","intercepts","residuals"))
-
-fitmeasures(fit.conf)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-fitmeasures(fit.deb )[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-fitmeasures(fit.fuer)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-fitmeasures(fit.str )[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]
-anova(fit.conf, fit.deb, fit.fuer, fit.str)
-```
-
-```{r}
-configural <- '
-# wave 01
-perc_merit1 =~NA*perc_effort_w1+lambda1*perc_effort_w1+perc_talent_w1 
-perc_nmerit1=~NA*perc_wpart_w1 +lambda1*perc_wpart_w1 +perc_netw_w1  
-pref_merit1 =~NA*pref_effort_w1+lambda1*pref_effort_w1+pref_talent_w1 
-pref_nmerit1=~NA*pref_wpart_w1 +lambda1*pref_wpart_w1 +pref_netw_w1
-# wave 02
-perc_merit2 =~NA*perc_effort_w2+lambda1*perc_effort_w2+perc_talent_w2 
-perc_nmerit2=~NA*perc_wpart_w2 +lambda1*perc_wpart_w2 +perc_netw_w2  
-pref_merit2 =~NA*pref_effort_w2+lambda1*pref_effort_w2+pref_talent_w2 
-pref_nmerit2=~NA*pref_wpart_w2 +lambda1*pref_wpart_w2 +pref_netw_w2
-# wave 03
-perc_merit3 =~NA*perc_effort_w3+lambda1*perc_effort_w3+perc_talent_w3 
-perc_nmerit3=~NA*perc_wpart_w3 +lambda1*perc_wpart_w3 +perc_netw_w3  
-pref_merit3 =~NA*pref_effort_w3+lambda1*pref_effort_w3+pref_talent_w3 
-pref_nmerit3=~NA*pref_wpart_w3 +lambda1*pref_wpart_w3 +pref_netw_w3
-
-# INTERCEPTS: Por cada variable latente, debo fijar el intercepto del primer indicador.
-# wave 01--------------------------
-perc_effort_w1 ~ i1*1   # Percepcion merito        
-perc_talent_w1 ~ 1      #      
-perc_wpart_w1 ~ i1*1    # Percepcion no-merito   
-perc_netw_w1 ~ 1        #
-pref_effort_w1 ~ i1*1   # Preferencia merito    
-pref_talent_w1 ~ 1      #  
-pref_wpart_w1 ~ i1*1    # Preferencia no-merito   
-pref_netw_w1 ~ 1        #
-
-# wave 02--------------------------
-perc_effort_w2 ~ i1*1   # Percepcion merito                    
-perc_talent_w2 ~ 1      #               
-perc_wpart_w2 ~ i1*1    # Percepcion no-merito              
-perc_netw_w2 ~ 1        #       
-pref_effort_w2 ~ i1*1   # Preferencia merito                
-pref_talent_w2 ~ 1      #           
-pref_wpart_w2 ~ i1*1    # Preferencia no-merito              
-pref_netw_w2 ~ 1        #
-
-# wave 03--------------------------
-perc_effort_w3 ~ i1*1   # Percepcion merito                     
-perc_talent_w3 ~ 1      #                            
-perc_wpart_w3 ~ i1*1    # Percepcion no-merito                
-perc_netw_w3 ~ 1        #                  
-pref_effort_w3 ~ i1*1   # Preferencia merito                   
-pref_talent_w3 ~ 1      #                        
-pref_wpart_w3 ~ i1*1    # Preferencia no-merito               
-pref_netw_w3 ~ 1        #
-
-# UNIQUE VARIANCES and COVARIANCES.
-
-# LATENT VARIABLES MEANS: Para cada variable latente fijar la media
-# Fix T1 common factor mean to zero
-perc_merit1 ~ 0*1
-# Freely estimate T2-T3 common factor means
-perc_merit2 ~ 1
-perc_merit3 ~ 1
-
-perc_nmerit1 ~ 0*1
-perc_nmerit2 ~ 1
-perc_nmerit3 ~ 1
-
-pref_merit1 ~ 0*1
-pref_merit2 ~ 1
-pref_merit3 ~ 1
-
-pref_nmerit1 ~ 0*1
-pref_nmerit2 ~ 1
-pref_nmerit3 ~ 1
-
-# LATENT VARIABLE VARIANCES AND COVARIANCES: correlaciones entre variables latentes
-
-'
-
-fit_conf <- cfa(model = configural,data =wide02)
-summary(fit_conf, standardized=TRUE)
-```
-
-* Invarianza Longitudinal en lavaan
-https://quantdev.ssri.psu.edu/tutorials/intro-basics-longitudinal-measurement-invariance
-
-* https://users.ugent.be/~yrosseel/lavaan/multiplegroup6Dec2012.pdf
-
-* Invarianza con indicadores categóricos
-https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5121102/pdf/nihms757737.pdf
-
-* Parametrización tetha: 
-https://groups.google.com/forum/#!topic/lavaan/nfdatPgLLhc
-
-* Colapsing categories:
-https://www.tandfonline.com/doi/abs/10.1080/10705511.2018.1547640?journalCode=hsem20

### 2020-04-06 production/prod_analysis-cfa.rmd
update durante reunion

@@
 ---
-title: "Analysis"
+title: "Analysis PPM-S"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 
-# Analysis PPM-S
-
 This section, documents all the procedures related to the data analysis and measurement models.
+(description)
 
 # Wave 01 {.tabset .tabset-fade .tabset-pills}
+
+Design: this wave ....
 
@@
 
-4. **Version 04:** Complete sample (N=`r dim(dat04)[1]`)
+4. **Complete Sample:** Complete sample (N=`r dim(dat04)[1]`)
 
@@
 
-## Version 04 (N=`r dim(dat04)[1]`)
+## Complete sample (N=`r dim(dat04)[1]`)
 

### 2020-04-07 production/prod_prep-data-A.rmd
data preparation

@@
----
-title: "Data preparation"
+title: ---"Data preparation"
 date: "`r format(Sys.time(), '%d %B, %Y')`"

### 2020-04-07 docs/paper.rmd
data_proc
dfSummary
traduccion
 + paper

--- 2020-03-20 docs/paper.rmd
@@
 output:
-  html_document:
-    number_sections: true
-    toc: true
-    toc_float: true
+  html_document: default
 linkcolor: blue
-bibliography: "../input/bib/EconomiaMoral.bib"
-css: "../input/bib/style.css"
-
+bibliography: input/bib/EconomiaMoral.bib
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = TRUE)
+knitr::opts_chunk$set(echo = TRUE)
 ```
+<div style="text-align: justify">
 
@@
 Más relevante aún se vuelve la legitimidad de la desigualdad en un contexto mundial de crecientes protestas sociales en contra de la misma.
-En el caso Chileno, estas movilizaciones han alcanzado una gran envergadura, relevando la importancia de evaluar los niveles de legitimidad de la desigualdad existente, a la luz de distintas ideologías que podrían justificarla.
+En el caso Chileno, estas movilizaciones han alcanzado una gran envergadura, relevando la importancia de evaluar los niveles de legitimidad de la desigualdad existente, a la luz de distintas ideologías que según la teoría podrían justificar.
 Dentro de estas ideologías que legitiman la desigualdad, la literatura sobre la meritocracia, presenta un notorio crecimiento [@MadeiraPrimesConsequencesSystematic2019].
 Tradicionalmente la meritocracia es entendida como una forma de distribución social de recursos donde el mérito individual (i.e.
-el esfuerzo y el talento) determinan la distribucion de los privilegios [@breenClassInequalityMeritocracy1999;@youngRiseMeritocracy1994; @OhsfeldtBeliefsInequalityAmericans1987; @Jostrolestereotypingsystemjustification1994], sin la interferencia de factores estructurales como la herencia o los contactos [@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995; @yairMeritocracy2007; @landWeSatTable2006;@youngRiseMeritocracy1994].
+el esfuerzo y el talento) determinan la distribución de los privilegios [@breenClassInequalityMeritocracy1999;@youngRiseMeritocracy1994; @OhsfeldtBeliefsInequalityAmericans1987; @Jostrolestereotypingsystemjustification1994], sin la interferencia de factores estructurales como la herencia o los contactos [@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995; @yairMeritocracy2007; @landWeSatTable2006;@youngRiseMeritocracy1994].
 
-Los estudios relativos a la meritocracia se pueden clasificar entre estudios de caracter objetivo y estudios decaracter subjetivo.
-Los estudios de caracter objetivo, suelen dedicarse a la observar de la movilidad social, evaluando críticamente hasta qué punto las sociedades son efectivamente meritocráticas y su distribución no está influida por contactos o ventajas externas [@alonDiversityOpportunityShifting2007; @dubetRepensarJusticiaSocial2017; @eriksonIntergenerationalClassMobility2010; @rosanvallonSocieteEgaux2011; @barozetValorHistoricoPituto2006; @mcnameeMeritocracyMyth2004; @goldthorpeMythEducationbasedMeritocracy2003] Los segundos de caracter subjetivo, ponen el enfacis no en la observacion de la distribucion material de los recursos sino en las ideas y discursos en torno a la distribucion social, mirando la meritocracia desde el enfoque de la economia moral.
+Los estudios relativos a la meritocracia se pueden clasificar entre estudios de carácter objetivo y estudios de carácter subjetivo.
+Los estudios de carácter objetivo, suelen dedicarse a la observar de la movilidad social, evaluando críticamente hasta qué punto las sociedades son efectivamente meritocráticas y su distribución no está influida por contactos o ventajas externas [@alonDiversityOpportunityShifting2007; @dubetRepensarJusticiaSocial2017; @eriksonIntergenerationalClassMobility2010; @rosanvallonSocieteEgaux2011; @barozetValorHistoricoPituto2006; @mcnameeMeritocracyMyth2004; @goldthorpeMythEducationbasedMeritocracy2003] Los segundos de carácter subjetivo, ponen el énfasis no en la observación de la distribución material de los recursos sino en las ideas y discursos en torno a la distribución social, mirando la meritocracia desde el enfoque de la economía moral.
 Desde esta perspectiva de la economía moral, estudian las creencias meritocráticas y su relación con distintos factores, como el ingreso, el contexto social o los prejuicios a grupos vulnerables [@KunovichSystemsDistributionSense2007;@NewmanFalseConsciousnessClass2015; @ReynoldsPerceptionsmeritocracyland2014; @Duru-BellatWhoMeritocracyIndividual2012].
@@
 ## Antecedentes generales, la economía moral y la meritocracia
-Las ciencias sociales, desde la perspectiva de la economía moral, han enriquecido la comprensión de las relaciones económicas, aportando a la lógica racional imperante en las teorias economicas, aspectos subjetivos como las normas, los procesos históricos y culturales que estructuran las prácticas tanto las económicas [@mauMoralEconomyWelfare2004; @sachwehMoralEconomyInequality2012; @polanyiGranTransformacionCritica1997; @zelizerEconomicLivesHow2010 ;@fourcadeMoralViewsMarket2007] como la produccion y reproduccion de la desigualdad [@lamontWhatMissingCultural2014].En este marco teorico, se ha  planteado que las creencias meritocráticas pueden legitimar la desigualdad existente [@hadjarMeritokratieAlsLegitimationsprinzip2008].
-En la misma linea, se ha planteado que la creencia en la meritocracia, que adjudica causas individuales a la pobreza, se relaciona con la atribucion de prejuicios hacia grupos de bajo estatus [@MadeiraPrimesConsequencesSystematic2019] y con ideas conservadoras de autoritarismo social [@AzevedoNeoliberalIdeologyJustification2019].
+Las ciencias sociales, desde la perspectiva de la economía moral, han enriquecido la comprensión de las relaciones económicas, aportando a la lógica racional imperante en las teorías económicas, aspectos subjetivos como las normas, los procesos históricos y culturales que estructuran las prácticas tanto las económicas [@mauMoralEconomyWelfare2004; @sachwehMoralEconomyInequality2012; @polanyiGranTransformacionCritica1997; @zelizerEconomicLivesHow2010 ;@fourcadeMoralViewsMarket2007] como la produccion y reproduccion de la desigualdad [@lamontWhatMissingCultural2014].En este marco teórico, se ha  planteado que las creencias meritocráticas pueden legitimar la desigualdad existente [@hadjarMeritokratieAlsLegitimationsprinzip2008].
+En la misma línea, se ha planteado que la creencia en la meritocracia, que adjudica causas individuales a la pobreza, se relaciona con la atribución de prejuicios hacia grupos de bajo estatus [@MadeiraPrimesConsequencesSystematic2019] y con ideas conservadoras de autoritarismo social [@AzevedoNeoliberalIdeologyJustification2019].
 
@@
 Por percepciones meritocráticas refiere al modo en que el entrevistado describe la realidad.
-Las preferencias refieren a un aspecto normativo, es decir, que tan de acuerdo se está con que la meritocracia es un buen sistema de distribucion.
+Las preferencias refieren a un aspecto normativo, es decir, que tan de acuerdo se está con que la meritocracia es un buen sistema de distribución.
 Esta distinción es fundamental, puesto que ambas variables poseen bastante relación entre sí, por lo cual es de suponer que trabajar con estas variables sin controlarlas entre sí nos puede llevar a tener ideas causales erróneas por espuriedad.
@@
 la ventaja de tener padres ricos) es contrario a creer en el rol del esfuerzo.
-De este modo, las creencias no meritocraticas son cosnideradas es el extremo negativo del continuo, mientras creer que la trayectoria personal depende del esfuerzo y el talento del individuo, corresponde al lado positivo del continuo.
+De este modo, las creencias no meritocráticas son consideradas es el extremo negativo del continuo, mientras creer que la trayectoria personal depende del esfuerzo y el talento del individuo, corresponde al lado positivo del continuo.
 Dicho planteamiento deriva teoría que señala que una sociedad meritocrática ideal se define por la distribución de riquezas en base a esfuerzo y no a fuerzas estructurales [@yairMeritocracy2007; @landWeSatTable2006].
-Aunque dicho planteamiento es correcto para evaluar que tan meritocratica es una sociedad, no necesariamente este pensamiento se refleja en el modo en como se escructuran los discursos en los imaginarios de la poblacion.
-De hecho, otros investigadores han evaluado la dimensionalidad de las creencias meritocráticas, generando evidencia sobre la multidimensionalidad de este constructo, dando cuenta de que aspectos meritocraticos y no meritocraticos no forman parte de la misma dimension dentro de la subjetividad de los encuestados.
+Aunque dicho planteamiento es correcto para evaluar qué tan meritocrática es una sociedad, no necesariamente este pensamiento se refleja en el modo en cómo se estructuran los discursos en los imaginarios de la población.
+De hecho, otros investigadores han evaluado la dimensionalidad de las creencias meritocráticas, generando evidencia sobre la multidimensionalidad de este constructo, dando cuenta de que aspectos meritocráticos y no meritocráticos no forman parte de la misma dimensión dentro de la subjetividad de los encuestados.
 Entre de las investigaciones que han evaluado dicha dimensionalidad se encuentran los trabajos de @KunovichSystemsDistributionSense2007 y @ReynoldsPerceptionsmeritocracyland2014.
@@
 Este estudio cuenta con una encuesta panel de 3 olas.
-En la primera de los olas, se incorporo la escala en distintos ordenes.
-El primer orden, pregunta primero por todas las percepciones y despues por todas las preferencias, el segundo orden, pregunta seguidamente por percepciones y preferencias en torno a un mismo tema y, el tercer orden incorpora los items de manera aleatoria.
-En la segunda y tercera ola, se utilizo el orden
+En la primera de los olas, se incorporó la escala en distintos órdenes.
+El primer orden, pregunta primero por todas las percepciones y después por todas las preferencias, el segundo orden, pregunta seguidamente por percepciones y preferencias en torno a un mismo tema y, el tercer orden incorpora los ítems de manera aleatoria.
+En la segunda y tercera ola, se utilizó el orden
 
@@
 
+Todos los modelos independiente de los órdenes obtuvieron un ajuste moderado con CFI superiores a .95 y RMSEA inferiores a .8, por contraparte ni un modelo logró un chi-square no significativo, aunque tanto el modelo aleatorio como el primero obtuvieron un adecuado chi-square ratio menor a 3.
 El primero de los órdenes fue el que mejor ajuste obtuvo (CFI=0.998, TLI= 0.995,RMSEA=0.037, χ2(df=14)=28,03, p = 0.014), seguido por el orden aleatorio de los ítems (CFI=0.992, TLI=0.984,RMSEA=0.051, χ2(df=14)=39.09, p < 0.001).
 Por su parte la escala ordenada por temáticas genera un efecto framing en el cual la relación entre las percepciones y las preferencias parecen sobreestimadas, lo cual afecta el ajuste, siendo este deficitario, con un χ2 ratio mayor a tres y un RMSEA que se puede encontrar por sobre .8 según el intervalo de confianza correspondiente(CFI=0.984, TLI=0.968,RMSEA=.071, χ2(df=14)=64.156, p < 0.001).
+
+A partir de los análisis de invarianza se puede decir que los distintos órdenes no son invariantes en términos configuracionales [@vandenbergReviewSynthesisMeasurement2000].
+El modelo que incorpora la diferenciación de los grupos en la estimación del modelo, este presenta un ajuste moderado similar al de los modelos por separados, aunque posee un ratio de chi-square deficiente por lo que se descarta la invarianza.
+Esta varianza configuracional se debe a la sobre estimación que se genera en el orden uno de las correlaciones entre los indicadores de un mismo factor los cuales aparecen juntos para el encuestado, mientras que en el modelo 2 se presentan separados, disminuyendo la correlación entre los ítem de un mismo factor.Podríamos decir que las correlaciones entre los ítems son sub o sobre-estimadas según el orden en el que aparezcan.
+Coherentemente el modelo aleatorio tiene un ajuste intermedio entre ambos.
+Considerando la ventaja del orden, respecto aleatorio despeja la posibilidad del efecto framing, es decir, de que el resultado del modelo se deba al orden de las preguntas, se ha desidido seguir utilizando el orden 3.
 
@@
 | **p**     |           .000          |             .000            |       .000      |
-| **χ2/gl** |          65.38          |             4.90            |       3.35      |
+| **χ2/gl** |          65.38          |             4.90            |       3.32      |
 
-En la segunda ola se utilizó el primer orden de los ítems ya que este fue el que presentó un mejor resultado.
-EL modelo teorico propuesto de cuatro factores ajustó, como se observa en la tabla 4, mejor que el modelo de contraste de 1 factor.
+EL modelo teórico propuesto de cuatro factores ajustó, como se observa en la tabla 4, mejor que el modelo de contraste de 1 factor.
 El modelo teórico ajustó de manera relativamente adecuada, pues muestra indicadores óptimos para CFI= 0.987, TLI = 0.975 y RMSEA=.041, aunque posee indicadores deficientes para la prueba χ2(df=14)=67.6, p-value=.000.
@@
 
-* agregar cuales fueron las cargas sujeridas.
+* agregar cuáles fueron las cargas sugeridas.
 
@@
 
-Esta escala cuenta con validez de constructo, ya que los distintos índices formados según factor, correlacionan coherentemente la escala "Get a head", utilizada por el programa internacional de encuestas sociales (ISSP), y con la escala de valores igualitarios utilizada por la encuesta internacional de valores.
-**(indicar validez del constructo y bibliografía que  lo use).**
+Esta escala cuenta con validez de constructo, ya que los distintos índices formados según factor correlacionan coherentemente con la escala "Get a head", utilizada por el programa internacional de encuestas sociales (ISSP), y con la escala de valores igualitarios utilizada por la encuesta internacional de valores.
 
@@
 | Preferencia <br>No Meritocratica |   -0.191***   | 0.083***   |
-
-**dar vuelta la tabla, get primero**
 
@@
 
-**contradicción a considerar: la validez por criterio, en la cual se evalúa la relación con una variable validada, depende de la calidad de la medida.
-nosotros lo relacionamos con lo que dice que esta mal.get-
-
-# Invarianza longitudinal
-
 # Discusión
-
- - el texto incorpora una discusión que parte de las cargas cruzadas, dándoles sentido (que es un hallazgo propio del texto) y dando cuenta de cómo esta estructura es equivalente en varias encuestas, posteriormente justifica la existencia de cierto desajuste en base a una hipótesis plausible (diferencia entre autoaplicado y por teléfono)
-
--  habla de la validez predictiva, habla de bajas correlaciones y plantea una posibilidad de un nuevo estudio con un cambio
-
-- plantea una limitación, su estudio no es experimental y podría estar influido por más por el tipo de encuesta que se contestó.
 
@@
 
-En un contexto de crecientes movilizaciones sociales y creciente deslegitimidad de los gobiernos, es necesario contar con instrumentos válidos para la medición de las creencias meritocráticas, las cuales, suelen funcionar como legitimadoras de la desigualdad.
-Por ello, es necesario contar con instrumentos breves y validados, para incorporar la medición de esta variable en otros estudios relacionados con temas de legitimidad.
-Utilizando los datos recolectados en una encuesta panel representativa de Chile, este artículo expone la elaboración y la validación de una escala de creencias meritocráticas de 4 factores: percepciones meritocráticas, percepciones no-meritocráticas, preferencias meritocráticas y preferencias no meritocráticas.
-La escala de percepciones y preferencias meritocráticas es un instrumento breve, el cual permite medir de manera relativamente sencilla las creencias meritocráticas, las cuales son fundamentales para abordar temas de legitimación de la desigualdad, preferencias redistributivas o prejuicios hacia grupos vulnerables.
+## Fiabilidad
+
+### Correlación entre los mismos items segun distintas olas.
+
+| **Pensando en la sociedad chilena, <br>¿En qué medida se encuentra usted de acuerdo o en desacuerdo<br>con cada una de las siguientes afirmaciones?** |  <br>Ola 1 - Ola 2  |  <br>Ola 2 - Ola 3 | <br>Ola 1 - Ola 3 |
+|-------------------------------------------------------------------------------------------------------------------------------------------------------|:-------------------:|:------------------:|:-----------------:|
+| Quienes más se esfuerzan logran obtener mayores <br>recompensas que quienes se esfuerzan menos.
+|         0.33        |        0.38        |        0.30       |
+| Quienes poseen más talento logran obtener mayores<br>recompensas que quienes poseen menos talento.
+|         0.30        |        0.28        |        0.27       |
+| Quienes tienen padres ricos logran salir adelante.
+|         0.33        |        0.36        |        0.38       |
+| Quienes tienen buenos contactos logran salir adelante.
+|         0.30        |        0.30        |        0.31       |
+| Quienes más se esfuerzan deberían obtener mayores<br>recompensas que quienes se esfuerzan menos.
+|         0.27        |        0.32        |        0.22       |
+| Quienes poseen más talento deberían obtener mayores<br>recompensas que quienes poseen menos talento.
+|         0.31        |        0.39        |        0.31       |
+| Está bien que quienes tienen padres ricos salganadelante.
+|         0.40        |        0.41        |        0.39       |
+| Está bien que quienes tienen buenos contactos salgan adelante.
+|         0.39        |        0.41        |        0.42       |
+
+ Pese a que existen bajas correlaciones, hay evidencia para apoyar la hipótesis de estabilidad de la escala a lo largo de las mediciones.
+Las correlaciones entre los mismos ítems en las distintas olas son de un tamaño pequeño a moderado según los parámetros de [@cohenStatisticalPowerAnalysis2013].
+Estas bajas  correlaciones indican inestabilidad en las respuestas de los encuestados, no obstante existe evidencia para señalar una invarianza fuerte de las escalas en las distintas mediciones.
+En primer lugar, podemos decir que no existe diferencia entre el ajuste de los tres  nos indica que no existe una diferencia significativa entre los ajustes al modelo de las tres olas, lo cual nos señala la existencia de invarianza configuracional.
+
+         Df   AIC   BIC  Chisq Chisq diff Df diff Pr(>Chisq)    
+fit.conf 42 35745 36222 110.91                                  
+fit.deb  50 35738 36172 119.39      8.485       8     0.3876    
+fit.fuer 58 35732 36124 129.94     10.546       8     0.2288    
+fit.str  74 35772 36079 202.02     72.081      16  4.298e-09 ***
+
+El análisis de invarianza longitudinal muestra resultados moderadamente buenos. hay equivalencia configuracional,  métrica, pero no escalar escalar
+
+<div/>
 

### 2020-04-08 production/prod_prep-data-A.rmd

--- 2020-04-05 production/prod_prep-data-A.rmd
@@
 
-# Data preparation
+# Data preparation according order of the items
 
@@
 
-### Preparation of group 1: order according to the factors.
+## Preparation of group 1: order according to the factors.
 
@@
 
-### Preparation of group 2: order according to topics (i.e Effort).
+## Preparation of group 2: order according to topics (i.e Effort).
 
@@
 
-### Preparation of group 3: randomized order.
+## Preparation of group 3: randomized order.
 
@@
 
+## save data
 ```{r}
@@
 
+load data
 ```{r}
@@
 w01$Intro <- set_labels(w01$Intro,  labels = c("I ACCEPT take part in this investigation","I NOT ACCEPT take part in this investigation."))
-
 ```
@@
 w02$Intro <- set_labels(w02$Intro,  labels = c("I ACCEPT take part in this investigation","I NOT ACCEPT take part in this investigation."))
-
 ```
 
-Frequency of respondents who Accept and Finish wave 1.
+Frequency of respondents who Accept and Finish wave 2.
 ```{r}
@@
 
-Frequency of respondents who Accept and Finish wave 1.
+Frequency of respondents who Accept and Finish wave 2.
 ```{r }
@@
 
-## 
 ```{r}

### 2020-04-10 production/prod_analysis-cfa.rmd
update-1004-2020

@@
 title: "Analysis PPM-S"
-date: "`r format(Sys.time(), '%d %B, %Y')`"
+date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
 css: "../input/bib/style.css"
@@
     toc: true
-    toc_float: true
+    toc_float:
+        collapsed: false
     toc_depth: 2
@@
 
-This section, documents all the procedures related to the data analysis and measurement models.
-(description)
-
 # Wave 01 {.tabset .tabset-fade .tabset-pills}
 
-Design: this wave ....
+On the first wave of the study we conducted the application of the Perceptions and Preferences Meritocracy Scale (PPM-S), in which we decided to test three different modalities for the application of the scale.
+The design of the study considered a sample of 2200 cases in Chile, using quotas for sex, age and educational level.
+The most important feature of this study consists in that we randomly assigned the participants to three different groups, in which the subjects complete one of the three combination of the questions.
+The key feature of these combinations relies on the display order of the questions regarding to each indicator of the scale.
+On this respect, we have three subsamples that correspond to each application modality that can be analysed separately, but also all together as a full sample.
+
+The scale has eight indicators, divided in four questions for perception and four for preferences, which are also divided in meritocratic and non-meritocratic.
+The meritocratic indicators refers to topics such as effort and talent, and the non-meritocratic refers to family wealth and networks.
+The three combinations relies in how these indicators are displayed to the respondent.
+The first combination corresponds to the group that answered the questions that were sorted by the Perceptions/Preferences dimension.
+the second combination corresponds to the group in which the questions are displayed following the a topic-sorted way, this means that the effort, talent, rich family and networks questions for perception and preference were displayed together.
+The last and third group corresponds to a completely randomized display order of the questions.
+
+The rationale of the design relies on the possibility to test how different display modalities could affect the hypothesized factor structure of the scale.
+On this regard, we conducted the analysis on two stages.
+On the first stage we tested the hypothesized factor structure on the three groups corresponding to each application modality and the full sample using a Confirmatory Factor Analysis (CFA).
+On the second stage, we conducted a Multi-group confirmatory factor analysis (MGCFA) using the groups created by the randomization.
+Furthermore, we explore the possibility of a second order factor structure, in which we tested two models.
+The first alternative corresponds to a second-order Confirmatory Factor Analysis assuming the Perception/Preference dimension as a second order factor.
+On the other hand, we tested a model in which we assume that the second order factor structure follows the Meritocratic/Non-meritocratic constructs.
+
+---
 
 ```{r eval=FALSE, include=FALSE}
-rmarkdown::render(input = "production/prod_analysis-cfa.Rmd",output_format = "html_document")
+rmarkdown::render(input = "production/prod_analysis-cfa.Rmd",output_format = "html_document"); browseURL(url = "production/prod_analysis-cfa.html")
 ```
@@
 library(lavaan)
+library(semPlot)
 library(stargazer)
@@
 **Load data**
+
 ```{r}
+load(file = "input/data/proc/dat01.RData")
+load(file = "input/data/proc/dat02.RData")
+load(file = "input/data/proc/dat03.RData")
+load(file = "input/data/proc/dat04.RData")
+```
+
+```{r eval=FALSE, include=FALSE}
 data01 <- sjlabelled::read_spss(path = "input/data/original/Estudio_3_ola1.sav",verbose = FALSE)
-
 dat01 <- data01 %>% filter(Intro==1,Finished==1) %>% select(starts_with("meritv01"))   %>% na.omit()
@@
 dat04[dat04==0] <- NA
+dat04 <- dat04 %>% select(starts_with(match = "perc_"),starts_with(match = "pref_"),grupo) %>% na.omit()
 ```
@@
 |-----------------------------|------------------------|
-| 1. Percepción esfuerzo      | Perception effort      |
-| 2. Percepción talento       | Perception talent      |
-| 3. Percepción familia rica  | Perception rich family |
-| 4. Percepción redes         | Perception networks    |
-| 5. Preferencia esfuerzo     | Preference effort      |
-| 6. Preferencia talento      | Preference talent      |
-| 7. Preferencia familia rica | Preference rich family |
-| 8. Preferencia redes        | Preference networks    |
+| A. Percepción esfuerzo      | Perception effort      |
+| B. Percepción talento       | Perception talent      |
+| C. Percepción familia rica  | Perception rich family |
+| D. Percepción redes         | Perception networks    |
+| E. Preferencia esfuerzo     | Preference effort      |
+| F. Preferencia talento      | Preference talent      |
+| G. Preferencia familia rica | Preference rich family |
+| H. Preferencia redes        | Preference networks    |
 Table: Mode 01: Fixed order by perception
@@
           digits = 2,
-          covariate.labels = c("1. pc.effort", "2. pc.talent","3. pc.wpart","4. pc.netw","5. pf.effort","6. pf.talent","7. pf.wpart", "8. pf.netw"))
-
+          covariate.labels = c("A. pc.effort", "B. pc.talent","C. pc.wpart","D. pc.netw","E. pf.effort","F. pf.talent","G. pf.wpart", "H. pf.netw"))
 ```
@@
 cor01<-polychoric(dat01)
-rownames(cor01$rho) <- c("1. pc_effort", "2. pc_talent","3. pc_wpart","4. pc_netw","5. pf_effort","6. pf_talent","7. pf_wpart", "8. pf_netw")
-colnames(cor01$rho) <- c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)")
+rownames(cor01$rho) <- c("A. pc.effort", "B. pc.talent","C. pc.wpart","D. pc.netw","E. pf.effort","F. pf.talent","G. pf.wpart", "H. pf.netw")
+colnames(cor01$rho) <- c("(A)","(B)","(C)","(D)","(E)","(F)","(G)", "(H)")
 corrplot(cor01$rho,method = "color" ,type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
@@
 ```{r}
-model01 <- 'perc_merit=~meritv01_perc_effort+meritv01_perc_talent
-            perc_nmerit=~meritv01_perc_wpart+meritv01_perc_netw
-            pref_merit=~meritv01_pref_effort+meritv01_pref_talent
-            pref_nmerit=~meritv01_pref_wpart+meritv01_pref_netw'
+source(file = "production/prod_polychoric.R")
+corr.test(cor01$rho)
+sjt.corr(data = cor01$rho,triangle = "lower",digits = 2)
+sjt.corr(data = dat01,triangle = "lower",digits = 2,var.labels = " ")
+
+corr <- psych::polychoric(data,  na.rm = TRUE)
+
+```
+
+```{r}
+model01 <- 'perc_merit=~perc_effort+perc_talent
+            perc_nmerit=~perc_wpart+perc_netw
+            pref_merit=~pref_effort+pref_talent
+            pref_nmerit=~pref_wpart+pref_netw'
 
@@
 fit1_o <- cfa(model = model01,data = dat01,
-            ordered = c("meritv01_perc_effort","meritv01_perc_talent",
-                        "meritv01_perc_wpart","meritv01_perc_netw",
-                        "meritv01_pref_effort","meritv01_pref_talent",
-                        "meritv01_pref_wpart","meritv01_pref_netw"))
+            ordered = c("perc_effort","perc_talent",
+                        "perc_wpart","perc_netw",
+                        "pref_effort","pref_talent",
+                        "pref_wpart","pref_netw"))
 
@@
 |-----------------------------|------------------------|
-| 1. Percepción esfuerzo      | Perception effort      |
-| 2. Preferencia esfuerzo     | Preference effort      |
-| 3. Percepción talento       | Perception talent      |
-| 4. Preferencia talento      | Preference talent      |
-| 5. Percepción familia rica  | Perception rich family |
-| 6. Preferencia familia rica | Preference rich family |
-| 7. Percepción redes         | Perception networks    |
-| 8. Preferencia redes        | Preference networks    |
+| A. Percepción esfuerzo      | Perception effort      |
+| E. Preferencia esfuerzo     | Preference effort      |
+| B. Percepción talento       | Perception talent      |
+| F. Preferencia talento      | Preference talent      |
+| C. Percepción familia rica  | Perception rich family |
+| G. Preferencia familia rica | Preference rich family |
+| D. Percepción redes         | Perception networks    |
+| H. Preferencia redes        | Preference networks    |
 Table: Mode 02: Fixed order, by topic.
@@
           digits = 2,
-          covariate.labels = c("1. pc.effort","2. pf.effort","3. pc.talent","4. pf.talent","5. pc.wpart","6. pf.wpart","7. pc.netw","8 pc.netw"))
+          covariate.labels = c("A. pc.effort","E. pf.effort","B. pc.talent","F. pf.talent","C. pc.wpart","G. pf.wpart","D. pc.netw","H. pc.netw"))
 ```
@@
 cor02<-polychoric(dat02)
-rownames(cor02$rho) <-c("1. pc_effort", "2. pc_talent","3. pc_wpart","4. pc_netw","5. pf_effort","6. pf_talent","7. pf_wpart", "8. pf_netw")
-colnames(cor02$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)")
+rownames(cor02$rho) <-c("A. pc.effort","E. pf.effort","B. pc.talent","F. pf.talent","C. pc.wpart","G. pf.wpart","D. pc.netw","H. pc.netw")
+colnames(cor02$rho) <-c("(A)", "(E)","(B)","(F)","(C)","(G)","(D)", "(H)")
 corrplot(cor02$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
@@
 ```{r,results='hold'}
-model02 <- 'perc_merit=~meritv02_perc_effort+meritv02_perc_talent
-            perc_nmerit=~meritv02_perc_wpart+meritv02_perc_netw
-            pref_merit=~meritv02_pref_effort+meritv02_pref_talent
-            pref_nmerit=~meritv02_pref_wpart+meritv02_pref_netw'
+model02 <- 'perc_merit=~perc_effort+perc_talent
+            perc_nmerit=~perc_wpart+perc_netw
+            pref_merit=~pref_effort+pref_talent
+            pref_nmerit=~pref_wpart+pref_netw'
 
@@
 
-fit2_o <- cfa(model = model02,data = dat02,ordered = c("meritv02_perc_effort","meritv02_perc_talent",
-                                                       "meritv02_perc_wpart"  ,"meritv02_perc_netw",
-                                                       "meritv02_pref_effort" ,"meritv02_pref_talent",
-                                                       "meritv02_pref_wpart"  ,"meritv02_pref_netw"))
+fit2_o <- cfa(model = model02,data = dat02,ordered = c("perc_effort","perc_talent",
+                                                       "perc_wpart"  ,"perc_netw",
+                                                       "pref_effort" ,"pref_talent",
+                                                       "pref_wpart"  ,"pref_netw"))
 
@@
           digits = 2,
-          covariate.labels = c("1. pc.effort", "2. pc.talent","3. pc.wpart","4. pc.netw","5. pf.effort","6. pf.talent","7. pf.wpart", "8. pf.netw"))
+          covariate.labels = c("A. pc.effort", "B. pc.talent","C. pc.wpart","D. pc.netw","E. pf.effort","F. pf.talent","G. pf.wpart", "H. pf.netw"))
 ```
@@
 cor03<-polychoric(dat03)
-rownames(cor03$rho) <-c("1. pc_effort", "2. pc_talent","3. pc_wpart","4. pc_netw","5. pf_effort","6. pf_talent","7. pf_wpart", "8. pf_netw")
-colnames(cor03$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)")
+rownames(cor03$rho) <-c("A. pc.effort", "B. pc.talent","C. pc.wpart","D. pc.netw","E. pf.effort","F. pf.talent","G. pf.wpart", "H. pf.netw")
+colnames(cor03$rho) <-c("(A)","(B)","(C)","(D)","(E)","(F)","(G)", "(H)")
 corrplot(cor03$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
@@
 model03 <- '
-perc_merit=~meritv03_perc_effort+meritv03_perc_talent 
-perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw  
-pref_merit=~meritv03_pref_effort+meritv03_pref_talent 
-pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw'
+perc_merit=~perc_effort+perc_talent
+perc_nmerit=~perc_wpart+perc_netw
+pref_merit=~pref_effort+pref_talent
+pref_nmerit=~pref_wpart+pref_netw'
 
@@
 
-fit3_o <- cfa(model = model03,data = dat03,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
-                                                       "meritv03_perc_wpart","meritv03_perc_netw",
-                                                       "meritv03_pref_effort","meritv03_pref_talent",
-                                                       "meritv03_pref_wpart","meritv03_pref_netw"),std.lv=FALSE)
+fit3_o <- cfa(model = model03,data = dat03,ordered = c("perc_effort","perc_talent",
+                                                       "perc_wpart","perc_netw",
+                                                       "pref_effort","pref_talent",
+                                                       "pref_wpart","pref_netw"),std.lv=FALSE)
 #---------------------------------------------------#
@@
 model03.tau <- '
-perc_merit=~ r1*meritv03_perc_effort+r1*meritv03_perc_talent 
-perc_nmerit=~r2*meritv03_perc_wpart+ r2*meritv03_perc_netw  
-pref_merit=~ r3*meritv03_pref_effort+r3*meritv03_pref_talent 
-pref_nmerit=~r4*meritv03_pref_wpart+ r4*meritv03_pref_netw'
+perc_merit=~ r1*perc_effort+r1*perc_talent
+perc_nmerit=~r2*perc_wpart+ r2*perc_netw
+pref_merit=~ r3*pref_effort+r3*pref_talent
+pref_nmerit=~r4*pref_wpart+ r4*pref_netw'
 
@@
 
-fit3_o.tau <- cfa(model = model03.tau,data = dat03,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
-                                                               "meritv03_perc_wpart","meritv03_perc_netw",
-                                                               "meritv03_pref_effort","meritv03_pref_talent",
-                                                               "meritv03_pref_wpart","meritv03_pref_netw"), std.lv=FALSE)
+fit3_o.tau <- cfa(model = model03.tau,data = dat03,ordered = c("perc_effort","perc_talent",
+                                                               "perc_wpart","perc_netw",
+                                                               "pref_effort","pref_talent",
+                                                               "pref_wpart","pref_netw"), std.lv=FALSE)
 #---------------------------------------------------#
@@
 cor04<-polychoric(select(dat04,perc_effort,perc_talent,perc_wpart,perc_netw,pref_effort,pref_talent,pref_wpart,pref_netw))
-rownames(cor04$rho) <-c("1. pc_effort", "2. pc_talent","3. pc_wpart","4. pc_netw","5. pf_effort","6. pf_talent","7. pf_wpart", "8. pf_netw")
-colnames(cor04$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)")
+rownames(cor04$rho) <-c("A. pc.effort", "B. pc.talent","C. pc.wpart","D. pc.netw","E. pf.effort","F. pf.talent","G. pf.wpart", "H. pf.netw")
+colnames(cor04$rho) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)", "(H)")
 corrplot(cor04$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
@@
 
-# Summary fit indices wave 01
+# Summary
 
+```{r semPlot, fig.cap="Measurement model PPM-S", fig.height=8, fig.width=12}
+model01 <- 'perc_merit=~perc_effort+perc_talent
+            perc_nmerit=~perc_wpart+perc_netw
+            pref_merit=~pref_effort+pref_talent
+            pref_nmerit=~pref_wpart+pref_netw'
+
+plot01 <- cfa(model = model01,data = dat01,estimator="MLR",std.lv=TRUE)
+
+nodeNames <-c("Perception effort","Perception talent","Perception rich family","Perception networks",
+              "Preference effort","Preference talent","Preference rich family","Preference networks",
+              "Percepction meritocratic","Percepction non meritocratic","Preference meritocratic","Preference non meritocratic")
+
+semPaths(plot01,
+         style = "lisrel",
+         fade=FALSE,
+         edge.color = "black",
+         nCharNodes = 0,
+         curvePivot = FALSE,
+         curve = 2,
+         rotation = 4,
+         layout = "tree3",
+         cardinal = "lat cov",
+         legend.cex = 0.6,
+         label.cex =1,
+         reorder = FALSE,
+         nodeNames = nodeNames,
+         nodeLabels = c("A","B","C","D","E","F","G","H","1","2","3","4"),
+         label.font = 6,
+         edge.label.font= 1,
+         asize = 3, #arrow size
+         edge.width =1,
+         sizeMan=5,
+         sizeLat = 10)
+```
+
+## Fit indices wave 01
 ```{r}
@@
 sum_fit$mod <- c("Model 1","Model 1","Model 2","Model 2","Model 3","Model 3","Model 4","Model 4")
-sum_fit$est <- c("Continous","Ordinal","Continous","Ordinal","Continous","Ordinal","Continous","Ordinal")
+sum_fit$est <- c("MLR","DWLS","MLR","DWLS","MLR","DWLS","MLR","DWLS")
 sum_fit <- select(sum_fit,mod,est,everything())
+colnames <- c("Model","Estimator","$\\chi^2$","df","CFI","CFI.sca","CFI.rob","RMSEA","RMSEA.sca","RMSEA.rob")
 
-sumtable01<- kable(sum_fit,digits = 3,format = "html",row.names = F,booktabs=T, caption = "Summary fit indices wave 01") %>%     kable_styling(full_width = F)  %>% 
+sumtable01<- kable(sum_fit,digits = 3,format = "html",row.names = F,booktabs=T, caption = "Summary fit indices wave 01",col.names = colnames,escape = FALSE) %>%
+  kable_styling(full_width = F)  %>%
       collapse_rows(columns = 1,valign = "middle")  %>%
@@
 
-```{r echo=FALSE,results='asis'}
+```{r results='asis'}
 stargazer(w02con,type = "html",
@@
 
-```{r echo=FALSE,results='asis'}
+```{r results='asis'}
 stargazer(w02pob,type = "html",
@@
 
-```{r echo=FALSE,results='asis'}
+```{r results='asis'}
 stargazer(w02des,type = "html",
@@
 
-```{r echo=TRUE,results='asis'}
+```{r results='asis'}
 stargazer(dat01,type = "html",
@@
 ```{r}
-sum_fitw02<- bind_rows(fitmeasures(fitw02a_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                       fitmeasures(fitw02a_o)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                       fitmeasures(fitw02b_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                       fitmeasures(fitw02b_o)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                       fitmeasures(fitw02c_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                       fitmeasures(fitw02c_c)[c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                       fitmeasures(fitw02_c)[ c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                       fitmeasures(fitw02_o)[ c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],)
+sum_fitw02<- bind_rows(fitmeasures(fitw02a_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+                       fitmeasures(fitw02a_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+                       fitmeasures(fitw02b_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+                       fitmeasures(fitw02b_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+                       fitmeasures(fitw02c_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+                       fitmeasures(fitw02c_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+                       fitmeasures(fitw02_c)[ c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
+                       fitmeasures(fitw02_o)[ c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],)
 
 sum_fitw02$mod <- c("Model 1","Model 1","Model 2","Model 2","Model 3","Model 3","Model 4","Model 4")
-sum_fitw02$est <- c("Continuous","Ordinal","Continuous","Ordinal","Continuous","Ordinal","Continuous","Ordinal")
+sum_fitw02$est <- c("MLR","DWLS","MLR","DWLS","MLR","DWLS","MLR","DWLS")
 sum_fitw02 <- select(sum_fitw02,mod,est,everything())
+colnames <- c("Model","Estimator","$\\chi^2$","df","CFI","CFI.sca","CFI.rob","RMSEA","RMSEA.sca","RMSEA.rob")
 
-kable(sum_fitw02,digits = 3,format = "html",row.names = F,booktabs=T, caption = "Summary fit indices wave 02") %>% 
+kable(sum_fitw02,digits = 3,format = "html",row.names = F,booktabs=T, caption = "Summary fit indices wave 02", col.names = colnames, escape = F) %>%
   kable_styling(full_width = F) %>%

### 2020-04-10 production/prod_prep-data-A.rmd
update-1004-2020

@@
 ```{r eval=FALSE, include=FALSE}
-rmarkdown::render(input = "production/prep-data-A.Rmd",output_format = "html_document",output_dir = ".")
+rmarkdown::render(input = "production/prod_prep-data-A.Rmd",output_format = "html_document",output_dir = ".")
 ```
@@
 dat04$perc_netw   <- rowSums(dat04[,c(matches(match = "perc_netw"  ,vars = names(dat04)))],na.rm = TRUE)
-
 dat04$pref_effort <- rowSums(dat04[,c(matches(match = "pref_effort",vars = names(dat04)))],na.rm = TRUE)
@@
 In the first order (group 1), questions of the same factor are then presented to them, in the second order (group 2) they were separated and in the third order (group 3) at random."
-
 dat04$perc_effort <- set_labels(dat04$perc_effort,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
@@
 less."
-
 dat04$perc_talent <- set_labels(dat04$perc_talent,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
@@
 less talent."
-
 dat04$perc_wpart <- set_labels(dat04$perc_wpart,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
@@
 ahead."
-
 dat04$perc_netw <- set_labels(dat04$perc_netw,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
@@
 ahead."
-
 dat04$pref_effort <- set_labels(dat04$pref_effort,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
@@
 they try less."
-
 dat04$pref_talent <- set_labels(dat04$pref_talent,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
@@
 less talent."
-
 dat04$pref_wpart <- set_labels(dat04$pref_wpart,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
@@
 contacts get ahead."
-
 dat04$pref_netw <- set_labels(dat04$pref_netw,  labels = c("strongly disagree", "disagreement", "Neither agree nor disagree","Agree", "totally agree"))
@@
 ```{r}
-dat01 <- dat04 %>% filter(grupo==1) %>%  na.omit()
+dat01 <- dat04 %>% filter(grupo==1) %>% select(-grupo) %>%  na.omit() 
 ```
@@
 ```{r}
-dat03 <- dat04 %>% filter(grupo==3) %>%  na.omit()
+dat03 <- dat04 %>% filter(grupo==3) %>% select(-grupo) %>%  na.omit()
 ```
@@
 If you do not want to participate in this study, press the alternative I NOT ACCEPT take part in this investigation."
-
 w01$Intro <- set_labels(w01$Intro,  labels = c("I ACCEPT take part in this investigation","I NOT ACCEPT take part in this investigation."))
@@
 w01c$perc_netw_w1   <- rowSums(w01c[,c(matches(match = "perc_netw"  ,vars = names(w01c)))],na.rm = TRUE)
-
 w01c$pref_effort_w1 <- rowSums(w01c[,c(matches(match = "pref_effort",vars = names(w01c)))],na.rm = TRUE)
@@
 If you do not want to participate in this study, press the alternative I NOT ACCEPT take part in this investigation."
-
 w02$Intro <- set_labels(w02$Intro,  labels = c("I ACCEPT take part in this investigation","I NOT ACCEPT take part in this investigation."))
@@
 w02b$ID <- stringr::str_split_fixed(w02b$ticket, "_", 4)[,1]
-
 ```
@@
 w02c <- w02b
-
 w02c$treat <- NA
@@
 w02c$treat[w02c$FL_6_DO_Tratamiento_desigualdad==1] <- 3
-
 w02c$treat <- factor(w02c$treat,levels = c(1,2,3),labels = c("control","pobreza","desiguald"))
-
 w02c <- w02c %>% rename(perc_effort_w2=meritv03_perc_effort,
@@
                         pref_netw_w2  =meritv03_pref_netw)
-
 w02c <- w02c %>% select(perc_effort_w2,perc_talent_w2,perc_wpart_w2,perc_netw_w2,
@@
 w03c <- w03b
-
 w03c <- w03c %>% rename(perc_effort_w3=meritv03_perc_effort,
@@
                         sexo_w3=sexo,edad_w3=edad,edcep_w3=edcep, ID)
-
 #-----------------------------------------------------------------------#
@@
 library(data.table)
-
 wide <- setDT(wide02)
-
 long01<- melt(wide,variable.name = "wave",
@@
                              "sexo","edad","edcep"))
-
 ```

### 2020-04-10 production/prod_polychoric.R

@@ -1,0 +1,1066 @@
+library(sjPlot)
+poly<- function (data, na.deletion = c("listwise", "pairwise"), corr.method = c("pearson", 
+                                                                         "spearman", "kendall"), title = NULL, var.labels = NULL, 
+          wrap.labels = 40, show.p = TRUE, p.numeric = FALSE, fade.ns = TRUE, 
+          val.rm = NULL, digits = 3, triangle = "both", string.diag = NULL, 
+          CSS = NULL, file = NULL,encoding=NULL, use.viewer = TRUE, 
+          remove.spaces = TRUE) 
+{
+  opt <- getOption("p_zero")
+  if (is.null(opt) || opt == FALSE) {
+    p_zero <- ""
+  }
+  else {
+    p_zero <- "0"
+  }
+  na.deletion <- match.arg(na.deletion)
+  corr.method <- match.arg(corr.method)
+  encoding <- get.encoding(encoding, data)
+  if (is.null(triangle)) {
+    triangle <- "both"
+  }
+  else if (triangle == "u" || triangle == "upper") {
+    triangle <- "upper"
+  }
+  else if (triangle == "l" || triangle == "lower") {
+    triangle <- "lower"
+  }
+  else triangle <- "both"
+  if (is.null(var.labels) && is.data.frame(data)) {
+    var.labels <- sjlabelled::get_label(data, def.value = colnames(data))
+  }
+  if (corr.method != "pearson" && corr.method != "spearman" && 
+      corr.method != "kendall") {
+    stop("argument 'corr.method' must be one of: pearson, spearman or kendall")
+  }
+  if (is.matrix(data)) {
+    corr <- data
+    cpvalues <- NULL
+  }
+  else {
+    if (na.deletion == "listwise") {
+      data <- stats::na.omit(data)
+      # corr <- stats::cor(data, method = corr.method)
+      corr <- psych::polychoric(data,  na.rm = TRUE)$rho
+    }
+    else {
+      corr <- stats::cor(data, method = corr.method, use = "pairwise.complete.obs")
+      # corr <- psych::polychoric(data,  na.rm = TRUE)
+    }
+    computePValues <- function(df) {
+      cp <- c()
+      for (i in 1:ncol(df)) {
+        pv <- c()
+        for (j in 1:ncol(df)) {
+          test <- suppressWarnings(stats::cor.test(df[[i]], 
+                                                   df[[j]], alternative = "two.sided", method = corr.method))
+          
[…]

### 2020-04-10 docs/paper.rmd
update-1004-2020
* nuevo texto en prod_analysys
* ahora correo todo junto el sitio, es decir, se vincula el proceo de data_prep con analysis.
* arreglé algunos problemas con el código de paper.rmd

@@
 output:
-  html_document: default
+    html_document:
+        number_sections: true
+        toc: true
+        toc_float: true
+        smart: false
 linkcolor: blue
-bibliography: input/bib/EconomiaMoral.bib
+bibliography: ../input/bib/EconomiaMoral.bib
 ---
@@
 
-## Introducción
+# Introducción
 

### 2020-04-10 production/prod_prep-data-A.rmd
update-1004-2020
* nuevo texto en prod_analysys
* ahora correo todo junto el sitio, es decir, se vincula el proceo de data_prep con analysis.
* arreglé algunos problemas con el código de paper.rmd

@@
         toc: true
-    toc_float: false
+        smart: false
         toc_depth: 2
@@
 ```{r eval=FALSE, include=FALSE}
-rmarkdown::render(input = "production/prod_prep-data-A.Rmd",output_format = "html_document",output_dir = ".")
+rmarkdown::render(input = "production/prod_prep-data-A.Rmd",output_format = "html_document",output_dir = "production/");browseURL(url = "production/prod_prep-data-A.html")
 ```

### 2020-04-11 production/prod_analysis-cfa.rmd

@@
 corrplot(cor01$rho,method = "color" ,type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
-```
-
-```{r}
-source(file = "production/prod_polychoric.R")
-corr.test(cor01$rho)
-sjt.corr(data = cor01$rho,triangle = "lower",digits = 2)
-sjt.corr(data = dat01,triangle = "lower",digits = 2,var.labels = " ")
-
-corr <- psych::polychoric(data,  na.rm = TRUE)
-
 ```
@@
 # Summary
+
+## Correlations
+
+```{r compare corr, echo=TRUE, out.width=c('33%', '33%', '33%'),fig.cap="Mode 01, 02 and 03",fig.show='hold'}
+corrplot(cor01$rho,method = "color" ,type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
+corrplot(cor02$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
+corrplot(cor03$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
+```
+
+## Measurement model
 

### 2020-04-11 production/prod_analysis-long_invariance.rmd

@@
 ---
+
+```{r eval=FALSE, include=FALSE}
+rmarkdown::render(input = "production/prod_analysis-long_invariance.Rmd",output_format = "html_document"); browseURL(url = "production/prod_analysis-long_invariance.html")
+```
+
 # Longitudinal analysis

### 2020-04-11 docs/index.Rmd
Review changes and add github activity feed to index.

@@
         smart: false
+        includes:
+          in_header: head_github-activity.html
 ---
@@
 
+### Reproducibility and documentation
+
+The repository of the project in Github is structured according to the [IPO reproducibility protocol](https://juancarloscastillo.github.io/ipo/).
+
+<br>
+
+## [Github activity feed](https://github.com/juancarloscastillo/github-activity)
+
+<div id="feed"></div>
+
+<br>
+
 ## References

### 2020-04-12 production/prod_analysis-long_invariance.rmd
update-12-04-2020
* correción de inconsistencias entre prod_data y long_invariance.
* las bases de datos producidas por el .rmd de procesamiento no eran las mismas que en long_invariance.
* Se modificó el código para vincular ambos flujos.

@@
 
+```{r include=FALSE}
+knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = TRUE)
+knitr::opts_knit$set(root.dir = "../")
+options(scipen=999)
+rm(list=ls())
+options(knitr.kable.NA = '')
+```
+
 # Longitudinal analysis
@@
 library(kableExtra)
+```
+
+```{r}
 load(file = "input/data/proc/datawide_123.RData")
@@
 load(file = "input/data/proc/datawide_123.RData") # Datos formato wide 3 olas
+load(file = "input/data/proc/dat04.RData")
 

### 2020-04-12 docs/prod_polychoric.R

@@ -1,1066 +1,0 @@
-library(sjPlot)
-poly<- function (data, na.deletion = c("listwise", "pairwise"), corr.method = c("pearson", 
-                                                                         "spearman", "kendall"), title = NULL, var.labels = NULL, 
-          wrap.labels = 40, show.p = TRUE, p.numeric = FALSE, fade.ns = TRUE, 
-          val.rm = NULL, digits = 3, triangle = "both", string.diag = NULL, 
-          CSS = NULL, file = NULL,encoding=NULL, use.viewer = TRUE, 
-          remove.spaces = TRUE) 
-{
-  opt <- getOption("p_zero")
-  if (is.null(opt) || opt == FALSE) {
-    p_zero <- ""
-  }
-  else {
-    p_zero <- "0"
-  }
-  na.deletion <- match.arg(na.deletion)
-  corr.method <- match.arg(corr.method)
-  encoding <- get.encoding(encoding, data)
-  if (is.null(triangle)) {
-    triangle <- "both"
-  }
-  else if (triangle == "u" || triangle == "upper") {
-    triangle <- "upper"
-  }
-  else if (triangle == "l" || triangle == "lower") {
-    triangle <- "lower"
-  }
-  else triangle <- "both"
-  if (is.null(var.labels) && is.data.frame(data)) {
-    var.labels <- sjlabelled::get_label(data, def.value = colnames(data))
-  }
-  if (corr.method != "pearson" && corr.method != "spearman" && 
-      corr.method != "kendall") {
-    stop("argument 'corr.method' must be one of: pearson, spearman or kendall")
-  }
-  if (is.matrix(data)) {
-    corr <- data
-    cpvalues <- NULL
-  }
-  else {
-    if (na.deletion == "listwise") {
-      data <- stats::na.omit(data)
-      # corr <- stats::cor(data, method = corr.method)
-      corr <- psych::polychoric(data,  na.rm = TRUE)$rho
-    }
-    else {
-      corr <- stats::cor(data, method = corr.method, use = "pairwise.complete.obs")
-      # corr <- psych::polychoric(data,  na.rm = TRUE)
-    }
-    computePValues <- function(df) {
-      cp <- c()
-      for (i in 1:ncol(df)) {
-        pv <- c()
-        for (j in 1:ncol(df)) {
-          test <- suppressWarnings(stats::cor.test(df[[i]], 
-                                                   df[[j]], alternative = "two.sided", method = corr.method))
-          
[…]

### 2020-04-12 docs/paper.Rmd
Update paper

--- 2020-04-07 docs/paper.rmd
@@
 ---
-title: "Perceptions and Preferences for Meritocracy Scale (PPM-S)"
+title: "Perceptions and Preferences for Meritocracy"
 linestretch: '1.5'
@@
 output:
-  html_document: default
+    html_document:
+        number_sections: true
+        toc: true
+        toc_float: true
+        smart: false
 linkcolor: blue
-bibliography: input/bib/EconomiaMoral.bib
+bibliography:
+  - ../input/bib/EconomiaMoral.bib
+  - ../input/bib/meritocracy.bib
 ---
@@
 
-## Introducción
+# Introducción
 
-Tanto las organizaciones internacionales, como parte del mundo académico, presentan una creciente preocupación por el aumento de la desigualdad social y los efectos negativos que esta posee en torno a la legitimidad de nuestras sociedades [ONU,2016; @schlozmanUnheavenlyChorusUnequal2012; @CastilloMeritocraciadesigualdadeconomica2019].
-Más relevante aún se vuelve la legitimidad de la desigualdad en un contexto mundial de crecientes protestas sociales en contra de la misma.
-En el caso Chileno, estas movilizaciones han alcanzado una gran envergadura, relevando la importancia de evaluar los niveles de legitimidad de la desigualdad existente, a la luz de distintas ideologías que según la teoría podrían justificar.
-Dentro de estas ideologías que legitiman la desigualdad, la literatura sobre la meritocracia, presenta un notorio crecimiento [@MadeiraPrimesConsequencesSystematic2019].
-Tradicionalmente la meritocracia es entendida como una forma de distribución social de recursos donde el mérito individual (i.e.
-el esfuerzo y el talento) determinan la distribución de los privilegios [@breenClassInequalityMeritocracy1999;@youngRiseMeritocracy1994; @OhsfeldtBeliefsInequalityAmericans1987; @Jostrolestereotypingsystemjustification1994], sin la interferencia de factores estructurales como la herencia o los contactos [@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995; @yairMeritocracy2007; @landWeSatTable2006;@youngRiseMeritocracy1994].
+La desigualdad económica se ha vuelto un tema que genera creciente preocupación y malestar alrededor del mundo.
+Esto se ha expresado en una serie de protestas como la emblemática "occupy wall street" el año 2011, así como también en una serie de análisis críticos respecto del desarrollo del capitalismo y sus consecuencias [@piketty_Capital_2014].
+En este contexto, el estudio de las visiones, preferencias y percepciones respecto de la desigualdad han adquirido relevancia en las ciencias sociales, en temas como las preferencias redistributivas [@alesina_preferences_2011; @Dimick2018ModelsOtherRegardingPreferences] la legitimación de la desigualdad económica [@schroder_is_2017] y el funcionamiento de la meritocracia [@duru-bellat_whos_2012; @mijs_unfulfillable_2016; @reynolds_perceptions_2014].
 
-Los estudios relativos a la meritocracia se pueden clasificar entre estudios de carácter objetivo y estudios de carácter subjetivo.
-Los estudios de carácter objetivo, suelen dedicarse a la observar de la movilidad social, evaluando críticamente hasta qué punto las sociedades son efectivamente meritocráticas y su distribución no está influida por contactos o ventajas externas [@alonDiversityOpportunityShifting2007; @dubetRepensarJusticiaSocial2017; @eriksonIntergenerationalClassMobility2010; @rosanvallonSocieteEgaux2011; @barozetValorHistoricoPituto2006; @mcnameeMeritocracyMyth2004; @goldthorpeMythEducationbasedMeritocracy2003] Los segundos de carácter subjetivo, ponen el énfasis no en la observación de la distribución material de los recursos sino en las ideas y discursos en torno a la distribución social, mirando la meritocracia desde el enfoque de la economía moral.
-Desde esta perspectiva de la economía moral, estudian las creencias meritocráticas y su relación con distintos factores, como el ingreso, el contexto social o los prejuicios a grupos vulnerables [@KunovichSystemsDistributionSense2007;@NewmanFalseConsciousnessClass2015; @ReynoldsPerceptionsmeritocracyland2014; @Duru-BellatWhoMeritocracyIndividual2012].
-Pese a que el concepto de creencias meritocráticas es fundamental para esta segunda línea de trabajo, las distintas escalas utilizadas en esta literatura no poseen suficientes evidencias de validez a su favor, lo que releva la importancia de aportar con propuestas para medir este constructo.
+En general, la meritocracia se define como un sistema de distribución de recursos y recompensas basados en el mérito individual, que en su concepción original es una suma de talento y esfuerzo [@young_rise_1962] y que pone en un lugar secundario interferencia de factores estructurales como la herencia o los contactos [@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995; @yairMeritocracy2007; @landWeSatTable2006;@youngRiseMeritocracy1994].
+Una serie de estudios han realizado críticas a la realización de este estándar moral de distribución, planteando que es una promesa incumplida dada la influencia preponderante de otros elementos más allá del mérito en el estatus individual [@arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019].
+Por otro lado, desde la psicología social y la sociología se han estudiado las características y consecuencias de las creencias en la meritocracia, en general basados en la hipótesis que mayor creencia en la meritocracia lleva a una mayor legitimación de las desigualdades [@hadjar_meritokratie_2008; @MadeiraPrimesConsequencesSystematic2019].
 
-Para estudiar las creencias meritocráticas se han utilizado múltiples escalas que suponen una estructura unifactorial de las creencias meritocráticas, sin distinguir entre elementos individuales y estructurales, ni diferenciar las preferencias de las percepciones [@KunovichSystemsDistributionSense2007;@NewmanFalseConsciousnessClass2015; @ReynoldsPerceptionsmeritocracyland2014].
-Más aún, se ha optado por índices unidimensionales incluso cuando se ha tenido evidencia en contra de ellos [@KunovichSystemsDistributionSense2007;@ReynoldsPerceptionsmeritocracyland2014].
-La mala especificación de las dimensiones del constructo puede llevar a conclusiones erradas sobre las creencias meritocráticas y su rol en la legitimidad de las desigualdades, por lo que es necesario hacer evaluaciones tanto sobre la dimensionalidad del constructo como de los ítems que permiten aproximarse adecuadamente al mismo.
+Para poder dar cuenta de los niveles de creencia en la meritocracia los estudios a la fecha generalmente han utilizado algunos indicadores de encuestas ya existentes, y en el menor de los casos se han creado instrumentos ad-hoc.
+Sin embargo, y como mostraremos más adelante, las formas de medición de meritocracia varían extremadamente entre estudios.
+Muchas veces fenómenos similares se asocian a indicadores distintos, y también ocurre que fenómenos distintos son medidos con indicadores similares, todo lo cual hace dificulta la comparabilidad entre estudios y el poder avanzar en la comprensión del concepto de meritocracia.
 
-El objetivo de este proyecto fue desarrollar una escala de creencias en la meritocracia, que presentará una adecuada estructura de medición, que además buscamos que fuese estable en el tiempo.
-Por ello, nuestras preguntas de investigación refieren a la estructura del constructo y a la mantención de su validez a nivel temporal.
+Basados en el análisis crítico de las formas de medición de meritocracia a la fecha, el presente estudio propone un instrumento para medir y relacionar dos aspectos claves en el estudio de la meritocracia: percepciones y preferencias.
+Además, como un segundo eje de análisis considera la generación de indicadores respecto de aspectos meritocráticos y anti-meritocráticos, demostrando que no son los dos polos de un mismo continuo como muchos estudios anteriores parecen sugerir.
 
-## Antecedentes generales, la economía moral y la meritocracia
-Las ciencias sociales, desde la perspectiva de la economía moral, han enriquecido la comprensión de las relaciones económicas, aportando a la lógica racional imperante en las teorías económicas, aspectos subjetivos como las normas, los procesos históricos y culturales que estructuran las prácticas tanto las económicas [@mauMoralEconomyWelfare2004; @sachwehMoralEconomyInequality2012; @polanyiGranTransformacionCritica1997; @zelizerEconomicLivesHow2010 ;@fourcadeMoralViewsMarket2007] como la produccion y reproduccion de la desigualdad [@lamontWhatMissingCultural2014].En este marco teórico, se ha  planteado que las creencias meritocráticas pueden legitimar la desigualdad existente [@hadjarMeritokratieAlsLegitimationsprinzip2008].
-En la misma línea, se ha planteado que la creencia en la meritocracia, que adjudica causas individuales a la pobreza, se relaciona con la atribución de prejuicios hacia grupos de bajo estatus [@MadeiraPrimesConsequencesSystematic2019] y con ideas conservadoras de autoritarismo social [@AzevedoNeoliberalIdeologyJustification2019].
+# La medición de los aspectos subjetivos de la meritocracia
 
-## Evidencia de la dimensionalidad de las creencias meritocráticas. 
-La mayoría de las investigaciones que han trabajado con las creencias meritocráticas, la han abordado como un concepto unidimensional.
-No obstante, algunos autores han destacado la diferencia existente entre las preferencias y las percepciones meritocráticas [@Duru-BellatWhoMeritocracyIndividual2012; @CastilloMeritocraciadesigualdadeconomica2019].
-Por percepciones meritocráticas refiere al modo en que el entrevistado describe la realidad.
-Las preferencias refieren a un aspecto normativo, es decir, que tan de acuerdo se está con que la meritocracia es un buen sistema de distribución.
-Esta distinción es fundamental, puesto que ambas variables poseen bastante relación entre sí, por lo cual es de suponer que trabajar con estas variables sin controlarlas entre sí nos puede llevar a tener ideas causales erróneas por espuriedad.
-Así, por ejemplo, puede ser que las conclusiones que indican una mayor percepción de meritocracia en base a los ingresos, posean coeficientes sobreestimados y errores no normales, por la falta de variables relevantes.
+A continuación se presenta una revisión de una serie de investigaciones que se han abocado al estudio de la meritocracia y que para ello han hecho una propuesta de medición.
+El primer eje de análisis tiene que ver con el uso del concepto "creencias" para referir a distintos aspectos subjetivos relacionados con meritocracia.
+El segundo eje tiene que ver con el uso de indicadores sobre aspectos anti-meritocráticos como el polo opuesto de los meritocráticos.
 
-Además, de la distinción entre percepciones y preferencias, es necesario hacer la distinción entre elementos meritocráticos y elementos no meritocráticos.
-Al respecto algunos autores, que trabajan solamente con las percepciones meritocráticas, suponen que creer en el peso de factores externos (e.g.
-la ventaja de tener padres ricos) es contrario a creer en el rol del esfuerzo.
-De este modo, las creencias no meritocráticas son consideradas es el extremo negativo del continuo, mientras creer que la trayectoria personal depende del esfuerzo y el talento del individuo, corresponde al lado positivo del continuo.
-Dicho planteamiento deriva teoría que señala que una sociedad meritocrática ideal se define por la distribución de riquezas en base a esfuerzo y no a fuerzas estructurales [@yairMeritocracy2007; @landWeSatTable2006].
-Aunque dicho planteamiento es correcto para evaluar qué tan meritocrática es una sociedad, no necesariamente este pensamiento se refleja en el modo en cómo se estructuran los discursos en los imaginarios de la población.
-De hecho, otros investigadores han evaluado la dimensionalidad de las creencias meritocráticas, generando evidencia sobre la multidimensionalidad de este constructo, dando cuenta de que aspectos meritocráticos y no meritocráticos no forman parte de la misma dimensión dentro de la subjetividad de los encuestados.
-Entre de las investigaciones que han evaluado dicha dimensionalidad se encuentran los trabajos de @KunovichSystemsDistributionSense2007 y @ReynoldsPerceptionsmeritocracyland2014.
+## La caja negra de las creencias en la meritocracia
 
-La investigación multinivel internacional de @KunovichSystemsDistributionSense2007, que busca explicar las preferencias meritocráticas en el pago utilizando la encuesta ISSP, realizó análisis factoriales para probar la unidimensionalidad sin encontrar resultados satisfactorios, llegando a la conclusión de que las creencias en la meritocracia son más bien un conjunto diverso de ideas agregadas respecto al esfuerzo y la recompensa, razón por la cual decide utilizar un índice sumatorio pese a tener correlaciones insatisfactorias, las cuales estaban más cargadas a los ítems relacionados con aspectos no meritocráticos.
-En vista de la valioza evidencia producida por esta investigacion, podemos decir que, pese a que se ha trabajado como tal, las creencias meritocraticas parecen no ajustar a un modelo unidimensional.
+Para ejemplificar la confusión del uso de creencias en meritocracia consideremos en primer lugar un artículo reciente de Mijs, donde se expresa la siguiente definición: "when I discuss meritocracy beliefs, I am referring to citizens’ belief in the importance of hard work relative to structural factors." [@mijs_paradox_2019, pg.9].
+In the operationalization, this is associated with the following indicator: "how important you think it is for getting ahead in life: (a) hard work", scored in a 1 to 5 likert scale.
+There are several assumptions behind this decision that are worth discussing:
 
-Por su parte, el artículo de @ReynoldsPerceptionsmeritocracyland2014, utilizando la batería del “get ahead” de la ISSP, plantea que las creencias meritocráticas poseen 3 dimensiones, las cuales refieren a la “percepción de meritocracia” (i.e.
-importancia atribuida al éxito y al talento para seguir adelante) a la “percepción de la importancia de las redes para salir adelante (i.e tener padres ricos, tener contactos) y la percepción de la discriminación como obstáculo para salir adelante (i.e la importancia del sexo y de la raza de la persona).
-Pese a esta evidencia, el autor decide hacer un índice unidimensional, donde se incorpora la primera dimensión como el lado positivo del continuo y un agregado de las segundas como el lado negativo del continuo, posteriormente el autor divide este continuo en 4 grupos.
+a.
+Dimensionality: the item used by Mijs is part of an battery present in several international surveys usually called "get ahead", which presents a series of indicators related to what people consider important to get ahead in life: hard work, education, ambition, wealthy family, right connections, religion, race and gender.
+For Mijs, other aspects as education, that could be associated to talent, are not meritocratic: "Hard work is arguably the
+most meritocratic part of Michael Young’s equation, ‘Merit = Intelligence + Effort’, for the simple fact that intelligence itself is conditioned by a nonmeritocratic factor: who your parents happen to be" (p.5).
+This is a very strong assumption: effort would not depend on parents influence, and talent is not meritocratic.
+A similar approach only considering effort as the main element of meritocracy is taken by @solt_economic_2016.
+We suggest here is that questions such as whether talent is not meritocratic (or less meritocratic than effort) should be open to empirical scrutiny and not disregarded upfront based on scholars' assumptions.
+Besides, the use of single indicators for measurement a construct is based on the assumption of a measurement without error.
 
-En suma, considerando la evidencia de estos autores, se deduce la multidimensionalidad del constructo, destacándose la diferencia entre los factores de percepciones meritocráticas y las percepciones no meritocráticas.
+b.
+Beliefs: the "get ahead" battery refers to "how important you think it is", considered by Mijs as a belief in meritocracy.
+Nevertheless, another version of this same battery asks about "how important you think it _should_ be".
+Which one of both is the "belief"?
+If we consider belief as related to a normative content, actually the more close version is the "should be", being the other one more close to the perception of the situation.
+Such distinction was already made by @duru-bellat_whos_2012, who used the item "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" for "desired" meritocracy, whereas for "perceived" meritocracy they use two items: “Would you say that in your country, people are rewarded for their efforts?” and “...
+people are rewarded for their skills?”.
+Is the belief in meritocracy a perception or a preference?
+A comprehensive measurement should be in line with what @duru-bellat_whos_2012 did, in order to open up possibilities of analyzing whether perceptions and preferences are actually the same (i.e.
+correlation close to 1) or they are different aspects of the same phenomenon.
 
-En consideración de lo discutido, se hipotetiza que las creencias en la meritocracia poseen una estructura multidimensional de 4 factores, los cuales son el resultado del cruce entre las diferencias de percepciones o preferencias y aspectos meritocráticos o no meritocráticos, como se indica en la tabla 1.
-Así, probaremos un modelo en que las dimensiones sean percepciones meritocráticas, percepciones no meritocráticas, preferencias meritocráticas y preferencias no meritocráticas.
-Este modelo de medida posee dos ventajas, en primer lugar, no parte del supuesto de la dicotomía entre las creencias meritocráticas y no meritocráticas permitiendo que los sujetos puntúan alto o bajo en las dos variables.
-Igualmente, es positivo que exista la distinción entre preferencias y percepciones puesto que esto nos permitirá predicciones más robustas y más precisas respecto a las creencias de la meritocracia.
+c.
+Non-meritocratic aspects: @mijs_paradox_2019 make reference to some non-meritocracti aspects as talent, letting this out of the measurement.
+A different approach was followed by @kunovich_systems_2007, who decide to include non-meritocratic elements in the measurement of meritocracy.
+Using the items’ battery listing a number of reasons about “How important should be in deciding pay...” (as @duru-bellat_whos_2012 for desired meritocracy), he decided that reasons as education and responsibility are meritocratic and pointed 1 if considered essential, whereas reasons such as having a family and children were pointed 1 if they were considered "not important at all" (i.e.
+reverse coded).
+A similar aproach was taken by @newman_false_2015, reverse-coding non-meritocratic items.
+The assumption that meritocratic and non-meritocratic elements are the poles of the same continuum was analyzed by @reynolds_perceptions_2014 using the same "get ahead" perceptions' battery items.
+They consider education, ambition and hard work as meritocratic and other reasons such as wealthy family and right connections and non-mertitocratic.
+Nevertheless, despite making this distinction the author ends up substracting one dimension from the other, assuming that they are two poles of the same continuum as @kunovich_systems_2007 did.
+
+d.
+Accounting for measurement error: finally, most of the studies in meritocracy so far have not incorporated the issue of measurement error @brown_confirmatory_2006, using single indicators and/or composite indexes for measuring meritocracy.
+This means assuming that the construct is measured perfectly by the indicators chosen, going as far as proposing that "we decided to rely on a summative index.
+In choosing this strategy of index construction, we argue that *support for meritocracy is not a latent variable* [@kunovich_systems_2007 p.653-654].
+Some advances were done by @reynolds_perceptions_2014 by doing a principal component analysis of meritocratic and non-meritocratic dimensions, but somewhat contradictorily they end up in a sum index despite proving multidimensionality.
+
+## An instrument proposal
+
+Based on the previous limitations in the measurement of meritocracy presented in the previous section, in this paper we propose and test an instrument with the following characteristics:
+
+- Multidimensional, incorporating previous distinctions between preferences and perceptions as well as between meritocratic and non-meritocratic aspects.
+- Multiple indicators for each dimension, in order to account for measurement error in a confirmatory factor analysis context.
+- Based on previous indicators as far as possible, for the sake of comparability between studies
+- Brief, as to be used in a regular survey.
+
+The proposed measurement model is based on two axis of analysis, as depicted in Table 1:
+
+Table 1: Model of perceptions and preferences for meritocracy and non-meritocracy
 
@@
 
-# Sección de indicadores 
-Si bien no existe un consenso respecto a cuáles son los indicadores más adecuados para medir creencias en la meritocracia, los indicadores más comunes son preguntas del tipo, “qué tan importante es … para salir adelante”, las cuales incluyen características como tener padres ricos, trabajar duro, ser talentoso, el sexo, la raza, entre otras.
-El problema con este indicador presente en la International Social Survey Program, es que habla de un criterio de importancia y no de suficiencia.
-Así, por ejemplo, alguien puede creer que el esfuerzo es fundamental para salir adelante, y creer al mismo tiempo que no todas las personas que se esfuerzan más logran obtener más recursos que quienes se esfuerzan menos.
-Al respecto, hemos encontrado evidencia sistemática de que si bien una gran mayoría de las personas de chile consideran que el esfuerzo es fundamental para salir adelante, no creen que las personas no son recompensadas por su esfuerzo en su país.
+The first one distinguishes between different types of "beliefs", using instead the terms perceptions and preferences for meritocracy.
+Perceptions refers to the extent to which people see meritocracy working in their society, which in terms of measurement relates to the "reasons to get ahead" battery, whereas preferences refer to normative expectations that are usually linked to a "should" expression (e.g.
+whether hard work should be related to payment).
+The second axis consider the distinction between meritocratic and non-meritocratic dimensions.
+This aspect has been usually treated as different ends of a same continuum in part of the previous research, an assumption that requieres empirical scrutiny.
+These non-meritocratic elements usually refer to the use of personal contacts or family advantages to get ahead in life.
 
-Otro indicador de esa misma encuesta, pregunta respecto a si cree que es importante a considerar en el pago de una persona distintas características, incluyendo características relativas al mérito, como educación o responsabilidad, y algunas relativas a las necesidades, como el tener hijos.
-No obstante, este modo de evaluar las creencias respecto a la meritocracia, si bien posee un aspecto no meritocrático, este no refiere a las desigualdades heredables que son lo propiamente contrario al ideal meritocrático (Yair, 2007; Land, 2006; newman; Reynolds and Xian, 2014).
-Además, este indicador no incluye aspectos de percepción, lo cual, como dijimos, puede generar problemas de espuriedad.
+Regarding the selection of indicators, most of them some adaptations of previous items that where streamlined in order to a better fit with the specific dimensions.
+For meritocratic indicators we use effort and talent as the main components of the traditional concept as defined by @young_rise_1962, whereas for non-meritocratic dimensions we use having rich parents and good contacts.
 
-En vista de la inexistencia de un consenso respecto a la manera correcta de medir las creencias meritocráticas proponemos nuevos indicadores que buscan ser más enfáticos en el fraseo respecto a la diferencia de preferencias y percepciones.
-Así, las preguntas que utilizamos son las presentes en la tabla 2.
-Estos indicadores, para referirse a las percepciones sitúan al entrevistado en la realidad de su país.
+![](../input/images/tabla2.png)
 
-![](input/images/tabla2.png)
+**Hypotheses**
 
-# Muestra del Estudio
+* H1. The perception of meritocracy is a latent variable based on indicators of the importance attributed to talent and the effort to get ahead in life.
+* H2. The non-meritocratic perception is a latent variable that derives from two indicators related to the agreement with the statement that people with contacts and rich parents manage to get ahead.
+* H3. Meritocratic preferences behave as the latent variable based on a normative value of effort and talent.
+* H4. Non-meritocratic preferences behave as a latent variable based on the normative value of the use of personal contacts and having rich parents.
 
-# Resultados: Estructura factorial, efecto framing e invarianza longitudinal
+# Methodology
+
+## Data
+
+## Variables
+
+## Methods
+
+# Results
+
+---
+
 Este estudio cuenta con una encuesta panel de 3 olas.

### 2020-04-12 docs/previous_versions/paper_version1.Rmd
Update paper

--- 2020-04-07 docs/paper.rmd
@@
 
-## Introducción
+# Introducción
 
-Tanto las organizaciones internacionales, como parte del mundo académico, presentan una creciente preocupación por el aumento de la desigualdad social y los efectos negativos que esta posee en torno a la legitimidad de nuestras sociedades [ONU,2016; @schlozmanUnheavenlyChorusUnequal2012; @CastilloMeritocraciadesigualdadeconomica2019].
-Más relevante aún se vuelve la legitimidad de la desigualdad en un contexto mundial de crecientes protestas sociales en contra de la misma.
-En el caso Chileno, estas movilizaciones han alcanzado una gran envergadura, relevando la importancia de evaluar los niveles de legitimidad de la desigualdad existente, a la luz de distintas ideologías que según la teoría podrían justificar.
-Dentro de estas ideologías que legitiman la desigualdad, la literatura sobre la meritocracia, presenta un notorio crecimiento [@MadeiraPrimesConsequencesSystematic2019].
+La desigualdad económica se ha vuelto un tema que genera creciente preocupación y malestar alrededor del mundo.
+Esto se ha expresado en una serie de protestas como la emblemática "occupy wall street" el año 2011, así como también en una serie de análisis críticos respecto del desarrollo del capitalismo y sus consecuencias [@piketty_Capital_2014].
+En este contexto, el estudio de las visiones, preferencias y percepciones respecto de la desigualdad han adquirido relevancia en las ciencias sociales, en temas como las preferencias redistributivas [@alesina_preferences_2011; @Dimick2018ModelsOtherRegardingPreferences] la legitimación de la desigualdad económica [@schroder_is_2017] y el funcionamiento de la meritocracia [@duru-bellat_whos_2012; @mijs_unfulfillable_2016; @reynolds_perceptions_2014].
+
+En general, la meritocracia se define como un sistema de distribución de recursos y recompensas basados en el mérito individual, que en su concepción original es una suma de talento y esfuerzo [@young_rise_1962].
+Una serie de estudios han realizado críticas a la realización de este estándar moral de distribución, planteando que es una promesa incumplida dada la influencia preponderante de otros elementos más allá del mérito en el estatus individual [@arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019].
+Por otro lado, desde la psicología social y la sociología se han estudiado las características y consecuencias de las creencias en la meritocracia, en general basados en la hipótesis que mayor creencia en la meritocracia lleva a una mayor legitimación de las desigualdades [@hadjar_meritokratie_2008; @MadeiraPrimesConsequencesSystematic2019].
+
+estas ideologías que legitiman la desigualdad, la literatura sobre la meritocracia, presenta un notorio crecimiento [@MadeiraPrimesConsequencesSystematic2019].
 Tradicionalmente la meritocracia es entendida como una forma de distribución social de recursos donde el mérito individual (i.e.

### 2020-04-12 docs/paper.Rmd
update paper

@@
 correlation close to 1) or they are different aspects of the same phenomenon.
+As @son_hing_merit_2011-1 have pointed out, "People can believe that outcomes ought to be distributed on the basis of merit and yet vary in their perceptions of whether this is how society currently operates" (p.
+435).
+Accordingly, we suggest that the use of the term belief is confusing, being more accurate to speak about meritocratic perceptions and preferences.
 
 c.
-Non-meritocratic aspects: @mijs_paradox_2019 make reference to some non-meritocracti aspects as talent, letting this out of the measurement.
+Non-meritocratic aspects: @mijs_paradox_2019 make reference to some non-meritocratic aspects as talent, letting this out of the measurement.
 A different approach was followed by @kunovich_systems_2007, who decide to include non-meritocratic elements in the measurement of meritocracy.
@@
 reverse coded).
-A similar aproach was taken by @newman_false_2015, reverse-coding non-meritocratic items.
+A similar approach was taken by @newman_false_2015, reverse-coding non-meritocratic items, similarly to what occurs with the "Preference for the Merit Principle Scale" [@davey_preference_1999].
 The assumption that meritocratic and non-meritocratic elements are the poles of the same continuum was analyzed by @reynolds_perceptions_2014 using the same "get ahead" perceptions' battery items.
@@
 - Brief, as to be used in a regular survey.
+In this point it differs for instance from the proposal of the "Preference for the Merit Principle Scale" [@davey_preference_1999], as they use 15 items just for one dimension (besides the problem of reverse-coding non-meritocratic items).
 
@@
 
-The first one distinguishes between different types of "beliefs", using instead the terms perceptions and preferences for meritocracy.
+The first axis distinguishes between different types of "beliefs", using instead the terms perceptions and preferences for meritocracy [@duru-bellat_whos_2012; @son_hing_merit_2011-1].
 Perceptions refers to the extent to which people see meritocracy working in their society, which in terms of measurement relates to the "reasons to get ahead" battery, whereas preferences refer to normative expectations that are usually linked to a "should" expression (e.g.
 whether hard work should be related to payment).
-The second axis consider the distinction between meritocratic and non-meritocratic dimensions.
+The second axis consider the distinction between meritocratic and non-meritocratic dimensions [@reynolds_perceptions_2014].
 This aspect has been usually treated as different ends of a same continuum in part of the previous research, an assumption that requieres empirical scrutiny.
@@
 Regarding the selection of indicators, most of them some adaptations of previous items that where streamlined in order to a better fit with the specific dimensions.
-For meritocratic indicators we use effort and talent as the main components of the traditional concept as defined by @young_rise_1962, whereas for non-meritocratic dimensions we use having rich parents and good contacts.
+For meritocratic indicators we use effort and talent as the main components of the traditional concept of merit as defined by @young_rise_1962, whereas for non-meritocratic dimensions we use having rich parents and good contacts.
 
@@
 * H1. The perception of meritocracy is a latent variable based on indicators of the importance attributed to talent and the effort to get ahead in life.
+
 * H2. The non-meritocratic perception is a latent variable that derives from two indicators related to the agreement with the statement that people with contacts and rich parents manage to get ahead.
+
 * H3. Meritocratic preferences behave as the latent variable based on a normative value of effort and talent.
+
 * H4. Non-meritocratic preferences behave as a latent variable based on the normative value of the use of personal contacts and having rich parents.

### 2020-04-12 docs/preanalisis.tex
update paper

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Pre-registration},
  colorlinks=true,
  linkcolor=blue,
  filecolor=Maroon,
  citecolor=Blue,
  urlcolor=Blue,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=0.78in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{caption}
\captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
\usepackage{graphicx}
\usepackage{float}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}

\title{Pre-registration}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{2}
\tableofcontents
}
\setstretch{1.5}
\hypertarget{perceptions-and-preferences-for-meritocracy-scale-ppm-s}{%
\section{Perceptions and Preferences for Meritocracy Scale
(PPM-S)}\label{perceptions-and-preferences-for-meritocracy-scale-ppm-s}}

(link to \href{docs/preanalisis.pdf}{PDF file})

\hypertarget{research-questions}{%
\section{Research questions}\label{research-questions}}

Is it possible to identify and measure perceptions and preferences for
meritocracy as two distinct but related factors?

\hypertarget{general-description-and-hypotheses}{%
\section{General description and
hypotheses}\label{general-description-and-hypotheses}}

This research builds upon previous studies attempting to develop a
comprehensive approach to measuring meritocratic beliefs (Duru-Bellat \&
Tenret, \protect\hyperlink{ref-Duru-bellat2012}{2012}; Kunovich \&
Slomczynski, \protect\hyperlink{ref-Kunovich2007}{2007}; Newman et al.,
\protect\hyperlink{ref-Newman2015}{2015}; Reynolds \& Xian,
\protect\hyperlink{ref-Reynolds2014}{2014}). The proposed measurement
model is based on two axis of analysis, as depicted in Table 1. The
first one distinguishes between different types of ``beliefs'', using
instead the terms perceptions and preferences for meritocracy.
Perceptions refers to the extent to which people see meritocracy working
in their society, which in terms of measurement relates to the ``reasons
to get ahead'' battery, whereas preferences refer to normative
expectations that are usually linked to a ``should'' expression
(e.g.~whether hard work should be related to payment). The second axis
consider the distinction between meritocratic and non-meritocratic
dimensions. This aspect has been usually treated as different ends of a
same continuum in part of the previous research, an assumption that
requieres empirical scrutiny. These non-meritocratic elements usually
refer to the use of personal contacts or family advantages to get ahead
in life.

Table 1: Model of perceptions and preferences for meritocracy and
non-meritocracy

\begin{longtable}[]{@{}lll@{}}
\toprule
& Perceptions & Preferences\tabularnewline
\midrule
\endhead
\emph{Meritocracy} & &\tabularnewline
\emph{Non-meritocracy} & &\tabularnewline
\bottomrule
\end{longtable}

\textbf{Hypotheses}

\begin{itemize}
\tightlist
\item
  H1. The perception of meritocracy is a latent variable based on
  indicators of the importance attributed to talent and the effort to
  get ahead in life.
\item
  H2. The non-meritocratic perception is a latent variable that derives
  from two indicators related to the agreement with the statement that
  people with contacts and rich parents manage to get ahead.
\item
  H3. Meritocratic preferences behave as the latent variable based on a
  normative value of effort and talent.
\item
  H4. Non-meritocratic preferences behave as a latent variable based on
  the normative value of the use of personal contacts and having rich
  parents.
\end{itemize}

\hypertarget{type-of-study}{%
\section{Type of study}\label{type-of-study}}

The scale will be included in an online three-wave panel survey study.
This study covers among other things an experiment of the effect of
information about poverty and inequality on opportunity beliefs. The
scale will be presented in wave 1 and wave 3.

\hypertarget{sample}{%
\section{Sample}\label{sample}}

\hypertarget{detailed-data-and-explanations-of-its-use.}{%
\subsection{Detailed data and explanations of its
use.}\label{detailed-data-and-explanations-of-its-use.}}

The data will come from a representative sample of the large cities of
Chile (Gran Santiago, Valparaíso-Viña del Mar, Concepción and
Antofagasta). In order to ensure representativeness, quotas for sex,
education and age will be provided to the agency in charge of the online
fieldwork.

\hypertarget{magnitude-of-the-sample-and-explanation-of-the-magnitude.}{%
\subsection{Magnitude of the sample and explanation of the
magnitude.}\label{magnitude-of-the-sample-and-explanation-of-the-magnitude.}}

The sample will have an initial number for the first wave of 2100 cases,
which are expected to be at least 1800 in the second wave. From the
second to the third wave it has been considered as ``free fall'', from
which we expect to retain about 1500 cases.

The sample size was calculated from power analysis that used information
from a pilot study that was applied in June 2019. This pilot study
consisted of an online experiment applied to 949 subjects, randomly
separating participants to two stimuli and a placebo, leaving a third of
the sample in each condition. From these data, the low power of a
two-tailed test was calculated, considering an average treatment effect
of 0.26, with a standard deviation of the result of the treatment group
of 1.23 and a significance level of 0.05. Based on multiple comparison
correlations (Bonferroni), we adjusted alpha for six tests (three
results * two comparisons between conditions and placebo). The results
indicated that the number of subjects exposed to each stimulus should be
580, so we will work with three samples of 600 cases, which gives us a
total of 1800 cases for the second wave in which the experiment will be
applied. To guarantee the above and in consideration of a high
non-response rate, we will start with a sample of 2100 subjects.

\hypertarget{data-collection-procedure}{%
\subsubsection{Data collection
procedure:}\label{data-collection-procedure}}

Respondents will be invited to participate in the online survey under
the incentive provided by the external company. Respondents have 4
business days to answer the survey, with a maximum of 20 minutes per
survey. In order to achieve greater representativeness of the sample,
the quota method will be used, that is, the program will only allow
respondents with the required demographic characteristics. The quotas
used were age, sex and education. It should be noted that the time
between wave and wave is around 7 and 9 days.

\hypertarget{stop-rule}{%
\subsubsection{Stop Rule:}\label{stop-rule}}

The rule of detention is to reach the number of respondents indicated in
the quotas. However, there may be modifications depending on the
contingencies of the field work and the response rates.

\hypertarget{variables-to-use}{%
\section{Variables to use:}\label{variables-to-use}}

All the variables used in the scale use the following question and
answer options:

\begin{itemize}
\tightlist
\item
  To what extent do you agree or disagree with each of the following
  affirmations?

  \begin{itemize}
  \tightlist
  \item
    Response scale:

    \begin{itemize}
    \tightlist
    \item
      Strongly disagree
    \item
      Disagree
    \item
      Neither agree nor disagree
    \item
      Agree
    \item
      Totally agree
    \end{itemize}
  \end{itemize}
\end{itemize}

\hypertarget{variable-manipulation}{%
\subsubsection{Variable manipulation}\label{variable-manipulation}}

There will not be variable manipulation.

\hypertarget{indexes}{%
\subsubsection{Indexes}\label{indexes}}

The investigation of the validity of this scale does not rely on indices
but rather confirmatory factor analysis will be used to estimate the
value of the latent variables underlying the indicators.

\hypertarget{blinding-of-information}{%
\subsubsection{Blinding of information}\label{blinding-of-information}}

No special blinding is used for this study.

\hypertarget{randomization}{%
\subsubsection{Randomization}\label{randomization}}

To test the items' order effects we present the items in three different
order conditions which will be assigned randomly. The conditions are the
following:

\begin{itemize}
\tightlist
\item
  Order 1: the order in which items are presented in Table 1
\item
  Order 2: the items are ordered based on topics (work harder, talent,
  rich parents, connections), and for each of them they have to respond
  first about perceptions and then about preferences. For instance: a)
  those who work harder get greater rewards, b) those who work arder
  should get greater rewards \ldots{}
\item
  Order 3: the order is randomized within this condition
\end{itemize}

\hypertarget{analysis-plan}{%
\section{Analysis plan}\label{analysis-plan}}

To evaluate the hypotheses, confirmatory factor analysis (CFA) will be
used as we rely on a theory regarding the underlying four factors. The
method of estimation will be Weighted Least Squares Mean Variance
Approximation (WLMSV) for categorical ordered indicators. For the
analysis we will use the R package Lavaan (Rosseel,
\protect\hyperlink{ref-Rosseel2012}{2012}).

\hypertarget{inference-criteria}{%
\subsubsection{Inference criteria}\label{inference-criteria}}

The values that will be used as evaluation criteria for the goodness of
the adjustment of the model were taken from the proposal of Brown
(\protect\hyperlink{ref-Brown2008}{2008}) and are the following:

\begin{itemize}
\tightlist
\item
  Chi-square:\textgreater{} 0.05
\item
  Chi-square ratio:\textgreater{} 3
\item
  Comparative adjustment goodness index (CFI):\textgreater{} 0.95
\item
  Tucker-Lewis Index (TLI):\textgreater{} 0.95
\item
  Root of the average squared residual approximation \textless0.08.
\end{itemize}

\hypertarget{secondary-analysis}{%
\subsubsection{Secondary analysis}\label{secondary-analysis}}

To evaluate the metric stability of the measurement model (Davidov et
al., \protect\hyperlink{ref-Davidovetal2014}{2014}) we will conduct a
longitudinal invariance test using the first and third wave of our
study. Following Liu et al.
(\protect\hyperlink{ref-liuTestingMeasurementInvariance2017}{2017}), we
will test a series of four hierarchical models: Configural, Weak, Strong
and Strict invariance models for ordered-categorical indicators based on
the assumption that a five category likert scale cannot be treated as a
continous variable because can lead to biased parameter estimates.

\hypertarget{data-exclusion}{%
\subsubsection{Data Exclusion}\label{data-exclusion}}

All cases will be used as long as they do not show missing values in any
of the items on the scale. No imputation criteria will be used.

\hypertarget{ethics.}{%
\subsubsection{Ethics.}\label{ethics.}}

The experiment and survey instruments are approved by the IRB of the
University of Chile.

\hypertarget{bibliography}{%
\section*{Bibliography}\label{bibliography}}
\addcontentsline{toc}{section}{Bibliography}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Brown2008}{}%
Brown, T. A. (2008). \emph{Confirmatory Factor Analysis for Applied
Research}. \url{https://doi.org/10.1198/tas.2008.s98}

\leavevmode\hypertarget{ref-Davidovetal2014}{}%
Davidov, E., Meuleman, B., Cieciuch, J., Schmidt, P., \& Billiet, J.
(2014). Measurement Equivalence in Cross-National Research. \emph{Annual
Review of Sociology}, \emph{40}(1), 55--75.
\url{https://doi.org/10.1146/annurev-soc-071913-043137}

\leavevmode\hypertarget{ref-Duru-bellat2012}{}%
Duru-Bellat, A. M., \& Tenret, E. (2012). Who's for Meritocracy?
Individual and Contextual Variations in the Faith. \emph{Comparative
Education Review}, \emph{56}(2), 223--247.

\leavevmode\hypertarget{ref-Kunovich2007}{}%
Kunovich, S., \& Slomczynski, K. M. (2007). Systems of distribution and
a sense of equity: A multilevel analysis of meritocratic attitudes in
post-industrial societies. \emph{European Sociological Review},
\emph{23}(5), 649--663. \url{https://doi.org/10.1093/esr/jcm026}

\leavevmode\hypertarget{ref-liuTestingMeasurementInvariance2017}{}%
Liu, Y., Millsap, R. E., West, S. G., Tein, J.-Y., Tanaka, R., \& Grimm,
K. J. (2017). Testing measurement invariance in longitudinal data with
ordered-categorical measures. \emph{Psychological Methods},
\emph{22}(3), 486--506. \url{https://doi.org/10.1037/met0000075}

\leavevmode\hypertarget{ref-Newman2015}{}%
Newman, B. J., Johnston, C. D., \& Lown, P. L. (2015). False
Consciousness or Class Awareness? Local Income Inequality, Personal
Economic Position, and Belief in American Meritocracy. \emph{American
Journal of Political Science}, \emph{59}(2), 326--340.
\url{https://doi.org/10.1111/ajps.12153}

\leavevmode\hypertarget{ref-Reynolds2014}{}%
Reynolds, J., \& Xian, H. (2014). Perceptions of meritocracy in the land
of opportunity. \emph{Research in Social Stratification and Mobility},
\emph{36}, 121--137. \url{https://doi.org/10.1016/j.rssm.2014.03.001}

\leavevmode\hypertarget{ref-Rosseel2012}{}%
Rosseel, Y. (2012). Lavaan: An R package for structural equation
modeling and more. Version 0.5--12 (BETA). \emph{Journal of Statistical
Software}, \emph{48}(2), 1--36.

\end{document}

### 2020-04-21 docs/paper.Rmd
* Actualización: 1. data proc 2. data analysis: se agrega un readme de production 3. Esquemas y figuras 4. render-site: eliminar cache y residuales del render 5. Test de paper con tables+figuras 6. Se elimino output pdf de preanalisis

@@
 title: "Perceptions and Preferences for Meritocracy"
+css: "../input/bib/style.css"
 linestretch: '1.5'
@@
   - ../input/bib/meritocracy.bib
+editor_options: 
+  chunk_output_type: console
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = FALSE)
+knitr::opts_knit$set(root.dir = "../")
+options(scipen=999)
+rm(list=ls())
+options(knitr.kable.NA = '')
 ```
+
+```{r message=FALSE, warning=FALSE, include=FALSE}
+library(knitr)
+library(kableExtra)
+library(dplyr)
+```
+
 <div style="text-align: justify">
@@
 
+The study design consist in two main stages.
+The first stage was the questionnaire design, in which we collect a serie of background variables, including sex, age and educational level.
+On this regard, one of the main aspects of our design is that we tested three different application modalities, in order to observe how the display order of the indicators could affect the model fit.
+We decide to randomize the assignment of the subject to each application modality, and as a result we obtain three groups within the full sample.
+The first combination corresponds to the group that answered the questions that were sorted by the Perceptions/Preferences dimension.
+The second combination corresponds to the group in which the questions are displayed following a topic-sorted way, this means that the effort, talent, rich family and networks questions for perception and preference were displayed together.
+The last and third group corresponds to a completely randomized display order of the questions.
+The figure below describes the complete survey flow.
+
+![Figure: Study design](../input/images/app_mod.png)
+
+To contact the subjects, we followed an online programmed survey approach using _Qualtrics_, and the field work was conducted by _Netquest_, which is a company with wide experience applying online survey, and for the case of Chile, they have a consolidated pool of individuals that systematically respond online surveys.
+On this regard, the company have their own incentive system, in which the subjects accumulates specific currencies that they can exchange for a wide range of goods.
+
 ## Data
+
+The field work was conducted during November and December of 2019 in Chile, obtaining a final sample of 2200 individuals, but considering that we use cases with complete information, the final analytical sample has 2141 observations.
 
@@
 
+<!-- [Descriptive Table for the PPM-S, by group or the full sample] -->
+
+```{r echo=FALSE}
+load(file = "output/tables/sumdat04.RData");sumdat04
+```
+
 ## Methods
 
+CFA
+
+MGCFA
+
+@rprogram2019
+@Rosseel2012
+
 # Results
+
+![](../output/images/meas01.png)
+
+```{r echo=FALSE}
+load(file = "output/tables/tb.fm01.RData");tb.fm01
+```
 

### 2020-04-21 production/prod_prep-data-A.rmd
* Actualización: 1. data proc 2. data analysis: se agrega un readme de production 3. Esquemas y figuras 4. render-site: eliminar cache y residuales del render 5. Test de paper con tables+figuras 6. Se elimino output pdf de preanalisis

@@
 title: "Data preparation"
-date: "`r format(Sys.time(), '%d %B, %Y')`"
+date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
 css: "../input/bib/style.css"
@@
     toc: true
-        smart: false
+    toc_float:
+        collapsed: false
     toc_depth: 2
@@
 agree or disagree with each of
-the following statements? - Is
-well that those who have good
-contacts get ahead."
+the following statements? - It's fine that those who have rich parents get ahead" 
 dat04$pref_netw <- set_labels(dat04$pref_netw,  labels = c("strongly disagree", "disagreement", "Neither agree nor disagree","Agree", "totally agree"))

### 2020-04-22 docs/paper.Rmd
Methods
Methods data variables.

--- 2020-04-12 docs/paper.Rmd
@@
 ## Data
+The Netquest company ([website](www.netquest.cl)) applied the online survey during December and January.
+The online survey distributed by the company was scheduled on the platform Qualtrics ([website](www.Qualtrics.com)).
+The sample was selected from a non-probabilistic design in three large cities in Chile.
+The quota method based on age, sex and educational level was used.
+Quotas were generated based on the survey of the Public Studies Center (CEP, 2019), which has a high prestige in the country.
+A total sample of 2141 people was collected, excluding those who did not answer the questions on the scale and those who did not accept informed consent.
+The representativeness of the sample is not satisfactory because, although the proportion by age and sex is consistent with the estimates of the CEP survey, a bias towards high educational levels is evident (as detailed in Annex __X__).
+This limitation must be considered when making inferences about the population.
+
+> Anexo x
+
+|                     | Sample  |   CEP  |
+|---------------------|:-------:|:------:|
+| __gender__          |         |        |
+| Men                 | 49,82%  | 50,52% |
+| Women               | 50.18%  | 49,47% |
+| __Age__             |         |        |
+| 18 - 24             | 18,55%  | 18,17% |
+| 25 - 34             | 18,86%  | 17,48% |
+| 35 - 44             | 19.09%  | 19,98% |
+| 45 - 54             | 17,96%  | 19,23% |
+| 55 - or more        | 25,54%  | 25.11% |
+| __Education__       |         |        |
+| Primary  or less    | 2,93%   | 15,88% |
+| Hig school          | 43,23%  | 37,04% |
+| Non university      | 32,63%  | 28,93% |
+| university or more  | 21,21%  | 18,13% |
+
+Based on the recommendations of Wilson, Strite and Loiacono (2017) it has been considered pertinent transparency that the order the items of the meritocracy scale in the survey was random for a third of the sample.
+With the objective of evaluating the effect of the order of the indicators, the respondents were randomly divided into three groups as explained in __Diagram 1__.
+The scale was presented to the first group in the order that appears in Table 2, in such a way that the items of the same factor are found together.
+The second group was presented with the scale with the items of the same factor separated from each other, in such a way that they were consulted at the same time for their perceptions and preferences of a particular topic (e.g.
+distribution according to effort).
+The third group was presented with the scale with the randomized items to assess the possible bias produced by the order in which the questions are presented ((Budd, 1987) [https://psycnet.apa.org/record/1988-31448-001]; Davis and Venkatesh 1996; Goodhue and Loiacono, 2002; Schriesheim et al., 1989).
+
+>Diagrama de flujo de la encuesta, orden de la escala.
 
@@
 
+The proposed scale of perceptions and preferences about meritocracy consists of 8 indicators that are grouped into the 4 dimensions indicated.
+The questions are adaptations of the "get ahead" scale.
+the terms of of this scale were used (i.e talent, effort, contacts and wealthy parents) which are included in different investigations on meritocracy [@mijs_paradox_2019; @duru-bellat_whos_2012; @reynolds_perceptions_2014].
+Unlike this scale, this proposal adds the difference between perceptions and preferences.
+In addition, a comparative component is incorporated into the question that refers to differences between people with little or fed up with some attribute (e.g.
+effort, contacts).The eight items ordered according to dimensions are presented in Table 2.
+These 8 likert-type items have 5 response alternatives ranging from "I completely agree" (5) to "I completely disagree"(1).
+According to the recommendations of Xia and Yang ([2018](https://link.springer.com/article/10.3758/s13428-018-1055-2)) for categorical data, we present below a table with the descriptive table and another with correlations between the items on the scale.
+
+```{r echo=FALSE,results='asis', warning=FALSE, message=FALSE}
+load(file = "../input/data/proc/dat04.RData")
+library(stargazer)
+stargazer(dat04,type = "html",
+          title = "Descriptive statistics",
+          font.size = "small",
+          median = TRUE,
+          header=FALSE,
+          digit.separator = "",
+          digits = 2,
+          covariate.labels = c("A. pc.effort", "B. pc.talent","C. pc.wpart","D. pc.netw","E. pf.effort","F. pf.talent","G. pf.wpart", "H. pf.netw"))
+```
+
 ## Methods
+
+We tested theoretical model of the scale of perceptions and preferences about meritocracy using confirmatory factor analysis (CFA) with for latent factor.
+CFA was conducted using the lavaan package (version 0.6-3; Rosseel, 2020) with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement (Kline, 2016; Rosseel, 2020).
+As recommended by Brown (2008), we assessed model fit by jointly considering the comparative fit index and Tucker-Lewis Index (CFI and TLI; acceptable fit > 0.95), Root of the average squared residual approximation (RMSEA; acceptable fit < 0.08),  Chi-square: (p-value; acceptable fit > 0.05, and Chi-square ratio:> 3).
 

### 2020-04-22 docs/preanalisis.Rmd

--- 2020-03-21 docs/preanalisis.Rmd
@@
 output:
-  pdf_document:
-    keep_tex: false
-    template: null
-    number_sections: true
-    toc: true
+  # pdf_document:
+  #   keep_tex: false
+  #   template: null
+  #   number_sections: true
+  #   toc: true
   html_document:

### 2020-04-22 production/prod_analysis-cfa.rmd

@@
 date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
-css: "../input/bib/style.css"
+css: "../input/css/style.css"
 output:
@@
 ```{r eval=FALSE, include=FALSE}
-rmarkdown::render(input = "production/prod_analysis-cfa.Rmd",output_format = "html_document"); browseURL(url = "production/prod_analysis-cfa.html")
+rmarkdown::render(input = "production/prod_analysis-cfa.rmd",output_format = "html_document"); browseURL(url = "production/prod_analysis-cfa.html")
 ```
@@
 library(kableExtra)
+library(rvest)
+source('production/prod_polychoric.R')
 ```
@@
 
+```{r}
+labs<- c("A. Who the more they try they manage to get bigger rewards that those who striveless.",
+         "B. Who possess more talent they manage to obtain greater rewards than those who possess less talent.",
+         "C. Who they have rich parents manage to get out ahead.",
+         "D. Who they have good contacts they manage to get out ahead.",
+         "E. Who the more they try they should get greater rewards than those who they try less." ,
+         "F. Who possess more talent they should get greater rewards than those who possess less talent." ,
+         "G. It's fine that those who have rich parents get ahead",
+         "H. Is well that those who have good contacts get ahead." )
+cnametab<- c("","(A)","(B)","(C)","(D)","(E)","(F)","(G)", "(H)","(1)","(2)","(3)","(4)","(1)","(2)","(3)","(4)")
+```
+
+```{r tabla01, eval=TRUE, include=TRUE}
+polyw01<- poly(sjlabelled::remove_all_labels(dat01),
+               digits = 2,
+               triangle = "l",
+               fade.ns = F,
+               remove.spaces = T,
+               string.diag = c("$-$","$-$","$-$","$-$","$-$","$-$","$-$","$-$"))
+matpolyw01 <- as.data.frame(read_html(x = polyw01[["page.content"]]) %>% html_table(fill=FALSE)) 
+matpolyw01 <- matpolyw01[-c(9),]
+colnames(matpolyw01) <- c("vars","(A)","(B)","(C)","(D)","(E)","(F)","(G)","(H)")
+
+load01<- data.frame(round(cbind(inspect(fit1_c,what="std")$lambda,inspect(fit1_o,what="std")$lambda),digits = 2))
+load01[load01==c(0.00)] <- NA; names(load01) <- c(1:4,1:4)
+
+table01<- bind_cols(x = matpolyw01,y = load01)
+# %>% mutate(vars=labs)
+
+kable(table01,escape = TRUE,align = "l",col.names = cnametab,format = "html") %>% 
+  footnote(general = paste0(paste0("$N$= ",nobs(fit1_c)),"; p<0.05=$*$; p<0.01=$**$; p<0.001=$***$"),footnote_as_chunk = T) %>%
+  add_header_above(c(" " = 1,"Polychoric correlations"=8, "MLR" = 4, "DWLS" = 4))%>%
+  add_header_above(c(" " = 9,"Factor loadings"=8)) %>%
+  column_spec(column = c(1),width ="7in",include_thead = T) %>%
+  column_spec(column = c(2:17),width ="0.5in",include_thead = T) %>% 
+  kable_styling(full_width = FALSE)
+```
+
 ## Version 02  (N=`r dim(dat02)[1]`)
@@
 
+```{r tabla02, echo=FALSE, include=TRUE}
+polyw02<- poly(sjlabelled::remove_all_labels(dat02),
+               digits = 2,
+               triangle = "l",
+               fade.ns = F,
+               remove.spaces = T,
+               string.diag = c("$-$","$-$","$-$","$-$","$-$","$-$","$-$","$-$"))
+matpolyw02 <- as.data.frame(read_html(x = polyw02[["page.content"]]) %>% html_table(fill=FALSE)) 
+matpolyw02 <- matpolyw02[-c(9),] 
+colnames(matpolyw02) <- c("vars","(A)","(B)","(C)","(D)","(E)","(F)","(G)", "(H)")
+
+load02<- data.frame(round(cbind(inspect(fit2_c,what="std")$lambda,inspect(fit2_o,what="std")$lambda),digits = 2))
+load02[load02==c(0.00)] <- NA; names(load02) <- c(1:4,1:4)
+
+table02<- bind_cols(x = matpolyw02,y = load02)
+# %>% mutate(vars=labs)
+
+kable(table02,escape = FALSE,align = "l",col.names = cnametab)  %>% 
+  footnote(general = paste0(paste0("$N$= ",nobs(fit2_c)),"; p<0.05=$*$; p<0.01=$**$; p<0.001=$***$"),footnote_as_chunk = T) %>%
+  add_header_above(c(" " = 1,"Polychoric correlations"=8, "MLR" = 4, "DWLS" = 4))%>% 
+  add_header_above(c(" " = 9,"Factor loadings"=8)) %>% 
+  column_spec(column = c(1),width ="7in",include_thead = T) %>%
+  column_spec(column = c(2:17),width ="0.5in",include_thead = T) %>% 
+  kable_styling(full_width = FALSE) %>% 
+  kable_styling(full_width = F)
+```
+
 ## Version 03 (N=`r dim(dat03)[1]`)
@@
 
+```{r}
+polyw03<- poly(sjlabelled::remove_all_labels(dat03),
+               digits = 2,
+               triangle = "l",
+               fade.ns = F,
+               remove.spaces = T,
+               string.diag = c("$-$","$-$","$-$","$-$","$-$","$-$","$-$","$-$"))
+matpolyw03 <- as.data.frame(read_html(x = polyw03[["page.content"]]) %>% html_table(fill=FALSE)) 
+matpolyw03 <- matpolyw03[-c(9),]
+
+colnames(matpolyw03) <- c("vars","(A)","(B)","(C)","(D)","(E)","(F)","(G)", "(H)")
+
+load03<- data.frame(round(cbind(inspect(fit3_c,what="std")$lambda,inspect(fit3_o,what="std")$lambda),digits = 2))
+load03[load03==c(0.00)] <- NA; names(load03) <- c(1:4,1:4)
+
+table03<- bind_cols(x = matpolyw03,y = load03)
+# %>% mutate(vars=labs)
+
+kable(table03,escape = FALSE,align = "l",col.names = cnametab) %>% 
+  footnote(general = paste0(paste0("$N$= ",nobs(fit3_c)),"; p<0.05=$*$; p<0.01=$**$; p<0.001=$***$"),footnote_as_chunk = T) %>%
+  add_header_above(c(" " = 1,"Polychoric correlations"=8, "MLR" = 4, "DWLS" = 4))%>% 
+  add_header_above(c(" " = 9,"Factor loadings"=8)) %>% 
+  column_spec(column = c(1),width ="7in",include_thead = T) %>%
+  column_spec(column = c(2:17),width ="0.5in",include_thead = T) %>% 
+  kable_styling(full_width = F)
+```
+
 ### Tau-equivalence
@@
 > **NOTE:** The model with equal factor loadings (in Meritocracy Perception) does not significantly deteriorate the model fit.
+
+```{r tabla03, include=FALSE}
+polyw03tau<- poly(sjlabelled::remove_all_labels(dat03),
+               digits = 2,
+               triangle = "l",
+               fade.ns = F,
+               remove.spaces = T,
+               string.diag = c("$-$","$-$","$-$","$-$","$-$","$-$","$-$","$-$"))
+matpolyw03tau <- as.data.frame(read_html(x = polyw03tau[["page.content"]]) %>% html_table(fill=FALSE)) 
+matpolyw03tau <- matpolyw03tau[-c(9),]
+
+colnames(matpolyw03tau) <- c("vars","(A)","(B)","(C)","(D)","(E)","(F)","(G)", "(H)") 
+load03tau<- data.frame(round(cbind(inspect(fit3_c.tau,what="std")$lambda,inspect(fit3_o.tau,what="std")$lambda),digits = 2))
+load03tau[load03tau==c(0.00)] <- NA; names(load03tau) <- c(1:4,1:4)
+
+table03tau<- bind_cols(x = matpolyw03tau,y = load03tau) 
+# %>% mutate(vars=labs)
+
+kable(table03tau,escape = FALSE,align = "l",col.names = cnametab) %>% 
+  footnote(general = paste0(paste0("$N$= ",nobs(fit3_c.tau)),"; p<0.05=$*$; p<0.01=$**$; p<0.001=$***$"),footnote_as_chunk = T) %>%
+  add_header_above(c(" " = 1,"Polychoric correlations"=8, "MLR" = 4, "DWLS" = 4))%>% 
+  add_header_above(c(" " = 9,"Factor loadings"=8)) %>% 
+  column_spec(column = c(1),width ="7in",include_thead = T) %>%
+  column_spec(column = c(2:17),width ="0.5in",include_thead = T) %>% 
+  kable_styling(full_width = F)
+```
 
@@
 
+```{r tabla04, include=FALSE}
+polyw04<- poly(sjlabelled::remove_all_labels(dat04[,-c(1)]),
+               digits = 2,
+               triangle = "l",
+               fade.ns = F,
+               remove.spaces = T,
+               string.diag = c("$-$","$-$","$-$","$-$","$-$","$-$","$-$","$-$"))
+matpolyw04 <- as.data.frame(read_html(x = polyw04[["page.content"]]) %>% html_table(fill=FALSE)) 
+matpolyw04 <- matpolyw04[-c(9),]
+
+colnames(matpolyw04) <- c("vars","(A)","(B)","(C)","(D)","(E)","(F)","(G)", "(H)") 
+load04<- data.frame(round(cbind(inspect(fit4_c,what="std")$lambda,inspect(fit4_o,what="std")$lambda),digits = 2))
+load04[load04==c(0.00)] <- NA; names(load04) <- c(1:4,1:4)
+
+table04<- bind_cols(x = matpolyw04,y = load04) 
+# %>% mutate(vars=labs)
+
+kable(table04,escape = FALSE,align = "l",col.names = cnametab) %>% 
+  footnote(general = paste0(paste0("$N$= ",nobs(fit4_c)),"; p<0.05=$*$; p<0.01=$**$; p<0.001=$***$"),footnote_as_chunk = T) %>%
+  add_header_above(c(" " = 1,"Polychoric correlations"=8, "MLR" = 4, "DWLS" = 4))%>% 
+  add_header_above(c(" " = 9,"Factor loadings"=8)) %>% 
+  column_spec(column = c(1),width ="7in",include_thead = T) %>%
+  column_spec(column = c(2:17),width ="0.5in",include_thead = T) %>% 
+  kable_styling(full_width = F)
+```
+
+# Tables
+
+```{r tb_desc-corr}
+load04<- data.frame(round(inspect(fit4_o,what="std")$lambda,digits = 2))
+load04[load04==c(0.00)] <- NA; names(load04) <- c(1:4)
+
+colnames(matpolyw04) <- c("Variable","A","B","C","D","E","F","G", "H")
+
+sum04b  <- psych::describe(dat04[,-c(1)], fast = T) %>% select(mean,sd,min,max) %>% data.frame() %>% round(digits = 2)
+table04b<- bind_cols(x = sum04b,y = matpolyw04) %>% select(Variable, everything()) %>% mutate(Variable=labs) %>% rename("Mean"=mean, "SD"=sd, "Min"=min,"Max"=max)
+
+cap01 <- "Descriptive statistic"
+sumdat04<- kable(table04b, escape = FALSE,align = "l",linesep = "", caption = cap01) %>% 
+           column_spec(column = 1,width = "10cm",width_min = "12cm",include_thead = TRUE) %>% 
+           footnote(general = paste0(paste0("$N$= ","2141"),"; p<0.05=$*$; p<0.01=$**$; p<0.001=$***$"),footnote_as_chunk = T) %>% 
+           add_header_above(header = c(" "=5,"Polychoric correlations"= 8)) %>% 
+           kable_styling(full_width = TRUE,font_size = 8,position = "center");sumdat04
+save(sumdat04,file = "output/tables/sumdat04.RData")
+```
+
+```{r tb_loads-fitm}
+tb.load<- data.frame(round(cbind(inspect(fit1_o,what="std")$lambda,
+                                 inspect(fit2_o,what="std")$lambda,
+                                 inspect(fit3_o,what="std")$lambda),
+                           digits = 2))
+tb.load[tb.load==c(0.00)] <- NA; names(tb.load) <- c(paste0("M1.",1:4),paste0("M2.",1:4),paste0("M3.",1:4))
+
+for (i in names(tb.load)) {
+  # tb.load[,i] <- sjlabelled::as_character(tb.load[,i])
+  tb.load[,i] <- sprintf(tb.load[,i], fmt = '%#.2f')
+}
+tb.load[tb.load=="NA"] <- ""
+
+#-------#
+fm01<- data.frame(t(data.frame(fitmeasures(fit1_o, output ="matrix")[c("chisq.scaled","df","cfi.scaled","rmsea.scaled"),]))); row.names(fm01) ="M1.1"
+fm02<- data.frame(t(data.frame(fitmeasures(fit2_o, output ="matrix")[c("chisq.scaled","df","cfi.scaled","rmsea.scaled"),]))); row.names(fm02) ="M2.1"
+fm03<- data.frame(t(data.frame(fitmeasures(fit3_o, output ="matrix")[c("chisq.scaled","df","cfi.scaled","rmsea.scaled"),]))); row.names(fm03) ="M3.1"
+
+#------chi2, df------#
+fm04<- round(rbind(fm01,fm02,fm03),3)
+fm04.1 <- fm04 %>% select(chisq.scaled,df) 
+fm04.1$chisq.scaled <- round(x = fm04.1$chisq.scaled,digits = 2)
+fm04.1$df <- round(x = fm04.1$df,digits = 0)
+fm04.1$chi2df <- paste0(fm04.1$chisq.scaled,"(",fm04.1$df,")")
+fm04.1 <- select(fm04.1,"chi2df")
+for (i in names(fm04.1)) {
+  fm04.1[,i] <- as.character(fm04.1[,i])
+}
+
+#------CFI, RMSEA------#
+fm04.2 <- fm04 %>% select(cfi.scaled,rmsea.scaled) 
+for (i in names(fm04.2)) {
+  fm04.2[,i] <- sprintf(fm04.2[,i], fmt = '%#.3f')
+}
+
+fm.df      <- bind_cols(fm04.1,fm04.2)
+fm.df$nobs <- c(nobs(fit1_o),nobs(fit2_o),nobs(fit3_o)) 
+fm.df <- data.frame(t(fm.df)); colnames(fm.df) <- c("M1.1","M2.1","M3.1")
+#------ merge ------#
+tb.fm<- bind_rows(tb.load,fm.df)
+tb.fm<- tb.fm %>% mutate(vars=c(labs,"$\\chi^2\\text{(df)}$","$\\text{CFI}$","$\\text{RMSEA}$","$N$")) %>% select(vars,everything())
+#--table---#
+tb.col  <- c("Variables",1:4,1:4,1:4)
+tb.foot <- paste0("; p<0.05=$*$; p<0.01=$**$; p<0.001=$***$.
+Standardised factor loadings using DWLS estimator.
+CFI = Comparative fit index (scaled), RMSEA = Root mean square error of approximation (scaled)")
+tb.caption <- "Factor loadings and fit measures"
+
+tb.fm01<- kable(tb.fm,escape = FALSE,align = "lcccccccccccc",col.names = tb.col, caption = tb.caption) %>% 
+          kable_styling(full_width = F,font_size = 9) %>% 
+          column_spec(column = 1,width = "10cm",width_min = "12cm",include_thead = TRUE) %>% 
+          column_spec(column = 2:13,width = "1cm",width_min = "1.3cm",include_thead = TRUE) %>% 
+          add_header_above(header = c(" "=1,"Version 1"= 4,"Version 2"= 4,"Version 3"= 4)) %>% 
+          add_header_above(header = c(" "=1,"Factor loadings"= 12)) %>% 
+          row_spec(row = 8,hline_after = TRUE) %>%
+          add_indent(c(9:12)) %>% 
+          footnote(general =tb.foot ,footnote_as_chunk = T);tb.fm01
+save(tb.fm01,file = "output/tables/tb.fm01.RData")
+```
+
 # Summary
 
-## Correlations
+## Correlations plots
 
@@
               "Preference effort","Preference talent","Preference rich family","Preference networks",
-              "Percepction meritocratic","Percepction non meritocratic","Preference meritocratic","Preference non meritocratic")
+              "Perception meritocratic","Perception non meritocratic","Preference meritocratic","Preference non meritocratic")
 
@@
          sizeMan=5,
-         sizeLat = 10)
+         sizeLat = 10,
+         filetype ="png",
+         width = 12,
+         height= 10,
+         filename = "output/images/meas01")
 ```
+
+![](../output/images/meas01.png)
 
@@
 # as_image(sumtable01,file = "output/sumtablew01.png")
-save_kable(sumtable01,file = "output/tables/sumtablew01.html",self_contained = FALSE)
+# save_kable(sumtable01,file = "output/tables/sumtablew01.html",self_contained = FALSE)
 ```
@@
 model04 <- '
-perc_merit =~ perc_effort+perc_talent
-perc_nmerit=~perc_wpart +perc_netw
-pref_merit =~ pref_effort+pref_talent
-pref_nmerit=~pref_wpart +pref_netw'
+perc_merit =~ perc_effort + perc_talent
+perc_nmerit=~ perc_wpart  + perc_netw
+pref_merit =~ pref_effort + pref_talent
+pref_nmerit=~ pref_wpart  + pref_netw'
 
@@
       format = "markdown",digits = 3,caption = "Fit indices")
+gr01=standardizedsolution(fit4_c_mg) %>% filter(op=="=~") %>% select(group,lhs,rhs,est.std)
+gr02=standardizedsolution(fit4_o_mg) %>% filter(op=="=~") %>% select(group,lhs,rhs,est.std)
+gr01$group=dplyr::recode(gr01$group, `1` = "mod03", `2` = "mod02", `3` = "mod01")
+gr02$group=dplyr::recode(gr02$group, `1` = "mod03", `2` = "mod02", `3` = "mod01")
 
 cnames <- c("Grupo","Factor","Indicator","Loading (MLR)","Loading (DWLS)")
-kable(left_join(x = standardizedsolution(fit4_c_mg) %>% filter(op=="=~") %>% select(group,lhs,rhs,est.std),
-                y = standardizedsolution(fit4_o_mg) %>% filter(op=="=~") %>% select(group,lhs,rhs,est.std),
+kable(left_join(x =gr01 ,
+                y =gr02 ,
                 c("group","lhs","rhs")),
@@
 
-<!-- > **NOTE:** Cargas factoriales de segundo orden son bajas. El ajuste del modelo no es bueno. -->
-
 > **NOTE:** Second order factor loads are low. The fit of the model is not good.
@@
 
-<!-- > **NOTE:** El modelo estimado con MLR con factores de segundo orden "meritocracia/ no meritocracia" no converge.
-EL modelo estimado a través de DWLS converge, posee cargas muy bajas y una muy alta (merit~pref_merit).
-Fit es malo (RMSEA >.10) -->
-
 > **NOTE:**  The model estimated with MLR with second order factors "meritocracy / non meritocracy" does not converge.
@@
 
-# Wave 02 {.tabset .tabset-fade .tabset-pills}
+<!-- # Wave 02 {.tabset .tabset-fade .tabset-pills} -->
 
-Each sample correspond to one of the three conditions generated by the randomized assignment.
+<!-- Each sample correspond to one of the three conditions generated by the randomized assignment. -->
 
-1. **Control group:** the group that has been assigned to the control group (placebo condition)
+<!-- 1. **Control group:** the group that has been assigned to the control group (placebo condition) -->
 
-2. **Poverty condition sample** the group that has been assigned to the Poverty treatment condition.
+<!-- 2. **Poverty condition sample** the group that has been assigned to the Poverty treatment condition. -->
 
-3. **Inequality condition sample:** the group that has been assigned to the Inequality treatment condition.
+<!-- 3. **Inequality condition sample:** the group that has been assigned to the Inequality treatment condition. -->
 
-4. **Complete sample:** Control, Poverty and Inequality treatment.
+<!-- 4. **Complete sample:** Control, Poverty and Inequality treatment. -->
 
-```{r}
-data02 <- sjlabelled::read_spss(path = "input/data/original/Estudio_3_ola2.sav",verbose = FALSE)
-load(file = "input/data/proc/datawide.RData")
-dat01 <- data02 %>% select(starts_with("meritv03_p"))
-```
+<!-- ```{r} -->
+<!-- data02 <- sjlabelled::read_spss(path = "input/data/original/Estudio_3_ola2.sav",verbose = FALSE) -->
+<!-- load(file = "input/data/proc/datawide.RData") -->
+<!-- dat01 <- data02 %>% select(starts_with("meritv03_p")) -->
+<!-- ``` -->
 
-**ATE on observed indicators**
+<!-- **ATE on observed indicators** -->
 
-```{r, out.width="75%",results='asis'}
-m01<- lm(perc_effort_w2~treat,data = wide01)
-m02<- lm(perc_talent_w2~treat,data = wide01)
-m03<- lm(perc_wpart_w2~treat, data = wide01)
-m04<- lm(perc_netw_w2~treat,  data = wide01)
-#--------------------------------------------#
-m05<- lm(pref_effort_w2~treat,data = wide01)
-m06<- lm(pref_talent_w2~treat,data = wide01)
-m07<- lm(pref_wpart_w2 ~treat, data = wide01)
-m08<- lm(pref_netw_w2  ~treat,  data = wide01)
-#--------------------------------------------#
-stargazer::stargazer(m01,m02,m03,m04,type = "html")
-stargazer::stargazer(m05,m06,m07,m08,type = "html")
-```
+<!-- ```{r, out.width="75%",results='asis'} -->
+<!-- m01<- lm(perc_effort_w2~treat,data = wide01) -->
+<!-- m02<- lm(perc_talent_w2~treat,data = wide01) -->
+<!-- m03<- lm(perc_wpart_w2~treat, data = wide01) -->
+<!-- m04<- lm(perc_netw_w2~treat,  data = wide01) -->
+<!-- #--------------------------------------------# -->
+<!-- m05<- lm(pref_effort_w2~treat,data = wide01) -->
+<!-- m06<- lm(pref_talent_w2~treat,data = wide01) -->
+<!-- m07<- lm(pref_wpart_w2 ~treat, data = wide01) -->
+<!-- m08<- lm(pref_netw_w2  ~treat,  data = wide01) -->
+<!-- #--------------------------------------------# -->
+<!-- stargazer::stargazer(m01,m02,m03,m04,type = "html") -->
+<!-- stargazer::stargazer(m05,m06,m07,m08,type = "html") -->
+<!-- ``` -->
 
-<!-- > **NOTE**: Los tratamientos de información no se asocian de manera estadísticamente significativa con ninguno de los Indicatores de la escala de meritocracia.   -->
+<!-- > **NOTE**: There are no statistically significant effects of the treatment conditions on the observed indicators of the meritocracy scale. -->
 
-> **NOTE**: There are no statistically significant effects of the treatment conditions on the observed indicators of the meritocracy scale.
+<!-- ## Control group -->
 
-## Control group
+<!-- ```{r} -->
+<!-- w02con <- data02 %>% filter(FL_6_DO_Control==1) %>% select(starts_with("meritv03_p")) -->
+<!-- ``` -->
 
-```{r}
-w02con <- data02 %>% filter(FL_6_DO_Control==1) %>% select(starts_with("meritv03_p"))
-```
+<!-- ```{r results='asis'} -->
+<!-- stargazer(w02con,type = "html", -->
+<!--           title = "Descriptive statistics", -->
+<!--           font.size = "small", -->
+<!--           median = TRUE, -->
+<!--           header=FALSE, -->
+<!--           digit.separator = "", -->
+<!--           digits = 2) -->
+<!-- ``` -->
 
-```{r results='asis'}
-stargazer(w02con,type = "html",
-          title = "Descriptive statistics",
-          font.size = "small",
-          median = TRUE,
-          header=FALSE,
-          digit.separator = "",
-          digits = 2)
-```
+<!-- ```{r} -->
+<!-- corw02_1<-polychoric(w02con) -->
+<!-- rownames(corw02_1$rho) <-c("1. pc_effort", "2. pc_talent","3. pc_wpart","4. pc_netw","5. pf_effort","6. pf_talent","7. pf_wpart", "8. pf_netw") -->
+<!-- colnames(corw02_1$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)") -->
+<!-- corrplot(corw02_1$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE) -->
+<!-- ``` -->
 
-```{r}
-corw02_1<-polychoric(w02con)
-rownames(corw02_1$rho) <-c("1. pc_effort", "2. pc_talent","3. pc_wpart","4. pc_netw","5. pf_effort","6. pf_talent","7. pf_wpart", "8. pf_netw")
-colnames(corw02_1$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)")
-corrplot(corw02_1$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
-```
+<!-- ```{r,results='hold'} -->
+<!-- model03 <- ' -->
+<!-- perc_merit=~meritv03_perc_effort+meritv03_perc_talent -->
+<!-- perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw -->
+<!-- pref_merit=~meritv03_pref_effort+meritv03_pref_talent -->
+<!-- pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw' -->
 
-```{r,results='hold'}
-model03 <- '
-perc_merit=~meritv03_perc_effort+meritv03_perc_talent
-perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw
-pref_merit=~meritv03_pref_effort+meritv03_pref_talent
-pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw'
+<!-- fitw02a_c <- cfa(model = model03,data = w02con,estimator="MLR") # Continuous/ estimator ML Robust -->
 
-fitw02a_c <- cfa(model = model03,data = w02con,estimator="MLR") # Continuous/ estimator ML Robust
+<!-- fitw02a_o <- cfa(model = model03,data = w02con,ordered = c("meritv03_perc_effort","meritv03_perc_talent", -->
+<!--                                                            "meritv03_perc_wpart","meritv03_perc_netw", -->
+<!--                                                            "meritv03_pref_effort","meritv03_pref_talent", -->
+<!--                                                            "meritv03_pref_wpart","meritv03_pref_netw")) -->
+<!-- #---------------------------------------------------# -->
+<!-- kable(rbind(Continuous=fitmeasures(fitw02a_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")], -->
+<!--             Ordinal= fitmeasures(fitw02a_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]), -->
+<!--       format = "markdown",digits = 3) -->
 
-fitw02a_o <- cfa(model = model03,data = w02con,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
-                                                           "meritv03_perc_wpart","meritv03_perc_netw",
-                                                           "meritv03_pref_effort","meritv03_pref_talent",
-                                                           "meritv03_pref_wpart","meritv03_pref_netw"))
-#---------------------------------------------------#
-kable(rbind(Continuous=fitmeasures(fitw02a_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-            Ordinal= fitmeasures(fitw02a_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
-      format = "markdown",digits = 3)
+<!-- cnames <- c("Factor","Indicator","Loading (MLR)","Loading (DWLS)") -->
+<!-- kable(left_join(x = standardizedsolution(fitw02a_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std), -->
+<!--                 y = standardizedsolution(fitw02a_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std), -->
+<!--                 c("lhs","rhs")), -->
+<!--       format = "markdown",digits = 2,col.names = cnames) -->
+<!-- ``` -->
 
-cnames <- c("Factor","Indicator","Loading (MLR)","Loading (DWLS)")
-kable(left_join(x = standardizedsolution(fitw02a_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
-                y = standardizedsolution(fitw02a_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
-                c("lhs","rhs")),
-      format = "markdown",digits = 2,col.names = cnames)
-```
+<!-- ## Poverty condition -->
 
-## Poverty condition
+<!-- ```{r} -->
+<!-- w02pob <- data02 %>% filter(FL_6_DO_Tratamiento==1) %>% select(starts_with("meritv03_p")) -->
+<!-- ``` -->
 
-```{r}
-w02pob <- data02 %>% filter(FL_6_DO_Tratamiento==1) %>% select(starts_with("meritv03_p"))
-```
+<!-- ```{r results='asis'} -->
+<!-- stargazer(w02pob,type = "html", -->
+<!--           title = "Descriptive statistics", -->
+<!--           font.size = "small", -->
+<!--           median = TRUE, -->
+<!--           header=FALSE, -->
+<!--           digit.separator = "", -->
+<!--           digits = 2) -->
+<!-- ``` -->
 
-```{r results='asis'}
-stargazer(w02pob,type = "html",
-          title = "Descriptive statistics",
-          font.size = "small",
-          median = TRUE,
-          header=FALSE,
-          digit.separator = "",
-          digits = 2)
-```
+<!-- ```{r} -->
+<!-- corw02_2<-polychoric(w02pob) -->
+<!-- rownames(corw02_2$rho) <-c("1. pc_effort", "2. pc_talent","3. pc_wpart","4. pc_netw","5. pf_effort","6. pf_talent","7. pf_wpart", "8. pf_netw") -->
+<!-- colnames(corw02_2$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)") -->
+<!-- corrplot(corw02_2$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE) -->
+<!-- ``` -->
 
-```{r}
-corw02_2<-polychoric(w02pob)
-rownames(corw02_2$rho) <-c("1. pc_effort", "2. pc_talent","3. pc_wpart","4. pc_netw","5. pf_effort","6. pf_talent","7. pf_wpart", "8. pf_netw")
-colnames(corw02_2$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)")
-corrplot(corw02_2$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
-```
+<!-- ```{r,results='hold'} -->
+<!-- model03 <- ' -->
+<!-- perc_merit=~meritv03_perc_effort+meritv03_perc_talent -->
+<!-- perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw -->
+<!-- pref_merit=~meritv03_pref_effort+meritv03_pref_talent -->
+<!-- pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw' -->
 
-```{r,results='hold'}
-model03 <- '
-perc_merit=~meritv03_perc_effort+meritv03_perc_talent
-perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw
-pref_merit=~meritv03_pref_effort+meritv03_pref_talent
-pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw'
+<!-- fitw02b_c <- cfa(model = model03,data = w02pob,estimator="MLR") # Continuous/ estimator ML Robust -->
 
-fitw02b_c <- cfa(model = model03,data = w02pob,estimator="MLR") # Continuous/ estimator ML Robust
+<!-- fitw02b_o <- cfa(model = model03,data = w02pob,ordered = c("meritv03_perc_effort","meritv03_perc_talent", -->
+<!--                                                            "meritv03_perc_wpart","meritv03_perc_netw", -->
+<!--                                                            "meritv03_pref_effort","meritv03_pref_talent", -->
+<!--                                                            "meritv03_pref_wpart","meritv03_pref_netw")) -->
 
-fitw02b_o <- cfa(model = model03,data = w02pob,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
-                                                           "meritv03_perc_wpart","meritv03_perc_netw",
-                                                           "meritv03_pref_effort","meritv03_pref_talent",
-                                                           "meritv03_pref_wpart","meritv03_pref_netw"))
+<!-- #---------------------------------------------------# -->
+<!-- kable(rbind(Continuous=fitmeasures(fitw02b_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")], -->
+<!--             Ordinal= fitmeasures(fitw02b_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]), -->
+<!--       format = "markdown",digits = 3) -->
 
-#---------------------------------------------------#
-kable(rbind(Continuous=fitmeasures(fitw02b_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-            Ordinal= fitmeasures(fitw02b_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
-      format = "markdown",digits = 3)
+<!-- cnames <- c("Factor","Indicator","Loading (MLR)","Loading (DWLS)") -->
+<!-- kable(left_join(x = standardizedsolution(fitw02b_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std), -->
+<!--                 y = standardizedsolution(fitw02b_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std), -->
+<!--                 c("lhs","rhs")), -->
+<!--       format = "markdown",digits = 2,col.names = cnames) -->
+<!-- ``` -->
 
-cnames <- c("Factor","Indicator","Loading (MLR)","Loading (DWLS)")
-kable(left_join(x = standardizedsolution(fitw02b_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
-                y = standardizedsolution(fitw02b_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
-                c("lhs","rhs")),
-      format = "markdown",digits = 2,col.names = cnames)
-```
+<!-- ## Inequality condition -->
 
-## Inequality condition
+<!-- ```{r} -->
+<!-- w02des <- data02 %>% filter(FL_6_DO_Tratamiento_desigualdad==1) %>% select(starts_with("meritv03_p")) -->
+<!-- ``` -->
 
-```{r}
-w02des <- data02 %>% filter(FL_6_DO_Tratamiento_desigualdad==1) %>% select(starts_with("meritv03_p"))
-```
+<!-- ```{r results='asis'} -->
+<!-- stargazer(w02des,type = "html", -->
+<!--           title = "Descriptive statistics", -->
+<!--           font.size = "small", -->
+<!--           median = TRUE, -->
+<!--           header=FALSE, -->
+<!--           digit.separator = "", -->
+<!--           digits = 2) -->
+<!-- ``` -->
 
-```{r results='asis'}
-stargazer(w02des,type = "html",
-          title = "Descriptive statistics",
-          font.size = "small",
-          median = TRUE,
-          header=FALSE,
-          digit.separator = "",
-          digits = 2)
-```
+<!-- ```{r} -->
+<!-- corw02_3<-polychoric(w02des) -->
+<!-- rownames(corw02_3$rho) <-c("1. pc_effort", "2. pc_talent","3. pc_wpart","4. pc_netw","5. pf_effort","6. pf_talent","7. pf_wpart", "8. pf_netw") -->
+<!-- colnames(corw02_3$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)") -->
+<!-- corrplot(corw02_3$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE) -->
+<!-- ``` -->
 
-```{r}
-corw02_3<-polychoric(w02des)
-rownames(corw02_3$rho) <-c("1. pc_effort", "2. pc_talent","3. pc_wpart","4. pc_netw","5. pf_effort","6. pf_talent","7. pf_wpart", "8. pf_netw")
-colnames(corw02_3$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)")
-corrplot(corw02_3$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
-```
+<!-- ```{r,results='hold'} -->
+<!-- model03 <- ' -->
+<!-- perc_merit=~meritv03_perc_effort+meritv03_perc_talent -->
+<!-- perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw -->
+<!-- pref_merit=~meritv03_pref_effort+meritv03_pref_talent -->
+<!-- pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw' -->
 
-```{r,results='hold'}
-model03 <- '
-perc_merit=~meritv03_perc_effort+meritv03_perc_talent
-perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw
-pref_merit=~meritv03_pref_effort+meritv03_pref_talent
-pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw'
+<!-- fitw02c_c <- cfa(model = model03,data = w02des,estimator="MLR") # Continuous/ estimator ML Robust -->
 
-fitw02c_c <- cfa(model = model03,data = w02des,estimator="MLR") # Continuous/ estimator ML Robust
+<!-- fitw02c_o <- cfa(model = model03,data = w02des,ordered = c("meritv03_perc_effort","meritv03_perc_talent", -->
+<!--                                                            "meritv03_perc_wpart","meritv03_perc_netw", -->
+<!--                                                            "meritv03_pref_effort","meritv03_pref_talent", -->
+<!--                                                            "meritv03_pref_wpart","meritv03_pref_netw")) -->
+<!-- #---------------------------------------------------# -->
+<!-- kable(rbind(Continuous=fitmeasures(fitw02c_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")], -->
+<!--             Ordinal= fitmeasures(fitw02c_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]), -->
+<!--       format = "markdown",digits = 3) -->
 
-fitw02c_o <- cfa(model = model03,data = w02des,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
-                                                           "meritv03_perc_wpart","meritv03_perc_netw",
-                                                           "meritv03_pref_effort","meritv03_pref_talent",
-                                                           "meritv03_pref_wpart","meritv03_pref_netw"))
-#---------------------------------------------------#
-kable(rbind(Continuous=fitmeasures(fitw02c_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-            Ordinal= fitmeasures(fitw02c_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
-      format = "markdown",digits = 3)
+<!-- cnames <- c("Factor","Indicator","Loading (MLR)","Loading (DWLS)") -->
+<!-- kable(left_join(x = standardizedsolution(fitw02c_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std), -->
+<!--                 y = standardizedsolution(fitw02c_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std), -->
+<!--                 c("lhs","rhs")), -->
+<!--       format = "markdown",digits = 2,col.names = cnames) -->
+<!-- ``` -->
 
-cnames <- c("Factor","Indicator","Loading (MLR)","Loading (DWLS)")
-kable(left_join(x = standardizedsolution(fitw02c_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
-                y = standardizedsolution(fitw02c_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
-                c("lhs","rhs")),
-      format = "markdown",digits = 2,col.names = cnames)
-```
+<!-- ## Complete sample -->
 
-## Complete sample
+<!-- ```{r results='asis'} -->
+<!-- stargazer(dat01,type = "html", -->
+<!--           title = "Descriptive statistics", -->
+<!--           font.size = "small", -->
+<!--           median = TRUE, -->
+<!--           header=FALSE, -->
+<!--           digit.separator = "", -->
+<!--           digits = 2) -->
+<!-- ``` -->
 
-```{r results='asis'}
-stargazer(dat01,type = "html",
-          title = "Descriptive statistics",
-          font.size = "small",
-          median = TRUE,
-          header=FALSE,
-          digit.separator = "",
-          digits = 2)
-```
+<!-- ```{r} -->
+<!-- dat01cor<- select(dat01,meritv03_perc_effort,meritv03_perc_talent, -->
+<!--                         meritv03_perc_wpart,meritv03_perc_netw, -->
+<!--                         meritv03_pref_effort,meritv03_pref_talent, -->
+<!--                         meritv03_pref_wpart,meritv03_pref_netw) -->
+<!-- cor01<-polychoric(dat01cor) -->
+<!-- rownames(cor01$rho) <-c("1. pc_effort", "2. pc_talent","3. pc_wpart","4. pc_netw","5. pf_effort","6. pf_talent","7. pf_wpart", "8. pf_netw") -->
+<!-- colnames(cor01$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)") -->
+<!-- corrplot(cor01$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE) -->
+<!-- ``` -->
 
-```{r}
-dat01cor<- select(dat01,meritv03_perc_effort,meritv03_perc_talent,
-                        meritv03_perc_wpart,meritv03_perc_netw,
-                        meritv03_pref_effort,meritv03_pref_talent,
-                        meritv03_pref_wpart,meritv03_pref_netw)
-cor01<-polychoric(dat01cor)
-rownames(cor01$rho) <-c("1. pc_effort", "2. pc_talent","3. pc_wpart","4. pc_netw","5. pf_effort","6. pf_talent","7. pf_wpart", "8. pf_netw")
-colnames(cor01$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)")
-corrplot(cor01$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
-```
+<!-- ```{r,results='hold'} -->
+<!-- model03 <- ' -->
+<!-- perc_merit=~meritv03_perc_effort+meritv03_perc_talent -->
+<!-- perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw -->
+<!-- pref_merit=~meritv03_pref_effort+meritv03_pref_talent -->
+<!-- pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw' -->
 
-```{r,results='hold'}
-model03 <- '
-perc_merit=~meritv03_perc_effort+meritv03_perc_talent
-perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw
-pref_merit=~meritv03_pref_effort+meritv03_pref_talent
-pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw'
+<!-- fitw02_c <- cfa(model = model03,data = dat01,estimator="MLR",std.lv=FALSE) # Continuous/ estimator ML Robust -->
 
-fitw02_c <- cfa(model = model03,data = dat01,estimator="MLR",std.lv=FALSE) # Continuous/ estimator ML Robust
+<!-- fitw02_o <- cfa(model = model03,data = dat01,ordered = c("meritv03_perc_effort","meritv03_perc_talent", -->
+<!--                                                          "meritv03_perc_wpart","meritv03_perc_netw", -->
+<!--                                                          "meritv03_pref_effort","meritv03_pref_talent", -->
+<!--                                                          "meritv03_pref_wpart","meritv03_pref_netw")) -->
+<!-- #---------------------------------------------------# -->
+<!-- kable(rbind(Continuous=fitmeasures(fitw02_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")], -->
+<!--             Ordinal= fitmeasures(fitw02_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]), -->
+<!--       format = "markdown",digits = 3) -->
 
-fitw02_o <- cfa(model = model03,data = dat01,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
-                                                         "meritv03_perc_wpart","meritv03_perc_netw",
-                                                         "meritv03_pref_effort","meritv03_pref_talent",
-                                                         "meritv03_pref_wpart","meritv03_pref_netw"))
-#---------------------------------------------------#
-kable(rbind(Continuous=fitmeasures(fitw02_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-            Ordinal= fitmeasures(fitw02_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
-      format = "markdown",digits = 3)
+<!-- cnames <- c("Factor","Indicator","Loading (MLR)","Loading (DWLS)") -->
+<!-- kable(left_join(x = standardizedsolution(fitw02_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std), -->
+<!--                 y = standardizedsolution(fitw02_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std), -->
+<!--                 c("lhs","rhs")), -->
+<!--       format = "markdown",digits = 2,col.names = cnames) -->
+<!-- ``` -->
 
-cnames <- c("Factor","Indicator","Loading (MLR)","Loading (DWLS)")
-kable(left_join(x = standardizedsolution(fitw02_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
-                y = standardizedsolution(fitw02_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
-                c("lhs","rhs")),
-      format = "markdown",digits = 2,col.names = cnames)
-```
+<!-- > **NOTE**: El ajuste de muestra completa en wave 02 con estimator MLR (robusto) es bastante bueno (CFI, RMSEA).
+El ajuste con DWLS es bastante bueno, CFI mejora, RMSEA aumenta un poco (0.59).
+-->
 
-> **NOTE**: El ajuste de muestra completa en wave 02 con estimator MLR (robusto) es bastante bueno (CFI, RMSEA). El ajuste con DWLS es bastante bueno, CFI mejora, RMSEA aumenta un poco (0.59).
+<!-- ### Second order CFA -->
 
-### Second order CFA
+<!-- ```{r} -->
+<!-- modw2 <- ' -->
+<!-- perc_merit=~meritv03_perc_effort+meritv03_perc_talent -->
+<!-- perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw -->
+<!-- pref_merit=~meritv03_pref_effort+meritv03_pref_talent -->
+<!-- pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw -->
+<!-- merit =~perc_merit+pref_merit -->
+<!-- nmerit=~perc_nmerit+pref_nmerit' -->
 
-```{r}
-modw2 <- '
-perc_merit=~meritv03_perc_effort+meritv03_perc_talent
-perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw
-pref_merit=~meritv03_pref_effort+meritv03_pref_talent
-pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw
-merit =~perc_merit+pref_merit
-nmerit=~perc_nmerit+pref_nmerit'
+<!-- # mod2_merit_c <- cfa(model = modw2,data = dat01) # Continuous/ estimator ML Robust -->
 
-# mod2_merit_c <- cfa(model = modw2,data = dat01) # Continuous/ estimator ML Robust
+<!-- mod2_merit_o <- cfa(model = modw2,data = dat01,ordered = c("meritv03_perc_effort","meritv03_perc_talent", -->
+<!--                                                            "meritv03_perc_wpart","meritv03_perc_netw", -->
+<!--                                                            "meritv03_pref_effort","meritv03_pref_talent", -->
+<!--                                                            "meritv03_pref_wpart","meritv03_pref_netw")) -->
 
-mod2_merit_o <- cfa(model = modw2,data = dat01,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
-                                                           "meritv03_perc_wpart","meritv03_perc_netw",
-                                                           "meritv03_pref_effort","meritv03_pref_talent",
-                                                           "meritv03_pref_wpart","meritv03_pref_netw"))
+<!-- #---------------------------------------------------# -->
+<!-- kable(rbind(Ordinal=fitmeasures(mod2_merit_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]), -->
+<!--       format = "markdown",digits = 3) -->
 
-#---------------------------------------------------#
-kable(rbind(Ordinal=fitmeasures(mod2_merit_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
-      format = "markdown",digits = 3)
+<!-- cnames <- c("Factor","Indicator","Loading (DWLS)") -->
+<!-- kable(standardizedsolution(mod2_merit_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std), -->
+<!--       format = "markdown",digits = 2,col.names = cnames) -->
+<!-- ``` -->
 
-cnames <- c("Factor","Indicator","Loading (DWLS)")
-kable(standardizedsolution(mod2_merit_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
-      format = "markdown",digits = 2,col.names = cnames)
-```
+<!-- > * Second order factor using MLR does not converge. -->
+<!-- > * Second order factor using DWLS converges, but loads are not good and the fit is not. -->
 
-<!-- > * Factor de segundo orden usando MLR no converge. -->
-<!-- > * Factor de segundo orden usando DWLS converge, pero las cargas no son buenas y el ajuste tampoco. -->
+<!-- # Summary fit indices wave 02 -->
 
-> * Second order factor using MLR does not converge.
-> * Second order factor using DWLS converges, but loads are not good and the fit is not.
+<!-- ```{r} -->
+<!-- sum_fitw02<- bind_rows(fitmeasures(fitw02a_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")], -->
+<!--                        fitmeasures(fitw02a_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")], -->
+<!--                        fitmeasures(fitw02b_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")], -->
+<!--                        fitmeasures(fitw02b_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")], -->
+<!--                        fitmeasures(fitw02c_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")], -->
+<!--                        fitmeasures(fitw02c_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")], -->
+<!--                        fitmeasures(fitw02_c)[ c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")], -->
+<!--                        fitmeasures(fitw02_o)[ c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],) -->
 
-# Summary fit indices wave 02
+<!-- sum_fitw02$mod <- c("Model 1","Model 1","Model 2","Model 2","Model 3","Model 3","Model 4","Model 4") -->
+<!-- sum_fitw02$est <- c("MLR","DWLS","MLR","DWLS","MLR","DWLS","MLR","DWLS") -->
+<!-- sum_fitw02 <- select(sum_fitw02,mod,est,everything()) -->
+<!-- colnames <- c("Model","Estimator","$\\chi^2$","df","CFI","CFI.sca","CFI.rob","RMSEA","RMSEA.sca","RMSEA.rob") -->
 
-```{r}
-sum_fitw02<- bind_rows(fitmeasures(fitw02a_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                       fitmeasures(fitw02a_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                       fitmeasures(fitw02b_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                       fitmeasures(fitw02b_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                       fitmeasures(fitw02c_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                       fitmeasures(fitw02c_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                       fitmeasures(fitw02_c)[ c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                       fitmeasures(fitw02_o)[ c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],)
+<!-- kable(sum_fitw02,digits = 3,format = "html",row.names = F,booktabs=T, caption = "Summary fit indices wave 02", col.names = colnames, escape = F) %>% -->
+<!--   kable_styling(full_width = F) %>% -->
+<!--   collapse_rows(columns = 1,valign = "middle") %>% -->
+<!--   footnote(number = c("Model 1: Control group sample", -->
+<!--                       "Model 2: Poverty treatment condition sample", -->
+<!--                       "Model 3: Inequality treatment condition sample", -->
+<!--                       "Model 4: Complete sample")) -->
+<!-- ``` -->
 
-sum_fitw02$mod <- c("Model 1","Model 1","Model 2","Model 2","Model 3","Model 3","Model 4","Model 4")
-sum_fitw02$est <- c("MLR","DWLS","MLR","DWLS","MLR","DWLS","MLR","DWLS")
-sum_fitw02 <- select(sum_fitw02,mod,est,everything())
-colnames <- c("Model","Estimator","$\\chi^2$","df","CFI","CFI.sca","CFI.rob","RMSEA","RMSEA.sca","RMSEA.rob")
+<!-- # Wave 03 -->
 
-kable(sum_fitw02,digits = 3,format = "html",row.names = F,booktabs=T, caption = "Summary fit indices wave 02", col.names = colnames, escape = F) %>%
-  kable_styling(full_width = F) %>%
-  collapse_rows(columns = 1,valign = "middle") %>%
-  footnote(number = c("Model 1: Control group sample",
-                      "Model 2: Poverty treatment condition sample",
-                      "Model 3: Inequality treatment condition sample",
-                      "Model 4: Complete sample"))
-```
+<!-- ```{r} -->
+<!-- data03 <- sjlabelled::read_spss(path = "input/data/original/Estudio_3_ola3.sav",verbose = FALSE) -->
+<!-- dat01 <- data03 %>% select(starts_with("meritv03_p")) -->
+<!-- ``` -->
 
-# Wave 03
+<!-- ```{r echo=TRUE,results='asis'} -->
+<!-- stargazer(dat01,type = "html", -->
+<!--           title = "Descriptive statistics", -->
+<!--           font.size = "small", -->
+<!--           median = TRUE, -->
+<!--           header=FALSE, -->
+<!--           digit.separator = "", -->
+<!--           digits = 2) -->
+<!-- ``` -->
 
-```{r}
-data03 <- sjlabelled::read_spss(path = "input/data/original/Estudio_3_ola3.sav",verbose = FALSE)
-dat01 <- data03 %>% select(starts_with("meritv03_p"))
-```
+<!-- ```{r} -->
+<!-- model03 <- ' -->
+<!-- perc_merit=~meritv03_perc_effort+meritv03_perc_talent -->
+<!-- perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw -->
+<!-- pref_merit=~meritv03_pref_effort+meritv03_pref_talent -->
+<!-- pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw' -->
 
-```{r echo=TRUE,results='asis'}
-stargazer(dat01,type = "html",
-          title = "Descriptive statistics",
-          font.size = "small",
-          median = TRUE,
-          header=FALSE,
-          digit.separator = "",
-          digits = 2)
-```
+<!-- fitw03c_c <- cfa(model = model03,data = dat01,estimator="MLR") # Continuous/ estimator ML Robust -->
 
-```{r}
-model03 <- '
-perc_merit=~meritv03_perc_effort+meritv03_perc_talent
-perc_nmerit=~meritv03_perc_wpart+meritv03_perc_netw
-pref_merit=~meritv03_pref_effort+meritv03_pref_talent
-pref_nmerit=~meritv03_pref_wpart+meritv03_pref_netw'
+<!-- fitw03c_o <- cfa(model = model03,data = dat01,ordered = c("meritv03_perc_effort","meritv03_perc_talent", -->
+<!--                                                           "meritv03_perc_wpart","meritv03_perc_netw", -->
+<!--                                                           "meritv03_pref_effort","meritv03_pref_talent", -->
+<!--                                                           "meritv03_pref_wpart","meritv03_pref_netw")) -->
+<!-- #---------------------------------------------------# -->
+<!-- kable(rbind(Continuous=fitmeasures(fitw03c_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")], -->
+<!--             Ordinal= fitmeasures(fitw03c_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]), -->
+<!--       format = "markdown",digits = 3) -->
 
-fitw03c_c <- cfa(model = model03,data = dat01,estimator="MLR") # Continuous/ estimator ML Robust
+<!-- cnames <- c("Factor","Indicator","Loading (MLR)","Loading (DWLS)") -->
+<!-- kable(left_join(x = standardizedsolution(fitw03c_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std), -->
+<!--                 y = standardizedsolution(fitw03c_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std), -->
+<!--                 c("lhs","rhs")), -->
+<!--       format = "markdown",digits = 2,col.names = cnames) -->
+<!-- ``` -->
+<!-- **summary** -->
 
-fitw03c_o <- cfa(model = model03,data = dat01,ordered = c("meritv03_perc_effort","meritv03_perc_talent",
-                                                          "meritv03_perc_wpart","meritv03_perc_netw",
-                                                          "meritv03_pref_effort","meritv03_pref_talent",
-                                                          "meritv03_pref_wpart","meritv03_pref_netw"))
-#---------------------------------------------------#
-kable(rbind(Continuous=fitmeasures(fitw03c_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-            Ordinal= fitmeasures(fitw03c_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
-      format = "markdown",digits = 3)
+<!-- > * Adjustment using MLR estimator is good (CFI = 0.971, RMSEA = 0.052). The chi2 (robust) is 62,421 -->
+<!-- > * Fit using DWLS estimator is good (CFI = 0.974, RMSEA = 0.067). The chi2 (robust) is 106,111 -->
 
-cnames <- c("Factor","Indicator","Loading (MLR)","Loading (DWLS)")
-kable(left_join(x = standardizedsolution(fitw03c_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
-                y = standardizedsolution(fitw03c_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
-                c("lhs","rhs")),
-      format = "markdown",digits = 2,col.names = cnames)
-```
-**summary**
+<!-- # Summary fit indices wave 03 -->
 
-<!-- > * Ajuste usando estimator MLR  es bueno (CFI = 0.971, RMSEA = 0.052). El chi2 (robusto) es 62.421   -->
-<!-- > * Ajuste usando estimator DWLS es bueno (CFI = 0.974, RMSEA = 0.067). El chi2 (robusto) es 106.111 -->
+<!-- ```{r} -->
+<!-- sum_fitw03<- bind_rows(fitmeasures(fitw03c_c)[ c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")], -->
+<!--                        fitmeasures(fitw03c_o)[ c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]) -->
 
-> * Adjustment using MLR estimator is good (CFI = 0.971, RMSEA = 0.052). The chi2 (robust) is 62,421
-> * Fit using DWLS estimator is good (CFI = 0.974, RMSEA = 0.067). The chi2 (robust) is 106,111
+<!-- sum_fitw03$mod <- c("Model 1","Model 1") -->
+<!-- sum_fitw03$est <- c("Continuous","Ordinal") -->
+<!-- sum_fitw03 <- select(sum_fitw03,mod,est,everything()) -->
 
-# Summary fit indices wave 03
-
-```{r}
-sum_fitw03<- bind_rows(fitmeasures(fitw03c_c)[ c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
-                       fitmeasures(fitw03c_o)[ c("cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")])
-
-sum_fitw03$mod <- c("Model 1","Model 1")
-sum_fitw03$est <- c("Continuous","Ordinal")
-sum_fitw03 <- select(sum_fitw03,mod,est,everything())
-
-kable(sum_fitw03,digits = 3,format = "html",row.names = F,booktabs=T,caption = "Summary fit indices wave 03") %>%
-  kable_styling(full_width = F) %>%
-  collapse_rows(columns = 1,valign = "middle") %>%
-  footnote(number = c("Model 1: Complete sample"))
-```
+<!-- kable(sum_fitw03,digits = 3,format = "html",row.names = F,booktabs=T,caption = "Summary fit indices wave 03") %>% -->
+<!--   kable_styling(full_width = F) %>% -->
+<!--   collapse_rows(columns = 1,valign = "middle") %>% -->
+<!--   footnote(number = c("Model 1: Complete sample")) -->
+<!-- ``` -->

### 2020-04-22 docs/paper.Rmd

@@
 
+The study design consist in two main stages.
+The first stage was the questionnaire design, in which we collect a serie of background variables, including sex, age and educational level.
+On this regard, one of the main aspects of our design is that we tested three different application modalities, in order to observe how the display order of the indicators could affect the model fit.
+We decide to randomize the assignment of the subject to each application modality, and as a result we obtain three groups within the full sample.
+The first combination corresponds to the group that answered the questions that were sorted by the Perceptions/Preferences dimension.
+The second combination corresponds to the group in which the questions are displayed following a topic-sorted way, this means that the effort, talent, rich family and networks questions for perception and preference were displayed together.
+The last and third group corresponds to a completely randomized display order of the questions.
+The figure below describes the complete survey flow.
+
+![Figure: Study design](../input/images/app_mod.png)
+
+To contact the subjects, we followed an online programmed survey approach using _Qualtrics_, and the field work was conducted by _Netquest_, which is a company with wide experience applying online survey, and for the case of Chile, they have a consolidated pool of individuals that systematically respond online surveys.
+On this regard, the company have their own incentive system, in which the subjects accumulates specific currencies that they can exchange for a wide range of goods.
+
 ## Data
@@
 According to the recommendations of Xia and Yang ([2018](https://link.springer.com/article/10.3758/s13428-018-1055-2)) for categorical data, we present below a table with the descriptive table and another with correlations between the items on the scale.
+=======
+The field work was conducted during November and December of 2019 in Chile, obtaining a final sample of 2200 individuals, but considering that we use cases with complete information, the final analytical sample has 2141 observations.
+
+## Variables
+
+<!-- [Descriptive Table for the PPM-S, by group or the full sample] -->
+
+```{r echo=FALSE}
+load(file = "output/tables/sumdat04.RData");sumdat04
+```
 
@@
 
+CFA
+
+MGCFA
+
+@rprogram2019
+@Rosseel2012
+
 ## Methods
@@
 # Results
+
+![](../output/images/meas01.png)
+
+```{r echo=FALSE}
+load(file = "output/tables/tb.fm01.RData");tb.fm01
+```
 

### 2020-04-23 docs/index.Rmd
Update paper
Render completo luego de pull-requests

@@
 
-last website update: `r format(Sys.time(), "%A %d %B %Y %X")`
+Last update: `r format(Sys.time(), "%A %d %B %Y %X")`
 
@@
 
-### Structure of the project website:
-
 - Pre-analysis: here is the description of the pre-registration of the analysis plan in the Open Science Framework, you can see it directly [here:](https://osf.io/z45y2).
@@
 
-- The other three tabs follow the logic of the [IPO protocol](https://github.com/juancarloscastillo/ipo) for academic workflows, based on three main components = **I**nput -> **P**roduction -> **O**utput.
+### Structure of the project website:
+
+- The tabs follow the logic of the [IPO protocol](https://github.com/juancarloscastillo/ipo) for academic workflows, based on three main components = **I**nput -> **P**roduction -> **O**utput.
 Here we show only one the main parts of each component: Data -> Analysis -> Paper.

### 2020-04-23 docs/paper.Rmd
Render completo luego de pull-requests

@@
 title: "Perceptions and Preferences for Meritocracy"
-css: "../input/bib/style.css"
+css: "../input/css/custom.css"
 linestretch: '1.5'
@@
 Accordingly, we suggest that the use of the term belief is confusing, being more accurate to speak about meritocratic perceptions and preferences.
+Uso conjunto de percepción y preferencias.
 
@@
 ## Data
+
 The Netquest company ([website](www.netquest.cl)) applied the online survey during December and January.
@@
 
->Diagrama de flujo de la encuesta, orden de la escala.
+!!! Diagrama de flujo de la encuesta, orden de la escala.
 
@@
 According to the recommendations of Xia and Yang ([2018](https://link.springer.com/article/10.3758/s13428-018-1055-2)) for categorical data, we present below a table with the descriptive table and another with correlations between the items on the scale.
-=======
+
 The field work was conducted during November and December of 2019 in Chile, obtaining a final sample of 2200 individuals, but considering that we use cases with complete information, the final analytical sample has 2141 observations.
-
-## Variables
 
@@
 ```{r echo=FALSE,results='asis', warning=FALSE, message=FALSE}
-load(file = "../input/data/proc/dat04.RData")
+load(file = "input/data/proc/dat04.RData")
 library(stargazer)
@@
 
+# Conclusiones
+
+uso conjunto percepciones-preferencias
+
 <div/>

### 2020-05-04 docs/paper.Rmd
Revisión paper, solicita correcciones

@@
 Esto se ha expresado en una serie de protestas como la emblemática "occupy wall street" el año 2011, así como también en una serie de análisis críticos respecto del desarrollo del capitalismo y sus consecuencias [@piketty_Capital_2014].
-En este contexto, el estudio de las visiones, preferencias y percepciones respecto de la desigualdad han adquirido relevancia en las ciencias sociales, en temas como las preferencias redistributivas [@alesina_preferences_2011; @Dimick2018ModelsOtherRegardingPreferences] la legitimación de la desigualdad económica [@schroder_is_2017] y el funcionamiento de la meritocracia [@duru-bellat_whos_2012; @mijs_unfulfillable_2016; @reynolds_perceptions_2014].
+En este contexto, el estudio de las visiones, preferencias y percepciones respecto de la desigualdad han adquirido relevancia en las ciencias sociales, en temas como las preferencias redistributivas [@alesina_Fairness_2005; @dimick_Models_2018] la legitimación de la desigualdad económica [@schroder_Income_2017]  y el funcionamiento de la meritocracia [@duru-bellat_whos_2012; @mijs_unfulfillable_2016; @reynolds_perceptions_2014].
 
-En general, la meritocracia se define como un sistema de distribución de recursos y recompensas basados en el mérito individual, que en su concepción original es una suma de talento y esfuerzo [@young_rise_1962] y que pone en un lugar secundario interferencia de factores estructurales como la herencia o los contactos [@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995; @yairMeritocracy2007; @landWeSatTable2006;@youngRiseMeritocracy1994].
+En general, la meritocracia se define como un sistema de distribución de recursos y recompensas basados en el mérito individual, que en su concepción original es una suma de talento y esfuerzo [@young_rise_1962] y que pone en un lugar secundario interferencia de factores estructurales como la herencia o los contactos [@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995; @yairMeritocracy2007; @landWeSatTable2006; @youngRiseMeritocracy1994].
 Una serie de estudios han realizado críticas a la realización de este estándar moral de distribución, planteando que es una promesa incumplida dada la influencia preponderante de otros elementos más allá del mérito en el estatus individual [@arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019].
@@
 Además, como un segundo eje de análisis considera la generación de indicadores respecto de aspectos meritocráticos y anti-meritocráticos, demostrando que no son los dos polos de un mismo continuo como muchos estudios anteriores parecen sugerir.
+La propuesta de medición además está orientada a generar un instrumento lo más breve posible de manera que pueda ser utilizado en encuestas de opinión.
 
@@
 correlation close to 1) or they are different aspects of the same phenomenon.
+Therefore, two main drawbacks of the term "beliefs" are that (i) it hides different operationalizations of the concept of meritocracy, and (ii) it blurs the distinction between perceptions and preferences.
+We argue that perceptions and preferences should be disentangled and used together, specially when analyzing preferences.
 As @son_hing_merit_2011-1 have pointed out, "People can believe that outcomes ought to be distributed on the basis of merit and yet vary in their perceptions of whether this is how society currently operates" (p.
 435).
-Accordingly, we suggest that the use of the term belief is confusing, being more accurate to speak about meritocratic perceptions and preferences.
-Uso conjunto de percepción y preferencias.
+In other words, preferences should be interpreted in the context of perception: a large preference for meritocracy certainly means something totally different for someone perceiving high meritocracy than for someone perceiving low meritocracy.
 
@@
 
-The proposed measurement model is based on two axis of analysis, as depicted in Table 1:
+The proposed measurement framework is based on two axis of analysis, as depicted in Table 1:
 
@@
 For meritocratic indicators we use effort and talent as the main components of the traditional concept of merit as defined by @young_rise_1962, whereas for non-meritocratic dimensions we use having rich parents and good contacts.
-
-![](../input/images/tabla2.png)
+The description of the specific items is presented in the methodology section.
 
@@
 
-The study design consist in two main stages.
-The first stage was the questionnaire design, in which we collect a serie of background variables, including sex, age and educational level.
-On this regard, one of the main aspects of our design is that we tested three different application modalities, in order to observe how the display order of the indicators could affect the model fit.
-We decide to randomize the assignment of the subject to each application modality, and as a result we obtain three groups within the full sample.
-The first combination corresponds to the group that answered the questions that were sorted by the Perceptions/Preferences dimension.
-The second combination corresponds to the group in which the questions are displayed following a topic-sorted way, this means that the effort, talent, rich family and networks questions for perception and preference were displayed together.
-The last and third group corresponds to a completely randomized display order of the questions.
-The figure below describes the complete survey flow.
+## Data collection
 
-![Figure: Study design](../input/images/app_mod.png)
-
-To contact the subjects, we followed an online programmed survey approach using _Qualtrics_, and the field work was conducted by _Netquest_, which is a company with wide experience applying online survey, and for the case of Chile, they have a consolidated pool of individuals that systematically respond online surveys.
-On this regard, the company have their own incentive system, in which the subjects accumulates specific currencies that they can exchange for a wide range of goods.
-
-## Data
-
-The Netquest company ([website](www.netquest.cl)) applied the online survey during December and January.
-The online survey distributed by the company was scheduled on the platform Qualtrics ([website](www.Qualtrics.com)).
+The data was obtained through an online questionnaire which was part of a larger study on meritocracy and preferences developed in Chile in 2019 and funded by the national scientific agency FONDECYT.
+The questionnaire was programmed in Qualtrics and the fieldwork was in charge of an external online survey agency ([netquest.cl](www.netquest.cl))during December 2019 and January 2020.
 The sample was selected from a non-probabilistic design in three large cities in Chile.
 The quota method based on age, sex and educational level was used.
-Quotas were generated based on the survey of the Public Studies Center (CEP, 2019), which has a high prestige in the country.
+Quotas were generated based on the survey of the Public Studies Center (CEP, 2019), which has a high prestige in the country and is also the counterpart agency of ISSP (International Social Survey Programme) in Chile.
 A total sample of 2141 people was collected, excluding those who did not answer the questions on the scale and those who did not accept informed consent.
-The representativeness of the sample is not satisfactory because, although the proportion by age and sex is consistent with the estimates of the CEP survey, a bias towards high educational levels is evident (as detailed in Annex __X__).
-This limitation must be considered when making inferences about the population.
+As it usually occurs online samples, there were some limitations in achieving the quotas for lower educational levels.
 
-> Anexo x
+## Instrument design
 
-|                     | Sample  |   CEP  |
-|---------------------|:-------:|:------:|
-| __gender__          |         |        |
-| Men                 | 49,82%  | 50,52% |
-| Women               | 50.18%  | 49,47% |
-| __Age__             |         |        |
-| 18 - 24             | 18,55%  | 18,17% |
-| 25 - 34             | 18,86%  | 17,48% |
-| 35 - 44             | 19.09%  | 19,98% |
-| 45 - 54             | 17,96%  | 19,23% |
-| 55 - or more        | 25,54%  | 25.11% |
-| __Education__       |         |        |
-| Primary  or less    | 2,93%   | 15,88% |
-| Hig school          | 43,23%  | 37,04% |
-| Non university      | 32,63%  | 28,93% |
-| university or more  | 21,21%  | 18,13% |
+The proposed scale of perceptions and preferences about meritocracy consists of 8 indicators that are grouped into 4 dimensions referred above: Perceptions (meritocratic/nonmeritocratic) and preferences (meritocratic/non-meritocratic).
+In order to achieve at least some comparability with previous studies, the questions were adapted from the items battery "reasons to get ahead" (ISSP/GSS), which are mostly used for operationalizing meritocracy in previous empirical studies [@mijs_paradox_2019; @duru-bellat_whos_2012; @reynolds_perceptions_2014].
+The eight items ordered according to dimensions are presented in Table 2.
+These 8 likert-type items have 5 response alternatives ranging from "Completely disagree"(1) to "Completely agree" (5).
 
-Based on the recommendations of Wilson, Strite and Loiacono (2017) it has been considered pertinent transparency that the order the items of the meritocracy scale in the survey was random for a third of the sample.
+![](../input/images/tabla2.png)
+
+XX ojo con esta imagen, titularla en el documento, no en la imagen misma
+
+## Administration sets
+
 With the objective of evaluating the effect of the order of the indicators, the respondents were randomly divided into three groups as explained in __Diagram 1__.
@@
 distribution according to effort).
-The third group was presented with the scale with the randomized items to assess the possible bias produced by the order in which the questions are presented ((Budd, 1987) [https://psycnet.apa.org/record/1988-31448-001]; Davis and Venkatesh 1996; Goodhue and Loiacono, 2002; Schriesheim et al., 1989).
+Finally, in the third group the items were completely randomized.
 
-!!! Diagrama de flujo de la encuesta, orden de la escala.
+XX agregar número de sujetos por grupo
 
-## Variables
+XX referenciar y titular figuras y tablas compatible con html/pdf
 
-The proposed scale of perceptions and preferences about meritocracy consists of 8 indicators that are grouped into the 4 dimensions indicated.
-The questions are adaptations of the "get ahead" scale.
-the terms of of this scale were used (i.e talent, effort, contacts and wealthy parents) which are included in different investigations on meritocracy [@mijs_paradox_2019; @duru-bellat_whos_2012; @reynolds_perceptions_2014].
-Unlike this scale, this proposal adds the difference between perceptions and preferences.
-In addition, a comparative component is incorporated into the question that refers to differences between people with little or fed up with some attribute (e.g.
-effort, contacts).The eight items ordered according to dimensions are presented in Table 2.
-These 8 likert-type items have 5 response alternatives ranging from "I completely agree" (5) to "I completely disagree"(1).
-According to the recommendations of Xia and Yang ([2018](https://link.springer.com/article/10.3758/s13428-018-1055-2)) for categorical data, we present below a table with the descriptive table and another with correlations between the items on the scale.
+![Figure: Study design](../input/images/app_mod.png)
 
-The field work was conducted during November and December of 2019 in Chile, obtaining a final sample of 2200 individuals, but considering that we use cases with complete information, the final analytical sample has 2141 observations.
+## Methods
+
+For testing the scale's underlying constructs we estimate confirmatory factor analysis models (CFA). The model estimates one factor for each dimension, as represented in the following figure:
+
+![](../output/images/meas01.png)
+
+CFA was conducted using the `lavaan` package (version 0.6-3; Rosseel, 2020) with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement (Kline, 2016; Rosseel, 2020).
+As recommended by Brown (2008), we assessed model fit by jointly considering the comparative fit index and Tucker-Lewis Index (CFI and TLI; acceptable fit > 0.95), Root of the average squared residual approximation (RMSEA; acceptable fit < 0.08),  Chi-square: (p-value; acceptable fit > 0.05, and Chi-square ratio:> 3).
+
+# Results
+
+## Descriptive analysis
+
+Table X presents de frequencies and correlations among the items:
+
+XX producir tabla y llamar directamente desde la fuente o como imagen
 
@@
 
-CFA
+<table style="text-align:center"><caption><strong>Descriptive statistics</strong></caption>
+<tr><td colspan="9" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left">Statistic</td><td>N</td><td>Mean</td><td>St.
+Dev.</td><td>Min</td><td>Pctl(25)</td><td>Median</td><td>Pctl(75)</td><td>Max</td></tr>
+<tr><td colspan="9" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left">A.
+pc.effort</td><td>2141</td><td>3.20</td><td>1.38</td><td>1</td><td>2</td><td>3</td><td>4</td><td>5</td></tr>
+<tr><td style="text-align:left">B. pc.talent</td><td>2141</td><td>3.02</td><td>1.16</td><td>1</td><td>2</td><td>3</td><td>4</td><td>5</td></tr>
+<tr><td style="text-align:left">C. pc.wpart</td><td>2141</td><td>3.66</td><td>1.36</td><td>1</td><td>3</td><td>4</td><td>5</td><td>5</td></tr>
+<tr><td style="text-align:left">D. pc.netw</td><td>2141</td><td>3.79</td><td>1.24</td><td>1</td><td>3</td><td>4</td><td>5</td><td>5</td></tr>
+<tr><td style="text-align:left">E. pf.effort</td><td>2141</td><td>3.89</td><td>1.25</td><td>1</td><td>3</td><td>4</td><td>5</td><td>5</td></tr>
+<tr><td style="text-align:left">F. pf.talent</td><td>2141</td><td>3.24</td><td>1.19</td><td>1</td><td>2</td><td>3</td><td>4</td><td>5</td></tr>
+<tr><td style="text-align:left">G. pf.wpart</td><td>2141</td><td>2.69</td><td>1.18</td><td>1</td><td>2</td><td>3</td><td>4</td><td>5</td></tr>
+<tr><td style="text-align:left">H. pf.netw</td><td>2141</td><td>2.41</td><td>1.11</td><td>1</td><td>2</td><td>2</td><td>3</td><td>5</td></tr>
+<tr><td style="text-align:left">grupo</td><td>2141</td><td>2.00</td><td>0.82</td><td>1</td><td>1</td><td>2</td><td>3</td><td>3</td></tr>
+<tr><td colspan="9" style="border-bottom: 1px solid black"></td></tr></table>
 
-MGCFA
-
-@rprogram2019
-@Rosseel2012
-
-## Methods
-
-We tested theoretical model of the scale of perceptions and preferences about meritocracy using confirmatory factor analysis (CFA) with for latent factor.
-CFA was conducted using the lavaan package (version 0.6-3; Rosseel, 2020) with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement (Kline, 2016; Rosseel, 2020).
-As recommended by Brown (2008), we assessed model fit by jointly considering the comparative fit index and Tucker-Lewis Index (CFI and TLI; acceptable fit > 0.95), Root of the average squared residual approximation (RMSEA; acceptable fit < 0.08),  Chi-square: (p-value; acceptable fit > 0.05, and Chi-square ratio:> 3).
-
-# Results
-
-![](../output/images/meas01.png)
+XX Agregar gráficos de frecuencia , puede ser tipo trellis 2 x 4, arriba percepción abajo preferencias; la idea es generar un input más claro para la siguiente idea: En general se observa que hay una mayor percepción de aspectos no meritocráticos que meritocráticos, mientras que en el caso de preferencias ocurre lo opuesto.
+En cuanto a las preferencias, llama la atención el rol preponderante del esfuerzo por sobre el talento.
 
@@
 
----
+XX Sobre tablas en general, entonces:
+- dejar tabla de descriptivos que ahora aparece junto a la de correlaciones
+- agregar gráficos
+- reemplazar tabla de correlaciones por un corrplot, creo que se aprecia más claramente el sentido de las asociaciones
 
-Este estudio cuenta con una encuesta panel de 3 olas.
-En la primera de los olas, se incorporó la escala en distintos órdenes.
-El primer orden, pregunta primero por todas las percepciones y después por todas las preferencias, el segundo orden, pregunta seguidamente por percepciones y preferencias en torno a un mismo tema y, el tercer orden incorpora los ítems de manera aleatoria.
-En la segunda y tercera ola, se utilizó el orden
+## Models estimation
 
-Para poder evaluar la estructura factorial de la escala propuesta, se realizará un análisis factorial confirmatorio, evaluando el ajuste del modelo propuesto a partir de los estadísticos CFA, TLI, RMSEA y χ2 ratio.
-Los criterios de ajuste utilizados fueron anteriormente reportados en un preregistro de la investigación en el sitio Open Science FrameWork.
-Basado en los criterios de Kline (2016)  y de Brown (2008)  esperamos tanto para la prueba CFA como TLI, valores por sobre .95, mientras que para la prueba RMSEA, se consideran aceptables valores bajo .08 e idóneos bajo .05.
-Por su parte, respecto al χ2, se esperan valores p no significativos como parámetro ideal y se considera aceptable un χ2 ratio menor a 3.
+XX En tabla de cargas factoriales, los fit aparecen corridos hacia la izquierda
 
@@
 
-Todos los modelos independiente de los órdenes obtuvieron un ajuste moderado con CFI superiores a .95 y RMSEA inferiores a .8, por contraparte ni un modelo logró un chi-square no significativo, aunque tanto el modelo aleatorio como el primero obtuvieron un adecuado chi-square ratio menor a 3.
-El primero de los órdenes fue el que mejor ajuste obtuvo (CFI=0.998, TLI= 0.995,RMSEA=0.037, χ2(df=14)=28,03, p = 0.014), seguido por el orden aleatorio de los ítems (CFI=0.992, TLI=0.984,RMSEA=0.051, χ2(df=14)=39.09, p < 0.001).
-Por su parte la escala ordenada por temáticas genera un efecto framing en el cual la relación entre las percepciones y las preferencias parecen sobreestimadas, lo cual afecta el ajuste, siendo este deficitario, con un χ2 ratio mayor a tres y un RMSEA que se puede encontrar por sobre .8 según el intervalo de confianza correspondiente(CFI=0.984, TLI=0.968,RMSEA=.071, χ2(df=14)=64.156, p < 0.001).
+Todos los modelos independiente de los órdenes obtuvieron un ajuste adecuado.
+con CFI superiores a .95 y RMSEA inferiores a .8, por contraparte ni un modelo logró un chi-square no significativo, aunque tanto el modelo aleatorio como el primero obtuvieron un adecuado chi-square ratio menor a 3.
+El primero de los órdenes fue el que obtuvo mejor ajuste (CFI=0.998, TLI= 0.995,RMSEA=0.037, χ2(df=14)=28,03, p = 0.014), seguido por el orden aleatorio de los ítems (CFI=0.992, TLI=0.984,RMSEA=0.051, χ2(df=14)=39.09, p < 0.001).
+Por su parte, la escala ordenada por temáticas parece generar un efecto framing en el cual la relación entre las percepciones y las preferencias parecen sobreestimadas, afectando de esta manera el ajuste (CFI=0.984, TLI=0.968,RMSEA=.071, χ2(df=14)=64.156, p < 0.001).
+
+XX explicar mejor lo que viene: invarianza de qué y por qué se realiza.
 
@@
 Coherentemente el modelo aleatorio tiene un ajuste intermedio entre ambos.
-Considerando la ventaja del orden, respecto aleatorio despeja la posibilidad del efecto framing, es decir, de que el resultado del modelo se deba al orden de las preguntas, se ha desidido seguir utilizando el orden 3.
 
@@
 
-# Validez del constructo
-
-## Validez convergente
-indicar la construcción de los índices -- hacer un anexo.
-
-Esta escala cuenta con validez de constructo, ya que los distintos índices formados según factor correlacionan coherentemente con la escala "Get a head", utilizada por el programa internacional de encuestas sociales (ISSP), y con la escala de valores igualitarios utilizada por la encuesta internacional de valores.
-
-|                                  | Igualitarismo | Get a head |
-|----------------------------------|:-------------:|:----------:|
-| Percepción <br>Meritocrática     |   -0.143***   | 0.228***   |
-| Percepción <br>No Meritocrática  |    0.130***   | -0.210***  |
-| Preferencia <br>Meritocrática    |   -0.055**    | 0.001      |
-| Preferencia <br>No Meritocratica |   -0.191***   | 0.083***   |
-
-Existe coherencia entre el índice "Get a Head" y el índice de preferencias y percepciones meritocráticas.
-Como era esperable, la relación es positiva entre este índice y la percepción de meritocracia de la escala propuesta (*r*=0.228).
-En la misma línea, el índice Get a head, correlaciona de manera negativa con la percepción de no meritocracia (c=-0.210;p=0.000).
-Además, las relaciones con las preferencias son considerablemente menores, e incluso no es significativa respecto a las preferencias meritocráticas.
-En suma, el "Get a head" que es comúnmente utilizado para medir percepciones, posee relación solo con las variables de percepción y no así con las de preferencias.
-
-Por su parte, el índice de Igualitarismo que refleja cuán problemática consideran la desigualdad los entrevistados, posee también relaciones coherentes.
-En primer lugar, es necesario destacar que solo posee una relación muy pequeña con la variable de preferencias meritocráticas, lo cual es positivo pues dicha dimensión de la escala no busca reflejar igualitarismo sino de preferencias meritocráticas.
-En segundo lugar, el índice se relaciona de manera negativa con la preferencia no meritocrática, lo cual es coherente en tanto alguien que valore la igualdad debería tender a no estar de acuerdo con la desigualdad de oportunidades.
-En tercer lugar, considerar la desigualdad como un problema, desde el índice de igualitarismo, posee como requisito creer que existe desigualdad, al respecto, es coherente que quienes consideran que la desigualdad existente es problemática, perciban igualmente que el éxito no depende sólo del esfuerzo y el talento (c=-0.143;p=0.000) sino que está influido por aspectos no meritocráticos (c=0.130;p=).
-
-# Discusión
-
 # Conclusión
 
-## Fiabilidad
-
-### Correlación entre los mismos items segun distintas olas.
-
-| **Pensando en la sociedad chilena, <br>¿En qué medida se encuentra usted de acuerdo o en desacuerdo<br>con cada una de las siguientes afirmaciones?** |  <br>Ola 1 - Ola 2  |  <br>Ola 2 - Ola 3 | <br>Ola 1 - Ola 3 |
-|-------------------------------------------------------------------------------------------------------------------------------------------------------|:-------------------:|:------------------:|:-----------------:|
-| Quienes más se esfuerzan logran obtener mayores <br>recompensas que quienes se esfuerzan menos.
-|         0.33        |        0.38        |        0.30       |
-| Quienes poseen más talento logran obtener mayores<br>recompensas que quienes poseen menos talento.
-|         0.30        |        0.28        |        0.27       |
-| Quienes tienen padres ricos logran salir adelante.
-|         0.33        |        0.36        |        0.38       |
-| Quienes tienen buenos contactos logran salir adelante.
-|         0.30        |        0.30        |        0.31       |
-| Quienes más se esfuerzan deberían obtener mayores<br>recompensas que quienes se esfuerzan menos.
-|         0.27        |        0.32        |        0.22       |
-| Quienes poseen más talento deberían obtener mayores<br>recompensas que quienes poseen menos talento.
-|         0.31        |        0.39        |        0.31       |
-| Está bien que quienes tienen padres ricos salganadelante.
-|         0.40        |        0.41        |        0.39       |
-| Está bien que quienes tienen buenos contactos salgan adelante.
-|         0.39        |        0.41        |        0.42       |
-
- Pese a que existen bajas correlaciones, hay evidencia para apoyar la hipótesis de estabilidad de la escala a lo largo de las mediciones.
-Las correlaciones entre los mismos ítems en las distintas olas son de un tamaño pequeño a moderado según los parámetros de [@cohenStatisticalPowerAnalysis2013].
-Estas bajas  correlaciones indican inestabilidad en las respuestas de los encuestados, no obstante existe evidencia para señalar una invarianza fuerte de las escalas en las distintas mediciones.
-En primer lugar, podemos decir que no existe diferencia entre el ajuste de los tres  nos indica que no existe una diferencia significativa entre los ajustes al modelo de las tres olas, lo cual nos señala la existencia de invarianza configuracional.
-
-         Df   AIC   BIC  Chisq Chisq diff Df diff Pr(>Chisq)
-fit.conf 42 35745 36222 110.91
-fit.deb  50 35738 36172 119.39      8.485       8     0.3876
-fit.fuer 58 35732 36124 129.94     10.546       8     0.2288
-fit.str  74 35772 36079 202.02     72.081      16  4.298e-09 ***
-
-El análisis de invarianza longitudinal muestra resultados moderadamente buenos. hay equivalencia configuracional,  métrica, pero no escalar escalar
+Considerando la ventaja del orden, respecto aleatorio despeja la posibilidad del efecto framing, es decir, de que el resultado del modelo se deba al orden de las preguntas, se ha decidido seguir utilizando el orden 3.
 
@@
 
+# Anexos
+
+|                     | Sample  |   CEP  |
+|---------------------|:-------:|:------:|
+| __gender__          |         |        |
+| Men                 | 49,82%  | 50,52% |
+| Women               | 50.18%  | 49,47% |
+| __Age__             |         |        |
+| 18 - 24             | 18,55%  | 18,17% |
+| 25 - 34             | 18,86%  | 17,48% |
+| 35 - 44             | 19.09%  | 19,98% |
+| 45 - 54             | 17,96%  | 19,23% |
+| 55 - or more        | 25,54%  | 25.11% |
+| __Education__       |         |        |
+| Primary  or less    | 2,93%   | 15,88% |
+| Hig school          | 43,23%  | 37,04% |
+| Non university      | 32,63%  | 28,93% |
+| university or more  | 21,21%  | 18,13% |
+
 # Bibliografía

### 2020-05-06 docs/paper.Rmd
ajustes paper
Quedo pendiente (Lo vere hoy):

- hacer titulos para tablas html/pdf
- centrar ajustes en tabla de cargas factoriales

@@
 
-![](../input/images/tabla2.png)
+![Tabla 2: Dimensiones](../input/images/tabla2.png)
 
-XX ojo con esta imagen, titularla en el documento, no en la imagen misma
+6 XX ojo con esta imagen, titularla en el documento, no en la imagen misma
 
@@
 
-XX agregar número de sujetos por grupo
+7 XX agregar número de sujetos por grupo
 
-XX referenciar y titular figuras y tablas compatible con html/pdf
+8 XX referenciar y titular figuras y tablas compatible con html/pdf
 
-![Figure: Study design](../input/images/app_mod.png)
+![Diagrama 1: Study design](../input/images/app_mod.png)
 
@@
 
-XX producir tabla y llamar directamente desde la fuente o como imagen
+1 XX producir tabla y llamar directamente desde la fuente o como imagen
 
 <!-- [Descriptive Table for the PPM-S, by group or the full sample] -->
-
-```{r echo=FALSE}
-load(file = "output/tables/sumdat04.RData");sumdat04
-```
 
@@
 
-<table style="text-align:center"><caption><strong>Descriptive statistics</strong></caption>
-<tr><td colspan="9" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left">Statistic</td><td>N</td><td>Mean</td><td>St.
-Dev.</td><td>Min</td><td>Pctl(25)</td><td>Median</td><td>Pctl(75)</td><td>Max</td></tr>
-<tr><td colspan="9" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left">A.
-pc.effort</td><td>2141</td><td>3.20</td><td>1.38</td><td>1</td><td>2</td><td>3</td><td>4</td><td>5</td></tr>
-<tr><td style="text-align:left">B. pc.talent</td><td>2141</td><td>3.02</td><td>1.16</td><td>1</td><td>2</td><td>3</td><td>4</td><td>5</td></tr>
-<tr><td style="text-align:left">C. pc.wpart</td><td>2141</td><td>3.66</td><td>1.36</td><td>1</td><td>3</td><td>4</td><td>5</td><td>5</td></tr>
-<tr><td style="text-align:left">D. pc.netw</td><td>2141</td><td>3.79</td><td>1.24</td><td>1</td><td>3</td><td>4</td><td>5</td><td>5</td></tr>
-<tr><td style="text-align:left">E. pf.effort</td><td>2141</td><td>3.89</td><td>1.25</td><td>1</td><td>3</td><td>4</td><td>5</td><td>5</td></tr>
-<tr><td style="text-align:left">F. pf.talent</td><td>2141</td><td>3.24</td><td>1.19</td><td>1</td><td>2</td><td>3</td><td>4</td><td>5</td></tr>
-<tr><td style="text-align:left">G. pf.wpart</td><td>2141</td><td>2.69</td><td>1.18</td><td>1</td><td>2</td><td>3</td><td>4</td><td>5</td></tr>
-<tr><td style="text-align:left">H. pf.netw</td><td>2141</td><td>2.41</td><td>1.11</td><td>1</td><td>2</td><td>2</td><td>3</td><td>5</td></tr>
-<tr><td style="text-align:left">grupo</td><td>2141</td><td>2.00</td><td>0.82</td><td>1</td><td>1</td><td>2</td><td>3</td><td>3</td></tr>
-<tr><td colspan="9" style="border-bottom: 1px solid black"></td></tr></table>
+```{r echo=FALSE}
+#load(file = "output/tables/sumdat04.RData");sumdat04
+```
 
-XX Agregar gráficos de frecuencia , puede ser tipo trellis 2 x 4, arriba percepción abajo preferencias; la idea es generar un input más claro para la siguiente idea: En general se observa que hay una mayor percepción de aspectos no meritocráticos que meritocráticos, mientras que en el caso de preferencias ocurre lo opuesto.
+```{r echo=FALSE}
+library(ggplot2)
+load(file = "output/images/grafico_descriptivo.RData")
+ggplot(tmp, aes(x = variable, y = value)) + geom_boxplot() +  facet_wrap(~ tipo)
+```
+
+```{r echo=FALSE}
+
+load(file = "output/images/correlation.RData")
+library(corrplot)
+corrplot(cor04$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
+```
+
+2 XX Agregar gráficos de frecuencia , puede ser tipo trellis 2 x 4, arriba percepción abajo preferencias; la idea es generar un input más claro para la siguiente idea: En general se observa que hay una mayor percepción de aspectos no meritocráticos que meritocráticos, mientras que en el caso de preferencias ocurre lo opuesto.
 En cuanto a las preferencias, llama la atención el rol preponderante del esfuerzo por sobre el talento.
@@
 
-XX Sobre tablas en general, entonces:
+3 - XX Sobre tablas en general, entonces:
 - dejar tabla de descriptivos que ahora aparece junto a la de correlaciones
-- agregar gráficos
-- reemplazar tabla de correlaciones por un corrplot, creo que se aprecia más claramente el sentido de las asociaciones
+- agregar gráficos (listo)
+- reemplazar tabla de correlaciones por un corrplot, creo que se aprecia más claramente el sentido de las asociaciones (listo)
 
@@
 
-XX En tabla de cargas factoriales, los fit aparecen corridos hacia la izquierda
+4 XX En tabla de cargas factoriales, los fit aparecen corridos hacia la izquierda
 
@@
 
-XX explicar mejor lo que viene: invarianza de qué y por qué se realiza.
+5 XX explicar mejor lo que viene: invarianza de qué y por qué se realiza.
 
-A partir de los análisis de invarianza se puede decir que los distintos órdenes no son invariantes en términos configuracionales [@vandenbergReviewSynthesisMeasurement2000].
-El modelo que incorpora la diferenciación de los grupos en la estimación del modelo, este presenta un ajuste moderado similar al de los modelos por separados, aunque posee un ratio de chi-square deficiente por lo que se descarta la invarianza.
-Esta varianza configuracional se debe a la sobre estimación que se genera en el orden uno de las correlaciones entre los indicadores de un mismo factor los cuales aparecen juntos para el encuestado, mientras que en el modelo 2 se presentan separados, disminuyendo la correlación entre los ítem de un mismo factor.Podríamos decir que las correlaciones entre los ítems son sub o sobre-estimadas según el orden en el que aparezcan.
-Coherentemente el modelo aleatorio tiene un ajuste intermedio entre ambos.
+Si bien  todas las pruebas obtuvieron indicadores relativamente adecuados, las diferencias mencionadas entre los ajustes de los modelos según orden son estadísticamente significativas, lo cual fue evaluado a partir de un análisis de invarianza.
+Se concluye que entre los tres órdenes no existe invarianza configuracional, es decir, no poseen la misma dimensionalidad y por ello no ajustan igualmente al modelo teórico [@vandenbergReviewSynthesisMeasurement2000].
+Esto se debe al efecto producido por la aparición conjunta de los ítems de un mismo factor en el orden 1, lo cual aumenta el ajuste del modelo.
+Además, en el orden 2, al preguntarse seguidamente por la percepción y la preferencia en torno al mismo indicador, aumentan las relaciones cruzadas disminuyendo el ajuste del modelo.
+De modo coherente, el modelo aleatorio presentó un ajuste intermedio entre el orden 1 y el orden 2.
 

### 2020-05-06 production/prod_analysis-cfa.rmd
Render completo luego de pull-requests
ajustes paper
Quedo pendiente (Lo vere hoy):

- hacer titulos para tablas html/pdf
- centrar ajustes en tabla de cargas factoriales

@@
 colnames(cor04$rho) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)", "(H)")
+
+save(cor04, file = "output/images/correlation.RData")
+
 corrplot(cor04$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
@@
               "Perception meritocratic","Perception non meritocratic","Preference meritocratic","Preference non meritocratic")
+
+dir.create("output/images")
 
@@
 <!-- ``` -->
+```{r}
+
+dat_per <- dat04 %>% select("Effort"=perc_effort,"Talent"=perc_talent, "Rich Parents"=perc_wpart, "Contacts"=perc_netw)    
+dat_pref <- dat04 %>% select("Effort"=pref_effort,"Talent"=pref_talent,"Rich Parents"=pref_wpart, "Contacts"=pref_netw)
+
+tmp_per<- melt(dat_per)
+tmp_per$tipo<- "Perceptions"
+
+tmp_prf<- melt(dat_pref)
+tmp_prf$tipo<- "Preferences"
+
+tmp <-rbind(tmp_per,tmp_prf)
+
+save(tmp, file = "output/images/grafico_descriptivo.RData")
+
+ggplot(tmp, aes(x = variable, y = value)) + geom_boxplot() +  facet_wrap(~ tipo)
+
+```

### 2020-05-07 docs/paper.Rmd
Ajustes
Incluye:
- Nueva bibliografia
- nuevo parrafo al paper
- Explicacion invarianza entre ordenes.
- Grafico descriptivo de la escala
- Corrplot
- html con rev.biblio (provisoria)
-

Falta
- centrar fit (creo que no se puede)
- separar descriptivos de la tabla de correlaciones.
- titulos
- llamar imagenes guardadas (no he podido guardarlas)

@@
 La desigualdad económica se ha vuelto un tema que genera creciente preocupación y malestar alrededor del mundo.
-Esto se ha expresado en una serie de protestas como la emblemática "occupy wall street" el año 2011, así como también en una serie de análisis críticos respecto del desarrollo del capitalismo y sus consecuencias [@piketty_Capital_2014].
-En este contexto, el estudio de las visiones, preferencias y percepciones respecto de la desigualdad han adquirido relevancia en las ciencias sociales, en temas como las preferencias redistributivas [@alesina_Fairness_2005; @dimick_Models_2018] la legitimación de la desigualdad económica [@schroder_Income_2017]  y el funcionamiento de la meritocracia [@duru-bellat_whos_2012; @mijs_unfulfillable_2016; @reynolds_perceptions_2014].
+Esto se ha expresado en una serie de protestas como la emblemática "occupy wall street" el año 2011, así como también en una serie de análisis críticos respecto del desarrollo del capitalismo y sus consecuencias [@PikettyCapitalTwentyFirstCentury2014b].
+En este contexto, el estudio de las visiones, preferencias y percepciones respecto de la desigualdad han adquirido relevancia en las ciencias sociales, en temas como las preferencias redistributivas [@alesina_Fairness_2005; @dimick_Models_2018] la legitimación de la desigualdad económica [@schroder_Income_2017]  y el funcionamiento de la meritocracia [@Duru-BellatWhoMeritocracyIndividual2012b; @MijsParadoxInequalityIncome2019b; @ReynoldsPerceptionsMeritocracyLand2014b].
 
 En general, la meritocracia se define como un sistema de distribución de recursos y recompensas basados en el mérito individual, que en su concepción original es una suma de talento y esfuerzo [@young_rise_1962] y que pone en un lugar secundario interferencia de factores estructurales como la herencia o los contactos [@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995; @yairMeritocracy2007; @landWeSatTable2006; @youngRiseMeritocracy1994].
-Una serie de estudios han realizado críticas a la realización de este estándar moral de distribución, planteando que es una promesa incumplida dada la influencia preponderante de otros elementos más allá del mérito en el estatus individual [@arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019].
-Por otro lado, desde la psicología social y la sociología se han estudiado las características y consecuencias de las creencias en la meritocracia, en general basados en la hipótesis que mayor creencia en la meritocracia lleva a una mayor legitimación de las desigualdades [@hadjar_meritokratie_2008; @MadeiraPrimesConsequencesSystematic2019].
+Una serie de estudios han realizado críticas a la realización de este estándar moral de distribución, planteando que es una promesa incumplida dada la influencia preponderante de otros elementos más allá del mérito en el estatus individual [@WitteveenReconsideringmeritocraticpower2020a; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019].
+Por otro lado, desde la psicología social y la sociología se han estudiado las características y consecuencias de las creencias en la meritocracia, en general basados en la hipótesis que mayor creencia en la meritocracia lleva a una mayor legitimación de las desigualdades [@PremingerMeritocracyserviceethnocracy2020; @TrumpWhenwhyeconomic2020; @Meredithsocietydivisibleblessed2020; @hadjar_meritokratie_2008; @MadeiraPrimesConsequencesSystematic2019].
+
+Debido al rol que cumplen las creencias meritocráticas dentro del pensamiento neoliberal [@Bay-ChengTrackingHomoOeconomicus2015], han surgido múltiples investigaciones que evalúan la relación entre creencias meritocráticas y diversos ámbitos sociales de la actualidad.
+Se han desarrollado estudios que vinculan la meritocracia con el reforzamiento de estereotipos socioeconómicos, de género y de etnias [@MadeiraPrimesConsequencesSystematic2019; @GirerdNeoliberalismIdeologicalBarrier2020; @PremingerMeritocracyserviceethnocracy2020].
+Igualmente, se han desarrollado líneas de investigación que evalúan el efecto de las creencias meritocráticas en el contexto educativo [@GenerettStoriesWeTell2020; @ReayPerilsPenaltiesMeritocracy2020; @OwensENGINESSOCIALMOBILITY2020].
+y en el contexto organizacional de las empresas [@PerezAdvancingcareersmerit2020; @AielloNewevidencefirmuniversity2019]
 
@@
 This is a very strong assumption: effort would not depend on parents influence, and talent is not meritocratic.
-A similar approach only considering effort as the main element of meritocracy is taken by @solt_economic_2016.
+A similar approach only considering effort as the main element of meritocracy is taken by other authors ho only included effort as an indicator of meritocratic beliefs [@MaitnerEmotionalreactionsunequal2015;@MajorPerceiveddiscriminationworldview2007; @BubakPerceptionsmeritocracynote2019; @solt_economic_2016].
 We suggest here is that questions such as whether talent is not meritocratic (or less meritocratic than effort) should be open to empirical scrutiny and not disregarded upfront based on scholars' assumptions.
@@
 
-![Tabla 2: Dimensiones](../input/images/tabla2.png)
+### Table 2
+![](../input/images/tabla2.png)
 
-6 XX ojo con esta imagen, titularla en el documento, no en la imagen misma
+1 XX ojo con esta imagen, titularla en el documento, no en la imagen misma
 
@@
 
-With the objective of evaluating the effect of the order of the indicators, the respondents were randomly divided into three groups as explained in __Diagram 1__.
-The scale was presented to the first group in the order that appears in Table 2, in such a way that the items of the same factor are found together.
-The second group was presented with the scale with the items of the same factor separated from each other, in such a way that they were consulted at the same time for their perceptions and preferences of a particular topic (e.g.
+With the objective of evaluating the effect of the order of the indicators, the respondents (_n = 2141_) were randomly divided into three groups as explained in __Diagram 1__.
+The scale was presented to the first group (_n = 712_) in the order that appears in Table 2, in such a way that the items of the same factor are found together.
+The second group (_n = 717_)was presented with the scale with the items of the same factor separated from each other, in such a way that they were consulted at the same time for their perceptions and preferences of a particular topic (e.g.
 distribution according to effort).
-Finally, in the third group the items were completely randomized.
+Finally, in the third group (_n = 712_) the items were completely randomized.
 
-7 XX agregar número de sujetos por grupo
+2 XX agregar número de sujetos por grupo
 
-8 XX referenciar y titular figuras y tablas compatible con html/pdf
+3 XX referenciar y titular figuras y tablas compatible con html/pdf
 
-![Diagrama 1: Study design](../input/images/app_mod.png)
+### Diagram 1 
+![](../input/images/app_mod.png)
 
@@
 
-1 XX producir tabla y llamar directamente desde la fuente o como imagen
+4 XX producir tabla y llamar directamente desde la fuente o como imagen [__Pendiente__]
 
@@
 
-```{r echo=FALSE,results='asis', warning=FALSE, message=FALSE}
-load(file = "input/data/proc/dat04.RData")
-library(stargazer)
-stargazer(dat04,type = "html",
-          title = "Descriptive statistics",
-          font.size = "small",
-          median = TRUE,
-          header=FALSE,
-          digit.separator = "",
-          digits = 2,
-          covariate.labels = c("A. pc.effort", "B. pc.talent","C. pc.wpart","D. pc.netw","E. pf.effort","F. pf.talent","G. pf.wpart", "H. pf.netw"))
-```
-
 ```{r echo=FALSE}
-#load(file = "output/tables/sumdat04.RData");sumdat04
+load(file = "output/tables/sumdat04.RData");sumdat04
 ```
@@
 
+5 XX Agregar gráficos de frecuencia , puede ser tipo trellis 2 x 4, arriba percepción abajo preferencias; la idea es generar un input más claro para la siguiente idea:
+
+En general se observa que hay una mayor percepción de aspectos no meritocráticos que meritocráticos, mientras que en el caso de preferencias ocurre lo opuesto.
+En cuanto a las preferencias, llama la atención el rol preponderante del esfuerzo por sobre el talento.
+
 ```{r echo=FALSE}
@@
 
-2 XX Agregar gráficos de frecuencia , puede ser tipo trellis 2 x 4, arriba percepción abajo preferencias; la idea es generar un input más claro para la siguiente idea: En general se observa que hay una mayor percepción de aspectos no meritocráticos que meritocráticos, mientras que en el caso de preferencias ocurre lo opuesto.
-En cuanto a las preferencias, llama la atención el rol preponderante del esfuerzo por sobre el talento.
-
 ```{r echo=FALSE}
@@
 
-3 - XX Sobre tablas en general, entonces:
-- dejar tabla de descriptivos que ahora aparece junto a la de correlaciones
+6 - XX Sobre tablas en general, entonces:
+- dejar tabla de descriptivos que ahora aparece junto a la de correlaciones [_Pendiente_ ]
 - agregar gráficos (listo)
@@
 
-4 XX En tabla de cargas factoriales, los fit aparecen corridos hacia la izquierda
+7 XX En tabla de cargas factoriales, los fit aparecen corridos hacia la izquierda [__Pendiente__]
 
@@
 
-5 XX explicar mejor lo que viene: invarianza de qué y por qué se realiza.
+8 XX explicar mejor lo que viene: invarianza de qué y por qué se realiza.
 

### 2020-05-07 production/prod_analysis-cfa.rmd
Ajustes
Incluye:
- Nueva bibliografia
- nuevo parrafo al paper
- Explicacion invarianza entre ordenes.
- Grafico descriptivo de la escala
- Corrplot
- html con rev.biblio (provisoria)
-

Falta
- centrar fit (creo que no se puede)
- separar descriptivos de la tabla de correlaciones.
- titulos
- llamar imagenes guardadas (no he podido guardarlas)

--- 2020-04-22 production/prod_analysis-cfa.rmd
@@
 colnames(cor04$rho) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)", "(H)")
+save(cor04, file = "output/images/correlation.RData")
 corrplot(cor04$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
@@
 <!-- ``` -->
+
+```{r}
+library(ggplot2)
+library(reshape2)
+dat_per <- dat04 %>% select("Effort"=perc_effort,"Talent"=perc_talent, "Rich Parents"=perc_wpart, "Contacts"=perc_netw)    
+dat_pref <- dat04 %>% select("Effort"=pref_effort,"Talent"=pref_talent,"Rich Parents"=pref_wpart, "Contacts"=pref_netw)
+
+tmp_per<- melt(dat_per)
+tmp_per$tipo<- "Perceptions"
+
+tmp_prf<- melt(dat_pref)
+tmp_prf$tipo<- "Preferences"
+
+tmp <-rbind(tmp_per,tmp_prf)
+
+save(tmp, file = "output/images/grafico_descriptivo.RData")
+
+ggplot(tmp, aes(x = variable, y = value)) + geom_boxplot() +  facet_wrap(~ tipo)
+
+```

### 2020-05-08 docs/paper.Rmd
paper
paper

@@
   html_document:
+    theme: flatly
     number_sections: true
-        toc: true
-        toc_float: true
-        smart: false
+    always_allow_html: yes
+  pdf_document:
+    keep_tex: false
+    template: null
+    number_sections: true
 linkcolor: blue
+
 bibliography:
@@
 library(dplyr)
+#install.packages("captioner")
+library(captioner)
+```
+
+```{r message=FALSE, warning=FALSE, include=FALSE}
+table_nums<-captioner(prefix = "Table")
+plot_nums<-captioner(prefix = "Figure")
+diam_nums<-captioner(prefix = "Diagram")
+anx_nums<-captioner(prefix = "Annexed")
+
+table_nums("ta", "Dimensions of meritocratic beliefs.")
+table_nums("tb", "Items according to dimension.")
+table_nums("tc", "Descriptive statistics of the scale.")
+table_nums("td", "Factor loads and model fit.")
+table_nums("te", "Comparison between model fit.")
+
+diam_nums("da", "Survey flow")
+diam_nums("db", "Theorical model")
+
+plot_nums("fa", "Descriptive boxplot of the scale.")
+plot_nums("fb", "Polychoric correlation.")
+
+anx_nums("aa", "Representativeness of the sample.")
 ```
@@
 
-Table 1: Model of perceptions and preferences for meritocracy and non-meritocracy
+`r table_nums("ta")`
 
@@
 
-### Table 2
+`r table_nums("tb")`
+
 ![](../input/images/tabla2.png)
@@
 
-### Diagram 1 
+`r diam_nums("da")`
+
 ![](../input/images/app_mod.png)
@@
 For testing the scale's underlying constructs we estimate confirmatory factor analysis models (CFA). The model estimates one factor for each dimension, as represented in the following figure:
+
+`r diam_nums("db")`
 
@@
 
-4 XX producir tabla y llamar directamente desde la fuente o como imagen [__Pendiente__]
-
 <!-- [Descriptive Table for the PPM-S, by group or the full sample] -->
+`r table_nums("tc")`
 
 ```{r echo=FALSE}
-load(file = "output/tables/sumdat04.RData");sumdat04
+#load(file = "output/tables/sumdat04.RData");sumdat04
 ```
+
+4 XX producir tabla y llamar directamente desde la fuente o como imagen [__Pendiente__]
+5 XX Agregar gráficos de frecuencia , puede ser tipo trellis 2 x 4, arriba percepción abajo preferencias; la idea es generar un input más claro para la siguiente idea:
+
+`r plot_nums("fa")`
 
@@
 
-5 XX Agregar gráficos de frecuencia , puede ser tipo trellis 2 x 4, arriba percepción abajo preferencias; la idea es generar un input más claro para la siguiente idea:
-
 En general se observa que hay una mayor percepción de aspectos no meritocráticos que meritocráticos, mientras que en el caso de preferencias ocurre lo opuesto.
@@
 
+`r plot_nums("fb")`
+
 ```{r echo=FALSE}
-
 load(file = "output/images/correlation.RData")
@@
 corrplot(cor04$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
-```
-
-```{r echo=FALSE}
-load(file = "output/tables/tb.fm01.RData");tb.fm01
 ```
@@
 
-Table 3: **Fit indicators**
+`r table_nums("td")`
 
-|           | Model order 1 | Model order 2 | Model order 3 |
-|:---------:|:-------------:|:-------------:|:-------------:|
-| **n**     |      697      |      702      |      697      |
-| **CFI**   |     0.997     |     0.984     |     0.991     |
-| **TLI**   |     0.992     |     0.968     |     0.984     |
-| **RMSEA** |     0.037     |     0.071     |     0.051     |
-| **χ2**    |     28.03     |     64.156    |     39.090    |
-| **p**     |      .000     |      .000     |      .000     |
-| **χ2/gl** | 2,00          | 4,58          | 2,78          |
+```{r echo=FALSE}
+#load(file = "output/tables/tb.fm01.RData");tb.fm01
+```
 
@@
 De modo coherente, el modelo aleatorio presentó un ajuste intermedio entre el orden 1 y el orden 2.
+
+`r table_nums("te")`
 
@@
 # Anexos
+`r anx_nums("aa")`
 

### 2020-05-09 production/prod_analysis-cfa.rmd
Plotlikert
El grafico se genera en analisis, es enviado a output/images y llamado desde paper. subo ademas el .css que comente.

--- 2020-04-22 production/prod_analysis-cfa.rmd
@@
 <!-- ``` -->
+
+```{r}
+library(ggplot2)
+library(reshape2)
+library(sjPlot)
+
+dat_merit <- dat04 %>% select("Effort"=perc_effort,"Talent"=perc_talent, "Rich Parents"=perc_wpart, "Contacts"=perc_netw, "Effort"=pref_effort,"Talent"=pref_talent,"Rich Parents"=pref_wpart, "Contacts"=pref_netw)
+
+plotlikert<-plot_likert(dat_merit, c(1,1,1,1,2,2,2,2),
+            groups.titles = c("Perceptions","Preferences"),
+            geom.colors ="PuBu", 
+            axis.labels=c("Effort","Talent","Rich Parents","Contacts"),
+            catcount = 4,
+            cat.neutral = 3, 
+            grid.range  =  c ( 1.2 , 1.4 ),
+            values  =  "sum.outside",
+            reverse.scale=TRUE)
+
+ggsave(plotlikert,filename = "output/images/plotlikert.png",device = "png",width = 30,height = 15,dpi = "retina",units = "cm")
+
+```

### 2020-05-20 docs/paper.Rmd
Plotlikert
El grafico se genera en analisis, es enviado a output/images y llamado desde paper. subo ademas el .css que comente.
Rederizado
 termino pendientes sobre las tablas.

aun existen problemas de compativilidad html/pdf, funcionan pero hay que mejorar los estilos. Tengo algunas ideas de como solucionar esto.

subo todos los documentos por que no se que es necesario o no para el renderizado.

@@
 output:
+  pdf_document:
+    keep_tex: false
+    template: null
+    number_sections: true
   html_document:
@@
     always_allow_html: yes
-  pdf_document:
-    keep_tex: false
-    template: null
-    number_sections: true
 linkcolor: blue
-
 bibliography:
@@
 
-```{r message=FALSE, warning=FALSE, include=FALSE}
+```{r message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
 library(knitr)
+#install.packages("KableExtra")
 library(kableExtra)
@@
 library(captioner)
+```
+
+```{r echo=FALSE }
+table_format = if(is_html_output()) {
+  "html"
+} else if(is_latex_output()) {
+  "latex"
+}
+
 ```
@@
 anx_nums<-captioner(prefix = "Annexed")
-
 table_nums("ta", "Dimensions of meritocratic beliefs.")
@@
 table_nums("te", "Comparison between model fit.")
-
 diam_nums("da", "Survey flow")
 diam_nums("db", "Theorical model")
-
 plot_nums("fa", "Descriptive boxplot of the scale.")
 plot_nums("fb", "Polychoric correlation.")
-
 anx_nums("aa", "Representativeness of the sample.")
@@
 
+.
+.
+.
+
 The proposed measurement framework is based on two axis of analysis, as depicted in Table 1:
@@
 
-1 XX ojo con esta imagen, titularla en el documento, no en la imagen misma
-
 ## Administration sets
@@
 Finally, in the third group (_n = 712_) the items were completely randomized.
-
-2 XX agregar número de sujetos por grupo
-
-3 XX referenciar y titular figuras y tablas compatible con html/pdf
 
@@
 
-Table X presents de frequencies and correlations among the items:
+```{r echo=FALSE}
 
-<!-- [Descriptive Table for the PPM-S, by group or the full sample] -->
-`r table_nums("tc")`
+load(file = "input/data/proc/dat04.RData")
 
-```{r echo=FALSE}
-#load(file = "output/tables/sumdat04.RData");sumdat04
+dat04b<- dat04 %>% rename("Perception Effort" =perc_effort,
+                          "Perception Talent" = perc_talent,
+                          "Perception rich parents" = perc_wpart,
+                         "Perception contacts" = perc_netw,
+                          "Preferences Effort" = pref_effort,
+                          "Preferences Talent" =  pref_talent, 
+                         "Preferences rich parents" =  pref_wpart ,
+                          "Preferences contacts" = pref_netw  )
+
+sum04b  <- psych::describe(dat04b[,-c(1)], fast = T) %>% select(mean,sd,min,max) %>% data.frame() %>% round(digits = 2)
+
+table04b<- sum04b %>% rename("Mean"=mean, "SD"=sd, "Min"=min,"Max"=max)
+
+cap01 <- "Descriptive statistic"
+sumdat04<- kable(table04b, table_format, booktabs= T, escape = FALSE,align = "l",linesep = "", caption = cap01
+) %>% 
+          kable_styling(full_width = F,font_size = 8,position = "center");sumdat04
+
 ```
@@
 4 XX producir tabla y llamar directamente desde la fuente o como imagen [__Pendiente__]
-5 XX Agregar gráficos de frecuencia , puede ser tipo trellis 2 x 4, arriba percepción abajo preferencias; la idea es generar un input más claro para la siguiente idea:
 
@@
 
-```{r echo=FALSE}
-library(ggplot2)
-load(file = "output/images/grafico_descriptivo.RData")
-ggplot(tmp, aes(x = variable, y = value)) + geom_boxplot() +  facet_wrap(~ tipo)
-```
+![](../output/images/plotlikert.png)
 
@@
 - dejar tabla de descriptivos que ahora aparece junto a la de correlaciones [_Pendiente_ ]
-- agregar gráficos (listo)
-- reemplazar tabla de correlaciones por un corrplot, creo que se aprecia más claramente el sentido de las asociaciones (listo)
 
@@
 
-`r table_nums("td")`
+```{r echo=FALSE}
 
-```{r echo=FALSE}
-#load(file = "output/tables/tb.fm01.RData");tb.fm01
+load(file = "output/tables/tb.fm.RData")
+
+tb.col  <- c("Variables",1:4,1:4,1:4)
+tb.caption <- "Factor loadings and fit measures" 
+tb.foot <- paste0("; p<0.05=$*$; p<0.01=$**$; p<0.001=$***$.
+Standardised factor loadings using DWLS estimator.
+CFI = Comparative fit index (scaled), RMSEA = Root mean square error of approximation (scaled)")
+
+tb.fm <- tb.fm %>% 
+    mutate(vars = recode(vars, 
+                         "A. Who the more they try they manage to get bigger rewards that those who striveless."= "Perception Effort",
+                         "B. Who possess more talent they manage to obtain greater rewards than those who possess less talent."= "Perception Talent",
+                         "C. Who they have rich parents manage to get out ahead."="Perception rich parents",
+                         "D. Who they have good contacts they manage to get out ahead."="Perception contacts",
+                         "E. Who the more they try they should get greater rewards than those who they try less." = "Preferences Effort",
+                         "F. Who possess more talent they should get greater rewards than those who possess less talent."= "Preferences Talent",
+                         "G. It's fine that those who have rich parents get ahead" ="Preferences rich parents",
+                         "H. Is well that those who have good contacts get ahead." = "Preferences contacts"))
+
+tb.fm01<- kable(tb.fm, table_format, booktabs = T, escape = FALSE,align = "lcccccccccccc",col.names = tb.col, caption = tb.caption) %>% 
+          kable_styling(position = "center", full_width = F,font_size = 6) %>% 
+          add_header_above(header = c(" "=1,"Version 1"= 4,"Version 2"= 4,"Version 3"= 4)) %>% 
+          add_header_above(header = c(" "=1,"Factor loadings"= 12)) %>% 
+          row_spec(row = 8,hline_after = TRUE) %>%
+          add_indent(c(9:12))
+
+tb.fm2<-as.data.frame(tb.fm[1:8,])
+
+tb.fm02<- kable(tb.fm2, table_format, booktabs = T, escape = FALSE,align = "lcccccccccccc", col.names = tb.col, caption = tb.caption) %>% 
+          kable_styling(position = "center", full_width = F,font_size = 8) %>% 
+          add_header_above(header = c(" "=1,"Version 1"= 4,"Version 2"= 4,"Version 3"= 4)) %>% 
+          add_header_above(header = c(" "=1,"Factor loadings"= 12)) %>% 
+          row_spec(row = 8,hline_after = F)
+
+tb.fm02
+tb.fm01
 ```
@@
 con CFI superiores a .95 y RMSEA inferiores a .8, por contraparte ni un modelo logró un chi-square no significativo, aunque tanto el modelo aleatorio como el primero obtuvieron un adecuado chi-square ratio menor a 3.
-El primero de los órdenes fue el que obtuvo mejor ajuste (CFI=0.998, TLI= 0.995,RMSEA=0.037, χ2(df=14)=28,03, p = 0.014), seguido por el orden aleatorio de los ítems (CFI=0.992, TLI=0.984,RMSEA=0.051, χ2(df=14)=39.09, p < 0.001).
-Por su parte, la escala ordenada por temáticas parece generar un efecto framing en el cual la relación entre las percepciones y las preferencias parecen sobreestimadas, afectando de esta manera el ajuste (CFI=0.984, TLI=0.968,RMSEA=.071, χ2(df=14)=64.156, p < 0.001).
+El primero de los órdenes fue el que obtuvo mejor ajuste (CFI=0.998, TLI= 0.995,RMSEA=0.037, x2(df=14)=28,03, p = 0.014), seguido por el orden aleatorio de los ítems (CFI=0.992, TLI=0.984,RMSEA=0.051, x2(df=14)=39.09, p < 0.001).
+Por su parte, la escala ordenada por temáticas parece generar un efecto framing en el cual la relación entre las percepciones y las preferencias parecen sobreestimadas, afectando de esta manera el ajuste (CFI=0.984, TLI=0.968,RMSEA=.071, x2(df=14)=64.156, p < 0.001).
 
@@
 | **RMSEA** |          0.226          |            0.047            |      0.036      |
-| **χ2**    |         1830.839        |            68.661           |      40.250     |
+| **x2**    |         1830.839        |            68.661           |      40.250     |
 | **p**     |           .000          |             .000            |       .000      |
-| **χ2/gl** |          65.38          |             4.90            |       3.32      |
+| **x2/gl** |          65.38          |             4.90            |       3.32      |
 
 EL modelo teórico propuesto de cuatro factores ajustó, como se observa en la tabla 4, mejor que el modelo de contraste de 1 factor.
-El modelo teórico ajustó de manera relativamente adecuada, pues muestra indicadores óptimos para CFI= 0.987, TLI = 0.975 y RMSEA=.041, aunque posee indicadores deficientes para la prueba χ2(df=14)=67.6, p-value=.000.
+El modelo teórico ajustó de manera relativamente adecuada, pues muestra indicadores óptimos para CFI= 0.987, TLI = 0.975 y RMSEA=.041, aunque posee indicadores deficientes para la prueba x2(df=14)=67.6, p-value=.000.
 Para evaluar posibles mejoras de la escala, se analizaron las relaciones propuestas por los índices de modificación.
 Estos indican la existencia de dos cargas cruzadas no especificadas.
-Cuando se generó un modelo siguiendo estas recomendaciones, hubo una mejora considerable del modelo, aunque el nuevo modelo tampoco obtuvo un X2 ratio menor a 3 y obtuvo cargas factoriales muy bajas ʎ < 0.15, por lo tanto, siguiendo las recomendación de Brown (2006) de solo aceptar las propuesta de los índices de modificación cuando se posee teoria y evidencia sólida, se ha decidido no incorporar estos parámetros al modelo.
+Cuando se generó un modelo siguiendo estas recomendaciones, hubo una mejora considerable del modelo, aunque el nuevo modelo tampoco obtuvo un X2 ratio menor a 3 y obtuvo cargas factoriales muy bajas (cargas)< 0.15, por lo tanto, siguiendo las recomendación de Brown (2006) de solo aceptar las propuesta de los índices de modificación cuando se posee teoria y evidencia sólida, se ha decidido no incorporar estos parámetros al modelo.
 
@@
 
-<div/>
+# Anexos
 
-# Anexos
 `r anx_nums("aa")`

### 2020-05-20 docs/paper_y.tex
Rederizado
 termino pendientes sobre las tablas.

aun existen problemas de compativilidad html/pdf, funcionan pero hay que mejorar los estilos. Tengo algunas ideas de como solucionar esto.

subo todos los documentos por que no se que es necesario o no para el renderizado.

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Perceptions and Preferences for Meritocracy},
  colorlinks=true,
  linkcolor=blue,
  filecolor=Maroon,
  citecolor=Blue,
  urlcolor=Blue,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}

\title{Perceptions and Preferences for Meritocracy}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

\setstretch{1.5}
\hypertarget{introducciuxf3n}{%
\section{Introducción}\label{introducciuxf3n}}

La desigualdad económica se ha vuelto un tema que genera creciente
preocupación y malestar alrededor del mundo. Esto se ha expresado en una
serie de protestas como la emblemática ``occupy wall street'' el año
2011, así como también en una serie de análisis críticos respecto del
desarrollo del capitalismo y sus consecuencias (Piketty
\protect\hyperlink{ref-PikettyCapitalTwentyFirstCentury2014b}{2014}). En
este contexto, el estudio de las visiones, preferencias y percepciones
respecto de la desigualdad han adquirido relevancia en las ciencias
sociales, en temas como las preferencias redistributivas (Alesina and
Angeletos \protect\hyperlink{ref-alesina_Fairness_2005}{2005}; Dimick,
Rueda, and Stegmueller \protect\hyperlink{ref-dimick_Models_2018}{2018})
la legitimación de la desigualdad económica (Schröder
\protect\hyperlink{ref-schroder_Income_2017}{2017}) y el funcionamiento
de la meritocracia (Duru-Bellat and Tenret
\protect\hyperlink{ref-Duru-BellatWhoMeritocracyIndividual2012b}{2012}\protect\hyperlink{ref-Duru-BellatWhoMeritocracyIndividual2012b}{a};
Mijs
\protect\hyperlink{ref-MijsParadoxInequalityIncome2019b}{2019}\protect\hyperlink{ref-MijsParadoxInequalityIncome2019b}{a};
Reynolds and Xian
\protect\hyperlink{ref-ReynoldsPerceptionsMeritocracyLand2014b}{2014}\protect\hyperlink{ref-ReynoldsPerceptionsMeritocracyLand2014b}{a}).

En general, la meritocracia se define como un sistema de distribución de
recursos y recompensas basados en el mérito individual, que en su
concepción original es una suma de talento y esfuerzo (Young
\protect\hyperlink{ref-young_rise_1962}{1962}) y que pone en un lugar
secundario interferencia de factores estructurales como la herencia o
los contactos (Breen and Goldthorpe
\protect\hyperlink{ref-breenClassInequalityMeritocracy1999}{1999};
Saunders \protect\hyperlink{ref-saundersMightBritainBe1995}{1995}; Yair
\protect\hyperlink{ref-yairMeritocracy2007}{2007}; Land
\protect\hyperlink{ref-landWeSatTable2006}{2006}; Young
\protect\hyperlink{ref-youngRiseMeritocracy1994}{1994}). Una serie de
estudios han realizado críticas a la realización de este estándar moral
de distribución, planteando que es una promesa incumplida dada la
influencia preponderante de otros elementos más allá del mérito en el
estatus individual (Witteveen and Attewell
\protect\hyperlink{ref-WitteveenReconsideringmeritocraticpower2020a}{2020};
Arrow, Bowles, and Durlauf
\protect\hyperlink{ref-arrow_meritocracy_2000}{2000}; Goldthorpe
\protect\hyperlink{ref-goldthorpe_myth_2003}{2003}; Markovits
\protect\hyperlink{ref-markovits_Meritocracy_2019}{2019}). Por otro
lado, desde la psicología social y la sociología se han estudiado las
características y consecuencias de las creencias en la meritocracia, en
general basados en la hipótesis que mayor creencia en la meritocracia
lleva a una mayor legitimación de las desigualdades (Preminger
\protect\hyperlink{ref-PremingerMeritocracyserviceethnocracy2020}{2020};
Trump \protect\hyperlink{ref-TrumpWhenwhyeconomic2020}{2020}; Meredith
\protect\hyperlink{ref-Meredithsocietydivisibleblessed2020}{2020};
Hadjar \protect\hyperlink{ref-hadjar_meritokratie_2008}{2008}; Madeira
et al.
\protect\hyperlink{ref-MadeiraPrimesConsequencesSystematic2019}{2019}).

Debido al rol que cumplen las creencias meritocráticas dentro del
pensamiento neoliberal (Bay-Cheng et al.
\protect\hyperlink{ref-Bay-ChengTrackingHomoOeconomicus2015}{2015}), han
surgido múltiples investigaciones que evalúan la relación entre
creencias meritocráticas y diversos ámbitos sociales de la actualidad.
Se han desarrollado estudios que vinculan la meritocracia con el
reforzamiento de estereotipos socioeconómicos, de género y de etnias
(Madeira et al.
\protect\hyperlink{ref-MadeiraPrimesConsequencesSystematic2019}{2019};
Girerd and Bonnot
\protect\hyperlink{ref-GirerdNeoliberalismIdeologicalBarrier2020}{2020};
Preminger
\protect\hyperlink{ref-PremingerMeritocracyserviceethnocracy2020}{2020}).
Igualmente, se han desarrollado líneas de investigación que evalúan el
efecto de las creencias meritocráticas en el contexto educativo
(Generett and Olson
\protect\hyperlink{ref-GenerettStoriesWeTell2020}{2020}; Reay
\protect\hyperlink{ref-ReayPerilsPenaltiesMeritocracy2020}{2020}; Owens
and de St Croix
\protect\hyperlink{ref-OwensENGINESSOCIALMOBILITY2020}{2020}). y en el
contexto organizacional de las empresas (Pérez and Sabelis
\protect\hyperlink{ref-PerezAdvancingcareersmerit2020}{2020}; Aiello,
Cardamone, and Pupo
\protect\hyperlink{ref-AielloNewevidencefirmuniversity2019}{2019})

Para poder dar cuenta de los niveles de creencia en la meritocracia los
estudios a la fecha generalmente han utilizado algunos indicadores de
encuestas ya existentes, y en el menor de los casos se han creado
instrumentos ad-hoc. Sin embargo, y como mostraremos más adelante, las
formas de medición de meritocracia varían extremadamente entre estudios.
Muchas veces fenómenos similares se asocian a indicadores distintos, y
también ocurre que fenómenos distintos son medidos con indicadores
similares, todo lo cual hace dificulta la comparabilidad entre estudios
y el poder avanzar en la comprensión del concepto de meritocracia.

Basados en el análisis crítico de las formas de medición de meritocracia
a la fecha, el presente estudio propone un instrumento para medir y
relacionar dos aspectos claves en el estudio de la meritocracia:
percepciones y preferencias. Además, como un segundo eje de análisis
considera la generación de indicadores respecto de aspectos
meritocráticos y anti-meritocráticos, demostrando que no son los dos
polos de un mismo continuo como muchos estudios anteriores parecen
sugerir. La propuesta de medición además está orientada a generar un
instrumento lo más breve posible de manera que pueda ser utilizado en
encuestas de opinión.

\hypertarget{la-mediciuxf3n-de-los-aspectos-subjetivos-de-la-meritocracia}{%
\section{La medición de los aspectos subjetivos de la
meritocracia}\label{la-mediciuxf3n-de-los-aspectos-subjetivos-de-la-meritocracia}}

A continuación se presenta una revisión de una serie de investigaciones
que se han abocado al estudio de la meritocracia y que para ello han
hecho una propuesta de medición. El primer eje de análisis tiene que ver
con el uso del concepto ``creencias'' para referir a distintos aspectos
subjetivos relacionados con meritocracia. El segundo eje tiene que ver
con el uso de indicadores sobre aspectos anti-meritocráticos como el
polo opuesto de los meritocráticos.

\hypertarget{la-caja-negra-de-las-creencias-en-la-meritocracia}{%
\subsection{La caja negra de las creencias en la
meritocracia}\label{la-caja-negra-de-las-creencias-en-la-meritocracia}}

Para ejemplificar la confusión del uso de creencias en meritocracia
consideremos en primer lugar un artículo reciente de Mijs, donde se
expresa la siguiente definición: ``when I discuss meritocracy beliefs, I
am referring to citizens' belief in the importance of hard work relative
to structural factors.'' (Mijs
\protect\hyperlink{ref-mijs_paradox_2019}{2019}\protect\hyperlink{ref-mijs_paradox_2019}{b},
pg.9). In the operationalization, this is associated with the following
indicator: ``how important you think it is for getting ahead in life:
(a) hard work'', scored in a 1 to 5 likert scale. There are several
assumptions behind this decision that are worth discussing:

\begin{enumerate}
\def\labelenumi{\alph{enumi}.}
\item
  Dimensionality: the item used by Mijs is part of an battery present in
  several international surveys usually called ``get ahead'', which
  presents a series of indicators related to what people consider
  important to get ahead in life: hard work, education, ambition,
  wealthy family, right connections, religion, race and gender. For
  Mijs, other aspects as education, that could be associated to talent,
  are not meritocratic: ``Hard work is arguably the most meritocratic
  part of Michael Young's equation, `Merit = Intelligence + Effort', for
  the simple fact that intelligence itself is conditioned by a
  nonmeritocratic factor: who your parents happen to be'' (p.5). This is
  a very strong assumption: effort would not depend on parents
  influence, and talent is not meritocratic. A similar approach only
  considering effort as the main element of meritocracy is taken by
  other authors ho only included effort as an indicator of meritocratic
  beliefs (Maitner
  \protect\hyperlink{ref-MaitnerEmotionalreactionsunequal2015}{2015};
  Major et al.
  \protect\hyperlink{ref-MajorPerceiveddiscriminationworldview2007}{2007};
  Bubak
  \protect\hyperlink{ref-BubakPerceptionsmeritocracynote2019}{2019};
  {\textbf{???}}). We suggest here is that questions such as whether
  talent is not meritocratic (or less meritocratic than effort) should
  be open to empirical scrutiny and not disregarded upfront based on
  scholars' assumptions. Besides, the use of single indicators for
  measurement a construct is based on the assumption of a measurement
  without error.
\item
  Beliefs: the ``get ahead'' battery refers to ``how important you think
  it is'', considered by Mijs as a belief in meritocracy. Nevertheless,
  another version of this same battery asks about ``how important you
  think it \emph{should} be''. Which one of both is the ``belief''? If
  we consider belief as related to a normative content, actually the
  more close version is the ``should be'', being the other one more
  close to the perception of the situation. Such distinction was already
  made by Duru-Bellat and Tenret
  (\protect\hyperlink{ref-duru-bellat_whos_2012}{2012}\protect\hyperlink{ref-duru-bellat_whos_2012}{b}),
  who used the item ``how important should the number of years spent in
  education and training be in deciding how much money people ought to
  earn?'' for ``desired'' meritocracy, whereas for ``perceived''
  meritocracy they use two items: ``Would you say that in your country,
  people are rewarded for their efforts?'' and ``\ldots{} people are
  rewarded for their skills?''. Is the belief in meritocracy a
  perception or a preference? A comprehensive measurement should be in
  line with what Duru-Bellat and Tenret
  (\protect\hyperlink{ref-duru-bellat_whos_2012}{2012}\protect\hyperlink{ref-duru-bellat_whos_2012}{b})
  did, in order to open up possibilities of analyzing whether
  perceptions and preferences are actually the same (i.e.~correlation
  close to 1) or they are different aspects of the same phenomenon.
  Therefore, two main drawbacks of the term ``beliefs'' are that (i) it
  hides different operationalizations of the concept of meritocracy, and
  (ii) it blurs the distinction between perceptions and preferences. We
  argue that perceptions and preferences should be disentangled and used
  together, specially when analyzing preferences. As Son Hing et al.
  (\protect\hyperlink{ref-son_hing_merit_2011-1}{2011}) have pointed
  out, ``People can believe that outcomes ought to be distributed on the
  basis of merit and yet vary in their perceptions of whether this is
  how society currently operates'' (p.~435). In other words, preferences
  should be interpreted in the context of perception: a large preference
  for meritocracy certainly means something totally different for
  someone perceiving high meritocracy than for someone perceiving low
  meritocracy.
\item
  Non-meritocratic aspects: Mijs
  (\protect\hyperlink{ref-mijs_paradox_2019}{2019}\protect\hyperlink{ref-mijs_paradox_2019}{b})
  make reference to some non-meritocratic aspects as talent, letting
  this out of the measurement. A different approach was followed by
  Kunovich and Slomczynski
  (\protect\hyperlink{ref-kunovich_systems_2007}{2007}), who decide to
  include non-meritocratic elements in the measurement of meritocracy.
  Using the items' battery listing a number of reasons about ``How
  important should be in deciding pay\ldots{}'' (as Duru-Bellat and
  Tenret
  (\protect\hyperlink{ref-duru-bellat_whos_2012}{2012}\protect\hyperlink{ref-duru-bellat_whos_2012}{b})
  for desired meritocracy), he decided that reasons as education and
  responsibility are meritocratic and pointed 1 if considered essential,
  whereas reasons such as having a family and children were pointed 1 if
  they were considered ``not important at all'' (i.e.~reverse coded). A
  similar approach was taken by Newman, Johnston, and Lown
  (\protect\hyperlink{ref-newman_false_2015}{2015}), reverse-coding
  non-meritocratic items, similarly to what occurs with the ``Preference
  for the Merit Principle Scale'' (Davey et al.
  \protect\hyperlink{ref-davey_preference_1999}{1999}). The assumption
  that meritocratic and non-meritocratic elements are the poles of the
  same continuum was analyzed by Reynolds and Xian
  (\protect\hyperlink{ref-reynolds_perceptions_2014}{2014}\protect\hyperlink{ref-reynolds_perceptions_2014}{b})
  using the same ``get ahead'' perceptions' battery items. They consider
  education, ambition and hard work as meritocratic and other reasons
  such as wealthy family and right connections and non-mertitocratic.
  Nevertheless, despite making this distinction the author ends up
  substracting one dimension from the other, assuming that they are two
  poles of the same continuum as Kunovich and Slomczynski
  (\protect\hyperlink{ref-kunovich_systems_2007}{2007}) did.
\item
  Accounting for measurement error: finally, most of the studies in
  meritocracy so far have not incorporated the issue of measurement
  error ({\textbf{???}}), using single indicators and/or composite
  indexes for measuring meritocracy. This means assuming that the
  construct is measured perfectly by the indicators chosen, going as far
  as proposing that "we decided to rely on a summative index. In
  choosing this strategy of index construction, we argue that
  \emph{support for meritocracy is not a latent variable} (Kunovich and
  Slomczynski \protect\hyperlink{ref-kunovich_systems_2007}{2007},
  653--54). Some advances were done by Reynolds and Xian
  (\protect\hyperlink{ref-reynolds_perceptions_2014}{2014}\protect\hyperlink{ref-reynolds_perceptions_2014}{b})
  by doing a principal component analysis of meritocratic and
  non-meritocratic dimensions, but somewhat contradictorily they end up
  in a sum index despite proving multidimensionality.
\end{enumerate}

\hypertarget{an-instrument-proposal}{%
\subsection{An instrument proposal}\label{an-instrument-proposal}}

Based on the previous limitations in the measurement of meritocracy
presented in the previous section, in this paper we propose and test an
instrument with the following characteristics:

\begin{itemize}
\tightlist
\item
  Multidimensional, incorporating previous distinctions between
  preferences and perceptions as well as between meritocratic and
  non-meritocratic aspects.
\item
  Multiple indicators for each dimension, in order to account for
  measurement error in a confirmatory factor analysis context.
\item
  Based on previous indicators as far as possible, for the sake of
  comparability between studies
\item
  Brief, as to be used in a regular survey. In this point it differs for
  instance from the proposal of the ``Preference for the Merit Principle
  Scale'' (Davey et al.
  \protect\hyperlink{ref-davey_preference_1999}{1999}), as they use 15
  items just for one dimension (besides the problem of reverse-coding
  non-meritocratic items).
\end{itemize}

The proposed measurement framework is based on two axis of analysis, as
depicted in Table 1:

The first axis distinguishes between different types of ``beliefs'',
using instead the terms perceptions and preferences for meritocracy
(Duru-Bellat and Tenret
\protect\hyperlink{ref-duru-bellat_whos_2012}{2012}\protect\hyperlink{ref-duru-bellat_whos_2012}{b};
Son Hing et al. \protect\hyperlink{ref-son_hing_merit_2011-1}{2011}).
Perceptions refers to the extent to which people see meritocracy working
in their society, which in terms of measurement relates to the ``reasons
to get ahead'' battery, whereas preferences refer to normative
expectations that are usually linked to a ``should'' expression
(e.g.~whether hard work should be related to payment). The second axis
consider the distinction between meritocratic and non-meritocratic
dimensions (Reynolds and Xian
\protect\hyperlink{ref-reynolds_perceptions_2014}{2014}\protect\hyperlink{ref-reynolds_perceptions_2014}{b}).
This aspect has been usually treated as different ends of a same
continuum in part of the previous research, an assumption that requieres
empirical scrutiny. These non-meritocratic elements usually refer to the
use of personal contacts or family advantages to get ahead in life.

Regarding the selection of indicators, most of them some adaptations of
previous items that where streamlined in order to a better fit with the
specific dimensions. For meritocratic indicators we use effort and
talent as the main components of the traditional concept of merit as
defined by Young (\protect\hyperlink{ref-young_rise_1962}{1962}),
whereas for non-meritocratic dimensions we use having rich parents and
good contacts. The description of the specific items is presented in the
methodology section.

\textbf{Hypotheses}

\begin{itemize}
\item
  H1. The perception of meritocracy is a latent variable based on
  indicators of the importance attributed to talent and the effort to
  get ahead in life.
\item
  H2. The non-meritocratic perception is a latent variable that derives
  from two indicators related to the agreement with the statement that
  people with contacts and rich parents manage to get ahead.
\item
  H3. Meritocratic preferences behave as the latent variable based on a
  normative value of effort and talent.
\item
  H4. Non-meritocratic preferences behave as a latent variable based on
  the normative value of the use of personal contacts and having rich
  parents.
\end{itemize}

\hypertarget{methodology}{%
\section{Methodology}\label{methodology}}

\hypertarget{data-collection}{%
\subsection{Data collection}\label{data-collection}}

The data was obtained through an online questionnaire which was part of
a larger study on meritocracy and preferences developed in Chile in 2019
and funded by the national scientific agency FONDECYT. The questionnaire
was programmed in Qualtrics and the fieldwork was in charge of an
external online survey agency
(\href{www.netquest.cl}{netquest.cl})during December 2019 and January
2020. The sample was selected from a non-probabilistic design in three
large cities in Chile. The quota method based on age, sex and
educational level was used. Quotas were generated based on the survey of
the Public Studies Center (CEP, 2019), which has a high prestige in the
country and is also the counterpart agency of ISSP (International Social
Survey Programme) in Chile. A total sample of 2141 people was collected,
excluding those who did not answer the questions on the scale and those
who did not accept informed consent. As it usually occurs online
samples, there were some limitations in achieving the quotas for lower
educational levels.

\hypertarget{instrument-design}{%
\subsection{Instrument design}\label{instrument-design}}

The proposed scale of perceptions and preferences about meritocracy
consists of 8 indicators that are grouped into 4 dimensions referred
above: Perceptions (meritocratic/nonmeritocratic) and preferences
(meritocratic/non-meritocratic). In order to achieve at least some
comparability with previous studies, the questions were adapted from the
items battery ``reasons to get ahead'' (ISSP/GSS), which are mostly used
for operationalizing meritocracy in previous empirical studies (Mijs
\protect\hyperlink{ref-mijs_paradox_2019}{2019}\protect\hyperlink{ref-mijs_paradox_2019}{b};
Duru-Bellat and Tenret
\protect\hyperlink{ref-duru-bellat_whos_2012}{2012}\protect\hyperlink{ref-duru-bellat_whos_2012}{b};
Reynolds and Xian
\protect\hyperlink{ref-reynolds_perceptions_2014}{2014}\protect\hyperlink{ref-reynolds_perceptions_2014}{b}).
The eight items ordered according to dimensions are presented in Table
2. These 8 likert-type items have 5 response alternatives ranging from
``Completely disagree''(1) to ``Completely agree'' (5).

\includegraphics{../input/images/tabla2.png}

1 XX ojo con esta imagen, titularla en el documento, no en la imagen
misma

\hypertarget{administration-sets}{%
\subsection{Administration sets}\label{administration-sets}}

With the objective of evaluating the effect of the order of the
indicators, the respondents (\emph{n = 2141}) were randomly divided into
three groups as explained in \textbf{Diagram 1}. The scale was presented
to the first group (\emph{n = 712}) in the order that appears in Table
2, in such a way that the items of the same factor are found together.
The second group (\emph{n = 717})was presented with the scale with the
items of the same factor separated from each other, in such a way that
they were consulted at the same time for their perceptions and
preferences of a particular topic (e.g.~distribution according to
effort). Finally, in the third group (\emph{n = 712}) the items were
completely randomized.

2 XX agregar número de sujetos por grupo

3 XX referenciar y titular figuras y tablas compatible con html/pdf

\includegraphics{../input/images/app_mod.png}

\hypertarget{methods}{%
\subsection{Methods}\label{methods}}

For testing the scale's underlying constructs we estimate confirmatory
factor analysis models (CFA). The model estimates one factor for each
dimension, as represented in the following figure:

\includegraphics{../output/images/meas01.png}

CFA was conducted using the \texttt{lavaan} package (version 0.6-3;
Rosseel, 2020) with diagonally weighted least squares (DWLS) estimation
due to the items' ordinal level of measurement (Kline, 2016; Rosseel,
2020). As recommended by Brown (2008), we assessed model fit by jointly
considering the comparative fit index and Tucker-Lewis Index (CFI and
TLI; acceptable fit \textgreater{} 0.95), Root of the average squared
residual approximation (RMSEA; acceptable fit \textless{} 0.08),
Chi-square: (p-value; acceptable fit \textgreater{} 0.05, and Chi-square
ratio:\textgreater{} 3).

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{descriptive-analysis}{%
\subsection{Descriptive analysis}\label{descriptive-analysis}}

4 XX producir tabla y llamar directamente desde la fuente o como imagen
{[}\textbf{Pendiente}{]} 5 XX Agregar gráficos de frecuencia , puede ser
tipo trellis 2 x 4, arriba percepción abajo preferencias; la idea es
generar un input más claro para la siguiente idea:

\includegraphics{../output/images/plotlikert.png}

En general se observa que hay una mayor percepción de aspectos no
meritocráticos que meritocráticos, mientras que en el caso de
preferencias ocurre lo opuesto. En cuanto a las preferencias, llama la
atención el rol preponderante del esfuerzo por sobre el talento.

\includegraphics{paper_files/figure-latex/unnamed-chunk-3-1.pdf}

6 - XX Sobre tablas en general, entonces: - dejar tabla de descriptivos
que ahora aparece junto a la de correlaciones {[}\emph{Pendiente} {]} -
agregar gráficos (listo) - reemplazar tabla de correlaciones por un
corrplot, creo que se aprecia más claramente el sentido de las
asociaciones (listo)

\hypertarget{models-estimation}{%
\subsection{Models estimation}\label{models-estimation}}

7 XX En tabla de cargas factoriales, los fit aparecen corridos hacia la
izquierda {[}\textbf{Pendiente}{]}

En una primera instancia se probó el ajuste del modelo con tres
distintos órdenes de los ítems, los cuales fueron aplicados a un tercio
de la muestra de la primera ola cada uno. El primer orden de los ítems
corresponde al orden de de la tabla 2, en el que primero se pregunta a
los encuestados por todas las percepciones y después por todas las
preferencias, en el segundo orden se les pregunta seguidamente por
percepciones y preferencias respecto a un mismo tema, por ejemplo, al
papel del esfuerzo, y en el tercero, los ítems aparecen de manera
aleatoria.

Todos los modelos independiente de los órdenes obtuvieron un ajuste
adecuado. con CFI superiores a .95 y RMSEA inferiores a .8, por
contraparte ni un modelo logró un chi-square no significativo, aunque
tanto el modelo aleatorio como el primero obtuvieron un adecuado
chi-square ratio menor a 3. El primero de los órdenes fue el que obtuvo
mejor ajuste (CFI=0.998, TLI= 0.995,RMSEA=0.037, x2(df=14)=28,03, p =
0.014), seguido por el orden aleatorio de los ítems (CFI=0.992,
TLI=0.984,RMSEA=0.051, x2(df=14)=39.09, p \textless{} 0.001). Por su
parte, la escala ordenada por temáticas parece generar un efecto framing
en el cual la relación entre las percepciones y las preferencias parecen
sobreestimadas, afectando de esta manera el ajuste (CFI=0.984,
TLI=0.968,RMSEA=.071, x2(df=14)=64.156, p \textless{} 0.001).

8 XX explicar mejor lo que viene: invarianza de qué y por qué se
realiza.

Si bien todas las pruebas obtuvieron indicadores relativamente
adecuados, las diferencias mencionadas entre los ajustes de los modelos
según orden son estadísticamente significativas, lo cual fue evaluado a
partir de un análisis de invarianza. Se concluye que entre los tres
órdenes no existe invarianza configuracional, es decir, no poseen la
misma dimensionalidad y por ello no ajustan igualmente al modelo teórico
({\textbf{???}}). Esto se debe al efecto producido por la aparición
conjunta de los ítems de un mismo factor en el orden 1, lo cual aumenta
el ajuste del modelo. Además, en el orden 2, al preguntarse seguidamente
por la percepción y la preferencia en torno al mismo indicador, aumentan
las relaciones cruzadas disminuyendo el ajuste del modelo. De modo
coherente, el modelo aleatorio presentó un ajuste intermedio entre el
orden 1 y el orden 2.

EL modelo teórico propuesto de cuatro factores ajustó, como se observa
en la tabla 4, mejor que el modelo de contraste de 1 factor. El modelo
teórico ajustó de manera relativamente adecuada, pues muestra
indicadores óptimos para CFI= 0.987, TLI = 0.975 y RMSEA=.041, aunque
posee indicadores deficientes para la prueba x2(df=14)=67.6,
p-value=.000. Para evaluar posibles mejoras de la escala, se analizaron
las relaciones propuestas por los índices de modificación. Estos indican
la existencia de dos cargas cruzadas no especificadas. Cuando se generó
un modelo siguiendo estas recomendaciones, hubo una mejora considerable
del modelo, aunque el nuevo modelo tampoco obtuvo un X2 ratio menor a 3
y obtuvo cargas factoriales muy bajas (cargas)\textless{} 0.15, por lo
tanto, siguiendo las recomendación de Brown (2006) de solo aceptar las
propuesta de los índices de modificación cuando se posee teoria y
evidencia sólida, se ha decidido no incorporar estos parámetros al
modelo.

\begin{itemize}
\tightlist
\item
  agregar cuáles fueron las cargas sugeridas.
\end{itemize}

\hypertarget{conclusiuxf3n}{%
\section{Conclusión}\label{conclusiuxf3n}}

Considerando la ventaja del orden, respecto aleatorio despeja la
posibilidad del efecto framing, es decir, de que el resultado del modelo
se deba al orden de las preguntas, se ha decidido seguir utilizando el
orden 3.

\hypertarget{conclusiones}{%
\section{Conclusiones}\label{conclusiones}}

uso conjunto percepciones-preferencias

\hypertarget{anexos}{%
\section{Anexos}\label{anexos}}

\hypertarget{bibliografuxeda}{%
\section*{Bibliografía}\label{bibliografuxeda}}
\addcontentsline{toc}{section}{Bibliografía}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-AielloNewevidencefirmuniversity2019}{}%
Aiello, Francesco, Paola Cardamone, and Valeria Pupo. 2019. ``New
Evidence on the Firm-University Linkages in Europe. The Role of
Meritocratic Management Practices.'' \emph{International Review of
Applied Economics} 33 (6): 813--28.
\url{https://doi.org/10.1080/02692171.2019.1608917}.

\leavevmode\hypertarget{ref-alesina_Fairness_2005}{}%
Alesina, A., and G. Angeletos. 2005. ``Fairness and Redistribution.''
\emph{American Economic Review}, 960--80.

\leavevmode\hypertarget{ref-arrow_meritocracy_2000}{}%
Arrow, Kenneth J., Samuel Bowles, and Steven N. Durlauf, eds. 2000.
\emph{Meritocracy and Economic Inequality}. Princeton, N.J: Princeton
University Press.

\leavevmode\hypertarget{ref-Bay-ChengTrackingHomoOeconomicus2015}{}%
Bay-Cheng, Laina Y., Caroline C. Fitz, Natalie M. Alizaga, and Alyssa N.
Zucker. 2015. ``Tracking Homo Oeconomicus: Development of the Neoliberal
Beliefs Inventory.'' \emph{Journal of Social and Political Psychology} 3
(1): 71--88. \url{https://doi.org/10.5964/jspp.v3i1.366}.

\leavevmode\hypertarget{ref-breenClassInequalityMeritocracy1999}{}%
Breen, Richard, and John H. Goldthorpe. 1999. ``Class Inequality and
Meritocracy: A Critique of Saunders and an Alternative Analysis1.''
\emph{The British Journal of Sociology} 50 (1): 1--27.
\url{https://doi.org/10.1111/j.1468-4446.1999.00001.x}.

\leavevmode\hypertarget{ref-BubakPerceptionsmeritocracynote2019}{}%
Bubak, Oldrich. 2019. ``Perceptions of Meritocracy: A Note on China.''
\emph{Asian Journal of Comparative Politics} 4 (2): 192--209.
\url{https://doi.org/10.1177/2057891118806065}.

\leavevmode\hypertarget{ref-davey_preference_1999}{}%
Davey, L. M., D. R. Bobocel, L. S. Son Hing, and M. P. Zanna. 1999.
``Preference for the Merit Principle Scale: An Individual Difference
Measure of Distributive Justice Preferences.'' \emph{Social Justice
Research} 12 (3): 223--40.

\leavevmode\hypertarget{ref-dimick_Models_2018}{}%
Dimick, Matthew, David Rueda, and Daniel Stegmueller. 2018. ``Models of
Other-Regarding Preferences, Inequality, and Redistribution.''
\emph{Annual Review of Political Science} 21 (1): 441--60.
\url{https://doi.org/10.1146/annurev-polisci-091515-030034}.

\leavevmode\hypertarget{ref-Duru-BellatWhoMeritocracyIndividual2012b}{}%
Duru-Bellat, Marie, and Elise Tenret. 2012a. ``Who's for Meritocracy?
Individual and Contextual Variations in the Faith.'' \emph{Comparative
Education Review} 56 (2): 223--47. \url{https://doi.org/10.1086/661290}.

\leavevmode\hypertarget{ref-duru-bellat_whos_2012}{}%
---------. 2012b. ``Who's for Meritocracy? Individual and Contextual
Variations in the Faith.'' \emph{Comparative Education Review} 56 (2):
223--47. \url{https://doi.org/10.1086/661290}.

\leavevmode\hypertarget{ref-GenerettStoriesWeTell2020}{}%
Generett, Gretchen Givens, and Amy M. Olson. 2020. ``The Stories We
Tell: How Merit Narratives Undermine Success for Urban Youth.''
\emph{Urban Education} 55 (3): 394--423.
\url{https://doi.org/10.1177/0042085918817342}.

\leavevmode\hypertarget{ref-GirerdNeoliberalismIdeologicalBarrier2020}{}%
Girerd, Lola, and Virginie Bonnot. 2020. ``Neoliberalism: An Ideological
Barrier to Feminist Identification and Collective Action.'' \emph{Social
Justice Research} 33 (1): 81--109.
\url{https://doi.org/10.1007/s11211-020-00347-8}.

\leavevmode\hypertarget{ref-goldthorpe_myth_2003}{}%
Goldthorpe, John. 2003. ``The Myth of Education-Based Meritocracy.''
\emph{New Economy} 10 (4): 234--39.
\url{https://doi.org/10.1046/j.1468-0041.2003.00324.x}.

\leavevmode\hypertarget{ref-hadjar_meritokratie_2008}{}%
Hadjar, Andreas. 2008. \emph{Meritokratie Als Legitimationsprinzip}.
Wiesbaden: VS Verlag.

\leavevmode\hypertarget{ref-kunovich_systems_2007}{}%
Kunovich, Sheri, and Kazimierz M. Slomczynski. 2007. ``Systems of
Distribution and a Sense of Equity: A Multilevel Analysis of
Meritocratic Attitudes in Post-Industrial Societies.'' \emph{European
Sociological Review} 23 (5): 649--63.
\url{https://doi.org/10.1093/esr/jcm026}.

\leavevmode\hypertarget{ref-landWeSatTable2006}{}%
Land, Hilary. 2006. ``We Sat down at the Table of Privilege and
Complained About the Food \textsuperscript{1}.'' \emph{The Political
Quarterly} 77 (s1): 45--60.
\url{https://doi.org/10.1111/j.1467-923X.2006.00780.x}.

\leavevmode\hypertarget{ref-MadeiraPrimesConsequencesSystematic2019}{}%
Madeira, Ana Filipa, Rui Costa-Lopes, John F. Dovidio, Gonçalo Freitas,
and Mafalda F. Mascarenhas. 2019. ``Primes and Consequences: A
Systematic Review of Meritocracy in Intergroup Relations.''
\emph{Frontiers in Psychology} 10 (September).
\url{https://doi.org/10.3389/fpsyg.2019.02007}.

\leavevmode\hypertarget{ref-MaitnerEmotionalreactionsunequal2015}{}%
Maitner, Angela T. 2015. ``Emotional Reactions to Unequal Payment: The
Impact of Meritocratic Ideology and Salary Negotiability.'' \emph{Group
Processes \& Intergroup Relations} 18 (2): 153--72.
\url{https://doi.org/10.1177/1368430214542255}.

\leavevmode\hypertarget{ref-MajorPerceiveddiscriminationworldview2007}{}%
Major, Brenda, Cheryl R. Kaiser, Laurie T. O'Brien, and Shannon K.
McCoy. 2007. ``Perceived Discrimination as Worldview Threat or Worldview
Confirmation: Implications for Self-Esteem.'' \emph{Journal of
Personality and Social Psychology} 92 (6): 1068--86.
\url{https://doi.org/10.1037/0022-3514.92.6.1068}.

\leavevmode\hypertarget{ref-markovits_Meritocracy_2019}{}%
Markovits, Daniel. 2019. \emph{The Meritocracy Trap: How America's
Foundational Myth Feeds Inequality, Dismantles the Middle Class, and
Devours the Elite}. New York: Penguin Press.

\leavevmode\hypertarget{ref-Meredithsocietydivisibleblessed2020}{}%
Meredith, Stephen. 2020. ``A `Society \ldots{} Divisible into the
Blessed and the Unblessed': Michael Young and Meritocracy in Postwar
Britain: Michael Young and Meritocracy in Postwar Britain.'' \emph{The
Political Quarterly}, March.
\url{https://doi.org/10.1111/1467-923X.12837}.

\leavevmode\hypertarget{ref-MijsParadoxInequalityIncome2019b}{}%
Mijs, Jonathan J B. 2019a. ``The Paradox of Inequality: Income
Inequality and Belief in Meritocracy Go Hand in Hand.''
\emph{Socio-Economic Review}, January.
\url{https://doi.org/10.1093/ser/mwy051}.

\leavevmode\hypertarget{ref-mijs_paradox_2019}{}%
---------. 2019b. ``The Paradox of Inequality: Income Inequality and
Belief in Meritocracy Go Hand in Hand.'' \emph{Socio-Economic Review},
January. \url{https://doi.org/10.1093/ser/mwy051}.

\leavevmode\hypertarget{ref-newman_false_2015}{}%
Newman, Benjamin J., Christopher D. Johnston, and Patrick L. Lown. 2015.
``False Consciousness or Class Awareness? Local Income Inequality,
Personal Economic Position, and Belief in American Meritocracy.''
\emph{American Journal of Political Science} 59 (2): 326--40.
\url{https://doi.org/10.1111/ajps.12153}.

\leavevmode\hypertarget{ref-OwensENGINESSOCIALMOBILITY2020}{}%
Owens, John, and Tania de St Croix. 2020. ``ENGINES OF SOCIAL MOBILITY?
NAVIGATING MERITOCRATIC EDUCATION DISCOURSE IN AN UNEQUAL SOCIETY.''
\emph{British Journal of Educational Studies}, January, 1--21.
\url{https://doi.org/10.1080/00071005.2019.1708863}.

\leavevmode\hypertarget{ref-PerezAdvancingcareersmerit2020}{}%
Pérez, Andrés, and Ida Sabelis. 2020. ``Advancing Careers Through
`Merit': A Rationalized-Sensemaking Narrative in Hierarchical
Organizations.'' \emph{Culture and Organization} 26 (4): 315--32.
\url{https://doi.org/10.1080/14759551.2019.1601723}.

\leavevmode\hypertarget{ref-PikettyCapitalTwentyFirstCentury2014b}{}%
Piketty, Thomas. 2014. \emph{Capital in the Twenty-First Century}.
Translated by Arthur Goldhammer. Cambridge Massachusetts: The Belknap
Press of Harvard University Press.

\leavevmode\hypertarget{ref-PremingerMeritocracyserviceethnocracy2020}{}%
Preminger, Jonathan. 2020. ``Meritocracy in the Service of Ethnocracy.''
\emph{Citizenship Studies} 24 (2): 247--63.
\url{https://doi.org/10.1080/13621025.2020.1720604}.

\leavevmode\hypertarget{ref-ReayPerilsPenaltiesMeritocracy2020}{}%
Reay, Diane. 2020. ``The Perils and Penalties of Meritocracy:
Sanctioning Inequalities and Legitimating Prejudice.'' \emph{The
Political Quarterly}, March, 1467--923X.12829.
\url{https://doi.org/10.1111/1467-923X.12829}.

\leavevmode\hypertarget{ref-ReynoldsPerceptionsMeritocracyLand2014b}{}%
Reynolds, Jeremy, and He Xian. 2014a. ``Perceptions of Meritocracy in
the Land of Opportunity.'' \emph{Research in Social Stratification and
Mobility} 36 (June): 121--37.
\url{https://doi.org/10.1016/j.rssm.2014.03.001}.

\leavevmode\hypertarget{ref-reynolds_perceptions_2014}{}%
---------. 2014b. ``Perceptions of Meritocracy in the Land of
Opportunity.'' \emph{Research in Social Stratification and Mobility} 36
(June): 121--37. \url{https://doi.org/10.1016/j.rssm.2014.03.001}.

\leavevmode\hypertarget{ref-saundersMightBritainBe1995}{}%
Saunders, Peter. 1995. ``Might Britain Be a Meritocracy?''
\emph{Sociology} 29 (1): 23--41.
\url{https://doi.org/10.1177/0038038595029001003}.

\leavevmode\hypertarget{ref-schroder_Income_2017}{}%
Schröder, Martin. 2017. ``Is Income Inequality Related to Tolerance for
Inequality?'' \emph{Social Justice Research} 30 (1): 23--47.
\url{https://doi.org/10.1007/s11211-016-0276-8}.

\leavevmode\hypertarget{ref-son_hing_merit_2011-1}{}%
Son Hing, Leanne S., D. Ramona, Mark P. Zanna, Donna M. Garcia,
Stephanie S. Gee, and Katie Orazietti. 2011. ``The Merit of
Meritocracy.'' \emph{Journal of Personality and Social Psychology} 101
(3): 433--50. \url{https://doi.org/10.1037/a0024618}.

\leavevmode\hypertarget{ref-TrumpWhenwhyeconomic2020}{}%
Trump, Kris-Stella. 2020. ``When and Why Is Economic Inequality Seen as
Fair.'' \emph{Current Opinion in Behavioral Sciences} 34 (August):
46--51. \url{https://doi.org/10.1016/j.cobeha.2019.12.001}.

\leavevmode\hypertarget{ref-WitteveenReconsideringmeritocraticpower2020a}{}%
Witteveen, Dirk, and Paul Attewell. 2020. ``Reconsidering the
`Meritocratic Power of a College Degree'.'' \emph{Research in Social
Stratification and Mobility} 66 (April): 100479.
\url{https://doi.org/10.1016/j.rssm.2020.100479}.

\leavevmode\hypertarget{ref-yairMeritocracy2007}{}%
Yair, Gad. 2007. ``Meritocracy.'' In \emph{The Blackwell Encyclopedia of
Sociology}, edited by George Ritzer. Oxford, UK: John Wiley \& Sons,
Ltd. \url{https://doi.org/10.1002/9781405165518.wbeosm082}.

\leavevmode\hypertarget{ref-young_rise_1962}{}%
Young, M. 1962. \emph{The Rise of the Meritocracy}. Baltimore: Penguin
Books.

\leavevmode\hypertarget{ref-youngRiseMeritocracy1994}{}%
Young, Michael Dunlop. 1994. \emph{The Rise of the Meritocracy}. New
Brunswick, N.J., U.S.A: Transaction Publishers.

\end{document}

### 2020-05-20 production/prod_analysis-cfa.rmd
Rederizado
 termino pendientes sobre las tablas.

aun existen problemas de compativilidad html/pdf, funcionan pero hay que mejorar los estilos. Tengo algunas ideas de como solucionar esto.

subo todos los documentos por que no se que es necesario o no para el renderizado.

@@
 date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
-css: "../input/css/style.css"
+css: "../input/css/custom.css"
 output:
@@
 ---
+
+```{r}
+table_format = if(is_html_output()) {
+  "html"
+} else if(is_latex_output()) {
+  "latex"
+}
+```
 
@@
 fm.df$nobs <- c(nobs(fit1_o),nobs(fit2_o),nobs(fit3_o)) 
-fm.df <- data.frame(t(fm.df)); colnames(fm.df) <- c("M1.1","M2.1","M3.1")
+fm.df <- data.frame(t(fm.df)); colnames(fm.df) <- c("M1.2","M2.2","M3.2")
+
 #------ merge ------#
@@
 save(tb.fm01,file = "output/tables/tb.fm01.RData")
+
+save(tb.fm,file = "output/tables/tb.fm.RData")
+save(fm.df,file = "output/tables/fm.df.RData")
+save(tb.load,file = "output/tables/tb.load.RData")
+
 ```
@@
 library(sjPlot)
-
-dat_merit <- dat04 %>% select("Effort"=perc_effort,"Talent"=perc_talent, "Rich Parents"=perc_wpart, "Contacts"=perc_netw, "Effort"=pref_effort,"Talent"=pref_talent,"Rich Parents"=pref_wpart, "Contacts"=pref_netw)
+library(dplyr)
+
+#install.packages("cowplot")
+library(cowplot)
+
+dat_merit <- dat04 %>% select("a.Effort"=perc_effort,"a.Talent"=perc_talent, "a.Rich Parents"=perc_wpart, "a.Contacts"=perc_netw, "b.Effort"=pref_effort,"b.Talent"=pref_talent,"b.Rich Parents"=pref_wpart, "b.Contacts"=pref_netw)
 
@@
             reverse.scale=TRUE)
+plotlikert
 

### 2020-05-26 docs/paper.Rmd
Render final
Se mejoran las tablas, los titulos y l pdf.
ademas se agrega la opcion pdf dentro de las pestañas del _site.yml

@@
 title: "Perceptions and Preferences for Meritocracy"
+author:
+- Juan Carlos Castillo 
+- Julio Iturra 
+- Francisco Meneses 
 css: "../input/css/custom.css"
@@
 output:
+  html_document:
+    theme: flatly
+    number_sections: true
+    always_allow_html: yes
   pdf_document:
@@
     number_sections: true
-  html_document:
-    theme: flatly
-    number_sections: true
-    always_allow_html: yes
 linkcolor: blue
@@
   chunk_output_type: console
+geometry: margin=0.78in
 ---
+
+```{r include=FALSE}
+# for render in pdf run rmarkdown::render_site("docs/paper.Rmd", output_format = "all")
+# clean #in the yml
+```
 
@@
 library(captioner)
+#install.packages("estimatr")
+library(estimatr)
+library(texreg)
+
 ```
@@
 
+table_format2 = if(is_html_output()) {
+  T
+} else if(is_latex_output()) {
+  F
+}
+
+```
+
+```{r include=FALSE}
+table_nums <- captioner(prefix = "Table")
+fig_nums   <- captioner(prefix = "Figure")
 ```
@@
 ```{r message=FALSE, warning=FALSE, include=FALSE}
-table_nums<-captioner(prefix = "Table")
-plot_nums<-captioner(prefix = "Figure")
-diam_nums<-captioner(prefix = "Diagram")
 anx_nums<-captioner(prefix = "Annexed")
+
 table_nums("ta", "Dimensions of meritocratic beliefs.")
@@
 table_nums("te", "Comparison between model fit.")
-diam_nums("da", "Survey flow")
-diam_nums("db", "Theorical model")
-plot_nums("fa", "Descriptive boxplot of the scale.")
-plot_nums("fb", "Polychoric correlation.")
+
+fig_nums("fa", "Descriptive boxplot of the scale.")
+fig_nums("fb", "Polychoric correlation.")
+anx_nums("aa", "Representativeness of the sample.")
+
+anx_nums<-captioner(prefix = "Annexed")
 anx_nums("aa", "Representativeness of the sample.")
@@
 
-.
-.
-.
-
 The proposed measurement framework is based on two axis of analysis, as depicted in Table 1:
@@
 
-`r table_nums("tb")`
+```{r include=FALSE}
+cap <- "Items according to dimension."
+if (knitr::is_html_output()) {tb <- fig_nums("tb",cap)} else {tb <- cap}
+```
 
-![](../input/images/tabla2.png)
+![`r tb`](../input/images/tabla2.png)
 
@@
 
-`r diam_nums("da")`
+```{r include=FALSE}
+cap <- "Survey flow"
+if (knitr::is_html_output()) {diam1 <- fig_nums("diam1",cap)} else {diam1 <- cap}
+```
 
-![](../input/images/app_mod.png)
+![`r diam1`](../input/images/app_mod.png)
 
@@
 
-`r diam_nums("db")`
+```{r include=FALSE}
+cap <- "Theorical model"
+if (knitr::is_html_output()) {diam2 <- fig_nums("diam2",cap)} else {diam2 <- cap}
+```
 
-![](../output/images/meas01.png)
+![`r diam2`](../output/images/meas01.png)
 
@@
 ## Descriptive analysis
+
+```{r include=FALSE}
+cap <- "Descriptive statistics of the scale."
+if (knitr::is_html_output()) {tades <- table_nums("tades",cap)} else {tades <- cap}
+```
+
+`r tades`
 
@@
 ) %>% 
-          kable_styling(full_width = F,font_size = 8,position = "center");sumdat04
+          kable_styling(full_width = table_format2,font_size = 12,position = "center");sumdat04
 
@@
 
-4 XX producir tabla y llamar directamente desde la fuente o como imagen [__Pendiente__]
+```{r include=FALSE}
+cap <- "Descriptive plot"
+if (knitr::is_html_output()) {liker <- fig_nums("liker",cap)} else {liker <- cap}
+```
 
-`r plot_nums("fa")`
-
-![](../output/images/plotlikert.png)
+![`r liker`](../output/images/plotlikert.png)
 
@@
 
-`r plot_nums("fb")`
+`r fig_nums("fb")`
 
@@
 
-6 - XX Sobre tablas en general, entonces:
-- dejar tabla de descriptivos que ahora aparece junto a la de correlaciones [_Pendiente_ ]
-
 ## Models estimation
-
-7 XX En tabla de cargas factoriales, los fit aparecen corridos hacia la izquierda [__Pendiente__]
 
@@
 El primer orden de los ítems corresponde al orden de de la tabla 2, en el que primero se pregunta a los encuestados por todas las percepciones y después por todas las preferencias, en el segundo orden se les pregunta seguidamente por percepciones y preferencias respecto a un mismo tema, por ejemplo, al papel del esfuerzo, y en el tercero, los ítems aparecen de manera aleatoria.
+
+```{r include=FALSE}
+cap <- "Factor loads and model fit."
+if (knitr::is_html_output()) {loadfit <- table_nums("loadfit",cap)} else {loadfit <- cap}
+```
+
+`r loadfit`
 
@@
 tb.fm01<- kable(tb.fm, table_format, booktabs = T, escape = FALSE,align = "lcccccccccccc",col.names = tb.col, caption = tb.caption) %>% 
-          kable_styling(position = "center", full_width = F,font_size = 6) %>% 
+          kable_styling(position = "center", full_width = table_format2,font_size = 9) %>% 
           add_header_above(header = c(" "=1,"Version 1"= 4,"Version 2"= 4,"Version 3"= 4)) %>% 
@@
 tb.fm02<- kable(tb.fm2, table_format, booktabs = T, escape = FALSE,align = "lcccccccccccc", col.names = tb.col, caption = tb.caption) %>% 
-          kable_styling(position = "center", full_width = F,font_size = 8) %>% 
+          kable_styling(position = "center", full_width = table_format2,font_size = 11) %>% 
           add_header_above(header = c(" "=1,"Version 1"= 4,"Version 2"= 4,"Version 3"= 4)) %>% 
           add_header_above(header = c(" "=1,"Factor loadings"= 12)) %>% 
-          row_spec(row = 8,hline_after = F)
+          row_spec(row = 8,hline_after = )
 
+tb.fm01
 tb.fm02
-tb.fm01
 ```
@@
 Por su parte, la escala ordenada por temáticas parece generar un efecto framing en el cual la relación entre las percepciones y las preferencias parecen sobreestimadas, afectando de esta manera el ajuste (CFI=0.984, TLI=0.968,RMSEA=.071, x2(df=14)=64.156, p < 0.001).
-
-8 XX explicar mejor lo que viene: invarianza de qué y por qué se realiza.
 
@@
 `r anx_nums("aa")`
-
 |                     | Sample  |   CEP  |

### 2020-05-26 production/prod_analysis-cfa.rmd
Render final
Se mejoran las tablas, los titulos y l pdf.
ademas se agrega la opcion pdf dentro de las pestañas del _site.yml

@@
 fm04.1 <- fm04 %>% select(chisq.scaled,df) 
-fm04.1$chisq.scaled <- round(x = fm04.1$chisq.scaled,digits = 2)
+fm04.1$chisq.scaled <- round(x = fm04.1$chisq.scaled,digits = 1)
 fm04.1$df <- round(x = fm04.1$df,digits = 0)

### 2020-05-28 docs/paper.Rmd
Todo listo
- tablas y titulos en ambos formatos.
- graficos llamados como imagenes

@@
 geometry: margin=0.78in
+header-includes:
+  # - \usepackage[spanish,es-tabla,es-nodecimaldot]{babel}
+  - \usepackage{caption}
+  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
+  - \usepackage{graphicx}
+  - \usepackage{float}
+  - \usepackage{booktabs}
+  - \usepackage{longtable}
+  - \usepackage{array}
+  - \usepackage{multirow}
+  - \usepackage{wrapfig}
+  - \usepackage{float}
+  - \usepackage{colortbl}
+  - \usepackage{pdflscape}
+  - \usepackage{tabu}
+  - \usepackage{threeparttable}
 ---
@@
 
-table_nums("ta", "Dimensions of meritocratic beliefs.")
-table_nums("tb", "Items according to dimension.")
-table_nums("tc", "Descriptive statistics of the scale.")
-table_nums("td", "Factor loads and model fit.")
-table_nums("te", "Comparison between model fit.")
-
-fig_nums("fa", "Descriptive boxplot of the scale.")
-fig_nums("fb", "Polychoric correlation.")
-anx_nums("aa", "Representativeness of the sample.")
-
-anx_nums<-captioner(prefix = "Annexed")
-anx_nums("aa", "Representativeness of the sample.")
 ```
@@
 
-`r table_nums("ta")`
+```{r include=FALSE}
+cap <- "Dimensions of meritocratic beliefs."
+if (knitr::is_html_output()) {ta <- table_nums("ta",cap)} else {ta <- cap}
+```
 
@@
 | _Non-meritocracy_ 	|             	|             	|
+Table:`r ta`
 
@@
 cap <- "Items according to dimension."
-if (knitr::is_html_output()) {tb <- fig_nums("tb",cap)} else {tb <- cap}
+if (knitr::is_html_output()) {tb <- table_nums("tb",cap)} else {tb <- cap}
 ```
 
-![`r tb`](../input/images/tabla2.png)
+```{r echo=FALSE}
+ tab01 <- read.csv(file = "input/data/original/materials/table01.csv",header = 1,sep = ";")
+
+ cnames <- c("Dimension", "Factor","Statement (english)" , "Statement (spanish)")
+
+ kable(tab01,format = table_format,booktabs = T, linesep = "",col.names = cnames, caption = tb) %>%
+   kable_styling(full_width = T,latex_options = c("hold_position","scale_down"),position = "center",font_size = 9) %>%
+   column_spec(column = 1,width = "1.5cm",) %>%
+   column_spec(column = 2,width = "2 cm") %>%
+   collapse_rows(columns = 1:2,valign = "middle")
+
+```
 
@@
 
-`r tades`
-
 ```{r echo=FALSE}
@@
 
-dat04b<- dat04 %>% rename("Perception Effort" =perc_effort,
-                          "Perception Talent" = perc_talent,
-                          "Perception rich parents" = perc_wpart,
-                         "Perception contacts" = perc_netw,
-                          "Preferences Effort" = pref_effort,
-                          "Preferences Talent" =  pref_talent, 
-                         "Preferences rich parents" =  pref_wpart ,
-                          "Preferences contacts" = pref_netw  )
+dat04b<- dat04 %>% rename("A.Perception Effort" =perc_effort,
+                          "B.Perception Talent" = perc_talent,
+                          "c.Perception rich parents" = perc_wpart,
+                         "D.Perception contacts" = perc_netw,
+                          "E.Preferences Effort" = pref_effort,
+                          "F.Preferences Talent" =  pref_talent, 
+                         "G.Preferences rich parents" =  pref_wpart ,
+                          "H.Preferences contacts" = pref_netw  )
 
@@
 
-cap01 <- "Descriptive statistic"
-sumdat04<- kable(table04b, table_format, booktabs= T, escape = FALSE,align = "l",linesep = "", caption = cap01
+sumdat04<- kable(table04b, table_format, booktabs= T, escape = FALSE,align = "l",linesep = "", caption = tades
 ) %>% 
@@
 
-`r fig_nums("fb")`
+```{r include=FALSE}
+cap <- "Descriptive plot"
+if (knitr::is_html_output()) {fb <- fig_nums("fb",cap)} else {fb <- cap}
+```
 
-```{r echo=FALSE}
-load(file = "output/images/correlation.RData")
-library(corrplot)
-corrplot(cor04$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
-```
+![`r fb`](../output/images/Rplot.png)
 
@@
 ```
-
-`r loadfit`
 
@@
 
-tb.fm01<- kable(tb.fm, table_format, booktabs = T, escape = FALSE,align = "lcccccccccccc",col.names = tb.col, caption = tb.caption) %>% 
+tb.fm01<- kable(tb.fm, table_format, booktabs = T,linesep = "", escape = FALSE,align = "lcccccccccccc",col.names = tb.col, caption = tb.caption) %>% 
           kable_styling(position = "center", full_width = table_format2,font_size = 9) %>% 
@@
 
-tb.fm2<-as.data.frame(tb.fm[1:8,])
+tb.fm01
 
-tb.fm02<- kable(tb.fm2, table_format, booktabs = T, escape = FALSE,align = "lcccccccccccc", col.names = tb.col, caption = tb.caption) %>% 
-          kable_styling(position = "center", full_width = table_format2,font_size = 11) %>% 
-          add_header_above(header = c(" "=1,"Version 1"= 4,"Version 2"= 4,"Version 3"= 4)) %>% 
-          add_header_above(header = c(" "=1,"Factor loadings"= 12)) %>% 
-          row_spec(row = 8,hline_after = )
-
-tb.fm01
-tb.fm02
 ```
@@
 con CFI superiores a .95 y RMSEA inferiores a .8, por contraparte ni un modelo logró un chi-square no significativo, aunque tanto el modelo aleatorio como el primero obtuvieron un adecuado chi-square ratio menor a 3.
-El primero de los órdenes fue el que obtuvo mejor ajuste (CFI=0.998, TLI= 0.995,RMSEA=0.037, x2(df=14)=28,03, p = 0.014), seguido por el orden aleatorio de los ítems (CFI=0.992, TLI=0.984,RMSEA=0.051, x2(df=14)=39.09, p < 0.001).
-Por su parte, la escala ordenada por temáticas parece generar un efecto framing en el cual la relación entre las percepciones y las preferencias parecen sobreestimadas, afectando de esta manera el ajuste (CFI=0.984, TLI=0.968,RMSEA=.071, x2(df=14)=64.156, p < 0.001).
+El primero de los órdenes fue el que obtuvo mejor ajuste (CFI=0.998, TLI= 0.995,RMSEA=0.037, $\chi2$(df=14)=28,03, p = 0.014), seguido por el orden aleatorio de los ítems (CFI=0.992, TLI=0.984,RMSEA=0.051, $\chi2$(df=14)=39.09, p < 0.001).
+Por su parte, la escala ordenada por temáticas parece generar un efecto framing en el cual la relación entre las percepciones y las preferencias parecen sobreestimadas, afectando de esta manera el ajuste (CFI=0.984, TLI=0.968,RMSEA=.071, $\chi2$(df=14)=64.156, p < 0.001).
 
@@
 
-`r table_nums("te")`
+```{r include=FALSE}
+cap <- "Items according to dimension."
+if (knitr::is_html_output()) {te <- table_nums("te",cap)} else {te <- cap}
+```
 
@@
 | **RMSEA** |          0.226          |            0.047            |      0.036      |
-| **x2**    |         1830.839        |            68.661           |      40.250     |
+| **$\chi2$**    |         1830.839        |            68.661           |      40.250     |
 | **p**     |           .000          |             .000            |       .000      |
-| **x2/gl** |          65.38          |             4.90            |       3.32      |
+| **$\chi2$/df** |          65.38          |             4.90            |       3.32      |
+Table:`r te`
 
 EL modelo teórico propuesto de cuatro factores ajustó, como se observa en la tabla 4, mejor que el modelo de contraste de 1 factor.
-El modelo teórico ajustó de manera relativamente adecuada, pues muestra indicadores óptimos para CFI= 0.987, TLI = 0.975 y RMSEA=.041, aunque posee indicadores deficientes para la prueba x2(df=14)=67.6, p-value=.000.
+El modelo teórico ajustó de manera relativamente adecuada, pues muestra indicadores óptimos para CFI= 0.987, TLI = 0.975 y RMSEA=.041, aunque posee indicadores deficientes para la prueba $\chi2$(df=14)=67.6, p-value=.000.
 Para evaluar posibles mejoras de la escala, se analizaron las relaciones propuestas por los índices de modificación.
 Estos indican la existencia de dos cargas cruzadas no especificadas.
-Cuando se generó un modelo siguiendo estas recomendaciones, hubo una mejora considerable del modelo, aunque el nuevo modelo tampoco obtuvo un X2 ratio menor a 3 y obtuvo cargas factoriales muy bajas (cargas)< 0.15, por lo tanto, siguiendo las recomendación de Brown (2006) de solo aceptar las propuesta de los índices de modificación cuando se posee teoria y evidencia sólida, se ha decidido no incorporar estos parámetros al modelo.
-
-* agregar cuáles fueron las cargas sugeridas.
+Cuando se generó un modelo siguiendo estas recomendaciones, hubo una mejora considerable del modelo, aunque el nuevo modelo tampoco obtuvo un $\chi2$ ratio menor a 3 y obtuvo cargas factoriales muy bajas ($\tau$)< 0.15, por lo tanto, siguiendo las recomendación de Brown (2006) de solo aceptar las propuesta de los índices de modificación cuando se posee teoria y evidencia sólida, se ha decidido no incorporar estos parámetros al modelo.
 
@@
 
-`r anx_nums("aa")`
+```{r include=FALSE}
+cap <- "Representativeness of the sample."
+if (knitr::is_html_output()) {aax <- table_nums("aax",cap)} else {aax <- cap}
+```
+
 |                     | Sample  |   CEP  |
@@
 | university or more  | 21,21%  | 18,13% |
+Table:`r aax`
 

### 2020-05-28 render-site.R

@@ -19,7 +19,7 @@ file.copy(file.path("production",production), "docs",overwrite = TRUE)# copy dat
 
 rmarkdown::render_site("docs",quiet = TRUE) # Render site
 
-# before you run this line, check if your R Markdown files have a .rmd or .Rmd extension
+# before you run this line, check if your R Markdown files have a .rmd or .Rmd extension 
 # on this case we use both
 ext <- c(grep("^prod_.*\\.rmd$",  x = dir(path = "docs"),value = T), # for .rmd
          grep("^prod_.*\\.Rmd$",  x = dir(path = "docs"),value = T), # for .Rmd

### 2020-05-29 docs/paper.Rmd
Renderizado

@@
 title: "Perceptions and Preferences for Meritocracy"
-author:
-- Juan Carlos Castillo 
-- Julio Iturra 
-- Francisco Meneses 
 css: "../input/css/custom.css"
@@
 
+\pagebreak
+
 Para poder dar cuenta de los niveles de creencia en la meritocracia los estudios a la fecha generalmente han utilizado algunos indicadores de encuestas ya existentes, y en el menor de los casos se han creado instrumentos ad-hoc.
@@
 Nevertheless, despite making this distinction the author ends up substracting one dimension from the other, assuming that they are two poles of the same continuum as @kunovich_systems_2007 did.
+
+\pagebreak
 
@@
 
+\pagebreak
+
 ## Instrument design
@@
 
- kable(tab01,format = table_format,booktabs = T, linesep = "",col.names = cnames, caption = tb) %>%
+ kable(tab01,table_format,booktabs = T, linesep = "",col.names = cnames, caption = tb) %>%
    kable_styling(full_width = T,latex_options = c("hold_position","scale_down"),position = "center",font_size = 9) %>%
@@
 
+\pagebreak
+
 ## Administration sets
 
-With the objective of evaluating the effect of the order of the indicators, the respondents (_n = 2141_) were randomly divided into three groups as explained in __Diagram 1__.
+With the objective of evaluating the effect of the order of the indicators, the respondents (_n = 2141_) were randomly divided into three groups as explained in __Figure 1__.
 The scale was presented to the first group (_n = 712_) in the order that appears in Table 2, in such a way that the items of the same factor are found together.
@@
 
-![`r diam1`](../input/images/app_mod.png)
+![`r diam1`](../input/images/app_mod.png){ width=80% }
+
+\pagebreak
 
@@
 
+Como puede observarse en la tabla 3, los indicadores poseen valores que van desde el 1 al 5.
+Se observan promedios desde 2,41 correspondiente a la preferencia de distribucion según contactos, hasta 3,89, correspondiente a la preferencia de distribucion según esfuerzo.
+Ambos indicadores son coherentes con la adhesión  a las creencias meritocraticas, privilegiando la distribucion meritocratica por sobre aquella que se debe a ventajas externas.
+
 ```{r include=FALSE}
@@
 ```{r echo=FALSE}
-
 load(file = "input/data/proc/dat04.RData")
-
 dat04b<- dat04 %>% rename("A.Perception Effort" =perc_effort,
                           "B.Perception Talent" = perc_talent,
-                          "c.Perception rich parents" = perc_wpart,
+                          "C.Perception rich parents" = perc_wpart,
                          "D.Perception contacts" = perc_netw,
@@
                           "H.Preferences contacts" = pref_netw  )
+sum04b  <- psych::describe(dat04b[,-c(1)], fast = T) %>% select(mean,sd,min,max) %>% data.frame() %>% round(digits = 2)
+table04b<- sum04b %>% rename("Mean"=mean, "SD"=sd, "Min"=min,"Max"=max)
+sumdat04<- kable(table04b,format =table_format, booktabs= T, escape = FALSE,align = "l",linesep = "", caption = tades
+) %>% kable_styling(full_width = table_format2, font_size = 9,position = "center");sumdat04
+```
 
-sum04b  <- psych::describe(dat04b[,-c(1)], fast = T) %>% select(mean,sd,min,max) %>% data.frame() %>% round(digits = 2)
+\pagebreak
 
-table04b<- sum04b %>% rename("Mean"=mean, "SD"=sd, "Min"=min,"Max"=max)
-
-sumdat04<- kable(table04b, table_format, booktabs= T, escape = FALSE,align = "l",linesep = "", caption = tades
-) %>% 
-          kable_styling(full_width = table_format2,font_size = 12,position = "center");sumdat04
-
-```
+A partir del grafico liker  se observa que, en general, hay una mayor percepción de aspectos no meritocráticos que meritocráticos, mientras que en el caso de preferencias ocurre lo opuesto.
+En cuanto a las preferencias, llama la atención el rol preponderante del esfuerzo por sobre el talento.
 
@@
 
-![`r liker`](../output/images/plotlikert.png)
+<center>
 
-En general se observa que hay una mayor percepción de aspectos no meritocráticos que meritocráticos, mientras que en el caso de preferencias ocurre lo opuesto.
-En cuanto a las preferencias, llama la atención el rol preponderante del esfuerzo por sobre el talento.
+![`r liker`](../output/images/plotlikert.png){ width=70% }
+
+</center>
+
+\pagebreak
+
+El grafico de correlaciones policoricas evidencia relaciones de moderada a alta intensidad entre los indicadores que corresponden al mismo factor, mientras que se observan correlaciones de baja intensidad en el resto de los cruces.
+Destacan además las relaciones entre las percepciones y preferencias de tipo meritocraticas, lo que no ocurre con los indicadores no meritocraticos.
 
 ```{r include=FALSE}
-cap <- "Descriptive plot"
+cap <- "polychoric correlation plot"
 if (knitr::is_html_output()) {fb <- fig_nums("fb",cap)} else {fb <- cap}
@@
 
-![`r fb`](../output/images/Rplot.png)
+<center>
+
+![`r fb`](../output/images/corpoly.png){ width=70% }
+</center>
+
+En suma, los análisis descriptivos señalan una relativa adhesión a la moral meritocratica, las cuales, a modo general, se expresan en una mayor preferencia de distribucion según criterios meritocraticos y una menor preferencia de criterios no meritocraticos.
+Igualmente, es observable una relativamente baja percepcion de distribucion meritocratica.
+Además se observa una relación coherente entre los indicadores según lo propuesto en el modelo teórico en el preregistro.
+
+\pagebreak
 
@@
 tb.caption <- "Factor loadings and fit measures" 
-tb.foot <- paste0("; p<0.05=$*$; p<0.01=$**$; p<0.001=$***$.
-Standardised factor loadings using DWLS estimator.
-CFI = Comparative fit index (scaled), RMSEA = Root mean square error of approximation (scaled)")
+tb.foot <- paste0("Standardised factor loadings using DWLS estimator")
 
@@
 
-tb.fm01<- kable(tb.fm, table_format, booktabs = T,linesep = "", escape = FALSE,align = "lcccccccccccc",col.names = tb.col, caption = tb.caption) %>% 
-          kable_styling(position = "center", full_width = table_format2,font_size = 9) %>% 
+tb.fm01<- kable(tb.fm, format = table_format, booktabs = T,linesep = "", escape = FALSE,align = "lcccccccccccc",col.names = tb.col, caption = loadfit) %>% 
+          kable_styling(full_width = table_format2,latex_options = c("hold_position","scale_down"), position = "center",font_size = 9) %>% 
           add_header_above(header = c(" "=1,"Version 1"= 4,"Version 2"= 4,"Version 3"= 4)) %>% 
@@
           row_spec(row = 8,hline_after = TRUE) %>%
-          add_indent(c(9:12))
-
-tb.fm01
+          add_indent(c(9:12))  %>%
+          footnote(general = tb.foot, number = c("CFI = Comparative fit index (scaled); ", " RMSEA = Root mean square error of approximation (scaled); "));tb.fm01
 
@@
 De modo coherente, el modelo aleatorio presentó un ajuste intermedio entre el orden 1 y el orden 2.
+
+\pagebreak
 
@@
 
-# Conclusiones
+uso conjunto percepciones-preferencias
 
-uso conjunto percepciones-preferencias
+\pagebreak
 

### 2020-05-29 production/prod_analysis-cfa.rmd
Renderizado

@@
 
-```{r}
-table_format = if(is_html_output()) {
-  "html"
-} else if(is_latex_output()) {
-  "latex"
-}
-```
-
 ```{r eval=FALSE, include=FALSE}
@@
 load(file = "input/data/proc/dat04.RData")
+```
+
+```{r}
+table_format = if(is_html_output()) {
+  "html"
+} else if(is_latex_output()) {
+  "latex"
+}
+
 ```
@@
 cor01<-polychoric(dat01)
-rownames(cor01$rho) <- c("A. pc.effort", "B. pc.talent","C. pc.wpart","D. pc.netw","E. pf.effort","F. pf.talent","G. pf.wpart", "H. pf.netw")
+rownames(cor01$rho) <- c("A.Perception Effort", "B.Perception Talent","C.Perception Rich parents","D.Perception contacts","E.Preferences Effort","F.Preferences Talent","G.Preferences rich parents", "H.Preferences contacts" )
 colnames(cor01$rho) <- c("(A)","(B)","(C)","(D)","(E)","(F)","(G)", "(H)")
+
 corrplot(cor01$rho,method = "color" ,type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
+
 ```
@@
 cor04<-polychoric(select(dat04,perc_effort,perc_talent,perc_wpart,perc_netw,pref_effort,pref_talent,pref_wpart,pref_netw))
-rownames(cor04$rho) <-c("A. pc.effort", "B. pc.talent","C. pc.wpart","D. pc.netw","E. pf.effort","F. pf.talent","G. pf.wpart", "H. pf.netw")
+rownames(cor04$rho) <-c("A.Perception Effort", "B.Perception Talent","C.Perception Rich parents","D.Perception contacts","E.Preferences Effort","F.Preferences Talent","G.Preferences rich parents", "H.Preferences contacts" )
 colnames(cor04$rho) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)", "(H)")
@@
 corrplot(cor04$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
+
+png("output/images/corpoly.png",
+    width = 15,height = 15,units="cm",
+    pointsize=12,bg="white",res=300)
+
+corrplot(cor01$rho,method = "color" ,type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
+
+dev.off()
 ```
@@
             geom.colors ="PuBu", 
+            geom.size = 0.8,
             axis.labels=c("Effort","Talent","Rich Parents","Contacts"),
@@
 
-ggsave(plotlikert,filename = "output/images/plotlikert.png",device = "png",width = 30,height = 15,dpi = "retina",units = "cm")
-
-```
+ggsave(plotlikert,filename = "output/images/plotlikert.png",device = "png",width = 18,height = 15,dpi = "retina",units = "cm")
+
+```

### 2020-05-29 production/prod_prep-data-A.rmd
Renderizado

@@
 ```{r}
-library(datasets)
-library(data.table)
-wide <- setDT(wide02)
-long01<- melt(wide,variable.name = "wave",
-              measure = patterns("^perc_effort_","^perc_talent_","^perc_wpart_","^perc_netw_",
-                                       "^pref_effort_","^pref_talent_","^pref_wpart_","^pref_netw_",
-                                       "^sexo_","^edad_","^edcep_"),
-              value.name = c("perc_effort","perc_talent","perc_wpart","perc_netw",
-                             "pref_effort","pref_talent","pref_wpart","pref_netw",
-                             "sexo","edad","edcep"))
+#library(datasets)
+#library(data.table)
+#wide <- setDT(wide02)
+#long01<- melt(wide,variable.name = "wave",
+#              measure = patterns("^perc_effort_","^perc_talent_","^perc_wpart_","^perc_netw_",
+#                                       "^pref_effort_","^pref_talent_","^pref_wpart_","^pref_netw_",
+#                                       "^sexo_","^edad_","^edcep_"),
+#              value.name = c("perc_effort","perc_talent","perc_wpart","perc_netw",
+#                             "pref_effort","pref_talent","pref_wpart","pref_netw",
+#                             "sexo","edad","edcep"))
 ```
@@
 ```{r}
-save(long01,file = "input/data/proc/datalong.RData")
+#save(long01,file = "input/data/proc/datalong.RData")
 ```

### 2020-05-29 render-site.R

@@ -12,7 +12,7 @@
 
 # INSTRUCTIONS-------------------------:
 # 1. set your root dir at "merit-scale"
-# 2. Run this code.
+# 2. Run this code.4
 
 production<- list.files(path = "production",pattern = "prod_")        # a list with the names of the files to copy
 file.copy(file.path("production",production), "docs",overwrite = TRUE)# copy data proc and analysis files
@@ -29,7 +29,7 @@ ext <- c(grep("^prod_.*\\.rmd$",  x = dir(path = "docs"),value = T), # for .rmd
          ".tex",".log",".RData",".Rhistory")                         # for .tex .log .RData .Rhistory
 
 for (i in 1:length(ext)) {
-  file.remove(paste0("docs/",dir(path="docs", pattern=ext[i]))) # delete files from /docs folder
+  file.remove(paste0("docs/",dir(path="docs", pattern=ext[i]))) # delete files from /docs folder4
   unlink(paste0("docs/",(ext[i])),recursive = TRUE)             # Remove the cache files (first)
 }
 

### 2020-06-01 production/prod_analysis-cfa.rmd
Todo listo
Ya estan listas las tablas con los titulos y los demas detalles de formato

@@
 
-nodeNames <-c("Perception effort","Perception talent","Perception rich family","Perception networks",
-              "Preference effort","Preference talent","Preference rich family","Preference networks",
+nodeNames <-c("Perception effort","Perception talent","Perception rich family","Perception contacts",
+              "Preference effort","Preference talent","Preference rich family","Preference contacts",
               "Perception meritocratic","Perception non meritocratic","Preference meritocratic","Preference non meritocratic")

### 2020-06-10 production/prod_analysis-cfa.rmd

--- 2020-05-26 production/prod_analysis-cfa.rmd
@@
 library(rvest)
+library(sjlabelled)
 source('production/prod_polychoric.R')
@@
 colnames(cor04$rho) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)", "(H)")
-save(cor04, file = "output/images/correlation.RData")
 corrplot(cor04$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
@@
 fm04.1 <- fm04 %>% select(chisq.scaled,df) 
-fm04.1$chisq.scaled <- round(x = fm04.1$chisq.scaled,digits = 1)
+fm04.1$chisq.scaled <- round(x = fm04.1$chisq.scaled,digits = 2)
 fm04.1$df <- round(x = fm04.1$df,digits = 0)
@@
 fm.df$nobs <- c(nobs(fit1_o),nobs(fit2_o),nobs(fit3_o)) 
-fm.df <- data.frame(t(fm.df)); colnames(fm.df) <- c("M1.2","M2.2","M3.2")
-
+fm.df <- data.frame(t(fm.df)); colnames(fm.df) <- c("M1.1","M2.1","M3.1")
 #------ merge ------#
@@
 save(tb.fm01,file = "output/tables/tb.fm01.RData")
-
-save(tb.fm,file = "output/tables/tb.fm.RData")
-save(fm.df,file = "output/tables/fm.df.RData")
-save(tb.load,file = "output/tables/tb.load.RData")
-
 ```
@@
 dir.create("output/images")
-
 semPaths(plot01,
@@
          filename = "output/images/meas01")
+
 ```
@@
 
-```{r}
-library(ggplot2)
-library(reshape2)
-library(sjPlot)
-library(dplyr)
-
-#install.packages("cowplot")
-library(cowplot)
-
-dat_merit <- dat04 %>% select("a.Effort"=perc_effort,"a.Talent"=perc_talent, "a.Rich Parents"=perc_wpart, "a.Contacts"=perc_netw, "b.Effort"=pref_effort,"b.Talent"=pref_talent,"b.Rich Parents"=pref_wpart, "b.Contacts"=pref_netw)
-
-plotlikert<-plot_likert(dat_merit, c(1,1,1,1,2,2,2,2),
-            groups.titles = c("Perceptions","Preferences"),
-            geom.colors ="PuBu", 
-            axis.labels=c("Effort","Talent","Rich Parents","Contacts"),
-            catcount = 4,
-            cat.neutral = 3, 
-            grid.range  =  c ( 1.2 , 1.4 ),
-            values  =  "sum.outside",
-            reverse.scale=TRUE)
-plotlikert
-
-ggsave(plotlikert,filename = "output/images/plotlikert.png",device = "png",width = 30,height = 15,dpi = "retina",units = "cm")
-
-```
+# Anexo 1: Sample
+
+```{r}
+load(file = "input/data/original/issp2019CL.RData")
+w01 <- sjlabelled::read_spss("input/data/original/Estudio_3_ola1.sav",verbose = F)
+```
+
+```{r}
+GS<- c("CERRILLOS",
+"CERRO NAVIA",
+"CONCHALÍ",
+"EL BOSQUE",
+"ESTACIÓN CENTRAL",
+"HUECHURABA",
+"INDEPENDENCIA",
+"LA CISTERNA",
+"LA FLORIDA",
+"LA GRANJA",
+"LA PINTANA",
+"LA REINA",
+"LAS CONDES",
+"LO BARNECHEA",
+"LO ESPEJO",
+"LO PRADO",
+"MACUL",
+"MAIPÚ",
+"ÑUÑOA",
+"PEDRO AGUIRRE CERDA",
+"PEÑALOLEN",
+"PROVIDENCIA",
+"PUDAHUEL",
+"PUENTE ALTO",
+"QUILICURA",
+"QUINTA NORMAL",
+"RECOLETA",
+"RENCA",
+"SAN BERNARDO",
+"SAN JOAQUÍN",
+"SAN MIGUEL",
+"SAN RAMÓN",
+"SANTIAGO",
+"VITACURA")
+
+cl19$COMUNA <- toupper(as_character(cl19$COMUNA))
+
+# cl19$DS_P1 # 1= hombre, 2= mujer
+# cl19$DS_P2_EXACTA  
+# cl19$DS_P4
+
+cep1 <- cl19 %>% 
+  select(COMUNA,DS_P1,DS_P2_EXACTA,DS_P4, FACTOR) %>% 
+  filter(COMUNA %in% GS)
+```
+
+```{r}
+cep1$edad_cat<- as.numeric(cep1$DS_P2_EXACTA) # Edad
+cep1$edad_cat[cep1$edad_cat %in%c(18:24)] <- 1   #"18 a 24"
+cep1$edad_cat[cep1$edad_cat %in%c(25:34)] <- 2 # "25 a 34"
+cep1$edad_cat[cep1$edad_cat %in%c(35:44)] <- 3 # "35 a 44"
+cep1$edad_cat[cep1$edad_cat %in%c(45:54)] <- 4 # "45 a 54"
+cep1$edad_cat[cep1$edad_cat >=  55]       <- 5 # "55 o más"
+
+cep1$educ <- cep1$DS_P4 # Educación
+cep1$educ <- as_numeric(cep1$educ)
+cep1$educ <- as.numeric(cep1$educ)
+cep1$educ[cep1$educ %in%c(88,99)] <- NA
+
+cep1$educ <- recode(cep1$educ, "0" = "1" , "1"="1", "2" = "1", "3"="2","4"="2", "5"="3","6"="3", "7" = "3", "8"="4" , "9"="4"   )
+
+cep1$sexo <- as.numeric(cep1$DS_P1) # 1= hombre, 2= mujer
+
+```
+
+```{r}
+
+# Gender
+
+cep1$Hombres<-ifelse(cep1$sexo==1,100,0)
+cep1$Mujeres<-ifelse(cep1$sexo==2,100,0)
+w01$Hombres<-ifelse(w01$sexo==1,100,0)
+w01$Mujeres<-ifelse(w01$sexo==2,100,0)
+
+cep1$ed_a<-ifelse(cep1$edad_cat==1,100,0)
+cep1$ed_b<-ifelse(cep1$edad_cat==2,100,0)
+cep1$ed_c<-ifelse(cep1$edad_cat==3,100,0)
+cep1$ed_d<-ifelse(cep1$edad_cat==4,100,0)
+cep1$ed_e<-ifelse(cep1$edad_cat==5,100,0)
+w01$ed_a<-ifelse(w01$edad==1,100,0)
+w01$ed_b<-ifelse(w01$edad==2,100,0)
+w01$ed_c<-ifelse(w01$edad==3,100,0)
+w01$ed_d<-ifelse(w01$edad==4,100,0)
+w01$ed_e<-ifelse(w01$edad==5,100,0)
+
+cep1$educ_a<-ifelse(cep1$educ==1,100,0)
+cep1$educ_b<-ifelse(cep1$educ==2,100,0)
+cep1$educ_c<-ifelse(cep1$educ==3,100,0)
+cep1$educ_d<-ifelse(cep1$educ==4,100,0)
+w01$educ_a<-ifelse(w01$educat==1,100,0)
+w01$educ_b<-ifelse(w01$educat==2,100,0)
+w01$educ_c<-ifelse(w01$educat==3,100,0)
+w01$educ_d<-ifelse(w01$educat==4,100,0)
+
+Men<-round(c(mean(w01$Hombres, na.rm = T), weighted.mean(cep1$Hombres,cep1$FACTOR, na.rm = T)),digits = 2)
+Women<-round(c(mean(w01$Mujeres, na.rm = T), weighted.mean(cep1$Mujeres,cep1$FACTOR, na.rm = T)),digits = 2)
+
+years.18.to.24<-round(c(mean(w01$ed_a, na.rm = T), weighted.mean(cep1$ed_a,cep1$FACTOR, na.rm = T)),digits = 2)
+years.25.to.34<-round(c(mean(w01$ed_b, na.rm = T), weighted.mean(cep1$ed_b,cep1$FACTOR, na.rm = T)),digits = 2)
+years.35.to.44<-round(c(mean(w01$ed_c, na.rm = T), weighted.mean(cep1$ed_c,cep1$FACTOR, na.rm = T)),digits = 2)
+years.45.to.54<-round(c(mean(w01$ed_d, na.rm = T), weighted.mean(cep1$ed_d,cep1$FACTOR, na.rm = T)),digits = 2)
+years.55.or.more<-round(c(mean(w01$ed_e, na.rm = T), weighted.mean(cep1$ed_e,cep1$FACTOR, na.rm = T)),digits = 2)
+
+Primary <-round(c(mean(w01$educ_a, na.rm = T), weighted.mean(cep1$educ_a,cep1$FACTOR, na.rm = T)),digits = 2)
+Secondary<-round(c(mean(w01$educ_b, na.rm = T), weighted.mean(cep1$educ_b,cep1$FACTOR, na.rm = T)),digits = 2)
+Technique<-round(c(mean(w01$educ_c, na.rm = T), weighted.mean(cep1$educ_c,cep1$FACTOR, na.rm = T)),digits = 2)
+University<-round(c(mean(w01$educ_d, na.rm = T), weighted.mean(cep1$educ_d,cep1$FACTOR, na.rm = T)),digits = 2)
+
+sample_cep<- rbind("__Gender__",Men,Women,"__Age__",years.18.to.24,years.25.to.34,years.35.to.44,years.45.to.54,years.55.or.more, "__Education__", Primary, Secondary, Technique, University)
+
+cnames <- c("Sample", "CEP") 
+knitr::kable(sample_cep, digits = 0, format = "markdown",col.names = cnames)
+
+```
+
+# Anexo 2: Fiability
+
+```{r}
+load(file = "input/data/proc/dat01.RData")
+load(file = "input/data/proc/dat02.RData")
+load(file = "input/data/proc/dat03.RData")
+load(file = "input/data/proc/dat04.RData")
+```
+
+```{r}
+lav.mod.1 <- 'perc_merit =~ perc_effort+perc_talent
+perc_nmerit=~perc_wpart +perc_netw
+pref_merit =~ pref_effort+pref_talent
+pref_nmerit=~pref_wpart +pref_netw'
+
+cor01a<-polychoric(dat01)
+cor02a<-polychoric(dat02)
+cor03a<-polychoric(dat03)
+cor04a <-polychoric(select(dat04,perc_effort,perc_talent,perc_wpart,perc_netw,pref_effort,pref_talent,pref_wpart,pref_netw))
+
+#Omega for group
+lav.jen.1<- cfa(lav.mod.1,sample.cov=cor01a$rho,sample.nobs=500,orthogonal=TRUE,std.lv=TRUE) # Total sample
+lav.jen.2<- cfa(lav.mod.1,sample.cov=cor02a$rho,sample.nobs=500,orthogonal=TRUE,std.lv=TRUE) # Total sample
+lav.jen.3<- cfa(lav.mod.1,sample.cov=cor03a$rho,sample.nobs=500,orthogonal=TRUE,std.lv=TRUE) # Total sample
+lav.jen.4<- cfa(lav.mod.1,sample.cov=cor04a$rho,sample.nobs=500,orthogonal=TRUE,std.lv=TRUE) # Total sample
+
+o1 <- omegaFromSem(lav.jen.1,cor04a$rho, plot = F)
+o2<-omegaFromSem(lav.jen.2,cor04a$rho, plot = F)
+o3<-omegaFromSem(lav.jen.3,cor04a$rho, plot = F)
+o4<-omegaFromSem(lav.jen.4,cor04a$rho, plot = F)
+
+Group_1 <- c(o1$omega.group$total)
+Group_2 <- c(o2$omega.group$total)
+Group_3 <- c(o3$omega.group$total)
+Total_sample <- c(o4$omega.group$total)
+
+Omega<- rbind(Group_1,Group_2,Group_3,Total_sample)
+
+cnames <- c("Perception Meritocratic","Perception Non Meritocratic", "Preference Meritocratic","Preference Non Meritocratic")
+knitr::kable(Omega, format = "markdown",digits = 2, col.names = cnames)
+
+```

### 2020-06-10 production/prod_analysis-cfa.rmd

@@
 
+```{r}
+table_format = if(is_html_output()) {
+  "html"
+} else if(is_latex_output()) {
+  "latex"
+}
+
+```
+
 ```{r eval=FALSE, include=FALSE}
@@
 cor04<-polychoric(select(dat04,perc_effort,perc_talent,perc_wpart,perc_netw,pref_effort,pref_talent,pref_wpart,pref_netw))
-rownames(cor04$rho) <-c("A. pc.effort", "B. pc.talent","C. pc.wpart","D. pc.netw","E. pf.effort","F. pf.talent","G. pf.wpart", "H. pf.netw")
+rownames(cor04$rho) <-c("A.Perception Effort", "B.Perception Talent","C.Perception Rich parents","D.Perception contacts","E.Preferences Effort","F.Preferences Talent","G.Preferences rich parents", "H.Preferences contacts" )
 colnames(cor04$rho) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)", "(H)")
+save(cor04, file = "output/images/correlation.RData")
 corrplot(cor04$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
+
+png("output/images/corpoly.png",
+    width = 15,height = 15,units="cm",
+    pointsize=12,bg="white",res=300)
+
+corrplot(cor01$rho,method = "color" ,type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
+
+dev.off()
 ```
@@
 
+```{r}
+library(ggplot2)
+library(reshape2)
+library(sjPlot)
+library(dplyr)
+
+#install.packages("cowplot")
+library(cowplot)
+
+dat_merit <- dat04 %>% select("a.Effort"=perc_effort,"a.Talent"=perc_talent, "a.Rich Parents"=perc_wpart, "a.Contacts"=perc_netw, "b.Effort"=pref_effort,"b.Talent"=pref_talent,"b.Rich Parents"=pref_wpart, "b.Contacts"=pref_netw)
+
+plotlikert<-plot_likert(dat_merit, c(1,1,1,1,2,2,2,2),
+            groups.titles = c("Perceptions","Preferences"),
+            geom.colors ="PuBu", 
+            geom.size = 0.8,
+            axis.labels=c("Effort","Talent","Rich Parents","Contacts"),
+            catcount = 4,
+            cat.neutral = 3, 
+            grid.range  =  c ( 1.2 , 1.4 ),
+            values  =  "sum.outside",
+            reverse.scale=TRUE)
+plotlikert
+
+ggsave(plotlikert,filename = "output/images/plotlikert.png",device = "png",width = 18,height = 15,dpi = "retina",units = "cm")
+
+```
+
+=======
 # Anexo 1: Sample

### 2020-06-10 docs/paper.Rmd

@@
 #install.packages("estimatr")
-library(estimatr)
-library(texreg)
-
 ```
@@
 fig_nums   <- captioner(prefix = "Figure")
-```
-
-```{r message=FALSE, warning=FALSE, include=FALSE}
 anx_nums<-captioner(prefix = "Annexed")
-
 ```
@@
 ![`r fb`](../output/images/corpoly.png){ width=70% }
+
 </center>
@@
 tb.caption <- "Factor loadings and fit measures" 
-tb.foot <- paste0("Standardised factor loadings using DWLS estimator")
+tb.foot <- paste0("Standardised factor loadings using DWLS estimator ; CFI = Comparative fit index (scaled); RMSEA = Root mean square error of approximation (scaled)")
 
@@
           add_indent(c(9:12))  %>%
-          footnote(general = tb.foot, number = c("CFI = Comparative fit index (scaled); ", " RMSEA = Root mean square error of approximation (scaled); "));tb.fm01
+          footnote(general = tb.foot,threeparttable = T);tb.fm01
 

### 2020-06-10 input/preanalysis/preanalisis.Rmd

--- 2020-04-22 docs/preanalisis.Rmd
@@
 
-# tab01 <- read.csv(file = "input/data/original/materials/table01.csv",header = 1,sep = ";")
-#
-# cnames <- c("Dimension", "Factor","Statement (english)" , "Statement (spanish)")
-#
-# caption01 <- capTab("Meritocracy perception and preference")
-# kable(tab01,format = table_format,booktabs = T, linesep = "",col.names = cnames, caption = caption01) %>%
-#   kable_styling(full_width = T,latex_options = c("hold_position","scale_down"),position = "center",font_size = 9) %>%
-#   column_spec(column = 1,width = "1.5cm",) %>%
-#   column_spec(column = 2,width = "2 cm") %>%
-#   collapse_rows(columns = 1:2,valign = "middle")
+ #tab01 <- read.csv(file = "input/data/original/materials/table01.csv",header = 1,sep = ";")
+
+ cnames <- c("Dimension", "Factor","Statement (english)" , "Statement (spanish)")
+
+ caption01 <- capTab("Meritocracy perception and preference")
+ kable(tab01,format = table_format,booktabs = T, linesep = "",col.names = cnames, caption = caption01) %>%
+   kable_styling(full_width = T,latex_options = c("hold_position","scale_down"),position = "center",font_size = 9) %>%
+   column_spec(column = 1,width = "1.5cm",) %>%
+   column_spec(column = 2,width = "2 cm") %>%
+   collapse_rows(columns = 1:2,valign = "middle")
 

### 2020-06-10 docs/paper.Rmd
ajustes
- aumento tamañode las  letras en las  tablas

- cambio _config.yml para imagenes

- agrego bibliografia sugerida

@@
 435).
-In other words, preferences should be interpreted in the context of perception: a large preference for meritocracy certainly means something totally different for someone perceiving high meritocracy than for someone perceiving low meritocracy.
+In other words, preferences should be interpreted in the context of perception: a large preference for meritocracy certainly means something totally different for someone perceiving high meritocracy than for someone perceiving low meritocracy.Incorporating this conceptual difference could improve the accuracy of empirical analyzes that work with the concept of meritocratic beliefs in general [eg.
+@MeritocraticNarrativesCulturaOrientations].
 
@@
  kable(tab01,table_format,booktabs = T, linesep = "",col.names = cnames, caption = tb) %>%
-   kable_styling(full_width = T,latex_options = c("hold_position","scale_down"),position = "center",font_size = 9) %>%
+   kable_styling(full_width = T,latex_options = c("hold_position","scale_down"),position = "center",font_size = 12) %>%
    column_spec(column = 1,width = "1.5cm",) %>%
@@
 
-![`r diam1`](../input/images/app_mod.png){ width=80% }
+![`r diam1`](../input/images/app_mod.png){ width=90% }
 
@@
 sumdat04<- kable(table04b,format =table_format, booktabs= T, escape = FALSE,align = "l",linesep = "", caption = tades
-) %>% kable_styling(full_width = table_format2, font_size = 9,position = "center");sumdat04
+) %>% kable_styling(full_width = table_format2, font_size = 12,position = "center");sumdat04
 ```
@@
 tb.fm01<- kable(tb.fm, format = table_format, booktabs = T,linesep = "", escape = FALSE,align = "lcccccccccccc",col.names = tb.col, caption = loadfit) %>% 
-          kable_styling(full_width = table_format2,latex_options = c("hold_position","scale_down"), position = "center",font_size = 9) %>% 
+          kable_styling(full_width = table_format2,latex_options = c("hold_position","scale_down"), position = "center",font_size = 12) %>% 
           add_header_above(header = c(" "=1,"Version 1"= 4,"Version 2"= 4,"Version 3"= 4)) %>% 

### 2020-07-26 docs/paper.Rmd
update paper

--- 2020-06-10 docs/paper.Rmd
@@
 ---
-title: "Perceptions and Preferences for Meritocracy"
+title: "Measuring Perceptions and Preferences for Meritocracy"
 css: "../input/css/custom.css"
@@
 link-citations: yes
+author: "Juan Carlos Castillo, Julio Iturra & Francisco Meneses <br><br>"
+abstract: "Economic and social inequality have raised growing concerns and crises across societies.
+One of social science concepts associated to the maintenance of inequality is the belief in meritocracy, which would legitimize economic differences based on effort and talent.
+Despite its wide use,  empirical research on meritocracy is something relatively novel.
+A number of studies have relied mostly in secondary data to operationalize meritocracy, with a large variation in the use and interpretation of survey items.
+Starting from a review of studies that measure meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding the conceptualization and measurement of meritocracy.
+Based on this critical analysis, we propose  an items battery called *Perceptions and preferences for meritocracy scale*, which is tested with confirmatory factor analysis with data from an online survey study (N=2,141).
+The results support the proposed conceptual structure which not only distinguishes between perceptions and preferences, but also between meritocratic and non-meritocratic dimensions.
+The discussion highlights the relevance of considering these different dimensions in order to advance in the study of meritocracy."
 output:
-  html_document:
-    theme: flatly
-    number_sections: true
-    always_allow_html: yes
-  pdf_document:
-    keep_tex: false
-    template: null
-    number_sections: true
+  bookdown::html_document2
 linkcolor: blue
@@
 knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = FALSE)
-knitr::opts_knit$set(root.dir = "../")
+# knitr::opts_knit$set(root.dir = "../")
 options(scipen=999)
@@
 ```{r message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
-library(knitr)
-#install.packages("KableExtra")
-library(kableExtra)
-library(dplyr)
-#install.packages("captioner")
-library(captioner)
-#install.packages("estimatr")
+pacman::p_load(knitr,kableExtra, dplyr, captioner)
 ```
@@
 
-En general, la meritocracia se define como un sistema de distribución de recursos y recompensas basados en el mérito individual, que en su concepción original es una suma de talento y esfuerzo [@young_rise_1962] y que pone en un lugar secundario interferencia de factores estructurales como la herencia o los contactos [@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995; @yairMeritocracy2007; @landWeSatTable2006; @youngRiseMeritocracy1994].
+En general, la meritocracia se define como un sistema de distribución de recursos y recompensas basados en el mérito individual, que en su concepción original es una suma de talento y esfuerzo [@young_rise_1962].
+Esta concepción tradicional de mérito pone en un lugar secundario la posible interferencia de factores estructurales o no meritocráticos como la herencia, los contactos personales, o la suerte [@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995; @yairMeritocracy2007; @landWeSatTable2006; @youngRiseMeritocracy1994].
 Una serie de estudios han realizado críticas a la realización de este estándar moral de distribución, planteando que es una promesa incumplida dada la influencia preponderante de otros elementos más allá del mérito en el estatus individual [@WitteveenReconsideringmeritocraticpower2020a; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019].
@@
 Debido al rol que cumplen las creencias meritocráticas dentro del pensamiento neoliberal [@Bay-ChengTrackingHomoOeconomicus2015], han surgido múltiples investigaciones que evalúan la relación entre creencias meritocráticas y diversos ámbitos sociales de la actualidad.
-Se han desarrollado estudios que vinculan la meritocracia con el reforzamiento de estereotipos socioeconómicos, de género y de etnias [@MadeiraPrimesConsequencesSystematic2019; @GirerdNeoliberalismIdeologicalBarrier2020; @PremingerMeritocracyserviceethnocracy2020].
-Igualmente, se han desarrollado líneas de investigación que evalúan el efecto de las creencias meritocráticas en el contexto educativo [@GenerettStoriesWeTell2020; @ReayPerilsPenaltiesMeritocracy2020; @OwensENGINESSOCIALMOBILITY2020].
-y en el contexto organizacional de las empresas [@PerezAdvancingcareersmerit2020; @AielloNewevidencefirmuniversity2019]
-
-\pagebreak
+Por ejemplo, se han desarrollado estudios que vinculan la meritocracia al reforzamiento de estereotipos socioeconómicos, de género y de etnias [@MadeiraPrimesConsequencesSystematic2019; @GirerdNeoliberalismIdeologicalBarrier2020; @PremingerMeritocracyserviceethnocracy2020], así como también líneas de investigación que evalúan el efecto de las creencias meritocráticas en el contexto educativo [@GenerettStoriesWeTell2020; @ReayPerilsPenaltiesMeritocracy2020; @OwensENGINESSOCIALMOBILITY2020] y en el contexto organizacional de las empresas [@PerezAdvancingcareersmerit2020; @AielloNewevidencefirmuniversity2019].
 
@@
 Sin embargo, y como mostraremos más adelante, las formas de medición de meritocracia varían extremadamente entre estudios.
-Muchas veces fenómenos similares se asocian a indicadores distintos, y también ocurre que fenómenos distintos son medidos con indicadores similares, todo lo cual hace dificulta la comparabilidad entre estudios y el poder avanzar en la comprensión del concepto de meritocracia.
+Muchas veces fenómenos similares se asocian a indicadores distintos, y también ocurre que fenómenos distintos son medidos con indicadores similares, todo lo cual hace dificulta la comparabilidad entre estudios y el poder avanzar en la comprensión y estudio de la meritocracia.
 
-Basados en el análisis crítico de las formas de medición de meritocracia a la fecha, el presente estudio propone un instrumento para medir y relacionar dos aspectos claves en el estudio de la meritocracia: percepciones y preferencias.
+Basados en el análisis crítico de las formas de medición de meritocracia a la fecha, el presente artículo propone un instrumento para medir y relacionar dos aspectos claves en el estudio de la meritocracia: percepciones y preferencias.
 Además, como un segundo eje de análisis considera la generación de indicadores respecto de aspectos meritocráticos y anti-meritocráticos, demostrando que no son los dos polos de un mismo continuo como muchos estudios anteriores parecen sugerir.
-La propuesta de medición además está orientada a generar un instrumento lo más breve posible de manera que pueda ser utilizado en encuestas de opinión.
+La propuesta de medición además está orientada a generar un instrumento lo más breve posible de manera que pueda ser utilizado en encuestas de opinión pública y así ser asociado a otros fenómenos sociales.
 
@@
 
-## La caja negra de las creencias en la meritocracia
+## The black-box of meritocratic beliefs
 
-Para ejemplificar la confusión del uso de creencias en meritocracia consideremos en primer lugar un artículo reciente de Mijs, donde se expresa la siguiente definición: "when I discuss meritocracy beliefs, I am referring to citizens’ belief in the importance of hard work relative to structural factors." [@mijs_paradox_2019, pg.9].
+Several approaches to the empirical study of meritocracy based on public opinion surveys make reference to the concept of _beliefs_, but behind this concept there are usually different meanings and operationalizations.
+To illustrate this point in the following we will start with the proposal from a recent paper by @mijs_paradox_2019, which we will take as a reference to discuss previous studies.
+
+The meritocratic beliefs' definition of Mijs is the following: "when I discuss meritocracy beliefs, I am referring to citizens’ belief in the importance of hard work relative to structural factors." [@mijs_paradox_2019, pg.9].
 In the operationalization, this is associated with the following indicator: "how important you think it is for getting ahead in life: (a) hard work", scored in a 1 to 5 likert scale.
@@
 
-a.
-Dimensionality: the item used by Mijs is part of an battery present in several international surveys usually called "get ahead", which presents a series of indicators related to what people consider important to get ahead in life: hard work, education, ambition, wealthy family, right connections, religion, race and gender.
-For Mijs, other aspects as education, that could be associated to talent, are not meritocratic: "Hard work is arguably the
-most meritocratic part of Michael Young’s equation, ‘Merit = Intelligence + Effort’, for the simple fact that intelligence itself is conditioned by a nonmeritocratic factor: who your parents happen to be" (p.5).
-This is a very strong assumption: effort would not depend on parents influence, and talent is not meritocratic.
-A similar approach only considering effort as the main element of meritocracy is taken by other authors ho only included effort as an indicator of meritocratic beliefs [@MaitnerEmotionalreactionsunequal2015;@MajorPerceiveddiscriminationworldview2007; @BubakPerceptionsmeritocracynote2019; @solt_economic_2016].
-We suggest here is that questions such as whether talent is not meritocratic (or less meritocratic than effort) should be open to empirical scrutiny and not disregarded upfront based on scholars' assumptions.
-Besides, the use of single indicators for measurement a construct is based on the assumption of a measurement without error.
+a. _Dimensionality_
 
-b.
-Beliefs: the "get ahead" battery refers to "how important you think it is", considered by Mijs as a belief in meritocracy.
-Nevertheless, another version of this same battery asks about "how important you think it _should_ be".
-Which one of both is the "belief"?
-If we consider belief as related to a normative content, actually the more close version is the "should be", being the other one more close to the perception of the situation.
-Such distinction was already made by @duru-bellat_whos_2012, who used the item "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" for "desired" meritocracy, whereas for "perceived" meritocracy they use two items: “Would you say that in your country, people are rewarded for their efforts?” and “...
+The item used by Mijs is part of an items' battery present in several international surveys, usually called "reasons to get ahead".
+This battery presents a series of indicators related to what people consider important to get ahead in life: hard work, education, ambition, wealthy family, right connections, religion, race and gender.
+Therefore, for Mijs other aspects as education, that could be associated to talent, are not meritocratic.
+As he points out: "Hard work is arguably the most meritocratic part of Michael Young’s equation, ‘Merit = Intelligence + Effort’, for the simple fact that intelligence itself is conditioned by a nonmeritocratic factor: who your parents happen to be" (p.5).
+
+In Mijs' proposal we can observe a couple of strong assumptions: effort would not depend on parents influence, and talent is not meritocratic (contrary to Michael Youngs original conceptualization).
+The problem of whether talent is or not meritocratic is something that actually should (and can) be tested, it cannot be simply ruled out particularly in empirical research.
+This conceptual and measurement issue it is possible to find in other studies that assume that effort is the main and only aspect of meritocracy [@MaitnerEmotionalreactionsunequal2015;@MajorPerceiveddiscriminationworldview2007; @BubakPerceptionsmeritocracynote2019; @solt_economic_2016].
+
+_b. Beliefs_
+
+The "resons to get ahead" battery refers to "how important you think it is", considered by Mijs as a belief in meritocracy for the item regarding effort.
+Nevertheless, another version of this same battery in the survey asks about "how important you think it _should_ be".
+Which one of both is a "belief"?
+According to @janmaat_subjective_2013-1 : "Perceptions refer to subjective estimates of existing inequality (i.e.
+thoughts about what is).
+Beliefs are here defined as normative ideas about just inequality (i.e.
+thoughts about what should be)"(p.359).
+Therefore, the referred paper would be actually using the term beliefs referring to perceptions.
+The same occurs in @reynolds_perceptions_2014, who explicitly use the term beliefs to talk about perceptions, whereas other authors use different terms as attitudes (@kunovich_systems_2007).
+The first attempt to shed light on this confusion was made by @duru-bellat_whos_2012, who used the item "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" for "desired" meritocracy (beliefs), whereas for "perceived" meritocracy they use two items: “Would you say that in your country, people are rewarded for their efforts?” and “...
 people are rewarded for their skills?”.
-Is the belief in meritocracy a perception or a preference?
-A comprehensive measurement should be in line with what @duru-bellat_whos_2012 did, in order to open up possibilities of analyzing whether perceptions and preferences are actually the same (i.e.
+
+Is the belief in meritocracy a perception or a desire/preference?
+In order to expand the analitical conceptual framework, we believe that both dimensions should be included in the analysis, as proposed by @duru-bellat_whos_2012.
+This opens up possibilities of analyzing whether perceptions and preferences are actually the same (i.e.
 correlation close to 1) or they are different aspects of the same phenomenon.
-Therefore, two main drawbacks of the term "beliefs" are that (i) it hides different operationalizations of the concept of meritocracy, and (ii) it blurs the distinction between perceptions and preferences.
-We argue that perceptions and preferences should be disentangled and used together, specially when analyzing preferences.
 As @son_hing_merit_2011-1 have pointed out, "People can believe that outcomes ought to be distributed on the basis of merit and yet vary in their perceptions of whether this is how society currently operates" (p.
 435).
-In other words, preferences should be interpreted in the context of perception: a large preference for meritocracy certainly means something totally different for someone perceiving high meritocracy than for someone perceiving low meritocracy.
+In other words, normative beliefs should be interpreted in the context of perception: a large normative belief in meritocracy certainly means something totally different for someone perceiving high meritocracy than for someone perceiving low meritocracy.
+In order to avoid the confusion generated by the term "belief", we propose the terms meritocratic preferences and meritocratic perceptions, as they better reflect the two dimension under scrutiny.
 
-c.
-Non-meritocratic aspects: @mijs_paradox_2019 make reference to some non-meritocratic aspects as talent, letting this out of the measurement.
-A different approach was followed by @kunovich_systems_2007, who decide to include non-meritocratic elements in the measurement of meritocracy.
+_c. Non-meritocratic aspects_
+
+@mijs_paradox_2019 makes reference to some non-meritocratic aspects as talent, which is ruled out of the operationalization of meritocracy.
+A different approach was followed by @kunovich_systems_2007, who decide to include some non-meritocratic elements.
 Using the items’ battery listing a number of reasons about “How important should be in deciding pay...” (as @duru-bellat_whos_2012 for desired meritocracy), he decided that reasons as education and responsibility are meritocratic and pointed 1 if considered essential, whereas reasons such as having a family and children were pointed 1 if they were considered "not important at all" (i.e.
@@
 A similar approach was taken by @newman_false_2015, reverse-coding non-meritocratic items, similarly to what occurs with the "Preference for the Merit Principle Scale" [@davey_preference_1999].
+
 The assumption that meritocratic and non-meritocratic elements are the poles of the same continuum was analyzed by @reynolds_perceptions_2014 using the same "get ahead" perceptions' battery items.
 They consider education, ambition and hard work as meritocratic and other reasons such as wealthy family and right connections and non-mertitocratic.
-Nevertheless, despite making this distinction the author ends up substracting one dimension from the other, assuming that they are two poles of the same continuum as @kunovich_systems_2007 did.
+Nevertheless, despite making this distinction the author ends up subtracting one dimension from the other, assuming that they are two poles of the same continuum as @kunovich_systems_2007 did.
+Still, taking into account this research perspective, we suggest that non-meritocratic aspects should be part of a meritocratic measurement but taken independently and not adding or subtracting from meritocratic ones unless it is empirically proved that they belong to the same conceptual dimension.
 
-\pagebreak
+_d. Accounting for measurement error_
 
-d.
-Accounting for measurement error: finally, most of the studies in meritocracy so far have not incorporated the issue of measurement error @brown_confirmatory_2006, using single indicators and/or composite indexes for measuring meritocracy.
-This means assuming that the construct is measured perfectly by the indicators chosen, going as far as proposing that "we decided to rely on a summative index.
+Finally, most of the studies in meritocracy so far have not incorporated the issue of measurement error [@brown_confirmatory_2006, @bollen_structural_1989], using single indicators and/or simple average indexes for measuring meritocracy.
+Such strategy assumes that the construct is measured perfectly by the indicators chosen, going as far as proposing that "...
 In choosing this strategy of index construction, we argue that *support for meritocracy is not a latent variable* [@kunovich_systems_2007 p.653-654].
@@
 
-Based on the previous limitations in the measurement of meritocracy presented in the previous section, in this paper we propose and test an instrument with the following characteristics:
+Based on the previous limitations and confussions in the measurement of meritocracy presented in the previous section, in this paper we propose and test an instrument with the following characteristics:
 
 - Multidimensional, incorporating previous distinctions between preferences and perceptions as well as between meritocratic and non-meritocratic aspects.
+
 - Multiple indicators for each dimension, in order to account for measurement error in a confirmatory factor analysis context.
+
 - Based on previous indicators as far as possible, for the sake of comparability between studies
+
 - Brief, as to be used in a regular survey.
@@
 
-The proposed measurement framework is based on two axis of analysis, as depicted in Table 1:
+The proposed measurement framework is depicted in Figure  \@ref(fig:merit-model):
 
-```{r include=FALSE}
-cap <- "Dimensions of meritocratic beliefs."
-if (knitr::is_html_output()) {ta <- table_nums("ta",cap)} else {ta <- cap}
+```{r merit-model, echo=FALSE, fig.cap = "Model of perception and preferences for meritocracy and non-meritocracy", fig.align='center'}
+
+knitr::include_graphics('../input/images/generalf.png')
 ```
 
-|                  	  | Perceptions 	| Preferences 	|
-|:-----------------	  |-------------	|-------------	|
-| _Meritocracy_     	|             	|             	|
-| _Non-meritocracy_ 	|             	|             	|
-Table:`r ta`
-
-The first axis distinguishes between different types of "beliefs", using instead the terms perceptions and preferences for meritocracy [@duru-bellat_whos_2012; @son_hing_merit_2011-1].
-Perceptions refers to the extent to which people see meritocracy working in their society, which in terms of measurement relates to the "reasons to get ahead" battery, whereas preferences refer to normative expectations that are usually linked to a "should" expression (e.g.
+The columns Perceptions and Preferences represent the distinction between this two concepts, usually confused under the label "beliefs".
+Perceptions refers to the extent to which people see that meritocracy functions or applies in their society, which in terms of measurement relates to items such as "I think hard work is important to get ahead in society", whereas preferences refer to normative expectations that are usually linked to a "should" expression (e.g.
 whether hard work should be related to payment).
-The second axis consider the distinction between meritocratic and non-meritocratic dimensions [@reynolds_perceptions_2014].
+The rows in the table of Figure 1 consider the distinction between meritocratic and non-meritocratic dimensions [@reynolds_perceptions_2014].
 This aspect has been usually treated as different ends of a same continuum in part of the previous research, an assumption that requieres empirical scrutiny.
@@
 
-Regarding the selection of indicators, most of them some adaptations of previous items that where streamlined in order to a better fit with the specific dimensions.
+Regarding the selection of indicators, most of them are taken or adapted from previous studies for the sake of comparability.
 For meritocratic indicators we use effort and talent as the main components of the traditional concept of merit as defined by @young_rise_1962, whereas for non-meritocratic dimensions we use having rich parents and good contacts.
@@
 
-**Hypotheses**
+The research hypotheses behind this conceptualization are the following:
 
@@
 
-* H3. Meritocratic preferences behave as the latent variable based on a normative value of effort and talent.
+* H3. Meritocratic preferences are a latent variable based on a normative value of effort and talent.
 
-* H4. Non-meritocratic preferences behave as a latent variable based on the normative value of the use of personal contacts and having rich parents.
+* H4. Non-meritocratic preferences are a latent variable based on the normative value of the use of personal contacts and having rich parents.
 
@@
 The data was obtained through an online questionnaire which was part of a larger study on meritocracy and preferences developed in Chile in 2019 and funded by the national scientific agency FONDECYT.
-The questionnaire was programmed in Qualtrics and the fieldwork was in charge of an external online survey agency ([netquest.cl](www.netquest.cl))during December 2019 and January 2020.
+The questionnaire was programmed in Qualtrics and the fieldwork was in charge of an external online survey agency ([netquest.cl](www.netquest.cl)) during December 2019 and January 2020.
 The sample was selected from a non-probabilistic design in three large cities in Chile.
@@
 Quotas were generated based on the survey of the Public Studies Center (CEP, 2019), which has a high prestige in the country and is also the counterpart agency of ISSP (International Social Survey Programme) in Chile.
-A total sample of 2141 people was collected, excluding those who did not answer the questions on the scale and those who did not accept informed consent.
+A total sample of 2,141 people was collected, excluding those who did not answer the questions on the scale and those who did not accept informed consent.
 As it usually occurs online samples, there were some limitations in achieving the quotas for lower educational levels.
-
-\pagebreak
 
@@
 
-The proposed scale of perceptions and preferences about meritocracy consists of 8 indicators that are grouped into 4 dimensions referred above: Perceptions (meritocratic/nonmeritocratic) and preferences (meritocratic/non-meritocratic).
+The proposed scale of perceptions and preferences about meritocracy consists of 8 indicators that are grouped into the 4 dimensions referred above: Perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic).
 In order to achieve at least some comparability with previous studies, the questions were adapted from the items battery "reasons to get ahead" (ISSP/GSS), which are mostly used for operationalizing meritocracy in previous empirical studies [@mijs_paradox_2019; @duru-bellat_whos_2012; @reynolds_perceptions_2014].
@@
 ```{r echo=FALSE}
- tab01 <- read.csv(file = "input/data/original/materials/table01.csv",header = 1,sep = ";")
+ tab01 <- read.csv(file = "../input/data/original/materials/table01.csv",header = 1,sep = ",")
 
@@
  kable(tab01,table_format,booktabs = T, linesep = "",col.names = cnames, caption = tb) %>%
-   kable_styling(full_width = T,latex_options = c("hold_position","scale_down"),position = "center",font_size = 9) %>%
+   kable_styling(full_width = T,latex_options = c("hold_position","scale_down"),position = "center",font_size = 14, bootstrap_options=c("striped", "bordered")) %>%
    column_spec(column = 1,width = "1.5cm",) %>%
@@
 
-With the objective of evaluating the effect of the order of the indicators, the respondents (_n = 2141_) were randomly divided into three groups as explained in __Figure 1__.
-The scale was presented to the first group (_n = 712_) in the order that appears in Table 2, in such a way that the items of the same factor are found together.
-The second group (_n = 717_)was presented with the scale with the items of the same factor separated from each other, in such a way that they were consulted at the same time for their perceptions and preferences of a particular topic (e.g.
-distribution according to effort).
-Finally, in the third group (_n = 712_) the items were completely randomized.
+With the objective of evaluating the effect of the indicators ordering, the respondents (_n = 2141_) were randomly divided into three different order versions as explained in __Figure 1__.
+The scale was presented to the first group (_n = 712_) in the order that appears in Table 2.
+For the second group (_n = 717_), the order of the items was organized according to the topics of the items, e.g.
+for the topic of hard work the item about perception was followed by the item about preference, and te same for the rest of the topics.
+Finally, for the third group (_n = 712_) the items were completely randomized.
 
@@
 ![`r diam1`](../input/images/app_mod.png){ width=80% }
-
-\pagebreak
 
@@
 
+The study was pre-registered ... (link) ***
+
 # Results
@@
 
-Como puede observarse en la tabla 3, los indicadores poseen valores que van desde el 1 al 5.
-Se observan promedios desde 2,41 correspondiente a la preferencia de distribucion según contactos, hasta 3,89, correspondiente a la preferencia de distribucion según esfuerzo.
-Ambos indicadores son coherentes con la adhesión  a las creencias meritocraticas, privilegiando la distribucion meritocratica por sobre aquella que se debe a ventajas externas.
+Como puede observarse en la tabla 3, los indicadores poseen valores que van desde el 1 (totalmente en desacuerdo) al 5 (totalmente de acuerdo).
+Se observan promedios desde 2.41 correspondiente a preferencia-contactos, hasta 3.89 correspondiente a la preferencia-esfuerzo.
+Ambos indicadores son coherentes con la adhesión general a la meritocracia, privilegiando aspectos individuales como el esfuerzo.
 
@@
 ```{r echo=FALSE}
-load(file = "input/data/proc/dat04.RData")
+load(file = "../input/data/proc/dat04.RData")
 dat04b<- dat04 %>% rename("A.Perception Effort" =perc_effort,
@@
 sumdat04<- kable(table04b,format =table_format, booktabs= T, escape = FALSE,align = "l",linesep = "", caption = tades
-) %>% kable_styling(full_width = table_format2, font_size = 9,position = "center");sumdat04
+) %>% kable_styling(full_width = table_format2, font_size = 14,position = "center");sumdat04
 ```
 
-\pagebreak
-
-A partir del grafico liker  se observa que, en general, hay una mayor percepción de aspectos no meritocráticos que meritocráticos, mientras que en el caso de preferencias ocurre lo opuesto.
+El gráfico XX muestra ...
+A partir del gráfico se observa que en general hay una mayor percepción de elementos no meritocráticos que meritocráticos, mientras que en el caso de preferencias ocurre lo opuesto.
 En cuanto a las preferencias, llama la atención el rol preponderante del esfuerzo por sobre el talento.
@@
 
-![`r liker`](../output/images/plotlikert.png){ width=70% }
+![`r liker`](../output/images/plotlikert.png){ width=60% }
 
@@
 
-\pagebreak
-
-El grafico de correlaciones policoricas evidencia relaciones de moderada a alta intensidad entre los indicadores que corresponden al mismo factor, mientras que se observan correlaciones de baja intensidad en el resto de los cruces.
-Destacan además las relaciones entre las percepciones y preferencias de tipo meritocraticas, lo que no ocurre con los indicadores no meritocraticos.
+El Grafico XX muestra ....
+Se observan relaciones de moderada a alta intensidad entre los indicadores que corresponden al mismo factor (por ejemplo, percepción de meritocracia esfuerzo con talento r=0.56), mientras que se observan correlaciones de baja intensidad entre el resto de los cruces.
+Destacan además las relaciones entre las percepciones y preferencias de tipo meritocráticas, lo que no ocurre con los indicadores no meritocraticos.
 
@@
 
-![`r fb`](../output/images/corpoly.png){ width=70% }
+![`r fb`](../output/images/corpoly.png){ width=60% }
 
@@
 
-En suma, los análisis descriptivos señalan una relativa adhesión a la moral meritocratica, las cuales, a modo general, se expresan en una mayor preferencia de distribucion según criterios meritocraticos y una menor preferencia de criterios no meritocraticos.
-Igualmente, es observable una relativamente baja percepcion de distribucion meritocratica.
-Además se observa una relación coherente entre los indicadores según lo propuesto en el modelo teórico en el preregistro.
+En suma, los análisis descriptivos señalan una relativa adhesión a la moral meritocratica, las cuales, a modo general, se expresan en una mayor preferencia por criterios meritocraticos y una menor por no meritocráticos.
+Igualmente, es observable una relativamente baja percepción de meritocracia.
+Se observa además una relación coherente entre los indicadores según lo propuesto en el modelo teórico en el preregistro del estudio, es decir, los pares de items asociados a un factor específico muestran correlaciones con un tamaño de efecto grande (por ejemplo, preferencias meritocrácticas por los items asociados a esfuerzo y a talento).
+En particular las asociaciones entre esfuerzo y talento son relevantes, ya que desestiman previos supuestos sobre que el talento no sería un criterio meritocrático [@mijs_paradox_2019], de otra manera la correlación sería cero o negativa.
+Junto a ello, vemos que no existe una correlacion negativa entre aspectos meritocráticos y no meritocráticos, desestimando los supuestos de estudios previos que  señalaban que estas dimensiones serían los polos opuestos de un mismo continuo [@reynolds_perceptions_2014].
 
-\pagebreak
+## Confirmatory Factor Analysis
 
-## Models estimation
+This section tests the item's battery and the corresponding conceptual framework under test.
+For this, we first estimate a confirmatory factor analysis model for the whole sample, and secondly we test the order effects applying the same model to each of the three order versions.
 
-En una primera instancia se probó el ajuste del modelo con tres distintos órdenes de los ítems, los cuales fueron aplicados a un tercio de la muestra de la primera ola cada uno.
-El primer orden de los ítems corresponde al orden de de la tabla 2, en el que primero se pregunta a los encuestados por todas las percepciones y después por todas las preferencias, en el segundo orden se les pregunta seguidamente por percepciones y preferencias respecto a un mismo tema, por ejemplo, al papel del esfuerzo, y en el tercero, los ítems aparecen de manera aleatoria.
+### Full sample CFA
+
+Figure X shows the results of the estimation for the four-factor model:
+
+Los modelos confirmatorios fueron estimados en primer lugar para cada uno de los tres administration-sets. Los resultados se presentan en la Tabla XX
 
@@
 
-load(file = "output/tables/tb.fm.RData")
+load(file = "../output/tables/tb.fm.RData")
 
@@
 
-Todos los modelos independiente de los órdenes obtuvieron un ajuste adecuado.
+Como se puede observar en la Tabla XX, todos los modelos independiente de los órdenes obtuvieron un ajuste adecuado.
 con CFI superiores a .95 y RMSEA inferiores a .8, por contraparte ni un modelo logró un chi-square no significativo, aunque tanto el modelo aleatorio como el primero obtuvieron un adecuado chi-square ratio menor a 3.

### 2020-07-26 docs/paper.tex
update paper

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Measuring Perceptions and Preferences for Meritocracy},
  colorlinks=true,
  linkcolor=blue,
  filecolor=Maroon,
  citecolor=Blue,
  urlcolor=Blue,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=0.78in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{caption}
\captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
\usepackage{graphicx}
\usepackage{float}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}

\title{Measuring Perceptions and Preferences for Meritocracy}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

\setstretch{1.5}
\emph{Abstract}

Economic and social inequality have raised growing concerns and crises
across societies. One of social science concepts associated to the
maintenance of inequality is the belief in meritocracy, which would
legitimize economic differences based on effort and talent. Despite its
wide use, empirical research on meritocracy is something relatively
novel. A number of studies have relied mostly in secondary data to
operationalize meritocracy, with a large variation in the use and
interpretation of survey items. Starting from a review of studies that
measure meritocracy, this article identifies a series of drawbacks and
inconsistencies within and between studies regarding the
conceptualization and measurement of meritocracy. Based on this critical
analysis, we propose an items battery called ``Perceptions and
preferences for meritocracy scale'', which is tested with confirmatory
factor analysis with data from an online survey study (N=2,141). The
results support the proposed conceptual structure which not only
distinguishes between perceptions and preferences, but also between
meritocratic and non-meritocratic dimensions. The discussion highlights
the relevance of considering these different dimensions in order to
advance in the study of meritocracy.

\hypertarget{introducciuxf3n}{%
\section{Introducción}\label{introducciuxf3n}}

La desigualdad económica se ha vuelto un tema que genera creciente
preocupación y malestar alrededor del mundo. Esto se ha expresado en una
serie de protestas como la emblemática ``occupy wall street'' el año
2011, así como también en una serie de análisis críticos respecto del
desarrollo del capitalismo y sus consecuencias ({\textbf{???}}). En este
contexto, el estudio de las visiones, preferencias y percepciones
respecto de la desigualdad han adquirido relevancia en las ciencias
sociales, en temas como las preferencias redistributivas (Alesina and
Angeletos \protect\hyperlink{ref-alesina_Fairness_2005}{2005}; Dimick,
Rueda, and Stegmueller \protect\hyperlink{ref-dimick_Models_2018}{2018})
la legitimación de la desigualdad económica (Schröder
\protect\hyperlink{ref-schroder_Income_2017}{2017}) y el funcionamiento
de la meritocracia ({\textbf{???}}; {\textbf{???}}; {\textbf{???}}).

En general, la meritocracia se define como un sistema de distribución de
recursos y recompensas basados en el mérito individual, que en su
concepción original es una suma de talento y esfuerzo (Young
\protect\hyperlink{ref-young_rise_1962}{1962}). Esta concepción
tradicional de mérito pone en un lugar secundario la posible
interferencia de factores estructurales o no meritocráticos como la
herencia, los contactos personales, o la suerte (Breen and Goldthorpe
\protect\hyperlink{ref-breenClassInequalityMeritocracy1999}{1999};
Saunders \protect\hyperlink{ref-saundersMightBritainBe1995}{1995}; Yair
\protect\hyperlink{ref-yairMeritocracy2007}{2007}; Land
\protect\hyperlink{ref-landWeSatTable2006}{2006}; Young
\protect\hyperlink{ref-youngRiseMeritocracy1994}{1994}). Una serie de
estudios han realizado críticas a la realización de este estándar moral
de distribución, planteando que es una promesa incumplida dada la
influencia preponderante de otros elementos más allá del mérito en el
estatus individual ({\textbf{???}}; Arrow, Bowles, and Durlauf
\protect\hyperlink{ref-arrow_meritocracy_2000}{2000}; Goldthorpe
\protect\hyperlink{ref-goldthorpe_myth_2003}{2003}; Markovits
\protect\hyperlink{ref-markovits_Meritocracy_2019}{2019}). Por otro
lado, desde la psicología social y la sociología se han estudiado las
características y consecuencias de las creencias en la meritocracia, en
general basados en la hipótesis que mayor creencia en la meritocracia
lleva a una mayor legitimación de las desigualdades ({\textbf{???}};
{\textbf{???}}; {\textbf{???}}; Hadjar
\protect\hyperlink{ref-hadjar_meritokratie_2008}{2008}; Madeira et al.
\protect\hyperlink{ref-MadeiraPrimesConsequencesSystematic2019}{2019}).

Debido al rol que cumplen las creencias meritocráticas dentro del
pensamiento neoliberal ({\textbf{???}}), han surgido múltiples
investigaciones que evalúan la relación entre creencias meritocráticas y
diversos ámbitos sociales de la actualidad. Por ejemplo, se han
desarrollado estudios que vinculan la meritocracia al reforzamiento de
estereotipos socioeconómicos, de género y de etnias (Madeira et al.
\protect\hyperlink{ref-MadeiraPrimesConsequencesSystematic2019}{2019};
{\textbf{???}}; {\textbf{???}}), así como también líneas de
investigación que evalúan el efecto de las creencias meritocráticas en
el contexto educativo ({\textbf{???}}; {\textbf{???}}; {\textbf{???}}) y
en el contexto organizacional de las empresas ({\textbf{???}};
{\textbf{???}}).

Para poder dar cuenta de los niveles de creencia en la meritocracia los
estudios a la fecha generalmente han utilizado algunos indicadores de
encuestas ya existentes, y en el menor de los casos se han creado
instrumentos ad-hoc. Sin embargo, y como mostraremos más adelante, las
formas de medición de meritocracia varían extremadamente entre estudios.
Muchas veces fenómenos similares se asocian a indicadores distintos, y
también ocurre que fenómenos distintos son medidos con indicadores
similares, todo lo cual hace dificulta la comparabilidad entre estudios
y el poder avanzar en la comprensión y estudio de la meritocracia.

Basados en el análisis crítico de las formas de medición de meritocracia
a la fecha, el presente artículo propone un instrumento para medir y
relacionar dos aspectos claves en el estudio de la meritocracia:
percepciones y preferencias. Además, como un segundo eje de análisis
considera la generación de indicadores respecto de aspectos
meritocráticos y anti-meritocráticos, demostrando que no son los dos
polos de un mismo continuo como muchos estudios anteriores parecen
sugerir. La propuesta de medición además está orientada a generar un
instrumento lo más breve posible de manera que pueda ser utilizado en
encuestas de opinión pública y así ser asociado a otros fenómenos
sociales.

\hypertarget{la-mediciuxf3n-de-los-aspectos-subjetivos-de-la-meritocracia}{%
\section{La medición de los aspectos subjetivos de la
meritocracia}\label{la-mediciuxf3n-de-los-aspectos-subjetivos-de-la-meritocracia}}

A continuación se presenta una revisión de una serie de investigaciones
que se han abocado al estudio de la meritocracia y que para ello han
hecho una propuesta de medición. El primer eje de análisis tiene que ver
con el uso del concepto ``creencias'' para referir a distintos aspectos
subjetivos relacionados con meritocracia. El segundo eje tiene que ver
con el uso de indicadores sobre aspectos anti-meritocráticos como el
polo opuesto de los meritocráticos.

\hypertarget{the-black-box-of-meritocratic-beliefs}{%
\subsection{The black-box of meritocratic
beliefs}\label{the-black-box-of-meritocratic-beliefs}}

Several approaches to the empirical study of meritocracy based on public
opinion surveys make reference to the concept of \emph{beliefs}, but
behind this concept there are usually different meanings and
operationalizations. To illustrate this point in the following we will
start with the proposal from a recent paper by Mijs
(\protect\hyperlink{ref-mijs_paradox_2019}{2019}), which we will take as
a reference to discuss previous studies.

The meritocratic beliefs' definition of Mijs is the following: ``when I
discuss meritocracy beliefs, I am referring to citizens' belief in the
importance of hard work relative to structural factors.'' (Mijs
\protect\hyperlink{ref-mijs_paradox_2019}{2019}, pg.9). In the
operationalization, this is associated with the following indicator:
``how important you think it is for getting ahead in life: (a) hard
work'', scored in a 1 to 5 likert scale. There are several assumptions
behind this decision that are worth discussing:

\begin{enumerate}
\def\labelenumi{\alph{enumi}.}
\tightlist
\item
  \emph{Dimensionality}
\end{enumerate}

The item used by Mijs is part of an items' battery present in several
international surveys, usually called ``reasons to get ahead''. This
battery presents a series of indicators related to what people consider
important to get ahead in life: hard work, education, ambition, wealthy
family, right connections, religion, race and gender. Therefore, for
Mijs other aspects as education, that could be associated to talent, are
not meritocratic. As he points out: ``Hard work is arguably the most
meritocratic part of Michael Young's equation, `Merit = Intelligence +
Effort', for the simple fact that intelligence itself is conditioned by
a nonmeritocratic factor: who your parents happen to be'' (p.5).

In Mijs' proposal we can observe a couple of strong assumptions: effort
would not depend on parents influence, and talent is not meritocratic
(contrary to Michael Youngs original conceptualization). The problem of
whether talent is or not meritocratic is something that actually should
(and can) be tested, it cannot be simply ruled out particularly in
empirical research. This conceptual and measurement issue it is possible
to find in other studies that assume that effort is the main and only
aspect of meritocracy ({\textbf{???}}; {\textbf{???}}; {\textbf{???}};
{\textbf{???}}).

\emph{b. Beliefs}

The ``resons to get ahead'' battery refers to ``how important you think
it is'', considered by Mijs as a belief in meritocracy for the item
regarding effort. Nevertheless, another version of this same battery in
the survey asks about ``how important you think it \emph{should} be''.
Which one of both is a ``belief''? According to ({\textbf{???}}) :
``Perceptions refer to subjective estimates of existing inequality
(i.e.~thoughts about what is). Beliefs are here defined as normative
ideas about just inequality (i.e.~thoughts about what should
be)''(p.359). Therefore, the referred paper would be actually using the
term beliefs referring to perceptions. The same occurs in Reynolds and
Xian (\protect\hyperlink{ref-reynolds_perceptions_2014}{2014}), who
explicitly use the term beliefs to talk about perceptions, whereas other
authors use different terms as attitudes (Kunovich and Slomczynski
(\protect\hyperlink{ref-kunovich_systems_2007}{2007})). The first
attempt to shed light on this confusion was made by Duru-Bellat and
Tenret (\protect\hyperlink{ref-duru-bellat_whos_2012}{2012}), who used
the item ``how important should the number of years spent in education
and training be in deciding how much money people ought to earn?'' for
``desired'' meritocracy (beliefs), whereas for ``perceived'' meritocracy
they use two items: ``Would you say that in your country, people are
rewarded for their efforts?'' and ``\ldots{} people are rewarded for
their skills?''.

Is the belief in meritocracy a perception or a desire/preference? In
order to expand the analitical conceptual framework, we believe that
both dimensions should be included in the analysis, as proposed by
Duru-Bellat and Tenret
(\protect\hyperlink{ref-duru-bellat_whos_2012}{2012}). This opens up
possibilities of analyzing whether perceptions and preferences are
actually the same (i.e.~correlation close to 1) or they are different
aspects of the same phenomenon. As Son Hing et al.
(\protect\hyperlink{ref-son_hing_merit_2011-1}{2011}) have pointed out,
``People can believe that outcomes ought to be distributed on the basis
of merit and yet vary in their perceptions of whether this is how
society currently operates'' (p.~435). In other words, normative beliefs
should be interpreted in the context of perception: a large normative
belief in meritocracy certainly means something totally different for
someone perceiving high meritocracy than for someone perceiving low
meritocracy. In order to avoid the confusion generated by the term
``belief'', we propose the terms meritocratic preferences and
meritocratic perceptions, as they better reflect the two dimension under
scrutiny.

\emph{c.~Non-meritocratic aspects}

Mijs (\protect\hyperlink{ref-mijs_paradox_2019}{2019}) makes reference
to some non-meritocratic aspects as talent, which is ruled out of the
operationalization of meritocracy. A different approach was followed by
Kunovich and Slomczynski
(\protect\hyperlink{ref-kunovich_systems_2007}{2007}), who decide to
include some non-meritocratic elements. Using the items' battery listing
a number of reasons about ``How important should be in deciding
pay\ldots{}'' (as Duru-Bellat and Tenret
(\protect\hyperlink{ref-duru-bellat_whos_2012}{2012}) for desired
meritocracy), he decided that reasons as education and responsibility
are meritocratic and pointed 1 if considered essential, whereas reasons
such as having a family and children were pointed 1 if they were
considered ``not important at all'' (i.e.~reverse coded). A similar
approach was taken by Newman, Johnston, and Lown
(\protect\hyperlink{ref-newman_false_2015}{2015}), reverse-coding
non-meritocratic items, similarly to what occurs with the ``Preference
for the Merit Principle Scale'' (Davey et al.
\protect\hyperlink{ref-davey_preference_1999}{1999}).

The assumption that meritocratic and non-meritocratic elements are the
poles of the same continuum was analyzed by Reynolds and Xian
(\protect\hyperlink{ref-reynolds_perceptions_2014}{2014}) using the same
``get ahead'' perceptions' battery items. They consider education,
ambition and hard work as meritocratic and other reasons such as wealthy
family and right connections and non-mertitocratic. Nevertheless,
despite making this distinction the author ends up subtracting one
dimension from the other, assuming that they are two poles of the same
continuum as Kunovich and Slomczynski
(\protect\hyperlink{ref-kunovich_systems_2007}{2007}) did. Still, taking
into account this research perspective, we suggest that non-meritocratic
aspects should be part of a meritocratic measurement but taken
independently and not adding or subtracting from meritocratic ones
unless it is empirically proved that they belong to the same conceptual
dimension.

\emph{d.~Accounting for measurement error}

Finally, most of the studies in meritocracy so far have not incorporated
the issue of measurement error ({\textbf{???}}), using single indicators
and/or simple average indexes for measuring meritocracy. Such strategy
assumes that the construct is measured perfectly by the indicators
chosen, going as far as proposing that "\ldots{} In choosing this
strategy of index construction, we argue that \emph{support for
meritocracy is not a latent variable} (Kunovich and Slomczynski
\protect\hyperlink{ref-kunovich_systems_2007}{2007}, 653--54). Some
advances were done by Reynolds and Xian
(\protect\hyperlink{ref-reynolds_perceptions_2014}{2014}) by doing a
principal component analysis of meritocratic and non-meritocratic
dimensions, but somewhat contradictorily they end up in a sum index
despite proving multidimensionality.

\hypertarget{an-instrument-proposal}{%
\subsection{An instrument proposal}\label{an-instrument-proposal}}

Based on the previous limitations and confussions in the measurement of
meritocracy presented in the previous section, in this paper we propose
and test an instrument with the following characteristics:

\begin{itemize}
\item
  Multidimensional, incorporating previous distinctions between
  preferences and perceptions as well as between meritocratic and
  non-meritocratic aspects.
\item
  Multiple indicators for each dimension, in order to account for
  measurement error in a confirmatory factor analysis context.
\item
  Based on previous indicators as far as possible, for the sake of
  comparability between studies
\item
  Brief, as to be used in a regular survey. In this point it differs for
  instance from the proposal of the ``Preference for the Merit Principle
  Scale'' (Davey et al.
  \protect\hyperlink{ref-davey_preference_1999}{1999}), as they use 15
  items just for one dimension (besides the problem of reverse-coding
  non-meritocratic items).
\end{itemize}

The proposed measurement framework is depicted in Figure 1:

\includegraphics[width=0.6\textwidth,height=\textheight]{../input/images/generalf.png}

The columns Perceptions and Preferences represent the distinction
between this two concepts, usually confused under the label ``beliefs''.
Perceptions refers to the extent to which people see that meritocracy
functions or applies in their society, which in terms of measurement
relates to items such as ``I think hard work is important to get ahead
in society'', whereas preferences refer to normative expectations that
are usually linked to a ``should'' expression (e.g.~whether hard work
should be related to payment). The rows in the table of Figure 1
consider the distinction between meritocratic and non-meritocratic
dimensions (Reynolds and Xian
\protect\hyperlink{ref-reynolds_perceptions_2014}{2014}). This aspect
has been usually treated as different ends of a same continuum in part
of the previous research, an assumption that requieres empirical
scrutiny. These non-meritocratic elements usually refer to the use of
personal contacts or family advantages to get ahead in life.

Regarding the selection of indicators, most of them are taken or adapted
from previous studies for the sake of comparability. For meritocratic
indicators we use effort and talent as the main components of the
traditional concept of merit as defined by Young
(\protect\hyperlink{ref-young_rise_1962}{1962}), whereas for
non-meritocratic dimensions we use having rich parents and good
contacts. The description of the specific items is presented in the
methodology section.

The research hypotheses behind this conceptualization are the following:

\begin{itemize}
\item
  H1. The perception of meritocracy is a latent variable based on
  indicators of the importance attributed to talent and the effort to
  get ahead in life.
\item
  H2. The non-meritocratic perception is a latent variable that derives
  from two indicators related to the agreement with the statement that
  people with contacts and rich parents manage to get ahead.
\item
  H3. Meritocratic preferences are a latent variable based on a
  normative value of effort and talent.
\item
  H4. Non-meritocratic preferences are a latent variable based on the
  normative value of the use of personal contacts and having rich
  parents.
\end{itemize}

\hypertarget{methodology}{%
\section{Methodology}\label{methodology}}

\hypertarget{data-collection}{%
\subsection{Data collection}\label{data-collection}}

The data was obtained through an online questionnaire which was part of
a larger study on meritocracy and preferences developed in Chile in 2019
and funded by the national scientific agency FONDECYT. The questionnaire
was programmed in Qualtrics and the fieldwork was in charge of an
external online survey agency (\href{www.netquest.cl}{netquest.cl})
during December 2019 and January 2020. The sample was selected from a
non-probabilistic design in three large cities in Chile. The quota
method based on age, sex and educational level was used. Quotas were
generated based on the survey of the Public Studies Center (CEP, 2019),
which has a high prestige in the country and is also the counterpart
agency of ISSP (International Social Survey Programme) in Chile. A total
sample of 2,141 people was collected, excluding those who did not answer
the questions on the scale and those who did not accept informed
consent. As it usually occurs online samples, there were some
limitations in achieving the quotas for lower educational levels.

\hypertarget{instrument-design}{%
\subsection{Instrument design}\label{instrument-design}}

The proposed scale of perceptions and preferences about meritocracy
consists of 8 indicators that are grouped into the 4 dimensions referred
above: Perceptions (meritocratic/non-meritocratic) and preferences
(meritocratic/non-meritocratic). In order to achieve at least some
comparability with previous studies, the questions were adapted from the
items battery ``reasons to get ahead'' (ISSP/GSS), which are mostly used
for operationalizing meritocracy in previous empirical studies (Mijs
\protect\hyperlink{ref-mijs_paradox_2019}{2019}; Duru-Bellat and Tenret
\protect\hyperlink{ref-duru-bellat_whos_2012}{2012}; Reynolds and Xian
\protect\hyperlink{ref-reynolds_perceptions_2014}{2014}). The eight
items ordered according to dimensions are presented in Table 2. These 8
likert-type items have 5 response alternatives ranging from ``Completely
disagree''(1) to ``Completely agree'' (5).

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-6}Items according to dimension.}
\centering
\resizebox{\linewidth}{!}{
\fontsize{14}{16}\selectfont
\begin{tabu} to \linewidth {>{\raggedright\arraybackslash}p{1.5cm}>{\raggedright\arraybackslash}p{2 cm}>{\raggedright}X>{\raggedright}X}
\toprule
Dimension & Factor & Statement (english) & Statement (spanish)\\
\midrule
 &  & Those who try harder get greater rewards than those who work less. & Quienes más se esfuerzan logran obtener mayores recompensas que quienes se esfuerzan menos.\\
\cmidrule{3-4}
 & \multirow{-2}{2 cm}{\raggedright\arraybackslash Meritocratic} & Those who have more talent achieve greater rewards than those who have less talent. & Quienes poseen más talento logran obtener mayores recompensas que quienes poseen menos talento.\\
\cmidrule{2-4}
 &  & Those who have rich parents succeed. & Quienes tienen padres ricos logran salir adelante.\\
\cmidrule{3-4}
\multirow{-4}{1.5cm}{\raggedright\arraybackslash Perception} & \multirow{-2}{2 cm}{\raggedright\arraybackslash Non meritocratic} & Those who have good contacts succeed. & Quienes tienen buenos contactos logran salir adelante.\\
\cmidrule{1-4}
 &  & Those who try harder should get greater rewards than those who work less. & Quienes más se esfuerzan deberían obtener mayores recompensas que quienes se esfuerzan menos.\\
\cmidrule{3-4}
 & \multirow{-2}{2 cm}{\raggedright\arraybackslash Meritocratic} & Those who have more talent should get greater rewards than those who have less talent. & Quienes poseen más talento deberían obtener mayores recompensas que quienes poseen menos talento.\\
\cmidrule{2-4}
 &  & It's fine that those with rich parents get ahead. & Está bien que quienes tienen padres ricos salgan adelante.\\
\cmidrule{3-4}
\multirow{-4}{1.5cm}{\raggedright\arraybackslash Preference} & \multirow{-2}{2 cm}{\raggedright\arraybackslash Non meritocratic} & It's fine that those who have good contacts get ahead. & Está bien que quienes tienen buenos contactos salgan adelante.\\
\bottomrule
\end{tabu}}
\end{table}

\pagebreak

\hypertarget{administration-sets}{%
\subsection{Administration sets}\label{administration-sets}}

With the objective of evaluating the effect of the indicators ordering,
the respondents (\emph{n = 2141}) were randomly divided into three
different order versions as explained in \textbf{Figure 1}. The scale
was presented to the first group (\emph{n = 712}) in the order that
appears in Table 2. For the second group (\emph{n = 717}), the order of
the items was organized according to the topics of the items, e.g.~for
the topic of hard work the item about perception was followed by the
item about preference, and te same for the rest of the topics. Finally,
for the third group (\emph{n = 712}) the items were completely
randomized.

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth,height=\textheight]{../input/images/app_mod.png}
\caption{Survey flow}
\end{figure}

\hypertarget{methods}{%
\subsection{Methods}\label{methods}}

For testing the scale's underlying constructs we estimate confirmatory
factor analysis models (CFA). The model estimates one factor for each
dimension, as represented in the following figure:

\begin{figure}
\centering
\includegraphics{../output/images/meas01.png}
\caption{Theorical model}
\end{figure}

CFA was conducted using the \texttt{lavaan} package (version 0.6-3;
Rosseel, 2020) with diagonally weighted least squares (DWLS) estimation
due to the items' ordinal level of measurement (Kline, 2016; Rosseel,
2020). As recommended by Brown (2008), we assessed model fit by jointly
considering the comparative fit index and Tucker-Lewis Index (CFI and
TLI; acceptable fit \textgreater{} 0.95), Root of the average squared
residual approximation (RMSEA; acceptable fit \textless{} 0.08),
Chi-square: (p-value; acceptable fit \textgreater{} 0.05, and Chi-square
ratio:\textgreater{} 3).

The study was pre-registered \ldots{} (link) ***

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{descriptive-analysis}{%
\subsection{Descriptive analysis}\label{descriptive-analysis}}

Como puede observarse en la tabla 3, los indicadores poseen valores que
van desde el 1 (totalmente en desacuerdo) al 5 (totalmente de acuerdo).
Se observan promedios desde 2.41 correspondiente a
preferencia-contactos, hasta 3.89 correspondiente a la
preferencia-esfuerzo. Ambos indicadores son coherentes con la adhesión
general a la meritocracia, privilegiando aspectos individuales como el
esfuerzo.

\begin{table}

\caption{\label{tab:unnamed-chunk-10}Descriptive statistics of the scale.}
\centering
\fontsize{9}{11}\selectfont
\begin{tabular}[t]{lllll}
\toprule
  & Mean & SD & Min & Max\\
\midrule
A.Perception Effort & 3.20 & 1.38 & 1 & 5\\
B.Perception Talent & 3.02 & 1.16 & 1 & 5\\
C.Perception rich parents & 3.66 & 1.36 & 1 & 5\\
D.Perception contacts & 3.79 & 1.24 & 1 & 5\\
E.Preferences Effort & 3.89 & 1.25 & 1 & 5\\
F.Preferences Talent & 3.24 & 1.19 & 1 & 5\\
G.Preferences rich parents & 2.69 & 1.18 & 1 & 5\\
H.Preferences contacts & 2.41 & 1.11 & 1 & 5\\
\bottomrule
\end{tabular}
\end{table}

El gráfico XX muestra \ldots{} A partir del gráfico se observa que en
general hay una mayor percepción de elementos no meritocráticos que
meritocráticos, mientras que en el caso de preferencias ocurre lo
opuesto. En cuanto a las preferencias, llama la atención el rol
preponderante del esfuerzo por sobre el talento.

\begin{figure}
\centering
\includegraphics[width=11\textwidth,height=\textheight]{../output/images/plotlikert.png}
\caption{Descriptive plot}
\end{figure}

\pagebreak

El Grafico XX muestra \ldots. Se observan relaciones de moderada a alta
intensidad entre los indicadores que corresponden al mismo factor (por
ejemplo, percepción de meritocracia esfuerzo con talento r=0.56),
mientras que se observan correlaciones de baja intensidad entre el resto
de los cruces. Destacan además las relaciones entre las percepciones y
preferencias de tipo meritocráticas, lo que no ocurre con los
indicadores no meritocraticos.

\begin{figure}
\centering
\includegraphics[width=0.9\textwidth,height=\textheight]{../output/images/corpoly.png}
\caption{polychoric correlation plot}
\end{figure}

En suma, los análisis descriptivos señalan una relativa adhesión a la
moral meritocratica, las cuales, a modo general, se expresan en una
mayor preferencia por criterios meritocraticos y una menor por no
meritocráticos. Igualmente, es observable una relativamente baja
percepción de meritocracia. Se observa además una relación coherente
entre los indicadores según lo propuesto en el modelo teórico en el
preregistro del estudio, es decir, los pares de items asociados a un
factor específico muestran correlaciones con un tamaño de efecto grande
(por ejemplo, preferencias meritocrácticas por los items asociados a
esfuerzo y a talento). En particular las asociaciones entre esfuerzo y
talento son relevantes, ya que desestiman previos supuestos sobre que el
talento no sería un criterio meritocrático (Mijs
\protect\hyperlink{ref-mijs_paradox_2019}{2019}), de otra manera la
correlación sería cero o negativa. Junto a ello, vemos que no existe una
correlacion negativa entre aspectos meritocráticos y no meritocráticos,
desestimando los supuestos de estudios previos que señalaban que estas
dimensiones serían los polos opuestos de un mismo continuo (Reynolds and
Xian \protect\hyperlink{ref-reynolds_perceptions_2014}{2014}).

\hypertarget{confirmatory-factor-analysis}{%
\subsection{Confirmatory Factor
Analysis}\label{confirmatory-factor-analysis}}

This section tests the item's battery and the corresponding conceptual
framework under test. For this, we first estimate a confirmatory factor
analysis model for the whole sample, and secondly we test the order
effects applying the same model to each of the three order versions.

\hypertarget{full-sample-cfa}{%
\subsubsection{Full sample CFA}\label{full-sample-cfa}}

Figure X shows the results of the estimation for the four-factor model:

Los modelos confirmatorios fueron estimados en primer lugar para cada
uno de los tres administration-sets. Los resultados se presentan en la
Tabla XX

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-14}Factor loads and model fit.}
\centering
\resizebox{\linewidth}{!}{
\fontsize{9}{11}\selectfont
\begin{threeparttable}
\begin{tabular}[t]{lcccccccccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{12}{c}{Factor loadings} \\
\cmidrule(l{3pt}r{3pt}){2-13}
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Version 1} & \multicolumn{4}{c}{Version 2} & \multicolumn{4}{c}{Version 3} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9} \cmidrule(l{3pt}r{3pt}){10-13}
Variables & 1 & 2 & 3 & 4 & 1 & 2 & 3 & 4 & 1 & 2 & 3 & 4\\
\midrule
Perception Effort & 0.69 &  &  &  & 0.76 &  &  &  & 0.70 &  &  & \\
Perception Talent & 0.81 &  &  &  & 0.72 &  &  &  & 0.65 &  &  & \\
Perception rich parents &  & 0.85 &  &  &  & 0.84 &  &  &  & 0.81 &  & \\
Perception contacts &  & 0.94 &  &  &  & 0.81 &  &  &  & 0.89 &  & \\
Preferences Effort &  &  & 0.85 &  &  &  & 0.82 &  &  &  & 0.66 & \\
Preferences Talent &  &  & 0.64 &  &  &  & 0.65 &  &  &  & 0.59 & \\
Preferences rich parents &  &  &  & 0.55 &  &  &  & 1.04 &  &  &  & 0.78\\
Preferences contacts &  &  &  & 1.26 &  &  &  & 0.52 &  &  &  & 0.77\\
\hline
\hspace{1em}$\chi^2\text{(df)}$ &  & 42.3(14) &  &  &  & 107.6(14) &  &  &  & 63.3(14) &  & \\
\hspace{1em}$\text{CFI}$ &  & 0.993 &  &  &  & 0.961 &  &  &  & 0.979 &  & \\
\hspace{1em}$\text{RMSEA}$ &  & 0.053 &  &  &  & 0.097 &  &  &  & 0.070 &  & \\
\hspace{1em}$N$ &  & 712 &  &  &  & 717 &  &  &  & 712 &  & \\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item \textit{Note: } 
\item Standardised factor loadings using DWLS estimator ; CFI = Comparative fit index (scaled); RMSEA = Root mean square error of approximation (scaled)
\end{tablenotes}
\end{threeparttable}}
\end{table}

Como se puede observar en la Tabla XX, todos los modelos independiente
de los órdenes obtuvieron un ajuste adecuado. con CFI superiores a .95 y
RMSEA inferiores a .8, por contraparte ni un modelo logró un chi-square
no significativo, aunque tanto el modelo aleatorio como el primero
obtuvieron un adecuado chi-square ratio menor a 3. El primero de los
órdenes fue el que obtuvo mejor ajuste (CFI=0.998, TLI=
0.995,RMSEA=0.037, \(\chi2\)(df=14)=28,03, p = 0.014), seguido por el
orden aleatorio de los ítems (CFI=0.992, TLI=0.984,RMSEA=0.051,
\(\chi2\)(df=14)=39.09, p \textless{} 0.001). Por su parte, la escala
ordenada por temáticas parece generar un efecto framing en el cual la
relación entre las percepciones y las preferencias parecen
sobreestimadas, afectando de esta manera el ajuste (CFI=0.984,
TLI=0.968,RMSEA=.071, \(\chi2\)(df=14)=64.156, p \textless{} 0.001).

Si bien todas las pruebas obtuvieron indicadores relativamente
adecuados, las diferencias mencionadas entre los ajustes de los modelos
según orden son estadísticamente significativas, lo cual fue evaluado a
partir de un análisis de invarianza. Se concluye que entre los tres
órdenes no existe invarianza configuracional, es decir, no poseen la
misma dimensionalidad y por ello no ajustan igualmente al modelo teórico
({\textbf{???}}). Esto se debe al efecto producido por la aparición
conjunta de los ítems de un mismo factor en el orden 1, lo cual aumenta
el ajuste del modelo. Además, en el orden 2, al preguntarse seguidamente
por la percepción y la preferencia en torno al mismo indicador, aumentan
las relaciones cruzadas disminuyendo el ajuste del modelo. De modo
coherente, el modelo aleatorio presentó un ajuste intermedio entre el
orden 1 y el orden 2.

\pagebreak

\begin{longtable}[]{@{}cccc@{}}
\caption{Items according to dimension.}\tabularnewline
\toprule
& Contrast Model 1 Factor & theoretical model 4 Factors & Model with
M.I.\tabularnewline
\midrule
\endfirsthead
\toprule
& Contrast Model 1 Factor & theoretical model 4 Factors & Model with
M.I.\tabularnewline
\midrule
\endhead
\textbf{n} & 1769 & 1769 & 1769\tabularnewline
\textbf{CFI} & 0.595 & 0.988 & 0.994\tabularnewline
\textbf{TLI} & 0.433 & 0.976 & 0.985\tabularnewline
\textbf{RMSEA} & 0.226 & 0.047 & 0.036\tabularnewline
\textbf{\(\chi2\)} & 1830.839 & 68.661 & 40.250\tabularnewline
\textbf{p} & .000 & .000 & .000\tabularnewline
\textbf{\(\chi2\)/df} & 65.38 & 4.90 & 3.32\tabularnewline
\bottomrule
\end{longtable}

EL modelo teórico propuesto de cuatro factores ajustó, como se observa
en la tabla 4, mejor que el modelo de contraste de 1 factor. El modelo
teórico ajustó de manera relativamente adecuada, pues muestra
indicadores óptimos para CFI= 0.987, TLI = 0.975 y RMSEA=.041, aunque
posee indicadores deficientes para la prueba \(\chi2\)(df=14)=67.6,
p-value=.000. Para evaluar posibles mejoras de la escala, se analizaron
las relaciones propuestas por los índices de modificación. Estos indican
la existencia de dos cargas cruzadas no especificadas. Cuando se generó
un modelo siguiendo estas recomendaciones, hubo una mejora considerable
del modelo, aunque el nuevo modelo tampoco obtuvo un \(\chi2\) ratio
menor a 3 y obtuvo cargas factoriales muy bajas (\(\tau\))\textless{}
0.15, por lo tanto, siguiendo las recomendación de Brown (2006) de solo
aceptar las propuesta de los índices de modificación cuando se posee
teoria y evidencia sólida, se ha decidido no incorporar estos parámetros
al modelo.

\hypertarget{conclusiuxf3n}{%
\section{Conclusión}\label{conclusiuxf3n}}

Considerando la ventaja del orden, respecto aleatorio despeja la
posibilidad del efecto framing, es decir, de que el resultado del modelo
se deba al orden de las preguntas, se ha decidido seguir utilizando el
orden 3.

uso conjunto percepciones-preferencias

\pagebreak

\hypertarget{anexos}{%
\section{Anexos}\label{anexos}}

\begin{longtable}[]{@{}lcc@{}}
\caption{Representativeness of the sample.}\tabularnewline
\toprule
& Sample & CEP\tabularnewline
\midrule
\endfirsthead
\toprule
& Sample & CEP\tabularnewline
\midrule
\endhead
\textbf{gender} & &\tabularnewline
Men & 49,82\% & 50,52\%\tabularnewline
Women & 50.18\% & 49,47\%\tabularnewline
\textbf{Age} & &\tabularnewline
18 - 24 & 18,55\% & 18,17\%\tabularnewline
25 - 34 & 18,86\% & 17,48\%\tabularnewline
35 - 44 & 19.09\% & 19,98\%\tabularnewline
45 - 54 & 17,96\% & 19,23\%\tabularnewline
55 - or more & 25,54\% & 25.11\%\tabularnewline
\textbf{Education} & &\tabularnewline
Primary or less & 2,93\% & 15,88\%\tabularnewline
Hig school & 43,23\% & 37,04\%\tabularnewline
Non university & 32,63\% & 28,93\%\tabularnewline
university or more & 21,21\% & 18,13\%\tabularnewline
\bottomrule
\end{longtable}

\hypertarget{bibliografuxeda}{%
\section*{Bibliografía}\label{bibliografuxeda}}
\addcontentsline{toc}{section}{Bibliografía}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-alesina_Fairness_2005}{}%
Alesina, A., and G. Angeletos. 2005. ``Fairness and Redistribution.''
\emph{American Economic Review}, 960--80.

\leavevmode\hypertarget{ref-arrow_meritocracy_2000}{}%
Arrow, Kenneth J., Samuel Bowles, and Steven N. Durlauf, eds. 2000.
\emph{Meritocracy and Economic Inequality}. Princeton, N.J: Princeton
University Press.

\leavevmode\hypertarget{ref-breenClassInequalityMeritocracy1999}{}%
Breen, Richard, and John H. Goldthorpe. 1999. ``Class Inequality and
Meritocracy: A Critique of Saunders and an Alternative Analysis1.''
\emph{The British Journal of Sociology} 50 (1): 1--27.
\url{https://doi.org/10.1111/j.1468-4446.1999.00001.x}.

\leavevmode\hypertarget{ref-davey_preference_1999}{}%
Davey, L. M., D. R. Bobocel, L. S. Son Hing, and M. P. Zanna. 1999.
``Preference for the Merit Principle Scale: An Individual Difference
Measure of Distributive Justice Preferences.'' \emph{Social Justice
Research} 12 (3): 223--40.

\leavevmode\hypertarget{ref-dimick_Models_2018}{}%
Dimick, Matthew, David Rueda, and Daniel Stegmueller. 2018. ``Models of
Other-Regarding Preferences, Inequality, and Redistribution.''
\emph{Annual Review of Political Science} 21 (1): 441--60.
\url{https://doi.org/10.1146/annurev-polisci-091515-030034}.

\leavevmode\hypertarget{ref-duru-bellat_whos_2012}{}%
Duru-Bellat, Marie, and Elise Tenret. 2012. ``Who's for Meritocracy?
Individual and Contextual Variations in the Faith.'' \emph{Comparative
Education Review} 56 (2): 223--47. \url{https://doi.org/10.1086/661290}.

\leavevmode\hypertarget{ref-goldthorpe_myth_2003}{}%
Goldthorpe, John. 2003. ``The Myth of Education-Based Meritocracy.''
\emph{New Economy} 10 (4): 234--39.
\url{https://doi.org/10.1046/j.1468-0041.2003.00324.x}.

\leavevmode\hypertarget{ref-hadjar_meritokratie_2008}{}%
Hadjar, Andreas. 2008. \emph{Meritokratie Als Legitimationsprinzip}.
Wiesbaden: VS Verlag.

\leavevmode\hypertarget{ref-kunovich_systems_2007}{}%
Kunovich, Sheri, and Kazimierz M. Slomczynski. 2007. ``Systems of
Distribution and a Sense of Equity: A Multilevel Analysis of
Meritocratic Attitudes in Post-Industrial Societies.'' \emph{European
Sociological Review} 23 (5): 649--63.
\url{https://doi.org/10.1093/esr/jcm026}.

\leavevmode\hypertarget{ref-landWeSatTable2006}{}%
Land, Hilary. 2006. ``We Sat down at the Table of Privilege and
Complained About the Food \textsuperscript{1}.'' \emph{The Political
Quarterly} 77 (s1): 45--60.
\url{https://doi.org/10.1111/j.1467-923X.2006.00780.x}.

\leavevmode\hypertarget{ref-MadeiraPrimesConsequencesSystematic2019}{}%
Madeira, Ana Filipa, Rui Costa-Lopes, John F. Dovidio, Gonçalo Freitas,
and Mafalda F. Mascarenhas. 2019. ``Primes and Consequences: A
Systematic Review of Meritocracy in Intergroup Relations.''
\emph{Frontiers in Psychology} 10 (September).
\url{https://doi.org/10.3389/fpsyg.2019.02007}.

\leavevmode\hypertarget{ref-markovits_Meritocracy_2019}{}%
Markovits, Daniel. 2019. \emph{The Meritocracy Trap: How America's
Foundational Myth Feeds Inequality, Dismantles the Middle Class, and
Devours the Elite}. New York: Penguin Press.

\leavevmode\hypertarget{ref-mijs_paradox_2019}{}%
Mijs, Jonathan J B. 2019. ``The Paradox of Inequality: Income Inequality
and Belief in Meritocracy Go Hand in Hand.'' \emph{Socio-Economic
Review}, January. \url{https://doi.org/10.1093/ser/mwy051}.

\leavevmode\hypertarget{ref-newman_false_2015}{}%
Newman, Benjamin J., Christopher D. Johnston, and Patrick L. Lown. 2015.
``False Consciousness or Class Awareness? Local Income Inequality,
Personal Economic Position, and Belief in American Meritocracy.''
\emph{American Journal of Political Science} 59 (2): 326--40.
\url{https://doi.org/10.1111/ajps.12153}.

\leavevmode\hypertarget{ref-reynolds_perceptions_2014}{}%
Reynolds, Jeremy, and He Xian. 2014. ``Perceptions of Meritocracy in the
Land of Opportunity.'' \emph{Research in Social Stratification and
Mobility} 36 (June): 121--37.
\url{https://doi.org/10.1016/j.rssm.2014.03.001}.

\leavevmode\hypertarget{ref-saundersMightBritainBe1995}{}%
Saunders, Peter. 1995. ``Might Britain Be a Meritocracy?''
\emph{Sociology} 29 (1): 23--41.
\url{https://doi.org/10.1177/0038038595029001003}.

\leavevmode\hypertarget{ref-schroder_Income_2017}{}%
Schröder, Martin. 2017. ``Is Income Inequality Related to Tolerance for
Inequality?'' \emph{Social Justice Research} 30 (1): 23--47.
\url{https://doi.org/10.1007/s11211-016-0276-8}.

\leavevmode\hypertarget{ref-son_hing_merit_2011-1}{}%
Son Hing, Leanne S., D. Ramona, Mark P. Zanna, Donna M. Garcia,
Stephanie S. Gee, and Katie Orazietti. 2011. ``The Merit of
Meritocracy.'' \emph{Journal of Personality and Social Psychology} 101
(3): 433--50. \url{https://doi.org/10.1037/a0024618}.

\leavevmode\hypertarget{ref-yairMeritocracy2007}{}%
Yair, Gad. 2007. ``Meritocracy.'' In \emph{The Blackwell Encyclopedia of
Sociology}, edited by George Ritzer. Oxford, UK: John Wiley \& Sons,
Ltd. \url{https://doi.org/10.1002/9781405165518.wbeosm082}.

\leavevmode\hypertarget{ref-young_rise_1962}{}%
Young, M. 1962. \emph{The Rise of the Meritocracy}. Baltimore: Penguin
Books.

\leavevmode\hypertarget{ref-youngRiseMeritocracy1994}{}%
Young, Michael Dunlop. 1994. \emph{The Rise of the Meritocracy}. New
Brunswick, N.J., U.S.A: Transaction Publishers.

\end{document}

### 2020-07-26 docs/paper.Rmd
add custom.css to docs

@@
 title: "Measuring Perceptions and Preferences for Meritocracy"
-css: "../input/css/custom.css"
+css: "custom.css"
 linestretch: '1.5'

### 2020-07-28 docs/paper.Rmd
update-double-render

@@
 output:
-  bookdown::html_document2
+  bookdown::html_document2:
+    number_sections: false
+  bookdown::pdf_document2:
+    template: null
+    toc: false 
 linkcolor: blue
@@
   # - \usepackage[spanish,es-tabla,es-nodecimaldot]{babel}
+  - \usepackage{times}           # Times New Roman
   - \usepackage{caption}
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = FALSE)
-# knitr::opts_knit$set(root.dir = "../")
+knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = FALSE,out.width = '75%',fig.pos= "H")
+# knitr::opts_knit$set(base.url = "../") #relative path for .html output file
+# knitr::opts_knit$set(root.dir = "../") #relative path for chunks within .rmd files
 options(scipen=999)
@@
 options(knitr.kable.NA = '')
+options(knitr.graphics.error = FALSE)
+Sys.setlocale("LC_ALL", "ES_ES.UTF-8")
 ```
@@
 ```{r message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
-pacman::p_load(knitr,kableExtra, dplyr, captioner)
+pacman::p_load(knitr,kableExtra, dplyr)
 ```
@@
 
-```{r include=FALSE}
-table_nums <- captioner(prefix = "Table")
-fig_nums   <- captioner(prefix = "Figure")
-anx_nums<-captioner(prefix = "Annexed")
-```
-
 <div style="text-align: justify">
 
-# Introducción
+# (PART\*) Introducción {-}
 
@@
 
-## The black-box of meritocratic beliefs
+## The black-box of meritocratic beliefs {-#blackbox}
 
@@
 
-## An instrument proposal
+## An instrument proposal {-#instrumentprop}
 
@@
 ```{r merit-model, echo=FALSE, fig.cap = "Model of perception and preferences for meritocracy and non-meritocracy", fig.align='center'}
-
-knitr::include_graphics('../input/images/generalf.png')
+knitr::include_graphics("../input/images/generalf.png")
 ```
@@
 
-## Data collection
+## Data collection {-#data}
 
@@
 In order to achieve at least some comparability with previous studies, the questions were adapted from the items battery "reasons to get ahead" (ISSP/GSS), which are mostly used for operationalizing meritocracy in previous empirical studies [@mijs_paradox_2019; @duru-bellat_whos_2012; @reynolds_perceptions_2014].
-The eight items ordered according to dimensions are presented in Table 2.
+The eight items ordered according to dimensions are presented in Table \@ref(tab:table-indicadores).
 These 8 likert-type items have 5 response alternatives ranging from "Completely disagree"(1) to "Completely agree" (5).
@@
 cap <- "Items according to dimension."
-if (knitr::is_html_output()) {tb <- table_nums("tb",cap)} else {tb <- cap}
+# if (knitr::is_html_output()) {tb <- table_nums("tb",cap)} else {tb <- cap}
 ```
 
-```{r echo=FALSE}
- tab01 <- read.csv(file = "../input/data/original/materials/table01.csv",header = 1,sep = ",")
-
+```{r table-indicadores, echo=FALSE}
+tab01 <- read.csv(file = "../input/data/original/materials/table01.csv",header = 1,sep = ",",encoding = "UTF-8")
 cnames <- c("Dimension", "Factor","Statement (english)" , "Statement (spanish)")
 
- kable(tab01,table_format,booktabs = T, linesep = "",col.names = cnames, caption = tb) %>%
-   kable_styling(full_width = T,latex_options = c("hold_position","scale_down"),position = "center",font_size = 14, bootstrap_options=c("striped", "bordered")) %>%
-   column_spec(column = 1,width = "1.5cm",) %>%
+kable(tab01,table_format,booktabs = T, linesep = "",col.names = cnames, caption = cap) %>%
+  kable_styling(
+    full_width = T,
+    latex_options = c("hold_position", "scale_down"),
+    position = "center",
+    font_size = 10,
+    bootstrap_options=c("striped", "bordered")) %>%
+  column_spec(column = 1, width = "1.5cm", ) %>%
   column_spec(column = 2,width = "2 cm") %>%
   collapse_rows(columns = 1:2,valign = "middle")
-
 ```
@@
 
-With the objective of evaluating the effect of the indicators ordering, the respondents (_n = 2141_) were randomly divided into three different order versions as explained in __Figure 1__.
+With the objective of evaluating the effect of the indicators ordering, the respondents (_n = 2141_) were randomly divided into three different order versions as explained in Figure \@ref(fig:appmod) .
 The scale was presented to the first group (_n = 712_) in the order that appears in Table 2.
@@
 
-```{r include=FALSE}
-cap <- "Survey flow"
-if (knitr::is_html_output()) {diam1 <- fig_nums("diam1",cap)} else {diam1 <- cap}
+```{r appmod, echo=FALSE, fig.cap = "Survey flow", fig.align='center'}
+knitr::include_graphics('../input/images/app_mod.png')
 ```
 
-![`r diam1`](../input/images/app_mod.png){ width=80% }
-
-## Methods
+# Methods
 
@@
 
-```{r include=FALSE}
-cap <- "Theorical model"
-if (knitr::is_html_output()) {diam2 <- fig_nums("diam2",cap)} else {diam2 <- cap}
+```{r meas01, echo=FALSE, fig.cap = "Theorical model", fig.align='center'}
+knitr::include_graphics('../output/images/meas01.png')
 ```
-
-![`r diam2`](../output/images/meas01.png)
 
@@
 
-Como puede observarse en la tabla 3, los indicadores poseen valores que van desde el 1 (totalmente en desacuerdo) al 5 (totalmente de acuerdo).
+Como puede observarse en la tabla \@ref(tab:desc01), los indicadores poseen valores que van desde el 1 (totalmente en desacuerdo) al 5 (totalmente de acuerdo).
 Se observan promedios desde 2.41 correspondiente a preferencia-contactos, hasta 3.89 correspondiente a la preferencia-esfuerzo.
@@
 
-```{r include=FALSE}
+```{r desc01, echo=FALSE}
 cap <- "Descriptive statistics of the scale."
-if (knitr::is_html_output()) {tades <- table_nums("tades",cap)} else {tades <- cap}
-```
-
-```{r echo=FALSE}
 load(file = "../input/data/proc/dat04.RData")
@@
 table04b<- sum04b %>% rename("Mean"=mean, "SD"=sd, "Min"=min,"Max"=max)
-sumdat04<- kable(table04b,format =table_format, booktabs= T, escape = FALSE,align = "l",linesep = "", caption = tades
-) %>% kable_styling(full_width = table_format2, font_size = 14,position = "center");sumdat04
+sumdat04<- kable(table04b,format =table_format, booktabs= T, escape = FALSE,align = "l",linesep = "", caption = cap
+) %>% kable_styling(full_width = table_format2, font_size = 10,position = "center",latex_options = c("hold_position"));sumdat04
 ```
@@
 El gráfico XX muestra ...
-A partir del gráfico se observa que en general hay una mayor percepción de elementos no meritocráticos que meritocráticos, mientras que en el caso de preferencias ocurre lo opuesto.
+A partir del la Figura \@ref(fig:plotlikert) se observa que en general hay una mayor percepción de elementos no meritocráticos que meritocráticos, mientras que en el caso de preferencias ocurre lo opuesto.
 En cuanto a las preferencias, llama la atención el rol preponderante del esfuerzo por sobre el talento.
 
-```{r include=FALSE}
-cap <- "Descriptive plot"
-if (knitr::is_html_output()) {liker <- fig_nums("liker",cap)} else {liker <- cap}
+```{r plotlikert, echo=FALSE, fig.cap = "Descriptive plot", fig.align='center'}
+knitr::include_graphics('../output/images/plotlikert.png')
 ```
 
-<center>
-
-![`r liker`](../output/images/plotlikert.png){ width=60% }
-
-</center>
-
-El Grafico XX muestra ....
+El Grafico \@ref(fig:corpoly) muestra ....
 Se observan relaciones de moderada a alta intensidad entre los indicadores que corresponden al mismo factor (por ejemplo, percepción de meritocracia esfuerzo con talento r=0.56), mientras que se observan correlaciones de baja intensidad entre el resto de los cruces.
@@
 
-```{r include=FALSE}
-cap <- "polychoric correlation plot"
-if (knitr::is_html_output()) {fb <- fig_nums("fb",cap)} else {fb <- cap}
+```{r corpoly, echo=FALSE, fig.cap = "Polychoric correlation plot", fig.align='center'}
+knitr::include_graphics('../output/images/corpoly.png')
 ```
-
-<center>
-
-![`r fb`](../output/images/corpoly.png){ width=60% }
-
-</center>
 
@@
 
-Los modelos confirmatorios fueron estimados en primer lugar para cada uno de los tres administration-sets. Los resultados se presentan en la Tabla XX
+> Julio: Esta sección debe ser revisada:
 
-```{r include=FALSE}
+- CFA general
+- Fit x modelo
+
+[Aquí SEM plot con cargas y ajuste global del modelo]
+
+Los modelos confirmatorios fueron estimados en primer lugar para cada uno de los tres administration-sets. Los resultados se presentan en la Tabla \@ref(tab:tab-loads)
+
+```{r tab-loads, echo=FALSE}
+load(file = "../output/tables/tb.fm.RData")
 cap <- "Factor loads and model fit."
-if (knitr::is_html_output()) {loadfit <- table_nums("loadfit",cap)} else {loadfit <- cap}
-```
-
-```{r echo=FALSE}
-
-load(file = "../output/tables/tb.fm.RData")
-
 tb.col  <- c("Variables",1:4,1:4,1:4)
 tb.caption <- "Factor loadings and fit measures"
-tb.foot <- paste0("Standardised factor loadings using DWLS estimator ; CFI = Comparative fit index (scaled); RMSEA = Root mean square error of approximation (scaled)")
+tb.foot <- paste0("Standardized factor loadings using DWLS estimator ; CFI = Comparative fit index (scaled); RMSEA = Root mean square error of approximation (scaled)")
 
@@
 
-tb.fm01<- kable(tb.fm, format = table_format, booktabs = T,linesep = "", escape = FALSE,align = "lcccccccccccc",col.names = tb.col, caption = loadfit) %>%
+tb.fm01<- kable(tb.fm, format = table_format, booktabs = T,linesep = "", escape = FALSE,align = "lcccccccccccc",col.names = tb.col, caption = cap) %>%
           kable_styling(full_width = table_format2,latex_options = c("hold_position","scale_down"), position = "center",font_size = 9) %>%
@@
           footnote(general = tb.foot,threeparttable = T);tb.fm01
-
 ```
@@
 
+> Julio: ¿Esto se debe mantener?
+
 ```{r include=FALSE}
 cap <- "Items according to dimension."
-if (knitr::is_html_output()) {te <- table_nums("te",cap)} else {te <- cap}
+# if (knitr::is_html_output()) {te <- table_nums("te",cap)} else {te <- cap}
 ```
@@
 | **$\chi2$/df** |          65.38          |             4.90            |       3.32      |
-Table:`r te`
+Table:`r cap`
 
@@
 
-# Anexos
+# References
+
+<div id="refs"> </div>  <!-- Although <div> is an HTML tag, this method also works for other output formats such as PDF. -->
+
+# Appendix {-}
 
@@
 cap <- "Representativeness of the sample."
-if (knitr::is_html_output()) {aax <- table_nums("aax",cap)} else {aax <- cap}
+# if (knitr::is_html_output()) {aax <- table_nums("aax",cap)} else {aax <- cap}
 ```
@@
 |---------------------|:-------:|:------:|
-| __gender__          |         |        |
+| __Gender__          |         |        |
 | Men                 | 49,82%  | 50,52% |
@@
 | university or more  | 21,21%  | 18,13% |
-Table:`r aax`
-
-# Bibliografía
+Table:`r cap`

### 2020-07-28 production/prod_analysis-cfa.rmd
update-double-render

@@
 output:
-  html_document:
+  bookdown::html_document2:
     number_sections: true

### 2020-07-28 render-site.R

@@ -17,7 +17,7 @@
 production<- list.files(path = "production",pattern = "prod_")        # a list with the names of the files to copy
 file.copy(file.path("production",production), "docs",overwrite = TRUE)# copy data proc and analysis files
 
-rmarkdown::render_site("docs",quiet = TRUE) # Render site
+rmarkdown::render_site("docs",quiet = FALSE) # Render site
 
 # before you run this line, check if your R Markdown files have a .rmd or .Rmd extension 
 # on this case we use both

### 2020-07-30 docs/paper.Rmd
likert, corplot, tablefit, sempath, bib, preanalisis

@@
 La desigualdad económica se ha vuelto un tema que genera creciente preocupación y malestar alrededor del mundo.
-Esto se ha expresado en una serie de protestas como la emblemática "occupy wall street" el año 2011, así como también en una serie de análisis críticos respecto del desarrollo del capitalismo y sus consecuencias [@PikettyCapitalTwentyFirstCentury2014b].
-En este contexto, el estudio de las visiones, preferencias y percepciones respecto de la desigualdad han adquirido relevancia en las ciencias sociales, en temas como las preferencias redistributivas [@alesina_Fairness_2005; @dimick_Models_2018] la legitimación de la desigualdad económica [@schroder_Income_2017]  y el funcionamiento de la meritocracia [@Duru-BellatWhoMeritocracyIndividual2012b; @MijsParadoxInequalityIncome2019b; @ReynoldsPerceptionsMeritocracyLand2014b].
+Esto se ha expresado en una serie de protestas como la emblemática "occupy wall street" el año 2011, así como también en una serie de análisis críticos respecto del desarrollo del capitalismo y sus consecuencias [@piketty_Capital_2014].
+En este contexto, el estudio de las visiones, preferencias y percepciones respecto de la desigualdad han adquirido relevancia en las ciencias sociales, en temas como las preferencias redistributivas [@alesina_Fairness_2005; @dimick_Models_2018] la legitimación de la desigualdad económica [@schroder_Income_2017]  y el funcionamiento de la meritocracia [@duru-bellat_whos_2012; @mijs_paradox_2019; @reynolds_perceptions_2014].
 
@@
 Esta concepción tradicional de mérito pone en un lugar secundario la posible interferencia de factores estructurales o no meritocráticos como la herencia, los contactos personales, o la suerte [@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995; @yairMeritocracy2007; @landWeSatTable2006; @youngRiseMeritocracy1994].
-Una serie de estudios han realizado críticas a la realización de este estándar moral de distribución, planteando que es una promesa incumplida dada la influencia preponderante de otros elementos más allá del mérito en el estatus individual [@WitteveenReconsideringmeritocraticpower2020a; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019].
-Por otro lado, desde la psicología social y la sociología se han estudiado las características y consecuencias de las creencias en la meritocracia, en general basados en la hipótesis que mayor creencia en la meritocracia lleva a una mayor legitimación de las desigualdades [@PremingerMeritocracyserviceethnocracy2020; @TrumpWhenwhyeconomic2020; @Meredithsocietydivisibleblessed2020; @hadjar_meritokratie_2008; @MadeiraPrimesConsequencesSystematic2019].
+Una serie de estudios han planteado críticas a la realización de este estándar moral de distribución, argumentando que es una promesa incumplida dada la influencia preponderante de otros elementos más allá del mérito en el estatus individual [@witteveenReconsideringMeritocraticPower2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019].
+Por otro lado, desde la psicología social y la sociología se han estudiado las características y consecuencias de las creencias en la meritocracia, en general basados en la hipótesis que mayor creencia en la meritocracia lleva a una mayor legitimación de las desigualdades [@premingerMeritocracyServiceEthnocracy2020; @trumpWhenWhyEconomic2020; @hadjar_meritokratie_2008; @MadeiraPrimesConsequencesSystematic2019].
 
-Debido al rol que cumplen las creencias meritocráticas dentro del pensamiento neoliberal [@Bay-ChengTrackingHomoOeconomicus2015], han surgido múltiples investigaciones que evalúan la relación entre creencias meritocráticas y diversos ámbitos sociales de la actualidad.
-Por ejemplo, se han desarrollado estudios que vinculan la meritocracia al reforzamiento de estereotipos socioeconómicos, de género y de etnias [@MadeiraPrimesConsequencesSystematic2019; @GirerdNeoliberalismIdeologicalBarrier2020; @PremingerMeritocracyserviceethnocracy2020], así como también líneas de investigación que evalúan el efecto de las creencias meritocráticas en el contexto educativo [@GenerettStoriesWeTell2020; @ReayPerilsPenaltiesMeritocracy2020; @OwensENGINESSOCIALMOBILITY2020] y en el contexto organizacional de las empresas [@PerezAdvancingcareersmerit2020; @AielloNewevidencefirmuniversity2019].
+Debido al rol que cumplen las creencias meritocráticas dentro del pensamiento neoliberal [@bay-chengTrackingHomoOeconomicus2015], han surgido múltiples investigaciones que evalúan la relación entre creencias meritocráticas y diversos ámbitos sociales de la actualidad.
+Por ejemplo, se han desarrollado estudios que vinculan la meritocracia al reforzamiento de estereotipos socioeconómicos, de género y de etnias [@MadeiraPrimesConsequencesSystematic2019; @girerdNeoliberalismIdeologicalBarrier2020; @premingerMeritocracyServiceEthnocracy2020], así como también líneas de investigación que evalúan el efecto de las creencias meritocráticas en el contexto educativo [@generettStoriesWeTell2020; @owensENGINESSOCIALMOBILITY2020] y en el contexto organizacional de las empresas [@perezAdvancingCareersMerit2020; @aielloNewEvidenceFirmuniversity2019].
 
@@
 The problem of whether talent is or not meritocratic is something that actually should (and can) be tested, it cannot be simply ruled out particularly in empirical research.
-This conceptual and measurement issue it is possible to find in other studies that assume that effort is the main and only aspect of meritocracy [@MaitnerEmotionalreactionsunequal2015;@MajorPerceiveddiscriminationworldview2007; @BubakPerceptionsmeritocracynote2019; @solt_economic_2016].
+This conceptual and measurement issue it is possible to find in other studies that assume that effort is the main and only aspect of meritocracy [@girerdNeoliberalismIdeologicalBarrier2020a; @bubakPerceptionsMeritocracyNote2019].
 
@@
 Which one of both is a "belief"?
-According to @janmaat_subjective_2013-1 : "Perceptions refer to subjective estimates of existing inequality (i.e.
+According to @janmaatSubjectiveInequalityReview2013: "Perceptions refer to subjective estimates of existing inequality (i.e.
 thoughts about what is).
@@
 
-Finally, most of the studies in meritocracy so far have not incorporated the issue of measurement error [@brown_confirmatory_2006, @bollen_structural_1989], using single indicators and/or simple average indexes for measuring meritocracy.
+Finally, most of the studies in meritocracy so far have not incorporated the issue of measurement error [@brownBookReviewTimothy2010, @bollenStructuralEquationsLatent1989], using single indicators and/or simple average indexes for measuring meritocracy.
 Such strategy assumes that the construct is measured perfectly by the indicators chosen, going as far as proposing that "...
@@
 
-The study was pre-registered ... (link) ***
+A pre-registration was made in the OSF platform, [Perceptions and Preferences for Meritocracy Scale
+Public registration](https://osf.io/z45y2) In it, it was indicated, before carrying out the analyzes, the four-dimensional theoretical model, the level of measurement of the variables, the statistical tests to be used with their respective evaluation parameters, along with other important aspects of the research design.
 

### 2020-07-30 docs/paper.tex
likert, corplot, tablefit, sempath, bib, preanalisis

@@
   pdftitle={Measuring Perceptions and Preferences for Meritocracy},
+  pdfauthor={Juan Carlos Castillo, Julio Iturra \& Francisco Meneses},
   colorlinks=true,
@@
 \makesavenoteenv{longtable}
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
 \setcounter{secnumdepth}{5}
+\usepackage{times}
 \usepackage{caption}
@@
 \title{Measuring Perceptions and Preferences for Meritocracy}
-\author{}
-\date{\vspace{-2.5em}}
+\author{Juan Carlos Castillo, Julio Iturra \& Francisco Meneses}
+\date{}
 
@@
 \maketitle
+\begin{abstract}
+Economic and social inequality have raised growing concerns and crises across societies.
+One of social science concepts associated to the maintenance of inequality is the belief in meritocracy, which would legitimize economic differences based on effort and talent.
+Despite its wide use, empirical research on meritocracy is something relatively novel.
+A number of studies have relied mostly in secondary data to operationalize meritocracy, with a large variation in the use and interpretation of survey items.
+Starting from a review of studies that measure meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding the conceptualization and measurement of meritocracy.
+Based on this critical analysis, we propose an items battery called \emph{Perceptions and preferences for meritocracy scale}, which is tested with confirmatory factor analysis with data from an online survey study (N=2,141).
+The results support the proposed conceptual structure which not only distinguishes between perceptions and preferences, but also between meritocratic and non-meritocratic dimensions.
+The discussion highlights the relevance of considering these different dimensions in order to advance in the study of meritocracy.
+\end{abstract}
 
 \setstretch{1.5}
-\emph{Abstract}
+\hypertarget{part-introducciuxf3n}{%
+\part*{Introducción}\label{part-introducciuxf3n}}
+\addcontentsline{toc}{part}{Introducción}
 
-Economic and social inequality have raised growing concerns and crises
-across societies. One of social science concepts associated to the
-maintenance of inequality is the belief in meritocracy, which would
-legitimize economic differences based on effort and talent. Despite its
-wide use, empirical research on meritocracy is something relatively
-novel. A number of studies have relied mostly in secondary data to
-operationalize meritocracy, with a large variation in the use and
-interpretation of survey items. Starting from a review of studies that
-measure meritocracy, this article identifies a series of drawbacks and
-inconsistencies within and between studies regarding the
-conceptualization and measurement of meritocracy. Based on this critical
-analysis, we propose an items battery called ``Perceptions and
-preferences for meritocracy scale'', which is tested with confirmatory
-factor analysis with data from an online survey study (N=2,141). The
-results support the proposed conceptual structure which not only
-distinguishes between perceptions and preferences, but also between
-meritocratic and non-meritocratic dimensions. The discussion highlights
-the relevance of considering these different dimensions in order to
-advance in the study of meritocracy.
+La desigualdad económica se ha vuelto un tema que genera creciente preocupación y malestar alrededor del mundo.
+Esto se ha expresado en una serie de protestas como la emblemática ``occupy wall street'' el año 2011, así como también en una serie de análisis críticos respecto del desarrollo del capitalismo y sus consecuencias ({\textbf{???}}).
+En este contexto, el estudio de las visiones, preferencias y percepciones respecto de la desigualdad han adquirido relevancia en las ciencias sociales, en temas como las preferencias redistributivas (Alesina and Angeletos \protect\hyperlink{ref-alesina_Fairness_2005}{2005}; Dimick, Rueda, and Stegmueller \protect\hyperlink{ref-dimick_Models_2018}{2018}) la legitimación de la desigualdad económica (Schröder \protect\hyperlink{ref-schroder_Income_2017}{2017}) y el funcionamiento de la meritocracia ({\textbf{???}}; {\textbf{???}}; {\textbf{???}}).
 
-\hypertarget{introducciuxf3n}{%
-\section{Introducción}\label{introducciuxf3n}}
-
-La desigualdad económica se ha vuelto un tema que genera creciente
-preocupación y malestar alrededor del mundo. Esto se ha expresado en una
-serie de protestas como la emblemática ``occupy wall street'' el año
-2011, así como también en una serie de análisis críticos respecto del
-desarrollo del capitalismo y sus consecuencias ({\textbf{???}}). En este
-contexto, el estudio de las visiones, preferencias y percepciones
-respecto de la desigualdad han adquirido relevancia en las ciencias
-sociales, en temas como las preferencias redistributivas (Alesina and
-Angeletos \protect\hyperlink{ref-alesina_Fairness_2005}{2005}; Dimick,
-Rueda, and Stegmueller \protect\hyperlink{ref-dimick_Models_2018}{2018})
-la legitimación de la desigualdad económica (Schröder
-\protect\hyperlink{ref-schroder_Income_2017}{2017}) y el funcionamiento
-de la meritocracia ({\textbf{???}}; {\textbf{???}}; {\textbf{???}}).
-
-En general, la meritocracia se define como un sistema de distribución de
-recursos y recompensas basados en el mérito individual, que en su
-concepción original es una suma de talento y esfuerzo (Young
-\protect\hyperlink{ref-young_rise_1962}{1962}). Esta concepción
-tradicional de mérito pone en un lugar secundario la posible
-interferencia de factores estructurales o no meritocráticos como la
-herencia, los contactos personales, o la suerte (Breen and Goldthorpe
-\protect\hyperlink{ref-breenClassInequalityMeritocracy1999}{1999};
-Saunders \protect\hyperlink{ref-saundersMightBritainBe1995}{1995}; Yair
-\protect\hyperlink{ref-yairMeritocracy2007}{2007}; Land
-\protect\hyperlink{ref-landWeSatTable2006}{2006}; Young
-\protect\hyperlink{ref-youngRiseMeritocracy1994}{1994}). Una serie de
-estudios han realizado críticas a la realización de este estándar moral
-de distribución, planteando que es una promesa incumplida dada la
-influencia preponderante de otros elementos más allá del mérito en el
-estatus individual ({\textbf{???}}; Arrow, Bowles, and Durlauf
-\protect\hyperlink{ref-arrow_meritocracy_2000}{2000}; Goldthorpe
-\protect\hyperlink{ref-goldthorpe_myth_2003}{2003}; Markovits
-\protect\hyperlink{ref-markovits_Meritocracy_2019}{2019}). Por otro
-lado, desde la psicología social y la sociología se han estudiado las
-características y consecuencias de las creencias en la meritocracia, en
-general basados en la hipótesis que mayor creencia en la meritocracia
-lleva a una mayor legitimación de las desigualdades ({\textbf{???}};
-{\textbf{???}}; {\textbf{???}}; Hadjar
-\protect\hyperlink{ref-hadjar_meritokratie_2008}{2008}; Madeira et al.
+En general, la meritocracia se define como un sistema de distribución de recursos y recompensas basados en el mérito individual, que en su concepción original es una suma de talento y esfuerzo (Young \protect\hyperlink{ref-young_rise_1962}{1962}).
+Esta concepción tradicional de mérito pone en un lugar secundario la posible interferencia de factores estructurales o no meritocráticos como la herencia, los contactos personales, o la suerte (Breen and Goldthorpe \protect\hyperlink{ref-breenClassInequalityMeritocracy1999}{1999}; Saunders \protect\hyperlink{ref-saundersMightBritainBe1995}{1995}; Yair \protect\hyperlink{ref-yairMeritocracy2007}{2007}; Land \protect\hyperlink{ref-landWeSatTable2006}{2006}; Young \protect\hyperlink{ref-youngRiseMeritocracy1994}{1994}).
+Una serie de estudios han realizado críticas a la realización de este estándar moral de distribución, planteando que es una promesa incumplida dada la influencia preponderante de otros elementos más allá del mérito en el estatus individual ({\textbf{???}}; Arrow, Bowles, and Durlauf \protect\hyperlink{ref-arrow_meritocracy_2000}{2000}; Goldthorpe \protect\hyperlink{ref-goldthorpe_myth_2003}{2003}; Markovits \protect\hyperlink{ref-markovits_Meritocracy_2019}{2019}).
+Por otro lado, desde la psicología social y la sociología se han estudiado las características y consecuencias de las creencias en la meritocracia, en general basados en la hipótesis que mayor creencia en la meritocracia lleva a una mayor legitimación de las desigualdades ({\textbf{???}}; {\textbf{???}}; {\textbf{???}}; Hadjar \protect\hyperlink{ref-hadjar_meritokratie_2008}{2008}; Madeira et al.
 \protect\hyperlink{ref-MadeiraPrimesConsequencesSystematic2019}{2019}).
 
-Debido al rol que cumplen las creencias meritocráticas dentro del
-pensamiento neoliberal ({\textbf{???}}), han surgido múltiples
-investigaciones que evalúan la relación entre creencias meritocráticas y
-diversos ámbitos sociales de la actualidad. Por ejemplo, se han
-desarrollado estudios que vinculan la meritocracia al reforzamiento de
-estereotipos socioeconómicos, de género y de etnias (Madeira et al.
-\protect\hyperlink{ref-MadeiraPrimesConsequencesSystematic2019}{2019};
-{\textbf{???}}; {\textbf{???}}), así como también líneas de
-investigación que evalúan el efecto de las creencias meritocráticas en
-el contexto educativo ({\textbf{???}}; {\textbf{???}}; {\textbf{???}}) y
-en el contexto organizacional de las empresas ({\textbf{???}};
-{\textbf{???}}).
+Debido al rol que cumplen las creencias meritocráticas dentro del pensamiento neoliberal ({\textbf{???}}), han surgido múltiples investigaciones que evalúan la relación entre creencias meritocráticas y diversos ámbitos sociales de la actualidad.
+Por ejemplo, se han desarrollado estudios que vinculan la meritocracia al reforzamiento de estereotipos socioeconómicos, de género y de etnias (Madeira et al.
+\protect\hyperlink{ref-MadeiraPrimesConsequencesSystematic2019}{2019}; {\textbf{???}}; {\textbf{???}}), así como también líneas de investigación que evalúan el efecto de las creencias meritocráticas en el contexto educativo ({\textbf{???}}; {\textbf{???}}; {\textbf{???}}) y en el contexto organizacional de las empresas ({\textbf{???}}; {\textbf{???}}).
 
-Para poder dar cuenta de los niveles de creencia en la meritocracia los
-estudios a la fecha generalmente han utilizado algunos indicadores de
-encuestas ya existentes, y en el menor de los casos se han creado
-instrumentos ad-hoc. Sin embargo, y como mostraremos más adelante, las
-formas de medición de meritocracia varían extremadamente entre estudios.
-Muchas veces fenómenos similares se asocian a indicadores distintos, y
-también ocurre que fenómenos distintos son medidos con indicadores
-similares, todo lo cual hace dificulta la comparabilidad entre estudios
-y el poder avanzar en la comprensión y estudio de la meritocracia.
+Para poder dar cuenta de los niveles de creencia en la meritocracia los estudios a la fecha generalmente han utilizado algunos indicadores de encuestas ya existentes, y en el menor de los casos se han creado instrumentos ad-hoc.
+Sin embargo, y como mostraremos más adelante, las formas de medición de meritocracia varían extremadamente entre estudios.
+Muchas veces fenómenos similares se asocian a indicadores distintos, y también ocurre que fenómenos distintos son medidos con indicadores similares, todo lo cual hace dificulta la comparabilidad entre estudios y el poder avanzar en la comprensión y estudio de la meritocracia.
 
-Basados en el análisis crítico de las formas de medición de meritocracia
-a la fecha, el presente artículo propone un instrumento para medir y
-relacionar dos aspectos claves en el estudio de la meritocracia:
-percepciones y preferencias. Además, como un segundo eje de análisis
-considera la generación de indicadores respecto de aspectos
-meritocráticos y anti-meritocráticos, demostrando que no son los dos
-polos de un mismo continuo como muchos estudios anteriores parecen
-sugerir. La propuesta de medición además está orientada a generar un
-instrumento lo más breve posible de manera que pueda ser utilizado en
-encuestas de opinión pública y así ser asociado a otros fenómenos
-sociales.
+Basados en el análisis crítico de las formas de medición de meritocracia a la fecha, el presente artículo propone un instrumento para medir y relacionar dos aspectos claves en el estudio de la meritocracia: percepciones y preferencias.
+Además, como un segundo eje de análisis considera la generación de indicadores respecto de aspectos meritocráticos y anti-meritocráticos, demostrando que no son los dos polos de un mismo continuo como muchos estudios anteriores parecen sugerir.
+La propuesta de medición además está orientada a generar un instrumento lo más breve posible de manera que pueda ser utilizado en encuestas de opinión pública y así ser asociado a otros fenómenos sociales.
 
 \hypertarget{la-mediciuxf3n-de-los-aspectos-subjetivos-de-la-meritocracia}{%
-\section{La medición de los aspectos subjetivos de la
-meritocracia}\label{la-mediciuxf3n-de-los-aspectos-subjetivos-de-la-meritocracia}}
+\section{La medición de los aspectos subjetivos de la meritocracia}\label{la-mediciuxf3n-de-los-aspectos-subjetivos-de-la-meritocracia}}
 
-A continuación se presenta una revisión de una serie de investigaciones
-que se han abocado al estudio de la meritocracia y que para ello han
-hecho una propuesta de medición. El primer eje de análisis tiene que ver
-con el uso del concepto ``creencias'' para referir a distintos aspectos
-subjetivos relacionados con meritocracia. El segundo eje tiene que ver
-con el uso de indicadores sobre aspectos anti-meritocráticos como el
-polo opuesto de los meritocráticos.
+A continuación se presenta una revisión de una serie de investigaciones que se han abocado al estudio de la meritocracia y que para ello han hecho una propuesta de medición.
+El primer eje de análisis tiene que ver con el uso del concepto ``creencias'' para referir a distintos aspectos subjetivos relacionados con meritocracia.
+El segundo eje tiene que ver con el uso de indicadores sobre aspectos anti-meritocráticos como el polo opuesto de los meritocráticos.
 
-\hypertarget{the-black-box-of-meritocratic-beliefs}{%
-\subsection{The black-box of meritocratic
-beliefs}\label{the-black-box-of-meritocratic-beliefs}}
+\hypertarget{blackbox}{%
+\subsection*{The black-box of meritocratic beliefs}\label{blackbox}}
+\addcontentsline{toc}{subsection}{The black-box of meritocratic beliefs}
 
-Several approaches to the empirical study of meritocracy based on public
-opinion surveys make reference to the concept of \emph{beliefs}, but
-behind this concept there are usually different meanings and
-operationalizations. To illustrate this point in the following we will
-start with the proposal from a recent paper by Mijs
-(\protect\hyperlink{ref-mijs_paradox_2019}{2019}), which we will take as
-a reference to discuss previous studies.
+Several approaches to the empirical study of meritocracy based on public opinion surveys make reference to the concept of \emph{beliefs}, but behind this concept there are usually different meanings and operationalizations.
+To illustrate this point in the following we will start with the proposal from a recent paper by Mijs (\protect\hyperlink{ref-mijs_paradox_2019}{2019}), which we will take as a reference to discuss previous studies.
 
-The meritocratic beliefs' definition of Mijs is the following: ``when I
-discuss meritocracy beliefs, I am referring to citizens' belief in the
-importance of hard work relative to structural factors.'' (Mijs
-\protect\hyperlink{ref-mijs_paradox_2019}{2019}, pg.9). In the
-operationalization, this is associated with the following indicator:
-``how important you think it is for getting ahead in life: (a) hard
-work'', scored in a 1 to 5 likert scale. There are several assumptions
-behind this decision that are worth discussing:
+The meritocratic beliefs' definition of Mijs is the following: ``when I discuss meritocracy beliefs, I am referring to citizens' belief in the importance of hard work relative to structural factors.'' (Mijs \protect\hyperlink{ref-mijs_paradox_2019}{2019}, pg.9).
+In the operationalization, this is associated with the following indicator: ``how important you think it is for getting ahead in life: (a) hard work'', scored in a 1 to 5 likert scale.
+There are several assumptions behind this decision that are worth discussing:
 
@@
 
-The item used by Mijs is part of an items' battery present in several
-international surveys, usually called ``reasons to get ahead''. This
-battery presents a series of indicators related to what people consider
-important to get ahead in life: hard work, education, ambition, wealthy
-family, right connections, religion, race and gender. Therefore, for
-Mijs other aspects as education, that could be associated to talent, are
-not meritocratic. As he points out: ``Hard work is arguably the most
-meritocratic part of Michael Young's equation, `Merit = Intelligence +
-Effort', for the simple fact that intelligence itself is conditioned by
-a nonmeritocratic factor: who your parents happen to be'' (p.5).
+The item used by Mijs is part of an items' battery present in several international surveys, usually called ``reasons to get ahead''.
+This battery presents a series of indicators related to what people consider important to get ahead in life: hard work, education, ambition, wealthy family, right connections, religion, race and gender.
+Therefore, for Mijs other aspects as education, that could be associated to talent, are not meritocratic.
+As he points out: ``Hard work is arguably the most meritocratic part of Michael Young's equation, `Merit = Intelligence + Effort', for the simple fact that intelligence itself is conditioned by a nonmeritocratic factor: who your parents happen to be'' (p.5).
 
-In Mijs' proposal we can observe a couple of strong assumptions: effort
-would not depend on parents influence, and talent is not meritocratic
-(contrary to Michael Youngs original conceptualization). The problem of
-whether talent is or not meritocratic is something that actually should
-(and can) be tested, it cannot be simply ruled out particularly in
-empirical research. This conceptual and measurement issue it is possible
-to find in other studies that assume that effort is the main and only
-aspect of meritocracy ({\textbf{???}}; {\textbf{???}}; {\textbf{???}};
-{\textbf{???}}).
+In Mijs' proposal we can observe a couple of strong assumptions: effort would not depend on parents influence, and talent is not meritocratic (contrary to Michael Youngs original conceptualization).
+The problem of whether talent is or not meritocratic is something that actually should (and can) be tested, it cannot be simply ruled out particularly in empirical research.
+This conceptual and measurement issue it is possible to find in other studies that assume that effort is the main and only aspect of meritocracy ({\textbf{???}}; {\textbf{???}}; {\textbf{???}}; {\textbf{???}}).
 
@@
 
-The ``resons to get ahead'' battery refers to ``how important you think
-it is'', considered by Mijs as a belief in meritocracy for the item
-regarding effort. Nevertheless, another version of this same battery in
-the survey asks about ``how important you think it \emph{should} be''.
-Which one of both is a ``belief''? According to ({\textbf{???}}) :
-``Perceptions refer to subjective estimates of existing inequality
-(i.e.~thoughts about what is). Beliefs are here defined as normative
-ideas about just inequality (i.e.~thoughts about what should
-be)''(p.359). Therefore, the referred paper would be actually using the
-term beliefs referring to perceptions. The same occurs in Reynolds and
-Xian (\protect\hyperlink{ref-reynolds_perceptions_2014}{2014}), who
-explicitly use the term beliefs to talk about perceptions, whereas other
-authors use different terms as attitudes (Kunovich and Slomczynski
-(\protect\hyperlink{ref-kunovich_systems_2007}{2007})). The first
-attempt to shed light on this confusion was made by Duru-Bellat and
-Tenret (\protect\hyperlink{ref-duru-bellat_whos_2012}{2012}), who used
-the item ``how important should the number of years spent in education
-and training be in deciding how much money people ought to earn?'' for
-``desired'' meritocracy (beliefs), whereas for ``perceived'' meritocracy
-they use two items: ``Would you say that in your country, people are
-rewarded for their efforts?'' and ``\ldots{} people are rewarded for
-their skills?''.
+The ``resons to get ahead'' battery refers to ``how important you think it is'', considered by Mijs as a belief in meritocracy for the item regarding effort.
+Nevertheless, another version of this same battery in the survey asks about ``how important you think it \emph{should} be''.
+Which one of both is a ``belief''?
+According to ({\textbf{???}}) : ``Perceptions refer to subjective estimates of existing inequality (i.e.~thoughts about what is).
+Beliefs are here defined as normative ideas about just inequality (i.e.~thoughts about what should be)''(p.359).
+Therefore, the referred paper would be actually using the term beliefs referring to perceptions.
+The same occurs in Reynolds and Xian (\protect\hyperlink{ref-reynolds_perceptions_2014}{2014}), who explicitly use the term beliefs to talk about perceptions, whereas other authors use different terms as attitudes (Kunovich and Slomczynski (\protect\hyperlink{ref-kunovich_systems_2007}{2007})).
+The first attempt to shed light on this confusion was made by Duru-Bellat and Tenret (\protect\hyperlink{ref-duru-bellat_whos_2012}{2012}), who used the item ``how important should the number of years spent in education and training be in deciding how much money people ought to earn?'' for ``desired'' meritocracy (beliefs), whereas for ``perceived'' meritocracy they use two items: ``Would you say that in your country, people are rewarded for their efforts?'' and ``\ldots{} people are rewarded for their skills?''.
 
-Is the belief in meritocracy a perception or a desire/preference? In
-order to expand the analitical conceptual framework, we believe that
-both dimensions should be included in the analysis, as proposed by
-Duru-Bellat and Tenret
-(\protect\hyperlink{ref-duru-bellat_whos_2012}{2012}). This opens up
-possibilities of analyzing whether perceptions and preferences are
-actually the same (i.e.~correlation close to 1) or they are different
-aspects of the same phenomenon. As Son Hing et al.
-(\protect\hyperlink{ref-son_hing_merit_2011-1}{2011}) have pointed out,
-``People can believe that outcomes ought to be distributed on the basis
-of merit and yet vary in their perceptions of whether this is how
-society currently operates'' (p.~435). In other words, normative beliefs
-should be interpreted in the context of perception: a large normative
-belief in meritocracy certainly means something totally different for
-someone perceiving high meritocracy than for someone perceiving low
-meritocracy. In order to avoid the confusion generated by the term
-``belief'', we propose the terms meritocratic preferences and
-meritocratic perceptions, as they better reflect the two dimension under
-scrutiny.
+Is the belief in meritocracy a perception or a desire/preference?
+In order to expand the analitical conceptual framework, we believe that both dimensions should be included in the analysis, as proposed by Duru-Bellat and Tenret (\protect\hyperlink{ref-duru-bellat_whos_2012}{2012}).
+This opens up possibilities of analyzing whether perceptions and preferences are actually the same (i.e.~correlation close to 1) or they are different aspects of the same phenomenon.
+As Son Hing et al.
+(\protect\hyperlink{ref-son_hing_merit_2011-1}{2011}) have pointed out, ``People can believe that outcomes ought to be distributed on the basis of merit and yet vary in their perceptions of whether this is how society currently operates'' (p.~435).
+In other words, normative beliefs should be interpreted in the context of perception: a large normative belief in meritocracy certainly means something totally different for someone perceiving high meritocracy than for someone perceiving low meritocracy.
+In order to avoid the confusion generated by the term ``belief'', we propose the terms meritocratic preferences and meritocratic perceptions, as they better reflect the two dimension under scrutiny.
 
@@
 
-Mijs (\protect\hyperlink{ref-mijs_paradox_2019}{2019}) makes reference
-to some non-meritocratic aspects as talent, which is ruled out of the
-operationalization of meritocracy. A different approach was followed by
-Kunovich and Slomczynski
-(\protect\hyperlink{ref-kunovich_systems_2007}{2007}), who decide to
-include some non-meritocratic elements. Using the items' battery listing
-a number of reasons about ``How important should be in deciding
-pay\ldots{}'' (as Duru-Bellat and Tenret
-(\protect\hyperlink{ref-duru-bellat_whos_2012}{2012}) for desired
-meritocracy), he decided that reasons as education and responsibility
-are meritocratic and pointed 1 if considered essential, whereas reasons
-such as having a family and children were pointed 1 if they were
-considered ``not important at all'' (i.e.~reverse coded). A similar
-approach was taken by Newman, Johnston, and Lown
-(\protect\hyperlink{ref-newman_false_2015}{2015}), reverse-coding
-non-meritocratic items, similarly to what occurs with the ``Preference
-for the Merit Principle Scale'' (Davey et al.
+Mijs (\protect\hyperlink{ref-mijs_paradox_2019}{2019}) makes reference to some non-meritocratic aspects as talent, which is ruled out of the operationalization of meritocracy.
+A different approach was followed by Kunovich and Slomczynski (\protect\hyperlink{ref-kunovich_systems_2007}{2007}), who decide to include some non-meritocratic elements.
+Using the items' battery listing a number of reasons about ``How important should be in deciding pay\ldots{}'' (as Duru-Bellat and Tenret (\protect\hyperlink{ref-duru-bellat_whos_2012}{2012}) for desired meritocracy), he decided that reasons as education and responsibility are meritocratic and pointed 1 if considered essential, whereas reasons such as having a family and children were pointed 1 if they were considered ``not important at all'' (i.e.~reverse coded).
+A similar approach was taken by Newman, Johnston, and Lown (\protect\hyperlink{ref-newman_false_2015}{2015}), reverse-coding non-meritocratic items, similarly to what occurs with the ``Preference for the Merit Principle Scale'' (Davey et al.
 \protect\hyperlink{ref-davey_preference_1999}{1999}).
 
-The assumption that meritocratic and non-meritocratic elements are the
-poles of the same continuum was analyzed by Reynolds and Xian
-(\protect\hyperlink{ref-reynolds_perceptions_2014}{2014}) using the same
-``get ahead'' perceptions' battery items. They consider education,
-ambition and hard work as meritocratic and other reasons such as wealthy
-family and right connections and non-mertitocratic. Nevertheless,
-despite making this distinction the author ends up subtracting one
-dimension from the other, assuming that they are two poles of the same
-continuum as Kunovich and Slomczynski
-(\protect\hyperlink{ref-kunovich_systems_2007}{2007}) did. Still, taking
-into account this research perspective, we suggest that non-meritocratic
-aspects should be part of a meritocratic measurement but taken
-independently and not adding or subtracting from meritocratic ones
-unless it is empirically proved that they belong to the same conceptual
-dimension.
+The assumption that meritocratic and non-meritocratic elements are the poles of the same continuum was analyzed by Reynolds and Xian (\protect\hyperlink{ref-reynolds_perceptions_2014}{2014}) using the same ``get ahead'' perceptions' battery items.
+They consider education, ambition and hard work as meritocratic and other reasons such as wealthy family and right connections and non-mertitocratic.
+Nevertheless, despite making this distinction the author ends up subtracting one dimension from the other, assuming that they are two poles of the same continuum as Kunovich and Slomczynski (\protect\hyperlink{ref-kunovich_systems_2007}{2007}) did.
+Still, taking into account this research perspective, we suggest that non-meritocratic aspects should be part of a meritocratic measurement but taken independently and not adding or subtracting from meritocratic ones unless it is empirically proved that they belong to the same conceptual dimension.
 
@@
 
-Finally, most of the studies in meritocracy so far have not incorporated
-the issue of measurement error ({\textbf{???}}), using single indicators
-and/or simple average indexes for measuring meritocracy. Such strategy
-assumes that the construct is measured perfectly by the indicators
-chosen, going as far as proposing that "\ldots{} In choosing this
-strategy of index construction, we argue that \emph{support for
-meritocracy is not a latent variable} (Kunovich and Slomczynski
-\protect\hyperlink{ref-kunovich_systems_2007}{2007}, 653--54). Some
-advances were done by Reynolds and Xian
-(\protect\hyperlink{ref-reynolds_perceptions_2014}{2014}) by doing a
-principal component analysis of meritocratic and non-meritocratic
-dimensions, but somewhat contradictorily they end up in a sum index
-despite proving multidimensionality.
+Finally, most of the studies in meritocracy so far have not incorporated the issue of measurement error ({\textbf{???}}), using single indicators and/or simple average indexes for measuring meritocracy.
+Such strategy assumes that the construct is measured perfectly by the indicators chosen, going as far as proposing that "\ldots{} In choosing this strategy of index construction, we argue that \emph{support for meritocracy is not a latent variable} (Kunovich and Slomczynski \protect\hyperlink{ref-kunovich_systems_2007}{2007}, 653--54).
+Some advances were done by Reynolds and Xian (\protect\hyperlink{ref-reynolds_perceptions_2014}{2014}) by doing a principal component analysis of meritocratic and non-meritocratic dimensions, but somewhat contradictorily they end up in a sum index despite proving multidimensionality.
 
-\hypertarget{an-instrument-proposal}{%
-\subsection{An instrument proposal}\label{an-instrument-proposal}}
+\hypertarget{instrumentprop}{%
+\subsection*{An instrument proposal}\label{instrumentprop}}
+\addcontentsline{toc}{subsection}{An instrument proposal}
 
-Based on the previous limitations and confussions in the measurement of
-meritocracy presented in the previous section, in this paper we propose
-and test an instrument with the following characteristics:
+Based on the previous limitations and confussions in the measurement of meritocracy presented in the previous section, in this paper we propose and test an instrument with the following characteristics:
 
@@
 \item
-  Multidimensional, incorporating previous distinctions between
-  preferences and perceptions as well as between meritocratic and
-  non-meritocratic aspects.
+  Multidimensional, incorporating previous distinctions between preferences and perceptions as well as between meritocratic and non-meritocratic aspects.
 \item
-  Multiple indicators for each dimension, in order to account for
-  measurement error in a confirmatory factor analysis context.
+  Multiple indicators for each dimension, in order to account for measurement error in a confirmatory factor analysis context.
 \item
-  Based on previous indicators as far as possible, for the sake of
-  comparability between studies
+  Based on previous indicators as far as possible, for the sake of comparability between studies
 \item
-  Brief, as to be used in a regular survey. In this point it differs for
-  instance from the proposal of the ``Preference for the Merit Principle
-  Scale'' (Davey et al.
-  \protect\hyperlink{ref-davey_preference_1999}{1999}), as they use 15
-  items just for one dimension (besides the problem of reverse-coding
-  non-meritocratic items).
+  Brief, as to be used in a regular survey.
+In this point it differs for instance from the proposal of the ``Preference for the Merit Principle Scale'' (Davey et al.
+\protect\hyperlink{ref-davey_preference_1999}{1999}), as they use 15 items just for one dimension (besides the problem of reverse-coding non-meritocratic items).
 \end{itemize}
 
-The proposed measurement framework is depicted in Figure 1:
+The proposed measurement framework is depicted in Figure \ref{fig:merit-model}:
 
-\includegraphics[width=0.6\textwidth,height=\textheight]{../input/images/generalf.png}
+\begin{figure}[H]
 
-The columns Perceptions and Preferences represent the distinction
-between this two concepts, usually confused under the label ``beliefs''.
-Perceptions refers to the extent to which people see that meritocracy
-functions or applies in their society, which in terms of measurement
-relates to items such as ``I think hard work is important to get ahead
-in society'', whereas preferences refer to normative expectations that
-are usually linked to a ``should'' expression (e.g.~whether hard work
-should be related to payment). The rows in the table of Figure 1
-consider the distinction between meritocratic and non-meritocratic
-dimensions (Reynolds and Xian
-\protect\hyperlink{ref-reynolds_perceptions_2014}{2014}). This aspect
-has been usually treated as different ends of a same continuum in part
-of the previous research, an assumption that requieres empirical
-scrutiny. These non-meritocratic elements usually refer to the use of
-personal contacts or family advantages to get ahead in life.
+{\centering \includegraphics[width=0.75\linewidth]{../input/images/generalf}
 
-Regarding the selection of indicators, most of them are taken or adapted
-from previous studies for the sake of comparability. For meritocratic
-indicators we use effort and talent as the main components of the
-traditional concept of merit as defined by Young
-(\protect\hyperlink{ref-young_rise_1962}{1962}), whereas for
-non-meritocratic dimensions we use having rich parents and good
-contacts. The description of the specific items is presented in the
-methodology section.
+}
+
+\caption{Model of perception and preferences for meritocracy and non-meritocracy}\label{fig:merit-model}
+\end{figure}
+
+The columns Perceptions and Preferences represent the distinction between this two concepts, usually confused under the label ``beliefs''.
+Perceptions refers to the extent to which people see that meritocracy functions or applies in their society, which in terms of measurement relates to items such as ``I think hard work is important to get ahead in society'', whereas preferences refer to normative expectations that are usually linked to a ``should'' expression (e.g.~whether hard work should be related to payment).
+The rows in the table of Figure 1 consider the distinction between meritocratic and non-meritocratic dimensions (Reynolds and Xian \protect\hyperlink{ref-reynolds_perceptions_2014}{2014}).
+This aspect has been usually treated as different ends of a same continuum in part of the previous research, an assumption that requieres empirical scrutiny.
+These non-meritocratic elements usually refer to the use of personal contacts or family advantages to get ahead in life.
+
+Regarding the selection of indicators, most of them are taken or adapted from previous studies for the sake of comparability.
+For meritocratic indicators we use effort and talent as the main components of the traditional concept of merit as defined by Young (\protect\hyperlink{ref-young_rise_1962}{1962}), whereas for non-meritocratic dimensions we use having rich parents and good contacts.
+The description of the specific items is presented in the methodology section.
 
@@
 \item
-  H1. The perception of meritocracy is a latent variable based on
-  indicators of the importance attributed to talent and the effort to
-  get ahead in life.
+  H1. The perception of meritocracy is a latent variable based on indicators of the importance attributed to talent and the effort to get ahead in life.
 \item
-  H2. The non-meritocratic perception is a latent variable that derives
-  from two indicators related to the agreement with the statement that
-  people with contacts and rich parents manage to get ahead.
+  H2. The non-meritocratic perception is a latent variable that derives from two indicators related to the agreement with the statement that people with contacts and rich parents manage to get ahead.
 \item
-  H3. Meritocratic preferences are a latent variable based on a
-  normative value of effort and talent.
+  H3. Meritocratic preferences are a latent variable based on a normative value of effort and talent.
 \item
-  H4. Non-meritocratic preferences are a latent variable based on the
-  normative value of the use of personal contacts and having rich
-  parents.
+  H4. Non-meritocratic preferences are a latent variable based on the normative value of the use of personal contacts and having rich parents.
 \end{itemize}
@@
 
-\hypertarget{data-collection}{%
-\subsection{Data collection}\label{data-collection}}
+\hypertarget{data}{%
+\subsection*{Data collection}\label{data}}
+\addcontentsline{toc}{subsection}{Data collection}
 
-The data was obtained through an online questionnaire which was part of
-a larger study on meritocracy and preferences developed in Chile in 2019
-and funded by the national scientific agency FONDECYT. The questionnaire
-was programmed in Qualtrics and the fieldwork was in charge of an
-external online survey agency (\href{www.netquest.cl}{netquest.cl})
-during December 2019 and January 2020. The sample was selected from a
-non-probabilistic design in three large cities in Chile. The quota
-method based on age, sex and educational level was used. Quotas were
-generated based on the survey of the Public Studies Center (CEP, 2019),
-which has a high prestige in the country and is also the counterpart
-agency of ISSP (International Social Survey Programme) in Chile. A total
-sample of 2,141 people was collected, excluding those who did not answer
-the questions on the scale and those who did not accept informed
-consent. As it usually occurs online samples, there were some
-limitations in achieving the quotas for lower educational levels.
+The data was obtained through an online questionnaire which was part of a larger study on meritocracy and preferences developed in Chile in 2019 and funded by the national scientific agency FONDECYT.
+The questionnaire was programmed in Qualtrics and the fieldwork was in charge of an external online survey agency (\href{www.netquest.cl}{netquest.cl}) during December 2019 and January 2020.
+The sample was selected from a non-probabilistic design in three large cities in Chile.
+The quota method based on age, sex and educational level was used.
+Quotas were generated based on the survey of the Public Studies Center (CEP, 2019), which has a high prestige in the country and is also the counterpart agency of ISSP (International Social Survey Programme) in Chile.
+A total sample of 2,141 people was collected, excluding those who did not answer the questions on the scale and those who did not accept informed consent.
+As it usually occurs online samples, there were some limitations in achieving the quotas for lower educational levels.
 
@@
 
-The proposed scale of perceptions and preferences about meritocracy
-consists of 8 indicators that are grouped into the 4 dimensions referred
-above: Perceptions (meritocratic/non-meritocratic) and preferences
-(meritocratic/non-meritocratic). In order to achieve at least some
-comparability with previous studies, the questions were adapted from the
-items battery ``reasons to get ahead'' (ISSP/GSS), which are mostly used
-for operationalizing meritocracy in previous empirical studies (Mijs
-\protect\hyperlink{ref-mijs_paradox_2019}{2019}; Duru-Bellat and Tenret
-\protect\hyperlink{ref-duru-bellat_whos_2012}{2012}; Reynolds and Xian
-\protect\hyperlink{ref-reynolds_perceptions_2014}{2014}). The eight
-items ordered according to dimensions are presented in Table 2. These 8
-likert-type items have 5 response alternatives ranging from ``Completely
-disagree''(1) to ``Completely agree'' (5).
+The proposed scale of perceptions and preferences about meritocracy consists of 8 indicators that are grouped into the 4 dimensions referred above: Perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic).
+In order to achieve at least some comparability with previous studies, the questions were adapted from the items battery ``reasons to get ahead'' (ISSP/GSS), which are mostly used for operationalizing meritocracy in previous empirical studies (Mijs \protect\hyperlink{ref-mijs_paradox_2019}{2019}; Duru-Bellat and Tenret \protect\hyperlink{ref-duru-bellat_whos_2012}{2012}; Reynolds and Xian \protect\hyperlink{ref-reynolds_perceptions_2014}{2014}).
+The eight items ordered according to dimensions are presented in Table \ref{tab:table-indicadores}.
+These 8 likert-type items have 5 response alternatives ranging from ``Completely disagree''(1) to ``Completely agree'' (5).
 
@@
 
-\caption{\label{tab:unnamed-chunk-6}Items according to dimension.}
+\caption{\label{tab:table-indicadores}Items according to dimension.}
 \centering
 \resizebox{\linewidth}{!}{
-\fontsize{14}{16}\selectfont
+\fontsize{10}{12}\selectfont
 \begin{tabu} to \linewidth {>{\raggedright\arraybackslash}p{1.5cm}>{\raggedright\arraybackslash}p{2 cm}>{\raggedright}X>{\raggedright}X}
@@
 
-With the objective of evaluating the effect of the indicators ordering,
-the respondents (\emph{n = 2141}) were randomly divided into three
-different order versions as explained in \textbf{Figure 1}. The scale
-was presented to the first group (\emph{n = 712}) in the order that
-appears in Table 2. For the second group (\emph{n = 717}), the order of
-the items was organized according to the topics of the items, e.g.~for
-the topic of hard work the item about perception was followed by the
-item about preference, and te same for the rest of the topics. Finally,
-for the third group (\emph{n = 712}) the items were completely
-randomized.
+With the objective of evaluating the effect of the indicators ordering, the respondents (\emph{n = 2141}) were randomly divided into three different order versions as explained in Figure \ref{fig:appmod} .
+The scale was presented to the first group (\emph{n = 712}) in the order that appears in Table 2.
+For the second group (\emph{n = 717}), the order of the items was organized according to the topics of the items, e.g.~for the topic of hard work the item about perception was followed by the item about preference, and te same for the rest of the topics.
+Finally, for the third group (\emph{n = 712}) the items were completely randomized.
 
-\begin{figure}
-\centering
-\includegraphics[width=0.8\textwidth,height=\textheight]{../input/images/app_mod.png}
-\caption{Survey flow}
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.75\linewidth]{../input/images/app_mod}
+
+}
+
+\caption{Survey flow}\label{fig:appmod}
 \end{figure}
@@
 \hypertarget{methods}{%
-\subsection{Methods}\label{methods}}
+\section{Methods}\label{methods}}
 
-For testing the scale's underlying constructs we estimate confirmatory
-factor analysis models (CFA). The model estimates one factor for each
-dimension, as represented in the following figure:
+For testing the scale's underlying constructs we estimate confirmatory factor analysis models (CFA). The model estimates one factor for each dimension, as represented in the following figure:
 
-\begin{figure}
-\centering
-\includegraphics{../output/images/meas01.png}
-\caption{Theorical model}
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.75\linewidth]{../output/images/meas01}
+
+}
+
+\caption{Theorical model}\label{fig:meas01}
 \end{figure}
 
-CFA was conducted using the \texttt{lavaan} package (version 0.6-3;
-Rosseel, 2020) with diagonally weighted least squares (DWLS) estimation
-due to the items' ordinal level of measurement (Kline, 2016; Rosseel,
-2020). As recommended by Brown (2008), we assessed model fit by jointly
-considering the comparative fit index and Tucker-Lewis Index (CFI and
-TLI; acceptable fit \textgreater{} 0.95), Root of the average squared
-residual approximation (RMSEA; acceptable fit \textless{} 0.08),
-Chi-square: (p-value; acceptable fit \textgreater{} 0.05, and Chi-square
-ratio:\textgreater{} 3).
+CFA was conducted using the \texttt{lavaan} package (version 0.6-3; Rosseel, 2020) with diagonally weighted least squares (DWLS) estimation due to the items' ordinal level of measurement (Kline, 2016; Rosseel, 2020).
+As recommended by Brown (2008), we assessed model fit by jointly considering the comparative fit index and Tucker-Lewis Index (CFI and TLI; acceptable fit \textgreater{} 0.95), Root of the average squared residual approximation (RMSEA; acceptable fit \textless{} 0.08), Chi-square: (p-value; acceptable fit \textgreater{} 0.05, and Chi-square ratio:\textgreater{} 3).
 
@@
 
-Como puede observarse en la tabla 3, los indicadores poseen valores que
-van desde el 1 (totalmente en desacuerdo) al 5 (totalmente de acuerdo).
-Se observan promedios desde 2.41 correspondiente a
-preferencia-contactos, hasta 3.89 correspondiente a la
-preferencia-esfuerzo. Ambos indicadores son coherentes con la adhesión
-general a la meritocracia, privilegiando aspectos individuales como el
-esfuerzo.
+Como puede observarse en la tabla \ref{tab:desc01}, los indicadores poseen valores que van desde el 1 (totalmente en desacuerdo) al 5 (totalmente de acuerdo).
+Se observan promedios desde 2.41 correspondiente a preferencia-contactos, hasta 3.89 correspondiente a la preferencia-esfuerzo.
+Ambos indicadores son coherentes con la adhesión general a la meritocracia, privilegiando aspectos individuales como el esfuerzo.
 
-\begin{table}
+\begin{table}[!h]
 
-\caption{\label{tab:unnamed-chunk-10}Descriptive statistics of the scale.}
+\caption{\label{tab:desc01}Descriptive statistics of the scale.}
 \centering
-\fontsize{9}{11}\selectfont
+\fontsize{10}{12}\selectfont
 \begin{tabular}[t]{lllll}
@@
 
-El gráfico XX muestra \ldots{} A partir del gráfico se observa que en
-general hay una mayor percepción de elementos no meritocráticos que
-meritocráticos, mientras que en el caso de preferencias ocurre lo
-opuesto. En cuanto a las preferencias, llama la atención el rol
-preponderante del esfuerzo por sobre el talento.
+El gráfico XX muestra \ldots{}
+A partir del la Figura \ref{fig:plotlikert} se observa que en general hay una mayor percepción de elementos no meritocráticos que meritocráticos, mientras que en el caso de preferencias ocurre lo opuesto.
+En cuanto a las preferencias, llama la atención el rol preponderante del esfuerzo por sobre el talento.
 
-\begin{figure}
-\centering
-\includegraphics[width=11\textwidth,height=\textheight]{../output/images/plotlikert.png}
-\caption{Descriptive plot}
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.75\linewidth]{../output/images/plotlikert}
+
+}
+
+\caption{Descriptive plot}\label{fig:plotlikert}
 \end{figure}
 
-\pagebreak
+El Grafico \ref{fig:corpoly} muestra \ldots.
+Se observan relaciones de moderada a alta intensidad entre los indicadores que corresponden al mismo factor (por ejemplo, percepción de meritocracia esfuerzo con talento r=0.56), mientras que se observan correlaciones de baja intensidad entre el resto de los cruces.
+Destacan además las relaciones entre las percepciones y preferencias de tipo meritocráticas, lo que no ocurre con los indicadores no meritocraticos.
 
-El Grafico XX muestra \ldots. Se observan relaciones de moderada a alta
-intensidad entre los indicadores que corresponden al mismo factor (por
-ejemplo, percepción de meritocracia esfuerzo con talento r=0.56),
-mientras que se observan correlaciones de baja intensidad entre el resto
-de los cruces. Destacan además las relaciones entre las percepciones y
-preferencias de tipo meritocráticas, lo que no ocurre con los
-indicadores no meritocraticos.
+\begin{figure}[H]
 
-\begin{figure}
-\centering
-\includegraphics[width=0.9\textwidth,height=\textheight]{../output/images/corpoly.png}
-\caption{polychoric correlation plot}
+{\centering \includegraphics[width=0.75\linewidth]{../output/images/corpoly}
+
+}
+
+\caption{Polychoric correlation plot}\label{fig:corpoly}
 \end{figure}
 
-En suma, los análisis descriptivos señalan una relativa adhesión a la
-moral meritocratica, las cuales, a modo general, se expresan en una
-mayor preferencia por criterios meritocraticos y una menor por no
-meritocráticos. Igualmente, es observable una relativamente baja
-percepción de meritocracia. Se observa además una relación coherente
-entre los indicadores según lo propuesto en el modelo teórico en el
-preregistro del estudio, es decir, los pares de items asociados a un
-factor específico muestran correlaciones con un tamaño de efecto grande
-(por ejemplo, preferencias meritocrácticas por los items asociados a
-esfuerzo y a talento). En particular las asociaciones entre esfuerzo y
-talento son relevantes, ya que desestiman previos supuestos sobre que el
-talento no sería un criterio meritocrático (Mijs
-\protect\hyperlink{ref-mijs_paradox_2019}{2019}), de otra manera la
-correlación sería cero o negativa. Junto a ello, vemos que no existe una
-correlacion negativa entre aspectos meritocráticos y no meritocráticos,
-desestimando los supuestos de estudios previos que señalaban que estas
-dimensiones serían los polos opuestos de un mismo continuo (Reynolds and
-Xian \protect\hyperlink{ref-reynolds_perceptions_2014}{2014}).
+En suma, los análisis descriptivos señalan una relativa adhesión a la moral meritocratica, las cuales, a modo general, se expresan en una mayor preferencia por criterios meritocraticos y una menor por no meritocráticos.
+Igualmente, es observable una relativamente baja percepción de meritocracia.
+Se observa además una relación coherente entre los indicadores según lo propuesto en el modelo teórico en el preregistro del estudio, es decir, los pares de items asociados a un factor específico muestran correlaciones con un tamaño de efecto grande (por ejemplo, preferencias meritocrácticas por los items asociados a esfuerzo y a talento).
+En particular las asociaciones entre esfuerzo y talento son relevantes, ya que desestiman previos supuestos sobre que el talento no sería un criterio meritocrático (Mijs \protect\hyperlink{ref-mijs_paradox_2019}{2019}), de otra manera la correlación sería cero o negativa.
+Junto a ello, vemos que no existe una correlacion negativa entre aspectos meritocráticos y no meritocráticos, desestimando los supuestos de estudios previos que señalaban que estas dimensiones serían los polos opuestos de un mismo continuo (Reynolds and Xian \protect\hyperlink{ref-reynolds_perceptions_2014}{2014}).
 
 \hypertarget{confirmatory-factor-analysis}{%
-\subsection{Confirmatory Factor
-Analysis}\label{confirmatory-factor-analysis}}
+\subsection{Confirmatory Factor Analysis}\label{confirmatory-factor-analysis}}
 
-This section tests the item's battery and the corresponding conceptual
-framework under test. For this, we first estimate a confirmatory factor
-analysis model for the whole sample, and secondly we test the order
-effects applying the same model to each of the three order versions.
+This section tests the item's battery and the corresponding conceptual framework under test.
+For this, we first estimate a confirmatory factor analysis model for the whole sample, and secondly we test the order effects applying the same model to each of the three order versions.
 
@@
 
-Los modelos confirmatorios fueron estimados en primer lugar para cada
-uno de los tres administration-sets. Los resultados se presentan en la
-Tabla XX
+\begin{quote}
+Julio: Esta sección debe ser revisada:
+\end{quote}
+
+\begin{itemize}
+\tightlist
+\item
+  CFA general
+\item
+  Fit x modelo
+\end{itemize}
+
+{[}Aquí SEM plot con cargas y ajuste global del modelo{]}
+
+Los modelos confirmatorios fueron estimados en primer lugar para cada uno de los tres administration-sets. Los resultados se presentan en la Tabla \ref{tab:tab-loads}
 
@@
 
-\caption{\label{tab:unnamed-chunk-14}Factor loads and model fit.}
+\caption{\label{tab:tab-loads}Factor loads and model fit.}
 \centering
@@
 \item \textit{Note: } 
-\item Standardised factor loadings using DWLS estimator ; CFI = Comparative fit index (scaled); RMSEA = Root mean square error of approximation (scaled)
+\item Standardized factor loadings using DWLS estimator ; CFI = Comparative fit index (scaled); RMSEA = Root mean square error of approximation (scaled)
 \end{tablenotes}
@@
 
-Como se puede observar en la Tabla XX, todos los modelos independiente
-de los órdenes obtuvieron un ajuste adecuado. con CFI superiores a .95 y
-RMSEA inferiores a .8, por contraparte ni un modelo logró un chi-square
-no significativo, aunque tanto el modelo aleatorio como el primero
-obtuvieron un adecuado chi-square ratio menor a 3. El primero de los
-órdenes fue el que obtuvo mejor ajuste (CFI=0.998, TLI=
-0.995,RMSEA=0.037, \(\chi2\)(df=14)=28,03, p = 0.014), seguido por el
-orden aleatorio de los ítems (CFI=0.992, TLI=0.984,RMSEA=0.051,
-\(\chi2\)(df=14)=39.09, p \textless{} 0.001). Por su parte, la escala
-ordenada por temáticas parece generar un efecto framing en el cual la
-relación entre las percepciones y las preferencias parecen
-sobreestimadas, afectando de esta manera el ajuste (CFI=0.984,
-TLI=0.968,RMSEA=.071, \(\chi2\)(df=14)=64.156, p \textless{} 0.001).
+Como se puede observar en la Tabla XX, todos los modelos independiente de los órdenes obtuvieron un ajuste adecuado.
+con CFI superiores a .95 y RMSEA inferiores a .8, por contraparte ni un modelo logró un chi-square no significativo, aunque tanto el modelo aleatorio como el primero obtuvieron un adecuado chi-square ratio menor a 3.
+El primero de los órdenes fue el que obtuvo mejor ajuste (CFI=0.998, TLI= 0.995,RMSEA=0.037, \(\chi2\)(df=14)=28,03, p = 0.014), seguido por el orden aleatorio de los ítems (CFI=0.992, TLI=0.984,RMSEA=0.051, \(\chi2\)(df=14)=39.09, p \textless{} 0.001).
+Por su parte, la escala ordenada por temáticas parece generar un efecto framing en el cual la relación entre las percepciones y las preferencias parecen sobreestimadas, afectando de esta manera el ajuste (CFI=0.984, TLI=0.968,RMSEA=.071, \(\chi2\)(df=14)=64.156, p \textless{} 0.001).
 
-Si bien todas las pruebas obtuvieron indicadores relativamente
-adecuados, las diferencias mencionadas entre los ajustes de los modelos
-según orden son estadísticamente significativas, lo cual fue evaluado a
-partir de un análisis de invarianza. Se concluye que entre los tres
-órdenes no existe invarianza configuracional, es decir, no poseen la
-misma dimensionalidad y por ello no ajustan igualmente al modelo teórico
-({\textbf{???}}). Esto se debe al efecto producido por la aparición
-conjunta de los ítems de un mismo factor en el orden 1, lo cual aumenta
-el ajuste del modelo. Además, en el orden 2, al preguntarse seguidamente
-por la percepción y la preferencia en torno al mismo indicador, aumentan
-las relaciones cruzadas disminuyendo el ajuste del modelo. De modo
-coherente, el modelo aleatorio presentó un ajuste intermedio entre el
-orden 1 y el orden 2.
+Si bien todas las pruebas obtuvieron indicadores relativamente adecuados, las diferencias mencionadas entre los ajustes de los modelos según orden son estadísticamente significativas, lo cual fue evaluado a partir de un análisis de invarianza.
+Se concluye que entre los tres órdenes no existe invarianza configuracional, es decir, no poseen la misma dimensionalidad y por ello no ajustan igualmente al modelo teórico ({\textbf{???}}).
+Esto se debe al efecto producido por la aparición conjunta de los ítems de un mismo factor en el orden 1, lo cual aumenta el ajuste del modelo.
+Además, en el orden 2, al preguntarse seguidamente por la percepción y la preferencia en torno al mismo indicador, aumentan las relaciones cruzadas disminuyendo el ajuste del modelo.
+De modo coherente, el modelo aleatorio presentó un ajuste intermedio entre el orden 1 y el orden 2.
 
 \pagebreak
+
+\begin{quote}
+Julio: ¿Esto se debe mantener?
+\end{quote}
 
@@
 \toprule
-& Contrast Model 1 Factor & theoretical model 4 Factors & Model with
-M.I.\tabularnewline
+& Contrast Model 1 Factor & theoretical model 4 Factors & Model with M.I.\tabularnewline
 \midrule
@@
 
-EL modelo teórico propuesto de cuatro factores ajustó, como se observa
-en la tabla 4, mejor que el modelo de contraste de 1 factor. El modelo
-teórico ajustó de manera relativamente adecuada, pues muestra
-indicadores óptimos para CFI= 0.987, TLI = 0.975 y RMSEA=.041, aunque
-posee indicadores deficientes para la prueba \(\chi2\)(df=14)=67.6,
-p-value=.000. Para evaluar posibles mejoras de la escala, se analizaron
-las relaciones propuestas por los índices de modificación. Estos indican
-la existencia de dos cargas cruzadas no especificadas. Cuando se generó
-un modelo siguiendo estas recomendaciones, hubo una mejora considerable
-del modelo, aunque el nuevo modelo tampoco obtuvo un \(\chi2\) ratio
-menor a 3 y obtuvo cargas factoriales muy bajas (\(\tau\))\textless{}
-0.15, por lo tanto, siguiendo las recomendación de Brown (2006) de solo
-aceptar las propuesta de los índices de modificación cuando se posee
-teoria y evidencia sólida, se ha decidido no incorporar estos parámetros
-al modelo.
+EL modelo teórico propuesto de cuatro factores ajustó, como se observa en la tabla 4, mejor que el modelo de contraste de 1 factor.
+El modelo teórico ajustó de manera relativamente adecuada, pues muestra indicadores óptimos para CFI= 0.987, TLI = 0.975 y RMSEA=.041, aunque posee indicadores deficientes para la prueba \(\chi2\)(df=14)=67.6, p-value=.000.
+Para evaluar posibles mejoras de la escala, se analizaron las relaciones propuestas por los índices de modificación.
+Estos indican la existencia de dos cargas cruzadas no especificadas.
+Cuando se generó un modelo siguiendo estas recomendaciones, hubo una mejora considerable del modelo, aunque el nuevo modelo tampoco obtuvo un \(\chi2\) ratio menor a 3 y obtuvo cargas factoriales muy bajas (\(\tau\))\textless{} 0.15, por lo tanto, siguiendo las recomendación de Brown (2006) de solo aceptar las propuesta de los índices de modificación cuando se posee teoria y evidencia sólida, se ha decidido no incorporar estos parámetros al modelo.
 
@@
 
-Considerando la ventaja del orden, respecto aleatorio despeja la
-posibilidad del efecto framing, es decir, de que el resultado del modelo
-se deba al orden de las preguntas, se ha decidido seguir utilizando el
-orden 3.
+Considerando la ventaja del orden, respecto aleatorio despeja la posibilidad del efecto framing, es decir, de que el resultado del modelo se deba al orden de las preguntas, se ha decidido seguir utilizando el orden 3.
 
@@
 
-\hypertarget{anexos}{%
-\section{Anexos}\label{anexos}}
+\hypertarget{references}{%
+\section{References}\label{references}}
+
+\hypertarget{refs}{}
+\leavevmode\hypertarget{ref-alesina_Fairness_2005}{}%
+Alesina, A., and G. Angeletos. 2005. ``Fairness and Redistribution.'' \emph{American Economic Review}, 960--80.
+
+\leavevmode\hypertarget{ref-arrow_meritocracy_2000}{}%
+Arrow, Kenneth J., Samuel Bowles, and Steven N. Durlauf, eds. 2000. \emph{Meritocracy and Economic Inequality}. Princeton, N.J: Princeton University Press.
+
+\leavevmode\hypertarget{ref-breenClassInequalityMeritocracy1999}{}%
+Breen, Richard, and John H.
+Goldthorpe.
+1999.
+``Class Inequality and Meritocracy: A Critique of Saunders and an Alternative Analysis1.'' \emph{The British Journal of Sociology} 50 (1): 1--27.
+\url{https://doi.org/10.1111/j.1468-4446.1999.00001.x}.
+
+\leavevmode\hypertarget{ref-davey_preference_1999}{}%
+Davey, L.
+M., D.
+R.
+Bobocel, L.
+S.
+Son Hing, and M.
+P.
+Zanna.
+1999.
+``Preference for the Merit Principle Scale: An Individual Difference Measure of Distributive Justice Preferences.'' \emph{Social Justice Research} 12 (3): 223--40.
+
+\leavevmode\hypertarget{ref-dimick_Models_2018}{}%
+Dimick, Matthew, David Rueda, and Daniel Stegmueller.
+2018.
+``Models of Other-Regarding Preferences, Inequality, and Redistribution.'' \emph{Annual Review of Political Science} 21 (1): 441--60.
+\url{https://doi.org/10.1146/annurev-polisci-091515-030034}.
+
+\leavevmode\hypertarget{ref-duru-bellat_whos_2012}{}%
+Duru-Bellat, Marie, and Elise Tenret.
+2012.
+``Who's for Meritocracy?
+Individual and Contextual Variations in the Faith.'' \emph{Comparative Education Review} 56 (2): 223--47.
+\url{https://doi.org/10.1086/661290}.
+
+\leavevmode\hypertarget{ref-goldthorpe_myth_2003}{}%
+Goldthorpe, John. 2003. ``The Myth of Education-Based Meritocracy.'' \emph{New Economy} 10 (4): 234--39. \url{https://doi.org/10.1046/j.1468-0041.2003.00324.x}.
+
+\leavevmode\hypertarget{ref-hadjar_meritokratie_2008}{}%
+Hadjar, Andreas. 2008. \emph{Meritokratie Als Legitimationsprinzip}. Wiesbaden: VS Verlag.
+
+\leavevmode\hypertarget{ref-kunovich_systems_2007}{}%
+Kunovich, Sheri, and Kazimierz M.
+Slomczynski.
+2007.
+``Systems of Distribution and a Sense of Equity: A Multilevel Analysis of Meritocratic Attitudes in Post-Industrial Societies.'' \emph{European Sociological Review} 23 (5): 649--63.
+\url{https://doi.org/10.1093/esr/jcm026}.
+
+\leavevmode\hypertarget{ref-landWeSatTable2006}{}%
+Land, Hilary.
+2006.
+``We Sat down at the Table of Privilege and Complained About the Food \textsuperscript{1}.'' \emph{The Political Quarterly} 77 (s1): 45--60.
+\url{https://doi.org/10.1111/j.1467-923X.2006.00780.x}.
+
+\leavevmode\hypertarget{ref-MadeiraPrimesConsequencesSystematic2019}{}%
+Madeira, Ana Filipa, Rui Costa-Lopes, John F.
+Dovidio, Gonçalo Freitas, and Mafalda F.
+Mascarenhas.
+2019.
+``Primes and Consequences: A Systematic Review of Meritocracy in Intergroup Relations.'' \emph{Frontiers in Psychology} 10 (September).
+\url{https://doi.org/10.3389/fpsyg.2019.02007}.
+
+\leavevmode\hypertarget{ref-markovits_Meritocracy_2019}{}%
+Markovits, Daniel. 2019. \emph{The Meritocracy Trap: How America's Foundational Myth Feeds Inequality, Dismantles the Middle Class, and Devours the Elite}. New York: Penguin Press.
+
+\leavevmode\hypertarget{ref-mijs_paradox_2019}{}%
+Mijs, Jonathan J B. 2019. ``The Paradox of Inequality: Income Inequality and Belief in Meritocracy Go Hand in Hand.'' \emph{Socio-Economic Review}, January. \url{https://doi.org/10.1093/ser/mwy051}.
+
+\leavevmode\hypertarget{ref-newman_false_2015}{}%
+Newman, Benjamin J., Christopher D.
+Johnston, and Patrick L.
+Lown.
+2015.
+``False Consciousness or Class Awareness?
+Local Income Inequality, Personal Economic Position, and Belief in American Meritocracy.'' \emph{American Journal of Political Science} 59 (2): 326--40.
+\url{https://doi.org/10.1111/ajps.12153}.
+
+\leavevmode\hypertarget{ref-reynolds_perceptions_2014}{}%
+Reynolds, Jeremy, and He Xian.
+2014.
+``Perceptions of Meritocracy in the Land of Opportunity.'' \emph{Research in Social Stratification and Mobility} 36 (June): 121--37.
+\url{https://doi.org/10.1016/j.rssm.2014.03.001}.
+
+\leavevmode\hypertarget{ref-saundersMightBritainBe1995}{}%
+Saunders, Peter. 1995. ``Might Britain Be a Meritocracy?'' \emph{Sociology} 29 (1): 23--41. \url{https://doi.org/10.1177/0038038595029001003}.
+
+\leavevmode\hypertarget{ref-schroder_Income_2017}{}%
+Schröder, Martin. 2017. ``Is Income Inequality Related to Tolerance for Inequality?'' \emph{Social Justice Research} 30 (1): 23--47. \url{https://doi.org/10.1007/s11211-016-0276-8}.
+
+\leavevmode\hypertarget{ref-son_hing_merit_2011-1}{}%
+Son Hing, Leanne S., D.
+Ramona, Mark P.
+Zanna, Donna M.
+Garcia, Stephanie S.
+Gee, and Katie Orazietti.
+2011.
+``The Merit of Meritocracy.'' \emph{Journal of Personality and Social Psychology} 101 (3): 433--50.
+\url{https://doi.org/10.1037/a0024618}.
+
+\leavevmode\hypertarget{ref-yairMeritocracy2007}{}%
+Yair, Gad.
+2007.
+``Meritocracy.'' In \emph{The Blackwell Encyclopedia of Sociology}, edited by George Ritzer.
+Oxford, UK: John Wiley \& Sons, Ltd.
+\url{https://doi.org/10.1002/9781405165518.wbeosm082}.
+
+\leavevmode\hypertarget{ref-young_rise_1962}{}%
+Young, M. 1962. \emph{The Rise of the Meritocracy}. Baltimore: Penguin Books.
+
+\leavevmode\hypertarget{ref-youngRiseMeritocracy1994}{}%
+Young, Michael Dunlop. 1994. \emph{The Rise of the Meritocracy}. New Brunswick, N.J., U.S.A: Transaction Publishers.
+
+\hypertarget{appendix}{%
+\section*{Appendix}\label{appendix}}
+\addcontentsline{toc}{section}{Appendix}
 
@@
 \endhead
-\textbf{gender} & &\tabularnewline
+\textbf{Gender} & &\tabularnewline
 Men & 49,82\% & 50,52\%\tabularnewline
@@
 
-\hypertarget{bibliografuxeda}{%
-\section*{Bibliografía}\label{bibliografuxeda}}
-\addcontentsline{toc}{section}{Bibliografía}
-
-\hypertarget{refs}{}
-\leavevmode\hypertarget{ref-alesina_Fairness_2005}{}%
-Alesina, A., and G. Angeletos. 2005. ``Fairness and Redistribution.''
-\emph{American Economic Review}, 960--80.
-
-\leavevmode\hypertarget{ref-arrow_meritocracy_2000}{}%
-Arrow, Kenneth J., Samuel Bowles, and Steven N. Durlauf, eds. 2000.
-\emph{Meritocracy and Economic Inequality}. Princeton, N.J: Princeton
-University Press.
-
-\leavevmode\hypertarget{ref-breenClassInequalityMeritocracy1999}{}%
-Breen, Richard, and John H. Goldthorpe. 1999. ``Class Inequality and
-Meritocracy: A Critique of Saunders and an Alternative Analysis1.''
-\emph{The British Journal of Sociology} 50 (1): 1--27.
-\url{https://doi.org/10.1111/j.1468-4446.1999.00001.x}.
-
-\leavevmode\hypertarget{ref-davey_preference_1999}{}%
-Davey, L. M., D. R. Bobocel, L. S. Son Hing, and M. P. Zanna. 1999.
-``Preference for the Merit Principle Scale: An Individual Difference
-Measure of Distributive Justice Preferences.'' \emph{Social Justice
-Research} 12 (3): 223--40.
-
-\leavevmode\hypertarget{ref-dimick_Models_2018}{}%
-Dimick, Matthew, David Rueda, and Daniel Stegmueller. 2018. ``Models of
-Other-Regarding Preferences, Inequality, and Redistribution.''
-\emph{Annual Review of Political Science} 21 (1): 441--60.
-\url{https://doi.org/10.1146/annurev-polisci-091515-030034}.
-
-\leavevmode\hypertarget{ref-duru-bellat_whos_2012}{}%
-Duru-Bellat, Marie, and Elise Tenret. 2012. ``Who's for Meritocracy?
-Individual and Contextual Variations in the Faith.'' \emph{Comparative
-Education Review} 56 (2): 223--47. \url{https://doi.org/10.1086/661290}.
-
-\leavevmode\hypertarget{ref-goldthorpe_myth_2003}{}%
-Goldthorpe, John. 2003. ``The Myth of Education-Based Meritocracy.''
-\emph{New Economy} 10 (4): 234--39.
-\url{https://doi.org/10.1046/j.1468-0041.2003.00324.x}.
-
-\leavevmode\hypertarget{ref-hadjar_meritokratie_2008}{}%
-Hadjar, Andreas. 2008. \emph{Meritokratie Als Legitimationsprinzip}.
-Wiesbaden: VS Verlag.
-
-\leavevmode\hypertarget{ref-kunovich_systems_2007}{}%
-Kunovich, Sheri, and Kazimierz M. Slomczynski. 2007. ``Systems of
-Distribution and a Sense of Equity: A Multilevel Analysis of
-Meritocratic Attitudes in Post-Industrial Societies.'' \emph{European
-Sociological Review} 23 (5): 649--63.
-\url{https://doi.org/10.1093/esr/jcm026}.
-
-\leavevmode\hypertarget{ref-landWeSatTable2006}{}%
-Land, Hilary. 2006. ``We Sat down at the Table of Privilege and
-Complained About the Food \textsuperscript{1}.'' \emph{The Political
-Quarterly} 77 (s1): 45--60.
-\url{https://doi.org/10.1111/j.1467-923X.2006.00780.x}.
-
-\leavevmode\hypertarget{ref-MadeiraPrimesConsequencesSystematic2019}{}%
-Madeira, Ana Filipa, Rui Costa-Lopes, John F. Dovidio, Gonçalo Freitas,
-and Mafalda F. Mascarenhas. 2019. ``Primes and Consequences: A
-Systematic Review of Meritocracy in Intergroup Relations.''
-\emph{Frontiers in Psychology} 10 (September).
-\url{https://doi.org/10.3389/fpsyg.2019.02007}.
-
-\leavevmode\hypertarget{ref-markovits_Meritocracy_2019}{}%
-Markovits, Daniel. 2019. \emph{The Meritocracy Trap: How America's
-Foundational Myth Feeds Inequality, Dismantles the Middle Class, and
-Devours the Elite}. New York: Penguin Press.
-
-\leavevmode\hypertarget{ref-mijs_paradox_2019}{}%
-Mijs, Jonathan J B. 2019. ``The Paradox of Inequality: Income Inequality
-and Belief in Meritocracy Go Hand in Hand.'' \emph{Socio-Economic
-Review}, January. \url{https://doi.org/10.1093/ser/mwy051}.
-
-\leavevmode\hypertarget{ref-newman_false_2015}{}%
-Newman, Benjamin J., Christopher D. Johnston, and Patrick L. Lown. 2015.
-``False Consciousness or Class Awareness? Local Income Inequality,
-Personal Economic Position, and Belief in American Meritocracy.''
-\emph{American Journal of Political Science} 59 (2): 326--40.
-\url{https://doi.org/10.1111/ajps.12153}.
-
-\leavevmode\hypertarget{ref-reynolds_perceptions_2014}{}%
-Reynolds, Jeremy, and He Xian. 2014. ``Perceptions of Meritocracy in the
-Land of Opportunity.'' \emph{Research in Social Stratification and
-Mobility} 36 (June): 121--37.
-\url{https://doi.org/10.1016/j.rssm.2014.03.001}.
-
-\leavevmode\hypertarget{ref-saundersMightBritainBe1995}{}%
-Saunders, Peter. 1995. ``Might Britain Be a Meritocracy?''
-\emph{Sociology} 29 (1): 23--41.
-\url{https://doi.org/10.1177/0038038595029001003}.
-
-\leavevmode\hypertarget{ref-schroder_Income_2017}{}%
-Schröder, Martin. 2017. ``Is Income Inequality Related to Tolerance for
-Inequality?'' \emph{Social Justice Research} 30 (1): 23--47.
-\url{https://doi.org/10.1007/s11211-016-0276-8}.
-
-\leavevmode\hypertarget{ref-son_hing_merit_2011-1}{}%
-Son Hing, Leanne S., D. Ramona, Mark P. Zanna, Donna M. Garcia,
-Stephanie S. Gee, and Katie Orazietti. 2011. ``The Merit of
-Meritocracy.'' \emph{Journal of Personality and Social Psychology} 101
-(3): 433--50. \url{https://doi.org/10.1037/a0024618}.
-
-\leavevmode\hypertarget{ref-yairMeritocracy2007}{}%
-Yair, Gad. 2007. ``Meritocracy.'' In \emph{The Blackwell Encyclopedia of
-Sociology}, edited by George Ritzer. Oxford, UK: John Wiley \& Sons,
-Ltd. \url{https://doi.org/10.1002/9781405165518.wbeosm082}.
-
-\leavevmode\hypertarget{ref-young_rise_1962}{}%
-Young, M. 1962. \emph{The Rise of the Meritocracy}. Baltimore: Penguin
-Books.
-
-\leavevmode\hypertarget{ref-youngRiseMeritocracy1994}{}%
-Young, Michael Dunlop. 1994. \emph{The Rise of the Meritocracy}. New
-Brunswick, N.J., U.S.A: Transaction Publishers.
-
 \end{document}

### 2020-07-30 production/prod_analysis-cfa.rmd
likert, corplot, tablefit, sempath, bib, preanalisis

@@
 table01<- bind_cols(x = matpolyw01,y = load01)
-# %>% mutate(vars=labs)
+### %>% mutate(vars=labs)
 
@@
 table02<- bind_cols(x = matpolyw02,y = load02)
-# %>% mutate(vars=labs)
+## %>% mutate(vars=labs)
 
@@
 table03<- bind_cols(x = matpolyw03,y = load03)
-# %>% mutate(vars=labs)
+   ## %>% mutate(vars=labs)
 
@@
 table03tau<- bind_cols(x = matpolyw03tau,y = load03tau) 
-# %>% mutate(vars=labs)
+### %>% mutate(vars=labs)
 
@@
 save(cor04, file = "output/images/correlation.RData")
-corrplot(cor04$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
+corrplot(cor04$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = TRUE)
 
@@
 
-corrplot(cor01$rho,method = "color" ,type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
+corrplot(cor01$rho,method = "color" ,type = "upper", tl.col = "black", addCoef.col = "black",diag = TRUE)
 
@@
       format = "markdown",digits = 2,col.names = cnames, caption = "Factor loadings")
+```
+
+```{r}
+
+nodeNames <-c("Perception effort","Perception talent","Perception rich family","Perception contacts",
+              "Preference effort","Preference talent","Preference rich family","Preference contacts",
+              "Perception meritocratic","Perception non meritocratic","Preference meritocratic","Preference non meritocratic")
+
+semPaths(fit4_o , 
+whatLabels = "std",
+what="col",
+label.cex=1,
+edge.label.cex=1,
+residuals=F, 
+edge.color = "black",
+style = "lisrel",
+fade=F, 
+nCharNodes = 0,
+ curvePivot = FALSE,
+         curve = 2,
+rotation = 4,
+layout = "tree3",
+         cardinal = "lat cov",  
+legend.cex = 0.6,
+         label.cex =1,
+label.font = 6,
+         edge.label.font= 1,
+         asize = 3,
+edge.width =1,
+         sizeMan=5,
+         sizeLat = 10, width = 12,
+         height= 15, nodeNames = nodeNames, intercepts = F, reorder = F, thresholds = F, nodeLabels = c("A","B","C","D","E","F","G","H","1","2","3","4"), filetype ="png", filename = "output/images/meas01") #arrow size)
+
 ```
@@
 
+```{r}
+sum_fit<- bind_rows(fitmeasures(fit4_o)[c("chisq.scaled","df","cfi.scaled","rmsea","rmsea.scaled")],
+                    fitmeasures(fit1_o)[c("chisq.scaled","df","cfi.scaled","rmsea","rmsea.scaled")],
+                    fitmeasures(fit2_o)[c("chisq.scaled","df","cfi.scaled","rmsea","rmsea.scaled")],
+                    fitmeasures(fit3_o)[c("chisq.scaled","df","cfi.scaled","rmsea","rmsea.scaled")])
+sum_fit$mod <- c("Model 1","Model 2","Model 3","Model 4")
+sum_fit$est <- c("DWLS","DWLS","DWLS","DWLS")
+sum_fit <- select(sum_fit,mod,est,everything())
+colnames <- c("Model","Estimator","$\\chi^2$","df","CFI.sca","RMSEA","RMSEA.sca")
+
+sumtable02<- kable(sum_fit,digits = 3,format = "html",row.names = F,booktabs=T, caption = "Summary fit indices",col.names = colnames,escape = FALSE) %>%
+  kable_styling(full_width = F)  %>%
+      collapse_rows(columns = 1,valign = "middle")  %>%
+      footnote(number = c("Model 1: Complete sample",
+                          "Model 2: fixed order by percepction/preference",
+                          "Model 3:  fixed order by topic (i.e: effort)",
+                          "Model 4: Randomized order"));sumtable02
+```
+
 ## Other analyses
@@
 
-plotlikert<-plot_likert(dat_merit, c(1,1,1,1,2,2,2,2),
+dat_merit_2<- dat_merit %>% sjmisc::rec(rec="rev") %>% select("a.Effort"=a.Effort_r,"a.Talent"=a.Talent_r, "a.Rich Parents"="a.Rich Parents_r", "a.Contacts"=a.Contacts_r, "b.Effort"=b.Effort_r,"b.Talent"=b.Talent_r,"b.Rich Parents"="b.Rich Parents_r", "b.Contacts"=b.Contacts_r)
+
+plotlikert<-plot_likert(dat_merit_2, c(1,1,1,1,2,2,2,2),
             groups.titles = c("Perceptions","Preferences"),
@@
             grid.range  =  c ( 1.2 , 1.4 ),
-            values  =  "sum.outside",
-            reverse.scale=TRUE)
+            values  =  "sum.outside", reverse.colors = T)
 plotlikert
 
-ggsave(plotlikert,filename = "output/images/plotlikert.png",device = "png",width = 18,height = 15,dpi = "retina",units = "cm")
+ggsave(plotlikert,filename = "output/images/plotlikert.png",device = "png",width = 25,height = 15,dpi = "retina",units = "cm")
 

### 2020-07-30 docs/paper.Rmd
seccion cfa

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = FALSE,out.width = '75%',fig.pos= "H")
+knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = TRUE,out.width = '75%',fig.pos= "H")
 # knitr::opts_knit$set(base.url = "../") #relative path for .html output file
@@
 
-Figure X shows the results of the estimation for the four-factor model:
+In \@ref(fig:meas02) the results of the estimation for the four-factor model with complete sample.
+This model have a good fit (CFI.sca = 0.977, RMSEA.sca = 0.079, $\chi2$(df = 14) = 28,03, p = 0.014) and it presents factorial loads with weights that vary between 0.63 and 0.88.
+Furthermore, considering the correlation between the latent variables, it is observed that the most related variables are meritocratic perceptions and preferences, which have a positive correlation of 0.59, while non-meritocratic perceptions and preferences do not present a considerable correlation.
 
-> Julio: Esta sección debe ser revisada:
-
-- CFA general
-- Fit x modelo
-
-[Aquí SEM plot con cargas y ajuste global del modelo]
-
-Los modelos confirmatorios fueron estimados en primer lugar para cada uno de los tres administration-sets. Los resultados se presentan en la Tabla \@ref(tab:tab-loads)
-
-```{r tab-loads, echo=FALSE}
-load(file = "../output/tables/tb.fm.RData")
-cap <- "Factor loads and model fit."
-tb.col  <- c("Variables",1:4,1:4,1:4)
-tb.caption <- "Factor loadings and fit measures"
-tb.foot <- paste0("Standardized factor loadings using DWLS estimator ; CFI = Comparative fit index (scaled); RMSEA = Root mean square error of approximation (scaled)")
-
-tb.fm <- tb.fm %>%
-    mutate(vars = recode(vars,
-                         "A. Who the more they try they manage to get bigger rewards that those who striveless."= "Perception Effort",
-                         "B. Who possess more talent they manage to obtain greater rewards than those who possess less talent."= "Perception Talent",
-                         "C. Who they have rich parents manage to get out ahead."="Perception rich parents",
-                         "D. Who they have good contacts they manage to get out ahead."="Perception contacts",
-                         "E. Who the more they try they should get greater rewards than those who they try less." = "Preferences Effort",
-                         "F. Who possess more talent they should get greater rewards than those who possess less talent."= "Preferences Talent",
-                         "G. It's fine that those who have rich parents get ahead" ="Preferences rich parents",
-                         "H. Is well that those who have good contacts get ahead." = "Preferences contacts"))
-
-tb.fm01<- kable(tb.fm, format = table_format, booktabs = T,linesep = "", escape = FALSE,align = "lcccccccccccc",col.names = tb.col, caption = cap) %>%
-          kable_styling(full_width = table_format2,latex_options = c("hold_position","scale_down"), position = "center",font_size = 9) %>%
-          add_header_above(header = c(" "=1,"Version 1"= 4,"Version 2"= 4,"Version 3"= 4)) %>%
-          add_header_above(header = c(" "=1,"Factor loadings"= 12)) %>%
-          row_spec(row = 8,hline_after = TRUE) %>%
-          add_indent(c(9:12))  %>%
-          footnote(general = tb.foot,threeparttable = T);tb.fm01
+```{r meas02, echo=FALSE, fig.cap = "Theorical model", fig.align='center'}
+knitr::include_graphics('../output/images/meas02.png')
 ```
 
-Como se puede observar en la Tabla XX, todos los modelos independiente de los órdenes obtuvieron un ajuste adecuado.
+Como se puede observar en la Tabla \@ref(tab:tab-loads), todos los modelos, independiente del grupo y el órden de los items, obtuvieron un ajuste adecuado.
 con CFI superiores a .95 y RMSEA inferiores a .8, por contraparte ni un modelo logró un chi-square no significativo, aunque tanto el modelo aleatorio como el primero obtuvieron un adecuado chi-square ratio menor a 3.
 El primero de los órdenes fue el que obtuvo mejor ajuste (CFI=0.998, TLI= 0.995,RMSEA=0.037, $\chi2$(df=14)=28,03, p = 0.014), seguido por el orden aleatorio de los ítems (CFI=0.992, TLI=0.984,RMSEA=0.051, $\chi2$(df=14)=39.09, p < 0.001).
-Por su parte, la escala ordenada por temáticas parece generar un efecto framing en el cual la relación entre las percepciones y las preferencias parecen sobreestimadas, afectando de esta manera el ajuste (CFI=0.984, TLI=0.968,RMSEA=.071, $\chi2$(df=14)=64.156, p < 0.001).
+Por su parte, la escala ordenada por temáticas parece generar un efecto framing en el cual la relación entre las percepciones y las preferencias parecen sobreestimadas, afectando de esta manera el ajuste (CFI=0.984, TLI=0.968, RMSEA=.071, $\chi2$(df=14)=64.156, p < 0.001).
+
+```{r tab03, echo=FALSE}
+library(lavaan)
+load(file = "../output/tables/fit4_o.RData")
+load(file = "../output/tables/fit3_o.RData")
+load(file = "../output/tables/fit2_o.RData")
+load(file = "../output/tables/fit1_o.RData")
+
+sum_fit<- bind_rows(fitmeasures(fit4_o)[c("chisq.scaled","df","cfi.scaled","rmsea","rmsea.scaled")],
+                    fitmeasures(fit1_o)[c("chisq.scaled","df","cfi.scaled","rmsea","rmsea.scaled")],
+                    fitmeasures(fit2_o)[c("chisq.scaled","df","cfi.scaled","rmsea","rmsea.scaled")],
+                    fitmeasures(fit3_o)[c("chisq.scaled","df","cfi.scaled","rmsea","rmsea.scaled")])
+sum_fit$mod <- c("Model 1","Model 2","Model 3","Model 4")
+sum_fit$est <- c("DWLS","DWLS","DWLS","DWLS")
+sum_fit <- select(sum_fit,mod,est,everything())
+colnames <- c("Model","Estimator","$\\chi^2$","df","CFI.sca","RMSEA","RMSEA.sca")
+
+sumtable02<- kable(sum_fit,digits = 3,format = "html",row.names = F,booktabs=T, caption = "Summary fit indices according group",col.names = colnames,escape = FALSE) %>%
+  kable_styling(full_width = F)  %>%
+      collapse_rows(columns = 1,valign = "middle")  %>%
+      footnote(number = c("Model 1: Complete sample",
+                          "Model 2: fixed order by percepction/preference",
+                          "Model 3:  fixed order by topic (i.e: effort)",
+                          "Model 4: Randomized order"));sumtable02
+```
 
@@
 
+```{r tab-loads, echo=FALSE}
+load(file = "../output/tables/tb.fm.RData")
+cap <- "Factor loads and model fit."
+tb.col  <- c("Variables",1:4,1:4,1:4)
+tb.caption <- "Factor loadings and fit measures"
+tb.foot <- paste0("Standardized factor loadings using DWLS estimator ; CFI = Comparative fit index (scaled); RMSEA = Root mean square error of approximation (scaled)")
+
+tb.fm <- tb.fm %>%
+    mutate(vars = recode(vars,
+                         "A. Who the more they try they manage to get bigger rewards that those who striveless."= "Perception Effort",
+                         "B. Who possess more talent they manage to obtain greater rewards than those who possess less talent."= "Perception Talent",
+                         "C. Who they have rich parents manage to get out ahead."="Perception rich parents",
+                         "D. Who they have good contacts they manage to get out ahead."="Perception contacts",
+                         "E. Who the more they try they should get greater rewards than those who they try less." = "Preferences Effort",
+                         "F. Who possess more talent they should get greater rewards than those who possess less talent."= "Preferences Talent",
+                         "G. It's fine that those who have rich parents get ahead" ="Preferences rich parents",
+                         "H. Is well that those who have good contacts get ahead." = "Preferences contacts"))
+
+tb.fm01<- kable(tb.fm, format = table_format, booktabs = T,linesep = "", escape = FALSE,align = "lcccccccccccc",col.names = tb.col, caption = cap) %>%
+          kable_styling(full_width = table_format2,latex_options = c("hold_position","scale_down"), position = "center",font_size = 9) %>%
+          add_header_above(header = c(" "=1,"Version 1"= 4,"Version 2"= 4,"Version 3"= 4)) %>%
+          add_header_above(header = c(" "=1,"Factor loadings"= 12)) %>%
+          row_spec(row = 8,hline_after = TRUE) %>%
+          add_indent(c(9:12))  %>%
+          footnote(general = tb.foot,threeparttable = T);tb.fm01
+```
+
 ```{r include=FALSE}

### 2020-07-30 production/prod_analysis-cfa.rmd
seccion cfa

@@
          sizeLat = 10, width = 12,
-         height= 15, nodeNames = nodeNames, intercepts = F, reorder = F, thresholds = F, nodeLabels = c("A","B","C","D","E","F","G","H","1","2","3","4"), filetype ="png", filename = "output/images/meas01") #arrow size)
+         height= 15, nodeNames = nodeNames, intercepts = F, reorder = F, thresholds = F, nodeLabels = c("A","B","C","D","E","F","G","H","1","2","3","4"), filetype ="png", filename = "output/images/meas02") #arrow size)
 
@@
 ```{r}
+
+save(fit4_o,file = "output/tables/fit4_o.RData")
+save(fit3_o,file = "output/tables/fit3_o.RData")
+save(fit2_o,file = "output/tables/fit2_o.RData")
+save(fit1_o,file = "output/tables/fit1_o.RData")
+
 sum_fit<- bind_rows(fitmeasures(fit4_o)[c("chisq.scaled","df","cfi.scaled","rmsea","rmsea.scaled")],

### 2020-07-31 docs/paper.Rmd
update-julio
correción de tablas, títulos y figuras.

--- 2020-07-30 docs/paper.Rmd
@@
 output:
-  bookdown::html_document2:
-    number_sections: false
   bookdown::pdf_document2:
@@
     toc: false 
+  bookdown::html_document2:
+    number_sections: false
 linkcolor: blue
@@
 
-```{r include=FALSE}
+```{r eval=FALSE, include=FALSE}
 # for render in pdf run rmarkdown::render_site("docs/paper.Rmd", output_format = "all")
 # clean #in the yml
+rmarkdown::render("docs/paper.Rmd", output_format = "bookdown::pdf_document2")
 ```
@@
 ```{r message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
-pacman::p_load(knitr,kableExtra, dplyr)
+pacman::p_load(knitr, kableExtra, dplyr, lavaan)
 ```
@@
 
-<div style="text-align: justify">
-
-# (PART\*) Introducción {-}
+# Introducción
 
@@
 
-* H1. The perception of meritocracy is a latent variable based on indicators of the importance attributed to talent and the effort to get ahead in life.
+* $H_1$. The perception of meritocracy is a latent variable based on indicators of the importance attributed to talent and the effort to get ahead in life.
 
-* H2. The non-meritocratic perception is a latent variable that derives from two indicators related to the agreement with the statement that people with contacts and rich parents manage to get ahead.
+* $H_2$.
+The non-meritocratic perception is a latent variable that derives from two indicators related to the agreement with the statement that people with contacts and rich parents manage to get ahead.
 
-* H3. Meritocratic preferences are a latent variable based on a normative value of effort and talent.
+* $H_3$. Meritocratic preferences are a latent variable based on a normative value of effort and talent.
 
-* H4. Non-meritocratic preferences are a latent variable based on the normative value of the use of personal contacts and having rich parents.
+* $H_4$. Non-meritocratic preferences are a latent variable based on the normative value of the use of personal contacts and having rich parents.
 
@@
 load(file = "../input/data/proc/dat04.RData")
-dat04b<- dat04 %>% rename("A.Perception Effort" =perc_effort,
-                          "B.Perception Talent" = perc_talent,
-                          "C.Perception rich parents" = perc_wpart,
-                         "D.Perception contacts" = perc_netw,
-                          "E.Preferences Effort" = pref_effort,
-                          "F.Preferences Talent" =  pref_talent,
-                         "G.Preferences rich parents" =  pref_wpart ,
-                          "H.Preferences contacts" = pref_netw  )
+dat04b<- dat04 %>% rename("A. Perception Effort" =perc_effort,
+                          "B. Perception Talent" = perc_talent,
+                          "C. Perception rich parents" = perc_wpart,
+                          "D. Perception contacts" = perc_netw,
+                          "E. Preferences Effort" = pref_effort,
+                          "F. Preferences Talent" =  pref_talent,
+                          "G. Preferences rich parents" =  pref_wpart ,
+                          "H. Preferences contacts" = pref_netw  )
 sum04b  <- psych::describe(dat04b[,-c(1)], fast = T) %>% select(mean,sd,min,max) %>% data.frame() %>% round(digits = 2)
@@
 
-El gráfico XX muestra ...
 A partir del la Figura \@ref(fig:plotlikert) se observa que en general hay una mayor percepción de elementos no meritocráticos que meritocráticos, mientras que en el caso de preferencias ocurre lo opuesto.
@@
 
-El Grafico \@ref(fig:corpoly) muestra ....
-Se observan relaciones de moderada a alta intensidad entre los indicadores que corresponden al mismo factor (por ejemplo, percepción de meritocracia esfuerzo con talento r=0.56), mientras que se observan correlaciones de baja intensidad entre el resto de los cruces.
+En la Figura \@ref(fig:corpoly) se observan relaciones de moderada a alta intensidad entre los indicadores que corresponden al mismo factor (por ejemplo, percepción de meritocracia esfuerzo con talento r=0.56), mientras que se observan correlaciones de baja intensidad entre el resto de los cruces.
 Destacan además las relaciones entre las percepciones y preferencias de tipo meritocráticas, lo que no ocurre con los indicadores no meritocraticos.
@@
 
-Figure X shows the results of the estimation for the four-factor model:
+In \@ref(fig:meas02) the results of the estimation for the four-factor model with complete sample.
+This model have a good fit (CFI = 0.977, RMSEA = 0.079, $\chi2$(df = 14) = 28,03, p = 0.014) and it presents factorial loads with weights that vary between 0.63 and 0.88.
+Furthermore, considering the correlation between the latent variables, it is observed that the most related variables are meritocratic perceptions and preferences, which have a positive correlation of 0.59, while non-meritocratic perceptions and preferences do not present a considerable correlation.
 
-> Julio: Esta sección debe ser revisada:
+```{r meas02, echo=FALSE, fig.cap = "Theorical model", fig.align='center'}
+knitr::include_graphics('../output/images/meas02.png')
+```
 
-- CFA general
-- Fit x modelo
+Como se puede observar en la Tabla \@ref(tab:tab-loads), todos los modelos, independiente del grupo y el órden de los items, obtuvieron un ajuste adecuado.
+con CFI superiores a .95 y RMSEA inferiores a .8, por contraparte ni un modelo logró un chi-square no significativo, aunque tanto el modelo aleatorio como el primero obtuvieron un adecuado chi-square ratio menor a 3.
+El primero de los órdenes fue el que obtuvo mejor ajuste (CFI=0.998, TLI= 0.995,RMSEA=0.037, $\chi2$(df=14)=28,03, p = 0.014), seguido por el orden aleatorio de los ítems (CFI=0.992, TLI=0.984,RMSEA=0.051, $\chi2$(df=14)=39.09, p < 0.001).
+Por su parte, la escala ordenada por temáticas parece generar un efecto framing en el cual la relación entre las percepciones y las preferencias parecen sobreestimadas, afectando de esta manera el ajuste (CFI=0.984, TLI=0.968, RMSEA=.071, $\chi2$(df=14)=64.156, p < 0.001).
 
-[Aquí SEM plot con cargas y ajuste global del modelo]
+```{r tab03, echo=FALSE,eval=TRUE}
+load(file = "../output/tables/fit4_o.RData")
+load(file = "../output/tables/fit3_o.RData")
+load(file = "../output/tables/fit2_o.RData")
+load(file = "../output/tables/fit1_o.RData")
 
-Los modelos confirmatorios fueron estimados en primer lugar para cada uno de los tres administration-sets. Los resultados se presentan en la Tabla \@ref(tab:tab-loads)
+sum_fit<- bind_rows(fitmeasures(fit4_o)[c("chisq.scaled","df","cfi.scaled","rmsea.scaled")],
+                    fitmeasures(fit1_o)[c("chisq.scaled","df","cfi.scaled","rmsea.scaled")],
+                    fitmeasures(fit2_o)[c("chisq.scaled","df","cfi.scaled","rmsea.scaled")],
+                    fitmeasures(fit3_o)[c("chisq.scaled","df","cfi.scaled","rmsea.scaled")])
+sum_fit$mod <- c("Model 1","Model 2","Model 3","Model 4")
+sum_fit$est <- c("DWLS","DWLS","DWLS","DWLS")
+sum_fit$nobs <- c(nobs(fit4_o),nobs(fit1_o),nobs(fit2_o),nobs(fit3_o))
+sum_fit <- select(sum_fit,mod,nobs,est,everything())
+colnames <- c("Model","$N$","Estimator","$\\chi^2$","df","CFI","RMSEA")
+
+sumtable02<- kable(sum_fit,digits = 3,format = table_format,row.names = F,booktabs=T, align = c("lcccccc"),
+                   caption = "Summary fit indices according group",
+                   col.names = colnames,
+                   escape = FALSE) %>%
+  kable_styling(full_width = F,
+                font_size = 10)  %>%
+      footnote(number = c("Model 1: Complete sample",
+                          "Model 2: fixed order by percepction/preference",
+                          "Model 3:  fixed order by topic (i.e: effort)",
+                          "Model 4: Randomized order"));sumtable02
+```
+
+Si bien  todas las pruebas obtuvieron indicadores relativamente adecuados, las diferencias mencionadas entre los ajustes de los modelos según orden son estadísticamente significativas, lo cual fue evaluado a partir de un análisis de invarianza.
+Se concluye que entre los tres órdenes no existe invarianza configuracional, es decir, no poseen la misma dimensionalidad y por ello no ajustan igualmente al modelo teórico [@vandenbergReviewSynthesisMeasurement2000].
+Esto se debe al efecto producido por la aparición conjunta de los ítems de un mismo factor en el orden 1, lo cual aumenta el ajuste del modelo.
+Además, en el orden 2, al preguntarse seguidamente por la percepción y la preferencia en torno al mismo indicador, aumentan las relaciones cruzadas disminuyendo el ajuste del modelo.
+De modo coherente, el modelo aleatorio presentó un ajuste intermedio entre el orden 1 y el orden 2.
+
+\pagebreak
+
+|           | Contrast Model 1 Factor | theoretical model 4 Factors | Model with M.I. |
+|:---------:|:-----------------------:|:---------------------------:|:---------------:|
+| **n**     |           1769          |             1769            |       1769      |
+| **CFI**   |          0.595          |            0.988            |      0.994      |
+| **TLI**   |          0.433          |            0.976            |      0.985      |
+| **RMSEA** |          0.226          |            0.047            |      0.036      |
+| **$\chi2$**    |         1830.839        |            68.661           |      40.250     |
+| **p**     |           .000          |             .000            |       .000      |
+| **$\chi2$/df** |          65.38          |             4.90            |       3.32      |
+Table: `r as.character(paste("(\\#tab:items-dim)", "Items according to dimension."))`.
+
+EL modelo teórico propuesto de cuatro factores ajustó, como se observa en la tabla 4, mejor que el modelo de contraste de 1 factor.
+El modelo teórico ajustó de manera relativamente adecuada, pues muestra indicadores óptimos para CFI= 0.987, TLI = 0.975 y RMSEA=.041, aunque posee indicadores deficientes para la prueba $\chi2$(df=14)=67.6, p-value=.000.
+Para evaluar posibles mejoras de la escala, se analizaron las relaciones propuestas por los índices de modificación.
+Estos indican la existencia de dos cargas cruzadas no especificadas.
+Cuando se generó un modelo siguiendo estas recomendaciones, hubo una mejora considerable del modelo, aunque el nuevo modelo tampoco obtuvo un $\chi2$ ratio menor a 3 y obtuvo cargas factoriales muy bajas ($\tau$)< 0.15, por lo tanto, siguiendo las recomendación de Brown (2006) de solo aceptar las propuesta de los índices de modificación cuando se posee teoria y evidencia sólida, se ha decidido no incorporar estos parámetros al modelo.
+
+# Conclusión
+
+Considerando la ventaja del orden, respecto aleatorio despeja la posibilidad del efecto framing, es decir, de que el resultado del modelo se deba al orden de las preguntas, se ha decidido seguir utilizando el orden 3.
+
+uso conjunto percepciones-preferencias
+
+\pagebreak
+
+# References
+
+<div id="refs"> </div>  <!-- Although <div> is an HTML tag, this method also works for other output formats such as PDF. -->
+
+# (APPENDIX) Appendix {-}
+
+`r if (knitr::is_latex_output()){ '\\appendix'}`
+
+`r if (knitr::is_latex_output()){ '\\section{Appendix}'}`
 
@@
 
-Como se puede observar en la Tabla XX, todos los modelos independiente de los órdenes obtuvieron un ajuste adecuado.
-con CFI superiores a .95 y RMSEA inferiores a .8, por contraparte ni un modelo logró un chi-square no significativo, aunque tanto el modelo aleatorio como el primero obtuvieron un adecuado chi-square ratio menor a 3.
-El primero de los órdenes fue el que obtuvo mejor ajuste (CFI=0.998, TLI= 0.995,RMSEA=0.037, $\chi2$(df=14)=28,03, p = 0.014), seguido por el orden aleatorio de los ítems (CFI=0.992, TLI=0.984,RMSEA=0.051, $\chi2$(df=14)=39.09, p < 0.001).
-Por su parte, la escala ordenada por temáticas parece generar un efecto framing en el cual la relación entre las percepciones y las preferencias parecen sobreestimadas, afectando de esta manera el ajuste (CFI=0.984, TLI=0.968,RMSEA=.071, $\chi2$(df=14)=64.156, p < 0.001).
-
-Si bien  todas las pruebas obtuvieron indicadores relativamente adecuados, las diferencias mencionadas entre los ajustes de los modelos según orden son estadísticamente significativas, lo cual fue evaluado a partir de un análisis de invarianza.
-Se concluye que entre los tres órdenes no existe invarianza configuracional, es decir, no poseen la misma dimensionalidad y por ello no ajustan igualmente al modelo teórico [@vandenbergReviewSynthesisMeasurement2000].
-Esto se debe al efecto producido por la aparición conjunta de los ítems de un mismo factor en el orden 1, lo cual aumenta el ajuste del modelo.
-Además, en el orden 2, al preguntarse seguidamente por la percepción y la preferencia en torno al mismo indicador, aumentan las relaciones cruzadas disminuyendo el ajuste del modelo.
-De modo coherente, el modelo aleatorio presentó un ajuste intermedio entre el orden 1 y el orden 2.
-
 \pagebreak
-
-> Julio: ¿Esto se debe mantener?
-
-```{r include=FALSE}
-cap <- "Items according to dimension."
-# if (knitr::is_html_output()) {te <- table_nums("te",cap)} else {te <- cap}
-```
-
-|           | Contrast Model 1 Factor | theoretical model 4 Factors | Model with M.I. |
-|:---------:|:-----------------------:|:---------------------------:|:---------------:|
-| **n**     |           1769          |             1769            |       1769      |
-| **CFI**   |          0.595          |            0.988            |      0.994      |
-| **TLI**   |          0.433          |            0.976            |      0.985      |
-| **RMSEA** |          0.226          |            0.047            |      0.036      |
-| **$\chi2$**    |         1830.839        |            68.661           |      40.250     |
-| **p**     |           .000          |             .000            |       .000      |
-| **$\chi2$/df** |          65.38          |             4.90            |       3.32      |
-Table:`r cap`
-
-EL modelo teórico propuesto de cuatro factores ajustó, como se observa en la tabla 4, mejor que el modelo de contraste de 1 factor.
-El modelo teórico ajustó de manera relativamente adecuada, pues muestra indicadores óptimos para CFI= 0.987, TLI = 0.975 y RMSEA=.041, aunque posee indicadores deficientes para la prueba $\chi2$(df=14)=67.6, p-value=.000.
-Para evaluar posibles mejoras de la escala, se analizaron las relaciones propuestas por los índices de modificación.
-Estos indican la existencia de dos cargas cruzadas no especificadas.
-Cuando se generó un modelo siguiendo estas recomendaciones, hubo una mejora considerable del modelo, aunque el nuevo modelo tampoco obtuvo un $\chi2$ ratio menor a 3 y obtuvo cargas factoriales muy bajas ($\tau$)< 0.15, por lo tanto, siguiendo las recomendación de Brown (2006) de solo aceptar las propuesta de los índices de modificación cuando se posee teoria y evidencia sólida, se ha decidido no incorporar estos parámetros al modelo.
-
-# Conclusión
-
-Considerando la ventaja del orden, respecto aleatorio despeja la posibilidad del efecto framing, es decir, de que el resultado del modelo se deba al orden de las preguntas, se ha decidido seguir utilizando el orden 3.
-
-uso conjunto percepciones-preferencias
-
-\pagebreak
-
-# References
-
-<div id="refs"> </div>  <!-- Although <div> is an HTML tag, this method also works for other output formats such as PDF. -->
-
-# Appendix {-}
-
-```{r include=FALSE}
-cap <- "Representativeness of the sample."
-# if (knitr::is_html_output()) {aax <- table_nums("aax",cap)} else {aax <- cap}
-```
 
@@
 | university or more  | 21,21%  | 18,13% |
-Table:`r cap`
+Table: `r as.character(paste("(\\#tab:rep-samp)", "Representativeness of the sample"))`.

### 2020-07-31 production/prod_analysis-cfa.rmd
update-julio
correción de tablas, títulos y figuras.

@@
 
-semPaths(fit4_o , 
+semPaths(
+  fit4_o ,
   whatLabels = "std",
-what="col",
-label.cex=1,
-edge.label.cex=1,
-residuals=F, 
+  # what = "col",
+  label.cex = 1,
+  edge.label.cex = 0.7,
+  residuals = F,
+  optimizeLatRes =T,
   edge.color = "black",
   style = "lisrel",
-fade=F, 
+  fade = F,
   nCharNodes = 0,
   curvePivot = FALSE,
-         curve = 2,
+  curve = 3,
   rotation = 4,
-layout = "tree3",
+  layout = "tree2",
   cardinal = "lat cov",
   legend.cex = 0.6,
-         label.cex =1,
+  label.cex = 1,
   label.font = 6,
-         edge.label.font= 1,
+  edge.label.font = 7,
   asize = 3,
-edge.width =1,
-         sizeMan=5,
-         sizeLat = 10, width = 12,
-         height= 15, nodeNames = nodeNames, intercepts = F, reorder = F, thresholds = F, nodeLabels = c("A","B","C","D","E","F","G","H","1","2","3","4"), filetype ="png", filename = "output/images/meas02") #arrow size)
+  edge.width = 1,
+  sizeMan = 5,
+  sizeLat = 10,
+  width = 20,
+  height = 15,
+  nodeNames = nodeNames,
+  intercepts = F,
+  reorder = F,
+  thresholds = F,
+  fixedStyle =1,
+  nodeLabels = c("A", "B", "C", "D", "E", "F", "G", "H", "1", "2", "3", "4"),
+  filetype = "png",
+  filename = "output/images/meas02"
+) #arrow size)
 
@@
 
-dat_merit_2<- dat_merit %>% sjmisc::rec(rec="rev") %>% select("a.Effort"=a.Effort_r,"a.Talent"=a.Talent_r, "a.Rich Parents"="a.Rich Parents_r", "a.Contacts"=a.Contacts_r, "b.Effort"=b.Effort_r,"b.Talent"=b.Talent_r,"b.Rich Parents"="b.Rich Parents_r", "b.Contacts"=b.Contacts_r)
-
-plotlikert<-plot_likert(dat_merit_2, c(1,1,1,1,2,2,2,2),
-            groups.titles = c("Perceptions","Preferences"),
-            geom.colors ="PuBu", 
+dat_merit_2<- dat_merit %>% sjmisc::rec(rec="rev") %>% 
+  select(
+    "a.Effort" = a.Effort_r,
+    "a.Talent" = a.Talent_r,
+    "a.Rich Parents" = "a.Rich Parents_r",
+    "a.Contacts" = a.Contacts_r,
+    "b.Effort" = b.Effort_r,
+    "b.Talent" = b.Talent_r,
+    "b.Rich Parents" = "b.Rich Parents_r",
+    "b.Contacts" = b.Contacts_r)
+
+plotlikert<-plot_likert(dat_merit_2, 
+                        c(1, 1, 1, 1, 2, 2, 2, 2),
+                        groups.titles = c("Perceptions", "Preferences"),
+                        geom.colors = "PuBu",
                         geom.size = 0.8,
-            axis.labels=c("Effort","Talent","Rich Parents","Contacts"),
+                        axis.labels = c("Effort", "Talent", "Rich Parents", "Contacts"),
                         catcount = 4,
                         cat.neutral = 3,
-            grid.range  =  c ( 1.2 , 1.4 ),
-            values  =  "sum.outside", reverse.colors = T)
+                        grid.range  =  c (1.2 , 1.4),
+                        values  =  "sum.outside", 
+                        reverse.colors = T,
+                        reverse.scale = F)
+
 plotlikert
 
-ggsave(plotlikert,filename = "output/images/plotlikert.png",device = "png",width = 25,height = 15,dpi = "retina",units = "cm")
-
+ggsave(plotlikert,filename = "output/images/plotlikert.png",device = "png",width = 30,height = 15,dpi = "retina",units = "cm")
 ```

### 2020-09-10 docs/index.Rmd

@@
 output:
-    html_document:
+    bookdown::html_document2:
             smart: false
@@
               in_header: head_github-activity.html
+            number_sections: false          
 ---

### 2020-09-10 docs/paper.Rmd

@@
 rmarkdown::render("docs/paper.Rmd", output_format = "bookdown::pdf_document2")
+rmarkdown::render("docs/paper.Rmd", output_format = "bookdown::html_document2")
 ```
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = FALSE,out.width = '75%',fig.pos= "H")
+knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = FALSE,out.width = '85%',fig.pos= "H")
 # knitr::opts_knit$set(base.url = "../") #relative path for .html output file
@@
     full_width = T,
-    latex_options = c("hold_position", "scale_down"),
+    latex_options = c("hold_position"),
     position = "center",
@@
 sumdat04<- kable(table04b,format =table_format, booktabs= T, escape = FALSE,align = "l",linesep = "", caption = cap
-) %>% kable_styling(full_width = table_format2, font_size = 10,position = "center",latex_options = c("hold_position"));sumdat04
+) %>% kable_styling(full_width = table_format2, 
+                    font_size = 10,
+                    position = "center", 
+                    latex_options = c("hold_position"));sumdat04
 ```
@@
 
-```{r meas02, echo=FALSE, fig.cap = "Theorical model", fig.align='center'}
+```{r meas02, echo=FALSE, fig.cap = "Confirmatory factor analysis of the Perceptions an Preferences for Meritocracy Scale", fig.align='center'}
 knitr::include_graphics('../output/images/meas02.png')

### 2020-09-10 production/prod_analysis-cfa.rmd

@@
 library(sjlabelled)
+library(ggplot2)
 source('production/prod_polychoric.R')
@@
 cor04<-polychoric(select(dat04,perc_effort,perc_talent,perc_wpart,perc_netw,pref_effort,pref_talent,pref_wpart,pref_netw))
-rownames(cor04$rho) <-c("A.Perception Effort", "B.Perception Talent","C.Perception Rich parents","D.Perception contacts","E.Preferences Effort","F.Preferences Talent","G.Preferences rich parents", "H.Preferences contacts" )
+windowsFonts(A = windowsFont("Times New Roman"))
+rownames(cor04$rho) <-c(
+    "A. Perception Effort",
+    "B. Perception Talent",
+    "C. Perception Rich parents",
+    "D. Perception Contacts",
+    "E. Preferences Effort",
+    "F. Preferences Talent",
+    "G. Preferences Rich parents",
+    "H. Preferences Contacts")
 colnames(cor04$rho) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)", "(H)")
 save(cor04, file = "output/images/correlation.RData")
-corrplot(cor04$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = TRUE)
+corrplot(
+  cor04$rho,
+  method = "color",
+  type = "upper",
+  tl.col = "black",
+  addCoef.col = "black",
+  diag = TRUE,
+  family = "A", 
+  number.font = 6,
+  tl.cex =0.75,
+  number.cex = 1)
 
 png("output/images/corpoly.png",
-    width = 15,height = 15,units="cm",
+    width = 18,height = 13,units="cm",
     pointsize=12,bg="white",res=300)
 
-corrplot(cor01$rho,method = "color" ,type = "upper", tl.col = "black", addCoef.col = "black",diag = TRUE)
+corrplot(
+  cor04$rho,
+  method = "color",
+  type = "upper",
+  tl.col = "black",
+  addCoef.col = "black",
+  diag = TRUE,
+  family = "A", 
+  number.font = 6,
+  tl.cex =0.75,
+  number.cex = 1)
 
@@
   sizeLat = 10,
-  width = 20,
+  width = 22, 
   height = 15,
-  nodeNames = nodeNames,
+  # nodeNames = nodeNames,
   intercepts = F, 
@@
 ) #arrow size)
-
+# text(2.5, -1, labels = "Hola")
 ```
@@
 ggsave(plotlikert,filename = "output/images/plotlikert.png",device = "png",width = 30,height = 15,dpi = "retina",units = "cm")
+```
+
+```{r}
+datli03 <- dat04 %>% select(starts_with("perc_"),
+                            starts_with("pref_")) %>%  sjmisc::rec(rec="rev",append = F)
+
+datli03 <- remove_label(datli03)
+
+set_theme(base = theme_bw(),
+          theme.font = "serif", 
+          axis.textsize.x = 1.2,
+          axis.textsize.y = 1.2, 
+          geom.label.size = 4, 
+          legend.just = 0.35, 
+          legend.size = 1.2,
+          legend.backgroundcol = "white")
+# http://www.sthda.com/english/wiki/colors-in-r  = PALETAS de COLORES 
+plot01<- plot_likert(
+  datli03,
+  c(rep("Perception", 4), rep("Preference", 4)),
+  values  =  "sum.outside",
+  show.prc.sign = F,
+  sort.groups = FALSE,
+  grid.range = c(1.0, 1.0),
+  geom.colors = "Blues",
+  rel_heights = c(4, 4),
+  wrap.labels = 100,
+  reverse.scale = FALSE,
+  cat.neutral = 3, 
+  legend.pos = "bottom", 
+  group.legend.options = list(reverse = TRUE),
+  axis.labels = c("Effort", "Talent", "Rich parents", "Contacts"))
+plot01
+
+ggsave(plot01,filename = "output/images/plotlikert.png",device = "png",width = 40,height = 25,dpi = "retina",units = "cm")
 ```

### 2020-09-10 production/prod_prep-data.rmd

@@
 In the first order (group 1), questions of the same factor are then presented to them, in the second order (group 2) they were separated and in the third order (group 3) at random."
-dat04$perc_effort <- set_labels(dat04$perc_effort,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
+dat04$perc_effort <- set_labels(dat04$perc_effort,  labels = c("Strongly disagree" = 1, "Disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "Totally agree" = 5))
 label(dat04$perc_effort) <- "Thinking of Chilean society,
@@
 less."
-dat04$perc_talent <- set_labels(dat04$perc_talent,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
+dat04$perc_talent <- set_labels(dat04$perc_talent,  labels = c("Strongly disagree" = 1, "Disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "Totally agree" = 5))
 label(dat04$perc_talent) <- "Thinking of Chilean society,
@@
 less talent."
-dat04$perc_wpart <- set_labels(dat04$perc_wpart,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
+dat04$perc_wpart <- set_labels(dat04$perc_wpart,  labels = c("Strongly disagree" = 1, "Disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "Totally agree" = 5))
 label(dat04$perc_wpart) <- "Thinking of Chilean society,
@@
 ahead."
-dat04$perc_netw <- set_labels(dat04$perc_netw,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
+dat04$perc_netw <- set_labels(dat04$perc_netw,  labels = c("Strongly disagree" = 1, "Disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "Totally agree" = 5))
 label(dat04$perc_netw) <- "Thinking of Chilean society,
@@
 ahead."
-dat04$pref_effort <- set_labels(dat04$pref_effort,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
+dat04$pref_effort <- set_labels(dat04$pref_effort,  labels = c("Strongly disagree" = 1, "Disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "Totally agree" = 5))
 label(dat04$pref_effort) <- "Thinking of Chilean society,
@@
 they try less."
-dat04$pref_talent <- set_labels(dat04$pref_talent,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
+dat04$pref_talent <- set_labels(dat04$pref_talent,  labels = c("Strongly disagree" = 1, "Disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "Totally agree" = 5))
 label(dat04$pref_talent) <- "Thinking of Chilean society,
@@
 less talent."
-dat04$pref_wpart <- set_labels(dat04$pref_wpart,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
+dat04$pref_wpart <- set_labels(dat04$pref_wpart,  labels = c("Strongly disagree" = 1, "Disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "Totally agree" = 5))
 label(dat04$pref_wpart) <- "Thinking of Chilean society,
@@
 the following statements? - It's fine that those who have rich parents get ahead" 
-dat04$pref_netw <- set_labels(dat04$pref_netw,  labels = c("strongly disagree", "disagreement", "Neither agree nor disagree","Agree", "totally agree"))
+dat04$pref_netw <- set_labels(dat04$pref_netw,  labels = c("Strongly disagree" = 1, "Disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "Totally agree" = 5))
 label(dat04$pref_netw) <- "Thinking of Chilean society,

### 2020-09-10 docs/paper.Rmd
correcciones finales en texto

@@
 A number of studies have relied mostly in secondary data to operationalize meritocracy, with a large variation in the use and interpretation of survey items.
-Starting from a review of studies that measure meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding the conceptualization and measurement of meritocracy.
-Based on this critical analysis, we propose  an items battery called *Perceptions and preferences for meritocracy scale*, which is tested with confirmatory factor analysis with data from an online survey study (N=2,141).
+Starting from a review of studies that measure meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding the conceptualization and indicators of meritocracy.
+Based on this critical analysis, we propose an item's battery called *Perceptions and preferences for meritocracy scale*, which is tested with confirmatory factor analysis with data from an online survey study (N=2,141).
 The results support the proposed conceptual structure which not only distinguishes between perceptions and preferences, but also between meritocratic and non-meritocratic dimensions.
@@
 
-Based on the previous limitations and confussions in the measurement of meritocracy presented in the previous section, in this paper we propose and test an instrument with the following characteristics:
+Based on the previous limitations and confusions in the measurement of meritocracy presented in the previous section, in this paper we propose and test an instrument with the following characteristics:
 
@@
 The questionnaire was programmed in Qualtrics and the fieldwork was in charge of an external online survey agency ([netquest.cl](www.netquest.cl)) during December 2019 and January 2020.
-The sample was selected from a non-probabilistic design in three large cities in Chile.
-The quota method based on age, sex and educational level was used.
-Quotas were generated based on the survey of the Public Studies Center (CEP, 2019), which has a high prestige in the country and is also the counterpart agency of ISSP (International Social Survey Programme) in Chile.
+The sample was selected from a non-probabilistic quota design in three large cities in Chile.
+The quotas were generated based on the survey of the Public Studies Center (CEP, 2019), which has a high prestige in the country and is also the counterpart agency of ISSP (International Social Survey Programme) in Chile.
 A total sample of 2,141 people was collected, excluding those who did not answer the questions on the scale and those who did not accept informed consent.
@@
 
-CFA was conducted using the `lavaan` package (version 0.6-3; Rosseel, 2020) with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement (Kline, 2016; Rosseel, 2020).
+CFA was conducted using the `lavaan` R package (version 0.6-3; Rosseel, 2020) with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement (Kline, 2016; Rosseel, 2020).
 As recommended by Brown (2008), we assessed model fit by jointly considering the comparative fit index and Tucker-Lewis Index (CFI and TLI; acceptable fit > 0.95), Root of the average squared residual approximation (RMSEA; acceptable fit < 0.08),  Chi-square: (p-value; acceptable fit > 0.05, and Chi-square ratio:> 3).
 
-A pre-registration was made in the OSF platform, [Perceptions and Preferences for Meritocracy Scale
-Public registration](https://osf.io/z45y2) In it, it was indicated, before carrying out the analyzes, the four-dimensional theoretical model, the level of measurement of the variables, the statistical tests to be used with their respective evaluation parameters, along with other important aspects of the research design.
+A pre-registration was made in the OSF platform, available in the following link: [https://osf.io/z45y2](https://osf.io/z45y2).
+In this pre-registration are the hypothesis regarding the four-dimensional conceptual model underlying the scale,  the variables measurement levels, the statistical tests to be performed with their respective evaluation parameters, along with other important aspects of the research design.
 
@@
 Se observan promedios desde 2.41 correspondiente a preferencia-contactos, hasta 3.89 correspondiente a la preferencia-esfuerzo.
-Ambos indicadores son coherentes con la adhesión general a la meritocracia, privilegiando aspectos individuales como el esfuerzo.
+Ambos indicadores son coherentes con la adhesión general a la meritocracia reportada en estudios previos, privilegiando aspectos individuales como el esfuerzo.
 
@@
 
-A partir del la Figura \@ref(fig:plotlikert) se observa que en general hay una mayor percepción de elementos no meritocráticos que meritocráticos, mientras que en el caso de preferencias ocurre lo opuesto.
-En cuanto a las preferencias, llama la atención el rol preponderante del esfuerzo por sobre el talento.
+Los gráficos presentados en la Figura \@ref(fig:plotlikert) presentan información desagregada y comparable de las distintas categorías de respuesta de cada item.
+Se observa que en general hay una mayor percepción de elementos no meritocráticos que meritocráticos, mientras que en el caso de preferencias ocurre lo opuesto.
+En cuanto a las preferencias, llama la atención el rol preponderante del esfuerzo por sobre el talento como criterio de preferencia meritocrático.
 
@@
 
-En la Figura \@ref(fig:corpoly) se observan relaciones de moderada a alta intensidad entre los indicadores que corresponden al mismo factor (por ejemplo, percepción de meritocracia esfuerzo con talento r=0.56), mientras que se observan correlaciones de baja intensidad entre el resto de los cruces.
+En la Figura \@ref(fig:corpoly) se observan relaciones de moderada a alta intensidad entre los indicadores que corresponden al mismo factor (por ejemplo, percepción de meritocracia por esfuerzo y por talento, r=0.56).
 Destacan además las relaciones entre las percepciones y preferencias de tipo meritocráticas, lo que no ocurre con los indicadores no meritocraticos.
@@
 
-En suma, los análisis descriptivos señalan una relativa adhesión a la moral meritocratica, las cuales, a modo general, se expresan en una mayor preferencia por criterios meritocraticos y una menor por no meritocráticos.
+En suma, los análisis descriptivos señalan una relativa adhesión a la moral meritocratica, las cuales se expresan en una mayor preferencia por criterios meritocraticos y una menor por no meritocráticos, aún cuando la preferencia por elementos no-meritocraticos alcanza una proporción de más de 20%.
 Igualmente, es observable una relativamente baja percepción de meritocracia.
@@
 En particular las asociaciones entre esfuerzo y talento son relevantes, ya que desestiman previos supuestos sobre que el talento no sería un criterio meritocrático [@mijs_paradox_2019], de otra manera la correlación sería cero o negativa.
-Junto a ello, vemos que no existe una correlacion negativa entre aspectos meritocráticos y no meritocráticos, desestimando los supuestos de estudios previos que  señalaban que estas dimensiones serían los polos opuestos de un mismo continuo [@reynolds_perceptions_2014].
+Junto a ello, vemos que no existe una correlación negativa entre aspectos meritocráticos y no meritocráticos, desestimando los supuestos de estudios previos que  señalaban que estas dimensiones serían los polos opuestos de un mismo continuo [@reynolds_perceptions_2014].
 
@@
 
-This section tests the item's battery and the corresponding conceptual framework under test.
+This section estimates the fit of the conceptual model behind the Perceptions and Preferences for Meritocracy Scale.
 For this, we first estimate a confirmatory factor analysis model for the whole sample, and secondly we test the order effects applying the same model to each of the three order versions.
@@
 
-In \@ref(fig:meas02) the results of the estimation for the four-factor model with complete sample.
-This model have a good fit (CFI = 0.977, RMSEA = 0.079, $\chi2$(df = 14) = 28,03, p = 0.014) and it presents factorial loads with weights that vary between 0.63 and 0.88.
-Furthermore, considering the correlation between the latent variables, it is observed that the most related variables are meritocratic perceptions and preferences, which have a positive correlation of 0.59, while non-meritocratic perceptions and preferences do not present a considerable correlation.
+\@ref(fig:meas02) SHOWS the results of the estimation for the four-factor model with complete sample.
+This model presents adequate fit indicators (CFI = 0.977, RMSEA = 0.079, $\chi2$(df = 14) = 28,03, p = 0.014) and it presents factorial loads with weights that vary between 0.63 and 0.88.
+Furthermore, considering the correlation between the latent variables, it is observed that the variables with the highest correlations are meritocratic perceptions and preferences, which have a positive correlation of 0.59, while non-meritocratic perceptions and preferences do not present a considerable correlation.
 
@@
 Como se puede observar en la Tabla \@ref(tab:tab-loads), todos los modelos, independiente del grupo y el órden de los items, obtuvieron un ajuste adecuado.
-con CFI superiores a .95 y RMSEA inferiores a .8, por contraparte ni un modelo logró un chi-square no significativo, aunque tanto el modelo aleatorio como el primero obtuvieron un adecuado chi-square ratio menor a 3.
+con CFI superiores a .95 y RMSEA inferiores a .8.
+Sin embargo ningún modelo logró un chi-square no significativo (algo esperable en muestras grandes), aunque tanto el modelo aleatorio como el primero obtuvieron un adecuado chi-square ratio menor a 3.
 El primero de los órdenes fue el que obtuvo mejor ajuste (CFI=0.998, TLI= 0.995,RMSEA=0.037, $\chi2$(df=14)=28,03, p = 0.014), seguido por el orden aleatorio de los ítems (CFI=0.992, TLI=0.984,RMSEA=0.051, $\chi2$(df=14)=39.09, p < 0.001).
@@
 Estos indican la existencia de dos cargas cruzadas no especificadas.
-Cuando se generó un modelo siguiendo estas recomendaciones, hubo una mejora considerable del modelo, aunque el nuevo modelo tampoco obtuvo un $\chi2$ ratio menor a 3 y obtuvo cargas factoriales muy bajas ($\tau$)< 0.15, por lo tanto, siguiendo las recomendación de Brown (2006) de solo aceptar las propuesta de los índices de modificación cuando se posee teoria y evidencia sólida, se ha decidido no incorporar estos parámetros al modelo.
+Cuando se generó un modelo siguiendo estas recomendaciones, hubo una mejora considerable del modelo, aunque el nuevo modelo tampoco obtuvo un $\chi2$ ratio menor a 3 y obtuvo cargas factoriales muy bajas ($\tau$)< 0.15, por lo tanto, siguiendo las recomendación de Brown (2006) de solo aceptar las propuesta de los índices de modificación cuando se posee teoría y evidencia sólida, se ha decidido no incorporar estos parámetros al modelo.
 

### 2020-09-15 docs/paper.Rmd
Revisión y conclusiones

@@
 abstract: "Economic and social inequality have raised growing concerns and crises across societies.
-One of social science concepts associated to the maintenance of inequality is the belief in meritocracy, which would legitimize economic differences based on effort and talent.
+One of social science concepts associated to the maintenance of inequality is the belief in meritocracy, which would legitimize economic differences based on criteria such as effort and talent.
 Despite its wide use, empirical research on meritocracy is something relatively novel.
@@
 Starting from a review of studies that measure meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding the conceptualization and indicators of meritocracy.
-Based on this critical analysis, we propose an item's battery called *Perceptions and preferences for meritocracy scale*, which is tested with confirmatory factor analysis with data from an online survey study (N=2,141).
+Based on this critical analysis, we propose an item's battery called *Perceptions and Preferences for Meritocracy Scale*, which is tested with confirmatory factor analysis with data from an online survey study (N=2,141).
 The results support the proposed conceptual structure which not only distinguishes between perceptions and preferences, but also between meritocratic and non-meritocratic dimensions.
@@
 En general, la meritocracia se define como un sistema de distribución de recursos y recompensas basados en el mérito individual, que en su concepción original es una suma de talento y esfuerzo [@young_rise_1962].
-Esta concepción tradicional de mérito pone en un lugar secundario la posible interferencia de factores estructurales o no meritocráticos como la herencia, los contactos personales, o la suerte [@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995; @yairMeritocracy2007; @landWeSatTable2006; @youngRiseMeritocracy1994].
-Una serie de estudios han planteado críticas a la realización de este estándar moral de distribución, argumentando que es una promesa incumplida dada la influencia preponderante de otros elementos más allá del mérito en el estatus individual [@witteveenReconsideringMeritocraticPower2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019].
-Por otro lado, desde la psicología social y la sociología se han estudiado las características y consecuencias de las creencias en la meritocracia, en general basados en la hipótesis que mayor creencia en la meritocracia lleva a una mayor legitimación de las desigualdades [@premingerMeritocracyServiceEthnocracy2020; @trumpWhenWhyEconomic2020; @hadjar_meritokratie_2008; @MadeiraPrimesConsequencesSystematic2019].
+Esta concepción tradicional de mérito sitúa en un lugar secundario la posible interferencia de factores estructurales o no meritocráticos como la herencia, los contactos personales, o la suerte [@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995; @yairMeritocracy2007; @landWeSatTable2006; @youngRiseMeritocracy1994].
+Desde la psicología social y la sociología se han estudiado las características y consecuencias de las creencias en la meritocracia, en general basados en la hipótesis que mayor creencia en la meritocracia lleva a una mayor legitimación de las desigualdades [@premingerMeritocracyServiceEthnocracy2020; @trumpWhenWhyEconomic2020; @hadjar_meritokratie_2008; @MadeiraPrimesConsequencesSystematic2019].
+Tales investigaciones han planteado críticas a la realización de este estándar moral de distribución, argumentando que sería una promesa incumplida dada la influencia preponderante de otros elementos más allá del mérito en el estatus individual [@witteveenReconsideringMeritocraticPower2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019, @khan_Privilege_2013].
 
@@
 
-Para poder dar cuenta de los niveles de creencia en la meritocracia los estudios a la fecha generalmente han utilizado algunos indicadores de encuestas ya existentes, y en el menor de los casos se han creado instrumentos ad-hoc.
+Para poder dar cuenta de los niveles de creencia en la meritocracia los estudios a la fecha generalmente han utilizado algunos indicadores de encuestas ya existentes, y en menor medida se han creado instrumentos ad-hoc.
 Sin embargo, y como mostraremos más adelante, las formas de medición de meritocracia varían extremadamente entre estudios.
-Muchas veces fenómenos similares se asocian a indicadores distintos, y también ocurre que fenómenos distintos son medidos con indicadores similares, todo lo cual hace dificulta la comparabilidad entre estudios y el poder avanzar en la comprensión y estudio de la meritocracia.
+Muchas veces fenómenos similares se asocian a indicadores distintos, y también ocurre que fenómenos distintos son medidos con indicadores similares, todo lo cual dificulta la comparabilidad entre investigaciones y avanzar en la comprensión y estudio de la meritocracia.
 
@@
 Además, como un segundo eje de análisis considera la generación de indicadores respecto de aspectos meritocráticos y anti-meritocráticos, demostrando que no son los dos polos de un mismo continuo como muchos estudios anteriores parecen sugerir.
-La propuesta de medición además está orientada a generar un instrumento lo más breve posible de manera que pueda ser utilizado en encuestas de opinión pública y así ser asociado a otros fenómenos sociales.
+La propuesta de medición además está orientada a generar un instrumento lo más breve posible de manera que pueda ser utilizado en encuestas de opinión pública y así ser estudiado en relación a otros fenómenos sociales.
 
@@
 Several approaches to the empirical study of meritocracy based on public opinion surveys make reference to the concept of _beliefs_, but behind this concept there are usually different meanings and operationalizations.
-To illustrate this point in the following we will start with the proposal from a recent paper by @mijs_paradox_2019, which we will take as a reference to discuss previous studies.
+To illustrate this point in the following we will start with the proposal from a recent paper by @mijs_paradox_2019, which we will take as a reference to discuss previous studies on meritocracy.
 
@@
 In the operationalization, this is associated with the following indicator: "how important you think it is for getting ahead in life: (a) hard work", scored in a 1 to 5 likert scale.
-There are several assumptions behind this decision that are worth discussing:
+There are several assumptions behind this decision that are worth discussing and that are related to previous assumptions in the study of meritocracy.
 
@@
 
-The item used by Mijs is part of an items' battery present in several international surveys, usually called "reasons to get ahead".
+The item used by @mijs_paradox_2019 is part of an items' battery present in several international surveys, usually called "reasons to get ahead".
 This battery presents a series of indicators related to what people consider important to get ahead in life: hard work, education, ambition, wealthy family, right connections, religion, race and gender.
-Therefore, for Mijs other aspects as education, that could be associated to talent, are not meritocratic.
+Therefore, for Mijs other aspects such as education, that could be associated to talent, are not meritocratic.
 As he points out: "Hard work is arguably the most meritocratic part of Michael Young’s equation, ‘Merit = Intelligence + Effort’, for the simple fact that intelligence itself is conditioned by a nonmeritocratic factor: who your parents happen to be" (p.5).
@@
 In Mijs' proposal we can observe a couple of strong assumptions: effort would not depend on parents influence, and talent is not meritocratic (contrary to Michael Youngs original conceptualization).
-The problem of whether talent is or not meritocratic is something that actually should (and can) be tested, it cannot be simply ruled out particularly in empirical research.
-This conceptual and measurement issue it is possible to find in other studies that assume that effort is the main and only aspect of meritocracy [@girerdNeoliberalismIdeologicalBarrier2020a; @bubakPerceptionsMeritocracyNote2019].
+The problem of whether talent is or not meritocratic is certainly an interesting point to discuss, but for those working based on empirical data it is something that should be empirically tested.
+Is effort the only dimension behind the concept of meritocracy, or is it a multidimensional concept that includes other elements as talent (as in its original conceptualization)?
+This conceptual and measurement unidimensional assumption it is possible to find in other studies that assume that effort is the main and only aspect of meritocracy [@girerdNeoliberalismIdeologicalBarrier2020a; @bubakPerceptionsMeritocracyNote2019].
 
@@
 
-The "resons to get ahead" battery refers to "how important you think it is", considered by Mijs as a belief in meritocracy for the item regarding effort.
-Nevertheless, another version of this same battery in the survey asks about "how important you think it _should_ be".
-Which one of both is a "belief"?
+The "reasons to get ahead" battery refers to "how important you think it is", considered by @mijs_paradox_2019 as a belief in meritocracy for the item regarding effort.
+Nevertheless, another version of this same battery used in several surveys is "how important you think it _should_ be".
+Therefore, the question raised here is: Which one of both is a "belief"?
 According to @janmaatSubjectiveInequalityReview2013: "Perceptions refer to subjective estimates of existing inequality (i.e.
@@
 thoughts about what should be)"(p.359).
-Therefore, the referred paper would be actually using the term beliefs referring to perceptions.
-The same occurs in @reynolds_perceptions_2014, who explicitly use the term beliefs to talk about perceptions, whereas other authors use different terms as attitudes (@kunovich_systems_2007).
+Therefore, the referred paper would be using the term beliefs (what should be) referring actually to perceptions (what is).
+The same occurs in studies as the one by @reynolds_perceptions_2014, who explicitly use the term beliefs to talk about perceptions, whereas other authors use general terms as attitudes (@kunovich_systems_2007).
 The first attempt to shed light on this confusion was made by @duru-bellat_whos_2012, who used the item "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" for "desired" meritocracy (beliefs), whereas for "perceived" meritocracy they use two items: “Would you say that in your country, people are rewarded for their efforts?” and “...
@@
 Is the belief in meritocracy a perception or a desire/preference?
-In order to expand the analitical conceptual framework, we believe that both dimensions should be included in the analysis, as proposed by @duru-bellat_whos_2012.
-This opens up possibilities of analyzing whether perceptions and preferences are actually the same (i.e.
-correlation close to 1) or they are different aspects of the same phenomenon.
+In order to expand the analytical conceptual framework, we believe that both dimensions should be included in the analysis, as proposed by @duru-bellat_whos_2012.
+This opens up possibilities of analyzing whether perceptions and preferences are actually related  (i.e.
+correlation close to 1) or whether they are different aspects of the same phenomenon.
 As @son_hing_merit_2011-1 have pointed out, "People can believe that outcomes ought to be distributed on the basis of merit and yet vary in their perceptions of whether this is how society currently operates" (p.
 435).
-In other words, normative beliefs should be interpreted in the context of perception: a large normative belief in meritocracy certainly means something totally different for someone perceiving high meritocracy than for someone perceiving low meritocracy.
+In other words, normative beliefs should be interpreted taken perception into account: a large normative belief in meritocracy certainly means something totally different for someone perceiving high meritocracy than for someone perceiving low meritocracy.
 In order to avoid the confusion generated by the term "belief", we propose the terms meritocratic preferences and meritocratic perceptions, as they better reflect the two dimension under scrutiny.
@@
 A different approach was followed by @kunovich_systems_2007, who decide to include some non-meritocratic elements.
-Using the items’ battery listing a number of reasons about “How important should be in deciding pay...” (as @duru-bellat_whos_2012 for desired meritocracy), he decided that reasons as education and responsibility are meritocratic and pointed 1 if considered essential, whereas reasons such as having a family and children were pointed 1 if they were considered "not important at all" (i.e.
+Using the items’ battery listing a number of reasons about “How important should be in deciding pay...” (as @duru-bellat_whos_2012 for desired meritocracy), they decided that reasons as education and responsibility are meritocratic and pointed 1 if considered essential, whereas reasons such as having a family and children were pointed 1 if they were considered "not important at all" (i.e.
 reverse coded).
-A similar approach was taken by @newman_false_2015, reverse-coding non-meritocratic items, similarly to what occurs with the "Preference for the Merit Principle Scale" [@davey_preference_1999].
+A similar approach was taken by @newman_false_2015, reverse-coding non-meritocratic items, the same principle applied in the "Preference for the Merit Principle Scale" [@davey_preference_1999].
 
-The assumption that meritocratic and non-meritocratic elements are the poles of the same continuum was analyzed by @reynolds_perceptions_2014 using the same "get ahead" perceptions' battery items.
+The assumption that meritocratic and non-meritocratic elements are the poles of the same continuum was analyzed by @reynolds_perceptions_2014 using the  "get ahead" perceptions' battery items mentioned above.
 They consider education, ambition and hard work as meritocratic and other reasons such as wealthy family and right connections and non-mertitocratic.
@@
 Finally, most of the studies in meritocracy so far have not incorporated the issue of measurement error [@brownBookReviewTimothy2010, @bollenStructuralEquationsLatent1989], using single indicators and/or simple average indexes for measuring meritocracy.
-Such strategy assumes that the construct is measured perfectly by the indicators chosen, going as far as proposing that "...
+Such strategy assumes that the latent construct is measured perfectly by the indicators chosen, going as far as proposing that "...
 In choosing this strategy of index construction, we argue that *support for meritocracy is not a latent variable* [@kunovich_systems_2007 p.653-654].
-Some advances were done by @reynolds_perceptions_2014 by doing a principal component analysis of meritocratic and non-meritocratic dimensions, but somewhat contradictorily they end up in a sum index despite proving multidimensionality.
+Some advances were done by @reynolds_perceptions_2014 by doing a principal component analysis of meritocratic and non-meritocratic dimensions, but somewhat contradictorily, they end up in a sum index despite proving a multidimensional latent structure.
 
@@
 
-Based on the previous limitations and confusions in the measurement of meritocracy presented in the previous section, in this paper we propose and test an instrument with the following characteristics:
+Based on the previous limitations and assumptions in the measurement of meritocracy presented in the previous section, in this paper we propose and test an instrument with the following characteristics:
 
-- Multidimensional, incorporating previous distinctions between preferences and perceptions as well as between meritocratic and non-meritocratic aspects.
+- _Multidimensional_, incorporating previous distinctions between preferences and perceptions as well as between meritocratic and non-meritocratic aspects.
 
-- Multiple indicators for each dimension, in order to account for measurement error in a confirmatory factor analysis context.
+- Multiple indicators for each dimension, in order to _account for measurement error_ in a confirmatory factor analysis context.
 
-- Based on previous indicators as far as possible, for the sake of comparability between studies
+- Based on _previous indicators_ as far as possible, for the sake of comparability between studies
 
-- Brief, as to be used in a regular survey.
+- _Brief_, as to be used in regular public opinion surveys.
 In this point it differs for instance from the proposal of the "Preference for the Merit Principle Scale" [@davey_preference_1999], as they use 15 items just for one dimension (besides the problem of reverse-coding non-meritocratic items).
@@
 The columns Perceptions and Preferences represent the distinction between this two concepts, usually confused under the label "beliefs".
-Perceptions refers to the extent to which people see that meritocracy functions or applies in their society, which in terms of measurement relates to items such as "I think hard work is important to get ahead in society", whereas preferences refer to normative expectations that are usually linked to a "should" expression (e.g.
+Perceptions refers to the extent to which people observe that meritocracy functions or applies in their society, which in terms of measurement relates to items such as "I think hard work is important to get ahead in society", whereas preferences refer to normative expectations that are usually linked to a "should" expression (e.g.
 whether hard work should be related to payment).
@@
 
-The research hypotheses behind this conceptualization are the following:
+The research hypotheses behind this conceptualization and measurement model are the following:
 
@@
 The proposed scale of perceptions and preferences about meritocracy consists of 8 indicators that are grouped into the 4 dimensions referred above: Perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic).
-In order to achieve at least some comparability with previous studies, the questions were adapted from the items battery "reasons to get ahead" (ISSP/GSS), which are mostly used for operationalizing meritocracy in previous empirical studies [@mijs_paradox_2019; @duru-bellat_whos_2012; @reynolds_perceptions_2014].
+In order to achieve at least some comparability with previous studies, the questions were adapted from the items battery "reasons to get ahead" (ISSP/GSS), which are mostly used for operationalizing meritocracy [@mijs_paradox_2019; @duru-bellat_whos_2012; @reynolds_perceptions_2014].
 The eight items ordered according to dimensions are presented in Table \@ref(tab:table-indicadores).
@@
 
-With the objective of evaluating the effect of the indicators ordering, the respondents (_n = 2141_) were randomly divided into three different order versions as explained in Figure \@ref(fig:appmod) .
+With the objective of evaluating the effect of the indicators ordering, the respondents (_n = 2141_) were randomly divided into three different order versions as explained in Figure \@ref(fig:appmod).
 The scale was presented to the first group (_n = 712_) in the order that appears in Table 2.
@@
 
-Considerando la ventaja del orden, respecto aleatorio despeja la posibilidad del efecto framing, es decir, de que el resultado del modelo se deba al orden de las preguntas, se ha decidido seguir utilizando el orden 3.
+Studies that attempt to characterize and compare societies regarding support for meritocratic beliefs have used different approaches.
+As most studies use secondary survey data, they tend to grasp in the available indicators assuming that they represent an underlying construct.
+Nevertheless, a review of these studies reveal several non-tested assumptions, as well as the use of similar indicators to represent different constructs and dimensions of meritocracy.
+As the existence of heterogeneous approaches certainly have consequences for advancing in the study of meritocracy, this paper presented a comprehensive conceptual framework for the empirical study of meritocracy built upon previous research, which is then tested again new survey data.
 
-uso conjunto percepciones-preferencias
+The results of the analysis indicate that is possible to distinguish between different dimensions in the study of meritocracy based on the proposed 8-items scale Perceptions And Preferences for Meritocracy.
+On the one hand, perceptions and preferences proved to be two related but different dimensions, usually confounded in previous research under the label of "beliefs".
+On the other hand, meritocratic and non-meritocratic dimensions are not poles of a same continuum as previosuly assumed.
+
+The four-dimensional framework and structure of the Perceptions and Preferences for Meritocracy scale opens several avenues for future research.
+For instance, distinguishing perceptions from preferences will allow to evaluate the extent to which different societies are accustomed or satisfied with the performance of meritocracy, in terms of difference of what is perceived from what is preferred.
+On the other hand, given that non-meritocratic elements are not necessarily related to the meritocratic ones, this open the door to assess the legitimacy of practices such as the use of personal contacts and their interference (or not) with meritocratic ideals in different societies.
+The consideration of structural factors, both individual and societal, will allow to advance in hypotheses of meritocratic legitimacy from a more comprehensive view than the used today based only in meritocratic perceptions.
+Furthermore, the impacts of different configurations of the four dimensional framework on practices and behaviors such as corruption, civic involvement and political views is an area that requires additional research particularly in times of economic crisis and growing inequality, and to which the proposed instrument aims at contributing.
 

### 2020-09-30 docs/paper.Rmd

@@
 Desde la psicología social y la sociología se han estudiado las características y consecuencias de las creencias en la meritocracia, en general basados en la hipótesis que mayor creencia en la meritocracia lleva a una mayor legitimación de las desigualdades [@premingerMeritocracyServiceEthnocracy2020; @trumpWhenWhyEconomic2020; @hadjar_meritokratie_2008; @MadeiraPrimesConsequencesSystematic2019].
-Tales investigaciones han planteado críticas a la realización de este estándar moral de distribución, argumentando que sería una promesa incumplida dada la influencia preponderante de otros elementos más allá del mérito en el estatus individual [@witteveenReconsideringMeritocraticPower2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019, @khan_Privilege_2013].
+Tales investigaciones han planteado críticas a la realización de este estándar moral de distribución, argumentando que sería una promesa incumplida dada la influencia preponderante de otros elementos más allá del mérito en el estatus individual [@witteveenReconsideringMeritocraticPower2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019; @khan_Privilege_2013].
 
@@
 
-Finally, most of the studies in meritocracy so far have not incorporated the issue of measurement error [@brownBookReviewTimothy2010, @bollenStructuralEquationsLatent1989], using single indicators and/or simple average indexes for measuring meritocracy.
+Finally, most of the studies in meritocracy so far have not incorporated the issue of measurement error [@brownBookReviewTimothy2010; @bollenStructuralEquationsLatent1989], using single indicators and/or simple average indexes for measuring meritocracy.
 Such strategy assumes that the latent construct is measured perfectly by the indicators chosen, going as far as proposing that "...
@@
 ```{r meas02, echo=FALSE, fig.cap = "Confirmatory factor analysis of the Perceptions an Preferences for Meritocracy Scale", fig.align='center'}
-knitr::include_graphics('../output/images/meas02.png')
+knitr::include_graphics('../output/images/meas03.png')
 ```
@@
 Por su parte, la escala ordenada por temáticas parece generar un efecto framing en el cual la relación entre las percepciones y las preferencias parecen sobreestimadas, afectando de esta manera el ajuste (CFI=0.984, TLI=0.968, RMSEA=.071, $\chi2$(df=14)=64.156, p < 0.001).
+
+```{r latcor, echo=FALSE, fig.cap = "Latent variables correlation", fig.align='center'}
+knitr::include_graphics('../output/images/latcor.png')
+```
 

### 2020-09-30 docs/prod_analysis-cfa.rmd

--- 2020-09-10 production/prod_analysis-cfa.rmd
@@
 
-```{r}
+```{r meas02}
 
@@
 
+```{r meas03-0}
+nodeNames <-c("Perception effort","Perception talent","Perception rich family","Perception contacts",
+              "Preference effort","Preference talent","Preference rich family","Preference contacts",
+              "Perception \n meritocratic","Perception \n non meritocratic","Preference \n meritocratic","Preference \n non meritocratic")
+
+graphics.off()
+fm03<- data.frame(v1=fitmeasures(fit4_o, output ="matrix")[c("chisq.scaled","df","cfi.scaled","rmsea.scaled"),])
+fm03 <- round(fm03,3)
+par(mai = c(2,2,2,2)) # Set the margin on all sides to 2
+par(mar = c(5, 5, 5, 5)) # Set the margin on all sides to 6
+layout(matrix(c(1, # semPlot
+                1, # semPlot
+                1,
+                1,
+                1,
+                2),# ajuste
+              nrow=6,
+              byrow=TRUE))
+
+layout.show(n = 2)
+semPaths(
+  fit4_o ,
+  # whatLabels = "std", # Indica si aparecen los valores
+  what = "mod",
+  label.cex = 1.3, # Tamaño de las etiquetas dentro de los nodos.
+  edge.label.cex =  0.8, #  0.8 Tamaño de los valores estimados.
+  residuals = T,
+  optimizeLatRes =F,
+  edge.color = "black",
+  style = "lisrel",
+  nCharNodes = 0,
+  curvePivot = F,
+  curve = 2.5,
+  rotation = 4,
+  layout = "tree2",
+  cardinal = "lat cov",
+  legend.cex = 0.6, # 0.6
+  label.cex = 1,
+  label.font = 6,
+  edge.label.font = 14,
+  asize = 3,
+  edge.width = 1.2,
+  sizeMan = 8,# largo manifest
+  sizeMan2= 8,# alto manifest
+  sizeLat = 5.5, # largo de los circulos
+  sizeLat2 = 25, # alto de los circulos
+  residScale = 10,
+  width = 21.5, # 17
+  height = 10, # 10
+  # nodeNames = nodeNames,
+  intercepts = F,
+  reorder = T,
+  thresholds = F,
+  fixedStyle =1,
+  node.height = 1,
+  node.width = 4,
+  label.scale = F,
+  shapeMan = "rectangle",
+  shapeLat = "ellipse",
+  nodeLabels = nodeNames,
+  details = T)
+
+ld<- standardizedsolution(fit4_o) %>% select(lhs,op,rhs,est.std) %>% filter(op=="=~") 
+ld$est.std<- sprintf("%.2f", ld$est.std)
+
+#pegar cargas factoriales al plot
+text(x = -0.1,y =  1.00, ld$est.std[1],font = 5,cex = 1.20)
+text(x = -0.1,y =  0.85, ld$est.std[2],font = 5,cex = 1.20)
+text(x = -0.1,y =  0.43, ld$est.std[3],font = 5,cex = 1.20)
+text(x = -0.1,y =  0.26, ld$est.std[4],font = 5,cex = 1.20)
+text(x = -0.1,y = -0.15, ld$est.std[5],font = 5,cex = 1.20)
+text(x = -0.1,y = -0.30, ld$est.std[6],font = 5,cex = 1.20)
+text(x = -0.1,y = -0.73, ld$est.std[7],font = 5,cex = 1.20)
+text(x = -0.1,y = -0.88, ld$est.std[8],font = 5,cex = 1.20)
+
+library(draw) #para hacer rectangulo de fit
+
+adj_y <- 0.6
+adj_x <- 0.2
+drawBox(x = 3.5, y = 1.2-adj_y, width = 7.5, height = 2,fillColor = "grey",opacity = 0)# marca espacio
+drawBox(x = 3.5, y = 1.4-adj_y, width = 5.5, height = 1,fillColor = "grey",opacity = 0.5) # crea rectangulo
+
+# paste0("Estimator DWLS", ", N=",nobs(fit4_o))
+# paste0("Model fit: ", "χ²","(",fm03[2,],")=",fm03[1,],"***","; CFI=",fm03[3,],"; RMSEA=",fm03[4,])
+# paste0("***p<0.001")
+
+#pegar texto fit en rectangulo
+
+size <- 10
+
+drawText(x = 1.4+adj_x, y = 1.75-adj_y, text = paste0("Estimator: DWLS", ", N=",nobs(fit4_o)),family = "serif",size = size)
+drawText(x = 2.3+adj_x, y = 1.59-adj_y, text = paste0("Model fit: ", "χ²","(",fm03[2,],")=",fm03[1,],"***","; CFI=",fm03[3,],"; RMSEA=",fm03[4,]),family = "serif",size = size)
+drawText(x = 1.0+adj_x, y = 1.43-adj_y, text = paste0("***p<0.001"),family = "serif",size = size)
+
+#guardar en png
+drawExport("output/images/meas03.png",units = "cm",width = 18,height = 15,ppi = 300) 
+```
+
+```{r lat-corr}
+latcor<- lavInspect(fit4_o,what = "cor.lv")
+windowsFonts(A = windowsFont("Times New Roman"))
+rownames(latcor) <-c(
+  "A. Meritocratic Perception",
+  "B. Non-Meritocratic Perception",
+  "C. Meritocratic Preferences",
+  "D. Non-Meritocratic Preferences"
+)
+colnames(latcor) <-c("(A)", "(B)","(C)","(D)")
+png("output/images/latcor.png",
+    width = 18,height = 13,units="cm",
+    pointsize=12,bg="white",res=300)
+corrplot(
+  latcor,
+  method = "color",
+  type = "upper",
+  tl.col = "black",
+  addCoef.col = "black",
+  diag = TRUE,
+  family = "A", 
+  number.font = 6,
+  tl.cex =0.75,
+  number.cex = 1)
+dev.off()
+```
+
 ```{r tabla04, include=FALSE}

### 2020-09-30 docs/paper.Rmd
paper para proofreading

--- 2020-09-15 docs/paper.Rmd
@@
 author: "Juan Carlos Castillo, Julio Iturra & Francisco Meneses <br><br>"
-abstract: "Economic and social inequality have raised growing concerns and crises across societies.
-One of social science concepts associated to the maintenance of inequality is the belief in meritocracy, which would legitimize economic differences based on criteria such as effort and talent.
-Despite its wide use, empirical research on meritocracy is something relatively novel.
-A number of studies have relied mostly in secondary data to operationalize meritocracy, with a large variation in the use and interpretation of survey items.
+abstract: "Economic and social inequality have raised growing concerns and crises across contemporary societies.
+One of the mechanisms developed by the social sciences to explain the maintenance of inequality is the belief in meritocracy, which would legitimize economic disparities based on differences of effort and talent.
+Despite its wide use, empirical research on meritocracy is relatively novel and since it is based on multiple approaches and techniques its analysis can be vague or confusing.
+Particularly problematic is that a number of studies have relied mostly on secondary data to operationalize meritocracy, with a large variation in the use and interpretation of survey items.
 Starting from a review of studies that measure meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding the conceptualization and indicators of meritocracy.
@@
 
-# Introducción
+# Introduction
 
-La desigualdad económica se ha vuelto un tema que genera creciente preocupación y malestar alrededor del mundo.
-Esto se ha expresado en una serie de protestas como la emblemática "occupy wall street" el año 2011, así como también en una serie de análisis críticos respecto del desarrollo del capitalismo y sus consecuencias [@piketty_Capital_2014].
-En este contexto, el estudio de las visiones, preferencias y percepciones respecto de la desigualdad han adquirido relevancia en las ciencias sociales, en temas como las preferencias redistributivas [@alesina_Fairness_2005; @dimick_Models_2018] la legitimación de la desigualdad económica [@schroder_Income_2017]  y el funcionamiento de la meritocracia [@duru-bellat_whos_2012; @mijs_paradox_2019; @reynolds_perceptions_2014].
+Economic inequality has become an issue of growing concern around the world.
+This has been expressed in a series of protests, as well as diverse critical analyses regarding the development of capitalism and its consequences [Streeck, 2014; @piketty_Capital_2014].
+In this context, the study of visions, preferences and perceptions of inequality has acquired relevance in the social sciences, in topics such as redistributive preferences [@alesina_Fairness_2005]; Dimick_Models_2018], the legitimization of economic inequality [@schroder_Income_2017] and the functioning of meritocracy [@duru-bellat_whos_2012; @mijs_paradox_2019; @reynolds_perceptions_2014; Atria et al., 2020].
 
-En general, la meritocracia se define como un sistema de distribución de recursos y recompensas basados en el mérito individual, que en su concepción original es una suma de talento y esfuerzo [@young_rise_1962].
-Esta concepción tradicional de mérito sitúa en un lugar secundario la posible interferencia de factores estructurales o no meritocráticos como la herencia, los contactos personales, o la suerte [@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995; @yairMeritocracy2007; @landWeSatTable2006; @youngRiseMeritocracy1994].
-Desde la psicología social y la sociología se han estudiado las características y consecuencias de las creencias en la meritocracia, en general basados en la hipótesis que mayor creencia en la meritocracia lleva a una mayor legitimación de las desigualdades [@premingerMeritocracyServiceEthnocracy2020; @trumpWhenWhyEconomic2020; @hadjar_meritokratie_2008; @MadeiraPrimesConsequencesSystematic2019].
-Tales investigaciones han planteado críticas a la realización de este estándar moral de distribución, argumentando que sería una promesa incumplida dada la influencia preponderante de otros elementos más allá del mérito en el estatus individual [@witteveenReconsideringMeritocraticPower2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019, @khan_Privilege_2013].
+In general, meritocracy is defined as a system of distribution of resources and rewards based on individual merit, which in its original conception is a sum of talent and effort [@young_rise_1962].
+This traditional conception of merit places in a secondary position the possible interference of structural or nonmeritocratic factors such as inheritance, personal contacts, or luck [@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995; @yairMeritocracy2007; @landWeSatTable2006; @youngRiseMeritocracy1994].
+Social psychology and sociology have studied the characteristics and consequences of beliefs in meritocracy, in general with the hypothesis that greater belief in meritocracy leads to greater legitimization of inequalities [@premingerMeritocracyServiceEthnocracy2020; @trumpWhenWhyEconomic2020; @hadjar_meritokratie_2008; @MadeiraPrimesConsequencesSystematic2019].
+Such research has raised criticism of the realization of this moral standard of distribution, arguing that it would be an unfulfilled promise given the preponderant influence of other elements beyond merit on individual status [Sandel, 2020; @witteveenReconsideringMeritocraticPower2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019, @khan_Privilege_2013].
 
-Debido al rol que cumplen las creencias meritocráticas dentro del pensamiento neoliberal [@bay-chengTrackingHomoOeconomicus2015], han surgido múltiples investigaciones que evalúan la relación entre creencias meritocráticas y diversos ámbitos sociales de la actualidad.
-Por ejemplo, se han desarrollado estudios que vinculan la meritocracia al reforzamiento de estereotipos socioeconómicos, de género y de etnias [@MadeiraPrimesConsequencesSystematic2019; @girerdNeoliberalismIdeologicalBarrier2020; @premingerMeritocracyServiceEthnocracy2020], así como también líneas de investigación que evalúan el efecto de las creencias meritocráticas en el contexto educativo [@generettStoriesWeTell2020; @owensENGINESSOCIALMOBILITY2020] y en el contexto organizacional de las empresas [@perezAdvancingCareersMerit2020; @aielloNewEvidenceFirmuniversity2019].
+Due to the role that meritocratic beliefs play as a justification for individual achievement or failure in contemporary societies [@bay-chengTrackingHomoOeconomicus2015], multiple investigations have emerged that evaluate the relationship between meritocratic beliefs and diverse social circumstances today.
+For example, studies have been developed that link meritocracy to the reinforcement of socio-economic, gender and ethnic stereotypes [@MadeiraPrimesConsequencesSystematic2019; @girerdNeoliberalismIdeologicalBarrier2020; premingerMeritocracyServiceEthnocracy2020], as well as lines of research that evaluate the effect of meritocratic beliefs in the educational context [@generettStoriesWeTell2020; @owensENGINESSOCIALMOBILITY2020] and in the organizational dimension of companies [@perezAdvancingCareersMerit2020; @aielloNewEvidenceFirmuniversity2019].
 
-Para poder dar cuenta de los niveles de creencia en la meritocracia los estudios a la fecha generalmente han utilizado algunos indicadores de encuestas ya existentes, y en menor medida se han creado instrumentos ad-hoc.
-Sin embargo, y como mostraremos más adelante, las formas de medición de meritocracia varían extremadamente entre estudios.
-Muchas veces fenómenos similares se asocian a indicadores distintos, y también ocurre que fenómenos distintos son medidos con indicadores similares, todo lo cual dificulta la comparabilidad entre investigaciones y avanzar en la comprensión y estudio de la meritocracia.
+To account for the levels of belief in meritocracy, studies have generally used some indicators from existing surveys, and to a lesser extent have created ad-hoc instruments.
+However, as we will show later, the forms of measuring meritocracy vary extremely among studies.
+Many times similar phenomena are associated with different indicators, and it also happens that different phenomena are measured with similar indicators, all of which make difficult not only the comparability between studies but also to acheive more precision in understanding the effects of meritocratic beliefs in different fields.
 
-Basados en el análisis crítico de las formas de medición de meritocracia a la fecha, el presente artículo propone un instrumento para medir y relacionar dos aspectos claves en el estudio de la meritocracia: percepciones y preferencias.
-Además, como un segundo eje de análisis considera la generación de indicadores respecto de aspectos meritocráticos y anti-meritocráticos, demostrando que no son los dos polos de un mismo continuo como muchos estudios anteriores parecen sugerir.
-La propuesta de medición además está orientada a generar un instrumento lo más breve posible de manera que pueda ser utilizado en encuestas de opinión pública y así ser estudiado en relación a otros fenómenos sociales.
-
-# La medición de los aspectos subjetivos de la meritocracia
-
-A continuación se presenta una revisión de una serie de investigaciones que se han abocado al estudio de la meritocracia y que para ello han hecho una propuesta de medición.
-El primer eje de análisis tiene que ver con el uso del concepto "creencias" para referir a distintos aspectos subjetivos relacionados con meritocracia.
-El segundo eje tiene que ver con el uso de indicadores sobre aspectos anti-meritocráticos como el polo opuesto de los meritocráticos.
+Based on the critical analysis of different approaches to the measurement of meritocracy to date, this article proposes a tool to measure and relate two key aspects in the study of meritocracy: perceptions and preferences.
+Furthermore, as a second axis of analysis it considers the elaboration of indicators regarding meritocratic and non-meritocratic aspects, demonstrating that they are not the two poles of the same continuum as many previous studies seem to suggest.
+The measurement proposal is also oriented to generate an instrument as brief as possible so that it can be used in public opinion surveys, integrating meritocratic beliefs in the study of different social phenomena.
 
@@
 
-Several approaches to the empirical study of meritocracy based on public opinion surveys make reference to the concept of _beliefs_, but behind this concept there are usually different meanings and operationalizations.
-To illustrate this point in the following we will start with the proposal from a recent paper by @mijs_paradox_2019, which we will take as a reference to discuss previous studies on meritocracy.
+Several approaches to the empirical study of meritocracy based on public opinion surveys refer to the concept of _beliefs_, but behind this concept there are usually different meanings and operationalizations.
+To illustrate this point in the following we will take as a main reference a recent paper about meritocracy by @mijs_paradox_2019, which makes several assumptions about the measurement of meritocracy that are also present in previous studies.
 
@@
 In the operationalization, this is associated with the following indicator: "how important you think it is for getting ahead in life: (a) hard work", scored in a 1 to 5 likert scale.
-There are several assumptions behind this decision that are worth discussing and that are related to previous assumptions in the study of meritocracy.
+There are several assumptions behind this decision that are worth discussing in light of the conceptual meaning of meritocracy and its possibilities of operationalization.
 
-a. _Dimensionality_
+a. _Unidimensionality_
 
 The item used by @mijs_paradox_2019 is part of an items' battery present in several international surveys, usually called "reasons to get ahead".
-This battery presents a series of indicators related to what people consider important to get ahead in life: hard work, education, ambition, wealthy family, right connections, religion, race and gender.
+This battery displays a series of indicators related to what people consider important to get ahead in life: hard work, education, ambition, wealthy family, right connections, religion, race and gender.
 Therefore, for Mijs other aspects such as education, that could be associated to talent, are not meritocratic.
@@
 
-In Mijs' proposal we can observe a couple of strong assumptions: effort would not depend on parents influence, and talent is not meritocratic (contrary to Michael Youngs original conceptualization).
+In this measure of meritocratic belief we can observe a couple of strong assumptions: effort would not depend on parents influence, and talent is not meritocratic (contrary to Michael Youngs original conceptualization).
 The problem of whether talent is or not meritocratic is certainly an interesting point to discuss, but for those working based on empirical data it is something that should be empirically tested.
@@
 This conceptual and measurement unidimensional assumption it is possible to find in other studies that assume that effort is the main and only aspect of meritocracy [@girerdNeoliberalismIdeologicalBarrier2020a; @bubakPerceptionsMeritocracyNote2019].
+Recently, Mijs & Hoy (2020) considered a measurement of talent as an indicator of meritocracy.
 
@@
 
-The "reasons to get ahead" battery refers to "how important you think it is", considered by @mijs_paradox_2019 as a belief in meritocracy for the item regarding effort.
+The most used item battery for the operationalization of meritocracy using survey data is the "reasons to get ahead", which asks about "how important you think it is" and then lists several reasons such as effort, education, parents wealth and contacts.
 Nevertheless, another version of this same battery used in several surveys is "how important you think it _should_ be".
 Therefore, the question raised here is: Which one of both is a "belief"?
-According to @janmaatSubjectiveInequalityReview2013: "Perceptions refer to subjective estimates of existing inequality (i.e.
+
+The term belief has an ambiguous character in the literature, conceived as "idea-elements" for Converse (1964) or "considerations" for Zaller (1992).
+As Kluegel & Smith pointed out about the scope of beliefs: "This usage encompasses such more specific social-psychological concepts as values, perceptions, and attitudes" (p.30).
+Therefore, it covers almos anything related to subjective factors.
+But a relevant distinction in this field was made by @janmaatSubjectiveInequalityReview2013: "Perceptions refer to subjective estimates of existing inequality (i.e.
 thoughts about what is).
@@
 thoughts about what should be)"(p.359).
-Therefore, the referred paper would be using the term beliefs (what should be) referring actually to perceptions (what is).
-The same occurs in studies as the one by @reynolds_perceptions_2014, who explicitly use the term beliefs to talk about perceptions, whereas other authors use general terms as attitudes (@kunovich_systems_2007).
-The first attempt to shed light on this confusion was made by @duru-bellat_whos_2012, who used the item "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" for "desired" meritocracy (beliefs), whereas for "perceived" meritocracy they use two items: “Would you say that in your country, people are rewarded for their efforts?” and “...
+Several papers dealing with meritocracy use the term beliefs (what should be) referring actually to perceptions (what is).
+This happens not only in mijs_paradox_2019, but also in @reynolds_perceptions_2014, who explicitly use the term beliefs to talk about perceptions, whereas other authors use general terms as attitudes (@kunovich_systems_2007).
+The first attempt to shed light on this issue was made by @duru-bellat_whos_2012, who used the item "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" for "desired" meritocracy (beliefs), whereas for "perceived" meritocracy they used two items: “Would you say that in your country, people are rewarded for their efforts?” and “...
 people are rewarded for their skills?”.
 
-Is the belief in meritocracy a perception or a desire/preference?
+Is the belief in meritocracy a perception or a preference with normative meaning?
 In order to expand the analytical conceptual framework, we believe that both dimensions should be included in the analysis, as proposed by @duru-bellat_whos_2012.
@@
 435).
-In other words, normative beliefs should be interpreted taken perception into account: a large normative belief in meritocracy certainly means something totally different for someone perceiving high meritocracy than for someone perceiving low meritocracy.
-In order to avoid the confusion generated by the term "belief", we propose the terms meritocratic preferences and meritocratic perceptions, as they better reflect the two dimension under scrutiny.
+In other words, normative beliefs should be interpreted taking perceptions into account: a large normative belief in meritocracy certainly means something totally different for someone perceiving high meritocracy than for someone perceiving low meritocracy.
+Avoiding the confusion generated by the term "belief", we propose the terms meritocratic preferences (normative dimension) and meritocratic perceptions (factual dimention), as they better reflect the two faces of meritocracy under scrutiny.
 
@@
 
-@mijs_paradox_2019 makes reference to some non-meritocratic aspects as talent, which is ruled out of the operationalization of meritocracy.
-A different approach was followed by @kunovich_systems_2007, who decide to include some non-meritocratic elements.
-Using the items’ battery listing a number of reasons about “How important should be in deciding pay...” (as @duru-bellat_whos_2012 for desired meritocracy), they decided that reasons as education and responsibility are meritocratic and pointed 1 if considered essential, whereas reasons such as having a family and children were pointed 1 if they were considered "not important at all" (i.e.
+Unlike studies that consider talent a non-meritocratic aspect and therefore rule it out of the operationalization of meritocracy, @kunovich_systems_2007 follow a different approach by including some non-meritocratic elements.
+Using the items’ battery listing a number of reasons about “How important should be in deciding pay...” (as @duru-bellat_whos_2012 for desired meritocracy), they argue that reasons such as education and responsibility are meritocratic and pointed 1 if considered essential, whereas reasons such as having a family and children were pointed 1 if they were considered "not important at all" (i.e.
 reverse coded).
@@
 The assumption that meritocratic and non-meritocratic elements are the poles of the same continuum was analyzed by @reynolds_perceptions_2014 using the  "get ahead" perceptions' battery items mentioned above.
-They consider education, ambition and hard work as meritocratic and other reasons such as wealthy family and right connections and non-mertitocratic.
+They consider education, ambition and hard work as meritocratic and other reasons such as wealthy family and right connections as non-meritocratic.
 Nevertheless, despite making this distinction the author ends up subtracting one dimension from the other, assuming that they are two poles of the same continuum as @kunovich_systems_2007 did.
-Still, taking into account this research perspective, we suggest that non-meritocratic aspects should be part of a meritocratic measurement but taken independently and not adding or subtracting from meritocratic ones unless it is empirically proved that they belong to the same conceptual dimension.
+Still, taking into account this view, we suggest that non-meritocratic aspects should be part of a meritocratic measurement but taken independently and not adding or subtracting from meritocratic ones unless it is empirically proved that they belong to the same conceptual dimension.
 
@@
 
-Finally, most of the studies in meritocracy so far have not incorporated the issue of measurement error [@brownBookReviewTimothy2010, @bollenStructuralEquationsLatent1989], using single indicators and/or simple average indexes for measuring meritocracy.
+Finally, most of the studies in meritocracy so far have not incorporated the issue of measurement error [@brownBookReviewTimothy2010, @bollenStructuralEquationsLatent1989] (Ansolabehere, S., Rodden, J., & Snyder Jr, J.
+M., 2008), using single indicators and/or simple average indexes for measuring meritocracy.
 Such strategy assumes that the latent construct is measured perfectly by the indicators chosen, going as far as proposing that "...
@@
 
-Based on the previous limitations and assumptions in the measurement of meritocracy presented in the previous section, in this paper we propose and test an instrument with the following characteristics:
+Based on the previous assumptions and limitations in the measurement of meritocracy, we propose and test an instrument with the following characteristics:
 
@@
 
-The columns Perceptions and Preferences represent the distinction between this two concepts, usually confused under the label "beliefs".
-Perceptions refers to the extent to which people observe that meritocracy functions or applies in their society, which in terms of measurement relates to items such as "I think hard work is important to get ahead in society", whereas preferences refer to normative expectations that are usually linked to a "should" expression (e.g.
+The columns Perceptions and Preferences represent the distinction between these two concepts, usually confused under the label "beliefs".
+Perceptions refers to the extent to which people observe that meritocracy works or applies in their society, which in terms of measurement relates to items such as "I think hard work is important to get ahead in society", whereas preferences refer to normative expectations that are usually linked to a "should" expression (e.g.
 whether hard work should be related to payment).
 The rows in the table of Figure 1 consider the distinction between meritocratic and non-meritocratic dimensions [@reynolds_perceptions_2014].
-This aspect has been usually treated as different ends of a same continuum in part of the previous research, an assumption that requieres empirical scrutiny.
+This aspect has been usually treated as different ends of a same continuum in part of the previous research, an assumption that requires empirical scrutiny.
 These non-meritocratic elements usually refer to the use of personal contacts or family advantages to get ahead in life.
@@
 Regarding the selection of indicators, most of them are taken or adapted from previous studies for the sake of comparability.
-For meritocratic indicators we use effort and talent as the main components of the traditional concept of merit as defined by @young_rise_1962, whereas for non-meritocratic dimensions we use having rich parents and good contacts.
+For meritocratic indicators we use effort and talent as the main components of the traditional concept of merit as defined by @young_rise_1962 (Atria et al.
+2020) , whereas for non-meritocratic dimensions we use having rich parents and good contacts.
 The description of the specific items is presented in the methodology section.
@@
 
-* $H_1$. The perception of meritocracy is a latent variable based on indicators of the importance attributed to talent and the effort to get ahead in life.
+* $H_1$. The perception of meritocracy is a latent variable based on indicators of the importance attributed to talent and effort to get ahead in life.
 
@@
 
-The data was obtained through an online questionnaire which was part of a larger study on meritocracy and preferences developed in Chile in 2019 and funded by the national scientific agency FONDECYT.
-The questionnaire was programmed in Qualtrics and the fieldwork was in charge of an external online survey agency ([netquest.cl](www.netquest.cl)) during December 2019 and January 2020.
+The data was obtained through an online questionnaire which was part of a larger study on meritocracy and preferences developed in Chile in 2019 and funded by the national scientific agency (ANID).
+The questionnaire was programmed in Qualtrics and the fieldwork was in charge of an external online survey agency ([netquest.cl](www.netquest.cl)) between December 2019 and January 2020.
 The sample was selected from a non-probabilistic quota design in three large cities in Chile.
-The quotas were generated based on the survey of the Public Studies Center (CEP, 2019), which has a high prestige in the country and is also the counterpart agency of ISSP (International Social Survey Programme) in Chile.
+The quotas for gender, age and education level were generated based on the survey of the Public Studies Center (CEP, 2019), which has a high prestige in the country and is also the counterpart agency of ISSP (International Social Survey Programme) in Chile.
 A total sample of 2,141 people was collected, excluding those who did not answer the questions on the scale and those who did not accept informed consent.
-As it usually occurs online samples, there were some limitations in achieving the quotas for lower educational levels.
+There are not significant differences between our effective sample and the population for most socio-demographics characteristics, with the exception of educational level (see Table XX in appendix).
+As it usually occurs with online samples, there were some limitations in achieving the quotas for lower educational levels.
 
@@
 
-With the objective of evaluating the effect of the indicators ordering, the respondents (_n = 2141_) were randomly divided into three different order versions as explained in Figure \@ref(fig:appmod).
+With the objective of evaluating the effect of the indicators ordering, the respondents (_n = 2141_) were randomly divided into three different set-order versions as explained in Figure \@ref(fig:appmod).
 The scale was presented to the first group (_n = 712_) in the order that appears in Table 2.
 For the second group (_n = 717_), the order of the items was organized according to the topics of the items, e.g.
-for the topic of hard work the item about perception was followed by the item about preference, and te same for the rest of the topics.
+for the topic of hard work the item about perception was followed by the item about preference, and the same for the rest of the topics.
 Finally, for the third group (_n = 712_) the items were completely randomized.
@@
 A pre-registration was made in the OSF platform, available in the following link: [https://osf.io/z45y2](https://osf.io/z45y2).
-In this pre-registration are the hypothesis regarding the four-dimensional conceptual model underlying the scale,  the variables measurement levels, the statistical tests to be performed with their respective evaluation parameters, along with other important aspects of the research design.
+In this pre-registration are the hypothesis regarding the four-dimensional conceptual model underlying the scale, the variables measurement levels, and the statistical tests to be performed with their respective evaluation parameters along with other important aspects of the research design.
 
@@
 
-Como puede observarse en la tabla \@ref(tab:desc01), los indicadores poseen valores que van desde el 1 (totalmente en desacuerdo) al 5 (totalmente de acuerdo).
-Se observan promedios desde 2.41 correspondiente a preferencia-contactos, hasta 3.89 correspondiente a la preferencia-esfuerzo.
-Ambos indicadores son coherentes con la adhesión general a la meritocracia reportada en estudios previos, privilegiando aspectos individuales como el esfuerzo.
+As can be seen in the table \ref(tab:desc01), the indicators have values ranging from 1 (totally disagree) to 5 (totally agree).
+Averages are observed from 2.41, corresponding to preference-contacts, to 3.89, corresponding to preference-effort.
+Both indicators are consistent with the general adherence to meritocracy reported in previous studies, privileging individual aspects such as effort (Frei et al 2020, Bucca, 2016).
 
@@
 table04b<- sum04b %>% rename("Mean"=mean, "SD"=sd, "Min"=min,"Max"=max)
-sumdat04<- kable(table04b,format =table_format, booktabs= T, escape = FALSE,align = "l",linesep = "", caption = cap
+sumdat04<- kable(table04b,format =table_format, booktabs= T, exhaust = FALSE,align = "l",linesep = "", caption = cap
 ) %>% kable_styling(full_width = table_format2,
@@
 
-Los gráficos presentados en la Figura \@ref(fig:plotlikert) presentan información desagregada y comparable de las distintas categorías de respuesta de cada item.
-Se observa que en general hay una mayor percepción de elementos no meritocráticos que meritocráticos, mientras que en el caso de preferencias ocurre lo opuesto.
-En cuanto a las preferencias, llama la atención el rol preponderante del esfuerzo por sobre el talento como criterio de preferencia meritocrático.
+The graphs presented in Figure \@ref(fig:plotlikert) present disaggregated and comparable information of the different response categories for each item.
+It can be seen that in general there is more agreement in the perception of non meritocratic items than meritocratic ones, while in the case of preferences the opposite occurs.
+As far as preferences are concerned, the preponderant role of effort over talent as a criterion of meritocratic preference is noteworthy.
 
@@
 
-En la Figura \@ref(fig:corpoly) se observan relaciones de moderada a alta intensidad entre los indicadores que corresponden al mismo factor (por ejemplo, percepción de meritocracia por esfuerzo y por talento, r=0.56).
-Destacan además las relaciones entre las percepciones y preferencias de tipo meritocráticas, lo que no ocurre con los indicadores no meritocraticos.
+In Figure Ref(fig:corpoly), moderate to high intensity relationships are observed between indicators that correspond to the same factor (for example, perception of meritocracy by effort and by talent, r=0.56).
+The relationships between meritocratic perceptions and preferences are also noteworthy, which is not the case for nonmeritocratic indicators.
 
@@
 
-En suma, los análisis descriptivos señalan una relativa adhesión a la moral meritocratica, las cuales se expresan en una mayor preferencia por criterios meritocraticos y una menor por no meritocráticos, aún cuando la preferencia por elementos no-meritocraticos alcanza una proporción de más de 20%.
-Igualmente, es observable una relativamente baja percepción de meritocracia.
-Se observa además una relación coherente entre los indicadores según lo propuesto en el modelo teórico en el preregistro del estudio, es decir, los pares de items asociados a un factor específico muestran correlaciones con un tamaño de efecto grande (por ejemplo, preferencias meritocrácticas por los items asociados a esfuerzo y a talento).
-En particular las asociaciones entre esfuerzo y talento son relevantes, ya que desestiman previos supuestos sobre que el talento no sería un criterio meritocrático [@mijs_paradox_2019], de otra manera la correlación sería cero o negativa.
-Junto a ello, vemos que no existe una correlación negativa entre aspectos meritocráticos y no meritocráticos, desestimando los supuestos de estudios previos que  señalaban que estas dimensiones serían los polos opuestos de un mismo continuo [@reynolds_perceptions_2014].
+In sum, descriptive analyses point to a relative adherence to meritocracy, which is expressed in a greater preference for meritocratic criteria and a lesser preference for non-meritocratic ones, even when the preference for non-meritocratic elements reaches a proportion of more than 20%.
+Likewise, a relatively low perception of meritocracy is observable.
+In addition, a coherent relationship between the indicators as proposed in the theoretical model is observed in the pre-registration of the study; that is, the pairs of items associated with a specific factor show correlations with a large effect size (for example, meritocratic preferences for items associated with effort and talent).
+In particular, the associations between effort and talent are relevant, since they dismiss previous assumptions that talent would not be a meritocratic criterion [@mijs_paradox_2019], otherwise the correlation would be zero or negative.
+In addition, we see that there is no negative correlation between meritocratic and non-meritocratic aspects, dismissing the assumptions of previous studies that these dimensions would be the opposite poles of the same continuum [@reynolds_perceptions_2014].
 
@@
 
-\@ref(fig:meas02) SHOWS the results of the estimation for the four-factor model with complete sample.
+\@ref(fig:meas02) shows the results of the estimation for the four-factor model with complete sample.
 This model presents adequate fit indicators (CFI = 0.977, RMSEA = 0.079, $\chi2$(df = 14) = 28,03, p = 0.014) and it presents factorial loads with weights that vary between 0.63 and 0.88.
@@
 
-Como se puede observar en la Tabla \@ref(tab:tab-loads), todos los modelos, independiente del grupo y el órden de los items, obtuvieron un ajuste adecuado.
-con CFI superiores a .95 y RMSEA inferiores a .8.
-Sin embargo ningún modelo logró un chi-square no significativo (algo esperable en muestras grandes), aunque tanto el modelo aleatorio como el primero obtuvieron un adecuado chi-square ratio menor a 3.
-El primero de los órdenes fue el que obtuvo mejor ajuste (CFI=0.998, TLI= 0.995,RMSEA=0.037, $\chi2$(df=14)=28,03, p = 0.014), seguido por el orden aleatorio de los ítems (CFI=0.992, TLI=0.984,RMSEA=0.051, $\chi2$(df=14)=39.09, p < 0.001).
-Por su parte, la escala ordenada por temáticas parece generar un efecto framing en el cual la relación entre las percepciones y las preferencias parecen sobreestimadas, afectando de esta manera el ajuste (CFI=0.984, TLI=0.968, RMSEA=.071, $\chi2$(df=14)=64.156, p < 0.001).
+As can be seen in the Table \ref(tab:tab-loads), all models, regardless of the group and order of items, obtained an adequate fit, with CFI's above .95 and RMSEA's below .8.
+However, no model achieved a non-significant chi-square (something to be expected in large samples), although both the randomized and the first model obtained an adequate chi-square ratio of less than 3.
+The first order was the one that obtained the best fit (CFI=0.998, TLI=0.995,RMSEA=0.037, $\chi2$(df=14)=28.03, p = 0.014), followed by the random order of items (CFI=0.992, TLI=0.984,RMSEA=0.051, $\chi2$(df=14)=39.09, p < 0.001).
+On the other hand, the scale ordered by themes seems to generate a framing effect in which the relationship between perceptions and preferences seems overestimated, thus affecting the adjustment (CFI=0.984, TLI=0.968, RMSEA=.071, $\chi2$(df=14)=64.156, p < 0.001).
 
 ```{r tab03, echo=FALSE,eval=TRUE}
-load(file = "../output/tables/fit4_o.RData")
-load(file = "../output/tables/fit3_o.RData")
-load(file = "../output/tables/fit2_o.RData")
-load(file = "../output/tables/fit1_o.RData")
-
-sum_fit<- bind_rows(fitmeasures(fit4_o)[c("chisq.scaled","df","cfi.scaled","rmsea.scaled")],
-                    fitmeasures(fit1_o)[c("chisq.scaled","df","cfi.scaled","rmsea.scaled")],
-                    fitmeasures(fit2_o)[c("chisq.scaled","df","cfi.scaled","rmsea.scaled")],
-                    fitmeasures(fit3_o)[c("chisq.scaled","df","cfi.scaled","rmsea.scaled")])
-sum_fit$mod <- c("Model 1","Model 2","Model 3","Model 4")
-sum_fit$est <- c("DWLS","DWLS","DWLS","DWLS")
-sum_fit$nobs <- c(nobs(fit4_o),nobs(fit1_o),nobs(fit2_o),nobs(fit3_o))
-sum_fit <- select(sum_fit,mod,nobs,est,everything())
-colnames <- c("Model","$N$","Estimator","$\\chi^2$","df","CFI","RMSEA")
-
-sumtable02<- kable(sum_fit,digits = 3,format = table_format,row.names = F,booktabs=T, align = c("lcccccc"),
-                   caption = "Summary fit indices according group",
-                   col.names = colnames,
-                   escape = FALSE) %>%
-  kable_styling(full_width = F,
-                font_size = 10)  %>%
-      footnote(number = c("Model 1: Complete sample",
-                          "Model 2: fixed order by percepction/preference",
-                          "Model 3:  fixed order by topic (i.e: effort)",
-                          "Model 4: Randomized order"));sumtable02
+(acá recuperar syntax)
 ```
 
-Si bien  todas las pruebas obtuvieron indicadores relativamente adecuados, las diferencias mencionadas entre los ajustes de los modelos según orden son estadísticamente significativas, lo cual fue evaluado a partir de un análisis de invarianza.
-Se concluye que entre los tres órdenes no existe invarianza configuracional, es decir, no poseen la misma dimensionalidad y por ello no ajustan igualmente al modelo teórico [@vandenbergReviewSynthesisMeasurement2000].
-Esto se debe al efecto producido por la aparición conjunta de los ítems de un mismo factor en el orden 1, lo cual aumenta el ajuste del modelo.
-Además, en el orden 2, al preguntarse seguidamente por la percepción y la preferencia en torno al mismo indicador, aumentan las relaciones cruzadas disminuyendo el ajuste del modelo.
-De modo coherente, el modelo aleatorio presentó un ajuste intermedio entre el orden 1 y el orden 2.
+Although all tests obtained relatively adequate indicators, the mentioned differences between the adjustments of the models according to order are statistically significant, which was evaluated from an invariance analysis.
+It is concluded that among the three orders there is no configurtional invariance, that is, they do not have the same dimensionality and therefore do not equally fit the theoretical model [@vandenbergReviewSynthesisMeasurement2000].
+This is due to the effect produced by the joint appearance of the items of the same factor in order 1, which increases the fit of the model.
+Furthermore, in order 2, when asking about the perception and preference around the same indicator, the cross relations increase and the model fit decreases.
+Consistently, the random model presented an intermediate adjustment between order 1 and order 2.
 
@@
 
-EL modelo teórico propuesto de cuatro factores ajustó, como se observa en la tabla 4, mejor que el modelo de contraste de 1 factor.
-El modelo teórico ajustó de manera relativamente adecuada, pues muestra indicadores óptimos para CFI= 0.987, TLI = 0.975 y RMSEA=.041, aunque posee indicadores deficientes para la prueba $\chi2$(df=14)=67.6, p-value=.000.
-Para evaluar posibles mejoras de la escala, se analizaron las relaciones propuestas por los índices de modificación.
-Estos indican la existencia de dos cargas cruzadas no especificadas.
-Cuando se generó un modelo siguiendo estas recomendaciones, hubo una mejora considerable del modelo, aunque el nuevo modelo tampoco obtuvo un $\chi2$ ratio menor a 3 y obtuvo cargas factoriales muy bajas ($\tau$)< 0.15, por lo tanto, siguiendo las recomendación de Brown (2006) de solo aceptar las propuesta de los índices de modificación cuando se posee teoría y evidencia sólida, se ha decidido no incorporar estos parámetros al modelo.
+As shown in table 4, the proposed theoretical four-factor model fits better than the one-factor contrast model.
+The theoretical model adjusted relatively well, as it shows optimal indicators for IFC=0.987, TLI=0.975 and RMSEA=.041, although it has poor indicators for the test $\chi2$(df=14)=67.6, p-value=.000.
+To evaluate possible improvements in the scale, the relationships proposed by the modification indexes were analyzed.
+These indicate the existence of two unspecified cross-loads.
+When a model was generated following these recommendations, there was a considerable improvement in the model, although the new model also did not obtain a $\chi2$ ratio less than 3 and obtained very low factorial loads ($\tau$)< 0.15.
+Therefore, following the recommendation of Brown (2006) to only accept the proposals of modification rates when theory and solid evidence are available, it has been decided not to incorporate these parameters into the model.
 
@@
 As most studies use secondary survey data, they tend to grasp in the available indicators assuming that they represent an underlying construct.
-Nevertheless, a review of these studies reveal several non-tested assumptions, as well as the use of similar indicators to represent different constructs and dimensions of meritocracy.
-As the existence of heterogeneous approaches certainly have consequences for advancing in the study of meritocracy, this paper presented a comprehensive conceptual framework for the empirical study of meritocracy built upon previous research, which is then tested again new survey data.
+A review of these studies reveals several non-tested assumptions, as well as the use of similar indicators to represent different constructs and dimensions of meritocracy.
+As the existence of heterogeneous approaches certainly have consequences for advancing in the study of meritocracy, this paper presented a comprehensive conceptual framework for the empirical study of meritocracy built upon previous research, which is then tested against new survey data.
 
-The results of the analysis indicate that is possible to distinguish between different dimensions in the study of meritocracy based on the proposed 8-items scale Perceptions And Preferences for Meritocracy.
-On the one hand, perceptions and preferences proved to be two related but different dimensions, usually confounded in previous research under the label of "beliefs".
-On the other hand, meritocratic and non-meritocratic dimensions are not poles of a same continuum as previosuly assumed.
+We identified four critical aspects regarding the measurement of meritocracy in previous studies: unidimensionality, the ambiguous use of “beliefs”, the use of non-meritocratic indicators as opposed to meritocratic and the consideration of measurement error.
+The proposed 8-items scale Perceptions And Preferences for Meritocracy was designed and tested in order to deal with these four issues.
+The results indicate that, on the one hand, perceptions and preferences proved to be two related but different dimensions, usually confounded in previous research under the label of "beliefs".
+On the other hand, meritocratic and non-meritocratic dimensions would not be the poles of the same continuum as some previous studies assumed.
 
 The four-dimensional framework and structure of the Perceptions and Preferences for Meritocracy scale opens several avenues for future research.
-For instance, distinguishing perceptions from preferences will allow to evaluate the extent to which different societies are accustomed or satisfied with the performance of meritocracy, in terms of difference of what is perceived from what is preferred.
-On the other hand, given that non-meritocratic elements are not necessarily related to the meritocratic ones, this open the door to assess the legitimacy of practices such as the use of personal contacts and their interference (or not) with meritocratic ideals in different societies.
-The consideration of structural factors, both individual and societal, will allow to advance in hypotheses of meritocratic legitimacy from a more comprehensive view than the used today based only in meritocratic perceptions.
+For instance, distinguishing perceptions from preferences will allow us to evaluate the extent to which different societies are accustomed or satisfied with the performance of meritocracy, in terms of difference of what is perceived from what is preferred.
+On the other hand, given that non-meritocratic elements are not necessarily related to the meritocratic ones, it makes it possible to assess the legitimacy of practices such as the use of personal contacts and their interference (or not) with meritocratic ideals in different societies.
+The consideration of structural factors, both individual and societal, will allow us to advance in hypotheses of meritocratic legitimacy from a more comprehensive view than the one used today based only on meritocratic perceptions.
 Furthermore, the impacts of different configurations of the four dimensional framework on practices and behaviors such as corruption, civic involvement and political views is an area that requires additional research particularly in times of economic crisis and growing inequality, and to which the proposed instrument aims at contributing.
@@
 Table: `r as.character(paste("(\\#tab:rep-samp)", "Representativeness of the sample"))`.
+<!--stackedit_data:
+eyJoaXN0b3J5IjpbLTkyNTUxNzA1Ml19
+-->

### 2020-10-10 docs/paper.Rmd
update-paper

@@
 author: "Juan Carlos Castillo, Julio Iturra & Francisco Meneses <br><br>"
-abstract: "Economic and social inequality have raised growing concerns and crises across contemporary societies.
-One of the mechanisms developed by the social sciences to explain the maintenance of inequality is the belief in meritocracy, which would legitimize economic disparities based on differences of effort and talent.
-Despite its wide use, empirical research on meritocracy is relatively novel and since it is based on multiple approaches and techniques its analysis can be vague or confusing.
-Particularly problematic is that a number of studies have relied mostly on secondary data to operationalize meritocracy, with a large variation in the use and interpretation of survey items.
-Starting from a review of studies that measure meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding the conceptualization and indicators of meritocracy.
-Based on this critical analysis, we propose an item's battery called *Perceptions and Preferences for Meritocracy Scale*, which is tested with confirmatory factor analysis with data from an online survey study (N=2,141).
-The results support the proposed conceptual structure which not only distinguishes between perceptions and preferences, but also between meritocratic and non-meritocratic dimensions.
-The discussion highlights the relevance of considering these different dimensions in order to advance in the study of meritocracy."
+abstract: "Economic and social inequalities have generated growing concern and crises across contemporary societies.
+One of the mechanisms proposed by the social sciences to explain the maintenance of inequality is the belief in meritocracy, which would legitimize economic disparities based on differences in effort and talent.
+Despite its wide use as a concept, empirical research on meritocracy is relatively novel, and as it is based upon multiple approaches and techniques, its interpretation can be difficult and confusing.
+Adding to these difficulties, is the fact that many such studies have relied upon secondary data to operationalize meritocracy, with significant variation in the use and interpretation of survey items.
+Based on a review of studies that measure meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding the conceptualization and indicators of meritocracy.
+Based on this critical analysis, we propose a scale named the *Perceptions and Preferences for Meritocracy Scale*, which we test with confirmatory factor analysis, using data from an online survey study (N=2,141).
+Our results support the proposed conceptual structure, which not only distinguishes between perceptions and preferences, but also between meritocratic and non-meritocratic dimensions.
+Our discussion highlights the importance of considering these different dimensions in order to advance the study of meritocracy."
 output:
@@
 This has been expressed in a series of protests, as well as diverse critical analyses regarding the development of capitalism and its consequences [Streeck, 2014; @piketty_Capital_2014].
-In this context, the study of visions, preferences and perceptions of inequality has acquired relevance in the social sciences, in topics such as redistributive preferences [@alesina_Fairness_2005]; Dimick_Models_2018], the legitimization of economic inequality [@schroder_Income_2017] and the functioning of meritocracy [@duru-bellat_whos_2012; @mijs_paradox_2019; @reynolds_perceptions_2014; Atria et al., 2020].
+In this context, the study of visions, preferences, and perceptions of inequality has increased relevance in the social sciences, in topics such as redistributive preferences [@alesina_Fairness_2005; @Dimick_Models_2018], the legitimization of economic inequality [@schroder_Income_2017] and the functioning of meritocracy [@duru-bellat_whos_2012; @mijs_paradox_2019; @reynolds_perceptions_2014; Atria et al., 2020].
 
-In general, meritocracy is defined as a system of distribution of resources and rewards based on individual merit, which in its original conception is a sum of talent and effort [@young_rise_1962].
-This traditional conception of merit places in a secondary position the possible interference of structural or nonmeritocratic factors such as inheritance, personal contacts, or luck [@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995; @yairMeritocracy2007; @landWeSatTable2006; @youngRiseMeritocracy1994].
-Social psychology and sociology have studied the characteristics and consequences of beliefs in meritocracy, in general with the hypothesis that greater belief in meritocracy leads to greater legitimization of inequalities [@premingerMeritocracyServiceEthnocracy2020; @trumpWhenWhyEconomic2020; @hadjar_meritokratie_2008; @MadeiraPrimesConsequencesSystematic2019].
-Such research has raised criticism of the realization of this moral standard of distribution, arguing that it would be an unfulfilled promise given the preponderant influence of other elements beyond merit on individual status [Sandel, 2020; @witteveenReconsideringMeritocraticPower2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019, @khan_Privilege_2013].
+In general, a meritocracy is defined as a system of distribution of resources and rewards based on individual merit, which in its original conception is a combination of talent and effort [@young_rise_1962].
+This traditional conception of merit, places in a secondary position the possible interference of structural or non-meritocratic factors, such as inheritance, personal contacts, and luck [@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995; @yairMeritocracy2007; @landWeSatTable2006; @youngRiseMeritocracy1994].
+Social psychology and sociology have studied the characteristics and consequences of beliefs in meritocracy, under the general hypothesis that a greater belief in meritocracy leads to greater legitimization of inequalities [@premingerMeritocracyServiceEthnocracy2020; @trumpWhenWhyEconomic2020; @hadjar_meritokratie_2008; @MadeiraPrimesConsequencesSystematic2019].
+Such research has raised criticism of meritocracy as a moral standard of distribution, arguing that it would be an unfulfilled promise, given the preponderant influence of non-meritocratic elements upon individual status [Sandel, 2020; @witteveenReconsideringMeritocraticPower2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019; @khan_Privilege_2013].
 
-Due to the role that meritocratic beliefs play as a justification for individual achievement or failure in contemporary societies [@bay-chengTrackingHomoOeconomicus2015], multiple investigations have emerged that evaluate the relationship between meritocratic beliefs and diverse social circumstances today.
-For example, studies have been developed that link meritocracy to the reinforcement of socio-economic, gender and ethnic stereotypes [@MadeiraPrimesConsequencesSystematic2019; @girerdNeoliberalismIdeologicalBarrier2020; premingerMeritocracyServiceEthnocracy2020], as well as lines of research that evaluate the effect of meritocratic beliefs in the educational context [@generettStoriesWeTell2020; @owensENGINESSOCIALMOBILITY2020] and in the organizational dimension of companies [@perezAdvancingCareersMerit2020; @aielloNewEvidenceFirmuniversity2019].
+Due to the role that meritocratic beliefs play in the justification of individual achievement or failure in contemporary societies [@bay-chengTrackingHomoOeconomicus2015], multiple investigations have evaluated the relationship between meritocratic beliefs and personal/contextual characteristics.
+For example, some studies have linked meritocracy to the reinforcement of socio-economic, gender, and ethnic stereotypes [@MadeiraPrimesConsequencesSystematic2019; @girerdNeoliberalismIdeologicalBarrier2020; @premingerMeritocracyServiceEthnocracy2020], as well as lines of research have evaluated the effects of meritocratic beliefs in educational [@generettStoriesWeTell2020; @owensENGINESSOCIALMOBILITY2020] and organizational contexts [@perezAdvancingCareersMerit2020; @aielloNewEvidenceFirmuniversity2019].
 
-To account for the levels of belief in meritocracy, studies have generally used some indicators from existing surveys, and to a lesser extent have created ad-hoc instruments.
-However, as we will show later, the forms of measuring meritocracy vary extremely among studies.
-Many times similar phenomena are associated with different indicators, and it also happens that different phenomena are measured with similar indicators, all of which make difficult not only the comparability between studies but also to acheive more precision in understanding the effects of meritocratic beliefs in different fields.
+To measure beliefs in meritocracy, studies have generally used indicators from existing surveys, and to a lesser extent, have created ad-hoc instruments.
+However, as we will show later, the methods of measuring meritocracy vary extensively among studies.
+In many cases, similar phenomena are associated with different indicators, and conversely, often different phenomena are measured with similar indicators.
+Such trends confound not only the comparability of studies, but also the ability to understanding the effects of meritocratic beliefs across different fields.
 
-Based on the critical analysis of different approaches to the measurement of meritocracy to date, this article proposes a tool to measure and relate two key aspects in the study of meritocracy: perceptions and preferences.
-Furthermore, as a second axis of analysis it considers the elaboration of indicators regarding meritocratic and non-meritocratic aspects, demonstrating that they are not the two poles of the same continuum as many previous studies seem to suggest.
-The measurement proposal is also oriented to generate an instrument as brief as possible so that it can be used in public opinion surveys, integrating meritocratic beliefs in the study of different social phenomena.
+Based on a critical analysis of different approaches to the measurement of meritocracy to date, this article proposes a tool to both measure and relate two key aspects in the study of meritocracy: perceptions and preferences.
+Furthermore, as a second axis of analysis, it considers the elaboration of indicators regarding meritocratic and non-meritocratic aspects, demonstrating that they are not two poles of the same continuum, as some previous studies would seem to suggest.
+The measurement system outlined is oriented to generate an instrument as simple and brief as possible, so that it can be used in public opinion surveys, thus integrating meritocratic beliefs in the study of different social phenomena.
 
@@
 
-Several approaches to the empirical study of meritocracy based on public opinion surveys refer to the concept of _beliefs_, but behind this concept there are usually different meanings and operationalizations.
-To illustrate this point in the following we will take as a main reference a recent paper about meritocracy by @mijs_paradox_2019, which makes several assumptions about the measurement of meritocracy that are also present in previous studies.
+Several approaches to the empirical study of meritocracy based on public opinion surveys, refer to the concept of _beliefs_, however different meanings and operationalizations usually underlie this concept.
+To illustrate this point, we will take as a main reference a recent paper about meritocracy by @mijs_paradox_2019, which makes several assumptions relating to the measurement of meritocracy that are also found in previous studies.
+The definition of meritocratic beliefs stated by Mijs is the following: "when I discuss meritocracy beliefs, I am referring to citizens’ belief in the importance of hard work relative to structural factors." [@mijs_paradox_2019, pg.9].
+In the subsequent operationalization, this is associated with the following question and indicator: "how important you think it is for getting ahead in life: (a) hard work", scored from 1 to 5 on a likert scale.
+There assumptions behind this definition are worth discussing in light of the conceptual meaning of meritocracy and its possibilities of operationalization.
 
-The meritocratic beliefs' definition of Mijs is the following: "when I discuss meritocracy beliefs, I am referring to citizens’ belief in the importance of hard work relative to structural factors." [@mijs_paradox_2019, pg.9].
-In the operationalization, this is associated with the following indicator: "how important you think it is for getting ahead in life: (a) hard work", scored in a 1 to 5 likert scale.
-There are several assumptions behind this decision that are worth discussing in light of the conceptual meaning of meritocracy and its possibilities of operationalization.
-
-a. _Unidimensionality_
+a. _Unidimensionality_; Is merit only effort?
 
 The item used by @mijs_paradox_2019 is part of an items' battery present in several international surveys, usually called "reasons to get ahead".
-This battery displays a series of indicators related to what people consider important to get ahead in life: hard work, education, ambition, wealthy family, right connections, religion, race and gender.
-Therefore, for Mijs other aspects such as education, that could be associated to talent, are not meritocratic.
-As he points out: "Hard work is arguably the most meritocratic part of Michael Young’s equation, ‘Merit = Intelligence + Effort’, for the simple fact that intelligence itself is conditioned by a nonmeritocratic factor: who your parents happen to be" (p.5).
+This battery displays a series of indicators related to what people consider important to get ahead in life: hard work, education, ambition, a wealthy family, the right connections, religion, race, and gender.
+Therefore for Mijs, other aspects that could be associated with talent, such as education, are not meritocratic.
+As he points out: "Hard work is arguably the most meritocratic part of Michael Young’s equation: ‘Merit = Intelligence + Effort’, for the simple fact that intelligence itself is influenced by a non-meritocratic factor: who your parents happen to be" (p.5).
 
-In this measure of meritocratic belief we can observe a couple of strong assumptions: effort would not depend on parents influence, and talent is not meritocratic (contrary to Michael Youngs original conceptualization).
-The problem of whether talent is or not meritocratic is certainly an interesting point to discuss, but for those working based on empirical data it is something that should be empirically tested.
-Is effort the only dimension behind the concept of meritocracy, or is it a multidimensional concept that includes other elements as talent (as in its original conceptualization)?
-This conceptual and measurement unidimensional assumption it is possible to find in other studies that assume that effort is the main and only aspect of meritocracy [@girerdNeoliberalismIdeologicalBarrier2020a; @bubakPerceptionsMeritocracyNote2019].
-Recently, Mijs & Hoy (2020) considered a measurement of talent as an indicator of meritocracy.
+In this measure of meritocratic belief, we can observe a couple of strong assumptions: effort would not depend on parental influence, and talent (as innate ability) is not meritocratic (contrary to Michael Youngs original conceptualization).
+The question of whether talent is or is not considered meritocratic is certainly an interesting topic to discuss from a philosophical point of view, but for those working on empirical studies it is something that should face empirical scrutiny.
+Is effort the only dimension behind the concept of meritocracy, or is it a multidimensional concept that includes other elements, such as talent (as in its original conceptualization)?
+This conceptual and measurement-based assumption of unidimensionality is found in other studies, that also assume effort to be the main and only aspect of meritocracy [@girerdNeoliberalismIdeologicalBarrier2020a; @bubakPerceptionsMeritocracyNote2019].
+Conversely, recently Mijs & Hoy (2020) also considered talent as an indicator of meritocracy.
 
@@
 
-The most used item battery for the operationalization of meritocracy using survey data is the "reasons to get ahead", which asks about "how important you think it is" and then lists several reasons such as effort, education, parents wealth and contacts.
-Nevertheless, another version of this same battery used in several surveys is "how important you think it _should_ be".
+The most commonly used survey items for the operationalization of meritocracy is the "reasons to get ahead" battery, which asks "how important you think it is" and then lists several factors, such as effort, education, parental wealth, and contacts.
+Another version of this same battery used in several surveys - sometimes along the previous one - asks about "how important you think it _should_ be".
 Therefore, the question raised here is: Which one of both is a "belief"?
+What _is_ or what _should_ be?
 
-The term belief has an ambiguous character in the literature, conceived as "idea-elements" for Converse (1964) or "considerations" for Zaller (1992).
+The term belief has an ambiguous character in the literature, conceived as "idea-elements" by Converse (1964) or "considerations" by Zaller (1992).
 As Kluegel & Smith pointed out about the scope of beliefs: "This usage encompasses such more specific social-psychological concepts as values, perceptions, and attitudes" (p.30).
-Therefore, it covers almos anything related to subjective factors.
-But a relevant distinction in this field was made by @janmaatSubjectiveInequalityReview2013: "Perceptions refer to subjective estimates of existing inequality (i.e.
+Therefore, beliefs cover almost anything related to subjective factors.
+To this regard, a relevant distinction in the field of inequality beliefs was made by @janmaatSubjectiveInequalityReview2013: "Perceptions refer to subjective estimates of existing inequality (i.e.
 thoughts about what is).
@@
 thoughts about what should be)"(p.359).
-Several papers dealing with meritocracy use the term beliefs (what should be) referring actually to perceptions (what is).
-This happens not only in mijs_paradox_2019, but also in @reynolds_perceptions_2014, who explicitly use the term beliefs to talk about perceptions, whereas other authors use general terms as attitudes (@kunovich_systems_2007).
-The first attempt to shed light on this issue was made by @duru-bellat_whos_2012, who used the item "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" for "desired" meritocracy (beliefs), whereas for "perceived" meritocracy they used two items: “Would you say that in your country, people are rewarded for their efforts?” and “...
+Several papers dealing with meritocracy use the term beliefs (i.e.
+what should be), while actually referring to perceptions (i.e.
+what is).
+This occurs not only in mijs_paradox_2019, but also in @reynolds_perceptions_2014, in which the term beliefs is used to talk about perceptions, whereas other authors use general terms such as attitudes (@kunovich_systems_2007).
+The first attempt to shed light on this issue was made by @duru-bellat_whos_2012, who used the question "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" as a proxy for "desired" meritocracy (beliefs).
+They then determined "perceived" meritocracy, using the questions: “Would you say that in your country, people are rewarded for their efforts?” and “...
 people are rewarded for their skills?”.
@@
 Is the belief in meritocracy a perception or a preference with normative meaning?
-In order to expand the analytical conceptual framework, we believe that both dimensions should be included in the analysis, as proposed by @duru-bellat_whos_2012.
-This opens up possibilities of analyzing whether perceptions and preferences are actually related  (i.e.
-correlation close to 1) or whether they are different aspects of the same phenomenon.
+In order to expand the analytical conceptual framework, we believe that both dimensions should be included in analyses, as proposed by @duru-bellat_whos_2012.
+This opens up the possibility of analyzing whether perceptions and preferences are actually related (i.e.
+have a correlation close to 1), or whether they are independent aspects of the same phenomenon.
 As @son_hing_merit_2011-1 have pointed out, "People can believe that outcomes ought to be distributed on the basis of merit and yet vary in their perceptions of whether this is how society currently operates" (p.
 435).
-In other words, normative beliefs should be interpreted taking perceptions into account: a large normative belief in meritocracy certainly means something totally different for someone perceiving high meritocracy than for someone perceiving low meritocracy.
-Avoiding the confusion generated by the term "belief", we propose the terms meritocratic preferences (normative dimension) and meritocratic perceptions (factual dimention), as they better reflect the two faces of meritocracy under scrutiny.
+In other words, normative beliefs should be considered while taking perceptions into account: a strong normative belief in meritocracy may mean something totally different to someone perceiving high meritocracy, when compared to someone perceiving low meritocracy.
+To avoid the confusion generated by the term "belief", we propose the terms meritocratic preferences (what should be), and meritocratic perceptions (what is), as they better reflect the two facets of meritocracy under scrutiny.
 
@@
 
-Unlike studies that consider talent a non-meritocratic aspect and therefore rule it out of the operationalization of meritocracy, @kunovich_systems_2007 follow a different approach by including some non-meritocratic elements.
-Using the items’ battery listing a number of reasons about “How important should be in deciding pay...” (as @duru-bellat_whos_2012 for desired meritocracy), they argue that reasons such as education and responsibility are meritocratic and pointed 1 if considered essential, whereas reasons such as having a family and children were pointed 1 if they were considered "not important at all" (i.e.
+Unlike studies that consider talent to be a non-meritocratic aspect, and therefore rule it out of the operationalization of meritocracy, @kunovich_systems_2007 follow a different approach, by including some non-meritocratic elements.
+Using an items’ battery listing a number of factors in relation to “How important each should be in deciding pay...” (as @duru-bellat_whos_2012 for desired meritocracy), they consider factors such as education and responsibility as meritocratic, giving them a value of 1 if considered “essential” in the scale response, whereas factors such as having a family or children are valued 1 when rated as "not important at all" (i.e.
 reverse coded).
-A similar approach was taken by @newman_false_2015, reverse-coding non-meritocratic items, the same principle applied in the "Preference for the Merit Principle Scale" [@davey_preference_1999].
+The assumption behind this approach is that rejecting a supposed non-meritocratic aspect (as having family and/or children) implies a larger belief in meritocracy.
+A similar approach of reverse-coding non-meritocratic items was taken by @newman_false_2015, using the same principles applied in the "Preference for the Merit Principle Scale" [@davey_preference_1999].
 
-The assumption that meritocratic and non-meritocratic elements are the poles of the same continuum was analyzed by @reynolds_perceptions_2014 using the  "get ahead" perceptions' battery items mentioned above.
-They consider education, ambition and hard work as meritocratic and other reasons such as wealthy family and right connections as non-meritocratic.
-Nevertheless, despite making this distinction the author ends up subtracting one dimension from the other, assuming that they are two poles of the same continuum as @kunovich_systems_2007 did.
-Still, taking into account this view, we suggest that non-meritocratic aspects should be part of a meritocratic measurement but taken independently and not adding or subtracting from meritocratic ones unless it is empirically proved that they belong to the same conceptual dimension.
+The assumption that meritocratic and non-meritocratic elements are poles of the same continuum was tested by @reynolds_perceptions_2014 using the "get ahead" perceptions' battery items mentioned above.
+They considered education, ambition, and hard work as meritocratic, and other factors such as family wealth and connections as non-meritocratic.
+Despite making and proving this distinction, however, the authors end up subtracting one dimension from the other, thus coming back to the assumption that they are two poles of the same continuum as @kunovich_systems_2007 did.
+We suggest that non-meritocratic factors should be part of meritocratic measurements, but should be considered independently, rather than adding to or subtracting from meritocratic factors, unless it is empirically proved that they belong to the same conceptual dimension.
 
@@
 
-Finally, most of the studies in meritocracy so far have not incorporated the issue of measurement error [@brownBookReviewTimothy2010, @bollenStructuralEquationsLatent1989] (Ansolabehere, S., Rodden, J., & Snyder Jr, J.
-M., 2008), using single indicators and/or simple average indexes for measuring meritocracy.
-Such strategy assumes that the latent construct is measured perfectly by the indicators chosen, going as far as proposing that "...
+Finally, most of the studies in meritocracy so far have not accounted for the issue of measurement error [@brownBookReviewTimothy2010, @bollenStructuralEquationsLatent1989] (Ansolabehere, S., Rodden, J., & Snyder Jr, J.
+M., 2008), when using single indicators and/or simple average indexes for measuring meritocracy.
+Such strategy assumes that the latent construct is measured perfectly by the indicators chosen, going as far as to propose that "...
 In choosing this strategy of index construction, we argue that *support for meritocracy is not a latent variable* [@kunovich_systems_2007 p.653-654].
-Some advances were done by @reynolds_perceptions_2014 by doing a principal component analysis of meritocratic and non-meritocratic dimensions, but somewhat contradictorily, they end up in a sum index despite proving a multidimensional latent structure.
+Some advances were made by @reynolds_perceptions_2014 through conducting a principal component analysis of meritocratic and non-meritocratic dimensions, but somewhat contradictorily, they chose a sum index, despite proving a multidimensional latent structure.
 
@@
 
-Based on the previous assumptions and limitations in the measurement of meritocracy, we propose and test an instrument with the following characteristics:
+Based on the previous assumptions and limitations of the measurements of meritocracy, we propose and test an instrument with the following characteristics:
 
-- _Multidimensional_, incorporating previous distinctions between preferences and perceptions as well as between meritocratic and non-meritocratic aspects.
+- _Multidimensionality_, incorporating previous distinctions between preferences and perceptions, as well as between meritocratic and non-meritocratic aspects.
 
@@
 
-- Based on _previous indicators_ as far as possible, for the sake of comparability between studies
+- Based on _previous indicators_ as much as possible, for the sake of comparability between studies.
 
 - _Brief_, as to be used in regular public opinion surveys.
-In this point it differs for instance from the proposal of the "Preference for the Merit Principle Scale" [@davey_preference_1999], as they use 15 items just for one dimension (besides the problem of reverse-coding non-meritocratic items).
+In this respect, it differs from the proposal of  "Preference for the Merit Principle Scale" [@davey_preference_1999], as they use 15 items for just one dimension (aside from the problem of reverse-coding non-meritocratic items).
 
@@
 
-The columns Perceptions and Preferences represent the distinction between these two concepts, usually confused under the label "beliefs".
-Perceptions refers to the extent to which people observe that meritocracy works or applies in their society, which in terms of measurement relates to items such as "I think hard work is important to get ahead in society", whereas preferences refer to normative expectations that are usually linked to a "should" expression (e.g.
+The columns "Perceptions" and "Preferences" represent the distinction between these two concepts, usually confused under the label "beliefs".
+Perceptions refer to the extent to which people observe that meritocracy works or applies in their society, which relates to items such as "I think hard work is important to get ahead in society".
+Preferences refer to normative expectations that are usually linked to a "should" expression (e.g.
 whether hard work should be related to payment).
-The rows in the table of Figure 1 consider the distinction between meritocratic and non-meritocratic dimensions [@reynolds_perceptions_2014].
-This aspect has been usually treated as different ends of a same continuum in part of the previous research, an assumption that requires empirical scrutiny.
-These non-meritocratic elements usually refer to the use of personal contacts or family advantages to get ahead in life.
+The rows in Figure 1 consider the distinction between meritocratic and non-meritocratic dimensions [@reynolds_perceptions_2014].
+These aspects have usually been treated as different ends of the same continuum in previous research, an assumption that requires empirical scrutiny.
+Non-meritocratic elements usually refer to the use of personal contacts or family advantages to get ahead in life.
 
-Regarding the selection of indicators, most of them are taken or adapted from previous studies for the sake of comparability.
-For meritocratic indicators we use effort and talent as the main components of the traditional concept of merit as defined by @young_rise_1962 (Atria et al.
-2020) , whereas for non-meritocratic dimensions we use having rich parents and good contacts.
-The description of the specific items is presented in the methodology section.
+Regarding the selection of indicators, most are taken or adapted from previous studies for the sake of comparability.
+For meritocratic indicators, we use effort and talent as the main components of the traditional concept of merit as defined by @young_rise_1962 (Atria et al.
+2020), whereas for non-meritocratic dimensions, we use wealthy parents and good contacts.
+Descriptions of the specific items are outlined in the methodology section.
 
-The research hypotheses behind this conceptualization and measurement model are the following:
+The research hypotheses behind this conceptualization and measurement model are as follows:
 
@@
 
-* $H_2$.
-The non-meritocratic perception is a latent variable that derives from two indicators related to the agreement with the statement that people with contacts and rich parents manage to get ahead.
+* $H_2$. The perception of non-meritocracy is a latent variable that derives from two indicators related to agreement with the statement that people with contacts and rich parents manage to get ahead.
 
-* $H_3$. Meritocratic preferences are a latent variable based on a normative value of effort and talent.
+* $H_3$. Meritocratic preferences are a latent variable based on the normative value of effort and talent.
 
-* $H_4$. Non-meritocratic preferences are a latent variable based on the normative value of the use of personal contacts and having rich parents.
+* $H_4$. Non-meritocratic preferences are a latent variable based on the normative value of the use of personal contacts and having wealthy parents.
 
@@
 
-The data was obtained through an online questionnaire which was part of a larger study on meritocracy and preferences developed in Chile in 2019 and funded by the national scientific agency (ANID).
-The questionnaire was programmed in Qualtrics and the fieldwork was in charge of an external online survey agency ([netquest.cl](www.netquest.cl)) between December 2019 and January 2020.
+The data was obtained through an online questionnaire, which was part of a larger study on meritocracy and preferences developed in Chile in 2019 funded by the national scientific agency ANID.
+The questionnaire was programmed in Qualtrics and the fieldwork was conducted by an external online survey agency ([netquest.cl](www.netquest.cl)) between December 2019 and January 2020.
 The sample was selected from a non-probabilistic quota design in three large cities in Chile.
-The quotas for gender, age and education level were generated based on the survey of the Public Studies Center (CEP, 2019), which has a high prestige in the country and is also the counterpart agency of ISSP (International Social Survey Programme) in Chile.
-A total sample of 2,141 people was collected, excluding those who did not answer the questions on the scale and those who did not accept informed consent.
-There are not significant differences between our effective sample and the population for most socio-demographics characteristics, with the exception of educational level (see Table XX in appendix).
-As it usually occurs with online samples, there were some limitations in achieving the quotas for lower educational levels.
+The quotas for gender, age, and education level were generated based on a survey by the Public Studies Center (CEP, 2019), which is a well regarded counterpart agency of the ISSP (International Social Survey Programme) in Chile.
+A total sample of 2,141 people was collected, excluding those who did not answer the questions on the scale, and those who did not accept informed consent.
+There were no significant differences between our sample and the wider population for most socio-demographic characteristics, with the exception of educational level (see Table \@ref(tab:rep-samp) in appendix).
+As is often the case with online surveys, there were some limitations in achieving the quotas for lower educational levels.
 
@@
 
-The proposed scale of perceptions and preferences about meritocracy consists of 8 indicators that are grouped into the 4 dimensions referred above: Perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic).
-In order to achieve at least some comparability with previous studies, the questions were adapted from the items battery "reasons to get ahead" (ISSP/GSS), which are mostly used for operationalizing meritocracy [@mijs_paradox_2019; @duru-bellat_whos_2012; @reynolds_perceptions_2014].
-The eight items ordered according to dimensions are presented in Table \@ref(tab:table-indicadores).
-These 8 likert-type items have 5 response alternatives ranging from "Completely disagree"(1) to "Completely agree" (5).
+The proposed scale of perceptions and preferences relating to meritocracy consisted of eight indicators that were grouped into the four dimensions listed earlier: Perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic).
+In order to achieve at least some comparability with previous studies, the questions were adapted from the items battery "reasons to get ahead" (ISSP/GSS), which has been widely used for operationalizing meritocracy in previous studies [@mijs_paradox_2019; @duru-bellat_whos_2012; @reynolds_perceptions_2014].
+The aforementioned eight items, ordered according to dimensions, are presented in Table \@ref(tab:table-indicadores).
+These eight likert-type items have five response alternatives, ranging from "Completely disagree"(1) to "Completely agree" (5).
 
@@
 
-With the objective of evaluating the effect of the indicators ordering, the respondents (_n = 2141_) were randomly divided into three different set-order versions as explained in Figure \@ref(fig:appmod).
+With the objective of evaluating the effect of indicator ordering, the respondents (_n = 2141_) were randomly divided into three different set-order versions, as explained in Figure \@ref(fig:appmod).
 The scale was presented to the first group (_n = 712_) in the order that appears in Table 2.
-For the second group (_n = 717_), the order of the items was organized according to the topics of the items, e.g.
-for the topic of hard work the item about perception was followed by the item about preference, and the same for the rest of the topics.
-Finally, for the third group (_n = 712_) the items were completely randomized.
+For the second group (_n = 717_), the order of the items was reorganized according to the topics of the items, e.g.
+for the topic of hard work, the item about perception was followed by the item about preference, and the same for the rest of the topics.
+Finally, for the third group (_n = 712_), the items were completely randomized.
 
@@
 
-For testing the scale's underlying constructs we estimate confirmatory factor analysis models (CFA). The model estimates one factor for each dimension, as represented in the following figure:
+To test the scale's underlying constructs, we employed confirmatory factor analysis models (CFA).
+The models estimated one factor for each of the proposed dimensions, as represented in the following figure:
 
-```{r meas01, echo=FALSE, fig.cap = "Theorical model", fig.align='center'}
+```{r meas01, echo=FALSE, fig.cap = "Theoretical model", fig.align='center'}
 knitr::include_graphics('../output/images/meas01.png')
@@
 
-CFA was conducted using the `lavaan` R package (version 0.6-3; Rosseel, 2020) with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement (Kline, 2016; Rosseel, 2020).
-As recommended by Brown (2008), we assessed model fit by jointly considering the comparative fit index and Tucker-Lewis Index (CFI and TLI; acceptable fit > 0.95), Root of the average squared residual approximation (RMSEA; acceptable fit < 0.08),  Chi-square: (p-value; acceptable fit > 0.05, and Chi-square ratio:> 3).
+CFA was conducted using the `lavaan` R package (version 0.6-3; Rosseel, 2020), with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement (Kline, 2016; Rosseel, 2020).
+As recommended by Brown (2008), we assessed model fit by jointly considering the comparative fit index and Tucker-Lewis Index (CFI and TLI; acceptable fit > 0.95), Root of the average squared residual approximation (RMSEA; acceptable fit < 0.08),  Chi-square: (p-value; acceptable fit > 0.05, and Chi-square ratio > 3).
 
-A pre-registration was made in the OSF platform, available in the following link: [https://osf.io/z45y2](https://osf.io/z45y2).
-In this pre-registration are the hypothesis regarding the four-dimensional conceptual model underlying the scale, the variables measurement levels, and the statistical tests to be performed with their respective evaluation parameters along with other important aspects of the research design.
+A pre-registration was made in the OSF platform, available at the following link: [https://osf.io/z45y2](https://osf.io/z45y2).
+Included in this pre-registration are the hypotheses regarding the four-dimensional conceptual model underlying the scale, the variable measurement levels, the statistical tests to be performed with their respective evaluation parameters, and other important aspects of the research design.
 
@@
 
-## Descriptive analysis
+## Descriptive analyses
 
-As can be seen in the table \ref(tab:desc01), the indicators have values ranging from 1 (totally disagree) to 5 (totally agree).
-Averages are observed from 2.41, corresponding to preference-contacts, to 3.89, corresponding to preference-effort.
-Both indicators are consistent with the general adherence to meritocracy reported in previous studies, privileging individual aspects such as effort (Frei et al 2020, Bucca, 2016).
+As can be seen in the Table \@ref(tab:desc01), the indicators have values ranging from 1 (totally disagree) to 5 (totally agree).
+Averages are observed from 2.41, corresponding to preference-contacts; to 3.89, corresponding to preference-effort.
+Both indicators are consistent with a general adherence to meritocracy as reported in previous studies, privileging individual aspects such as effort (Frei et al 2020, Bucca, 2016).
 
@@
 table04b<- sum04b %>% rename("Mean"=mean, "SD"=sd, "Min"=min,"Max"=max)
-sumdat04<- kable(table04b,format =table_format, booktabs= T, exhaust = FALSE,align = "l",linesep = "", caption = cap
-) %>% kable_styling(full_width = table_format2,
+sumdat04<- kable(table04b,format =table_format, 
+                 booktabs= T, 
+                 #exhaust = FALSE,
+                 align = "l",
+                 linesep = "", 
+                 caption = cap) %>% 
+  kable_styling(full_width = table_format2,
                 font_size = 10,
                 position = "center",
- latex_options = c("hold_position"));sumdat04
+                latex_options = c("HOLD_position"));sumdat04
 ```
 
-The graphs presented in Figure \@ref(fig:plotlikert) present disaggregated and comparable information of the different response categories for each item.
-It can be seen that in general there is more agreement in the perception of non meritocratic items than meritocratic ones, while in the case of preferences the opposite occurs.
+The graphs presented in Figure \@ref(fig:plotlikert) display disaggregated and comparable information of the different response categories for each item.
+It can be seen that, in general, there is more agreement in the perception of non-meritocratic items than in meritocratic items, while in the case of preferences, the opposite occurs.
 As far as preferences are concerned, the preponderant role of effort over talent as a criterion of meritocratic preference is noteworthy.
@@
 
-In Figure Ref(fig:corpoly), moderate to high intensity relationships are observed between indicators that correspond to the same factor (for example, perception of meritocracy by effort and by talent, r=0.56).
-The relationships between meritocratic perceptions and preferences are also noteworthy, which is not the case for nonmeritocratic indicators.
+In Figure \@ref(fig:corpoly), moderate to high intensity relationships are observed between indicators that correspond to the same factor (e.g., perception of meritocracy by effort and by talent, r=0.56).
+The relationships between meritocratic perceptions and preferences are also noteworthy, which is not the case for non-meritocratic indicators.
 
@@
 
-In sum, descriptive analyses point to a relative adherence to meritocracy, which is expressed in a greater preference for meritocratic criteria and a lesser preference for non-meritocratic ones, even when the preference for non-meritocratic elements reaches a proportion of more than 20%.
+In sum, descriptive analyses point to a relative adherence to meritocracy, which is expressed as a greater preference for meritocratic criteria, and a lesser preference for non-meritocratic criteria, even when the preference for non-meritocratic elements reaches a proportion of more than 20%.
 Likewise, a relatively low perception of meritocracy is observable.
-In addition, a coherent relationship between the indicators as proposed in the theoretical model is observed in the pre-registration of the study; that is, the pairs of items associated with a specific factor show correlations with a large effect size (for example, meritocratic preferences for items associated with effort and talent).
-In particular, the associations between effort and talent are relevant, since they dismiss previous assumptions that talent would not be a meritocratic criterion [@mijs_paradox_2019], otherwise the correlation would be zero or negative.
+In addition, we observed a coherent relationship between the indicators of the same dimension as proposed in the pre-registration of the study; that is, the pairs of items associated with a specific factor show correlations with a large effect size (for example, meritocratic preferences for items associated with effort and talent).
+In particular, the associations between effort and talent are relevant, challenging previous assumptions that talent would not be a meritocratic criterion [@mijs_paradox_2019], otherwise the correlation would be zero or negative.
 In addition, we see that there is no negative correlation between meritocratic and non-meritocratic aspects, dismissing the assumptions of previous studies that these dimensions would be the opposite poles of the same continuum [@reynolds_perceptions_2014].
@@
 This section estimates the fit of the conceptual model behind the Perceptions and Preferences for Meritocracy Scale.
-For this, we first estimate a confirmatory factor analysis model for the whole sample, and secondly we test the order effects applying the same model to each of the three order versions.
+For this, we first estimate a confirmatory factor analysis model for the whole sample, and secondly we test the order effects applying the same model to each of the three order permutations.
 
@@
 
-\@ref(fig:meas02) shows the results of the estimation for the four-factor model with complete sample.
-This model presents adequate fit indicators (CFI = 0.977, RMSEA = 0.079, $\chi2$(df = 14) = 28,03, p = 0.014) and it presents factorial loads with weights that vary between 0.63 and 0.88.
-Furthermore, considering the correlation between the latent variables, it is observed that the variables with the highest correlations are meritocratic perceptions and preferences, which have a positive correlation of 0.59, while non-meritocratic perceptions and preferences do not present a considerable correlation.
+Figure \@ref(fig:meas02) shows the results of the estimation for the four-factor model using the complete sample.
+This model shows adequate fit indicators (CFI = 0.977, RMSEA = 0.079, $\chi2$(df = 14) = 28,03, p = 0.014) and it presents factorial loads with weights that vary between 0.63 and 0.88.
+Considering the correlations between the latent variables, it is observed that the variables with the strongest correlations are meritocratic perceptions and preferences, which have a positive correlation of 0.59, while non-meritocratic perceptions and preferences do not show considerable correlation.
 
-```{r meas02, echo=FALSE, fig.cap = "Confirmatory factor analysis of the Perceptions an Preferences for Meritocracy Scale", fig.align='center'}
-knitr::include_graphics('../output/images/meas02.png')
+```{r meas02, echo=FALSE, fig.cap = "Confirmatory factor analysis of the Perceptions and Preferences for Meritocracy Scale", fig.align='center'}
+knitr::include_graphics('../output/images/meas03.png')
 ```
 
-As can be seen in the Table \ref(tab:tab-loads), all models, regardless of the group and order of items, obtained an adequate fit, with CFI's above .95 and RMSEA's below .8.
+As can be seen in the Table \@ref(tab:tab-loads), all models, regardless of the group and order of items, obtained an adequate fit, with CFI's above 0.95 and RMSEA's below 0.08.
 However, no model achieved a non-significant chi-square (something to be expected in large samples), although both the randomized and the first model obtained an adequate chi-square ratio of less than 3.
-The first order was the one that obtained the best fit (CFI=0.998, TLI=0.995,RMSEA=0.037, $\chi2$(df=14)=28.03, p = 0.014), followed by the random order of items (CFI=0.992, TLI=0.984,RMSEA=0.051, $\chi2$(df=14)=39.09, p < 0.001).
-On the other hand, the scale ordered by themes seems to generate a framing effect in which the relationship between perceptions and preferences seems overestimated, thus affecting the adjustment (CFI=0.984, TLI=0.968, RMSEA=.071, $\chi2$(df=14)=64.156, p < 0.001).
+The first order was the one that obtained the best fit (CFI=0.998, TLI=0.995, RMSEA=0.037, $\chi2$(df=14)=28.03, p = 0.014), followed by the random order of items (CFI=0.992, TLI=0.984,RMSEA=0.051, $\chi2$(df=14)=39.09, p < 0.001).
+On the other hand, the scale ordered by themes seems to generate a framing effect, in which the relationship between perceptions and preferences seemed to be overestimated, thus affecting the adjustment (CFI=0.984, TLI=0.968, RMSEA=.071, $\chi2$(df=14)=64.156, p < 0.001).
+
+```{r latcor, echo=FALSE, fig.cap = "Latent variables correlation", fig.align='center'}
+knitr::include_graphics('../output/images/latcor.png')
+```
 
 ```{r tab03, echo=FALSE,eval=TRUE}
-(acá recuperar syntax)
+load(file = "../output/tables/fit4_o.RData")
+load(file = "../output/tables/fit3_o.RData")
+load(file = "../output/tables/fit2_o.RData")
+load(file = "../output/tables/fit1_o.RData")
+
+sum_fit<- bind_rows(fitmeasures(fit4_o)[c("chisq.scaled","df","cfi.scaled","rmsea.scaled")],
+                    fitmeasures(fit1_o)[c("chisq.scaled","df","cfi.scaled","rmsea.scaled")],
+                    fitmeasures(fit2_o)[c("chisq.scaled","df","cfi.scaled","rmsea.scaled")],
+                    fitmeasures(fit3_o)[c("chisq.scaled","df","cfi.scaled","rmsea.scaled")])
+sum_fit$mod <- c("Model 1","Model 2","Model 3","Model 4")
+sum_fit$est <- c("DWLS","DWLS","DWLS","DWLS")
+sum_fit$nobs <- c(nobs(fit4_o),nobs(fit1_o),nobs(fit2_o),nobs(fit3_o))
+sum_fit <- select(sum_fit,mod,nobs,est,everything())
+colnames <- c("Model","$N$","Estimator","$\\chi^2$","df","CFI","RMSEA")
+
+sumtable02<- kable(sum_fit,digits = 3,format = table_format,row.names = F,booktabs=T, align = c("lcccccc"),
+                   caption = "Summary fit indices according group",
+                   col.names = colnames,
+                   escape = FALSE) %>%
+  kable_styling(full_width = F,
+                font_size = 10)  %>%
+      footnote(number = c("Model 1: Complete sample",
+                          "Model 2: fixed order by perception/preference",
+                          "Model 3:  fixed order by topic (i.e: effort)",
+                          "Model 4: Randomized order"));sumtable02
+
 ```
 
-Although all tests obtained relatively adequate indicators, the mentioned differences between the adjustments of the models according to order are statistically significant, which was evaluated from an invariance analysis.
-It is concluded that among the three orders there is no configurtional invariance, that is, they do not have the same dimensionality and therefore do not equally fit the theoretical model [@vandenbergReviewSynthesisMeasurement2000].
-This is due to the effect produced by the joint appearance of the items of the same factor in order 1, which increases the fit of the model.
-Furthermore, in order 2, when asking about the perception and preference around the same indicator, the cross relations increase and the model fit decreases.
-Consistently, the random model presented an intermediate adjustment between order 1 and order 2.
+Although all tests obtained relatively adequate indicators, the aforementioned differences between the adjustments of the models according to order are statistically significant, as evaluated by an invariance analysis.
+It was concluded that among the three orders there was no configurational invariance, that is, they did not have the same dimensionality, and therefore did not equally fit the theoretical model [@vandenbergReviewSynthesisMeasurement2000].
+This is due to the effect produced by the joint appearance of items relating to the same factor in order one, which improved the fit of the model.
+Furthermore, in order two, when asking about perception and preference around the same indicator, cross relations increased and the model fit was worsened.
+Consistently, the random model presented an intermediate case between order one and order two.
 
@@
 
-As shown in table 4, the proposed theoretical four-factor model fits better than the one-factor contrast model.
-The theoretical model adjusted relatively well, as it shows optimal indicators for IFC=0.987, TLI=0.975 and RMSEA=.041, although it has poor indicators for the test $\chi2$(df=14)=67.6, p-value=.000.
+As shown in Table \@ref(tab:items-dim), the proposed theoretical four-factor model fits better than the one-factor contrast model.
+The theoretical model adjusted relatively well, as it shows preferable indicators of IFC=0.987, TLI=0.975 and RMSEA=0.041, although it also has poor indicators of $\chi2$(df=14)=67.6 and p-value<0.001.
 To evaluate possible improvements in the scale, the relationships proposed by the modification indexes were analyzed.
 These indicate the existence of two unspecified cross-loads.
-When a model was generated following these recommendations, there was a considerable improvement in the model, although the new model also did not obtain a $\chi2$ ratio less than 3 and obtained very low factorial loads ($\tau$)< 0.15.
-Therefore, following the recommendation of Brown (2006) to only accept the proposals of modification rates when theory and solid evidence are available, it has been decided not to incorporate these parameters into the model.
+When a model was generated following these recommendations, there was a considerable improvement in the model, although the new model still did not obtain a $\chi2$ ratio of less than 3, and had very low factorial loads ($\tau$)< 0.15.
+Therefore, following the recommendation of Brown (2006) to only accept the proposals of modification when theory and solid evidence are available, we decided not to incorporate these parameters into the model.
 
-# Conclusión
+# Conclusion
 
-Studies that attempt to characterize and compare societies regarding support for meritocratic beliefs have used different approaches.
-As most studies use secondary survey data, they tend to grasp in the available indicators assuming that they represent an underlying construct.
+Studies that attempt to characterize and compare societies by their support for meritocratic beliefs have used differencing approaches.
+As most studies use secondary survey data, they tend to assume that the available indicators represent an underlying construct.
 A review of these studies reveals several non-tested assumptions, as well as the use of similar indicators to represent different constructs and dimensions of meritocracy.
-As the existence of heterogeneous approaches certainly have consequences for advancing in the study of meritocracy, this paper presented a comprehensive conceptual framework for the empirical study of meritocracy built upon previous research, which is then tested against new survey data.
+As the existence of heterogeneous approaches certainly has consequences for the advancement of the study of meritocracy, this paper presents a comprehensive conceptual framework for the empirical study of meritocracy, building upon previous research.
+This framework was then tested against new survey data.
 
-We identified four critical aspects regarding the measurement of meritocracy in previous studies: unidimensionality, the ambiguous use of “beliefs”, the use of non-meritocratic indicators as opposed to meritocratic and the consideration of measurement error.
-The proposed 8-items scale Perceptions And Preferences for Meritocracy was designed and tested in order to deal with these four issues.
-The results indicate that, on the one hand, perceptions and preferences proved to be two related but different dimensions, usually confounded in previous research under the label of "beliefs".
-On the other hand, meritocratic and non-meritocratic dimensions would not be the poles of the same continuum as some previous studies assumed.
+We identified four critical aspects regarding the measurement of meritocracy in previous studies: unidimensionality, the ambiguous use of the term “beliefs”, the use of non-meritocratic indicators as opposed to meritocratic indicators, and the consideration of measurement error.
+The proposed 8-item scale, “Perceptions and Preferences for Meritocracy”, was designed and tested in order to deal with these four issues.
+Our results indicate that perceptions and preferences seem to be two related but different dimensions, often confounded in previous research under the label of "beliefs".
+Additionally, meritocratic and non-meritocratic dimensions do not appear to constitute poles of the same continuum, as some previous studies have assumed.
 
 The four-dimensional framework and structure of the Perceptions and Preferences for Meritocracy scale opens several avenues for future research.
-For instance, distinguishing perceptions from preferences will allow us to evaluate the extent to which different societies are accustomed or satisfied with the performance of meritocracy, in terms of difference of what is perceived from what is preferred.
-On the other hand, given that non-meritocratic elements are not necessarily related to the meritocratic ones, it makes it possible to assess the legitimacy of practices such as the use of personal contacts and their interference (or not) with meritocratic ideals in different societies.
-The consideration of structural factors, both individual and societal, will allow us to advance in hypotheses of meritocratic legitimacy from a more comprehensive view than the one used today based only on meritocratic perceptions.
-Furthermore, the impacts of different configurations of the four dimensional framework on practices and behaviors such as corruption, civic involvement and political views is an area that requires additional research particularly in times of economic crisis and growing inequality, and to which the proposed instrument aims at contributing.
+For instance, distinguishing perceptions from preferences will allow us to evaluate the extent to which different societies are accustomed to, or satisfied with, the perceived level of meritocracy, in terms of differences between what is perceived and what is preferred.
+Additionally, given that non-meritocratic factors are not necessarily related to meritocratic ones, our framework makes it possible to assess the perceived legitimacy of practices such as the use of personal contacts and their interference (or not), within meritocratic ideals in different societies.
+The consideration of structural factors, both individual and societal, will allow us to advance future hypotheses relating to meritocratic legitimacy to a more comprehensive level than the one currently used, which is based solely on meritocratic perceptions.
+Furthermore, the impacts of different configurations of the four-dimensional framework on practices and behaviors such as corruption, civic involvement, and political alignment, is an area that requires additional research.
+Such future agendas could be specially relevant in times of economic crisis and growing inequalities, that could entail changes in the legitimation of the current distributive structure based on meritocratic ideals.
 
@@
     mutate(vars = recode(vars,
-                         "A. Who the more they try they manage to get bigger rewards that those who striveless."= "Perception Effort",
+                         "A. Who the more they try they manage to get bigger rewards that those who strive less."= "Perception Effort",
                          "B. Who possess more talent they manage to obtain greater rewards than those who possess less talent."= "Perception Talent",
@@
 | Primary  or less    | 2,93%   | 15,88% |
-| Hig school          | 43,23%  | 37,04% |
+| High school          | 43,23%  | 37,04% |
 | Non university      | 32,63%  | 28,93% |
@@
 Table: `r as.character(paste("(\\#tab:rep-samp)", "Representativeness of the sample"))`.
-<!--stackedit_data:
-eyJoaXN0b3J5IjpbLTkyNTUxNzA1Ml19
--->

### 2020-10-11 docs/paper.Rmd

@@
 Table: `r as.character(paste("(\\#tab:rep-samp)", "Representativeness of the sample"))`.
+=======
+---
+title: "Measuring Perceptions and Preferences for Meritocracy"
+css: "custom.css"
+linestretch: '1.5'
+link-citations: yes
+author: "Juan Carlos Castillo, Julio Iturra & Francisco Meneses <br><br>"
+abstract: "Economic and social inequality have raised growing concerns and crises across contemporary societies.
+One of the mechanisms developed by the social sciences to explain the maintenance of inequality is the belief in meritocracy, which would legitimize economic disparities based on differences of effort and talent.
+Despite its wide use, empirical research on meritocracy is relatively novel and since it is based on multiple approaches and techniques its analysis can be vague or confusing.
+Particularly problematic is that a number of studies have relied mostly on secondary data to operationalize meritocracy, with a large variation in the use and interpretation of survey items.
+Starting from a review of studies that measure meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding the conceptualization and indicators of meritocracy.
+Based on this critical analysis, we propose an item's battery called *Perceptions and Preferences for Meritocracy Scale*, which is tested with confirmatory factor analysis with data from an online survey study (N=2,141).
+The results support the proposed conceptual structure which not only distinguishes between perceptions and preferences, but also between meritocratic and non-meritocratic dimensions.
+The discussion highlights the relevance of considering these different dimensions in order to advance in the study of meritocracy."
+output:
+  bookdown::pdf_document2:
+    template: null
+    toc: false
+  bookdown::html_document2:
+    number_sections: false
+linkcolor: blue
+bibliography:
+  - ../input/bib/EconomiaMoral.bib
+  - ../input/bib/meritocracy.bib
+editor_options:
+  chunk_output_type: console
+geometry: margin=0.78in
+header-includes:
+  # - \usepackage[spanish,es-tabla,es-nodecimaldot]{babel}
+  - \usepackage{times}           # Times New Roman
+  - \usepackage{caption}
+  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
+  - \usepackage{graphicx}
+  - \usepackage{float}
+  - \usepackage{booktabs}
+  - \usepackage{longtable}
+  - \usepackage{array}
+  - \usepackage{multirow}
+  - \usepackage{wrapfig}
+  - \usepackage{float}
+  - \usepackage{colortbl}
+  - \usepackage{pdflscape}
+  - \usepackage{tabu}
+  - \usepackage{threeparttable}
+---
+
+```{r eval=FALSE, include=FALSE}
+# for render in pdf run rmarkdown::render_site("docs/paper.Rmd", output_format = "all")
+# clean #in the yml
+rmarkdown::render("docs/paper.Rmd", output_format = "bookdown::pdf_document2")
+rmarkdown::render("docs/paper.Rmd", output_format = "bookdown::html_document2")
+```
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = FALSE,out.width = '85%',fig.pos= "H")
+# knitr::opts_knit$set(base.url = "../") #relative path for .html output file
+# knitr::opts_knit$set(root.dir = "../") #relative path for chunks within .rmd files
+options(scipen=999)
+rm(list=ls())
+options(knitr.kable.NA = '')
+options(knitr.graphics.error = FALSE)
+Sys.setlocale("LC_ALL", "ES_ES.UTF-8")
+```
+
+```{r message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
+pacman::p_load(knitr, kableExtra, dplyr, lavaan)
+```
+
+```{r echo=FALSE }
+table_format = if(is_html_output()) {
+  "html"
+} else if(is_latex_output()) {
+  "latex"
+}
+
+table_format2 = if(is_html_output()) {
+  T
+} else if(is_latex_output()) {
+  F
+}
+
+```
+
+# Introduction
+
+Economic inequality has become an issue of growing concern around the world.
+This has been expressed in a series of protests, as well as diverse critical analyses regarding the development of capitalism and its consequences [Streeck, 2014; @piketty_Capital_2014].
+In this context, the study of visions, preferences and perceptions of inequality has acquired relevance in the social sciences, in topics such as redistributive preferences [@alesina_Fairness_2005]; Dimick_Models_2018], the legitimization of economic inequality [@schroder_Income_2017] and the functioning of meritocracy [@duru-bellat_whos_2012; @mijs_paradox_2019; @reynolds_perceptions_2014; Atria et al., 2020].
+
+In general, meritocracy is defined as a system of distribution of resources and rewards based on individual merit, which in its original conception is a sum of talent and effort [@young_rise_1962].
+This traditional conception of merit places in a secondary position the possible interference of structural or nonmeritocratic factors such as inheritance, personal contacts, or luck [@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995; @yairMeritocracy2007; @landWeSatTable2006; @youngRiseMeritocracy1994].
+Social psychology and sociology have studied the characteristics and consequences of beliefs in meritocracy, in general with the hypothesis that greater belief in meritocracy leads to greater legitimization of inequalities [@premingerMeritocracyServiceEthnocracy2020; @trumpWhenWhyEconomic2020; @hadjar_meritokratie_2008; @MadeiraPrimesConsequencesSystematic2019].
+Such research has raised criticism of the realization of this moral standard of distribution, arguing that it would be an unfulfilled promise given the preponderant influence of other elements beyond merit on individual status [Sandel, 2020; @witteveenReconsideringMeritocraticPower2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019, @khan_Privilege_2013].
+
+Due to the role that meritocratic beliefs play as a justification for individual achievement or failure in contemporary societies [@bay-chengTrackingHomoOeconomicus2015], multiple investigations have emerged that evaluate the relationship between meritocratic beliefs and diverse social circumstances today.
+For example, studies have been developed that link meritocracy to the reinforcement of socio-economic, gender and ethnic stereotypes [@MadeiraPrimesConsequencesSystematic2019; @girerdNeoliberalismIdeologicalBarrier2020; premingerMeritocracyServiceEthnocracy2020], as well as lines of research that evaluate the effect of meritocratic beliefs in the educational context [@generettStoriesWeTell2020; @owensENGINESSOCIALMOBILITY2020] and in the organizational dimension of companies [@perezAdvancingCareersMerit2020; @aielloNewEvidenceFirmuniversity2019].
+
+To account for the levels of belief in meritocracy, studies have generally used some indicators from existing surveys, and to a lesser extent have created ad-hoc instruments.
+However, as we will show later, the forms of measuring meritocracy vary extremely among studies.
+Many times similar phenomena are associated with different indicators, and it also happens that different phenomena are measured with similar indicators, all of which make difficult not only the comparability between studies but also to acheive more precision in understanding the effects of meritocratic beliefs in different fields.
+
+Based on the critical analysis of different approaches to the measurement of meritocracy to date, this article proposes a tool to measure and relate two key aspects in the study of meritocracy: perceptions and preferences.
+Furthermore, as a second axis of analysis it considers the elaboration of indicators regarding meritocratic and non-meritocratic aspects, demonstrating that they are not the two poles of the same continuum as many previous studies seem to suggest.
+The measurement proposal is also oriented to generate an instrument as brief as possible so that it can be used in public opinion surveys, integrating meritocratic beliefs in the study of different social phenomena.
+
+## The black-box of meritocratic beliefs {-#blackbox}
+
+Several approaches to the empirical study of meritocracy based on public opinion surveys refer to the concept of _beliefs_, but behind this concept there are usually different meanings and operationalizations.
+To illustrate this point in the following we will take as a main reference a recent paper about meritocracy by @mijs_paradox_2019, which makes several assumptions about the measurement of meritocracy that are also present in previous studies.
+
+The meritocratic beliefs' definition of Mijs is the following: "when I discuss meritocracy beliefs, I am referring to citizens’ belief in the importance of hard work relative to structural factors." [@mijs_paradox_2019, pg.9].
+In the operationalization, this is associated with the following indicator: "how important you think it is for getting ahead in life: (a) hard work", scored in a 1 to 5 likert scale.
+There are several assumptions behind this decision that are worth discussing in light of the conceptual meaning of meritocracy and its possibilities of operationalization.
+
+a. _Unidimensionality_
+
+The item used by @mijs_paradox_2019 is part of an items' battery present in several international surveys, usually called "reasons to get ahead".
+This battery displays a series of indicators related to what people consider important to get ahead in life: hard work, education, ambition, wealthy family, right connections, religion, race and gender.
+Therefore, for Mijs other aspects such as education, that could be associated to talent, are not meritocratic.
+As he points out: "Hard work is arguably the most meritocratic part of Michael Young’s equation, ‘Merit = Intelligence + Effort’, for the simple fact that intelligence itself is conditioned by a nonmeritocratic factor: who your parents happen to be" (p.5).
+
+In this measure of meritocratic belief we can observe a couple of strong assumptions: effort would not depend on parents influence, and talent is not meritocratic (contrary to Michael Youngs original conceptualization).
+The problem of whether talent is or not meritocratic is certainly an interesting point to discuss, but for those working based on empirical data it is something that should be empirically tested.
+Is effort the only dimension behind the concept of meritocracy, or is it a multidimensional concept that includes other elements as talent (as in its original conceptualization)?
+This conceptual and measurement unidimensional assumption it is possible to find in other studies that assume that effort is the main and only aspect of meritocracy [@girerdNeoliberalismIdeologicalBarrier2020a; @bubakPerceptionsMeritocracyNote2019].
+Recently, Mijs & Hoy (2020) considered a measurement of talent as an indicator of meritocracy.
+
+_b. Beliefs_
+
+The most used item battery for the operationalization of meritocracy using survey data is the "reasons to get ahead", which asks about "how important you think it is" and then lists several reasons such as effort, education, parents wealth and contacts.
+Nevertheless, another version of this same battery used in several surveys is "how important you think it _should_ be".
+Therefore, the question raised here is: Which one of both is a "belief"?
+
+The term belief has an ambiguous character in the literature, conceived as "idea-elements" for Converse (1964) or "considerations" for Zaller (1992).
+As Kluegel & Smith pointed out about the scope of beliefs: "This usage encompasses such more specific social-psychological concepts as values, perceptions, and attitudes" (p.30).
+Therefore, it covers almos anything related to subjective factors.
+But a relevant distinction in this field was made by @janmaatSubjectiveInequalityReview2013: "Perceptions refer to subjective estimates of existing inequality (i.e.
+thoughts about what is).
+Beliefs are here defined as normative ideas about just inequality (i.e.
+thoughts about what should be)"(p.359).
+Several papers dealing with meritocracy use the term beliefs (what should be) referring actually to perceptions (what is).
+This happens not only in mijs_paradox_2019, but also in @reynolds_perceptions_2014, who explicitly use the term beliefs to talk about perceptions, whereas other authors use general terms as attitudes (@kunovich_systems_2007).
+The first attempt to shed light on this issue was made by @duru-bellat_whos_2012, who used the item "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" for "desired" meritocracy (beliefs), whereas for "perceived" meritocracy they used two items: “Would you say that in your country, people are rewarded for their efforts?” and “...
+people are rewarded for their skills?”.
+
+Is the belief in meritocracy a perception or a preference with normative meaning?
+In order to expand the analytical conceptual framework, we believe that both dimensions should be included in the analysis, as proposed by @duru-bellat_whos_2012.
+This opens up possibilities of analyzing whether perceptions and preferences are actually related  (i.e.
+correlation close to 1) or whether they are different aspects of the same phenomenon.
+As @son_hing_merit_2011-1 have pointed out, "People can believe that outcomes ought to be distributed on the basis of merit and yet vary in their perceptions of whether this is how society currently operates" (p.
+435).
+In other words, normative beliefs should be interpreted taking perceptions into account: a large normative belief in meritocracy certainly means something totally different for someone perceiving high meritocracy than for someone perceiving low meritocracy.
+Avoiding the confusion generated by the term "belief", we propose the terms meritocratic preferences (normative dimension) and meritocratic perceptions (factual dimention), as they better reflect the two faces of meritocracy under scrutiny.
+
+_c. Non-meritocratic aspects_
+
+Unlike studies that consider talent a non-meritocratic aspect and therefore rule it out of the operationalization of meritocracy, @kunovich_systems_2007 follow a different approach by including some non-meritocratic elements.
+Using the items’ battery listing a number of reasons about “How important should be in deciding pay...” (as @duru-bellat_whos_2012 for desired meritocracy), they argue that reasons such as education and responsibility are meritocratic and pointed 1 if considered essential, whereas reasons such as having a family and children were pointed 1 if they were considered "not important at all" (i.e.
+reverse coded).
+A similar approach was taken by @newman_false_2015, reverse-coding non-meritocratic items, the same principle applied in the "Preference for the Merit Principle Scale" [@davey_preference_1999].
+
+The assumption that meritocratic and non-meritocratic elements are the poles of the same continuum was analyzed by @reynolds_perceptions_2014 using the  "get ahead" perceptions' battery items mentioned above.
+They consider education, ambition and hard work as meritocratic and other reasons such as wealthy family and right connections as non-meritocratic.
+Nevertheless, despite making this distinction the author ends up subtracting one dimension from the other, assuming that they are two poles of the same continuum as @kunovich_systems_2007 did.
+Still, taking into account this view, we suggest that non-meritocratic aspects should be part of a meritocratic measurement but taken independently and not adding or subtracting from meritocratic ones unless it is empirically proved that they belong to the same conceptual dimension.
+
+_d. Accounting for measurement error_
+
+Finally, most of the studies in meritocracy so far have not incorporated the issue of measurement error [@brownBookReviewTimothy2010, @bollenStructuralEquationsLatent1989] (Ansolabehere, S., Rodden, J., & Snyder Jr, J.
+M., 2008), using single indicators and/or simple average indexes for measuring meritocracy.
+Such strategy assumes that the latent construct is measured perfectly by the indicators chosen, going as far as proposing that "...
+In choosing this strategy of index construction, we argue that *support for meritocracy is not a latent variable* [@kunovich_systems_2007 p.653-654].
+Some advances were done by @reynolds_perceptions_2014 by doing a principal component analysis of meritocratic and non-meritocratic dimensions, but somewhat contradictorily, they end up in a sum index despite proving a multidimensional latent structure.
+
+## An instrument proposal {-#instrumentprop}
+
+Based on the previous assumptions and limitations in the measurement of meritocracy, we propose and test an instrument with the following characteristics:
+
+- _Multidimensional_, incorporating previous distinctions between preferences and perceptions as well as between meritocratic and non-meritocratic aspects.
+
+- Multiple indicators for each dimension, in order to _account for measurement error_ in a confirmatory factor analysis context.
+
+- Based on _previous indicators_ as far as possible, for the sake of comparability between studies
+
+- _Brief_, as to be used in regular public opinion surveys.
+In this point it differs for instance from the proposal of the "Preference for the Merit Principle Scale" [@davey_preference_1999], as they use 15 items just for one dimension (besides the problem of reverse-coding non-meritocratic items).
+
+The proposed measurement framework is depicted in Figure \@ref(fig:merit-model):
+
+```{r merit-model, echo=FALSE, fig.cap = "Model of perception and preferences for meritocracy and non-meritocracy", fig.align='center'}
+knitr::include_graphics("../input/images/generalf.png")
+```
+
+The columns Perceptions and Preferences represent the distinction between these two concepts, usually confused under the label "beliefs".
+Perceptions refers to the extent to which people observe that meritocracy works or applies in their society, which in terms of measurement relates to items such as "I think hard work is important to get ahead in society", whereas preferences refer to normative expectations that are usually linked to a "should" expression (e.g.
+whether hard work should be related to payment).
+The rows in the table of Figure 1 consider the distinction between meritocratic and non-meritocratic dimensions [@reynolds_perceptions_2014].
+This aspect has been usually treated as different ends of a same continuum in part of the previous research, an assumption that requires empirical scrutiny.
+These non-meritocratic elements usually refer to the use of personal contacts or family advantages to get ahead in life.
+
+Regarding the selection of indicators, most of them are taken or adapted from previous studies for the sake of comparability.
+For meritocratic indicators we use effort and talent as the main components of the traditional concept of merit as defined by @young_rise_1962 (Atria et al.
+2020) , whereas for non-meritocratic dimensions we use having rich parents and good contacts.
+The description of the specific items is presented in the methodology section.
+
+The research hypotheses behind this conceptualization and measurement model are the following:
+
+* $H_1$. The perception of meritocracy is a latent variable based on indicators of the importance attributed to talent and effort to get ahead in life.
+
+* $H_2$.
+The non-meritocratic perception is a latent variable that derives from two indicators related to the agreement with the statement that people with contacts and rich parents manage to get ahead.
+
+* $H_3$. Meritocratic preferences are a latent variable based on a normative value of effort and talent.
+
+* $H_4$. Non-meritocratic preferences are a latent variable based on the normative value of the use of personal contacts and having rich parents.
+
+# Methodology
+
+## Data collection {-#data}
+
+The data was obtained through an online questionnaire which was part of a larger study on meritocracy and preferences developed in Chile in 2019 and funded by the national scientific agency (ANID).
+The questionnaire was programmed in Qualtrics and the fieldwork was in charge of an external online survey agency ([netquest.cl](www.netquest.cl)) between December 2019 and January 2020.
+The sample was selected from a non-probabilistic quota design in three large cities in Chile.
+The quotas for gender, age and education level were generated based on the survey of the Public Studies Center (CEP, 2019), which has a high prestige in the country and is also the counterpart agency of ISSP (International Social Survey Programme) in Chile.
+A total sample of 2,141 people was collected, excluding those who did not answer the questions on the scale and those who did not accept informed consent.
+There are not significant differences between our effective sample and the population for most socio-demographics characteristics, with the exception of educational level (see Table XX in appendix).
+As it usually occurs with online samples, there were some limitations in achieving the quotas for lower educational levels.
+
+## Instrument design
+
+The proposed scale of perceptions and preferences about meritocracy consists of 8 indicators that are grouped into the 4 dimensions referred above: Perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic).
+In order to achieve at least some comparability with previous studies, the questions were adapted from the items battery "reasons to get ahead" (ISSP/GSS), which are mostly used for operationalizing meritocracy [@mijs_paradox_2019; @duru-bellat_whos_2012; @reynolds_perceptions_2014].
+The eight items ordered according to dimensions are presented in Table \@ref(tab:table-indicadores).
+These 8 likert-type items have 5 response alternatives ranging from "Completely disagree"(1) to "Completely agree" (5).
+
+```{r include=FALSE}
+cap <- "Items according to dimension."
+# if (knitr::is_html_output()) {tb <- table_nums("tb",cap)} else {tb <- cap}
+```
+
+```{r table-indicadores, echo=FALSE}
+tab01 <- read.csv(file = "../input/data/original/materials/table01.csv",header = 1,sep = ",",encoding = "UTF-8")
+cnames <- c("Dimension", "Factor","Statement (english)" , "Statement (spanish)")
+
+kable(tab01,table_format,booktabs = T, linesep = "",col.names = cnames, caption = cap) %>%
+  kable_styling(
+    full_width = T,
+    latex_options = c("hold_position"),
+    position = "center",
+    font_size = 10,
+    bootstrap_options=c("striped", "bordered")) %>%
+  column_spec(column = 1, width = "1.5cm", ) %>%
+  column_spec(column = 2,width = "2 cm") %>%
+  collapse_rows(columns = 1:2,valign = "middle")
+```
+
+\pagebreak
+
+## Administration sets
+
+With the objective of evaluating the effect of the indicators ordering, the respondents (_n = 2141_) were randomly divided into three different set-order versions as explained in Figure \@ref(fig:appmod).
+The scale was presented to the first group (_n = 712_) in the order that appears in Table 2.
+For the second group (_n = 717_), the order of the items was organized according to the topics of the items, e.g.
+for the topic of hard work the item about perception was followed by the item about preference, and the same for the rest of the topics.
+Finally, for the third group (_n = 712_) the items were completely randomized.
+
+```{r appmod, echo=FALSE, fig.cap = "Survey flow", fig.align='center'}
+knitr::include_graphics('../input/images/app_mod.png')
+```
+
+# Methods
+
+For testing the scale's underlying constructs we estimate confirmatory factor analysis models (CFA). The model estimates one factor for each dimension, as represented in the following figure:
+
+```{r meas01, echo=FALSE, fig.cap = "Theorical model", fig.align='center'}
+knitr::include_graphics('../output/images/meas01.png')
+```
+
+CFA was conducted using the `lavaan` R package (version 0.6-3; Rosseel, 2020) with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement (Kline, 2016; Rosseel, 2020).
+As recommended by Brown (2008), we assessed model fit by jointly considering the comparative fit index and Tucker-Lewis Index (CFI and TLI; acceptable fit > 0.95), Root of the average squared residual approximation (RMSEA; acceptable fit < 0.08),  Chi-square: (p-value; acceptable fit > 0.05, and Chi-square ratio:> 3).
+
+A pre-registration was made in the OSF platform, available in the following link: [https://osf.io/z45y2](https://osf.io/z45y2).
+In this pre-registration are the hypothesis regarding the four-dimensional conceptual model underlying the scale, the variables measurement levels, and the statistical tests to be performed with their respective evaluation parameters along with other important aspects of the research design.
+
+# Results
+
+## Descriptive analysis
+
+As can be seen in the table \ref(tab:desc01), the indicators have values ranging from 1 (totally disagree) to 5 (totally agree).
+Averages are observed from 2.41, corresponding to preference-contacts, to 3.89, corresponding to preference-effort.
+Both indicators are consistent with the general adherence to meritocracy reported in previous studies, privileging individual aspects such as effort (Frei et al 2020, Bucca, 2016).
+
+```{r desc01, echo=FALSE}
+cap <- "Descriptive statistics of the scale."
+load(file = "../input/data/proc/dat04.RData")
+dat04b<- dat04 %>% rename("A. Perception Effort" =perc_effort,
+ "B. Perception Talent" = perc_talent,
+ "C. Perception rich parents" = perc_wpart,
+ "D. Perception contacts" = perc_netw,
+ "E. Preferences Effort" = pref_effort,
+ "F. Preferences Talent" = pref_talent,
+ "G. Preferences rich parents" = pref_wpart ,
+ "H. Preferences contacts" = pref_netw )
+sum04b <- psych::describe(dat04b[,-c(1)], fast = T) %>% select(mean,sd,min,max) %>% data.frame() %>% round(digits = 2)
+table04b<- sum04b %>% rename("Mean"=mean, "SD"=sd, "Min"=min,"Max"=max)
+sumdat04<- kable(table04b,format =table_format, booktabs= T, exhaust = FALSE,align = "l",linesep = "", caption = cap
+) %>% kable_styling(full_width = table_format2,
+ font_size = 10,
+ position = "center",
+ latex_options = c("hold_position"));sumdat04
+```
+
+The graphs presented in Figure \@ref(fig:plotlikert) present disaggregated and comparable information of the different response categories for each item.
+It can be seen that in general there is more agreement in the perception of non meritocratic items than meritocratic ones, while in the case of preferences the opposite occurs.
+As far as preferences are concerned, the preponderant role of effort over talent as a criterion of meritocratic preference is noteworthy.
+
+```{r plotlikert, echo=FALSE, fig.cap = "Descriptive plot", fig.align='center'}
+knitr::include_graphics('../output/images/plotlikert.png')
+```
+
+In Figure Ref(fig:corpoly), moderate to high intensity relationships are observed between indicators that correspond to the same factor (for example, perception of meritocracy by effort and by talent, r=0.56).
+The relationships between meritocratic perceptions and preferences are also noteworthy, which is not the case for nonmeritocratic indicators.
+
+```{r corpoly, echo=FALSE, fig.cap = "Polychoric correlation plot", fig.align='center'}
+knitr::include_graphics('../output/images/corpoly.png')
+```
+
+In sum, descriptive analyses point to a relative adherence to meritocracy, which is expressed in a greater preference for meritocratic criteria and a lesser preference for non-meritocratic ones, even when the preference for non-meritocratic elements reaches a proportion of more than 20%.
+Likewise, a relatively low perception of meritocracy is observable.
+In addition, a coherent relationship between the indicators as proposed in the theoretical model is observed in the pre-registration of the study; that is, the pairs of items associated with a specific factor show correlations with a large effect size (for example, meritocratic preferences for items associated with effort and talent).
+In particular, the associations between effort and talent are relevant, since they dismiss previous assumptions that talent would not be a meritocratic criterion [@mijs_paradox_2019], otherwise the correlation would be zero or negative.
+In addition, we see that there is no negative correlation between meritocratic and non-meritocratic aspects, dismissing the assumptions of previous studies that these dimensions would be the opposite poles of the same continuum [@reynolds_perceptions_2014].
+
+## Confirmatory Factor Analysis
+
+This section estimates the fit of the conceptual model behind the Perceptions and Preferences for Meritocracy Scale.
+For this, we first estimate a confirmatory factor analysis model for the whole sample, and secondly we test the order effects applying the same model to each of the three order versions.
+
+### Full sample CFA
+
+\@ref(fig:meas02) shows the results of the estimation for the four-factor model with complete sample.
+This model presents adequate fit indicators (CFI = 0.977, RMSEA = 0.079, $\chi2$(df = 14) = 28,03, p = 0.014) and it presents factorial loads with weights that vary between 0.63 and 0.88.
+Furthermore, considering the correlation between the latent variables, it is observed that the variables with the highest correlations are meritocratic perceptions and preferences, which have a positive correlation of 0.59, while non-meritocratic perceptions and preferences do not present a considerable correlation.
+
+```{r meas02, echo=FALSE, fig.cap = "Confirmatory factor analysis of the Perceptions an Preferences for Meritocracy Scale", fig.align='center'}
+knitr::include_graphics('../output/images/meas02.png')
+```
+
+As can be seen in the Table \ref(tab:tab-loads), all models, regardless of the group and order of items, obtained an adequate fit, with CFI's above .95 and RMSEA's below .8.
+However, no model achieved a non-significant chi-square (something to be expected in large samples), although both the randomized and the first model obtained an adequate chi-square ratio of less than 3.
+The first order was the one that obtained the best fit (CFI=0.998, TLI=0.995,RMSEA=0.037, $\chi2$(df=14)=28.03, p = 0.014), followed by the random order of items (CFI=0.992, TLI=0.984,RMSEA=0.051, $\chi2$(df=14)=39.09, p < 0.001).
+On the other hand, the scale ordered by themes seems to generate a framing effect in which the relationship between perceptions and preferences seems overestimated, thus affecting the adjustment (CFI=0.984, TLI=0.968, RMSEA=.071, $\chi2$(df=14)=64.156, p < 0.001).
+
+```{r tab03, echo=FALSE,eval=TRUE}
+(acá recuperar syntax)
+```
+
+Although all tests obtained relatively adequate indicators, the mentioned differences between the adjustments of the models according to order are statistically significant, which was evaluated from an invariance analysis.
+It is concluded that among the three orders there is no configurtional invariance, that is, they do not have the same dimensionality and therefore do not equally fit the theoretical model [@vandenbergReviewSynthesisMeasurement2000].
+This is due to the effect produced by the joint appearance of the items of the same factor in order 1, which increases the fit of the model.
+Furthermore, in order 2, when asking about the perception and preference around the same indicator, the cross relations increase and the model fit decreases.
+Consistently, the random model presented an intermediate adjustment between order 1 and order 2.
+
+\pagebreak
+
+|           | Contrast Model 1 Factor | theoretical model 4 Factors | Model with M.I. |
+|:---------:|:-----------------------:|:---------------------------:|:---------------:|
+| **n**     |           1769          |             1769            |       1769      |
+| **CFI**   |          0.595          |            0.988            |      0.994      |
+| **TLI**   |          0.433          |            0.976            |      0.985      |
+| **RMSEA** |          0.226          |            0.047            |      0.036      |
+| **$\chi2$**    |         1830.839        |            68.661           |      40.250     |
+| **p**     |           .000          |             .000            |       .000      |
+| **$\chi2$/df** |          65.38          |             4.90            |       3.32      |
+Table: `r as.character(paste("(\\#tab:items-dim)", "Items according to dimension."))`.
+
+As shown in table 4, the proposed theoretical four-factor model fits better than the one-factor contrast model.
+The theoretical model adjusted relatively well, as it shows optimal indicators for IFC=0.987, TLI=0.975 and RMSEA=.041, although it has poor indicators for the test $\chi2$(df=14)=67.6, p-value=.000.
+To evaluate possible improvements in the scale, the relationships proposed by the modification indexes were analyzed.
+These indicate the existence of two unspecified cross-loads.
+When a model was generated following these recommendations, there was a considerable improvement in the model, although the new model also did not obtain a $\chi2$ ratio less than 3 and obtained very low factorial loads ($\tau$)< 0.15.
+Therefore, following the recommendation of Brown (2006) to only accept the proposals of modification rates when theory and solid evidence are available, it has been decided not to incorporate these parameters into the model.
+
+# Conclusión
+
+Studies that attempt to characterize and compare societies regarding support for meritocratic beliefs have used different approaches.
+As most studies use secondary survey data, they tend to grasp in the available indicators assuming that they represent an underlying construct.
+A review of these studies reveals several non-tested assumptions, as well as the use of similar indicators to represent different constructs and dimensions of meritocracy.
+As the existence of heterogeneous approaches certainly have consequences for advancing in the study of meritocracy, this paper presented a comprehensive conceptual framework for the empirical study of meritocracy built upon previous research, which is then tested against new survey data.
+
+We identified four critical aspects regarding the measurement of meritocracy in previous studies: unidimensionality, the ambiguous use of “beliefs”, the use of non-meritocratic indicators as opposed to meritocratic and the consideration of measurement error.
+The proposed 8-items scale Perceptions And Preferences for Meritocracy was designed and tested in order to deal with these four issues.
+The results indicate that, on the one hand, perceptions and preferences proved to be two related but different dimensions, usually confounded in previous research under the label of "beliefs".
+On the other hand, meritocratic and non-meritocratic dimensions would not be the poles of the same continuum as some previous studies assumed.
+
+The four-dimensional framework and structure of the Perceptions and Preferences for Meritocracy scale opens several avenues for future research.
+For instance, distinguishing perceptions from preferences will allow us to evaluate the extent to which different societies are accustomed or satisfied with the performance of meritocracy, in terms of difference of what is perceived from what is preferred.
+On the other hand, given that non-meritocratic elements are not necessarily related to the meritocratic ones, it makes it possible to assess the legitimacy of practices such as the use of personal contacts and their interference (or not) with meritocratic ideals in different societies.
+The consideration of structural factors, both individual and societal, will allow us to advance in hypotheses of meritocratic legitimacy from a more comprehensive view than the one used today based only on meritocratic perceptions.
+Furthermore, the impacts of different configurations of the four dimensional framework on practices and behaviors such as corruption, civic involvement and political views is an area that requires additional research particularly in times of economic crisis and growing inequality, and to which the proposed instrument aims at contributing.
+
+\pagebreak
+
+# References
+
+<div id="refs"> </div>  <!-- Although <div> is an HTML tag, this method also works for other output formats such as PDF. -->
+
+# (APPENDIX) Appendix {-}
+
+`r if (knitr::is_latex_output()){ '\\appendix'}`
+
+`r if (knitr::is_latex_output()){ '\\section{Appendix}'}`
+
+```{r tab-loads, echo=FALSE}
+load(file = "../output/tables/tb.fm.RData")
+cap <- "Factor loads and model fit."
+tb.col  <- c("Variables",1:4,1:4,1:4)
+tb.caption <- "Factor loadings and fit measures"
+tb.foot <- paste0("Standardized factor loadings using DWLS estimator ; CFI = Comparative fit index (scaled); RMSEA = Root mean square error of approximation (scaled)")
+
+tb.fm <- tb.fm %>%
+    mutate(vars = recode(vars,
+                         "A. Who the more they try they manage to get bigger rewards that those who striveless."= "Perception Effort",
+                         "B. Who possess more talent they manage to obtain greater rewards than those who possess less talent."= "Perception Talent",
+                         "C. Who they have rich parents manage to get out ahead."="Perception rich parents",
+                         "D. Who they have good contacts they manage to get out ahead."="Perception contacts",
+                         "E. Who the more they try they should get greater rewards than those who they try less." = "Preferences Effort",
+                         "F. Who possess more talent they should get greater rewards than those who possess less talent."= "Preferences Talent",
+                         "G. It's fine that those who have rich parents get ahead" ="Preferences rich parents",
+                         "H. Is well that those who have good contacts get ahead." = "Preferences contacts"))
+
+tb.fm01<- kable(tb.fm, format = table_format, booktabs = T,linesep = "", escape = FALSE,align = "lcccccccccccc",col.names = tb.col, caption = cap) %>%
+          kable_styling(full_width = table_format2,latex_options = c("hold_position","scale_down"), position = "center",font_size = 9) %>%
+          add_header_above(header = c(" "=1,"Version 1"= 4,"Version 2"= 4,"Version 3"= 4)) %>%
+          add_header_above(header = c(" "=1,"Factor loadings"= 12)) %>%
+          row_spec(row = 8,hline_after = TRUE) %>%
+          add_indent(c(9:12))  %>%
+          footnote(general = tb.foot,threeparttable = T);tb.fm01
+```
+
+\pagebreak
+
+|                     | Sample  |   CEP  |
+|---------------------|:-------:|:------:|
+| __Gender__          |         |        |
+| Men                 | 49,82%  | 50,52% |
+| Women               | 50.18%  | 49,47% |
+| __Age__             |         |        |
+| 18 - 24             | 18,55%  | 18,17% |
+| 25 - 34             | 18,86%  | 17,48% |
+| 35 - 44             | 19.09%  | 19,98% |
+| 45 - 54             | 17,96%  | 19,23% |
+| 55 - or more        | 25,54%  | 25.11% |
+| __Education__       |         |        |
+| Primary  or less    | 2,93%   | 15,88% |
+| Hig school          | 43,23%  | 37,04% |
+| Non university      | 32,63%  | 28,93% |
+| university or more  | 21,21%  | 18,13% |
+Table: `r as.character(paste("(\\#tab:rep-samp)", "Representativeness of the sample"))`.

### 2020-10-12 docs/paper.Rmd
keys

--- 2020-10-10 docs/paper.Rmd
@@
 Economic inequality has become an issue of growing concern around the world.
-This has been expressed in a series of protests, as well as diverse critical analyses regarding the development of capitalism and its consequences [Streeck, 2014; @piketty_Capital_2014].
-In this context, the study of visions, preferences, and perceptions of inequality has increased relevance in the social sciences, in topics such as redistributive preferences [@alesina_Fairness_2005; @Dimick_Models_2018], the legitimization of economic inequality [@schroder_Income_2017] and the functioning of meritocracy [@duru-bellat_whos_2012; @mijs_paradox_2019; @reynolds_perceptions_2014; Atria et al., 2020].
+This has been expressed in a series of protests, as well as diverse critical analyses regarding the development of capitalism and its consequences [@streeckPoliticsPublicDebt2014; @piketty_Capital_2014].
+In this context, the study of visions, preferences and perceptions of inequality has acquired relevance in the social sciences, in topics such as redistributive preferences [@alesina_Fairness_2005; @dimick_Models_2018], the legitimization of economic inequality [@schroder_Income_2017] and the functioning of meritocracy [@duru-bellat_whos_2012; @mijs_paradox_2019; @reynolds_perceptions_2014; @atriaEconomicElitesAttitudes2020].
 
@@
 Social psychology and sociology have studied the characteristics and consequences of beliefs in meritocracy, under the general hypothesis that a greater belief in meritocracy leads to greater legitimization of inequalities [@premingerMeritocracyServiceEthnocracy2020; @trumpWhenWhyEconomic2020; @hadjar_meritokratie_2008; @MadeiraPrimesConsequencesSystematic2019].
-Such research has raised criticism of meritocracy as a moral standard of distribution, arguing that it would be an unfulfilled promise, given the preponderant influence of non-meritocratic elements upon individual status [Sandel, 2020; @witteveenReconsideringMeritocraticPower2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019; @khan_Privilege_2013].
+Such research has raised criticism of meritocracy as a moral standard of distribution, arguing that it would be an unfulfilled promise, given the preponderant influence of non-meritocratic elements upon individual status [@sandelTyrannyMeritWhat2020a; @witteveenReconsideringMeritocraticPower2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019; @khan_Privilege_2013].
 
@@
 This conceptual and measurement-based assumption of unidimensionality is found in other studies, that also assume effort to be the main and only aspect of meritocracy [@girerdNeoliberalismIdeologicalBarrier2020a; @bubakPerceptionsMeritocracyNote2019].
-Conversely, recently Mijs & Hoy (2020) also considered talent as an indicator of meritocracy.
+Conversely, recently @mijsHowInformationEconomic2020 also considered talent as an indicator of meritocracy.
 
@@
 
-The term belief has an ambiguous character in the literature, conceived as "idea-elements" by Converse (1964) or "considerations" by Zaller (1992).
-As Kluegel & Smith pointed out about the scope of beliefs: "This usage encompasses such more specific social-psychological concepts as values, perceptions, and attitudes" (p.30).
+The term belief has an ambiguous character in the literature, conceived as "idea-elements" by @converseNatureBeliefSystems1964 or "considerations" by @zallerNatureOriginsMass1992.
+As @kluegelBeliefsInequalityAmericans2017 pointed out about the scope of beliefs: "This usage encompasses such more specific social-psychological concepts as values, perceptions, and attitudes" (p.30).
 Therefore, beliefs cover almost anything related to subjective factors.
@@
 
-Finally, most of the studies in meritocracy so far have not accounted for the issue of measurement error [@brownBookReviewTimothy2010, @bollenStructuralEquationsLatent1989] (Ansolabehere, S., Rodden, J., & Snyder Jr, J.
+Finally, most of the studies in meritocracy so far have not accounted for the issue of measurement error [@brownBookReviewTimothy2010; @bollenStructuralEquationsLatent1989] (Ansolabehere, S., Rodden, J., & Snyder Jr, J.
 M., 2008), when using single indicators and/or simple average indexes for measuring meritocracy.
@@
 Regarding the selection of indicators, most are taken or adapted from previous studies for the sake of comparability.
-For meritocratic indicators, we use effort and talent as the main components of the traditional concept of merit as defined by @young_rise_1962 (Atria et al.
-2020), whereas for non-meritocratic dimensions, we use wealthy parents and good contacts.
+For meritocratic indicators, we use effort and talent as the main components of the traditional concept of merit as defined by @young_rise_1962, whereas for non-meritocratic dimensions, we use wealthy parents and good contacts.
 Descriptions of the specific items are outlined in the methodology section.
@@
 The sample was selected from a non-probabilistic quota design in three large cities in Chile.
-The quotas for gender, age, and education level were generated based on a survey by the Public Studies Center (CEP, 2019), which is a well regarded counterpart agency of the ISSP (International Social Survey Programme) in Chile.
+The quotas for gender, age, and education level were generated based on a survey by the Public Studies Center [@cepEncuestaCEPMayo2019], which is a well regarded counterpart agency of the ISSP (International Social Survey Programme) in Chile.
 A total sample of 2,141 people was collected, excluding those who did not answer the questions on the scale, and those who did not accept informed consent.

### 2020-12-18 docs/paper.Rmd
update validez convergente

@@
 
-# Conclusion
+## Reliability and validity analyses
+
+We performed two further analyses in order to add evidence regarding the reliability and validity of the scale.
+Firstly we test the stability of the scale scores along time through longitudinal measurement invariance methods, and secondly we analyze the convergent validity by assessing the association of the merit scale with related concepts and correspondent measurements.
+
+### Longitudinal measurement invariance
+
+### Convergent validity
+
+_Data_
+
+For this analysis we examined information of the third wave of the study, which included additional measures that allow testing the correlations of the merit scale with some related constructs as opportunity beliefs and personal wherewithal.
+After listwise deletion of missing cases in socio-demographics information, a total of 1422 individuals took part in the third wave: 668 (46.97%) women and 754 (53.03%) men.
+The ages of 50.57% of them are 45 years or older, and 36.28% held a tertiary degree.1
+
+_Intruments_
+
+The following scales were included for testing their correlation with the the 8-item meritocracy scale:
+
+- Opportunity beliefs: We used two items of the social inequality module of the International Social Survey Programme.
+Studies use generally both indicators to measure meritocratic and non-meritocratic beliefs (McCall et al., 2017; Mijs, 2019).
+One item measures the importance of individual factors in determining life outcomes by asking for the importance of hard work for getting ahead in life (M=3.76, SD=0.93).
+The second item asks for the importance of coming from a wealthy family and captures the importance of structural factors in getting ahead (M=2.75, SD=1.27).
+Participants responded to each item using a 1 (not at all important) to 5 (essential) response scale.
+
+- Personal wherewithal: the questionnaire included The Neoliberal Beliefs Inventory (NBI), which consider four factors: [completar] (Bay-Cheng et al., 2015).
+We used the factor personal wherewithal that reflects meritocracy beliefs in terms of the importance of personal attributes as strength and skills to yield success.
+The construct consists of 8 items (e.g., “Any goal can be achieved with enough hard work and talent”, “I’ve benefited from working hard, so there’s no reason others can’t”, “Anyone who is willing to work hard can be succesful in Chile”).
+Participants answered the items using a 1 (totally disagree) to 6 (totally agree) scale.
+We simple computed mean scores [reportar alfa the cronbach], with higher scores indicating stronger support for meritocracy (M=3.32, SD=0.89, =0.89).
+
+Both instruments cover mainly the perceptual side of the concept.
+Therefore, in general we expect larger correlation with meritocratic perceptions than with preferences.
+In the case of opportunity beliefs, the hard work item should correlate positively with meritocratic perception whereas the importance for coming from a wealthy family is expected to correlate positively with non-meritocratic perceptions.
+Regarding the second instrument of personal wherewithal, we also anticipate a positive correlation with meritocratic perception and low to null correlations with the other dimensions of the merit scale.
+
+_Results_
+
+We used polyserial correlations to evaluate relationships of meritocractic and non-meritocratic preferences and perceptions with the items of the opportunity beliefs battery.
+For personal wherewithal,  we examined Pearson correlations.
+
+Table XX about here.
+
+Table X shows ... [completar ]. The findings provide evidence for convergent validity for meritocratic perceptions [re-escribir en el orden de los instrumentos ].
+As expected, we see a substantial positive correlation for the association between meritocratic perception and personal wherewithal.
+Finding for the association of non-meritocratic perception and non-meritocratic item of opportunity beliefs battery is also consistent with our expectation, insofar as the size of the correlation is relevant and the direction of it is positive.
+For the rest of correlations, we note weak correlations but most directions are consistent with our expectations.
+One exception is personal wherewithal, with a positive association for this construct.
+Correlations for meritocratic preference are also weak.
+
+Table XX. Correlations
+
+[insertar tabla imagen o markdown]
+
+Coming from a wealthy family
+Hard work
+Personal wherewithal
+Meritocratic
+Perception
+-0,1975
+0,1877
+0,3921
+Unmeritocratic
+Perception
+0,3547
+-0,0775
+-0,1844
+Meritocratic
+Preferences
+0,1078
+0,0779
+0,0819
+Unmeritocratic
+Preferences
+-0,0397
+0,0824
+0,2082
+
+# Conclusions
 

### 2020-12-21 docs/paper.Rmd
save-files

@@
 # clean #in the yml
-rmarkdown::render("docs/paper.Rmd", output_format = "bookdown::pdf_document2")
-rmarkdown::render("docs/paper.Rmd", output_format = "bookdown::html_document2")
+rmarkdown::render("paper.Rmd", output_format = "bookdown::pdf_document2")
+rmarkdown::render("paper.Rmd", output_format = "bookdown::html_document2")
 ```
@@
 knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = FALSE,out.width = '85%',fig.pos= "H")
-# knitr::opts_knit$set(base.url = "../") #relative path for .html output file
-# knitr::opts_knit$set(root.dir = "../") #relative path for chunks within .rmd files
+# knitr::opts_knit$set(base.url = NULL) #relative path for .html output file
+# knitr::opts_knit$set(root.dir = NULL) #relative path for chunks within .rmd files
 options(scipen=999)
@@
  "H. Preferences contacts" = pref_netw )
-sum04b <- psych::describe(dat04b[,-c(1)], fast = T) %>% select(mean,sd,min,max) %>% data.frame() %>% round(digits = 2)
+sum04b <- psych::describe(dat04b[,-c(1)], fast = T) %>% 
+  as.data.frame() %>%
+  select(mean,sd,min,max) %>% 
+  round(digits = 2)
 table04b<- sum04b %>% rename("Mean"=mean, "SD"=sd, "Min"=min,"Max"=max)

### 2020-12-21 production/prod_analysis-group_invariance.rmd
save-files

---
title: "Group invariance "
date: "`r format(Sys.time(), '%d %B, %Y')`"
css: "../input/bib/style.css"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: "hide"
editor_options: 
  chunk_output_type: console
---

```{r eval=FALSE, include=FALSE}
rmarkdown::render(input = "production/prod_analysis-group_invariance.Rmd",output_format = "html_document"); browseURL(url = "production/")
```

```{r include=FALSE}
knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = TRUE)
knitr::opts_knit$set(root.dir = "../")
options(scipen=999)
rm(list=ls())
options(knitr.kable.NA = '')
```

# Analysis

**load libraries**
```{r}
library(sjPlot)
library(dplyr)
library(lavaan)
library(semPlot)
library(semTools)
library(stargazer)
library(corrplot)
library(psych)
library(knitr)
library(kableExtra)

```

**load data**
```{r}
load(file = "input/data/proc/pov01_fact02.RData")
names(pov_fs)
```

```{r}
pov01 <- pov_fs %>% filter(dataset=="pvw01")
fs01  <- pov_fs %>% filter(dataset=="fsw02")
```

Meassurement model

```{r}
model01 <- '
perc_merit  =~ perc_effort + perc_talent
perc_nmerit =~ perc_wpart  + perc_netw
pref_merit  =~ pref_effort + pref_talent
pref_nmerit =~ pref_wpart  + pref_netw'
```

## Poverty Experiment Sample (wave 01)
```{r}
pov1_c <- cfa(model = model01,data = pov01,estimator="MLR",std.lv=FALSE) # Continuous/ estimator ML Robust

pov1_o <- cfa(model = model01,data = pov01,ordered = c("perc_effort","perc_talent",
                                                       "perc_wpart","perc_netw",
                                                       "pref_effort","pref_talent",
                                                       "pref_wpart","pref_netw"),std.lv=FALSE)
#---------------------------------------------------#
cnames <- c("Factor","Indicator","Loading (MLR)","Loading (DWLS)")
kable(left_join(x = standardizedsolution(pov1_c) %>% 
                  filter(op=="=~") %>% 
                  select(lhs,rhs,est.std),y = standardizedsolution(pov1_o) %>% 
                  filter(op=="=~") %>%
                  select(lhs,rhs,est.std),c("lhs","rhs")),
      format = "markdown",digits = 2,col.names = cnames, caption = "Factor loadings")
```

```{r eval=FALSE, include=FALSE}
fitmeasures(pov1_c, output ="matrix")[c("chisq.scaled","df","cfi.scaled","rmsea.scaled"),]
fitmeasures(pov1_o, output ="matrix")[c("chisq.scaled","df","cfi.scaled","rmsea.scaled"),]
```

## Factorial Survey sample (wave 02)

```{r}
fs1_c <- cfa(model = model01,data = fs01,estimator="MLR",std.lv=FALSE) # Continuous/ estimator ML Robust

fs1_o <- cfa(model = model01,data = fs01,ordered = c("perc_effort","perc_talent",
                                                      "perc_wpart","perc_netw",
                                                      "pref_effort","pref_talent",
                                                      "pref_wpart","pref_netw"),std.lv=FALSE)
#---------------------------------------------------#
cnames <- c("Factor","Indicator","Loading (MLR)","Loading (DWLS)")
kable(left_join(x = standardizedsolution(fs1_c) %>% 
                  filter(op=="=~") %>% 
                  select(lhs,rhs,est.std),y = standardizedsolution(fs1_o) %>% 
                  filter(op=="=~") %>%
                  select(lhs,rhs,est.std),c("lhs","rhs")),
      format = "markdown",digits = 2,col.names = cnames, caption = "Factor loadings")
```

```{r eval=FALSE, include=FALSE}
fitmeasures(fs1_c, output ="matrix")[c("chisq.scaled","df","cfi.scaled","rmsea.scaled"),]
fitmeasures(fs1_o, output ="matrix")[c("chisq.scaled","df","cfi.scaled","rmsea.scaled"),]
```

## Group CFA
```{r}
group_c <- cfa(model01, 
           data = pov_fs, 
           group = "dataset",estimator="MLR",std.lv=FALSE)

group_o <- cfa(model = model01,data = pov_fs,ordered = c("perc_effort","perc_talent",
                                                        "perc_wpart","perc_netw",
                                                        "pref_effort","pref_talent",
                                                        "pref_wpart","pref_netw"),std.lv=FALSE,group = "dataset")
```

```{r}
summary(group_c, standardized=TRUE,fit.measures=TRUE)
summary(group_o, standardized=TRUE,fit.measures=TRUE)
```

```{r eval=FALSE, include=FALSE}
fitmeasures(group_c, output ="matrix")[c("chisq.scaled","df","cfi.scaled","rmsea.scaled"),]
fitmeasures(group_o, output ="matrix")[c("chisq.scaled","df","cfi.scaled","rmsea.scaled"),]
```

# Tables
## CFA

```{r}
labs<- c("A. Who the more they try they manage to get bigger rewards that those who striveless.",
         "B. Who possess more talent they manage to obtain greater rewards than those who possess less talent.",
         "C. Who they have rich parents manage to get out ahead.",
         "D. Who they have good contacts they manage to get out ahead.",
         "E. Who the more they try they should get greater rewards than those who they try less." ,
         "F. Who possess more talent they should get greater rewards than those who possess less talent." ,
         "G. It's fine that those who have rich parents get ahead",
         "H. Is well that those who have good contacts get ahead." )
```

```{r}
loads01<- data.frame(round(inspect(fs1_o,what="std")$lambda,digits = 2))
loads01[loads01==c(0.00)] <- NA; names(loads01) <- c(paste0("M1.",1:4))

for (i in names(loads01)) {
  # tb.load[,i] <- sjlabelled::as_character(tb.load[,i])
  loads01[,i] <- sprintf(loads01[,i], fmt = '%#.2f')
}
loads01[loads01=="NA"] <- ""
```

```{r tb_loads-fit-fac02}
fm01<- data.frame(t(data.frame(fitmeasures(fs1_o, output ="matrix")[c("chisq.scaled","df","cfi.scaled","rmsea.scaled"),]))); row.names(fm01) ="M1.1"

#------chi2, df------#
fm04.1 <- fm01 %>% select(chisq.scaled,df) 
fm04.1$chisq.scaled <- round(x = fm04.1$chisq.scaled,digits = 1)
fm04.1$df <- round(x = fm04.1$df,digits = 0)
fm04.1$chi2df <- paste0(fm04.1$chisq.scaled,"(",fm04.1$df,")")
fm04.1 <- select(fm04.1,"chi2df")

#------CFI, RMSEA------#
fm04.2 <- fm01 %>% select(cfi.scaled,rmsea.scaled)
for (i in names(fm04.2)) {
  fm04.2[,i] <- sprintf(fm04.2[,i], fmt = '%#.3f')
}

#------ merge fit indices ------#
fm.df      <- bind_cols(fm04.1,fm04.2)
fm.df$nobs <- c(nobs(fs1_o)) 
fm.df      <- data.frame(t(fm.df))

#------ merge ------#
tb.fm<- bind_rows(loads01,fm.df)
tb.fm<- tb.fm %>% 
  mutate(vars=c(labs,"$\\chi^2\\text{(df)}$","$\\text{CFI}$","$\\text{RMSEA}$","$N$")) %>%
  select(vars,everything())
#--table---#
tb.col  <- c("Variables",1:4)
tb.foot <- paste0("; p<0.05=$^{*}$; p<0.01=$^{**}$; p<0.001=$^{***}$. Standardised factor loadings using DWLS estimator. CFI = Comparative fit index (scaled), RMSEA = Root mean square error of approximation (scaled)")
tb.caption <- "Factor loadings and fit measures"

tb.fm01<- kable(tb.fm,escape = FALSE,align = "lcccc",col.names = tb.col, caption = tb.caption) %>% 
          kable_classic(full_width = F,font_size = 9,bootstrap_options = c("striped", "hover"),html_font = "Sans") %>% 
          add_header_above(header = c(" "=1,"Factor loadings"= 4)) %>%
          column_spec(column = 1,width = "5cm") %>% 
          column_spec(column = c(2:5),width = "0.5cm") %>% 
          row_spec(row = 8,hline_after = TRUE) %>%
          add_indent(c(9:12)) %>% 
          footnote(general =tb.foot ,footnote_as_chunk = T);tb.fm01

save(tb.fm,file = "output/tables/tb.fact-f02.RData")
```

## Invariance

```{r}
inv01<- measurementInvariance(model=model01,data=pov_fs,group="dataset",estimator = "ML",strict=TRUE)
```

```{r}
conf  <- inv01$fit.configural
weak  <- inv01$fit.loadings
strong<- inv01$fit.intercepts
strict<- inv01$fit.residuals

tab01<- anova(conf,weak,strong,strict,SB.classic=TRUE) %>% 
  as_tibble() %>% 
  select("Chisq","Df","chisq_diff"=`Chisq diff`,"df_diff"=`Df diff`,"pvalue"=`Pr(>Chisq)`) %>% 
  mutate(stars=gtools::stars.pval(pvalue),
         chisqt=paste0(round(Chisq,2)," (",Df,") "),
         decision=ifelse(pvalue>0.05,yes = "Accept",no = "Reject"),
         model=c("Configural","Weak","Strong","Strict"))

fit.meas<- bind_rows(fitmeasures(inv01$fit.configural,output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
                     fitmeasures(inv01$fit.loadings,  output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
                     fitmeasures(inv01$fit.intercepts,output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
                     fitmeasures(inv01$fit.residuals, output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),])

# compute differences in chisq, df, cfi and rmsea (90%, lower.ci - upper.ci )
fit.meas<- fit.meas %>% 
  mutate(diff.chi2 = chisq    - lag(chisq,default = first(chisq)),
         diff.df   = df       - lag(df,   default = first(df)),
         diff.cfi  = cfi      - lag(cfi,  default = first(cfi)),
         diff.rmsea   = rmsea - lag(rmsea,default = first(rmsea))) %>%
  round(3) %>% 
  mutate(rmsea.ci=paste0(rmsea," \n ", "(",rmsea.ci.lower,"-",rmsea.ci.upper,")"))

tab.inv<- bind_cols(tab01,fit.meas) %>% 
  select(model,chisqt,cfi,rmsea.ci,diff.chi2,diff.df,diff.cfi,diff.rmsea,stars,decision) %>% 
  mutate(diff.chi2=paste0(diff.chi2," (",diff.df,") ",stars)) %>% 
  select(model,chisqt,cfi,rmsea.ci,diff.chi2,diff.cfi,diff.rmsea,decision)

#clean values
tab.inv[tab.inv == c("0 (0) ")] <- NA
tab.inv[tab.inv == c(0)] <- NA
```

```{r}
col.nam <- c("Model","$\\chi^2 (\\text{df})$","CFI","RMSEA (90% CI)",
             "$\\Delta \\chi^2 (\\Delta \\text{df}$)","$\\Delta \\text{CFI}$","$\\Delta \\text{RMSEA}$","Decision")
footnote <- paste0("$N$ = ",nobs(conf),"; group 1 $n$ = ",conf@Data@nobs[[1]],"; group 2 $n$ = ",conf@Data@nobs[[2]], "\n ***p < .001")

kable(tab.inv, col.names = col.nam, booktabs=TRUE) %>% 
  kable_classic(full_width = F, bootstrap_options = c("striped", "hover"),html_font = "Serif") %>% 
  footnote(general = footnote, footnote_as_chunk = T)

save(tab.inv,file = "output/tables/tb.inv-group.RData")
```

```{r eval=FALSE, include=FALSE}
### Check an alternative method
configural        <- cfa(model01, data=pov_fs, group = "dataset", estimator="ML")
weak.invariance   <- cfa(model01, data=pov_fs, group = "dataset",group.equal = "loadings", estimator="ML")
strong.invariance <- cfa(model01, data=pov_fs, group = "dataset",group.equal = c( "loadings", "intercepts"), estimator="ML")
strict.invariance <- cfa(model01, data=pov_fs, group = "dataset",group.equal = c( "loadings", "intercepts", "residuals"), estimator="ML")
```

```{r eval=FALSE, include=FALSE}
anova(configural,weak.invariance,strong.invariance,strict.invariance)
```

```{r eval=FALSE, include=FALSE}
fitmeasures(configural       )
fitmeasures(weak.invariance  )
fitmeasures(strong.invariance) 
fitmeasures(strict.invariance) 
```

* The results are the same in both methods

* Tabla de CFA con la segunda muestra
* Tabla resultados de invarianza entre muestras  
* Modelo de reporte: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5145197/

### 2020-12-21 docs/paper.Rmd
update sección invarianza

@@
   - ../input/bib/meritocracy.bib
+  - ../input/bib/invariance.bib
 editor_options:
   chunk_output_type: console
+  markdown:
+    wrap: 72
 geometry: margin=0.78in
@@
   - \usepackage{graphicx}
-  - \usepackage{float}
   - \usepackage{booktabs}
@@
   - \usepackage{threeparttable}
+  - \usepackage{threeparttablex}
+  - \usepackage[normalem]{ulem}
+  - \usepackage{makecell}
+  - \usepackage{xcolor}
 ---
@@
 
-Economic inequality has become an issue of growing concern around the world.
-This has been expressed in a series of protests, as well as diverse critical analyses regarding the development of capitalism and its consequences [@streeckPoliticsPublicDebt2014; @piketty_Capital_2014].
-In this context, the study of visions, preferences and perceptions of inequality has acquired relevance in the social sciences, in topics such as redistributive preferences [@alesina_Fairness_2005; @dimick_Models_2018], the legitimization of economic inequality [@schroder_Income_2017] and the functioning of meritocracy [@duru-bellat_whos_2012; @mijs_paradox_2019; @reynolds_perceptions_2014; @atriaEconomicElitesAttitudes2020].
+Economic inequality has become an issue of growing concern around the
+world. This has been expressed in a series of protests, as well as
+diverse critical analyses regarding the development of capitalism and
+its consequences [@streeckPoliticsPublicDebt2014;
+@piketty_Capital_2014]. In this context, the study of visions,
+preferences and perceptions of inequality has acquired relevance in the
+social sciences, in topics such as redistributive preferences
+[@alesina_Fairness_2005; @dimick_Models_2018], the legitimization of
+economic inequality [@schroder_Income_2017] and the functioning of
+meritocracy [@duru-bellat_whos_2012; @mijs_paradox_2019;
+@reynolds_perceptions_2014; @atriaEconomicElitesAttitudes2020].
 
-In general, a meritocracy is defined as a system of distribution of resources and rewards based on individual merit, which in its original conception is a combination of talent and effort [@young_rise_1962].
-This traditional conception of merit, places in a secondary position the possible interference of structural or non-meritocratic factors, such as inheritance, personal contacts, and luck [@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995; @yairMeritocracy2007; @landWeSatTable2006; @youngRiseMeritocracy1994].
-Social psychology and sociology have studied the characteristics and consequences of beliefs in meritocracy, under the general hypothesis that a greater belief in meritocracy leads to greater legitimization of inequalities [@premingerMeritocracyServiceEthnocracy2020; @trumpWhenWhyEconomic2020; @hadjar_meritokratie_2008; @MadeiraPrimesConsequencesSystematic2019].
-Such research has raised criticism of meritocracy as a moral standard of distribution, arguing that it would be an unfulfilled promise, given the preponderant influence of non-meritocratic elements upon individual status [@sandelTyrannyMeritWhat2020a; @witteveenReconsideringMeritocraticPower2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019; @khan_Privilege_2013].
+In general, a meritocracy is defined as a system of distribution of
+resources and rewards based on individual merit, which in its original
+conception is a combination of talent and effort [@young_rise_1962].
+This traditional conception of merit, places in a secondary position the
+possible interference of structural or non-meritocratic factors, such as
+inheritance, personal contacts, and luck
+[@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995;
+@yairMeritocracy2007; @landWeSatTable2006; @youngRiseMeritocracy1994].
+Social psychology and sociology have studied the characteristics and
+consequences of beliefs in meritocracy, under the general hypothesis
+that a greater belief in meritocracy leads to greater legitimization of
+inequalities [@premingerMeritocracyServiceEthnocracy2020;
+@trumpWhenWhyEconomic2020; @hadjar_meritokratie_2008;
+@MadeiraPrimesConsequencesSystematic2019]. Such research has raised
+criticism of meritocracy as a moral standard of distribution, arguing
+that it would be an unfulfilled promise, given the preponderant
+influence of non-meritocratic elements upon individual status
+[@sandelTyrannyMeritWhat2020a;
+@witteveenReconsideringMeritocraticPower2020; @arrow_meritocracy_2000;
+@goldthorpe_myth_2003; @markovits_Meritocracy_2019;
+@khan_Privilege_2013].
 
-Due to the role that meritocratic beliefs play in the justification of individual achievement or failure in contemporary societies [@bay-chengTrackingHomoOeconomicus2015], multiple investigations have evaluated the relationship between meritocratic beliefs and personal/contextual characteristics.
-For example, some studies have linked meritocracy to the reinforcement of socio-economic, gender, and ethnic stereotypes [@MadeiraPrimesConsequencesSystematic2019; @girerdNeoliberalismIdeologicalBarrier2020; @premingerMeritocracyServiceEthnocracy2020], as well as lines of research have evaluated the effects of meritocratic beliefs in educational [@generettStoriesWeTell2020; @owensENGINESSOCIALMOBILITY2020] and organizational contexts [@perezAdvancingCareersMerit2020; @aielloNewEvidenceFirmuniversity2019].
+Due to the role that meritocratic beliefs play in the justification of
+individual achievement or failure in contemporary societies
+[@bay-chengTrackingHomoOeconomicus2015], multiple investigations have
+evaluated the relationship between meritocratic beliefs and
+personal/contextual characteristics. For example, some studies have
+linked meritocracy to the reinforcement of socio-economic, gender, and
+ethnic stereotypes [@MadeiraPrimesConsequencesSystematic2019;
+@girerdNeoliberalismIdeologicalBarrier2020;
+@premingerMeritocracyServiceEthnocracy2020], as well as lines of
+research have evaluated the effects of meritocratic beliefs in
+educational [@generettStoriesWeTell2020;
+@owensENGINESSOCIALMOBILITY2020] and organizational contexts
+[@perezAdvancingCareersMerit2020; @aielloNewEvidenceFirmuniversity2019].
 
-To measure beliefs in meritocracy, studies have generally used indicators from existing surveys, and to a lesser extent, have created ad-hoc instruments.
-However, as we will show later, the methods of measuring meritocracy vary extensively among studies.
-In many cases, similar phenomena are associated with different indicators, and conversely, often different phenomena are measured with similar indicators.
-Such trends confound not only the comparability of studies, but also the ability to understanding the effects of meritocratic beliefs across different fields.
+To measure beliefs in meritocracy, studies have generally used
+indicators from existing surveys, and to a lesser extent, have created
+ad-hoc instruments. However, as we will show later, the methods of
+measuring meritocracy vary extensively among studies. In many cases,
+similar phenomena are associated with different indicators, and
+conversely, often different phenomena are measured with similar
+indicators. Such trends confound not only the comparability of studies,
+but also the ability to understanding the effects of meritocratic
+beliefs across different fields.
 
-Based on a critical analysis of different approaches to the measurement of meritocracy to date, this article proposes a tool to both measure and relate two key aspects in the study of meritocracy: perceptions and preferences.
-Furthermore, as a second axis of analysis, it considers the elaboration of indicators regarding meritocratic and non-meritocratic aspects, demonstrating that they are not two poles of the same continuum, as some previous studies would seem to suggest.
-The measurement system outlined is oriented to generate an instrument as simple and brief as possible, so that it can be used in public opinion surveys, thus integrating meritocratic beliefs in the study of different social phenomena.
+Based on a critical analysis of different approaches to the measurement
+of meritocracy to date, this article proposes a tool to both measure and
+relate two key aspects in the study of meritocracy: perceptions and
+preferences. Furthermore, as a second axis of analysis, it considers the
+elaboration of indicators regarding meritocratic and non-meritocratic
+aspects, demonstrating that they are not two poles of the same
+continuum, as some previous studies would seem to suggest. The
+measurement system outlined is oriented to generate an instrument as
+simple and brief as possible, so that it can be used in public opinion
+surveys, thus integrating meritocratic beliefs in the study of different
+social phenomena.
 
-## The black-box of meritocratic beliefs {-#blackbox}
+## The black-box of meritocratic beliefs {#blackbox .unnumbered}
 
-Several approaches to the empirical study of meritocracy based on public opinion surveys, refer to the concept of _beliefs_, however different meanings and operationalizations usually underlie this concept.
-To illustrate this point, we will take as a main reference a recent paper about meritocracy by @mijs_paradox_2019, which makes several assumptions relating to the measurement of meritocracy that are also found in previous studies.
-The definition of meritocratic beliefs stated by Mijs is the following: "when I discuss meritocracy beliefs, I am referring to citizens’ belief in the importance of hard work relative to structural factors." [@mijs_paradox_2019, pg.9].
-In the subsequent operationalization, this is associated with the following question and indicator: "how important you think it is for getting ahead in life: (a) hard work", scored from 1 to 5 on a likert scale.
-There assumptions behind this definition are worth discussing in light of the conceptual meaning of meritocracy and its possibilities of operationalization.
+Several approaches to the empirical study of meritocracy based on public
+opinion surveys, refer to the concept of *beliefs*, however different
+meanings and operationalizations usually underlie this concept. To
+illustrate this point, we will take as a main reference a recent paper
+about meritocracy by @mijs_paradox_2019, which makes several assumptions
+relating to the measurement of meritocracy that are also found in
+previous studies. The definition of meritocratic beliefs stated by Mijs
+is the following: "when I discuss meritocracy beliefs, I am referring to
+citizens' belief in the importance of hard work relative to structural
+factors." [@mijs_paradox_2019, pg.9]. In the subsequent
+operationalization, this is associated with the following question and
+indicator: "how important you think it is for getting ahead in life: (a)
+hard work", scored from 1 to 5 on a likert scale. There assumptions
+behind this definition are worth discussing in light of the conceptual
+meaning of meritocracy and its possibilities of operationalization.
 
-a. _Unidimensionality_; Is merit only effort?
+a.  *Unidimensionality*; Is merit only effort?
 
-The item used by @mijs_paradox_2019 is part of an items' battery present in several international surveys, usually called "reasons to get ahead".
-This battery displays a series of indicators related to what people consider important to get ahead in life: hard work, education, ambition, a wealthy family, the right connections, religion, race, and gender.
-Therefore for Mijs, other aspects that could be associated with talent, such as education, are not meritocratic.
-As he points out: "Hard work is arguably the most meritocratic part of Michael Young’s equation: ‘Merit = Intelligence + Effort’, for the simple fact that intelligence itself is influenced by a non-meritocratic factor: who your parents happen to be" (p.5).
+The item used by @mijs_paradox_2019 is part of an items' battery present
+in several international surveys, usually called "reasons to get ahead".
+This battery displays a series of indicators related to what people
+consider important to get ahead in life: hard work, education, ambition,
+a wealthy family, the right connections, religion, race, and gender.
+Therefore for Mijs, other aspects that could be associated with talent,
+such as education, are not meritocratic. As he points out: "Hard work is
+arguably the most meritocratic part of Michael Young's equation: 'Merit
+= Intelligence + Effort', for the simple fact that intelligence itself
+is influenced by a non-meritocratic factor: who your parents happen to
+be" (p.5).
 
-In this measure of meritocratic belief, we can observe a couple of strong assumptions: effort would not depend on parental influence, and talent (as innate ability) is not meritocratic (contrary to Michael Youngs original conceptualization).
-The question of whether talent is or is not considered meritocratic is certainly an interesting topic to discuss from a philosophical point of view, but for those working on empirical studies it is something that should face empirical scrutiny.
-Is effort the only dimension behind the concept of meritocracy, or is it a multidimensional concept that includes other elements, such as talent (as in its original conceptualization)?
-This conceptual and measurement-based assumption of unidimensionality is found in other studies, that also assume effort to be the main and only aspect of meritocracy [@girerdNeoliberalismIdeologicalBarrier2020a; @bubakPerceptionsMeritocracyNote2019].
-Conversely, recently @mijsHowInformationEconomic2020 also considered talent as an indicator of meritocracy.
+In this measure of meritocratic belief, we can observe a couple of
+strong assumptions: effort would not depend on parental influence, and
+talent (as innate ability) is not meritocratic (contrary to Michael
+Youngs original conceptualization). The question of whether talent is or
+is not considered meritocratic is certainly an interesting topic to
+discuss from a philosophical point of view, but for those working on
+empirical studies it is something that should face empirical scrutiny.
+Is effort the only dimension behind the concept of meritocracy, or is it
+a multidimensional concept that includes other elements, such as talent
+(as in its original conceptualization)? This conceptual and
+measurement-based assumption of unidimensionality is found in other
+studies, that also assume effort to be the main and only aspect of
+meritocracy [@girerdNeoliberalismIdeologicalBarrier2020a;
+@bubakPerceptionsMeritocracyNote2019]. Conversely, recently
+@mijsHowInformationEconomic2020 also considered talent as an indicator
+of meritocracy.
 
-_b. Beliefs_
+*b. Beliefs*
 
-The most commonly used survey items for the operationalization of meritocracy is the "reasons to get ahead" battery, which asks "how important you think it is" and then lists several factors, such as effort, education, parental wealth, and contacts.
-Another version of this same battery used in several surveys - sometimes along the previous one - asks about "how important you think it _should_ be".
-Therefore, the question raised here is: Which one of both is a "belief"?
-What _is_ or what _should_ be?
+The most commonly used survey items for the operationalization of
+meritocracy is the "reasons to get ahead" battery, which asks "how
+important you think it is" and then lists several factors, such as
+effort, education, parental wealth, and contacts. Another version of
+this same battery used in several surveys - sometimes along the previous
+one - asks about "how important you think it *should* be". Therefore,
+the question raised here is: Which one of both is a "belief"? What *is*
+or what *should* be?
 
-The term belief has an ambiguous character in the literature, conceived as "idea-elements" by @converseNatureBeliefSystems1964 or "considerations" by @zallerNatureOriginsMass1992.
-As @kluegelBeliefsInequalityAmericans2017 pointed out about the scope of beliefs: "This usage encompasses such more specific social-psychological concepts as values, perceptions, and attitudes" (p.30).
-Therefore, beliefs cover almost anything related to subjective factors.
-To this regard, a relevant distinction in the field of inequality beliefs was made by @janmaatSubjectiveInequalityReview2013: "Perceptions refer to subjective estimates of existing inequality (i.e.
-thoughts about what is).
-Beliefs are here defined as normative ideas about just inequality (i.e.
-thoughts about what should be)"(p.359).
-Several papers dealing with meritocracy use the term beliefs (i.e.
-what should be), while actually referring to perceptions (i.e.
-what is).
-This occurs not only in mijs_paradox_2019, but also in @reynolds_perceptions_2014, in which the term beliefs is used to talk about perceptions, whereas other authors use general terms such as attitudes (@kunovich_systems_2007).
-The first attempt to shed light on this issue was made by @duru-bellat_whos_2012, who used the question "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" as a proxy for "desired" meritocracy (beliefs).
-They then determined "perceived" meritocracy, using the questions: “Would you say that in your country, people are rewarded for their efforts?” and “...
-people are rewarded for their skills?”.
+The term belief has an ambiguous character in the literature, conceived
+as "idea-elements" by @converseNatureBeliefSystems1964 or
+"considerations" by @zallerNatureOriginsMass1992. As
+@kluegelBeliefsInequalityAmericans2017 pointed out about the scope of
+beliefs: "This usage encompasses such more specific social-psychological
+concepts as values, perceptions, and attitudes" (p.30). Therefore,
+beliefs cover almost anything related to subjective factors. To this
+regard, a relevant distinction in the field of inequality beliefs was
+made by @janmaatSubjectiveInequalityReview2013: "Perceptions refer to
+subjective estimates of existing inequality (i.e. thoughts about what
+is). Beliefs are here defined as normative ideas about just inequality
+(i.e. thoughts about what should be)"(p.359). Several papers dealing
+with meritocracy use the term beliefs (i.e. what should be), while
+actually referring to perceptions (i.e. what is). This occurs not only
+in mijs_paradox_2019, but also in @reynolds_perceptions_2014, in which
+the term beliefs is used to talk about perceptions, whereas other
+authors use general terms such as attitudes (@kunovich_systems_2007).
+The first attempt to shed light on this issue was made by
+@duru-bellat_whos_2012, who used the question "how important should the
+number of years spent in education and training be in deciding how much
+money people ought to earn?" as a proxy for "desired" meritocracy
+(beliefs). They then determined "perceived" meritocracy, using the
+questions: "Would you say that in your country, people are rewarded for
+their efforts?" and "... people are rewarded for their skills?".
 
-Is the belief in meritocracy a perception or a preference with normative meaning?
-In order to expand the analytical conceptual framework, we believe that both dimensions should be included in analyses, as proposed by @duru-bellat_whos_2012.
-This opens up the possibility of analyzing whether perceptions and preferences are actually related (i.e.
-have a correlation close to 1), or whether they are independent aspects of the same phenomenon.
-As @son_hing_merit_2011-1 have pointed out, "People can believe that outcomes ought to be distributed on the basis of merit and yet vary in their perceptions of whether this is how society currently operates" (p.
-435).
-In other words, normative beliefs should be considered while taking perceptions into account: a strong normative belief in meritocracy may mean something totally different to someone perceiving high meritocracy, when compared to someone perceiving low meritocracy.
-To avoid the confusion generated by the term "belief", we propose the terms meritocratic preferences (what should be), and meritocratic perceptions (what is), as they better reflect the two facets of meritocracy under scrutiny.
+Is the belief in meritocracy a perception or a preference with normative
+meaning? In order to expand the analytical conceptual framework, we
+believe that both dimensions should be included in analyses, as proposed
+by @duru-bellat_whos_2012. This opens up the possibility of analyzing
+whether perceptions and preferences are actually related (i.e. have a
+correlation close to 1), or whether they are independent aspects of the
+same phenomenon. As @son_hing_merit_2011-1 have pointed out, "People can
+believe that outcomes ought to be distributed on the basis of merit and
+yet vary in their perceptions of whether this is how society currently
+operates" (p. 435). In other words, normative beliefs should be
+considered while taking perceptions into account: a strong normative
+belief in meritocracy may mean something totally different to someone
+perceiving high meritocracy, when compared to someone perceiving low
+meritocracy. To avoid the confusion generated by the term "belief", we
+propose the terms meritocratic preferences (what should be), and
+meritocratic perceptions (what is), as they better reflect the two
+facets of meritocracy under scrutiny.
 
-_c. Non-meritocratic aspects_
+*c. Non-meritocratic aspects*
 
-Unlike studies that consider talent to be a non-meritocratic aspect, and therefore rule it out of the operationalization of meritocracy, @kunovich_systems_2007 follow a different approach, by including some non-meritocratic elements.
-Using an items’ battery listing a number of factors in relation to “How important each should be in deciding pay...” (as @duru-bellat_whos_2012 for desired meritocracy), they consider factors such as education and responsibility as meritocratic, giving them a value of 1 if considered “essential” in the scale response, whereas factors such as having a family or children are valued 1 when rated as "not important at all" (i.e.
-reverse coded).
-The assumption behind this approach is that rejecting a supposed non-meritocratic aspect (as having family and/or children) implies a larger belief in meritocracy.
-A similar approach of reverse-coding non-meritocratic items was taken by @newman_false_2015, using the same principles applied in the "Preference for the Merit Principle Scale" [@davey_preference_1999].
+Unlike studies that consider talent to be a non-meritocratic aspect, and
+therefore rule it out of the operationalization of meritocracy,
+@kunovich_systems_2007 follow a different approach, by including some
+non-meritocratic elements. Using an items' battery listing a number of
+factors in relation to "How important each should be in deciding pay..."
+(as @duru-bellat_whos_2012 for desired meritocracy), they consider
+factors such as education and responsibility as meritocratic, giving
+them a value of 1 if considered "essential" in the scale response,
+whereas factors such as having a family or children are valued 1 when
+rated as "not important at all" (i.e. reverse coded). The assumption
+behind this approach is that rejecting a supposed non-meritocratic
+aspect (as having family and/or children) implies a larger belief in
+meritocracy. A similar approach of reverse-coding non-meritocratic items
+was taken by @newman_false_2015, using the same principles applied in
+the "Preference for the Merit Principle Scale" [@davey_preference_1999].
 
-The assumption that meritocratic and non-meritocratic elements are poles of the same continuum was tested by @reynolds_perceptions_2014 using the "get ahead" perceptions' battery items mentioned above.
-They considered education, ambition, and hard work as meritocratic, and other factors such as family wealth and connections as non-meritocratic.
-Despite making and proving this distinction, however, the authors end up subtracting one dimension from the other, thus coming back to the assumption that they are two poles of the same continuum as @kunovich_systems_2007 did.
-We suggest that non-meritocratic factors should be part of meritocratic measurements, but should be considered independently, rather than adding to or subtracting from meritocratic factors, unless it is empirically proved that they belong to the same conceptual dimension.
+The assumption that meritocratic and non-meritocratic elements are poles
+of the same continuum was tested by @reynolds_perceptions_2014 using the
+"get ahead" perceptions' battery items mentioned above. They considered
+education, ambition, and hard work as meritocratic, and other factors
+such as family wealth and connections as non-meritocratic. Despite
+making and proving this distinction, however, the authors end up
+subtracting one dimension from the other, thus coming back to the
+assumption that they are two poles of the same continuum as
+@kunovich_systems_2007 did. We suggest that non-meritocratic factors
+should be part of meritocratic measurements, but should be considered
+independently, rather than adding to or subtracting from meritocratic
+factors, unless it is empirically proved that they belong to the same
+conceptual dimension.
 
-_d. Accounting for measurement error_
+*d. Accounting for measurement error*
 
-Finally, most of the studies in meritocracy so far have not accounted for the issue of measurement error [@brownBookReviewTimothy2010; @bollenStructuralEquationsLatent1989] (Ansolabehere, S., Rodden, J., & Snyder Jr, J.
-M., 2008), when using single indicators and/or simple average indexes for measuring meritocracy.
-Such strategy assumes that the latent construct is measured perfectly by the indicators chosen, going as far as to propose that "...
-In choosing this strategy of index construction, we argue that *support for meritocracy is not a latent variable* [@kunovich_systems_2007 p.653-654].
-Some advances were made by @reynolds_perceptions_2014 through conducting a principal component analysis of meritocratic and non-meritocratic dimensions, but somewhat contradictorily, they chose a sum index, despite proving a multidimensional latent structure.
+Finally, most of the studies in meritocracy so far have not accounted
+for the issue of measurement error [@brownBookReviewTimothy2010;
+@bollenStructuralEquationsLatent1989] (Ansolabehere, S., Rodden, J., &
+Snyder Jr, J. M., 2008), when using single indicators and/or simple
+average indexes for measuring meritocracy. Such strategy assumes that
+the latent construct is measured perfectly by the indicators chosen,
+going as far as to propose that "... In choosing this strategy of index
+construction, we argue that *support for meritocracy is not a latent
+variable* [@kunovich_systems_2007 p.653-654]. Some advances were made by
+@reynolds_perceptions_2014 through conducting a principal component
+analysis of meritocratic and non-meritocratic dimensions, but somewhat
+contradictorily, they chose a sum index, despite proving a
+multidimensional latent structure.
 
-## An instrument proposal {-#instrumentprop}
+## An instrument proposal {#instrumentprop .unnumbered}
 
-Based on the previous assumptions and limitations of the measurements of meritocracy, we propose and test an instrument with the following characteristics:
+Based on the previous assumptions and limitations of the measurements of
+meritocracy, we propose and test an instrument with the following
+characteristics:
 
-- _Multidimensionality_, incorporating previous distinctions between preferences and perceptions, as well as between meritocratic and non-meritocratic aspects.
+-   *Multidimensionality*, incorporating previous distinctions between
+    preferences and perceptions, as well as between meritocratic and
+    non-meritocratic aspects.
 
-- Multiple indicators for each dimension, in order to _account for measurement error_ in a confirmatory factor analysis context.
+-   Multiple indicators for each dimension, in order to *account for
+    measurement error* in a confirmatory factor analysis context.
 
-- Based on _previous indicators_ as much as possible, for the sake of comparability between studies.
+-   Based on *previous indicators* as much as possible, for the sake of
+    comparability between studies.
 
-- _Brief_, as to be used in regular public opinion surveys.
-In this respect, it differs from the proposal of  "Preference for the Merit Principle Scale" [@davey_preference_1999], as they use 15 items for just one dimension (aside from the problem of reverse-coding non-meritocratic items).
+-   *Brief*, as to be used in regular public opinion surveys. In this
+    respect, it differs from the proposal of "Preference for the Merit
+    Principle Scale" [@davey_preference_1999], as they use 15 items for
+    just one dimension (aside from the problem of reverse-coding
+    non-meritocratic items).
 
-The proposed measurement framework is depicted in Figure \@ref(fig:merit-model):
+The proposed measurement framework is depicted in Figure
+\@ref(fig:merit-model):
 
@@
 
-The columns "Perceptions" and "Preferences" represent the distinction between these two concepts, usually confused under the label "beliefs".
-Perceptions refer to the extent to which people observe that meritocracy works or applies in their society, which relates to items such as "I think hard work is important to get ahead in society".
-Preferences refer to normative expectations that are usually linked to a "should" expression (e.g.
-whether hard work should be related to payment).
-The rows in Figure 1 consider the distinction between meritocratic and non-meritocratic dimensions [@reynolds_perceptions_2014].
-These aspects have usually been treated as different ends of the same continuum in previous research, an assumption that requires empirical scrutiny.
-Non-meritocratic elements usually refer to the use of personal contacts or family advantages to get ahead in life.
+The columns "Perceptions" and "Preferences" represent the distinction
+between these two concepts, usually confused under the label "beliefs".
+Perceptions refer to the extent to which people observe that meritocracy
+works or applies in their society, which relates to items such as "I
+think hard work is important to get ahead in society". Preferences refer
+to normative expectations that are usually linked to a "should"
+expression (e.g. whether hard work should be related to payment). The
+rows in Figure 1 consider the distinction between meritocratic and
+non-meritocratic dimensions [@reynolds_perceptions_2014]. These aspects
+have usually been treated as different ends of the same continuum in
+previous research, an assumption that requires empirical scrutiny.
+Non-meritocratic elements usually refer to the use of personal contacts
+or family advantages to get ahead in life.
 
-Regarding the selection of indicators, most are taken or adapted from previous studies for the sake of comparability.
-For meritocratic indicators, we use effort and talent as the main components of the traditional concept of merit as defined by @young_rise_1962, whereas for non-meritocratic dimensions, we use wealthy parents and good contacts.
-Descriptions of the specific items are outlined in the methodology section.
+Regarding the selection of indicators, most are taken or adapted from
+previous studies for the sake of comparability. For meritocratic
+indicators, we use effort and talent as the main components of the
+traditional concept of merit as defined by @young_rise_1962, whereas for
+non-meritocratic dimensions, we use wealthy parents and good contacts.
+Descriptions of the specific items are outlined in the methodology
+section.
 
-The research hypotheses behind this conceptualization and measurement model are as follows:
+The research hypotheses behind this conceptualization and measurement
+model are as follows:
 
-* $H_1$. The perception of meritocracy is a latent variable based on indicators of the importance attributed to talent and effort to get ahead in life.
+-   $H_1$. The perception of meritocracy is a latent variable based on
+    indicators of the importance attributed to talent and effort to get
+    ahead in life.
 
-* $H_2$. The perception of non-meritocracy is a latent variable that derives from two indicators related to agreement with the statement that people with contacts and rich parents manage to get ahead.
+-   $H_2$. The perception of non-meritocracy is a latent variable that
+    derives from two indicators related to agreement with the statement
+    that people with contacts and rich parents manage to get ahead.
 
-* $H_3$. Meritocratic preferences are a latent variable based on the normative value of effort and talent.
+-   $H_3$. Meritocratic preferences are a latent variable based on the
+    normative value of effort and talent.
 
-* $H_4$. Non-meritocratic preferences are a latent variable based on the normative value of the use of personal contacts and having wealthy parents.
+-   $H_4$. Non-meritocratic preferences are a latent variable based on
+    the normative value of the use of personal contacts and having
+    wealthy parents.
 
@@
 
-## Data collection {-#data}
+## Data collection {#data .unnumbered}
 
-The data was obtained through an online questionnaire, which was part of a larger study on meritocracy and preferences developed in Chile in 2019 funded by the national scientific agency ANID.
-The questionnaire was programmed in Qualtrics and the fieldwork was conducted by an external online survey agency ([netquest.cl](www.netquest.cl)) between December 2019 and January 2020.
-The sample was selected from a non-probabilistic quota design in three large cities in Chile.
-The quotas for gender, age, and education level were generated based on a survey by the Public Studies Center [@cepEncuestaCEPMayo2019], which is a well regarded counterpart agency of the ISSP (International Social Survey Programme) in Chile.
-A total sample of 2,141 people was collected, excluding those who did not answer the questions on the scale, and those who did not accept informed consent.
-There were no significant differences between our sample and the wider population for most socio-demographic characteristics, with the exception of educational level (see Table \@ref(tab:rep-samp) in appendix).
-As is often the case with online surveys, there were some limitations in achieving the quotas for lower educational levels.
+The data was obtained through an online questionnaire, which was part of
+a larger study on meritocracy and preferences developed in Chile in 2019
+funded by the national scientific agency ANID. The questionnaire was
+programmed in Qualtrics and the fieldwork was conducted by an external
+online survey agency ([netquest.cl](www.netquest.cl)) between December
+2019 and January 2020. The sample was selected from a non-probabilistic
+quota design in three large cities in Chile. The quotas for gender, age,
+and education level were generated based on a survey by the Public
+Studies Center [@cepEncuestaCEPMayo2019], which is a well regarded
+counterpart agency of the ISSP (International Social Survey Programme)
+in Chile. A total sample of 2,141 people was collected, excluding those
+who did not answer the questions on the scale, and those who did not
+accept informed consent. There were no significant differences between
+our sample and the wider population for most socio-demographic
+characteristics, with the exception of educational level (see Table
+\@ref(tab:rep-samp) in appendix). As is often the case with online
+surveys, there were some limitations in achieving the quotas for lower
+educational levels.
 
@@
 
-The proposed scale of perceptions and preferences relating to meritocracy consisted of eight indicators that were grouped into the four dimensions listed earlier: Perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic).
-In order to achieve at least some comparability with previous studies, the questions were adapted from the items battery "reasons to get ahead" (ISSP/GSS), which has been widely used for operationalizing meritocracy in previous studies [@mijs_paradox_2019; @duru-bellat_whos_2012; @reynolds_perceptions_2014].
-The aforementioned eight items, ordered according to dimensions, are presented in Table \@ref(tab:table-indicadores).
-These eight likert-type items have five response alternatives, ranging from "Completely disagree"(1) to "Completely agree" (5).
+The proposed scale of perceptions and preferences relating to
+meritocracy consisted of eight indicators that were grouped into the
+four dimensions listed earlier: Perceptions
+(meritocratic/non-meritocratic) and preferences
+(meritocratic/non-meritocratic). In order to achieve at least some
+comparability with previous studies, the questions were adapted from the
+items battery "reasons to get ahead" (ISSP/GSS), which has been widely
+used for operationalizing meritocracy in previous studies
+[@mijs_paradox_2019; @duru-bellat_whos_2012;
+@reynolds_perceptions_2014]. The aforementioned eight items, ordered
+according to dimensions, are presented in Table
+\@ref(tab:table-indicadores). These eight likert-type items have five
+response alternatives, ranging from "Completely disagree"(1) to
+"Completely agree" (5).
 
@@
 
-With the objective of evaluating the effect of indicator ordering, the respondents (_n = 2141_) were randomly divided into three different set-order versions, as explained in Figure \@ref(fig:appmod).
-The scale was presented to the first group (_n = 712_) in the order that appears in Table 2.
-For the second group (_n = 717_), the order of the items was reorganized according to the topics of the items, e.g.
-for the topic of hard work, the item about perception was followed by the item about preference, and the same for the rest of the topics.
-Finally, for the third group (_n = 712_), the items were completely randomized.
+With the objective of evaluating the effect of indicator ordering, the
+respondents (*n = 2141*) were randomly divided into three different
+set-order versions, as explained in Figure \@ref(fig:appmod). The scale
+was presented to the first group (*n = 712*) in the order that appears
+in Table 2. For the second group (*n = 717*), the order of the items was
+reorganized according to the topics of the items, e.g. for the topic of
+hard work, the item about perception was followed by the item about
+preference, and the same for the rest of the topics. Finally, for the
+third group (*n = 712*), the items were completely randomized.
 
@@
 
-To test the scale's underlying constructs, we employed confirmatory factor analysis models (CFA).
-The models estimated one factor for each of the proposed dimensions, as represented in the following figure:
+To test the scale's underlying constructs, we employed confirmatory
+factor analysis models (CFA). The models estimated one factor for each
+of the proposed dimensions, as represented in the following figure:
 
@@
 
-CFA was conducted using the `lavaan` R package (version 0.6-3; Rosseel, 2020), with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement (Kline, 2016; Rosseel, 2020).
-As recommended by Brown (2008), we assessed model fit by jointly considering the comparative fit index and Tucker-Lewis Index (CFI and TLI; acceptable fit > 0.95), Root of the average squared residual approximation (RMSEA; acceptable fit < 0.08),  Chi-square: (p-value; acceptable fit > 0.05, and Chi-square ratio > 3).
+CFA was conducted using the `lavaan` R package (version 0.6-3; Rosseel,
+2020), with diagonally weighted least squares (DWLS) estimation due to
+the items' ordinal level of measurement (Kline, 2016; Rosseel, 2020). As
+recommended by Brown (2008), we assessed model fit by jointly
+considering the comparative fit index and Tucker-Lewis Index (CFI and
+TLI; acceptable fit \> 0.95), Root of the average squared residual
+approximation (RMSEA; acceptable fit \< 0.08), Chi-square: (p-value;
+acceptable fit \> 0.05, and Chi-square ratio \> 3).
 
-A pre-registration was made in the OSF platform, available at the following link: [https://osf.io/z45y2](https://osf.io/z45y2).
-Included in this pre-registration are the hypotheses regarding the four-dimensional conceptual model underlying the scale, the variable measurement levels, the statistical tests to be performed with their respective evaluation parameters, and other important aspects of the research design.
+A pre-registration was made in the OSF platform, available at the
+following link: <https://osf.io/z45y2>. Included in this
+pre-registration are the hypotheses regarding the four-dimensional
+conceptual model underlying the scale, the variable measurement levels,
+the statistical tests to be performed with their respective evaluation
+parameters, and other important aspects of the research design.
 
@@
 
-As can be seen in the Table \@ref(tab:desc01), the indicators have values ranging from 1 (totally disagree) to 5 (totally agree).
-Averages are observed from 2.41, corresponding to preference-contacts; to 3.89, corresponding to preference-effort.
-Both indicators are consistent with a general adherence to meritocracy as reported in previous studies, privileging individual aspects such as effort (Frei et al 2020, Bucca, 2016).
+As can be seen in the Table \@ref(tab:desc01), the indicators have
+values ranging from 1 (totally disagree) to 5 (totally agree). Averages
+are observed from 2.41, corresponding to preference-contacts; to 3.89,
+corresponding to preference-effort. Both indicators are consistent with
+a general adherence to meritocracy as reported in previous studies,
+privileging individual aspects such as effort (Frei et al 2020, Bucca,
+2016).
 
@@
 
-The graphs presented in Figure \@ref(fig:plotlikert) display disaggregated and comparable information of the different response categories for each item.
-It can be seen that, in general, there is more agreement in the perception of non-meritocratic items than in meritocratic items, while in the case of preferences, the opposite occurs.
-As far as preferences are concerned, the preponderant role of effort over talent as a criterion of meritocratic preference is noteworthy.
+The graphs presented in Figure \@ref(fig:plotlikert) display
+disaggregated and comparable information of the different response
+categories for each item. It can be seen that, in general, there is more
+agreement in the perception of non-meritocratic items than in
+meritocratic items, while in the case of preferences, the opposite
+occurs. As far as preferences are concerned, the preponderant role of
+effort over talent as a criterion of meritocratic preference is
+noteworthy.
 
@@
 
-In Figure \@ref(fig:corpoly), moderate to high intensity relationships are observed between indicators that correspond to the same factor (e.g., perception of meritocracy by effort and by talent, r=0.56).
-The relationships between meritocratic perceptions and preferences are also noteworthy, which is not the case for non-meritocratic indicators.
+In Figure \@ref(fig:corpoly), moderate to high intensity relationships
+are observed between indicators that correspond to the same factor
+(e.g., perception of meritocracy by effort and by talent, r=0.56). The
+relationships between meritocratic perceptions and preferences are also
+noteworthy, which is not the case for non-meritocratic indicators.
 
@@
 
-In sum, descriptive analyses point to a relative adherence to meritocracy, which is expressed as a greater preference for meritocratic criteria, and a lesser preference for non-meritocratic criteria, even when the preference for non-meritocratic elements reaches a proportion of more than 20%.
-Likewise, a relatively low perception of meritocracy is observable.
-In addition, we observed a coherent relationship between the indicators of the same dimension as proposed in the pre-registration of the study; that is, the pairs of items associated with a specific factor show correlations with a large effect size (for example, meritocratic preferences for items associated with effort and talent).
-In particular, the associations between effort and talent are relevant, challenging previous assumptions that talent would not be a meritocratic criterion [@mijs_paradox_2019], otherwise the correlation would be zero or negative.
-In addition, we see that there is no negative correlation between meritocratic and non-meritocratic aspects, dismissing the assumptions of previous studies that these dimensions would be the opposite poles of the same continuum [@reynolds_perceptions_2014].
+In sum, descriptive analyses point to a relative adherence to
+meritocracy, which is expressed as a greater preference for meritocratic
+criteria, and a lesser preference for non-meritocratic criteria, even
+when the preference for non-meritocratic elements reaches a proportion
+of more than 20%. Likewise, a relatively low perception of meritocracy
+is observable. In addition, we observed a coherent relationship between
+the indicators of the same dimension as proposed in the pre-registration
+of the study; that is, the pairs of items associated with a specific
+factor show correlations with a large effect size (for example,
+meritocratic preferences for items associated with effort and talent).
+In particular, the associations between effort and talent are relevant,
+challenging previous assumptions that talent would not be a meritocratic
+criterion [@mijs_paradox_2019], otherwise the correlation would be zero
+or negative. In addition, we see that there is no negative correlation
+between meritocratic and non-meritocratic aspects, dismissing the
+assumptions of previous studies that these dimensions would be the
+opposite poles of the same continuum [@reynolds_perceptions_2014].
 
@@
 
-This section estimates the fit of the conceptual model behind the Perceptions and Preferences for Meritocracy Scale.
-For this, we first estimate a confirmatory factor analysis model for the whole sample, and secondly we test the order effects applying the same model to each of the three order permutations.
+This section estimates the fit of the conceptual model behind the
+Perceptions and Preferences for Meritocracy Scale. For this, we first
+estimate a confirmatory factor analysis model for the whole sample, and
+secondly we test the order effects applying the same model to each of
+the three order permutations.
 
@@
 
-Figure \@ref(fig:meas02) shows the results of the estimation for the four-factor model using the complete sample.
-This model shows adequate fit indicators (CFI = 0.977, RMSEA = 0.079, $\chi2$(df = 14) = 28,03, p = 0.014) and it presents factorial loads with weights that vary between 0.63 and 0.88.
-Considering the correlations between the latent variables, it is observed that the variables with the strongest correlations are meritocratic perceptions and preferences, which have a positive correlation of 0.59, while non-meritocratic perceptions and preferences do not show considerable correlation.
+Figure \@ref(fig:meas02) shows the results of the estimation for the
+four-factor model using the complete sample. This model shows adequate
+fit indicators (CFI = 0.977, RMSEA = 0.079, $\chi2$(df = 14) = 28,03, p
+= 0.014) and it presents factorial loads with weights that vary between
+0.63 and 0.88. Considering the correlations between the latent
+variables, it is observed that the variables with the strongest
+correlations are meritocratic perceptions and preferences, which have a
+positive correlation of 0.59, while non-meritocratic perceptions and
+preferences do not show considerable correlation.
 
@@
 
-As can be seen in the Table \@ref(tab:tab-loads), all models, regardless of the group and order of items, obtained an adequate fit, with CFI's above 0.95 and RMSEA's below 0.08.
-However, no model achieved a non-significant chi-square (something to be expected in large samples), although both the randomized and the first model obtained an adequate chi-square ratio of less than 3.
-The first order was the one that obtained the best fit (CFI=0.998, TLI=0.995, RMSEA=0.037, $\chi2$(df=14)=28.03, p = 0.014), followed by the random order of items (CFI=0.992, TLI=0.984,RMSEA=0.051, $\chi2$(df=14)=39.09, p < 0.001).
-On the other hand, the scale ordered by themes seems to generate a framing effect, in which the relationship between perceptions and preferences seemed to be overestimated, thus affecting the adjustment (CFI=0.984, TLI=0.968, RMSEA=.071, $\chi2$(df=14)=64.156, p < 0.001).
+As can be seen in the Table \@ref(tab:tab-loads), all models, regardless
+of the group and order of items, obtained an adequate fit, with CFI's
+above 0.95 and RMSEA's below 0.08. However, no model achieved a
+non-significant chi-square (something to be expected in large samples),
+although both the randomized and the first model obtained an adequate
+chi-square ratio of less than 3. The first order was the one that
+obtained the best fit (CFI=0.998, TLI=0.995, RMSEA=0.037,
+$\chi2$(df=14)=28.03, p = 0.014), followed by the random order of items
+(CFI=0.992, TLI=0.984,RMSEA=0.051, $\chi2$(df=14)=39.09, p \< 0.001). On
+the other hand, the scale ordered by themes seems to generate a framing
+effect, in which the relationship between perceptions and preferences
+seemed to be overestimated, thus affecting the adjustment (CFI=0.984,
+TLI=0.968, RMSEA=.071, $\chi2$(df=14)=64.156, p \< 0.001).
 
@@
 
-Although all tests obtained relatively adequate indicators, the aforementioned differences between the adjustments of the models according to order are statistically significant, as evaluated by an invariance analysis.
-It was concluded that among the three orders there was no configurational invariance, that is, they did not have the same dimensionality, and therefore did not equally fit the theoretical model [@vandenbergReviewSynthesisMeasurement2000].
-This is due to the effect produced by the joint appearance of items relating to the same factor in order one, which improved the fit of the model.
-Furthermore, in order two, when asking about perception and preference around the same indicator, cross relations increased and the model fit was worsened.
-Consistently, the random model presented an intermediate case between order one and order two.
+Although all tests obtained relatively adequate indicators, the
+aforementioned differences between the adjustments of the models
+according to order are statistically significant, as evaluated by an
+invariance analysis. It was concluded that among the three orders there
+was no configurational invariance, that is, they did not have the same
+dimensionality, and therefore did not equally fit the theoretical model
+[@vandenbergReviewSynthesisMeasurement2000]. This is due to the effect
+produced by the joint appearance of items relating to the same factor in
+order one, which improved the fit of the model. Furthermore, in order
+two, when asking about perception and preference around the same
+indicator, cross relations increased and the model fit was worsened.
+Consistently, the random model presented an intermediate case between
+order one and order two.
 
@@
 |            | Contrast Model 1 Factor | theoretical model 4 Factors | Model with M.I. |
-|:---------:|:-----------------------:|:---------------------------:|:---------------:|
+|:----------:|:-----------------------:|:---------------------------:|:---------------:|
 |   **n**    |          1769           |            1769             |      1769       |
@@
 | **RMSEA**  |          0.226          |            0.047            |      0.036      |
-| **$\chi2$**    |         1830.839        |            68.661           |      40.250     |
+|  $\chi2$   |        1830.839         |           68.661            |     40.250      |
 |   **p**    |          .000           |            .000             |      .000       |
-| **$\chi2$/df** |          65.38          |             4.90            |       3.32      |
-Table: `r as.character(paste("(\\#tab:items-dim)", "Items according to dimension."))`.
+| $\chi2$/df |          65.38          |            4.90             |      3.32       |
 
-As shown in Table \@ref(tab:items-dim), the proposed theoretical four-factor model fits better than the one-factor contrast model.
-The theoretical model adjusted relatively well, as it shows preferable indicators of IFC=0.987, TLI=0.975 and RMSEA=0.041, although it also has poor indicators of $\chi2$(df=14)=67.6 and p-value<0.001.
-To evaluate possible improvements in the scale, the relationships proposed by the modification indexes were analyzed.
-These indicate the existence of two unspecified cross-loads.
-When a model was generated following these recommendations, there was a considerable improvement in the model, although the new model still did not obtain a $\chi2$ ratio of less than 3, and had very low factorial loads ($\tau$)< 0.15.
-Therefore, following the recommendation of Brown (2006) to only accept the proposals of modification when theory and solid evidence are available, we decided not to incorporate these parameters into the model.
+: `r as.character(paste("(\\#tab:items-dim)", "Items according to dimension."))`.
+
+As shown in Table \@ref(tab:items-dim), the proposed theoretical
+four-factor model fits better than the one-factor contrast model. The
+theoretical model adjusted relatively well, as it shows preferable
+indicators of IFC=0.987, TLI=0.975 and RMSEA=0.041, although it also has
+poor indicators of $\chi2$(df=14)=67.6 and p-value\<0.001. To evaluate
+possible improvements in the scale, the relationships proposed by the
+modification indexes were analyzed. These indicate the existence of two
+unspecified cross-loads. When a model was generated following these
+recommendations, there was a considerable improvement in the model,
+although the new model still did not obtain a $\chi2$ ratio of less than
+3, and had very low factorial loads ($\tau$)\< 0.15. Therefore,
+following the recommendation of Brown (2006) to only accept the
+proposals of modification when theory and solid evidence are available,
+we decided not to incorporate these parameters into the model.
 
@@
 
-We performed two further analyses in order to add evidence regarding the reliability and validity of the scale.
-Firstly we test the stability of the scale scores along time through longitudinal measurement invariance methods, and secondly we analyze the convergent validity by assessing the association of the merit scale with related concepts and correspondent measurements.
+We performed two further analyses in order to add evidence regarding the
+reliability and validity of the scale. Firstly we test the stability of
+the scale scores along time through longitudinal measurement invariance
+methods, and secondly we analyze the convergent validity by assessing
+the association of the merit scale with related concepts and
+correspondent measurements.
 
-### Longitudinal measurement invariance
+### Measurement invariance
+
+*Datos*
+
+Para el desarrollo de esta sección, se empleó una fuente de datos
+externa a los resultados presentados en el apartado anterior. Los datos
+provienen de un estudio realizado durante el primer semestre de 2020,
+dentro del cual se realizó una medición usando la escala a través de
+aleatorización completa de los indicadores, lo cual corresponde a la
+modalidad de aplicación del Grupo 3 presente en la Figura
+\@ref(fig:appmod). Después de la eliminación de casos perdidos por
+lista, obtuvimos una muestra analítica final de 1242 casos donde 605
+(48,71%) son mujeres y 637 (51,28%) son hombres. El 48,84% de ellos
+tienen 45 años o más, y el 33,97% poseen educación terciaria o superior.
+En la Tabla \@ref(tab:factloads2) se pueden observar los resultados del ajuste de la
+escala en esta muestra. Análisis El modelamiento para el testeo
+invarianza consiste en una serie de modelos anidados a los que
+progresivamente se le van incorporando restricciones sobre los
+parámetros del modelo de medición , permitiendo esta progresión
+localizar la fuente de una violación de invarianza, lo cual ha sido
+tomando mayor relevancia en los estudios sociológicos a través de
+encuestas, principalmente por la validez del uso de escalas de medición
+en estudios de carácter comparado en diversos contextos sociales y
+culturales [@Davidovetal2014]. En esta sección emplearemos un análisis
+de invarianza entre grupos, donde se espera que la estructura factorial
+sea independiente de la pertenencia a las la muestra correspondiente al
+estudio original y a los datos del estudio más reciente. Para el
+desarrollo del test de invarianza, los autores sostienen que el
+modelamiento debe realizarse a través de cuatro niveles o tipos de
+restricción progresiva [@Milfont2010; @millsap_Statistical_2011;
+@vandeschoot_checklist_2012]:
+
+1.  **Configural:** Se estima el Modelo solo de indicando la estructura
+    de este, razón por lo cual todos los parámetros se estiman
+    libremente.
+
+2.  **Débil:** Posteriormente, en el modelo se restringen las cargas
+    factoriales de cada ítem, haciéndolas idénticas en ambas mediciones.
+
+3.  **Fuerte:** Luego, se estima el modelo restringiendo las cargas
+    factoriales y los interceptos de cada ítem, haciéndolos idénticos
+    para cada ambas mediciones.
+
+4.  **Estricta:** Finalmente, se estima un modelo donde se restringen
+    las cargas factoriales, los interceptos y las varianzas de error
+    para cada ítem, haciéndolos idénticos para ambas mediciones.
+
+El primer paso para el testeo de invarianza es la estimación del modelo
+Configural, el cual se espera que cumpla adecuadamente con los criterios
+de ajuste global de un modelo de medición, debido a que es el modelo
+base en proceso de comparación. Convencionalmente el estadístico
+chi-cuadrado es empleado como medida global de ajuste de un modelo de
+medición. Sin embargo, al ser sensible al tamaño de la muestra, se
+recomienda emplear tres índices de ajuste adicionales: Comparative Fit
+Index (CFI), el cual debe tener un valor mayor a 0.9; Root Mean Square
+Error Approximation (RMSEA), el cual debe estar en el rango de valores
+de 0.05 y 0.08; y el Standardised Root Mean Square Residuals (SRMR) que
+debe ser menor a 0.08 (@vandeschoot_checklist_2012)
+
+La literatura sobre invarianza métrica sugiere dos aproximaciones
+complementarias para el testeo de invarianza. En nuestro caso,
+emplearemos la comparación basada en el ajuste incremental de los
+estadísticos de ajuste [@cheung_Evaluating_2002; @Milfont2010;
+@dimitrov_Testing_2010] y el test ANOVA para comparación de medias en
+modelos anidados [@newsom_Longitudinal_2015].
+
+```{r invargroup, eval=TRUE, echo=FALSE}
+load(file = "../output/tables/tb.inv-group.RData")
+col.nam <- c("Model","$\\chi^2 (\\text{df})$","CFI","RMSEA (90 CI)",
+             "$\\Delta \\chi^2 (\\Delta \\text{df}$)","$\\Delta \\text{CFI}$","$\\Delta \\text{RMSEA}$","Decision")
+footnote <- paste0("N = ",nobs(conf),"; Group 1, n = ",conf@Data@nobs[[1]],"; Group 2, n = ",conf@Data@nobs[[2]], "***p < .001")
+
+kable(tab.inv, col.names = col.nam, align = "l",
+      booktabs=TRUE,format = table_format,escape = FALSE,
+      caption = "Multiple Group meassurement invariance Perceptions and Preferences for Meritocracy") %>% 
+  kable_styling(full_width = FALSE) %>%
+  footnote(general = footnote, footnote_as_chunk = T)
+```
+
+Como se puede observar en la Table \@ref(tab:invargroup)
 
@@
 
-_Data_
+*Data*
 
-For this analysis we examined information of the third wave of the study, which included additional measures that allow testing the correlations of the merit scale with some related constructs as opportunity beliefs and personal wherewithal.
-After listwise deletion of missing cases in socio-demographics information, a total of 1422 individuals took part in the third wave: 668 (46.97%) women and 754 (53.03%) men.
-The ages of 50.57% of them are 45 years or older, and 36.28% held a tertiary degree.1
+For this analysis we examined information of the third wave of the
+study, which included additional measures that allow testing the
+correlations of the merit scale with some related constructs as
+opportunity beliefs and personal wherewithal. After listwise deletion of
+missing cases in socio-demographics information, a total of 1422
+individuals took part in the third wave: 668 (46.97%) women and 754
+(53.03%) men. The ages of 50.57% of them are 45 years or older, and
+36.28% held a tertiary degree.1
 
-_Intruments_
+*Intruments*
 
-The following scales were included for testing their correlation with the the 8-item meritocracy scale:
+The following scales were included for testing their correlation with
+the the 8-item meritocracy scale:
 
-- Opportunity beliefs: We used two items of the social inequality module of the International Social Survey Programme.
-Studies use generally both indicators to measure meritocratic and non-meritocratic beliefs (McCall et al., 2017; Mijs, 2019).
-One item measures the importance of individual factors in determining life outcomes by asking for the importance of hard work for getting ahead in life (M=3.76, SD=0.93).
-The second item asks for the importance of coming from a wealthy family and captures the importance of structural factors in getting ahead (M=2.75, SD=1.27).
-Participants responded to each item using a 1 (not at all important) to 5 (essential) response scale.
+-   Opportunity beliefs: We used two items of the social inequality
+    module of the International Social Survey Programme. Studies use
+    generally both indicators to measure meritocratic and
+    non-meritocratic beliefs (McCall et al., 2017; Mijs, 2019). One item
+    measures the importance of individual factors in determining life
+    outcomes by asking for the importance of hard work for getting ahead
+    in life (M=3.76, SD=0.93). The second item asks for the importance
+    of coming from a wealthy family and captures the importance of
+    structural factors in getting ahead (M=2.75, SD=1.27). Participants
+    responded to each item using a 1 (not at all important) to 5
+    (essential) response scale.
 
-- Personal wherewithal: the questionnaire included The Neoliberal Beliefs Inventory (NBI), which consider four factors: [completar] (Bay-Cheng et al., 2015).
-We used the factor personal wherewithal that reflects meritocracy beliefs in terms of the importance of personal attributes as strength and skills to yield success.
-The construct consists of 8 items (e.g., “Any goal can be achieved with enough hard work and talent”, “I’ve benefited from working hard, so there’s no reason others can’t”, “Anyone who is willing to work hard can be succesful in Chile”).
-Participants answered the items using a 1 (totally disagree) to 6 (totally agree) scale.
-We simple computed mean scores [reportar alfa the cronbach], with higher scores indicating stronger support for meritocracy (M=3.32, SD=0.89, =0.89).
+-   Personal wherewithal: the questionnaire included The Neoliberal
+    Beliefs Inventory (NBI), which consider four factors: [completar]
+    (Bay-Cheng et al., 2015). We used the factor personal wherewithal
+    that reflects meritocracy beliefs in terms of the importance of
+    personal attributes as strength and skills to yield success. The
+    construct consists of 8 items (e.g., "Any goal can be achieved with
+    enough hard work and talent", "I've benefited from working hard, so
+    there's no reason others can't", "Anyone who is willing to work hard
+    can be succesful in Chile"). Participants answered the items using a
+    1 (totally disagree) to 6 (totally agree) scale. We simple computed
+    mean scores [reportar alfa the cronbach], with higher scores
+    indicating stronger support for meritocracy (M=3.32, SD=0.89,
+    =0.89).
 
 Both instruments cover mainly the perceptual side of the concept.
-Therefore, in general we expect larger correlation with meritocratic perceptions than with preferences.
-In the case of opportunity beliefs, the hard work item should correlate positively with meritocratic perception whereas the importance for coming from a wealthy family is expected to correlate positively with non-meritocratic perceptions.
-Regarding the second instrument of personal wherewithal, we also anticipate a positive correlation with meritocratic perception and low to null correlations with the other dimensions of the merit scale.
+Therefore, in general we expect larger correlation with meritocratic
+perceptions than with preferences. In the case of opportunity beliefs,
+the hard work item should correlate positively with meritocratic
+perception whereas the importance for coming from a wealthy family is
+expected to correlate positively with non-meritocratic perceptions.
+Regarding the second instrument of personal wherewithal, we also
+anticipate a positive correlation with meritocratic perception and low
+to null correlations with the other dimensions of the merit scale.
 
-_Results_
+*Results*
 
-We used polyserial correlations to evaluate relationships of meritocractic and non-meritocratic preferences and perceptions with the items of the opportunity beliefs battery.
-For personal wherewithal,  we examined Pearson correlations.
+We used polyserial correlations to evaluate relationships of
+meritocractic and non-meritocratic preferences and perceptions with the
+items of the opportunity beliefs battery. For personal wherewithal, we
+examined Pearson correlations.
 
@@
 
-Table X shows ... [completar ]. The findings provide evidence for convergent validity for meritocratic perceptions [re-escribir en el orden de los instrumentos ].
-As expected, we see a substantial positive correlation for the association between meritocratic perception and personal wherewithal.
-Finding for the association of non-meritocratic perception and non-meritocratic item of opportunity beliefs battery is also consistent with our expectation, insofar as the size of the correlation is relevant and the direction of it is positive.
-For the rest of correlations, we note weak correlations but most directions are consistent with our expectations.
-One exception is personal wherewithal, with a positive association for this construct.
-Correlations for meritocratic preference are also weak.
+Table X shows ... [completar ]. The findings provide evidence for
+convergent validity for meritocratic perceptions [re-escribir en el
+orden de los instrumentos ]. As expected, we see a substantial positive
+correlation for the association between meritocratic perception and
+personal wherewithal. Finding for the association of non-meritocratic
+perception and non-meritocratic item of opportunity beliefs battery is
+also consistent with our expectation, insofar as the size of the
+correlation is relevant and the direction of it is positive. For the
+rest of correlations, we note weak correlations but most directions are
+consistent with our expectations. One exception is personal wherewithal,
+with a positive association for this construct. Correlations for
+meritocratic preference are also weak.
 
@@
 
-Coming from a wealthy family
-Hard work
-Personal wherewithal
-Meritocratic
-Perception
--0,1975
-0,1877
-0,3921
-Unmeritocratic
-Perception
-0,3547
--0,0775
--0,1844
-Meritocratic
-Preferences
-0,1078
-0,0779
-0,0819
-Unmeritocratic
-Preferences
--0,0397
-0,0824
-0,2082
+Coming from a wealthy family Hard work Personal wherewithal Meritocratic
+Perception -0,1975 0,1877 0,3921 Unmeritocratic Perception 0,3547
+-0,0775 -0,1844 Meritocratic Preferences 0,1078 0,0779 0,0819
+Unmeritocratic Preferences -0,0397 0,0824 0,2082
 
@@
 
-Studies that attempt to characterize and compare societies by their support for meritocratic beliefs have used differencing approaches.
-As most studies use secondary survey data, they tend to assume that the available indicators represent an underlying construct.
-A review of these studies reveals several non-tested assumptions, as well as the use of similar indicators to represent different constructs and dimensions of meritocracy.
-As the existence of heterogeneous approaches certainly has consequences for the advancement of the study of meritocracy, this paper presents a comprehensive conceptual framework for the empirical study of meritocracy, building upon previous research.
-This framework was then tested against new survey data.
+Studies that attempt to characterize and compare societies by their
+support for meritocratic beliefs have used differencing approaches. As
+most studies use secondary survey data, they tend to assume that the
+available indicators represent an underlying construct. A review of
+these studies reveals several non-tested assumptions, as well as the use
+of similar indicators to represent different constructs and dimensions
+of meritocracy. As the existence of heterogeneous approaches certainly
+has consequences for the advancement of the study of meritocracy, this
+paper presents a comprehensive conceptual framework for the empirical
+study of meritocracy, building upon previous research. This framework
+was then tested against new survey data.
 
-We identified four critical aspects regarding the measurement of meritocracy in previous studies: unidimensionality, the ambiguous use of the term “beliefs”, the use of non-meritocratic indicators as opposed to meritocratic indicators, and the consideration of measurement error.
-The proposed 8-item scale, “Perceptions and Preferences for Meritocracy”, was designed and tested in order to deal with these four issues.
-Our results indicate that perceptions and preferences seem to be two related but different dimensions, often confounded in previous research under the label of "beliefs".
-Additionally, meritocratic and non-meritocratic dimensions do not appear to constitute poles of the same continuum, as some previous studies have assumed.
+We identified four critical aspects regarding the measurement of
+meritocracy in previous studies: unidimensionality, the ambiguous use of
+the term "beliefs", the use of non-meritocratic indicators as opposed to
+meritocratic indicators, and the consideration of measurement error. The
+proposed 8-item scale, "Perceptions and Preferences for Meritocracy",
+was designed and tested in order to deal with these four issues. Our
+results indicate that perceptions and preferences seem to be two related
+but different dimensions, often confounded in previous research under
+the label of "beliefs". Additionally, meritocratic and non-meritocratic
+dimensions do not appear to constitute poles of the same continuum, as
+some previous studies have assumed.
 
-The four-dimensional framework and structure of the Perceptions and Preferences for Meritocracy scale opens several avenues for future research.
-For instance, distinguishing perceptions from preferences will allow us to evaluate the extent to which different societies are accustomed to, or satisfied with, the perceived level of meritocracy, in terms of differences between what is perceived and what is preferred.
-Additionally, given that non-meritocratic factors are not necessarily related to meritocratic ones, our framework makes it possible to assess the perceived legitimacy of practices such as the use of personal contacts and their interference (or not), within meritocratic ideals in different societies.
-The consideration of structural factors, both individual and societal, will allow us to advance future hypotheses relating to meritocratic legitimacy to a more comprehensive level than the one currently used, which is based solely on meritocratic perceptions.
-Furthermore, the impacts of different configurations of the four-dimensional framework on practices and behaviors such as corruption, civic involvement, and political alignment, is an area that requires additional research.
-Such future agendas could be specially relevant in times of economic crisis and growing inequalities, that could entail changes in the legitimation of the current distributive structure based on meritocratic ideals.
+The four-dimensional framework and structure of the Perceptions and
+Preferences for Meritocracy scale opens several avenues for future
+research. For instance, distinguishing perceptions from preferences will
+allow us to evaluate the extent to which different societies are
+accustomed to, or satisfied with, the perceived level of meritocracy, in
+terms of differences between what is perceived and what is preferred.
+Additionally, given that non-meritocratic factors are not necessarily
+related to meritocratic ones, our framework makes it possible to assess
+the perceived legitimacy of practices such as the use of personal
+contacts and their interference (or not), within meritocratic ideals in
+different societies. The consideration of structural factors, both
+individual and societal, will allow us to advance future hypotheses
+relating to meritocratic legitimacy to a more comprehensive level than
+the one currently used, which is based solely on meritocratic
+perceptions. Furthermore, the impacts of different configurations of the
+four-dimensional framework on practices and behaviors such as
+corruption, civic involvement, and political alignment, is an area that
+requires additional research. Such future agendas could be specially
+relevant in times of economic crisis and growing inequalities, that
+could entail changes in the legitimation of the current distributive
+structure based on meritocratic ideals.
 
@@
 
-<div id="refs"> </div>  <!-- Although <div> is an HTML tag, this method also works for other output formats such as PDF. -->
+::: {#refs}
+:::
 
-# (APPENDIX) Appendix {-}
+<!-- Although <div> is an HTML tag, this method also works for other output formats such as PDF. -->
+
+# (APPENDIX) Appendix {.unnumbered}
 
@@
 
+```{r factloads2}
+load(file = "../output/tables/tb.fact-f02.RData")
+tb.fm01<- kable(tb.fm,escape = FALSE,align = "lcccc",col.names = tb.col, caption = tb.caption) %>% 
+          kable_styling(full_width = F,font_size = 9) %>% 
+          add_header_above(header = c(" "=1,"Factor loadings"= 4)) %>%
+          column_spec(column = 1,width = "5cm") %>% 
+          column_spec(column = c(2:5),width = "0.5cm") %>% 
+          row_spec(row = 8,hline_after = TRUE) %>%
+          add_indent(c(9:12)) %>% 
+          footnote(general =tb.foot ,footnote_as_chunk = T);tb.fm01
+```
+
 \pagebreak
@@
 |                    | Sample |  CEP   |
-|---------------------|:-------:|:------:|
-| __Gender__          |         |        |
+|--------------------|:------:|:------:|
+| **Gender**         |        |        |
 | Men                | 49,82% | 50,52% |
 | Women              | 50.18% | 49,47% |
-| __Age__             |         |        |
+| **Age**            |        |        |
 | 18 - 24            | 18,55% | 18,17% |
@@
 | 55 - or more       | 25,54% | 25.11% |
-| __Education__       |         |        |
+| **Education**      |        |        |
 | Primary or less    | 2,93%  | 15,88% |
@@
 | university or more | 21,21% | 18,13% |
-Table: `r as.character(paste("(\\#tab:rep-samp)", "Representativeness of the sample"))`.
+
+: `r as.character(paste("(\\#tab:rep-samp)", "Representativeness of the sample"))`.

### 2020-12-21 production/prod_analysis-group_invariance.rmd
update sección invarianza

@@
 CFI = Comparative fit index (scaled), RMSEA = Root mean square error of approximation (scaled)")
-tb.caption <- "Factor loadings and fit measures"
+tb.caption <- "Factor loadings and fit measuresfor the second sample"
 
@@
 
-save(tb.fm,file = "output/tables/tb.fact-f02.RData")
+save(tb.fm,tb.col,tb.foot,tb.caption,file = "output/tables/tb.fact-f02.RData")
 ```
@@
 
-save(tab.inv,file = "output/tables/tb.inv-group.RData")
+save(tab.inv,conf,file = "output/tables/tb.inv-group.RData")
 ```

### 2020-12-21 docs/paper.Rmd
update- tablas invarianza cross ref

@@
 tienen 45 años o más, y el 33,97% poseen educación terciaria o superior.
-En la Tabla \@ref(tab:factloads2) se pueden observar los resultados del ajuste de la
-escala en esta muestra. Análisis El modelamiento para el testeo
-invarianza consiste en una serie de modelos anidados a los que
+En la Tabla \@ref(tab:fcloads-inv) se pueden observar los resultados del
+ajuste de la escala en esta muestra. Análisis El modelamiento para el
+testeo invarianza consiste en una serie de modelos anidados a los que
 progresivamente se le van incorporando restricciones sobre los
@@
              "$\\Delta \\chi^2 (\\Delta \\text{df}$)","$\\Delta \\text{CFI}$","$\\Delta \\text{RMSEA}$","Decision")
-footnote <- paste0("N = ",nobs(conf),"; Group 1, n = ",conf@Data@nobs[[1]],"; Group 2, n = ",conf@Data@nobs[[2]], "***p < .001")
+footnote <- paste0("N = ",nobs(conf),"; Group 1, n = ",conf@Data@nobs[[1]],"; Group 2, n = ",conf@Data@nobs[[2]], ", ***p < 0.001")
 
@@
       booktabs=TRUE,format = table_format,escape = FALSE,
-      caption = "Multiple Group meassurement invariance Perceptions and Preferences for Meritocracy") %>% 
+      caption = "Multiple Group meassurement invariance for Perceptions and Preferences for Meritocracy") %>% 
   kable_styling(full_width = FALSE) %>%
@@
 
-```{r factloads2}
+```{r fcloads-inv, echo=FALSE}
 load(file = "../output/tables/tb.fact-f02.RData")
-tb.fm01<- kable(tb.fm,escape = FALSE,align = "lcccc",col.names = tb.col, caption = tb.caption) %>% 
-          kable_styling(full_width = F,font_size = 9) %>% 
+tb.foot <- paste0("Standardized factor loadings using DWLS estimator ; CFI = Comparative fit index (scaled); RMSEA = Root mean square error of approximation (scaled)")
+tb.fm02<- kable(tb.fm2,escape = FALSE,align = "lcccc",col.names = tb.col, caption = tb.caption, booktabs = T,linesep = "",format = table_format) %>% 
+          kable_styling(full_width = table_format2,latex_options = c("hold_position"), position = "center",font_size = 9) %>% 
           add_header_above(header = c(" "=1,"Factor loadings"= 4)) %>%
-          column_spec(column = 1,width = "5cm") %>% 
-          column_spec(column = c(2:5),width = "0.5cm") %>% 
           row_spec(row = 8,hline_after = TRUE) %>%
           add_indent(c(9:12)) %>% 
-          footnote(general =tb.foot ,footnote_as_chunk = T);tb.fm01
+          footnote(general =tb.foot ,footnote_as_chunk = T);tb.fm02
 ```

### 2020-12-21 production/prod_analysis-group_invariance.rmd
update- tablas invarianza cross ref

@@
 #------ merge ------#
-tb.fm<- bind_rows(loads01,fm.df)
-tb.fm<- tb.fm %>% 
+tb.fm2<- bind_rows(loads01,fm.df)
+tb.fm2<- tb.fm2 %>% 
   mutate(vars=c(labs,"$\\chi^2\\text{(df)}$","$\\text{CFI}$","$\\text{RMSEA}$","$N$")) %>%
@@
 tb.col  <- c("Variables",1:4)
-tb.foot <- paste0("; p<0.05=$^{*}$; p<0.01=$^{**}$; p<0.001=$^{***}$.
-Standardised factor loadings using DWLS estimator.
-CFI = Comparative fit index (scaled), RMSEA = Root mean square error of approximation (scaled)")
-tb.caption <- "Factor loadings and fit measuresfor the second sample"
+tb.foot <- paste0("Standardised factor loadings using DWLS estimator. CFI = Comparative fit index (scaled), RMSEA = Root mean square error of approximation (scaled)")
+tb.caption <- "Factor loadings and fit measures for the second sample"
 
-tb.fm01<- kable(tb.fm,escape = FALSE,align = "lcccc",col.names = tb.col, caption = tb.caption) %>% 
+tb.fm01<- kable(tb.fm2,escape = FALSE,align = "lcccc",col.names = tb.col, caption = tb.caption) %>% 
           kable_classic(full_width = F,font_size = 9,bootstrap_options = c("striped", "hover"),html_font = "Sans") %>% 
@@
 
-save(tb.fm,tb.col,tb.foot,tb.caption,file = "output/tables/tb.fact-f02.RData")
+save(tb.fm2,tb.col,tb.foot,tb.caption,file = "output/tables/tb.fact-f02.RData")
 ```

### 2021-01-04 docs/paper.Rmd
Update paper.Rmd

--- 2020-12-21 docs/paper.Rmd
@@
   - ../input/bib/meritocracy.bib
+  - ../input/bib/invariance.bib
 editor_options:
@@
 # clean #in the yml
-rmarkdown::render("paper.Rmd", output_format = "bookdown::pdf_document2")
-rmarkdown::render("paper.Rmd", output_format = "bookdown::html_document2")
+rmarkdown::render("docs/paper.Rmd", output_format = "bookdown::pdf_document2")
+rmarkdown::render("docs/paper.Rmd", output_format = "bookdown::html_document2")
 ```
@@
 knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = FALSE,out.width = '85%',fig.pos= "H")
-# knitr::opts_knit$set(base.url = NULL) #relative path for .html output file
-# knitr::opts_knit$set(root.dir = NULL) #relative path for chunks within .rmd files
+# knitr::opts_knit$set(base.url = "../") #relative path for .html output file
+# knitr::opts_knit$set(root.dir = "../") #relative path for chunks within .rmd files
 options(scipen=999)
@@
  "H. Preferences contacts" = pref_netw )
-sum04b <- psych::describe(dat04b[,-c(1)], fast = T) %>% 
-  as.data.frame() %>%
-  select(mean,sd,min,max) %>% 
-  round(digits = 2)
+sum04b <- psych::describe(dat04b[,-c(1)], fast = T) %>% select(mean,sd,min,max) %>% data.frame() %>% round(digits = 2)
 table04b<- sum04b %>% rename("Mean"=mean, "SD"=sd, "Min"=min,"Max"=max)
@@
 
-### Longitudinal measurement invariance
+### Measurement invariance
+
+*Datos*
+
+Para el desarrollo de esta sección, se empleó una fuente de datos externa a los resultados presentados en el apartado anterior.
+Los datos provienen de un estudio realizado durante el primer semestre de 2020, dentro del cual se realizó una medición usando la escala a través de aleatorización completa de los indicadores, lo cual corresponde a la modalidad de aplicación del Grupo 3 presente en la Figura \@ref(fig:appmod).
+Después de la eliminación de casos perdidos por lista, obtuvimos una muestra analítica final de 1242 casos donde 605 (48,71%) son mujeres y 637 (51,28%) son hombres.
+El 48,84% de ellos tienen 45 años o más, y el 33,97% poseen educación terciaria o superior.
+En la Tabla \@ref(tab:fcloads-inv) se pueden observar los resultados del ajuste de la escala en esta muestra.
+Análisis El modelamiento para el testeo invarianza consiste en una serie de modelos anidados a los que progresivamente se le van incorporando restricciones sobre los parámetros del modelo de medición , permitiendo esta progresión localizar la fuente de una violación de invarianza, lo cual ha sido tomando mayor relevancia en los estudios sociológicos a través de encuestas, principalmente por la validez del uso de escalas de medición en estudios de carácter comparado en diversos contextos sociales y culturales [@Davidovetal2014].
+En esta sección emplearemos un análisis de invarianza entre grupos, donde se espera que la estructura factorial sea independiente de la pertenencia a las la muestra correspondiente al estudio original y a los datos del estudio más reciente.
+Para el desarrollo del test de invarianza, los autores sostienen que el modelamiento debe realizarse a través de cuatro niveles o tipos de restricción progresiva [@Milfont2010; @millsap_Statistical_2011; @vandeschoot_checklist_2012]:
+
+1.  **Configural:** Se estima el Modelo solo de indicando la estructura  de este, razón por lo cual todos los parámetros se estiman libremente.
+
+2.  **Débil:** Posteriormente, en el modelo se restringen las cargas factoriales de cada ítem, haciéndolas idénticas en ambas mediciones.
+
+3.  **Fuerte:** Luego, se estima el modelo restringiendo las cargas factoriales y los interceptos de cada ítem, haciéndolos idénticos para cada ambas mediciones.
+
+4.  **Estricta:** Finalmente, se estima un modelo donde se restringen las cargas factoriales, los interceptos y las varianzas de error para cada ítem, haciéndolos idénticos para ambas mediciones.
+
+El primer paso para el testeo de invarianza es la estimación del modelo Configural, el cual se espera que cumpla adecuadamente con los criterios de ajuste global de un modelo de medición, debido a que es el modelo base en proceso de comparación.
+Convencionalmente el estadístico chi-cuadrado es empleado como medida global de ajuste de un modelo de medición.
+Sin embargo, al ser sensible al tamaño de la muestra, se recomienda emplear tres índices de ajuste adicionales: Comparative Fit Index (CFI), el cual debe tener un valor mayor a 0.9; Root Mean Square Error Approximation (RMSEA), el cual debe estar en el rango de valores de 0.05 y 0.08; y el Standardised Root Mean Square Residuals (SRMR) que debe ser menor a 0.08 (@vandeschoot_checklist_2012)
+
+La literatura sobre invarianza métrica sugiere dos aproximaciones complementarias para el testeo de invarianza.
+En nuestro caso, emplearemos la comparación basada en el ajuste incremental de los estadísticos de ajuste [@cheung_Evaluating_2002; @Milfont2010; @dimitrov_Testing_2010] y el test ANOVA para comparación de medias en modelos anidados [@newsom_Longitudinal_2015].
+
+```{r invargroup, eval=TRUE, echo=FALSE}
+load(file = "../output/tables/tb.inv-group.RData")
+col.nam <- c("Model","$\\chi^2 (\\text{df})$","CFI","RMSEA (90 CI)",
+             "$\\Delta \\chi^2 (\\Delta \\text{df}$)","$\\Delta \\text{CFI}$","$\\Delta \\text{RMSEA}$","Decision")
+footnote <- paste0("N = ",nobs(conf),"; Group 1, n = ",conf@Data@nobs[[1]],"; Group 2, n = ",conf@Data@nobs[[2]], ", ***p < 0.001")
+
+kable(tab.inv, col.names = col.nam, align = "l",
+      booktabs=TRUE,format = table_format,escape = FALSE,
+      caption = "Multiple Group meassurement invariance for Perceptions and Preferences for Meritocracy") %>% 
+  kable_styling(full_width = FALSE) %>%
+  footnote(general = footnote, footnote_as_chunk = T)
+```
+
+Como se puede observar en la Table \@ref(tab:invargroup)
 

### 2021-01-06 production/prod_prep-data.rmd

@@
 ```
+
+# Factorial survey data:
+
+Prepare poverty experiment data (wave 01):
+
+```{r}
+w01c <- w01 %>% filter(Intro==1) %>% as.data.frame()
+w01c$ID<- stringr::str_split_fixed(w01c$ticket,"_", 4)[,1]
+w01c$grupo <- NA
+w01c$grupo[w01c$FL_21_DO_merit_perc_pref_julio19v01==1] <- 1
+w01c$grupo[w01c$FL_21_DO_merit_perc_pref_julio19v02==1] <- 2
+w01c$grupo[w01c$FL_21_DO_merit_perc_pref_julio19v03==1] <- 3
+w01c$perc_effort_w1 <- rowSums(w01c[,c(matches(match = "perc_effort",vars = names(w01c)))],na.rm = TRUE)
+w01c$perc_talent_w1 <- rowSums(w01c[,c(matches(match = "perc_talent",vars = names(w01c)))],na.rm = TRUE)
+w01c$perc_wpart_w1  <- rowSums(w01c[,c(matches(match = "perc_wpart" ,vars = names(w01c)))],na.rm = TRUE)
+w01c$perc_netw_w1   <- rowSums(w01c[,c(matches(match = "perc_netw"  ,vars = names(w01c)))],na.rm = TRUE)
+w01c$pref_effort_w1 <- rowSums(w01c[,c(matches(match = "pref_effort",vars = names(w01c)))],na.rm = TRUE)
+w01c$pref_talent_w1 <- rowSums(w01c[,c(matches(match = "pref_talent",vars = names(w01c)))],na.rm = TRUE)
+w01c$pref_wpart_w1  <- rowSums(w01c[,c(matches(match = "pref_wpart" ,vars = names(w01c)))],na.rm = TRUE)
+w01c$pref_netw_w1   <- rowSums(w01c[,c(matches(match = "pref_netw"  ,vars = names(w01c)))],na.rm = TRUE)
+w01c[w01c==0] <- NA
+w01c <- w01c %>% select(grupo,
+                        perc_effort_w1,perc_talent_w1,perc_wpart_w1,perc_netw_w1,
+                        pref_effort_w1,pref_talent_w1,pref_wpart_w1,pref_netw_w1,
+                        sexo_w1=sexo,edad_w1=edad,edcep_w1=edcep, ID)
+```
+
+```{r}
+w01fs <- w01c
+names(w01fs)<- stringr::str_replace_all(string = names(w01fs),pattern = "_w1",replacement = "") 
+```
+
+Prepare factorial survey data (wave 02):
+
+```{r}
+fs01 <- sjlabelled::read_spss("input/data/original/factorial_ola2.sav",verbose = F)
+```
+
+```{r}
+names(long01)
+fs02 <- fs01 %>% 
+  filter(consen==1) %>% 
+  mutate(grupo=4,ID=ResponseId,) %>% 
+  select(grupo,ID,"perc_effort"="merit_perc_effort",
+         "perc_talent" = "merit_perc_talent",
+         "perc_wpart" = "merit_perc_wpart",
+         "perc_netw"  = "merit_perc_netw",
+         "pref_effort" = "merit_pref_effort",
+         "pref_talent" = "merit_pref_talent",
+         "pref_wpart" = "merit_pref_wpart",
+         "pref_netw"  = "merit_pref_netw",
+         sexo,edad,
+         edcep=educ)
+```
+
+Identify dataset of origin:
+
+```{r}
+w01fs$dataset <- "pvw01"
+fs02$dataset  <- "fsw02"
+```
+
+Merge data sets:
+
+```{r}
+pov_fs<-bind_rows(fs02,w01fs) %>%
+  set_label(label = get_label(fs02)) %>%
+  set_labels(labels = get_labels(fs02))
+```
+
+```{r}
+save(pov_fs,file = "input/data/proc/pov01_fact02.RData")
+```

### 2021-01-06 docs/paper.Rmd

@@
       caption = "Multiple Group meassurement invariance for Perceptions and Preferences for Meritocracy") %>% 
-  kable_styling(full_width = FALSE) %>%
+  kable_styling(full_width = FALSE,latex_options = "HOLD_position") %>%
   footnote(general = footnote, footnote_as_chunk = T)
@@
 
+```{r fcloads-inv, echo=FALSE}
+load(file = "../output/tables/tb.fact-f02.RData")
+tb.foot <- paste0("Standardized factor loadings using DWLS estimator ; CFI = Comparative fit index (scaled); RMSEA = Root mean square error of approximation (scaled)")
+tb.fm02<- kable(tb.fm2,escape = FALSE,align = "lcccc",col.names = tb.col, caption = tb.caption, booktabs = T,linesep = "",format = table_format) %>% 
+          kable_styling(full_width = table_format2,latex_options = c("hold_position"), position = "center",font_size = 9) %>% 
+          add_header_above(header = c(" "=1,"Factor loadings"= 4)) %>%
+          row_spec(row = 8,hline_after = TRUE) %>%
+          add_indent(c(9:12)) %>% 
+          footnote(general =tb.foot ,footnote_as_chunk = T);tb.fm02
+```
+
 \pagebreak

### 2021-01-07 docs/paper.Rmd
update invariance and validity section

@@
 
-## Reliability and validity analyses
+## Additional validity analyses
 
 We performed two further analyses in order to add evidence regarding the reliability and validity of the scale.
-Firstly we test the stability of the scale scores along time through longitudinal measurement invariance methods, and secondly we analyze the convergent validity by assessing the association of the merit scale with related concepts and correspondent measurements.
-
-### Measurement invariance
-
-*Datos*
-
-Para el desarrollo de esta sección, se empleó una fuente de datos externa a los resultados presentados en el apartado anterior.
-Los datos provienen de un estudio realizado durante el primer semestre de 2020, dentro del cual se realizó una medición usando la escala a través de aleatorización completa de los indicadores, lo cual corresponde a la modalidad de aplicación del Grupo 3 presente en la Figura \@ref(fig:appmod).
-Después de la eliminación de casos perdidos por lista, obtuvimos una muestra analítica final de 1242 casos donde 605 (48,71%) son mujeres y 637 (51,28%) son hombres.
-El 48,84% de ellos tienen 45 años o más, y el 33,97% poseen educación terciaria o superior.
-En la Tabla \@ref(tab:fcloads-inv) se pueden observar los resultados del ajuste de la escala en esta muestra.
-Análisis El modelamiento para el testeo invarianza consiste en una serie de modelos anidados a los que progresivamente se le van incorporando restricciones sobre los parámetros del modelo de medición , permitiendo esta progresión localizar la fuente de una violación de invarianza, lo cual ha sido tomando mayor relevancia en los estudios sociológicos a través de encuestas, principalmente por la validez del uso de escalas de medición en estudios de carácter comparado en diversos contextos sociales y culturales [@Davidovetal2014].
-En esta sección emplearemos un análisis de invarianza entre grupos, donde se espera que la estructura factorial sea independiente de la pertenencia a las la muestra correspondiente al estudio original y a los datos del estudio más reciente.
-Para el desarrollo del test de invarianza, los autores sostienen que el modelamiento debe realizarse a través de cuatro niveles o tipos de restricción progresiva [@Milfont2010; @millsap_Statistical_2011; @vandeschoot_checklist_2012]:
-
-1.  **Configural:** Se estima el Modelo solo de indicando la estructura  de este, razón por lo cual todos los parámetros se estiman libremente.
-
-2.  **Débil:** Posteriormente, en el modelo se restringen las cargas factoriales de cada ítem, haciéndolas idénticas en ambas mediciones.
-
-3.  **Fuerte:** Luego, se estima el modelo restringiendo las cargas factoriales y los interceptos de cada ítem, haciéndolos idénticos para cada ambas mediciones.
-
-4.  **Estricta:** Finalmente, se estima un modelo donde se restringen las cargas factoriales, los interceptos y las varianzas de error para cada ítem, haciéndolos idénticos para ambas mediciones.
-
-El primer paso para el testeo de invarianza es la estimación del modelo Configural, el cual se espera que cumpla adecuadamente con los criterios de ajuste global de un modelo de medición, debido a que es el modelo base en proceso de comparación.
-Convencionalmente el estadístico chi-cuadrado es empleado como medida global de ajuste de un modelo de medición.
-Sin embargo, al ser sensible al tamaño de la muestra, se recomienda emplear tres índices de ajuste adicionales: Comparative Fit Index (CFI), el cual debe tener un valor mayor a 0.9; Root Mean Square Error Approximation (RMSEA), el cual debe estar en el rango de valores de 0.05 y 0.08; y el Standardised Root Mean Square Residuals (SRMR) que debe ser menor a 0.08 (@vandeschoot_checklist_2012)
-
-La literatura sobre invarianza métrica sugiere dos aproximaciones complementarias para el testeo de invarianza.
-En nuestro caso, emplearemos la comparación basada en el ajuste incremental de los estadísticos de ajuste [@cheung_Evaluating_2002; @Milfont2010; @dimitrov_Testing_2010] y el test ANOVA para comparación de medias en modelos anidados [@newsom_Longitudinal_2015].
-
-```{r invargroup, eval=TRUE, echo=FALSE}
-load(file = "../output/tables/tb.inv-group.RData")
-col.nam <- c("Model","$\\chi^2 (\\text{df})$","CFI","RMSEA (90 CI)",
-             "$\\Delta \\chi^2 (\\Delta \\text{df}$)","$\\Delta \\text{CFI}$","$\\Delta \\text{RMSEA}$","Decision")
-footnote <- paste0("N = ",nobs(conf),"; Group 1, n = ",conf@Data@nobs[[1]],"; Group 2, n = ",conf@Data@nobs[[2]], ", ***p < 0.001")
-
-kable(tab.inv, col.names = col.nam, align = "l",
-      booktabs=TRUE,format = table_format,escape = FALSE,
-      caption = "Multiple Group meassurement invariance for Perceptions and Preferences for Meritocracy") %>% 
-  kable_styling(full_width = FALSE,latex_options = "HOLD_position") %>%
-  footnote(general = footnote, footnote_as_chunk = T)
-```
-
-Como se puede observar en la Table \@ref(tab:invargroup)
+Firstly we analyze the convergent validity by assessing the association of the merit scale with related concepts and correspondent measurements.
+Secondly, we test the internal consistency  of the scale in different studies through measurement invariance methods.
 
@@
 
+### Measurement invariance
+
+El modelamiento para el testeo invarianza ha ido tomando mayor relevancia en los estudios sociológicos de encuestas, principalmente por el interés en la validez de escalas de medición en estudios de carácter comparado en diversos contextos sociales y culturales [@Davidovetal2014].
+En esta sección emplearemos un análisis de invarianza entre grupos, donde se espera que la estructura factorial propuesta sea independiente de la pertenencia a las la muestra correspondiente al estudio original y a los datos del estudio más reciente.
+
+El procedimiento de medición de invarianza consiste en una serie de modelos anidados a los que progresivamente se le van incorporando restricciones sobre los parámetros del modelo de medición.
+La literatura sugiere en general que este modelamiento debe realizarse a través de cuatro niveles o tipos de restricción progresiva [@Milfont2010; @millsap_Statistical_2011; @vandeschoot_checklist_2012]:
+
+1.  **Configural:** Se estima el Modelo solo indicando la estructura  factorial utilizada en el CFA.
+
+2.  **Débil:** se aplica una restricción de igualdad a las cargas factoriales en los distintos grupos, es decir, se fuerza a que las cargas sean idénticas en ambas mediciones.
+
+3.  **Fuerte:** se suman restricciones de igualdad a los interceptos de cada indicador.
+
+4.  **Estricta:** se suman restricciones a las varianzas de error de cada indicador.
+
+*Data*
+
+Para el desarrollo de este estudio se empleó una fuente de datos distinta a la utilizada hasta ahora, consistente en una encuesta online realizada durante el primer semestre de 2020, cuyas características en términos de aplicación y cobertura son equivalentes a la del estudio anterior.
+La muestra final obtenida contiene 1,242 casos, donde 605 (48,71%) son mujeres y 637 (51,28%) son hombres.
+El 48,84% de ellos tienen 45 años o más, y el 33,97% poseen educación terciaria o superior.
+
+En la Tabla \@ref(tab:fcloads-inv) se pueden observar los resultados del ajuste de la escala en esta muestra.
+
+_Instruments_
+
+., dentro del cual se realizó una medición usando la escala a través de aleatorización completa de los indicadores, lo cual corresponde a la modalidad de aplicación del Grupo 3 presente en la Figura \@ref(fig:appmod).
+
+_Results_
+
+El primer paso para el testeo de invarianza es la estimación del modelo Configural, el cual se espera que cumpla adecuadamente con los criterios de ajuste global de un modelo de medición, debido a que es el modelo base en proceso de comparación.
+Convencionalmente el estadístico chi-cuadrado es empleado como medida global de ajuste de un modelo de medición.
+Sin embargo, al ser sensible al tamaño de la muestra, se recomienda emplear tres índices de ajuste adicionales: Comparative Fit Index (CFI), el cual debe tener un valor mayor a 0.9; Root Mean Square Error Approximation (RMSEA), el cual debe estar en el rango de valores de 0.05 y 0.08; y el Standardised Root Mean Square Residuals (SRMR) que debe ser menor a 0.08 (@vandeschoot_checklist_2012)
+
+La literatura sobre invarianza métrica sugiere dos aproximaciones complementarias para el testeo de invarianza.
+En nuestro caso, emplearemos la comparación basada en el ajuste incremental de los estadísticos de ajuste [@cheung_Evaluating_2002; @Milfont2010; @dimitrov_Testing_2010] y el test ANOVA para comparación de medias en modelos anidados [@newsom_Longitudinal_2015].
+
+```{r invargroup, eval=TRUE, echo=FALSE}
+load(file = "../output/tables/tb.inv-group.RData")
+col.nam <- c("Model","$\\chi^2 (\\text{df})$","CFI","RMSEA (90 CI)",
+             "$\\Delta \\chi^2 (\\Delta \\text{df}$)","$\\Delta \\text{CFI}$","$\\Delta \\text{RMSEA}$","Decision")
+footnote <- paste0("N = ",nobs(conf),"; Group 1, n = ",conf@Data@nobs[[1]],"; Group 2, n = ",conf@Data@nobs[[2]], ", ***p < 0.001")
+
+kable(tab.inv, col.names = col.nam, align = "l",
+      booktabs=TRUE,format = table_format,escape = FALSE,
+      caption = "Multiple Group meassurement invariance for Perceptions and Preferences for Meritocracy") %>%
+  kable_styling(full_width = FALSE,latex_options = "HOLD_position") %>%
+  footnote(general = footnote, footnote_as_chunk = T)
+```
+
+Table \@ref(tab:invargroup) shows the results of the measurement invariance tests.
+When attending to the traditional invariance test of $\\Delta \\chi^2 (\\Delta \\text{df})$, the results  support the invariance at the strong level meaning that the fit of the factor model of the PPM-S scale is equivalent across samples when constraining factor loadings and intercepts to being equal.
+Such results is considered in general as evidence of invariance [citar paper Fischer en SJR], as strict forms of measurement invariance rarely hold [citar van der Schoot et al].
+Still, the comparability of latents means requires strict invariance which in this case does not hold using $\\Delta \\chi^2 (\\Delta \\text{df})$.
+Nevertheless the criteria of $\\Delta \\text{CFI}$ used for comparing models is close to the rejection criteria of >.01, whereas the $\\Delta \\text{RMSEA}$ is below the cut-off criteria suggested by this criteria [citar  @chen 2007 ].
+Therefore, the evidence for strict invariance is still not concluding.
+
 # Conclusions
@@
 Our results indicate that perceptions and preferences seem to be two related but different dimensions, often confounded in previous research under the label of "beliefs".
-Additionally, meritocratic and non-meritocratic dimensions do not appear to constitute poles of the same continuum, as some previous studies have assumed.
+Meritocratic and non-meritocratic dimensions do not appear to constitute poles of the same continuum, as some previous studies have assumed.
+Besides, we find evidence of convergent validity as well as measurement invariance that approach well the requirement of comparability between samples.
 

### 2021-01-11 docs/paper.Rmd
corregir mathjax

@@
 Table \@ref(tab:invargroup) shows the results of the measurement invariance tests.
-When attending to the traditional invariance test of $\\Delta \\chi^2 (\\Delta \\text{df})$, the results  support the invariance at the strong level meaning that the fit of the factor model of the PPM-S scale is equivalent across samples when constraining factor loadings and intercepts to being equal.
+When attending to the traditional invariance test of $\Delta \chi^2 (\Delta \text{df})$, the results  support the invariance at the strong level meaning that the fit of the factor model of the PPM-S scale is equivalent across samples when constraining factor loadings and intercepts to being equal.
 Such results is considered in general as evidence of invariance [citar paper Fischer en SJR], as strict forms of measurement invariance rarely hold [citar van der Schoot et al].
-Still, the comparability of latents means requires strict invariance which in this case does not hold using $\\Delta \\chi^2 (\\Delta \\text{df})$.
-Nevertheless the criteria of $\\Delta \\text{CFI}$ used for comparing models is close to the rejection criteria of >.01, whereas the $\\Delta \\text{RMSEA}$ is below the cut-off criteria suggested by this criteria [citar  @chen 2007 ].
+Still, the comparability of latents means requires strict invariance which in this case does not hold using $\Delta \chi^2 (\Delta \text{df})$.
+Nevertheless the criteria of $\Delta \text{CFI}$ used for comparing models is close to the rejection criteria of >.01, whereas the $\Delta \text{RMSEA}$ is below the cut-off criteria suggested by this criteria [citar  @chen 2007 ].
 Therefore, the evidence for strict invariance is still not concluding.

### 2021-01-14 docs/paper.Rmd
Cambios Convergencia.
Incorporo dos parrafos de analisis, uno por indicador.

Agrego las llaves y literatura a la biblioteca.

Agrego una tabla formato rmd

@@
 output:
+  bookdown::html_document2:
+    number_sections: false
   bookdown::pdf_document2:
@@
     toc: false
-  bookdown::html_document2:
-    number_sections: false
 linkcolor: blue
@@
 - Opportunity beliefs: We used two items of the social inequality module of the International Social Survey Programme.
-Studies use generally both indicators to measure meritocratic and non-meritocratic beliefs (McCall et al., 2017; Mijs, 2019).
+Studies use generally both indicators to measure meritocratic and non-meritocratic beliefs [@McCall2017Exposurerisinginequality; @mijs_paradox_2019].
 One item measures the importance of individual factors in determining life outcomes by asking for the importance of hard work for getting ahead in life (M=3.76, SD=0.93).
@@
 
-- Personal wherewithal: the questionnaire included The Neoliberal Beliefs Inventory (NBI), which consider four factors: [completar] (Bay-Cheng et al., 2015).
+- Personal wherewithal: the questionnaire included The Neoliberal Beliefs Inventory (NBI), which consider four factors: Government Interference preferences, Competition preferences, System Inequality perception and Personal Wherewithal [@bay-chengTrackingHomoOeconomicus2015].
 We used the factor personal wherewithal that reflects meritocracy beliefs in terms of the importance of personal attributes as strength and skills to yield success.
@@
 Participants answered the items using a 1 (totally disagree) to 6 (totally agree) scale.
-We simple computed mean scores [reportar alfa the cronbach], with higher scores indicating stronger support for meritocracy (M=3.32, SD=0.89, =0.89).
+We simple computed mean scores [reportar alfa the cronbach], with higher scores indicating stronger support for meritocracy (M=3.32, SD=0.88, $\alpha$ = 0.89).
 
@@
 
-Table XX about here.
+Table \@ref(tab:val-conv) table shows the polyserial and pearson correlations between the PPM-S and other common measures of meritocracy.
+The findings provide evidence for convergent validity for the scale.
+En términos generales como era esperable las preferencias poseen relaciones de menor intensidad con estas escalas que son utilizadas usualmente para medir percepciones.
+A continuación, evaluamos la relación de ambas escalas con las dimensiones de PPM-S
 
-Table X shows ... [completar ]. The findings provide evidence for convergent validity for meritocratic perceptions [re-escribir en el orden de los instrumentos ].
-As expected, we see a substantial positive correlation for the association between meritocratic perception and personal wherewithal.
-Finding for the association of non-meritocratic perception and non-meritocratic item of opportunity beliefs battery is also consistent with our expectation, insofar as the size of the correlation is relevant and the direction of it is positive.
-For the rest of correlations, we note weak correlations but most directions are consistent with our expectations.
-One exception is personal wherewithal, with a positive association for this construct.
-Correlations for meritocratic preference are also weak.
+|                                      |     Origen     |     Hard Work    |     Wherewithal    |
+|--------------------------------------|----------------|------------------|--------------------|
+|     Meritocratic     Perception      |     -0,1975    |     0,1877       |     0,3921         |
+|     Unmeritocratic    Perception     |     0,3547     |     -0,0775      |     -0,1844        |
+|     Meritocratic    Preferences      |     0,1078     |     0,0779       |     0,0819         |
+|     Unmeritocratic    Preferences    |     -0,0397    |     0,0824       |     0,2082         |
+Table: `r as.character(paste("(\\#tab:val-conv)", "Polyserial and Pearson correlation whit other merit scales"))`.
 
-Table XX. Correlations
+Los dos indicadores de la batería “salir a delante” se correlacionan coherentemente, pero con una intensidad de moderada a baja.
+Considerar importante tener padres con recursos para salir a delante posee una asociación positiva de moderada intensidad con la percepción de distribución no meritocrática (r = .354) y una relación negativa de baja intensidad con percepción meritocrática (-0.197), lo cual constituye una evidencia de valides considerando la relación entre los constructos.
+Por su parte, la importancia asignada trabajar duro para salir adelante solo se relaciona débilmente con la percepción de meritocrática (r = .187).
 
-[insertar tabla imagen o markdown]
-
-Coming from a wealthy family
-Hard work
-Personal wherewithal
-Meritocratic
-Perception
--0,1975
-0,1877
-0,3921
-Unmeritocratic
-Perception
-0,3547
--0,0775
--0,1844
-Meritocratic
-Preferences
-0,1078
-0,0779
-0,0819
-Unmeritocratic
-Preferences
--0,0397
-0,0824
-0,2082
+La dimensión de la escala NBI “Creer que los recursos personales producen el éxito” posee una relación positiva de moderada intensidad con la percepción de meritocracia (r = .392), lo cual constituye una evidencia de valides considerando que miden conceptos parcialmente similares.
+La escala de recursos personales también se asocia débil pero positivamente a las preferencias no meritocraticas (r = .208).
+Esta asociación es coherente tomando en cuenta que la creencia en la meritocracia se asocia a la justificación de distintas desigualdades [@mccoy_Priming_2007;@MadeiraPrimesConsequencesSystematic2019] y considerando que dentro de nuestra escala PPM-S la percepción de meritocracia y la preferencia no meritocracia también se relacionan positiva y débilmente.
 

### 2021-01-15 docs/paper.Rmd
Traduccion
Se analiza la validez convergente en orden segun las escalas utilizadas para el contraste.

Se traduce al ingles y mejora la redacción.

@@
 After listwise deletion of missing cases in socio-demographics information, a total of 1422 individuals took part in the third wave: 668 (46.97%) women and 754 (53.03%) men.
-The ages of 50.57% of them are 45 years or older, and 36.28% held a tertiary degree.1
+The ages of 50.57% of them are 45 years or older, and 36.28% held a tertiary degree.
 
@@
 
-The following scales were included for testing their correlation with the the 8-item meritocracy scale:
+The following scales were included for testing their correlations with the the 8-item meritocracy scale:
 
@@
 The second item asks for the importance of coming from a wealthy family and captures the importance of structural factors in getting ahead (M=2.75, SD=1.27).
-Participants responded to each item using a 1 (not at all important) to 5 (essential) response scale.
+Participants responded to each item using a 1 (not important at all) to 5 (essential) response scale.
+We call the first item \textit{hard work} and the second one \textit{social origin}.
 
@@
 Participants answered the items using a 1 (totally disagree) to 6 (totally agree) scale.
-We simple computed mean scores [reportar alfa the cronbach], with higher scores indicating stronger support for meritocracy (M=3.32, SD=0.88, $\alpha$ = 0.89).
+We simple computed mean scores, with higher scores indicating stronger support for meritocracy (M=3.32, SD=0.88, $\alpha$ = 0.89).
 
@@
 
-Table \@ref(tab:val-conv) table shows the polyserial and pearson correlations between the PPM-S and other common measures of meritocracy.
+Table \@ref(tab:val-conv) table shows the polyserial and pearson correlations between the PPM-S and common measures of meritocracy.
 The findings provide evidence for convergent validity for the scale.
-En términos generales como era esperable las preferencias poseen relaciones de menor intensidad con estas escalas que son utilizadas usualmente para medir percepciones.
-A continuación, evaluamos la relación de ambas escalas con las dimensiones de PPM-S
+As we expected, preferences show very weak correlations.
+Findings also indicate that correlations of perceptions with opportunity beliefs are consistent with our expectations.
+The item for social origin shows a positive and moderate association with meritocratic perception (r=.354), and a negative and low correlation with meritocratic perception (-0.197).
+Regarding the item of hard work, the correlation with meritocratic perception is positive but weak (r=.187).
+One explanation for this result is that the support for hard work is very high among respondents---only 8.6 percent indicate \textit{not very important} or \textit{noy importat at all}.
 
-|                                      |     Origen     |     Hard Work    |     Wherewithal    |
+  |                                      | Social Origin  |     Hard Work    |     Wherewithal    |
   |--------------------------------------|----------------|------------------|--------------------|
-|     Meritocratic     Perception      |     -0,1975    |     0,1877       |     0,3921         |
-|     Unmeritocratic    Perception     |     0,3547     |     -0,0775      |     -0,1844        |
-|     Meritocratic    Preferences      |     0,1078     |     0,0779       |     0,0819         |
-|     Unmeritocratic    Preferences    |     -0,0397    |     0,0824       |     0,2082         |
+  |     Meritocratic     Perception      |     -0.1975    |     0.1877       |     0.3921         |
+  |     Unmeritocratic    Perception     |     0.3547     |     -0.0775      |     -0.1844        |
+  |     Meritocratic    Preference       |     0.1078     |     0.0779       |     0.0819         |
+  |     Unmeritocratic    Preference     |     -0.0397    |     0.0824       |     0.2082         |
   Table: `r as.character(paste("(\\#tab:val-conv)", "Polyserial and Pearson correlation whit other merit scales"))`.
 
-Los dos indicadores de la batería “salir a delante” se correlacionan coherentemente, pero con una intensidad de moderada a baja.
-Considerar importante tener padres con recursos para salir a delante posee una asociación positiva de moderada intensidad con la percepción de distribución no meritocrática (r = .354) y una relación negativa de baja intensidad con percepción meritocrática (-0.197), lo cual constituye una evidencia de valides considerando la relación entre los constructos.
-Por su parte, la importancia asignada trabajar duro para salir adelante solo se relaciona débilmente con la percepción de meritocrática (r = .187).
-
-La dimensión de la escala NBI “Creer que los recursos personales producen el éxito” posee una relación positiva de moderada intensidad con la percepción de meritocracia (r = .392), lo cual constituye una evidencia de valides considerando que miden conceptos parcialmente similares.
-La escala de recursos personales también se asocia débil pero positivamente a las preferencias no meritocraticas (r = .208).
-Esta asociación es coherente tomando en cuenta que la creencia en la meritocracia se asocia a la justificación de distintas desigualdades [@mccoy_Priming_2007;@MadeiraPrimesConsequencesSystematic2019] y considerando que dentro de nuestra escala PPM-S la percepción de meritocracia y la preferencia no meritocracia también se relacionan positiva y débilmente.
+Findings the measure of NBI for \textit{personal wherewithal} also provide evidence for the convergent validity of PPM-S.
+Results indicate a positive correlation of this measure with meritocratic preference (r=.392), suggesting that both variables capture similar concepts.
+The NBI's factor depicts a positive association with unmeritocratic preference (r=.208).
+This finding is consistent with literature, inasmuch as the belief for meritocray is associated with the justification of inequality [@mccoy_Priming_2007;@MadeiraPrimesConsequencesSystematic2019].
+Furthermore, results of our confirmatory factor analysis indicate that meritocratic perception and unmeritocratic preferences shows a positive and weak correlation.
 

### 2021-01-18 docs/paper.Rmd

@@
     toc: false
+  bookdown::html_document2:
+    number_sections: false
 linkcolor: blue
@@
 The NBI's factor depicts a positive association with unmeritocratic preference (r=.208).
-This finding is consistent with literature, inasmuch as the belief for meritocray is associated with the justification of inequality [@mccoy_Priming_2007;@MadeiraPrimesConsequencesSystematic2019].
+This finding is consistent with literature, inasmuch as the belief for meritocray is associated with the justification of inequality [@McCall2017Exposurerisinginequality;@MadeiraPrimesConsequencesSystematic2019].
 Furthermore, results of our confirmatory factor analysis indicate that meritocratic perception and unmeritocratic preferences shows a positive and weak correlation.

### 2021-01-19 docs/paper.Rmd
versión para revisión Atria

@@
  "H. Preferences contacts" = pref_netw )
-sum04b <- psych::describe(dat04b[,-c(1)], fast = T) %>% select(mean,sd,min,max) %>% data.frame() %>% round(digits = 2)
+sum04b <- psych::describe(dat04b[,-c(1)], fast = T) %>% data.frame() %>% select(mean,sd,min,max) %>% round(digits = 2)
 table04b<- sum04b %>% rename("Mean"=mean, "SD"=sd, "Min"=min,"Max"=max)
@@
 
-We performed two further analyses in order to add evidence regarding the reliability and validity of the scale.
+We performed two further analyses in order to add evidence regarding the quality of the scale.
 Firstly we analyze the convergent validity by assessing the association of the merit scale with related concepts and correspondent measurements.
-Secondly, we test the internal consistency  of the scale in different studies through measurement invariance methods.
+Secondly, we test the internal consistency of the measurement model behind the scale through measurement invariance procedures.
 
@@
 
-For this analysis we examined information of the third wave of the study, which included additional measures that allow testing the correlations of the merit scale with some related constructs as opportunity beliefs and personal wherewithal.
+For this analysis we examined data from one wave of the same study described above, which was a three-wave panel survey.
+The last wave of the study included additional measures that allow testing the correlations of the merit scale with some related constructs as opportunity beliefs and personal wherewithal.
 After listwise deletion of missing cases in socio-demographics information, a total of 1422 individuals took part in the third wave: 668 (46.97%) women and 754 (53.03%) men.
@@
 
-1.  **Configural:** Se estima el Modelo solo indicando la estructura  factorial utilizada en el CFA.
+1.  **Configural:** se estima el modelo solo indicando la estructura  factorial utilizada en el CFA.
 
@@
 
-., dentro del cual se realizó una medición usando la escala a través de aleatorización completa de los indicadores, lo cual corresponde a la modalidad de aplicación del Grupo 3 presente en la Figura \@ref(fig:appmod).
+La aplicación de la PPM-S se realizó con aleatorización completa de los indicadores, lo cual corresponde a la modalidad de aplicación del Grupo 3 presente en la Figura \@ref(fig:appmod).
 
@@
 Nevertheless the criteria of $\Delta \text{CFI}$ used for comparing models is close to the rejection criteria of >.01, whereas the $\Delta \text{RMSEA}$ is below the cut-off criteria suggested by this criteria [citar  @chen 2007 ].
-Therefore, the evidence for strict invariance is still not concluding.
+Therefore, the tests for strict invariance show mixed evidence.
 

### 2021-01-19 docs/paper.Rmd
... y update lista autores

@@
 link-citations: yes
-author: "Juan Carlos Castillo, Julio Iturra & Francisco Meneses <br><br>"
+author: "Juan Carlos Castillo, Julio Iturra, Luis Maldonado, Francisco Meneses & Jorge Atria <br><br>"
 abstract: "Economic and social inequalities have generated growing concern and crises across contemporary societies.

### 2021-01-25 docs/paper.Rmd
icorpora correcciones Atria & otras

@@
 abstract: "Economic and social inequalities have generated growing concern and crises across contemporary societies.
-One of the mechanisms proposed by the social sciences to explain the maintenance of inequality is the belief in meritocracy, which would legitimize economic disparities based on differences in effort and talent.
-Despite its wide use as a concept, empirical research on meritocracy is relatively novel, and as it is based upon multiple approaches and techniques, its interpretation can be difficult and confusing.
-Adding to these difficulties, is the fact that many such studies have relied upon secondary data to operationalize meritocracy, with significant variation in the use and interpretation of survey items.
-Based on a review of studies that measure meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding the conceptualization and indicators of meritocracy.
-Based on this critical analysis, we propose a scale named the *Perceptions and Preferences for Meritocracy Scale*, which we test with confirmatory factor analysis, using data from an online survey study (N=2,141).
+One of the mechanisms proposed by social sciences to explain the persistence of inequality is the belief in meritocracy, which would legitimize economic disparities based on differences in effort and talent.
+Despite its wide use as a concept, empirical research on meritocracy is relatively novel and characterized by diverse conceptualization and measures that make the findings and their interpretation rather confusing.
+Most of the studies in the area have relied upon secondary data to operationalize meritocracy, with a wide variation in the use and interpretation of the same survey items.
+Taking into account the extant literature that uses measures of meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding conceptualization and operationalization.
+Based on this critical analysis, we propose a survey instrument called *Perceptions and Preferences for Meritocracy Scale*, which we test with confirmatory factor analysis using data from an online survey study (N=2,141).
 Our results support the proposed conceptual structure, which not only distinguishes between perceptions and preferences, but also between meritocratic and non-meritocratic dimensions.
@@
 }
-
 ```
@@
 This has been expressed in a series of protests, as well as diverse critical analyses regarding the development of capitalism and its consequences [@streeckPoliticsPublicDebt2014; @piketty_Capital_2014].
-In this context, the study of visions, preferences and perceptions of inequality has acquired relevance in the social sciences, in topics such as redistributive preferences [@alesina_Fairness_2005; @dimick_Models_2018], the legitimization of economic inequality [@schroder_Income_2017] and the functioning of meritocracy [@duru-bellat_whos_2012; @mijs_paradox_2019; @reynolds_perceptions_2014; @atriaEconomicElitesAttitudes2020].
+In this context, the study of views, preferences and perceptions of inequality has acquired relevance in the social sciences, in topics such as redistributive preferences [@alesina_Fairness_2005; @dimick_Models_2018], the legitimization of economic inequality [@schroder_Income_2017] and the functioning of meritocracy [@duru-bellat_whos_2012; @mijs_paradox_2019; @reynolds_perceptions_2014; @atriaEconomicElitesAttitudes2020].
 
-In general, a meritocracy is defined as a system of distribution of resources and rewards based on individual merit, which in its original conception is a combination of talent and effort [@young_rise_1962].
+In general, meritocracy is defined as a system of distribution of resources and rewards based on individual merit, which in its original conception is a combination of talent and effort [@young_rise_1962].
 This traditional conception of merit, places in a secondary position the possible interference of structural or non-meritocratic factors, such as inheritance, personal contacts, and luck [@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995; @yairMeritocracy2007; @landWeSatTable2006; @youngRiseMeritocracy1994].
@@
 
-Due to the role that meritocratic beliefs play in the justification of individual achievement or failure in contemporary societies [@bay-chengTrackingHomoOeconomicus2015], multiple investigations have evaluated the relationship between meritocratic beliefs and personal/contextual characteristics.
+Due to the role that meritocratic beliefs play in the justification of individual achievement or failure in contemporary societies [@hadjar_meritokratie_2008] [acá también sumar a Markovits(Trap) y Sandel, último libro], multiple investigations have evaluated the relationship between meritocratic beliefs and personal/contextual characteristics.
 For example, some studies have linked meritocracy to the reinforcement of socio-economic, gender, and ethnic stereotypes [@MadeiraPrimesConsequencesSystematic2019; @girerdNeoliberalismIdeologicalBarrier2020; @premingerMeritocracyServiceEthnocracy2020], as well as lines of research have evaluated the effects of meritocratic beliefs in educational [@generettStoriesWeTell2020; @owensENGINESSOCIALMOBILITY2020] and organizational contexts [@perezAdvancingCareersMerit2020; @aielloNewEvidenceFirmuniversity2019].
@@
 
-Based on a critical analysis of different approaches to the measurement of meritocracy to date, this article proposes a tool to both measure and relate two key aspects in the study of meritocracy: perceptions and preferences.
+Based on a critical analysis of different approaches to the measurement of meritocracy to date, this article proposes an instrument to both measure and relate two key aspects in the study of meritocracy: perceptions and preferences.
 Furthermore, as a second axis of analysis, it considers the elaboration of indicators regarding meritocratic and non-meritocratic aspects, demonstrating that they are not two poles of the same continuum, as some previous studies would seem to suggest.
-The measurement system outlined is oriented to generate an instrument as simple and brief as possible, so that it can be used in public opinion surveys, thus integrating meritocratic beliefs in the study of different social phenomena.
+The measurement system outlined is oriented to generate an instrument as simple and brief as possible, being suitable to be used in public opinion surveys and allowing to integrate meritocratic beliefs in the study of different social phenomena.
 
@@
 Several approaches to the empirical study of meritocracy based on public opinion surveys, refer to the concept of _beliefs_, however different meanings and operationalizations usually underlie this concept.
-To illustrate this point, we will take as a main reference a recent paper about meritocracy by @mijs_paradox_2019, which makes several assumptions relating to the measurement of meritocracy that are also found in previous studies.
-The definition of meritocratic beliefs stated by Mijs is the following: "when I discuss meritocracy beliefs, I am referring to citizens’ belief in the importance of hard work relative to structural factors." [@mijs_paradox_2019, pg.9].
+To illustrate this point, we will take as a main reference a recent paper about meritocracy by @mijs_paradox_2019, because of its novelty and its influence in current research about meritocracy, particularly in sociology.
+The definition of meritocratic beliefs stated in this paper is the following: "when I discuss meritocracy beliefs, I am referring to citizens’ belief in the importance of hard work relative to structural factors." [@mijs_paradox_2019, pg.9].
 In the subsequent operationalization, this is associated with the following question and indicator: "how important you think it is for getting ahead in life: (a) hard work", scored from 1 to 5 on a likert scale.
-There assumptions behind this definition are worth discussing in light of the conceptual meaning of meritocracy and its possibilities of operationalization.
+The assumptions behind such definition are worth discussing in light of the conceptual meaning of meritocracy and its possibilities of operationalization.
 
-a. _Unidimensionality_; Is merit only effort?
+a. _Conceptual components_: Is merit only effort?
 
@@
 This battery displays a series of indicators related to what people consider important to get ahead in life: hard work, education, ambition, a wealthy family, the right connections, religion, race, and gender.
-Therefore for Mijs, other aspects that could be associated with talent, such as education, are not meritocratic.
+Therefore, when considering only one of the items, it means that other aspects that could be associated with talent, such as education, would not be deemed as meritocratic.
 As he points out: "Hard work is arguably the most meritocratic part of Michael Young’s equation: ‘Merit = Intelligence + Effort’, for the simple fact that intelligence itself is influenced by a non-meritocratic factor: who your parents happen to be" (p.5).
 
-In this measure of meritocratic belief, we can observe a couple of strong assumptions: effort would not depend on parental influence, and talent (as innate ability) is not meritocratic (contrary to Michael Youngs original conceptualization).
-The question of whether talent is or is not considered meritocratic is certainly an interesting topic to discuss from a philosophical point of view, but for those working on empirical studies it is something that should face empirical scrutiny.
-Is effort the only dimension behind the concept of meritocracy, or is it a multidimensional concept that includes other elements, such as talent (as in its original conceptualization)?
-This conceptual and measurement-based assumption of unidimensionality is found in other studies, that also assume effort to be the main and only aspect of meritocracy [@girerdNeoliberalismIdeologicalBarrier2020a; @bubakPerceptionsMeritocracyNote2019].
-Conversely, recently @mijsHowInformationEconomic2020 also considered talent as an indicator of meritocracy.
+In this measure of meritocratic beliefs, we can observe a couple of strong conceptual assumptions: effort would not depend on parental influence, and talent (as innate ability) is not meritocratic (contrary to Michael Youngs original conceptualization).
+This conceptual and measurement-based assumption is found in other studies, that also assume effort to be the main and only aspect of meritocracy [@girerdNeoliberalismIdeologicalBarrier2020a; @bubakPerceptionsMeritocracyNote2019].
+Is effort the only dimension behind the concept of meritocracy, or includes other components, such as talent (as in its original conceptualization).
+The question of whether talent is or is not considered meritocratic is certainly an interesting topic to discuss from a philosophical point of view, but for those working on empirical studies it should be something that face empirical scrutiny.
+Besides, the consideration of talent as part of meritocracy certainly opens some interesting avenues of research, for instance in studies showing that for the elites meritocracy is more related to talent, whereas effort is something more characteristic of meritocracy of the middle and low classes ([citar Atria ABS]).
+Actually, in a posterior paper by @mijsHowInformationEconomic2020, somewhat contradictory with its previous definition, he considered talent as an indicator of meritocracy.
 
-_b. Beliefs_
+_b. Beliefs?_
 
-The most commonly used survey items for the operationalization of meritocracy is the "reasons to get ahead" battery, which asks "how important you think it is" and then lists several factors, such as effort, education, parental wealth, and contacts.
+The most commonly used items for operationalizing meritocracy in survey research is the "reasons to get ahead" battery, which asks "how important you think it is" and then lists several factors, such as effort, education, parental wealth, and contacts.
 Another version of this same battery used in several surveys - sometimes along the previous one - asks about "how important you think it _should_ be".
 Therefore, the question raised here is: Which one of both is a "belief"?
-What _is_ or what _should_ be?
+What _is_ or what it _should_ be?
 
@@
 what is).
-This occurs not only in mijs_paradox_2019, but also in @reynolds_perceptions_2014, in which the term beliefs is used to talk about perceptions, whereas other authors use general terms such as attitudes (@kunovich_systems_2007).
+This occurs not only in @mijs_paradox_2019, but also in @reynolds_perceptions_2014, in which the term beliefs is used to talk about perceptions, whereas other authors use general terms such as attitudes [@kunovich_systems_2007].
 The first attempt to shed light on this issue was made by @duru-bellat_whos_2012, who used the question "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" as a proxy for "desired" meritocracy (beliefs).
@@
 In other words, normative beliefs should be considered while taking perceptions into account: a strong normative belief in meritocracy may mean something totally different to someone perceiving high meritocracy, when compared to someone perceiving low meritocracy.
-To avoid the confusion generated by the term "belief", we propose the terms meritocratic preferences (what should be), and meritocratic perceptions (what is), as they better reflect the two facets of meritocracy under scrutiny.
+To avoid the confusion generated by the term "belief", we propose the terms meritocratic preferences ("what should be"), and meritocratic perceptions ("what is"), as they better reflect the two facets of meritocracy under scrutiny **[citar Castillo et al 2019]**.
 
@@
 reverse coded).
-The assumption behind this approach is that rejecting a supposed non-meritocratic aspect (as having family and/or children) implies a larger belief in meritocracy.
+The assumption behind this approach is that rejecting a supposed non-meritocratic aspect (as having family and/or children) implies a stronger belief in meritocracy.
 A similar approach of reverse-coding non-meritocratic items was taken by @newman_false_2015, using the same principles applied in the "Preference for the Merit Principle Scale" [@davey_preference_1999].
@@
 Despite making and proving this distinction, however, the authors end up subtracting one dimension from the other, thus coming back to the assumption that they are two poles of the same continuum as @kunovich_systems_2007 did.
-We suggest that non-meritocratic factors should be part of meritocratic measurements, but should be considered independently, rather than adding to or subtracting from meritocratic factors, unless it is empirically proved that they belong to the same conceptual dimension.
+We suggest that non-meritocratic factors should be part of meritocratic measurements, though it should be considered independently rather than adding to or subtracting from meritocratic factors, unless it is empirically proved that they belong to the same conceptual dimension.
 
@@
 
-Finally, most of the studies in meritocracy so far have not accounted for the issue of measurement error [@brownBookReviewTimothy2010; @bollenStructuralEquationsLatent1989] (Ansolabehere, S., Rodden, J., & Snyder Jr, J.
+Finally, most studies in meritocracy so far have not accounted for the issue of measurement error [@brownBookReviewTimothy2010; @bollenStructuralEquationsLatent1989] (Ansolabehere, S., Rodden, J., & Snyder Jr, J.
 M., 2008), when using single indicators and/or simple average indexes for measuring meritocracy.
 Such strategy assumes that the latent construct is measured perfectly by the indicators chosen, going as far as to propose that "...
-In choosing this strategy of index construction, we argue that *support for meritocracy is not a latent variable* [@kunovich_systems_2007 p.653-654].
-Some advances were made by @reynolds_perceptions_2014 through conducting a principal component analysis of meritocratic and non-meritocratic dimensions, but somewhat contradictorily, they chose a sum index, despite proving a multidimensional latent structure.
+In choosing this strategy of index construction, we argue that support for meritocracy is not a latent variable" [@kunovich_systems_2007 p.653-654].
+Some advances were made by @reynolds_perceptions_2014 through conducting a principal component analysis of meritocratic and non-meritocratic dimensions, but somewhat contradictorily, they chose a sum index despite proving a multidimensional latent structure.
 
@@
 ```
-
 The columns "Perceptions" and "Preferences" represent the distinction between these two concepts, usually confused under the label "beliefs".
@@
 The questionnaire was programmed in Qualtrics and the fieldwork was conducted by an external online survey agency ([netquest.cl](www.netquest.cl)) between December 2019 and January 2020.
-The sample was selected from a non-probabilistic quota design in three large cities in Chile.
+The sample was selected from a non-probabilistic quota design in three large cities in Chile (Santiago, Concepción, Antofagasta **[check]**).
 The quotas for gender, age, and education level were generated based on a survey by the Public Studies Center [@cepEncuestaCEPMayo2019], which is a well regarded counterpart agency of the ISSP (International Social Survey Programme) in Chile.
@@
 There were no significant differences between our sample and the wider population for most socio-demographic characteristics, with the exception of educational level (see Table \@ref(tab:rep-samp) in appendix).
-As is often the case with online surveys, there were some limitations in achieving the quotas for lower educational levels.
+As is often the case with online surveys, there were some limitations in achieving the quotas for lower educational levels **[citas]**.
 
@@
 tab01 <- read.csv(file = "../input/data/original/materials/table01.csv",header = 1,sep = ",",encoding = "UTF-8")
-cnames <- c("Dimension", "Factor","Statement (english)" , "Statement (spanish)")
+cnames <- c("Dimensions", "Factor","Statement (english)" , "Statement (spanish)")
 
@@
 Both indicators are consistent with a general adherence to meritocracy as reported in previous studies, privileging individual aspects such as effort (Frei et al 2020, Bucca, 2016).
+**[citar desde Bibtex]**
 
@@
 In Figure \@ref(fig:corpoly), moderate to high intensity relationships are observed between indicators that correspond to the same factor (e.g., perception of meritocracy by effort and by talent, r=0.56).
-The relationships between meritocratic perceptions and preferences are also noteworthy, which is not the case for non-meritocratic indicators.
+The relationships between meritocratic perceptions and preferences are also salient, which is not the case for non-meritocratic indicators.
 
@@
 This is due to the effect produced by the joint appearance of items relating to the same factor in order one, which improved the fit of the model.
-Furthermore, in order two, when asking about perception and preference around the same indicator, cross relations increased and the model fit was worsened.
+Furthermore, in order two, when asking about perception and preference around the same indicator, the correlations between the error terms increased and the model fit worsened.
 Consistently, the random model presented an intermediate case between order one and order two.
@@
 
-- Personal wherewithal: the questionnaire included The Neoliberal Beliefs Inventory (NBI), which consider four factors: Government Interference preferences, Competition preferences, System Inequality perception and Personal Wherewithal [@bay-chengTrackingHomoOeconomicus2015].
+- Personal wherewithal: the questionnaire included The Neoliberal Beliefs Inventory (NBI), which considers four factors: Government Interference preferences, Competition preferences, System Inequality perception and Personal Wherewithal [@bay-chengTrackingHomoOeconomicus2015].
 We used the factor personal wherewithal that reflects meritocracy beliefs in terms of the importance of personal attributes as strength and skills to yield success.
@@
 
-Table \@ref(tab:val-conv) table shows the polyserial and pearson correlations between the PPM-S and common measures of meritocracy.
+Table \@ref(tab:val-conv) shows the polyserial and pearson correlations between the PPM-S and common measures of meritocracy.
 The findings provide evidence for convergent validity for the scale.
@@
 Regarding the item of hard work, the correlation with meritocratic perception is positive but weak (r=.187).
-One explanation for this result is that the support for hard work is very high among respondents---only 8.6 percent indicate \textit{not very important} or \textit{noy importat at all}.
+One explanation for this result is that the support for hard work is very high among respondents---only 8.6 percent indicate \textit{not very important} or \textit{not importat at all}.
 
@@
 
-Findings the measure of NBI for \textit{personal wherewithal} also provide evidence for the convergent validity of PPM-S.
+The measure of NBI for \textit{personal wherewithal} also provides evidence for the convergent validity of PPM-S.
 Results indicate a positive correlation of this measure with meritocratic preference (r=.392), suggesting that both variables capture similar concepts.
@@
 This finding is consistent with literature, inasmuch as the belief for meritocray is associated with the justification of inequality [@McCall2017Exposurerisinginequality;@MadeiraPrimesConsequencesSystematic2019].
-Furthermore, results of our confirmatory factor analysis indicate that meritocratic perception and unmeritocratic preferences shows a positive and weak correlation.
+Furthermore, results of our confirmatory factor analysis indicate that meritocratic perceptions and unmeritocratic preferences show a positive and weak correlation.
 
@@
 El modelamiento para el testeo invarianza ha ido tomando mayor relevancia en los estudios sociológicos de encuestas, principalmente por el interés en la validez de escalas de medición en estudios de carácter comparado en diversos contextos sociales y culturales [@Davidovetal2014].
-En esta sección emplearemos un análisis de invarianza entre grupos, donde se espera que la estructura factorial propuesta sea independiente de la pertenencia a las la muestra correspondiente al estudio original y a los datos del estudio más reciente.
+En esta sección emplearemos un análisis de invarianza entre grupos, donde se espera que la estructura factorial propuesta sea independiente de la pertenencia a la muestra correspondiente al estudio original, comparándola con otro estudio disponible más reciente y donde también se incorporó la escala de meritocracia.
 
@@
 
-Para el desarrollo de este estudio se empleó una fuente de datos distinta a la utilizada hasta ahora, consistente en una encuesta online realizada durante el primer semestre de 2020, cuyas características en términos de aplicación y cobertura son equivalentes a la del estudio anterior.
+Para el desarrollo de este estudio se empleó una fuente de datos distinta a la utilizada hasta ahora, consistente en una encuesta online realizada durante el primer semestre de 2020, cuyas características en términos de aplicación y cobertura son equivalentes a las del estudio anterior.
 La muestra final obtenida contiene 1,242 casos, donde 605 (48,71%) son mujeres y 637 (51,28%) son hombres.
-El 48,84% de ellos tienen 45 años o más, y el 33,97% poseen educación terciaria o superior.
+El 48,84% de ellos tiene 45 años o más, y el 33,97% posee educación terciaria o superior.
 
@@
 
-El primer paso para el testeo de invarianza es la estimación del modelo Configural, el cual se espera que cumpla adecuadamente con los criterios de ajuste global de un modelo de medición, debido a que es el modelo base en proceso de comparación.
+El primer paso para el testeo de invarianza es la estimación del modelo Configural, el cual se espera que cumpla adecuadamente con los criterios de ajuste global de un modelo de medición, debido a que es el modelo base en el proceso de comparación.
 Convencionalmente el estadístico chi-cuadrado es empleado como medida global de ajuste de un modelo de medición.
-Sin embargo, al ser sensible al tamaño de la muestra, se recomienda emplear tres índices de ajuste adicionales: Comparative Fit Index (CFI), el cual debe tener un valor mayor a 0.9; Root Mean Square Error Approximation (RMSEA), el cual debe estar en el rango de valores de 0.05 y 0.08; y el Standardised Root Mean Square Residuals (SRMR) que debe ser menor a 0.08 (@vandeschoot_checklist_2012)
+Sin embargo, al ser sensible al tamaño de la muestra, se recomienda emplear tres índices de ajuste adicionales: Comparative Fit Index (CFI), el cual debe tener un valor mayor a 0.9; Root Mean Square Error Approximation (RMSEA), el cual debe estar en el rango de valores de 0.05 y 0.08; y el Standardised Root Mean Square Residuals (SRMR) que debe ser menor a 0.08 [@vandeschoot_checklist_2012].
 
@@
 When attending to the traditional invariance test of $\Delta \chi^2 (\Delta \text{df})$, the results  support the invariance at the strong level meaning that the fit of the factor model of the PPM-S scale is equivalent across samples when constraining factor loadings and intercepts to being equal.
-Such results is considered in general as evidence of invariance [citar paper Fischer en SJR], as strict forms of measurement invariance rarely hold [citar van der Schoot et al].
+Such result is considered in general as evidence of invariance [citar paper Fischer en SJR], as strict forms of measurement invariance rarely hold [citar van der Schoot et al].
 Still, the comparability of latents means requires strict invariance which in this case does not hold using $\Delta \chi^2 (\Delta \text{df})$.
@@
 
-Studies that attempt to characterize and compare societies by their support for meritocratic beliefs have used differencing approaches.
+Studies that attempt to characterize and compare societies by their support for meritocratic beliefs have used different approaches.
 As most studies use secondary survey data, they tend to assume that the available indicators represent an underlying construct.
@@
 Additionally, given that non-meritocratic factors are not necessarily related to meritocratic ones, our framework makes it possible to assess the perceived legitimacy of practices such as the use of personal contacts and their interference (or not), within meritocratic ideals in different societies.
-The consideration of structural factors, both individual and societal, will allow us to advance future hypotheses relating to meritocratic legitimacy to a more comprehensive level than the one currently used, which is based solely on meritocratic perceptions.
+The consideration of structural factors, both individual and societal, will allow us to advance future hypotheses relating to meritocratic legitimacy to a more comprehensive level than the curren one, which is based solely on meritocratic perceptions.
 Furthermore, the impacts of different configurations of the four-dimensional framework on practices and behaviors such as corruption, civic involvement, and political alignment, is an area that requires additional research.

### 2021-01-25 docs/prod_analysis-cfa.rmd
icorpora correcciones Atria & otras

@@
   add_header_above(c(" " = 9,"Factor loadings"=8)) %>%
-  column_spec(column = c(1),width ="7in",include_thead = T) %>%
-  column_spec(column = c(2:17),width ="0.5in",include_thead = T) %>% 
+  # # column_spec(column = c(1),width ="7in",include_thead = T) %>%
+  # # column_spec(column = c(2:17),width ="0.5in",include_thead = T) %>% 
   kable_styling(full_width = FALSE)
@@
   add_header_above(c(" " = 9,"Factor loadings"=8)) %>% 
-  column_spec(column = c(1),width ="7in",include_thead = T) %>%
-  column_spec(column = c(2:17),width ="0.5in",include_thead = T) %>% 
+  # column_spec(column = c(1),width ="7in",include_thead = T) %>%
+  # column_spec(column = c(2:17),width ="0.5in",include_thead = T) %>% 
   kable_styling(full_width = FALSE) %>% 
@@
   add_header_above(c(" " = 9,"Factor loadings"=8)) %>% 
-  column_spec(column = c(1),width ="7in",include_thead = T) %>%
-  column_spec(column = c(2:17),width ="0.5in",include_thead = T) %>% 
+  # column_spec(column = c(1),width ="7in",include_thead = T) %>%
+  # column_spec(column = c(2:17),width ="0.5in",include_thead = T) %>% 
   kable_styling(full_width = F)
@@
 
-```{r meas03-0}
+```{r meas03-0, eval=FALSE, include=TRUE}
 nodeNames <-c("Perception effort","Perception talent","Perception rich family","Perception contacts",
@@
 
-graphics.off()
+# graphics.off()
 fm03<- data.frame(v1=fitmeasures(fit4_o, output ="matrix")[c("chisq.scaled","df","cfi.scaled","rmsea.scaled"),])
@@
   sizeLat = 5.5, # largo de los circulos
-  sizeLat2 = 25, # alto de los circulos
+  sizeLat2 = 20, # alto de los circulos
   residScale = 10,
@@
 sumdat04<- kable(table04b, escape = FALSE,align = "l",linesep = "", caption = cap01) %>% 
-           column_spec(column = 1,width = "10cm",width_min = "12cm",include_thead = TRUE) %>% 
+           # column_spec(column = 1,width = "10cm",width_min = "12cm",include_thead = TRUE) %>% 
            footnote(general = paste0(paste0("$N$= ","2141"),"; p<0.05=$*$; p<0.01=$**$; p<0.001=$***$"),footnote_as_chunk = T) %>% 
@@
           kable_styling(full_width = F,font_size = 9) %>% 
-          column_spec(column = 1,width = "10cm",width_min = "12cm",include_thead = TRUE) %>% 
-          column_spec(column = 2:13,width = "1cm",width_min = "1.3cm",include_thead = TRUE) %>% 
+          # column_spec(column = 1,width = "10cm",width_min = "12cm",include_thead = TRUE) %>% 
+          # column_spec(column = 2:13,width = "1cm",width_min = "1.3cm",include_thead = TRUE) %>% 
           add_header_above(header = c(" "=1,"Version 1"= 4,"Version 2"= 4,"Version 3"= 4)) %>% 

### 2021-01-29 docs/prod_analysis-group_invariance.rmd

--- 2020-12-21 production/prod_analysis-group_invariance.rmd
@@
 ```{r tb_loads-fit-fac02}
-fm01<- data.frame(t(data.frame(fitmeasures(fs1_o, output ="matrix")[c("chisq.scaled","df","cfi.scaled","rmsea.scaled"),]))); row.names(fm01) ="M1.1"
+fm01<- data.frame(t(data.frame(fitmeasures(fs1_o, output ="matrix")[c("chisq","df","cfi","tli","rmsea"),]))); row.names(fm01) ="M1.1"
 
 #------chi2, df------#
-fm04.1 <- fm01 %>% select(chisq.scaled,df) 
-fm04.1$chisq.scaled <- round(x = fm04.1$chisq.scaled,digits = 1)
+fm04.1 <- fm01 %>% select(chisq,df) 
+fm04.1$chisq <- round(x = fm04.1$chisq,digits = 1)
 fm04.1$df <- round(x = fm04.1$df,digits = 0)
-fm04.1$chi2df <- paste0(fm04.1$chisq.scaled,"(",fm04.1$df,")")
+fm04.1$chi2df <- paste0(fm04.1$chisq,"(",fm04.1$df,")")
 fm04.1 <- select(fm04.1,"chi2df")
@@
 #------CFI, RMSEA------#
-fm04.2 <- fm01 %>% select(cfi.scaled,rmsea.scaled)
+fm04.2 <- fm01 %>% select(cfi,tli,rmsea)
 for (i in names(fm04.2)) {
@@
 tb.fm2<- tb.fm2 %>% 
-  mutate(vars=c(labs,"$\\chi^2\\text{(df)}$","$\\text{CFI}$","$\\text{RMSEA}$","$N$")) %>%
+  mutate(vars=c(labs,"$\\chi^2\\text{(df)}$","$\\text{CFI}$","$\\text{TLI}$","$\\text{RMSEA}$","$N$")) %>%
   select(vars,everything())

### 2021-01-29 production/prod_analysis-cfa.rmd

@@
 corrplot(cor01$rho,method = "color" ,type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
-
 ```
@@
 colnames(cor04$rho) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)", "(H)")
+cor04$rho[cor04$rho==1] <- NA
 save(cor04, file = "output/images/correlation.RData")
@@
   family = "A",
+  na.label = "-",
   number.font = 6,
@@
   family = "A", 
+  na.label = "-",
   number.font = 6,
@@
 
+```{r meas03-0, eval=FALSE, include=TRUE}
+nodeNames <-c("Perception effort","Perception talent","Perception rich family","Perception contacts",
+              "Preference effort","Preference talent","Preference rich family","Preference contacts",
+              "Perception \n meritocratic","Perception \n non meritocratic","Preference \n meritocratic","Preference \n non meritocratic")
+
+# graphics.off()
+fm03<- data.frame(v1=fitmeasures(fit4_o, output ="matrix")[c("chisq.scaled","df","cfi.scaled","rmsea.scaled"),])
+fm03 <- round(fm03,3)
+par(mai = c(2,2,2,2)) # Set the margin on all sides to 2
+par(mar = c(5, 5, 5, 5)) # Set the margin on all sides to 6
+layout(matrix(c(1, # semPlot
+                1, # semPlot
+                1,
+                1,
+                1,
+                2),# ajuste
+              nrow=6,
+              byrow=TRUE))
+
+layout.show(n = 2)
+semPaths(
+  fit4_o ,
+  # whatLabels = "std", # Indica si aparecen los valores
+  what = "mod",
+  label.cex = 1.3, # Tamaño de las etiquetas dentro de los nodos.
+  edge.label.cex =  0.8, #  0.8 Tamaño de los valores estimados.
+  residuals = T,
+  optimizeLatRes =F,
+  edge.color = "black",
+  style = "lisrel",
+  nCharNodes = 0,
+  curvePivot = F,
+  curve = 2.5,
+  rotation = 4,
+  layout = "tree2",
+  cardinal = "lat cov",
+  legend.cex = 0.6, # 0.6
+  label.cex = 1,
+  label.font = 6,
+  edge.label.font = 14,
+  asize = 3,
+  edge.width = 1.2,
+  sizeMan = 8,# largo manifest
+  sizeMan2= 8,# alto manifest
+  sizeLat = 5.5, # largo de los circulos
+  sizeLat2 = 20, # alto de los circulos
+  residScale = 10,
+  width = 21.5, # 17
+  height = 10, # 10
+  # nodeNames = nodeNames,
+  intercepts = F,
+  reorder = T,
+  thresholds = F,
+  fixedStyle =1,
+  node.height = 1,
+  node.width = 4,
+  label.scale = F,
+  shapeMan = "rectangle",
+  shapeLat = "ellipse",
+  nodeLabels = nodeNames,
+  details = T)
+
+ld<- standardizedsolution(fit4_o) %>% select(lhs,op,rhs,est.std) %>% filter(op=="=~") 
+ld$est.std<- sprintf("%.2f", ld$est.std)
+
+#pegar cargas factoriales al plot
+text(x = -0.1,y =  1.00, ld$est.std[1],font = 5,cex = 1.20)
+text(x = -0.1,y =  0.85, ld$est.std[2],font = 5,cex = 1.20)
+text(x = -0.1,y =  0.43, ld$est.std[3],font = 5,cex = 1.20)
+text(x = -0.1,y =  0.26, ld$est.std[4],font = 5,cex = 1.20)
+text(x = -0.1,y = -0.15, ld$est.std[5],font = 5,cex = 1.20)
+text(x = -0.1,y = -0.30, ld$est.std[6],font = 5,cex = 1.20)
+text(x = -0.1,y = -0.73, ld$est.std[7],font = 5,cex = 1.20)
+text(x = -0.1,y = -0.88, ld$est.std[8],font = 5,cex = 1.20)
+
+library(draw) #para hacer rectangulo de fit
+
+adj_y <- 0.6
+adj_x <- 0.2
+drawBox(x = 3.5, y = 1.2-adj_y, width = 7.5, height = 2,fillColor = "grey",opacity = 0)# marca espacio
+drawBox(x = 3.5, y = 1.4-adj_y, width = 5.5, height = 1,fillColor = "grey",opacity = 0.5) # crea rectangulo
+
+# paste0("Estimator DWLS", ", N=",nobs(fit4_o))
+# paste0("Model fit: ", "χ²","(",fm03[2,],")=",fm03[1,],"***","; CFI=",fm03[3,],"; RMSEA=",fm03[4,])
+# paste0("***p<0.001")
+
+#pegar texto fit en rectangulo
+
+size <- 10
+
+drawText(x = 1.4+adj_x, y = 1.75-adj_y, text = paste0("Estimator: DWLS", ", N=",nobs(fit4_o)),family = "serif",size = size)
+drawText(x = 2.3+adj_x, y = 1.59-adj_y, text = paste0("Model fit: ", "χ²","(",fm03[2,],")=",fm03[1,],"***","; CFI=",fm03[3,],"; RMSEA=",fm03[4,]),family = "serif",size = size)
+drawText(x = 1.0+adj_x, y = 1.43-adj_y, text = paste0("***p<0.001"),family = "serif",size = size)
+
+#guardar en png
+drawExport("output/images/meas03.png",units = "cm",width = 18,height = 15,ppi = 300) 
+```
+
+```{r lat-corr}
+latcor<- lavInspect(fit4_o,what = "cor.lv")
+windowsFonts(A = windowsFont("Times New Roman"))
+rownames(latcor) <-c(
+  "A. Meritocratic Perception",
+  "B. Non-Meritocratic Perception",
+  "C. Meritocratic Preferences",
+  "D. Non-Meritocratic Preferences"
+)
+colnames(latcor) <-c("(A)", "(B)","(C)","(D)")
+latcor[latcor==1] <- NA
+png("output/images/latcor.png",
+    width = 18,height = 13,units="cm",
+    pointsize=12,bg="white",res=300)
+corrplot(
+  latcor,
+  method = "color",
+  type = "upper",
+  tl.col = "black",
+  addCoef.col = "black",
+  diag = TRUE,
+  family = "A",
+  na.label = "-",
+  number.font = 6,
+  tl.cex =0.75,
+  number.cex = 1)
+dev.off()
+```
+
 ```{r tabla04, include=FALSE}
@@
 #-------#
-fm01<- data.frame(t(data.frame(fitmeasures(fit1_o, output ="matrix")[c("chisq.scaled","df","cfi.scaled","rmsea.scaled"),]))); row.names(fm01) ="M1.1"
-fm02<- data.frame(t(data.frame(fitmeasures(fit2_o, output ="matrix")[c("chisq.scaled","df","cfi.scaled","rmsea.scaled"),]))); row.names(fm02) ="M2.1"
-fm03<- data.frame(t(data.frame(fitmeasures(fit3_o, output ="matrix")[c("chisq.scaled","df","cfi.scaled","rmsea.scaled"),]))); row.names(fm03) ="M3.1"
+fm01<- data.frame(t(data.frame(fitmeasures(fit1_o, output ="matrix")[c("chisq","df","cfi","tli","rmsea"),]))); row.names(fm01) ="M1.1"
+fm02<- data.frame(t(data.frame(fitmeasures(fit2_o, output ="matrix")[c("chisq","df","cfi","tli","rmsea"),]))); row.names(fm02) ="M2.1"
+fm03<- data.frame(t(data.frame(fitmeasures(fit3_o, output ="matrix")[c("chisq","df","cfi","tli","rmsea"),]))); row.names(fm03) ="M3.1"
 
@@
 fm04<- round(rbind(fm01,fm02,fm03),3)
-fm04.1 <- fm04 %>% select(chisq.scaled,df) 
-fm04.1$chisq.scaled <- round(x = fm04.1$chisq.scaled,digits = 1)
+fm04.1 <- fm04 %>% select(chisq,df) 
+fm04.1$chisq <- round(x = fm04.1$chisq,digits = 1)
 fm04.1$df <- round(x = fm04.1$df,digits = 0)
-fm04.1$chi2df <- paste0(fm04.1$chisq.scaled,"(",fm04.1$df,")")
+fm04.1$chi2df <- paste0(fm04.1$chisq,"(",fm04.1$df,")")
 fm04.1 <- select(fm04.1,"chi2df")
@@
 #------CFI, RMSEA------#
-fm04.2 <- fm04 %>% select(cfi.scaled,rmsea.scaled) 
+fm04.2 <- fm04 %>% select(cfi,tli,rmsea) 
 for (i in names(fm04.2)) {
@@
 tb.fm<- bind_rows(tb.load,fm.df)
-tb.fm<- tb.fm %>% mutate(vars=c(labs,"$\\chi^2\\text{(df)}$","$\\text{CFI}$","$\\text{RMSEA}$","$N$")) %>% select(vars,everything())
+tb.fm<- tb.fm %>% mutate(vars=c(labs,"$\\chi^2\\text{(df)}$","$\\text{CFI}$","$\\text{TLI}$","$\\text{RMSEA}$","$N$")) %>% select(vars,everything())
 #--table---#

### 2021-01-29 docs/paper.Rmd
actualizacion

@@
 Despite making and proving this distinction, however, the authors end up subtracting one dimension from the other, thus coming back to the assumption that they are two poles of the same continuum as @kunovich_systems_2007 did.
+Similarly, @roex_attitudes_2019 used ISSP indicators for perceived meritocracy and non-meritocracy to build a single score by reverse coding the non-meritocratic items.
 We suggest that non-meritocratic factors should be part of meritocratic measurements, though it should be considered independently rather than adding to or subtracting from meritocratic factors, unless it is empirically proved that they belong to the same conceptual dimension.
@@
 ```
-The columns "Perceptions" and "Preferences" represent the distinction between these two concepts, usually confused under the label "beliefs".
+The columns "Perceptions" and "Preferences" represent the distinction between these two concepts, usually confused under the label "beliefs" [@castillo_Meritocracia_2019].
 Perceptions refer to the extent to which people observe that meritocracy works or applies in their society, which relates to items such as "I think hard work is important to get ahead in society".
@@
 
-With the objective of evaluating the effect of indicator ordering, the respondents (_n = 2141_) were randomly divided into three different set-order versions, as explained in Figure \@ref(fig:appmod).
-The scale was presented to the first group (_n = 712_) in the order that appears in Table 2.
-For the second group (_n = 717_), the order of the items was reorganized according to the topics of the items, e.g.
+With the objective of evaluating the effect of indicator ordering in the responses, three different versions of items' order were designed and randomly assigned, as depicted in Figure \@ref(fig:appmod).
+The scale was presented to the first group (_n = 712_) in the order that appears in Table \@ref(tab:table-indicadores) according to perceptions and preferences.
+For the second group (_n = 717_), the order  was reorganized according to perceptions and preferences over the same topic, e.g.
 for the topic of hard work, the item about perception was followed by the item about preference, and the same for the rest of the topics.
@@
 ```
-
 CFA was conducted using the `lavaan` R package (version 0.6-3; Rosseel, 2020), with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement (Kline, 2016; Rosseel, 2020).
@@
 Averages are observed from 2.41, corresponding to preference-contacts; to 3.89, corresponding to preference-effort.
-Both indicators are consistent with a general adherence to meritocracy as reported in previous studies, privileging individual aspects such as effort (Frei et al 2020, Bucca, 2016).
-**[citar desde Bibtex]**
+Both indicators are consistent with a general adherence to meritocracy as reported in previous studies, privileging individual aspects such as effort [@bucca_merit_2016; @frei_fruto_2020].
 
@@
 ```
-
 In sum, descriptive analyses point to a relative adherence to meritocracy, which is expressed as a greater preference for meritocratic criteria, and a lesser preference for non-meritocratic criteria, even when the preference for non-meritocratic elements reaches a proportion of more than 20%.
@@
 In particular, the associations between effort and talent are relevant, challenging previous assumptions that talent would not be a meritocratic criterion [@mijs_paradox_2019], otherwise the correlation would be zero or negative.
-In addition, we see that there is no negative correlation between meritocratic and non-meritocratic aspects, dismissing the assumptions of previous studies that these dimensions would be the opposite poles of the same continuum [@reynolds_perceptions_2014].
+In addition, we observe that there is no negative correlation between meritocratic and non-meritocratic aspects, undermining the assumptions of previous studies that suggested that these dimensions would be the opposite poles of one same continuum [@reynolds_perceptions_2014].
 
@@
 
-### Full sample CFA
-
 Figure \@ref(fig:meas02) shows the results of the estimation for the four-factor model using the complete sample.
 This model shows adequate fit indicators (CFI = 0.977, RMSEA = 0.079, $\chi2$(df = 14) = 28,03, p = 0.014) and it presents factorial loads with weights that vary between 0.63 and 0.88.
-Considering the correlations between the latent variables, it is observed that the variables with the strongest correlations are meritocratic perceptions and preferences, which have a positive correlation of 0.59, while non-meritocratic perceptions and preferences do not show considerable correlation.
 
@@
 ```
-
-As can be seen in the Table \@ref(tab:tab-loads), all models, regardless of the group and order of items, obtained an adequate fit, with CFI's above 0.95 and RMSEA's below 0.08.
-However, no model achieved a non-significant chi-square (something to be expected in large samples), although both the randomized and the first model obtained an adequate chi-square ratio of less than 3.
-The first order was the one that obtained the best fit (CFI=0.998, TLI=0.995, RMSEA=0.037, $\chi2$(df=14)=28.03, p = 0.014), followed by the random order of items (CFI=0.992, TLI=0.984,RMSEA=0.051, $\chi2$(df=14)=39.09, p < 0.001).
-On the other hand, the scale ordered by themes seems to generate a framing effect, in which the relationship between perceptions and preferences seemed to be overestimated, thus affecting the adjustment (CFI=0.984, TLI=0.968, RMSEA=.071, $\chi2$(df=14)=64.156, p < 0.001).
+Attending to the correlations between the latent variables as depicted in Figure \@ref(fig:latcor), it is observed that the variables with the strongest correlations are meritocratic perceptions and preferences, whit an $r=0.59$, while non-meritocratic perceptions and preferences show a correlation close to 0.
+Besides, and as commented before in the analysis of the single items correlations, the latent variables of meritocratic and non-meritocratic versions of both perceptions and preferences show very low correlations.
+This last finding gives further evidence regarding the lack of unidimensionality of meritocratic and non-meritocratic aspects: they do not appear as the opposite poles of the same continuum and based on this evidence the use of reverse-coding for non-meritocratic items as indicators of meritocracy should be avoided.
 
@@
 ```
+In a second step we test the effect of the items' order in the confirmatory model.
+Table \@ref(tab:tab03) shows the fit indicators of the models estimated for each of the three versions of the items' order described in the methodology, being the first one according to perceptions and preferences, the second one according to meritocratic and non-meritocratic aspects, and the third one completely randomized.
+Regardless of the group and order of items, all models obtained an adquate fit, with CFI's above 0.95 and RMSEA's below 0.08.
+However, none of the models achieved a non-significant chi-square, something expected in large samples as the one used here).
+Model 1 shows the fit indicators for the estimation of the CFA model to the first version order (perceptions-preferences), which was the one obtaining best fit (CFI=0.993, TLI=0.995, RMSEA=0.034, $\chi2$(df=14)=42.276), whereas Model 2 for the fixed order according to merit/non-merit items shows the comparatively worst indicators.
+The CFA fit indices for the completely randomized items' order (Model 3) is somehow an intermediate fit indexes when compared with the previous two.
+Still, it keeps all the indicators within the acceptable cut-off criteria and besides it controls for possible order effects in the administration of the instrument.
 
@@
 load(file = "../output/tables/fit1_o.RData")
-
-sum_fit<- bind_rows(fitmeasures(fit4_o)[c("chisq.scaled","df","cfi.scaled","rmsea.scaled")],
-                    fitmeasures(fit1_o)[c("chisq.scaled","df","cfi.scaled","rmsea.scaled")],
-                    fitmeasures(fit2_o)[c("chisq.scaled","df","cfi.scaled","rmsea.scaled")],
-                    fitmeasures(fit3_o)[c("chisq.scaled","df","cfi.scaled","rmsea.scaled")])
-sum_fit$mod <- c("Model 1","Model 2","Model 3","Model 4")
-sum_fit$est <- c("DWLS","DWLS","DWLS","DWLS")
-sum_fit$nobs <- c(nobs(fit4_o),nobs(fit1_o),nobs(fit2_o),nobs(fit3_o))
+sum_fit<- bind_rows(fitmeasures(fit1_o)[c("chisq","df","cfi","tli","rmsea")],
+                    fitmeasures(fit2_o)[c("chisq","df","cfi","tli","rmsea")],
+                    fitmeasures(fit3_o)[c("chisq","df","cfi","tli","rmsea")])
+sum_fit$mod <- c("Model 1","Model 2","Model 3")
+sum_fit$est <- c("DWLS","DWLS","DWLS")
+sum_fit$nobs <- c(nobs(fit1_o),nobs(fit2_o),nobs(fit3_o))
 sum_fit <- select(sum_fit,mod,nobs,est,everything())
-colnames <- c("Model","$N$","Estimator","$\\chi^2$","df","CFI","RMSEA")
-
+colnames <- c("Model","$N$","Estimator","$\\chi^2$","df","CFI","TLI","RMSEA")
 sumtable02<- kable(sum_fit,digits = 3,format = table_format,row.names = F,booktabs=T, align = c("lcccccc"),
@@
                 font_size = 10)  %>%
-      footnote(number = c("Model 1: Complete sample",
-                          "Model 2: fixed order by perception/preference",
-                          "Model 3:  fixed order by topic (i.e: effort)",
-                          "Model 4: Randomized order"));sumtable02
-
+      footnote(number = c("Model 1: fixed order by perception/preference",
+                          "Model 2: fixed order by merit-non merit",
+                          "Model 3: Randomized order"));sumtable02
 ```
-
-Although all tests obtained relatively adequate indicators, the aforementioned differences between the adjustments of the models according to order are statistically significant, as evaluated by an invariance analysis.
-It was concluded that among the three orders there was no configurational invariance, that is, they did not have the same dimensionality, and therefore did not equally fit the theoretical model [@vandenbergReviewSynthesisMeasurement2000].
-This is due to the effect produced by the joint appearance of items relating to the same factor in order one, which improved the fit of the model.
-Furthermore, in order two, when asking about perception and preference around the same indicator, the correlations between the error terms increased and the model fit worsened.
-Consistently, the random model presented an intermediate case between order one and order two.
-
-\pagebreak
-
-|           | Contrast Model 1 Factor | theoretical model 4 Factors | Model with M.I. |
-|:---------:|:-----------------------:|:---------------------------:|:---------------:|
-| **n**     |           1769          |             1769            |       1769      |
-| **CFI**   |          0.595          |            0.988            |      0.994      |
-| **TLI**   |          0.433          |            0.976            |      0.985      |
-| **RMSEA** |          0.226          |            0.047            |      0.036      |
-| **$\chi2$**    |         1830.839        |            68.661           |      40.250     |
-| **p**     |           .000          |             .000            |       .000      |
-| **$\chi2$/df** |          65.38          |             4.90            |       3.32      |
-Table: `r as.character(paste("(\\#tab:items-dim)", "Items according to dimension."))`.
-
-As shown in Table \@ref(tab:items-dim), the proposed theoretical four-factor model fits better than the one-factor contrast model.
-The theoretical model adjusted relatively well, as it shows preferable indicators of IFC=0.987, TLI=0.975 and RMSEA=0.041, although it also has poor indicators of $\chi2$(df=14)=67.6 and p-value<0.001.
-To evaluate possible improvements in the scale, the relationships proposed by the modification indexes were analyzed.
-These indicate the existence of two unspecified cross-loads.
-When a model was generated following these recommendations, there was a considerable improvement in the model, although the new model still did not obtain a $\chi2$ ratio of less than 3, and had very low factorial loads ($\tau$)< 0.15.
-Therefore, following the recommendation of Brown (2006) to only accept the proposals of modification when theory and solid evidence are available, we decided not to incorporate these parameters into the model.
 
@@
 We performed two further analyses in order to add evidence regarding the quality of the scale.
-Firstly we analyze the convergent validity by assessing the association of the merit scale with related concepts and correspondent measurements.
-Secondly, we test the internal consistency of the measurement model behind the scale through measurement invariance procedures.
+Firstly we analyzed the convergent validity by exploring the association of the merit scale with related concepts and itscorrespondent measurements.
+Secondly, we test the internal consistency of the measurement model behind the scale comparing with a different sample through measurement invariance procedures .
 
@@
 - Personal wherewithal: the questionnaire included The Neoliberal Beliefs Inventory (NBI), which considers four factors: Government Interference preferences, Competition preferences, System Inequality perception and Personal Wherewithal [@bay-chengTrackingHomoOeconomicus2015].
-We used the factor personal wherewithal that reflects meritocracy beliefs in terms of the importance of personal attributes as strength and skills to yield success.
+We used the factor _personal wherewithal_ that reflects meritocracy beliefs in terms of the importance of personal attributes as strength and skills to yield success.
 The construct consists of 8 items (e.g., “Any goal can be achieved with enough hard work and talent”, “I’ve benefited from working hard, so there’s no reason others can’t”, “Anyone who is willing to work hard can be succesful in Chile”).
@@
 
-Table \@ref(tab:val-conv) shows the polyserial and pearson correlations between the PPM-S and common measures of meritocracy.
+Table \@ref(tab:val-conv) shows the polyserial and pearson correlations between our meritocracy scale and common measures of meritocracy.
 The findings provide evidence for convergent validity for the scale.
@@
 Findings also indicate that correlations of perceptions with opportunity beliefs are consistent with our expectations.
-The item for social origin shows a positive and moderate association with meritocratic perception (r=.354), and a negative and low correlation with meritocratic perception (-0.197).
+The item for social origin shows a positive and moderate association with unmeritocratic perception (r=.354), and a negative and low correlation with meritocratic perception (-0.197).
 Regarding the item of hard work, the correlation with meritocratic perception is positive but weak (r=.187).
-One explanation for this result is that the support for hard work is very high among respondents---only 8.6 percent indicate \textit{not very important} or \textit{not importat at all}.
+One explanation for this result is that the support for hard work is very high among respondents---only 8.6 percent indicates _not very important_ or _not important at all_.
 
@@
 
-The measure of NBI for \textit{personal wherewithal} also provides evidence for the convergent validity of PPM-S.
+The measure of NBI for _personal wherewithal_ also provides evidence for the convergent validity of PPM-S.
 Results indicate a positive correlation of this measure with meritocratic preference (r=.392), suggesting that both variables capture similar concepts.
 The NBI's factor depicts a positive association with unmeritocratic preference (r=.208).
-This finding is consistent with literature, inasmuch as the belief for meritocray is associated with the justification of inequality [@McCall2017Exposurerisinginequality;@MadeiraPrimesConsequencesSystematic2019].
+This finding is consistent with literature, inasmuch as the belief for meritocray is associated with the justification of inequality [@McCall2017Exposurerisinginequality; @MadeiraPrimesConsequencesSystematic2019].
 Furthermore, results of our confirmatory factor analysis indicate that meritocratic perceptions and unmeritocratic preferences show a positive and weak correlation.
@@
 
-El modelamiento para el testeo invarianza ha ido tomando mayor relevancia en los estudios sociológicos de encuestas, principalmente por el interés en la validez de escalas de medición en estudios de carácter comparado en diversos contextos sociales y culturales [@Davidovetal2014].
-En esta sección emplearemos un análisis de invarianza entre grupos, donde se espera que la estructura factorial propuesta sea independiente de la pertenencia a la muestra correspondiente al estudio original, comparándola con otro estudio disponible más reciente y donde también se incorporó la escala de meritocracia.
+The modeling for invariance testing has been gaining more relevance in sociological survey studies, mainly due to the interest in the validity of measurement scales in comparative studies in various social and cultural contexts [@Davidovetal2014].
+In this section we will use an analysis of invariance between groups, where the proposed factorial structure is expected to be independent of belonging to the sample corresponding to the original study, comparing it with another more recent study available and where the meritocracy scale was also incorporated.
 
-El procedimiento de medición de invarianza consiste en una serie de modelos anidados a los que progresivamente se le van incorporando restricciones sobre los parámetros del modelo de medición.
-La literatura sugiere en general que este modelamiento debe realizarse a través de cuatro niveles o tipos de restricción progresiva [@Milfont2010; @millsap_Statistical_2011; @vandeschoot_checklist_2012]:
+The invariance measurement procedure consists of a series of nested models to which restrictions are progressively incorporated on the parameters of the measurement model.
+The literature generally suggests that this modeling should be done through four levels or types of progressive restriction [@Milfont2010; @millsap_Statistical_2011; @vandeschoot_checklist_2012]:
 
-1.  **Configural:** se estima el modelo solo indicando la estructura  factorial utilizada en el CFA.
+1. **Configural:** the model is estimated only indicating the factorial structure used in the CFA.
 
-2.  **Débil:** se aplica una restricción de igualdad a las cargas factoriales en los distintos grupos, es decir, se fuerza a que las cargas sean idénticas en ambas mediciones.
+2. **Weak:** an equality restriction is applied to the factor loadings in the different groups, that is, the loadings are forced to be identical in both measurements.
 
-3.  **Fuerte:** se suman restricciones de igualdad a los interceptos de cada indicador.
+3. **Strong:** equality restrictions are added to the intercepts of each indicator.
 
-4.  **Estricta:** se suman restricciones a las varianzas de error de cada indicador.
+4. **Strict:** equality restrictions are added to the error variances of each indicator.
 
@@
 
-Para el desarrollo de este estudio se empleó una fuente de datos distinta a la utilizada hasta ahora, consistente en una encuesta online realizada durante el primer semestre de 2020, cuyas características en términos de aplicación y cobertura son equivalentes a las del estudio anterior.
-La muestra final obtenida contiene 1,242 casos, donde 605 (48,71%) son mujeres y 637 (51,28%) son hombres.
-El 48,84% de ellos tiene 45 años o más, y el 33,97% posee educación terciaria o superior.
-
-En la Tabla \@ref(tab:fcloads-inv) se pueden observar los resultados del ajuste de la escala en esta muestra.
+For testing the invariance we compared the data described previously with a new data source coming from an online survey carried out during the first half of 2020.
+The characteristics of this sample in terms of application and coverage are equivalent to those of the previous study.
+The final sample obtained contains 1,242 cases, where 605 (48.71%) are women and 637 (51.28%) are men.
+48.84% of them are 45 years of age or older, and 33.97% have tertiary education or higher.
 
@@
 
-La aplicación de la PPM-S se realizó con aleatorización completa de los indicadores, lo cual corresponde a la modalidad de aplicación del Grupo 3 presente en la Figura \@ref(fig:appmod).
+The items of the meritocracy scale were identical to the original.
+The items were administrated in a randomized order (which corresponds to the application modality of Group 3 present in Figure \@ref(fig: appmod)).
 
@@
 
-El primer paso para el testeo de invarianza es la estimación del modelo Configural, el cual se espera que cumpla adecuadamente con los criterios de ajuste global de un modelo de medición, debido a que es el modelo base en el proceso de comparación.
-Convencionalmente el estadístico chi-cuadrado es empleado como medida global de ajuste de un modelo de medición.
-Sin embargo, al ser sensible al tamaño de la muestra, se recomienda emplear tres índices de ajuste adicionales: Comparative Fit Index (CFI), el cual debe tener un valor mayor a 0.9; Root Mean Square Error Approximation (RMSEA), el cual debe estar en el rango de valores de 0.05 y 0.08; y el Standardised Root Mean Square Residuals (SRMR) que debe ser menor a 0.08 [@vandeschoot_checklist_2012].
+Tabla \@ref(tab:fcloads-inv) shows de results of the CFA estimation of the meritocracy scale in this new sample.
+As observed, the loadings and the fit indices show similar results as the ones presented for the original study above, which gives a first base with wich to start the invariance analysis.
 
-La literatura sobre invarianza métrica sugiere dos aproximaciones complementarias para el testeo de invarianza.
-En nuestro caso, emplearemos la comparación basada en el ajuste incremental de los estadísticos de ajuste [@cheung_Evaluating_2002; @Milfont2010; @dimitrov_Testing_2010] y el test ANOVA para comparación de medias en modelos anidados [@newsom_Longitudinal_2015].
+The first step for invariance testing is the estimation of the configural model, which serves as the baseline for further comparison and it is expected to adequately meet the global fit criteria of a measurement model.
+Although the chi-square statistic is used as a global measure of fit,since its sensitivity to sample size it is recommended to use three additional fit indices: Comparative Fit Index (CFI), which should have a value greater than 0.95; Root Mean Square Error Approximation (RMSEA), which must be in the range of values ​​of 0.05 and 0.08; and the Standardized Root Mean Square Residuals (SRMR) which must be less than 0.08 [@vandeschoot_checklist_2012].
+Besides, the literature on measurement invariance suggests some complementary approaches for the evaluation of the fit, from which we will consider the incremental adjustment of the fit indexes [@cheung_Evaluating_2002; @Milfont2010; @dimitrov_Testing_2010] and the ANOVA test for means comparison in nested models [@newsom_Longitudinal_2015].
 
@@
 
-Table \@ref(tab:invargroup) shows the results of the measurement invariance tests.
+Table \@ref(tab:invargroup) shows the results of the measurement invariance estimation.
 When attending to the traditional invariance test of $\Delta \chi^2 (\Delta \text{df})$, the results  support the invariance at the strong level meaning that the fit of the factor model of the PPM-S scale is equivalent across samples when constraining factor loadings and intercepts to being equal.
 Such result is considered in general as evidence of invariance [citar paper Fischer en SJR], as strict forms of measurement invariance rarely hold [citar van der Schoot et al].
-Still, the comparability of latents means requires strict invariance which in this case does not hold using $\Delta \chi^2 (\Delta \text{df})$.
-Nevertheless the criteria of $\Delta \text{CFI}$ used for comparing models is close to the rejection criteria of >.01, whereas the $\Delta \text{RMSEA}$ is below the cut-off criteria suggested by this criteria [citar  @chen 2007 ].
-Therefore, the tests for strict invariance show mixed evidence.
+Still, the comparability of latents means requires strict invariance which in this case does not hold when considering $\Delta \chi^2 (\Delta \text{df})$.
+Nevertheless, the criteria of $\Delta \text{CFI}$ used for comparing models is close to the rejection criteria of >.01, whereas the $\Delta \text{RMSEA}$ fulfills the requirements of being below of the cut-off criteria as suggested by [ @chen 2007 ].
+Therefore, using this last standard the level for strict invariance holds for the meritocracy scale.
 
@@
 Studies that attempt to characterize and compare societies by their support for meritocratic beliefs have used different approaches.
-As most studies use secondary survey data, they tend to assume that the available indicators represent an underlying construct.
+As most studies use secondary survey data, they tend to assume that the available indicators represent an underlying meritocratic construct.
 A review of these studies reveals several non-tested assumptions, as well as the use of similar indicators to represent different constructs and dimensions of meritocracy.
-As the existence of heterogeneous approaches certainly has consequences for the advancement of the study of meritocracy, this paper presents a comprehensive conceptual framework for the empirical study of meritocracy, building upon previous research.
+As the existence of heterogeneous approaches certainly has consequences for the advancement of the study of meritocracy, this paper presented a comprehensive conceptual framework for the empirical study of meritocracy, building upon previous research.
 This framework was then tested against new survey data.
@@
 Meritocratic and non-meritocratic dimensions do not appear to constitute poles of the same continuum, as some previous studies have assumed.
-Besides, we find evidence of convergent validity as well as measurement invariance that approach well the requirement of comparability between samples.
+Regarding the possible effects of items order in the estimation of latent variable, we tested three different order versions and found evidence that suggest the use of the randomized version of the scale.
+Furthermore, we found evidence of convergent validity as well as measurement invariance that approach well the requirement of comparability between samples.
 
@@
 For instance, distinguishing perceptions from preferences will allow us to evaluate the extent to which different societies are accustomed to, or satisfied with, the perceived level of meritocracy, in terms of differences between what is perceived and what is preferred.
-Additionally, given that non-meritocratic factors are not necessarily related to meritocratic ones, our framework makes it possible to assess the perceived legitimacy of practices such as the use of personal contacts and their interference (or not), within meritocratic ideals in different societies.
-The consideration of structural factors, both individual and societal, will allow us to advance future hypotheses relating to meritocratic legitimacy to a more comprehensive level than the curren one, which is based solely on meritocratic perceptions.
+Additionally, given that non-meritocratic factors are not necessarily related to meritocratic ones, our framework makes it possible to assess the perceived legitimacy of practices such as the use of personal contacts and their interference (or not) with meritocratic ideals in different societies.
+The consideration of structural factors, both individual and societal, will allow us to advance future hypotheses relating to meritocratic legitimacy to a more comprehensive level than the current one, which is based solely on meritocratic perceptions.
 Furthermore, the impacts of different configurations of the four-dimensional framework on practices and behaviors such as corruption, civic involvement, and political alignment, is an area that requires additional research.

### 2021-01-29 docs/paper.Rmd
ultima pasada JC antes de vacaciones

@@
 In general, meritocracy is defined as a system of distribution of resources and rewards based on individual merit, which in its original conception is a combination of talent and effort [@young_rise_1962].
-This traditional conception of merit, places in a secondary position the possible interference of structural or non-meritocratic factors, such as inheritance, personal contacts, and luck [@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995; @yairMeritocracy2007; @landWeSatTable2006; @youngRiseMeritocracy1994].
-Social psychology and sociology have studied the characteristics and consequences of beliefs in meritocracy, under the general hypothesis that a greater belief in meritocracy leads to greater legitimization of inequalities [@premingerMeritocracyServiceEthnocracy2020; @trumpWhenWhyEconomic2020; @hadjar_meritokratie_2008; @MadeiraPrimesConsequencesSystematic2019].
-Such research has raised criticism of meritocracy as a moral standard of distribution, arguing that it would be an unfulfilled promise, given the preponderant influence of non-meritocratic elements upon individual status [@sandelTyrannyMeritWhat2020a; @witteveenReconsideringMeritocraticPower2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019; @khan_Privilege_2013].
+This traditional conception of merit places in a secondary position the possible interference of structural or non-meritocratic factors, such as inheritance, personal contacts, and luck [@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995; @yairMeritocracy2007; @landWeSatTable2006; @youngRiseMeritocracy1994].
+Social psychology and sociology have studied the characteristics and consequences of beliefs in meritocracy, under the general hypothesis that a greater belief in meritocracy emphasizes the role of individual over structural factors in personal achievements, leading to greater legitimization of inequalities [@premingerMeritocracyServiceEthnocracy2020; @trumpWhenWhyEconomic2020; @hadjar_meritokratie_2008; @MadeiraPrimesConsequencesSystematic2019].
+Such research has raised criticism of meritocracy as a moral standard of distribution, arguing that it would be an unfulfilled promise given the still preponderant weight of non-meritocratic elements upon individual status and social mobility[@sandelTyrannyMeritWhat2020a; @witteveenReconsideringMeritocraticPower2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019; @khan_Privilege_2013].
 
-Due to the role that meritocratic beliefs play in the justification of individual achievement or failure in contemporary societies [@hadjar_meritokratie_2008] [acá también sumar a Markovits(Trap) y Sandel, último libro], multiple investigations have evaluated the relationship between meritocratic beliefs and personal/contextual characteristics.
-For example, some studies have linked meritocracy to the reinforcement of socio-economic, gender, and ethnic stereotypes [@MadeiraPrimesConsequencesSystematic2019; @girerdNeoliberalismIdeologicalBarrier2020; @premingerMeritocracyServiceEthnocracy2020], as well as lines of research have evaluated the effects of meritocratic beliefs in educational [@generettStoriesWeTell2020; @owensENGINESSOCIALMOBILITY2020] and organizational contexts [@perezAdvancingCareersMerit2020; @aielloNewEvidenceFirmuniversity2019].
+Due to the role that meritocratic beliefs play in the justification of individual achievement or failure in contemporary societies [@hadjar_meritokratie_2008, @markovits_meritocracy_2019; @sandel_tyranny_2020], multiple studies have evaluated the relationship between meritocratic beliefs and personal and/or contextual characteristics.
+For example, some studies have linked meritocracy to the reinforcement of socio-economic, gender, and ethnic stereotypes [@MadeiraPrimesConsequencesSystematic2019; @girerdNeoliberalismIdeologicalBarrier2020; @premingerMeritocracyServiceEthnocracy2020], as well as the effects of meritocratic beliefs in educational [@generettStoriesWeTell2020; @owensENGINESSOCIALMOBILITY2020] and organizational contexts [@perezAdvancingCareersMerit2020; @aielloNewEvidenceFirmuniversity2019].
 
-To measure beliefs in meritocracy, studies have generally used indicators from existing surveys, and to a lesser extent, have created ad-hoc instruments.
+Most of the studies so far have used indicators from existing standard surveys (as the International Social Survey Programme - ISSP) to measure meritocracy.
 However, as we will show later, the methods of measuring meritocracy vary extensively among studies.
 In many cases, similar phenomena are associated with different indicators, and conversely, often different phenomena are measured with similar indicators.
-Such trends confound not only the comparability of studies, but also the ability to understanding the effects of meritocratic beliefs across different fields.
+Such trends not only limits the comparability of studies, but also the ability to understanding the effects of meritocratic beliefs across different fields.
 
 Based on a critical analysis of different approaches to the measurement of meritocracy to date, this article proposes an instrument to both measure and relate two key aspects in the study of meritocracy: perceptions and preferences.
-Furthermore, as a second axis of analysis, it considers the elaboration of indicators regarding meritocratic and non-meritocratic aspects, demonstrating that they are not two poles of the same continuum, as some previous studies would seem to suggest.
+Furthermore, as a second axis of analysis, it considers the elaboration of indicators regarding meritocratic and non-meritocratic aspects, demonstrating that they are not two poles of the same continuum as some previous studies would seem to suggest.
 The measurement system outlined is oriented to generate an instrument as simple and brief as possible, being suitable to be used in public opinion surveys and allowing to integrate meritocratic beliefs in the study of different social phenomena.
@@
 
-Several approaches to the empirical study of meritocracy based on public opinion surveys, refer to the concept of _beliefs_, however different meanings and operationalizations usually underlie this concept.
-To illustrate this point, we will take as a main reference a recent paper about meritocracy by @mijs_paradox_2019, because of its novelty and its influence in current research about meritocracy, particularly in sociology.
-The definition of meritocratic beliefs stated in this paper is the following: "when I discuss meritocracy beliefs, I am referring to citizens’ belief in the importance of hard work relative to structural factors." [@mijs_paradox_2019, pg.9].
+In the following we discuss four critical aspects in studies conceptualizing and measuring meritocracy and that are the basis of our scale proposal.
+
+a. _Conceptual components_: Is merit only effort?
+
+One recent definition of meritocracy by @mijs_paradox_2019 is the following: "when I discuss meritocracy beliefs, I am referring to citizens’ belief in the importance of hard work relative to structural factors." [@mijs_paradox_2019, pg.9].
 In the subsequent operationalization, this is associated with the following question and indicator: "how important you think it is for getting ahead in life: (a) hard work", scored from 1 to 5 on a likert scale.
 The assumptions behind such definition are worth discussing in light of the conceptual meaning of meritocracy and its possibilities of operationalization.
-
-a. _Conceptual components_: Is merit only effort?
 
@@
 
-In this measure of meritocratic beliefs, we can observe a couple of strong conceptual assumptions: effort would not depend on parental influence, and talent (as innate ability) is not meritocratic (contrary to Michael Youngs original conceptualization).
+In this measure of meritocratic beliefs, we can observe a couple of strong conceptual assumptions: effort would not depend on parental influence, and talent (as innate ability) is not meritocratic (contrary to Michael Youngs' original conceptualization).
 This conceptual and measurement-based assumption is found in other studies, that also assume effort to be the main and only aspect of meritocracy [@girerdNeoliberalismIdeologicalBarrier2020a; @bubakPerceptionsMeritocracyNote2019].
@@
 
-The most commonly used items for operationalizing meritocracy in survey research is the "reasons to get ahead" battery, which asks "how important you think it is" and then lists several factors, such as effort, education, parental wealth, and contacts.
-Another version of this same battery used in several surveys - sometimes along the previous one - asks about "how important you think it _should_ be".
+Several approaches to the empirical study of meritocracy based on public opinion surveys, refer to the concept of _beliefs_, however different meanings and operationalizations usually underlie this concept.
+To illustrate this point, lets consider the most commonly used items for operationalizing meritocratic beliefs in survey research (as ISSP), which is the "reasons to get ahead" battery.
+It consists in a series of items asking "how important you think it is for getting ahead in life" and then lists several factors, such as effort, education, parental wealth, and contacts.
+Another version of this same battery used in several surveys - sometimes along the previous one - asks about "how important you think it _should_ be ...", then listing the same concepts.
 Therefore, the question raised here is: Which one of both is a "belief"?
@@
 This opens up the possibility of analyzing whether perceptions and preferences are actually related (i.e.
-have a correlation close to 1), or whether they are independent aspects of the same phenomenon.
+have a high correlation), or whether they are independent aspects of the same phenomenon.
 As @son_hing_merit_2011-1 have pointed out, "People can believe that outcomes ought to be distributed on the basis of merit and yet vary in their perceptions of whether this is how society currently operates" (p.
@@
 In other words, normative beliefs should be considered while taking perceptions into account: a strong normative belief in meritocracy may mean something totally different to someone perceiving high meritocracy, when compared to someone perceiving low meritocracy.
-To avoid the confusion generated by the term "belief", we propose the terms meritocratic preferences ("what should be"), and meritocratic perceptions ("what is"), as they better reflect the two facets of meritocracy under scrutiny **[citar Castillo et al 2019]**.
+To avoid the confusion generated by the term "belief", we propose the terms meritocratic preferences ("what should be"), and meritocratic perceptions ("what is"), as they better reflect the two facets of meritocracy under scrutiny [@castillo_meritocracia_2019].
 
@@
 
-Unlike studies that consider talent to be a non-meritocratic aspect, and therefore rule it out of the operationalization of meritocracy, @kunovich_systems_2007 follow a different approach, by including some non-meritocratic elements.
-Using an items’ battery listing a number of factors in relation to “How important each should be in deciding pay...” (as @duru-bellat_whos_2012 for desired meritocracy), they consider factors such as education and responsibility as meritocratic, giving them a value of 1 if considered “essential” in the scale response, whereas factors such as having a family or children are valued 1 when rated as "not important at all" (i.e.
+A number of studies have dealt with aspects opposed to meritocracy, as for instance the use of personal advantages (as contacts or having a wealthy family) to get ahead in life.
+For instance, @kunovich_systems_2007 used an items’ battery listing a number of factors in relation to “How important each should be in deciding pay...” (as @duru-bellat_whos_2012 for desired meritocracy).
+They consider factors such as education and responsibility as meritocratic, giving them a value of 1 if considered “essential” in the scale response, whereas factors such as having a family or children are valued 1 when rated as "not important at all" (i.e.
 reverse coded).
@@
 Similarly, @roex_attitudes_2019 used ISSP indicators for perceived meritocracy and non-meritocracy to build a single score by reverse coding the non-meritocratic items.
-We suggest that non-meritocratic factors should be part of meritocratic measurements, though it should be considered independently rather than adding to or subtracting from meritocratic factors, unless it is empirically proved that they belong to the same conceptual dimension.
+Therefore, the treatment of non-meritocratic items has been rather inconsistent across studies and the assumption that they are the simple opposite of meritocracy requires further empirical assessment.
 
@@
 Finally, most studies in meritocracy so far have not accounted for the issue of measurement error [@brownBookReviewTimothy2010; @bollenStructuralEquationsLatent1989] (Ansolabehere, S., Rodden, J., & Snyder Jr, J.
-M., 2008), when using single indicators and/or simple average indexes for measuring meritocracy.
-Such strategy assumes that the latent construct is measured perfectly by the indicators chosen, going as far as to propose that "...
+M., 2008), as they use single indicators and/or simple average indexes for measuring meritocracy.
+Such strategy assumes that the latent construct is measured perfectly (i.e.
+no error or residual variance) by the indicators chosen, going as far as to propose that "...
 In choosing this strategy of index construction, we argue that support for meritocracy is not a latent variable" [@kunovich_systems_2007 p.653-654].
@@
 
-- Based on _previous indicators_ as much as possible, for the sake of comparability between studies.
+- Based on _previous indicators_ as much as possible, for the sake of keeping comparability between studies.
 
@@
 ```
-The columns "Perceptions" and "Preferences" represent the distinction between these two concepts, usually confused under the label "beliefs" [@castillo_Meritocracia_2019].
+The columns "Perceptions" and "Preferences" represent the distinction between these two concepts, usually confused under the label "beliefs" [@castillo_meritocracia_2019].
 Perceptions refer to the extent to which people observe that meritocracy works or applies in their society, which relates to items such as "I think hard work is important to get ahead in society".
@@
 
-* $H_2$. The perception of non-meritocracy is a latent variable that derives from two indicators related to agreement with the statement that people with contacts and rich parents manage to get ahead.
+* $H_2$. The perception of non-meritocracy is a latent variable that derives from  indicators related to agreement with the statement that people with contacts and rich parents manage to get ahead.
 
@@
 The questionnaire was programmed in Qualtrics and the fieldwork was conducted by an external online survey agency ([netquest.cl](www.netquest.cl)) between December 2019 and January 2020.
-The sample was selected from a non-probabilistic quota design in three large cities in Chile (Santiago, Concepción, Antofagasta **[check]**).
+The sample was selected from a non-probabilistic quota design in three large cities in Chile (Santiago, Concepción & Antofagasta).
 The quotas for gender, age, and education level were generated based on a survey by the Public Studies Center [@cepEncuestaCEPMayo2019], which is a well regarded counterpart agency of the ISSP (International Social Survey Programme) in Chile.
@@
 There were no significant differences between our sample and the wider population for most socio-demographic characteristics, with the exception of educational level (see Table \@ref(tab:rep-samp) in appendix).
-As is often the case with online surveys, there were some limitations in achieving the quotas for lower educational levels **[citas]**.
+As is often the case with online surveys, there were some limitations in achieving the quotas for lower educational levels **[citas - LUCHO]**.
 
@@
 The graphs presented in Figure \@ref(fig:plotlikert) display disaggregated and comparable information of the different response categories for each item.
-It can be seen that, in general, there is more agreement in the perception of non-meritocratic items than in meritocratic items, while in the case of preferences, the opposite occurs.
+As it can be observed, in general there is more agreement in the perception of non-meritocratic items than in meritocratic items, while in the case of preferences, the opposite occurs.
 As far as preferences are concerned, the preponderant role of effort over talent as a criterion of meritocratic preference is noteworthy.
@@
 Likewise, a relatively low perception of meritocracy is observable.
-In addition, we observed a coherent relationship between the indicators of the same dimension as proposed in the pre-registration of the study; that is, the pairs of items associated with a specific factor show correlations with a large effect size (for example, meritocratic preferences for items associated with effort and talent).
+In addition, we observed a coherent relationship between the indicators of the same dimension as proposed in the pre-registration of the study; that is, the pairs of items associated with a specific factor show correlations with a large effect size.
 In particular, the associations between effort and talent are relevant, challenging previous assumptions that talent would not be a meritocratic criterion [@mijs_paradox_2019], otherwise the correlation would be zero or negative.
@@
 
-The measure of NBI for _personal wherewithal_ also provides evidence for the convergent validity of PPM-S.
+The measure of NBI for _personal wherewithal_ also provides evidence for the convergent validity of the merit scale.
 Results indicate a positive correlation of this measure with meritocratic preference (r=.392), suggesting that both variables capture similar concepts.
@@
 Table \@ref(tab:invargroup) shows the results of the measurement invariance estimation.
-When attending to the traditional invariance test of $\Delta \chi^2 (\Delta \text{df})$, the results  support the invariance at the strong level meaning that the fit of the factor model of the PPM-S scale is equivalent across samples when constraining factor loadings and intercepts to being equal.
+When attending to the traditional invariance test of $\Delta \chi^2 (\Delta \text{df})$, the results  support the invariance at the strong level meaning that the fit of the factor model of the merit scale is equivalent across samples when constraining factor loadings and intercepts to being equal.
 Such result is considered in general as evidence of invariance [citar paper Fischer en SJR], as strict forms of measurement invariance rarely hold [citar van der Schoot et al].

### 2021-03-23 docs/paper.Rmd
Study 1 ISSP

@@
 
+The analysis is organized in three studies.
+Study 1 serves as a background to our instrument proposal by attempting to operationalize the meritocracy dimensions with available data.
+For this we use data from the International Social Survey Programme (ISSP).
+Although we are aware that the available indicators in this dataset do not allow a precise operationalization of the conceptual model, we know that this dataset is the most widely used in comparative meritocracy studies (as well as overall subjective inequality) and therefore it is certainly useful to look at its potentials and limitations to this regard.
+Study 2 is the core of this paper and consists of the proposal of a new scale for measuring meritocratic perceptions and preferences.
+Finally, in Study 3 we perform additional validity analysis of the meritocracy scale.
+
+## Study 1
+
+### Data
+
+The data used in the analysis corresponds to the last available wave of the social inequality module from the International Social Survey Programm (ISSP), which is the most specialized international comparative survey in perceptions, attitudes and beliefs about inequality related issues.
+This wave corresponds to the year 2009 and covers attitudes towards a series of topics dealing with social inequality across 26 countries and XXXX individuals.
+Although there is data available from this module for previous waves (1987, 1992 and 1999), unfortunately there are several variables that are important for this research that were not included, particularly in 1987 and 1999, reason why we only use the last wave.
+
+### Variables
+
+There are a series of indicators in the ISSP survey that in the following we attempt to classify in our meritocracy measurement scheme.
+However, it is important to mention  that the items were not originally thought as a specific measure of meritocracy.
+Still, many of the items have been used in previous research on meritocracy (and certainly will keep being used in the future) and before proposing a new measure we consider to give a first test to our proposed conceptualization using available data that not only covers several countries but also is validated by the research community.
+
+The variables are presented below organized in perceptions and preferences:
+
+- _Perception of meritocracy/non-meritocracy_: for operationalizing perceptions the closest set of ISSP's indicators comes from the question asking about perceptions for opportunities to get ahead, which are usually considered as "meritocratic beliefs" in previous research.
+The general heading of the battery is: _"To begin we have some questions about opportunities for getting ahead.
+Please tick one box for each of these to show how important you think it is for getting ahead in life."_ This is followed by a list of statements to be answered in a scale with the following levels numbered from 1 to 5: essential, very important, fairly important, not very important, not important at all.
+Below the items are listed classified according to the meritocracy/non-meritocracy character:
+
+  - Meritocratic perception:
+    - How important is having ambition?
+    - How important is hard work?
+
+  - Non-meritocratic perception
+    - How important is coming from a wealthy family?
+    - How important is having well-educated parents
+    - How important is knowing the right people
+    - How important is having political connections
+    - How important is a person's race?
+    - How important is being born a man or a woman?
+
+<br>
+  The classification is based on criteria of internal motivation (meritocratic) and structural constraints (non-meritocratic).
+There were two items from the battery that were excluded from the analysis as they would not fit into the classification.
+The first one was "having good education yourself", since it was not clear whether this is could be due to individual motivation or system opportunities, and the second was "giving bribes", as introduced elements of criminality that were beyond a non-meritocratic perception.
+
+- Preferences for meritocracy-non meritocracy: for operationalizing normative preferences we used a list of items related to reasons for pay.
+The ISSP question was: _In deciding how much people ought to earn, how important should each of these things be, in your opinion_, rated in the same _essential-non important at all_ scale as the questions for meritocratic preferences.
+The reasons are:
+
+  - Meritocratic preferences
+    - How well he or she does the job?
+    - How hard he or she works at the job?
+
+  - Non-meritocratic preferences
+    - What is needed to support a family?
+    - Whether the person has children to support?
+
+(here summary descriptive table - @julio)
+
+### Methods
+
+The estimation was performed using Confirmatory Factor Analysis (CFA).
+CFA was conducted using the `lavaan` R package (version 0.6-3; Rosseel, 2020), with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement (Kline, 2016; Rosseel, 2020).
+As recommended by Brown (2008), we assessed model fit by jointly considering the comparative fit index and Tucker-Lewis Index (CFI and TLI; acceptable fit > 0.95), Root of the average squared residual approximation (RMSEA; acceptable fit < 0.08),  Chi-square: (p-value; acceptable fit > 0.05, and Chi-square ratio > 3).
+
+### Results
+
+[diagrama SEM]
+
+[tabla CFA]
+
+## Study 2
+
 ## Data collection {-#data}
@@
 ```
-CFA was conducted using the `lavaan` R package (version 0.6-3; Rosseel, 2020), with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement (Kline, 2016; Rosseel, 2020).
-As recommended by Brown (2008), we assessed model fit by jointly considering the comparative fit index and Tucker-Lewis Index (CFI and TLI; acceptable fit > 0.95), Root of the average squared residual approximation (RMSEA; acceptable fit < 0.08),  Chi-square: (p-value; acceptable fit > 0.05, and Chi-square ratio > 3).
+As in Study 1, CFA was conducted using the `lavaan` R package (version 0.6-3; Rosseel, 2020), with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement (Kline, 2016; Rosseel, 2020).
+The fit indexes and cut-off criteria were the same as in Study 1.
 

### 2021-03-26 docs/paper.Rmd
commit01
update files

@@
 
-(here summary descriptive table - @julio)
+<!-- (here summary descriptive table - @julio) -->
+
+```{r desc.issp, results='asis'}
+load(file = "../output/tables/desc.issp.Rdata")
+sum.issp09<- kable(desc.issp,format =table_format,
+                 booktabs= T,
+                 align = "l",
+                 linesep = "",
+                 caption = "Descriptive statistics for the ISSP 2009 sample") %>%
+  kable_styling(full_width = table_format2,
+                font_size = 10,
+                position = "center",
+                latex_options = c("HOLD_position"));sum.issp09
+```
 
@@
 
-[diagrama SEM]
+<!-- [diagrama SEM] -->
 
-[tabla CFA]
+```{r meas02-issp, echo=FALSE, fig.cap = "Descriptive plot", fig.align='center'}
+knitr::include_graphics('../output/images/meas02-issp.PNG')
+```
+
+<!-- [tabla CFA] -->
+```{r tb.perpref1, echo=FALSE}
+load(file = "../output/tables/tb.perpref1.RData")
+tb.foot <- paste0("Standardized factor loadings using DWLS estimator ; CFI = Comparative fit index ;TLI = Tucker-Lewis index; RMSEA = Root mean square error of approximation")
+tb.col <- c("","Meritocratic","Non-meritocratic","Meritocratic","Non-meritocratic")
+tb.caption <- c(" Factor loadings and fit measures for ISSP 2009")
+tb.issp<- kable(tb.perpref1,escape = FALSE,align = "lcccc",col.names = tb.col, caption = tb.caption, booktabs = T,linesep = "",format = table_format) %>%
+          kable_styling(full_width = table_format2,latex_options = c("hold_position"), position = "center",font_size = 12) %>%
+          add_header_above(header = c(" "=1,"Perception"= 2,"Preference"= 2)) %>%
+          add_header_above(header = c(" "=1,"Factor loadings"= 4)) %>%
+          row_spec(row = 12,hline_after = TRUE) %>%
+          add_indent(c(13:17)) %>%
+          footnote(general =tb.foot ,footnote_as_chunk = T);tb.issp
+```
 

### 2021-04-01 docs/paper.Rmd

--- 2021-03-23 docs/paper.Rmd
@@
 
-(here summary descriptive table - @julio)
+<!-- (here summary descriptive table - @julio) -->
+
+Table \@ref(tab:desc-issp)
+
+```{r desc-issp, echo=FALSE, results='asis'}
+load(file = "../output/tables/desc.issp.Rdata")
+sum.issp09<- kable(desc.issp,format =table_format,
+                 booktabs= T,
+                 align = "l",
+                 linesep = "",
+                 caption = "Descriptive statistics for the ISSP 2009 sample") %>%
+  kable_styling(full_width = table_format2,
+                font_size = 10,
+                position = "center",
+                latex_options = c("HOLD_position"));sum.issp09
+```
 
@@
 
-[diagrama SEM]
+<!-- [diagrama SEM] -->
 
-[tabla CFA]
+Figure \@ref(fig:meas02-issp)
+
+```{r meas02-issp, echo=FALSE, fig.cap = "Descriptive plot", fig.align='center'}
+knitr::include_graphics('../output/images/meas02-issp.PNG')
+```
+
+<!-- [tabla CFA] -->
+
+Table \@ref(tab:tb-perpref1)
+
+```{r tb-perpref1, echo=FALSE}
+load(file = "../output/tables/tb.perpref1.RData")
+tb.foot <- paste0("Standardized factor loadings using DWLS estimator ; CFI = Comparative fit index ;TLI = Tucker-Lewis index; RMSEA = Root mean square error of approximation")
+tb.col <- c("","Meritocratic","Non-meritocratic","Meritocratic","Non-meritocratic")
+tb.caption <- c(" Factor loadings and fit measures for ISSP 2009")
+tb.issp<- kable(tb.perpref1,escape = FALSE,align = "lcccc",col.names = tb.col, caption = tb.caption, booktabs = T,linesep = "",format = table_format) %>%
+          kable_styling(full_width = table_format2,latex_options = c("hold_position"), position = "center",font_size = 12) %>%
+          add_header_above(header = c(" "=1,"Perception"= 2,"Preference"= 2)) %>%
+          add_header_above(header = c(" "=1,"Factor loadings"= 4)) %>%
+          row_spec(row = 12,hline_after = TRUE) %>%
+          add_indent(c(13:17)) %>%
+          footnote(general =tb.foot ,footnote_as_chunk = T);tb.issp
+```
 

### 2021-04-01 production/prod_analysis-cfa-study01.rmd

---
title: "Analysis PPM-S"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
css: "../input/css/custom.css"
output:
  bookdown::html_document2:
    number_sections: true
    toc: true
    toc_float:
        collapsed: false
    toc_depth: 2
    code_folding: "hide"
editor_options:
  chunk_output_type: console
---

```{r include=FALSE}
knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = TRUE)
knitr::opts_knit$set(root.dir = "../")
options(scipen=999)
rm(list=ls())
options(knitr.kable.NA = '')
```

**Libraries**
```{r}
library(sjPlot)
library(dplyr)
library(lavaan)
library(semPlot)
library(stargazer)
library(corrplot)
library(psych)
library(knitr)
library(kableExtra)
library(rvest)
library(sjlabelled)
library(ggplot2)
library(stargazer)
```

**Load data**

```{r}
load(file = "input/data/proc/getahead.rda")
load(file = "input/data/proc/detearn.rda")
```

```{r}
table_format = if(is_html_output()) {
  "html"
} else if(is_latex_output()) {
  "latex"
}

```

# Perception

## Descriptives
```{r}
# Summary table
names(getahead)

stargazer(getahead[, -c(1, 2, 11:18)], type = "text")

# Simple freq graph
plot_frq(getahead$wealthy,
        show.values = TRUE,
        coord.flip = TRUE)

# Likert graph
plot_likert(getahead %>% select(wealthy:gender)) # ajustar

# short var labels
items <- c("Wealthy family", "Educated parents","Ambition", 
           "Hard work", "Race","Gender", "Know people", "Pol. connect.")
# cambiar orden para la leyenda
labels = c("Essential",
           "Very important",
           "Not very important",
           "Not important at all",
           "(neutral)")

# ajustes globales sjplot

set_theme(
  theme = "blank",
  legend.inside = TRUE,
  legend.pos = "bottom right",
  legend.size = 0.8,
  geom.outline.size = 0,
  geom.label.size = 3)

# Gráfico final

plot_likert(getahead%>% select(wealthy:gender),
           axis.labels   = items,
           legend.labels = labels,
           cat.neutral   = 3,
           # identifica a indiferentes
           geom.colors   = c("#9ecae1", "#6baed6", "#4292c6",
                             "#2171b5"),
           # colorbrewer2.org para temas
           sort.frq      = "neg.asc",
           # sort descending)
           title         = "How important is for getting ahead ...",
           intercept.line.color = "white",
           # middle line white
           expand.grid   = FALSE,
           # no inner margins in plot
           show.n        = FALSE,
           # hide N's in axis labels
           grid.range    = 1.3
) # expand axis from 130% to 130%

# Save
# dev.copy(png, "./results/likert_ga.png")
# dev.off()

# Correlations

# Matrices & plot

cormat_ga = cor(getahead %>% select(wealthy:gender),
                use = "pairwise.complete.obs")
corrplot.mixed(cormat_ga)
# dev.copy(png,
#          "./results/corplot_ga.png",
#          width = 600,
#          height = 600)
# dev.off()

# Correlation table a latex
# sink("./results/cortab_ga.txt")
# cor2latex(
#   getahead[2:10],
#   stars   = TRUE,
#   adjust = "none",
#   apa     = TRUE,
#   heading = " ",
#   caption = "Reasons to get ahead")
# sink()
```

## Measurement models

### Continuos 
```{r}
# Estimación del modelo de 1er order
cfa_1 <- '
	  # latent variables
	  merit =~ hwork + ambition
	  parent=~ wealthy + pareduc
	  backg=~ race + gender
		network=~ people + polcone
	  '
fit_1 <- cfa(cfa_1, data = getahead, missing = "ML")
summary(fit_1, fit.measures = TRUE, standardized = TRUE)

# Path model
semPaths(
  fit_1,
  layout = "tree2",
  rotation = 2,
  intercepts = FALSE,
  style = "lisrel",
  curvePivot = TRUE,
  cardinal = TRUE
)
```

### Categorical

```{r}
# Estimación categorical 3 levels (5 levels convergence prob.)
cfa_1c <- '
		  # latent variables
		  merit=~ c3hwork + c3ambition
		  parent=~ c3wealthy + c3pareduc
		  backg=~ c3race + c3gender
			network=~ people + polcone
		  '
fit_1c <- cfa(
  cfa_1c,
  data = getahead,
  ordered = c(
    "c3hwork",
    "c3ambition",
    "c3wealthy",
    "c3pareduc",
    "c3race",
    "c3gender",
    "c3people",
    "c3polcone"
  )
)
summary(fit_1c, fit.measures = TRUE, standardized = TRUE) # works
mod = modificationindices(fit_1c)
subset(mod, mi > 3.84, select = lhs:mi)
```

```{r eval=FALSE, include=FALSE}
# Estimación del modelo 2nd order

cfa_2 <- '
	  # latent variables
	  merit =~ hwork + ambition
	  parent=~ wealthy + pareduc
	  backg=~ race + gender
		network=~ people + polcone
		unmerit =~parent + backg + network
	  '
fit_2 <- cfa(cfa_2, data = getahead, missing = "ML")
summary(fit_2, fit.measures = TRUE, standardized = TRUE)

# Path
semPaths(
  fit_2,
  layout = "tree",
  rotation = 2,
  intercepts = FALSE,
  style = "lisrel",
  curvePivot = TRUE,
  cardinal = FALSE
)
dev.copy(png,
         "./results/path_cfa2.png",
         width = 750,
         height = 600)
dev.off()
```

# Preference

## Descriprtives
```{r}    
# 2.2 Descriptives ---------------------- ######

load(file = "input/data/proc/detearn.rda")

# Summary table
names(detearn)
stargazer(detearn[,-c(1,2)], type = "text")

# Summary table to latex
stargazer(detearn[, -c(1, 2)])

# Simple freq graph
sjp.frq(detearn2$respons,
        show.values = TRUE,
        coord.flip = TRUE)

# Likert graph (uses detearn2 object, with original coding)

sjp.likert((detearn2[, -c(1, 2)])) # ajustar

# short var labels
items <- c("Responsibility",
           "Years education",
           "Family",
           "Children",
           "Well job",
           "Hard job")

# cambiar orden para la leyenda
labels = c("Essential",
           "Very important",
           "(neutral)"
           ,
           "Not very important",
           "Not important at all")

# ajustes globales sjplot
theme_set(
  theme = "blank",
  legend.inside = TRUE,
  legend.pos = "bottom right",
  legend.size = 0.8,
  geom.outline.size = 0,
  geom.label.size = 3
)

plot_likert((detearn2[, -c(1, 2)]),
           axis.labels   = items,
           legend.labels = labels,
           cat.neutral   = 3,# identifica a indiferentes
           geom.colors   = c("#9ecae1", "#6baed6", "#4292c6","#2171b5"),# colorbrewer2.org para temas
           sort.frq      = "neg.asc",# sort descending)
           title         = "In deciding how much people ought to earn, how important should each of these things be...",
           intercept.line.color = "white",# middle line white
           expand.grid   = FALSE,# no inner margins in plot
           show.n        = FALSE,# hide N's in axis labels
           grid.range    = 1.3
) # expand axis from 130% to 130%

# Save
dev.copy(png, "./results/likert_detearn.png")
dev.off()

# freq(detearn2$pareduc)   # check educated parents, ok

# Correlations
cormat = cor(detearn[5:8], use = "pairwise.complete.obs")
corrplot.mixed(cormat)

#save
dev.copy(png,
         "./results/corplot_detearn.png",
         width = 600,
         height = 600)
dev.off()
```

## Measurement models

```{r}
cfa_1 <- '
	  # latent variables
	  bmerit =~ welljob + hardjob
	  need=~ family + child
	  '
fit_1 <- cfa(cfa_1, data = detearn, missing = "ML")

summary(fit_1, fit.measures = TRUE, standardized = TRUE)

# Path model
semPaths(
  fit_1,
  "std",
  layout = "tree2",
  rotation = 2,
  intercepts = FALSE,
  style = "lisrel",
  curvePivot = TRUE,
  cardinal = TRUE,
  edge.color = "black",
  residScale = 10
)

dev.copy(png,
         "./results/path_detearn1.png",
         width = 750,
         height = 600)
dev.off()

# Categorical model
fit_1c <- cfa(cfa_1,
              data = detearn,
              ordered = c("welljob",
                          "hardjob", "family", "child"))

summary(fit_1c, fit.measures = TRUE, standardized = TRUE)  #probs
```

# Full model

## Descriptives

```{r}
load(file = "input/data/proc/merit_per_pref.rda")
names(merit_per_pref)
merit_per_pref %>% 
  dplyr::select(ambition,hwork,wealthy,pareduc,people,polcone,race,gender,welljob,hardjob,family,child) %>% 
  psych::describe() %>% 
  as.data.frame() %>% 
  dplyr::select("Mean"=mean,"SD"=sd,"Min"=min,"Max"=max) %>% 
  round(2) ->desc.issp

row.names(desc.issp) <- c(
  "Importance: hard work",
  "Importance: having ambition",
  "Importance: wealthy family",
  "Importance: educated parents",
  "Importance: race",
  "Importance: gender",
  "Importance: knowing people",
  "Importance: political connections",
  "Reasons for pay: well job",
  "Reasons for pay: hard job",
  "Reasons for pay: support family",
  "Reasons for pay: has children"
)

save(desc.issp,file = "output/tables/desc.issp.Rdata")
```

## Measurement model

```{r}    
# 3. MEASURING BELIEFS & PERCEPTIONS ---------------------------------------------------------------
# 3.1 Define object --------------------------------------------------------------------------------
names(merit_per_pref)
# 3.2 Measurement continuous -----------------------------------------------------------------------

# Estimación del modelo simple
cfa_perpref1 <- '
	  # latent variables
	  percmerit =~ hwork + ambition
	  percnmerit=~ wealthy + pareduc +race + gender +people + polcone
	  prefmerit =~ welljob + hardjob
	  prefnmerit=~ family + child
	  '
fit_perpref1 <- cfa(cfa_perpref1, data = merit_per_pref,
                    # missing = "ML",
                    ordered = c("wealthy","pareduc","ambition","hwork","people","polcone","race","gender",
                                "respons","yeduc","family","child","welljob","hardjob"))

names(merit_per_pref)
summary(fit_perpref1,
        fit.measures = TRUE,
        standardized = TRUE)
```

```{r}
plot_m_perpref1 <- '
	  # latent variables
	  percmerit =~ 1*hwork + ambition
	  percnmerit=~ 1*wealthy + pareduc +race + gender +people + polcone
	  prefmerit =~ 1*welljob + hardjob
	  prefnmerit=~ 1*family + child
	  '
plot_perpref1 <- cfa(plot_m_perpref1, data = merit_per_pref,
                    # missing = "ML"
                    )

nodeNames<- c(
  "Importance: hard work",
  "Importance: having ambition",
  "Importance: wealthy family",
  "Importance: educated parents",
  "Importance: race",
  "Importance: gender",
  "Importance: knowing people",
  "Importance: political connections",
  "Reasons for pay: well job",
  "Reasons for pay: hard job",
  "Reasons for pay: support family",
  "Reasons for pay: has children",
  "Perception \n meritocratic",
  "Perception \n non meritocratic",
  "Preference \n meritocratic",
  "Preference \n non meritocratic"
)

xfa <- 1*0.5

# Path model
semPaths(
  plot_perpref1 ,
  # whatLabels = "std", # Indica si aparecen los valores
  what = "mod",
  label.cex = 0.38, # Tamaño de las etiquetas dentro de los nodos.
  edge.label.cex =  0.8, #  0.8 Tamaño de los valores estimados.
  residuals = T,
  optimizeLatRes =T,
  edge.color = "black",
  style = "lisrel",
  nCharNodes = 0,
  curvePivot = F,
  curve = 3,
  rotation = 4,
  layout = "tree3",
  cardinal = "lat cov",
  legend.cex = 0.6, # 0.6
  label.cex = 4,
  label.font = 6,
  edge.label.font = 14,
  asize = 3,
  edge.width = 1.5,
  sizeMan = 6,# largo manifest
  sizeMan2= 3,# alto manifest
  sizeLat = 5, # largo de los circulos
  sizeLat2 = 11.5, # alto de los circulos
  residScale = 10,
  width =6, # 17
  height = 3, # 10
  # nodeNames = nodeNames,
  intercepts = F,
  reorder = T,
  thresholds = F,
  fixedStyle =1,
  node.height = 1,
  node.width = 3,
  label.scale = F,
  shapeMan = "rectangle",
  shapeLat = "ellipse",
  nodeLabels = nodeNames,
  details = T,
  mar = c(1,
          4, #left
          1, #top
          5), #right
  filetype ="pdf",
  filename = "output/images/meas-issp01"
  )->semplot01

```

```{r eval=FALSE, include=FALSE}
# Estimación del modelo c/ unmerit 2do orden
cfa_perpref2 <- '
	  # latent variables
	  merit =~ hwork + ambition
	  parent=~ wealthy + pareduc
	  backg=~ race + gender
		network=~ people + polcone
		unmerit=~ parent + backg + network # 2nd order
	  bmerit =~ welljob + hardjob
	  need=~ family + child
	  '
fit_perpref2 <- cfa(cfa_perpref2, data = merit_per_pref,
                    missing = "ML")
summary(fit_perpref2,
        fit.measures = TRUE,
        standardized = TRUE)

# Path model
semPaths(
  fit_perpref2,
  layout = "tree",
  rotation = 2,
  intercepts = FALSE,
  style = "lisrel",
  curvePivot = TRUE,
  sizeMan = 3.5,
  sizeLat = 6
)

dev.copy(png,
         "./results/path_perpref2.png",
         width = 800,
         height = 650)
dev.off()

```

- Meritocratic perception:
  - How important is having ambition?
  - How important is hard work?

- Non-meritocratic perception
  - How important is coming from a wealthy family?
  - How important is having well-educated parents
  - How important is knowing the right people
  - How important is having political connections
  - How important is a person's race?
  - How important is being born a man or a woman?

- Meritocratic preferences
  - How well he or she does the job?
  - How hard he or she works at the job?

- Non-meritocratic preferences
  - What is needed to support a family?
  - Whether the person has children to support?

```{r}
labs <- c(
  "How important is having ambition?",
  "How important is hard work?",
  "How important is coming from a wealthy family?",
  "How important is having welleducated parents",
  "How important is knowing the right people",
  "How important is having political connections",
  "How important is a person's race?",
  "How important is being born a man or a woman?",
  "How well he or she does the job?",
  "How hard he or she works at the job?",
  "What is needed to support a family?",
  "Whether the person has children to support?"
)

tb.load<- data.frame(round(cbind(inspect(fit_perpref1,what="std")$lambda),digits = 2))
tb.load[tb.load==c(0.00)] <- NA

for (i in names(tb.load)) {
  # tb.load[,i] <- sjlabelled::as_character(tb.load[,i])
  tb.load[,i] <- sprintf(tb.load[,i], fmt = '%#.2f')
}
tb.load[tb.load=="NA"] <- ""

#-------#
fm01<- data.frame(t(data.frame(fitmeasures(fit_perpref1, output ="matrix")[c("chisq","df","cfi","tli","rmsea"),]))); row.names(fm01) ="percnmerit"

#------chi2, df------#
fm04<- round(rbind(fm01),3)
fm04.1 <- fm04 %>% select(chisq,df) 
fm04.1$chisq <- round(x = fm04.1$chisq,digits = 1)
fm04.1$df <- round(x = fm04.1$df,digits = 0)
fm04.1$chi2df <- paste0(fm04.1$chisq,"(",fm04.1$df,")")
fm04.1 <- select(fm04.1,"chi2df")
for (i in names(fm04.1)) {
  fm04.1[,i] <- as.character(fm04.1[,i])
}

#------CFI, RMSEA------#
fm04.2 <- fm04 %>% select(cfi,tli,rmsea) 
for (i in names(fm04.2)) {
  fm04.2[,i] <- sprintf(fm04.2[,i], fmt = '%#.3f')
}

fm.df      <- bind_cols(fm04.1,fm04.2)
fm.df$nobs <- c(nobs(fit_perpref1)) 
fm.df <- data.frame(t(fm.df)); colnames(fm.df) <- c("percnmerit")

#------ merge ------#
tb.fm<- bind_rows(tb.load,fm.df)
tb.fm<- tb.fm %>% mutate(vars=c(labs,"$\\chi^2\\text{(df)}$","$\\text{CFI}$","$\\text{TLI}$","$\\text{RMSEA}$","$N$")) %>% select(vars,everything())
tb.perpref1 <- tb.fm
```

```{r}
save(tb.perpref1,file = "output/tables/tb.perpref1.RData")
```

# Other analysis

```{r}
# 3.3 Measurement categorical -pendiente- #####

# Estimación del modelo simple categórico
fit_1c <- cfa(
  cfa_1,
  data = merit_per_pref,
  ordered = c(
    "wealthy",
    "ambition",
    "hwork",
    "people",
    "polcone",
    "race",
    "gender",
    "pareduc",
    "welljob",
    "hardjob",
    "family",
    "child"
  )
)
summary(fit_1c, fit.measures = TRUE, standardized = TRUE)

# Estimación del modelo 2do orden categórico
fit_2c <- cfa(
  cfa_2,
  data = merit_per_pref,
  ordered = c(
    "wealthy",
    "ambition",
    "hwork",
    "people",
    "polcone",
    "race",
    "gender",
    "pareduc",
    "welljob",
    "hardjob",
    "family",
    "child"
  )
)
summary(fit_2c, fit.measures = TRUE, standardized = TRUE)
```

```{r eval=FALSE, include=FALSE}
# 3.4. Save factor scores --------------- #####

# Factor scores fit_perpref2

# Multiple imputation (para recuperar missings en scores) - REVIEW THIS, command lavPredict from 2017 should predict with missing values therefore saving from multiple imputation

# generate 5 multiple complete datasets
out <- mice(merit_per_pref[, c(3:10, 19:24)], m = 5)
D1 <- complete(out, 1)
D2 <- complete(out, 2)
D3 <- complete(out, 3)
D4 <- complete(out, 4)
D5 <- complete(out, 5)

# fit model for each complete dataset
fit1 <- cfa(cfa_perpref2, data = D1)
fit2 <- cfa(cfa_perpref2, data = D2)
fit3 <- cfa(cfa_perpref2, data = D3)
fit4 <- cfa(cfa_perpref2, data = D4)
fit5 <- cfa(cfa_perpref2, data = D5)

# predict scores for all models
p1 <- predict(fit1)
p2 <- predict(fit2)
p3 <- predict(fit3)
p4 <- predict(fit4)
p5 <- predict(fit5)

# compute 'average' across 5 sets of scores:
scores <- (p1 + p2 + p3 + p4 + p5) / 5

# Generate factor scores
scores_dat = as.data.frame(scores)
dim(scores_dat)
str(scores_dat)
summary(scores)

# Generate object with factor scores
dim(merit_per_pref)
dim(scores_dat)
merit_per_pref_sco = cbind(merit_per_pref, scores)

# Save object with factor scores
names(merit_per_pref_sco)
merit_per_pref_sco = merit_per_pref_sco[, c(1:2, 25:31)]
save(merit_per_pref_sco,
     file = "./objects/merit_per_pref_sco.rda")

# Descriptives factor scores
load("./objects/merit_per_pref_sco.rda")
# Correlations
names(merit_per_pref_sco)
cormat = cor(merit_per_pref_sco[c(3, 7:9)],
             use = "pairwise.complete.obs")
corrplot.mixed(cormat)
#save
dev.copy(png,
         "./results/corplot_scores.png",
         width = 600,
         height = 600)
dev.off()

# Descriptive scatter countries
load("./objects/merit_per_pref_sco.rda")
agg_merit = merit_per_pref_sco %>% group_by(v5) %>% summarise_all(funs(mean))

agg_merit$v5c = countrycode(agg_merit$v5,
                            'iso3n', 'iso3c')

set_label(agg_merit$merit) = "Meritocratic perception"
set_label(agg_merit$bmerit) = "Meritocratic preferences"

sjp.scatter(
  agg_merit$merit,
  agg_merit$bmerit,
  dot.labels = agg_merit$v5c,
  fit.line = TRUE,
  show.ci = TRUE
)

dev.copy(png,
         "./results/mperc_mprefs.png",
         width = 600,
         height = 600)
dev.off()

sjp.scatter(
  agg_merit$merit,
  agg_merit$unmerit,
  dot.labels = agg_merit$v5c,
  fit.line = TRUE,
  show.ci = TRUE
)

dev.copy(png,
         "./results/mperc_munperc.png",
         width = 600,
         height = 600)
dev.off()

sjp.scatter(
  agg_merit$bmerit,
  agg_merit$need,
  dot.labels = agg_merit$v5c,
  fit.line = TRUE,
  show.ci = TRUE
)

dev.copy(png,
         "./results/mperc_munperc.png",
         width = 600,
         height = 600)
dev.off()
```

```{r}
cfa_perpref1 <- '
	  # latent variables
	  percmerit =~ hwork + ambition
	  percnmerit=~ wealthy + pareduc +race + gender +people + polcone
	  prefmerit =~ welljob + hardjob
	  prefnmerit=~ family + child
	  '
fit_perpref1 <- cfa(cfa_perpref1, data = merit_per_pref,
                    # missing = "ML",
                    ordered = c("wealthy","pareduc","ambition","hwork","people","polcone","race","gender",
                                "respons","yeduc","family","child","welljob","hardjob"))
```

```{r eval=FALSE, include=FALSE}
### Check an alternative method
merit_per_pref$v5 <- as.factor(merit_per_pref$v5)
configural        <- cfa(cfa_perpref1, data=merit_per_pref, group = "v5", estimator="ML")
weak.invariance   <- cfa(cfa_perpref1, data=merit_per_pref, group = "v5",group.equal = "loadings", estimator="ML")
strong.invariance <- cfa(cfa_perpref1, data=merit_per_pref, group = "v5",group.equal = c( "loadings", "intercepts"), estimator="ML")
strict.invariance <- cfa(cfa_perpref1, data=merit_per_pref, group = "v5",group.equal = c( "loadings", "intercepts", "residuals"), estimator="ML")
```

```{r eval=FALSE, include=FALSE}
anova(configural,weak.invariance,strong.invariance,strict.invariance)
```

```{r eval=FALSE, include=FALSE}
fitmeasures(configural       )
fitmeasures(weak.invariance  )
fitmeasures(strong.invariance) 
fitmeasures(strict.invariance) 
```

```{r eval=FALSE, include=FALSE}
# 3.5 Invariance full model ---pendiente- --------

fit.indices = c("chisq", "df", "cfi", "tli", "rmsea")

# Continuous first order
fit.conf1 = cfa(cfa_1,
                data = merit_per_pref,
                group = "v5",
                std.lv = TRUE)

fitMeasures(fit.conf1, fit.indices)
summary(fit.conf1)
inspect(fit.conf1, "theta")
inspect(fit.conf1, "cov.lv")

lavaan::
partialInvariance(fit_perpref1, type = "strict", p.adjust = "holm")
```

### 2021-05-25 index.Rmd
- major update

--- 2020-01-10 index.Rmd
@@
 ---
-title: "Perceptions and Preferences for Meritocracy Scale (PPM-S)"
+title: "PPM-S"
 csl: "input/bib/apa6.csl"
 bibliography: "input/bib/zlibrary.bib"
-css: tyle.css
+author: " "
 output:
-  pdf_document:
-    keep_tex: false
-    template: null
-    number_sections: true
-    toc: true
-  html_document:
-    theme: flatly
-    number_sections: true
-    toc: true
-    toc_float: true
-geometry: margin=0.78in
-header-includes:
-  # - \usepackage[spanish,es-tabla,es-nodecimaldot]{babel}
-  - \usepackage{caption}
-  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
-  - \usepackage{graphicx}
-  - \usepackage{float}
-  - \usepackage{booktabs}
-  - \usepackage{longtable}
-  - \usepackage{array}
-  - \usepackage{multirow}
-  - \usepackage{wrapfig}
-  - \usepackage{float}
-  - \usepackage{colortbl}
-  - \usepackage{pdflscape}
-  - \usepackage{tabu}
-  - \usepackage{threeparttable}
-linestretch: '1.5'
-link-citations: yes
-linkcolor: blue
+    bookdown::html_document2:
+            smart: false
+            number_sections: false
 ---
 
-```{r eval=FALSE, echo=FALSE}
-# Para obtener doble output (html/pdf) correr directamente en consola/terminal:
-rmarkdown::render('index.rmd', output_format = 'all')
-```
+Last update: `r format(Sys.time(), "%A %d %B %Y %X")`
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
-
-```{r include=FALSE}
-#Determine the output format of the document
-outputFormat   = knitr::opts_knit$get("rmarkdown.pandoc.to")
-
-#Figure and Table Caption Numbering, for HTML do it manually
-capTabNo = 1; capFigNo = 1;
-
-#Function to add the Table Number
-capTab = function(x){
-  if(outputFormat == 'html'){
-    x = paste0("Table ",capTabNo,". ",x)
-    capTabNo <<- capTabNo + 1
-  }; x
-}
-
-#Function to add the Figure Number
-capFig = function(x){
-  if(outputFormat == 'html'){
-    x = paste0("**","Figura ",capFigNo,": ","**",x)
-    capFigNo <<- capFigNo + 1
-  }; x
-}
-```
-
-# Research questions
-
-Is it possible to identify and measure perceptions and preferences for meritocracy as two distinct but related factors?
-
-# General description and hypotheses
-
-This research builds upon previous studies attempting to develop a comprehensive approach to measuring meritocratic beliefs [@Duru-bellat2012; @Kunovich2007; @Newman2015; @Reynolds2014].
-The proposed measurement model is based on two axis of analysis, as depicted in Table 1.
-The first one distinguishes between different types of "beliefs", using instead the terms perceptions and preferences for meritocracy.
-Perceptions refers to the extent to which people see meritocracy working in their society, which in terms of measurement relates to the "reasons to get ahead" battery, whereas preferences refer to normative expectations that are usually linked to a "should" expression (e.g.
-whether hard work should be related to payment).
-The second axis consider the distinction between meritocratic and non-meritocratic dimensions.
-This aspect has been usually treated as different ends of a same continuum in part of the previous research, an assumption that requieres empirical scrutiny.
-These non-meritocratic elements usually refer to the use of personal contacts or family advantages to get ahead in life.
-
-Table 1: Model of perceptions and preferences for meritocracy and non-meritocracy
-
-|                  	  | Perceptions 	| Preferences 	|
-|:-----------------	  |-------------	|-------------	|
-| _Meritocracy_     	|             	|             	|
-| _Non-meritocracy_ 	|             	|             	|
-
-**Hypotheses**
-
-* H1. The perception of meritocracy is a latent variable based on indicators of the importance attributed to talent and the effort to get ahead in life.
-* H2. The non-meritocratic perception is a latent variable that derives from two indicators related to the agreement with the statement that people with contacts and rich parents manage to get ahead.
-* H3. Meritocratic preferences behave as the latent variable based on a normative value of effort and talent.
-* H4. Non-meritocratic preferences behave as a latent variable based on the normative value of the use of personal contacts and having rich parents.
-
-# Type of study
-
-The scale will be included in an online three-wave panel survey study.
-This study covers among other things an experiment of the effect of information about poverty and inequality on opportunity beliefs.
-The scale will be presented in wave 1 and wave 3.
-
-# Sample
-
-## Detailed data and explanations of its use.
-
-The data will come from a representative sample of the large cities of Chile (Gran Santiago, Valparaíso-Viña del Mar, Concepción and Antofagasta).
-In order to ensure representativeness, quotas for sex, education and age will be provided to the agency in charge of the online fieldwork.
-
-## Magnitude of the sample and explanation of the magnitude.
-
-The sample will have an initial number for the first wave of 2100 cases, which are expected to be at least 1800 in the second wave.
-From the second to the third wave it has been considered as “free fall”, from which we expect to retain about 1500 cases.
-
-The sample size was calculated from power analysis that used information from a pilot study that was applied in June 2019.
-This pilot study consisted of an online experiment applied to 949 subjects, randomly separating participants to two stimuli and a placebo, leaving a third of the sample in each condition.
-From these data, the low power of a two-tailed test was calculated, considering an average treatment effect of 0.26, with a standard deviation of the result of the treatment group of 1.23 and a significance level of 0.05.
-Based on multiple comparison correlations (Bonferroni), we adjusted alpha for six tests (three results * two comparisons between conditions and placebo).
-The results indicated that the number of subjects exposed to each stimulus should be 580, so we will work with three samples of 600 cases, which gives us a total of 1800 cases for the second wave in which the experiment will be applied.
-To guarantee the above and in consideration of a high non-response rate, we will start with a sample of 2100 subjects.
-
-### Data collection procedure:
-
-Respondents will be invited to participate in the online survey under the incentive provided by the external company.
-Respondents have 4 business days to answer the survey, with a maximum of 20 minutes per survey.
-In order to achieve greater representativeness of the sample, the quota method will be used, that is, the program will only allow respondents with the required demographic characteristics.
-The quotas used were age, sex and education.
-It should be noted that the time between wave and wave is around 7 and 9 days.
-
-### Stop Rule:
-
-The rule of detention is to reach the number of respondents indicated in the quotas. However, there may be modifications depending on the contingencies of the field work and the response rates.
-
-# Variables to use:
-
-All the variables used in the scale use the following question and answer options:
-
-* To what extent do you agree or disagree with each of the following affirmations?
-  - Response scale:
-    + Strongly disagree
-    + Disagree
-    + Neither agree nor disagree
-    + Agree
-    + Totally agree
-
-```{r include=FALSE}
-# Get output document type and use as the format argument in kable function
-library(knitr)
-library(kableExtra)
-
-table_format = if(is_html_output()) {
-  "html"
-} else if(is_latex_output()) {
-  "latex"
-}
-```
-
-```{r echo=FALSE}
-
-tab01 <- read.csv(file = "input/data/original/materials/table01.csv",header = 1,sep = ";")
-
-cnames <- c("Dimension", "Factor","Statement (english)" , "Statement (spanish)")
-
-caption01 <- capTab("Meritocracy perception and preference")
-kable(tab01,format = table_format,booktabs = T, linesep = "",col.names = cnames, caption = caption01) %>%
-  kable_styling(full_width = T,latex_options = c("hold_position","scale_down"),position = "center",font_size = 9) %>%
-  column_spec(column = 1,width = "1.5cm",) %>%
-  column_spec(column = 2,width = "2 cm") %>%
-  collapse_rows(columns = 1:2,valign = "middle")
+```{r, child="readme.md", eval=TRUE}
 
 ```
-
-### Variable manipulation
-
-There will not be variable manipulation.
-
-### Indexes
-
-The investigation of the validity of this scale does not rely on indices but rather confirmatory factor analysis will be used to estimate the value of the latent variables underlying the indicators.
-
-### Blinding of information
-
-No special blinding is used for this study.
-
-### Randomization
-
-To test the items' order effects we present the items in three different order conditions which will be assigned randomly. The conditions are the following:
-
-- Order 1: the order in which items are presented in Table 1
-- Order 2: the items are ordered based on topics (work harder, talent, rich parents, connections), and for each of them they have to respond first about perceptions and then about preferences.
-For instance: a) those who work harder get greater rewards, b) those who work arder should get greater rewards ...
-- Order 3: the order is randomized within this condition
-
-# Analysis plan
-
-To evaluate the hypotheses, confirmatory factor analysis (CFA) will be used as we rely on a theory regarding the underlying four factors.
-The method of estimation will be Weighted Least Squares Mean Variance Approximation (WLMSV) for categorical ordered indicators.
-For the analysis we will use the R package Lavaan [@Rosseel2012].
-
-### Inference criteria
-
-The values that will be used as evaluation criteria for the goodness of the adjustment of the model were taken from the proposal of @Brown2008 and are the following:
-
-* Chi-square:> 0.05
-* Chi-square ratio:> 3
-* Comparative adjustment goodness index (CFI):> 0.95
-* Tucker-Lewis Index (TLI):> 0.95
-* Root of the average squared residual approximation <0.08.
-
-### Secondary analysis
-
-To evaluate the metric stability of the measurement model [@Davidovetal2014] we will conduct a longitudinal invariance test using the first and third wave of our study.
-Following @liuTestingMeasurementInvariance2017, we will test a series of four hierarchical models: Configural, Weak, Strong and Strict invariance models for ordered-categorical indicators based on the assumption that a five category likert scale cannot be treated as a continous variable because can lead to biased parameter estimates.
-
-### Data Exclusion
-
-All cases will be used as long as they do not show missing values in any of the items on the scale. No imputation criteria will be used.
-
-### Ethics.
-
-The experiment and survey instruments are approved by the IRB of the University of Chile.
-
-# Bibliography

### 2021-05-25 papelera/docs/prod_analysis-cfa.rmd
- major update

--- 2021-01-29 production/prod_analysis-cfa.rmd
@@
 
-```{r tabla01, eval=TRUE, include=TRUE}
-polyw01<- poly(sjlabelled::remove_all_labels(dat01),
-               digits = 2,
-               triangle = "l",
-               fade.ns = F,
-               remove.spaces = T,
-             string.diag = c("$-$","$-$","$-$","$-$","$-$","$-$","$-$","$-$"))
-matpolyw01 <- as.data.frame(read_html(x = polyw01[["page.content"]]) %>% html_table(fill=FALSE)) 
-matpolyw01 <- matpolyw01[-c(9),]
-colnames(matpolyw01) <- c("vars","(A)","(B)","(C)","(D)","(E)","(F)","(G)","(H)")
-
-load01<- data.frame(round(cbind(inspect(fit1_c,what="std")$lambda,inspect(fit1_o,what="std")$lambda),digits = 2))
-load01[load01==c(0.00)] <- NA; names(load01) <- c(1:4,1:4)
-
-table01<- bind_cols(x = matpolyw01,y = load01)
-### %>% mutate(vars=labs)
-
-kable(table01,escape = TRUE,align = "l",col.names = cnametab,format = "html") %>% 
-  footnote(general = paste0(paste0("$N$= ",nobs(fit1_c)),"; p<0.05=$*$; p<0.01=$**$; p<0.001=$***$"),footnote_as_chunk = T) %>%
-  add_header_above(c(" " = 1,"Polychoric correlations"=8, "MLR" = 4, "DWLS" = 4))%>%
-  add_header_above(c(" " = 9,"Factor loadings"=8)) %>%
-  # # column_spec(column = c(1),width ="7in",include_thead = T) %>%
-  # # column_spec(column = c(2:17),width ="0.5in",include_thead = T) %>% 
-  kable_styling(full_width = FALSE)
-```
-
 ## Version 02  (N=`r dim(dat02)[1]`)
@@
 
-```{r tabla02, echo=FALSE, include=TRUE}
-polyw02<- poly(sjlabelled::remove_all_labels(dat02),
-               digits = 2,
-               triangle = "l",
-               fade.ns = F,
-               remove.spaces = T,
-               string.diag = c("$-$","$-$","$-$","$-$","$-$","$-$","$-$","$-$"))
-matpolyw02 <- as.data.frame(read_html(x = polyw02[["page.content"]]) %>% html_table(fill=FALSE)) 
-matpolyw02 <- matpolyw02[-c(9),] 
-colnames(matpolyw02) <- c("vars","(A)","(B)","(C)","(D)","(E)","(F)","(G)", "(H)")
-
-load02<- data.frame(round(cbind(inspect(fit2_c,what="std")$lambda,inspect(fit2_o,what="std")$lambda),digits = 2))
-load02[load02==c(0.00)] <- NA; names(load02) <- c(1:4,1:4)
-
-table02<- bind_cols(x = matpolyw02,y = load02)
-## %>% mutate(vars=labs)
-
-kable(table02,escape = FALSE,align = "l",col.names = cnametab)  %>% 
-  footnote(general = paste0(paste0("$N$= ",nobs(fit2_c)),"; p<0.05=$*$; p<0.01=$**$; p<0.001=$***$"),footnote_as_chunk = T) %>%
-  add_header_above(c(" " = 1,"Polychoric correlations"=8, "MLR" = 4, "DWLS" = 4))%>% 
-  add_header_above(c(" " = 9,"Factor loadings"=8)) %>% 
-  # column_spec(column = c(1),width ="7in",include_thead = T) %>%
-  # column_spec(column = c(2:17),width ="0.5in",include_thead = T) %>% 
-  kable_styling(full_width = FALSE) %>% 
-  kable_styling(full_width = F)
-```
-
 ## Version 03 (N=`r dim(dat03)[1]`)
@@
 
-```{r}
-polyw03<- poly(sjlabelled::remove_all_labels(dat03),
-               digits = 2,
-               triangle = "l",
-               fade.ns = F,
-               remove.spaces = T,
-               string.diag = c("$-$","$-$","$-$","$-$","$-$","$-$","$-$","$-$"))
-matpolyw03 <- as.data.frame(read_html(x = polyw03[["page.content"]]) %>% html_table(fill=FALSE)) 
-matpolyw03 <- matpolyw03[-c(9),]
-
-colnames(matpolyw03) <- c("vars","(A)","(B)","(C)","(D)","(E)","(F)","(G)", "(H)")
-
-load03<- data.frame(round(cbind(inspect(fit3_c,what="std")$lambda,inspect(fit3_o,what="std")$lambda),digits = 2))
-load03[load03==c(0.00)] <- NA; names(load03) <- c(1:4,1:4)
-
-table03<- bind_cols(x = matpolyw03,y = load03)
-   ## %>% mutate(vars=labs)
-
-kable(table03,escape = FALSE,align = "l",col.names = cnametab) %>% 
-  footnote(general = paste0(paste0("$N$= ",nobs(fit3_c)),"; p<0.05=$*$; p<0.01=$**$; p<0.001=$***$"),footnote_as_chunk = T) %>%
-  add_header_above(c(" " = 1,"Polychoric correlations"=8, "MLR" = 4, "DWLS" = 4))%>% 
-  add_header_above(c(" " = 9,"Factor loadings"=8)) %>% 
-  # column_spec(column = c(1),width ="7in",include_thead = T) %>%
-  # column_spec(column = c(2:17),width ="0.5in",include_thead = T) %>% 
-  kable_styling(full_width = F)
-```
-
 ### Tau-equivalence
@@
 > **NOTE:** The model with equal factor loadings (in Meritocracy Perception) does not significantly deteriorate the model fit.
-
-```{r tabla03, include=FALSE}
-polyw03tau<- poly(sjlabelled::remove_all_labels(dat03),
-               digits = 2,
-               triangle = "l",
-               fade.ns = F,
-               remove.spaces = T,
-               string.diag = c("$-$","$-$","$-$","$-$","$-$","$-$","$-$","$-$"))
-matpolyw03tau <- as.data.frame(read_html(x = polyw03tau[["page.content"]]) %>% html_table(fill=FALSE)) 
-matpolyw03tau <- matpolyw03tau[-c(9),]
-
-colnames(matpolyw03tau) <- c("vars","(A)","(B)","(C)","(D)","(E)","(F)","(G)", "(H)") 
-load03tau<- data.frame(round(cbind(inspect(fit3_c.tau,what="std")$lambda,inspect(fit3_o.tau,what="std")$lambda),digits = 2))
-load03tau[load03tau==c(0.00)] <- NA; names(load03tau) <- c(1:4,1:4)
-
-table03tau<- bind_cols(x = matpolyw03tau,y = load03tau) 
-### %>% mutate(vars=labs)
-
-kable(table03tau,escape = FALSE,align = "l",col.names = cnametab) %>% 
-  footnote(general = paste0(paste0("$N$= ",nobs(fit3_c.tau)),"; p<0.05=$*$; p<0.01=$**$; p<0.001=$***$"),footnote_as_chunk = T) %>%
-  add_header_above(c(" " = 1,"Polychoric correlations"=8, "MLR" = 4, "DWLS" = 4))%>% 
-  add_header_above(c(" " = 9,"Factor loadings"=8)) %>% 
-  # column_spec(column = c(1),width ="7in",include_thead = T) %>%
-  # column_spec(column = c(2:17),width ="0.5in",include_thead = T) %>% 
-  kable_styling(full_width = F)
-```
 
@@
 
-```{r tabla04, include=FALSE}
-polyw04<- poly(sjlabelled::remove_all_labels(dat04[,-c(1)]),
-               digits = 2,
-               triangle = "l",
-               fade.ns = F,
-               remove.spaces = T,
-               string.diag = c("$-$","$-$","$-$","$-$","$-$","$-$","$-$","$-$"))
-matpolyw04 <- as.data.frame(read_html(x = polyw04[["page.content"]]) %>% html_table(fill=FALSE)) 
-matpolyw04 <- matpolyw04[-c(9),]
-
-colnames(matpolyw04) <- c("vars","(A)","(B)","(C)","(D)","(E)","(F)","(G)", "(H)") 
-load04<- data.frame(round(cbind(inspect(fit4_c,what="std")$lambda,inspect(fit4_o,what="std")$lambda),digits = 2))
-load04[load04==c(0.00)] <- NA; names(load04) <- c(1:4,1:4)
-
-table04<- bind_cols(x = matpolyw04,y = load04)  
-  # %>% mutate(vars=labs)
-
-kable(table04,escape = FALSE,align = "l",col.names = cnametab) %>% 
-  footnote(general = paste0(paste0("$N$= ",nobs(fit4_c)),"; p<0.05=$*$; p<0.01=$**$; p<0.001=$***$"),footnote_as_chunk = T) %>%
-  add_header_above(c(" " = 1,"Polychoric correlations"=8, "MLR" = 4, "DWLS" = 4))%>% 
-  add_header_above(c(" " = 9,"Factor loadings"=8)) %>% 
-  # column_spec(column = c(1),width ="7in",include_thead = T) %>%
-  # column_spec(column = c(2:17),width ="0.5in",include_thead = T) %>% 
-  kable_styling(full_width = F)
-```
-
 # Tables
-
-```{r tb_desc-corr}
-load04<- data.frame(round(inspect(fit4_o,what="std")$lambda,digits = 2))
-load04[load04==c(0.00)] <- NA; names(load04) <- c(1:4)
-
-colnames(matpolyw04) <- c("Variable","A","B","C","D","E","F","G", "H")
-
-sum04b  <- psych::describe(dat04[,-c(1)], fast = T) %>% select(mean,sd,min,max) %>% data.frame() %>% round(digits = 2)
-table04b<- bind_cols(x = sum04b,y = matpolyw04) %>% select(Variable, everything()) %>% mutate(Variable=labs) %>% rename("Mean"=mean, "SD"=sd, "Min"=min,"Max"=max)
-
-cap01 <- "Descriptive statistic"
-sumdat04<- kable(table04b, escape = FALSE,align = "l",linesep = "", caption = cap01) %>% 
-           # column_spec(column = 1,width = "10cm",width_min = "12cm",include_thead = TRUE) %>% 
-           footnote(general = paste0(paste0("$N$= ","2141"),"; p<0.05=$*$; p<0.01=$**$; p<0.001=$***$"),footnote_as_chunk = T) %>% 
-           add_header_above(header = c(" "=5,"Polychoric correlations"= 8)) %>% 
-           kable_styling(full_width = TRUE,font_size = 8,position = "center");sumdat04
-save(sumdat04,file = "output/tables/sumdat04.RData")
-```
 

### 2021-05-25 papelera/processing/.prod_analysis-long_invariance.rmd
- major update

--- 2020-04-11 production/prod_analysis-long_invariance.rmd
@@
 rmarkdown::render(input = "production/prod_analysis-long_invariance.Rmd",output_format = "html_document"); browseURL(url = "production/prod_analysis-long_invariance.html")
+```
+
+```{r include=FALSE}
+knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = TRUE)
+knitr::opts_knit$set(root.dir = "../")
+options(scipen=999)
+rm(list=ls())
+options(knitr.kable.NA = '')
 ```
@@
 load(file = "input/data/proc/datawide_123.RData") # Datos formato wide 3 olas
+load(file = "input/data/proc/dat04.RData") # Datos formato wide 3 olas
 

### 2021-05-25 papelera/processing/analysis_study2.Rmd
- major update

--- 2021-05-25 papelera/docs/prod_analysis-cfa.rmd
@@
 ---
-title: "Analysis PPM-S"
+title: "Analysis Merit Scale - Study 2"
 date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
@@
 
-# Wave 01 {.tabset .tabset-fade .tabset-pills}
-
-On the first wave of the study we conducted the application of the Perceptions and Preferences Meritocracy Scale (PPM-S), in which we decided to test three different modalities for the application of the scale.
-The design of the study considered a sample of 2200 cases in Chile, using quotas for sex, age and educational level.
-The most important feature of this study consists in that we randomly assigned the participants to three different groups, in which the subjects complete one of the three combination of the questions.
-The key feature of these combinations relies on the display order of the questions regarding to each indicator of the scale.
-On this respect, we have three subsamples that correspond to each application modality that can be analysed separately, but also all together as a full sample.
-
-The scale has eight indicators, divided in four questions for perception and four for preferences, which are also divided in meritocratic and non-meritocratic.
-The meritocratic indicators refers to topics such as effort and talent, and the non-meritocratic refers to family wealth and networks.
-The three combinations relies in how these indicators are displayed to the respondent.
-The first combination corresponds to the group that answered the questions that were sorted by the Perceptions/Preferences dimension.
-the second combination corresponds to the group in which the questions are displayed following the a topic-sorted way, this means that the effort, talent, rich family and networks questions for perception and preference were displayed together.
-The last and third group corresponds to a completely randomized display order of the questions.
-
-The rationale of the design relies on the possibility to test how different display modalities could affect the hypothesized factor structure of the scale.
-On this regard, we conducted the analysis on two stages.
-On the first stage we tested the hypothesized factor structure on the three groups corresponding to each application modality and the full sample using a Confirmatory Factor Analysis (CFA).
-On the second stage, we conducted a Multi-group confirmatory factor analysis (MGCFA) using the groups created by the randomization.
-Furthermore, we explore the possibility of a second order factor structure, in which we tested two models.
-The first alternative corresponds to a second-order Confirmatory Factor Analysis assuming the Perception/Preference dimension as a second order factor.
-On the other hand, we tested a model in which we assume that the second order factor structure follows the Meritocratic/Non-meritocratic constructs.
-
----
-
-```{r eval=FALSE, include=FALSE}
-rmarkdown::render(input = "production/prod_analysis-cfa.rmd",output_format = "html_document"); browseURL(url = "production/prod_analysis-cfa.html")
-```
-
 ```{r include=FALSE}
+#knitr general options
+
 knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = TRUE)
@@
 
-**Libraries**
-```{r}
-library(sjPlot)
-library(dplyr)
-library(lavaan)
-library(semPlot)
-library(stargazer)
-library(corrplot)
-library(psych)
-library(knitr)
-library(kableExtra)
-library(rvest)
-library(sjlabelled)
-library(ggplot2)
-source('production/prod_polychoric.R')
-```
-
-**Load data**
-
-```{r}
-load(file = "input/data/proc/dat01.RData")
-load(file = "input/data/proc/dat02.RData")
-load(file = "input/data/proc/dat03.RData")
-load(file = "input/data/proc/dat04.RData")
-```
-
-```{r}
+```{r}
+# this in for use in case it is necessary to produce different tables for pdf and is_html_output
 table_format = if(is_html_output()) {
@@
 }
-
+```
+
+This is the code for generating tables and graphs based on the analysis of the of the Perceptions and Preferences for Meritocracy Scale.
+
+# Libraries
+
+```{r}
+pacman::p_load(sjPlot, dplyr, lavaan, semPlot, stargazer, corrplot, psych, knitr, kableExtra, rvest, sjlabelled, ggplot2)
+source('production/prod_polychoric.R') # for polychoric correlation analysis
+```
+
+# Data
+
+```{r}
+load(file = "input/data/proc/data_s2.RData")
 ```

### 2021-05-25 papelera/processing/merit_per_pref.Rmd
- major update

```{r }
# Meritocracy/non-meritocracy perceptions and preferences ISSP

# Juan Carlos Castillo, June 2017 - jc-castillo.com

setwd("/media/ntfs/Dropbox/proyectos y papers compartidos/Meritocracy and preferences/meritocracy_pref_international/loc_merit_meassure") # jc

# 0.DATA & LIBRARIES ---------------------- -

# 0.1 Libraries ------------------------- ####

  # Libraries
    pacman::p_load(sjPlot,
     sjmisc, # several functions,ej read_stata
     descr,
     memisc,  # codebook
     psych, # cor2latex
     corrplot, # graph correlations
     lavaan,
     stargazer,
     semTools,
     plyr, # count
     dplyr,
     car, # recode
     ggplot2,
		 Hmisc,
		 tidyverse,
		 data.table,
  	 stringr,
  	 countrycode,
  	 lme4,
     texreg,
     mice,
     ggrepel, # for scatter labels in sjp.scatter
     semPlot,
		 haven,
		 sjlabelled)

# 0.2 Original data preparation --------- ####

    #issp2009=read_stata("https://www.dropbox.com/s/gkq8lue7nnl9joy/ZA5400_v4-0-0.dta") # takes a while but it works
    issp2009=read_dta("../data/issp2009/ZA5400_v4-0-0.dta")
    names(issp2009) <- tolower(names(issp2009)) # change to small

  # Check data
    names(issp2009)

  # Check & generate IDs (v4: respondent, v5: country)
    tapply(issp2009$v4, issp2009$v5, mean, na.rm=TRUE)
    # Algunos paises (ej:826) tienen ids raros ...

  # Generar nueva id secuencial individuos por país para merges
    issp2009$id <- sequence(tabulate(issp2009$v5))
    tapply(issp2009$id, issp2009$v5, mean, na.rm=TRUE) # ok
    summary(issp2009$id)
    issp2009$idtem=issp2009$v5*10000
    issp2009$id=issp2009$idtem + issp2009$id

# End sec 0 -----------------------------

# 1. Measuring perceptions (getting-ahead) --

# 1.1 Data preparation, gen getahead.rda  #####

  names(issp2009)
  get_label(issp2009[8:18])

  #  v6: wealthy family (c 8)
  #  v7: educated parents (c 9)
  #  v9: ambition c11
  # v10: hard work c12
  # v11: right people
  # v12: political connections
  # v14: race c16
  # v16: gender c18
  # IDs: v5 (country) & id (individual id)

  # Recodes & relabels

    # Rename
    names(issp2009)
    getahead=issp2009[,c("v6","v7","v9", "v10","v11","v12",
             "v14","v16", # get ahead
    	  		 "v5","id")]

    getahead$wealthy=getahead$v6
    getahead$pareduc=getahead$v7
    getahead$ambition= getahead$v9
    getahead$hwork=getahead$v10
    getahead$people=getahead$v11
    getahead$polcone=getahead$v12
    getahead$race=getahead$v14
    getahead$gender=getahead$v16

    names(getahead)
    getahead=getahead[-c(1:8)]

  # recode values 8 9 to NA
    freq(getahead$wealthy)
    getahead[3:10][getahead[3:10] > 5] <- NA
    freq(getahead$gender) # check ok
    summary(getahead)

  # adjust labels
    get_labels(getahead$wealthy)
    set_na(getahead[ , 3:10]) <- c(8, 9)
    get_labels(getahead[3:10]) # check

  # Dataset with original code for sjp.likert graph
    getahead2=getahead

  # Reverse code for SEM models
    getahead[3:10]=6-getahead[3:10]

  # Generate variables with 3 values for getahead mess model categoric
    names(getahead)
    getahead_temp=subset(getahead[c(2,3:10)])
    names(getahead_temp)
    getahead_temp[2:9][getahead_temp[2:9] <  4] <- 3
    freq(getahead_temp$hwork)
    freq(getahead$hwork)
    colnames(getahead_temp)=paste("c3", colnames(getahead_temp),
    	sep="")
    names(getahead_temp)[names(getahead_temp)=="c3id"] <- "id"
    getahead_temp2=merge(getahead,getahead_temp, by=c("id"))
  	names(getahead_temp2) # ok
  	getahead=getahead_temp2

	# Save object
    save(getahead, file="./objects/getahead.rda")

# 1.2 Descriptives ---------------------- #####

    load("./objects/getahead.rda")

  # Summary table
    names(getahead)
    stargazer(getahead[,-c(1,2,11:18)], type = "text")

  # Summary table to latex
    sink("./results/desctab_ga.txt")
    stargazer(getahead[,-c(1,2,9:14)]);  sink()

  # Simple freq graph
    sjp.frq(getahead2$wealthy, show.values = TRUE,
    	coord.flip = TRUE)

  # Likert graph
    sjp.likert((getahead2[,-c(1,2)])) # ajustar

    # short var labels
      items <- c("Wealthy family", "Educated parents",
                 "Ambition", "Hard work", "Race",
                 "Gender","Know people", "Pol. connect." )

    # cambiar orden para la leyenda
      labels= c("Essential", "Very important",
        "Not very important",
        "Not important at all", "(neutral)")

    # ajustes globales sjplot

      sjp.setTheme(theme = "blank",legend.inside = TRUE,
        legend.pos = "bottom right", legend.size = 0.8,
        geom.outline.size=0, geom.label.size =3)

    # Gráfico final

      sjp.likert((getahead2[,-c(1,2)]),
        axis.labels   = items,
        legend.labels = labels,
        cat.neutral   = 3, # identifica a indiferentes
        geom.colors   = c("#9ecae1", "#6baed6","#4292c6",
                       "#2171b5"), # colorbrewer2.org para temas
        sort.frq      = "neg.asc", # sort descending)
        title         = "How important is for getting ahead ...",
        intercept.line.color = "white", # middle line white
        expand.grid   = FALSE, # no inner margins in plot
        show.n        = FALSE,    # hide N's in axis labels
        grid.range    = 1.3) # expand axis from 130% to 130%

      # Save
      dev.copy(png,"./results/likert_ga.png"); dev.off()

  # Correlations

    # Matrices & plot

      cormat_ga=cor(getahead[c(5,6,3,4,7:10)],
      	use="pairwise.complete.obs")
      corrplot.mixed(cormat_ga)
      dev.copy(png,"./results/corplot_ga.png", width=600, height=600); dev.off()

    # Correlation table a latex
         sink("./results/cortab_ga.txt")
      cor2latex(getahead[2:10],
            stars   = TRUE, adjust = "none",
            apa     = TRUE,heading = " ",
            caption ="Reasons to get ahead"); sink()

# 1.3 Measurement models--- ------------- #####

		load("../objects/getahead.rda")
    names(getahead)

	# Estimación del modelo de 1er order
	  cfa_1 <- '
	  # latent variables
	  merit =~ hwork + ambition
	  parent=~ wealthy + pareduc
	  backg=~ race + gender
		network=~ people + polcone
	  '
	  fit_1 <- cfa(cfa_1,data=getahead, missing = "ML")
	  summary(fit_1, fit.measures=TRUE,standardized=TRUE)

	  # Path model
	  semPaths(fit_1,layout = "tree2", rotation = 2,
	  	intercepts = FALSE, style = "lisrel", curvePivot = TRUE,
	  	cardinal = TRUE)
    dev.copy(png,"./results/path_cfa1.png",width=750,height=600)
    dev.off()

		# Estimación categorical 3 levels (5 levels convergence prob.)

	    cfa_1c <- '
		  # latent variables
		  merit=~ c3hwork + c3ambition
		  parent=~ c3wealthy + c3pareduc
		  backg=~ c3race + c3gender
			network=~ people + polcone
		  '
		  fit_1c <- cfa(cfa_1c,data=getahead,ordered = c("c3hwork",
		  	"c3ambition", "c3wealthy", "c3pareduc", "c3race",
		  	"c3gender",	"c3people", "c3polcone"))
		  summary(fit_1c, fit.measures=TRUE,standardized=TRUE) # works
		  mod=modificationindices(fit_1c)
		  subset(mod, mi>3.84, select=lhs:mi)

  # Estimación del modelo 2nd order

	  cfa_2 <- '
	  # latent variables
	  merit =~ hwork + ambition
	  parent=~ wealthy + pareduc
	  backg=~ race + gender
		network=~ people + polcone
		unmerit =~parent + backg + network
	  '
	  fit_2 <- cfa(cfa_2,data=getahead, missing = "ML")
	  summary(fit_2, fit.measures=TRUE,standardized=TRUE)

	  # Path
		  semPaths(fit_2,layout = "tree", rotation = 2,
		  	intercepts = FALSE,	style = "lisrel", curvePivot = TRUE,
		  	cardinal = FALSE)
		  dev.copy(png,"./results/path_cfa2.png",width=750,height=600)
	    dev.off()

# End sec 1 -----------------------------

# 2. MEASURING BELIEFS (detearn) ---------- -

# 2.1 Data preparation, detearn.rda ----- #####

		names(issp2009)
	  detearn=issp2009[,c("v47", "v48", "v49", "v50", "v51",
	   	"v52", "v5","id")]

    detearn$respons 	= detearn$v47
    detearn$yeduc  	  = detearn$v48
    detearn$family    = detearn$v49
    detearn$child     = detearn$v50
    detearn$welljob   = detearn$v51
    detearn$hardjob   = detearn$v52

	  names(detearn)
    detearn=detearn[-c(1:6)]

  # recode values 8 9 to NA
    summary(detearn)
    freq(detearn$respons)
    detearn[3:8][detearn[3:8] > 5] <- NA
    freq(detearn$respons) # check ok

  # Out Portugal (no answer in well job item)
    detearn[which(detearn$v5==620),]=NA

  # adjust labels ###PROB PORTUCAL!!!
    get_labels(detearn)
    set_na(detearn[ , 3:8]) <- c(8, 9)
    set_na(detearn$welljob)=c(1)
    summary(detearn)
    freq(detearn$welljob) ## prob portugal
    detearn$welljob[detearn$welljob ==0] = NA
    get_labels(detearn$welljob)
    labels(detearn$welljob)=labels(detearn$respons)### PROB NOT SOLVED

  # Dataset with original code for sjp.likert graph
    detearn2=detearn

  # Reverse code for SEM models
    detearn[3:8]=6-detearn[3:8]
    freq(detearn$respons)

	# Save object
    save(detearn, file="detearn.rda")

# 2.2 Descriptives ---------------------- ######

	load("objects/detearn.rda")

  # Summary table
    names(detearn)
    stargazer(detearn[,-c(1,2)], type = "text")

  # Summary table to latex
    stargazer(detearn[,-c(1,2)])

  # Simple freq graph
    sjp.frq(detearn2$respons, show.values = TRUE, coord.flip = TRUE)

  # Likert graph (uses detearn2 object, with original coding)

    sjp.likert((detearn2[,-c(1,2)])) # ajustar

    # short var labels
      items <- c("Responsibility", "Years education",
                 "Family", "Children", "Well job",
                 "Hard job")

    # cambiar orden para la leyenda
      labels= c("Essential", "Very important","(neutral)"
                ,"Not very important","Not important at all")

    # ajustes globales sjplot
      sjp.setTheme(theme = "blank",legend.inside = TRUE,
        legend.pos = "bottom right", legend.size = 0.8,
        geom.outline.size=0, geom.label.size =3)

      sjp.likert((detearn2[,-c(1,2)]),
        axis.labels   = items,
        legend.labels = labels,
        cat.neutral   = 3, # identifica a indiferentes
        geom.colors   = c("#9ecae1", "#6baed6","#4292c6",
                       "#2171b5"), # colorbrewer2.org para temas
        sort.frq      = "neg.asc", # sort descending)
        title         = "In deciding how much people ought to earn, how important should each of these things be...",
        intercept.line.color = "white", # middle line white
        expand.grid   = FALSE, # no inner margins in plot
        show.n        = FALSE,    # hide N's in axis labels
        grid.range    = 1.3) # expand axis from 130% to 130%

      # Save
      	dev.copy(png,"./results/likert_detearn.png"); dev.off()

      # freq(detearn2$pareduc)   # check educated parents, ok

  # Correlations
    cormat=cor(detearn[5:8], use="pairwise.complete.obs")
    corrplot.mixed(cormat)

    #save
     dev.copy(png,"./results/corplot_detearn.png",width=600,
      	height=600); dev.off()

# 2.3 Measurement models ---------------- ######

  # Continuous model

	  cfa_1 <- '
	  # latent variables
	  bmerit =~ welljob + hardjob
	  need=~ family + child
	  '
	  fit_1 <- cfa(cfa_1,data=detearn, missing = "ML")
	  summary(fit_1, fit.measures=TRUE,standardized=TRUE)

	   # Path model
		  semPaths(fit_1,"std",layout = "tree2", rotation = 2,
	  	intercepts = FALSE,	style = "lisrel", curvePivot = TRUE,
	  	cardinal = TRUE, edge.color = "black", residScale = 10)
    dev.copy(png,"./results/path_detearn1.png",width=750,height=600)
    dev.off()

	# Categorical model

    fit_1c <- cfa(cfa_1,data=detearn, ordered = c("welljob",
    	"hardjob","family","child"))
    summary(fit_1c, fit.measures=TRUE,standardized=TRUE)  #probs

# End sec 2 -----------------------------

# 3. MEASURING BELIEFS & PERCEPTIONS ------ -

# 3.1 Define object --------------------- -------------------------------

			load("./objects/getahead.rda")
			load("./objects/detearn.rda")
			names(getahead)
			names(detearn)

			merit_per_pref=merge(getahead,detearn, by=c("id","v5"))
			names(merit_per_pref)
# 3.2 Measurement continuous ------------ ------------------

	# Estimación del modelo simple
	  cfa_perpref1 <- '
	  # latent variables
	  merit =~ hwork + ambition
	  parent=~ wealthy + pareduc
	  backg=~ race + gender
		network=~ people + polcone
	  bmerit =~ welljob + hardjob
	  need=~ family + child
	  '
	  fit_perpref1 <- cfa(cfa_perpref1,data=merit_per_pref,
	  	missing = "ML")
	  summary(fit_perpref1, fit.measures=TRUE,standardized=TRUE)

 	  # Path model
		  semPaths(fit_perpref1,layout = "tree2", rotation = 2,
		  	intercepts = FALSE, style = "lisrel", curvePivot = TRUE,
		  	sizeMan = 3.5, sizeLat = 6)

	    dev.copy(png,"./results/path_perpref1.png",width=800,
	    	height=650);  dev.off()

  # Estimación del modelo c/ unmerit 2do orden
	  cfa_perpref2 <- '
	  # latent variables
	  merit =~ hwork + ambition
	  parent=~ wealthy + pareduc
	  backg=~ race + gender
		network=~ people + polcone
		unmerit=~ parent + backg + network # 2nd order
	  bmerit =~ welljob + hardjob
	  need=~ family + child
	  '
	  fit_perpref2 <- cfa(cfa_perpref2,data=merit_per_pref,
	  	missing = "ML")
	  summary(fit_perpref2, fit.measures=TRUE,standardized=TRUE)

 	  # Path model
		  semPaths(fit_perpref2,layout = "tree", rotation = 2,
		  	intercepts = FALSE, style = "lisrel", curvePivot = TRUE,
		  	sizeMan = 3.5, sizeLat = 6)

	    dev.copy(png,"./results/path_perpref2.png",width=800,
	    	height=650);  dev.off()

# 3.3 Measurement categorical -pendiente- #####

	# Estimación del modelo simple categórico
	  fit_1c <- cfa(cfa_1,data=merit_per_pref, ordered=c("wealthy",
	  	"ambition","hwork","people","polcone","race","gender",
	  	"pareduc","welljob","hardjob","family","child"))
	  summary(fit_1c, fit.measures=TRUE,standardized=TRUE)

	# Estimación del modelo 2do orden categórico
	  fit_2c <- cfa(cfa_2,data=merit_per_pref, ordered=c("wealthy",
	  	"ambition","hwork","people","polcone","race","gender",
	  	"pareduc","welljob","hardjob","family","child"))
	  summary(fit_2c, fit.measures=TRUE,standardized=TRUE)
# 3.4. Save factor scores --------------- #####

	# Factor scores fit_perpref2

		# Multiple imputation (para recuperar missings en scores) - REVIEW THIS, command lavPredict from 2017 should predict with missing values therefore saving from multiple imputation

		  # generate 5 multiple complete datasets
			out <- mice(merit_per_pref[,c(3:10,19:24)], m=5)
			D1 <- complete(out, 1)
			D2 <- complete(out, 2)
			D3 <- complete(out, 3)
			D4 <- complete(out, 4)
			D5 <- complete(out, 5)

			# fit model for each complete dataset
			fit1 <- cfa(cfa_perpref2, data=D1)
			fit2 <- cfa(cfa_perpref2, data=D2)
			fit3 <- cfa(cfa_perpref2, data=D3)
			fit4 <- cfa(cfa_perpref2, data=D4)
			fit5 <- cfa(cfa_perpref2, data=D5)

		# predict scores for all models
			p1 <- predict(fit1)
			p2 <- predict(fit2)
			p3 <- predict(fit3)
			p4 <- predict(fit4)
			p5 <- predict(fit5)

		# compute 'average' across 5 sets of scores:
			scores <- (p1 + p2 + p3 + p4 + p5)/5

	  # Generate factor scores
	    scores_dat=as.data.frame(scores)
	    dim(scores_dat)
	    str(scores_dat)
	    summary(scores)

		# Generate object with factor scores
	    dim(merit_per_pref)
	    dim(scores_dat)
		  merit_per_pref_sco=cbind(merit_per_pref,scores)

	# Save object with factor scores
		names(merit_per_pref_sco)
		merit_per_pref_sco=merit_per_pref_sco[,c(1:2,25:31)]
	  save(merit_per_pref_sco,
	  	file="./objects/merit_per_pref_sco.rda")

 	# Descriptives factor scores
	  load("./objects/merit_per_pref_sco.rda")
	   # Correlations
	  names(merit_per_pref_sco)
    cormat=cor(merit_per_pref_sco[c(3,7:9)],
    	use="pairwise.complete.obs")
    corrplot.mixed(cormat)
      #save
     dev.copy(png,"./results/corplot_scores.png",width=600,
      	height=600); dev.off()

  # Descriptive scatter countries
     load("./objects/merit_per_pref_sco.rda")
     agg_merit=merit_per_pref_sco %>% group_by(v5) %>% summarise_all(funs(mean))

     agg_merit$v5c=countrycode(agg_merit$v5,
       'iso3n','iso3c')

     set_label(agg_merit$merit)="Meritocratic perception"
     set_label(agg_merit$bmerit)="Meritocratic preferences"

     sjp.scatter(agg_merit$merit,agg_merit$bmerit, 
                 dot.labels = agg_merit$v5c,
                 fit.line = TRUE,
                 show.ci = TRUE)

     dev.copy(png,"./results/mperc_mprefs.png",width=600,
              height=600); dev.off()

     sjp.scatter(agg_merit$merit,agg_merit$unmerit, 
                 dot.labels = agg_merit$v5c,
                 fit.line = TRUE,
                 show.ci = TRUE)

     dev.copy(png,"./results/mperc_munperc.png",width=600,
              height=600); dev.off()

     sjp.scatter(agg_merit$bmerit,agg_merit$need, 
                 dot.labels = agg_merit$v5c,
                 fit.line = TRUE,
                 show.ci = TRUE)

     dev.copy(png,"./results/mperc_munperc.png",width=600,
              height=600); dev.off()

# 3.5 Invariance full model ---pendiente- --------

	fit.indices=c("chisq", "df", "cfi", "tli", "rmsea")

	# Continuous first order
	  fit.conf1=cfa(cfa_1, data=merit_per_pref, group="v5",std.lv = TRUE)
		fitMeasures(fit.conf1,fit.indices)
		summary(fit.conf1)
	  inspect(fit.conf1,"theta")
	  inspect(fit.conf1,"cov.lv")
	  partialInvariance(fit.conf1, type = "strict", p.adjust = "holm")

# End section 3 -------------------------
```

### 2021-05-25 paper.Rmd
- major update

--- 2021-03-23 docs/paper.Rmd
@@
 bibliography:
-  - ../input/bib/EconomiaMoral.bib
-  - ../input/bib/meritocracy.bib
-  - ../input/bib/invariance.bib
+#  - input/bib/EconomiaMoral.bib
+#  - input/bib/meritocracy.bib
+#  - input/bib/invariance.bib
 editor_options:
@@
 Sys.setlocale("LC_ALL", "ES_ES.UTF-8")
-```
-
-```{r message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
-pacman::p_load(knitr, kableExtra, dplyr, lavaan)
-```
-
-```{r echo=FALSE }
-table_format = if(is_html_output()) {
-  "html"
-} else if(is_latex_output()) {
-  "latex"
-}
-
-table_format2 = if(is_html_output()) {
-  T
-} else if(is_latex_output()) {
-  F
-}
 ```
@@
 ```{r merit-model, echo=FALSE, fig.cap = "Model of perception and preferences for meritocracy and non-meritocracy", fig.align='center'}
-knitr::include_graphics("../input/images/generalf.png")
+knitr::include_graphics("input/images/generalf.png")
 ```
@@
 
+pacman::p_load(knitr)
+```{r message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
+#RUN THIS BEFORE THE ANALYSIS
+
+table_format = if(is_html_output()) {
+  "html"
+} else if(is_latex_output()) {
+  "latex"
+}
+
+table_format2 = if(is_html_output()) {
+  T
+} else if(is_latex_output()) {
+  F
+}
+```
+
 The analysis is organized in three studies.
-Study 1 serves as a background to our instrument proposal by attempting to operationalize the meritocracy dimensions with available data.
+Study 1 serves as a background to our instrument proposal by attempting to operationalize the meritocracy dimensions with available secondary data.
 For this we use data from the International Social Survey Programme (ISSP).
@@
 Finally, in Study 3 we perform additional validity analysis of the meritocracy scale.
-
-## Study 1
-
-### Data
-
-The data used in the analysis corresponds to the last available wave of the social inequality module from the International Social Survey Programm (ISSP), which is the most specialized international comparative survey in perceptions, attitudes and beliefs about inequality related issues.
-This wave corresponds to the year 2009 and covers attitudes towards a series of topics dealing with social inequality across 26 countries and XXXX individuals.
-Although there is data available from this module for previous waves (1987, 1992 and 1999), unfortunately there are several variables that are important for this research that were not included, particularly in 1987 and 1999, reason why we only use the last wave.
-
-### Variables
-
-There are a series of indicators in the ISSP survey that in the following we attempt to classify in our meritocracy measurement scheme.
-However, it is important to mention  that the items were not originally thought as a specific measure of meritocracy.
-Still, many of the items have been used in previous research on meritocracy (and certainly will keep being used in the future) and before proposing a new measure we consider to give a first test to our proposed conceptualization using available data that not only covers several countries but also is validated by the research community.
-
-The variables are presented below organized in perceptions and preferences:
-
-- _Perception of meritocracy/non-meritocracy_: for operationalizing perceptions the closest set of ISSP's indicators comes from the question asking about perceptions for opportunities to get ahead, which are usually considered as "meritocratic beliefs" in previous research.
-The general heading of the battery is: _"To begin we have some questions about opportunities for getting ahead.
-Please tick one box for each of these to show how important you think it is for getting ahead in life."_ This is followed by a list of statements to be answered in a scale with the following levels numbered from 1 to 5: essential, very important, fairly important, not very important, not important at all.
-Below the items are listed classified according to the meritocracy/non-meritocracy character:
-
-  - Meritocratic perception:
-    - How important is having ambition?
-    - How important is hard work?
-
-  - Non-meritocratic perception
-    - How important is coming from a wealthy family?
-    - How important is having well-educated parents
-    - How important is knowing the right people
-    - How important is having political connections
-    - How important is a person's race?
-    - How important is being born a man or a woman?
-
-<br>
-  The classification is based on criteria of internal motivation (meritocratic) and structural constraints (non-meritocratic).
-There were two items from the battery that were excluded from the analysis as they would not fit into the classification.
-The first one was "having good education yourself", since it was not clear whether this is could be due to individual motivation or system opportunities, and the second was "giving bribes", as introduced elements of criminality that were beyond a non-meritocratic perception.
-
-- Preferences for meritocracy-non meritocracy: for operationalizing normative preferences we used a list of items related to reasons for pay.
-The ISSP question was: _In deciding how much people ought to earn, how important should each of these things be, in your opinion_, rated in the same _essential-non important at all_ scale as the questions for meritocratic preferences.
-The reasons are:
-
-  - Meritocratic preferences
-    - How well he or she does the job?
-    - How hard he or she works at the job?
-
-  - Non-meritocratic preferences
-    - What is needed to support a family?
-    - Whether the person has children to support?
-
-(here summary descriptive table - @julio)
-
-### Methods
-
-The estimation was performed using Confirmatory Factor Analysis (CFA).
-CFA was conducted using the `lavaan` R package (version 0.6-3; Rosseel, 2020), with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement (Kline, 2016; Rosseel, 2020).
-As recommended by Brown (2008), we assessed model fit by jointly considering the comparative fit index and Tucker-Lewis Index (CFI and TLI; acceptable fit > 0.95), Root of the average squared residual approximation (RMSEA; acceptable fit < 0.08),  Chi-square: (p-value; acceptable fit > 0.05, and Chi-square ratio > 3).
-
-### Results
-
-[diagrama SEM]
-
-[tabla CFA]
-
-## Study 2
-
-## Data collection {-#data}
-
-The data was obtained through an online questionnaire, which was part of a larger study on meritocracy and preferences developed in Chile in 2019 funded by the national scientific agency ANID.
-The questionnaire was programmed in Qualtrics and the fieldwork was conducted by an external online survey agency ([netquest.cl](www.netquest.cl)) between December 2019 and January 2020.
-The sample was selected from a non-probabilistic quota design in three large cities in Chile (Santiago, Concepción & Antofagasta).
-The quotas for gender, age, and education level were generated based on a survey by the Public Studies Center [@cepEncuestaCEPMayo2019], which is a well regarded counterpart agency of the ISSP (International Social Survey Programme) in Chile.
-A total sample of 2,141 people was collected, excluding those who did not answer the questions on the scale, and those who did not accept informed consent.
-There were no significant differences between our sample and the wider population for most socio-demographic characteristics, with the exception of educational level (see Table \@ref(tab:rep-samp) in appendix).
-As is often the case with online surveys, there were some limitations in achieving the quotas for lower educational levels **[citas - LUCHO]**.
-
-## Instrument design
-
-The proposed scale of perceptions and preferences relating to meritocracy consisted of eight indicators that were grouped into the four dimensions listed earlier: Perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic).
-In order to achieve at least some comparability with previous studies, the questions were adapted from the items battery "reasons to get ahead" (ISSP/GSS), which has been widely used for operationalizing meritocracy in previous studies [@mijs_paradox_2019; @duru-bellat_whos_2012; @reynolds_perceptions_2014].
-The aforementioned eight items, ordered according to dimensions, are presented in Table \@ref(tab:table-indicadores).
-These eight likert-type items have five response alternatives, ranging from "Completely disagree"(1) to "Completely agree" (5).
-
-```{r include=FALSE}
-cap <- "Items according to dimension."
-# if (knitr::is_html_output()) {tb <- table_nums("tb",cap)} else {tb <- cap}
-```
-
-```{r table-indicadores, echo=FALSE}
-tab01 <- read.csv(file = "../input/data/original/materials/table01.csv",header = 1,sep = ",",encoding = "UTF-8")
-cnames <- c("Dimensions", "Factor","Statement (english)" , "Statement (spanish)")
-
-kable(tab01,table_format,booktabs = T, linesep = "",col.names = cnames, caption = cap) %>%
-  kable_styling(
-    full_width = T,
-    latex_options = c("hold_position"),
-    position = "center",
-    font_size = 10,
-    bootstrap_options=c("striped", "bordered")) %>%
-  column_spec(column = 1, width = "1.5cm", ) %>%
-  column_spec(column = 2,width = "2 cm") %>%
-  collapse_rows(columns = 1:2,valign = "middle")
-```
-
-\pagebreak
-
-## Administration sets
-
-With the objective of evaluating the effect of indicator ordering in the responses, three different versions of items' order were designed and randomly assigned, as depicted in Figure \@ref(fig:appmod).
-The scale was presented to the first group (_n = 712_) in the order that appears in Table \@ref(tab:table-indicadores) according to perceptions and preferences.
-For the second group (_n = 717_), the order was reorganized according to perceptions and preferences over the same topic, e.g.
-for the topic of hard work, the item about perception was followed by the item about preference, and the same for the rest of the topics.
-Finally, for the third group (_n = 712_), the items were completely randomized.
-
-```{r appmod, echo=FALSE, fig.cap = "Survey flow", fig.align='center'}
-knitr::include_graphics('../input/images/app_mod.png')
-```
-
-# Methods
-
-To test the scale's underlying constructs, we employed confirmatory factor analysis models (CFA).
-The models estimated one factor for each of the proposed dimensions, as represented in the following figure:
-
-```{r meas01, echo=FALSE, fig.cap = "Theoretical model", fig.align='center'}
-knitr::include_graphics('../output/images/meas01.png')
-```
-As in Study 1, CFA was conducted using the `lavaan` R package (version 0.6-3; Rosseel, 2020), with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement (Kline, 2016; Rosseel, 2020).
-The fit indexes and cut-off criteria were the same as in Study 1.
-
-A pre-registration was made in the OSF platform, available at the following link: [https://osf.io/z45y2](https://osf.io/z45y2).
-Included in this pre-registration are the hypotheses regarding the four-dimensional conceptual model underlying the scale, the variable measurement levels, the statistical tests to be performed with their respective evaluation parameters, and other important aspects of the research design.
-
-# Results
-
-## Descriptive analyses
-
-As can be seen in the Table \@ref(tab:desc01), the indicators have values ranging from 1 (totally disagree) to 5 (totally agree).
-Averages are observed from 2.41, corresponding to preference-contacts; to 3.89, corresponding to preference-effort.
-Both indicators are consistent with a general adherence to meritocracy as reported in previous studies, privileging individual aspects such as effort [@bucca_merit_2016; @frei_fruto_2020].
-
-```{r desc01, echo=FALSE}
-cap <- "Descriptive statistics of the scale."
-load(file = "../input/data/proc/dat04.RData")
-dat04b<- dat04 %>% rename("A. Perception Effort" =perc_effort,
- "B. Perception Talent" = perc_talent,
- "C. Perception rich parents" = perc_wpart,
- "D. Perception contacts" = perc_netw,
- "E. Preferences Effort" = pref_effort,
- "F. Preferences Talent" = pref_talent,
- "G. Preferences rich parents" = pref_wpart ,
- "H. Preferences contacts" = pref_netw )
-sum04b <- psych::describe(dat04b[,-c(1)], fast = T) %>% data.frame() %>% select(mean,sd,min,max) %>% round(digits = 2)
-table04b<- sum04b %>% rename("Mean"=mean, "SD"=sd, "Min"=min,"Max"=max)
-sumdat04<- kable(table04b,format =table_format,
-                 booktabs= T,
-                 #exhaust = FALSE,
-                 align = "l",
-                 linesep = "",
-                 caption = cap) %>%
-  kable_styling(full_width = table_format2,
-                font_size = 10,
-                position = "center",
-                latex_options = c("HOLD_position"));sumdat04
-```
-
-The graphs presented in Figure \@ref(fig:plotlikert) display disaggregated and comparable information of the different response categories for each item.
-As it can be observed, in general there is more agreement in the perception of non-meritocratic items than in meritocratic items, while in the case of preferences, the opposite occurs.
-As far as preferences are concerned, the preponderant role of effort over talent as a criterion of meritocratic preference is noteworthy.
-
-```{r plotlikert, echo=FALSE, fig.cap = "Descriptive plot", fig.align='center'}
-knitr::include_graphics('../output/images/plotlikert.png')
-```
-
-In Figure \@ref(fig:corpoly), moderate to high intensity relationships are observed between indicators that correspond to the same factor (e.g., perception of meritocracy by effort and by talent, r=0.56).
-The relationships between meritocratic perceptions and preferences are also salient, which is not the case for non-meritocratic indicators.
-
-```{r corpoly, echo=FALSE, fig.cap = "Polychoric correlation plot", fig.align='center'}
-knitr::include_graphics('../output/images/corpoly.png')
-```
-In sum, descriptive analyses point to a relative adherence to meritocracy, which is expressed as a greater preference for meritocratic criteria, and a lesser preference for non-meritocratic criteria, even when the preference for non-meritocratic elements reaches a proportion of more than 20%.
-Likewise, a relatively low perception of meritocracy is observable.
-In addition, we observed a coherent relationship between the indicators of the same dimension as proposed in the pre-registration of the study; that is, the pairs of items associated with a specific factor show correlations with a large effect size.
-In particular, the associations between effort and talent are relevant, challenging previous assumptions that talent would not be a meritocratic criterion [@mijs_paradox_2019], otherwise the correlation would be zero or negative.
-In addition, we observe that there is no negative correlation between meritocratic and non-meritocratic aspects, undermining the assumptions of previous studies that suggested that these dimensions would be the opposite poles of one same continuum [@reynolds_perceptions_2014].
-
-## Confirmatory Factor Analysis
-
-This section estimates the fit of the conceptual model behind the Perceptions and Preferences for Meritocracy Scale.
-For this, we first estimate a confirmatory factor analysis model for the whole sample, and secondly we test the order effects applying the same model to each of the three order permutations.
-
-Figure \@ref(fig:meas02) shows the results of the estimation for the four-factor model using the complete sample.
-This model shows adequate fit indicators (CFI = 0.977, RMSEA = 0.079, $\chi2$(df = 14) = 28,03, p = 0.014) and it presents factorial loads with weights that vary between 0.63 and 0.88.
-
-```{r meas02, echo=FALSE, fig.cap = "Confirmatory factor analysis of the Perceptions and Preferences for Meritocracy Scale", fig.align='center'}
-knitr::include_graphics('../output/images/meas03.png')
-```
-Attending to the correlations between the latent variables as depicted in Figure \@ref(fig:latcor), it is observed that the variables with the strongest correlations are meritocratic perceptions and preferences, whit an $r=0.59$, while non-meritocratic perceptions and preferences show a correlation close to 0.
-Besides, and as commented before in the analysis of the single items correlations, the latent variables of meritocratic and non-meritocratic versions of both perceptions and preferences show very low correlations.
-This last finding gives further evidence regarding the lack of unidimensionality of meritocratic and non-meritocratic aspects: they do not appear as the opposite poles of the same continuum and based on this evidence the use of reverse-coding for non-meritocratic items as indicators of meritocracy should be avoided.
-
-```{r latcor, echo=FALSE, fig.cap = "Latent variables correlation", fig.align='center'}
-knitr::include_graphics('../output/images/latcor.png')
-```
-In a second step we test the effect of the items' order in the confirmatory model.
-Table \@ref(tab:tab03) shows the fit indicators of the models estimated for each of the three versions of the items' order described in the methodology, being the first one according to perceptions and preferences, the second one according to meritocratic and non-meritocratic aspects, and the third one completely randomized.
-Regardless of the group and order of items, all models obtained an adquate fit, with CFI's above 0.95 and RMSEA's below 0.08.
-However, none of the models achieved a non-significant chi-square, something expected in large samples as the one used here).
-Model 1 shows the fit indicators for the estimation of the CFA model to the first version order (perceptions-preferences), which was the one obtaining best fit (CFI=0.993, TLI=0.995, RMSEA=0.034, $\chi2$(df=14)=42.276), whereas Model 2 for the fixed order according to merit/non-merit items shows the comparatively worst indicators.
-The CFA fit indices for the completely randomized items' order (Model 3) is somehow an intermediate fit indexes when compared with the previous two.
-Still, it keeps all the indicators within the acceptable cut-off criteria and besides it controls for possible order effects in the administration of the instrument.
-
-```{r tab03, echo=FALSE,eval=TRUE}
-load(file = "../output/tables/fit4_o.RData")
-load(file = "../output/tables/fit3_o.RData")
-load(file = "../output/tables/fit2_o.RData")
-load(file = "../output/tables/fit1_o.RData")
-sum_fit<- bind_rows(fitmeasures(fit1_o)[c("chisq","df","cfi","tli","rmsea")],
-                    fitmeasures(fit2_o)[c("chisq","df","cfi","tli","rmsea")],
-                    fitmeasures(fit3_o)[c("chisq","df","cfi","tli","rmsea")])
-sum_fit$mod <- c("Model 1","Model 2","Model 3")
-sum_fit$est <- c("DWLS","DWLS","DWLS")
-sum_fit$nobs <- c(nobs(fit1_o),nobs(fit2_o),nobs(fit3_o))
-sum_fit <- select(sum_fit,mod,nobs,est,everything())
-colnames <- c("Model","$N$","Estimator","$\\chi^2$","df","CFI","TLI","RMSEA")
-sumtable02<- kable(sum_fit,digits = 3,format = table_format,row.names = F,booktabs=T, align = c("lcccccc"),
-                   caption = "Summary fit indices according group",
-                   col.names = colnames,
-                   escape = FALSE) %>%
-  kable_styling(full_width = F,
-                font_size = 10)  %>%
-      footnote(number = c("Model 1: fixed order by perception/preference",
-                          "Model 2: fixed order by merit-non merit",
-                          "Model 3: Randomized order"));sumtable02
-```
-
-## Additional validity analyses
-
-We performed two further analyses in order to add evidence regarding the quality of the scale.
-Firstly we analyzed the convergent validity by exploring the association of the merit scale with related concepts and itscorrespondent measurements.
-Secondly, we test the internal consistency of the measurement model behind the scale comparing with a different sample through measurement invariance procedures .
-
-### Convergent validity
-
-_Data_
-
-For this analysis we examined data from one wave of the same study described above, which was a three-wave panel survey.
-The last wave of the study included additional measures that allow testing the correlations of the merit scale with some related constructs as opportunity beliefs and personal wherewithal.
-After listwise deletion of missing cases in socio-demographics information, a total of 1422 individuals took part in the third wave: 668 (46.97%) women and 754 (53.03%) men.
-The ages of 50.57% of them are 45 years or older, and 36.28% held a tertiary degree.
-
-_Intruments_
-
-The following scales were included for testing their correlations with the the 8-item meritocracy scale:
-
-- Opportunity beliefs: We used two items of the social inequality module of the International Social Survey Programme.
-Studies use generally both indicators to measure meritocratic and non-meritocratic beliefs [@McCall2017Exposurerisinginequality; @mijs_paradox_2019].
-One item measures the importance of individual factors in determining life outcomes by asking for the importance of hard work for getting ahead in life (M=3.76, SD=0.93).
-The second item asks for the importance of coming from a wealthy family and captures the importance of structural factors in getting ahead (M=2.75, SD=1.27).
-Participants responded to each item using a 1 (not important at all) to 5 (essential) response scale.
-We call the first item \textit{hard work} and the second one \textit{social origin}.
-
-- Personal wherewithal: the questionnaire included The Neoliberal Beliefs Inventory (NBI), which considers four factors: Government Interference preferences, Competition preferences, System Inequality perception and Personal Wherewithal [@bay-chengTrackingHomoOeconomicus2015].
-We used the factor _personal wherewithal_ that reflects meritocracy beliefs in terms of the importance of personal attributes as strength and skills to yield success.
-The construct consists of 8 items (e.g., “Any goal can be achieved with enough hard work and talent”, “I’ve benefited from working hard, so there’s no reason others can’t”, “Anyone who is willing to work hard can be succesful in Chile”).
-Participants answered the items using a 1 (totally disagree) to 6 (totally agree) scale.
-We simple computed mean scores, with higher scores indicating stronger support for meritocracy (M=3.32, SD=0.88, $\alpha$ = 0.89).
-
-Both instruments cover mainly the perceptual side of the concept.
-Therefore, in general we expect larger correlation with meritocratic perceptions than with preferences.
-In the case of opportunity beliefs, the hard work item should correlate positively with meritocratic perception whereas the importance for coming from a wealthy family is expected to correlate positively with non-meritocratic perceptions.
-Regarding the second instrument of personal wherewithal, we also anticipate a positive correlation with meritocratic perception and low to null correlations with the other dimensions of the merit scale.
-
-_Results_
-
-We used polyserial correlations to evaluate relationships of meritocractic and non-meritocratic preferences and perceptions with the items of the opportunity beliefs battery.
-For personal wherewithal, we examined Pearson correlations.
-
-Table \@ref(tab:val-conv) shows the polyserial and pearson correlations between our meritocracy scale and common measures of meritocracy.
-The findings provide evidence for convergent validity for the scale.
-As we expected, preferences show very weak correlations.
-Findings also indicate that correlations of perceptions with opportunity beliefs are consistent with our expectations.
-The item for social origin shows a positive and moderate association with unmeritocratic perception (r=.354), and a negative and low correlation with meritocratic perception (-0.197).
-Regarding the item of hard work, the correlation with meritocratic perception is positive but weak (r=.187).
-One explanation for this result is that the support for hard work is very high among respondents---only 8.6 percent indicates _not very important_ or _not important at all_.
-
-  |                                      | Social Origin  |     Hard Work    |     Wherewithal    |
-  |--------------------------------------|----------------|------------------|--------------------|
-  |     Meritocratic     Perception      |     -0.1975    |     0.1877       |     0.3921         |
-  |     Unmeritocratic    Perception     |     0.3547     |     -0.0775      |     -0.1844        |
-  |     Meritocratic    Preference       |     0.1078     |     0.0779       |     0.0819         |
-  |     Unmeritocratic    Preference     |     -0.0397    |     0.0824       |     0.2082         |
-  Table: `r as.character(paste("(\\#tab:val-conv)", "Polyserial and Pearson correlation whit other merit scales"))`.
-
-The measure of NBI for _personal wherewithal_ also provides evidence for the convergent validity of the merit scale.
-Results indicate a positive correlation of this measure with meritocratic preference (r=.392), suggesting that both variables capture similar concepts.
-The NBI's factor depicts a positive association with unmeritocratic preference (r=.208).
-This finding is consistent with literature, inasmuch as the belief for meritocray is associated with the justification of inequality [@McCall2017Exposurerisinginequality; @MadeiraPrimesConsequencesSystematic2019].
-Furthermore, results of our confirmatory factor analysis indicate that meritocratic perceptions and unmeritocratic preferences show a positive and weak correlation.
-
-### Measurement invariance
-
-The modeling for invariance testing has been gaining more relevance in sociological survey studies, mainly due to the interest in the validity of measurement scales in comparative studies in various social and cultural contexts [@Davidovetal2014].
-In this section we will use an analysis of invariance between groups, where the proposed factorial structure is expected to be independent of belonging to the sample corresponding to the original study, comparing it with another more recent study available and where the meritocracy scale was also incorporated.
-
-The invariance measurement procedure consists of a series of nested models to which restrictions are progressively incorporated on the parameters of the measurement model.
-The literature generally suggests that this modeling should be done through four levels or types of progressive restriction [@Milfont2010; @millsap_Statistical_2011; @vandeschoot_checklist_2012]:
-
-1. **Configural:** the model is estimated only indicating the factorial structure used in the CFA.
-
-2. **Weak:** an equality restriction is applied to the factor loadings in the different groups, that is, the loadings are forced to be identical in both measurements.
-
-3. **Strong:** equality restrictions are added to the intercepts of each indicator.
-
-4. **Strict:** equality restrictions are added to the error variances of each indicator.
-
-*Data*
-
-For testing the invariance we compared the data described previously with a new data source coming from an online survey carried out during the first half of 2020.
-The characteristics of this sample in terms of application and coverage are equivalent to those of the previous study.
-The final sample obtained contains 1,242 cases, where 605 (48.71%) are women and 637 (51.28%) are men.
-48.84% of them are 45 years of age or older, and 33.97% have tertiary education or higher.
-
-_Instruments_
-
-The items of the meritocracy scale were identical to the original.
-The items were administrated in a randomized order (which corresponds to the application modality of Group 3 present in Figure \@ref(fig: appmod)).
-
-_Results_
-
-Tabla \@ref(tab:fcloads-inv) shows de results of the CFA estimation of the meritocracy scale in this new sample.
-As observed, the loadings and the fit indices show similar results as the ones presented for the original study above, which gives a first base with wich to start the invariance analysis.
-
-The first step for invariance testing is the estimation of the configural model, which serves as the baseline for further comparison and it is expected to adequately meet the global fit criteria of a measurement model.
-Although the chi-square statistic is used as a global measure of fit,since its sensitivity to sample size it is recommended to use three additional fit indices: Comparative Fit Index (CFI), which should have a value greater than 0.95; Root Mean Square Error Approximation (RMSEA), which must be in the range of values ​​of 0.05 and 0.08; and the Standardized Root Mean Square Residuals (SRMR) which must be less than 0.08 [@vandeschoot_checklist_2012].
-Besides, the literature on measurement invariance suggests some complementary approaches for the evaluation of the fit, from which we will consider the incremental adjustment of the fit indexes [@cheung_Evaluating_2002; @Milfont2010; @dimitrov_Testing_2010] and the ANOVA test for means comparison in nested models [@newsom_Longitudinal_2015].
-
-```{r invargroup, eval=TRUE, echo=FALSE}
-load(file = "../output/tables/tb.inv-group.RData")
-col.nam <- c("Model","$\\chi^2 (\\text{df})$","CFI","RMSEA (90 CI)",
-             "$\\Delta \\chi^2 (\\Delta \\text{df}$)","$\\Delta \\text{CFI}$","$\\Delta \\text{RMSEA}$","Decision")
-footnote <- paste0("N = ",nobs(conf),"; Group 1, n = ",conf@Data@nobs[[1]],"; Group 2, n = ",conf@Data@nobs[[2]], ", ***p < 0.001")
-
-kable(tab.inv, col.names = col.nam, align = "l",
-      booktabs=TRUE,format = table_format,escape = FALSE,
-      caption = "Multiple Group meassurement invariance for Perceptions and Preferences for Meritocracy") %>%
-  kable_styling(full_width = FALSE,latex_options = "HOLD_position") %>%
-  footnote(general = footnote, footnote_as_chunk = T)
-```
-
-Table \@ref(tab:invargroup) shows the results of the measurement invariance estimation.
-When attending to the traditional invariance test of $\Delta \chi^2 (\Delta \text{df})$, the results  support the invariance at the strong level meaning that the fit of the factor model of the merit scale is equivalent across samples when constraining factor loadings and intercepts to being equal.
-Such result is considered in general as evidence of invariance [citar paper Fischer en SJR], as strict forms of measurement invariance rarely hold [citar van der Schoot et al].
-Still, the comparability of latents means requires strict invariance which in this case does not hold when considering $\Delta \chi^2 (\Delta \text{df})$.
-Nevertheless, the criteria of $\Delta \text{CFI}$ used for comparing models is close to the rejection criteria of >.01, whereas the $\Delta \text{RMSEA}$ fulfills the requirements of being below of the cut-off criteria as suggested by [ @chen 2007 ].
-Therefore, using this last standard the level for strict invariance holds for the meritocracy scale.
 
@@
 
-```{r tab-loads, echo=FALSE}
-load(file = "../output/tables/tb.fm.RData")
+```{r tab-loads, echo=FALSE, eval=FALSE}
+pacman::p_load(dplyr)
+load(file = "output/tables/tb.fm.RData")
 cap <- "Factor loads and model fit."
@@
 
-```{r fcloads-inv, echo=FALSE}
-load(file = "../output/tables/tb.fact-f02.RData")
+```{r fcloads-inv, echo=FALSE, eval=FALSE}
+load(file = "output/tables/tb.fact-f02.RData")
 tb.foot <- paste0("Standardized factor loadings using DWLS estimator ; CFI = Comparative fit index (scaled); RMSEA = Root mean square error of approximation (scaled)")

### 2021-05-25 paper_study1.Rmd
- major update

--- 2021-04-01 docs/paper.Rmd
@@
 ```{r message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
-pacman::p_load(knitr, kableExtra, dplyr, lavaan)
-```
+#RUN THIS BEFORE THE ANALYSIS
+pacman::p_load(knitr)
 
-```{r echo=FALSE }
 table_format = if(is_html_output()) {
@@
 
-# Introduction
-
-Economic inequality has become an issue of growing concern around the world.
-This has been expressed in a series of protests, as well as diverse critical analyses regarding the development of capitalism and its consequences [@streeckPoliticsPublicDebt2014; @piketty_Capital_2014].
-In this context, the study of views, preferences and perceptions of inequality has acquired relevance in the social sciences, in topics such as redistributive preferences [@alesina_Fairness_2005; @dimick_Models_2018], the legitimization of economic inequality [@schroder_Income_2017] and the functioning of meritocracy [@duru-bellat_whos_2012; @mijs_paradox_2019; @reynolds_perceptions_2014; @atriaEconomicElitesAttitudes2020].
-
-In general, meritocracy is defined as a system of distribution of resources and rewards based on individual merit, which in its original conception is a combination of talent and effort [@young_rise_1962].
-This traditional conception of merit places in a secondary position the possible interference of structural or non-meritocratic factors, such as inheritance, personal contacts, and luck [@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995; @yairMeritocracy2007; @landWeSatTable2006; @youngRiseMeritocracy1994].
-Social psychology and sociology have studied the characteristics and consequences of beliefs in meritocracy, under the general hypothesis that a greater belief in meritocracy emphasizes the role of individual over structural factors in personal achievements, leading to greater legitimization of inequalities [@premingerMeritocracyServiceEthnocracy2020; @trumpWhenWhyEconomic2020; @hadjar_meritokratie_2008; @MadeiraPrimesConsequencesSystematic2019].
-Such research has raised criticism of meritocracy as a moral standard of distribution, arguing that it would be an unfulfilled promise given the still preponderant weight of non-meritocratic elements upon individual status and social mobility[@sandelTyrannyMeritWhat2020a; @witteveenReconsideringMeritocraticPower2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019; @khan_Privilege_2013].
-
-Due to the role that meritocratic beliefs play in the justification of individual achievement or failure in contemporary societies [@hadjar_meritokratie_2008, @markovits_meritocracy_2019; @sandel_tyranny_2020], multiple studies have evaluated the relationship between meritocratic beliefs and personal and/or contextual characteristics.
-For example, some studies have linked meritocracy to the reinforcement of socio-economic, gender, and ethnic stereotypes [@MadeiraPrimesConsequencesSystematic2019; @girerdNeoliberalismIdeologicalBarrier2020; @premingerMeritocracyServiceEthnocracy2020], as well as the effects of meritocratic beliefs in educational [@generettStoriesWeTell2020; @owensENGINESSOCIALMOBILITY2020] and organizational contexts [@perezAdvancingCareersMerit2020; @aielloNewEvidenceFirmuniversity2019].
-
-Most of the studies so far have used indicators from existing standard surveys (as the International Social Survey Programme - ISSP) to measure meritocracy.
-However, as we will show later, the methods of measuring meritocracy vary extensively among studies.
-In many cases, similar phenomena are associated with different indicators, and conversely, often different phenomena are measured with similar indicators.
-Such trends not only limits the comparability of studies, but also the ability to understanding the effects of meritocratic beliefs across different fields.
-
-Based on a critical analysis of different approaches to the measurement of meritocracy to date, this article proposes an instrument to both measure and relate two key aspects in the study of meritocracy: perceptions and preferences.
-Furthermore, as a second axis of analysis, it considers the elaboration of indicators regarding meritocratic and non-meritocratic aspects, demonstrating that they are not two poles of the same continuum as some previous studies would seem to suggest.
-The measurement system outlined is oriented to generate an instrument as simple and brief as possible, being suitable to be used in public opinion surveys and allowing to integrate meritocratic beliefs in the study of different social phenomena.
-
-## The black-box of meritocratic beliefs {-#blackbox}
-
-In the following we discuss four critical aspects in studies conceptualizing and measuring meritocracy and that are the basis of our scale proposal.
-
-a. _Conceptual components_: Is merit only effort?
-
-One recent definition of meritocracy by @mijs_paradox_2019 is the following: "when I discuss meritocracy beliefs, I am referring to citizens’ belief in the importance of hard work relative to structural factors." [@mijs_paradox_2019, pg.9].
-In the subsequent operationalization, this is associated with the following question and indicator: "how important you think it is for getting ahead in life: (a) hard work", scored from 1 to 5 on a likert scale.
-The assumptions behind such definition are worth discussing in light of the conceptual meaning of meritocracy and its possibilities of operationalization.
-
-The item used by @mijs_paradox_2019 is part of an items' battery present in several international surveys, usually called "reasons to get ahead".
-This battery displays a series of indicators related to what people consider important to get ahead in life: hard work, education, ambition, a wealthy family, the right connections, religion, race, and gender.
-Therefore, when considering only one of the items, it means that other aspects that could be associated with talent, such as education, would not be deemed as meritocratic.
-As he points out: "Hard work is arguably the most meritocratic part of Michael Young’s equation: ‘Merit = Intelligence + Effort’, for the simple fact that intelligence itself is influenced by a non-meritocratic factor: who your parents happen to be" (p.5).
-
-In this measure of meritocratic beliefs, we can observe a couple of strong conceptual assumptions: effort would not depend on parental influence, and talent (as innate ability) is not meritocratic (contrary to Michael Youngs' original conceptualization).
-This conceptual and measurement-based assumption is found in other studies, that also assume effort to be the main and only aspect of meritocracy [@girerdNeoliberalismIdeologicalBarrier2020a; @bubakPerceptionsMeritocracyNote2019].
-Is effort the only dimension behind the concept of meritocracy, or includes other components, such as talent (as in its original conceptualization).
-The question of whether talent is or is not considered meritocratic is certainly an interesting topic to discuss from a philosophical point of view, but for those working on empirical studies it should be something that face empirical scrutiny.
-Besides, the consideration of talent as part of meritocracy certainly opens some interesting avenues of research, for instance in studies showing that for the elites meritocracy is more related to talent, whereas effort is something more characteristic of meritocracy of the middle and low classes ([citar Atria ABS]).
-Actually, in a posterior paper by @mijsHowInformationEconomic2020, somewhat contradictory with its previous definition, he considered talent as an indicator of meritocracy.
-
-_b. Beliefs?_
-
-Several approaches to the empirical study of meritocracy based on public opinion surveys, refer to the concept of _beliefs_, however different meanings and operationalizations usually underlie this concept.
-To illustrate this point, lets consider the most commonly used items for operationalizing meritocratic beliefs in survey research (as ISSP), which is the "reasons to get ahead" battery.
-It consists in a series of items asking "how important you think it is for getting ahead in life" and then lists several factors, such as effort, education, parental wealth, and contacts.
-Another version of this same battery used in several surveys - sometimes along the previous one - asks about "how important you think it _should_ be ...", then listing the same concepts.
-Therefore, the question raised here is: Which one of both is a "belief"?
-What _is_ or what it _should_ be?
-
-The term belief has an ambiguous character in the literature, conceived as "idea-elements" by @converseNatureBeliefSystems1964 or "considerations" by @zallerNatureOriginsMass1992.
-As @kluegelBeliefsInequalityAmericans2017 pointed out about the scope of beliefs: "This usage encompasses such more specific social-psychological concepts as values, perceptions, and attitudes" (p.30).
-Therefore, beliefs cover almost anything related to subjective factors.
-To this regard, a relevant distinction in the field of inequality beliefs was made by @janmaatSubjectiveInequalityReview2013: "Perceptions refer to subjective estimates of existing inequality (i.e.
-thoughts about what is).
-Beliefs are here defined as normative ideas about just inequality (i.e.
-thoughts about what should be)"(p.359).
-Several papers dealing with meritocracy use the term beliefs (i.e.
-what should be), while actually referring to perceptions (i.e.
-what is).
-This occurs not only in @mijs_paradox_2019, but also in @reynolds_perceptions_2014, in which the term beliefs is used to talk about perceptions, whereas other authors use general terms such as attitudes [@kunovich_systems_2007].
-The first attempt to shed light on this issue was made by @duru-bellat_whos_2012, who used the question "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" as a proxy for "desired" meritocracy (beliefs).
-They then determined "perceived" meritocracy, using the questions: “Would you say that in your country, people are rewarded for their efforts?” and “...
-people are rewarded for their skills?”.
-
-Is the belief in meritocracy a perception or a preference with normative meaning?
-In order to expand the analytical conceptual framework, we believe that both dimensions should be included in analyses, as proposed by @duru-bellat_whos_2012.
-This opens up the possibility of analyzing whether perceptions and preferences are actually related (i.e.
-have a high correlation), or whether they are independent aspects of the same phenomenon.
-As @son_hing_merit_2011-1 have pointed out, "People can believe that outcomes ought to be distributed on the basis of merit and yet vary in their perceptions of whether this is how society currently operates" (p.
-435).
-In other words, normative beliefs should be considered while taking perceptions into account: a strong normative belief in meritocracy may mean something totally different to someone perceiving high meritocracy, when compared to someone perceiving low meritocracy.
-To avoid the confusion generated by the term "belief", we propose the terms meritocratic preferences ("what should be"), and meritocratic perceptions ("what is"), as they better reflect the two facets of meritocracy under scrutiny [@castillo_meritocracia_2019].
-
-_c. Non-meritocratic aspects_
-
-A number of studies have dealt with aspects opposed to meritocracy, as for instance the use of personal advantages (as contacts or having a wealthy family) to get ahead in life.
-For instance, @kunovich_systems_2007 used an items’ battery listing a number of factors in relation to “How important each should be in deciding pay...” (as @duru-bellat_whos_2012 for desired meritocracy).
-They consider factors such as education and responsibility as meritocratic, giving them a value of 1 if considered “essential” in the scale response, whereas factors such as having a family or children are valued 1 when rated as "not important at all" (i.e.
-reverse coded).
-The assumption behind this approach is that rejecting a supposed non-meritocratic aspect (as having family and/or children) implies a stronger belief in meritocracy.
-A similar approach of reverse-coding non-meritocratic items was taken by @newman_false_2015, using the same principles applied in the "Preference for the Merit Principle Scale" [@davey_preference_1999].
-
-The assumption that meritocratic and non-meritocratic elements are poles of the same continuum was tested by @reynolds_perceptions_2014 using the "get ahead" perceptions' battery items mentioned above.
-They considered education, ambition, and hard work as meritocratic, and other factors such as family wealth and connections as non-meritocratic.
-Despite making and proving this distinction, however, the authors end up subtracting one dimension from the other, thus coming back to the assumption that they are two poles of the same continuum as @kunovich_systems_2007 did.
-Similarly, @roex_attitudes_2019 used ISSP indicators for perceived meritocracy and non-meritocracy to build a single score by reverse coding the non-meritocratic items.
-Therefore, the treatment of non-meritocratic items has been rather inconsistent across studies and the assumption that they are the simple opposite of meritocracy requires further empirical assessment.
-
-_d. Accounting for measurement error_
-
-Finally, most studies in meritocracy so far have not accounted for the issue of measurement error [@brownBookReviewTimothy2010; @bollenStructuralEquationsLatent1989] (Ansolabehere, S., Rodden, J., & Snyder Jr, J.
-M., 2008), as they use single indicators and/or simple average indexes for measuring meritocracy.
-Such strategy assumes that the latent construct is measured perfectly (i.e.
-no error or residual variance) by the indicators chosen, going as far as to propose that "...
-In choosing this strategy of index construction, we argue that support for meritocracy is not a latent variable" [@kunovich_systems_2007 p.653-654].
-Some advances were made by @reynolds_perceptions_2014 through conducting a principal component analysis of meritocratic and non-meritocratic dimensions, but somewhat contradictorily, they chose a sum index despite proving a multidimensional latent structure.
-
-## An instrument proposal {-#instrumentprop}
-
-Based on the previous assumptions and limitations of the measurements of meritocracy, we propose and test an instrument with the following characteristics:
-
-- _Multidimensionality_, incorporating previous distinctions between preferences and perceptions, as well as between meritocratic and non-meritocratic aspects.
-
-- Multiple indicators for each dimension, in order to _account for measurement error_ in a confirmatory factor analysis context.
-
-- Based on _previous indicators_ as much as possible, for the sake of keeping comparability between studies.
-
-- _Brief_, as to be used in regular public opinion surveys.
-In this respect, it differs from the proposal of  "Preference for the Merit Principle Scale" [@davey_preference_1999], as they use 15 items for just one dimension (aside from the problem of reverse-coding non-meritocratic items).
-
-The proposed measurement framework is depicted in Figure \@ref(fig:merit-model):
-
-```{r merit-model, echo=FALSE, fig.cap = "Model of perception and preferences for meritocracy and non-meritocracy", fig.align='center'}
-knitr::include_graphics("../input/images/generalf.png")
-```
-The columns "Perceptions" and "Preferences" represent the distinction between these two concepts, usually confused under the label "beliefs" [@castillo_meritocracia_2019].
-Perceptions refer to the extent to which people observe that meritocracy works or applies in their society, which relates to items such as "I think hard work is important to get ahead in society".
-Preferences refer to normative expectations that are usually linked to a "should" expression (e.g.
-whether hard work should be related to payment).
-The rows in Figure 1 consider the distinction between meritocratic and non-meritocratic dimensions [@reynolds_perceptions_2014].
-These aspects have usually been treated as different ends of the same continuum in previous research, an assumption that requires empirical scrutiny.
-Non-meritocratic elements usually refer to the use of personal contacts or family advantages to get ahead in life.
-
-Regarding the selection of indicators, most are taken or adapted from previous studies for the sake of comparability.
-For meritocratic indicators, we use effort and talent as the main components of the traditional concept of merit as defined by @young_rise_1962, whereas for non-meritocratic dimensions, we use wealthy parents and good contacts.
-Descriptions of the specific items are outlined in the methodology section.
-
-The research hypotheses behind this conceptualization and measurement model are as follows:
-
-* $H_1$. The perception of meritocracy is a latent variable based on indicators of the importance attributed to talent and effort to get ahead in life.
-
-* $H_2$. The perception of non-meritocracy is a latent variable that derives from  indicators related to agreement with the statement that people with contacts and rich parents manage to get ahead.
-
-* $H_3$. Meritocratic preferences are a latent variable based on the normative value of effort and talent.
-
-* $H_4$. Non-meritocratic preferences are a latent variable based on the normative value of the use of personal contacts and having wealthy parents.
-
-# Methodology
-
-The analysis is organized in three studies.
-Study 1 serves as a background to our instrument proposal by attempting to operationalize the meritocracy dimensions with available data.
-For this we use data from the International Social Survey Programme (ISSP).
-Although we are aware that the available indicators in this dataset do not allow a precise operationalization of the conceptual model, we know that this dataset is the most widely used in comparative meritocracy studies (as well as overall subjective inequality) and therefore it is certainly useful to look at its potentials and limitations to this regard.
-Study 2 is the core of this paper and consists of the proposal of a new scale for measuring meritocratic perceptions and preferences.
-Finally, in Study 3 we perform additional validity analysis of the meritocracy scale.
-
-## Study 1
+## Study 1: Measuring meritocracy with international secondary data
 
@@
 
-The data used in the analysis corresponds to the last available wave of the social inequality module from the International Social Survey Programm (ISSP), which is the most specialized international comparative survey in perceptions, attitudes and beliefs about inequality related issues.
+The data used in the analysis corresponds to the last available wave of the social inequality module from the International Social Survey Programme (ISSP), which is the most specialized international comparative survey in perceptions, attitudes and beliefs about inequality related issues.
 This wave corresponds to the year 2009 and covers attitudes towards a series of topics dealing with social inequality across 26 countries and XXXX individuals.
-Although there is data available from this module for previous waves (1987, 1992 and 1999), unfortunately there are several variables that are important for this research that were not included, particularly in 1987 and 1999, reason why we only use the last wave.
+Although there is data available from this module for previous waves (1987, 1992 and 1999), unfortunately there are several variables that are important for this research that were not included, particularly in 1987 and 1999, reason why we only use the 2009 wave.
 
@@
 ```
-
-## Study 2
-
-## Data collection {-#data}
-
-The data was obtained through an online questionnaire, which was part of a larger study on meritocracy and preferences developed in Chile in 2019 funded by the national scientific agency ANID.
-The questionnaire was programmed in Qualtrics and the fieldwork was conducted by an external online survey agency ([netquest.cl](www.netquest.cl)) between December 2019 and January 2020.
-The sample was selected from a non-probabilistic quota design in three large cities in Chile (Santiago, Concepción & Antofagasta).
-The quotas for gender, age, and education level were generated based on a survey by the Public Studies Center [@cepEncuestaCEPMayo2019], which is a well regarded counterpart agency of the ISSP (International Social Survey Programme) in Chile.
-A total sample of 2,141 people was collected, excluding those who did not answer the questions on the scale, and those who did not accept informed consent.
-There were no significant differences between our sample and the wider population for most socio-demographic characteristics, with the exception of educational level (see Table \@ref(tab:rep-samp) in appendix).
-As is often the case with online surveys, there were some limitations in achieving the quotas for lower educational levels **[citas - LUCHO]**.
-
-## Instrument design
-
-The proposed scale of perceptions and preferences relating to meritocracy consisted of eight indicators that were grouped into the four dimensions listed earlier: Perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic).
-In order to achieve at least some comparability with previous studies, the questions were adapted from the items battery "reasons to get ahead" (ISSP/GSS), which has been widely used for operationalizing meritocracy in previous studies [@mijs_paradox_2019; @duru-bellat_whos_2012; @reynolds_perceptions_2014].
-The aforementioned eight items, ordered according to dimensions, are presented in Table \@ref(tab:table-indicadores).
-These eight likert-type items have five response alternatives, ranging from "Completely disagree"(1) to "Completely agree" (5).
-
-```{r include=FALSE}
-cap <- "Items according to dimension."
-# if (knitr::is_html_output()) {tb <- table_nums("tb",cap)} else {tb <- cap}
-```
-
-```{r table-indicadores, echo=FALSE}
-tab01 <- read.csv(file = "../input/data/original/materials/table01.csv",header = 1,sep = ",",encoding = "UTF-8")
-cnames <- c("Dimensions", "Factor","Statement (english)" , "Statement (spanish)")
-
-kable(tab01,table_format,booktabs = T, linesep = "",col.names = cnames, caption = cap) %>%
-  kable_styling(
-    full_width = T,
-    latex_options = c("hold_position"),
-    position = "center",
-    font_size = 10,
-    bootstrap_options=c("striped", "bordered")) %>%
-  column_spec(column = 1, width = "1.5cm", ) %>%
-  column_spec(column = 2,width = "2 cm") %>%
-  collapse_rows(columns = 1:2,valign = "middle")
-```
-
-\pagebreak
-
-## Administration sets
-
-With the objective of evaluating the effect of indicator ordering in the responses, three different versions of items' order were designed and randomly assigned, as depicted in Figure \@ref(fig:appmod).
-The scale was presented to the first group (_n = 712_) in the order that appears in Table \@ref(tab:table-indicadores) according to perceptions and preferences.
-For the second group (_n = 717_), the order was reorganized according to perceptions and preferences over the same topic, e.g.
-for the topic of hard work, the item about perception was followed by the item about preference, and the same for the rest of the topics.
-Finally, for the third group (_n = 712_), the items were completely randomized.
-
-```{r appmod, echo=FALSE, fig.cap = "Survey flow", fig.align='center'}
-knitr::include_graphics('../input/images/app_mod.png')
-```
-
-# Methods
-
-To test the scale's underlying constructs, we employed confirmatory factor analysis models (CFA).
-The models estimated one factor for each of the proposed dimensions, as represented in the following figure:
-
-```{r meas01, echo=FALSE, fig.cap = "Theoretical model", fig.align='center'}
-knitr::include_graphics('../output/images/meas01.png')
-```
-As in Study 1, CFA was conducted using the `lavaan` R package (version 0.6-3; Rosseel, 2020), with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement (Kline, 2016; Rosseel, 2020).
-The fit indexes and cut-off criteria were the same as in Study 1.
-
-A pre-registration was made in the OSF platform, available at the following link: [https://osf.io/z45y2](https://osf.io/z45y2).
-Included in this pre-registration are the hypotheses regarding the four-dimensional conceptual model underlying the scale, the variable measurement levels, the statistical tests to be performed with their respective evaluation parameters, and other important aspects of the research design.
-
-# Results
-
-## Descriptive analyses
-
-As can be seen in the Table \@ref(tab:desc01), the indicators have values ranging from 1 (totally disagree) to 5 (totally agree).
-Averages are observed from 2.41, corresponding to preference-contacts; to 3.89, corresponding to preference-effort.
-Both indicators are consistent with a general adherence to meritocracy as reported in previous studies, privileging individual aspects such as effort [@bucca_merit_2016; @frei_fruto_2020].
-
-```{r desc01, echo=FALSE}
-cap <- "Descriptive statistics of the scale."
-load(file = "../input/data/proc/dat04.RData")
-dat04b<- dat04 %>% rename("A. Perception Effort" =perc_effort,
- "B. Perception Talent" = perc_talent,
- "C. Perception rich parents" = perc_wpart,
- "D. Perception contacts" = perc_netw,
- "E. Preferences Effort" = pref_effort,
- "F. Preferences Talent" = pref_talent,
- "G. Preferences rich parents" = pref_wpart ,
- "H. Preferences contacts" = pref_netw )
-sum04b <- psych::describe(dat04b[,-c(1)], fast = T) %>% data.frame() %>% select(mean,sd,min,max) %>% round(digits = 2)
-table04b<- sum04b %>% rename("Mean"=mean, "SD"=sd, "Min"=min,"Max"=max)
-sumdat04<- kable(table04b,format =table_format,
-                 booktabs= T,
-                 #exhaust = FALSE,
-                 align = "l",
-                 linesep = "",
-                 caption = cap) %>%
-  kable_styling(full_width = table_format2,
-                font_size = 10,
-                position = "center",
-                latex_options = c("HOLD_position"));sumdat04
-```
-
-The graphs presented in Figure \@ref(fig:plotlikert) display disaggregated and comparable information of the different response categories for each item.
-As it can be observed, in general there is more agreement in the perception of non-meritocratic items than in meritocratic items, while in the case of preferences, the opposite occurs.
-As far as preferences are concerned, the preponderant role of effort over talent as a criterion of meritocratic preference is noteworthy.
-
-```{r plotlikert, echo=FALSE, fig.cap = "Descriptive plot", fig.align='center'}
-knitr::include_graphics('../output/images/plotlikert.png')
-```
-
-In Figure \@ref(fig:corpoly), moderate to high intensity relationships are observed between indicators that correspond to the same factor (e.g., perception of meritocracy by effort and by talent, r=0.56).
-The relationships between meritocratic perceptions and preferences are also salient, which is not the case for non-meritocratic indicators.
-
-```{r corpoly, echo=FALSE, fig.cap = "Polychoric correlation plot", fig.align='center'}
-knitr::include_graphics('../output/images/corpoly.png')
-```
-In sum, descriptive analyses point to a relative adherence to meritocracy, which is expressed as a greater preference for meritocratic criteria, and a lesser preference for non-meritocratic criteria, even when the preference for non-meritocratic elements reaches a proportion of more than 20%.
-Likewise, a relatively low perception of meritocracy is observable.
-In addition, we observed a coherent relationship between the indicators of the same dimension as proposed in the pre-registration of the study; that is, the pairs of items associated with a specific factor show correlations with a large effect size.
-In particular, the associations between effort and talent are relevant, challenging previous assumptions that talent would not be a meritocratic criterion [@mijs_paradox_2019], otherwise the correlation would be zero or negative.
-In addition, we observe that there is no negative correlation between meritocratic and non-meritocratic aspects, undermining the assumptions of previous studies that suggested that these dimensions would be the opposite poles of one same continuum [@reynolds_perceptions_2014].
-
-## Confirmatory Factor Analysis
-
-This section estimates the fit of the conceptual model behind the Perceptions and Preferences for Meritocracy Scale.
-For this, we first estimate a confirmatory factor analysis model for the whole sample, and secondly we test the order effects applying the same model to each of the three order permutations.
-
-Figure \@ref(fig:meas02) shows the results of the estimation for the four-factor model using the complete sample.
-This model shows adequate fit indicators (CFI = 0.977, RMSEA = 0.079, $\chi2$(df = 14) = 28,03, p = 0.014) and it presents factorial loads with weights that vary between 0.63 and 0.88.
-
-```{r meas02, echo=FALSE, fig.cap = "Confirmatory factor analysis of the Perceptions and Preferences for Meritocracy Scale", fig.align='center'}
-knitr::include_graphics('../output/images/meas03.png')
-```
-Attending to the correlations between the latent variables as depicted in Figure \@ref(fig:latcor), it is observed that the variables with the strongest correlations are meritocratic perceptions and preferences, whit an $r=0.59$, while non-meritocratic perceptions and preferences show a correlation close to 0.
-Besides, and as commented before in the analysis of the single items correlations, the latent variables of meritocratic and non-meritocratic versions of both perceptions and preferences show very low correlations.
-This last finding gives further evidence regarding the lack of unidimensionality of meritocratic and non-meritocratic aspects: they do not appear as the opposite poles of the same continuum and based on this evidence the use of reverse-coding for non-meritocratic items as indicators of meritocracy should be avoided.
-
-```{r latcor, echo=FALSE, fig.cap = "Latent variables correlation", fig.align='center'}
-knitr::include_graphics('../output/images/latcor.png')
-```
-In a second step we test the effect of the items' order in the confirmatory model.
-Table \@ref(tab:tab03) shows the fit indicators of the models estimated for each of the three versions of the items' order described in the methodology, being the first one according to perceptions and preferences, the second one according to meritocratic and non-meritocratic aspects, and the third one completely randomized.
-Regardless of the group and order of items, all models obtained an adquate fit, with CFI's above 0.95 and RMSEA's below 0.08.
-However, none of the models achieved a non-significant chi-square, something expected in large samples as the one used here).
-Model 1 shows the fit indicators for the estimation of the CFA model to the first version order (perceptions-preferences), which was the one obtaining best fit (CFI=0.993, TLI=0.995, RMSEA=0.034, $\chi2$(df=14)=42.276), whereas Model 2 for the fixed order according to merit/non-merit items shows the comparatively worst indicators.
-The CFA fit indices for the completely randomized items' order (Model 3) is somehow an intermediate fit indexes when compared with the previous two.
-Still, it keeps all the indicators within the acceptable cut-off criteria and besides it controls for possible order effects in the administration of the instrument.
-
-```{r tab03, echo=FALSE,eval=TRUE}
-load(file = "../output/tables/fit4_o.RData")
-load(file = "../output/tables/fit3_o.RData")
-load(file = "../output/tables/fit2_o.RData")
-load(file = "../output/tables/fit1_o.RData")
-sum_fit<- bind_rows(fitmeasures(fit1_o)[c("chisq","df","cfi","tli","rmsea")],
-                    fitmeasures(fit2_o)[c("chisq","df","cfi","tli","rmsea")],
-                    fitmeasures(fit3_o)[c("chisq","df","cfi","tli","rmsea")])
-sum_fit$mod <- c("Model 1","Model 2","Model 3")
-sum_fit$est <- c("DWLS","DWLS","DWLS")
-sum_fit$nobs <- c(nobs(fit1_o),nobs(fit2_o),nobs(fit3_o))
-sum_fit <- select(sum_fit,mod,nobs,est,everything())
-colnames <- c("Model","$N$","Estimator","$\\chi^2$","df","CFI","TLI","RMSEA")
-sumtable02<- kable(sum_fit,digits = 3,format = table_format,row.names = F,booktabs=T, align = c("lcccccc"),
-                   caption = "Summary fit indices according group",
-                   col.names = colnames,
-                   escape = FALSE) %>%
-  kable_styling(full_width = F,
-                font_size = 10)  %>%
-      footnote(number = c("Model 1: fixed order by perception/preference",
-                          "Model 2: fixed order by merit-non merit",
-                          "Model 3: Randomized order"));sumtable02
-```
-
-## Additional validity analyses
-
-We performed two further analyses in order to add evidence regarding the quality of the scale.
-Firstly we analyzed the convergent validity by exploring the association of the merit scale with related concepts and itscorrespondent measurements.
-Secondly, we test the internal consistency of the measurement model behind the scale comparing with a different sample through measurement invariance procedures .
-
-### Convergent validity
-
-_Data_
-
-For this analysis we examined data from one wave of the same study described above, which was a three-wave panel survey.
-The last wave of the study included additional measures that allow testing the correlations of the merit scale with some related constructs as opportunity beliefs and personal wherewithal.
-After listwise deletion of missing cases in socio-demographics information, a total of 1422 individuals took part in the third wave: 668 (46.97%) women and 754 (53.03%) men.
-The ages of 50.57% of them are 45 years or older, and 36.28% held a tertiary degree.
-
-_Intruments_
-
-The following scales were included for testing their correlations with the the 8-item meritocracy scale:
-
-- Opportunity beliefs: We used two items of the social inequality module of the International Social Survey Programme.
-Studies use generally both indicators to measure meritocratic and non-meritocratic beliefs [@McCall2017Exposurerisinginequality; @mijs_paradox_2019].
-One item measures the importance of individual factors in determining life outcomes by asking for the importance of hard work for getting ahead in life (M=3.76, SD=0.93).
-The second item asks for the importance of coming from a wealthy family and captures the importance of structural factors in getting ahead (M=2.75, SD=1.27).
-Participants responded to each item using a 1 (not important at all) to 5 (essential) response scale.
-We call the first item \textit{hard work} and the second one \textit{social origin}.
-
-- Personal wherewithal: the questionnaire included The Neoliberal Beliefs Inventory (NBI), which considers four factors: Government Interference preferences, Competition preferences, System Inequality perception and Personal Wherewithal [@bay-chengTrackingHomoOeconomicus2015].
-We used the factor _personal wherewithal_ that reflects meritocracy beliefs in terms of the importance of personal attributes as strength and skills to yield success.
-The construct consists of 8 items (e.g., “Any goal can be achieved with enough hard work and talent”, “I’ve benefited from working hard, so there’s no reason others can’t”, “Anyone who is willing to work hard can be succesful in Chile”).
-Participants answered the items using a 1 (totally disagree) to 6 (totally agree) scale.
-We simple computed mean scores, with higher scores indicating stronger support for meritocracy (M=3.32, SD=0.88, $\alpha$ = 0.89).
-
-Both instruments cover mainly the perceptual side of the concept.
-Therefore, in general we expect larger correlation with meritocratic perceptions than with preferences.
-In the case of opportunity beliefs, the hard work item should correlate positively with meritocratic perception whereas the importance for coming from a wealthy family is expected to correlate positively with non-meritocratic perceptions.
-Regarding the second instrument of personal wherewithal, we also anticipate a positive correlation with meritocratic perception and low to null correlations with the other dimensions of the merit scale.
-
-_Results_
-
-We used polyserial correlations to evaluate relationships of meritocractic and non-meritocratic preferences and perceptions with the items of the opportunity beliefs battery.
-For personal wherewithal, we examined Pearson correlations.
-
-Table \@ref(tab:val-conv) shows the polyserial and pearson correlations between our meritocracy scale and common measures of meritocracy.
-The findings provide evidence for convergent validity for the scale.
-As we expected, preferences show very weak correlations.
-Findings also indicate that correlations of perceptions with opportunity beliefs are consistent with our expectations.
-The item for social origin shows a positive and moderate association with unmeritocratic perception (r=.354), and a negative and low correlation with meritocratic perception (-0.197).
-Regarding the item of hard work, the correlation with meritocratic perception is positive but weak (r=.187).
-One explanation for this result is that the support for hard work is very high among respondents---only 8.6 percent indicates _not very important_ or _not important at all_.
-
-  |                                      | Social Origin  |     Hard Work    |     Wherewithal    |
-  |--------------------------------------|----------------|------------------|--------------------|
-  |     Meritocratic     Perception      |     -0.1975    |     0.1877       |     0.3921         |
-  |     Unmeritocratic    Perception     |     0.3547     |     -0.0775      |     -0.1844        |
-  |     Meritocratic    Preference       |     0.1078     |     0.0779       |     0.0819         |
-  |     Unmeritocratic    Preference     |     -0.0397    |     0.0824       |     0.2082         |
-  Table: `r as.character(paste("(\\#tab:val-conv)", "Polyserial and Pearson correlation whit other merit scales"))`.
-
-The measure of NBI for _personal wherewithal_ also provides evidence for the convergent validity of the merit scale.
-Results indicate a positive correlation of this measure with meritocratic preference (r=.392), suggesting that both variables capture similar concepts.
-The NBI's factor depicts a positive association with unmeritocratic preference (r=.208).
-This finding is consistent with literature, inasmuch as the belief for meritocray is associated with the justification of inequality [@McCall2017Exposurerisinginequality; @MadeiraPrimesConsequencesSystematic2019].
-Furthermore, results of our confirmatory factor analysis indicate that meritocratic perceptions and unmeritocratic preferences show a positive and weak correlation.
-
-### Measurement invariance
-
-The modeling for invariance testing has been gaining more relevance in sociological survey studies, mainly due to the interest in the validity of measurement scales in comparative studies in various social and cultural contexts [@Davidovetal2014].
-In this section we will use an analysis of invariance between groups, where the proposed factorial structure is expected to be independent of belonging to the sample corresponding to the original study, comparing it with another more recent study available and where the meritocracy scale was also incorporated.
-
-The invariance measurement procedure consists of a series of nested models to which restrictions are progressively incorporated on the parameters of the measurement model.
-The literature generally suggests that this modeling should be done through four levels or types of progressive restriction [@Milfont2010; @millsap_Statistical_2011; @vandeschoot_checklist_2012]:
-
-1. **Configural:** the model is estimated only indicating the factorial structure used in the CFA.
-
-2. **Weak:** an equality restriction is applied to the factor loadings in the different groups, that is, the loadings are forced to be identical in both measurements.
-
-3. **Strong:** equality restrictions are added to the intercepts of each indicator.
-
-4. **Strict:** equality restrictions are added to the error variances of each indicator.
-
-*Data*
-
-For testing the invariance we compared the data described previously with a new data source coming from an online survey carried out during the first half of 2020.
-The characteristics of this sample in terms of application and coverage are equivalent to those of the previous study.
-The final sample obtained contains 1,242 cases, where 605 (48.71%) are women and 637 (51.28%) are men.
-48.84% of them are 45 years of age or older, and 33.97% have tertiary education or higher.
-
-_Instruments_
-
-The items of the meritocracy scale were identical to the original.
-The items were administrated in a randomized order (which corresponds to the application modality of Group 3 present in Figure \@ref(fig: appmod)).
-
-_Results_
-
-Tabla \@ref(tab:fcloads-inv) shows de results of the CFA estimation of the meritocracy scale in this new sample.
-As observed, the loadings and the fit indices show similar results as the ones presented for the original study above, which gives a first base with wich to start the invariance analysis.
-
-The first step for invariance testing is the estimation of the configural model, which serves as the baseline for further comparison and it is expected to adequately meet the global fit criteria of a measurement model.
-Although the chi-square statistic is used as a global measure of fit,since its sensitivity to sample size it is recommended to use three additional fit indices: Comparative Fit Index (CFI), which should have a value greater than 0.95; Root Mean Square Error Approximation (RMSEA), which must be in the range of values ​​of 0.05 and 0.08; and the Standardized Root Mean Square Residuals (SRMR) which must be less than 0.08 [@vandeschoot_checklist_2012].
-Besides, the literature on measurement invariance suggests some complementary approaches for the evaluation of the fit, from which we will consider the incremental adjustment of the fit indexes [@cheung_Evaluating_2002; @Milfont2010; @dimitrov_Testing_2010] and the ANOVA test for means comparison in nested models [@newsom_Longitudinal_2015].
-
-```{r invargroup, eval=TRUE, echo=FALSE}
-load(file = "../output/tables/tb.inv-group.RData")
-col.nam <- c("Model","$\\chi^2 (\\text{df})$","CFI","RMSEA (90 CI)",
-             "$\\Delta \\chi^2 (\\Delta \\text{df}$)","$\\Delta \\text{CFI}$","$\\Delta \\text{RMSEA}$","Decision")
-footnote <- paste0("N = ",nobs(conf),"; Group 1, n = ",conf@Data@nobs[[1]],"; Group 2, n = ",conf@Data@nobs[[2]], ", ***p < 0.001")
-
-kable(tab.inv, col.names = col.nam, align = "l",
-      booktabs=TRUE,format = table_format,escape = FALSE,
-      caption = "Multiple Group meassurement invariance for Perceptions and Preferences for Meritocracy") %>%
-  kable_styling(full_width = FALSE,latex_options = "HOLD_position") %>%
-  footnote(general = footnote, footnote_as_chunk = T)
-```
-
-Table \@ref(tab:invargroup) shows the results of the measurement invariance estimation.
-When attending to the traditional invariance test of $\Delta \chi^2 (\Delta \text{df})$, the results  support the invariance at the strong level meaning that the fit of the factor model of the merit scale is equivalent across samples when constraining factor loadings and intercepts to being equal.
-Such result is considered in general as evidence of invariance [citar paper Fischer en SJR], as strict forms of measurement invariance rarely hold [citar van der Schoot et al].
-Still, the comparability of latents means requires strict invariance which in this case does not hold when considering $\Delta \chi^2 (\Delta \text{df})$.
-Nevertheless, the criteria of $\Delta \text{CFI}$ used for comparing models is close to the rejection criteria of >.01, whereas the $\Delta \text{RMSEA}$ fulfills the requirements of being below of the cut-off criteria as suggested by [ @chen 2007 ].
-Therefore, using this last standard the level for strict invariance holds for the meritocracy scale.
-
-# Conclusions
-
-Studies that attempt to characterize and compare societies by their support for meritocratic beliefs have used different approaches.
-As most studies use secondary survey data, they tend to assume that the available indicators represent an underlying meritocratic construct.
-A review of these studies reveals several non-tested assumptions, as well as the use of similar indicators to represent different constructs and dimensions of meritocracy.
-As the existence of heterogeneous approaches certainly has consequences for the advancement of the study of meritocracy, this paper presented a comprehensive conceptual framework for the empirical study of meritocracy, building upon previous research.
-This framework was then tested against new survey data.
-
-We identified four critical aspects regarding the measurement of meritocracy in previous studies: unidimensionality, the ambiguous use of the term “beliefs”, the use of non-meritocratic indicators as opposed to meritocratic indicators, and the consideration of measurement error.
-The proposed 8-item scale, “Perceptions and Preferences for Meritocracy”, was designed and tested in order to deal with these four issues.
-Our results indicate that perceptions and preferences seem to be two related but different dimensions, often confounded in previous research under the label of "beliefs".
-Meritocratic and non-meritocratic dimensions do not appear to constitute poles of the same continuum, as some previous studies have assumed.
-Regarding the possible effects of items order in the estimation of latent variable, we tested three different order versions and found evidence that suggest the use of the randomized version of the scale.
-Furthermore, we found evidence of convergent validity as well as measurement invariance that approach well the requirement of comparability between samples.
-
-The four-dimensional framework and structure of the Perceptions and Preferences for Meritocracy scale opens several avenues for future research.
-For instance, distinguishing perceptions from preferences will allow us to evaluate the extent to which different societies are accustomed to, or satisfied with, the perceived level of meritocracy, in terms of differences between what is perceived and what is preferred.
-Additionally, given that non-meritocratic factors are not necessarily related to meritocratic ones, our framework makes it possible to assess the perceived legitimacy of practices such as the use of personal contacts and their interference (or not) with meritocratic ideals in different societies.
-The consideration of structural factors, both individual and societal, will allow us to advance future hypotheses relating to meritocratic legitimacy to a more comprehensive level than the current one, which is based solely on meritocratic perceptions.
-Furthermore, the impacts of different configurations of the four-dimensional framework on practices and behaviors such as corruption, civic involvement, and political alignment, is an area that requires additional research.
-Such future agendas could be specially relevant in times of economic crisis and growing inequalities, that could entail changes in the legitimation of the current distributive structure based on meritocratic ideals.
-
-\pagebreak
-
-# References
-
-<div id="refs"> </div>  <!-- Although <div> is an HTML tag, this method also works for other output formats such as PDF. -->
-
-# (APPENDIX) Appendix {-}
-
-`r if (knitr::is_latex_output()){ '\\appendix'}`
-
-`r if (knitr::is_latex_output()){ '\\section{Appendix}'}`
-
-```{r tab-loads, echo=FALSE}
-load(file = "../output/tables/tb.fm.RData")
-cap <- "Factor loads and model fit."
-tb.col  <- c("Variables",1:4,1:4,1:4)
-tb.caption <- "Factor loadings and fit measures"
-tb.foot <- paste0("Standardized factor loadings using DWLS estimator ; CFI = Comparative fit index (scaled); RMSEA = Root mean square error of approximation (scaled)")
-
-tb.fm <- tb.fm %>%
-    mutate(vars = recode(vars,
-                         "A. Who the more they try they manage to get bigger rewards that those who strive less."= "Perception Effort",
-                         "B. Who possess more talent they manage to obtain greater rewards than those who possess less talent."= "Perception Talent",
-                         "C. Who they have rich parents manage to get out ahead."="Perception rich parents",
-                         "D. Who they have good contacts they manage to get out ahead."="Perception contacts",
-                         "E. Who the more they try they should get greater rewards than those who they try less." = "Preferences Effort",
-                         "F. Who possess more talent they should get greater rewards than those who possess less talent."= "Preferences Talent",
-                         "G. It's fine that those who have rich parents get ahead" ="Preferences rich parents",
-                         "H. Is well that those who have good contacts get ahead." = "Preferences contacts"))
-
-tb.fm01<- kable(tb.fm, format = table_format, booktabs = T,linesep = "", escape = FALSE,align = "lcccccccccccc",col.names = tb.col, caption = cap) %>%
-          kable_styling(full_width = table_format2,latex_options = c("hold_position","scale_down"), position = "center",font_size = 9) %>%
-          add_header_above(header = c(" "=1,"Version 1"= 4,"Version 2"= 4,"Version 3"= 4)) %>%
-          add_header_above(header = c(" "=1,"Factor loadings"= 12)) %>%
-          row_spec(row = 8,hline_after = TRUE) %>%
-          add_indent(c(9:12))  %>%
-          footnote(general = tb.foot,threeparttable = T);tb.fm01
-```
-
-```{r fcloads-inv, echo=FALSE}
-load(file = "../output/tables/tb.fact-f02.RData")
-tb.foot <- paste0("Standardized factor loadings using DWLS estimator ; CFI = Comparative fit index (scaled); RMSEA = Root mean square error of approximation (scaled)")
-tb.fm02<- kable(tb.fm2,escape = FALSE,align = "lcccc",col.names = tb.col, caption = tb.caption, booktabs = T,linesep = "",format = table_format) %>%
-          kable_styling(full_width = table_format2,latex_options = c("hold_position"), position = "center",font_size = 9) %>%
-          add_header_above(header = c(" "=1,"Factor loadings"= 4)) %>%
-          row_spec(row = 8,hline_after = TRUE) %>%
-          add_indent(c(9:12)) %>%
-          footnote(general =tb.foot ,footnote_as_chunk = T);tb.fm02
-```
-
-\pagebreak
-
-|                     | Sample  |   CEP  |
-|---------------------|:-------:|:------:|
-| __Gender__          |         |        |
-| Men                 | 49,82%  | 50,52% |
-| Women               | 50.18%  | 49,47% |
-| __Age__             |         |        |
-| 18 - 24             | 18,55%  | 18,17% |
-| 25 - 34             | 18,86%  | 17,48% |
-| 35 - 44             | 19.09%  | 19,98% |
-| 45 - 54             | 17,96%  | 19,23% |
-| 55 - or more        | 25,54%  | 25.11% |
-| __Education__       |         |        |
-| Primary  or less    | 2,93%   | 15,88% |
-| High school          | 43,23%  | 37,04% |
-| Non university      | 32,63%  | 28,93% |
-| university or more  | 21,21%  | 18,13% |
-Table: `r as.character(paste("(\\#tab:rep-samp)", "Representativeness of the sample"))`.

### 2021-05-25 paper_study2.Rmd
- major update

---
title: "Measuring Perceptions and Preferences for Meritocracy"
css: "custom.css"
linestretch: '1.5'
link-citations: yes
output:
  bookdown::pdf_document2:
    template: null
    toc: false
  bookdown::html_document2:
    number_sections: false
linkcolor: blue
bibliography:
#  - ../input/bib/EconomiaMoral.bib
#  - ../input/bib/meritocracy.bib
#  - ../input/bib/invariance.bib
editor_options:
  chunk_output_type: console
geometry: margin=0.78in
header-includes:
  # - \usepackage[spanish,es-tabla,es-nodecimaldot]{babel}
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
---

```{r eval=FALSE, include=FALSE}
# for render in pdf run rmarkdown::render_site("docs/paper.Rmd", output_format = "all")
# clean #in the yml
rmarkdown::render("paper_study2.Rmd", output_format = "bookdown::pdf_document2")
rmarkdown::render("paper.Rmd", output_format = "bookdown::html_document2")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = FALSE,out.width = '85%',fig.pos= "H")
# knitr::opts_knit$set(base.url = "../") #relative path for .html output file
# knitr::opts_knit$set(root.dir = "../") #relative path for chunks within .rmd files
options(scipen=999)
rm(list=ls())
options(knitr.kable.NA = '')
options(knitr.graphics.error = FALSE)
Sys.setlocale("LC_ALL", "ES_ES.UTF-8")
```

```{r message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
#RUN THIS BEFORE THE ANALYSIS
pacman::p_load(knitr)
table_format = if(is_html_output()) {
  "html"
} else if(is_latex_output()) {
  "latex"
}

table_format2 = if(is_html_output()) {
  T
} else if(is_latex_output()) {
  F
}
```

## Study 2: The Perceptions and Preferences for Meritocracy Scale

## Data collection {-#data}

The data was obtained through an online questionnaire, which was part of a larger study on meritocracy and preferences developed in Chile in 2019 funded by the national scientific agency ANID. The questionnaire was programmed in Qualtrics and the fieldwork was conducted by an external online survey agency ([netquest.cl](www.netquest.cl)) between December 2019 and January 2020. The sample was selected from a non-probabilistic quota design in three large cities in Chile (Santiago, Concepción & Antofagasta). The quotas for gender, age, and education level were generated based on a survey by the Public Studies Center [@cepEncuestaCEPMayo2019], which is a well regarded counterpart agency of the ISSP (International Social Survey Programme) in Chile. A total sample of 2,141 people was collected, excluding those who did not answer the questions on the scale, and those who did not accept informed consent. There were no significant differences between our sample and the wider population for most socio-demographic characteristics, with the exception of educational level (see Table \@ref(tab:rep-samp) in appendix). As is often the case with online surveys, there were some limitations in achieving the quotas for lower educational levels **[citas - LUCHO]**.

## Instrument design

The proposed scale of perceptions and preferences relating to meritocracy consisted of eight indicators that were grouped into the four dimensions listed earlier: Perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic). In order to achieve at least some comparability with previous studies, the questions were adapted from the items battery "reasons to get ahead" (ISSP/GSS), which has been widely used for operationalizing meritocracy in previous studies [@mijs_paradox_2019; @duru-bellat_whos_2012; @reynolds_perceptions_2014]. The aforementioned eight items, ordered according to dimensions, are presented in Table \@ref(tab:table-indicadores). These eight likert-type items have five response alternatives, ranging from "Completely disagree"(1) to "Completely agree" (5).

```{r include=FALSE}

# if (knitr::is_html_output()) {tb <- table_nums("tb",cap)} else {tb <- cap}
```

```{r table-indicadores, echo=FALSE}
pacman::p_load(knitr, kableExtra, dplyr)

tabitems <- read.csv(file = "input/tables/table01.csv",header = 1,sep = ",",encoding = "UTF-8") # call file generated externally for the table
cnames <- c("Component", "Dimensions","Item (english)" , "Item original (spanish)")
cap <- "Items of the perceptions and preferences for meritocracy scale."

kable(tabitems, table_format, booktabs = T, linesep = "",col.names = cnames, caption = cap) %>%
  kable_styling(
    full_width = T,
    latex_options = c("hold_position"),
    position = "center",
    font_size = 12,
    bootstrap_options=c("striped", "bordered")) %>%
  column_spec(column = 1, width = "1.5cm", ) %>%
  column_spec(column = 2,width = "2 cm") %>%
  collapse_rows(columns = 1:2,valign = "middle")
```

\pagebreak

## Administration sets

With the objective of evaluating the effect of indicator ordering in the responses, three different versions of items' order were designed and randomly assigned to the respondents, as depicted in Figure \@ref(fig:appmod). The scale was presented to the first group (_n = 712_) in the order that appears in Table \@ref(tab:table-indicadores) according to perceptions and preferences. For the second group (_n = 717_), the order was reorganized according to perceptions and preferences over the same topic, e.g. for the topic of hard work, the item about perception was followed by the item about preference, and the same for the rest of the topics. Finally, for the third group (_n = 712_), the items were presented as completely randomized.

```{r appmod, echo=FALSE, fig.cap = "Survey flow", fig.align='center'}
knitr::include_graphics('output/images/app_mod.png')
```

# Methods

To test the scale's underlying constructs, we employed confirmatory factor analysis models (CFA). The models estimated one factor for each of the four proposed dimensions presented in Table \@ref(tab:table-indicadores). As in Study 1, CFA was conducted using the `lavaan` R package (version 0.6-3; Rosseel, 2020), with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement (Kline, 2016; Rosseel, 2020). The fit indexes and cut-off criteria were the same as the ones used in Study 1.

A pre-registration was made in the OSF platform, available at the following link: [https://osf.io/z45y2](https://osf.io/z45y2). Included in this pre-registration are the hypotheses regarding the four-dimensional conceptual model underlying the scale, the variable measurement levels, the statistical tests to be performed with their respective evaluation parameters, and other important aspects of the research design.

# Results

## Descriptive analyses

The graphs presented in Figure \@ref(fig:plotlikert) display disaggregated and comparable information of the different response categories for each item. As it can be observed, in general there is more agreement in the perception of non-meritocratic items than in meritocratic ones, while in the case of preferences the opposite occurs. As far as preferences are concerned, the preponderant role of effort over talent as a criterion of meritocratic preference is noteworthy.

```{r produce-likert_s2}
pacman::p_load(dplyr)
load(file = "input/data/proc/data_s2.RData")
names(data_s2)
dat_likert_s2 <- data_s2 %>% select(!group) # exclude group variable
dat_likert_s2 <- sjmisc::rec(dat_likert_s2, rec="rev", append=FALSE)
names(dat_likert_s2) # reverse coding for likert graph

table(data_s2$perc_effort)
table(dat_likert_s2$perc_effort_r) # check ok.

plotlikert_s2<- sjPlot::plot_likert(dat_likert_s2,
                        c(1, 1, 1, 1, 2, 2, 2, 2),
                        groups.titles = c("Perceptions", "Preferences"),
                        geom.colors = "PuBu",
                        geom.size = 0.8,
                        axis.labels = c("Effort", "Talent", "Rich Parents", "Contacts"),
                        catcount = 4,
                        cat.neutral = 3,
                        grid.range  =  c (1.2 , 1.4),
                        values  =  "sum.outside",
                        reverse.colors = T,
                        reverse.scale = F,
                        show.n=FALSE)
```

```{r plotlikert, echo=FALSE, fig.cap = "Distribution of responses in the Merit Scale items", fig.align='center'}
plotlikert_s2
```

Attending now to the association among the scale items, Figure \@ref(fig:corpoly) shows items polychoric correlations. There are three main aspects to highlight from this correlation matrix.  Firstly, as expected the largest correlations are between indicators that correspond to the same factors behind the conceptual model (e.g., perception of meritocracy by effort and by talent, r=0.52). Secondly, among this correlations the highest are those between the non- meritocratic dimension buth in perceptions (r=0.73) and preferences (r=0.61). thirdly, both items for meritocratic preferences (E and F) are the ones that mostly correlate with the rest of the perceptual items, showing medium to high correlations. This is noteworthy because it indicates that even the perception of non-meritocracy would be related to larger meritocratic preferences. Finally, we observe that there are no considerable negative correlations between meritocratic and non-meritocratic aspects, undermining the assumptions of previous studies that suggested that these dimensions would be the opposite poles of one same continuum [@reynolds_perceptions_2014].

```{r corpoly, echo=FALSE, fig.cap = "Perceptions & preferences for meritcoracy items' polychoric correlations", fig.align='center'}
load(file = "input/data/proc/data_s2.RData")

# generate polychoric matrix
cor_s2<- data_s2 %>% select(perc_effort,perc_talent,perc_wpart,perc_netw,pref_effort,pref_talent,pref_wpart,pref_netw) %>% lavaan::lavCor(., ordered=names(.))

diag(cor_s2) = NA

# windowsFonts(A = windowsFont("Times New Roman"))

rownames(cor_s2) <-c(
    "A. Perception Effort",
    "B. Perception Talent",
    "C. Perception Rich parents",
    "D. Perception Contacts",
    "E. Preferences Effort",
    "F. Preferences Talent",
    "G. Preferences Rich parents",
    "H. Preferences Contacts")
colnames(cor_s2) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)", "(H)")

corrplot::corrplot(cor_s2,
  method = "color",
  addCoef.col = "#000390",
  type = "upper",
  tl.col = "black",
  col=colorRampPalette(c("white","#0068DC"))(8),
  bg = "white",
  na.label = "-",
  lower.col = "black")

```

## Confirmatory Factor Analysis

The present sections shows the results of the confirmatory factor analysis estimation. The model estimates four latent factors: perception meritocratic, perception non-meritocratic, preferences meritocratic and preferences non-meritocratic. Each factor is estimated based on two items of the scale as detailed in Table \@ref(tab:table-indicadores).

The first step in the analysis consists of comparing the model fit indicators for the three versions of the scale that were randomly assigned to the participants: order according to perceptions/preferences, order according to topics and random order (see Figure \@ref(fig:appmod)).

```{r model-generation-cfa}
# Libraries & data
pacman::p_load(lavaan)
load(file = "input/data/proc/data_s2.RData")
names(data_s2)

# model
model_cfa <- '
perc_merit =~ perc_effort+perc_talent
perc_nmerit=~perc_wpart +perc_netw
pref_merit =~ pref_effort+pref_talent
pref_nmerit=~pref_wpart +pref_netw
'

# estimation for each order set
fit_cfa1 <- cfa(model_cfa, data = data_s2[data_s2$group == 1,], ordered = TRUE, std.lv=FALSE)
fit_cfa2 <- cfa(model_cfa, data = data_s2[data_s2$group == 2,], ordered = TRUE, std.lv=FALSE)
fit_cfa3 <- cfa(model_cfa, data = data_s2[data_s2$group == 3,], ordered = TRUE, std.lv=FALSE)

```

```{r table_cfa_fits, echo=FALSE}
pacman::p_load(dplyr, kableExtra, knitr)

# extract fit indices from models and add to table
sum_fit<- dplyr::bind_rows(fitmeasures(fit_cfa1)[c("chisq","df","cfi","tli","rmsea")],
                    fitmeasures(fit_cfa2)[c("chisq","df","cfi","tli","rmsea")],
                    fitmeasures(fit_cfa3)[c("chisq","df","cfi","tli","rmsea")])

# Customize object
sum_fit$mod <- c("Version 1","Version 2","Version 3")
sum_fit$nobs <- c(nobs(fit_cfa1),nobs(fit_cfa2),nobs(fit_cfa3))
sum_fit$est <- c("DWLS","DWLS","DWLS")
sum_fit <- dplyr::select(sum_fit,mod,nobs,est,everything())
colnames <- c("Model","$N$","Estimator","$\\chi^2$","df","CFI","TLI","RMSEA")

# Create table
table_cfa_fits <-kable(sum_fit, format="html", digits=3, booktabs=T, col.names=colnames,  caption = "Summary fit indices according to order versions", escape = FALSE) %>% kable_styling(full_width = F, font_size = 10)
table_cfa_fits
```

Table \@ref(tab:table_cfa_fits) shows the fit indicators of the models estimated for each of the three versions of the items' order described in the methodology. Regardless of the version, all models obtained adequate fit indicators, with CFI's above 0.95 and RMSEA's below 0.08. However, none of the models achieved a non-significant chi-square, something expected in large samples as the one used here. The first version order (perceptions-preferences) was the one obtaining best fit (CFI=0.993, TLI=0.995, RMSEA=0.034, $\chi2$(df=14)=42.276), whereas version 2 with the fixed order according to merit/non-merit items shows the comparatively worst indicators. The CFA fit indices for the completely randomized items' order (Model 3) it keeps all the indicators within the acceptable cut-off criteria and besides it controls for possible order effects in the administration of the instrument. The model and parameter estimates for this version are depicted in Figure \@ref(fig:meas02):

```{r meas02, echo=FALSE, fig.cap = "Confirmatory factor analysis of the Perceptions and Preferences for Meritocracy Scale", fig.align='center'}

# estimation for diagram: summary(fit_cfa3, standardized = TRUE, fit.measures=TRUE)
knitr::include_graphics('output/images/cfa_randomized_study2.png')
```

Attending to the correlations between the latent variables as depicted in Figure \@ref(fig:meas02), it is observed that meritocratic preferences are correlated with preferences, both meritocratic ($r=0.457$) and non-meritocratic ($r=0.500$). The correlation between both types of perceptions ($r=-0.044$) and both types preferences ($r=0.185$) are low, as well as between non-meritocratic preferences and perceptions ($r=-0.059$). This last finding gives further evidence regarding the lack of unidimensionality of meritocratic and non-meritocratic aspects: they do not appear as the opposite poles of the same continuum and based on this evidence the use of reverse-coding for non-meritocratic items as indicators of meritocracy should be avoided.

### 2021-05-25 paper_study3.Rmd
- major update

--- 2021-03-23 docs/paper.Rmd
@@
 ```{r message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
-pacman::p_load(knitr, kableExtra, dplyr, lavaan)
-```
-
-```{r echo=FALSE }
+#RUN THIS BEFORE THE ANALYSIS
+pacman::p_load(knitr)
 table_format = if(is_html_output()) {
@@
 
-# Introduction
-
-Economic inequality has become an issue of growing concern around the world.
-This has been expressed in a series of protests, as well as diverse critical analyses regarding the development of capitalism and its consequences [@streeckPoliticsPublicDebt2014; @piketty_Capital_2014].
-In this context, the study of views, preferences and perceptions of inequality has acquired relevance in the social sciences, in topics such as redistributive preferences [@alesina_Fairness_2005; @dimick_Models_2018], the legitimization of economic inequality [@schroder_Income_2017] and the functioning of meritocracy [@duru-bellat_whos_2012; @mijs_paradox_2019; @reynolds_perceptions_2014; @atriaEconomicElitesAttitudes2020].
-
-In general, meritocracy is defined as a system of distribution of resources and rewards based on individual merit, which in its original conception is a combination of talent and effort [@young_rise_1962].
-This traditional conception of merit places in a secondary position the possible interference of structural or non-meritocratic factors, such as inheritance, personal contacts, and luck [@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995; @yairMeritocracy2007; @landWeSatTable2006; @youngRiseMeritocracy1994].
-Social psychology and sociology have studied the characteristics and consequences of beliefs in meritocracy, under the general hypothesis that a greater belief in meritocracy emphasizes the role of individual over structural factors in personal achievements, leading to greater legitimization of inequalities [@premingerMeritocracyServiceEthnocracy2020; @trumpWhenWhyEconomic2020; @hadjar_meritokratie_2008; @MadeiraPrimesConsequencesSystematic2019].
-Such research has raised criticism of meritocracy as a moral standard of distribution, arguing that it would be an unfulfilled promise given the still preponderant weight of non-meritocratic elements upon individual status and social mobility[@sandelTyrannyMeritWhat2020a; @witteveenReconsideringMeritocraticPower2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019; @khan_Privilege_2013].
-
-Due to the role that meritocratic beliefs play in the justification of individual achievement or failure in contemporary societies [@hadjar_meritokratie_2008, @markovits_meritocracy_2019; @sandel_tyranny_2020], multiple studies have evaluated the relationship between meritocratic beliefs and personal and/or contextual characteristics.
-For example, some studies have linked meritocracy to the reinforcement of socio-economic, gender, and ethnic stereotypes [@MadeiraPrimesConsequencesSystematic2019; @girerdNeoliberalismIdeologicalBarrier2020; @premingerMeritocracyServiceEthnocracy2020], as well as the effects of meritocratic beliefs in educational [@generettStoriesWeTell2020; @owensENGINESSOCIALMOBILITY2020] and organizational contexts [@perezAdvancingCareersMerit2020; @aielloNewEvidenceFirmuniversity2019].
-
-Most of the studies so far have used indicators from existing standard surveys (as the International Social Survey Programme - ISSP) to measure meritocracy.
-However, as we will show later, the methods of measuring meritocracy vary extensively among studies.
-In many cases, similar phenomena are associated with different indicators, and conversely, often different phenomena are measured with similar indicators.
-Such trends not only limits the comparability of studies, but also the ability to understanding the effects of meritocratic beliefs across different fields.
-
-Based on a critical analysis of different approaches to the measurement of meritocracy to date, this article proposes an instrument to both measure and relate two key aspects in the study of meritocracy: perceptions and preferences.
-Furthermore, as a second axis of analysis, it considers the elaboration of indicators regarding meritocratic and non-meritocratic aspects, demonstrating that they are not two poles of the same continuum as some previous studies would seem to suggest.
-The measurement system outlined is oriented to generate an instrument as simple and brief as possible, being suitable to be used in public opinion surveys and allowing to integrate meritocratic beliefs in the study of different social phenomena.
-
-## The black-box of meritocratic beliefs {-#blackbox}
-
-In the following we discuss four critical aspects in studies conceptualizing and measuring meritocracy and that are the basis of our scale proposal.
-
-a. _Conceptual components_: Is merit only effort?
-
-One recent definition of meritocracy by @mijs_paradox_2019 is the following: "when I discuss meritocracy beliefs, I am referring to citizens’ belief in the importance of hard work relative to structural factors." [@mijs_paradox_2019, pg.9].
-In the subsequent operationalization, this is associated with the following question and indicator: "how important you think it is for getting ahead in life: (a) hard work", scored from 1 to 5 on a likert scale.
-The assumptions behind such definition are worth discussing in light of the conceptual meaning of meritocracy and its possibilities of operationalization.
-
-The item used by @mijs_paradox_2019 is part of an items' battery present in several international surveys, usually called "reasons to get ahead".
-This battery displays a series of indicators related to what people consider important to get ahead in life: hard work, education, ambition, a wealthy family, the right connections, religion, race, and gender.
-Therefore, when considering only one of the items, it means that other aspects that could be associated with talent, such as education, would not be deemed as meritocratic.
-As he points out: "Hard work is arguably the most meritocratic part of Michael Young’s equation: ‘Merit = Intelligence + Effort’, for the simple fact that intelligence itself is influenced by a non-meritocratic factor: who your parents happen to be" (p.5).
-
-In this measure of meritocratic beliefs, we can observe a couple of strong conceptual assumptions: effort would not depend on parental influence, and talent (as innate ability) is not meritocratic (contrary to Michael Youngs' original conceptualization).
-This conceptual and measurement-based assumption is found in other studies, that also assume effort to be the main and only aspect of meritocracy [@girerdNeoliberalismIdeologicalBarrier2020a; @bubakPerceptionsMeritocracyNote2019].
-Is effort the only dimension behind the concept of meritocracy, or includes other components, such as talent (as in its original conceptualization).
-The question of whether talent is or is not considered meritocratic is certainly an interesting topic to discuss from a philosophical point of view, but for those working on empirical studies it should be something that face empirical scrutiny.
-Besides, the consideration of talent as part of meritocracy certainly opens some interesting avenues of research, for instance in studies showing that for the elites meritocracy is more related to talent, whereas effort is something more characteristic of meritocracy of the middle and low classes ([citar Atria ABS]).
-Actually, in a posterior paper by @mijsHowInformationEconomic2020, somewhat contradictory with its previous definition, he considered talent as an indicator of meritocracy.
-
-_b. Beliefs?_
-
-Several approaches to the empirical study of meritocracy based on public opinion surveys, refer to the concept of _beliefs_, however different meanings and operationalizations usually underlie this concept.
-To illustrate this point, lets consider the most commonly used items for operationalizing meritocratic beliefs in survey research (as ISSP), which is the "reasons to get ahead" battery.
-It consists in a series of items asking "how important you think it is for getting ahead in life" and then lists several factors, such as effort, education, parental wealth, and contacts.
-Another version of this same battery used in several surveys - sometimes along the previous one - asks about "how important you think it _should_ be ...", then listing the same concepts.
-Therefore, the question raised here is: Which one of both is a "belief"?
-What _is_ or what it _should_ be?
-
-The term belief has an ambiguous character in the literature, conceived as "idea-elements" by @converseNatureBeliefSystems1964 or "considerations" by @zallerNatureOriginsMass1992.
-As @kluegelBeliefsInequalityAmericans2017 pointed out about the scope of beliefs: "This usage encompasses such more specific social-psychological concepts as values, perceptions, and attitudes" (p.30).
-Therefore, beliefs cover almost anything related to subjective factors.
-To this regard, a relevant distinction in the field of inequality beliefs was made by @janmaatSubjectiveInequalityReview2013: "Perceptions refer to subjective estimates of existing inequality (i.e.
-thoughts about what is).
-Beliefs are here defined as normative ideas about just inequality (i.e.
-thoughts about what should be)"(p.359).
-Several papers dealing with meritocracy use the term beliefs (i.e.
-what should be), while actually referring to perceptions (i.e.
-what is).
-This occurs not only in @mijs_paradox_2019, but also in @reynolds_perceptions_2014, in which the term beliefs is used to talk about perceptions, whereas other authors use general terms such as attitudes [@kunovich_systems_2007].
-The first attempt to shed light on this issue was made by @duru-bellat_whos_2012, who used the question "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" as a proxy for "desired" meritocracy (beliefs).
-They then determined "perceived" meritocracy, using the questions: “Would you say that in your country, people are rewarded for their efforts?” and “...
-people are rewarded for their skills?”.
-
-Is the belief in meritocracy a perception or a preference with normative meaning?
-In order to expand the analytical conceptual framework, we believe that both dimensions should be included in analyses, as proposed by @duru-bellat_whos_2012.
-This opens up the possibility of analyzing whether perceptions and preferences are actually related (i.e.
-have a high correlation), or whether they are independent aspects of the same phenomenon.
-As @son_hing_merit_2011-1 have pointed out, "People can believe that outcomes ought to be distributed on the basis of merit and yet vary in their perceptions of whether this is how society currently operates" (p.
-435).
-In other words, normative beliefs should be considered while taking perceptions into account: a strong normative belief in meritocracy may mean something totally different to someone perceiving high meritocracy, when compared to someone perceiving low meritocracy.
-To avoid the confusion generated by the term "belief", we propose the terms meritocratic preferences ("what should be"), and meritocratic perceptions ("what is"), as they better reflect the two facets of meritocracy under scrutiny [@castillo_meritocracia_2019].
-
-_c. Non-meritocratic aspects_
-
-A number of studies have dealt with aspects opposed to meritocracy, as for instance the use of personal advantages (as contacts or having a wealthy family) to get ahead in life.
-For instance, @kunovich_systems_2007 used an items’ battery listing a number of factors in relation to “How important each should be in deciding pay...” (as @duru-bellat_whos_2012 for desired meritocracy).
-They consider factors such as education and responsibility as meritocratic, giving them a value of 1 if considered “essential” in the scale response, whereas factors such as having a family or children are valued 1 when rated as "not important at all" (i.e.
-reverse coded).
-The assumption behind this approach is that rejecting a supposed non-meritocratic aspect (as having family and/or children) implies a stronger belief in meritocracy.
-A similar approach of reverse-coding non-meritocratic items was taken by @newman_false_2015, using the same principles applied in the "Preference for the Merit Principle Scale" [@davey_preference_1999].
-
-The assumption that meritocratic and non-meritocratic elements are poles of the same continuum was tested by @reynolds_perceptions_2014 using the "get ahead" perceptions' battery items mentioned above.
-They considered education, ambition, and hard work as meritocratic, and other factors such as family wealth and connections as non-meritocratic.
-Despite making and proving this distinction, however, the authors end up subtracting one dimension from the other, thus coming back to the assumption that they are two poles of the same continuum as @kunovich_systems_2007 did.
-Similarly, @roex_attitudes_2019 used ISSP indicators for perceived meritocracy and non-meritocracy to build a single score by reverse coding the non-meritocratic items.
-Therefore, the treatment of non-meritocratic items has been rather inconsistent across studies and the assumption that they are the simple opposite of meritocracy requires further empirical assessment.
-
-_d. Accounting for measurement error_
-
-Finally, most studies in meritocracy so far have not accounted for the issue of measurement error [@brownBookReviewTimothy2010; @bollenStructuralEquationsLatent1989] (Ansolabehere, S., Rodden, J., & Snyder Jr, J.
-M., 2008), as they use single indicators and/or simple average indexes for measuring meritocracy.
-Such strategy assumes that the latent construct is measured perfectly (i.e.
-no error or residual variance) by the indicators chosen, going as far as to propose that "...
-In choosing this strategy of index construction, we argue that support for meritocracy is not a latent variable" [@kunovich_systems_2007 p.653-654].
-Some advances were made by @reynolds_perceptions_2014 through conducting a principal component analysis of meritocratic and non-meritocratic dimensions, but somewhat contradictorily, they chose a sum index despite proving a multidimensional latent structure.
-
-## An instrument proposal {-#instrumentprop}
-
-Based on the previous assumptions and limitations of the measurements of meritocracy, we propose and test an instrument with the following characteristics:
-
-- _Multidimensionality_, incorporating previous distinctions between preferences and perceptions, as well as between meritocratic and non-meritocratic aspects.
-
-- Multiple indicators for each dimension, in order to _account for measurement error_ in a confirmatory factor analysis context.
-
-- Based on _previous indicators_ as much as possible, for the sake of keeping comparability between studies.
-
-- _Brief_, as to be used in regular public opinion surveys.
-In this respect, it differs from the proposal of  "Preference for the Merit Principle Scale" [@davey_preference_1999], as they use 15 items for just one dimension (aside from the problem of reverse-coding non-meritocratic items).
-
-The proposed measurement framework is depicted in Figure \@ref(fig:merit-model):
-
-```{r merit-model, echo=FALSE, fig.cap = "Model of perception and preferences for meritocracy and non-meritocracy", fig.align='center'}
-knitr::include_graphics("../input/images/generalf.png")
-```
-The columns "Perceptions" and "Preferences" represent the distinction between these two concepts, usually confused under the label "beliefs" [@castillo_meritocracia_2019].
-Perceptions refer to the extent to which people observe that meritocracy works or applies in their society, which relates to items such as "I think hard work is important to get ahead in society".
-Preferences refer to normative expectations that are usually linked to a "should" expression (e.g.
-whether hard work should be related to payment).
-The rows in Figure 1 consider the distinction between meritocratic and non-meritocratic dimensions [@reynolds_perceptions_2014].
-These aspects have usually been treated as different ends of the same continuum in previous research, an assumption that requires empirical scrutiny.
-Non-meritocratic elements usually refer to the use of personal contacts or family advantages to get ahead in life.
-
-Regarding the selection of indicators, most are taken or adapted from previous studies for the sake of comparability.
-For meritocratic indicators, we use effort and talent as the main components of the traditional concept of merit as defined by @young_rise_1962, whereas for non-meritocratic dimensions, we use wealthy parents and good contacts.
-Descriptions of the specific items are outlined in the methodology section.
-
-The research hypotheses behind this conceptualization and measurement model are as follows:
-
-* $H_1$. The perception of meritocracy is a latent variable based on indicators of the importance attributed to talent and effort to get ahead in life.
-
-* $H_2$. The perception of non-meritocracy is a latent variable that derives from  indicators related to agreement with the statement that people with contacts and rich parents manage to get ahead.
-
-* $H_3$. Meritocratic preferences are a latent variable based on the normative value of effort and talent.
-
-* $H_4$. Non-meritocratic preferences are a latent variable based on the normative value of the use of personal contacts and having wealthy parents.
-
-# Methodology
-
-The analysis is organized in three studies.
-Study 1 serves as a background to our instrument proposal by attempting to operationalize the meritocracy dimensions with available data.
-For this we use data from the International Social Survey Programme (ISSP).
-Although we are aware that the available indicators in this dataset do not allow a precise operationalization of the conceptual model, we know that this dataset is the most widely used in comparative meritocracy studies (as well as overall subjective inequality) and therefore it is certainly useful to look at its potentials and limitations to this regard.
-Study 2 is the core of this paper and consists of the proposal of a new scale for measuring meritocratic perceptions and preferences.
-Finally, in Study 3 we perform additional validity analysis of the meritocracy scale.
-
-## Study 1
-
-### Data
-
-The data used in the analysis corresponds to the last available wave of the social inequality module from the International Social Survey Programm (ISSP), which is the most specialized international comparative survey in perceptions, attitudes and beliefs about inequality related issues.
-This wave corresponds to the year 2009 and covers attitudes towards a series of topics dealing with social inequality across 26 countries and XXXX individuals.
-Although there is data available from this module for previous waves (1987, 1992 and 1999), unfortunately there are several variables that are important for this research that were not included, particularly in 1987 and 1999, reason why we only use the last wave.
-
-### Variables
-
-There are a series of indicators in the ISSP survey that in the following we attempt to classify in our meritocracy measurement scheme.
-However, it is important to mention  that the items were not originally thought as a specific measure of meritocracy.
-Still, many of the items have been used in previous research on meritocracy (and certainly will keep being used in the future) and before proposing a new measure we consider to give a first test to our proposed conceptualization using available data that not only covers several countries but also is validated by the research community.
-
-The variables are presented below organized in perceptions and preferences:
-
-- _Perception of meritocracy/non-meritocracy_: for operationalizing perceptions the closest set of ISSP's indicators comes from the question asking about perceptions for opportunities to get ahead, which are usually considered as "meritocratic beliefs" in previous research.
-The general heading of the battery is: _"To begin we have some questions about opportunities for getting ahead.
-Please tick one box for each of these to show how important you think it is for getting ahead in life."_ This is followed by a list of statements to be answered in a scale with the following levels numbered from 1 to 5: essential, very important, fairly important, not very important, not important at all.
-Below the items are listed classified according to the meritocracy/non-meritocracy character:
-
-  - Meritocratic perception:
-    - How important is having ambition?
-    - How important is hard work?
-
-  - Non-meritocratic perception
-    - How important is coming from a wealthy family?
-    - How important is having well-educated parents
-    - How important is knowing the right people
-    - How important is having political connections
-    - How important is a person's race?
-    - How important is being born a man or a woman?
-
-<br>
-  The classification is based on criteria of internal motivation (meritocratic) and structural constraints (non-meritocratic).
-There were two items from the battery that were excluded from the analysis as they would not fit into the classification.
-The first one was "having good education yourself", since it was not clear whether this is could be due to individual motivation or system opportunities, and the second was "giving bribes", as introduced elements of criminality that were beyond a non-meritocratic perception.
-
-- Preferences for meritocracy-non meritocracy: for operationalizing normative preferences we used a list of items related to reasons for pay.
-The ISSP question was: _In deciding how much people ought to earn, how important should each of these things be, in your opinion_, rated in the same _essential-non important at all_ scale as the questions for meritocratic preferences.
-The reasons are:
-
-  - Meritocratic preferences
-    - How well he or she does the job?
-    - How hard he or she works at the job?
-
-  - Non-meritocratic preferences
-    - What is needed to support a family?
-    - Whether the person has children to support?
-
-(here summary descriptive table - @julio)
-
-### Methods
-
-The estimation was performed using Confirmatory Factor Analysis (CFA).
-CFA was conducted using the `lavaan` R package (version 0.6-3; Rosseel, 2020), with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement (Kline, 2016; Rosseel, 2020).
-As recommended by Brown (2008), we assessed model fit by jointly considering the comparative fit index and Tucker-Lewis Index (CFI and TLI; acceptable fit > 0.95), Root of the average squared residual approximation (RMSEA; acceptable fit < 0.08),  Chi-square: (p-value; acceptable fit > 0.05, and Chi-square ratio > 3).
-
-### Results
-
-[diagrama SEM]
-
-[tabla CFA]
-
-## Study 2
-
-## Data collection {-#data}
-
-The data was obtained through an online questionnaire, which was part of a larger study on meritocracy and preferences developed in Chile in 2019 funded by the national scientific agency ANID.
-The questionnaire was programmed in Qualtrics and the fieldwork was conducted by an external online survey agency ([netquest.cl](www.netquest.cl)) between December 2019 and January 2020.
-The sample was selected from a non-probabilistic quota design in three large cities in Chile (Santiago, Concepción & Antofagasta).
-The quotas for gender, age, and education level were generated based on a survey by the Public Studies Center [@cepEncuestaCEPMayo2019], which is a well regarded counterpart agency of the ISSP (International Social Survey Programme) in Chile.
-A total sample of 2,141 people was collected, excluding those who did not answer the questions on the scale, and those who did not accept informed consent.
-There were no significant differences between our sample and the wider population for most socio-demographic characteristics, with the exception of educational level (see Table \@ref(tab:rep-samp) in appendix).
-As is often the case with online surveys, there were some limitations in achieving the quotas for lower educational levels **[citas - LUCHO]**.
-
-## Instrument design
-
-The proposed scale of perceptions and preferences relating to meritocracy consisted of eight indicators that were grouped into the four dimensions listed earlier: Perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic).
-In order to achieve at least some comparability with previous studies, the questions were adapted from the items battery "reasons to get ahead" (ISSP/GSS), which has been widely used for operationalizing meritocracy in previous studies [@mijs_paradox_2019; @duru-bellat_whos_2012; @reynolds_perceptions_2014].
-The aforementioned eight items, ordered according to dimensions, are presented in Table \@ref(tab:table-indicadores).
-These eight likert-type items have five response alternatives, ranging from "Completely disagree"(1) to "Completely agree" (5).
-
-```{r include=FALSE}
-cap <- "Items according to dimension."
-# if (knitr::is_html_output()) {tb <- table_nums("tb",cap)} else {tb <- cap}
-```
-
-```{r table-indicadores, echo=FALSE}
-tab01 <- read.csv(file = "../input/data/original/materials/table01.csv",header = 1,sep = ",",encoding = "UTF-8")
-cnames <- c("Dimensions", "Factor","Statement (english)" , "Statement (spanish)")
-
-kable(tab01,table_format,booktabs = T, linesep = "",col.names = cnames, caption = cap) %>%
-  kable_styling(
-    full_width = T,
-    latex_options = c("hold_position"),
-    position = "center",
-    font_size = 10,
-    bootstrap_options=c("striped", "bordered")) %>%
-  column_spec(column = 1, width = "1.5cm", ) %>%
-  column_spec(column = 2,width = "2 cm") %>%
-  collapse_rows(columns = 1:2,valign = "middle")
-```
-
-\pagebreak
-
-## Administration sets
-
-With the objective of evaluating the effect of indicator ordering in the responses, three different versions of items' order were designed and randomly assigned, as depicted in Figure \@ref(fig:appmod).
-The scale was presented to the first group (_n = 712_) in the order that appears in Table \@ref(tab:table-indicadores) according to perceptions and preferences.
-For the second group (_n = 717_), the order was reorganized according to perceptions and preferences over the same topic, e.g.
-for the topic of hard work, the item about perception was followed by the item about preference, and the same for the rest of the topics.
-Finally, for the third group (_n = 712_), the items were completely randomized.
-
-```{r appmod, echo=FALSE, fig.cap = "Survey flow", fig.align='center'}
-knitr::include_graphics('../input/images/app_mod.png')
-```
-
-# Methods
-
-To test the scale's underlying constructs, we employed confirmatory factor analysis models (CFA).
-The models estimated one factor for each of the proposed dimensions, as represented in the following figure:
-
-```{r meas01, echo=FALSE, fig.cap = "Theoretical model", fig.align='center'}
-knitr::include_graphics('../output/images/meas01.png')
-```
-As in Study 1, CFA was conducted using the `lavaan` R package (version 0.6-3; Rosseel, 2020), with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement (Kline, 2016; Rosseel, 2020).
-The fit indexes and cut-off criteria were the same as in Study 1.
-
-A pre-registration was made in the OSF platform, available at the following link: [https://osf.io/z45y2](https://osf.io/z45y2).
-Included in this pre-registration are the hypotheses regarding the four-dimensional conceptual model underlying the scale, the variable measurement levels, the statistical tests to be performed with their respective evaluation parameters, and other important aspects of the research design.
-
-# Results
-
-## Descriptive analyses
-
-As can be seen in the Table \@ref(tab:desc01), the indicators have values ranging from 1 (totally disagree) to 5 (totally agree).
-Averages are observed from 2.41, corresponding to preference-contacts; to 3.89, corresponding to preference-effort.
-Both indicators are consistent with a general adherence to meritocracy as reported in previous studies, privileging individual aspects such as effort [@bucca_merit_2016; @frei_fruto_2020].
-
-```{r desc01, echo=FALSE}
-cap <- "Descriptive statistics of the scale."
-load(file = "../input/data/proc/dat04.RData")
-dat04b<- dat04 %>% rename("A. Perception Effort" =perc_effort,
- "B. Perception Talent" = perc_talent,
- "C. Perception rich parents" = perc_wpart,
- "D. Perception contacts" = perc_netw,
- "E. Preferences Effort" = pref_effort,
- "F. Preferences Talent" = pref_talent,
- "G. Preferences rich parents" = pref_wpart ,
- "H. Preferences contacts" = pref_netw )
-sum04b <- psych::describe(dat04b[,-c(1)], fast = T) %>% data.frame() %>% select(mean,sd,min,max) %>% round(digits = 2)
-table04b<- sum04b %>% rename("Mean"=mean, "SD"=sd, "Min"=min,"Max"=max)
-sumdat04<- kable(table04b,format =table_format,
-                 booktabs= T,
-                 #exhaust = FALSE,
-                 align = "l",
-                 linesep = "",
-                 caption = cap) %>%
-  kable_styling(full_width = table_format2,
-                font_size = 10,
-                position = "center",
-                latex_options = c("HOLD_position"));sumdat04
-```
-
-The graphs presented in Figure \@ref(fig:plotlikert) display disaggregated and comparable information of the different response categories for each item.
-As it can be observed, in general there is more agreement in the perception of non-meritocratic items than in meritocratic items, while in the case of preferences, the opposite occurs.
-As far as preferences are concerned, the preponderant role of effort over talent as a criterion of meritocratic preference is noteworthy.
-
-```{r plotlikert, echo=FALSE, fig.cap = "Descriptive plot", fig.align='center'}
-knitr::include_graphics('../output/images/plotlikert.png')
-```
-
-In Figure \@ref(fig:corpoly), moderate to high intensity relationships are observed between indicators that correspond to the same factor (e.g., perception of meritocracy by effort and by talent, r=0.56).
-The relationships between meritocratic perceptions and preferences are also salient, which is not the case for non-meritocratic indicators.
-
-```{r corpoly, echo=FALSE, fig.cap = "Polychoric correlation plot", fig.align='center'}
-knitr::include_graphics('../output/images/corpoly.png')
-```
-In sum, descriptive analyses point to a relative adherence to meritocracy, which is expressed as a greater preference for meritocratic criteria, and a lesser preference for non-meritocratic criteria, even when the preference for non-meritocratic elements reaches a proportion of more than 20%.
-Likewise, a relatively low perception of meritocracy is observable.
-In addition, we observed a coherent relationship between the indicators of the same dimension as proposed in the pre-registration of the study; that is, the pairs of items associated with a specific factor show correlations with a large effect size.
-In particular, the associations between effort and talent are relevant, challenging previous assumptions that talent would not be a meritocratic criterion [@mijs_paradox_2019], otherwise the correlation would be zero or negative.
-In addition, we observe that there is no negative correlation between meritocratic and non-meritocratic aspects, undermining the assumptions of previous studies that suggested that these dimensions would be the opposite poles of one same continuum [@reynolds_perceptions_2014].
-
-## Confirmatory Factor Analysis
-
-This section estimates the fit of the conceptual model behind the Perceptions and Preferences for Meritocracy Scale.
-For this, we first estimate a confirmatory factor analysis model for the whole sample, and secondly we test the order effects applying the same model to each of the three order permutations.
-
-Figure \@ref(fig:meas02) shows the results of the estimation for the four-factor model using the complete sample.
-This model shows adequate fit indicators (CFI = 0.977, RMSEA = 0.079, $\chi2$(df = 14) = 28,03, p = 0.014) and it presents factorial loads with weights that vary between 0.63 and 0.88.
-
-```{r meas02, echo=FALSE, fig.cap = "Confirmatory factor analysis of the Perceptions and Preferences for Meritocracy Scale", fig.align='center'}
-knitr::include_graphics('../output/images/meas03.png')
-```
-Attending to the correlations between the latent variables as depicted in Figure \@ref(fig:latcor), it is observed that the variables with the strongest correlations are meritocratic perceptions and preferences, whit an $r=0.59$, while non-meritocratic perceptions and preferences show a correlation close to 0.
-Besides, and as commented before in the analysis of the single items correlations, the latent variables of meritocratic and non-meritocratic versions of both perceptions and preferences show very low correlations.
-This last finding gives further evidence regarding the lack of unidimensionality of meritocratic and non-meritocratic aspects: they do not appear as the opposite poles of the same continuum and based on this evidence the use of reverse-coding for non-meritocratic items as indicators of meritocracy should be avoided.
-
-```{r latcor, echo=FALSE, fig.cap = "Latent variables correlation", fig.align='center'}
-knitr::include_graphics('../output/images/latcor.png')
-```
-In a second step we test the effect of the items' order in the confirmatory model.
-Table \@ref(tab:tab03) shows the fit indicators of the models estimated for each of the three versions of the items' order described in the methodology, being the first one according to perceptions and preferences, the second one according to meritocratic and non-meritocratic aspects, and the third one completely randomized.
-Regardless of the group and order of items, all models obtained an adquate fit, with CFI's above 0.95 and RMSEA's below 0.08.
-However, none of the models achieved a non-significant chi-square, something expected in large samples as the one used here).
-Model 1 shows the fit indicators for the estimation of the CFA model to the first version order (perceptions-preferences), which was the one obtaining best fit (CFI=0.993, TLI=0.995, RMSEA=0.034, $\chi2$(df=14)=42.276), whereas Model 2 for the fixed order according to merit/non-merit items shows the comparatively worst indicators.
-The CFA fit indices for the completely randomized items' order (Model 3) is somehow an intermediate fit indexes when compared with the previous two.
-Still, it keeps all the indicators within the acceptable cut-off criteria and besides it controls for possible order effects in the administration of the instrument.
-
-```{r tab03, echo=FALSE,eval=TRUE}
-load(file = "../output/tables/fit4_o.RData")
-load(file = "../output/tables/fit3_o.RData")
-load(file = "../output/tables/fit2_o.RData")
-load(file = "../output/tables/fit1_o.RData")
-sum_fit<- bind_rows(fitmeasures(fit1_o)[c("chisq","df","cfi","tli","rmsea")],
-                    fitmeasures(fit2_o)[c("chisq","df","cfi","tli","rmsea")],
-                    fitmeasures(fit3_o)[c("chisq","df","cfi","tli","rmsea")])
-sum_fit$mod <- c("Model 1","Model 2","Model 3")
-sum_fit$est <- c("DWLS","DWLS","DWLS")
-sum_fit$nobs <- c(nobs(fit1_o),nobs(fit2_o),nobs(fit3_o))
-sum_fit <- select(sum_fit,mod,nobs,est,everything())
-colnames <- c("Model","$N$","Estimator","$\\chi^2$","df","CFI","TLI","RMSEA")
-sumtable02<- kable(sum_fit,digits = 3,format = table_format,row.names = F,booktabs=T, align = c("lcccccc"),
-                   caption = "Summary fit indices according group",
-                   col.names = colnames,
-                   escape = FALSE) %>%
-  kable_styling(full_width = F,
-                font_size = 10)  %>%
-      footnote(number = c("Model 1: fixed order by perception/preference",
-                          "Model 2: fixed order by merit-non merit",
-                          "Model 3: Randomized order"));sumtable02
-```
-
-## Additional validity analyses
+## Study 3: Additional validity analyses
 
 We performed two further analyses in order to add evidence regarding the quality of the scale.
-Firstly we analyzed the convergent validity by exploring the association of the merit scale with related concepts and itscorrespondent measurements.
+Firstly we analyzed the convergent validity by exploring the association of the merit scale with related concepts and its correspondent measurements.
 Secondly, we test the internal consistency of the measurement model behind the scale comparing with a different sample through measurement invariance procedures .
@@
 Therefore, using this last standard the level for strict invariance holds for the meritocracy scale.
-
-# Conclusions
-
-Studies that attempt to characterize and compare societies by their support for meritocratic beliefs have used different approaches.
-As most studies use secondary survey data, they tend to assume that the available indicators represent an underlying meritocratic construct.
-A review of these studies reveals several non-tested assumptions, as well as the use of similar indicators to represent different constructs and dimensions of meritocracy.
-As the existence of heterogeneous approaches certainly has consequences for the advancement of the study of meritocracy, this paper presented a comprehensive conceptual framework for the empirical study of meritocracy, building upon previous research.
-This framework was then tested against new survey data.
-
-We identified four critical aspects regarding the measurement of meritocracy in previous studies: unidimensionality, the ambiguous use of the term “beliefs”, the use of non-meritocratic indicators as opposed to meritocratic indicators, and the consideration of measurement error.
-The proposed 8-item scale, “Perceptions and Preferences for Meritocracy”, was designed and tested in order to deal with these four issues.
-Our results indicate that perceptions and preferences seem to be two related but different dimensions, often confounded in previous research under the label of "beliefs".
-Meritocratic and non-meritocratic dimensions do not appear to constitute poles of the same continuum, as some previous studies have assumed.
-Regarding the possible effects of items order in the estimation of latent variable, we tested three different order versions and found evidence that suggest the use of the randomized version of the scale.
-Furthermore, we found evidence of convergent validity as well as measurement invariance that approach well the requirement of comparability between samples.
-
-The four-dimensional framework and structure of the Perceptions and Preferences for Meritocracy scale opens several avenues for future research.
-For instance, distinguishing perceptions from preferences will allow us to evaluate the extent to which different societies are accustomed to, or satisfied with, the perceived level of meritocracy, in terms of differences between what is perceived and what is preferred.
-Additionally, given that non-meritocratic factors are not necessarily related to meritocratic ones, our framework makes it possible to assess the perceived legitimacy of practices such as the use of personal contacts and their interference (or not) with meritocratic ideals in different societies.
-The consideration of structural factors, both individual and societal, will allow us to advance future hypotheses relating to meritocratic legitimacy to a more comprehensive level than the current one, which is based solely on meritocratic perceptions.
-Furthermore, the impacts of different configurations of the four-dimensional framework on practices and behaviors such as corruption, civic involvement, and political alignment, is an area that requires additional research.
-Such future agendas could be specially relevant in times of economic crisis and growing inequalities, that could entail changes in the legitimation of the current distributive structure based on meritocratic ideals.
-
-\pagebreak
-
-# References
-
-<div id="refs"> </div>  <!-- Although <div> is an HTML tag, this method also works for other output formats such as PDF. -->
-
-# (APPENDIX) Appendix {-}
-
-`r if (knitr::is_latex_output()){ '\\appendix'}`
-
-`r if (knitr::is_latex_output()){ '\\section{Appendix}'}`
-
-```{r tab-loads, echo=FALSE}
-load(file = "../output/tables/tb.fm.RData")
-cap <- "Factor loads and model fit."
-tb.col  <- c("Variables",1:4,1:4,1:4)
-tb.caption <- "Factor loadings and fit measures"
-tb.foot <- paste0("Standardized factor loadings using DWLS estimator ; CFI = Comparative fit index (scaled); RMSEA = Root mean square error of approximation (scaled)")
-
-tb.fm <- tb.fm %>%
-    mutate(vars = recode(vars,
-                         "A. Who the more they try they manage to get bigger rewards that those who strive less."= "Perception Effort",
-                         "B. Who possess more talent they manage to obtain greater rewards than those who possess less talent."= "Perception Talent",
-                         "C. Who they have rich parents manage to get out ahead."="Perception rich parents",
-                         "D. Who they have good contacts they manage to get out ahead."="Perception contacts",
-                         "E. Who the more they try they should get greater rewards than those who they try less." = "Preferences Effort",
-                         "F. Who possess more talent they should get greater rewards than those who possess less talent."= "Preferences Talent",
-                         "G. It's fine that those who have rich parents get ahead" ="Preferences rich parents",
-                         "H. Is well that those who have good contacts get ahead." = "Preferences contacts"))
-
-tb.fm01<- kable(tb.fm, format = table_format, booktabs = T,linesep = "", escape = FALSE,align = "lcccccccccccc",col.names = tb.col, caption = cap) %>%
-          kable_styling(full_width = table_format2,latex_options = c("hold_position","scale_down"), position = "center",font_size = 9) %>%
-          add_header_above(header = c(" "=1,"Version 1"= 4,"Version 2"= 4,"Version 3"= 4)) %>%
-          add_header_above(header = c(" "=1,"Factor loadings"= 12)) %>%
-          row_spec(row = 8,hline_after = TRUE) %>%
-          add_indent(c(9:12))  %>%
-          footnote(general = tb.foot,threeparttable = T);tb.fm01
-```
-
-```{r fcloads-inv, echo=FALSE}
-load(file = "../output/tables/tb.fact-f02.RData")
-tb.foot <- paste0("Standardized factor loadings using DWLS estimator ; CFI = Comparative fit index (scaled); RMSEA = Root mean square error of approximation (scaled)")
-tb.fm02<- kable(tb.fm2,escape = FALSE,align = "lcccc",col.names = tb.col, caption = tb.caption, booktabs = T,linesep = "",format = table_format) %>%
-          kable_styling(full_width = table_format2,latex_options = c("hold_position"), position = "center",font_size = 9) %>%
-          add_header_above(header = c(" "=1,"Factor loadings"= 4)) %>%
-          row_spec(row = 8,hline_after = TRUE) %>%
-          add_indent(c(9:12)) %>%
-          footnote(general =tb.foot ,footnote_as_chunk = T);tb.fm02
-```
-
-\pagebreak
-
-|                     | Sample  |   CEP  |
-|---------------------|:-------:|:------:|
-| __Gender__          |         |        |
-| Men                 | 49,82%  | 50,52% |
-| Women               | 50.18%  | 49,47% |
-| __Age__             |         |        |
-| 18 - 24             | 18,55%  | 18,17% |
-| 25 - 34             | 18,86%  | 17,48% |
-| 35 - 44             | 19.09%  | 19,98% |
-| 45 - 54             | 17,96%  | 19,23% |
-| 55 - or more        | 25,54%  | 25.11% |
-| __Education__       |         |        |
-| Primary  or less    | 2,93%   | 15,88% |
-| High school          | 43,23%  | 37,04% |
-| Non university      | 32,63%  | 28,93% |
-| university or more  | 21,21%  | 18,13% |
-Table: `r as.character(paste("(\\#tab:rep-samp)", "Representativeness of the sample"))`.

### 2021-05-25 processing/preparation_study2.Rmd
- major update

---
title: "Data preparation Study 2 - Merit Scale"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
css: "../input/bib/style.css"
output:
  bookdown::html_document2:
    number_sections: true
    toc: true
    toc_float:
        collapsed: false
    toc_depth: 2
    code_folding: "hide"
editor_options:
  chunk_output_type: console
---

```{r include=FALSE}
# Knitr options

knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = FALSE,results = "hold")
knitr::opts_knit$set(root.dir = "../")
options(scipen=999)
rm(list=ls())
options(knitr.kable.NA = '')
```

This is the preparation code of the data for Study 2,: the Perceptions and Preferences of Meritocracy Scale. The prepared data is `data_s2.RData`

# Libraries
```{r}
pacman::p_load(sjmisc, sjPlot, dplyr, tidyr, lavaan, stargazer, corrplot, psych, knitr, kableExtra, summarytools, sjlabelled, Hmisc)
```

# Load original data

```{r}
data <- sjlabelled::read_spss("input/data/original/Estudio_3_ola1.sav",verbose = F)
dim(data)
```

# Select variables subset

First we select those who that accepted to participate in the informed consent based on the response to the variable `Intro`:

```{r}
table(data$Intro)
```

There are 2236 who accepted to participate (1) and 221 who does not (221). Now we filter them:

```{r}
data_s2 <- data %>% filter(Intro==1)
```

Now we select the variables of the scale which are the ones with "meritv0" in the label are the ones used in the scale (8 items):

```{r}
find_var(data = data_s2,"meritv03")
```

There are 4 topics (effort, talents, parents' wealth & networks) both for perceptions and preferences, summing up 8 items in total.

Suffixes v01, v02 & v03 make reference to the different order sets in which items were administrated (see paper for more details). Besides, we select variables that indicate the order set, beggining with FL_21_DO

```{r}
data_s2 <- data_s2 %>%
  select(starts_with("meritv01"),
         starts_with("meritv02"),
         starts_with("meritv03"),
         starts_with("FL_21_DO"))

dim(data_s2)
sjmisc::descr(data_s2, show = "short") # labels too long: shorten

```

Change long labels from FL_21 ... variables

```{r}
data_s2 <- data_s2 %>% rename(
  "id_v1"=FL_21_DO_merit_perc_pref_julio19v01,
  "id_v2"=FL_21_DO_merit_perc_pref_julio19v02,
  "id_v3"=FL_21_DO_merit_perc_pref_julio19v03
  )
```

# Process variables for analysis

In the following we make a unique identification for each item across order sets, and create a variable that allows identifying to which administration set they belong.

First just check whether the set's id variables are working properly identifying the corresponding items:

```{r}
data_s2 %>% filter(id_v1==1) %>%sjmisc::descr(., show = "short")
data_s2 %>% filter(id_v2==1) %>%sjmisc::descr(., show = "short")
data_s2 %>% filter(id_v3==1) %>%sjmisc::descr(., show = "short")
```

Ok. Now recode de values on the version in order to be able to identify the groups afterwards

```{r}
data_s2$id_v2 <- car::recode(data_s2$id_v2, "1=2")
data_s2$id_v3 <- car::recode(data_s2$id_v3, "1=3")

```

Now we create the variables across response set versions. For this first we create a vector with the corresponding variable names, e.g. the three containing "perc_effort" (independent of the ser version). Then we use these vectors for creating the new variables with `coalesce`.

```{r}
#create vector with names
match_perc_effort = rlang::syms(grep("perc_effort", names(data_s2), value = TRUE))
match_perc_talent = rlang::syms(grep("perc_talent", names(data_s2), value = TRUE))
match_perc_wpart = rlang::syms(grep("perc_wpart", names(data_s2), value = TRUE))
match_perc_netw = rlang::syms(grep("perc_netw", names(data_s2), value = TRUE))
match_pref_effort = rlang::syms(grep("pref_effort", names(data_s2), value = TRUE))
match_pref_talent = rlang::syms(grep("pref_talent", names(data_s2), value = TRUE))
match_pref_wpart = rlang::syms(grep("pref_wpart", names(data_s2), value = TRUE))
match_pref_netw = rlang::syms(grep("pref_netw", names(data_s2), value = TRUE))
match_set_version = rlang::syms(grep("id_v", names(data_s2), value = TRUE))

#create variables
data_s2 <- data_s2 %>%
  mutate(perc_effort = coalesce(!!! match_perc_effort),
         perc_talent = coalesce(!!! match_perc_talent),
         perc_wpart = coalesce(!!! match_perc_wpart),
         perc_netw = coalesce(!!! match_perc_netw),
         pref_effort = coalesce(!!! match_pref_effort),
         pref_talent = coalesce(!!! match_pref_talent),
         pref_wpart = coalesce(!!! match_pref_wpart),
         pref_netw = coalesce(!!! match_pref_netw),
         group = coalesce(!!! match_set_version)
        )

#check:
data_s2 %>% select(perc_effort, meritv01_perc_effort, meritv02_perc_effort, meritv03_perc_effort) %>% slice_head(n=10)

sjmisc::descr(data_s2)
frq(data_s2$group)

# 75 NAs in group, probably those who not continued the survey, check:

data_s2 %>%  filter(is.na(group))

# Ok, remove NAs in groups

data_s2 <- data_s2 %>%  filter(!is.na(group))
```

Now select the final variables

```{r}
data_s2 <- data_s2 %>% select(perc_effort,perc_talent,perc_wpart, perc_netw,
                              pref_effort,pref_talent,pref_wpart, pref_netw,
                              group)

names(data_s2)
dim(data_s2)
```

## Labels

### Variable labels

```{r}

set_label(data_s2$group) <- "Order sets"

set_label(data_s2$perc_effort)  <- "More effort get greater rewards"
set_label(data_s2$perc_talent)  <- "More talent get greater rewards"
set_label(data_s2$perc_wpart)   <- "Rich parents to get ahead"
set_label(data_s2$perc_netw)    <- "Contacts to get ahead"

set_label(data_s2$pref_effort)  <- "More effort should get greater rewards"
set_label(data_s2$pref_talent)  <- "More talent should get greater rewards"
set_label(data_s2$pref_wpart)   <- "Rich parents should get ahead"
set_label(data_s2$pref_netw)    <- "Contacts should get ahead"

sjmisc::descr(data_s2)
sjmisc::frq(data_s2$perc_effort)

```

### Value labels
```{r}
get_labels(data_s2$perc_effort) # translate

# generate vector of variables to relabel:
perc_or_pref = rlang::syms(grep("perc_|pref_", names(data_s2), value = TRUE))
perc_or_pref

# label values

data_s2 <- data_s2  %>% replace_labels(labels = c("Strongly disagree" = 1, "Disagree" = 2, "Neither agree nor disagree"=3,"Agree"=4, "Totally agree" = 5))

data_s2$group<- set_labels(data_s2$group, labels = c("Order set 1"=1,"Order set 2"=2,"Order set 3"=3 ))

#check
get_labels(data_s2$perc_effort) # ok
get_labels(data_s2$group) # ok

sjmisc::descr(data_s2)

```

# Final summary
```{r}
sjmisc::descr(data_s2,
      show = c("label","range", "mean", "sd", "NA.prc", "n"))%>%
      kable(.,"markdown")

print(dfSummary(data_s2, valid.col = FALSE, graph.magnif = 0.75),
      max.tbl.height = 300, method = "render")

```

## save data

```{r}

save(data_s2,file = "input/data/proc/data_s2.RData")
```

### 2021-05-25 papelera/render-site.R

@@ -1,0 +1,61 @@
+
+# Render site ------------------------------------------------------------
+
+###### logic
+# there are some limitations for rendering websites in Rmarkdown/Githubpages
+# as all files should be in a same directory, which goes against the logic
+# of a clean and organized folder structure. Therefore, this script:
+# 1. bring all necessary files to the "docs" folder, where the site files are hosted
+# 2. renders the site from this location (docs folder)
+# 3. delete the Rmd files that where brought in step 1, leaving only the html files
+
+
+# INSTRUCTIONS-------------------------:
+# 1. set your root dir at "merit-scale"
+# 2. Run this code.4
+
+production<- list.files(path = "production",pattern = "prod_")        # a list with the names of the files to copy
+file.copy(file.path("production",production), "docs",overwrite = TRUE)# copy data proc and analysis files
+
+rmarkdown::render_site("docs",quiet = FALSE) # Render site
+
+# before you run this line, check if your R Markdown files have a .rmd or .Rmd extension 
+# on this case we use both
+ext <- c(grep("^prod_.*\\.rmd$",  x = dir(path = "docs"),value = T), # for .rmd
+         grep("^prod_.*\\.Rmd$",  x = dir(path = "docs"),value = T), # for .Rmd
+         grep("_cache",x = dir(path = "docs"),value = T),            # for files within the "_cache" folder
+         grep("^prod_.*\\_cache$",  x = dir(path = "docs"),value = T), # 
+         grep("^prod_.*\\.R$",    x = dir(path = "docs"),value = T), # for .R
+         ".tex",".log",".RData",".Rhistory")                         # for .tex .log .RData .Rhistory
+
+for (i in 1:length(ext)) {
+  file.remove(paste0("docs/",dir(path="docs", pattern=ext[i]))) # delete files from /docs folder4
+  unlink(paste0("docs/",(ext[i])),recursive = TRUE)             # Remove the cache files (first)
+}
+
+for (i in 1:length(ext)) {
+  unlink(paste0("docs/",(ext[i])),recursive = TRUE) # Remove the cache folders (second)
+}
+
+
+# [ALTERNATIVE:] delete files from /production folder (in case you previously rendered these files)
+prod.files <- c(grep("^prod_.*\\_cache$",x = dir(path = "production"),value = T),
+                grep("^prod_.*\\_files$",x = dir(path = "production"),value = T),
+                grep("^prod_.*\\.html$" ,x = dir(path = "production"),value = T),
[…]

### 2021-05-25 production/merit_per_pref.R

@@ -1,574 +1,0 @@
-# Meritocracy/non-meritocracy perceptions and preferences ISSP
-
-# Juan Carlos Castillo, June 2017 - jc-castillo.com
-
-setwd("/media/ntfs/Dropbox/proyectos y papers compartidos/Meritocracy and preferences/meritocracy_pref_international/loc_merit_meassure") # jc
-
-# 0.DATA & LIBRARIES ---------------------- -
-
-# 0.1 Libraries ------------------------- ####
-
-  # Libraries
-    pacman::p_load(sjPlot,
-     sjmisc, # several functions,ej read_stata
-     descr,
-     memisc,  # codebook
-     psych, # cor2latex
-     corrplot, # graph correlations
-     lavaan,
-     stargazer,
-     semTools,
-     plyr, # count
-     dplyr,
-     car, # recode
-     ggplot2,
-		 Hmisc,
-		 tidyverse,
-		 data.table,
-  	 stringr,
-  	 countrycode,
-  	 lme4,
-     texreg,
-     mice,
-     ggrepel, # for scatter labels in sjp.scatter
-     semPlot,
-		 haven,
-		 sjlabelled)
-
-# 0.2 Original data preparation --------- ####
-
-    #issp2009=read_stata("https://www.dropbox.com/s/gkq8lue7nnl9joy/ZA5400_v4-0-0.dta") # takes a while but it works
-    issp2009=read_dta("../data/issp2009/ZA5400_v4-0-0.dta")
-    names(issp2009) <- tolower(names(issp2009)) # change to small
-
-  # Check data
-    names(issp2009)
-
-  # Check & generate IDs (v4: respondent, v5: country)
-    tapply(issp2009$v4, issp2009$v5, mean, na.rm=TRUE)
-    # Algunos paises (ej:826) tienen ids raros ...
-
-  # Generar nueva id secuencial individuos por país para merges
-    issp2009$id <- sequence(tabulate(issp2009$v5))
-    tapply(issp2009$id, issp2009$v5, mean, na.rm=TRUE) # ok
-    summary(issp2009$id)
-    issp2009$idtem=issp2009$v5*10000
-    issp2009$id=issp2009$idtem + issp2009$id
-
-# End sec 0 -----------------------------
-
-# 1. Measuring perceptions (getting-ahead) --
-
-# 1.1 Data preparation, gen getahead.rda  #####
-
-  names(issp2009)
[…]

### 2021-05-25 render-site.R

@@ -1,61 +1,0 @@
-
-# Render site ------------------------------------------------------------
-
-###### logic
-# there are some limitations for rendering websites in Rmarkdown/Githubpages
-# as all files should be in a same directory, which goes against the logic
-# of a clean and organized folder structure. Therefore, this script:
-# 1. bring all necessary files to the "docs" folder, where the site files are hosted
-# 2. renders the site from this location (docs folder)
-# 3. delete the Rmd files that where brought in step 1, leaving only the html files
-
-
-# INSTRUCTIONS-------------------------:
-# 1. set your root dir at "merit-scale"
-# 2. Run this code.4
-
-production<- list.files(path = "production",pattern = "prod_")        # a list with the names of the files to copy
-file.copy(file.path("production",production), "docs",overwrite = TRUE)# copy data proc and analysis files
-
-rmarkdown::render_site("docs",quiet = FALSE) # Render site
-
-# before you run this line, check if your R Markdown files have a .rmd or .Rmd extension 
-# on this case we use both
-ext <- c(grep("^prod_.*\\.rmd$",  x = dir(path = "docs"),value = T), # for .rmd
-         grep("^prod_.*\\.Rmd$",  x = dir(path = "docs"),value = T), # for .Rmd
-         grep("_cache",x = dir(path = "docs"),value = T),            # for files within the "_cache" folder
-         grep("^prod_.*\\_cache$",  x = dir(path = "docs"),value = T), # 
-         grep("^prod_.*\\.R$",    x = dir(path = "docs"),value = T), # for .R
-         ".tex",".log",".RData",".Rhistory")                         # for .tex .log .RData .Rhistory
-
-for (i in 1:length(ext)) {
-  file.remove(paste0("docs/",dir(path="docs", pattern=ext[i]))) # delete files from /docs folder4
-  unlink(paste0("docs/",(ext[i])),recursive = TRUE)             # Remove the cache files (first)
-}
-
-for (i in 1:length(ext)) {
-  unlink(paste0("docs/",(ext[i])),recursive = TRUE) # Remove the cache folders (second)
-}
-
-
-# [ALTERNATIVE:] delete files from /production folder (in case you previously rendered these files)
-prod.files <- c(grep("^prod_.*\\_cache$",x = dir(path = "production"),value = T),
-                grep("^prod_.*\\_files$",x = dir(path = "production"),value = T),
-                grep("^prod_.*\\.html$" ,x = dir(path = "production"),value = T),
[…]

### 2021-06-02 paper_study1.Rmd
update

@@
 author: "Juan Carlos Castillo, Julio Iturra, Luis Maldonado, Francisco Meneses & Jorge Atria <br><br>"
-abstract: "Economic and social inequalities have generated growing concern and crises across contemporary societies.
-One of the mechanisms proposed by social sciences to explain the persistence of inequality is the belief in meritocracy, which would legitimize economic disparities based on differences in effort and talent.
-Despite its wide use as a concept, empirical research on meritocracy is relatively novel and characterized by diverse conceptualization and measures that make the findings and their interpretation rather confusing.
-Most of the studies in the area have relied upon secondary data to operationalize meritocracy, with a wide variation in the use and interpretation of the same survey items.
-Taking into account the extant literature that uses measures of meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding conceptualization and operationalization.
-Based on this critical analysis, we propose a survey instrument called *Perceptions and Preferences for Meritocracy Scale*, which we test with confirmatory factor analysis using data from an online survey study (N=2,141).
-Our results support the proposed conceptual structure, which not only distinguishes between perceptions and preferences, but also between meritocratic and non-meritocratic dimensions.
-Our discussion highlights the importance of considering these different dimensions in order to advance the study of meritocracy."
 output:
+  bookdown::html_document2:
+    number_sections: false
   bookdown::pdf_document2:
@@
     toc: false
-  bookdown::html_document2:
-    number_sections: false
 linkcolor: blue
-bibliography:
-  - ../input/bib/EconomiaMoral.bib
-  - ../input/bib/meritocracy.bib
-  - ../input/bib/invariance.bib
+# bibliography:
+  # - ../input/bib/EconomiaMoral.bib
+  # - ../input/bib/meritocracy.bib
+  # - ../input/bib/invariance.bib
 editor_options:
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = FALSE,out.width = '85%',fig.pos= "H")
+knitr::opts_chunk$set(
+	echo = FALSE,
+	fig.pos = "H",
+	message = FALSE,
+	warning = FALSE,
+	cache = FALSE,
+	out.width = "85%"
+)
 # knitr::opts_knit$set(base.url = "../") #relative path for .html output file
@@
 The data used in the analysis corresponds to the last available wave of the social inequality module from the International Social Survey Programme (ISSP), which is the most specialized international comparative survey in perceptions, attitudes and beliefs about inequality related issues.
-This wave corresponds to the year 2009 and covers attitudes towards a series of topics dealing with social inequality across 26 countries and XXXX individuals.
+This wave corresponds to the year 2009 and covers attitudes towards a series of topics dealing with social inequality across 41 countries and 56021 individuals.
 Although there is data available from this module for previous waves (1987, 1992 and 1999), unfortunately there are several variables that are important for this research that were not included, particularly in 1987 and 1999, reason why we only use the 2009 wave.
@@
 ```{r desc-issp, echo=FALSE, results='asis'}
-load(file = "../output/tables/desc.issp.Rdata")
-sum.issp09<- kable(desc.issp,format =table_format,
+load(file = "input/data/proc/merit_per_pref.rda")
+pacman::p_load(dplyr)
+merit_per_pref %>% 
+  dplyr::select(ambition,hwork,wealthy,pareduc,people,polcone,
+                race,gender,welljob,hardjob,family,child) %>% 
+  psych::describe() %>% 
+  as.data.frame() %>% 
+  dplyr::select("Mean"=mean,"SD"=sd,"Min"=min,"Max"=max) %>% 
+  round(2) ->desc.issp
+
+row.names(desc.issp) <- c(
+  "Importance: hard work",
+  "Importance: having ambition",
+  "Importance: wealthy family",
+  "Importance: educated parents",
+  "Importance: race",
+  "Importance: gender",
+  "Importance: knowing people",
+  "Importance: political connections",
+  "Reasons for pay: well job",
+  "Reasons for pay: hard job",
+  "Reasons for pay: support family",
+  "Reasons for pay: has children"
+)
+
+sum.issp09<- knitr::kable(desc.issp,format =table_format,
                  booktabs= T,
@@
                  caption = "Descriptive statistics for the ISSP 2009 sample") %>%
-  kable_styling(full_width = table_format2,
-                font_size = 10,
+  kableExtra::kable_styling(full_width = table_format2,
+                            font_size = 15,
                             position = "center",
@@
 
+```{r}
+load(file = "input/data/proc/merit_per_pref.rda")
+cfa_perpref1 <- '
+	  # latent variables
+	  percmerit =~ hwork + ambition
+	  percnmerit=~ wealthy + pareduc +race + gender +people + polcone
+	  prefmerit =~ welljob + hardjob
+	  prefnmerit=~ family + child
+	  '
+fit_perpref1 <- lavaan::cfa(cfa_perpref1, data = merit_per_pref,
+                            ordered = c("wealthy","pareduc","ambition","hwork",
+                                        "people","polcone","race","gender",
+                                        "respons","yeduc","family","child",
+                                        "welljob","hardjob"))
+```
+
 <!-- [diagrama SEM] -->
@@
 
-```{r meas02-issp, echo=FALSE, fig.cap = "Descriptive plot", fig.align='center'}
-knitr::include_graphics('../output/images/meas02-issp.PNG')
+```{r meas02-issp, echo=FALSE, fig.cap = "Descriptive plot", fig.align='center',fig.cap="Meassurement model for Perceptions and Preferences in meritocracy"}
+knitr::include_graphics('output/images/meas02-issp.PNG')
 ```
@@
 ```{r tb-perpref1, echo=FALSE}
-load(file = "../output/tables/tb.perpref1.RData")
+load(file = "input/data/proc/merit_per_pref.rda")
+pacman::p_load(dplyr,knitr,kableExtra)
+cfa_perpref1 <- '
+	  # latent variables
+	  percmerit =~ hwork + ambition
+	  percnmerit=~ wealthy + pareduc +race + gender +people + polcone
+	  prefmerit =~ welljob + hardjob
+	  prefnmerit=~ family + child
+	  '
+fit_perpref1 <- lavaan::cfa(cfa_perpref1, data = merit_per_pref,
+                            ordered = c("wealthy","pareduc","ambition","hwork",
+                                        "people","polcone","race","gender",
+                                        "respons","yeduc",
+                                        "family","child",
+                                        "welljob","hardjob"))
+labs <- c(
+  "How important is having ambition?",
+  "How important is hard work?",
+  "How important is coming from a wealthy family?",
+  "How important is having welleducated parents",
+  "How important is knowing the right people",
+  "How important is having political connections",
+  "How important is a person's race?",
+  "How important is being born a man or a woman?",
+  "How well he or she does the job?",
+  "How hard he or she works at the job?",
+  "What is needed to support a family?",
+  "Whether the person has children to support?"
+)
+
+tb.load<- data.frame(round(cbind(lavaan::inspect(fit_perpref1,
+                                                        what="std")$lambda),
+                           digits = 2))
+tb.load[tb.load==c(0.00)] <- NA
+
+for (i in names(tb.load)) {
+  # tb.load[,i] <- sjlabelled::as_character(tb.load[,i])
+  tb.load[,i] <- sprintf(tb.load[,i], fmt = '%#.2f')
+}
+tb.load[tb.load=="NA"] <- ""
+#-------#
+fm01<- data.frame(t(data.frame(lavaan::fitmeasures(fit_perpref1, output ="matrix")[c("chisq","df","cfi","tli","rmsea"),]))); row.names(fm01) ="percnmerit"
+
+#------chi2, df------#
+fm04<- round(rbind(fm01),3)
+fm04.1 <- fm04 %>% dplyr::select(chisq,df) 
+fm04.1$chisq <- round(x = fm04.1$chisq,digits = 1)
+fm04.1$df <- round(x = fm04.1$df,digits = 0)
+fm04.1$chi2df <- paste0(fm04.1$chisq,"(",fm04.1$df,")")
+fm04.1 <- dplyr::select(fm04.1,"chi2df")
+for (i in names(fm04.1)) {
+  fm04.1[,i] <- as.character(fm04.1[,i])
+}
+
+#------CFI, RMSEA------#
+fm04.2 <- fm04 %>% dplyr::select(cfi,tli,rmsea) 
+for (i in names(fm04.2)) {
+  fm04.2[,i] <- sprintf(fm04.2[,i], fmt = '%#.3f')
+}
+
+fm.df      <- dplyr::bind_cols(fm04.1,fm04.2)
+fm.df$nobs <- c(lavaan::nobs(fit_perpref1)) 
+fm.df <- data.frame(t(fm.df)); colnames(fm.df) <- c("percnmerit")
+
+#------ merge ------#
+tb.fm<- dplyr::bind_rows(tb.load,fm.df)
+tb.fm<- tb.fm %>% 
+  dplyr::mutate(vars=c(labs,"$\\chi^2\\text{(df)}$","$\\text{CFI}$",
+                       "$\\text{TLI}$","$\\text{RMSEA}$","$N$")) %>%
+  dplyr::select(vars,everything())
+tb.perpref1 <- tb.fm
+
+# Table --------------------------------------------------------------------.
 tb.foot <- paste0("Standardized factor loadings using DWLS estimator ; CFI = Comparative fit index ;TLI = Tucker-Lewis index; RMSEA = Root mean square error of approximation")
-tb.col <- c("","Meritocratic","Non-meritocratic","Meritocratic","Non-meritocratic")
+tb.col <- c("","Meritocratic","Non-meritocratic",
+            "Meritocratic","Non-meritocratic")
 tb.caption <- c(" Factor loadings and fit measures for ISSP 2009")
-tb.issp<- kable(tb.perpref1,escape = FALSE,align = "lcccc",col.names = tb.col, caption = tb.caption, booktabs = T,linesep = "",format = table_format) %>%
-          kable_styling(full_width = table_format2,latex_options = c("hold_position"), position = "center",font_size = 12) %>%
+tb.issp<- kable(tb.perpref1,escape = FALSE,
+                align = "lcccc",
+                col.names = tb.col, 
+                caption = tb.caption, 
+                booktabs = T,
+                linesep = "",
+                format = table_format) %>%
+          kable_styling(full_width = table_format2,
+                        latex_options = c("hold_position"), 
+                        position = "center",
+                        font_size = 12) %>%
           add_header_above(header = c(" "=1,"Perception"= 2,"Preference"= 2)) %>%

### 2021-06-02 processing/analysis_study1.Rmd
update

---
title: "Analysis Study 1"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
output:
  bookdown::html_document2:
    number_sections: true
    toc: true
    toc_float:
        collapsed: false
    toc_depth: 2
    code_folding: "hide"
editor_options:
  chunk_output_type: console
---

# Setup

```{r setup}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      echo = TRUE 
                      )
knitr::opts_knit$set(root.dir = "../")
options(scipen=9999) # desactivar notacion cientifica
rm(list=ls())
options(knitr.kable.NA = '')
```

```{css, echo=FALSE}
# /* expandir el espacio del documento*/
body .main-container {
      max-width: 1600px !important;
      width: 1600px !important;
    }
#/* expandir width de la TOC*/
div.tocify {
    width: 22% !important;
    max-width: 331px !important;
    max-height: 102% !important;
}
```

## Libraries

```{r}
if (!require("pacman")) install.packages("pacman") # instalar pacman
                            # cargar librerias
pacman::p_load(dplyr,sjPlot,lavaan,psych, knitr,kableExtra,ggplot2,stargazer,
               corrplot,semPlot)
```

```{r}
table_format = if(is_html_output()) {
  "html"
} else if(is_latex_output()) {
  "latex"
}

```

# Analysis

## Descriptives

- Meritocratic perception:
  - How important is having ambition?
  - How important is hard work?

- Non-meritocratic perception
  - How important is coming from a wealthy family?
  - How important is having well-educated parents
  - How important is knowing the right people
  - How important is having political connections
  - How important is a person's race?
  - How important is being born a man or a woman?

- Meritocratic preferences
  - How well he or she does the job?
  - How hard he or she works at the job?

- Non-meritocratic preferences
  - What is needed to support a family?
  - Whether the person has children to support?

```{r}
load(file = "input/data/proc/merit_per_pref.rda")
pacman::p_load(dplyr,sjPlot,lavaan,psych, knitr,kableExtra,ggplot2,stargazer,
               corrplot,semPlot)
names(merit_per_pref)
merit_per_pref %>% 
  dplyr::select(ambition,hwork,wealthy,pareduc,people,polcone,race,gender,welljob,hardjob,family,child) %>% 
  psych::describe() %>% 
  as.data.frame() %>% 
  dplyr::select("Mean"=mean,"SD"=sd,"Min"=min,"Max"=max) %>% 
  round(2) ->desc.issp

row.names(desc.issp) <- c(
  "Importance: hard work",
  "Importance: having ambition",
  "Importance: wealthy family",
  "Importance: educated parents",
  "Importance: race",
  "Importance: gender",
  "Importance: knowing people",
  "Importance: political connections",
  "Important for pay: well job",
  "Important for pay: hard job",
  "Important for pay: support family",
  "Important for pay: has children"
)

# save(desc.issp,file = "output/tables/desc.issp.Rdata")
```

## Measurement model

```{r}    
# 3. MEASURING BELIEFS & PERCEPTIONS 
load(file = "input/data/proc/merit_per_pref.rda")
cfa_perpref1 <- '
	  # latent variables
	  percmerit =~ hwork + ambition
	  percnmerit=~ wealthy + pareduc +race + gender +people + polcone
	  prefmerit =~ welljob + hardjob
	  prefnmerit=~ family + child
	  '
fit_perpref1 <- lavaan::cfa(cfa_perpref1, data = merit_per_pref,
                            ordered = c("wealthy","pareduc","ambition","hwork",
                                        "people","polcone","race","gender",
                                        "respons","yeduc",
                                        "family","child",
                                        "welljob","hardjob"))
summary(fit_perpref1,
        fit.measures = TRUE,
        standardized = TRUE)
```

```{r, fig.width=10,fig.height=8}
load(file = "input/data/proc/merit_per_pref.rda")
plot_m_perpref1 <- '
	  # latent variables
	  percmerit =~ 1*hwork + ambition
	  percnmerit=~ 1*wealthy + pareduc +race + gender +people + polcone
	  prefmerit =~ 1*welljob + hardjob
	  prefnmerit=~ 1*family + child
	  '
plot_perpref1 <- lavaan::cfa(plot_m_perpref1, data = merit_per_pref)

nodeNames<- c(
  "Importance: hard work",
  "Importance: having ambition",
  "Importance: wealthy family",
  "Importance: educated parents",
  "Importance: race",
  "Importance: gender",
  "Importance: knowing people",
  "Importance: political connections",
  "Important for pay: well job",
  "Important for pay: hard job",
  "Important for pay: support family",
  "Important for pay: has children",
  "Perception \n meritocratic",
  "Perception \n non meritocratic",
  "Preference \n meritocratic",
  "Preference \n non meritocratic"
)

xfa <- 1*0.5

# Path model
semPlot::semPaths(
  plot_perpref1 ,
  # whatLabels = "std", # Indica si aparecen los valores
  what = "mod",
  label.cex = 0.38, # Tamaño de las etiquetas dentro de los nodos.
  edge.label.cex =  0.8, #  0.8 Tamaño de los valores estimados.
  residuals = T,
  optimizeLatRes =T,
  edge.color = "black",
  style = "lisrel",
  nCharNodes = 0,
  curvePivot = F,
  curve = 3,
  rotation = 4,
  layout = "tree3",
  cardinal = "lat cov",
  legend.cex = 0.8, # 0.6
  label.cex = 4,
  label.font = 6,
  edge.label.font = 14,
  asize = 3,
  edge.width = 1.5,
  sizeMan = 6,# largo manifest
  sizeMan2= 3,# alto manifest
  sizeLat = 5, # largo de los circulos
  sizeLat2 = 11.5, # alto de los circulos
  residScale = 10,
  width =6, # 17
  height = 3, # 10
  # nodeNames = nodeNames,
  intercepts = F,
  reorder = T,
  thresholds = F,
  fixedStyle =1,
  node.height = 1,
  node.width = 3,
  label.scale = F,
  shapeMan = "rectangle",
  shapeLat = "ellipse",
  nodeLabels = nodeNames,
  details = T,
  mar = c(1,
          4, #left
          1, #top
          4), #right
  # filetype ="pdf",
  # filename = "output/images/meas-issp01"
  )->semplot01

```

```{r}
load(file = "input/data/proc/merit_per_pref.rda")
pacman::p_load(dplyr)
cfa_perpref1 <- '
	  # latent variables
	  percmerit =~ hwork + ambition
	  percnmerit=~ wealthy + pareduc +race + gender +people + polcone
	  prefmerit =~ welljob + hardjob
	  prefnmerit=~ family + child
	  '
fit_perpref1 <- lavaan::cfa(cfa_perpref1, data = merit_per_pref,
                            ordered = c("wealthy","pareduc","ambition","hwork",
                                        "people","polcone","race","gender",
                                        "respons","yeduc",
                                        "family","child",
                                        "welljob","hardjob"))
labs <- c(
  "How important is having ambition?",
  "How important is hard work?",
  "How important is coming from a wealthy family?",
  "How important is having welleducated parents",
  "How important is knowing the right people",
  "How important is having political connections",
  "How important is a person's race?",
  "How important is being born a man or a woman?",
  "How well he or she does the job?",
  "How hard he or she works at the job?",
  "What is needed to support a family?",
  "Whether the person has children to support?"
)

tb.load<- data.frame(round(cbind(lavaan::inspect(fit_perpref1,
                                                        what="std")$lambda),
                           digits = 2))
tb.load[tb.load==c(0.00)] <- NA

for (i in names(tb.load)) {
  # tb.load[,i] <- sjlabelled::as_character(tb.load[,i])
  tb.load[,i] <- sprintf(tb.load[,i], fmt = '%#.2f')
}
tb.load[tb.load=="NA"] <- ""
#-------#
fm01<- data.frame(t(data.frame(lavaan::fitmeasures(fit_perpref1, output ="matrix")[c("chisq","df","cfi","tli","rmsea"),]))); row.names(fm01) ="percnmerit"

#------chi2, df------#
fm04<- round(rbind(fm01),3)
fm04.1 <- fm04 %>% dplyr::select(chisq,df) 
fm04.1$chisq <- round(x = fm04.1$chisq,digits = 1)
fm04.1$df <- round(x = fm04.1$df,digits = 0)
fm04.1$chi2df <- paste0(fm04.1$chisq,"(",fm04.1$df,")")
fm04.1 <- dplyr::select(fm04.1,"chi2df")
for (i in names(fm04.1)) {
  fm04.1[,i] <- as.character(fm04.1[,i])
}

#------CFI, RMSEA------#
fm04.2 <- fm04 %>% dplyr::select(cfi,tli,rmsea) 
for (i in names(fm04.2)) {
  fm04.2[,i] <- sprintf(fm04.2[,i], fmt = '%#.3f')
}

fm.df      <- dplyr::bind_cols(fm04.1,fm04.2)
fm.df$nobs <- c(lavaan::nobs(fit_perpref1)) 
fm.df <- data.frame(t(fm.df)); colnames(fm.df) <- c("percnmerit")

#------ merge ------#
tb.fm<- dplyr::bind_rows(tb.load,fm.df)
tb.fm<- tb.fm %>% 
  dplyr::mutate(vars=c(labs,"$\\chi^2\\text{(df)}$","$\\text{CFI}$",
                       "$\\text{TLI}$","$\\text{RMSEA}$","$N$")) %>%
  dplyr::select(vars,everything())
tb.perpref1 <- tb.fm
```

```{r}
# save(tb.perpref1,file = "output/tables/tb.perpref1.RData")
```

### 2021-06-02 processing/preparation_study1.Rmd
update

---
title: "Data preparation: Study 1"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
output:
  bookdown::html_document2:
    number_sections: true
    toc: true
    toc_float:
        collapsed: false
    toc_depth: 2
    code_folding: "show"
editor_options:
  chunk_output_type: console
---

# Setup

```{r setup}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      echo = TRUE 
                      )
knitr::opts_knit$set(root.dir = "../")
options(scipen=9999) # desactivar notacion cientifica
```

```{css, echo=FALSE}
# /* expandir el espacio del documento*/
body .main-container {
      max-width: 1600px !important;
      width: 1600px !important;
    }
#/* expandir width de la TOC*/
div.tocify {
    width: 22% !important;
    max-width: 331px !important;
    max-height: 102% !important;
}
```

## Libraries
```{r}
if (!require("pacman")) install.packages("pacman") # install pacman
pacman::p_load(dplyr,lavaan,sjPlot,sjmisc,sjlabelled,haven,stringr,summarytools)
```

# Data

```{r}
issp2009 <- read_dta("../input/data/original/ZA5400_v4-0-0.dta")
```

```{r }
names(issp2009) <- tolower(names(issp2009))
# Check & generate IDs (v4: respondent, v5: country)
tapply(issp2009$v4, issp2009$v5, mean, na.rm = TRUE)
# Algunos paises (ej:826) tienen ids raros ...
# Generar nueva id secuencial individuos por país para merges
issp2009$id <- sequence(tabulate(issp2009$v5))
tapply(issp2009$id, issp2009$v5, mean, na.rm = TRUE) # ok
summary(issp2009$id)
issp2009$idtem = issp2009$v5 * 10000
issp2009$id = issp2009$idtem + issp2009$id
```

## Perceptions (getting-ahead)

*  v6: wealthy family
*  v7: educated parents
*  v9: ambition 
* v10: hard work 
* v11: right people
* v12: political connections
* v14: race 
* v16: gender 
* IDs: v5 (country) & id (individual id)

```{r}
# names(issp2009)          #Check names
get_label(issp2009[8:18])#Check variable labels

# 1. Select and generate new variables
getahead <- 
  issp2009 %>%
  select("id","v5","v6","v7","v9","v10","v11","v12","v14","v16") %>% 
  mutate(wealthy = v6,
         pareduc = v7,
         ambition = v9,
         hwork = v10,
         people = v11,
         polcone = v12,
         race = v14,
         gender = v16) %>% 
  select(-c("v6","v7","v9","v10","v11","v12","v14","v16"))

names(getahead)

#2. Recode missing values
# * 8 & 9 to NA
frq(getahead$wealthy)
getahead[3:10][getahead[3:10] > 5] <- NA
frq(getahead$gender) # OK
summary(getahead)    # OK

#3. value labels
get_labels(getahead$wealthy)
getahead[, 3:10] <- 
  set_na(x = getahead[, 3:10],
         na = c(8,9))
get_labels(getahead[3:10]) # OK
frq(getahead[3:10])        # OK

# 4. variable label

getahed.label <-  str_remove_all(string = get_label(getahead[3:10]),
                                 pattern = "Q1[abcdefghijk] ")
getahed.label

getahead[3:10] = 6 - getahead[3:10] # Reverse code for SEM models
getahead[3:10] <- set_label(getahead[3:10],label = getahed.label) # variable label
get_label(getahead[3:10]) # OK

# 5. final ckeck   
view_df(getahead,show.frq = T, show.prc = T)
```

## Preferences (detearn)

```{r}	    
# names(issp2009)

# 1. Select and generate new variables
detearn <- 
  issp2009 %>%
  select("id","v5","v47", "v48", "v49", "v50", "v51","v52") %>% 
  mutate(respons 	= v47,
         yeduc  	= v48,
         family   = v49,
         child    = v50,
         welljob  = v51,
         hardjob  = v52) %>% 
  select(-c("v47", "v48", "v49", "v50", "v51","v52"))

#2. Recode missing values
# * 8 & 9 to NA
summary(detearn)
frq(detearn$respons)
detearn[3:8][detearn[3:8] > 5] <- NA
frq(detearn$respons)

# Out Portugal (no answer in well job item)
detearn[which(detearn$v5 == 620), ] = NA

#3. value labels
get_labels(detearn[3:8])
detearn[, 3:8] <- 
  set_na(detearn[, 3:8],
         na =  c(8, 9)) 
summary(detearn)      # Value 0 in Portugal Answers
frq(detearn$welljob)  # Check

detearn$welljob[detearn$welljob == 0] = NA # recode to NA value 0 in welljob
get_labels(detearn$welljob)
frq(detearn$welljob)

# value label correction in welljob
detearn$welljob <- set_labels(x = detearn$welljob,
                              labels = get_labels(detearn$hardjob))
frq(detearn$welljob) # OK, solved
frq(detearn[3:8])    # Check value labels

# 4. Variable label
get_label(detearn[3:8])
detearn.label <-  str_remove_all(string = get_label(detearn[3:8]),
                                 pattern = "Q12[abcdef] ")

detearn[3:8] <-  6 - detearn[3:8] # Reverse code for SEM models
detearn[3:8] <- set_label(detearn[3:8],label = detearn.label) # variable label

frq(detearn[3:8])

# 5. final ckeck   
view_df(detearn,show.frq = T, show.prc = T)
```

# Merge

```{r}
merit_per_pref   <- left_join(getahead, detearn, by = c("id", "v5"))

```

# Summary
```{r,results='asis', echo=FALSE}
st_css() # esto configura el css de dfsummary
print(dfSummary(merit_per_pref),
      method = "render")
```

```{r}
save(merit_per_pref, file = "../input/data/proc/merit_per_pref.rda")
```

### 2021-06-03 paper_study3.Rmd
Estudio 3
Se ajusta preparación del S3, se crean 2 bases data_s3_conv y data_s3_inv, para convergencia e invarianza.

Se ajusta paper_s3, reproducible.

@@
 link-citations: yes
-author: "Juan Carlos Castillo, Julio Iturra, Luis Maldonado, Francisco Meneses & Jorge Atria <br><br>"
-abstract: "Economic and social inequalities have generated growing concern and crises across contemporary societies.
-One of the mechanisms proposed by social sciences to explain the persistence of inequality is the belief in meritocracy, which would legitimize economic disparities based on differences in effort and talent.
-Despite its wide use as a concept, empirical research on meritocracy is relatively novel and characterized by diverse conceptualization and measures that make the findings and their interpretation rather confusing.
-Most of the studies in the area have relied upon secondary data to operationalize meritocracy, with a wide variation in the use and interpretation of the same survey items.
-Taking into account the extant literature that uses measures of meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding conceptualization and operationalization.
-Based on this critical analysis, we propose a survey instrument called *Perceptions and Preferences for Meritocracy Scale*, which we test with confirmatory factor analysis using data from an online survey study (N=2,141).
-Our results support the proposed conceptual structure, which not only distinguishes between perceptions and preferences, but also between meritocratic and non-meritocratic dimensions.
-Our discussion highlights the importance of considering these different dimensions in order to advance the study of meritocracy."
 output:
@@
 linkcolor: blue
-bibliography:
-  - ../input/bib/EconomiaMoral.bib
-  - ../input/bib/meritocracy.bib
-  - ../input/bib/invariance.bib
+#bibliography:
+#  - ../input/bib/EconomiaMoral.bib
+#  - ../input/bib/meritocracy.bib
+#  - ../input/bib/invariance.bib
 editor_options:
@@
 
-  |                                      | Social Origin  |     Hard Work    |     Wherewithal    |
-  |--------------------------------------|----------------|------------------|--------------------|
-  |     Meritocratic     Perception      |     -0.1975    |     0.1877       |     0.3921         |
-  |     Unmeritocratic    Perception     |     0.3547     |     -0.0775      |     -0.1844        |
-  |     Meritocratic    Preference       |     0.1078     |     0.0779       |     0.0819         |
-  |     Unmeritocratic    Preference     |     -0.0397    |     0.0824       |     0.2082         |
-  Table: `r as.character(paste("(\\#tab:val-conv)", "Polyserial and Pearson correlation whit other merit scales"))`.
+```{r val-conv}
+pacman::p_load(kableExtra, polycor)
+
+load(file="input/data/proc/data_s3_conv.RData")
+
+pl1 <- polycor::polyserial(data_s3_conv$perc_merit,data_s3_conv$get_ah_1)
+pl2 <- polycor::polyserial(data_s3_conv$pref_merit,data_s3_conv$get_ah_1)
+pl3 <- polycor::polyserial(data_s3_conv$perc_nmerit,data_s3_conv$get_ah_1)
+pl4 <- polycor::polyserial(data_s3_conv$pref_nmerit,data_s3_conv$get_ah_1)
+valores1 <- round(c(pl1,pl2,pl3,pl4),4)
+
+pl1b <- polycor::polyserial(data_s3_conv$perc_merit,data_s3_conv$get_ah_5)
+pl2b <- polycor::polyserial(data_s3_conv$pref_merit,data_s3_conv$get_ah_5)
+pl3b <- polycor::polyserial(data_s3_conv$perc_nmerit,data_s3_conv$get_ah_5)
+pl4b <- polycor::polyserial(data_s3_conv$pref_nmerit,data_s3_conv$get_ah_5)
+valores2 <- round(c(pl1b,pl2b,pl3b,pl4b),4)
+
+pl1c <- cor(data_s3_conv$perc_merit,data_s3_conv$nbmerit)
+pl2c <- cor(data_s3_conv$pref_merit,data_s3_conv$nbmerit)
+pl3c <- cor(data_s3_conv$perc_nmerit,data_s3_conv$nbmerit)
+pl4c <- cor(data_s3_conv$pref_nmerit,data_s3_conv$nbmerit)
+valores3 <- round(c(pl1c,pl2c,pl3c,pl4c),4)
+
+variables= c("Meritocratic Perception", "Meritocratic Preferences", "Unmeritocratic Perception","Unmeritocratic Preferences")
+
+table01 <- cbind(variables, valores1, valores2, valores3)
+
+kable(table01, col.names = c("Merit-Scale","Social Origin ", "Hard Work " , "Wherewithal"), full_width = T, linesep = "", caption = c("Polyserial and Pearson correlation whit other merit scales")) %>% kable_styling(
+    full_width = T,
+    position = "center",
+    font_size = 14,
+    bootstrap_options=c("striped", "bordered")) %>%
+  column_spec(column = 1, width = "1.5cm", )
+
+```
 
@@
 
-```{r invargroup, eval=TRUE, echo=FALSE}
-load(file = "../output/tables/tb.inv-group.RData")
+# Analysis
+
+```{r invargroup}
+
+#---Load packages and data -------------------------------------#
+
+pacman::p_load(sjPlot,dplyr,lavaan,semPlot,semTools,stargazer,corrplot,psych,knitr,kableExtra,stats)
+
+load(file = "input/data/proc/pov01_fact02.RData")
+
+pov01 <- data_s3_inv %>% filter(dataset=="pvw01")
+fs01  <- data_s3_inv %>% filter(dataset=="fsw02")
+
+#-----Meassurement model--------------------------------------#
+
+model01 <- '
+perc_merit  =~ perc_effort + perc_talent
+perc_nmerit =~ perc_wpart  + perc_netw
+pref_merit  =~ pref_effort + pref_talent
+pref_nmerit =~ pref_wpart  + pref_netw'
+
+#-----Invariance---------------------------------------------#
+
+inv01<- measurementInvariance(model=model01,data=data_s3_inv,group="dataset",estimator = "ML",strict=TRUE)
+
+conf  <- inv01$fit.configural
+weak  <- inv01$fit.loadings
+strong<- inv01$fit.intercepts
+strict<- inv01$fit.residuals
+
+tab01<- anova(conf,weak,strong,strict,SB.classic=TRUE) %>% 
+  as_tibble() %>% 
+  select("Chisq","Df","chisq_diff"=`Chisq diff`,"df_diff"=`Df diff`,"pvalue"=`Pr(>Chisq)`) %>% 
+  mutate(stars=gtools::stars.pval(pvalue),
+         chisqt=paste0(round(Chisq,2)," (",Df,") "),
+         decision=ifelse(pvalue>0.05,yes = "Accept",no = "Reject"),
+         model=c("Configural","Weak","Strong","Strict"))
+
+fit.meas<- bind_rows(fitmeasures(inv01$fit.configural,output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
+                     fitmeasures(inv01$fit.loadings,  output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
+                     fitmeasures(inv01$fit.intercepts,output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
+                     fitmeasures(inv01$fit.residuals, output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),])
+
+# compute differences in chisq, df, cfi and rmsea (90%, lower.ci - upper.ci )
+fit.meas<- fit.meas %>% 
+  mutate(diff.chi2 = chisq    - lag(chisq,default = first(chisq)),
+         diff.df   = df       - lag(df,   default = first(df)),
+         diff.cfi  = cfi      - lag(cfi,  default = first(cfi)),
+         diff.rmsea   = rmsea - lag(rmsea,default = first(rmsea))) %>%
+  round(3) %>% 
+  mutate(rmsea.ci=paste0(rmsea," \n ", "(",rmsea.ci.lower,"-",rmsea.ci.upper,")"))
+
+tab.inv<- bind_cols(tab01,fit.meas) %>% 
+  select(model,chisqt,cfi,rmsea.ci,diff.chi2,diff.df,diff.cfi,diff.rmsea,stars,decision) %>% 
+  mutate(diff.chi2=paste0(diff.chi2," (",diff.df,") ",stars)) %>% 
+  select(model,chisqt,cfi,rmsea.ci,diff.chi2,diff.cfi,diff.rmsea,decision)
+
+#clean values
+tab.inv[tab.inv == c("0 (0) ")] <- NA
+tab.inv[tab.inv == c(0)] <- NA
+
+### Check an alternative method
+configural        <- cfa(model01, data=data_s3_inv, group = "dataset", estimator="ML")
+weak.invariance   <- cfa(model01, data=data_s3_inv, group = "dataset",group.equal = "loadings", estimator="ML")
+strong.invariance <- cfa(model01, data=data_s3_inv, group = "dataset",group.equal = c( "loadings", "intercepts"), estimator="ML")
+strict.invariance <- cfa(model01, data=data_s3_inv, group = "dataset",group.equal = c( "loadings", "intercepts", "residuals"), estimator="ML")
+
+table_format = if(is_html_output()) {
+  "html"
+} else if(is_latex_output()) {
+  "latex"
+}
+
 col.nam <- c("Model","$\\chi^2 (\\text{df})$","CFI","RMSEA (90 CI)",
              "$\\Delta \\chi^2 (\\Delta \\text{df}$)","$\\Delta \\text{CFI}$","$\\Delta \\text{RMSEA}$","Decision")
-footnote <- paste0("N = ",nobs(conf),"; Group 1, n = ",conf@Data@nobs[[1]],"; Group 2, n = ",conf@Data@nobs[[2]], ", ***p < 0.001")
+footnote <- paste0("N = ","; Group 1, n = ",conf@Data@nobs[[1]],"; Group 2, n = ",conf@Data@nobs[[2]], ", **p < 0.001")
 
@@
   footnote(general = footnote, footnote_as_chunk = T)
+
 ```

### 2021-06-03 processing/preparation_study3.rmd
Estudio 3
Se ajusta preparación del S3, se crean 2 bases data_s3_conv y data_s3_inv, para convergencia e invarianza.

Se ajusta paper_s3, reproducible.

--- 2021-01-06 production/prod_prep-data.rmd
@@
 ---
-title: "Data preparation"
+title: "Data preparation Study 3 - Merit Scale"
 date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
-css: "../input/bib/style.css"
 output:
@@
 
-```{r eval=FALSE, include=FALSE}
-rmarkdown::render(input = "production/prod_prep-data-A.Rmd",output_format = "html_document",output_dir = "production/");browseURL(url = "production/prod_prep-data-A.html")
-```
+```{r include=FALSE}
+# Knitr options
 
-```{r include=FALSE}
 knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = FALSE,results = "hold")
@@
 
-load packages
+This is the preparation code of the data for Study 3:  Additional validity analyses. The prepared data is `data_s3_conv.RData` in "Convergence section" and `data_s3_conv.RData` in "Invariance section"
+
+# Convergence
+
+## Libraries
 ```{r}
-library(sjPlot)
-library(dplyr)
-library(tidyr)
-library(lavaan)
-library(stargazer)
-library(corrplot)
-library(psych)
-library(knitr)
-library(kableExtra)
-library(summarytools)
-library(sjlabelled)
-library(Hmisc)
+pacman::p_load(haven, sjPlot, dplyr, mice, lavaan, stargazer, corrplot, polycor, knitr, kableExtra, summarytools, sjlabelled)
 ```
 
-# Data preparation according order of the items
+## Load original data
 
 ```{r}
-w01 <- sjlabelled::read_spss("input/data/original/Estudio_3_ola1.sav",verbose = F)
+dat04w3 <- sjlabelled::read_spss("input/data/original/Estudio_3_ola3.sav",verbose = F)
+dim(data) 
 ```
 
-##  homogenize names of the items on the scale.
+## Create Perceptions and Preferences for Meritocracy (PPM-S) factor scores
 
-The items on the scale had different orders according to the order in which they were exposed to the respondents.
-There are three order.
-The first one the orders keeps together the items of the same factor.
-The second order then asks for perceptions and preferences regarding the same topic (i.e Distribution of resources according to efforts).
-In the third order, the items are presented randomly to each respondent.
+The following code creates new variables corresponding to the four dimensions of the PPM scale.
+Based on the measurement model, four factorial scores are created.
+To retrieve cases the "mice" function is used
+
+```{r results='hide'}
+dat04w3
+
+dat04w3$meritv03_perc_effort <- as.numeric(dat04w3$meritv03_perc_effort)
+dat04w3$meritv03_perc_talent <- as.numeric(dat04w3$meritv03_perc_talent)
+dat04w3$meritv03_perc_netw <- as.numeric(dat04w3$meritv03_perc_netw)
+dat04w3$meritv03_perc_wpart <- as.numeric(dat04w3$meritv03_perc_wpart)
+dat04w3$meritv03_pref_effort <- as.numeric(dat04w3$meritv03_pref_effort)
+dat04w3$meritv03_pref_talent <- as.numeric(dat04w3$meritv03_pref_talent)
+dat04w3$meritv03_pref_wpart <- as.numeric(dat04w3$meritv03_pref_wpart)
+dat04w3$meritv03_pref_netw <- as.numeric(dat04w3$meritv03_pref_netw)
+
+model04 <- '
+perc_merit =~  meritv03_perc_effort + meritv03_perc_talent
+perc_nmerit=~ meritv03_perc_wpart + meritv03_perc_netw
+pref_merit =~ meritv03_pref_effort + meritv03_pref_talent
+pref_nmerit=~ meritv03_pref_wpart + meritv03_pref_netw'
+
+# generate 5 multiple complete datasets
+out <- mice(dat04w3[23:30], m = 5)
+
+D1 <- complete(out, 1)
+D2 <- complete(out, 2)
+D3 <- complete(out, 3)
+D4 <- complete(out, 4)
+D5 <- complete(out, 5)
+
+# fit model for each complete dataset
+fit1 <- cfa(model04, data = D1)
+fit2 <- cfa(model04, data = D2)
+fit3 <- cfa(model04, data = D3)
+fit4 <- cfa(model04, data = D4)
+fit5 <- cfa(model04, data = D5)
+
+# predict scores for all models
+p1 <- predict(fit1)
+p2 <- predict(fit2)
+p3 <- predict(fit3)
+p4 <- predict(fit4)
+p5 <- predict(fit5)
+
+# compute 'average' across 5 sets of scores:
+scores <- (p1 + p2 + p3 + p4 + p5) / 5
+
+#Generate factor scores
+scores_dat = as.data.frame(scores)
+
+# Adaptar base para agregar factor scores (sacar todos NA)
+dat04w3 = dat04w3[rowSums(is.na(dat04w3[23:28])) != 5, ]
+
+# Merge with factor scores
+dat04w3 = cbind(dat04w3, scores)
+
+# Check
+stargazer(dat04w3[, 78:81], type = "text")
+
+```
+
+## Create The Neoliberal Beliefs Inventory (NBI) factor scores
+
+The following code creates new variables corresponding to the four dimensions of the NBI scale.
+Based on the measurement model, four factorial scores are created.
+To retrieve cases the "mice" function is used
+```{r results='hide'}
+datnbmerit <- dat04w3 %>% select(nb_merit_1,
+nb_merit_2,
+nb_merit_3,
+nb_merit_4,
+nb_merit_5,
+nb_merit_6,
+nb_merit_7,
+nb_merit_8)
+psych::alpha(datnbmerit)
+
+dat04w3$nb_merit_1 <- as.numeric(dat04w3$nb_merit_1)
+dat04w3$nb_merit_2 <- as.numeric(dat04w3$nb_merit_2)
+dat04w3$nb_merit_3 <- as.numeric(dat04w3$nb_merit_3)
+dat04w3$nb_merit_4 <- as.numeric(dat04w3$nb_merit_4)
+dat04w3$nb_merit_5 <- as.numeric(dat04w3$nb_merit_5)
+dat04w3$nb_merit_6 <- as.numeric(dat04w3$nb_merit_6)
+dat04w3$nb_merit_7 <- as.numeric(dat04w3$nb_merit_7)
+dat04w3$nb_merit_8 <- as.numeric(dat04w3$nb_merit_8)
+
+model04 <- '
+nbmerit =~  nb_merit_1+
+nb_merit_2+
+nb_merit_3+
+nb_merit_4+
+nb_merit_5+
+nb_merit_6+
+nb_merit_7'
+
+# generate 5 multiple complete datasets
+out <- mice(dat04w3[57:64], m = 5)
+
+D1 <- complete(out, 1)
+D2 <- complete(out, 2)
+D3 <- complete(out, 3)
+D4 <- complete(out, 4)
+D5 <- complete(out, 5)
+
+# fit model for each complete dataset
+fit1 <- cfa(model04, data = D1)
+fit2 <- cfa(model04, data = D2)
+fit3 <- cfa(model04, data = D3)
+fit4 <- cfa(model04, data = D4)
+fit5 <- cfa(model04, data = D5)
+
+# predict scores for all models
+p1 <- predict(fit1)
+p2 <- predict(fit2)
+p3 <- predict(fit3)
+p4 <- predict(fit4)
+p5 <- predict(fit5)
+
+# compute 'average' across 5 sets of scores:
+scores <- (p1 + p2 + p3 + p4 + p5) / 5
+
+#Generate factor scores
+scores_dat = as.data.frame(scores)
+
+# Adaptar base para agregar factor scores (sacar todos NA)
+dat04w3 = dat04w3[rowSums(is.na(dat04w3[57:64])) != 5, ]
+
+# Merge with factor scores
+dat04w3 = cbind(dat04w3, scores)
+
+data_s3_conv <- dat04w3 %>% select(nbmerit, perc_merit,pref_merit,perc_nmerit,pref_nmerit,get_ah_1, get_ah_5) %>% na.omit()
+
+```
+
+## Labels
+
+### Variable labels
+
+Created variables are labeled and described
 
 ```{r}
-dat04 <- w01 %>% filter(Intro==1) %>%
-  select(starts_with("meritv01"),
-         starts_with("meritv02"),
-         starts_with("meritv03_p"),
-         starts_with("FL_21_DO"))
-dat04$grupo <- NA
-dat04$grupo[dat04$FL_21_DO_merit_perc_pref_julio19v01==1] <- 1
-dat04$grupo[dat04$FL_21_DO_merit_perc_pref_julio19v02==1] <- 2
-dat04$grupo[dat04$FL_21_DO_merit_perc_pref_julio19v03==1] <- 3
-dat04$perc_effort <- rowSums(dat04[,c(matches(match = "perc_effort",vars = names(dat04)))],na.rm = TRUE)
-dat04$perc_talent <- rowSums(dat04[,c(matches(match = "perc_talent",vars = names(dat04)))],na.rm = TRUE)
-dat04$perc_wpart  <- rowSums(dat04[,c(matches(match = "perc_wpart" ,vars = names(dat04)))],na.rm = TRUE)
-dat04$perc_netw   <- rowSums(dat04[,c(matches(match = "perc_netw"  ,vars = names(dat04)))],na.rm = TRUE)
-dat04$pref_effort <- rowSums(dat04[,c(matches(match = "pref_effort",vars = names(dat04)))],na.rm = TRUE)
-dat04$pref_talent <- rowSums(dat04[,c(matches(match = "pref_talent",vars = names(dat04)))],na.rm = TRUE)
-dat04$pref_wpart  <- rowSums(dat04[,c(matches(match = "pref_wpart" ,vars = names(dat04)))],na.rm = TRUE)
-dat04$pref_netw   <- rowSums(dat04[,c(matches(match = "pref_netw"  ,vars = names(dat04)))],na.rm = TRUE)
-dat04[dat04==0] <- NA
-dat04 <- dat04 %>% select(starts_with(match = "perc_"),starts_with(match = "pref_"),grupo) %>% na.omit()
+
+set_label(data_s3_conv$nbmerit)  <- "Personal wherewithal - NBI"
+set_label(data_s3_conv$perc_merit)  <- "Meritocratic perception"
+set_label(data_s3_conv$pref_merit)   <- "Meritocratic preferences"
+set_label(data_s3_conv$perc_wpart)    <- "Unmeritocratic perception"
+set_label(data_s3_conv$pref_nmerit)  <-  "Unmeritocratic preferences"
+
+set_label(data_s3_conv$get_ah_1)  <- "Importance of a rich family to get ahead"
+set_label(data_s3_conv$get_ah_5)   <- "Importance of hard work to get ahead"
+
+sjmisc::descr(data_s3_conv)
+
 ```
 
-## Translation of variable labels and value labels.
+## Final summary
 ```{r}
-label(dat04$grupo) <- "This variable classifies the interviewees according to the order in which the indicators are shown within the scale.
-There are three orders  of items on the scale.
-In the first order (group 1), questions of the same factor are then presented to them, in the second order (group 2) they were separated and in the third order (group 3) at random."
-dat04$perc_effort <- set_labels(dat04$perc_effort,  labels = c("Strongly disagree" = 1, "Disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "Totally agree" = 5))
-label(dat04$perc_effort) <- "Thinking of Chilean society,
-what measure do you find of
-agree or disagree with each of
-the following statements? - Who
-the more they try they manage to get bigger
-rewards that those who strive
-less."
-dat04$perc_talent <- set_labels(dat04$perc_talent,  labels = c("Strongly disagree" = 1, "Disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "Totally agree" = 5))
-label(dat04$perc_talent) <- "Thinking of Chilean society,
-what measure do you find of
-agree or disagree with each of
-the following statements? - Who
-possess more talent they manage to obtain
-greater rewards than those who possess
-less talent."
-dat04$perc_wpart <- set_labels(dat04$perc_wpart,  labels = c("Strongly disagree" = 1, "Disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "Totally agree" = 5))
-label(dat04$perc_wpart) <- "Thinking of Chilean society,
-what measure do you find of
-agree or disagree with each of
-the following statements? - Who
-they have rich parents manage to get out
-ahead."
-dat04$perc_netw <- set_labels(dat04$perc_netw,  labels = c("Strongly disagree" = 1, "Disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "Totally agree" = 5))
-label(dat04$perc_netw) <- "Thinking of Chilean society,
-what measure do you find of
-agree or disagree with each of
-the following statements? - Who
-they have good contacts they manage to get out
-ahead."
-dat04$pref_effort <- set_labels(dat04$pref_effort,  labels = c("Strongly disagree" = 1, "Disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "Totally agree" = 5))
-label(dat04$pref_effort) <- "Thinking of Chilean society,
-what measure do you find of
-agree or disagree with each of
-the following statements? - Who
-the more they try they should get
-greater rewards than those who
-they try less."
-dat04$pref_talent <- set_labels(dat04$pref_talent,  labels = c("Strongly disagree" = 1, "Disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "Totally agree" = 5))
-label(dat04$pref_talent) <- "Thinking of Chilean society,
-what measure do you find of
-agree or disagree with each of
-the following statements? - Who
-possess more talent they should get
-greater rewards than those who possess
-less talent."
-dat04$pref_wpart <- set_labels(dat04$pref_wpart,  labels = c("Strongly disagree" = 1, "Disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "Totally agree" = 5))
-label(dat04$pref_wpart) <- "Thinking of Chilean society,
-what measure do you find of
-agree or disagree with each of
-the following statements? - It's fine that those who have rich parents get ahead" 
-dat04$pref_netw <- set_labels(dat04$pref_netw,  labels = c("Strongly disagree" = 1, "Disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "Totally agree" = 5))
-label(dat04$pref_netw) <- "Thinking of Chilean society,
-what measure do you find of
-agree or disagree with each of
-the following statements? - Is
-well that those who have good
-contacts get ahead."
-```
+sjmisc::descr(data_s3_conv,
+      show = c("label","range", "mean", "sd", "NA.prc", "n"))%>%
+      kable(.,"markdown")
 
-```{r}
-dat04 <-  dat04 %>% select(grupo, perc_effort, perc_talent, perc_wpart, perc_netw, pref_effort, pref_talent, pref_wpart , pref_netw) %>% as.data.frame()
-```
+print(dfSummary(data_s3_conv, valid.col = FALSE, graph.magnif = 0.75),
+      max.tbl.height = 300, method = "render")
 
-```{r}
-print(dfSummary(dat04, valid.col = FALSE, graph.magnif = 0.75),
-      max.tbl.height = 300, method = "render")
-```
-
-## Preparation of group 1: order according to the factors.
-
-```{r}
-dat01 <- dat04 %>% filter(grupo==1) %>% select(-grupo) %>%  na.omit()
-```
-
-```{r}
-print(dfSummary(dat01, valid.col = FALSE, graph.magnif = 0.75),
-      max.tbl.height = 300, method = "render")
-```
-
-## Preparation of group 2: order according to topics (i.e Effort).
-
-```{r}
-dat02 <- dat04 %>% filter(grupo==2) %>% select(perc_effort, pref_effort, perc_talent,pref_talent, perc_wpart,pref_wpart, perc_netw, pref_netw) %>% na.omit()
-```
-
-```{r}
-print(dfSummary(dat02, valid.col = FALSE, graph.magnif = 0.75),
-      max.tbl.height = 300, method = "render")
-```
-
-## Preparation of group 3: randomized order.
-
-```{r}
-dat03 <- dat04 %>% filter(grupo==3) %>% select(-grupo) %>%  na.omit()
-```
-
-```{r}
-print(dfSummary(dat03, valid.col = FALSE, graph.magnif = 0.75),
-      max.tbl.height = 300, method = "render")
 ```
@@
 ## save data
+
 ```{r}
-save(dat01,file = "input/data/proc/dat01.RData")
-save(dat02,file = "input/data/proc/dat02.RData")
-save(dat03,file = "input/data/proc/dat03.RData")
-save(dat04,file = "input/data/proc/dat04.RData")
+
+save(data_s3_conv,file = "input/data/proc/data_s3_conv.RData")
 ```
 
-# Longitudinal data
+# Invariance
 
-load data
+## Libraries
+```{r}
+pacman::p_load(haven, sjPlot, dplyr, lavaan, stargazer, corrplot, knitr, kableExtra, summarytools, sjlabelled,Hmisc,psych,tidyr)
+
+```
+
+## Load longitudinal data
+
 ```{r}
@@
 
+A variable is created to indicate the order in which the items on the PPM scale are presented. The variables are renamed and those of interest are selected. The results are described.
 ```{r}
@@
 
+A variable is created to indicate the treatment to which the respondent was exposed. The variables are renamed and those of interest are selected. The results are described.
+
 ```{r}
@@
 
+The variables of the PPM scale are renamed indicating the wave. Those of interest are selected. The results are described.
 ```{r}
@@
 ```{r}
-wide01 <- inner_join(x = w01c,y = w02c, by = "ID") # Solamente nos quedamos con casos en ambas olas
+wide01 <- inner_join(x = w01c,y = w02c, by = "ID") 
 dim(wide01) #en w01 y w02
@@
 ```{r}
-wide02 <- inner_join(x = wide01,y = w03c, by = "ID") # Solamente nos quedamos con casos en ambas olas
+wide02 <- inner_join(x = wide01,y = w03c, by = "ID") 
 dim(wide02) #en w01, w02 y w03
@@
 ```{r}
-save(wide01,file = "input/data/proc/datawide.RData")
-save(wide02,file = "input/data/proc/datawide_123.RData")
+#save(wide01,file = "input/data/proc/datawide.RData")
+#save(wide02,file = "input/data/proc/datawide_123.RData")
 ```
 
-# Wide to long data
+## Wide to long data
 
@@
                              "sexo","edad","edcep"))
+#save(long01,file = "input/data/proc/datalong.RData")
+
 ```
 
-```{r}
-#save(long01,file = "input/data/proc/datalong.RData")
-```
-
-# Factorial survey data:
+## Factorial survey data:
 
@@
 ```{r}
-fs01 <- sjlabelled::read_spss("input/data/original/factorial_ola2.sav",verbose = F)
+# fs01 <- sjlabelled::read_spss("input/data/original/factorial_ola2.sav",verbose = F)
+load(file = "papelera/input/data/original/factorial_ola2.RData") 
 ```
@@
 ```{r}
-pov_fs<-bind_rows(fs02,w01fs) %>%
+data_s3_inv<-bind_rows(fs02,w01fs) %>%
   set_label(label = get_label(fs02)) %>%
@@
 ```{r}
-save(pov_fs,file = "input/data/proc/pov01_fact02.RData")
+save(data_s3_inv,file = "input/data/proc/data_s3_inv.RData")
 ```

### 2021-06-11 paper_study1.Rmd

@@
 
-```{r meas02-issp, echo=FALSE, fig.cap = "Descriptive plot", fig.align='center',fig.cap="Meassurement model for Perceptions and Preferences in meritocracy"}
-knitr::include_graphics('output/images/meas02-issp.PNG')
+```{r meas02-issp, echo=FALSE, fig.cap = "Descriptive plot", fig.align='center',fig.cap="Confirmatory factor analysis of the Perceptions and Preferences for Meritocracy"}
+knitr::include_graphics('output/images/cfa_issp_study1.PNG')
 ```

### 2021-06-15 paper_study1.Rmd
update-study1

--- 2021-06-02 paper_study1.Rmd
@@
 	warning = FALSE,
-	cache = FALSE,
-	out.width = "85%"
+	cache = FALSE
 )
@@
 
-```{r message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
-#RUN THIS BEFORE THE ANALYSIS
-pacman::p_load(knitr)
-
-table_format = if(is_html_output()) {
-  "html"
-} else if(is_latex_output()) {
-  "latex"
-}
-
-table_format2 = if(is_html_output()) {
-  T
-} else if(is_latex_output()) {
-  F
-}
-```
-
 ## Study 1: Measuring meritocracy with international secondary data
 
-### Data
+## Data
 
@@
 
-### Variables
+## Variables
 
@@
 Please tick one box for each of these to show how important you think it is for getting ahead in life."_ This is followed by a list of statements to be answered in a scale with the following levels numbered from 1 to 5: essential, very important, fairly important, not very important, not important at all.
-Below the items are listed classified according to the meritocracy/non-meritocracy character:
 
-  - Meritocratic perception:
-    - How important is having ambition?
-    - How important is hard work?
-
-  - Non-meritocratic perception
-    - How important is coming from a wealthy family?
-    - How important is having well-educated parents
-    - How important is knowing the right people
-    - How important is having political connections
-    - How important is a person's race?
-    - How important is being born a man or a woman?
-
-<br>
 The classification is based on criteria of internal motivation (meritocratic) and structural constraints (non-meritocratic).
@@
 The ISSP question was: _In deciding how much people ought to earn, how important should each of these things be, in your opinion_, rated in the same _essential-non important at all_ scale as the questions for meritocratic preferences.
-The reasons are:
 
-  - Meritocratic preferences
-    - How well he or she does the job?
-    - How hard he or she works at the job?
+Table \@ref(tab:table-issp) below summarise the items classified according to the proposal based on perception and preference regarding to the meritocratic/non-meritocratic dimensions:
 
-  - Non-meritocratic preferences
-    - What is needed to support a family?
-    - Whether the person has children to support?
+```{r table-issp, echo=FALSE}
+pacman::p_load(knitr, kableExtra, dplyr)
+table_format = if(knitr::is_html_output()) {
+  "html"
+} else if(knitr::is_latex_output()) {
+  "latex"
+}
+
+tableissp <- read.csv(file = "input/tables/table-issp.csv",header = 1,sep = ";",encoding = "UTF-8") # call file generated externally for the table
+cnames <- c("Component", "Dimensions","Item")
+cap <- "Items of the ISSP meritocratic perceptions and preferences measures"
+
+kable(tableissp, table_format, booktabs = T, linesep = "",col.names = cnames, caption = cap) %>%
+  kable_styling(
+    full_width = F,
+    latex_options = c("hold_position"),
+    position = "center",
+    font_size = 12,
+    bootstrap_options=c("striped", "bordered")) %>%
+  column_spec(column = 1, width = "1.5cm", ) %>%
+  column_spec(column = 2,width = "2 cm") %>%
+  collapse_rows(columns = 1:2,valign = "middle")
+```
+
+# Methods
+
+The estimation was performed using Confirmatory Factor Analysis (CFA).
+CFA was conducted using the `lavaan` R package (version 0.6-3; Rosseel, 2020), with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement (Kline, 2016; Rosseel, 2020).
+As recommended by Brown (2008), we assessed model fit by jointly considering the comparative fit index and Tucker-Lewis Index (CFI and TLI; acceptable fit > 0.95), Root of the average squared residual approximation (RMSEA; acceptable fit < 0.08),  Chi-square: (p-value; acceptable fit > 0.05, and Chi-square ratio > 3).
+
+# Results
+
+## Descriptive analyses
 
@@
 
-Table \@ref(tab:desc-issp)
+<!-- Table \@ref(tab:desc-issp) -->
 
-```{r desc-issp, echo=FALSE, results='asis'}
+```{r desc-issp, eval=FALSE, include=FALSE, results='asis'}
 load(file = "input/data/proc/merit_per_pref.rda")
 pacman::p_load(dplyr)
+table_format = if(knitr::is_html_output()) {
+  "html"
+} else if(knitr::is_latex_output()) {
+  "latex"
+}
+
+table_format2 = if(knitr::is_html_output()) {
+  T
+} else if(knitr::is_latex_output()) {
+  F
+}
+
 merit_per_pref %>% 
-  dplyr::select(ambition,hwork,wealthy,pareduc,people,polcone,
-                race,gender,welljob,hardjob,family,child) %>% 
+  dplyr::select(hwork,ambition,wealthy,pareduc,
+                race,gender,people,polcone,welljob,hardjob,family,child) %>% 
   psych::describe() %>% 
@@
 
-### Methods
+Figure \@ref(fig:likert-issp) shows the distribution of responses across the indicators for each perception and preference in the meritocratic and non-meritocratic dimensions.
+On one side, we see that there is a certain degree of importance attributed to factors such as hard work and ambition in the process of getting ahead, concentrating 94.8% and 92.8% in the fairly important to essential categories.
 
-The estimation was performed using Confirmatory Factor Analysis (CFA).
-CFA was conducted using the `lavaan` R package (version 0.6-3; Rosseel, 2020), with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement (Kline, 2016; Rosseel, 2020).
-As recommended by Brown (2008), we assessed model fit by jointly considering the comparative fit index and Tucker-Lewis Index (CFI and TLI; acceptable fit > 0.95), Root of the average squared residual approximation (RMSEA; acceptable fit < 0.08),  Chi-square: (p-value; acceptable fit > 0.05, and Chi-square ratio > 3).
+By contrast, for the non-meritocratic aspects, parental background showed a more dispersed opinion, where 61.2% considers that coming from a wealthy family is important, and 76% considers that having educated parents is important.
+On the other hand, the background dimension shows that people' opinions are generally in the opposite sense, in which 64.2% consider that a person's racial origin is not very important or not important at all in the process of getting ahead in life, while 35% consider it important, and this is nearly identical regarding the importance of gender.
+Lastly, we can observe differences in the networks dimension, in which 85.5% consider it to be important to know the right people in the process of getting ahead in life, contrasting with 50% who consider political connections to be important.
 
-### Results
+Regarding normative preferences, we observe how well a person performs his or her job and amount of effort involved concentrates 98% and 96.8% of the responses in the categories that attribute importance to these characteristics when evaluating the pay that people deserve.
+Finally, in relation to structural constrains, 83.3% attribute importance to the fact that supporting a family should be a relevant factor with respect to the pay obtained, while having children to support translates into 76.5% of importance in this dimension.
 
@@
 load(file = "input/data/proc/merit_per_pref.rda")
+pacman::p_load(dplyr)
+likert_issp<- merit_per_pref %>% 
+  dplyr::select(hwork,ambition,wealthy,pareduc,
+                race,gender,people,polcone,
+                welljob,hardjob,family,child)
+
+likert_issp <- sjlabelled::set_label(likert_issp,
+                                     label = c("Hard work",
+                                               "Having ambition",
+                                               "Wealthy family",
+                                               "Educated parents",
+                                               "Race",
+                                               "Gender",
+                                               "Knowing people",
+                                               "Political connections",
+                                               "Well job",
+                                               "Hard job",
+                                               "Support family",
+                                               "Has children"))
+
+likert_issp <- sjmisc::rec(likert_issp, rec="rev", append=FALSE)
+likert_issp <- 
+  sjlabelled::set_labels(likert_issp,
+                         labels = c("Essential","Very important",
+                                    "Fairly important","Not very important",
+                                    "Not important at all"))
+
+plot_likert_issp<- sjPlot::plot_likert(likert_issp,
+                        c(1, 1, 2, 2, 2, 2, 2, 2,3,3,4,4),
+                        groups.titles = c("Perception meritocratic", "Perception Non-meritocratic",
+                                          "Preference meritocratic", "Preference Non-meritocratic"),
+                        geom.colors = "PuBu",
+                        geom.size = 0.8,
+                        catcount = 5,
+                        # cat.neutral = 3,
+                        grid.range  =  c(1.2 , 1.4),
+                        values  =  "sum.outside",
+                        reverse.colors = T,
+                        reverse.scale = F,
+                        show.n=FALSE)
+
+```
+
+```{r likert-issp,fig.height=10,fig.width=12,fig.cap="Distribution of responses in the ISSP meritocracy items"}
+plot_likert_issp
+```
+
+In terms of the association between the indicators, Figure \@ref(fig:corr-issp) shows the polychoric correlation matrix.
+Firstly, we see that the correlations between the indicators are consistent with the structure of the initially proposed structure of dimensions distinguishing between the perception and preferences of meritocratic and non-meritocratic factors.
+In general, the associations are positive between the perception and preference dimensions, such as the importance of hard work to get ahead and the importance of how hard the work is ($r$=0.34), however, they tend to be weak.
+A further aspect to consider is the relatively low correlation between dimensions concerning meritocratic and non-meritocratic aspects, (e.g.
+importance of hard work with educated parents, $r$=0.15).
+However, with respect to the non-meritocratic perception indicators, we can see that the associations demonstrate dimensionality, however, it is not entirely clear.
+In this regard, the indicators referring to parental characteristics have a positive and stronger association ($r$=0.56), but this association is much weaker in comparison to the other indicators of Non-meritocratic perception (e.g.
+Wealthy family and Gender, $r$=0.01).
+The same holds true for the Gender and Race indicators ($r$=0.6), and also applies to both the indicators of knowing the right people and having political connections ($r$=0.55).
+On this point, we will return to it in the next section.
+
+```{r corr-issp,fig.cap = "Perceptions & preferences for ISSP  meritocracy items' polychoric correlations", fig.align='center',fig.height=6,fig.width=8}
+pacman::p_load(dplyr)
+cor_issp<- 
+  merit_per_pref %>%
+  dplyr::select(hwork,ambition,wealthy,pareduc,
+                race,gender,people,polcone,
+                welljob,hardjob,family,child) %>% 
+  lavaan::lavCor(., ordered=names(.))
+diag(cor_issp) = NA
+rownames(cor_issp) <- c("A. Hard work",
+                       "B. Having ambition",
+                       "C. Wealthy family",
+                       "D. Educated parents",
+                       "E. Race",
+                       "F. Gender",
+                       "G. Knowing people",
+                       "H. Political connections",
+                       "I. Well job",
+                       "J. Hard job",
+                       "K. Support family",
+                       "L. Has children")
+colnames(cor_issp) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)", 
+                       "(H)", "(I)","(J)","(K)","(L)")
+corrplot::corrplot(cor_issp,
+  method = "color",
+  addCoef.col = "#000390",
+  type = "upper",
+  tl.col = "black",
+  col=colorRampPalette(c("white","#0068DC"))(12),
+  bg = "white",
+  na.label = "-",
+  lower.col = "black")
+```
+
+## Confirmatory Factor Analysis
+
+This section presents the results of the confirmatory factor analysis.
+First, a four-dimensional model for meritocratic and non-meritocratic perception and preferences is estimated, considering all indicators of the non-meritocratic perception dimension as a single latent variable.
+Then, we estimate a model that follows the proposed four latent variable measurement approach, however, the non-meritocratic perception dimension is assumed as a second-order factor for the first order latent variables parental attributes, racial and gender background and networks.
+
+```{r sum-fit-issp}
+load(file = "input/data/proc/merit_per_pref.rda")
+pacman::p_load(dplyr, kableExtra, knitr)
+
 cfa_perpref1 <- '
@@
 	  '
-fit_perpref1 <- lavaan::cfa(cfa_perpref1, data = merit_per_pref,
-                            ordered = c("wealthy","pareduc","ambition","hwork",
-                                        "people","polcone","race","gender",
-                                        "respons","yeduc","family","child",
-                                        "welljob","hardjob"))
+fit_perpref1 <- lavaan::cfa(cfa_perpref1, data = merit_per_pref,ordered = T)
+
+cfa_perpref2 <- '
+	  # latent variables
+	  percmerit =~ hwork + ambition
+	  parents =~ wealthy + pareduc
+	  backgrd =~ race + gender
+	  networks=~ people + polcone
+	  percnmerit=~ parents + backgrd + networks
+	  prefmerit =~ welljob + hardjob
+	  prefnmerit=~ family + child
+	  '
+fit_perpref2 <- lavaan::cfa(cfa_perpref2, data = merit_per_pref,ordered = T)
+
+# extract fit indices from models and add to table
+sum_fit_issp<- dplyr::bind_rows(
+  lavaan::fitmeasures(fit_perpref1)[c("chisq","df","cfi","tli","rmsea")],
+  lavaan::fitmeasures(fit_perpref2)[c("chisq","df","cfi","tli","rmsea")]
+  )
+
+# Customize object
+sum_fit_issp$mod <- c("First order","Second order")
+sum_fit_issp$nobs <- c(lavaan::nobs(fit_perpref2),lavaan::nobs(fit_perpref2))
+sum_fit_issp$est <- c("DWLS","DWLS")
+sum_fit_issp <- dplyr::select(sum_fit_issp,mod,nobs,est,dplyr::everything())
+colnames <- c("Model","$N$","Estimator","$\\chi^2$","df","CFI","TLI","RMSEA")
+
+# Create table
+table_issp_fits <-
+  knitr::kable(sum_fit_issp, format="html", digits=3, booktabs=T, 
+               col.names=colnames,  
+               caption = "Summary fit indices according to model", 
+               escape = FALSE) %>%
+  kableExtra::kable_styling(full_width = F, 
+                            font_size = 15,
+                            bootstrap_options=c("striped", "bordered"))
+table_issp_fits
 ```
+
+Table \@ref(tab:sum-fit-issp) shows the four-factor model fit permitting comparison of first and second-order factors.
+The model restricted to four first-order factor dimensions shows a regular fit (CFI=0.959, TLI=0.944, RMSEA=0.098, χ2(df=48)= 21308.535; more detail se Table \@ref(tab:tb-perpref1)) in contrast to the model considering a second-order factor for non-meritocratic perception that has a slightly better fit based on the standard fit measures (CFI=0.959, TLI=0.944, RMSEA=0.098, χ2(df=48)= 21308.535).
 
@@
 
-Figure \@ref(fig:meas02-issp)
-
-```{r meas02-issp, echo=FALSE, fig.cap = "Descriptive plot", fig.align='center',fig.cap="Meassurement model for Perceptions and Preferences in meritocracy"}
-knitr::include_graphics('output/images/meas02-issp.PNG')
+```{r meas02-issp, echo=FALSE, fig.cap = "Descriptive plot", fig.align='center',fig.cap="Confirmatory factor analysis of the ISSP indicators of Perceptions and Preferences for Meritocracy"}
+knitr::include_graphics('output/images/cfa_issp_study1.PNG')
 ```
@@
 <!-- [tabla CFA] -->
+
+Figure \@ref(fig:meas02-issp) shows the overall results for this model, evidencing an improvement in factor loadings as well as in the overall fit of the measurement model.
+On the one hand, we see that perceptions are correlated with preferences, but this correlation is stronger for the meritocratic dimension of perception and preference ($r$=0.46) than for the non-meritocratic ($r$=0.27), which is much stronger when looking at the correlations between meritocratic and non-meritocratic, both in perception ($r$=0.04) and preference ($r$=0.13).
+These results show that, on one hand, the positive correlation between perception indicates that the two are not completely opposite poles of a continuum ($r$=0.30), which is reflected in the correlation between preferences ($r$=0.25).
+Furthermore, it is shown that the indicators of non-meritocratic perception refer to three other sub-dimensions and cannot be completely attributed to a an entirely common factor as has been employed in previous research.
+
+# Appendix
 
@@
 pacman::p_load(dplyr,knitr,kableExtra)
+options(knitr.kable.NA = '')
+table_format = if(knitr::is_html_output()) {
+  "html"
+} else if(knitr::is_latex_output()) {
+  "latex"
+}
+
+table_format2 = if(knitr::is_html_output()) {
+  T
+} else if(knitr::is_latex_output()) {
+  F
+}
+
 cfa_perpref1 <- '
@@
 labs <- c(
-  "How important is having ambition?",
-  "How important is hard work?",
-  "How important is coming from a wealthy family?",
-  "How important is having welleducated parents",
-  "How important is knowing the right people",
-  "How important is having political connections",
-  "How important is a person's race?",
-  "How important is being born a man or a woman?",
-  "How well he or she does the job?",
-  "How hard he or she works at the job?",
-  "What is needed to support a family?",
-  "Whether the person has children to support?"
+  "Importance: hard work",
+  "Importance: having ambition",
+  "Importance: wealthy family",
+  "Importance: educated parents",
+  "Importance: race",
+  "Importance: gender",
+  "Importance: knowing people",
+  "Importance: political connections",
+  "Reasons for pay: well job",
+  "Reasons for pay: hard job",
+  "Reasons for pay: support family",
+  "Reasons for pay: has children"
 )
@@
                                                         what="std")$lambda),
-                           digits = 2))
+                           digits = 3))
 tb.load[tb.load==c(0.00)] <- NA
@@
 for (i in names(tb.load)) {
-  # tb.load[,i] <- sjlabelled::as_character(tb.load[,i])
-  tb.load[,i] <- sprintf(tb.load[,i], fmt = '%#.2f')
+  tb.load[,i] <- sprintf(tb.load[,i], fmt = '%#.3f')
 }
@@
 tb.caption <- c(" Factor loadings and fit measures for ISSP 2009")
-tb.issp<- kable(tb.perpref1,escape = FALSE,
+tb.issp<- kable(tb.perpref1,escape = FALSE,digits = 3,
                 align = "lcccc",

### 2021-06-16 paper_study3.Rmd
correlaciones latentes
Se simplifica el código eliminando los puntajes factoriales

@@
     toc: false
-  bookdown::html_document2:
-    number_sections: false
 linkcolor: blue
 #bibliography:
-#  - ../input/bib/EconomiaMoral.bib
-#  - ../input/bib/meritocracy.bib
-#  - ../input/bib/invariance.bib
 editor_options:
@@
 ```{r val-conv}
-pacman::p_load(kableExtra, polycor)
+pacman::p_load(kableExtra, lavaan, dplyr,tidyr)
 
@@
 
-pl1 <- polycor::polyserial(data_s3_conv$perc_merit,data_s3_conv$get_ah_1)
-pl2 <- polycor::polyserial(data_s3_conv$pref_merit,data_s3_conv$get_ah_1)
-pl3 <- polycor::polyserial(data_s3_conv$perc_nmerit,data_s3_conv$get_ah_1)
-pl4 <- polycor::polyserial(data_s3_conv$pref_nmerit,data_s3_conv$get_ah_1)
-valores1 <- round(c(pl1,pl2,pl3,pl4),4)
+#
+model <- '
+perc_merit =~  meritv03_perc_effort + meritv03_perc_talent
+perc_nmerit=~ meritv03_perc_wpart + meritv03_perc_netw
+pref_merit =~ meritv03_pref_effort + meritv03_pref_talent
+pref_nmerit=~ meritv03_pref_wpart + meritv03_pref_netw
 
-pl1b <- polycor::polyserial(data_s3_conv$perc_merit,data_s3_conv$get_ah_5)
-pl2b <- polycor::polyserial(data_s3_conv$pref_merit,data_s3_conv$get_ah_5)
-pl3b <- polycor::polyserial(data_s3_conv$perc_nmerit,data_s3_conv$get_ah_5)
-pl4b <- polycor::polyserial(data_s3_conv$pref_nmerit,data_s3_conv$get_ah_5)
-valores2 <- round(c(pl1b,pl2b,pl3b,pl4b),4)
+nbmerit =~  nb_merit_1 + nb_merit_2 + nb_merit_3 + nb_merit_4 + nb_merit_5 + nb_merit_6 + nb_merit_7
 
-pl1c <- cor(data_s3_conv$perc_merit,data_s3_conv$nbmerit)
-pl2c <- cor(data_s3_conv$pref_merit,data_s3_conv$nbmerit)
-pl3c <- cor(data_s3_conv$perc_nmerit,data_s3_conv$nbmerit)
-pl4c <- cor(data_s3_conv$pref_nmerit,data_s3_conv$nbmerit)
-valores3 <- round(c(pl1c,pl2c,pl3c,pl4c),4)
+perc_merit ~~  get_ah_1
+perc_nmerit ~~ get_ah_1
+pref_merit ~~ get_ah_1
+pref_nmerit ~~ get_ah_1
+nbmerit ~~ get_ah_1
 
-variables= c("Meritocratic Perception", "Meritocratic Preferences", "Unmeritocratic Perception","Unmeritocratic Preferences")
+perc_merit ~~  get_ah_5
+perc_nmerit ~~ get_ah_5
+pref_merit ~~ get_ah_5
+pref_nmerit ~~ get_ah_5
+nbmerit ~~ get_ah_5
 
-table01 <- cbind(variables, valores1, valores2, valores3)
+perc_merit ~~  nbmerit
+perc_nmerit ~~ nbmerit
+pref_merit ~~ nbmerit
+pref_nmerit ~~ nbmerit
 
-kable(table01, col.names = c("Merit-Scale","Social Origin ", "Hard Work " , "Wherewithal"), full_width = T, linesep = "", caption = c("Polyserial and Pearson correlation whit other merit scales")) %>% kable_styling(
+'
+fit3_o1=lavaan::cfa(model,data = dat04w3, estimator = "DWLS", ordered = c("meritv03_perc_effort","meritv03_perc_talent",
+"meritv03_perc_wpart","meritv03_perc_netw","meritv03_pref_effort","meritv03_pref_talent","meritv03_pref_wpart","meritv03_pref_netw" ,"nb_merit_1" , "nb_merit_2" , "nb_merit_3" , "nb_merit_4" , "nb_merit_5" , "nb_merit_6" , "nb_merit_7","get_ah_1","get_ah_5"))
+
+# extract correlation matrix from lavaan model
+correlations = standardizedsolution(fit3_o1) %>% filter(op=="~~") %>% 
+                dplyr::select(lhs,rhs,est.std) %>% 
+                dplyr::filter(lhs== "perc_merit" | 
+                             lhs== "pref_merit" | 
+                             lhs== "perc_nmerit"| 
+                             lhs== "pref_nmerit") %>%
+                pivot_wider(names_from = "rhs", values_from = "est.std") %>% 
+                dplyr::select(lhs,get_ah_1, get_ah_5, nbmerit) %>% 
+                mutate(merit_scale=
+                      case_when(lhs == "perc_merit" ~  "Meritocratic perception",
+                                lhs=="perc_nmerit" ~ "Unmeritocratic perception", 
+                                lhs =="pref_merit" ~ "Meritocratic preferences" , 
+                                lhs =="pref_nmerit" ~ "Unmeritocratic preferences"))%>% 
+                select(merit_scale, get_ah_1, get_ah_5,nbmerit)
+
+# Make table 
+kable(correlations,col.names = c("Merit-Scale","Social Origin ", "Hard Work " , "Wherewithal"), full_width = T, linesep = "", caption = c("Correlation whit other merit scales")) %>% kable_styling(
     full_width = T,
@@
 
-load(file = "input/data/proc/pov01_fact02.RData")
+load(file = "input/data/proc/data_s3_inv.RData")
 

### 2021-06-16 processing/preparation_study3.rmd
correlaciones latentes
Se simplifica el código eliminando los puntajes factoriales

@@
 ```{r}
-pacman::p_load(haven, sjPlot, dplyr, mice, lavaan, stargazer, corrplot, polycor, knitr, kableExtra, summarytools, sjlabelled)
+pacman::p_load(haven, sjPlot, dplyr, tidyr, mice, lavaan, stargazer, corrplot, polycor, knitr, kableExtra, summarytools, sjlabelled)
 ```
@@
 dat04w3 <- sjlabelled::read_spss("input/data/original/Estudio_3_ola3.sav",verbose = F)
-dim(data) 
+dim(dat04w3) 
 ```
 
-## Create Perceptions and Preferences for Meritocracy (PPM-S) factor scores
+## Estimate the measurement model with the dimensions of the merit-scale and other meritocracy scales
 
@@
 ```{r results='hide'}
-dat04w3
 
-dat04w3$meritv03_perc_effort <- as.numeric(dat04w3$meritv03_perc_effort)
-dat04w3$meritv03_perc_talent <- as.numeric(dat04w3$meritv03_perc_talent)
-dat04w3$meritv03_perc_netw <- as.numeric(dat04w3$meritv03_perc_netw)
-dat04w3$meritv03_perc_wpart <- as.numeric(dat04w3$meritv03_perc_wpart)
-dat04w3$meritv03_pref_effort <- as.numeric(dat04w3$meritv03_pref_effort)
-dat04w3$meritv03_pref_talent <- as.numeric(dat04w3$meritv03_pref_talent)
-dat04w3$meritv03_pref_wpart <- as.numeric(dat04w3$meritv03_pref_wpart)
-dat04w3$meritv03_pref_netw <- as.numeric(dat04w3$meritv03_pref_netw)
-
-model04 <- '
+model <- '
 perc_merit =~  meritv03_perc_effort + meritv03_perc_talent
@@
 pref_merit =~ meritv03_pref_effort + meritv03_pref_talent
-pref_nmerit=~ meritv03_pref_wpart + meritv03_pref_netw'
+pref_nmerit=~ meritv03_pref_wpart + meritv03_pref_netw
 
-# generate 5 multiple complete datasets
-out <- mice(dat04w3[23:30], m = 5)
+nbmerit =~  nb_merit_1 + nb_merit_2 + nb_merit_3 + nb_merit_4 + nb_merit_5 + nb_merit_6 + nb_merit_7
 
-D1 <- complete(out, 1)
-D2 <- complete(out, 2)
-D3 <- complete(out, 3)
-D4 <- complete(out, 4)
-D5 <- complete(out, 5)
+perc_merit ~~  get_ah_1
+perc_nmerit ~~ get_ah_1
+pref_merit ~~ get_ah_1
+pref_nmerit ~~ get_ah_1
+nbmerit ~~ get_ah_1
 
-# fit model for each complete dataset
-fit1 <- cfa(model04, data = D1)
-fit2 <- cfa(model04, data = D2)
-fit3 <- cfa(model04, data = D3)
-fit4 <- cfa(model04, data = D4)
-fit5 <- cfa(model04, data = D5)
+perc_merit ~~  get_ah_5
+perc_nmerit ~~ get_ah_5
+pref_merit ~~ get_ah_5
+pref_nmerit ~~ get_ah_5
+nbmerit ~~ get_ah_5
 
-# predict scores for all models
-p1 <- predict(fit1)
-p2 <- predict(fit2)
-p3 <- predict(fit3)
-p4 <- predict(fit4)
-p5 <- predict(fit5)
+perc_merit ~~  nbmerit
+perc_nmerit ~~ nbmerit
+pref_merit ~~ nbmerit
+pref_nmerit ~~ nbmerit
 
-# compute 'average' across 5 sets of scores:
-scores <- (p1 + p2 + p3 + p4 + p5) / 5
+'
 
-#Generate factor scores
-scores_dat = as.data.frame(scores)
+fit3_c1=lavaan::cfa(model,data = dat04w3)
 
-# Adaptar base para agregar factor scores (sacar todos NA)
-dat04w3 = dat04w3[rowSums(is.na(dat04w3[23:28])) != 5, ]
+fit3_o1=lavaan::cfa(model,data = dat04w3, estimator = "DWLS", ordered = c("meritv03_perc_effort","meritv03_perc_talent",
+"meritv03_perc_wpart","meritv03_perc_netw","meritv03_pref_effort","meritv03_pref_talent","meritv03_pref_wpart","meritv03_pref_netw" ,"nb_merit_1" , "nb_merit_2" , "nb_merit_3" , "nb_merit_4" , "nb_merit_5" , "nb_merit_6" , "nb_merit_7","get_ah_1","get_ah_5"))
 
-# Merge with factor scores
-dat04w3 = cbind(dat04w3, scores)
+cargas1 = left_join(x = standardizedsolution(fit3_c1) %>% filter(op=="=~") %>% dplyr::select(lhs,rhs,est.std),y = standardizedsolution(fit3_o1) %>% filter(op=="=~") %>% dplyr::select(lhs,rhs,est.std),c("lhs","rhs"))
 
-# Check
-stargazer(dat04w3[, 78:81], type = "text")
+kable(cargas1,
+      format = "html",digits = 2, booktabs=T, caption = "Cargas factoriales",escape = FALSE) %>%  
+   kable_styling(full_width = T, font_size = 8,  bootstrap_options=c("striped", "bordered")) %>%
+     column_spec(column = 1,width = "2 cm") %>% 
+   column_spec(column = 2,width = "11 cm") %>%
+  column_spec(column = 3,width = "1 cm") %>%
+  column_spec(column = 4,width = "1 cm") %>%
+      collapse_rows(columns = 1,valign = "middle")
+
+pacman::p_load(dplyr, kableExtra, knitr)
+
+# extract fit indices from models and add to table
+sum_fit1<- dplyr::bind_rows(fitmeasures(fit3_c1)[c("chisq","df","cfi","tli","rmsea","srmr")],
+                    fitmeasures(fit3_o1)[c("chisq","df","cfi","tli","rmsea","srmr")])
+
+# Customize object
+sum_fit1$mod <- c("Continuous","Ordinals")
+sum_fit1$nobs <- c(nobs(fit3_c1),nobs(fit3_o1))
+sum_fit1$est <- c("MLR","DWLS")
+sum_fit1 <- dplyr::select(sum_fit1,mod,nobs,est,everything())
+colnames <- c("Treated as","$N$","Estimator","$\\chi^2$","df","CFI","TLI","RMSEA","SRMR")
+
+# Create table
+table_cfa_fits1 <-kable(sum_fit1, format="html", digits=3, booktabs=T, col.names=colnames, escape = FALSE) %>% kable_styling(full_width = F, font_size = 10)
+table_cfa_fits1
 
@@
 
-## Create The Neoliberal Beliefs Inventory (NBI) factor scores
-
-The following code creates new variables corresponding to the four dimensions of the NBI scale.
-Based on the measurement model, four factorial scores are created.
-To retrieve cases the "mice" function is used
-```{r results='hide'}
-datnbmerit <- dat04w3 %>% select(nb_merit_1,
-nb_merit_2,
-nb_merit_3,
-nb_merit_4,
-nb_merit_5,
-nb_merit_6,
-nb_merit_7,
-nb_merit_8)
-psych::alpha(datnbmerit)
-
-dat04w3$nb_merit_1 <- as.numeric(dat04w3$nb_merit_1)
-dat04w3$nb_merit_2 <- as.numeric(dat04w3$nb_merit_2)
-dat04w3$nb_merit_3 <- as.numeric(dat04w3$nb_merit_3)
-dat04w3$nb_merit_4 <- as.numeric(dat04w3$nb_merit_4)
-dat04w3$nb_merit_5 <- as.numeric(dat04w3$nb_merit_5)
-dat04w3$nb_merit_6 <- as.numeric(dat04w3$nb_merit_6)
-dat04w3$nb_merit_7 <- as.numeric(dat04w3$nb_merit_7)
-dat04w3$nb_merit_8 <- as.numeric(dat04w3$nb_merit_8)
-
-model04 <- '
-nbmerit =~  nb_merit_1+
-nb_merit_2+
-nb_merit_3+
-nb_merit_4+
-nb_merit_5+
-nb_merit_6+
-nb_merit_7'
-
-# generate 5 multiple complete datasets
-out <- mice(dat04w3[57:64], m = 5)
-
-D1 <- complete(out, 1)
-D2 <- complete(out, 2)
-D3 <- complete(out, 3)
-D4 <- complete(out, 4)
-D5 <- complete(out, 5)
-
-# fit model for each complete dataset
-fit1 <- cfa(model04, data = D1)
-fit2 <- cfa(model04, data = D2)
-fit3 <- cfa(model04, data = D3)
-fit4 <- cfa(model04, data = D4)
-fit5 <- cfa(model04, data = D5)
-
-# predict scores for all models
-p1 <- predict(fit1)
-p2 <- predict(fit2)
-p3 <- predict(fit3)
-p4 <- predict(fit4)
-p5 <- predict(fit5)
-
-# compute 'average' across 5 sets of scores:
-scores <- (p1 + p2 + p3 + p4 + p5) / 5
-
-#Generate factor scores
-scores_dat = as.data.frame(scores)
-
-# Adaptar base para agregar factor scores (sacar todos NA)
-dat04w3 = dat04w3[rowSums(is.na(dat04w3[57:64])) != 5, ]
-
-# Merge with factor scores
-dat04w3 = cbind(dat04w3, scores)
-
-data_s3_conv <- dat04w3 %>% select(nbmerit, perc_merit,pref_merit,perc_nmerit,pref_nmerit,get_ah_1, get_ah_5) %>% na.omit()
-
-```
-
-## Labels
-
-### Variable labels
-
-Created variables are labeled and described
+## Calculate the correlations between the scale of this work and other measures of meritocratic beliefs.
 
@@
 
-set_label(data_s3_conv$nbmerit)  <- "Personal wherewithal - NBI"
-set_label(data_s3_conv$perc_merit)  <- "Meritocratic perception"
-set_label(data_s3_conv$pref_merit)   <- "Meritocratic preferences"
-set_label(data_s3_conv$perc_wpart)    <- "Unmeritocratic perception"
-set_label(data_s3_conv$pref_nmerit)  <-  "Unmeritocratic preferences"
+correlaciones = standardizedsolution(fit3_o1) %>% filter(op=="~~") %>% 
+                dplyr::select(lhs,rhs,est.std) %>% 
+                dplyr::filter(lhs== "perc_merit" | 
+                             lhs== "pref_merit" | 
+                             lhs== "perc_nmerit"| 
+                             lhs== "pref_nmerit") %>%
+                pivot_wider(names_from = "rhs", values_from = "est.std") %>% 
+                dplyr::select(lhs,get_ah_1, get_ah_5, nbmerit) %>% 
+                mutate(merit_scale=
+                      case_when(lhs == "perc_merit" ~  "Meritocratic perception",
+                                lhs=="perc_nmerit" ~ "Unmeritocratic perception", 
+                                lhs =="pref_merit" ~ "Meritocratic preferences" , 
+                                lhs =="pref_nmerit" ~ "Unmeritocratic preferences"))%>% 
+                select(merit_scale, get_ah_1, get_ah_5,nbmerit)
 
-set_label(data_s3_conv$get_ah_1)  <- "Importance of a rich family to get ahead"
-set_label(data_s3_conv$get_ah_5)   <- "Importance of hard work to get ahead"
-
-sjmisc::descr(data_s3_conv)
-
+kable(correlaciones,col.names = c("Merit-Scale","Social Origin ", "Hard Work " , "Personal wherewithal - NBI"),
+      format = "html",digits = 2, booktabs=T,escape = FALSE) %>%  
+      kable_styling(full_width = T, font_size = 8,  bootstrap_options=c("striped", "bordered")) %>%
+      column_spec(column = 1,width = "2 cm") %>% 
+      column_spec(column = 2,width = "11 cm") %>%
+      column_spec(column = 3,width = "1 cm") %>%
+      collapse_rows(columns = 1,valign = "middle")
 ```
@@
 ```{r}
-sjmisc::descr(data_s3_conv,
-      show = c("label","range", "mean", "sd", "NA.prc", "n"))%>%
-      kable(.,"markdown")
+#sjmisc::descr(dat04w3,
+#      show = c("label","range", "mean", "sd", "NA.prc", "n"))%>%
+#      kable(.,"markdown")
 
-print(dfSummary(data_s3_conv, valid.col = FALSE, graph.magnif = 0.75),
+print(dfSummary(dat04w3, valid.col = FALSE, graph.magnif = 0.75),
       max.tbl.height = 300, method = "render")
@@
 
-save(data_s3_conv,file = "input/data/proc/data_s3_conv.RData")
+save(dat04w3,file = "input/data/proc/data_s3_conv.RData")
 ```

### 2021-07-01 paper_study1.Rmd

@@
 
-Table \@ref(tab:sum-fit-issp) shows the four-factor model fit permitting comparison of first and second-order factors.
-The model restricted to four first-order factor dimensions shows a regular fit (CFI=0.959, TLI=0.944, RMSEA=0.098, χ2(df=48)= 21308.535; more detail se Table \@ref(tab:tb-perpref1)) in contrast to the model considering a second-order factor for non-meritocratic perception that has a slightly better fit based on the standard fit measures (CFI=0.959, TLI=0.944, RMSEA=0.098, χ2(df=48)= 21308.535).
+Table \@ref(tab:sum-fit-issp) shows the four-factor model fit allowing comparison of first and second-order factors.
+The model restricted to four first-order factor dimensions shows a regular fit (CFI=0.959, TLI=0.944, RMSEA=0.098, χ2(df=48)= 21308.535; more detail in Table \@ref(tab:tb-perpref1)) in contrast to the model considering a second-order factor for non-meritocratic perception that has a slightly better fit based on the standard fit measures (CFI=0.959, TLI=0.944, RMSEA=0.098, χ2(df=48)= 21308.535).
 
@@
 Figure \@ref(fig:meas02-issp) shows the overall results for this model, evidencing an improvement in factor loadings as well as in the overall fit of the measurement model.
-On the one hand, we see that perceptions are correlated with preferences, but this correlation is stronger for the meritocratic dimension of perception and preference ($r$=0.46) than for the non-meritocratic ($r$=0.27), which is much stronger when looking at the correlations between meritocratic and non-meritocratic, both in perception ($r$=0.04) and preference ($r$=0.13).
+On one hand, we see that perceptions are correlated with preferences, but this correlation is stronger for the meritocratic dimension of perception and preference ($r$=0.46) than for the non-meritocratic ($r$=0.27), which is much stronger when looking at the correlations between meritocratic and non-meritocratic, both in perception ($r$=0.04) and preference ($r$=0.13).
 These results show that, on one hand, the positive correlation between perception indicates that the two are not completely opposite poles of a continuum ($r$=0.30), which is reflected in the correlation between preferences ($r$=0.25).

### 2021-07-02 paper_study1.Rmd
study1-01
- tabla ajustada de indicadores
- plot ajustado
- se incluye una correción en la descripción del plot likert

@@
 
-kable(tableissp, table_format, booktabs = T, linesep = "",col.names = cnames, caption = cap) %>%
-  kable_styling(
+knitr::kable(tableissp, table_format, booktabs = T, linesep = "",col.names = cnames, caption = cap) %>%
+  kableExtra::kable_styling(
     full_width = F,
@@
     position = "center",
-    font_size = 12,
+    font_size = 20,
     bootstrap_options=c("striped", "bordered")) %>%
-  column_spec(column = 1, width = "1.5cm", ) %>%
-  column_spec(column = 2,width = "2 cm") %>%
-  collapse_rows(columns = 1:2,valign = "middle")
+  kableExtra::column_spec(column = 1, width = "2cm", ) %>%
+  kableExtra::column_spec(column = 2,width = "5cm") %>%
+  kableExtra::column_spec(column = 3,width = "12cm") %>%
+  kableExtra::collapse_rows(columns = 1:2,valign = "middle")
 ```
@@
                  caption = "Descriptive statistics for the ISSP 2009 sample") %>%
-  kableExtra::kable_styling(full_width = table_format2,
-                            font_size = 15,
+  kableExtra::kable_styling(full_width = T,
+                            font_size = 20,
                             position = "center",
@@
 Figure \@ref(fig:likert-issp) shows the distribution of responses across the indicators for each perception and preference in the meritocratic and non-meritocratic dimensions.
-On one side, we see that there is a certain degree of importance attributed to factors such as hard work and ambition in the process of getting ahead, concentrating 94.8% and 92.8% in the fairly important to essential categories.
+On one side, we see that there is a certain degree of importance attributed to factors such as hard work and ambition in the process of getting ahead, concentrating 94.9% and 92.8% in the fairly important to essential categories.
 
 By contrast, for the non-meritocratic aspects, parental background showed a more dispersed opinion, where 61.2% considers that coming from a wealthy family is important, and 76% considers that having educated parents is important.
-On the other hand, the background dimension shows that people' opinions are generally in the opposite sense, in which 64.2% consider that a person's racial origin is not very important or not important at all in the process of getting ahead in life, while 35% consider it important, and this is nearly identical regarding the importance of gender.
-Lastly, we can observe differences in the networks dimension, in which 85.5% consider it to be important to know the right people in the process of getting ahead in life, contrasting with 50% who consider political connections to be important.
+On the other hand, the background dimension shows that people' opinions are generally in the opposite sense, in which 63.7% consider that a person's racial origin is not very important or not important at all in the process of getting ahead in life, while 35% consider it important, and this is nearly identical regarding the importance of gender.
+Lastly, we can observe differences in the networks dimension, in which 85.7% consider it to be important to know the right people in the process of getting ahead in life, contrasting with 50% who consider political connections to be important.
 
-Regarding normative preferences, we observe how well a person performs his or her job and amount of effort involved concentrates 98% and 96.8% of the responses in the categories that attribute importance to these characteristics when evaluating the pay that people deserve.
-Finally, in relation to structural constrains, 83.3% attribute importance to the fact that supporting a family should be a relevant factor with respect to the pay obtained, while having children to support translates into 76.5% of importance in this dimension.
+Regarding normative preferences, we observe how well a person performs his or her job and amount of effort involved concentrates 98% and 96.9% of the responses in the categories that attribute importance to these characteristics when evaluating the pay that people deserve.
+Finally, in relation to structural constrains, 83% attribute importance to the fact that supporting a family should be a relevant factor with respect to the pay obtained, while having children to support translates into 76.4% of importance in this dimension.
 
@@
                 race,gender,people,polcone,
-                welljob,hardjob,family,child)
+                welljob,hardjob,family,child) %>% na.omit()
 
@@
   sjlabelled::set_labels(likert_issp,
-                         labels = c("Essential","Very important",
-                                    "Fairly important","Not very important",
-                                    "Not important at all"))
+                         labels = c("Essential","Very important","Fairly important",
+                                    "Not very important","Not important at all"))
 
-plot_likert_issp<- sjPlot::plot_likert(likert_issp,
-                        c(1, 1, 2, 2, 2, 2, 2, 2,3,3,4,4),
-                        groups.titles = c("Perception meritocratic", "Perception Non-meritocratic",
-                                          "Preference meritocratic", "Preference Non-meritocratic"),
-                        geom.colors = "PuBu",
-                        geom.size = 0.8,
+size=1.1
+ggplot2::theme_set(ggplot2::theme(panel.background = ggplot2::element_rect(fill = "gray85",
+                                                                           colour = "gray85"),
+                                  panel.border = ggplot2::element_blank(),
+                                  axis.text.y = ggplot2::element_text(size = 12,
+                                                                      hjust = 1),
+                                  title = ggplot2::element_text(size = 12,
+                                                                face = "bold"),
+                                  legend.text = ggplot2::element_text(size = 12),
+                                  plot.caption = ggplot2::element_text(size = 10,
+                                                                       face = "plain",
+                                                                       hjust = 1)))
+p1<- likert_issp %>%
+  dplyr::select(hwork_r:ambition_r) %>%
+  sjPlot::plot_likert(geom.colors = "PuBu",
+                      title = "Perception meritocratic",
+                        geom.size = c(0.6*size),
                         catcount = 5,
-                        # cat.neutral = 3,
-                        grid.range  =  c(1.2 , 1.4),
+                        expand.grid = T,
                         values  =  "sum.outside",
@@
                         reverse.scale = F,
-                        show.n=FALSE)
+                        show.n = FALSE) + 
+  ggplot2::theme(legend.position = "none")
 
+p2<-likert_issp %>%
+  dplyr::select(wealthy_r:polcone_r) %>% 
+  sjPlot::plot_likert(geom.colors = "PuBu",
+                      title = "Perception Non-meritocratic",
+                        geom.size = c(0.78*size),
+                        catcount = 5,
+                        expand.grid = T,
+                        values  =  "sum.outside",
+                        reverse.colors = T,
+                        reverse.scale = F,
+                        show.n = FALSE)  + 
+  ggplot2::theme(legend.position = "none")
+
+p3<-likert_issp %>% 
+  dplyr::select(welljob_r:hardjob_r) %>% 
+  sjPlot::plot_likert(geom.colors = "PuBu",
+                      title = "Preference meritocratic",
+                        geom.size = c(0.6*size),
+                        catcount = 5,
+                        expand.grid = T,
+                        values  =  "sum.outside",
+                        reverse.colors = T,
+                        reverse.scale = F,
+                        show.n = FALSE) + 
+  ggplot2::theme(legend.position = "none")
+
+p4<-likert_issp %>% 
+  dplyr::select(family_r:child_r) %>% 
+  sjPlot::plot_likert(geom.colors = "PuBu",
+                      title = "Preference Non-meritocratic",
+                        geom.size = c(0.6*size),
+                        catcount = 5,
+                        expand.grid = T,
+                        values  =  "sum.outside",
+                        reverse.colors = F,
+                        reverse.scale = F,
+                        show.n = FALSE) +
+  ggplot2::labs(caption = paste0("n=",dim(likert_issp)[1]))+
+  ggplot2::theme(legend.position = "bottom") +
+  ggplot2::scale_fill_brewer(breaks=c(5:1),
+                             direction = -1,
+                             labels = c("Not important at all","Not very important",
+                                        "Fairly important","Very important","Essential"))
 ```
@@
 ```{r likert-issp,fig.height=10,fig.width=12,fig.cap="Distribution of responses in the ISSP meritocracy items"}
-plot_likert_issp
+# plot_likert_issp
+cowplot::plot_grid(p1,p2,p3,p4,
+                     align = "hv", 
+                     axis = "a",
+                     nrow = 4, 
+                     ncol = 1,
+                     rel_heights = c(1,2,1,1.3))
 ```

### 2021-07-02 paper_study1.Rmd
update plot
- footnote source label

@@
                         show.n = FALSE) +
-  ggplot2::labs(caption = paste0("n=",dim(likert_issp)[1]))+
+  ggplot2::labs(caption = paste0("Source: Authors calculation based on International Social Survey Programme - Social Inequality 2009"," (n=",dim(likert_issp)[1],")"))+
   ggplot2::theme(legend.position = "bottom") +
@@
                      ncol = 1,
-                     rel_heights = c(1,2,1,1.3))
+                     rel_heights = c(1,
+                                     2.1,
+                                     1,
+                                     1.3))
 ```

### 2021-07-02 paper_study1.Rmd
ajuste bins

--- 2021-07-02 paper_study1.Rmd
@@
 
-size=1.1
 ggplot2::theme_set(ggplot2::theme(panel.background = ggplot2::element_rect(fill = "gray85",
@@
                       title = "Perception meritocratic",
-                        geom.size = c(0.6*size),
+                        geom.size = c(0.62),
                         catcount = 5,
@@
                       title = "Perception Non-meritocratic",
-                        geom.size = c(0.78*size),
+                        geom.size = c(0.9),
                         catcount = 5,
@@
                       title = "Preference meritocratic",
-                        geom.size = c(0.6*size),
+                        geom.size = c(0.62),
                         catcount = 5,
@@
                       title = "Preference Non-meritocratic",
-                        geom.size = c(0.6*size),
+                        geom.size = c(0.62),
                         catcount = 5,
@@
                         show.n = FALSE) +
-  ggplot2::labs(caption = paste0("n=",dim(likert_issp)[1]))+
-  ggplot2::theme(legend.position = "bottom") +
-  ggplot2::scale_fill_brewer(breaks=c(5:1),
+  ggplot2::theme(legend.position = "none")
+
+leg<- ggpubr::as_ggplot(
+  ggpubr::get_legend(
+    dplyr::select(likert_issp, family_r:child_r) %>%
+      sjPlot::plot_likert(geom.colors = "PuBu", catcount = 5) +
+      ggplot2::scale_fill_brewer(
+        breaks = c(5:1),
         direction = -1,
-                             labels = c("Not important at all","Not very important",
-                                        "Fairly important","Very important","Essential"))
+        labels = c("Not important at all","Not very important","Fairly important",
+                   "Very important","Essential")) +
+      ggplot2::theme(legend.position = "bottom")
+    )
+  )
 ```
@@
 ```{r likert-issp,fig.height=10,fig.width=12,fig.cap="Distribution of responses in the ISSP meritocracy items"}
-# plot_likert_issp
-cowplot::plot_grid(p1,p2,p3,p4,
+cowplot::add_sub(
+  cowplot::plot_grid(p1,p2,p3,p4,leg,
     align = "hv",
     axis = "a",
-                     nrow = 4, 
+    nrow = 5,
     ncol = 1,
-                     rel_heights = c(1,2,1,1.3))
+    rel_heights = c(1,2,1,1,0.25)
+    ),
+  label = paste0(
+    "Source: Authors calculation based on International Social Survey Programme - Social Inequality 2009",
+    " (n=",dim(likert_issp)[1],")"
+    ),
+  size = 10,
+  hjust = 0.15
+  ) %>% 
+  cowplot::ggdraw() 
 ```

### 2021-07-02 paper_study1.Rmd
tables & plot
- fig out size de plot likert
- table stripped

@@
                                   panel.border = ggplot2::element_blank(),
-                                  axis.text.y = ggplot2::element_text(size = 12,
+                                  axis.text.y = ggplot2::element_text(size = 13,
                                                                       hjust = 1),
-                                  title = ggplot2::element_text(size = 12,
+                                  title = ggplot2::element_text(size = 13,
                                                                 face = "bold"),
@@
 
-```{r likert-issp,fig.height=10,fig.width=12,fig.cap="Distribution of responses in the ISSP meritocracy items"}
+```{r likert-issp,fig.height=11,fig.width=11,fig.cap="Distribution of responses in the ISSP meritocracy items", out.height='80%'}
 cowplot::add_sub(
@@
   size = 10,
-  hjust = 0.15
+  hjust = 0.25
   ) %>% 
@@
 
-Table \@ref(tab:tb-perpref1)
-
 ```{r tb-perpref1, echo=FALSE}

### 2021-07-03 paper_study1.Rmd

@@
                         values  =  "sum.outside",
-                        reverse.colors = F,
+                        reverse.colors = T,
                         reverse.scale = F,

### 2021-07-08 paper_study1.Rmd
ajustes
- cambio a data_s1.RData
- Modificación en data prep de study1. Se detalla mejor el procedimiento.
- check render desde consola  html (OK) y pdf (OK)

@@
 ```{r eval=FALSE, include=FALSE}
-# for render in pdf run rmarkdown::render_site("docs/paper.Rmd", output_format = "all")
-# clean #in the yml
-rmarkdown::render("docs/paper.Rmd", output_format = "bookdown::pdf_document2")
-rmarkdown::render("docs/paper.Rmd", output_format = "bookdown::html_document2")
+rmarkdown::render("paper_study1.Rmd", output_format = "bookdown::pdf_document2")
+rmarkdown::render("paper_study1.Rmd", output_format = "bookdown::html_document2")
 ```
@@
 	fig.pos = "H",
+	fig.align = "center",
 	message = FALSE,
 	warning = FALSE,
-	cache = FALSE
+	cache = FALSE, 
+	out.width = if(knitr::is_html_output()) {"80%"} 
+	            else if(knitr::is_latex_output()) {"80%"}
 )
-# knitr::opts_knit$set(base.url = "../") #relative path for .html output file
-# knitr::opts_knit$set(root.dir = "../") #relative path for chunks within .rmd files
 options(scipen=999)
@@
 
+fsize = if(knitr::is_html_output()) {
+  20
+} else if(knitr::is_latex_output()) {
+  12
+}
+
 tableissp <- read.csv(file = "input/tables/table-issp.csv",header = 1,sep = ";",encoding = "UTF-8") # call file generated externally for the table
@@
     position = "center",
-    font_size = 20,
+    font_size = fsize,
     bootstrap_options=c("striped", "bordered")) %>%
@@
   kableExtra::column_spec(column = 2,width = "5cm") %>%
-  kableExtra::column_spec(column = 3,width = "12cm") %>%
+  # kableExtra::column_spec(column = 3,width = "12cm") %>%
   kableExtra::collapse_rows(columns = 1:2,valign = "middle")
@@
 ```{r desc-issp, eval=FALSE, include=FALSE, results='asis'}
-load(file = "input/data/proc/merit_per_pref.rda")
+load(file = "input/data/proc/data_s1.RData")
 pacman::p_load(dplyr)
@@
 
-merit_per_pref %>% 
+data_s1 %>% 
   dplyr::select(hwork,ambition,wealthy,pareduc,
@@
 ```{r}
-load(file = "input/data/proc/merit_per_pref.rda")
+load(file = "input/data/proc/data_s1.RData")
 pacman::p_load(dplyr)
-likert_issp<- merit_per_pref %>% 
+likert_issp<- data_s1 %>% 
   dplyr::select(hwork,ambition,wealthy,pareduc,
@@
 
-```{r likert-issp,fig.height=11,fig.width=11,fig.cap="Distribution of responses in the ISSP meritocracy items", out.height='80%'}
+```{r likert-issp,fig.height=11,fig.width=11,fig.cap="Distribution of responses in the ISSP meritocracy items"}
 cowplot::add_sub(
@@
 
-```{r corr-issp,fig.cap = "Perceptions & preferences for ISSP  meritocracy items' polychoric correlations", fig.align='center',fig.height=6,fig.width=8}
+```{r corr-issp,fig.cap = "Perceptions and preferences for ISSP meritocracy items polychoric correlations", fig.align='center',fig.height=6,fig.width=8}
+load(file = "input/data/proc/data_s1.RData")
 pacman::p_load(dplyr)
 cor_issp<- 
-  merit_per_pref %>%
+  data_s1 %>%
   dplyr::select(hwork,ambition,wealthy,pareduc,
@@
 ```{r sum-fit-issp}
-load(file = "input/data/proc/merit_per_pref.rda")
+load(file = "input/data/proc/data_s1.RData")
 pacman::p_load(dplyr, kableExtra, knitr)
+
+table_format = if(knitr::is_html_output()) {
+  "html"
+} else if(knitr::is_latex_output()) {
+  "latex"
+}
+
+table_format2 = if(knitr::is_html_output()) {
+  T
+} else if(knitr::is_latex_output()) {
+  F
+}
+
+fsize = if(knitr::is_html_output()) {
+  20
+} else if(knitr::is_latex_output()) {
+  12
+}
 
@@
 	  '
-fit_perpref1 <- lavaan::cfa(cfa_perpref1, data = merit_per_pref,ordered = T)
+fit_perpref1 <- lavaan::cfa(cfa_perpref1, data = data_s1,ordered = T)
 
@@
 	  '
-fit_perpref2 <- lavaan::cfa(cfa_perpref2, data = merit_per_pref,ordered = T)
+fit_perpref2 <- lavaan::cfa(cfa_perpref2, data = data_s1,ordered = T)
 
@@
 table_issp_fits <-
-  knitr::kable(sum_fit_issp, format="html", digits=3, booktabs=T, 
+  knitr::kable(sum_fit_issp, format=table_format, digits=3, booktabs=T, 
                col.names=colnames,  
@@
                escape = FALSE) %>%
-  kableExtra::kable_styling(full_width = F, 
-                            font_size = 15,
+  kableExtra::kable_styling(font_size = fsize,
+                            full_width = F,
                             bootstrap_options=c("striped", "bordered"))
@@
 Table \@ref(tab:sum-fit-issp) shows the four-factor model fit allowing comparison of first and second-order factors.
-The model restricted to four first-order factor dimensions shows a regular fit (CFI=0.959, TLI=0.944, RMSEA=0.098, χ2(df=48)= 21308.535; more detail in Table \@ref(tab:tb-perpref1)) in contrast to the model considering a second-order factor for non-meritocratic perception that has a slightly better fit based on the standard fit measures (CFI=0.959, TLI=0.944, RMSEA=0.098, χ2(df=48)= 21308.535).
+The model restricted to four first-order factor dimensions shows a regular fit (CFI=0.959, TLI=0.944, RMSEA=0.098, $\chi^2$(df=48)= 21308.535; more detail in Table \@ref(tab:tb-perpref1)) in contrast to the model considering a second-order factor for non-meritocratic perception that has a slightly better fit based on the standard fit measures (CFI=0.959, TLI=0.944, RMSEA=0.098, $\chi^2$(df=48)= 21308.535).
 
@@
 
-```{r meas02-issp, echo=FALSE, fig.cap = "Descriptive plot", fig.align='center',fig.cap="Confirmatory factor analysis of the ISSP indicators of Perceptions and Preferences for Meritocracy"}
+```{r meas02-issp, echo=FALSE, fig.cap = "Descriptive plot", fig.align='center',fig.cap="Confirmatory factor analysis of the ISSP indicators of Perceptions and Preferences for Meritocracy",out.width='100%'}
 knitr::include_graphics('output/images/cfa_issp_study1.PNG')
@@
 ```{r tb-perpref1, echo=FALSE}
-load(file = "input/data/proc/merit_per_pref.rda")
+load(file = "input/data/proc/data_s1.RData")
 pacman::p_load(dplyr,knitr,kableExtra)
@@
 	  '
-fit_perpref1 <- lavaan::cfa(cfa_perpref1, data = merit_per_pref,
+fit_perpref1 <- lavaan::cfa(cfa_perpref1, data = data_s1,
                             ordered = c("wealthy","pareduc","ambition","hwork",
@@
                 format = table_format) %>%
-          kable_styling(full_width = table_format2,
-                        latex_options = c("hold_position"), 
+          kable_styling(latex_options = c("hold_position"),
+                        bootstrap_options = c("striped", "bordered"),
                         position = "center",
@@
           add_indent(c(13:17)) %>%
-          footnote(general =tb.foot ,footnote_as_chunk = T);tb.issp
+          footnote(general =tb.foot ,footnote_as_chunk = T,threeparttable = T);tb.issp
 ```

### 2021-07-08 processing/preparation_study1.Rmd
ajustes
- cambio a data_s1.RData
- Modificación en data prep de study1. Se detalla mejor el procedimiento.
- check render desde consola  html (OK) y pdf (OK)

@@
 
+This is the preparation code of the data for Study 1: Measuring meritocracy with international secondary data. The prepared data is `data_s1.RData`
+
 ## Libraries
@@
 ```{r}
-issp2009 <- read_dta("../input/data/original/ZA5400_v4-0-0.dta")
+issp2009 <- read_dta("input/data/original/ZA5400_v4-0-0.dta")
 ```
+
+First, we check country identification variables and generate a specific individual-country variable.
 
@@
 
+The following variables are those who are conceptualized as meritocratic and Non-meritocratic perception.
+
 *  v6: wealthy family
@@
 * IDs: v5 (country) & id (individual id)
+
+First, we rename the variables to a more substantial name.
+In second place, we check for missing values and declare them as NA.
+Then, we reverse code all the variables, because number 1 represents the "Essential" and 5 represents "Not important at all".
+Finally, whe adjust variable labels to a more intuitive one.
 
@@
 ## Preferences (detearn)
+
+* v47: responsibility job
+* v48: year on education
+* v49: support family
+* v50: has children to support
+* v51: well job
+* v52: hard job
+* IDs: v5 (country) & id (individual id)
+
+Same as the meritocratic perception.
+First, we rename the variables to a more substantial name.
+In second place, we check for missing values and declare them as NA.
+Then, we reversecode al the variables, because number 1 represents the "Esential" and 5 represents "Not important at all".
+Finally, whe adjust variable labels to a more intuitive one.
 
@@
 
+We merge the meritocratic perceptions and preferences variables into a single dataset.
+
 ```{r}
-merit_per_pref   <- left_join(getahead, detearn, by = c("id", "v5"))
-
+data_s1   <- left_join(getahead, detearn, by = c("id", "v5"))
 ```
 
-# Summary
+# Final summary
 ```{r,results='asis', echo=FALSE}
 st_css() # esto configura el css de dfsummary
-print(dfSummary(merit_per_pref),
+print(dfSummary(data_s1),
       method = "render")
@@
 ```{r}
-save(merit_per_pref, file = "../input/data/proc/merit_per_pref.rda")
+save(data_s1, file = "input/data/proc/data_s1.RData")
 ```

### 2021-07-08 paper_study1.Rmd
instructions
- mayor detalle de instrucciones dentro de chunks

@@
 pacman::p_load(knitr, kableExtra, dplyr)
-table_format = if(knitr::is_html_output()) {
-  "html"
+table_format = if(knitr::is_html_output()) { #conditional instructions for kable
+  "html" #if html set "html" in format
 } else if(knitr::is_latex_output()) {
-  "latex"
+  "latex"#if latex set "latex" in format
 }
 
-fsize = if(knitr::is_html_output()) {
-  20
+fsize = if(knitr::is_html_output()) { #conditional instructions for kable
+  20 # if html set table fontsize to 20
 } else if(knitr::is_latex_output()) {
-  12
+  12 # if latex set table fontsize to 12
 }
@@
 tableissp <- read.csv(file = "input/tables/table-issp.csv",header = 1,sep = ";",encoding = "UTF-8") # call file generated externally for the table
-cnames <- c("Component", "Dimensions","Item")
-cap <- "Items of the ISSP meritocratic perceptions and preferences measures"
+cnames <- c("Component", "Dimensions","Item") # table column names
+cap <- "Items of the ISSP meritocratic perceptions and preferences measures" # Table caption
 
@@
 pacman::p_load(dplyr)
-table_format = if(knitr::is_html_output()) {
-  "html"
+table_format = if(knitr::is_html_output()) { #conditional instructions for kable
+  "html" #if html set "html" in format
 } else if(knitr::is_latex_output()) {
-  "latex"
+  "latex"#if latex set "latex" in format
 }
 
-table_format2 = if(knitr::is_html_output()) {
-  T
+fsize = if(knitr::is_html_output()) { #conditional instructions for kable
+  20 # if html set table fontsize to 20
 } else if(knitr::is_latex_output()) {
-  F
+  12 # if latex set table fontsize to 12
 }
@@
                 race,gender,people,polcone,
-                welljob,hardjob,family,child) %>% na.omit()
-
+                welljob,hardjob,family,child) %>% na.omit() #Subset variables
+# Set labels for each variable
 likert_issp <- sjlabelled::set_label(likert_issp,
@@
                                                "Has children")) 
-
+#Reverse coding (ONLY) for plot
 likert_issp <- sjmisc::rec(likert_issp, rec="rev", append=FALSE)
+#Set value labels for plot
 likert_issp <- 
@@
 
+# Declare ggplot2 features
 ggplot2::theme_set(ggplot2::theme(panel.background = ggplot2::element_rect(fill = "gray85",
@@
                                                                        hjust = 1)))
+#Plot 1: Perception meritocratic
 p1<- likert_issp %>%
@@
   ggplot2::theme(legend.position = "none") 
-
+#Plot 2: Perception Non-meritocratic
 p2<-likert_issp %>%
@@
   ggplot2::theme(legend.position = "none")
-
+#Plot 3: Preference meritocratic
 p3<-likert_issp %>% 
@@
   ggplot2::theme(legend.position = "none")
-
+#Plot 4: Preference Non-meritocratic
 p4<-likert_issp %>% 
@@
 
+#Legend: Extract legend for the final plot using ggpubr
 leg<- ggpubr::as_ggplot(
@@
 ```{r likert-issp,fig.height=11,fig.width=11,fig.cap="Distribution of responses in the ISSP meritocracy items"}
+#Final plot
 cowplot::add_sub(
-  cowplot::plot_grid(p1,p2,p3,p4,leg,
+  cowplot::plot_grid(p1,p2,p3,p4,leg, #plot 1,2,3,4 and legend (bottom)
     align = "hv",
     axis = "a",
-    nrow = 5,
-    ncol = 1,
-    rel_heights = c(1,2,1,1,0.25)
+    nrow = 5, #5 plots
+    ncol = 1, #1 column
+    rel_heights = c(1,2,1,1,0.25) #fig size within plot
     ),
@@
     " (n=",dim(likert_issp)[1],")"
-    ),
-  size = 10,
-  hjust = 0.25
+    ),        # Caption: source and sample size 
+  size = 10,  # Caption font size
+  hjust = 0.25# Caption align 
   ) %>% 
@@
 pacman::p_load(dplyr)
+# create polychoric correlation matrix
 cor_issp<- 
@@
   lavaan::lavCor(., ordered=names(.)) 
-diag(cor_issp) = NA
+diag(cor_issp) = NA #set diagonal values to NA
+# Set Row names of the matrix
 rownames(cor_issp) <- c("A. Hard work", 
@@
                        "L. Has children")
+#set Column names of the matrix
 colnames(cor_issp) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)", 
                        "(H)", "(I)","(J)","(K)","(L)")
+
+#Plot the matrix using corrplot
 corrplot::corrplot(cor_issp,
@@
 pacman::p_load(dplyr, kableExtra, knitr)
-
-table_format = if(knitr::is_html_output()) {
-  "html"
+table_format = if(knitr::is_html_output()) { #conditional instructions for kable
+  "html" #if html set "html" in format
 } else if(knitr::is_latex_output()) {
-  "latex"
+  "latex"#if latex set "latex" in format
 }
 
-table_format2 = if(knitr::is_html_output()) {
-  T
+fsize = if(knitr::is_html_output()) { #conditional instructions for kable
+  20 # if html set table fontsize to 20
 } else if(knitr::is_latex_output()) {
-  F
+  12 # if latex set table fontsize to 12
 }
-
-fsize = if(knitr::is_html_output()) {
-  20
-} else if(knitr::is_latex_output()) {
-  12
-}
-
+# meassurement model: First order 
 cfa_perpref1 <- ' 
@@
 	  '
+# Model fit: First order 
 fit_perpref1 <- lavaan::cfa(cfa_perpref1, data = data_s1,ordered = T)
 
+# meassurement model: Second order 
 cfa_perpref2 <- '
@@
 	  '
+# Model fit: second order 
 fit_perpref2 <- lavaan::cfa(cfa_perpref2, data = data_s1,ordered = T)
@@
 options(knitr.kable.NA = '')
-table_format = if(knitr::is_html_output()) {
-  "html"
+table_format = if(knitr::is_html_output()) { #conditional instructions for kable
+  "html" #if html set "html" in format
 } else if(knitr::is_latex_output()) {
-  "latex"
+  "latex"#if latex set "latex" in format
 }
 
-table_format2 = if(knitr::is_html_output()) {
-  T
+fsize = if(knitr::is_html_output()) { #conditional instructions for kable
+  20 # if html set table fontsize to 20
 } else if(knitr::is_latex_output()) {
-  F
+  12 # if latex set table fontsize to 12
 }
 
+# First order model for Appendix
 cfa_perpref1 <- '
@@
 	  '
+# Fit First order model for Appendix
 fit_perpref1 <- lavaan::cfa(cfa_perpref1, data = data_s1,
@@
                                         "welljob","hardjob"))
+# Labels of First order model for Appendix
 labs <- c(
@@
 )
-
+#Extract factor loadings
 tb.load<- data.frame(round(cbind(lavaan::inspect(fit_perpref1,
@@
                            digits = 3))
-tb.load[tb.load==c(0.00)] <- NA
+tb.load[tb.load==c(0.00)] <- NA #set to NA factor loadings of 0
 
 for (i in names(tb.load)) {
-  tb.load[,i] <- sprintf(tb.load[,i], fmt = '%#.3f')
+  tb.load[,i] <- sprintf(tb.load[,i], fmt = '%#.3f') #set values as character
 }
-tb.load[tb.load=="NA"] <- ""
-#-------#
+tb.load[tb.load=="NA"] <- "" #set NA values to "" (empty space)
+#---extract fit measures-------#
 fm01<- data.frame(t(data.frame(lavaan::fitmeasures(fit_perpref1, output ="matrix")[c("chisq","df","cfi","tli","rmsea"),]))); row.names(fm01) ="percnmerit"
@@
 #------chi2, df------#
-fm04<- round(rbind(fm01),3)
-fm04.1 <- fm04 %>% dplyr::select(chisq,df) 
-fm04.1$chisq <- round(x = fm04.1$chisq,digits = 1)
-fm04.1$df <- round(x = fm04.1$df,digits = 0)
-fm04.1$chi2df <- paste0(fm04.1$chisq,"(",fm04.1$df,")")
-fm04.1 <- dplyr::select(fm04.1,"chi2df")
+fm04<- round(rbind(fm01),3) #select and round to 4 digits
+fm04.1 <- fm04 %>% dplyr::select(chisq,df) #select chi2 and degrees of freedom
+fm04.1$chisq <- round(x = fm04.1$chisq,digits = 1) # round Chi2 to 1 digit
+fm04.1$df <- round(x = fm04.1$df,digits = 0) #round Df to 0 digits
+fm04.1$chi2df <- paste0(fm04.1$chisq,"(",fm04.1$df,")") #Paste chi2(df)
+fm04.1 <- dplyr::select(fm04.1,"chi2df") #select final column
 for (i in names(fm04.1)) {
-  fm04.1[,i] <- as.character(fm04.1[,i])
+  fm04.1[,i] <- as.character(fm04.1[,i]) # transform all to character
 }
@@
 #------CFI, RMSEA------#
-fm04.2 <- fm04 %>% dplyr::select(cfi,tli,rmsea) 
-for (i in names(fm04.2)) {
+fm04.2 <- fm04 %>% dplyr::select(cfi,tli,rmsea) # select CFI, TLI and RMSEA
+for (i in names(fm04.2)) {                      # transform all to character
   fm04.2[,i] <- sprintf(fm04.2[,i], fmt = '%#.3f')
@@
 
-fm.df      <- dplyr::bind_cols(fm04.1,fm04.2)
-fm.df$nobs <- c(lavaan::nobs(fit_perpref1)) 
-fm.df <- data.frame(t(fm.df)); colnames(fm.df) <- c("percnmerit")
+fm.df      <- dplyr::bind_cols(fm04.1,fm04.2) # factor loading and labels
+fm.df$nobs <- c(lavaan::nobs(fit_perpref1))   # add number of observations
+fm.df <- data.frame(t(fm.df)); colnames(fm.df) <- c("percnmerit") #transpose and set colnames
 
 #------ merge ------#
-tb.fm<- dplyr::bind_rows(tb.load,fm.df)
+tb.fm<- dplyr::bind_rows(tb.load,fm.df) #merge factor loadings and fit indices
 tb.fm<- tb.fm %>% 
@@
                        "$\\text{TLI}$","$\\text{RMSEA}$","$N$")) %>%
-  dplyr::select(vars,everything())
-tb.perpref1 <- tb.fm
+  dplyr::select(vars,everything()) #Change names of Fit meassures to latex symbols
+tb.perpref1 <- tb.fm #final matrix for Table
 

### 2021-08-11 index.Rmd
unir estudios & pendientes

@@
 csl: "input/bib/apa6.csl"
-bibliography: "input/bib/zlibrary.bib"
+bibliography: "input/bib/meritocracy.bib"
 author: " "

### 2021-08-11 paper.Rmd
unir estudios & pendientes

@@
 output:
-  bookdown::pdf_document2:
-    template: null
-    toc: false
+#  bookdown::pdf_document2:
+#    template: null
+#    toc: false
   bookdown::html_document2:
@@
 bibliography:
-#  - input/bib/EconomiaMoral.bib
-#  - input/bib/meritocracy.bib
-#  - input/bib/invariance.bib
+  - input/bib/meritocracy.bib
 editor_options:
@@
 
-Economic inequality has become an issue of growing concern around the world.
-This has been expressed in a series of protests, as well as diverse critical analyses regarding the development of capitalism and its consequences [@streeckPoliticsPublicDebt2014; @piketty_Capital_2014].
+Economic inequality and income concentration have become an issue of growing concern in the last years.
+This has been expressed in a series of social movements as well as diverse critical analyses regarding the development of capitalism and its consequences [@streeckPoliticsPublicDebt2014; @piketty_Capital_2014].
 In this context, the study of views, preferences and perceptions of inequality has acquired relevance in the social sciences, in topics such as redistributive preferences [@alesina_Fairness_2005; @dimick_Models_2018], the legitimization of economic inequality [@schroder_Income_2017] and the functioning of meritocracy [@duru-bellat_whos_2012; @mijs_paradox_2019; @reynolds_perceptions_2014; @atriaEconomicElitesAttitudes2020].
+Within this context, the ideal of meritocracy has been strongly challenged as an unfulfilled promise of modern societies that allows the perpetuation of social inequalities [@sandel_tyranny_2020; @goldthorpe_myth_2003].
 
@@
 Social psychology and sociology have studied the characteristics and consequences of beliefs in meritocracy, under the general hypothesis that a greater belief in meritocracy emphasizes the role of individual over structural factors in personal achievements, leading to greater legitimization of inequalities [@premingerMeritocracyServiceEthnocracy2020; @trumpWhenWhyEconomic2020; @hadjar_meritokratie_2008; @MadeiraPrimesConsequencesSystematic2019].
-Such research has raised criticism of meritocracy as a moral standard of distribution, arguing that it would be an unfulfilled promise given the still preponderant weight of non-meritocratic elements upon individual status and social mobility[@sandelTyrannyMeritWhat2020a; @witteveenReconsideringMeritocraticPower2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019; @khan_Privilege_2013].
+Such research has raised criticism of meritocracy as a moral standard of distribution given the preponderant weight of non-meritocratic elements upon individual status and social mobility[@sandelTyrannyMeritWhat2020a; @witteveenReconsideringMeritocraticPower2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019; @khan_Privilege_2013].
 
-Due to the role that meritocratic beliefs play in the justification of individual achievement or failure in contemporary societies [@hadjar_meritokratie_2008, @markovits_meritocracy_2019; @sandel_tyranny_2020], multiple studies have evaluated the relationship between meritocratic beliefs and personal and/or contextual characteristics.
+Due to the role that meritocratic beliefs play in the justification of individual achievement (or failure) in contemporary societies [@hadjar_meritokratie_2008, @markovits_meritocracy_2019; @sandel_tyranny_2020], multiple studies have evaluated the relationship between meritocratic beliefs and personal and/or contextual characteristics.
 For example, some studies have linked meritocracy to the reinforcement of socio-economic, gender, and ethnic stereotypes [@MadeiraPrimesConsequencesSystematic2019; @girerdNeoliberalismIdeologicalBarrier2020; @premingerMeritocracyServiceEthnocracy2020], as well as the effects of meritocratic beliefs in educational [@generettStoriesWeTell2020; @owensENGINESSOCIALMOBILITY2020] and organizational contexts [@perezAdvancingCareersMerit2020; @aielloNewEvidenceFirmuniversity2019].
-
-Most of the studies so far have used indicators from existing standard surveys (as the International Social Survey Programme - ISSP) to measure meritocracy.
-However, as we will show later, the methods of measuring meritocracy vary extensively among studies.
+Most of such studies so far have used indicators from existing standard surveys (as the International Social Survey Programme - ISSP) to measure meritocracy.
+However, as we will show later, the concepts used as well as the instruments for  measuring meritocracy vary extensively among studies.
 In many cases, similar phenomena are associated with different indicators, and conversely, often different phenomena are measured with similar indicators.
-Such trends not only limits the comparability of studies, but also the ability to understanding the effects of meritocratic beliefs across different fields.
+Such trends not only limits the comparability of studies, but also the ability to understanding the effects of meritocratic beliefs across different lines of research.
 
@@
 In this measure of meritocratic beliefs, we can observe a couple of strong conceptual assumptions: effort would not depend on parental influence, and talent (as innate ability) is not meritocratic (contrary to Michael Youngs' original conceptualization).
-This conceptual and measurement-based assumption is found in other studies, that also assume effort to be the main and only aspect of meritocracy [@girerdNeoliberalismIdeologicalBarrier2020a; @bubakPerceptionsMeritocracyNote2019].
-Is effort the only dimension behind the concept of meritocracy, or includes other components, such as talent (as in its original conceptualization).
+This conceptual and measurement-based assumption is found in other studies, that also assume effort to be the main and only aspect of meritocracy [@girerdNeoliberalismIdeologicalBarrier2020a; @bubakPerceptionsMeritocracyNote2019], raising the question: Is effort the only dimension behind the concept of meritocracy?
 The question of whether talent is or is not considered meritocratic is certainly an interesting topic to discuss from a philosophical point of view, but for those working on empirical studies it should be something that face empirical scrutiny.
 Besides, the consideration of talent as part of meritocracy certainly opens some interesting avenues of research, for instance in studies showing that for the elites meritocracy is more related to talent, whereas effort is something more characteristic of meritocracy of the middle and low classes ([citar Atria ABS]).
-Actually, in a posterior paper by @mijsHowInformationEconomic2020, somewhat contradictory with its previous definition, he considered talent as an indicator of meritocracy.
 
@@
 what is).
-This occurs not only in @mijs_paradox_2019, but also in @reynolds_perceptions_2014, in which the term beliefs is used to talk about perceptions, whereas other authors use general terms such as attitudes [@kunovich_systems_2007].
+This occurs for instance in @reynolds_perceptions_2014, in which the term beliefs is used to talk about perceptions, whereas other authors use general terms such as attitudes [@kunovich_systems_2007].
 The first attempt to shed light on this issue was made by @duru-bellat_whos_2012, who used the question "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" as a proxy for "desired" meritocracy (beliefs).
@@
 
-The proposed measurement framework is depicted in Figure \@ref(fig:merit-model):
+The proposed conceptual and measurement framework is depicted in Figure \@ref(fig:merit-model):
 
@@
 The columns "Perceptions" and "Preferences" represent the distinction between these two concepts, usually confused under the label "beliefs" [@castillo_meritocracia_2019].
-Perceptions refer to the extent to which people observe that meritocracy works or applies in their society, which relates to items such as "I think hard work is important to get ahead in society".
+Perceptions refer to the extent to which people observe that meritocracy functions or applies in their society, which relates to items such as "I think hard work is important to get ahead in society".
 Preferences refer to normative expectations that are usually linked to a "should" expression (e.g.
@@
 
+## Study 1: Measuring meritocracy with international secondary data
+
+## Data
+
+The data is the last available wave of the social inequality module from the International Social Survey Programme (ISSP), which is the most specialized international comparative survey in perceptions, attitudes and beliefs about inequality related issues.
+This wave corresponds to the year 2009 and covers attitudes towards a series of topics dealing with social inequality across 41 countries and 56021 individuals.
+Although there is data available from this module for previous waves (1987, 1992 and 1999), unfortunately there are several variables that are important for this research that were not included, particularly in 1987 and 1999, reason why we only use the 2009 wave.
+
+## Variables
+
+There are a series of indicators in the ISSP survey that in the following we attempt to classify in our meritocracy measurement scheme.
+However, it is important to mention  that the items were not originally designed as a specific measure of meritocracy.
+Still, as this database is one of the most widely used for international social research, many of the items have been used in previous research on meritocracy (and certainly will keep being used in the future).
+Therefore, before proposing a new measure of meritocracy we  give a first test to our proposed conceptualization using available data that not only covers several countries but also is validated by the research community.
+
+The variables are presented below organized in perceptions and preferences:
+
+- _Perception of meritocracy/non-meritocracy_: for operationalizing perceptions the closest set of ISSP's indicators comes from the question asking about perceptions for opportunities to get ahead, which are usually considered as "meritocratic beliefs" in previous research.
+The general heading of the battery is: _"To begin we have some questions about opportunities for getting ahead.
+Please tick one box for each of these to show how important you think it is for getting ahead in life."_ This is followed by a list of statements to be answered in a scale with the following levels numbered from 1 to 5: essential, very important, fairly important, not very important, not important at all.
+
+The classification is based on criteria of internal motivation (meritocratic) and structural constraints (non-meritocratic).
+There were two items from the battery that were excluded from the analysis as they would not fit into the classification.
+The first one was "having good education yourself", since it was not clear whether this is could be due to individual motivation or system opportunities, and the second was "giving bribes", as introduced elements of criminality that were beyond a non-meritocratic perception.
+
+- Preferences for meritocracy-non meritocracy: for the operationalization of normative preferences we used a list of items related to reasons for pay.
+The ISSP question was: _In deciding how much people ought to earn, how important should each of these things be, in your opinion_, rated in the same _essential-non important at all_ scale as the questions for meritocratic preferences.
+
+Table \@ref(tab:table-issp) below summarise the items classified according to the proposal based on perception and preference regarding to the meritocratic/non-meritocratic dimensions:
+
+```{r table-issp, echo=FALSE}
+pacman::p_load(knitr, kableExtra, dplyr)
+table_format = if(knitr::is_html_output()) {
+  "html"
+} else if(knitr::is_latex_output()) {
+  "latex"
+}
+
+tableissp <- read.csv(file = "input/tables/table-issp.csv",header = 1,sep = ";",encoding = "UTF-8") # call file generated externally for the table
+cnames <- c("Component", "Dimensions","Item")
+cap <- "Items of the ISSP meritocratic perceptions and preferences measures"
+
+knitr::kable(tableissp, table_format, booktabs = T, linesep = "",col.names = cnames, caption = cap) %>%
+  kableExtra::kable_styling(
+    full_width = F,
+    latex_options = c("hold_position"),
+    position = "center",
+    font_size = 20,
+    bootstrap_options=c("striped", "bordered")) %>%
+  kableExtra::column_spec(column = 1, width = "2cm", ) %>%
+  kableExtra::column_spec(column = 2,width = "5cm") %>%
+  kableExtra::column_spec(column = 3,width = "12cm") %>%
+  kableExtra::collapse_rows(columns = 1:2,valign = "middle")
+```
+
+# Methods
+
+The estimation was performed using Confirmatory Factor Analysis (CFA).
+CFA was conducted using the `lavaan` R package (version 0.6-3; Rosseel, 2020), with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement (Kline, 2016; Rosseel, 2020).
+As recommended by Brown (2008), we assessed model fit by jointly considering the comparative fit index and Tucker-Lewis Index (CFI and TLI; acceptable fit > 0.95), Root of the average squared residual approximation (RMSEA; acceptable fit < 0.08),  Chi-square: (p-value; acceptable fit > 0.05, and Chi-square ratio > 3).
+
+# Results
+
+## Descriptive analyses
+
+<!-- (here summary descriptive table - @julio) -->
+
+<!-- Table \@ref(tab:desc-issp) -->
+
+```{r desc-issp, eval=FALSE, include=FALSE, results='asis'}
+load(file = "input/data/proc/merit_per_pref.rda")
+pacman::p_load(dplyr)
+table_format = if(knitr::is_html_output()) {
+  "html"
+} else if(knitr::is_latex_output()) {
+  "latex"
+}
+
+table_format2 = if(knitr::is_html_output()) {
+  T
+} else if(knitr::is_latex_output()) {
+  F
+}
+
+merit_per_pref %>%
+  dplyr::select(hwork,ambition,wealthy,pareduc,
+                race,gender,people,polcone,welljob,hardjob,family,child) %>%
+  psych::describe() %>%
+  as.data.frame() %>%
+  dplyr::select("Mean"=mean,"SD"=sd,"Min"=min,"Max"=max) %>%
+  round(2) ->desc.issp
+
+row.names(desc.issp) <- c(
+  "Importance: hard work",
+  "Importance: having ambition",
+  "Importance: wealthy family",
+  "Importance: educated parents",
+  "Importance: race",
+  "Importance: gender",
+  "Importance: knowing people",
+  "Importance: political connections",
+  "Reasons for pay: well job",
+  "Reasons for pay: hard job",
+  "Reasons for pay: support family",
+  "Reasons for pay: has children"
+)
+
+sum.issp09<- knitr::kable(desc.issp,format =table_format,
+                 booktabs= T,
+                 align = "l",
+                 linesep = "",
+                 caption = "Descriptive statistics for the ISSP 2009 sample") %>%
+  kableExtra::kable_styling(full_width = T,
+                            font_size = 20,
+                            position = "center",
+                            latex_options = c("HOLD_position"));sum.issp09
+```
+
+Figure \@ref(fig:likert-issp) shows the distribution of responses across the indicators for each perception and preference in the meritocratic and non-meritocratic dimensions.
+On one side, we see that there is a certain degree of importance attributed to factors such as hard work and ambition in the process of getting ahead, concentrating 94.9% and 92.8% in the fairly important to essential categories.
+
+By contrast, for the non-meritocratic aspects, parental background showed a more dispersed opinion, where 61.2% considers that coming from a wealthy family is important, and 76% considers that having educated parents is important.
+On the other hand, the background dimension shows that people' opinions are generally in the opposite sense, in which 63.7% consider that a person's racial origin is not very important or not important at all in the process of getting ahead in life, while 35% consider it important, and this is nearly identical regarding the importance of gender.
+Lastly, we can observe differences in the networks dimension, in which 85.7% consider it to be important to know the right people in the process of getting ahead in life, contrasting with 50% who consider political connections to be important.
+
+Regarding normative preferences, we observe how well a person performs his or her job and amount of effort involved concentrates 98% and 96.9% of the responses in the categories that attribute importance to these characteristics when evaluating the pay that people deserve.
+Finally, in relation to structural constrains, 83% attribute importance to the fact that supporting a family should be a relevant factor with respect to the pay obtained, while having children to support translates into 76.4% of importance in this dimension.
+
+```{r}
+load(file = "input/data/proc/merit_per_pref.rda")
+pacman::p_load(dplyr)
+likert_issp<- merit_per_pref %>%
+  dplyr::select(hwork,ambition,wealthy,pareduc,
+                race,gender,people,polcone,
+                welljob,hardjob,family,child) %>% na.omit()
+
+likert_issp <- sjlabelled::set_label(likert_issp,
+                                     label = c("Hard work",
+                                               "Having ambition",
+                                               "Wealthy family",
+                                               "Educated parents",
+                                               "Race",
+                                               "Gender",
+                                               "Knowing people",
+                                               "Political connections",
+                                               "Well job",
+                                               "Hard job",
+                                               "Support family",
+                                               "Has children"))
+
+likert_issp <- sjmisc::rec(likert_issp, rec="rev", append=FALSE)
+likert_issp <-
+  sjlabelled::set_labels(likert_issp,
+                         labels = c("Essential","Very important","Fairly important",
+                                    "Not very important","Not important at all"))
+
+ggplot2::theme_set(ggplot2::theme(panel.background = ggplot2::element_rect(fill = "gray85",
+                                                                           colour = "gray85"),
+                                  panel.border = ggplot2::element_blank(),
+                                  axis.text.y = ggplot2::element_text(size = 13,
+                                                                      hjust = 1),
+                                  title = ggplot2::element_text(size = 13,
+                                                                face = "bold"),
+                                  legend.text = ggplot2::element_text(size = 12),
+                                  plot.caption = ggplot2::element_text(size = 10,
+                                                                       face = "plain",
+                                                                       hjust = 1)))
+p1<- likert_issp %>%
+  dplyr::select(hwork_r:ambition_r) %>%
+  sjPlot::plot_likert(geom.colors = "PuBu",
+                      title = "Perception meritocratic",
+                        geom.size = c(0.62),
+                        catcount = 5,
+                        expand.grid = T,
+                        values  =  "sum.outside",
+                        reverse.colors = T,
+                        reverse.scale = F,
+                        show.n = FALSE) +
+  ggplot2::theme(legend.position = "none")
+
+p2<-likert_issp %>%
+  dplyr::select(wealthy_r:polcone_r) %>%
+  sjPlot::plot_likert(geom.colors = "PuBu",
+                      title = "Perception Non-meritocratic",
+                        geom.size = c(0.9),
+                        catcount = 5,
+                        expand.grid = T,
+                        values  =  "sum.outside",
+                        reverse.colors = T,
+                        reverse.scale = F,
+                        show.n = FALSE)  +
+  ggplot2::theme(legend.position = "none")
+
+p3<-likert_issp %>%
+  dplyr::select(welljob_r:hardjob_r) %>%
+  sjPlot::plot_likert(geom.colors = "PuBu",
+                      title = "Preference meritocratic",
+                        geom.size = c(0.62),
+                        catcount = 5,
+                        expand.grid = T,
+                        values  =  "sum.outside",
+                        reverse.colors = T,
+                        reverse.scale = F,
+                        show.n = FALSE) +
+  ggplot2::theme(legend.position = "none")
+
+p4<-likert_issp %>%
+  dplyr::select(family_r:child_r) %>%
+  sjPlot::plot_likert(geom.colors = "PuBu",
+                      title = "Preference Non-meritocratic",
+                        geom.size = c(0.62),
+                        catcount = 5,
+                        expand.grid = T,
+                        values  =  "sum.outside",
+                        reverse.colors = T,
+                        reverse.scale = F,
+                        show.n = FALSE) +
+  ggplot2::theme(legend.position = "none")
+
+leg<- ggpubr::as_ggplot(
+  ggpubr::get_legend(
+    dplyr::select(likert_issp, family_r:child_r) %>%
+      sjPlot::plot_likert(geom.colors = "PuBu", catcount = 5) +
+      ggplot2::scale_fill_brewer(
+        breaks = c(5:1),
+        direction = -1,
+        labels = c("Not important at all","Not very important","Fairly important",
+                   "Very important","Essential")) +
+      ggplot2::theme(legend.position = "bottom")
+    )
+  )
+```
+
+```{r likert-issp,fig.height=11,fig.width=11,fig.cap="Distribution of responses in the ISSP meritocracy items", out.height='80%'}
+cowplot::add_sub(
+  cowplot::plot_grid(p1,p2,p3,p4,leg,
+    align = "hv",
+    axis = "a",
+    nrow = 5,
+    ncol = 1,
+    rel_heights = c(1,2,1,1,0.25)
+    ),
+  label = paste0(
+    "Source: Authors calculation based on International Social Survey Programme - Social Inequality 2009",
+    " (n=",dim(likert_issp)[1],")"
+    ),
+  size = 10,
+  hjust = 0.25
+  ) %>%
+  cowplot::ggdraw()
+```
+
+In terms of the association between the indicators, Figure \@ref(fig:corr-issp) shows the polychoric correlation matrix.
+Firstly, we see that the correlations between the indicators are consistent with the structure of the initially proposed structure of dimensions distinguishing between the perception and preferences of meritocratic and non-meritocratic factors.
+In general, the associations are positive between the perception and preference dimensions, such as the importance of hard work to get ahead and the importance of how hard the work is ($r$=0.34), however, they tend to be weak.
+A further aspect to consider is the relatively low correlation between dimensions concerning meritocratic and non-meritocratic aspects, (e.g.
+importance of hard work with educated parents, $r$=0.15).
+However, with respect to the non-meritocratic perception indicators, we can see that the associations demonstrate dimensionality, however, it is not entirely clear.
+In this regard, the indicators referring to parental characteristics have a positive and stronger association ($r$=0.56), but this association is much weaker in comparison to the other indicators of Non-meritocratic perception (e.g.
+Wealthy family and Gender, $r$=0.01).
+The same holds true for the Gender and Race indicators ($r$=0.6), and also applies to both the indicators of knowing the right people and having political connections ($r$=0.55).
+On this point, we will return to it in the next section.
+
+```{r corr-issp,fig.cap = "Perceptions & preferences for ISSP  meritocracy items' polychoric correlations", fig.align='center',fig.height=6,fig.width=8}
+pacman::p_load(dplyr)
+cor_issp<-
+  merit_per_pref %>%
+  dplyr::select(hwork,ambition,wealthy,pareduc,
+                race,gender,people,polcone,
+                welljob,hardjob,family,child) %>%
+  lavaan::lavCor(., ordered=names(.))
+diag(cor_issp) = NA
+rownames(cor_issp) <- c("A. Hard work",
+                       "B. Having ambition",
+                       "C. Wealthy family",
+                       "D. Educated parents",
+                       "E. Race",
+                       "F. Gender",
+                       "G. Knowing people",
+                       "H. Political connections",
+                       "I. Well job",
+                       "J. Hard job",
+                       "K. Support family",
+                       "L. Has children")
+colnames(cor_issp) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)",
+                       "(H)", "(I)","(J)","(K)","(L)")
+corrplot::corrplot(cor_issp,
+  method = "color",
+  addCoef.col = "#000390",
+  type = "upper",
+  tl.col = "black",
+  col=colorRampPalette(c("white","#0068DC"))(12),
+  bg = "white",
+  na.label = "-",
+  lower.col = "black")
+```
+
+## Confirmatory Factor Analysis
+
+This section presents the results of the confirmatory factor analysis.
+First, a four-dimensional model for meritocratic and non-meritocratic perception and preferences is estimated, considering all indicators of the non-meritocratic perception dimension as a single latent variable.
+Then, we estimate a model that follows the proposed four latent variable measurement approach, however, the non-meritocratic perception dimension is assumed as a second-order factor for the first order latent variables parental attributes, racial and gender background and networks.
+
+```{r sum-fit-issp}
+load(file = "input/data/proc/merit_per_pref.rda")
+pacman::p_load(dplyr, kableExtra, knitr)
+
+cfa_perpref1 <- '
+	  # latent variables
+	  percmerit =~ hwork + ambition
+	  percnmerit=~ wealthy + pareduc +race + gender +people + polcone
+	  prefmerit =~ welljob + hardjob
+	  prefnmerit=~ family + child
+	  '
+fit_perpref1 <- lavaan::cfa(cfa_perpref1, data = merit_per_pref,ordered = T)
+
+cfa_perpref2 <- '
+	  # latent variables
+	  percmerit =~ hwork + ambition
+	  parents =~ wealthy + pareduc
+	  backgrd =~ race + gender
+	  networks=~ people + polcone
+	  percnmerit=~ parents + backgrd + networks
+	  prefmerit =~ welljob + hardjob
+	  prefnmerit=~ family + child
+	  '
+fit_perpref2 <- lavaan::cfa(cfa_perpref2, data = merit_per_pref,ordered = T)
+
+# extract fit indices from models and add to table
+sum_fit_issp<- dplyr::bind_rows(
+  lavaan::fitmeasures(fit_perpref1)[c("chisq","df","cfi","tli","rmsea")],
+  lavaan::fitmeasures(fit_perpref2)[c("chisq","df","cfi","tli","rmsea")]
+  )
+
+# Customize object
+sum_fit_issp$mod <- c("First order","Second order")
+sum_fit_issp$nobs <- c(lavaan::nobs(fit_perpref2),lavaan::nobs(fit_perpref2))
+sum_fit_issp$est <- c("DWLS","DWLS")
+sum_fit_issp <- dplyr::select(sum_fit_issp,mod,nobs,est,dplyr::everything())
+colnames <- c("Model","$N$","Estimator","$\\chi^2$","df","CFI","TLI","RMSEA")
+
+# Create table
+table_issp_fits <-
+  knitr::kable(sum_fit_issp, format="html", digits=3, booktabs=T,
+               col.names=colnames,
+               caption = "Summary fit indices according to model",
+               escape = FALSE) %>%
+  kableExtra::kable_styling(full_width = F,
+                            font_size = 15,
+                            bootstrap_options=c("striped", "bordered"))
+table_issp_fits
+```
+
+Table \@ref(tab:sum-fit-issp) shows the four-factor model fit allowing comparison of first and second-order factors.
+The model restricted to four first-order factor dimensions shows a regular fit (CFI=0.959, TLI=0.944, RMSEA=0.098, χ2(df=48)= 21308.535; more detail in Table \@ref(tab:tb-perpref1)) in contrast to the model considering a second-order factor for non-meritocratic perception that has a slightly better fit based on the standard fit measures (CFI=0.959, TLI=0.944, RMSEA=0.098, χ2(df=48)= 21308.535).
+
+<!-- [diagrama SEM] -->
+
+```{r meas02-issp, echo=FALSE, fig.cap = "Descriptive plot", fig.align='center',fig.cap="Confirmatory factor analysis of the ISSP indicators of Perceptions and Preferences for Meritocracy"}
+knitr::include_graphics('output/images/cfa_issp_study1.PNG')
+```
+
+<!-- [tabla CFA] -->
+
+Figure \@ref(fig:meas02-issp) shows the overall results for this model, evidencing an improvement in factor loadings as well as in the overall fit of the measurement model.
+On one hand, we see that perceptions are correlated with preferences, but this correlation is stronger for the meritocratic dimension of perception and preference ($r$=0.46) than for the non-meritocratic ($r$=0.27), which is much stronger when looking at the correlations between meritocratic and non-meritocratic, both in perception ($r$=0.04) and preference ($r$=0.13).
+These results show that, on one hand, the positive correlation between perception indicates that the two are not completely opposite poles of a continuum ($r$=0.30), which is reflected in the correlation between preferences ($r$=0.25).
+Furthermore, it is shown that the indicators of non-meritocratic perception refer to three other sub-dimensions and cannot be completely attributed to a an entirely common factor as has been employed in previous research.
+
+## Study 2: The Perceptions and Preferences for Meritocracy Scale
+
+### Data {-#data}
+
+The data was obtained through an online questionnaire, which was part of a larger study on meritocracy and preferences developed in Chile in 2019 funded by the national scientific agency ANID.
+The questionnaire was programmed in Qualtrics and the fieldwork was conducted by an external online survey agency ([netquest.cl](www.netquest.cl)) between December 2019 and January 2020.
+The sample was selected from a non-probabilistic quota design in three large cities in Chile (Santiago, Concepción & Antofagasta).
+The quotas for gender, age, and education level were generated based on a survey by the Public Studies Center [@cepEncuestaCEPMayo2019], which is a well regarded counterpart agency of the ISSP (International Social Survey Programme) in Chile.
+A total sample of 2,141 people was collected, excluding those who did not answer the questions on the scale, and those who did not accept informed consent.
+There were no significant differences between our sample and the wider population for most socio-demographic characteristics, with the exception of educational level (see Table \@ref(tab:rep-samp) in appendix).
+As is often the case with online surveys, there were some limitations in achieving the quotas for lower educational levels **[citas - LUCHO]**.
+
+### Study design
+
+_Instrument_
+
+The proposed scale of perceptions and preferences relating to meritocracy consisted of eight indicators that were grouped into the four dimensions listed earlier: Perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic).
+In order to achieve at least some comparability with previous studies, the questions were adapted from the items battery "reasons to get ahead" (ISSP/GSS), which has been widely used for operationalizing meritocracy in previous studies [@mijs_paradox_2019; @duru-bellat_whos_2012; @reynolds_perceptions_2014].
+The aforementioned eight items, ordered according to dimensions, are presented in Table \@ref(tab:table-indicadores).
+These eight likert-type items have five response alternatives, ranging from "Completely disagree"(1) to "Completely agree" (5).
+
+```{r include=FALSE}
+
+# if (knitr::is_html_output()) {tb <- table_nums("tb",cap)} else {tb <- cap}
+```
+
+```{r table-indicadores, echo=FALSE}
+pacman::p_load(knitr, kableExtra, dplyr)
+
+tabitems <- read.csv(file = "input/tables/table01.csv",header = 1,sep = ",",encoding = "UTF-8") # call file generated externally for the table
+cnames <- c("Component", "Dimensions","Item (english)" , "Item original (spanish)")
+cap <- "Items of the perceptions and preferences for meritocracy scale."
+
+kable(tabitems, table_format, booktabs = T, linesep = "",col.names = cnames, caption = cap) %>%
+  kable_styling(
+    full_width = T,
+    latex_options = c("hold_position"),
+    position = "center",
+    font_size = 12,
+    bootstrap_options=c("striped", "bordered")) %>%
+  column_spec(column = 1, width = "1.5cm", ) %>%
+  column_spec(column = 2,width = "2 cm") %>%
+  collapse_rows(columns = 1:2,valign = "middle")
+```
+
+\pagebreak
+
+_Administration sets_
+
+With the objective of evaluating the effect of indicator ordering in the responses, three different versions of items' order were designed and randomly assigned to the respondents, as depicted in Figure \@ref(fig:appmod).
+The scale was presented to the first group (_n = 712_) in the order that appears in Table \@ref(tab:table-indicadores) according to perceptions and preferences.
+For the second group (_n = 717_), the order was reorganized according to perceptions and preferences over the same topic, e.g.
+for the topic of hard work, the item about perception was followed by the item about preference, and the same for the rest of the topics.
+Finally, for the third group (_n = 712_), the items were presented as completely randomized.
+
+```{r appmod, echo=FALSE, fig.cap = "Survey flow", fig.align='center'}
+knitr::include_graphics('output/images/app_mod.png')
+```
+
+_Methods_
+
+To test the scale's underlying constructs, we employed confirmatory factor analysis models (CFA).
+The models estimated one factor for each of the four proposed dimensions presented in Table \@ref(tab:table-indicadores).
+As in Study 1, CFA was conducted using the `lavaan` R package (version 0.6-3; Rosseel, 2020), with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement (Kline, 2016; Rosseel, 2020).
+The fit indexes and cut-off criteria were the same as the ones used in Study 1.
+
+A pre-registration was made in the OSF platform, available at the following link: [https://osf.io/z45y2](https://osf.io/z45y2).
+Included in this pre-registration are the hypotheses regarding the four-dimensional conceptual model underlying the scale, the variable measurement levels, the statistical tests to be performed with their respective evaluation parameters, and other important aspects of the research design.
+
+### Results
+
+_Descriptive analyses_
+
+The graphs presented in Figure \@ref(fig:plotlikert) display disaggregated and comparable information of the different response categories for each item.
+As it can be observed, in general there is more agreement in the perception of non-meritocratic items than in meritocratic ones, while in the case of preferences the opposite occurs.
+As far as preferences are concerned, the preponderant role of effort over talent as a criterion of meritocratic preference is noteworthy.
+
+```{r produce-likert_s2}
+pacman::p_load(dplyr)
+load(file = "input/data/proc/data_s2.RData")
+names(data_s2)
+dat_likert_s2 <- data_s2 %>% select(!group) # exclude group variable
+dat_likert_s2 <- sjmisc::rec(dat_likert_s2, rec="rev", append=FALSE)
+names(dat_likert_s2) # reverse coding for likert graph
+
+table(data_s2$perc_effort)
+table(dat_likert_s2$perc_effort_r) # check ok.
+
+ # plotlikert_s2<- sjPlot::plot_likert(dat_likert_s2,
+ #                        c(1, 1, 1, 1, 2, 2, 2, 2),
+ #                        groups.titles = c("Perceptions", "Preferences"),
+ #                        geom.colors = "PuBu",
+ #                        geom.size = 0.8,
+ #                        axis.labels = c("Effort", "Talent", "Rich Parents", "Contacts"),
+ #                        catcount = 4,
+ #                        cat.neutral = 3,
+ #                        grid.range  =  c (1.2 , 1.4),
+ #                        values  =  "sum.outside",
+ #                        reverse.colors = T,
+ #                        reverse.scale = F,
+ #                        show.n=FALSE)
+```
+
+```{r plotlikert, echo=FALSE, fig.cap = "Distribution of responses in the Merit Scale items", fig.align='center'}
+# plotlikert_s2
+```
+
+Attending now to the association among the scale items, Figure \@ref(fig:corpoly) shows items polychoric correlations.
+There are three main aspects to highlight from this correlation matrix.
+Firstly, as expected the largest correlations are between indicators that correspond to the same factors behind the conceptual model (e.g., perception of meritocracy by effort and by talent, r=0.52).
+Secondly, among this correlations the highest are those between the non- meritocratic dimension buth in perceptions (r=0.73) and preferences (r=0.61).
+thirdly, both items for meritocratic preferences (E and F) are the ones that mostly correlate with the rest of the perceptual items, showing medium to high correlations.
+This is noteworthy because it indicates that even the perception of non-meritocracy would be related to larger meritocratic preferences.
+Finally, we observe that there are no considerable negative correlations between meritocratic and non-meritocratic aspects, undermining the assumptions of previous studies that suggested that these dimensions would be the opposite poles of one same continuum [@reynolds_perceptions_2014].
+
+```{r corpoly, echo=FALSE, fig.cap = "Perceptions & preferences for meritcoracy items' polychoric correlations", fig.align='center'}
+load(file = "input/data/proc/data_s2.RData")
+
+# generate polychoric matrix
+cor_s2<- data_s2 %>% select(perc_effort,perc_talent,perc_wpart,perc_netw,pref_effort,pref_talent,pref_wpart,pref_netw) %>% lavaan::lavCor(., ordered=names(.))
+
+diag(cor_s2) = NA
+
+# windowsFonts(A = windowsFont("Times New Roman"))
+
+rownames(cor_s2) <-c(
+    "A. Perception Effort",
+    "B. Perception Talent",
+    "C. Perception Rich parents",
+    "D. Perception Contacts",
+    "E. Preferences Effort",
+    "F. Preferences Talent",
+    "G. Preferences Rich parents",
+    "H. Preferences Contacts")
+colnames(cor_s2) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)", "(H)")
+
+corrplot::corrplot(cor_s2,
+  method = "color",
+  addCoef.col = "#000390",
+  type = "upper",
+  tl.col = "black",
+  col=colorRampPalette(c("white","#0068DC"))(8),
+  bg = "white",
+  na.label = "-",
+  lower.col = "black")
+
+```
+
+_Confirmatory Factor Analysis_
+
+The present sections shows the results of the confirmatory factor analysis estimation.
+The model estimates four latent factors: perception meritocratic, perception non-meritocratic, preferences meritocratic and preferences non-meritocratic.
+Each factor is estimated based on two items of the scale as detailed in Table \@ref(tab:table-indicadores).
+
+The first step in the analysis consists of comparing the model fit indicators for the three versions of the scale that were randomly assigned to the participants: order according to perceptions/preferences, order according to topics and random order (see Figure \@ref(fig:appmod)).
+
+```{r model-generation-cfa}
+# Libraries & data
+pacman::p_load(lavaan)
+load(file = "input/data/proc/data_s2.RData")
+names(data_s2)
+
+# model
+model_cfa <- '
+perc_merit =~ perc_effort+perc_talent
+perc_nmerit=~perc_wpart +perc_netw
+pref_merit =~ pref_effort+pref_talent
+pref_nmerit=~pref_wpart +pref_netw
+'
+
+# estimation for each order set
+fit_cfa1 <- cfa(model_cfa, data = data_s2[data_s2$group == 1,], ordered = TRUE, std.lv=FALSE)
+fit_cfa2 <- cfa(model_cfa, data = data_s2[data_s2$group == 2,], ordered = TRUE, std.lv=FALSE)
+fit_cfa3 <- cfa(model_cfa, data = data_s2[data_s2$group == 3,], ordered = TRUE, std.lv=FALSE)
+
+```
+
+```{r table_cfa_fits, echo=FALSE}
+pacman::p_load(dplyr, kableExtra, knitr)
+
+# extract fit indices from models and add to table
+sum_fit<- dplyr::bind_rows(fitmeasures(fit_cfa1)[c("chisq","df","cfi","tli","rmsea")],
+                    fitmeasures(fit_cfa2)[c("chisq","df","cfi","tli","rmsea")],
+                    fitmeasures(fit_cfa3)[c("chisq","df","cfi","tli","rmsea")])
+
+# Customize object
+sum_fit$mod <- c("Version 1","Version 2","Version 3")
+sum_fit$nobs <- c(nobs(fit_cfa1),nobs(fit_cfa2),nobs(fit_cfa3))
+sum_fit$est <- c("DWLS","DWLS","DWLS")
+sum_fit <- dplyr::select(sum_fit,mod,nobs,est,everything())
+colnames <- c("Model","$N$","Estimator","$\\chi^2$","df","CFI","TLI","RMSEA")
+
+# Create table
+table_cfa_fits <-kable(sum_fit, format="html", digits=3, booktabs=T, col.names=colnames,  caption = "Summary fit indices according to order versions", escape = FALSE) %>% kable_styling(full_width = F, font_size = 10)
+table_cfa_fits
+```
+
+Table \@ref(tab:table_cfa_fits) shows the fit indicators of the models estimated for each of the three versions of the items' order described in the methodology.
+Regardless of the version, all models obtained adequate fit indicators, with CFI's above 0.95 and RMSEA's below 0.08.
+However, none of the models achieved a non-significant chi-square, something expected in large samples as the one used here.
+The first version order (perceptions-preferences) was the one obtaining best fit (CFI=0.993, TLI=0.995, RMSEA=0.034, $\chi2$(df=14)=42.276), whereas version 2 with the fixed order according to merit/non-merit items shows the comparatively worst indicators.
+The CFA fit indices for the completely randomized items' order (Model 3) it keeps all the indicators within the acceptable cut-off criteria and besides it controls for possible order effects in the administration of the instrument.
+The model and parameter estimates for this version are depicted in Figure \@ref(fig:meas02):
+
+```{r meas02, echo=FALSE, fig.cap = "Confirmatory factor analysis of the Perceptions and Preferences for Meritocracy Scale", fig.align='center'}
+
+# estimation for diagram: summary(fit_cfa3, standardized = TRUE, fit.measures=TRUE)
+knitr::include_graphics('output/images/cfa_randomized_study2.png')
+```
+
+Attending to the correlations between the latent variables as depicted in Figure \@ref(fig:meas02), it is observed that meritocratic preferences are correlated with preferences, both meritocratic ($r=0.457$) and non-meritocratic ($r=0.500$).
+The correlation between both types of perceptions ($r=-0.044$) and both types preferences ($r=0.185$) are low, as well as between non-meritocratic preferences and perceptions ($r=-0.059$).
+This last finding gives further evidence regarding the lack of unidimensionality of meritocratic and non-meritocratic aspects: they do not appear as the opposite poles of the same continuum and based on this evidence the use of reverse-coding for non-meritocratic items as indicators of meritocracy should be avoided.
+
+## Study 3: Additional validity analyses
+
+We performed two further analyses in order to add evidence regarding the quality of the scale.
+Firstly we analyzed the convergent validity by exploring the association of the merit scale with related concepts and its correspondent measurements.
+Secondly, we test the internal consistency of the measurement model behind the scale comparing with a different sample through measurement invariance procedures .
+
+### Convergent validity
+
+_Data_
+
+For this analysis we examined data from one wave of the same study described above, which was a three-wave panel survey.
+The last wave of the study included additional measures that allow testing the correlations of the merit scale with some related constructs as opportunity beliefs and personal wherewithal.
+After listwise deletion of missing cases in socio-demographics information, a total of 1422 individuals took part in the third wave: 668 (46.97%) women and 754 (53.03%) men.
+The ages of 50.57% of them are 45 years or older, and 36.28% held a tertiary degree.
+
+_Intruments_
+
+The following scales were included for testing their correlations with the the 8-item meritocracy scale:
+
+- Opportunity beliefs: We used two items of the social inequality module of the International Social Survey Programme.
+Studies use generally both indicators to measure meritocratic and non-meritocratic beliefs [@McCall2017Exposurerisinginequality; @mijs_paradox_2019].
+One item measures the importance of individual factors in determining life outcomes by asking for the importance of hard work for getting ahead in life (M=3.76, SD=0.93).
+The second item asks for the importance of coming from a wealthy family and captures the importance of structural factors in getting ahead (M=2.75, SD=1.27).
+Participants responded to each item using a 1 (not important at all) to 5 (essential) response scale.
+We call the first item \textit{hard work} and the second one \textit{social origin}.
+
+- Personal wherewithal: the questionnaire included The Neoliberal Beliefs Inventory (NBI), which considers four factors: Government Interference preferences, Competition preferences, System Inequality perception and Personal Wherewithal [@bay-chengTrackingHomoOeconomicus2015].
+We used the factor _personal wherewithal_ that reflects meritocracy beliefs in terms of the importance of personal attributes as strength and skills to yield success.
+The construct consists of 8 items (e.g., “Any goal can be achieved with enough hard work and talent”, “I’ve benefited from working hard, so there’s no reason others can’t”, “Anyone who is willing to work hard can be succesful in Chile”).
+Participants answered the items using a 1 (totally disagree) to 6 (totally agree) scale.
+We simple computed mean scores, with higher scores indicating stronger support for meritocracy (M=3.32, SD=0.88, $\alpha$ = 0.89).
+
+Both instruments cover mainly the perceptual side of the concept.
+Therefore, in general we expect larger correlation with meritocratic perceptions than with preferences.
+In the case of opportunity beliefs, the hard work item should correlate positively with meritocratic perception whereas the importance for coming from a wealthy family is expected to correlate positively with non-meritocratic perceptions.
+Regarding the second instrument of personal wherewithal, we also anticipate a positive correlation with meritocratic perception and low to null correlations with the other dimensions of the merit scale.
+
+_Results_
+
+We used polyserial correlations to evaluate relationships of meritocractic and non-meritocratic preferences and perceptions with the items of the opportunity beliefs battery.
+For personal wherewithal, we examined Pearson correlations.
+
+Table \@ref(tab:val-conv) shows the polyserial and pearson correlations between our meritocracy scale and common measures of meritocracy.
+The findings provide evidence for convergent validity for the scale.
+As we expected, preferences show very weak correlations.
+Findings also indicate that correlations of perceptions with opportunity beliefs are consistent with our expectations.
+The item for social origin shows a positive and moderate association with unmeritocratic perception (r=.354), and a negative and low correlation with meritocratic perception (-0.197).
+Regarding the item of hard work, the correlation with meritocratic perception is positive but weak (r=.187).
+One explanation for this result is that the support for hard work is very high among respondents---only 8.6 percent indicates _not very important_ or _not important at all_.
+
+```{r val-conv}
+pacman::p_load(kableExtra, polycor)
+
+load(file="input/data/proc/data_s3_conv.RData")
+
+pl1 <- polycor::polyserial(data_s3_conv$perc_merit,data_s3_conv$get_ah_1)
+pl2 <- polycor::polyserial(data_s3_conv$pref_merit,data_s3_conv$get_ah_1)
+pl3 <- polycor::polyserial(data_s3_conv$perc_nmerit,data_s3_conv$get_ah_1)
+pl4 <- polycor::polyserial(data_s3_conv$pref_nmerit,data_s3_conv$get_ah_1)
+valores1 <- round(c(pl1,pl2,pl3,pl4),4)
+
+pl1b <- polycor::polyserial(data_s3_conv$perc_merit,data_s3_conv$get_ah_5)
+pl2b <- polycor::polyserial(data_s3_conv$pref_merit,data_s3_conv$get_ah_5)
+pl3b <- polycor::polyserial(data_s3_conv$perc_nmerit,data_s3_conv$get_ah_5)
+pl4b <- polycor::polyserial(data_s3_conv$pref_nmerit,data_s3_conv$get_ah_5)
+valores2 <- round(c(pl1b,pl2b,pl3b,pl4b),4)
+
+pl1c <- cor(data_s3_conv$perc_merit,data_s3_conv$nbmerit)
+pl2c <- cor(data_s3_conv$pref_merit,data_s3_conv$nbmerit)
+pl3c <- cor(data_s3_conv$perc_nmerit,data_s3_conv$nbmerit)
+pl4c <- cor(data_s3_conv$pref_nmerit,data_s3_conv$nbmerit)
+valores3 <- round(c(pl1c,pl2c,pl3c,pl4c),4)
+
+variables= c("Meritocratic Perception", "Meritocratic Preferences", "Unmeritocratic Perception","Unmeritocratic Preferences")
+
+table01 <- cbind(variables, valores1, valores2, valores3)
+
+kable(table01, col.names = c("Merit-Scale","Social Origin ", "Hard Work " , "Wherewithal"), full_width = T, linesep = "", caption = c("Polyserial and Pearson correlation whit other merit scales")) %>% kable_styling(
+    full_width = T,
+    position = "center",
+    font_size = 14,
+    bootstrap_options=c("striped", "bordered")) %>%
+  column_spec(column = 1, width = "1.5cm", )
+
+```
+
+The measure of NBI for _personal wherewithal_ also provides evidence for the convergent validity of the merit scale.
+Results indicate a positive correlation of this measure with meritocratic preference (r=.392), suggesting that both variables capture similar concepts.
+The NBI's factor depicts a positive association with unmeritocratic preference (r=.208).
+This finding is consistent with literature, inasmuch as the belief for meritocray is associated with the justification of inequality [@McCall2017Exposurerisinginequality; @MadeiraPrimesConsequencesSystematic2019].
+Furthermore, results of our confirmatory factor analysis indicate that meritocratic perceptions and unmeritocratic preferences show a positive and weak correlation.
+
+### Measurement invariance
+
+The modeling for invariance testing has been gaining more relevance in sociological survey studies, mainly due to the interest in the validity of measurement scales in comparative studies in various social and cultural contexts [@Davidovetal2014].
+In this section we will use an analysis of invariance between groups, where the proposed factorial structure is expected to be independent of belonging to the sample corresponding to the original study, comparing it with another more recent study available and where the meritocracy scale was also incorporated.
+
+The invariance measurement procedure consists of a series of nested models to which restrictions are progressively incorporated on the parameters of the measurement model.
+The literature generally suggests that this modeling should be done through four levels or types of progressive restriction [@Milfont2010; @millsap_Statistical_2011; @vandeschoot_checklist_2012]:
+
+1. **Configural:** the model is estimated only indicating the factorial structure used in the CFA.
+
+2. **Weak:** an equality restriction is applied to the factor loadings in the different groups, that is, the loadings are forced to be identical in both measurements.
+
+3. **Strong:** equality restrictions are added to the intercepts of each indicator.
+
+4. **Strict:** equality restrictions are added to the error variances of each indicator.
+
+*Data*
+
+For testing the invariance we compared the data described previously with a new data source coming from an online survey carried out during the first half of 2020.
+The characteristics of this sample in terms of application and coverage are equivalent to those of the previous study.
+The final sample obtained contains 1,242 cases, where 605 (48.71%) are women and 637 (51.28%) are men.
+48.84% of them are 45 years of age or older, and 33.97% have tertiary education or higher.
+
+_Instruments_
+
+The items of the meritocracy scale were identical to the original.
+The items were administrated in a randomized order (which corresponds to the application modality of Group 3 present in Figure \@ref(fig: appmod)).
+
+_Results_
+
+Tabla \@ref(tab:fcloads-inv) shows de results of the CFA estimation of the meritocracy scale in this new sample.
+As observed, the loadings and the fit indices show similar results as the ones presented for the original study above, which gives a first base with wich to start the invariance analysis.
+
+The first step for invariance testing is the estimation of the configural model, which serves as the baseline for further comparison and it is expected to adequately meet the global fit criteria of a measurement model.
+Although the chi-square statistic is used as a global measure of fit,since its sensitivity to sample size it is recommended to use three additional fit indices: Comparative Fit Index (CFI), which should have a value greater than 0.95; Root Mean Square Error Approximation (RMSEA), which must be in the range of values ​​of 0.05 and 0.08; and the Standardized Root Mean Square Residuals (SRMR) which must be less than 0.08 [@vandeschoot_checklist_2012].
+Besides, the literature on measurement invariance suggests some complementary approaches for the evaluation of the fit, from which we will consider the incremental adjustment of the fit indexes [@cheung_Evaluating_2002; @Milfont2010; @dimitrov_Testing_2010] and the ANOVA test for means comparison in nested models [@newsom_Longitudinal_2015].
+
+# Analysis
+
+```{r invargroup}
+
+#---Load packages and data -------------------------------------#
+
+pacman::p_load(sjPlot,dplyr,lavaan,semPlot,semTools,stargazer,corrplot,psych,knitr,kableExtra,stats)
+
+load(file = "input/data/proc/data_s3_inv.RData")
+
+pov01 <- data_s3_inv %>% filter(dataset=="pvw01")
+fs01  <- data_s3_inv %>% filter(dataset=="fsw02")
+
+#-----Meassurement model--------------------------------------#
+
+model01 <- '
+perc_merit  =~ perc_effort + perc_talent
+perc_nmerit =~ perc_wpart  + perc_netw
+pref_merit  =~ pref_effort + pref_talent
+pref_nmerit =~ pref_wpart  + pref_netw'
+
+#-----Invariance---------------------------------------------#
+
+inv01<- measurementInvariance(model=model01,data=data_s3_inv,group="dataset",estimator = "ML",strict=TRUE)
+
+conf  <- inv01$fit.configural
+weak  <- inv01$fit.loadings
+strong<- inv01$fit.intercepts
+strict<- inv01$fit.residuals
+
+tab01<- anova(conf,weak,strong,strict,SB.classic=TRUE) %>%
+  as_tibble() %>%
+  select("Chisq","Df","chisq_diff"=`Chisq diff`,"df_diff"=`Df diff`,"pvalue"=`Pr(>Chisq)`) %>%
+  mutate(stars=gtools::stars.pval(pvalue),
+         chisqt=paste0(round(Chisq,2)," (",Df,") "),
+         decision=ifelse(pvalue>0.05,yes = "Accept",no = "Reject"),
+         model=c("Configural","Weak","Strong","Strict"))
+
+fit.meas<- bind_rows(fitmeasures(inv01$fit.configural,output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
+                     fitmeasures(inv01$fit.loadings,  output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
+                     fitmeasures(inv01$fit.intercepts,output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
+                     fitmeasures(inv01$fit.residuals, output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),])
+
+# compute differences in chisq, df, cfi and rmsea (90%, lower.ci - upper.ci )
+fit.meas<- fit.meas %>%
+  mutate(diff.chi2 = chisq    - lag(chisq,default = first(chisq)),
+         diff.df   = df       - lag(df,   default = first(df)),
+         diff.cfi  = cfi      - lag(cfi,  default = first(cfi)),
+         diff.rmsea   = rmsea - lag(rmsea,default = first(rmsea))) %>%
+  round(3) %>%
+  mutate(rmsea.ci=paste0(rmsea," \n ", "(",rmsea.ci.lower,"-",rmsea.ci.upper,")"))
+
+tab.inv<- bind_cols(tab01,fit.meas) %>%
+  select(model,chisqt,cfi,rmsea.ci,diff.chi2,diff.df,diff.cfi,diff.rmsea,stars,decision) %>%
+  mutate(diff.chi2=paste0(diff.chi2," (",diff.df,") ",stars)) %>%
+  select(model,chisqt,cfi,rmsea.ci,diff.chi2,diff.cfi,diff.rmsea,decision)
+
+#clean values
+tab.inv[tab.inv == c("0 (0) ")] <- NA
+tab.inv[tab.inv == c(0)] <- NA
+
+### Check an alternative method
+configural        <- cfa(model01, data=data_s3_inv, group = "dataset", estimator="ML")
+weak.invariance   <- cfa(model01, data=data_s3_inv, group = "dataset",group.equal = "loadings", estimator="ML")
+strong.invariance <- cfa(model01, data=data_s3_inv, group = "dataset",group.equal = c( "loadings", "intercepts"), estimator="ML")
+strict.invariance <- cfa(model01, data=data_s3_inv, group = "dataset",group.equal = c( "loadings", "intercepts", "residuals"), estimator="ML")
+
+table_format = if(is_html_output()) {
+  "html"
+} else if(is_latex_output()) {
+  "latex"
+}
+
+col.nam <- c("Model","$\\chi^2 (\\text{df})$","CFI","RMSEA (90 CI)",
+             "$\\Delta \\chi^2 (\\Delta \\text{df}$)","$\\Delta \\text{CFI}$","$\\Delta \\text{RMSEA}$","Decision")
+footnote <- paste0("N = ","; Group 1, n = ",conf@Data@nobs[[1]],"; Group 2, n = ",conf@Data@nobs[[2]], ", **p < 0.001")
+
+kable(tab.inv, col.names = col.nam, align = "l",
+      booktabs=TRUE,format = table_format,escape = FALSE,
+      caption = "Multiple Group meassurement invariance for Perceptions and Preferences for Meritocracy") %>%
+  kable_styling(full_width = FALSE,latex_options = "HOLD_position") %>%
+  footnote(general = footnote, footnote_as_chunk = T)
+
+```
+
+Table \@ref(tab:invargroup) shows the results of the measurement invariance estimation.
+When attending to the traditional invariance test of $\Delta \chi^2 (\Delta \text{df})$, the results  support the invariance at the strong level meaning that the fit of the factor model of the merit scale is equivalent across samples when constraining factor loadings and intercepts to being equal.
+Such result is considered in general as evidence of invariance [citar paper Fischer en SJR], as strict forms of measurement invariance rarely hold [citar van der Schoot et al].
+Still, the comparability of latents means requires strict invariance which in this case does not hold when considering $\Delta \chi^2 (\Delta \text{df})$.
+Nevertheless, the criteria of $\Delta \text{CFI}$ used for comparing models is close to the rejection criteria of >.01, whereas the $\Delta \text{RMSEA}$ fulfills the requirements of being below of the cut-off criteria as suggested by [ @chen 2007 ].
+Therefore, using this last standard the level for strict invariance holds for the meritocracy scale.
+
 # Conclusions
@@
 Such future agendas could be specially relevant in times of economic crisis and growing inequalities, that could entail changes in the legitimation of the current distributive structure based on meritocratic ideals.
-
-\pagebreak
 
@@
           kable_styling(full_width = table_format2,latex_options = c("hold_position","scale_down"), position = "center",font_size = 9) %>%
-          add_header_above(header = c(" "=1,"Version 1"= 4,"Version 2"= 4,"Version 3"= 4)) %>%
-          add_header_above(header = c(" "=1,"Factor loadings"= 12)) %>%
+#          add_header_above(header = c(" "=1,"Version 1"= 4,"Version 2"= 4,"Version 3"= 4)) %>%
+#          add_header_above(header = c(" "=1,"Factor loadings"= 12)) %>%
           row_spec(row = 8,hline_after = TRUE) %>%
@@
           kable_styling(full_width = table_format2,latex_options = c("hold_position"), position = "center",font_size = 9) %>%
-          add_header_above(header = c(" "=1,"Factor loadings"= 4)) %>%
+#          add_header_above(header = c(" "=1,"Factor loadings"= 4)) %>%
           row_spec(row = 8,hline_after = TRUE) %>%
@@
 Table: `r as.character(paste("(\\#tab:rep-samp)", "Representativeness of the sample"))`.
+
+# Appendix sudy 1
+
+```{r tb-perpref1, echo=FALSE}
+load(file = "input/data/proc/merit_per_pref.rda")
+pacman::p_load(dplyr,knitr,kableExtra)
+options(knitr.kable.NA = '')
+table_format = if(knitr::is_html_output()) {
+  "html"
+} else if(knitr::is_latex_output()) {
+  "latex"
+}
+
+table_format2 = if(knitr::is_html_output()) {
+  T
+} else if(knitr::is_latex_output()) {
+  F
+}
+
+cfa_perpref1 <- '
+	  # latent variables
+	  percmerit =~ hwork + ambition
+	  percnmerit=~ wealthy + pareduc +race + gender +people + polcone
+	  prefmerit =~ welljob + hardjob
+	  prefnmerit=~ family + child
+	  '
+fit_perpref1 <- lavaan::cfa(cfa_perpref1, data = merit_per_pref,
+                            ordered = c("wealthy","pareduc","ambition","hwork",
+                                        "people","polcone","race","gender",
+                                        "respons","yeduc",
+                                        "family","child",
+                                        "welljob","hardjob"))
+labs <- c(
+  "Importance: hard work",
+  "Importance: having ambition",
+  "Importance: wealthy family",
+  "Importance: educated parents",
+  "Importance: race",
+  "Importance: gender",
+  "Importance: knowing people",
+  "Importance: political connections",
+  "Reasons for pay: well job",
+  "Reasons for pay: hard job",
+  "Reasons for pay: support family",
+  "Reasons for pay: has children"
+)
+
+tb.load<- data.frame(round(cbind(lavaan::inspect(fit_perpref1,
+                                                        what="std")$lambda),
+                           digits = 3))
+tb.load[tb.load==c(0.00)] <- NA
+
+for (i in names(tb.load)) {
+  tb.load[,i] <- sprintf(tb.load[,i], fmt = '%#.3f')
+}
+tb.load[tb.load=="NA"] <- ""
+#-------#
+fm01<- data.frame(t(data.frame(lavaan::fitmeasures(fit_perpref1, output ="matrix")[c("chisq","df","cfi","tli","rmsea"),]))); row.names(fm01) ="percnmerit"
+
+#------chi2, df------#
+fm04<- round(rbind(fm01),3)
+fm04.1 <- fm04 %>% dplyr::select(chisq,df)
+fm04.1$chisq <- round(x = fm04.1$chisq,digits = 1)
+fm04.1$df <- round(x = fm04.1$df,digits = 0)
+fm04.1$chi2df <- paste0(fm04.1$chisq,"(",fm04.1$df,")")
+fm04.1 <- dplyr::select(fm04.1,"chi2df")
+for (i in names(fm04.1)) {
+  fm04.1[,i] <- as.character(fm04.1[,i])
+}
+
+#------CFI, RMSEA------#
+fm04.2 <- fm04 %>% dplyr::select(cfi,tli,rmsea)
+for (i in names(fm04.2)) {
+  fm04.2[,i] <- sprintf(fm04.2[,i], fmt = '%#.3f')
+}
+
+fm.df      <- dplyr::bind_cols(fm04.1,fm04.2)
+fm.df$nobs <- c(lavaan::nobs(fit_perpref1))
+fm.df <- data.frame(t(fm.df)); colnames(fm.df) <- c("percnmerit")
+
+#------ merge ------#
+tb.fm<- dplyr::bind_rows(tb.load,fm.df)
+tb.fm<- tb.fm %>%
+  dplyr::mutate(vars=c(labs,"$\\chi^2\\text{(df)}$","$\\text{CFI}$",
+                       "$\\text{TLI}$","$\\text{RMSEA}$","$N$")) %>%
+  dplyr::select(vars,everything())
+tb.perpref1 <- tb.fm
+
+# Table --------------------------------------------------------------------.
+tb.foot <- paste0("Standardized factor loadings using DWLS estimator ; CFI = Comparative fit index ;TLI = Tucker-Lewis index; RMSEA = Root mean square error of approximation")
+tb.col <- c("","Meritocratic","Non-meritocratic",
+            "Meritocratic","Non-meritocratic")
+tb.caption <- c(" Factor loadings and fit measures for ISSP 2009")
+tb.issp<- kable(tb.perpref1,escape = FALSE,digits = 3,
+                align = "lcccc",
+                col.names = tb.col,
+                caption = tb.caption,
+                booktabs = T,
+                linesep = "",
+                format = table_format) %>%
+          kable_styling(full_width = table_format2,
+                        latex_options = c("hold_position"),
+                        position = "center",
+                        font_size = 12) %>%
+#          add_header_above(header = c(" "=1,"Perception"= 2,"Preference"= 2)) %>%
+#          add_header_above(header = c(" "=1,"Factor loadings"= 4)) %>%
+          row_spec(row = 12,hline_after = TRUE) %>%
+          add_indent(c(13:17)) %>%
+          footnote(general =tb.foot ,footnote_as_chunk = T);tb.issp
+```

### 2021-08-19 paper.Rmd
study 1 fixes
- modificaciones en preparacion de study 1, nombre de data procesada como data_s1.Rdata
- Modificacion de chunks en paper.Rmd autocontenidos
- Propuesta de orden de títulos
- Referenciadas tablas y figuras
- Propuesta de orden de anexos para study 1 y 3
- Plot likert ajustado para study 2

@@
 
-pacman::p_load(knitr)
 ```{r message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
 #RUN THIS BEFORE THE ANALYSIS
-
+pacman::p_load(knitr)
 table_format = if(is_html_output()) {
@@
 
-## Data
+### Data
 
@@
 
-## Variables
+### Variables
 
@@
 
-# Methods
+### Methods
 
@@
 
-# Results
-
-## Descriptive analyses
+### Results
+
+#### Descriptive analyses
 
@@
 ```{r desc-issp, eval=FALSE, include=FALSE, results='asis'}
-load(file = "input/data/proc/merit_per_pref.rda")
+load(file = "input/data/proc/data_s1.RData")
 pacman::p_load(dplyr)
-table_format = if(knitr::is_html_output()) {
-  "html"
+table_format = if(knitr::is_html_output()) { #conditional instructions for kable
+  "html" #if html set "html" in format
 } else if(knitr::is_latex_output()) {
-  "latex"
-}
-
-table_format2 = if(knitr::is_html_output()) {
-  T
+  "latex"#if latex set "latex" in format
+}
+
+fsize = if(knitr::is_html_output()) { #conditional instructions for kable
+  20 # if html set table fontsize to 20
 } else if(knitr::is_latex_output()) {
-  F
-}
-
-merit_per_pref %>%
+  12 # if latex set table fontsize to 12
+}
+
+data_s1 %>% 
   dplyr::select(hwork,ambition,wealthy,pareduc,
@@
 
-```{r}
-load(file = "input/data/proc/merit_per_pref.rda")
+```{r build-likert}
+load(file = "input/data/proc/data_s1.RData")
 pacman::p_load(dplyr)
-likert_issp<- merit_per_pref %>%
+likert_issp<- data_s1 %>% 
   dplyr::select(hwork,ambition,wealthy,pareduc,
                 race,gender,people,polcone,
-                welljob,hardjob,family,child) %>% na.omit()
-
+                welljob,hardjob,family,child) %>% na.omit() #Subset variables
+# Set labels for each variable
 likert_issp <- sjlabelled::set_label(likert_issp,
@@
 ```{r likert-issp,fig.height=11,fig.width=11,fig.cap="Distribution of responses in the ISSP meritocracy items", out.height='80%'}
+#Final plot
 cowplot::add_sub(
-  cowplot::plot_grid(p1,p2,p3,p4,leg,
+  cowplot::plot_grid(p1,p2,p3,p4,leg, #plot 1,2,3,4 and legend (bottom)
     align = "hv",
     axis = "a",
-    nrow = 5,
-    ncol = 1,
-    rel_heights = c(1,2,1,1,0.25)
+    nrow = 5, #5 plots
+    ncol = 1, #1 column
+    rel_heights = c(1,2,1,1,0.25) #fig size within plot
     ),
@@
 ```{r corr-issp,fig.cap = "Perceptions & preferences for ISSP  meritocracy items' polychoric correlations", fig.align='center',fig.height=6,fig.width=8}
+load(file = "input/data/proc/data_s1.RData")
 pacman::p_load(dplyr)
+# create polychoric correlation matrix
 cor_issp<- 
-  merit_per_pref %>%
+  data_s1 %>%
   dplyr::select(hwork,ambition,wealthy,pareduc,
@@
 
-## Confirmatory Factor Analysis
+#### Confirmatory Factor Analysis
 
@@
 ```{r sum-fit-issp}
-load(file = "input/data/proc/merit_per_pref.rda")
+load(file = "input/data/proc/data_s1.RData")
 pacman::p_load(dplyr, kableExtra, knitr)
-
+table_format = if(knitr::is_html_output()) { #conditional instructions for kable
+  "html" #if html set "html" in format
+} else if(knitr::is_latex_output()) {
+  "latex"#if latex set "latex" in format
+}
+
+fsize = if(knitr::is_html_output()) { #conditional instructions for kable
+  20 # if html set table fontsize to 20
+} else if(knitr::is_latex_output()) {
+  12 # if latex set table fontsize to 12
+}
+# meassurement model: First order 
 cfa_perpref1 <- ' 
@@
 	  '
-fit_perpref1 <- lavaan::cfa(cfa_perpref1, data = merit_per_pref,ordered = T)
-
+# Model fit: First order 
+fit_perpref1 <- lavaan::cfa(cfa_perpref1, data = data_s1,ordered = T)
+
+# meassurement model: Second order 
 cfa_perpref2 <- '
@@
 	  '
-fit_perpref2 <- lavaan::cfa(cfa_perpref2, data = merit_per_pref,ordered = T)
+# Model fit: second order 
+fit_perpref2 <- lavaan::cfa(cfa_perpref2, data = data_s1,ordered = T)
 
@@
 table_issp_fits <-
-  knitr::kable(sum_fit_issp, format="html", digits=3, booktabs=T,
+  knitr::kable(sum_fit_issp, format=table_format, digits=3, booktabs=T, 
                col.names=colnames,  
@@
                escape = FALSE) %>%
-  kableExtra::kable_styling(full_width = F,
-                            font_size = 15,
+  kableExtra::kable_styling(font_size = fsize,
+                            full_width = F,
                             bootstrap_options=c("striped", "bordered"))
@@
 
-_Instrument_
+#### Instrument
 
@@
 
-_Administration sets_
+#### Administration sets
 
@@
 
-_Methods_
+### Methods
 
@@
 
-_Descriptive analyses_
+#### Descriptive analyses
 
@@
 
-```{r produce-likert_s2}
+```{r produce-likert_s2, include=FALSE}
 pacman::p_load(dplyr)
@@
 table(dat_likert_s2$perc_effort_r) # check ok.
-
+# update.packages("sjPlot") # please install version 2.8.9 or above
 # plotlikert_s2<- sjPlot::plot_likert(dat_likert_s2,
@@
 #                        show.n=FALSE)
-```
-
-```{r plotlikert, echo=FALSE, fig.cap = "Distribution of responses in the Merit Scale items", fig.align='center'}
-# plotlikert_s2
+ggplot2::theme_set(ggplot2::theme(panel.background = ggplot2::element_rect(fill = "gray85",
+                                                                           colour = "gray85"),
+                                  panel.border = ggplot2::element_blank(),
+                                  axis.text.y = ggplot2::element_text(size = 13,
+                                                                      hjust = 1),
+                                  title = ggplot2::element_text(size = 13,
+                                                                face = "bold"),
+                                  legend.text = ggplot2::element_text(size = 12),
+                                  plot.caption = ggplot2::element_text(size = 10,
+                                                                       face = "plain",
+                                                                       hjust = 1)))
+
+#Plot 1: Perception 
+p1<- 
+  dplyr::select(dat_likert_s2,perc_effort_r:perc_netw_r) %>%
+  sjPlot::plot_likert(geom.colors = "PuBu",
+                      title = c("Perceptions"),                      
+                      geom.size = 0.8,
+                      axis.labels = c("Effort", "Talent", "Rich Parents", "Contacts"),
+                      catcount = 4,
+                      cat.neutral = 3,
+                      grid.range  =  c (1.2 , 1.4),
+                      values  =  "sum.outside",
+                      reverse.colors = T,
+                      reverse.scale = F,
+                      show.n = FALSE) + 
+  ggplot2::theme(legend.position = "none")
+
+#Plot 2: Preferences 
+p2<-
+  dplyr::select(dat_likert_s2,pref_effort_r:pref_netw_r) %>%
+  sjPlot::plot_likert(geom.colors = "PuBu",
+                      title = c("Preferences"),                      
+                      geom.size = 0.8,
+                      axis.labels = c("Effort", "Talent", "Rich Parents", "Contacts"),
+                      catcount = 4,
+                      cat.neutral = 3,
+                      grid.range  =  c (1.2 , 1.4),
+                      values  =  "sum.outside",
+                      reverse.colors = T,
+                      reverse.scale = F,
+                      show.n = FALSE)  + 
+  ggplot2::theme(legend.position = "none")
+
+#Legend: Extract legend for the final plot using ggpubr
+leg<- ggpubr::as_ggplot(
+  ggpubr::get_legend(
+  dplyr::select(dat_likert_s2,pref_effort_r:pref_netw_r) %>%
+      sjPlot::plot_likert(geom.colors = "PuBu", 
+                          catcount = 4,
+                          cat.neutral = 3) +
+    ggplot2::theme(legend.position = "bottom") +
+    ggplot2::scale_fill_manual(
+      values = c("#b3b3b3ff", "#f1eef6ff", "#bdc9e1ff","#74a9cfff","#0570b0ff"),
+      breaks=c("neutral",4,3,2,1),
+      labels=c("Neither agree nor disagree","Strongly disagree","Disagree",
+               "Agree","Totally agree")
+      )
+    )
+  )
+```
+
+```{r plotlikert, echo=FALSE, fig.align='center', fig.cap="Distribution of responses in the Merit Scale items", fig.height=8, fig.width=11}
+#Final plot
+cowplot::add_sub(
+  cowplot::plot_grid(p1,p2,leg, #plot 1,2 and legend (bottom)
+    align = "hv",
+    axis = "a",
+    nrow = 3, #2 plots
+    ncol = 1, #1 column
+    rel_heights = c(1,1,0.25) #fig size within plot
+    ),
+  label = paste0("n=",dim(dat_likert_s2)[1]
+    ),        # Caption: source and sample size 
+  size = 12,  # Caption font size
+  hjust = -8# Caption align 
+  ) %>% 
+  cowplot::ggdraw() 
 ```
@@
 
-_Confirmatory Factor Analysis_
+#### Confirmatory Factor Analysis
 
@@
 
-```{r table_cfa_fits, echo=FALSE}
+```{r table-cfafits, echo=FALSE}
 pacman::p_load(dplyr, kableExtra, knitr)
@@
 # Create table
-table_cfa_fits <-kable(sum_fit, format="html", digits=3, booktabs=T, col.names=colnames,  caption = "Summary fit indices according to order versions", escape = FALSE) %>% kable_styling(full_width = F, font_size = 10)
+table_cfa_fits <-kable(sum_fit, format="html", digits=3, booktabs=T, col.names=colnames,  caption = "Summary fit indices according to order versions", escape = FALSE) %>% kable_styling(full_width = F, font_size = 12,bootstrap_options=c("striped", "bordered"))
 table_cfa_fits
@@
 
-Table \@ref(tab:table_cfa_fits) shows the fit indicators of the models estimated for each of the three versions of the items' order described in the methodology.
+Table \@ref(tab:table-cfafits) shows the fit indicators of the models estimated for each of the three versions of the items' order described in the methodology.
 Regardless of the version, all models obtained adequate fit indicators, with CFI's above 0.95 and RMSEA's below 0.08.
@@
 
-_Data_
+#### Data
 
@@
 
-_Intruments_
+#### Variables
 
@@
 
-_Results_
+#### Results
 
@@
 
-Table \@ref(tab:val-conv) shows the polyserial and pearson correlations between our meritocracy scale and common measures of meritocracy.
+Table \@ref(tab:val-conv) shows the polyserial and Pearson correlations between our meritocracy scale and common measures of meritocracy.
 The findings provide evidence for convergent validity for the scale.
@@
   column_spec(column = 1, width = "1.5cm", )
-
 ```
@@
 
-*Data*
+#### Data
 
@@
 
-_Instruments_
+#### Variables
 
 The items of the meritocracy scale were identical to the original.
-The items were administrated in a randomized order (which corresponds to the application modality of Group 3 present in Figure \@ref(fig: appmod)).
-
-_Results_
-
-Tabla \@ref(tab:fcloads-inv) shows de results of the CFA estimation of the meritocracy scale in this new sample.
+The items were administrated in a randomized order (which corresponds to the application modality of Group 3 present in Figure \@ref(fig:appmod)).
+
+#### Results
+
+Table \@ref(tab:fcloads-inv) shows the results of the CFA estimation of the meritocracy scale in this new sample.
 As observed, the loadings and the fit indices show similar results as the ones presented for the original study above, which gives a first base with wich to start the invariance analysis.
@@
 
-# Analysis
-
-```{r invargroup}
-
+```{r invargroup, echo=FALSE}
 #---Load packages and data -------------------------------------#
-
-pacman::p_load(sjPlot,dplyr,lavaan,semPlot,semTools,stargazer,corrplot,psych,knitr,kableExtra,stats)
-
+pacman::p_load(sjPlot,dplyr,lavaan,semPlot,semTools,stargazer,
+               corrplot,psych,knitr,kableExtra,stats)
 load(file = "input/data/proc/data_s3_inv.RData")
-
 pov01 <- data_s3_inv %>% filter(dataset=="pvw01")
@@
 #-----Meassurement model--------------------------------------#
-
 model01 <- '
@@
 #-----Invariance---------------------------------------------#
-
-inv01<- measurementInvariance(model=model01,data=data_s3_inv,group="dataset",estimator = "ML",strict=TRUE)
+inv01<- measurementInvariance(model=model01,data=data_s3_inv,group="dataset",estimator = "ML",strict=TRUE,quiet = T)
 
@@
       caption = "Multiple Group meassurement invariance for Perceptions and Preferences for Meritocracy") %>%
-  kable_styling(full_width = FALSE,latex_options = "HOLD_position") %>%
+  kable_styling(full_width = FALSE,latex_options = "HOLD_position",bootstrap_options=c("striped", "bordered")) %>%
   footnote(general = footnote, footnote_as_chunk = T)
@@
 
-```{r tab-loads, echo=FALSE, eval=FALSE}
-pacman::p_load(dplyr)
-load(file = "output/tables/tb.fm.RData")
-cap <- "Factor loads and model fit."
-tb.col  <- c("Variables",1:4,1:4,1:4)
-tb.caption <- "Factor loadings and fit measures"
-tb.foot <- paste0("Standardized factor loadings using DWLS estimator ; CFI = Comparative fit index (scaled); RMSEA = Root mean square error of approximation (scaled)")
-
-tb.fm <- tb.fm %>%
-    mutate(vars = recode(vars,
-                         "A. Who the more they try they manage to get bigger rewards that those who strive less."= "Perception Effort",
-                         "B. Who possess more talent they manage to obtain greater rewards than those who possess less talent."= "Perception Talent",
-                         "C. Who they have rich parents manage to get out ahead."="Perception rich parents",
-                         "D. Who they have good contacts they manage to get out ahead."="Perception contacts",
-                         "E. Who the more they try they should get greater rewards than those who they try less." = "Preferences Effort",
-                         "F. Who possess more talent they should get greater rewards than those who possess less talent."= "Preferences Talent",
-                         "G. It's fine that those who have rich parents get ahead" ="Preferences rich parents",
-                         "H. Is well that those who have good contacts get ahead." = "Preferences contacts"))
-
-tb.fm01<- kable(tb.fm, format = table_format, booktabs = T,linesep = "", escape = FALSE,align = "lcccccccccccc",col.names = tb.col, caption = cap) %>%
-          kable_styling(full_width = table_format2,latex_options = c("hold_position","scale_down"), position = "center",font_size = 9) %>%
-#          add_header_above(header = c(" "=1,"Version 1"= 4,"Version 2"= 4,"Version 3"= 4)) %>%
-#          add_header_above(header = c(" "=1,"Factor loadings"= 12)) %>%
-          row_spec(row = 8,hline_after = TRUE) %>%
-          add_indent(c(9:12))  %>%
-          footnote(general = tb.foot,threeparttable = T);tb.fm01
-```
-
-```{r fcloads-inv, echo=FALSE, eval=FALSE}
-load(file = "output/tables/tb.fact-f02.RData")
-tb.foot <- paste0("Standardized factor loadings using DWLS estimator ; CFI = Comparative fit index (scaled); RMSEA = Root mean square error of approximation (scaled)")
-tb.fm02<- kable(tb.fm2,escape = FALSE,align = "lcccc",col.names = tb.col, caption = tb.caption, booktabs = T,linesep = "",format = table_format) %>%
-          kable_styling(full_width = table_format2,latex_options = c("hold_position"), position = "center",font_size = 9) %>%
-#          add_header_above(header = c(" "=1,"Factor loadings"= 4)) %>%
-          row_spec(row = 8,hline_after = TRUE) %>%
-          add_indent(c(9:12)) %>%
-          footnote(general =tb.foot ,footnote_as_chunk = T);tb.fm02
-```
-
-\pagebreak
+## Appendix study 1
 
@@
 | university or more  | 21,21%  | 18,13% |
-Table: `r as.character(paste("(\\#tab:rep-samp)", "Representativeness of the sample"))`.
-
-# Appendix sudy 1
+Table: `r as.character(paste("(\\#tab:rep-samp)", "Representativeness of the study 1 sample"))`.
 
 ```{r tb-perpref1, echo=FALSE}
-load(file = "input/data/proc/merit_per_pref.rda")
+load(file = "input/data/proc/data_s1.RData")
 pacman::p_load(dplyr,knitr,kableExtra)
 options(knitr.kable.NA = '')
-table_format = if(knitr::is_html_output()) {
-  "html"
+table_format = if(knitr::is_html_output()) { #conditional instructions for kable
+  "html" #if html set "html" in format
 } else if(knitr::is_latex_output()) {
-  "latex"
-}
-
-table_format2 = if(knitr::is_html_output()) {
-  T
+  "latex"#if latex set "latex" in format
+}
+
+fsize = if(knitr::is_html_output()) { #conditional instructions for kable
+  20 # if html set table fontsize to 20
 } else if(knitr::is_latex_output()) {
-  F
-}
-
+  12 # if latex set table fontsize to 12
+}
+
+# First order model for Appendix
 cfa_perpref1 <- '
@@
 	  '
-fit_perpref1 <- lavaan::cfa(cfa_perpref1, data = merit_per_pref,
+# Fit First order model for Appendix
+fit_perpref1 <- lavaan::cfa(cfa_perpref1, data = data_s1,
                             ordered = c("wealthy","pareduc","ambition","hwork",
@@
                            digits = 3))
-tb.load[tb.load==c(0.00)] <- NA
+tb.load[tb.load==c(0.00)] <- NA #set to NA factor loadings of 0
 
 for (i in names(tb.load)) {
-  tb.load[,i] <- sprintf(tb.load[,i], fmt = '%#.3f')
-}
-tb.load[tb.load=="NA"] <- ""
-#-------#
+  tb.load[,i] <- sprintf(tb.load[,i], fmt = '%#.3f') #set values as character
+}
+tb.load[tb.load=="NA"] <- "" #set NA values to "" (empty space)
+#---extract fit measures-------#
 fm01<- data.frame(t(data.frame(lavaan::fitmeasures(fit_perpref1, output ="matrix")[c("chisq","df","cfi","tli","rmsea"),]))); row.names(fm01) ="percnmerit"
@@
 #------chi2, df------#
-fm04<- round(rbind(fm01),3)
-fm04.1 <- fm04 %>% dplyr::select(chisq,df)
-fm04.1$chisq <- round(x = fm04.1$chisq,digits = 1)
-fm04.1$df <- round(x = fm04.1$df,digits = 0)
-fm04.1$chi2df <- paste0(fm04.1$chisq,"(",fm04.1$df,")")
-fm04.1 <- dplyr::select(fm04.1,"chi2df")
+fm04<- round(rbind(fm01),3) #select and round to 4 digits
+fm04.1 <- fm04 %>% dplyr::select(chisq,df) #select chi2 and degrees of freedom
+fm04.1$chisq <- round(x = fm04.1$chisq,digits = 1) # round Chi2 to 1 digit
+fm04.1$df <- round(x = fm04.1$df,digits = 0) #round Df to 0 digits
+fm04.1$chi2df <- paste0(fm04.1$chisq,"(",fm04.1$df,")") #Paste chi2(df)
+fm04.1 <- dplyr::select(fm04.1,"chi2df") #select final column
 for (i in names(fm04.1)) {
-  fm04.1[,i] <- as.character(fm04.1[,i])
+  fm04.1[,i] <- as.character(fm04.1[,i]) # transform all to character
 }
@@
 #------CFI, RMSEA------#
-fm04.2 <- fm04 %>% dplyr::select(cfi,tli,rmsea)
-for (i in names(fm04.2)) {
+fm04.2 <- fm04 %>% dplyr::select(cfi,tli,rmsea) # select CFI, TLI and RMSEA
+for (i in names(fm04.2)) {                      # transform all to character
   fm04.2[,i] <- sprintf(fm04.2[,i], fmt = '%#.3f')
@@
 
-fm.df      <- dplyr::bind_cols(fm04.1,fm04.2)
-fm.df$nobs <- c(lavaan::nobs(fit_perpref1))
-fm.df <- data.frame(t(fm.df)); colnames(fm.df) <- c("percnmerit")
+fm.df      <- dplyr::bind_cols(fm04.1,fm04.2) # factor loading and labels
+fm.df$nobs <- c(lavaan::nobs(fit_perpref1))   # add number of observations
+fm.df <- data.frame(t(fm.df)); colnames(fm.df) <- c("percnmerit") #transpose and set colnames
 
 #------ merge ------#
-tb.fm<- dplyr::bind_rows(tb.load,fm.df)
+tb.fm<- dplyr::bind_rows(tb.load,fm.df) #merge factor loadings and fit indices
 tb.fm<- tb.fm %>% 
@@
             "Meritocratic","Non-meritocratic")
-tb.caption <- c(" Factor loadings and fit measures for ISSP 2009")
+tb.caption <- c(" Factor loadings and fit measures of the four first-order factor model in study 1")
 tb.issp<- kable(tb.perpref1,escape = FALSE,digits = 3,
@@
                 format = table_format) %>%
+          kable_styling(latex_options = c("hold_position"),
+                        bootstrap_options = c("striped", "bordered"),
+                        position = "center",
+                        font_size = 12) %>%
+          add_header_above(header = c(" "=1,"Perception"= 2,"Preference"= 2)) %>%
+          add_header_above(header = c(" "=1,"Factor loadings"= 4)) %>%
+          row_spec(row = 12,hline_after = TRUE) %>%
+          add_indent(c(13:17)) %>%
+          footnote(general =tb.foot ,footnote_as_chunk = T,threeparttable = T);tb.issp
+```
+
+```{r tab-loads, echo=FALSE, eval=FALSE}
+pacman::p_load(dplyr)
+load(file = "output/tables/tb.fm.RData")
+cap <- "Factor loads and model fit."
+tb.col  <- c("Variables",1:4,1:4,1:4)
+tb.caption <- "Factor loadings and fit measures"
+tb.foot <- paste0("Standardized factor loadings using DWLS estimator ; CFI = Comparative fit index (scaled); RMSEA = Root mean square error of approximation (scaled)")
+tb.fm <- tb.fm %>%
+    mutate(vars = recode(vars,
+                         "A. Who the more they try they manage to get bigger rewards that those who strive less."= "Perception Effort",
+                         "B. Who possess more talent they manage to obtain greater rewards than those who possess less talent."= "Perception Talent",
+                         "C. Who they have rich parents manage to get out ahead."="Perception rich parents",
+                         "D. Who they have good contacts they manage to get out ahead."="Perception contacts",
+                         "E. Who the more they try they should get greater rewards than those who they try less." = "Preferences Effort",
+                         "F. Who possess more talent they should get greater rewards than those who possess less talent."= "Preferences Talent",
+                         "G. It's fine that those who have rich parents get ahead" ="Preferences rich parents",
+                         "H. Is well that those who have good contacts get ahead." = "Preferences contacts"))
+tb.fm01<- kable(tb.fm, format = table_format, booktabs = T,linesep = "", escape = FALSE,align = "lcccccccccccc",col.names = tb.col, caption = cap) %>%
+          kable_styling(full_width = table_format2,latex_options = c("hold_position","scale_down"), position = "center",font_size = 9) %>%
+#          add_header_above(header = c(" "=1,"Version 1"= 4,"Version 2"= 4,"Version 3"= 4)) %>%
+#          add_header_above(header = c(" "=1,"Factor loadings"= 12)) %>%
+          row_spec(row = 8,hline_after = TRUE) %>%
+          add_indent(c(9:12))  %>%
+          footnote(general = tb.foot,threeparttable = T);tb.fm01
+```
+
+## Appendix study 3
+
+```{r fcloads-inv, echo=FALSE}
+table_format = if(knitr::is_html_output()) { #conditional instructions for kable
+  "html" #if html set "html" in format
+} else if(knitr::is_latex_output()) {
+  "latex"#if latex set "latex" in format
+}
+table_format2 = if(knitr::is_html_output()) {
+  T
+} else if(knitr::is_latex_output()) {
+  F
+}
+# Note: 
+# This CFA corresponds to the sample from the Factorial Survey Study 
+load(file = "input/data/proc/data_s3_inv.RData") #load data study 3
+fs01  <- data_s3_inv %>% dplyr::filter(dataset=="fsw02") #subset sample
+
+# Set model
+model01 <- '
+perc_merit  =~ perc_effort + perc_talent
+perc_nmerit =~ perc_wpart  + perc_netw
+pref_merit  =~ pref_effort + pref_talent
+pref_nmerit =~ pref_wpart  + pref_netw'
+
+# Model fit
+fs1_o <- lavaan::cfa(model = model01,
+             data = fs01,
+             ordered = c("perc_effort","perc_talent",
+                         "perc_wpart","perc_netw",
+                         "pref_effort","pref_talent",
+                         "pref_wpart","pref_netw"),std.lv=FALSE)
+
+# Extract information for table
+fm01<- data.frame(t(data.frame(lavaan::fitmeasures(fs1_o, output ="matrix")[c("chisq.scaled","df","cfi.scaled","rmsea.scaled"),]))); row.names(fm01) ="M1.1"
+
+#------chi2, df------#
+fm04.1 <- fm01 %>% dplyr::select(chisq.scaled,df) 
+fm04.1$chisq.scaled <- round(x = fm04.1$chisq.scaled,digits = 1)
+fm04.1$df <- round(x = fm04.1$df,digits = 0)
+fm04.1$chi2df <- paste0(fm04.1$chisq.scaled,"(",fm04.1$df,")")
+fm04.1 <- dplyr::select(fm04.1,"chi2df")
+
+#------CFI, RMSEA------#
+fm04.2 <- fm01 %>% dplyr::select(cfi.scaled,rmsea.scaled)
+for (i in names(fm04.2)) {
+  fm04.2[,i] <- sprintf(fm04.2[,i], fmt = '%#.3f')
+}
+
+#------ merge fit indices ------#
+fm.df      <- dplyr::bind_cols(fm04.1,fm04.2)
+fm.df$nobs <- c(lavaan::nobs(fs1_o)) 
+fm.df      <- data.frame(t(fm.df))
+
+#------ extract factor loadings ------#
+loads01<- data.frame(round(inspect(fs1_o,what="std")$lambda,digits = 3))
+loads01[loads01==c(0.00)] <- NA; names(loads01) <- c(paste0("M1.",1:4))
+
+for (i in names(loads01)) {
+  # tb.load[,i] <- sjlabelled::as_character(tb.load[,i])
+  loads01[,i] <- sprintf(loads01[,i], fmt = '%#.3f')
+}
+loads01[loads01=="NA"] <- ""
+
+#------ labels for table ------#
+labs <-c(
+    "A. Perception Effort",
+    "B. Perception Talent",
+    "C. Perception Rich parents",
+    "D. Perception Contacts",
+    "E. Preferences Effort",
+    "F. Preferences Talent",
+    "G. Preferences Rich parents",
+    "H. Preferences Contacts")
+
+#------ merge ------#
+tb.fm2<- bind_rows(loads01,fm.df)
+tb.fm2<- tb.fm2 %>% 
+  mutate(vars=c(labs,"$\\chi^2\\text{(df)}$","$\\text{CFI}$","$\\text{RMSEA}$","$N$")) %>%
+  select(vars,everything())
+# load(file = "output/tables/tb.fact-f02.RData")
+tb.col  <- c(" ",
+             "Meritocratic","Non-meritocratic",
+             "Meritocratic","Non-meritocratic")
+tb.caption <- "Factor loadings and fit measures for the second sample for study 3" 
+tb.foot <- paste0("Standardized factor loadings using DWLS estimator ; CFI = Comparative fit index (scaled); RMSEA = Root mean square error of approximation (scaled)")
+tb.fm02<- kable(tb.fm2,escape = FALSE,align = "lcccc",col.names = tb.col, caption = tb.caption, booktabs = T,linesep = "",format = table_format) %>%
           kable_styling(full_width = table_format2,
@@
                         position = "center",
-                        font_size = 12) %>%
-#          add_header_above(header = c(" "=1,"Perception"= 2,"Preference"= 2)) %>%
-#          add_header_above(header = c(" "=1,"Factor loadings"= 4)) %>%
-          row_spec(row = 12,hline_after = TRUE) %>%
-          add_indent(c(13:17)) %>%
-          footnote(general =tb.foot ,footnote_as_chunk = T);tb.issp
-```
+                        font_size = 12,
+                        bootstrap_options = c("striped", "bordered")) %>%
+          add_header_above(header = c(" "=1,"Perception"= 2,"Preference"= 2)) %>%
+          add_header_above(header = c(" "=1,"Factor loadings"= 4)) %>%
+          row_spec(row = 8,hline_after = TRUE) %>%
+          add_indent(c(9:12)) %>%
+          footnote(general =tb.foot ,footnote_as_chunk = T);tb.fm02
+```

### 2021-08-19 paper.Rmd
colors
- caption y color en plot

@@
 whether hard work should be related to payment).
-The rows in Figure 1 consider the distinction between meritocratic and non-meritocratic dimensions [@reynolds_perceptions_2014].
+The rows in Figure \@ref(fig:merit-model) consider the distinction between meritocratic and non-meritocratic dimensions [@reynolds_perceptions_2014].
 These aspects have usually been treated as different ends of the same continuum in previous research, an assumption that requires empirical scrutiny.
@@
     ),
-  label = paste0("n=",dim(dat_likert_s2)[1]
+  label = paste0("Source: Authors calculations based on Study 1 sample"," (n=",dim(na.omit(dat_likert_s2))[1],")"
     ),        # Caption: source and sample size 
   size = 12,  # Caption font size
-  hjust = -8# Caption align 
+  hjust = -0.15# Caption align 
   ) %>% 

### 2021-09-02 processing/preparation_study1.Rmd

@@
 ```
-
 This is the preparation code of the data for Study 1: Measuring meritocracy with international secondary data. The prepared data is `data_s1.RData`

### 2021-09-02 paper.Rmd
chunks
- hide chunks

@@
 #------ extract factor loadings ------#
-loads01<- data.frame(round(inspect(fs1_o,what="std")$lambda,digits = 3))
+loads01<- data.frame(round(lavaan::inspect(fs1_o,what="std")$lambda,digits = 3))
 loads01[loads01==c(0.00)] <- NA; names(loads01) <- c(paste0("M1.",1:4))

### 2021-09-02 paper.Rmd
update
- check chunks
- render paper en html OK

@@
 pacman::p_load(knitr, kableExtra, dplyr)
-
+table_format = if(knitr::is_html_output()) { #conditional instructions for kable
+  "html" #if html set "html" in format
+} else if(knitr::is_latex_output()) {
+  "latex"#if latex set "latex" in format
+}
 tabitems <- read.csv(file = "input/tables/table01.csv",header = 1,sep = ",",encoding = "UTF-8") # call file generated externally for the table
@@
 load(file = "input/data/proc/data_s2.RData")
-
+pacman::p_load(dplyr)
 # generate polychoric matrix
-cor_s2<- data_s2 %>% select(perc_effort,perc_talent,perc_wpart,perc_netw,pref_effort,pref_talent,pref_wpart,pref_netw) %>% lavaan::lavCor(., ordered=names(.))
+cor_s2<- data_s2 %>% dplyr::select(perc_effort,perc_talent,perc_wpart,perc_netw,pref_effort,pref_talent,pref_wpart,pref_netw) %>% lavaan::lavCor(., ordered=names(.))
 
@@
 
-```{r tab-loads, echo=FALSE, eval=FALSE}
-pacman::p_load(dplyr)
-load(file = "output/tables/tb.fm.RData")
-cap <- "Factor loads and model fit."
-tb.col  <- c("Variables",1:4,1:4,1:4)
-tb.caption <- "Factor loadings and fit measures"
-tb.foot <- paste0("Standardized factor loadings using DWLS estimator ; CFI = Comparative fit index (scaled); RMSEA = Root mean square error of approximation (scaled)")
-tb.fm <- tb.fm %>%
-    mutate(vars = recode(vars,
-                         "A. Who the more they try they manage to get bigger rewards that those who strive less."= "Perception Effort",
-                         "B. Who possess more talent they manage to obtain greater rewards than those who possess less talent."= "Perception Talent",
-                         "C. Who they have rich parents manage to get out ahead."="Perception rich parents",
-                         "D. Who they have good contacts they manage to get out ahead."="Perception contacts",
-                         "E. Who the more they try they should get greater rewards than those who they try less." = "Preferences Effort",
-                         "F. Who possess more talent they should get greater rewards than those who possess less talent."= "Preferences Talent",
-                         "G. It's fine that those who have rich parents get ahead" ="Preferences rich parents",
-                         "H. Is well that those who have good contacts get ahead." = "Preferences contacts"))
-tb.fm01<- kable(tb.fm, format = table_format, booktabs = T,linesep = "", escape = FALSE,align = "lcccccccccccc",col.names = tb.col, caption = cap) %>%
-          kable_styling(full_width = table_format2,latex_options = c("hold_position","scale_down"), position = "center",font_size = 9) %>%
-#          add_header_above(header = c(" "=1,"Version 1"= 4,"Version 2"= 4,"Version 3"= 4)) %>%
-#          add_header_above(header = c(" "=1,"Factor loadings"= 12)) %>%
-          row_spec(row = 8,hline_after = TRUE) %>%
-          add_indent(c(9:12))  %>%
-          footnote(general = tb.foot,threeparttable = T);tb.fm01
-```
-
 ## Appendix study 3
@@
 load(file = "input/data/proc/data_s3_inv.RData") #load data study 3
+pacman::p_load(dplyr,knitr,kableExtra)
 fs01  <- data_s3_inv %>% dplyr::filter(dataset=="fsw02") #subset sample
@@
 for (i in names(loads01)) {
-  # tb.load[,i] <- sjlabelled::as_character(tb.load[,i])
   loads01[,i] <- sprintf(loads01[,i], fmt = '%#.3f')
@@
   select(vars,everything())
-# load(file = "output/tables/tb.fact-f02.RData")
 tb.col  <- c(" ",

### 2021-09-03 paper.Rmd
version for SJR

--- 2021-09-02 paper.Rmd
@@
 author: "Juan Carlos Castillo, Julio Iturra, Luis Maldonado, Francisco Meneses & Jorge Atria <br><br>"
-abstract: "Economic and social inequalities have generated growing concern and crises across contemporary societies.
+abstract: "Economic and social inequalities have generated growing concerns and crises across contemporary societies.
 One of the mechanisms proposed by social sciences to explain the persistence of inequality is the belief in meritocracy, which would legitimize economic disparities based on differences in effort and talent.
-Despite its wide use as a concept, empirical research on meritocracy is relatively novel and characterized by diverse conceptualization and measures that make the findings and their interpretation rather confusing.
+Despite its wide use as a concept, empirical research on meritocracy is relatively novel and characterized by diverse conceptualization and measures that make the findings and their interpretation rather inconsistent.
 Most of the studies in the area have relied upon secondary data to operationalize meritocracy, with a wide variation in the use and interpretation of the same survey items.
 Taking into account the extant literature that uses measures of meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding conceptualization and operationalization.
-Based on this critical analysis, we propose a survey instrument called *Perceptions and Preferences for Meritocracy Scale*, which we test with confirmatory factor analysis using data from an online survey study (N=2,141).
-Our results support the proposed conceptual structure, which not only distinguishes between perceptions and preferences, but also between meritocratic and non-meritocratic dimensions.
-Our discussion highlights the importance of considering these different dimensions in order to advance the study of meritocracy."
+Based on this critical analysis, we propose a conceptual framework for studying perceptions and preferences for meritocracy and non-meritocracy, which is then tested through confirmatory analysis using ISSP (International Social Survey Programme) data as well as a novel survey designed specifically with this purpose(N=2,141).
+Our results support the conceptual framework and its operationalizarion, although with a better fit for the proposed scale than for the ISSP survey.
+Our discussion highlights the importance of considering different dimensions in order to advance in the study of meritocracy."
 output:
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(
-	echo = FALSE,
-	fig.pos = "H",
-	message = FALSE,
-	warning = FALSE,
-	cache = FALSE,
-	out.width = "85%"
-)
+knitr::opts_chunk$set(echo=FALSE, warning = FALSE,message = FALSE, cache = FALSE,out.width = '85%',fig.pos= "H")
 # knitr::opts_knit$set(base.url = "../") #relative path for .html output file
@@
 
-Economic inequality and income concentration have become an issue of growing concern in the last years.
-This has been expressed in a series of social movements as well as diverse critical analyses regarding the development of capitalism and its consequences [@streeckPoliticsPublicDebt2014; @piketty_Capital_2014].
+Economic inequality and income concentration have become topics of growing concern over the last years, acquiring even more salience in times of pandemic and economic crises.
+This has led to a series of social movements in different societies as well as diverse critical analyses regarding the development of capitalism and its consequences [@streeckPoliticsPublicDebt2014; @piketty_Capital_2014].
 In this context, the study of views, preferences and perceptions of inequality has acquired relevance in the social sciences, in topics such as redistributive preferences [@alesina_Fairness_2005; @dimick_Models_2018], the legitimization of economic inequality [@schroder_Income_2017] and the functioning of meritocracy [@duru-bellat_whos_2012; @mijs_paradox_2019; @reynolds_perceptions_2014; @atriaEconomicElitesAttitudes2020].
@@
 For example, some studies have linked meritocracy to the reinforcement of socio-economic, gender, and ethnic stereotypes [@MadeiraPrimesConsequencesSystematic2019; @girerdNeoliberalismIdeologicalBarrier2020; @premingerMeritocracyServiceEthnocracy2020], as well as the effects of meritocratic beliefs in educational [@generettStoriesWeTell2020; @owensENGINESSOCIALMOBILITY2020] and organizational contexts [@perezAdvancingCareersMerit2020; @aielloNewEvidenceFirmuniversity2019].
-Most of such studies so far have used indicators from existing standard surveys (as the International Social Survey Programme - ISSP) to measure meritocracy.
+Most of such studies so far have used indicators from existing standard social surveys, such as the International Social Survey Programme (ISSP), to measure meritocracy.
 However, as we will show later, the concepts used as well as the instruments for  measuring meritocracy vary extensively among studies.
 In many cases, similar phenomena are associated with different indicators, and conversely, often different phenomena are measured with similar indicators.
-Such trends not only limits the comparability of studies, but also the ability to understanding the effects of meritocratic beliefs across different lines of research.
-
-Based on a critical analysis of different approaches to the measurement of meritocracy to date, this article proposes an instrument to both measure and relate two key aspects in the study of meritocracy: perceptions and preferences.
-Furthermore, as a second axis of analysis, it considers the elaboration of indicators regarding meritocratic and non-meritocratic aspects, demonstrating that they are not two poles of the same continuum as some previous studies would seem to suggest.
-The measurement system outlined is oriented to generate an instrument as simple and brief as possible, being suitable to be used in public opinion surveys and allowing to integrate meritocratic beliefs in the study of different social phenomena.
+Such trends not only limits the comparability of studies, but also the ability to understanding the effects of meritocratic beliefs across different disciplines and lines of research.
+
+Based on a critical analysis of different approaches to the measurement of meritocracy to date, this article proposes an instrument to both measure and relate two main aspects in the study of meritocracy: perceptions and preferences.
+Furthermore, it distinguishes between meritocratic and non-meritocratic dimensions, as they are not necessarily two poles of the same continuum as previous studies seem to suggest.
+This conceptual framework is oriented to develop a measurement instrument as simple and brief as possible, being suitable to be used in public opinion surveys and allowing to integrate meritocratic beliefs in the study of different social phenomena.
 
@@
 
-In the following we discuss four critical aspects in studies conceptualizing and measuring meritocracy and that are the basis of our scale proposal.
+In the following we discuss four critical aspects in studies conceptualizing and measuring meritocracy and that are the basis of the conceptual and measurement proposal.
 
@@
 
-In this measure of meritocratic beliefs, we can observe a couple of strong conceptual assumptions: effort would not depend on parental influence, and talent (as innate ability) is not meritocratic (contrary to Michael Youngs' original conceptualization).
+In this conceptual and measurement approach of meritocratic beliefs, we can observe a couple of strong conceptual assumptions: a) effort would not depend on parental influence, and b) talent (as innate ability) is not meritocratic (contrary to Michael Youngs' original conceptualization).
 This conceptual and measurement-based assumption is found in other studies, that also assume effort to be the main and only aspect of meritocracy [@girerdNeoliberalismIdeologicalBarrier2020a; @bubakPerceptionsMeritocracyNote2019], raising the question: Is effort the only dimension behind the concept of meritocracy?
-The question of whether talent is or is not considered meritocratic is certainly an interesting topic to discuss from a philosophical point of view, but for those working on empirical studies it should be something that face empirical scrutiny.
-Besides, the consideration of talent as part of meritocracy certainly opens some interesting avenues of research, for instance in studies showing that for the elites meritocracy is more related to talent, whereas effort is something more characteristic of meritocracy of the middle and low classes ([citar Atria ABS]).
+The question of whether talent (as inteligence and ability) is or is not considered meritocratic is certainly an interesting topic to discuss from a philosophical point of view, but for those working on empirical studies it should face empirical scrutiny.
+Besides, the consideration of talent as part of meritocracy certainly opens some interesting avenues of research, for instance in studies showing that for the elites meritocracy is more related to talent, whereas effort is something more characteristic of meritocracy of the middle and low classes ([### citar Atria ABS]).
 
@@
 
-Several approaches to the empirical study of meritocracy based on public opinion surveys, refer to the concept of _beliefs_, however different meanings and operationalizations usually underlie this concept.
+Several approaches to the empirical study of meritocracy based on public opinion surveys refer to the concept of _beliefs_, but with wide differences in meanings and operationalizations.
 To illustrate this point, lets consider the most commonly used items for operationalizing meritocratic beliefs in survey research (as ISSP), which is the "reasons to get ahead" battery.
 It consists in a series of items asking "how important you think it is for getting ahead in life" and then lists several factors, such as effort, education, parental wealth, and contacts.
-Another version of this same battery used in several surveys - sometimes along the previous one - asks about "how important you think it _should_ be ...", then listing the same concepts.
-Therefore, the question raised here is: Which one of both is a "belief"?
-What _is_ or what it _should_ be?
+Another version of this same battery used in other surveys - sometimes together with the previous one - asking about "how important you think it _should_ be ...", and then listing the same concepts.
+Therefore, the question raised here is: Which one of both is a "belief": what _is_ or what it _should_ be?
 
@@
 what is).
-This occurs for instance in @reynolds_perceptions_2014, in which the term beliefs is used to talk about perceptions, whereas other authors use general terms such as attitudes [@kunovich_systems_2007].
-The first attempt to shed light on this issue was made by @duru-bellat_whos_2012, who used the question "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" as a proxy for "desired" meritocracy (beliefs).
+This occurs for instance in @reynolds_perceptions_2014, in which the term beliefs is used to talk about what @janmaatSubjectiveInequalityReview2013 refer as perceptions, whereas other authors use general terms such as attitudes [@kunovich_systems_2007].
+The first attempt to shed light on this issue in meritocracy research was made by @duru-bellat_whos_2012, who used the question "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" as a proxy for "desired" meritocracy (normative beliefs).
 They then determined "perceived" meritocracy, using the questions: “Would you say that in your country, people are rewarded for their efforts?” and “...
@@
 Is the belief in meritocracy a perception or a preference with normative meaning?
-In order to expand the analytical conceptual framework, we believe that both dimensions should be included in analyses, as proposed by @duru-bellat_whos_2012.
+In order to expand the analytical conceptual framework, we believe that both dimensions should be included as proposed by @duru-bellat_whos_2012.
 This opens up the possibility of analyzing whether perceptions and preferences are actually related (i.e.
@@
 
-A number of studies have dealt with aspects opposed to meritocracy, as for instance the use of personal advantages (as contacts or having a wealthy family) to get ahead in life.
+Some research in meritocracy considers aspects usually opposed to effort and talent for personal achievement, as for instance the use of personal advantages (as contacts or having a wealthy family) to get ahead in life.
 For instance, @kunovich_systems_2007 used an items’ battery listing a number of factors in relation to “How important each should be in deciding pay...” (as @duru-bellat_whos_2012 for desired meritocracy).
@@
 Similarly, @roex_attitudes_2019 used ISSP indicators for perceived meritocracy and non-meritocracy to build a single score by reverse coding the non-meritocratic items.
-Therefore, the treatment of non-meritocratic items has been rather inconsistent across studies and the assumption that they are the simple opposite of meritocracy requires further empirical assessment.
+Therefore, the treatment of non-meritocratic items has been rather inconsistent across studies and the assumption that they are the simple opposite of meritocracy certainly requires further empirical assessment.
 
@@
 
-Finally, most studies in meritocracy so far have not accounted for the issue of measurement error [@brownBookReviewTimothy2010; @bollenStructuralEquationsLatent1989] (Ansolabehere, S., Rodden, J., & Snyder Jr, J.
-M., 2008), as they use single indicators and/or simple average indexes for measuring meritocracy.
+Finally, most studies in meritocracy so far have not properly considered the issue of latent structures and measurement error [@brownBookReviewTimothy2010; @bollenStructuralEquationsLatent1989] (Ansolabehere, S., Rodden, J., & Snyder Jr, J.
+M., 2008), as they mostly use single indicators and/or simple average indexes for measuring meritocracy.
 Such strategy assumes that the latent construct is measured perfectly (i.e.
-no error or residual variance) by the indicators chosen, going as far as to propose that "...
+no error or residual variance) by the selected indicators, going as far as to propose that "...
 In choosing this strategy of index construction, we argue that support for meritocracy is not a latent variable" [@kunovich_systems_2007 p.653-654].
-Some advances were made by @reynolds_perceptions_2014 through conducting a principal component analysis of meritocratic and non-meritocratic dimensions, but somewhat contradictorily, they chose a sum index despite proving a multidimensional latent structure.
-
-## An instrument proposal {-#instrumentprop}
-
-Based on the previous assumptions and limitations of the measurements of meritocracy, we propose and test an instrument with the following characteristics:
+Although some advances were made by @reynolds_perceptions_2014 through conducting a principal component analysis of meritocratic and non-meritocratic dimensions, somewhat surprisingly they finally chose to build a sum index despite proving a multidimensional latent structure.
+
+## Conceptual proposal for studying meritocracy {-#instrumentprop}
+
+Based on the previous assumptions and limitations identified in the empirical study of  meritocracy, we propose a conceptual and measurement framework with the following characteristics:
 
@@
 The columns "Perceptions" and "Preferences" represent the distinction between these two concepts, usually confused under the label "beliefs" [@castillo_meritocracia_2019].
-Perceptions refer to the extent to which people observe that meritocracy functions or applies in their society, which relates to items such as "I think hard work is important to get ahead in society".
+Perceptions refer to the extent to which people observe that meritocracy functions or applies in their society, which relates to items such as "Hard work is important to get ahead in society".
 Preferences refer to normative expectations that are usually linked to a "should" expression (e.g.
 whether hard work should be related to payment).
-The rows in Figure \@ref(fig:merit-model) consider the distinction between meritocratic and non-meritocratic dimensions [@reynolds_perceptions_2014].
-These aspects have usually been treated as different ends of the same continuum in previous research, an assumption that requires empirical scrutiny.
-Non-meritocratic elements usually refer to the use of personal contacts or family advantages to get ahead in life.
-
-Regarding the selection of indicators, most are taken or adapted from previous studies for the sake of comparability.
-For meritocratic indicators, we use effort and talent as the main components of the traditional concept of merit as defined by @young_rise_1962, whereas for non-meritocratic dimensions, we use wealthy parents and good contacts.
-Descriptions of the specific items are outlined in the methodology section.
-
-The research hypotheses behind this conceptualization and measurement model are as follows:
-
-* $H_1$. The perception of meritocracy is a latent variable based on indicators of the importance attributed to talent and effort to get ahead in life.
-
-* $H_2$. The perception of non-meritocracy is a latent variable that derives from  indicators related to agreement with the statement that people with contacts and rich parents manage to get ahead.
-
-* $H_3$. Meritocratic preferences are a latent variable based on the normative value of effort and talent.
-
-* $H_4$. Non-meritocratic preferences are a latent variable based on the normative value of the use of personal contacts and having wealthy parents.
+The rows in Figure \@ref(fig:merit-model) consider the distinction between meritocratic and non-meritocratic dimensions [@reynolds_perceptions_2014], usually treated as different ends of the same continuum in previous research.
+Non-meritocratic elements refer to the use of resources such as personal contacts or family advantages to get ahead in life.
 
@@
   }
+
 ```
@@
 The analysis is organized in three studies.
-Study 1 serves as a background to our instrument proposal by attempting to operationalize the meritocracy dimensions with available secondary data.
-For this we use data from the International Social Survey Programme (ISSP).
-Although we are aware that the available indicators in this dataset do not allow a precise operationalization of the conceptual model, we know that this dataset is the most widely used in comparative meritocracy studies (as well as overall subjective inequality) and therefore it is certainly useful to look at its potentials and limitations to this regard.
+Study 1 serves as a background to our instrument proposal by attempting to operationalize the meritocracy dimensions with available secondary data from the inequality module of the International Social Survey Programme (ISSP).
+Although we are aware that the available indicators in this dataset do not allow a precise operationalization of the conceptual model, we know that it is the most widely used data in comparative meritocracy studies (as well as overall subjective inequality) and therefore it is certainly useful to look at its potentials and limitations to this regard.
 Study 2 is the core of this paper and consists of the proposal of a new scale for measuring meritocratic perceptions and preferences.
@@
 
-The data is the last available wave of the social inequality module from the International Social Survey Programme (ISSP), which is the most specialized international comparative survey in perceptions, attitudes and beliefs about inequality related issues.
+The data cooresponds to the last available wave of the social inequality module from the International Social Survey Programme (ISSP), which is the most specialized international comparative survey in perceptions, attitudes and beliefs about inequality related issues.
 This wave corresponds to the year 2009 and covers attitudes towards a series of topics dealing with social inequality across 41 countries and 56021 individuals.
-Although there is data available from this module for previous waves (1987, 1992 and 1999), unfortunately there are several variables that are important for this research that were not included, particularly in 1987 and 1999, reason why we only use the 2009 wave.
+Although there is data available from this module for previous waves (1987, 1992 and 1999), unfortunately there are several variables that are important for this research that were not included, particularly in 1987 and 1999, reason why we only use the 2009 wave (during the development of this research the 2019 wave became available, but so far only with a restricted set of countries).
 
@@
 However, it is important to mention  that the items were not originally designed as a specific measure of meritocracy.
-Still, as this database is one of the most widely used for international social research, many of the items have been used in previous research on meritocracy (and certainly will keep being used in the future).
-Therefore, before proposing a new measure of meritocracy we  give a first test to our proposed conceptualization using available data that not only covers several countries but also is validated by the research community.
-
 The variables are presented below organized in perceptions and preferences:
-
-- _Perception of meritocracy/non-meritocracy_: for operationalizing perceptions the closest set of ISSP's indicators comes from the question asking about perceptions for opportunities to get ahead, which are usually considered as "meritocratic beliefs" in previous research.
-The general heading of the battery is: _"To begin we have some questions about opportunities for getting ahead.
-Please tick one box for each of these to show how important you think it is for getting ahead in life."_ This is followed by a list of statements to be answered in a scale with the following levels numbered from 1 to 5: essential, very important, fairly important, not very important, not important at all.
-
-The classification is based on criteria of internal motivation (meritocratic) and structural constraints (non-meritocratic).
-There were two items from the battery that were excluded from the analysis as they would not fit into the classification.
-The first one was "having good education yourself", since it was not clear whether this is could be due to individual motivation or system opportunities, and the second was "giving bribes", as introduced elements of criminality that were beyond a non-meritocratic perception.
-
-- Preferences for meritocracy-non meritocracy: for the operationalization of normative preferences we used a list of items related to reasons for pay.
-The ISSP question was: _In deciding how much people ought to earn, how important should each of these things be, in your opinion_, rated in the same _essential-non important at all_ scale as the questions for meritocratic preferences.
-
-Table \@ref(tab:table-issp) below summarise the items classified according to the proposal based on perception and preference regarding to the meritocratic/non-meritocratic dimensions:
 
@@
 
+- _Perception of meritocracy/non-meritocracy_: for operationalizing perceptions the closest set of ISSP's indicators comes from the question asking about perceptions for opportunities to get ahead, which are usually considered as "meritocratic beliefs" in previous research.
+The general heading of the battery is: _"To begin we have some questions about opportunities for getting ahead.
+Please tick one box for each of these to show how important you think it is for getting ahead in life."_ This is followed by a list of statements to be rated from 1 to 5: essential, very important, fairly important, not very important, not important at all.
+
+The classification of the items is based on criteria of internal motivation (meritocratic) and structural constraints (non-meritocratic).
+There were two items from the battery that were excluded from the analysis as they would not fit into the classification.
+The first one was "having good education yourself", since it was not clear whether this is could be due to individual motivation or system opportunities, and the second was "giving bribes", as introduced elements of criminality that were beyond a non-meritocratic perception.
+
+- Preferences for meritocracy-non meritocracy: for the operationalization of normative preferences we used a list of items related to reasons for pay battery.
+The ISSP question was: _In deciding how much people ought to earn, how important should each of these things be, in your opinion_, rated in the same _essential-non important at all_ scale (1 to 5) as the questions for meritocratic perceptions.
+
 ### Methods
@@
 The estimation was performed using Confirmatory Factor Analysis (CFA).
-CFA was conducted using the `lavaan` R package (version 0.6-3; Rosseel, 2020), with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement (Kline, 2016; Rosseel, 2020).
-As recommended by Brown (2008), we assessed model fit by jointly considering the comparative fit index and Tucker-Lewis Index (CFI and TLI; acceptable fit > 0.95), Root of the average squared residual approximation (RMSEA; acceptable fit < 0.08),  Chi-square: (p-value; acceptable fit > 0.05, and Chi-square ratio > 3).
+CFA was conducted using the `lavaan` R package (version 0.6-3; ### Rosseel, 2020), with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement (### Kline, 2016; Rosseel, 2020).
+As recommended by ### Brown (2008), we assessed model fit by jointly considering the comparative fit index and Tucker-Lewis Index (CFI and TLI; acceptable fit > 0.95), Root of the average squared residual approximation (RMSEA; acceptable fit < 0.08),  Chi-square: (p-value; acceptable fit > 0.05, and Chi-square ratio > 3).
 
@@
 #### Descriptive analyses
-
-<!-- (here summary descriptive table - @julio) -->
-
-<!-- Table \@ref(tab:desc-issp) -->
 
@@
 
-Figure \@ref(fig:likert-issp) shows the distribution of responses across the indicators for each perception and preference in the meritocratic and non-meritocratic dimensions.
-On one side, we see that there is a certain degree of importance attributed to factors such as hard work and ambition in the process of getting ahead, concentrating 94.9% and 92.8% in the fairly important to essential categories.
-
-By contrast, for the non-meritocratic aspects, parental background showed a more dispersed opinion, where 61.2% considers that coming from a wealthy family is important, and 76% considers that having educated parents is important.
-On the other hand, the background dimension shows that people' opinions are generally in the opposite sense, in which 63.7% consider that a person's racial origin is not very important or not important at all in the process of getting ahead in life, while 35% consider it important, and this is nearly identical regarding the importance of gender.
-Lastly, we can observe differences in the networks dimension, in which 85.7% consider it to be important to know the right people in the process of getting ahead in life, contrasting with 50% who consider political connections to be important.
-
-Regarding normative preferences, we observe how well a person performs his or her job and amount of effort involved concentrates 98% and 96.9% of the responses in the categories that attribute importance to these characteristics when evaluating the pay that people deserve.
-Finally, in relation to structural constrains, 83% attribute importance to the fact that supporting a family should be a relevant factor with respect to the pay obtained, while having children to support translates into 76.4% of importance in this dimension.
+Figure \@ref(fig:likert-issp) shows the distribution of responses across the selected items in their corresponding dimensions.
+On the one side, we see that there is a high degree of importance attributed to factors such as hard work and ambition in the process of getting ahead, concentrating 94.9% and 92.8% in the fairly important to essential categories.
+By contrast, the perception non-meritocratic aspects is rated lower that the meritocratic ones, particularly for gender, race and political connections.
+Regarding normative preferences, we observe that the meritocratic ones are deemed as important (from fairly to essential) for nearly the whole sample, support that decreases slightly for the non-meritocratic ones.
+Still, we have to consider that the preferences for non-meritocratic aspects in this case refer to the distributive principle of need rather than personal background (as in the perceptions of non-meritocracy), which certainly is one of the limitations of this set of items for the operationalization of the proposed conceptual framework with this dataset.
 
@@
 In terms of the association between the indicators, Figure \@ref(fig:corr-issp) shows the polychoric correlation matrix.
-Firstly, we see that the correlations between the indicators are consistent with the structure of the initially proposed structure of dimensions distinguishing between the perception and preferences of meritocratic and non-meritocratic factors.
-In general, the associations are positive between the perception and preference dimensions, such as the importance of hard work to get ahead and the importance of how hard the work is ($r$=0.34), however, they tend to be weak.
-A further aspect to consider is the relatively low correlation between dimensions concerning meritocratic and non-meritocratic aspects, (e.g.
-importance of hard work with educated parents, $r$=0.15).
-However, with respect to the non-meritocratic perception indicators, we can see that the associations demonstrate dimensionality, however, it is not entirely clear.
-In this regard, the indicators referring to parental characteristics have a positive and stronger association ($r$=0.56), but this association is much weaker in comparison to the other indicators of Non-meritocratic perception (e.g.
-Wealthy family and Gender, $r$=0.01).
-The same holds true for the Gender and Race indicators ($r$=0.6), and also applies to both the indicators of knowing the right people and having political connections ($r$=0.55).
-On this point, we will return to it in the next section.
+Firstly, we observe in general that the moderat to high associations are between the pairs of items representing each of the dimensions as described in \@ref(tab:table-issp).
+The exception here is the dimension of non-meritocratic perception, in which there are six items which appear mostly associated by pairs according to their specific topics (family background, personal background and connections).
+Still, between this same set of items there are moderate correlations (around 0.3), which could anticipate a single latent factor underlying non meritocratic perception and that is tested next through confirmatory factor analysis.
 
@@
 
-This section presents the results of the confirmatory factor analysis.
 First, a four-dimensional model for meritocratic and non-meritocratic perception and preferences is estimated, considering all indicators of the non-meritocratic perception dimension as a single latent variable.
@@
 
-Table \@ref(tab:sum-fit-issp) shows the four-factor model fit allowing comparison of first and second-order factors.
-The model restricted to four first-order factor dimensions shows a regular fit (CFI=0.959, TLI=0.944, RMSEA=0.098, χ2(df=48)= 21308.535; more detail in Table \@ref(tab:tb-perpref1)) in contrast to the model considering a second-order factor for non-meritocratic perception that has a slightly better fit based on the standard fit measures (CFI=0.959, TLI=0.944, RMSEA=0.098, χ2(df=48)= 21308.535).
-
-<!-- [diagrama SEM] -->
+Table \@ref(tab:sum-fit-issp) shows the results of the estimation of two confirmatory models.
+The first one (First Order) corresponds to a model that estimate four factors, each for one of the dimensions presented in \@ref(tab:desc-issp), showing only regular fit indicators (CFI=0.959, TLI=0.944, RMSEA=0.098, χ2(df=48)= 21308.535; more detail in Table \@ref(tab:tb-perpref1)).
+Further analysis showed that the sources for poor fit were mostly related to the items of the non-meritocratic perceptions, which as mentioned above displayed correlations among them not taken into account in this model specification.
+Attending to this constraints, we specified a second model that keeps the basic four-factor structure but generates an additional model for the non-meritocratic perceptions as depicted in \@ref(fig:meas02-issp).
+In this model, the pairs of items in this dimension form three latent factors which at the same time generates a second-order factor of non-meritocratic perceptions, improving the fit of the model.
 
@@
 ```
-
-<!-- [tabla CFA] -->
-
-Figure \@ref(fig:meas02-issp) shows the overall results for this model, evidencing an improvement in factor loadings as well as in the overall fit of the measurement model.
-On one hand, we see that perceptions are correlated with preferences, but this correlation is stronger for the meritocratic dimension of perception and preference ($r$=0.46) than for the non-meritocratic ($r$=0.27), which is much stronger when looking at the correlations between meritocratic and non-meritocratic, both in perception ($r$=0.04) and preference ($r$=0.13).
-These results show that, on one hand, the positive correlation between perception indicates that the two are not completely opposite poles of a continuum ($r$=0.30), which is reflected in the correlation between preferences ($r$=0.25).
-Furthermore, it is shown that the indicators of non-meritocratic perception refer to three other sub-dimensions and cannot be completely attributed to a an entirely common factor as has been employed in previous research.
+Regarding the correlations between the factors in the Figure \@ref(fig:meas02-issp), we observe that perceptions are correlated with preferences, but more strongly for the meritocratic ($r$=0.46) than for the non-meritocratic ($r$=0.27) dimensions.
+Secondly, both perceptual dimensions (meritocratic and non-meritocratic) depict a moderate correlation ($r$=0.30), suggesting that they are not the opposite poles of the same  continuum.
+Something similar occurs for the correlation between the two dimensions of preferences ($r$=0.25).
+From this information we can preliminary conclude that the wide use of reverse-coding for non-meritocratic items attempting to measure meritocracy is not an adequate operationalization of the construct.
+All in all, despite the good fit of the second order model there are still several limitations regarding the content of the items which we attempt to overcome next by developing a new instrument.
 
@@
 
-The data was obtained through an online questionnaire, which was part of a larger study on meritocracy and preferences developed in Chile in 2019 funded by the national scientific agency ANID.
+The data was obtained through an online survey which was part of a larger study on meritocracy and preferences developed in Chile in 2019, funded by the national scientific agency ANID.
 The questionnaire was programmed in Qualtrics and the fieldwork was conducted by an external online survey agency ([netquest.cl](www.netquest.cl)) between December 2019 and January 2020.
 The sample was selected from a non-probabilistic quota design in three large cities in Chile (Santiago, Concepción & Antofagasta).
-The quotas for gender, age, and education level were generated based on a survey by the Public Studies Center [@cepEncuestaCEPMayo2019], which is a well regarded counterpart agency of the ISSP (International Social Survey Programme) in Chile.
-A total sample of 2,141 people was collected, excluding those who did not answer the questions on the scale, and those who did not accept informed consent.
-There were no significant differences between our sample and the wider population for most socio-demographic characteristics, with the exception of educational level (see Table \@ref(tab:rep-samp) in appendix).
-As is often the case with online surveys, there were some limitations in achieving the quotas for lower educational levels **[citas - LUCHO]**.
+The quotas for gender, age, and educational levels were generated based on a survey by the Public Studies Center [@cepEncuestaCEPMayo2019], which is a well regarded counterpart agency of the ISSP (International Social Survey Programme) in Chile.
+A total sample of 2,141 people was collected, excluding those who did not signed the informed consent as well as those not answering the meritocracy response battery.
+There were no significant differences between our sample and the wider population for most socio-demographic characteristics, with the exception of educational level (see Table \@ref(tab:rep-samp) in Appendix).
+As it is often the case with online surveys, there were some limitations in achieving the quotas for lower educational levels **[citas - LUCHO]**.
 
@@
 These eight likert-type items have five response alternatives, ranging from "Completely disagree"(1) to "Completely agree" (5).
-
-```{r include=FALSE}
-
-# if (knitr::is_html_output()) {tb <- table_nums("tb",cap)} else {tb <- cap}
-```
 
@@
 As far as preferences are concerned, the preponderant role of effort over talent as a criterion of meritocratic preference is noteworthy.
+All in all the descriptive results tend to show a sort of critical view of meritocracy, perceiving the operation of non-meritocratic aspects over meritocratic ones, whereas in the preferences the  opposite occurs.
 
@@
 ```
-
 Attending now to the association among the scale items, Figure \@ref(fig:corpoly) shows items polychoric correlations.
@@
 Firstly, as expected the largest correlations are between indicators that correspond to the same factors behind the conceptual model (e.g., perception of meritocracy by effort and by talent, r=0.52).
-Secondly, among this correlations the highest are those between the non- meritocratic dimension buth in perceptions (r=0.73) and preferences (r=0.61).
-thirdly, both items for meritocratic preferences (E and F) are the ones that mostly correlate with the rest of the perceptual items, showing medium to high correlations.
-This is noteworthy because it indicates that even the perception of non-meritocracy would be related to larger meritocratic preferences.
-Finally, we observe that there are no considerable negative correlations between meritocratic and non-meritocratic aspects, undermining the assumptions of previous studies that suggested that these dimensions would be the opposite poles of one same continuum [@reynolds_perceptions_2014].
+Secondly, among this correlations the highest are those between the non-meritocratic dimension but in perceptions (r=0.73) and preferences (r=0.61).
+Thirdly, both items for meritocratic preferences (E and F) are the ones that mostly correlate with the rest of the perceptual items, showing medium to high correlations.
+This is noteworthy because it indicates that the perception of non-meritocracy would be related to larger meritocratic preferences.
+Finally, and similar to what was found in Study 1 with ISSP data, we observe that there are no considerable negative correlations between meritocratic and non-meritocratic aspects, undermining the assumptions of previous studies that suggested that these dimensions would be the opposite poles of one same continuum [@reynolds_perceptions_2014].
 
@@
     lower.col = "black")
-
 ```
@@
 
-The first step in the analysis consists of comparing the model fit indicators for the three versions of the scale that were randomly assigned to the participants: order according to perceptions/preferences, order according to topics and random order (see Figure \@ref(fig:appmod)).
+The first step in the analysis consists of comparing the model fit indicators for the three versions of the scale that were randomly assigned to the participants: order according to perceptions/preferences, order according to topics, and random order (see Figure \@ref(fig:appmod)).
 
@@
   load(file = "input/data/proc/data_s2.RData")
-names(data_s2)
+  # names(data_s2)
 
@@
   fit_cfa3 <- cfa(model_cfa, data = data_s2[data_s2$group == 3,], ordered = TRUE, std.lv=FALSE)
-
 ```
@@
 
-Attending to the correlations between the latent variables as depicted in Figure \@ref(fig:meas02), it is observed that meritocratic preferences are correlated with preferences, both meritocratic ($r=0.457$) and non-meritocratic ($r=0.500$).
+Attending to the correlations between the latent variables as depicted in Figure \@ref(fig:meas02), meritocratic preferences are moderate to highly correlated with perceptions, both meritocratic ($r=0.457$) and non-meritocratic ($r=0.500$).
 The correlation between both types of perceptions ($r=-0.044$) and both types preferences ($r=0.185$) are low, as well as between non-meritocratic preferences and perceptions ($r=-0.059$).
-This last finding gives further evidence regarding the lack of unidimensionality of meritocratic and non-meritocratic aspects: they do not appear as the opposite poles of the same continuum and based on this evidence the use of reverse-coding for non-meritocratic items as indicators of meritocracy should be avoided.
+This last finding gives further evidence regarding the lack of unidimensionality of meritocratic and non-meritocratic aspects as assumed by previous studies.
 
@@
 Firstly we analyzed the convergent validity by exploring the association of the merit scale with related concepts and its correspondent measurements.
-Secondly, we test the internal consistency of the measurement model behind the scale comparing with a different sample through measurement invariance procedures .
+Secondly, we test the internal consistency of the measurement model behind the scale comparing with a different sample through measurement invariance procedures.
 
@@
 
-For this analysis we examined data from one wave of the same study described above, which was a three-wave panel survey.
-The last wave of the study included additional measures that allow testing the correlations of the merit scale with some related constructs as opportunity beliefs and personal wherewithal.
+For this analysis we examined data from the last of the three waves of the online panel survey used in Study 2.
+This wave included additional measures that allow testing the correlations of the merit scale with some related constructs as opportunity beliefs and personal wherewithal.
 After listwise deletion of missing cases in socio-demographics information, a total of 1422 individuals took part in the third wave: 668 (46.97%) women and 754 (53.03%) men.
@@
 Participants responded to each item using a 1 (not important at all) to 5 (essential) response scale.
-We call the first item \textit{hard work} and the second one \textit{social origin}.
+We call the first item _hard work_ and the second one _social origin_.
 
@@
 
-We used polyserial correlations to evaluate relationships of meritocractic and non-meritocratic preferences and perceptions with the items of the opportunity beliefs battery.
+We used polychoric correlations to evaluate relationships of meritocractic and non-meritocratic preferences and perceptions with the items of the opportunity beliefs battery.
 For personal wherewithal, we examined Pearson correlations.
 
-Table \@ref(tab:val-conv) shows the polyserial and Pearson correlations between our meritocracy scale and common measures of meritocracy.
-The findings provide evidence for convergent validity for the scale.
+Table \@ref(tab:val-conv) shows the polychoric and Pearson correlations between our meritocracy scale and common measures of meritocracy.
 As we expected, preferences show very weak correlations.
 Findings also indicate that correlations of perceptions with opportunity beliefs are consistent with our expectations.
-The item for social origin shows a positive and moderate association with unmeritocratic perception (r=.354), and a negative and low correlation with meritocratic perception (-0.197).
-Regarding the item of hard work, the correlation with meritocratic perception is positive but weak (r=.187).
-One explanation for this result is that the support for hard work is very high among respondents---only 8.6 percent indicates _not very important_ or _not important at all_.
+The item for social origin shows a positive and moderate association with unmeritocratic perception (r=.361), and a negative and low correlation with meritocratic perception (-0.186).
+Regarding the item of hard work, the correlation with meritocratic perception is positive but weak (r=.201).
+One explanation for this result is that the support for hard work is very high among respondents and therefore with low variability, as only 8.6 percent indicates _not very important_ or _not important at all_.
 
@@
 The measure of NBI for _personal wherewithal_ also provides evidence for the convergent validity of the merit scale.
-Results indicate a positive correlation of this measure with meritocratic preference (r=.392), suggesting that both variables capture similar concepts.
-The NBI's factor depicts a positive association with unmeritocratic preference (r=.208).
+Results indicate a positive correlation of this measure with meritocratic perception (r=.406), suggesting that both variables capture similar concepts.
+The NBI's factor depicts a positive association with unmeritocratic preference (r=.225).
 This finding is consistent with literature, inasmuch as the belief for meritocray is associated with the justification of inequality [@McCall2017Exposurerisinginequality; @MadeiraPrimesConsequencesSystematic2019].
@@
 Table \@ref(tab:fcloads-inv) shows the results of the CFA estimation of the meritocracy scale in this new sample.
-As observed, the loadings and the fit indices show similar results as the ones presented for the original study above, which gives a first base with wich to start the invariance analysis.
+As observed, the loadings and the fit indices show similar results as the ones presented for the original study above, which gives a first base from which to start the invariance analysis.
 
 The first step for invariance testing is the estimation of the configural model, which serves as the baseline for further comparison and it is expected to adequately meet the global fit criteria of a measurement model.
-Although the chi-square statistic is used as a global measure of fit,since its sensitivity to sample size it is recommended to use three additional fit indices: Comparative Fit Index (CFI), which should have a value greater than 0.95; Root Mean Square Error Approximation (RMSEA), which must be in the range of values ​​of 0.05 and 0.08; and the Standardized Root Mean Square Residuals (SRMR) which must be less than 0.08 [@vandeschoot_checklist_2012].
+Although the chi-square statistic is used as a global measure of fit, since its sensitivity to sample size it is recommended to use three additional fit indices: Comparative Fit Index (CFI), which should have a value greater than 0.95; Root Mean Square Error Approximation (RMSEA), which must be lower than 0.06; and the Standardized Root Mean Square Residuals (SRMR) which must be less than 0.08 [@vandeschoot_checklist_2012].
 Besides, the literature on measurement invariance suggests some complementary approaches for the evaluation of the fit, from which we will consider the incremental adjustment of the fit indexes [@cheung_Evaluating_2002; @Milfont2010; @dimitrov_Testing_2010] and the ANOVA test for means comparison in nested models [@newsom_Longitudinal_2015].
@@
     footnote(general = footnote, footnote_as_chunk = T)
-
 ```
@@
 As the existence of heterogeneous approaches certainly has consequences for the advancement of the study of meritocracy, this paper presented a comprehensive conceptual framework for the empirical study of meritocracy, building upon previous research.
-This framework was then tested against new survey data.
+This framework was then tested against available and new survey data.
 
@@
 
-The four-dimensional framework and structure of the Perceptions and Preferences for Meritocracy scale opens several avenues for future research.
+The four-dimensional conceptual framework and its operationalization in the Perceptions and Preferences for Meritocracy scale opens several avenues for future research.
 For instance, distinguishing perceptions from preferences will allow us to evaluate the extent to which different societies are accustomed to, or satisfied with, the perceived level of meritocracy, in terms of differences between what is perceived and what is preferred.
@@
 The consideration of structural factors, both individual and societal, will allow us to advance future hypotheses relating to meritocratic legitimacy to a more comprehensive level than the current one, which is based solely on meritocratic perceptions.
-Furthermore, the impacts of different configurations of the four-dimensional framework on practices and behaviors such as corruption, civic involvement, and political alignment, is an area that requires additional research.
+Furthermore, the impact of different configurations of the four-dimensional framework on practices and behaviors such as corruption, civic involvement, and political alignment, is an area that requires additional research.
 Such future agendas could be specially relevant in times of economic crisis and growing inequalities, that could entail changes in the legitimation of the current distributive structure based on meritocratic ideals.
@@
                         font_size = 12) %>%
-          add_header_above(header = c(" "=1,"Perception"= 2,"Preference"= 2)) %>%
-          add_header_above(header = c(" "=1,"Factor loadings"= 4)) %>%
+          add_header_above(header = c(" "=2,"Perception"= 2,"Preference"= 2)) %>%
+          add_header_above(header = c(" "=2,"Factor loadings"= 4)) %>%
           row_spec(row = 12,hline_after = TRUE) %>%
@@
 ```{r fcloads-inv, echo=FALSE}
+library(dplyr,kable, kableExtra)
 table_format = if(knitr::is_html_output()) { #conditional instructions for kable
@@
                         bootstrap_options = c("striped", "bordered")) %>%
-          add_header_above(header = c(" "=1,"Perception"= 2,"Preference"= 2)) %>%
-          add_header_above(header = c(" "=1,"Factor loadings"= 4)) %>%
+          add_header_above(header = c(" "=2,"Perception"= 2,"Preference"= 2)) %>%
+          add_header_above(header = c(" "=2,"Factor loadings"= 4)) %>%
           row_spec(row = 8,hline_after = TRUE) %>%

### 2021-09-07 paper.Rmd
fix-indent
- indent para chunks en atom

@@
 ```{r meas02, echo=FALSE, fig.cap = "Confirmatory factor analysis of the Perceptions and Preferences for Meritocracy Scale", fig.align='center'}
-
  # estimation for diagram: summary(fit_cfa3, standardized = TRUE, fit.measures=TRUE)

### 2021-09-08 paper.Rmd
render-pdf
- ajustes posición de tablas
- ajuste figuras tamaño y align
- indent code folding para Atom
- test render pdf, html y word

@@
 Taking into account the extant literature that uses measures of meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding conceptualization and operationalization.
-Based on this critical analysis, we propose a conceptual framework for studying perceptions and preferences for meritocracy and non-meritocracy, which is then tested through confirmatory analysis using ISSP (International Social Survey Programme) data as well as a novel survey designed specifically with this purpose(N=2,141).
+Based on this critical analysis, we propose a conceptual framework for studying perceptions and preferences for meritocracy and non-meritocracy, which is then tested through confirmatory analysis using ISSP (International Social Survey Programme) data as well as a novel survey designed specifically with this purpose (N=2,141).
 Our results support the conceptual framework and its operationalizarion, although with a better fit for the proposed scale than for the ISSP survey.
@@
 output:
-#  bookdown::pdf_document2:
-#    template: null
-#    toc: false
+  bookdown::pdf_document2:
+     template: null
+     toc: false
+     number_sections: false 
   bookdown::html_document2:
+    number_sections: false
+  bookdown::word_document2:
     number_sections: false  
@@
   - \usepackage{threeparttable}
+  - \usepackage{titlesec}
+  # - \setcounter{secnumdepth}{4} # activate for lvl 4 heading numbering
+  - \titleformat{\paragraph}
+    {\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}
+    \titlespacing*{\paragraph}
+    {0pt}{3.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}
+always_allow_html: true #for word output    
 ---
@@
  rmarkdown::render("paper.Rmd", output_format = "bookdown::html_document2")
+ rmarkdown::render("paper.Rmd", output_format = "bookdown::word_document2")
 ```
@@
 ```{r setup, include=FALSE}
- knitr::opts_chunk$set(echo=FALSE, warning = FALSE,message = FALSE, cache = FALSE,out.width = '85%',fig.pos= "H")
+ knitr::opts_chunk$set(echo=FALSE, warning = FALSE,message = FALSE, cache = FALSE,out.width = '85%',fig.pos= "H", fig.align = 'center')
  # knitr::opts_knit$set(base.url = "../") #relative path for .html output file
@@
   }
-
 ```
@@
 ```{r table-issp, echo=FALSE}
+  if (!require("pacman")) install.packages("pacman")
+  # remotes::install_github('haozhu233/kableExtra@a6af5c0')  # for problems with collapse_rows()
   pacman::p_load(knitr, kableExtra, dplyr)
@@
       position = "center",
-      font_size = 20,
+      font_size = 10,
       bootstrap_options=c("striped", "bordered")) %>%
@@
     kableExtra::column_spec(column = 2,width = "5cm") %>%
-    kableExtra::column_spec(column = 3,width = "12cm") %>%
+    kableExtra::column_spec(column = 3,width = "8cm") %>%
     kableExtra::collapse_rows(columns = 1:2,valign = "middle")
@@
   load(file = "input/data/proc/data_s1.RData")
-  pacman::p_load(dplyr)
+  if (!require("pacman")) install.packages("pacman")
+  pacman::p_load(dplyr, psych, knitr, kableExtra)
   table_format = if(knitr::is_html_output()) { #conditional instructions for kable
@@
     "latex"#if latex set "latex" in format
-  }
-
-  fsize = if(knitr::is_html_output()) { #conditional instructions for kable
-    20 # if html set table fontsize to 20
-  } else if(knitr::is_latex_output()) {
-    12 # if latex set table fontsize to 12
   }
@@
     kableExtra::kable_styling(full_width = T,
-                              font_size = 20,
+                              font_size = 10,
                               position = "center",
@@
   load(file = "input/data/proc/data_s1.RData")
-  pacman::p_load(dplyr)
+  if (!require("pacman")) install.packages("pacman")
+  pacman::p_load(dplyr,sjPlot, ggplot2, sjlabelled,sjmisc, ggpubr)
   likert_issp<- data_s1 %>%
@@
     sjPlot::plot_likert(geom.colors = "PuBu",
+                        grid.range  =  c (1.2 , 1.2),                        
                         title = "Perception meritocratic",
@@
     sjPlot::plot_likert(geom.colors = "PuBu",
+                        grid.range  =  c (1.2 , 1.2),                        
                         title = "Perception Non-meritocratic",
@@
     sjPlot::plot_likert(geom.colors = "PuBu",
+                        grid.range  =  c (1.2 , 1.2),                        
                         title = "Preference meritocratic",
@@
     sjPlot::plot_likert(geom.colors = "PuBu",
+                        grid.range  =  c (1.2 , 1.2),                        
                         title = "Preference Non-meritocratic",
@@
 
-```{r likert-issp,fig.height=11,fig.width=11,fig.cap="Distribution of responses in the ISSP meritocracy items", out.height='80%'}
+```{r likert-issp,fig.height=11,fig.width=11,fig.cap="Distribution of responses in the ISSP meritocracy items", out.width='80%'}
+  if (!require("pacman")) install.packages("pacman")
+  pacman::p_load(cowplot, dplyr)
   #Final plot
@@
 
-```{r corr-issp,fig.cap = "Perceptions & preferences for ISSP  meritocracy items' polychoric correlations", fig.align='center',fig.height=6,fig.width=8}
+```{r corr-issp,fig.cap = "Perceptions and preferences for ISSP  meritocracy items' polychoric correlations", fig.align='center',fig.height=6,fig.width=8}
   load(file = "input/data/proc/data_s1.RData")
-  pacman::p_load(dplyr)
+  if (!require("pacman")) install.packages("pacman")
+  pacman::p_load(dplyr, corrplot, lavaan)
   # create polychoric correlation matrix
@@
     bg = "white",
-    na.label = "-",
-    lower.col = "black")
+    na.label = "-")
 ```
@@
   load(file = "input/data/proc/data_s1.RData")
-  pacman::p_load(dplyr, kableExtra, knitr)
+  if (!require("pacman")) install.packages("pacman")
+  pacman::p_load(dplyr, kableExtra, knitr,lavaan)
   table_format = if(knitr::is_html_output()) { #conditional instructions for kable
@@
 
-  fsize = if(knitr::is_html_output()) { #conditional instructions for kable
-    20 # if html set table fontsize to 20
-  } else if(knitr::is_latex_output()) {
-    12 # if latex set table fontsize to 12
-  }
   # meassurement model: First order
@@
                  escape = FALSE) %>%
-    kableExtra::kable_styling(font_size = fsize,
+    kableExtra::kable_styling(font_size = 10,
                               full_width = F,
+                              latex_options = "HOLD_position",
                               bootstrap_options=c("striped", "bordered"))
@@
 Table \@ref(tab:sum-fit-issp) shows the results of the estimation of two confirmatory models.
-The first one (First Order) corresponds to a model that estimate four factors, each for one of the dimensions presented in \@ref(tab:desc-issp), showing only regular fit indicators (CFI=0.959, TLI=0.944, RMSEA=0.098, χ2(df=48)= 21308.535; more detail in Table \@ref(tab:tb-perpref1)).
+The first one (First Order) corresponds to a model that estimate four factors, each for one of the dimensions presented in \@ref(tab:table-issp), showing only regular fit indicators (CFI=0.959, TLI=0.944, RMSEA=0.098, $\chi^2$(df=48)= 21308.535; more detail in Table \@ref(tab:tb-perpref1)).
 Further analysis showed that the sources for poor fit were mostly related to the items of the non-meritocratic perceptions, which as mentioned above displayed correlations among them not taken into account in this model specification.
@@
 ```
-Regarding the correlations between the factors in the Figure \@ref(fig:meas02-issp), we observe that perceptions are correlated with preferences, but more strongly for the meritocratic ($r$=0.46) than for the non-meritocratic ($r$=0.27) dimensions.
-Secondly, both perceptual dimensions (meritocratic and non-meritocratic) depict a moderate correlation ($r$=0.30), suggesting that they are not the opposite poles of the same  continuum.
-Something similar occurs for the correlation between the two dimensions of preferences ($r$=0.25).
+Regarding the correlations between the factors in the Figure \@ref(fig:meas02-issp), we observe that perceptions are correlated with preferences, but more strongly for the meritocratic ($r=0.46$) than for the non-meritocratic ($r=0.27$) dimensions.
+Secondly, both perceptual dimensions (meritocratic and non-meritocratic) depict a moderate correlation ($r=0.30$), suggesting that they are not the opposite poles of the same  continuum.
+Something similar occurs for the correlation between the two dimensions of preferences ($r=0.25$).
 From this information we can preliminary conclude that the wide use of reverse-coding for non-meritocratic items attempting to measure meritocracy is not an adequate operationalization of the construct.
@@
 The data was obtained through an online survey which was part of a larger study on meritocracy and preferences developed in Chile in 2019, funded by the national scientific agency ANID.
-The questionnaire was programmed in Qualtrics and the fieldwork was conducted by an external online survey agency ([netquest.cl](www.netquest.cl)) between December 2019 and January 2020.
+The questionnaire was programmed in Qualtrics and the fieldwork was conducted by an external online survey agency ([netquest.com](www.netquest.com)) between December 2019 and January 2020.
 The sample was selected from a non-probabilistic quota design in three large cities in Chile (Santiago, Concepción & Antofagasta).
@@
 ```{r table-indicadores, echo=FALSE}
+ if (!require("pacman")) install.packages("pacman")
+ # remotes::install_github('haozhu233/kableExtra@a6af5c0')  # for problems with collapse_rows()
  pacman::p_load(knitr, kableExtra, dplyr)
-
+ table_format = if (knitr::is_html_output()) {
+   #conditional instructions for kable
+   "html" #if html set "html" in format
+ } else if (knitr::is_latex_output()) {
+   "latex"#if latex set "latex" in format
+ }
  tabitems <- read.csv(file = "input/tables/table01.csv",header = 1,sep = ",",encoding = "UTF-8") # call file  generated externally for the table
@@
 
- kable(tabitems, table_format, booktabs = T, linesep = "",col.names = cnames, caption = cap) %>%
-   kable_styling(
+ knitr::kable(tabitems, table_format, booktabs = T, linesep = "",col.names = cnames, caption = cap) %>%
+   kableExtra::kable_styling(
      full_width = T,
@@
      position = "center",
-     font_size = 12,
+     font_size = 10,
      bootstrap_options=c("striped", "bordered")) %>%
-   column_spec(column = 1, width = "1.5cm", ) %>%
-   column_spec(column = 2,width = "2 cm") %>%
-   collapse_rows(columns = 1:2,valign = "middle")
+   kableExtra::column_spec(column = 1, width = "1.5cm", ) %>%
+   kableExtra::column_spec(column = 2,width = "2 cm") %>%
+   kableExtra::collapse_rows(columns = 1:2,valign = "middle")
 ```
@@
 ```{r produce-likert_s2, include=FALSE}
-  pacman::p_load(dplyr)
+  if (!require("pacman")) install.packages("pacman")
+  pacman::p_load(dplyr,sjPlot, ggplot2, sjlabelled,sjmisc, ggpubr)
   load(file = "input/data/proc/data_s2.RData")
   names(data_s2)
-  dat_likert_s2 <- data_s2 %>% select(!group) # exclude group variable
+  dat_likert_s2 <- data_s2 %>% dplyr::select(!group) # exclude group variable
   dat_likert_s2 <- sjmisc::rec(dat_likert_s2, rec="rev", append=FALSE)
@@
   # update.packages("sjPlot") # please install version 2.8.9 or above
-  # plotlikert_s2<- sjPlot::plot_likert(dat_likert_s2,
-  #                        c(1, 1, 1, 1, 2, 2, 2, 2),
-  #                        groups.titles = c("Perceptions", "Preferences"),
-  #                        geom.colors = "PuBu",
-  #                        geom.size = 0.8,
-  #                        axis.labels = c("Effort", "Talent", "Rich Parents", "Contacts"),
-  #                        catcount = 4,
-  #                        cat.neutral = 3,
-  #                        grid.range  =  c (1.2 , 1.4),
-  #                        values  =  "sum.outside",
-  #                        reverse.colors = T,
-  #                        reverse.scale = F,
-  #                        show.n=FALSE)
   ggplot2::theme_set(ggplot2::theme(panel.background = ggplot2::element_rect(fill = "gray85",
@@
                         cat.neutral = 3,
-                        grid.range  =  c (1.2 , 1.4),
+                        grid.range  =  c (1.2 , 1.2),
                         values  =  "sum.outside",
@@
 ```{r plotlikert, echo=FALSE, fig.align='center', fig.cap="Distribution of responses in the Merit Scale items", fig.height=8, fig.width=11}
+  if (!require("pacman")) install.packages("pacman")
+  pacman::p_load(cowplot)
   #Final plot
@@
       ),
-    label = paste0("Source: Authors calculations based on Study 1 sample"," (n=",dim(na.omit(dat_likert_s2))[1],")"
+    label = paste0("Source: Authors calculations based on Study 2 sample"," (n=",dim(na.omit(dat_likert_s2))[1],")"
       ),        # Caption: source and sample size
@@
 There are three main aspects to highlight from this correlation matrix.
-Firstly, as expected the largest correlations are between indicators that correspond to the same factors behind the conceptual model (e.g., perception of meritocracy by effort and by talent, r=0.52).
-Secondly, among this correlations the highest are those between the non-meritocratic dimension but in perceptions (r=0.73) and preferences (r=0.61).
+Firstly, as expected the largest correlations are between indicators that correspond to the same factors behind the conceptual model (e.g., perception of meritocracy by effort and by talent, $r$=0.52).
+Secondly, among this correlations the highest are those between the non-meritocratic dimension but in perceptions ($r$=0.73) and preferences ($r$=0.61).
 Thirdly, both items for meritocratic preferences (E and F) are the ones that mostly correlate with the rest of the perceptual items, showing medium to high correlations.
@@
 
-```{r corpoly, echo=FALSE, fig.cap = "Perceptions & preferences for meritcoracy items' polychoric correlations", fig.align='center'}
+```{r corpoly, echo=FALSE, fig.cap = "Perceptions and preferences for meritcoracy items' polychoric correlations", fig.align='center'}
+  if (!require("pacman")) install.packages("pacman")
   load(file = "input/data/proc/data_s2.RData")
-
+  pacman::p_load(dplyr, corrplot, lavaan)
   # generate polychoric matrix
-  cor_s2<- data_s2 %>% select(perc_effort,perc_talent,perc_wpart,perc_netw,pref_effort,pref_talent,pref_wpart,pref_netw) %>% lavaan::lavCor(., ordered=names(.))
+  cor_s2<- data_s2 %>% 
+    dplyr::select(perc_effort,perc_talent,perc_wpart,perc_netw,pref_effort,pref_talent,pref_wpart,pref_netw) %>% 
+    lavaan::lavCor(., ordered=names(.))
 
   diag(cor_s2) = NA
-
-  # windowsFonts(A = windowsFont("Times New Roman"))
 
@@
   # Libraries & data
+  if (!require("pacman")) install.packages("pacman")  
   pacman::p_load(lavaan)
@@
   # estimation for each order set
-  fit_cfa1 <- cfa(model_cfa, data = data_s2[data_s2$group == 1,], ordered = TRUE, std.lv=FALSE)
-  fit_cfa2 <- cfa(model_cfa, data = data_s2[data_s2$group == 2,], ordered = TRUE, std.lv=FALSE)
-  fit_cfa3 <- cfa(model_cfa, data = data_s2[data_s2$group == 3,], ordered = TRUE, std.lv=FALSE)
+  fit_cfa1 <- lavaan::cfa(model_cfa, data = data_s2[data_s2$group == 1,], ordered = TRUE, std.lv=FALSE)
+  fit_cfa2 <- lavaan::cfa(model_cfa, data = data_s2[data_s2$group == 2,], ordered = TRUE, std.lv=FALSE)
+  fit_cfa3 <- lavaan::cfa(model_cfa, data = data_s2[data_s2$group == 3,], ordered = TRUE, std.lv=FALSE)
 ```
@@
 ```{r table-cfafits, echo=FALSE}
+  if (!require("pacman")) install.packages("pacman")
   pacman::p_load(dplyr, kableExtra, knitr)
-
+  table_format = if(knitr::is_html_output()) { #conditional instructions for kable
+     "html" #if html set "html" in format
+   } else if(knitr::is_latex_output()) {
+     "latex"#if latex set "latex" in format
+   }
   # extract fit indices from models and add to table
-  sum_fit<- dplyr::bind_rows(fitmeasures(fit_cfa1)[c("chisq","df","cfi","tli","rmsea")],
-                      fitmeasures(fit_cfa2)[c("chisq","df","cfi","tli","rmsea")],
-                      fitmeasures(fit_cfa3)[c("chisq","df","cfi","tli","rmsea")])
+  sum_fit<- dplyr::bind_rows(lavaan::fitmeasures(fit_cfa1)[c("chisq","df","cfi","tli","rmsea")],
+                             lavaan::fitmeasures(fit_cfa2)[c("chisq","df","cfi","tli","rmsea")],
+                             lavaan::fitmeasures(fit_cfa3)[c("chisq","df","cfi","tli","rmsea")])
 
@@
   # Create table
-  table_cfa_fits <-kable(sum_fit, format="html", digits=3, booktabs=T, col.names=colnames,  caption = "Summary fit indices according to order versions", escape = FALSE) %>% kable_styling(full_width = F, font_size = 12,bootstrap_options=c("striped", "bordered"))
+  table_cfa_fits <-kable(sum_fit, format=table_format, 
+                         digits=3, 
+                         booktabs=T, 
+                         col.names=colnames,  
+                         caption = "Summary fit indices according to order versions", 
+                         escape = FALSE) %>%
+    kable_styling(full_width = F, 
+                  font_size = 10,
+                  latex_options = "HOLD_position",
+                  bootstrap_options=c("striped", "bordered"))
   table_cfa_fits
@@
 Findings also indicate that correlations of perceptions with opportunity beliefs are consistent with our expectations.
-The item for social origin shows a positive and moderate association with unmeritocratic perception (r=.361), and a negative and low correlation with meritocratic perception (-0.186).
-Regarding the item of hard work, the correlation with meritocratic perception is positive but weak (r=.201).
+The item for social origin shows a positive and moderate association with unmeritocratic perception ($r=0.361$), and a negative and low correlation with meritocratic perception ($r=-0.186$).
+Regarding the item of hard work, the correlation with meritocratic perception is positive but weak ($r=.201$).
 One explanation for this result is that the support for hard work is very high among respondents and therefore with low variability, as only 8.6 percent indicates _not very important_ or _not important at all_.
@@
 ```{r val-conv}
+  if (!require("pacman")) install.packages("pacman")
   pacman::p_load(kableExtra, polycor)
-
+  table_format = if(knitr::is_html_output()) { #conditional instructions for kable
+     "html" #if html set "html" in format
+   } else if(knitr::is_latex_output()) {
+     "latex"#if latex set "latex" in format
+   }
   load(file="input/data/proc/data_s3_conv.RData")
@@
 
-  variables= c("Meritocratic Perception", "Meritocratic Preferences", "Unmeritocratic Perception","Unmeritocratic Preferences")
+  variables= c("Meritocratic Perception", "Meritocratic Preferences","Non-meritocratic Perception","Non-meritocratic Preferences")
 
@@
 
-  kable(table01, col.names = c("Merit-Scale","Social Origin ", "Hard Work " , "Wherewithal"), full_width = T, linesep = "", caption = c("Polyserial and Pearson correlation whit other merit scales")) %>% kable_styling(
-      full_width = T,
+  knitr::kable(table01, 
+        col.names = c("Merit-Scale","Social Origin ", "Hard Work " , "Wherewithal"), 
+        linesep = "", 
+        format = table_format,
+        booktabs=T,
+        digits = 3,
+        caption = c("Polyserial and Pearson correlation whit other merit scales")) %>% 
+    kableExtra::kable_styling(
+      full_width = F,
       position = "center",
-      font_size = 14,
-      bootstrap_options=c("striped", "bordered")) %>%
-    column_spec(column = 1, width = "1.5cm", )
+      font_size = 10,
+      bootstrap_options = c("striped", "bordered")) %>%
+    kableExtra::column_spec(column = 1, width = "4 cm")
 ```
@@
 The measure of NBI for _personal wherewithal_ also provides evidence for the convergent validity of the merit scale.
-Results indicate a positive correlation of this measure with meritocratic perception (r=.406), suggesting that both variables capture similar concepts.
-The NBI's factor depicts a positive association with unmeritocratic preference (r=.225).
+Results indicate a positive correlation of this measure with meritocratic perception ($r=0.406$), suggesting that both variables capture similar concepts.
+The NBI's factor depicts a positive association with unmeritocratic preference ($r=0.225$).
 This finding is consistent with literature, inasmuch as the belief for meritocray is associated with the justification of inequality [@McCall2017Exposurerisinginequality; @MadeiraPrimesConsequencesSystematic2019].
@@
 ```{r invargroup, echo=FALSE}
+  if (!require("pacman")) install.packages("pacman")
   #---Load packages and data -------------------------------------#
-  pacman::p_load(sjPlot,dplyr,lavaan,semPlot,semTools,stargazer,
-                 corrplot,psych,knitr,kableExtra,stats)
+  pacman::p_load(dplyr,lavaan,semTools,knitr,kableExtra,stats,gtools)
+  options(knitr.kable.NA = '')
   load(file = "input/data/proc/data_s3_inv.RData")
-  pov01 <- data_s3_inv %>% filter(dataset=="pvw01")
-  fs01  <- data_s3_inv %>% filter(dataset=="fsw02")
+  pov01 <- data_s3_inv %>% dplyr::filter(dataset=="pvw01")
+  fs01  <- data_s3_inv %>% dplyr::filter(dataset=="fsw02")
 
@@
   #-----Invariance---------------------------------------------#
-  inv01<- measurementInvariance(model=model01,data=data_s3_inv,group="dataset",estimator = "ML",strict=TRUE,quiet = T)
+  inv01<- semTools::measurementInvariance(model=model01,data=data_s3_inv,group="dataset",estimator = "ML",strict=TRUE,quiet = T)
 
@@
 
-  tab01<- anova(conf,weak,strong,strict,SB.classic=TRUE) %>%
-    as_tibble() %>%
-    select("Chisq","Df","chisq_diff"=`Chisq diff`,"df_diff"=`Df diff`,"pvalue"=`Pr(>Chisq)`) %>%
-    mutate(stars=gtools::stars.pval(pvalue),
+  tab01<- lavaan::anova(conf,weak,strong,strict,SB.classic=TRUE) %>%
+    dplyr::as_tibble() %>%
+    dplyr::select("Chisq","Df","chisq_diff"=`Chisq diff`,"df_diff"=`Df diff`,"pvalue"=`Pr(>Chisq)`) %>%
+    dplyr::mutate(stars=gtools::stars.pval(pvalue),
            chisqt=paste0(round(Chisq,2)," (",Df,") "),
@@
 
-  fit.meas<- bind_rows(fitmeasures(inv01$fit.configural,output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
-                       fitmeasures(inv01$fit.loadings,  output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
-                       fitmeasures(inv01$fit.intercepts,output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
-                       fitmeasures(inv01$fit.residuals, output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),])
+  fit.meas<- dplyr::bind_rows(lavaan::fitmeasures(inv01$fit.configural,output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
+                              lavaan::fitmeasures(inv01$fit.loadings,  output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
+                              lavaan::fitmeasures(inv01$fit.intercepts,output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
+                              lavaan::fitmeasures(inv01$fit.residuals, output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),])
 
@@
   fit.meas<- fit.meas %>%
-    mutate(diff.chi2 = chisq    - lag(chisq,default = first(chisq)),
-           diff.df   = df       - lag(df,   default = first(df)),
-           diff.cfi  = cfi      - lag(cfi,  default = first(cfi)),
-           diff.rmsea   = rmsea - lag(rmsea,default = first(rmsea))) %>%
+    dplyr::mutate(diff.chi2 = chisq    - lag(chisq,default = dplyr::first(chisq)),
+           diff.df   = df       - lag(df,   default = dplyr::first(df)),
+           diff.cfi  = cfi      - lag(cfi,  default = dplyr::first(cfi)),
+           diff.rmsea   = rmsea - lag(rmsea,default = dplyr::first(rmsea))) %>%
     round(3) %>%
-    mutate(rmsea.ci=paste0(rmsea," \n ", "(",rmsea.ci.lower,"-",rmsea.ci.upper,")"))
-
-  tab.inv<- bind_cols(tab01,fit.meas) %>%
-    select(model,chisqt,cfi,rmsea.ci,diff.chi2,diff.df,diff.cfi,diff.rmsea,stars,decision) %>%
-    mutate(diff.chi2=paste0(diff.chi2," (",diff.df,") ",stars)) %>%
-    select(model,chisqt,cfi,rmsea.ci,diff.chi2,diff.cfi,diff.rmsea,decision)
+    dplyr::mutate(rmsea.ci=paste0(rmsea," \n ", "(",rmsea.ci.lower,"-",rmsea.ci.upper,")"))
+
+  tab.inv<- dplyr::bind_cols(tab01,fit.meas) %>%
+    dplyr::select(model,chisqt,cfi,rmsea.ci,diff.chi2,diff.df,diff.cfi,diff.rmsea,stars,decision) %>%
+    dplyr::mutate(diff.chi2=paste0(diff.chi2," (",diff.df,") ",stars)) %>%
+    dplyr::select(model,chisqt,cfi,rmsea.ci,diff.chi2,diff.cfi,diff.rmsea,decision)
 
@@
   ### Check an alternative method
-  configural        <- cfa(model01, data=data_s3_inv, group = "dataset", estimator="ML")
-  weak.invariance   <- cfa(model01, data=data_s3_inv, group = "dataset",group.equal = "loadings", estimator="ML")
-  strong.invariance <- cfa(model01, data=data_s3_inv, group = "dataset",group.equal = c( "loadings", "intercepts"), estimator="ML")
-  strict.invariance <- cfa(model01, data=data_s3_inv, group = "dataset",group.equal = c( "loadings", "intercepts", "residuals"), estimator="ML")
+  configural        <- lavaan::cfa(model01, data=data_s3_inv, group = "dataset", estimator="ML")
+  weak.invariance   <- lavaan::cfa(model01, data=data_s3_inv, group = "dataset",group.equal = "loadings", estimator="ML")
+  strong.invariance <- lavaan::cfa(model01, data=data_s3_inv, group = "dataset",group.equal = c( "loadings", "intercepts"), estimator="ML")
+  strict.invariance <- lavaan::cfa(model01, data=data_s3_inv, group = "dataset",group.equal = c( "loadings", "intercepts", "residuals"), estimator="ML")
 
@@
                "$\\Delta \\chi^2 (\\Delta \\text{df}$)","$\\Delta \\text{CFI}$","$\\Delta \\text{RMSEA}$","Decision")
-  footnote <- paste0("N = ","; Group 1, n = ",conf@Data@nobs[[1]],"; Group 2, n = ",conf@Data@nobs[[2]], ", **p < 0.001")
-
-  kable(tab.inv, col.names = col.nam, align = "l",
+  footnote <- paste0("N = ","; Group 1, n = ",conf@Data@nobs[[1]],"; Group 2, n = ",conf@Data@nobs[[2]], ", ***p < 0.001")
+
+  knitr::kable(tab.inv, col.names = col.nam, align = "l",
         booktabs=TRUE,format = table_format,escape = FALSE,
         caption = "Multiple Group meassurement invariance for Perceptions and Preferences for Meritocracy") %>%
-    kable_styling(full_width = FALSE,latex_options = "HOLD_position",bootstrap_options=c("striped", "bordered")) %>%
-    footnote(general = footnote, footnote_as_chunk = T)
+    kableExtra::kable_styling(full_width = FALSE,latex_options = "HOLD_position",bootstrap_options=c("striped", "bordered")) %>%
+    kableExtra::footnote(general = footnote, footnote_as_chunk = T)
 ```
@@
 <div id="refs"> </div>  <!-- Although <div> is an HTML tag, this method also works for other output formats such as PDF. -->
+
+\pagebreak
 
@@
 
+\pagebreak
+
 ```{r tb-perpref1, echo=FALSE}
+ if (!require("pacman")) install.packages("pacman")
  load(file = "input/data/proc/data_s1.RData")
@@
    "latex"#if latex set "latex" in format
- }
-
- fsize = if(knitr::is_html_output()) { #conditional instructions for kable
-   20 # if html set table fontsize to 20
- } else if(knitr::is_latex_output()) {
-   12 # if latex set table fontsize to 12
  }
@@
  fm.df$nobs <- c(lavaan::nobs(fit_perpref1))   # add number of observations
- fm.df <- data.frame(t(fm.df)); colnames(fm.df) <- c("percnmerit") #transpose and set colnames
+ fm.df <- data.frame(t(fm.df)); colnames(fm.df) <- c("percmerit") #transpose and set colnames
 
@@
                  linesep = "",
-                 format = table_format) %>%
+                 format = table_format,
+                 row.names=F) %>%
            kable_styling(latex_options = c("hold_position"),
@@
                          position = "center",
-                         font_size = 12) %>%
-           add_header_above(header = c(" "=2,"Perception"= 2,"Preference"= 2)) %>%
-           add_header_above(header = c(" "=2,"Factor loadings"= 4)) %>%
+                         font_size = 10) %>%
+           add_header_above(header = c(" "=1,"Perception"= 2,"Preference"= 2)) %>%
+           add_header_above(header = c(" "=1,"Factor loadings"= 4)) %>%
            row_spec(row = 12,hline_after = TRUE) %>%
@@
 ```{r fcloads-inv, echo=FALSE}
- library(dplyr,kable, kableExtra)
+ if (!require("pacman")) install.packages("pacman")
+ pacman::p_load(dplyr, knitr, kableExtra,lavaan)
+ options(knitr.kable.NA = '')
  table_format = if(knitr::is_html_output()) { #conditional instructions for kable
@@
  # Extract information for table
- fm01<- data.frame(t(data.frame(lavaan::fitmeasures(fs1_o, output ="matrix")[c("chisq.scaled","df","cfi.scaled","rmsea.scaled"),])));  row.names(fm01) ="M1.1"
+ fm01<- data.frame(t(data.frame(lavaan::fitmeasures(fs1_o, output ="matrix")[c("chisq","df","cfi","tli","rmsea"),])));  row.names(fm01) ="M1.1"
 
  #------chi2, df------#
- fm04.1 <- fm01 %>% dplyr::select(chisq.scaled,df)
- fm04.1$chisq.scaled <- round(x = fm04.1$chisq.scaled,digits = 1)
+ fm04.1 <- fm01 %>% dplyr::select(chisq,df) 
+ fm04.1$chisq <- round(x = fm04.1$chisq,digits = 1)
  fm04.1$df <- round(x = fm04.1$df,digits = 0)
- fm04.1$chi2df <- paste0(fm04.1$chisq.scaled,"(",fm04.1$df,")")
+ fm04.1$chi2df <- paste0(fm04.1$chisq,"(",fm04.1$df,")")
  fm04.1 <- dplyr::select(fm04.1,"chi2df")
@@
  #------CFI, RMSEA------#
- fm04.2 <- fm01 %>% dplyr::select(cfi.scaled,rmsea.scaled)
+ fm04.2 <- fm01 %>% dplyr::select(cfi,tli,rmsea)
  for (i in names(fm04.2)) {
@@
  #------ extract factor loadings ------#
- loads01<- data.frame(round(lavaan::inspect(fs1_o,what="std")$lambda,digits = 3))
+ loads01<- data.frame(round(inspect(fs1_o,what="std")$lambda,digits = 3))
  loads01[loads01==c(0.00)] <- NA; names(loads01) <- c(paste0("M1.",1:4))
@@
  for (i in names(loads01)) {
-   # tb.load[,i] <- sjlabelled::as_character(tb.load[,i])
    loads01[,i] <- sprintf(loads01[,i], fmt = '%#.3f')
@@
  #------ merge ------#
- tb.fm2<- bind_rows(loads01,fm.df)
+ tb.fm2<- dplyr::bind_rows(loads01,fm.df)
  tb.fm2<- tb.fm2 %>% 
-   mutate(vars=c(labs,"$\\chi^2\\text{(df)}$","$\\text{CFI}$","$\\text{RMSEA}$","$N$")) %>%
-   select(vars,everything())
- # load(file = "output/tables/tb.fact-f02.RData")
+   dplyr::mutate(vars=c(labs,"$\\chi^2\\text{(df)}$","$\\text{CFI}$","$\\text{TLI}$","$\\text{RMSEA}$","$N$")) %>%
+   dplyr::select(vars,everything())
+
  tb.col  <- c(" ",
@@
  tb.caption <- "Factor loadings and fit measures for the second sample for study 3" 
- tb.foot <- paste0("Standardized factor loadings using DWLS estimator ; CFI = Comparative fit index (scaled); RMSEA = Root mean square  error of approximation (scaled)")
- tb.fm02<- kable(tb.fm2,escape = FALSE,align = "lcccc",col.names = tb.col, caption = tb.caption, booktabs = T,linesep = "",format =  table_format) %>%
-           kable_styling(full_width = table_format2,
+ tb.foot <- paste0("Standardized factor loadings using DWLS estimator ; CFI = Comparative fit index ;TLI = Tucker-Lewis index; RMSEA =  Root mean square error of approximation")
+ tb.fm02<- knitr::kable(tb.fm2,escape = FALSE,
+                        align = "lcccc",col.names = tb.col, 
+                        caption = tb.caption, 
+                        booktabs = T,linesep = "",
+                        format = table_format,
+                        row.names = F) %>%
+           kableExtra::kable_styling(full_width = table_format2,
                          latex_options = c("hold_position"), 
                          position = "center",
-                         font_size = 12,
+                         font_size = 10,
                          bootstrap_options = c("striped", "bordered")) %>%
-           add_header_above(header = c(" "=2,"Perception"= 2,"Preference"= 2)) %>%
-           add_header_above(header = c(" "=2,"Factor loadings"= 4)) %>%
-           row_spec(row = 8,hline_after = TRUE) %>%
-           add_indent(c(9:12)) %>%
-           footnote(general =tb.foot ,footnote_as_chunk = T);tb.fm02
-```
+           kableExtra::add_header_above(header = c(" "=1,"Perception"= 2,"Preference"= 2)) %>%
+           kableExtra::add_header_above(header = c(" "=1,"Factor loadings"= 4)) %>%
+           kableExtra::row_spec(row = 8,hline_after = TRUE) %>%
+           kableExtra::add_indent(c(9:13)) %>%
+           kableExtra::footnote(general =tb.foot ,footnote_as_chunk = T,threeparttable = T);tb.fm02
+```

### 2021-09-08 paper.Rmd
citationkeys
- ajuste de bib
- modificacion de keys en text
- csl apa7 incluido
- render en word (tablas manuales)

@@
 output:
+  bookdown::word_document2:
+    number_sections: false
+  bookdown::html_document2:
+    number_sections: false
   bookdown::pdf_document2:
@@
      number_sections: false
-  bookdown::html_document2:
-    number_sections: false
-  bookdown::word_document2:
-    number_sections: false  
 linkcolor: blue
@@
   - input/bib/meritocracy.bib
+csl: input/bib/apa7-singlespace.csl  
 editor_options:
@@
   - \usepackage{titlesec}
-  # - \setcounter{secnumdepth}{4} # activate for lvl 4 heading numbering
   - \titleformat{\paragraph}
@@
     {0pt}{3.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}
-always_allow_html: true #for word output    
+always_allow_html: true # for word output
+fontsize: 10pt
 ---
@@
  rmarkdown::render("paper.Rmd", output_format = "bookdown::html_document2")
- rmarkdown::render("paper.Rmd", output_format = "bookdown::word_document2")
+ # rmarkdown::render("paper.Rmd", output_format = "bookdown::word_document2")
 ```
@@
  options(scipen=999)
+ options(kableExtra.auto_format = FALSE)
  rm(list=ls())
@@
 Economic inequality and income concentration have become topics of growing concern over the last years, acquiring even more salience in times of pandemic and economic crises.
-This has led to a series of social movements in different societies as well as diverse critical analyses regarding the development of capitalism and its consequences [@streeckPoliticsPublicDebt2014; @piketty_Capital_2014].
-In this context, the study of views, preferences and perceptions of inequality has acquired relevance in the social sciences, in topics such as redistributive preferences [@alesina_Fairness_2005; @dimick_Models_2018], the legitimization of economic inequality [@schroder_Income_2017] and the functioning of meritocracy [@duru-bellat_whos_2012; @mijs_paradox_2019; @reynolds_perceptions_2014; @atriaEconomicElitesAttitudes2020].
+This has led to a series of social movements in different societies as well as diverse critical analyses regarding the development of capitalism and its consequences [@streeck_politics_2014; @piketty_capital_2014].
+In this context, the study of views, preferences and perceptions of inequality has acquired relevance in the social sciences, in topics such as redistributive preferences [@alesina_fairness_2005; @dimick_models_2018], the legitimization of economic inequality [@schroder_income_2017] and the functioning of meritocracy [@duru-bellat_who_2012; @mijs_paradox_2019; @reynolds_perceptions_2014; @atria_economic_2020].
 Within this context, the ideal of meritocracy has been strongly challenged as an unfulfilled promise of modern societies that allows the perpetuation of social inequalities [@sandel_tyranny_2020; @goldthorpe_myth_2003].
@@
 In general, meritocracy is defined as a system of distribution of resources and rewards based on individual merit, which in its original conception is a combination of talent and effort [@young_rise_1962].
-This traditional conception of merit places in a secondary position the possible interference of structural or non-meritocratic factors, such as inheritance, personal contacts, and luck [@breenClassInequalityMeritocracy1999; @saundersMightBritainBe1995; @yairMeritocracy2007; @landWeSatTable2006; @youngRiseMeritocracy1994].
-Social psychology and sociology have studied the characteristics and consequences of beliefs in meritocracy, under the general hypothesis that a greater belief in meritocracy emphasizes the role of individual over structural factors in personal achievements, leading to greater legitimization of inequalities [@premingerMeritocracyServiceEthnocracy2020; @trumpWhenWhyEconomic2020; @hadjar_meritokratie_2008; @MadeiraPrimesConsequencesSystematic2019].
-Such research has raised criticism of meritocracy as a moral standard of distribution given the preponderant weight of non-meritocratic elements upon individual status and social mobility[@sandelTyrannyMeritWhat2020a; @witteveenReconsideringMeritocraticPower2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_Meritocracy_2019; @khan_Privilege_2013].
+This traditional conception of merit places in a secondary position the possible interference of structural or non-meritocratic factors, such as inheritance, personal contacts, and luck [@breen_class_1999; @saunders_might_1995; @yair_meritocracy_2007; @land_we_2006; @young_rise_1962].
+Social psychology and sociology have studied the characteristics and consequences of beliefs in meritocracy, under the general hypothesis that a greater belief in meritocracy emphasizes the role of individual over structural factors in personal achievements, leading to greater legitimization of inequalities [@preminger_meritocracy_2020; @trump_when_2020; @hadjar_meritokratie_2008; @madeira_primes_2019].
+Such research has raised criticism of meritocracy as a moral standard of distribution given the preponderant weight of non-meritocratic elements upon individual status and social mobility[@sandel_tyranny_2020; @witteveen_reconsidering_2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_meritocracy_2019; @khan_privilege_2013].
 
 Due to the role that meritocratic beliefs play in the justification of individual achievement (or failure) in contemporary societies [@hadjar_meritokratie_2008, @markovits_meritocracy_2019; @sandel_tyranny_2020], multiple studies have evaluated the relationship between meritocratic beliefs and personal and/or contextual characteristics.
-For example, some studies have linked meritocracy to the reinforcement of socio-economic, gender, and ethnic stereotypes [@MadeiraPrimesConsequencesSystematic2019; @girerdNeoliberalismIdeologicalBarrier2020; @premingerMeritocracyServiceEthnocracy2020], as well as the effects of meritocratic beliefs in educational [@generettStoriesWeTell2020; @owensENGINESSOCIALMOBILITY2020] and organizational contexts [@perezAdvancingCareersMerit2020; @aielloNewEvidenceFirmuniversity2019].
+For example, some studies have linked meritocracy to the reinforcement of socio-economic, gender, and ethnic stereotypes [@madeira_primes_2019; @girerd_neoliberalism_2020; @girerd_neoliberalism_2020], as well as the effects of meritocratic beliefs in educational [@generett_stories_2020; @owens_engines_2020] and organizational contexts [@perez_advancing_2020; @aiello_new_2019].
 Most of such studies so far have used indicators from existing standard social surveys, such as the International Social Survey Programme (ISSP), to measure meritocracy.
@@
 In this conceptual and measurement approach of meritocratic beliefs, we can observe a couple of strong conceptual assumptions: a) effort would not depend on parental influence, and b) talent (as innate ability) is not meritocratic (contrary to Michael Youngs' original conceptualization).
-This conceptual and measurement-based assumption is found in other studies, that also assume effort to be the main and only aspect of meritocracy [@girerdNeoliberalismIdeologicalBarrier2020a; @bubakPerceptionsMeritocracyNote2019], raising the question: Is effort the only dimension behind the concept of meritocracy?
+This conceptual and measurement-based assumption is found in other studies, that also assume effort to be the main and only aspect of meritocracy [@girerd_neoliberalism_2020; @bubak_perceptions_2019], raising the question: Is effort the only dimension behind the concept of meritocracy?
 The question of whether talent (as inteligence and ability) is or is not considered meritocratic is certainly an interesting topic to discuss from a philosophical point of view, but for those working on empirical studies it should face empirical scrutiny.
-Besides, the consideration of talent as part of meritocracy certainly opens some interesting avenues of research, for instance in studies showing that for the elites meritocracy is more related to talent, whereas effort is something more characteristic of meritocracy of the middle and low classes ([### citar Atria ABS]).
+Besides, the consideration of talent as part of meritocracy certainly opens some interesting avenues of research, for instance in studies showing that for the elites meritocracy is more related to talent, whereas effort is something more characteristic of meritocracy of the middle and low classes [@atria_economic_2020].
 
@@
 
-The term belief has an ambiguous character in the literature, conceived as "idea-elements" by @converseNatureBeliefSystems1964 or "considerations" by @zallerNatureOriginsMass1992.
-As @kluegelBeliefsInequalityAmericans2017 pointed out about the scope of beliefs: "This usage encompasses such more specific social-psychological concepts as values, perceptions, and attitudes" (p.30).
+The term belief has an ambiguous character in the literature, conceived as "idea-elements" by @converse_nature_1964 or "considerations" by @zaller_nature_1992.
+As @kluegel_beliefs_1986 pointed out about the scope of beliefs: "This usage encompasses such more specific social-psychological concepts as values, perceptions, and attitudes" (p.30).
 Therefore, beliefs cover almost anything related to subjective factors.
-To this regard, a relevant distinction in the field of inequality beliefs was made by @janmaatSubjectiveInequalityReview2013: "Perceptions refer to subjective estimates of existing inequality (i.e.
+To this regard, a relevant distinction in the field of inequality beliefs was made by @janmaat_subjective_2013: "Perceptions refer to subjective estimates of existing inequality (i.e.
 thoughts about what is).
@@
 what is).
-This occurs for instance in @reynolds_perceptions_2014, in which the term beliefs is used to talk about what @janmaatSubjectiveInequalityReview2013 refer as perceptions, whereas other authors use general terms such as attitudes [@kunovich_systems_2007].
-The first attempt to shed light on this issue in meritocracy research was made by @duru-bellat_whos_2012, who used the question "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" as a proxy for "desired" meritocracy (normative beliefs).
+This occurs for instance in @reynolds_perceptions_2014, in which the term beliefs is used to talk about what @janmaat_subjective_2013 refer as perceptions, whereas other authors use general terms such as attitudes [@kunovich_systems_2007].
+The first attempt to shed light on this issue in meritocracy research was made by @duru-bellat_who_2012, who used the question "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" as a proxy for "desired" meritocracy (normative beliefs).
 They then determined "perceived" meritocracy, using the questions: “Would you say that in your country, people are rewarded for their efforts?” and “...
@@
 Is the belief in meritocracy a perception or a preference with normative meaning?
-In order to expand the analytical conceptual framework, we believe that both dimensions should be included as proposed by @duru-bellat_whos_2012.
+In order to expand the analytical conceptual framework, we believe that both dimensions should be included as proposed by @duru-bellat_who_2012.
 This opens up the possibility of analyzing whether perceptions and preferences are actually related (i.e.
 have a high correlation), or whether they are independent aspects of the same phenomenon.
-As @son_hing_merit_2011-1 have pointed out, "People can believe that outcomes ought to be distributed on the basis of merit and yet vary in their perceptions of whether this is how society currently operates" (p.
+As @sonhing_merit_2011 have pointed out, "People can believe that outcomes ought to be distributed on the basis of merit and yet vary in their perceptions of whether this is how society currently operates" (p.
 435).
@@
 Some research in meritocracy considers aspects usually opposed to effort and talent for personal achievement, as for instance the use of personal advantages (as contacts or having a wealthy family) to get ahead in life.
-For instance, @kunovich_systems_2007 used an items’ battery listing a number of factors in relation to “How important each should be in deciding pay...” (as @duru-bellat_whos_2012 for desired meritocracy).
+For instance, @kunovich_systems_2007 used an items’ battery listing a number of factors in relation to “How important each should be in deciding pay...” (as @duru-bellat_who_2012 for desired meritocracy).
 They consider factors such as education and responsibility as meritocratic, giving them a value of 1 if considered “essential” in the scale response, whereas factors such as having a family or children are valued 1 when rated as "not important at all" (i.e.
@@
 Despite making and proving this distinction, however, the authors end up subtracting one dimension from the other, thus coming back to the assumption that they are two poles of the same continuum as @kunovich_systems_2007 did.
-Similarly, @roex_attitudes_2019 used ISSP indicators for perceived meritocracy and non-meritocracy to build a single score by reverse coding the non-meritocratic items.
+Similarly, @roex_attitudes_2018 used ISSP indicators for perceived meritocracy and non-meritocracy to build a single score by reverse coding the non-meritocratic items.
 Therefore, the treatment of non-meritocratic items has been rather inconsistent across studies and the assumption that they are the simple opposite of meritocracy certainly requires further empirical assessment.
@@
 
-Finally, most studies in meritocracy so far have not properly considered the issue of latent structures and measurement error [@brownBookReviewTimothy2010; @bollenStructuralEquationsLatent1989] (Ansolabehere, S., Rodden, J., & Snyder Jr, J.
-M., 2008), as they mostly use single indicators and/or simple average indexes for measuring meritocracy.
+Finally, most studies in meritocracy so far have not properly considered the issue of latent structures and measurement error [@brown_confirmatory_2015; @bollen_structural_1989; @ansolabehere_strength_2008], as they mostly use single indicators and/or simple average indexes for measuring meritocracy.
 Such strategy assumes that the latent construct is measured perfectly (i.e.
@@
 ```
+
 The columns "Perceptions" and "Preferences" represent the distinction between these two concepts, usually confused under the label "beliefs" [@castillo_meritocracia_2019].
@@
 The estimation was performed using Confirmatory Factor Analysis (CFA).
-CFA was conducted using the `lavaan` R package (version 0.6-3; ### Rosseel, 2020), with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement (### Kline, 2016; Rosseel, 2020).
-As recommended by ### Brown (2008), we assessed model fit by jointly considering the comparative fit index and Tucker-Lewis Index (CFI and TLI; acceptable fit > 0.95), Root of the average squared residual approximation (RMSEA; acceptable fit < 0.08),  Chi-square: (p-value; acceptable fit > 0.05, and Chi-square ratio > 3).
+CFA was conducted using the `lavaan` R package (version 0.6-3; @rosseel_lavaan_2012), with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement [@kline_principles_2016; @rosseel_lavaan_2012].
+As recommended by @brown_confirmatory_2015, we assessed model fit by jointly considering the comparative fit index and Tucker-Lewis Index (CFI and TLI; acceptable fit > 0.95), Root of the average squared residual approximation (RMSEA; acceptable fit < 0.08),  Chi-square: (p-value; acceptable fit > 0.05, and Chi-square ratio > 3).
 
@@
 ```
+
 Regarding the correlations between the factors in the Figure \@ref(fig:meas02-issp), we observe that perceptions are correlated with preferences, but more strongly for the meritocratic ($r=0.46$) than for the non-meritocratic ($r=0.27$) dimensions.
@@
 The sample was selected from a non-probabilistic quota design in three large cities in Chile (Santiago, Concepción & Antofagasta).
-The quotas for gender, age, and educational levels were generated based on a survey by the Public Studies Center [@cepEncuestaCEPMayo2019], which is a well regarded counterpart agency of the ISSP (International Social Survey Programme) in Chile.
+The quotas for gender, age, and educational levels were generated based on a survey by the Public Studies Center [@cep_encuesta_2019], which is a well regarded counterpart agency of the ISSP (International Social Survey Programme) in Chile.
 A total sample of 2,141 people was collected, excluding those who did not signed the informed consent as well as those not answering the meritocracy response battery.
 There were no significant differences between our sample and the wider population for most socio-demographic characteristics, with the exception of educational level (see Table \@ref(tab:rep-samp) in Appendix).
-As it is often the case with online surveys, there were some limitations in achieving the quotas for lower educational levels **[citas - LUCHO]**.
+As it is often the case with online surveys, there were some limitations in achieving the quotas for lower educational levels [@zhang_quota_2018; @boas_recruiting_2020].
 
@@
 The proposed scale of perceptions and preferences relating to meritocracy consisted of eight indicators that were grouped into the four dimensions listed earlier: Perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic).
-In order to achieve at least some comparability with previous studies, the questions were adapted from the items battery "reasons to get ahead" (ISSP/GSS), which has been widely used for operationalizing meritocracy in previous studies [@mijs_paradox_2019; @duru-bellat_whos_2012; @reynolds_perceptions_2014].
+In order to achieve at least some comparability with previous studies, the questions were adapted from the items battery "reasons to get ahead" (ISSP/GSS), which has been widely used for operationalizing meritocracy in previous studies [@mijs_paradox_2019; @duru-bellat_who_2012; @reynolds_perceptions_2014].
 The aforementioned eight items, ordered according to dimensions, are presented in Table \@ref(tab:table-indicadores).
@@
 The models estimated one factor for each of the four proposed dimensions presented in Table \@ref(tab:table-indicadores).
-As in Study 1, CFA was conducted using the `lavaan` R package (version 0.6-3; Rosseel, 2020), with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement (Kline, 2016; Rosseel, 2020).
+As in Study 1, CFA was conducted using the `lavaan` R package (version 0.6-3; @rosseel_lavaan_2012), with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement [@kline_principles_2016; @rosseel_lavaan_2012].
 The fit indexes and cut-off criteria were the same as the ones used in Study 1.
@@
 ```
+
 Attending now to the association among the scale items, Figure \@ref(fig:corpoly) shows items polychoric correlations.
 There are three main aspects to highlight from this correlation matrix.
-Firstly, as expected the largest correlations are between indicators that correspond to the same factors behind the conceptual model (e.g., perception of meritocracy by effort and by talent, $r$=0.52).
-Secondly, among this correlations the highest are those between the non-meritocratic dimension but in perceptions ($r$=0.73) and preferences ($r$=0.61).
+Firstly, as expected the largest correlations are between indicators that correspond to the same factors behind the conceptual model (e.g., perception of meritocracy by effort and by talent, $r=0.52$).
+Secondly, among this correlations the highest are those between the non-meritocratic dimension but in perceptions ($r=0.73$) and preferences ($r=0.61$).
 Thirdly, both items for meritocratic preferences (E and F) are the ones that mostly correlate with the rest of the perceptual items, showing medium to high correlations.
@@
 - Opportunity beliefs: We used two items of the social inequality module of the International Social Survey Programme.
-Studies use generally both indicators to measure meritocratic and non-meritocratic beliefs [@McCall2017Exposurerisinginequality; @mijs_paradox_2019].
+Studies use generally both indicators to measure meritocratic and non-meritocratic beliefs [@mccall_exposure_2017; @mijs_paradox_2019].
 One item measures the importance of individual factors in determining life outcomes by asking for the importance of hard work for getting ahead in life (M=3.76, SD=0.93).
@@
 
-- Personal wherewithal: the questionnaire included The Neoliberal Beliefs Inventory (NBI), which considers four factors: Government Interference preferences, Competition preferences, System Inequality perception and Personal Wherewithal [@bay-chengTrackingHomoOeconomicus2015].
+- Personal wherewithal: the questionnaire included The Neoliberal Beliefs Inventory (NBI), which considers four factors: Government Interference preferences, Competition preferences, System Inequality perception and Personal Wherewithal [@bay-cheng_tracking_2015].
 We used the factor _personal wherewithal_ that reflects meritocracy beliefs in terms of the importance of personal attributes as strength and skills to yield success.
@@
 The item for social origin shows a positive and moderate association with unmeritocratic perception ($r=0.361$), and a negative and low correlation with meritocratic perception ($r=-0.186$).
-Regarding the item of hard work, the correlation with meritocratic perception is positive but weak ($r=.201$).
+Regarding the item of hard work, the correlation with meritocratic perception is positive but weak ($r=0.201$).
 One explanation for this result is that the support for hard work is very high among respondents and therefore with low variability, as only 8.6 percent indicates _not very important_ or _not important at all_.
@@
       bootstrap_options = c("striped", "bordered")) %>%
-    kableExtra::column_spec(column = 1, width = "4 cm")
+    kableExtra::column_spec(column = 1, width = "4.5 cm")
 ```
@@
 The NBI's factor depicts a positive association with unmeritocratic preference ($r=0.225$).
-This finding is consistent with literature, inasmuch as the belief for meritocray is associated with the justification of inequality [@McCall2017Exposurerisinginequality; @MadeiraPrimesConsequencesSystematic2019].
+This finding is consistent with literature, inasmuch as the belief for meritocray is associated with the justification of inequality [@mccall_exposure_2017; @madeira_primes_2019].
 Furthermore, results of our confirmatory factor analysis indicate that meritocratic perceptions and unmeritocratic preferences show a positive and weak correlation.
@@
 
-The modeling for invariance testing has been gaining more relevance in sociological survey studies, mainly due to the interest in the validity of measurement scales in comparative studies in various social and cultural contexts [@Davidovetal2014].
+The modeling for invariance testing has been gaining more relevance in sociological survey studies, mainly due to the interest in the validity of measurement scales in comparative studies in various social and cultural contexts [@davidov_measurement_2014].
 In this section we will use an analysis of invariance between groups, where the proposed factorial structure is expected to be independent of belonging to the sample corresponding to the original study, comparing it with another more recent study available and where the meritocracy scale was also incorporated.
@@
 The invariance measurement procedure consists of a series of nested models to which restrictions are progressively incorporated on the parameters of the measurement model.
-The literature generally suggests that this modeling should be done through four levels or types of progressive restriction [@Milfont2010; @millsap_Statistical_2011; @vandeschoot_checklist_2012]:
+The literature generally suggests that this modeling should be done through four levels or types of progressive restriction [@milfont_testing_2010; @millsap_statistical_2011; @vandeschoot_checklist_2012]:
 
@@
 Although the chi-square statistic is used as a global measure of fit, since its sensitivity to sample size it is recommended to use three additional fit indices: Comparative Fit Index (CFI), which should have a value greater than 0.95; Root Mean Square Error Approximation (RMSEA), which must be lower than 0.06; and the Standardized Root Mean Square Residuals (SRMR) which must be less than 0.08 [@vandeschoot_checklist_2012].
-Besides, the literature on measurement invariance suggests some complementary approaches for the evaluation of the fit, from which we will consider the incremental adjustment of the fit indexes [@cheung_Evaluating_2002; @Milfont2010; @dimitrov_Testing_2010] and the ANOVA test for means comparison in nested models [@newsom_Longitudinal_2015].
+Besides, the literature on measurement invariance suggests some complementary approaches for the evaluation of the fit, from which we will consider the incremental adjustment of the fit indexes [@cheung_evaluating_2002; @milfont_testing_2010; @dimitrov_testing_2010] and the ANOVA test for means comparison in nested models [@newsom_longitudinal_2015].
 
@@
 When attending to the traditional invariance test of $\Delta \chi^2 (\Delta \text{df})$, the results  support the invariance at the strong level meaning that the fit of the factor model of the merit scale is equivalent across samples when constraining factor loadings and intercepts to being equal.
-Such result is considered in general as evidence of invariance [citar paper Fischer en SJR], as strict forms of measurement invariance rarely hold [citar van der Schoot et al].
+Such result is considered in general as evidence of invariance [@fischer_are_2011], as strict forms of measurement invariance rarely hold [@vandeschoot_checklist_2012].
 Still, the comparability of latents means requires strict invariance which in this case does not hold when considering $\Delta \chi^2 (\Delta \text{df})$.
-Nevertheless, the criteria of $\Delta \text{CFI}$ used for comparing models is close to the rejection criteria of >.01, whereas the $\Delta \text{RMSEA}$ fulfills the requirements of being below of the cut-off criteria as suggested by [ @chen 2007 ].
+Nevertheless, the criteria of $\Delta \text{CFI}$ used for comparing models is close to the rejection criteria of >.01, whereas the $\Delta \text{RMSEA}$ fulfills the requirements of being below of the cut-off criteria as suggested by [@chen_sensitivity_2007].
 Therefore, using this last standard the level for strict invariance holds for the meritocracy scale.

### 2021-09-10 paper.Rmd
Checklist avance
Se ajustan aspectos de forma como títulos, abreviaturas y agradecimientos, así como se revisaron citas. Falta ver el formato de las imágenes y los títulos de las imágenes. Subo esto mientras.

@@
 output:
-  bookdown::word_document2:
+  bookdown::pdf_document2:
+     template: null
+     toc: false
      number_sections: false
@@
     number_sections: false
-  bookdown::pdf_document2:
-     template: null
-     toc: false
+  bookdown::word_document2:
     number_sections: false
@@
 
+\newpage
+
+# Acknowledgments
+
+The authors recognize the support of ANID/FONDECYT grant 1160921 and of  the Center for Conflict and Social Cohesion Studies (COES)  grant ANID/FONDAP 1513009.
+
+\newpage
+
 # Introduction
@@
 The analysis is organized in three studies.
-Study 1 serves as a background to our instrument proposal by attempting to operationalize the meritocracy dimensions with available secondary data from the inequality module of the International Social Survey Programme (ISSP).
+Study 1 serves as a background to our instrument proposal by attempting to operationalize the meritocracy dimensions with available secondary data from the inequality module of the ISSP.
 Although we are aware that the available indicators in this dataset do not allow a precise operationalization of the conceptual model, we know that it is the most widely used data in comparative meritocracy studies (as well as overall subjective inequality) and therefore it is certainly useful to look at its potentials and limitations to this regard.
@@
 
-## Study 1: Measuring meritocracy with international secondary data
-
-### Data
-
-The data cooresponds to the last available wave of the social inequality module from the International Social Survey Programme (ISSP), which is the most specialized international comparative survey in perceptions, attitudes and beliefs about inequality related issues.
+# Study 1: Measuring meritocracy with international secondary data
+
+## Data
+
+The data cooresponds to the last available wave of the social inequality module from the ISSP, which is the most specialized international comparative survey in perceptions, attitudes and beliefs about inequality related issues.
 This wave corresponds to the year 2009 and covers attitudes towards a series of topics dealing with social inequality across 41 countries and 56021 individuals.
@@
 
-### Methods
+## Methods
 
@@
 
-### Results
-
-#### Descriptive analyses
+## Results
+
+### Descriptive analyses
 
@@
 
-#### Confirmatory Factor Analysis
+### Confirmatory Factor Analysis
 
@@
 
-## Study 2: The Perceptions and Preferences for Meritocracy Scale
-
-### Data {-#data}
+# Study 2: The Perceptions and Preferences for Meritocracy Scale
+
+## Data {-#data}
 
@@
 The sample was selected from a non-probabilistic quota design in three large cities in Chile (Santiago, Concepción & Antofagasta).
-The quotas for gender, age, and educational levels were generated based on a survey by the Public Studies Center [@cep_encuesta_2019], which is a well regarded counterpart agency of the ISSP (International Social Survey Programme) in Chile.
+The quotas for gender, age, and educational levels were generated based on a survey by the Public Studies Center [@cep_encuesta_2019], which is a well regarded counterpart agency of the ISSP in Chile.
 A total sample of 2,141 people was collected, excluding those who did not signed the informed consent as well as those not answering the meritocracy response battery.
@@
 
-### Study design
-
-#### Instrument
+## Study design
+
+### Instrument
 
@@
 
-#### Administration sets
+### Administration sets
 
@@
 
-## Study 3: Additional validity analyses
+# Study 3: Additional validity analyses
 
@@
 
-### Convergent validity
-
-#### Data
+## Convergent validity
+
+### Data
 
@@
 
-#### Variables
+### Variables
 
@@
 
-#### Results
+### Results
 
@@
 
-### Measurement invariance
+## Measurement invariance
 
@@
 
-#### Data
+### Data
 

### 2021-09-15 paper.Rmd
Ajustes edición Checklistsjr
 +  Se crean algunos archivos de imagen, para dejar en png todo lo que aparece en el pdf.

+  Se cambian los nombres de los archivos según la nomesclatura del [Checklist](https://docs.google.com/spreadsheets/d/1O8GbZ07JKsqoukfPmQ6ugieUS5mXB24alwzlNrwRIFs/edit#gid=0) "Name your figure files with "Fig" and the figure number, e.g., Fig1.eps."

+ se indican las partes de las figuras con minusculas

+ Se cambia figura por Fig.  en el titulo.

Por ahora no pude resolver dos problemas. No puedo poner en negrita Fig. 1 en el titulo de las figuras.  y por otro no puedo elim

@@
   - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
+  - \usepackage[figurename=Fig.]{caption}
   - \usepackage{graphicx}
@@
 ```{r merit-model, echo=FALSE, fig.cap = "Model of perception and preferences for meritocracy and non-meritocracy", fig.align='center'}
- knitr::include_graphics("input/images/generalf.png")
+ knitr::include_graphics("input/images/Fig1.generalf.png")
 ```
@@
 ```{r desc-issp, eval=FALSE, include=FALSE, results='asis'}
+
   load(file = "input/data/proc/data_s1.RData")
@@
                         grid.range  =  c (1.2 , 1.2),
-                        title = "Perception meritocratic",
+                        title = "a. Perception meritocratic",
                           geom.size = c(0.62),
@@
                         grid.range  =  c (1.2 , 1.2),
-                        title = "Perception Non-meritocratic",
+                        title = "b. Perception Non-meritocratic",
                           geom.size = c(0.9),
@@
                         grid.range  =  c (1.2 , 1.2),
-                        title = "Preference meritocratic",
+                        title = "c. Preference meritocratic",
                           geom.size = c(0.62),
@@
                         grid.range  =  c (1.2 , 1.2),
-                        title = "Preference Non-meritocratic",
+                        title = "d. Preference Non-meritocratic",
                           geom.size = c(0.62),
@@
 ```{r likert-issp,fig.height=11,fig.width=11,fig.cap="Distribution of responses in the ISSP meritocracy items", out.width='80%'}
+
+#dev.copy(png,'output/images/Fig2.plotlikert-study1.png') # save image
+
   if (!require("pacman")) install.packages("pacman")
@@
     cowplot::ggdraw()
+#dev.off()
 ```
@@
 ```{r corr-issp,fig.cap = "Perceptions and preferences for ISSP  meritocracy items' polychoric correlations", fig.align='center',fig.height=6,fig.width=8}
+
+#dev.copy(png,'output/images/Fig3.plotcor-sutdy1.png') # save image
+
   load(file = "input/data/proc/data_s1.RData")
@@
     na.label = "-")
+
+#dev.off()
+
 ```
@@
 ```{r meas02-issp, echo=FALSE, fig.cap = "Descriptive plot", fig.align='center',fig.cap="Confirmatory factor analysis of the ISSP indicators of Perceptions and Preferences for Meritocracy"}
- knitr::include_graphics('output/images/cfa_issp_study1.PNG')
+ knitr::include_graphics('output/images/Fig4.cfa_issp_study1.PNG')
 ```
@@
 ```{r appmod, echo=FALSE, fig.cap = "Survey flow", fig.align='center'}
- knitr::include_graphics('output/images/app_mod.png')
+ knitr::include_graphics('output/images/Fig5.app_mod.png')
 ```
@@
     sjPlot::plot_likert(geom.colors = "PuBu",
-                        title = c("Perceptions"),
+                        title = c("a. Perceptions"),
                         geom.size = 0.8,
@@
     sjPlot::plot_likert(geom.colors = "PuBu",
-                        title = c("Preferences"),
+                        title = c("b. Preferences"),
                         geom.size = 0.8,
@@
 ```{r plotlikert, echo=FALSE, fig.align='center', fig.cap="Distribution of responses in the Merit Scale items", fig.height=8, fig.width=11}
+
+#dev.copy(png,'output/images/Fig6.plotlikertmerit-scale.png') # save image
+
   if (!require("pacman")) install.packages("pacman")
@@
     cowplot::ggdraw()
+  #dev.off()
+
 ```
@@
 ```{r corpoly, echo=FALSE, fig.cap = "Perceptions and preferences for meritcoracy items' polychoric correlations", fig.align='center'}
+
+#dev.copy(png,'output/images/Fig6.plotcor-sutdy2.png') # save image  
   if (!require("pacman")) install.packages("pacman")
@@
     na.label = "-")
+#dev.off()
 ```
@@
  # estimation for diagram: summary(fit_cfa3, standardized = TRUE, fit.measures=TRUE)
- knitr::include_graphics('output/images/cfa_randomized_study2.png')
+ knitr::include_graphics('output/images/Fig8.cfa_randomized_study2.png')
 ```

### 2021-09-22 paper.Rmd
ajustes grammar

@@
      toc: false
-     number_sections: false
-  bookdown::html_document2:
-    number_sections: false
-  bookdown::word_document2:
      number_sections: false
@@
 header-includes:
-  # - \usepackage[spanish,es-tabla,es-nodecimaldot]{babel}
   - \usepackage{times}           # Times New Roman
@@
   - \usepackage{threeparttable}
-  - \usepackage{titlesec}
-  - \titleformat{\paragraph}
-    {\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}
-    \titlespacing*{\paragraph}
-    {0pt}{3.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}
 always_allow_html: true # for word output
@@
 
-\newpage
+\pagebreak
 
@@
 This has led to a series of social movements in different societies as well as diverse critical analyses regarding the development of capitalism and its consequences [@streeck_politics_2014; @piketty_capital_2014].
-In this context, the study of views, preferences and perceptions of inequality has acquired relevance in the social sciences, in topics such as redistributive preferences [@alesina_fairness_2005; @dimick_models_2018], the legitimization of economic inequality [@schroder_income_2017] and the functioning of meritocracy [@duru-bellat_who_2012; @mijs_paradox_2019; @reynolds_perceptions_2014; @atria_economic_2020].
+In this context, the study of views, preferences, and perceptions of inequality has acquired relevance in the social sciences, in topics such as redistributive preferences [@alesina_fairness_2005; @dimick_models_2018], the legitimization of economic inequality [@schroder_income_2017] and the functioning of meritocracy [@duru-bellat_who_2012; @mijs_paradox_2019; @reynolds_perceptions_2014; @atria_economic_2020].
 Within this context, the ideal of meritocracy has been strongly challenged as an unfulfilled promise of modern societies that allows the perpetuation of social inequalities [@sandel_tyranny_2020; @goldthorpe_myth_2003].
@@
 This traditional conception of merit places in a secondary position the possible interference of structural or non-meritocratic factors, such as inheritance, personal contacts, and luck [@breen_class_1999; @saunders_might_1995; @yair_meritocracy_2007; @land_we_2006; @young_rise_1962].
-Social psychology and sociology have studied the characteristics and consequences of beliefs in meritocracy, under the general hypothesis that a greater belief in meritocracy emphasizes the role of individual over structural factors in personal achievements, leading to greater legitimization of inequalities [@preminger_meritocracy_2020; @trump_when_2020; @hadjar_meritokratie_2008; @madeira_primes_2019].
-Such research has raised criticism of meritocracy as a moral standard of distribution given the preponderant weight of non-meritocratic elements upon individual status and social mobility[@sandel_tyranny_2020; @witteveen_reconsidering_2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_meritocracy_2019; @khan_privilege_2013].
-
-Due to the role that meritocratic beliefs play in the justification of individual achievement (or failure) in contemporary societies [@hadjar_meritokratie_2008, @markovits_meritocracy_2019; @sandel_tyranny_2020], multiple studies have evaluated the relationship between meritocratic beliefs and personal and/or contextual characteristics.
+Social psychology and sociology have studied the characteristics and consequences of beliefs in meritocracy, under the general hypothesis that a greater belief in meritocracy emphasizes the role of the individual over structural factors in personal achievements, leading to greater legitimization of inequalities [@preminger_meritocracy_2020; @trump_when_2020; @hadjar_meritokratie_2008; @madeira_primes_2019].
+Such research has raised criticism of meritocracy as a moral standard of distribution given the preponderant weight of non-meritocratic elements upon the individual status and social mobility [@sandel_tyranny_2020; @witteveen_reconsidering_2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_meritocracy_2019; @khan_privilege_2013].
+
+Due to the role that meritocratic beliefs play in the justification of individual achievement (or failure) in contemporary societies [@hadjar_meritokratie_2008; @markovits_meritocracy_2019; @sandel_tyranny_2020], multiple studies have evaluated the relationship between meritocratic beliefs and personal and/or contextual characteristics.
 For example, some studies have linked meritocracy to the reinforcement of socio-economic, gender, and ethnic stereotypes [@madeira_primes_2019; @girerd_neoliberalism_2020; @girerd_neoliberalism_2020], as well as the effects of meritocratic beliefs in educational [@generett_stories_2020; @owens_engines_2020] and organizational contexts [@perez_advancing_2020; @aiello_new_2019].
@@
 In many cases, similar phenomena are associated with different indicators, and conversely, often different phenomena are measured with similar indicators.
-Such trends not only limits the comparability of studies, but also the ability to understanding the effects of meritocratic beliefs across different disciplines and lines of research.
+Such trends not only limit the comparability of studies, but also the ability to understanding the effects of meritocratic beliefs across different disciplines and lines of research.
 
@@
 
-In the following we discuss four critical aspects in studies conceptualizing and measuring meritocracy and that are the basis of the conceptual and measurement proposal.
+In the following, we discuss four critical aspects in studies conceptualizing and measuring meritocracy, based upon which we develop a conceptual and measurement proposal.
 
@@
 
-One recent definition of meritocracy by @mijs_paradox_2019 is the following: "when I discuss meritocracy beliefs, I am referring to citizens’ belief in the importance of hard work relative to structural factors." [@mijs_paradox_2019, pg.9].
+One recent definition of meritocracy by @mijs_paradox_2019 is the following: "When I discuss meritocracy beliefs, I am referring to citizens' belief in the importance of hard work relative to structural factors." [@mijs_paradox_2019, pg.9].
 In the subsequent operationalization, this is associated with the following question and indicator: "how important you think it is for getting ahead in life: (a) hard work", scored from 1 to 5 on a likert scale.
@@
 Therefore, when considering only one of the items, it means that other aspects that could be associated with talent, such as education, would not be deemed as meritocratic.
-As he points out: "Hard work is arguably the most meritocratic part of Michael Young’s equation: ‘Merit = Intelligence + Effort’, for the simple fact that intelligence itself is influenced by a non-meritocratic factor: who your parents happen to be" (p.5).
+As he points out: "Hard work is arguably the most meritocratic part of Michael Young's equation: 'Merit = Intelligence + Effort', for the simple fact that intelligence itself is influenced by a non-meritocratic factor: who your parents happen to be" (p.5).
 
 In this conceptual and measurement approach of meritocratic beliefs, we can observe a couple of strong conceptual assumptions: a) effort would not depend on parental influence, and b) talent (as innate ability) is not meritocratic (contrary to Michael Youngs' original conceptualization).
-This conceptual and measurement-based assumption is found in other studies, that also assume effort to be the main and only aspect of meritocracy [@girerd_neoliberalism_2020; @bubak_perceptions_2019], raising the question: Is effort the only dimension behind the concept of meritocracy?
-The question of whether talent (as inteligence and ability) is or is not considered meritocratic is certainly an interesting topic to discuss from a philosophical point of view, but for those working on empirical studies it should face empirical scrutiny.
-Besides, the consideration of talent as part of meritocracy certainly opens some interesting avenues of research, for instance in studies showing that for the elites meritocracy is more related to talent, whereas effort is something more characteristic of meritocracy of the middle and low classes [@atria_economic_2020].
+This conceptual and measurement-based assumption is found in other studies, which also assumes effort to be the main and only aspect of meritocracy [@girerd_neoliberalism_2020; @bubak_perceptions_2019], raising the question: Is effort the only dimension behind the concept of meritocracy?
+The question of whether talent (as intelligence and ability) is or is not considered meritocratic is certainly an interesting topic to discuss from a philosophical point of view, but for those working on empirical studies, it should face empirical scrutiny.
+Besides, the consideration of talent as part of meritocracy certainly opens some interesting avenues of research, for instance in studies showing that for the elites meritocracy is more related to talent, whereas effort is something more characteristic of the meritocracy of the middle and low classes [@atria_economic_2020].
 
@@
 Several approaches to the empirical study of meritocracy based on public opinion surveys refer to the concept of _beliefs_, but with wide differences in meanings and operationalizations.
-To illustrate this point, lets consider the most commonly used items for operationalizing meritocratic beliefs in survey research (as ISSP), which is the "reasons to get ahead" battery.
-It consists in a series of items asking "how important you think it is for getting ahead in life" and then lists several factors, such as effort, education, parental wealth, and contacts.
+To illustrate this point let us consider the most commonly used items for operationalizing meritocratic beliefs in survey research (as ISSP), which is the "reasons to get ahead" battery.
+It consists of a series of items asking "how important you think it is for getting ahead in life" and then lists several factors, such as effort, education, parental wealth, and contacts.
 Another version of this same battery used in other surveys - sometimes together with the previous one - asking about "how important you think it _should_ be ...", and then listing the same concepts.
@@
 what is).
-This occurs for instance in @reynolds_perceptions_2014, in which the term beliefs is used to talk about what @janmaat_subjective_2013 refer as perceptions, whereas other authors use general terms such as attitudes [@kunovich_systems_2007].
+This occurs for instance in @reynolds_perceptions_2014, in which the term belief is used to talk about what @janmaat_subjective_2013 refers to as perceptions, whereas other authors use general terms such as attitudes [@kunovich_systems_2007].
 The first attempt to shed light on this issue in meritocracy research was made by @duru-bellat_who_2012, who used the question "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" as a proxy for "desired" meritocracy (normative beliefs).
-They then determined "perceived" meritocracy, using the questions: “Would you say that in your country, people are rewarded for their efforts?” and “...
-people are rewarded for their skills?”.
+They then determined "perceived" meritocracy, using the questions: "Would you say that in your country, people are rewarded for their efforts?" and "...
+people are rewarded for their skills?".
 
@@
 have a high correlation), or whether they are independent aspects of the same phenomenon.
-As @sonhing_merit_2011 have pointed out, "People can believe that outcomes ought to be distributed on the basis of merit and yet vary in their perceptions of whether this is how society currently operates" (p.
+As @sonhing_merit_2011 has pointed out, "People can believe that outcomes ought to be distributed on the basis of merit and yet vary in their perceptions of whether this is how society currently operates" (p.
 435).
-In other words, normative beliefs should be considered while taking perceptions into account: a strong normative belief in meritocracy may mean something totally different to someone perceiving high meritocracy, when compared to someone perceiving low meritocracy.
+In other words, normative beliefs should be considered while taking perceptions into account: a strong normative belief in meritocracy may mean something totally different to someone perceiving high meritocracy than to someone perceiving low meritocracy.
 To avoid the confusion generated by the term "belief", we propose the terms meritocratic preferences ("what should be"), and meritocratic perceptions ("what is"), as they better reflect the two facets of meritocracy under scrutiny [@castillo_meritocracia_2019].
@@
 Some research in meritocracy considers aspects usually opposed to effort and talent for personal achievement, as for instance the use of personal advantages (as contacts or having a wealthy family) to get ahead in life.
-For instance, @kunovich_systems_2007 used an items’ battery listing a number of factors in relation to “How important each should be in deciding pay...” (as @duru-bellat_who_2012 for desired meritocracy).
-They consider factors such as education and responsibility as meritocratic, giving them a value of 1 if considered “essential” in the scale response, whereas factors such as having a family or children are valued 1 when rated as "not important at all" (i.e.
+For instance, @kunovich_systems_2007 used an items' battery listing a number of factors in relation to "How important each should be in deciding pay..." (as @duru-bellat_who_2012 for desired meritocracy).
+They consider factors such as education and responsibility as meritocratic, giving them a value of 1 is considered "essential" in the scale response, whereas factors such as having a family or children are valued 1 when rated as "not important at all" (i.e.
 reverse coded).
@@
 The columns "Perceptions" and "Preferences" represent the distinction between these two concepts, usually confused under the label "beliefs" [@castillo_meritocracia_2019].
-Perceptions refer to the extent to which people observe that meritocracy functions or applies in their society, which relates to items such as "Hard work is important to get ahead in society".
+Perceptions refer to the extent to which people observe that meritocracy functions or apply in their society, which relates to items such as "Hard work is important to get ahead in society".
 Preferences refer to normative expectations that are usually linked to a "should" expression (e.g.
@@
 The rows in Figure \@ref(fig:merit-model) consider the distinction between meritocratic and non-meritocratic dimensions [@reynolds_perceptions_2014], usually treated as different ends of the same continuum in previous research.
-Non-meritocratic elements refer to the use of resources such as personal contacts or family advantages to get ahead in life.
+Non-meritocratic elements refer to the use of resources such as personal contacts or family advantages to getting ahead in life.
 
@@
 
-The analysis is organized in three studies.
-Study 1 serves as a background to our instrument proposal by attempting to operationalize the meritocracy dimensions with available secondary data from the inequality module of the ISSP.
+The analysis is organized into three studies.
+Study 1 serves as a background to our instrument proposal by attempting to operationalize the meritocracy dimensions with available secondary data from the inequality module of the International Social Survey Programme (ISSP).
 Although we are aware that the available indicators in this dataset do not allow a precise operationalization of the conceptual model, we know that it is the most widely used data in comparative meritocracy studies (as well as overall subjective inequality) and therefore it is certainly useful to look at its potentials and limitations to this regard.
 Study 2 is the core of this paper and consists of the proposal of a new scale for measuring meritocratic perceptions and preferences.
-Finally, in Study 3 we perform additional validity analysis of the meritocracy scale.
+Finally, in Study 3 we perform an additional validity analysis of the meritocracy scale.
 
@@
 
-The data cooresponds to the last available wave of the social inequality module from the ISSP, which is the most specialized international comparative survey in perceptions, attitudes and beliefs about inequality related issues.
+The data corresponds to the last available wave of the social inequality module from the International Social Survey Programme (ISSP), which is the most specialized international comparative survey in perceptions, attitudes, and beliefs about inequality-related issues.
 This wave corresponds to the year 2009 and covers attitudes towards a series of topics dealing with social inequality across 41 countries and 56021 individuals.
-Although there is data available from this module for previous waves (1987, 1992 and 1999), unfortunately there are several variables that are important for this research that were not included, particularly in 1987 and 1999, reason why we only use the 2009 wave (during the development of this research the 2019 wave became available, but so far only with a restricted set of countries).
+Although there is data available from this module for previous waves (1987, 1992, and 1999), unfortunately, there are several variables that are important for this research which were not included, particularly in 1987 and 1999, the reason why we only use the 2009 wave (during the development of this research the 2019 wave became available, but so far only with a restricted set of countries).
 
@@
 
-- Preferences for meritocracy-non meritocracy: for the operationalization of normative preferences we used a list of items related to reasons for pay battery.
+- _Preferences for meritocracy-non meritocracy_: for the operationalization of normative preferences we used a list of items related to reasons for pay battery.
 The ISSP question was: _In deciding how much people ought to earn, how important should each of these things be, in your opinion_, rated in the same _essential-non important at all_ scale (1 to 5) as the questions for meritocratic perceptions.
@@
 The estimation was performed using Confirmatory Factor Analysis (CFA).
-CFA was conducted using the `lavaan` R package (version 0.6-3; @rosseel_lavaan_2012), with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement [@kline_principles_2016; @rosseel_lavaan_2012].
+CFA was conducted using the `lavaan` R package (version 0.6-3; @rosseel_lavaan_2012), with diagonally weighted least squares (DWLS) estimation due to the items' ordinal level of measurement [@kline_principles_2016; @rosseel_lavaan_2012].
 As recommended by @brown_confirmatory_2015, we assessed model fit by jointly considering the comparative fit index and Tucker-Lewis Index (CFI and TLI; acceptable fit > 0.95), Root of the average squared residual approximation (RMSEA; acceptable fit < 0.08),  Chi-square: (p-value; acceptable fit > 0.05, and Chi-square ratio > 3).
@@
 ### Descriptive analyses
-
-```{r desc-issp, eval=FALSE, include=FALSE, results='asis'}
-
-  load(file = "input/data/proc/data_s1.RData")
-  if (!require("pacman")) install.packages("pacman")
-  pacman::p_load(dplyr, psych, knitr, kableExtra)
-  table_format = if(knitr::is_html_output()) { #conditional instructions for kable
-    "html" #if html set "html" in format
-  } else if(knitr::is_latex_output()) {
-    "latex"#if latex set "latex" in format
-  }
-
-  data_s1 %>%
-    dplyr::select(hwork,ambition,wealthy,pareduc,
-                  race,gender,people,polcone,welljob,hardjob,family,child) %>%
-    psych::describe() %>%
-    as.data.frame() %>%
-    dplyr::select("Mean"=mean,"SD"=sd,"Min"=min,"Max"=max) %>%
-    round(2) ->desc.issp
-
-  row.names(desc.issp) <- c(
-    "Importance: hard work",
-    "Importance: having ambition",
-    "Importance: wealthy family",
-    "Importance: educated parents",
-    "Importance: race",
-    "Importance: gender",
-    "Importance: knowing people",
-    "Importance: political connections",
-    "Reasons for pay: well job",
-    "Reasons for pay: hard job",
-    "Reasons for pay: support family",
-    "Reasons for pay: has children"
-  )
-
-  sum.issp09<- knitr::kable(desc.issp,format =table_format,
-                   booktabs= T,
-                   align = "l",
-                   linesep = "",
-                   caption = "Descriptive statistics for the ISSP 2009 sample") %>%
-    kableExtra::kable_styling(full_width = T,
-                              font_size = 10,
-                              position = "center",
-                              latex_options = c("HOLD_position"));sum.issp09
-```
 
@@
 On the one side, we see that there is a high degree of importance attributed to factors such as hard work and ambition in the process of getting ahead, concentrating 94.9% and 92.8% in the fairly important to essential categories.
-By contrast, the perception non-meritocratic aspects is rated lower that the meritocratic ones, particularly for gender, race and political connections.
+By contrast, the perception of non-meritocratic aspects is rated lower than the meritocratic ones, particularly for gender, race, and political connections.
 Regarding normative preferences, we observe that the meritocratic ones are deemed as important (from fairly to essential) for nearly the whole sample, support that decreases slightly for the non-meritocratic ones.
-Still, we have to consider that the preferences for non-meritocratic aspects in this case refer to the distributive principle of need rather than personal background (as in the perceptions of non-meritocracy), which certainly is one of the limitations of this set of items for the operationalization of the proposed conceptual framework with this dataset.
+Still, we have to consider that the preferences for non-meritocratic aspects, in this case, refer to the distributive principle of need rather than personal background (as in the perceptions of non-meritocracy), which certainly is one of the limitations of this set of items for the operationalization of the proposed conceptual framework with this dataset.
 
@@
 ```{r likert-issp,fig.height=11,fig.width=11,fig.cap="Distribution of responses in the ISSP meritocracy items", out.width='80%'}
-
-#dev.copy(png,'output/images/Fig2.plotlikert-study1.png') # save image
-
+  # dev.copy(png,'output/images/Fig2.plotlikert-study1.png', width = 800, height = 600) # save image
   if (!require("pacman")) install.packages("pacman")
@@
     cowplot::ggdraw()
-#dev.off()
+    # dev.off()
 ```
@@
 In terms of the association between the indicators, Figure \@ref(fig:corr-issp) shows the polychoric correlation matrix.
-Firstly, we observe in general that the moderat to high associations are between the pairs of items representing each of the dimensions as described in \@ref(tab:table-issp).
-The exception here is the dimension of non-meritocratic perception, in which there are six items which appear mostly associated by pairs according to their specific topics (family background, personal background and connections).
-Still, between this same set of items there are moderate correlations (around 0.3), which could anticipate a single latent factor underlying non meritocratic perception and that is tested next through confirmatory factor analysis.
+Firstly, we observe in general that the moderate to high associations are between the pairs of items representing each of the dimensions as described in \@ref(tab:table-issp).
+The exception here is the dimension of non-meritocratic perception, in which there are six items that appear mostly associated by pairs according to their specific topics (family background, personal background, and connections).
+Still, between this same set of items, there are moderate correlations (around 0.3) which could anticipate a single latent factor underlying non-meritocratic perception and that is tested next through confirmatory factor analysis.
 
 ```{r corr-issp,fig.cap = "Perceptions and preferences for ISSP  meritocracy items' polychoric correlations", fig.align='center',fig.height=6,fig.width=8}
-
-#dev.copy(png,'output/images/Fig3.plotcor-sutdy1.png') # save image
-
+  # dev.copy(png,'output/images/Fig3.plotcor-study1.png', width = 800, height = 600) # save image
   load(file = "input/data/proc/data_s1.RData")
@@
     na.label = "-")
-
-#dev.off()
-
+    # dev.off()
 ```
@@
 First, a four-dimensional model for meritocratic and non-meritocratic perception and preferences is estimated, considering all indicators of the non-meritocratic perception dimension as a single latent variable.
-Then, we estimate a model that follows the proposed four latent variable measurement approach, however, the non-meritocratic perception dimension is assumed as a second-order factor for the first order latent variables parental attributes, racial and gender background and networks.
+Then, we estimate a model that follows the proposed four latent variable measurement approach, however, the non-meritocratic perception dimension is assumed as a second-order factor for the first-order latent variables parental attributes, racial and gender background, and networks.
 
@@
 Further analysis showed that the sources for poor fit were mostly related to the items of the non-meritocratic perceptions, which as mentioned above displayed correlations among them not taken into account in this model specification.
-Attending to this constraints, we specified a second model that keeps the basic four-factor structure but generates an additional model for the non-meritocratic perceptions as depicted in \@ref(fig:meas02-issp).
+Attending to these constraints, we specified a second model that keeps the basic four-factor structure but generates an additional model for the non-meritocratic perceptions as depicted in \@ref(fig:meas02-issp).
 In this model, the pairs of items in this dimension form three latent factors which at the same time generates a second-order factor of non-meritocratic perceptions, improving the fit of the model.
@@
 Something similar occurs for the correlation between the two dimensions of preferences ($r=0.25$).
-From this information we can preliminary conclude that the wide use of reverse-coding for non-meritocratic items attempting to measure meritocracy is not an adequate operationalization of the construct.
-All in all, despite the good fit of the second order model there are still several limitations regarding the content of the items which we attempt to overcome next by developing a new instrument.
+From this information, we can preliminary conclude that the wide use of reverse-coding for non-meritocratic items attempting to measure meritocracy is not an adequate operationalization of the construct.
+All in all, despite the good fit of the second-order model there are still several limitations regarding the content of the items which we attempt to overcome next by developing a new instrument.
 
@@
 
-## Data {-#data}
+## Data
 
@@
 The sample was selected from a non-probabilistic quota design in three large cities in Chile (Santiago, Concepción & Antofagasta).
-The quotas for gender, age, and educational levels were generated based on a survey by the Public Studies Center [@cep_encuesta_2019], which is a well regarded counterpart agency of the ISSP in Chile.
-A total sample of 2,141 people was collected, excluding those who did not signed the informed consent as well as those not answering the meritocracy response battery.
+The quotas for gender, age, and educational levels were generated based on a survey by the Public Studies Center [@cep_encuesta_2019], which is a well-regarded counterpart agency of the ISSP (International Social Survey Programme) in Chile.
+A total sample of 2,141 people was collected, excluding those who did not sign the informed consent as well as those not answering the meritocracy response battery.
 There were no significant differences between our sample and the wider population for most socio-demographic characteristics, with the exception of educational level (see Table \@ref(tab:rep-samp) in Appendix).
@@
 In order to achieve at least some comparability with previous studies, the questions were adapted from the items battery "reasons to get ahead" (ISSP/GSS), which has been widely used for operationalizing meritocracy in previous studies [@mijs_paradox_2019; @duru-bellat_who_2012; @reynolds_perceptions_2014].
-The aforementioned eight items, ordered according to dimensions, are presented in Table \@ref(tab:table-indicadores).
+The aforementioned eight items ordered according to dimensions are presented in Table \@ref(tab:table-indicadores).
 These eight likert-type items have five response alternatives, ranging from "Completely disagree"(1) to "Completely agree" (5).
@@
  tabitems <- read.csv(file = "input/tables/table01.csv",header = 1,sep = ",",encoding = "UTF-8") # call file  generated externally for the table
- cnames <- c("Component", "Dimensions","Item (english)" , "Item original (spanish)")
+ cnames <- c("Component", "Dimensions","Item (English)" , "Item original (Spanish)")
  cap <- "Items of the perceptions and preferences for meritocracy scale."
@@
 For the second group (_n = 717_), the order was reorganized according to perceptions and preferences over the same topic, e.g.
-for the topic of hard work, the item about perception was followed by the item about preference, and the same for the rest of the topics.
+for the topic of hard work, the item about perception was followed by the item about preference and the same for the rest of the topics.
 Finally, for the third group (_n = 712_), the items were presented as completely randomized.
@@
 The models estimated one factor for each of the four proposed dimensions presented in Table \@ref(tab:table-indicadores).
-As in Study 1, CFA was conducted using the `lavaan` R package (version 0.6-3; @rosseel_lavaan_2012), with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement [@kline_principles_2016; @rosseel_lavaan_2012].
+As in Study 1, CFA was conducted using the `lavaan` R package (version 0.6-3; @rosseel_lavaan_2012), with diagonally weighted least squares (DWLS) estimation due to the items' ordinal level of measurement [@kline_principles_2016; @rosseel_lavaan_2012].
 The fit indexes and cut-off criteria were the same as the ones used in Study 1.
@@
 The graphs presented in Figure \@ref(fig:plotlikert) display disaggregated and comparable information of the different response categories for each item.
-As it can be observed, in general there is more agreement in the perception of non-meritocratic items than in meritocratic ones, while in the case of preferences the opposite occurs.
+As it can be observed, in general, there is more agreement in the perception of non-meritocratic items than in meritocratic ones, while in the case of preferences the opposite occurs.
 As far as preferences are concerned, the preponderant role of effort over talent as a criterion of meritocratic preference is noteworthy.
@@
 ```{r plotlikert, echo=FALSE, fig.align='center', fig.cap="Distribution of responses in the Merit Scale items", fig.height=8, fig.width=11}
-
-#dev.copy(png,'output/images/Fig6.plotlikertmerit-scale.png') # save image
-
+  # dev.copy(png,'output/images/Fig6.plotlikert-study2.png', width = 820, height = 600)
   if (!require("pacman")) install.packages("pacman")
@@
     cowplot::ggdraw()
-  #dev.off()
-
-```
-
-Attending now to the association among the scale items, Figure \@ref(fig:corpoly) shows items polychoric correlations.
+  # dev.off()
+```
+
+Attending now to the association among the scale items, Figure \@ref(fig:corpoly) shows the items' polychoric correlations.
 There are three main aspects to highlight from this correlation matrix.
@@
 ```{r corpoly, echo=FALSE, fig.cap = "Perceptions and preferences for meritcoracy items' polychoric correlations", fig.align='center'}
-
-#dev.copy(png,'output/images/Fig6.plotcor-sutdy2.png') # save image  
+  # dev.copy(png,'output/images/Fig7.plotcor-study2.png',width = 800, height = 600) # save image
   if (!require("pacman")) install.packages("pacman")
@@
     na.label = "-")
-#dev.off()
+  # dev.off()
 ```
@@
 
-The present sections shows the results of the confirmatory factor analysis estimation.
-The model estimates four latent factors: perception meritocratic, perception non-meritocratic, preferences meritocratic and preferences non-meritocratic.
+The present section describes the results of the confirmatory factor analysis estimation.
+The model estimates four latent factors: perception meritocratic, perception non-meritocratic, preferences meritocratic, and preferences non-meritocratic.
 Each factor is estimated based on two items of the scale as detailed in Table \@ref(tab:table-indicadores).
@@
 Firstly we analyzed the convergent validity by exploring the association of the merit scale with related concepts and its correspondent measurements.
-Secondly, we test the internal consistency of the measurement model behind the scale comparing with a different sample through measurement invariance procedures.
+Secondly, we test the internal consistency of the measurement model behind the scale comparing it with a different sample through measurement invariance procedures.
 
@@
 
-For this analysis we examined data from the last of the three waves of the online panel survey used in Study 2.
+For this analysis, we examined data from the last of the three waves of the online panel survey used in Study 2.
 This wave included additional measures that allow testing the correlations of the merit scale with some related constructs as opportunity beliefs and personal wherewithal.
-After listwise deletion of missing cases in socio-demographics information, a total of 1422 individuals took part in the third wave: 668 (46.97%) women and 754 (53.03%) men.
+After a listwise deletion of missing cases in socio-demographic information, a total of 1422 individuals took part in the third wave: 668 (46.97%) women and 754 (53.03%) men.
 The ages of 50.57% of them are 45 years or older, and 36.28% held a tertiary degree.
@@
 
-The following scales were included for testing their correlations with the the 8-item meritocracy scale:
+The following scales were included for testing their correlations with the 8-item meritocracy scale:
 
@@
 
-- Personal wherewithal: the questionnaire included The Neoliberal Beliefs Inventory (NBI), which considers four factors: Government Interference preferences, Competition preferences, System Inequality perception and Personal Wherewithal [@bay-cheng_tracking_2015].
+- Personal wherewithal: the questionnaire included The Neoliberal Beliefs Inventory (NBI), which considers four factors: Government Interference preferences, Competition preferences, System Inequality perception, and Personal Wherewithal [@bay-cheng_tracking_2015].
 We used the factor _personal wherewithal_ that reflects meritocracy beliefs in terms of the importance of personal attributes as strength and skills to yield success.
-The construct consists of 8 items (e.g., “Any goal can be achieved with enough hard work and talent”, “I’ve benefited from working hard, so there’s no reason others can’t”, “Anyone who is willing to work hard can be succesful in Chile”).
+The construct consists of 8 items (e.g., "Any goal can be achieved with enough hard work and talent", "I've benefited from working hard, so there's no reason others can't", "Anyone who is willing to work hard can be successful in Chile").
 Participants answered the items using a 1 (totally disagree) to 6 (totally agree) scale.
-We simple computed mean scores, with higher scores indicating stronger support for meritocracy (M=3.32, SD=0.88, $\alpha$ = 0.89).
+We simply computed mean scores, with higher scores indicating stronger support for meritocracy (M=3.32, SD=0.88, $\alpha$ = 0.89).
 
 Both instruments cover mainly the perceptual side of the concept.
-Therefore, in general we expect larger correlation with meritocratic perceptions than with preferences.
-In the case of opportunity beliefs, the hard work item should correlate positively with meritocratic perception whereas the importance for coming from a wealthy family is expected to correlate positively with non-meritocratic perceptions.
+Therefore, in general, we expect a larger correlation with meritocratic perceptions than with preferences.
+In the case of opportunity beliefs, the hard work item should correlate positively with meritocratic perception whereas the importance of coming from a wealthy family is expected to correlate positively with non-meritocratic perceptions.
 Regarding the second instrument of personal wherewithal, we also anticipate a positive correlation with meritocratic perception and low to null correlations with the other dimensions of the merit scale.
@@
 
-We used polychoric correlations to evaluate relationships of meritocractic and non-meritocratic preferences and perceptions with the items of the opportunity beliefs battery.
+We used polychoric correlations to evaluate relationships of meritocratic and non-meritocratic preferences and perceptions with the items of the opportunity beliefs battery.
 For personal wherewithal, we examined Pearson correlations.
@@
 The NBI's factor depicts a positive association with unmeritocratic preference ($r=0.225$).
-This finding is consistent with literature, inasmuch as the belief for meritocray is associated with the justification of inequality [@mccall_exposure_2017; @madeira_primes_2019].
+This finding is consistent with the literature, in as much as the belief for meritocracy is associated with the justification of inequality [@mccall_exposure_2017; @madeira_primes_2019].
 Furthermore, results of our confirmatory factor analysis indicate that meritocratic perceptions and unmeritocratic preferences show a positive and weak correlation.
@@
 
-The first step for invariance testing is the estimation of the configural model, which serves as the baseline for further comparison and it is expected to adequately meet the global fit criteria of a measurement model.
+The first step for invariance testing is the estimation of the configural model, which serves as the baseline for further comparison and is expected to adequately meet the global fit criteria of a measurement model.
 Although the chi-square statistic is used as a global measure of fit, since its sensitivity to sample size it is recommended to use three additional fit indices: Comparative Fit Index (CFI), which should have a value greater than 0.95; Root Mean Square Error Approximation (RMSEA), which must be lower than 0.06; and the Standardized Root Mean Square Residuals (SRMR) which must be less than 0.08 [@vandeschoot_checklist_2012].
@@
 Such result is considered in general as evidence of invariance [@fischer_are_2011], as strict forms of measurement invariance rarely hold [@vandeschoot_checklist_2012].
-Still, the comparability of latents means requires strict invariance which in this case does not hold when considering $\Delta \chi^2 (\Delta \text{df})$.
-Nevertheless, the criteria of $\Delta \text{CFI}$ used for comparing models is close to the rejection criteria of >.01, whereas the $\Delta \text{RMSEA}$ fulfills the requirements of being below of the cut-off criteria as suggested by [@chen_sensitivity_2007].
+Still, the comparability of latent means requires strict invariance which in this case does not hold when considering $\Delta \chi^2 (\Delta \text{df})$.
+Nevertheless, the criteria of $\Delta \text{CFI}$ used for comparing models is close to the rejection criteria of >.01, whereas the $\Delta \text{RMSEA}$ fulfils the requirements of being below the cut-off criteria as suggested by [@chen_sensitivity_2007].
 Therefore, using this last standard the level for strict invariance holds for the meritocracy scale.
@@
 
-We identified four critical aspects regarding the measurement of meritocracy in previous studies: unidimensionality, the ambiguous use of the term “beliefs”, the use of non-meritocratic indicators as opposed to meritocratic indicators, and the consideration of measurement error.
-The proposed 8-item scale, “Perceptions and Preferences for Meritocracy”, was designed and tested in order to deal with these four issues.
+We identified four critical aspects regarding the measurement of meritocracy in previous studies: unidimensionality, the ambiguous use of the term "beliefs", the use of non-meritocratic indicators as opposed to meritocratic indicators, and the consideration of measurement error.
+The proposed 8-item scale, "Perceptions and Preferences for Meritocracy", was designed and tested in order to deal with these four issues.
 Our results indicate that perceptions and preferences seem to be two related but different dimensions, often confounded in previous research under the label of "beliefs".
 Meritocratic and non-meritocratic dimensions do not appear to constitute poles of the same continuum, as some previous studies have assumed.
-Regarding the possible effects of items order in the estimation of latent variable, we tested three different order versions and found evidence that suggest the use of the randomized version of the scale.
+Regarding the possible effects of the items' order in the estimation of the latent variable, we tested three different order versions and found evidence that suggests the use of the randomized version of the scale.
 Furthermore, we found evidence of convergent validity as well as measurement invariance that approach well the requirement of comparability between samples.
 
-The four-dimensional conceptual framework and its operationalization in the Perceptions and Preferences for Meritocracy scale opens several avenues for future research.
+The four-dimensional conceptual framework and its operationalization in the Perceptions and Preferences for Meritocracy scale open several avenues for future research.
 For instance, distinguishing perceptions from preferences will allow us to evaluate the extent to which different societies are accustomed to, or satisfied with, the perceived level of meritocracy, in terms of differences between what is perceived and what is preferred.
@@
 Furthermore, the impact of different configurations of the four-dimensional framework on practices and behaviors such as corruption, civic involvement, and political alignment, is an area that requires additional research.
-Such future agendas could be specially relevant in times of economic crisis and growing inequalities, that could entail changes in the legitimation of the current distributive structure based on meritocratic ideals.
+Such future agendas could be especially relevant in times of economic crisis and growing inequalities, that could entail changes in the legitimation of the current distributive structure based on meritocratic ideals.
 

### 2021-09-23 paper.Rmd

@@
 
-_b. Beliefs?_
+b. _Beliefs?_
 
@@
 
-_c. Non-meritocratic aspects_
+c. _Non-meritocratic aspects_
 
@@
 
-_d. Accounting for measurement error_
+d. _Accounting for measurement error_
 
@@
         caption = "Multiple Group meassurement invariance for Perceptions and Preferences for Meritocracy") %>%
-    kableExtra::kable_styling(full_width = FALSE,latex_options = "HOLD_position",bootstrap_options=c("striped", "bordered")) %>%
+    kableExtra::kable_styling(full_width = FALSE,
+                              latex_options = "HOLD_position",
+                              bootstrap_options=c("striped", "bordered"),
+                              font_size = 10) %>%
     kableExtra::footnote(general = footnote, footnote_as_chunk = T)

### 2021-10-07 paper-blinded.Rmd
version final sometida a SJR, es paper-blinded, y la información de autores se traspasa a title-page

--- 2021-09-23 paper.Rmd
@@
 link-citations: yes
-author: "Juan Carlos Castillo, Julio Iturra, Luis Maldonado, Francisco Meneses & Jorge Atria <br><br>"
+# author: "Juan Carlos Castillo, Julio Iturra, Luis Maldonado, Francisco Meneses & Jorge Atria <br><br>"
 abstract: "Economic and social inequalities have generated growing concerns and crises across contemporary societies.
 One of the mechanisms proposed by social sciences to explain the persistence of inequality is the belief in meritocracy, which would legitimize economic disparities based on differences in effort and talent.
-Despite its wide use as a concept, empirical research on meritocracy is relatively novel and characterized by diverse conceptualization and measures that make the findings and their interpretation rather inconsistent.
+Despite its wide use as a concept, empirical research on meritocracy is relatively novel and characterized by diverse conceptualizations and measures that make the findings and their interpretation rather inconsistent.
 Most of the studies in the area have relied upon secondary data to operationalize meritocracy, with a wide variation in the use and interpretation of the same survey items.
-Taking into account the extant literature that uses measures of meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding conceptualization and operationalization.
-Based on this critical analysis, we propose a conceptual framework for studying perceptions and preferences for meritocracy and non-meritocracy, which is then tested through confirmatory analysis using ISSP (International Social Survey Programme) data as well as a novel survey designed specifically with this purpose (N=2,141).
+Taking into account the extant literature that uses measures of meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding the conceptualization and operationalization of meritocracy beliefs.
+Based on this critical analysis, we propose a conceptual framework for measuring perceptions and preferences for meritocracy and non-meritocracy, which is then tested through confirmatory analysis using ISSP (International Social Survey Programme) data as well as a novel scale designed with this purpose (N=2,141).
 Our results support the conceptual framework and its operationalizarion, although with a better fit for the proposed scale than for the ISSP survey.
@@
 
+<!---
 \pagebreak
@@
 \pagebreak
+
+--->
 
@@
 The question of whether talent (as intelligence and ability) is or is not considered meritocratic is certainly an interesting topic to discuss from a philosophical point of view, but for those working on empirical studies, it should face empirical scrutiny.
-Besides, the consideration of talent as part of meritocracy certainly opens some interesting avenues of research, for instance in studies showing that for the elites meritocracy is more related to talent, whereas effort is something more characteristic of the meritocracy of the middle and low classes [@atria_economic_2020].
+Besides, the consideration of talent as part of meritocracy certainly opens some interesting avenues of research, as for instance some studies suggest that for the elites meritocracy is actually related to talent, whereas effort is something more characteristic of the meritocracy of the middle and low classes [@atria_economic_2020].
 
@@
 Several approaches to the empirical study of meritocracy based on public opinion surveys refer to the concept of _beliefs_, but with wide differences in meanings and operationalizations.
-To illustrate this point let us consider the most commonly used items for operationalizing meritocratic beliefs in survey research (as ISSP), which is the "reasons to get ahead" battery.
-It consists of a series of items asking "how important you think it is for getting ahead in life" and then lists several factors, such as effort, education, parental wealth, and contacts.
-Another version of this same battery used in other surveys - sometimes together with the previous one - asking about "how important you think it _should_ be ...", and then listing the same concepts.
-Therefore, the question raised here is: Which one of both is a "belief": what _is_ or what it _should_ be?
+To illustrate this point let us consider the commonly used "reasons to get ahead" battery, mentioned above.
+Some versions of this battery ask "how important you think it is for getting ahead in life" and then lists several factors, whereas another version of this same battery - sometimes presented along the previous one - asks about "how important you think it _should_ be ...", and then listing the same concepts.
+Therefore, the question raised here is: Which one of both is a "belief": what _is_ (the first one) or what it _should_ be (the second one)?
 
@@
 As @kluegel_beliefs_1986 pointed out about the scope of beliefs: "This usage encompasses such more specific social-psychological concepts as values, perceptions, and attitudes" (p.30).
-Therefore, beliefs cover almost anything related to subjective factors.
+Therefore, beliefs use to cover almost anything related to subjective factors.
 To this regard, a relevant distinction in the field of inequality beliefs was made by @janmaat_subjective_2013: "Perceptions refer to subjective estimates of existing inequality (i.e.
@@
 They considered education, ambition, and hard work as meritocratic, and other factors such as family wealth and connections as non-meritocratic.
-Despite making and proving this distinction, however, the authors end up subtracting one dimension from the other, thus coming back to the assumption that they are two poles of the same continuum as @kunovich_systems_2007 did.
+Despite making and proving this distinction, however, the authors end up subtracting one dimension from the other, thus coming back to the assumption that they are two poles of the same continuum, as @kunovich_systems_2007 also did.
 Similarly, @roex_attitudes_2018 used ISSP indicators for perceived meritocracy and non-meritocracy to build a single score by reverse coding the non-meritocratic items.
@@
 In choosing this strategy of index construction, we argue that support for meritocracy is not a latent variable" [@kunovich_systems_2007 p.653-654].
-Although some advances were made by @reynolds_perceptions_2014 through conducting a principal component analysis of meritocratic and non-meritocratic dimensions, somewhat surprisingly they finally chose to build a sum index despite proving a multidimensional latent structure.
-
-## Conceptual proposal for studying meritocracy {-#instrumentprop}
+Although some advances were made by @reynolds_perceptions_2014 through conducting a principal component analysis of meritocratic and non-meritocratic dimensions, they finally chose to build a sum index despite proving a multidimensional latent structure.
+
+## A conceptual proposal for studying meritocracy {-#instrumentprop}
 
@@
 
-```{r merit-model, echo=FALSE, fig.cap = "Model of perception and preferences for meritocracy and non-meritocracy", fig.align='center'}
+```{r merit-model, echo=FALSE, fig.cap = "Conceptual Model of perception and preferences for meritocracy and non-meritocracy", fig.align='center'}
  knitr::include_graphics("input/images/Fig1.generalf.png")
@@
 
-There are a series of indicators in the ISSP survey that in the following we attempt to classify in our meritocracy measurement scheme.
-However, it is important to mention that the items were not originally designed as a specific measure of meritocracy.
-The variables are presented below organized in perceptions and preferences:
+There are a series of indicators in the ISSP survey that in the following we attempt to classify in our meritocracy conceptual scheme, as presented below in Table \@ref(tab:table-issp):
 
@@
 
-First, a four-dimensional model for meritocratic and non-meritocratic perception and preferences is estimated, considering all indicators of the non-meritocratic perception dimension as a single latent variable.
-Then, we estimate a model that follows the proposed four latent variable measurement approach, however, the non-meritocratic perception dimension is assumed as a second-order factor for the first-order latent variables parental attributes, racial and gender background, and networks.
+Table \@ref(tab:sum-fit-issp) shows the results of the estimation of two confirmatory models.
+The first one (First Order) corresponds to a model that estimate four factors, each for one of the dimensions presented in \@ref(tab:table-issp), showing only regular fit indicators (CFI=0.959, TLI=0.944, RMSEA=0.098, $\chi^2$(df=48)= 21308.535; more detail in Table \@ref(tab:tb-perpref1)).
+Further analysis showed that the sources for poor fit were mostly related to the items of the non-meritocratic perceptions, which as mentioned above displayed correlations among them not taken into account in this model specification.
 
@@
 
-Table \@ref(tab:sum-fit-issp) shows the results of the estimation of two confirmatory models.
-The first one (First Order) corresponds to a model that estimate four factors, each for one of the dimensions presented in \@ref(tab:table-issp), showing only regular fit indicators (CFI=0.959, TLI=0.944, RMSEA=0.098, $\chi^2$(df=48)= 21308.535; more detail in Table \@ref(tab:tb-perpref1)).
-Further analysis showed that the sources for poor fit were mostly related to the items of the non-meritocratic perceptions, which as mentioned above displayed correlations among them not taken into account in this model specification.
-Attending to these constraints, we specified a second model that keeps the basic four-factor structure but generates an additional model for the non-meritocratic perceptions as depicted in \@ref(fig:meas02-issp).
+Attending to the sources of poor fit from the first estimated confirmatory model, we specified a second model that keeps the basic four-factor structure but generates an additional model for the non-meritocratic perceptions as depicted in \@ref(fig:meas02-issp).
 In this model, the pairs of items in this dimension form three latent factors which at the same time generates a second-order factor of non-meritocratic perceptions, improving the fit of the model.
 
-```{r meas02-issp, echo=FALSE, fig.cap = "Descriptive plot", fig.align='center',fig.cap="Confirmatory factor analysis of the ISSP indicators of Perceptions and Preferences for Meritocracy"}
+```{r meas02-issp, echo=FALSE, fig.cap = "Descriptive plot", fig.align='center',fig.cap="Second-order confirmatory factor analysis model using ISSP indicators of Perceptions and Preferences for Meritocracy"}
  knitr::include_graphics('output/images/Fig4.cfa_issp_study1.PNG')
@@
 From this information, we can preliminary conclude that the wide use of reverse-coding for non-meritocratic items attempting to measure meritocracy is not an adequate operationalization of the construct.
-All in all, despite the good fit of the second-order model there are still several limitations regarding the content of the items which we attempt to overcome next by developing a new instrument.
+All in all, despite the good fit of the second-order model there are still several limitations regarding the content of the items which we attempt to overcome next by developing and testing new scale for measuring meritocracy.
 
@@
 The items of the meritocracy scale were identical to the original.
-The items were administrated in a randomized order (which corresponds to the application modality of Group 3 present in Figure \@ref(fig:appmod)).
+The items were administrated in a randomized order (which corresponds to the application modality of Group 3 as showed in Figure \@ref(fig:appmod)).
 
@@
 
-Table \@ref(tab:fcloads-inv) shows the results of the CFA estimation of the meritocracy scale in this new sample.
+Table \@ref(tab:fcloads-inv) in the Appendix shows the results of the CFA estimation of the meritocracy scale in this new sample.
 As observed, the loadings and the fit indices show similar results as the ones presented for the original study above, which gives a first base from which to start the invariance analysis.
@@
 Nevertheless, the criteria of $\Delta \text{CFI}$ used for comparing models is close to the rejection criteria of >.01, whereas the $\Delta \text{RMSEA}$ fulfils the requirements of being below the cut-off criteria as suggested by [@chen_sensitivity_2007].
-Therefore, using this last standard the level for strict invariance holds for the meritocracy scale.
+Therefore, using this last standard, the level for strict invariance would hold for the meritocracy scale.
 
@@
 
-The four-dimensional conceptual framework and its operationalization in the Perceptions and Preferences for Meritocracy scale open several avenues for future research.
+The four-dimensional conceptual framework and its operationalization in the Perceptions and Preferences for Meritocracy scale opens several avenues for future research.
 For instance, distinguishing perceptions from preferences will allow us to evaluate the extent to which different societies are accustomed to, or satisfied with, the perceived level of meritocracy, in terms of differences between what is perceived and what is preferred.
@@
 
+# Transparency statement
+
+This research follows a series of open science guidelines as the pre-registration of Study 2 (scale development) as well as the availability of data and codes (R) in an public repository in Github.
+This document was generated with the R package Rmarkdown and contains the code for all the tables and figures in order to make it reproducible.
+
 # References
@@
 
-# (APPENDIX) Appendix {-}
+# Appendix {-}
 
@@
 
-## Appendix study 1
+**Study 1**
 
@@
 
-## Appendix study 3
+**Study 3**
 

### 2021-10-07 paper.Rmd
version final sometida a SJR, es paper-blinded, y la información de autores se traspasa a title-page

@@
 link-citations: yes
-author: "Juan Carlos Castillo, Julio Iturra, Luis Maldonado, Francisco Meneses & Jorge Atria <br><br>"
+# author: "Juan Carlos Castillo, Julio Iturra, Luis Maldonado, Francisco Meneses & Jorge Atria <br><br>"
 abstract: "Economic and social inequalities have generated growing concerns and crises across contemporary societies.

### 2021-12-22 castillo-et-al-2021-prerint-meritocracy-scale.Rmd
generate preprint for socarxiv

--- 2021-10-07 paper-blinded.Rmd
@@
 link-citations: yes
-# author: "Juan Carlos Castillo, Julio Iturra, Luis Maldonado, Francisco Meneses & Jorge Atria <br><br>"
+date: "September 2021"
+author: 
+  - Juan Carlos Castillo, Universidad de Chile - jc-castillo.com
+  - Julio Iturra, COES
+  - Luis Maldonado, P. Universidad Católica de Chile
+  - Francisco Meneses, Universidad de Chile
+  - Jorge Atria, Universidad Diego Portales
 abstract: "Economic and social inequalities have generated growing concerns and crises across contemporary societies.
@@
 header-includes:
+  - \renewcommand{\and}{\\}  # this is for line breaks between authors 
   - \usepackage{times}           # Times New Roman
@@
 ```{r setup, include=FALSE}
- knitr::opts_chunk$set(echo=FALSE, warning = FALSE,message = FALSE, cache = FALSE,out.width = '85%',fig.pos= "H", fig.align = 'center')
+ knitr::opts_chunk$set(echo=FALSE, warning = FALSE,message = FALSE, cache = TRUE,out.width = '85%',fig.pos= "H", fig.align = 'center')
  # knitr::opts_knit$set(base.url = "../") #relative path for .html output file
@@
 
-<!---
-\pagebreak
-
 # Acknowledgments
 
-The authors recognize the support of ANID/FONDECYT grant 1160921 and of  the Center for Conflict and Social Cohesion Studies (COES)  grant ANID/FONDAP 1513009.
-
-\pagebreak
-
---->
+The authors recognize the support of ANID/FONDECYT grant 1210847 and of  the Center for Conflict and Social Cohesion Studies (COES) grant ANID/FONDAP 1513009.
 

### 2022-05-19 paper-blinded.Rmd
ediciones SIR

@@
   - \usepackage[figurename=Fig.]{caption}
+  - \usepackage[labelfont=bf]{caption}
   - \usepackage{graphicx}
@@
  # clean #in the yml
- rmarkdown::render("paper.Rmd", output_format = "bookdown::pdf_document2")
- rmarkdown::render("paper.Rmd", output_format = "bookdown::html_document2")
+ rmarkdown::render("paper-blinded.Rmd", output_format = "bookdown::pdf_document2")
+ rmarkdown::render("paper-blinded.Rmd", output_format = "bookdown::html_document2")
  # rmarkdown::render("paper.Rmd", output_format = "bookdown::word_document2")
@@
 
-1. **Configural:** the model is estimated only indicating the factorial structure used in the CFA.
-
-2. **Weak:** an equality restriction is applied to the factor loadings in the different groups, that is, the loadings are forced to be identical in both measurements.
-
-3. **Strong:** equality restrictions are added to the intercepts of each indicator.
-
-4. **Strict:** equality restrictions are added to the error variances of each indicator.
+1. *Configural:* the model is estimated only indicating the factorial structure used in the CFA.
+
+2. *Weak:* an equality restriction is applied to the factor loadings in the different groups, that is, the loadings are forced to be identical in both measurements.
+
+3. *Strong:* equality restrictions are added to the intercepts of each indicator.
+
+4. *Strict:* equality restrictions are added to the error variances of each indicator.
 

### 2022-12-26 papelera/previous-submission/sjr/Title Page.Rmd
modificaciones paper-blinded para nuevo envío a SIR - primera parte hasta methods

---
title: "Measuring Perceptions and Preferences for Meritocracy"
linestretch: '1.5'
output:
  bookdown::pdf_document2:
     template: null
     toc: false
     number_sections: false
geometry: margin=0.78in
header-includes:
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage[figurename=Fig.]{caption}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
fontsize: 10pt
---

Juan Carlos Castillo, Department of Sociology, Universidad de Chile (
juancastillov@uchile.cl - [jc-castillo.com](https://juancarloscastillo.github.io/jc-castillo/) -  Ignacio Carrera Pinto 1045 Ñuñoa, Santiago de Chile )

Luis Maldonado, Institute of Sociology, Pontificia Universidad Católica de Chile

Jorge Atria, Department of Sociology, Universidad Diego Portales

Julio Iturra, Center for Social Conflict and Cohesion Studies

Francisco Meneses, Department of Sociology, Universidad de Chile

### 2022-12-26 paper-blinded.Rmd
modificaciones paper-blinded para nuevo envío a SIR - primera parte hasta methods

--- 2021-10-07 paper-blinded.Rmd
@@
 abstract: "Economic and social inequalities have generated growing concerns and crises across contemporary societies.
-One of the mechanisms proposed by social sciences to explain the persistence of inequality is the belief in meritocracy, which would legitimize economic disparities based on differences in effort and talent.
+The belief in meritocracy has been suggested as one of the factors that explain the persistence of inequality, as it would legitimize economic disparities based on differences in effort and talent.
 Despite its wide use as a concept, empirical research on meritocracy is relatively novel and characterized by diverse conceptualizations and measures that make the findings and their interpretation rather inconsistent.
-Most of the studies in the area have relied upon secondary data to operationalize meritocracy, with a wide variation in the use and interpretation of the same survey items.
-Taking into account the extant literature that uses measures of meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding the conceptualization and operationalization of meritocracy beliefs.
-Based on this critical analysis, we propose a conceptual framework for measuring perceptions and preferences for meritocracy and non-meritocracy, which is then tested through confirmatory analysis using ISSP (International Social Survey Programme) data as well as a novel scale designed with this purpose (N=2,141).
-Our results support the conceptual framework and its operationalizarion, although with a better fit for the proposed scale than for the ISSP survey.
+Most of the studies in the area have relied upon secondary data from different surveys to operationalize meritocracy, with a wide variation in the use and interpretation of the same measuring items.
+Based on a critical analysis of concepts and operationalizations of meritocracy in survey research, we propose a conceptual framework for measuring perceptions and preferences for meritocracy and non-meritocracy in a minimalistic version (few items as possible), in order to be considered in comparative survey research studies.
+The proposal is first tested through confirmatory factor analysis (CFA) using available ISSP (International Social Survey Programme) inequality module items.
+Based on these results, in a second stage we developed a improved scale that better suits the perceptions/preferences distinction, which was applied in a Chilean sample (N=2,141).
+The results of the CFA estimation give evidence of a better fit for the new scale.
 Our discussion highlights the importance of considering different dimensions in order to advance in the study of meritocracy."
+
 output:
@@
 
-Economic inequality and income concentration have become topics of growing concern over the last years, acquiring even more salience in times of pandemic and economic crises.
-This has led to a series of social movements in different societies as well as diverse critical analyses regarding the development of capitalism and its consequences [@streeck_politics_2014; @piketty_capital_2014].
-In this context, the study of views, preferences, and perceptions of inequality has acquired relevance in the social sciences, in topics such as redistributive preferences [@alesina_fairness_2005; @dimick_models_2018], the legitimization of economic inequality [@schroder_income_2017] and the functioning of meritocracy [@duru-bellat_who_2012; @mijs_paradox_2019; @reynolds_perceptions_2014; @atria_economic_2020].
-Within this context, the ideal of meritocracy has been strongly challenged as an unfulfilled promise of modern societies that allows the perpetuation of social inequalities [@sandel_tyranny_2020; @goldthorpe_myth_2003].
-
-In general, meritocracy is defined as a system of distribution of resources and rewards based on individual merit, which in its original conception is a combination of talent and effort [@young_rise_1962].
+Economic inequality and income concentration have become topics of growing concern over the last years, leading to a series of social movements in different societies as well as diverse critical analyses regarding the development of capitalism and its consequences [@streeck_politics_2014; @piketty_capital_2014].
+In such a context, the study of views, preferences, and perceptions of inequality has acquired relevance in the social sciences, in topics such as redistributive preferences [@alesina_fairness_2005; @dimick_models_2018], the legitimization of economic inequality [@schroder_income_2017] and the functioning of meritocracy [@duru-bellat_who_2012; @mijs_paradox_2019; @reynolds_perceptions_2014; @atria_economic_2020].
+Within this research area, and particularly in times of economic and health crises, the ideal of meritocracy has been strongly challenged as an unfulfilled promise of modern societies that allows the perpetuation of social inequalities [@sandel_tyranny_2020; @goldthorpe_myth_2003].
+
+In general, meritocracy has been defined as a system of distribution of resources and rewards based on individual merit, which in its original conception is a combination of talent and effort [@young_rise_1962].
 This traditional conception of merit places in a secondary position the possible interference of structural or non-meritocratic factors, such as inheritance, personal contacts, and luck [@breen_class_1999; @saunders_might_1995; @yair_meritocracy_2007; @land_we_2006; @young_rise_1962].
@@
 
-Based on a critical analysis of different approaches to the measurement of meritocracy to date, this article proposes an instrument to both measure and relate two main aspects in the study of meritocracy: perceptions and preferences.
+Based on a critical analysis of different approaches to the measurement of meritocracy to date, this article proposes a conceptual and measurement framework that considers two of the main aspects in the study of meritocracy: perceptions and preferences.
 Furthermore, it distinguishes between meritocratic and non-meritocratic dimensions, as they are not necessarily two poles of the same continuum as previous studies seem to suggest.
-This conceptual framework is oriented to develop a measurement instrument as simple and brief as possible, being suitable to be used in public opinion surveys and allowing to integrate meritocratic beliefs in the study of different social phenomena.
+This conceptual framework is oriented to develop a measurement instrument as simple and brief as possible, being suitable to be used in public opinion surveys and thus allowing to integrate meritocratic beliefs in the study of different social phenomena.
 
@@
 Besides, the consideration of talent as part of meritocracy certainly opens some interesting avenues of research, as for instance some studies suggest that for the elites meritocracy is actually related to talent, whereas effort is something more characteristic of the meritocracy of the middle and low classes [@atria_economic_2020].
+Therefore, we propose that a concept (and measure) of meritocracy should include both talent and effort as in its original conception.
 
@@
 - Based on _previous indicators_ as much as possible, for the sake of keeping comparability between studies.
+As most of the survey research on meritocracy so far has been made with the inequality module of the ISSP survey, we attempt to rescue as much as possible the original wording of some of its items in the new scale proposal.
 
@@
 In this respect, it differs from the proposal of  "Preference for the Merit Principle Scale" [@davey_preference_1999], as they use 15 items for just one dimension (aside from the problem of reverse-coding non-meritocratic items).
+There is a clear trade-off when using a reduced amount of items: it increases the probability of use in public opinion surveys that usually have very limited space, but it reduces the measurement quality since this requires several items per dimension.
+As we will show later in the methods section, we opted by a minimalistic approach of two items per conceptual dimension.
+We discuss the limitations of such decision in the conclusions.
 

### 2022-12-29 paper-blinded.Rmd
agrega discusion a cada estudio, modifica estudio 3 y redaccion general

@@
 
-Some research in meritocracy considers aspects usually opposed to effort and talent for personal achievement, as for instance the use of personal advantages (as contacts or having a wealthy family) to get ahead in life.
+Some research in meritocracy considers aspects usually opposed to effort and talent for personal achievement, as for instance family status and the use of personal advantages (as contacts or having a wealthy family) to get ahead in life.
+This distinction refers to the classical achieved and ascribed status dimension from @linton_study_1936.
 For instance, @kunovich_systems_2007 used an items' battery listing a number of factors in relation to "How important each should be in deciding pay..." (as @duru-bellat_who_2012 for desired meritocracy).
@@
 reverse coded).
-The assumption behind this approach is that rejecting a supposed non-meritocratic aspect (as having family and/or children) implies a stronger belief in meritocracy.
+The assumption behind this approach is that rejecting a supposed ascribed or non-meritocratic^[Although the use of the concept "non-meritocratic" is very broad, we prefer to use it instead of something like ascription, given that this last one leaves out aspects such as the use of personal relationships.@@] aspect (as having family and/or children) implies a stronger belief in meritocracy.
 A similar approach of reverse-coding non-meritocratic items was taken by @newman_false_2015, using the same principles applied in the "Preference for the Merit Principle Scale" [@davey_preference_1999].
@@
 The assumption that meritocratic and non-meritocratic elements are poles of the same continuum was tested by @reynolds_perceptions_2014 using the "get ahead" perceptions' battery items mentioned above.
-They considered education, ambition, and hard work as meritocratic, and other factors such as family wealth and connections as non-meritocratic.
-Despite making and proving this distinction, however, the authors end up subtracting one dimension from the other, thus coming back to the assumption that they are two poles of the same continuum, as @kunovich_systems_2007 also did.
+They considered education, ambition, and hard work as meritocratic, whereas factors such as family wealth and connections where classified as non-meritocratic.
+Despite making this distinction, the authors end up subtracting one dimension from the other, thus coming back to the assumption that they are two poles of the same continuum, as @kunovich_systems_2007 also did.
 Similarly, @roex_attitudes_2018 used ISSP indicators for perceived meritocracy and non-meritocracy to build a single score by reverse coding the non-meritocratic items.
-Therefore, the treatment of non-meritocratic items has been rather inconsistent across studies and the assumption that they are the simple opposite of meritocracy certainly requires further empirical assessment.
+Therefore, the treatment of non-meritocratic items has been rather inconsistent across studies and the assumption that they are the simple opposite of meritocracy certainly requires further conceptual elaboration and empirical assessment.
+For instance, such distinction could open the door to study to what extent different and apparently contradictory views of meritocracy can coexist, to what extent individuals and societies differ to this regard, and of course the associated factors to such differences.
 
@@
 
+There is a relevant trade-off to consider when it comes to measurement error estimation in survey scales.
+Including multiple items per conceptual dimension opens the possibility to validity studies (at least in terms of construct) by using factor analysis techniques, particularly confirmatory ones.
+However, increasing the amount of items reduces the probability of usability of the scale in survey research given space limitations.
+This validity-usability conundrum is not easy to solve and, as we will show later in the methods section, we opted by a minimalistic approach oriented to the usability of the scale but that still gives the possibility of confirmatory factor analysis, although with some limitations.
+
 ## A conceptual proposal for studying meritocracy {-#instrumentprop}
@@
 - Based on _previous indicators_ as much as possible, for the sake of keeping comparability between studies.
-As most of the survey research on meritocracy so far has been made with the inequality module of the ISSP survey, we attempt to rescue as much as possible the original wording of some of its items in the new scale proposal.
+As most of the survey research on meritocracy so far has been made with the inequality module of the ISSP survey, we attempted to rescue as much as possible the original wording of some of its items in the new scale proposal.
 
 - _Brief_, as to be used in regular public opinion surveys.
-In this respect, it differs from the proposal of  "Preference for the Merit Principle Scale" [@davey_preference_1999], as they use 15 items for just one dimension (aside from the problem of reverse-coding non-meritocratic items).
-There is a clear trade-off when using a reduced amount of items: it increases the probability of use in public opinion surveys that usually have very limited space, but it reduces the measurement quality since this requires several items per dimension.
-As we will show later in the methods section, we opted by a minimalistic approach of two items per conceptual dimension.
-We discuss the limitations of such decision in the conclusions.
+To this regard, it differs from the proposal of  "Preference for the Merit Principle Scale" [@davey_preference_1999], as they use 15 items for just one dimension (aside from the problem of reverse-coding non-meritocratic items).
 
@@
 The rows in Figure \@ref(fig:merit-model) consider the distinction between meritocratic and non-meritocratic dimensions [@reynolds_perceptions_2014], usually treated as different ends of the same continuum in previous research.
-Non-meritocratic elements refer to the use of resources such as personal contacts or family advantages to getting ahead in life.
+Non-meritocratic elements refer to the use of resources such as personal contacts and/or family advantages to getting ahead in life.
 
@@
 The analysis is organized into three studies.
-Study 1 serves as a background to our instrument proposal by attempting to operationalize the meritocracy dimensions with available secondary data from the inequality module of the International Social Survey Programme (ISSP).
-Although we are aware that the available indicators in this dataset do not allow a precise operationalization of the conceptual model, we know that it is the most widely used data in comparative meritocracy studies (as well as overall subjective inequality) and therefore it is certainly useful to look at its potentials and limitations to this regard.
+Study 1 serves as a background to our instrument proposal by attempting to operationalize the meritocracy conceptual model with available secondary data from the inequality module of the International Social Survey Programme (ISSP).
+Although we are aware that the available indicators in this dataset do not allow a precise operationalization of the conceptual model, we know that it is the most widely used dataset in comparative meritocracy studies (as well as overall subjective inequality) and therefore it is certainly useful to look at its potentials and limitations to this regard.
 Study 2 is the core of this paper and consists of the proposal of a new scale for measuring meritocratic perceptions and preferences.
@@
 The data corresponds to the last available wave of the social inequality module from the International Social Survey Programme (ISSP), which is the most specialized international comparative survey in perceptions, attitudes, and beliefs about inequality-related issues.
-This wave corresponds to the year 2009 and covers attitudes towards a series of topics dealing with social inequality across 41 countries and 56021 individuals.
-Although there is data available from this module for previous waves (1987, 1992, and 1999), unfortunately, there are several variables that are important for this research which were not included, particularly in 1987 and 1999, the reason why we only use the 2009 wave (during the development of this research the 2019 wave became available, but so far only with a restricted set of countries).
+This wave corresponds to the year 2009 and covers attitudes towards a series of topics dealing with social inequality across 41 countries and 56,021 individuals.
+Although there is data available from this module for previous waves (1987, 1992, and 1999), unfortunately, there are several variables that are important for this research which were not included, particularly in 1987 and 1999, reason why we only use the 2009 wave (during the development of this research the 2019 wave became available, but so far with fewer countries than the previous one).
 
@@
 
-- _Perception of meritocracy/non-meritocracy_: for operationalizing perceptions the closest set of ISSP's indicators comes from the question asking about perceptions for opportunities to get ahead, which are usually considered as "meritocratic beliefs" in previous research.
+- _Perception of meritocracy/non-meritocracy_: for operationalizing perceptions the closest set of ISSP's indicators comes from the question asking about perceptions for opportunities to get ahead, which are usually considered as "meritocratic beliefs" in previous studies.
 The general heading of the battery is: _"To begin we have some questions about opportunities for getting ahead.
@@
 
-The estimation was performed using Confirmatory Factor Analysis (CFA).
+The estimation of the measurement model was performed using Confirmatory Factor Analysis (CFA).
 CFA was conducted using the `lavaan` R package (version 0.6-3; @rosseel_lavaan_2012), with diagonally weighted least squares (DWLS) estimation due to the items' ordinal level of measurement [@kline_principles_2016; @rosseel_lavaan_2012].
@@
 
-Attending to the sources of poor fit from the first estimated confirmatory model, we specified a second model that keeps the basic four-factor structure but generates an additional model for the non-meritocratic perceptions as depicted in \@ref(fig:meas02-issp).
+Attending to the sources of poor fit from the first estimated confirmatory model, we specified a second model that keeps the basic four-factor structure but generates an additional model for the non-meritocratic perceptions as depicted in Figure \@ref(fig:meas02-issp).
 In this model, the pairs of items in this dimension form three latent factors which at the same time generates a second-order factor of non-meritocratic perceptions, improving the fit of the model.
@@
 Something similar occurs for the correlation between the two dimensions of preferences ($r=0.25$).
-From this information, we can preliminary conclude that the wide use of reverse-coding for non-meritocratic items attempting to measure meritocracy is not an adequate operationalization of the construct.
-All in all, despite the good fit of the second-order model there are still several limitations regarding the content of the items which we attempt to overcome next by developing and testing new scale for measuring meritocracy.
+
+## Discussion study 1
+
+Although it was possible to get an adequate fit of the model to the pooled data from ISSP, there are several aspects to take into account when using these items for measuring meritocracy.
+First of all, the items were not designed based on this conceptual approach and a first estimation approach (first order factor model) reveals some problems of fit.
+Secondly, some of the items in similar dimensions refer to different concepts, as for instance meritocratic perceptions are related to ambition and hard work, whereas meritocratic preferences deal with how well a job is done.
+A better operationalization would require more alignment between the different sides of the same concept.
+And thirdly, the scale is rather unbalanced given he larger amount of non-meritocratic perception items, producing correlated errors.
+All these limitations are considered in the design of Study 2.
+
+With regard to the meritocratic / non-meritocratic distinction in previous studies, we can preliminary conclude that the wide use of reverse-coding for non-meritocratic items attempting to measure meritocracy is not an adequate operationalization of the construct.
+Meritocracy and non-meritocracy are related but different constructs, meaning that it is theoretically possible to find high and lower level of both.
+This finding opens a series of avenues for future research that are developed in the conclusions.
 
@@
 
-The data was obtained through an online survey which was part of a larger study on meritocracy and preferences developed in Chile in 2019, funded by the national scientific agency ANID.
+The data was obtained through an online survey which was part of a larger study on meritocracy and preferences developed in Chile in 2019, funded by the Chilean National Scientific Agency (ANID).
 The questionnaire was programmed in Qualtrics and the fieldwork was conducted by an external online survey agency ([netquest.com](www.netquest.com)) between December 2019 and January 2020.
@@
 As it is often the case with online surveys, there were some limitations in achieving the quotas for lower educational levels [@zhang_quota_2018; @boas_recruiting_2020].
+This survey was designed as a three-wave panel and the data used in this analysis corresponds to the first wave.
 
@@
 
-The proposed scale of perceptions and preferences relating to meritocracy consisted of eight indicators that were grouped into the four dimensions listed earlier: Perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic).
-In order to achieve at least some comparability with previous studies, the questions were adapted from the items battery "reasons to get ahead" (ISSP/GSS), which has been widely used for operationalizing meritocracy in previous studies [@mijs_paradox_2019; @duru-bellat_who_2012; @reynolds_perceptions_2014].
-The aforementioned eight items ordered according to dimensions are presented in Table \@ref(tab:table-indicadores).
+The proposed scale of perceptions and preferences relating to meritocracy consisted of eight indicators, two for each of the four dimensions listed earlier: Perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic).
+This design is oriented by two main criteria.
+The first one is usability in survey research, for which the length of the scale is restricted due to space reasons in such studies.
+Nevertheless, at the same time the scale should offer possibilities of accounting for measurement error through factor analysis techniques.
+Although a rule of thumb in latent  measurement suggest at least three indicators per dimension, we still opted for two in order to count with a brief scale.
+Although such decision posits some limitation for modeling measurement error, still allows for factor analysis and is a common approach in previous social survey studies with limited items (for instance see @davidov_measurement_2009).
+
+The second criteria is to expand the possibilities of comparability with previous studies on meritocracy, by considering some of the items previously used in this research area.
+To this regard, some of the items were adapted from the items battery "reasons to get ahead" (ISSP/GSS), which has been widely used for operationalizing meritocracy in international research [@mijs_paradox_2019; @duru-bellat_who_2012; @reynolds_perceptions_2014].
+The scale items ordered according to their respective dimensions are presented in Table \@ref(tab:table-indicadores).
 These eight likert-type items have five response alternatives, ranging from "Completely disagree"(1) to "Completely agree" (5).
@@
 There are three main aspects to highlight from this correlation matrix.
-Firstly, as expected the largest correlations are between indicators that correspond to the same factors behind the conceptual model (e.g., perception of meritocracy by effort and by talent, $r=0.52$).
-Secondly, among this correlations the highest are those between the non-meritocratic dimension but in perceptions ($r=0.73$) and preferences ($r=0.61$).
+Firstly, and as expected,  the largest correlations are between indicators that correspond to the same factors behind the conceptual model (e.g., perception of meritocracy by effort and by talent, $r=0.52$).
+Secondly, among these correlations the highest are those between the non-meritocratic dimension but in perceptions ($r=0.73$) and preferences ($r=0.61$).
 Thirdly, both items for meritocratic preferences (E and F) are the ones that mostly correlate with the rest of the perceptual items, showing medium to high correlations.
-This is noteworthy because it indicates that the perception of non-meritocracy would be related to larger meritocratic preferences.
+This is noteworthy because it would indicate that the perception of non-meritocracy would be related to larger meritocratic preferences.
 Finally, and similar to what was found in Study 1 with ISSP data, we observe that there are no considerable negative correlations between meritocratic and non-meritocratic aspects, undermining the assumptions of previous studies that suggested that these dimensions would be the opposite poles of one same continuum [@reynolds_perceptions_2014].
@@
 
+## Discussion study 2
+
+The estimation performed with different administration sets for the meritocracy scale gives evidence in line with the conceptual model.
+The four dimensions appear as independent but related, although with some differences when compared with the estimation with ISSP data in Study 1.
+The main general difference is that with ISSP data the four latent dimensions where positively correlated, whereas now some correlations are statistically not different from zero.
+For instance, perceptions and preferences for meritocracy keep being related, which is not the case for the non-meritocratic pair.
+In principle this makes more sense in theoretical terms, as if someones perceives that meritocracy functions well, probably will support it more.
+On the other side, if someones perceives non-meritocracy, this should lead to a larger preference for meritocracy as it would constitute a demand for lager meritocracy when this one is not perceived to work.
+
+Despite the good general working of the scale, there are some limitations in the interpretation of the results given the few items (two) per dimension.
+As it was mentioned before, the option of a minimalistic approach in terms of items has advantages and disadvantages: advantages in terms of usability in social surveys with limited space, and disadvantages in terms of measurement validity in the estimation of the latent constructs.
+Being aware of this trade-off, in the following section we offer additional validity evidence as a way of complementing some of the weaknesses identified.
+
 # Study 3: Additional validity analyses
 
-We performed two further analyses in order to add evidence regarding the quality of the scale.
-Firstly we analyzed the convergent validity by exploring the association of the merit scale with related concepts and its correspondent measurements.
-Secondly, we test the internal consistency of the measurement model behind the scale comparing it with a different sample through measurement invariance procedures.
-
-## Convergent validity
-
-### Data
-
-For this analysis, we examined data from the last of the three waves of the online panel survey used in Study 2.
-This wave included additional measures that allow testing the correlations of the merit scale with some related constructs as opportunity beliefs and personal wherewithal.
-After a listwise deletion of missing cases in socio-demographic information, a total of 1422 individuals took part in the third wave: 668 (46.97%) women and 754 (53.03%) men.
-The ages of 50.57% of them are 45 years or older, and 36.28% held a tertiary degree.
-
-### Variables
-
-The following scales were included for testing their correlations with the 8-item meritocracy scale:
-
-- Opportunity beliefs: We used two items of the social inequality module of the International Social Survey Programme.
-Studies use generally both indicators to measure meritocratic and non-meritocratic beliefs [@mccall_exposure_2017; @mijs_paradox_2019].
-One item measures the importance of individual factors in determining life outcomes by asking for the importance of hard work for getting ahead in life (M=3.76, SD=0.93).
-The second item asks for the importance of coming from a wealthy family and captures the importance of structural factors in getting ahead (M=2.75, SD=1.27).
-Participants responded to each item using a 1 (not important at all) to 5 (essential) response scale.
-We call the first item _hard work_ and the second one _social origin_.
-
-- Personal wherewithal: the questionnaire included The Neoliberal Beliefs Inventory (NBI), which considers four factors: Government Interference preferences, Competition preferences, System Inequality perception, and Personal Wherewithal [@bay-cheng_tracking_2015].
-We used the factor _personal wherewithal_ that reflects meritocracy beliefs in terms of the importance of personal attributes as strength and skills to yield success.
-The construct consists of 8 items (e.g., "Any goal can be achieved with enough hard work and talent", "I've benefited from working hard, so there's no reason others can't", "Anyone who is willing to work hard can be successful in Chile").
-Participants answered the items using a 1 (totally disagree) to 6 (totally agree) scale.
-We simply computed mean scores, with higher scores indicating stronger support for meritocracy (M=3.32, SD=0.88, $\alpha$ = 0.89).
-
-Both instruments cover mainly the perceptual side of the concept.
-Therefore, in general, we expect a larger correlation with meritocratic perceptions than with preferences.
-In the case of opportunity beliefs, the hard work item should correlate positively with meritocratic perception whereas the importance of coming from a wealthy family is expected to correlate positively with non-meritocratic perceptions.
-Regarding the second instrument of personal wherewithal, we also anticipate a positive correlation with meritocratic perception and low to null correlations with the other dimensions of the merit scale.
-
-### Results
-
-We used polychoric correlations to evaluate relationships of meritocratic and non-meritocratic preferences and perceptions with the items of the opportunity beliefs battery.
-For personal wherewithal, we examined Pearson correlations.
-
-Table \@ref(tab:val-conv) shows the polychoric and Pearson correlations between our meritocracy scale and common measures of meritocracy.
-As we expected, preferences show very weak correlations.
-Findings also indicate that correlations of perceptions with opportunity beliefs are consistent with our expectations.
-The item for social origin shows a positive and moderate association with unmeritocratic perception ($r=0.361$), and a negative and low correlation with meritocratic perception ($r=-0.186$).
-Regarding the item of hard work, the correlation with meritocratic perception is positive but weak ($r=0.201$).
-One explanation for this result is that the support for hard work is very high among respondents and therefore with low variability, as only 8.6 percent indicates _not very important_ or _not important at all_.
-
-```{r val-conv}
-  if (!require("pacman")) install.packages("pacman")
-  pacman::p_load(kableExtra, polycor)
-  table_format = if(knitr::is_html_output()) { #conditional instructions for kable
-     "html" #if html set "html" in format
-   } else if(knitr::is_latex_output()) {
-     "latex"#if latex set "latex" in format
-   }
-  load(file="input/data/proc/data_s3_conv.RData")
-
-  pl1 <- polycor::polyserial(data_s3_conv$perc_merit,data_s3_conv$get_ah_1)
-  pl2 <- polycor::polyserial(data_s3_conv$pref_merit,data_s3_conv$get_ah_1)
-  pl3 <- polycor::polyserial(data_s3_conv$perc_nmerit,data_s3_conv$get_ah_1)
-  pl4 <- polycor::polyserial(data_s3_conv$pref_nmerit,data_s3_conv$get_ah_1)
-  valores1 <- round(c(pl1,pl2,pl3,pl4),4)
-
-  pl1b <- polycor::polyserial(data_s3_conv$perc_merit,data_s3_conv$get_ah_5)
-  pl2b <- polycor::polyserial(data_s3_conv$pref_merit,data_s3_conv$get_ah_5)
-  pl3b <- polycor::polyserial(data_s3_conv$perc_nmerit,data_s3_conv$get_ah_5)
-  pl4b <- polycor::polyserial(data_s3_conv$pref_nmerit,data_s3_conv$get_ah_5)
-  valores2 <- round(c(pl1b,pl2b,pl3b,pl4b),4)
-
-  pl1c <- cor(data_s3_conv$perc_merit,data_s3_conv$nbmerit)
-  pl2c <- cor(data_s3_conv$pref_merit,data_s3_conv$nbmerit)
-  pl3c <- cor(data_s3_conv$perc_nmerit,data_s3_conv$nbmerit)
-  pl4c <- cor(data_s3_conv$pref_nmerit,data_s3_conv$nbmerit)
-  valores3 <- round(c(pl1c,pl2c,pl3c,pl4c),4)
-
-  variables= c("Meritocratic Perception", "Meritocratic Preferences","Non-meritocratic Perception","Non-meritocratic Preferences")
-
-  table01 <- cbind(variables, valores1, valores2, valores3)
-
-  knitr::kable(table01,
-        col.names = c("Merit-Scale","Social Origin ", "Hard Work " , "Wherewithal"),
-        linesep = "",
-        format = table_format,
-        booktabs=T,
-        digits = 3,
-        caption = c("Polyserial and Pearson correlation whit other merit scales")) %>%
-    kableExtra::kable_styling(
-      full_width = F,
-      position = "center",
-      font_size = 10,
-      bootstrap_options = c("striped", "bordered")) %>%
-    kableExtra::column_spec(column = 1, width = "4.5 cm")
-```
-
-The measure of NBI for _personal wherewithal_ also provides evidence for the convergent validity of the merit scale.
-Results indicate a positive correlation of this measure with meritocratic perception ($r=0.406$), suggesting that both variables capture similar concepts.
-The NBI's factor depicts a positive association with unmeritocratic preference ($r=0.225$).
-This finding is consistent with the literature, in as much as the belief for meritocracy is associated with the justification of inequality [@mccall_exposure_2017; @madeira_primes_2019].
-Furthermore, results of our confirmatory factor analysis indicate that meritocratic perceptions and unmeritocratic preferences show a positive and weak correlation.
-
-## Measurement invariance
-
 The modeling for invariance testing has been gaining more relevance in sociological survey studies, mainly due to the interest in the validity of measurement scales in comparative studies in various social and cultural contexts [@davidov_measurement_2014].
-In this section we will use an analysis of invariance between groups, where the proposed factorial structure is expected to be independent of belonging to the sample corresponding to the original study, comparing it with another more recent study available and where the meritocracy scale was also incorporated.
+In this section we present the results of an analysis of invariance between groups, testing whether the proposed factorial structure for the meritocracy scale is applicable in a different dataset.
 
@@
 
-For testing the invariance we compared the data described previously with a new data source coming from an online survey carried out during the first half of 2020.
+For testing the invariance we compared the data described previously in Study 2 to a different  dataset coming from an online survey carried out during the first half of 2020.
 The characteristics of this sample in terms of application and coverage are equivalent to those of the previous study.
@@
 
-The items of the meritocracy scale were identical to the original.
-The items were administrated in a randomized order (which corresponds to the application modality of Group 3 as showed in Figure \@ref(fig:appmod)).
+The items of the meritocracy scale were identical to the original and they  were administrated in a randomized order (which corresponds to the application modality of Group 3 in Study 2, as showed in Figure \@ref(fig:appmod)).
 
 ### Results
-
-Table \@ref(tab:fcloads-inv) in the Appendix shows the results of the CFA estimation of the meritocracy scale in this new sample.
-As observed, the loadings and the fit indices show similar results as the ones presented for the original study above, which gives a first base from which to start the invariance analysis.
 
@@
 Such result is considered in general as evidence of invariance [@fischer_are_2011], as strict forms of measurement invariance rarely hold [@vandeschoot_checklist_2012].
-Still, the comparability of latent means requires strict invariance which in this case does not hold when considering $\Delta \chi^2 (\Delta \text{df})$.
-Nevertheless, the criteria of $\Delta \text{CFI}$ used for comparing models is close to the rejection criteria of >.01, whereas the $\Delta \text{RMSEA}$ fulfils the requirements of being below the cut-off criteria as suggested by [@chen_sensitivity_2007].
+Still, the comparability of latent means requires strict invariance which in this case does not hold when considering just $\Delta \chi^2 (\Delta \text{df})$.
+Nevertheless, the criteria of $\Delta \text{CFI}$ used for comparing models is close to the rejection criteria of >.01, whereas the $\Delta \text{RMSEA}$ fulfills the requirements of being below the cut-off criteria as suggested by [@chen_sensitivity_2007].
 Therefore, using this last standard, the level for strict invariance would hold for the meritocracy scale.
+
+## Discusion Study 3
+
+Based on the measurement invariance analysis for two datasets, we found evidence supporting the equivalence of the scale when applied in different samples.
+This gives a more robust ground when using this measure for comparing groups, as for instance country comparison.
+Although the results ar promising, still the analysis was performed in two samples in one country (Chile) and it requires further examination with social surveys in other societies.
 
@@
 As the existence of heterogeneous approaches certainly has consequences for the advancement of the study of meritocracy, this paper presented a comprehensive conceptual framework for the empirical study of meritocracy, building upon previous research.
-This framework was then tested against available and new survey data.
+This framework was operacionalized in a survey scale and tested against available and new survey data.
 
@@
 Our results indicate that perceptions and preferences seem to be two related but different dimensions, often confounded in previous research under the label of "beliefs".
-Meritocratic and non-meritocratic dimensions do not appear to constitute poles of the same continuum, as some previous studies have assumed.
+At the same time, meritocratic and non-meritocratic dimensions do not appear to constitute poles of the same continuum, as some previous studies have assumed.
 Regarding the possible effects of the items' order in the estimation of the latent variable, we tested three different order versions and found evidence that suggests the use of the randomized version of the scale.
-Furthermore, we found evidence of convergent validity as well as measurement invariance that approach well the requirement of comparability between samples.
+Furthermore, we found evidence of measurement invariance that approach well the requirement of comparability between samples.
+
+One of the criteria in designing the merit scale was to develop a short instrument that could be used in social surveys, assuming that using only two items per dimension posits limitations for the confirmatory analysis estimated here.
+We preferred this confirmatory approach at it offers several advantages in terms of information about the quality of the measurement, and we offered additional validity evidence with measurement invariance analysis.
+Despite the advantages, a more simple and straightforward perspective is always available, as for instance using only the average of the items per dimension.
+Whereas a short scale it is always appreciated in social survey studies, for other audiences could not reach measurement standards.
+This may require further expansions and/or consider using a different instrument as the "Preference for the Merit Principle Scale" [@davey_preference_1999], which measures only one dimension of meritocracy with fifteen items.
 

### 2022-12-30 paper-blinded.Rmd
ediciones SIR

@@
 Our discussion highlights the importance of considering different dimensions in order to advance in the study of meritocracy."
-
 output:
@@
 reverse coded).
-The assumption behind this approach is that rejecting a supposed ascribed or non-meritocratic^[Although the use of the concept "non-meritocratic" is very broad, we prefer to use it instead of something like ascription, given that this last one leaves out aspects such as the use of personal relationships.@@] aspect (as having family and/or children) implies a stronger belief in meritocracy.
+The assumption behind this approach is that rejecting a supposed ascribed or non-meritocratic^[Although the use of the concept "non-meritocratic" is very broad, we prefer to use it instead of something like ascription, given that this last one leaves out aspects such as the use of personal relationships.] aspect (as having family and/or children) implies a stronger belief in meritocracy.
 A similar approach of reverse-coding non-meritocratic items was taken by @newman_false_2015, using the same principles applied in the "Preference for the Merit Principle Scale" [@davey_preference_1999].
@@
 
+\break
+
 The classification of the items is based on criteria of internal motivation (meritocratic) and structural constraints (non-meritocratic).
@@
 The estimation of the measurement model was performed using Confirmatory Factor Analysis (CFA).
-CFA was conducted using the `lavaan` R package (version 0.6-3; @rosseel_lavaan_2012), with diagonally weighted least squares (DWLS) estimation due to the items' ordinal level of measurement [@kline_principles_2016; @rosseel_lavaan_2012].
+CFA was conducted using the `lavaan` R package (version 0.6-3; @rosseel_lavaan_2012a), with diagonally weighted least squares (DWLS) estimation due to the items' ordinal level of measurement [@kline_principles_2016; @rosseel_lavaan_2012a].
 As recommended by @brown_confirmatory_2015, we assessed model fit by jointly considering the comparative fit index and Tucker-Lewis Index (CFI and TLI; acceptable fit > 0.95), Root of the average squared residual approximation (RMSEA; acceptable fit < 0.08),  Chi-square: (p-value; acceptable fit > 0.05, and Chi-square ratio > 3).
@@
 Table \@ref(tab:sum-fit-issp) shows the results of the estimation of two confirmatory models.
-The first one (First Order) corresponds to a model that estimate four factors, each for one of the dimensions presented in \@ref(tab:table-issp), showing only regular fit indicators (CFI=0.959, TLI=0.944, RMSEA=0.098, $\chi^2$(df=48)= 21308.535; more detail in Table \@ref(tab:tb-perpref1)).
+The first one (First Order) corresponds to a model that estimate four factors, each for one of the dimensions presented in \@ref(tab:table-issp), showing only regular fit indicators (CFI=0.959, TLI=0.944, RMSEA=0.098, $\chi^2$(df=48)= 21308.535).
 Further analysis showed that the sources for poor fit were mostly related to the items of the non-meritocratic perceptions, which as mentioned above displayed correlations among them not taken into account in this model specification.
@@
 
+Based on the measurement validation, we asses if the proposed measurement strategy can be used in cross-national comparisons.
+In this line, Figure \@ref(fig:scat1) depicts the correlation of factor scores for Meritocratic Perception and Meritocratic Preferences at the country level.
+For instance, Bulgaria (BGR) and the United States (USA) appropriately represent the positive association between perception and preferences.
+In other words, countries that describe merit in their society as an essential factor to get ahead in life also consider that effort and talent should be determinant factors in how much an individual earns.
+On the other hand, in countries with lower perceived meritocracy, preferences are also low, where Denmark (DNK) and Venezuela (VEN) are representative societies of this association.
+
+Another finding depicted by Figure \@ref(fig:scat1) is the positive association of Non-meritocratic perception with Meritocratic preferences, showing that countries like Bulgaria (BGR) and Philipines (PHL) represent cases with higher Non-meritocratic perceptions and higher preferences for meritocracy.
+In other words, in societies where people perceive higher importance of non-meritocratic aspects to get ahead in life, also preferer that merit should be a determinant factor in wages.
+On the other side, countries like Sweden (SWE) or Denmark (DNK), both Scandinavian societies with strong welfare systems, represent contexts in which non-meritocratic aspects are not described as important, and merit is not considered as a factor that should be determinant in wages.
+
+```{r scat1, fig.cap="Cross-national comparison of non-meritocratic and meritocratic perceptions with meritocratic preference",fig.height=6, fig.width=12}
+  load(file = "input/data/proc/data_s1.RData")
+  if (!require("pacman")) install.packages("pacman")
+  pacman::p_load(dplyr, kableExtra, knitr,lavaan,countrycode,ggplot2,haven,
+                 gridExtra)
+  table_format = if(knitr::is_html_output()) { #conditional instructions for kable
+    "html" #if html set "html" in format
+  } else if(knitr::is_latex_output()) {
+    "latex"#if latex set "latex" in format
+  }
+
+  # meassurement model: Second order
+  cfa_perpref2 <- '
+  	  # latent variables
+  	  percmerit =~ hwork + ambition
+  	  parents =~ wealthy + pareduc
+  	  backgrd =~ race + gender
+  	  networks=~ people + polcone
+  	  percnmerit=~ parents + backgrd + networks
+  	  prefmerit =~ welljob + hardjob
+  	  prefnmerit=~ family + child
+  	  '
+  # Model fit: second order
+
+  fit_perpref2 <- lavaan::cfa(cfa_perpref2, data = data_s1,ordered = F)
+
+  df_proc<- data_s1 %>% select(id,v5,wealthy:hardjob)  
+  df_proc$ccode <- as.numeric(data_s1$v5)
+
+  predicted2 <- lavPredict(fit_perpref2,newdata = data_s1)
+
+  idx <- lavInspect(fit_perpref2, "case.idx")
+  fscores <- lavPredict(fit_perpref2)
+  ## loop over factors
+  for (fs in colnames(fscores)) {
+    df_proc[idx, fs] <- fscores[ , fs]
+  }
+
+  df_proc <-  df_proc %>%  select(id,ccode,percmerit,prefmerit,prefnmerit) %>% na.omit()
+
+  df_proc$clab <- countrycode(sourcevar = df_proc$ccode,
+                            origin = "iso3n",
+                            destination = "iso3c")
+
+avg_count <-  
+  df_proc %>% group_by(clab) %>% summarise(prefmerit=mean(prefmerit),
+                                            percmerit=mean(percmerit),
+                                            prefnmerit=mean(prefnmerit))
+
+# create figure 1: perception merit by preferences merit
+figure1 <- 
+ggplot(avg_count, aes(x=percmerit, y=prefmerit)) +
+  # geom_point() + 
+  geom_text(label=as.character(avg_count$clab)) +
+  geom_smooth(method = "lm") +
+  xlab("Meritocratic perception")+
+  ylab("Meritocratic preference")  
+# create figure 2: perception non-merit by preferences merit
+figure2 <- 
+ggplot(avg_count, aes(x=prefnmerit, y=prefmerit)) +
+  # geom_point() + 
+  geom_text(label=as.character(avg_count$clab)) +
+  geom_smooth(method = "lm") +
+  xlab("Non-meritocratic perception")+
+  ylab("Meritocratic preference")
+
+grid.arrange(figure1, figure2, ncol = 2,
+             bottom = grid::textGrob(label = paste0("Source: Authors calculation based on International Social Survey Programme - Social Inequality 2009",
+      " (n=",dim(df_proc)[1],")"),
+                                     hjust = -0.07,
+                                     gp = grid::gpar(fontface = 1, fontsize = 9)))
+```
+
 ## Discussion study 1
@@
 The models estimated one factor for each of the four proposed dimensions presented in Table \@ref(tab:table-indicadores).
-As in Study 1, CFA was conducted using the `lavaan` R package (version 0.6-3; @rosseel_lavaan_2012), with diagonally weighted least squares (DWLS) estimation due to the items' ordinal level of measurement [@kline_principles_2016; @rosseel_lavaan_2012].
+As in Study 1, CFA was conducted using the `lavaan` R package (version 0.6-3; @rosseel_lavaan_2012a), with diagonally weighted least squares (DWLS) estimation due to the items' ordinal level of measurement [@kline_principles_2016; @rosseel_lavaan_2012a].
 The fit indexes and cut-off criteria were the same as the ones used in Study 1.
@@
 
-# Appendix {-}
+<!-- # Appendix {-} -->
 
@@
 
-```{r tb-perpref1, echo=FALSE}
+```{r tb-perpref1, eval=FALSE, include=FALSE}
  if (!require("pacman")) install.packages("pacman")
@@
 
-**Study 3**
-
-```{r fcloads-inv, echo=FALSE}
+<!-- **Study 3** -->
+
+```{r fcloads-inv, eval=FALSE, include=FALSE}
  if (!require("pacman")) install.packages("pacman")

### 2023-01-03 paper-blinded.Rmd

@@
       " (n=",dim(df_proc)[1],")"),
-                                     hjust = -0.07,
+                                     hjust = 0.1,
                                      gp = grid::gpar(fontface = 1, fontsize = 9)))

### 2023-01-25 paper-blinded.Rmd
ajustes comentarios

@@
 output:
+  bookdown::pdf_document2:
+     template: null
+     toc: false
+     number_sections: false
   bookdown::word_document2:
@@
   bookdown::html_document2:
-    number_sections: false
-  bookdown::pdf_document2:
-     template: null
-     toc: false
     number_sections: false
@@
 The question of whether talent (as intelligence and ability) is or is not considered meritocratic is certainly an interesting topic to discuss from a philosophical point of view, but for those working on empirical studies, it should face empirical scrutiny.
-Besides, the consideration of talent as part of meritocracy certainly opens some interesting avenues of research, as for instance some studies suggest that for the elites meritocracy is actually related to talent, whereas effort is something more characteristic of the meritocracy of the middle and low classes [@atria_economic_2020].
+Besides, the consideration of talent as part of meritocracy certainly opens some interesting avenues of research, as for instance some studies suggest that for the elites meritocracy is actually related to talent, whereas effort is something more characteristic of the meritocracy of the middle and lower classes [@atria_economic_2020].
 Therefore, we propose that a concept (and measure) of meritocracy should include both talent and effort as in its original conception.
@@
 In choosing this strategy of index construction, we argue that support for meritocracy is not a latent variable" [@kunovich_systems_2007 p.653-654].
-Although some advances were made by @reynolds_perceptions_2014 through conducting a principal component analysis of meritocratic and non-meritocratic dimensions, they finally chose to build a sum index despite proving a multidimensional latent structure.
+Although some advances were made by Reynolds and Xian [-@reynolds_perceptions_2014] through conducting a principal component analysis of meritocratic and non-meritocratic dimensions, they finally chose to build a sum index despite proving a multidimensional latent structure.
 
@@
 
-- _Brief_, as to be used in regular public opinion surveys.
+- _Brief_, as to be used in regular public opinion surveys [@davidov_measurement_2009].
 To this regard, it differs from the proposal of  "Preference for the Merit Principle Scale" [@davey_preference_1999], as they use 15 items for just one dimension (aside from the problem of reverse-coding non-meritocratic items).
@@
 Although there is data available from this module for previous waves (1987, 1992, and 1999), unfortunately, there are several variables that are important for this research which were not included, particularly in 1987 and 1999, reason why we only use the 2009 wave (during the development of this research the 2019 wave became available, but so far with fewer countries than the previous one).
+Details about the sample of countries can be seen on Table \@ref(tab:countries).
 
@@
 Attending to the sources of poor fit from the first estimated confirmatory model, we specified a second model that keeps the basic four-factor structure but generates an additional model for the non-meritocratic perceptions as depicted in Figure \@ref(fig:meas02-issp).
-In this model, the pairs of items in this dimension form three latent factors which at the same time generates a second-order factor of non-meritocratic perceptions, improving the fit of the model.
+In this model, the pairs of items in this dimension form three latent factors which at the same time generate a second-order factor of non-meritocratic perceptions, improving the fit of the model.
 
@@
 Another finding depicted by Figure \@ref(fig:scat1) is the positive association of Non-meritocratic perception with Meritocratic preferences, showing that countries like Bulgaria (BGR) and Philipines (PHL) represent cases with higher Non-meritocratic perceptions and higher preferences for meritocracy.
-In other words, in societies where people perceive higher importance of non-meritocratic aspects to get ahead in life, also preferer that merit should be a determinant factor in wages.
+In other words, in societies where people perceive higher importance of non-meritocratic aspects to get ahead in life, people also prefer that merit should be a determinant factor in wages.
 On the other side, countries like Sweden (SWE) or Denmark (DNK), both Scandinavian societies with strong welfare systems, represent contexts in which non-meritocratic aspects are not described as important, and merit is not considered as a factor that should be determinant in wages.
@@
 Nevertheless, at the same time the scale should offer possibilities of accounting for measurement error through factor analysis techniques.
-Although a rule of thumb in latent  measurement suggest at least three indicators per dimension, we still opted for two in order to count with a brief scale.
+Although a rule of thumb in latent  measurement suggests at least three indicators per dimension, we still opted for two in order to count with a brief scale.
 Although such decision posits some limitation for modeling measurement error, still allows for factor analysis and is a common approach in previous social survey studies with limited items (for instance see @davidov_measurement_2009).
@@
 The four dimensions appear as independent but related, although with some differences when compared with the estimation with ISSP data in Study 1.
-The main general difference is that with ISSP data the four latent dimensions where positively correlated, whereas now some correlations are statistically not different from zero.
+The main general difference is that with ISSP data the four latent dimensions were positively correlated, whereas now some correlations are statistically not different from zero.
 For instance, perceptions and preferences for meritocracy keep being related, which is not the case for the non-meritocratic pair.
@@
 This gives a more robust ground when using this measure for comparing groups, as for instance country comparison.
-Although the results ar promising, still the analysis was performed in two samples in one country (Chile) and it requires further examination with social surveys in other societies.
+Although the results are promising, still the analysis was performed in two samples in one country (Chile) and it requires further examination with social surveys in other societies.
 
@@
 Regarding the possible effects of the items' order in the estimation of the latent variable, we tested three different order versions and found evidence that suggests the use of the randomized version of the scale.
-Furthermore, we found evidence of measurement invariance that approach well the requirement of comparability between samples.
+Furthermore, we found evidence of measurement invariance that approaches well the requirement of comparability between samples.
 
@@
 **Study 1**
+
+```{r countries, echo=FALSE}
+  load(file = "input/data/proc/data_s1.RData")
+  if (!require("pacman")) install.packages("pacman")
+  pacman::p_load(dplyr, kableExtra, knitr,lavaan,countrycode,ggplot2,haven,
+                 gridExtra)
+  table_format = if(knitr::is_html_output()) { #conditional instructions for kable
+    "html" #if html set "html" in format
+  } else if(knitr::is_latex_output()) {
+    "latex"#if latex set "latex" in format
+  }
+  load(file = "input/data/proc/data_s1.RData")
+  if (!require("pacman")) install.packages("pacman")
+  pacman::p_load(dplyr, kableExtra, knitr,lavaan,countrycode,ggplot2,haven,
+                 gridExtra)
+  table_format = if(knitr::is_html_output()) { #conditional instructions for kable
+    "html" #if html set "html" in format
+  } else if(knitr::is_latex_output()) {
+    "latex"#if latex set "latex" in format
+  }
+ # meassurement model: Second order
+  cfa_perpref2 <- '
+  	  # latent variables
+  	  percmerit =~ hwork + ambition
+  	  parents =~ wealthy + pareduc
+  	  backgrd =~ race + gender
+  	  networks=~ people + polcone
+  	  percnmerit=~ parents + backgrd + networks
+  	  prefmerit =~ welljob + hardjob
+  	  prefnmerit=~ family + child
+  	  '
+  # Model fit: second order
+
+  fit_perpref2 <- lavaan::cfa(cfa_perpref2, data = data_s1,ordered = F)
+
+  df_proc<- data_s1 %>% select(id,v5,wealthy:hardjob)  
+  df_proc$ccode <- as.numeric(data_s1$v5)
+
+  predicted2 <- lavPredict(fit_perpref2,newdata = data_s1)
+
+  idx <- lavInspect(fit_perpref2, "case.idx")
+  fscores <- lavPredict(fit_perpref2)
+  ## loop over factors
+  for (fs in colnames(fscores)) {
+    df_proc[idx, fs] <- fscores[ , fs]
+  }
+
+  df_proc <-  df_proc %>%  select(id,ccode,percmerit,prefmerit,prefnmerit) %>% na.omit()
+
+  df_proc$clab <- countrycode(sourcevar = df_proc$ccode,
+                            origin = "iso3n",
+                            destination = "country.name")
+avg_count <-  
+  df_proc %>% group_by(clab) %>% summarise(N=n())
+avg_count <- arrange(avg_count)
+
+# sum(avg_count[, 'N'])
+  col.nam <- c("Country","N", "Country", "N")
+  footnote <- paste0("N = ", sum(avg_count[, 'N']),". Source: International Social Survey Programme (2009).")
+
+  knitr::kable(bind_cols(avg_count[1:20,],avg_count[21:40,]),
+               col.names = col.nam, align = c("l","l","l","l"),linesep = "",
+        booktabs=TRUE,format = table_format,escape = FALSE,
+        caption = "Sample of Study 1 ") %>%
+    kableExtra::kable_styling(full_width = FALSE,
+                              latex_options = "HOLD_position",
+                              bootstrap_options=c("striped", "bordered"),
+                              font_size = 10) %>%
+    kableExtra::column_spec(column = c(2,4),width = "1cm") %>% 
+    kableExtra::column_spec(column = c(1,3),width = "2.5cm") %>% 
+    column_spec(3, border_left = T) %>% 
+    kableExtra::footnote(general = footnote, footnote_as_chunk = T,threeparttable = T)
+```
+
+\pagebreak
+
+**Study 2**
 
@@
 | university or more  | 21,21%  | 18,13% |
-Table: `r as.character(paste("(\\#tab:rep-samp)", "Representativeness of the study 1 sample"))`.
-
-\pagebreak
+Table: `r as.character(paste("(\\#tab:rep-samp)", "Representativeness of the study 2 sample"))`.
 

### 2023-02-01 paper-blinded.Rmd
grammarly

@@
 Despite its wide use as a concept, empirical research on meritocracy is relatively novel and characterized by diverse conceptualizations and measures that make the findings and their interpretation rather inconsistent.
-Most of the studies in the area have relied upon secondary data from different surveys to operationalize meritocracy, with a wide variation in the use and interpretation of the same measuring items.
+Most of the studies in the area have relied upon secondary data from different surveys to operationalize meritocracy, with a large variation in the use and interpretation of the same measuring items.
 Based on a critical analysis of concepts and operationalizations of meritocracy in survey research, we propose a conceptual framework for measuring perceptions and preferences for meritocracy and non-meritocracy in a minimalistic version (few items as possible), in order to be considered in comparative survey research studies.
 The proposal is first tested through confirmatory factor analysis (CFA) using available ISSP (International Social Survey Programme) inequality module items.
-Based on these results, in a second stage we developed a improved scale that better suits the perceptions/preferences distinction, which was applied in a Chilean sample (N=2,141).
+Based on these results, in a second stage we propose  an  improved scale that better suits the perceptions/preferences distinction, which was applied in a Chilean sample (N=2,141).
 The results of the CFA estimation give evidence of a better fit for the new scale.
@@
 
-Due to the role that meritocratic beliefs play in the justification of individual achievement (or failure) in contemporary societies [@hadjar_meritokratie_2008; @markovits_meritocracy_2019; @sandel_tyranny_2020], multiple studies have evaluated the relationship between meritocratic beliefs and personal and/or contextual characteristics.
+Due to the role that meritocratic beliefs play in the justification of individual achievement (or failure) in contemporary societies [@hadjar_meritokratie_2008; @markovits_meritocracy_2019; @sandel_tyranny_2020], multiple studies have evaluated the relationship between meritocratic beliefs and personal and (or) contextual characteristics.
 For example, some studies have linked meritocracy to the reinforcement of socio-economic, gender, and ethnic stereotypes [@madeira_primes_2019; @girerd_neoliberalism_2020; @girerd_neoliberalism_2020], as well as the effects of meritocratic beliefs in educational [@generett_stories_2020; @owens_engines_2020] and organizational contexts [@perez_advancing_2020; @aiello_new_2019].
 Most of such studies so far have used indicators from existing standard social surveys, such as the International Social Survey Programme (ISSP), to measure meritocracy.
-However, as we will show later, the concepts used as well as the instruments for measuring meritocracy vary extensively among studies.
-In many cases, similar phenomena are associated with different indicators, and conversely, often different phenomena are measured with similar indicators.
-Such trends not only limit the comparability of studies, but also the ability to understanding the effects of meritocratic beliefs across different disciplines and lines of research.
+However, as we will show later, the concepts and instruments used for measuring meritocracy vary extensively among studies.
+In many cases, similar phenomena are associated with different indicators, and conversely, different phenomena are measured with similar indicators.
+Such trends limit the comparability of studies and the ability to understand the effects of meritocratic beliefs across different disciplines and lines of research.
 
 Based on a critical analysis of different approaches to the measurement of meritocracy to date, this article proposes a conceptual and measurement framework that considers two of the main aspects in the study of meritocracy: perceptions and preferences.
-Furthermore, it distinguishes between meritocratic and non-meritocratic dimensions, as they are not necessarily two poles of the same continuum as previous studies seem to suggest.
-This conceptual framework is oriented to develop a measurement instrument as simple and brief as possible, being suitable to be used in public opinion surveys and thus allowing to integrate meritocratic beliefs in the study of different social phenomena.
-
-## The black-box of meritocratic beliefs {-#blackbox}
+Furthermore, it distinguishes between meritocratic and non-meritocratic dimensions, as they are not necessarily two poles of the same continuum as previous studies suggest.
+This conceptual framework is oriented to develop a measurement instrument as simple and brief as possible, suitable for public opinion surveys and thus allowing to integrate meritocratic beliefs in studying different social phenomena.
+
+## The black box of meritocratic beliefs {-#blackbox}
 
@@
 In the subsequent operationalization, this is associated with the following question and indicator: "how important you think it is for getting ahead in life: (a) hard work", scored from 1 to 5 on a likert scale.
-The assumptions behind such definition are worth discussing in light of the conceptual meaning of meritocracy and its possibilities of operationalization.
-
-The item used by @mijs_paradox_2019 is part of an items' battery present in several international surveys, usually called "reasons to get ahead".
-This battery displays a series of indicators related to what people consider important to get ahead in life: hard work, education, ambition, a wealthy family, the right connections, religion, race, and gender.
-Therefore, when considering only one of the items, it means that other aspects that could be associated with talent, such as education, would not be deemed as meritocratic.
+The assumptions behind such a definition are worth discussing in light of the conceptual meaning of meritocracy and its possibilities of operationalization.
+
+The item used by @mijs_paradox_2019 is part of an item's battery which is available in several international surveys, usually called "reasons to get ahead." This battery displays a series of indicators related to what people consider important to get ahead: hard work, education, ambition, a wealthy family, the right connections, religion, race, and gender.
+Therefore, considering only one of the items means that other aspects associated with talent, such as education, would not be deemed meritocratic.
 As he points out: "Hard work is arguably the most meritocratic part of Michael Young's equation: 'Merit = Intelligence + Effort', for the simple fact that intelligence itself is influenced by a non-meritocratic factor: who your parents happen to be" (p.5).
@@
 In this conceptual and measurement approach of meritocratic beliefs, we can observe a couple of strong conceptual assumptions: a) effort would not depend on parental influence, and b) talent (as innate ability) is not meritocratic (contrary to Michael Youngs' original conceptualization).
-This conceptual and measurement-based assumption is found in other studies, which also assumes effort to be the main and only aspect of meritocracy [@girerd_neoliberalism_2020; @bubak_perceptions_2019], raising the question: Is effort the only dimension behind the concept of meritocracy?
-The question of whether talent (as intelligence and ability) is or is not considered meritocratic is certainly an interesting topic to discuss from a philosophical point of view, but for those working on empirical studies, it should face empirical scrutiny.
-Besides, the consideration of talent as part of meritocracy certainly opens some interesting avenues of research, as for instance some studies suggest that for the elites meritocracy is actually related to talent, whereas effort is something more characteristic of the meritocracy of the middle and lower classes [@atria_economic_2020].
+This conceptual and measurement-based assumption is found in other studies, which also assume effort to be the main and only aspect of meritocracy [@girerd_neoliberalism_2020; @bubak_perceptions_2019], raising the question: Is effort the only dimension behind the concept of meritocracy?
+Whether talent (as intelligence and ability) is or is not considered meritocratic is an interesting topic to discuss from a philosophical point of view, but it also requires empirical evidence.
+Furthermore, considering talent as part of meritocracy certainly opens some interesting avenues of research, as for instance some studies suggest that for the elites meritocracy is related to talent, whereas effort is more characteristic of the meritocracy of the middle and lower classes [@atria_economic_2020].
 Therefore, we propose that a concept (and measure) of meritocracy should include both talent and effort as in its original conception.
@@
 Several papers dealing with meritocracy use the term beliefs (i.e.
-what should be), while actually referring to perceptions (i.e.
+what should be), while referring to perceptions (i.e.
 what is).
-This occurs for instance in @reynolds_perceptions_2014, in which the term belief is used to talk about what @janmaat_subjective_2013 refers to as perceptions, whereas other authors use general terms such as attitudes [@kunovich_systems_2007].
+For instance,  in @reynolds_perceptions_2014, the term belief is used to talk about what @janmaat_subjective_2013 refers to as perceptions, whereas other authors use general terms such as attitudes [@kunovich_systems_2007].
 The first attempt to shed light on this issue in meritocracy research was made by @duru-bellat_who_2012, who used the question "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" as a proxy for "desired" meritocracy (normative beliefs).
@@
 Is the belief in meritocracy a perception or a preference with normative meaning?
-In order to expand the analytical conceptual framework, we believe that both dimensions should be included as proposed by @duru-bellat_who_2012.
-This opens up the possibility of analyzing whether perceptions and preferences are actually related (i.e.
-have a high correlation), or whether they are independent aspects of the same phenomenon.
-As @sonhing_merit_2011 has pointed out, "People can believe that outcomes ought to be distributed on the basis of merit and yet vary in their perceptions of whether this is how society currently operates" (p.
+To expand the analytical conceptual framework, we believe that both dimensions should be considered, as proposed by @duru-bellat_who_2012.
+Such an approach opens up the possibility of analyzing whether perceptions and preferences are related (i.e., have a high correlation), or constitute independent aspects of the same phenomenon.
+As @sonhing_merit_2011 has pointed out, "People can believe that outcomes ought to be distributed based on merit and yet vary in their perceptions of whether this is how society currently operates" (p.
 435).
-In other words, normative beliefs should be considered while taking perceptions into account: a strong normative belief in meritocracy may mean something totally different to someone perceiving high meritocracy than to someone perceiving low meritocracy.
+In other words, normative beliefs should be considered while taking perceptions into account: a strong normative belief in meritocracy may mean something different to someone perceiving high meritocracy than to someone perceiving low meritocracy.
 To avoid the confusion generated by the term "belief", we propose the terms meritocratic preferences ("what should be"), and meritocratic perceptions ("what is"), as they better reflect the two facets of meritocracy under scrutiny [@castillo_meritocracia_2019].
@@
 
-Some research in meritocracy considers aspects usually opposed to effort and talent for personal achievement, as for instance family status and the use of personal advantages (as contacts or having a wealthy family) to get ahead in life.
+Some research in meritocracy considers aspects usually opposed to effort and talent for personal achievement, for instance, family status and the use of personal advantages (as contacts or having a wealthy family) to get ahead in life.
 This distinction refers to the classical achieved and ascribed status dimension from @linton_study_1936.
-For instance, @kunovich_systems_2007 used an items' battery listing a number of factors in relation to "How important each should be in deciding pay..." (as @duru-bellat_who_2012 for desired meritocracy).
-They consider factors such as education and responsibility as meritocratic, giving them a value of 1 is considered "essential" in the scale response, whereas factors such as having a family or children are valued 1 when rated as "not important at all" (i.e.
+For instance, @kunovich_systems_2007 used an items' battery following the question “How important each should be in deciding pay..." (as @duru-bellat_who_2012 for desired meritocracy).
+They consider factors such as education and responsibility as meritocratic, giving them a value of 1 is considered "essential" in the scale response, whereas factors such as having a family or children are valued as 1 when rated as "not important at all" (i.e.
 reverse coded).
-The assumption behind this approach is that rejecting a supposed ascribed or non-meritocratic^[Although the use of the concept "non-meritocratic" is very broad, we prefer to use it instead of something like ascription, given that this last one leaves out aspects such as the use of personal relationships.] aspect (as having family and/or children) implies a stronger belief in meritocracy.
+The assumption behind this approach is that rejecting a supposed ascribed or non-meritocratic aspect (as having family and/or children) implies a stronger belief in meritocracy.
 A similar approach of reverse-coding non-meritocratic items was taken by @newman_false_2015, using the same principles applied in the "Preference for the Merit Principle Scale" [@davey_preference_1999].
@@
 The assumption that meritocratic and non-meritocratic elements are poles of the same continuum was tested by @reynolds_perceptions_2014 using the "get ahead" perceptions' battery items mentioned above.
-They considered education, ambition, and hard work as meritocratic, whereas factors such as family wealth and connections where classified as non-meritocratic.
+They considered education, ambition, and hard work as meritocratic, whereas factors such as family wealth and connections were classified as non-meritocratic.
 Despite making this distinction, the authors end up subtracting one dimension from the other, thus coming back to the assumption that they are two poles of the same continuum, as @kunovich_systems_2007 also did.
@@
 For instance, such distinction could open the door to study to what extent different and apparently contradictory views of meritocracy can coexist, to what extent individuals and societies differ to this regard, and of course the associated factors to such differences.
+Although the use of the concept "non-meritocratic" is very broad, we prefer to use it instead of something like ascription, given that this last one leaves out aspects such as the use of personal relationships.
 
@@
 
-Finally, most studies in meritocracy so far have not properly considered the issue of latent structures and measurement error [@brown_confirmatory_2015; @bollen_structural_1989; @ansolabehere_strength_2008], as they mostly use single indicators and/or simple average indexes for measuring meritocracy.
-Such strategy assumes that the latent construct is measured perfectly (i.e.
-no error or residual variance) by the selected indicators, going as far as to propose that "...
+Finally, most meritocracy studies have not adequately considered the issue of latent structures and measurement error [@brown_confirmatory_2015; @bollen_structural_1989; @ansolabehere_strength_2008], as they mostly use single indicators and/or simple average indexes for measuring meritocracy.
+Such a strategy assumes that the latent construct is measured perfectly (i.e., no error or residual variance) by the selected indicators, going as far as to propose that "...
 In choosing this strategy of index construction, we argue that support for meritocracy is not a latent variable" [@kunovich_systems_2007 p.653-654].
-Although some advances were made by Reynolds and Xian [-@reynolds_perceptions_2014] through conducting a principal component analysis of meritocratic and non-meritocratic dimensions, they finally chose to build a sum index despite proving a multidimensional latent structure.
-
-There is a relevant trade-off to consider when it comes to measurement error estimation in survey scales.
+Although Reynolds and Xian [-@reynolds_perceptions_2014] made advances by conducting a principal component analysis of meritocratic and non-meritocratic dimensions, they finally chose to build a sum index despite proving a multidimensional latent structure.
+
+There is a relevant trade-off regarding measurement error estimation in survey scales.
 Including multiple items per conceptual dimension opens the possibility to validity studies (at least in terms of construct) by using factor analysis techniques, particularly confirmatory ones.
@@
 
-- Multiple indicators for each dimension, in order to _account for measurement error_ in a confirmatory factor analysis context.
-
-- Based on _previous indicators_ as much as possible, for the sake of keeping comparability between studies.
+- Multiple indicators for each dimension to _account for measurement error_ in a confirmatory factor analysis context.
+
+- Based on _previous indicators_ as much as possible to keep comparability between studies.
 As most of the survey research on meritocracy so far has been made with the inequality module of the ISSP survey, we attempted to rescue as much as possible the original wording of some of its items in the new scale proposal.
@@
 - _Brief_, as to be used in regular public opinion surveys [@davidov_measurement_2009].
-To this regard, it differs from the proposal of  "Preference for the Merit Principle Scale" [@davey_preference_1999], as they use 15 items for just one dimension (aside from the problem of reverse-coding non-meritocratic items).
+In this regard, it differs from the proposal of  "Preference for the Merit Principle Scale" [@davey_preference_1999], as they use 15 items for just one dimension (aside from the problem of reverse-coding non-meritocratic items).
 
@@
 
-The data corresponds to the last available wave of the social inequality module from the International Social Survey Programme (ISSP), which is the most specialized international comparative survey in perceptions, attitudes, and beliefs about inequality-related issues.
-This wave corresponds to the year 2009 and covers attitudes towards a series of topics dealing with social inequality across 41 countries and 56,021 individuals.
-Although there is data available from this module for previous waves (1987, 1992, and 1999), unfortunately, there are several variables that are important for this research which were not included, particularly in 1987 and 1999, reason why we only use the 2009 wave (during the development of this research the 2019 wave became available, but so far with fewer countries than the previous one).
-Details about the sample of countries can be seen on Table \@ref(tab:countries).
+The data corresponds to the last available wave of the social inequality module from the International Social Survey Programme (ISSP), the most specialized international comparative survey in perceptions, attitudes, and beliefs about inequality-related issues.
+This wave corresponds to 2009 and covers attitudes towards a series of topics dealing with social inequality across 41 countries and 56,021 individuals.
+Although there is data available from this module for previous waves (1987, 1992, and 1999), unfortunately, there are several variables that are important for this research that were not included, particularly in 1987 and 1999; the reason why we only use the 2009 wave (during the development of this research the 2019 wave became available, but so far with fewer countries than the previous one).
+Table \@ref(tab:countries) shows details about the sample of countries.
 
@@
 Figure \@ref(fig:likert-issp) shows the distribution of responses across the selected items in their corresponding dimensions.
-On the one side, we see that there is a high degree of importance attributed to factors such as hard work and ambition in the process of getting ahead, concentrating 94.9% and 92.8% in the fairly important to essential categories.
+On the one side, we see a high degree of importance attributed to factors such as hard work and ambition for getting ahead, concentrating 94.9% and 92.8% in the fairly important to essential categories.
 By contrast, the perception of non-meritocratic aspects is rated lower than the meritocratic ones, particularly for gender, race, and political connections.
-Regarding normative preferences, we observe that the meritocratic ones are deemed as important (from fairly to essential) for nearly the whole sample, support that decreases slightly for the non-meritocratic ones.
+Regarding normative preferences, we observe that the meritocratic ones are deemed as important (from fairly to essential) for nearly the whole sample, decreasing slightly for the non-meritocratic ones.
 Still, we have to consider that the preferences for non-meritocratic aspects, in this case, refer to the distributive principle of need rather than personal background (as in the perceptions of non-meritocracy), which certainly is one of the limitations of this set of items for the operationalization of the proposed conceptual framework with this dataset.
@@
 The exception here is the dimension of non-meritocratic perception, in which there are six items that appear mostly associated by pairs according to their specific topics (family background, personal background, and connections).
-Still, between this same set of items, there are moderate correlations (around 0.3) which could anticipate a single latent factor underlying non-meritocratic perception and that is tested next through confirmatory factor analysis.
+Still, between this same set of items, moderate correlations (around 0.3) could anticipate a single latent factor underlying non-meritocratic perception, which is tested next through confirmatory factor analysis.
 
@@
 Table \@ref(tab:sum-fit-issp) shows the results of the estimation of two confirmatory models.
-The first one (First Order) corresponds to a model that estimate four factors, each for one of the dimensions presented in \@ref(tab:table-issp), showing only regular fit indicators (CFI=0.959, TLI=0.944, RMSEA=0.098, $\chi^2$(df=48)= 21308.535).
+The first one (First Order) corresponds to a model that estimate four factors, each for one of the dimensions presented in \@ref(tab:table-issp), with only regular fit indicators (CFI=0.959, TLI=0.944, RMSEA=0.098, $\chi^2$(df=48)= 21308.535).
 Further analysis showed that the sources for poor fit were mostly related to the items of the non-meritocratic perceptions, which as mentioned above displayed correlations among them not taken into account in this model specification.
@@
 Although it was possible to get an adequate fit of the model to the pooled data from ISSP, there are several aspects to take into account when using these items for measuring meritocracy.
-First of all, the items were not designed based on this conceptual approach and a first estimation approach (first order factor model) reveals some problems of fit.
+First of all, these items were not specifically designed based on a conceptual approach to meritocracy, but it still is possible to find a certain fit of the model to the data after minor adjustments.
 Secondly, some of the items in similar dimensions refer to different concepts, as for instance meritocratic perceptions are related to ambition and hard work, whereas meritocratic preferences deal with how well a job is done.
 A better operationalization would require more alignment between the different sides of the same concept.
-And thirdly, the scale is rather unbalanced given he larger amount of non-meritocratic perception items, producing correlated errors.
-All these limitations are considered in the design of Study 2.
-
-With regard to the meritocratic / non-meritocratic distinction in previous studies, we can preliminary conclude that the wide use of reverse-coding for non-meritocratic items attempting to measure meritocracy is not an adequate operationalization of the construct.
-Meritocracy and non-meritocracy are related but different constructs, meaning that it is theoretically possible to find high and lower level of both.
-This finding opens a series of avenues for future research that are developed in the conclusions.
+And thirdly, the scale is rather unbalanced given the larger amount of non-meritocratic perception items, producing correlated errors.
+All these limitations are taken into account next in the design of Study 2.
+
+Concerning the meritocratic / non-meritocratic distinction in previous studies, we can preliminary conclude that the wide use of reverse-coding for non-meritocratic items attempting to measure meritocracy is not an adequate operationalization of the construct.
+Meritocracy and non-meritocracy are related but different concepts, meaning that it is theoretically possible to find different combinations at individual and contextual levels.
+This finding opens a series of avenues for future research that are considered in the conclusions.
 
@@
 
-The data was obtained through an online survey which was part of a larger study on meritocracy and preferences developed in Chile in 2019, funded by the Chilean National Scientific Agency (ANID).
+The data comes from an online survey which was part of a larger study on meritocracy and preferences developed in Chile in 2019, funded by the Chilean National Scientific Agency (ANID).
 The questionnaire was programmed in Qualtrics and the fieldwork was conducted by an external online survey agency ([netquest.com](www.netquest.com)) between December 2019 and January 2020.
@@
 The quotas for gender, age, and educational levels were generated based on a survey by the Public Studies Center [@cep_encuesta_2019], which is a well-regarded counterpart agency of the ISSP (International Social Survey Programme) in Chile.
-A total sample of 2,141 people was collected, excluding those who did not sign the informed consent as well as those not answering the meritocracy response battery.
-There were no significant differences between our sample and the wider population for most socio-demographic characteristics, with the exception of educational level (see Table \@ref(tab:rep-samp) in Appendix).
+A total sample of 2,141 people was collected, excluding those who did not sign the informed consent and those not answering the meritocracy response battery.
+There were no significant differences between our sample and the wider population for most socio-demographic characteristics, except for the educational level (see Table \@ref(tab:rep-samp) in Appendix).
 As it is often the case with online surveys, there were some limitations in achieving the quotas for lower educational levels [@zhang_quota_2018; @boas_recruiting_2020].
@@
 
-The proposed scale of perceptions and preferences relating to meritocracy consisted of eight indicators, two for each of the four dimensions listed earlier: Perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic).
-This design is oriented by two main criteria.
-The first one is usability in survey research, for which the length of the scale is restricted due to space reasons in such studies.
+The proposed scale of meritocratic perceptions and preferences consisted of eight indicators, two for each of the four dimensions listed earlier: Perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic).
+Two main criteria oriented this design.
+The first one was usability in social survey research, for which the scale length must attend spacer restrictions in such studies.
 Nevertheless, at the same time the scale should offer possibilities of accounting for measurement error through factor analysis techniques.
-Although a rule of thumb in latent  measurement suggests at least three indicators per dimension, we still opted for two in order to count with a brief scale.
+Although a rule of thumb in latent measurement suggests at least three indicators per dimension, we still opted for two to count with a brief scale.
 Although such decision posits some limitation for modeling measurement error, still allows for factor analysis and is a common approach in previous social survey studies with limited items (for instance see @davidov_measurement_2009).
@@
 With the objective of evaluating the effect of indicator ordering in the responses, three different versions of items' order were designed and randomly assigned to the respondents, as depicted in Figure \@ref(fig:appmod).
-The scale was presented to the first group (_n = 712_) in the order that appears in Table \@ref(tab:table-indicadores) according to perceptions and preferences.
-For the second group (_n = 717_), the order was reorganized according to perceptions and preferences over the same topic, e.g.
-for the topic of hard work, the item about perception was followed by the item about preference and the same for the rest of the topics.
+The scale was presented to the first group (_n = 712_) in the order shown in Table \@ref(tab:table-indicadores) according to perceptions and preferences.
+For the second group (_n = 717_), the order was reorganized according to perceptions and preferences over the same topic, e.g., for the topic of hard work, the item about perception was followed by the item about preference, and the same for the rest of the topics.
 Finally, for the third group (_n = 712_), the items were presented as completely randomized.
@@
 The models estimated one factor for each of the four proposed dimensions presented in Table \@ref(tab:table-indicadores).
-As in Study 1, CFA was conducted using the `lavaan` R package (version 0.6-3; @rosseel_lavaan_2012a), with diagonally weighted least squares (DWLS) estimation due to the items' ordinal level of measurement [@kline_principles_2016; @rosseel_lavaan_2012a].
+As in Study 1, CFA was performed with the `lavaan` R package (version 0.6-3; @rosseel_lavaan_2012a), with diagonally weighted least squares (DWLS) estimation due to the items' ordinal level of measurement [@kline_principles_2016; @rosseel_lavaan_2012a].
 The fit indexes and cut-off criteria were the same as the ones used in Study 1.
@@
 A pre-registration was made in the OSF platform, available at the following link: [https://osf.io/z45y2](https://osf.io/z45y2).
-Included in this pre-registration are the hypotheses regarding the four-dimensional conceptual model underlying the scale, the variable measurement levels, the statistical tests to be performed with their respective evaluation parameters, and other important aspects of the research design.
+This pre-registration includes the hypotheses regarding the four-dimensional conceptual model underlying the scale, the variable measurement levels, the statistical tests to be performed with their respective evaluation parameters, and other important aspects of the research design.
 
@@
 The graphs presented in Figure \@ref(fig:plotlikert) display disaggregated and comparable information of the different response categories for each item.
-As it can be observed, in general, there is more agreement in the perception of non-meritocratic items than in meritocratic ones, while in the case of preferences the opposite occurs.
-As far as preferences are concerned, the preponderant role of effort over talent as a criterion of meritocratic preference is noteworthy.
-All in all the descriptive results tend to show a sort of critical view of meritocracy, perceiving the operation of non-meritocratic aspects over meritocratic ones, whereas in the preferences the opposite occurs.
+Generally, there is more agreement in the perception of non-meritocratic items than in meritocratic ones, while in the preferences, the opposite occurs.
+Regarding preferences, the predominant role of effort over talent as a criterion of meritocratic preference is noteworthy.
+All in all, the descriptive results show a critical view of meritocracy, perceiving the operation of non-meritocratic aspects over meritocratic ones, whereas in the preferences the opposite occurs.
 
@@
 Firstly, and as expected, the largest correlations are between indicators that correspond to the same factors behind the conceptual model (e.g., perception of meritocracy by effort and by talent, $r=0.52$).
-Secondly, among these correlations the highest are those between the non-meritocratic dimension but in perceptions ($r=0.73$) and preferences ($r=0.61$).
-Thirdly, both items for meritocratic preferences (E and F) are the ones that mostly correlate with the rest of the perceptual items, showing medium to high correlations.
-This is noteworthy because it would indicate that the perception of non-meritocracy would be related to larger meritocratic preferences.
-Finally, and similar to what was found in Study 1 with ISSP data, we observe that there are no considerable negative correlations between meritocratic and non-meritocratic aspects, undermining the assumptions of previous studies that suggested that these dimensions would be the opposite poles of one same continuum [@reynolds_perceptions_2014].
+Secondly, among these correlations, the highest are those between the non-meritocratic dimension but in perceptions ($r=0.73$) and preferences ($r=0.61$).
+Thirdly, both items for meritocratic preferences (E and F) are the ones that mostly correlate with the rest of the perceptual items, showing medium to high correlations and indicating that the perception of non-meritocracy would be related to larger meritocratic preferences.
+Finally, similar to the results of Study 1 with ISSP data, we observe no considerable negative correlations between meritocratic and non-meritocratic aspects, undermining previous studies' assumptions that these dimensions would be the opposite poles of one same continuum [@reynolds_perceptions_2014].
 
@@
 The model estimates four latent factors: perception meritocratic, perception non-meritocratic, preferences meritocratic, and preferences non-meritocratic.
-Each factor is estimated based on two items of the scale as detailed in Table \@ref(tab:table-indicadores).
-
-The first step in the analysis consists of comparing the model fit indicators for the three versions of the scale that were randomly assigned to the participants: order according to perceptions/preferences, order according to topics, and random order (see Figure \@ref(fig:appmod)).
+Each factor is estimated based on two scale items as detailed in Table \@ref(tab:table-indicadores).
+
+The first step in the analysis consists of comparing the model fit indicators for the three versions of the scale that were randomly assigned to the participants: order according to perceptions/preferences, order according to topics, and a complete random items’ order (see Figure \@ref(fig:appmod)).
 
@@
 Regardless of the version, all models obtained adequate fit indicators, with CFI's above 0.95 and RMSEA's below 0.08.
-However, none of the models achieved a non-significant chi-square, something expected in large samples as the one used here.
-The first version order (perceptions-preferences) was the one obtaining best fit (CFI=0.993, TLI=0.995, RMSEA=0.034, $\chi2$(df=14)=42.276), whereas version 2 with the fixed order according to merit/non-merit items shows the comparatively worst indicators.
-The CFA fit indices for the completely randomized items' order (Model 3) it keeps all the indicators within the acceptable cut-off criteria and besides it controls for possible order effects in the administration of the instrument.
+However, none of the models achieved a non-significant chi-square, as expected in large samples like the one used here.
+The first version order (perceptions-preferences) was the one obtaining the best fit (CFI=0.993, TLI=0.995, RMSEA=0.034, $\chi2$(df=14)=42.276), whereas version 2 with the fixed order according to merit/non-merit items shows the comparatively worst indicators.
+The CFA fit indices for the completely randomized items' order (Model 3) keep all the indicators within the acceptable cut-off criteria and controls for possible order effects in the administration of the instrument.
 The model and parameter estimates for this version are depicted in Figure \@ref(fig:meas02):
@@
 Attending to the correlations between the latent variables as depicted in Figure \@ref(fig:meas02), meritocratic preferences are moderate to highly correlated with perceptions, both meritocratic ($r=0.457$) and non-meritocratic ($r=0.500$).
-The correlation between both types of perceptions ($r=-0.044$) and both types preferences ($r=0.185$) are low, as well as between non-meritocratic preferences and perceptions ($r=-0.059$).
+The correlation between both types of perceptions ($r=-0.044$) and both types of preferences ($r=0.185$) was low, as well as between non-meritocratic preferences and perceptions ($r=-0.059$).
 This last finding gives further evidence regarding the lack of unidimensionality of meritocratic and non-meritocratic aspects as assumed by previous studies.
@@
 The estimation performed with different administration sets for the meritocracy scale gives evidence in line with the conceptual model.
-The four dimensions appear as independent but related, although with some differences when compared with the estimation with ISSP data in Study 1.
-The main general difference is that with ISSP data the four latent dimensions were positively correlated, whereas now some correlations are statistically not different from zero.
+The four dimensions appear independent but related, with some differences when compared with the estimation with ISSP data in Study 1.
+The main general difference is that with ISSP data, the four latent dimensions were positively correlated, whereas now some correlations are statistically not different from zero.
 For instance, perceptions and preferences for meritocracy keep being related, which is not the case for the non-meritocratic pair.
-In principle this makes more sense in theoretical terms, as if someones perceives that meritocracy functions well, probably will support it more.
-On the other side, if someones perceives non-meritocracy, this should lead to a larger preference for meritocracy as it would constitute a demand for lager meritocracy when this one is not perceived to work.
-
-Despite the good general working of the scale, there are some limitations in the interpretation of the results given the few items (two) per dimension.
-As it was mentioned before, the option of a minimalistic approach in terms of items has advantages and disadvantages: advantages in terms of usability in social surveys with limited space, and disadvantages in terms of measurement validity in the estimation of the latent constructs.
-Being aware of this trade-off, in the following section we offer additional validity evidence as a way of complementing some of the weaknesses identified.
+Theoretically, this makes more sense, as if someone perceives that meritocracy functions well, they probably will support it more.
+On the other side, if someone perceives non-meritocracy, this should lead to a more prominent preference for meritocracy as it would constitute a demand for a larger meritocracy when it is not perceived to work.
+
+Despite the good general working of the scale, there are some limitations in interpreting the results given the few items (two) per dimension.
+As mentioned before, the option of a minimalistic approach in terms of items has advantages and disadvantages: advantages in terms of usability in social surveys with limited space and disadvantages in terms of measurement validity in the estimation of the latent constructs.
+Aware of this trade-off, next we present additional evidence of validity to complement some of the weaknesses identified.
 
@@
 The modeling for invariance testing has been gaining more relevance in sociological survey studies, mainly due to the interest in the validity of measurement scales in comparative studies in various social and cultural contexts [@davidov_measurement_2014].
-In this section we present the results of an analysis of invariance between groups, testing whether the proposed factorial structure for the meritocracy scale is applicable in a different dataset.
-
-The invariance measurement procedure consists of a series of nested models to which restrictions are progressively incorporated on the parameters of the measurement model.
-The literature generally suggests that this modeling should be done through four levels or types of progressive restriction [@milfont_testing_2010; @millsap_statistical_2011; @vandeschoot_checklist_2012]:
+In this section, we present the results of an analysis of invariance between groups, testing whether the proposed factorial structure for the meritocracy scale is applicable in a different dataset.
+
+The invariance measurement procedure consists of a series of nested models in which restrictions are progressively incorporated into the parameters of the measurement model.
+The literature generally suggests that this modeling approach should consider four levels or types of progressive restriction [@milfont_testing_2010; @millsap_statistical_2011; @vandeschoot_checklist_2012]:
 
@@
 
-2. **Weak:** an equality restriction is applied to the factor loadings in the different groups, that is, the loadings are forced to be identical in both measurements.
-
-3. **Strong:** equality restrictions are added to the intercepts of each indicator.
-
-4. **Strict:** equality restrictions are added to the error variances of each indicator.
+2. **Weak:** an equality restriction  applied to the factor loadings in the different groups; that is, the loadings are forced to be identical in both measurements.
+
+3. **Strong:** equality restrictions added to the intercepts of each indicator.
+
+4. **Strict:** equality restrictions added to the error variances of each indicator.
 
@@
 
-For testing the invariance we compared the data described previously in Study 2 to a different  dataset coming from an online survey carried out during the first half of 2020.
+For testing the invariance, we compared the data described previously in Study 2 to a different dataset from an online survey carried out during the first half of 2020.
 The characteristics of this sample in terms of application and coverage are equivalent to those of the previous study.
-The final sample obtained contains 1,242 cases, where 605 (48.71%) are women and 637 (51.28%) are men.
-48.84% of them are 45 years of age or older, and 33.97% have tertiary education or higher.
+The final sample contains 1,242 cases, where 605 (48.71%) are women, and 637 (51.28%) are men.
+48.84% are 45 or older, and 33.97% have tertiary education or higher.
 
@@
 
-The items of the meritocracy scale were identical to the original and they  were administrated in a randomized order (which corresponds to the application modality of Group 3 in Study 2, as showed in Figure \@ref(fig:appmod)).
+The items of the meritocracy scale were identical to the original scale of Study 2 administrated in a randomized order (which corresponds to the application modality of Group 3 in Study 2, as shown in Figure \@ref(fig:appmod)).
 
@@
 
-The first step for invariance testing is the estimation of the configural model, which serves as the baseline for further comparison and is expected to adequately meet the global fit criteria of a measurement model.
-Although the chi-square statistic is used as a global measure of fit, since its sensitivity to sample size it is recommended to use three additional fit indices: Comparative Fit Index (CFI), which should have a value greater than 0.95; Root Mean Square Error Approximation (RMSEA), which must be lower than 0.06; and the Standardized Root Mean Square Residuals (SRMR) which must be less than 0.08 [@vandeschoot_checklist_2012].
-Besides, the literature on measurement invariance suggests some complementary approaches for the evaluation of the fit, from which we will consider the incremental adjustment of the fit indexes [@cheung_evaluating_2002; @milfont_testing_2010; @dimitrov_testing_2010] and the ANOVA test for means comparison in nested models [@newsom_longitudinal_2015].
+The first step for invariance testing is estimating the configural model, which serves as the baseline for further comparison and is expected to adequately meet the global fit criteria of a measurement model.
+Although the chi-square statistic is used as a global measure of fit, it is usually complemented with other indexes given its high sensitivity to sample size: the Comparative Fit Index (CFI), which should have a value greater than 0.95; the Root Mean Square Error Approximation (RMSEA), which must be lower than 0.06; and the Standardized Root Mean Square Residuals (SRMR) which must be less than 0.08 [@vandeschoot_checklist_2012].
+Besides, the literature on measurement invariance suggests some complementary approaches for evaluamerit scale the fit, from which we will consider the incremental adjustment of the fit indexes [@cheung_evaluating_2002; @milfont_testing_2010; @dimitrov_testing_2010] and the ANOVA test for means comparison in nested models [@newsom_longitudinal_2015].
 
@@
 Based on the measurement invariance analysis for two datasets, we found evidence supporting the equivalence of the scale when applied in different samples.
-This gives a more robust ground when using this measure for comparing groups, as for instance country comparison.
-Although the results are promising, still the analysis was performed in two samples in one country (Chile) and it requires further examination with social surveys in other societies.
+This gives a more robust ground when using this measure for comparing groups, as for instance in country comparisons.
+Although the results are promising, still the analysis was performed in two samples in a single country (Chile) and it requires further examination with social surveys in other societies.
 
@@
 As most studies use secondary survey data, they tend to assume that the available indicators represent an underlying meritocratic construct.
-A review of these studies reveals several non-tested assumptions, as well as the use of similar indicators to represent different constructs and dimensions of meritocracy.
+A review of these studies reveals several non-tested assumptions and the use of similar indicators to represent different constructs and dimensions of meritocracy.
 As the existence of heterogeneous approaches certainly has consequences for the advancement of the study of meritocracy, this paper presented a comprehensive conceptual framework for the empirical study of meritocracy, building upon previous research.
@@
 The proposed 8-item scale, "Perceptions and Preferences for Meritocracy", was designed and tested in order to deal with these four issues.
-Our results indicate that perceptions and preferences seem to be two related but different dimensions, often confounded in previous research under the label of "beliefs".
+Our results indicate that perceptions and preferences are two related but different dimensions, often confounded in the previous research under the label of "beliefs".
 At the same time, meritocratic and non-meritocratic dimensions do not appear to constitute poles of the same continuum, as some previous studies have assumed.
@@
 One of the criteria in designing the merit scale was to develop a short instrument that could be used in social surveys, assuming that using only two items per dimension posits limitations for the confirmatory analysis estimated here.
-We preferred this confirmatory approach at it offers several advantages in terms of information about the quality of the measurement, and we offered additional validity evidence with measurement invariance analysis.
+We preferred this confirmatory-minimalistic approach at it offers several advantages in terms of information about the quality of the measurement, and we offered additional validity evidence with measurement invariance analysis.
 Despite the advantages, a more simple and straightforward perspective is always available, as for instance using only the average of the items per dimension.
 Whereas a short scale it is always appreciated in social survey studies, for other audiences could not reach measurement standards.
-This may require further expansions and/or consider using a different instrument as the "Preference for the Merit Principle Scale" [@davey_preference_1999], which measures only one dimension of meritocracy with fifteen items.
+This may require further expansions and/or consider using a different instrument as the "Preference for the Merit Principle Scale" [@davey_preference_1999] (which measures only one dimension of meritocracy with fifteen items).
 
@@
 Additionally, given that non-meritocratic factors are not necessarily related to meritocratic ones, our framework makes it possible to assess the perceived legitimacy of practices such as the use of personal contacts and their interference (or not) with meritocratic ideals in different societies.
-The consideration of structural factors, both individual and societal, will allow us to advance future hypotheses relating to meritocratic legitimacy to a more comprehensive level than the current one, which is based solely on meritocratic perceptions.
-Furthermore, the impact of different configurations of the four-dimensional framework on practices and behaviors such as corruption, civic involvement, and political alignment, is an area that requires additional research.
+Considering individual and societal structural factors affecting meritocracy will allow us to advance future hypotheses relating to meritocratic legitimacy to a more comprehensive level than the current one, which is based mostly on meritocratic perceptions.
+Furthermore, the impact of different configurations of the four-dimensional framework on practices and behaviors, such as corruption, civic involvement, and political alignment, are areas open to further research.
 Such future agendas could be especially relevant in times of economic crisis and growing inequalities, that could entail changes in the legitimation of the current distributive structure based on meritocratic ideals.
@@
 
-This research follows a series of open science guidelines as the pre-registration of Study 2 (scale development) as well as the availability of data and codes (R) in an public repository in Github.
+This research follows a series of open science guidelines as the pre-registration of Study 2 (scale development) as well as the availability of data and codes (R) in a public repository in Github.
 This document was generated with the R package Rmarkdown and contains the code for all the tables and figures in order to make it reproducible.
@@
 ```
+
+ing Perceptions and Preferences for Meritocracy"
+css: "custom.css"
+linestretch: '1.5'
+link-citations: yes
+# author: "Juan Carlos Castillo, Julio Iturra, Luis Maldonado, Francisco Meneses & Jorge Atria <br><br>"
+abstract: "Economic and social inequalities have generated growing concerns and crises across contemporary societies.
+The belief in meritocracy has been suggested as one of the factors that explain the persistence of inequality, as it would legitimize economic disparities based on differences in effort and talent.
+Despite its wide use as a concept, empirical research on meritocracy is relatively novel and characterized by diverse conceptualizations and measures that make the findings and their interpretation rather inconsistent.
+Most of the studies in the area have relied upon secondary data from different surveys to operationalize meritocracy, with a wide variation in the use and interpretation of the same measuring items.
+Based on a critical analysis of concepts and operationalizations of meritocracy in survey research, we propose a conceptual framework for measuring perceptions and preferences for meritocracy and non-meritocracy in a minimalistic version (few items as possible), in order to be considered in comparative survey research studies.
+The proposal is first tested through confirmatory factor analysis (CFA) using available ISSP (International Social Survey Programme) inequality module items.
+Based on these results, in a second stage we developed a improved scale that better suits the perceptions/preferences distinction, which was applied in a Chilean sample (N=2,141).
+The results of the CFA estimation give evidence of a better fit for the new scale.
+Our discussion highlights the importance of considering different dimensions in order to advance in the study of meritocracy."
+output:
+  bookdown::pdf_document2:
+     template: null
+     toc: false
+     number_sections: false
+  bookdown::word_document2:
+    number_sections: false
+  bookdown::html_document2:
+    number_sections: false
+linkcolor: blue
+bibliography:
+  - input/bib/meritocracy.bib
+csl: input/bib/apa7-singlespace.csl
+editor_options:
+  chunk_output_type: console
+geometry: margin=0.78in
+header-includes:
+  - \usepackage{times}           # Times New Roman
+  - \usepackage{caption}
+  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
+  - \usepackage[figurename=Fig.]{caption}
+  - \usepackage{graphicx}
+  - \usepackage{float}
+  - \usepackage{booktabs}
+  - \usepackage{longtable}
+  - \usepackage{array}
+  - \usepackage{multirow}
+  - \usepackage{wrapfig}
+  - \usepackage{float}
+  - \usepackage{colortbl}
+  - \usepackage{pdflscape}
+  - \usepackage{tabu}
+  - \usepackage{threeparttable}
+always_allow_html: true # for word output
+fontsize: 10pt
+---
+
+```{r eval=FALSE, include=FALSE}
+ # for render in pdf run rmarkdown::render_site("docs/paper.Rmd", output_format = "all")
+ # clean #in the yml
+ rmarkdown::render("paper-blinded.Rmd", output_format = "bookdown::pdf_document2")
+ rmarkdown::render("paper-blinded.Rmd", output_format = "bookdown::html_document2")
+ # rmarkdown::render("paper.Rmd", output_format = "bookdown::word_document2")
+```
+
+```{r setup, include=FALSE}
+ knitr::opts_chunk$set(echo=FALSE, warning = FALSE,message = FALSE, cache = FALSE,out.width = '85%',fig.pos= "H", fig.align = 'center')
+ # knitr::opts_knit$set(base.url = "../") #relative path for .html output file
+ # knitr::opts_knit$set(root.dir = "../") #relative path for chunks within .rmd files
+ options(scipen=999)
+ options(kableExtra.auto_format = FALSE)
+ rm(list=ls())
+ options(knitr.kable.NA = '')
+ options(knitr.graphics.error = FALSE)
+ Sys.setlocale("LC_ALL", "ES_ES.UTF-8")
+```
+
+<!---
+\pagebreak
+
+# Acknowledgments
+
+The authors recognize the support of ANID/FONDECYT grant 1160921 and of  the Center for Conflict and Social Cohesion Studies (COES)  grant ANID/FONDAP 1513009.
+
+\pagebreak
+
+--->
+
+# Introduction
+
+Economic inequality and income concentration have become topics of growing concern over the last years, leading to a series of social movements in different societies as well as diverse critical analyses regarding the development of capitalism and its consequences [@streeck_politics_2014; @piketty_capital_2014].
+In such a context, the study of views, preferences, and perceptions of inequality has acquired relevance in the social sciences, in topics such as redistributive preferences [@alesina_fairness_2005; @dimick_models_2018], the legitimization of economic inequality [@schroder_income_2017] and the functioning of meritocracy [@duru-bellat_who_2012; @mijs_paradox_2019; @reynolds_perceptions_2014; @atria_economic_2020].
+Within this research area, and particularly in times of economic and health crises, the ideal of meritocracy has been strongly challenged as an unfulfilled promise of modern societies that allows the perpetuation of social inequalities [@sandel_tyranny_2020; @goldthorpe_myth_2003].
+
+In general, meritocracy has been defined as a system of distribution of resources and rewards based on individual merit, which in its original conception is a combination of talent and effort [@young_rise_1962].
+This traditional conception of merit places in a secondary position the possible interference of structural or non-meritocratic factors, such as inheritance, personal contacts, and luck [@breen_class_1999; @saunders_might_1995; @yair_meritocracy_2007; @land_we_2006; @young_rise_1962].
+Social psychology and sociology have studied the characteristics and consequences of beliefs in meritocracy, under the general hypothesis that a greater belief in meritocracy emphasizes the role of the individual over structural factors in personal achievements, leading to greater legitimization of inequalities [@preminger_meritocracy_2020; @trump_when_2020; @hadjar_meritokratie_2008; @madeira_primes_2019].
+Such research has raised criticism of meritocracy as a moral standard of distribution given the preponderant weight of non-meritocratic elements upon the individual status and social mobility [@sandel_tyranny_2020; @witteveen_reconsidering_2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_meritocracy_2019; @khan_privilege_2013].
+
+Due to the role that meritocratic beliefs play in the justification of individual achievement (or failure) in contemporary societies [@hadjar_meritokratie_2008; @markovits_meritocracy_2019; @sandel_tyranny_2020], multiple studies have evaluated the relationship between meritocratic beliefs and personal and/or contextual characteristics.
+For example, some studies have linked meritocracy to the reinforcement of socio-economic, gender, and ethnic stereotypes [@madeira_primes_2019; @girerd_neoliberalism_2020; @girerd_neoliberalism_2020], as well as the effects of meritocratic beliefs in educational [@generett_stories_2020; @owens_engines_2020] and organizational contexts [@perez_advancing_2020; @aiello_new_2019].
+Most of such studies so far have used indicators from existing standard social surveys, such as the International Social Survey Programme (ISSP), to measure meritocracy.
+However, as we will show later, the concepts used as well as the instruments for measuring meritocracy vary extensively among studies.
+In many cases, similar phenomena are associated with different indicators, and conversely, often different phenomena are measured with similar indicators.
+Such trends not only limit the comparability of studies, but also the ability to understanding the effects of meritocratic beliefs across different disciplines and lines of research.
+
+Based on a critical analysis of different approaches to the measurement of meritocracy to date, this article proposes a conceptual and measurement framework that considers two of the main aspects in the study of meritocracy: perceptions and preferences.
+Furthermore, it distinguishes between meritocratic and non-meritocratic dimensions, as they are not necessarily two poles of the same continuum as previous studies seem to suggest.
+This conceptual framework is oriented to develop a measurement instrument as simple and brief as possible, being suitable to be used in public opinion surveys and thus allowing to integrate meritocratic beliefs in the study of different social phenomena.
+
+## The black-box of meritocratic beliefs {-#blackbox}
+
+In the following, we discuss four critical aspects in studies conceptualizing and measuring meritocracy, based upon which we develop a conceptual and measurement proposal.
+
+a. _Conceptual components_: Is merit only effort?
+
+One recent definition of meritocracy by @mijs_paradox_2019 is the following: "When I discuss meritocracy beliefs, I am referring to citizens' belief in the importance of hard work relative to structural factors." [@mijs_paradox_2019, pg.9].
+In the subsequent operationalization, this is associated with the following question and indicator: "how important you think it is for getting ahead in life: (a) hard work", scored from 1 to 5 on a likert scale.
+The assumptions behind such definition are worth discussing in light of the conceptual meaning of meritocracy and its possibilities of operationalization.
+
+The item used by @mijs_paradox_2019 is part of an items' battery present in several international surveys, usually called "reasons to get ahead".
+This battery displays a series of indicators related to what people consider important to get ahead in life: hard work, education, ambition, a wealthy family, the right connections, religion, race, and gender.
+Therefore, when considering only one of the items, it means that other aspects that could be associated with talent, such as education, would not be deemed as meritocratic.
+As he points out: "Hard work is arguably the most meritocratic part of Michael Young's equation: 'Merit = Intelligence + Effort', for the simple fact that intelligence itself is influenced by a non-meritocratic factor: who your parents happen to be" (p.5).
+
+In this conceptual and measurement approach of meritocratic beliefs, we can observe a couple of strong conceptual assumptions: a) effort would not depend on parental influence, and b) talent (as innate ability) is not meritocratic (contrary to Michael Youngs' original conceptualization).
+This conceptual and measurement-based assumption is found in other studies, which also assumes effort to be the main and only aspect of meritocracy [@girerd_neoliberalism_2020; @bubak_perceptions_2019], raising the question: Is effort the only dimension behind the concept of meritocracy?
+The question of whether talent (as intelligence and ability) is or is not considered meritocratic is certainly an interesting topic to discuss from a philosophical point of view, but for those working on empirical studies, it should face empirical scrutiny.
+Besides, the consideration of talent as part of meritocracy certainly opens some interesting avenues of research, as for instance some studies suggest that for the elites meritocracy is actually related to talent, whereas effort is something more characteristic of the meritocracy of the middle and lower classes [@atria_economic_2020].
+Therefore, we propose that a concept (and measure) of meritocracy should include both talent and effort as in its original conception.
+
+b. _Beliefs?_
+
+Several approaches to the empirical study of meritocracy based on public opinion surveys refer to the concept of _beliefs_, but with wide differences in meanings and operationalizations.
+To illustrate this point let us consider the commonly used "reasons to get ahead" battery, mentioned above.
+Some versions of this battery ask "how important you think it is for getting ahead in life" and then lists several factors, whereas another version of this same battery - sometimes presented along the previous one - asks about "how important you think it _should_ be ...", and then listing the same concepts.
+Therefore, the question raised here is: Which one of both is a "belief": what _is_ (the first one) or what it _should_ be (the second one)?
+
+The term belief has an ambiguous character in the literature, conceived as "idea-elements" by @converse_nature_1964 or "considerations" by @zaller_nature_1992.
+As @kluegel_beliefs_1986 pointed out about the scope of beliefs: "This usage encompasses such more specific social-psychological concepts as values, perceptions, and attitudes" (p.30).
+Therefore, beliefs use to cover almost anything related to subjective factors.
+To this regard, a relevant distinction in the field of inequality beliefs was made by @janmaat_subjective_2013: "Perceptions refer to subjective estimates of existing inequality (i.e.
+thoughts about what is).
+Beliefs are here defined as normative ideas about just inequality (i.e.
+thoughts about what should be)"(p.359).
+Several papers dealing with meritocracy use the term beliefs (i.e.
+what should be), while actually referring to perceptions (i.e.
+what is).
+This occurs for instance in @reynolds_perceptions_2014, in which the term belief is used to talk about what @janmaat_subjective_2013 refers to as perceptions, whereas other authors use general terms such as attitudes [@kunovich_systems_2007].
+The first attempt to shed light on this issue in meritocracy research was made by @duru-bellat_who_2012, who used the question "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" as a proxy for "desired" meritocracy (normative beliefs).
+They then determined "perceived" meritocracy, using the questions: "Would you say that in your country, people are rewarded for their efforts?" and "...
+people are rewarded for their skills?".
+
+Is the belief in meritocracy a perception or a preference with normative meaning?
+In order to expand the analytical conceptual framework, we believe that both dimensions should be included as proposed by @duru-bellat_who_2012.
+This opens up the possibility of analyzing whether perceptions and preferences are actually related (i.e.
+have a high correlation), or whether they are independent aspects of the same phenomenon.
+As @sonhing_merit_2011 has pointed out, "People can believe that outcomes ought to be distributed on the basis of merit and yet vary in their perceptions of whether this is how society currently operates" (p.
+435).
+In other words, normative beliefs should be considered while taking perceptions into account: a strong normative belief in meritocracy may mean something totally different to someone perceiving high meritocracy than to someone perceiving low meritocracy.
+To avoid the confusion generated by the term "belief", we propose the terms meritocratic preferences ("what should be"), and meritocratic perceptions ("what is"), as they better reflect the two facets of meritocracy under scrutiny [@castillo_meritocracia_2019].
+
+c. _Non-meritocratic aspects_
+
+Some research in meritocracy considers aspects usually opposed to effort and talent for personal achievement, as for instance family status and the use of personal advantages (as contacts or having a wealthy family) to get ahead in life.
+This distinction refers to the classical achieved and ascribed status dimension from @linton_study_1936.
+For instance, @kunovich_systems_2007 used an items' battery listing a number of factors in relation to "How important each should be in deciding pay..." (as @duru-bellat_who_2012 for desired meritocracy).
+They consider factors such as education and responsibility as meritocratic, giving them a value of 1 is considered "essential" in the scale response, whereas factors such as having a family or children are valued 1 when rated as "not important at all" (i.e.
+reverse coded).
+The assumption behind this approach is that rejecting a supposed ascribed or non-meritocratic^[Although the use of the concept "non-meritocratic" is very broad, we prefer to use it instead of something like ascription, given that this last one leaves out aspects such as the use of personal relationships.] aspect (as having family and/or children) implies a stronger belief in meritocracy.
+A similar approach of reverse-coding non-meritocratic items was taken by @newman_false_2015, using the same principles applied in the "Preference for the Merit Principle Scale" [@davey_preference_1999].
+
+The assumption that meritocratic and non-meritocratic elements are poles of the same continuum was tested by @reynolds_perceptions_2014 using the "get ahead" perceptions' battery items mentioned above.
+They considered education, ambition, and hard work as meritocratic, whereas factors such as family wealth and connections where classified as non-meritocratic.
+Despite making this distinction, the authors end up subtracting one dimension from the other, thus coming back to the assumption that they are two poles of the same continuum, as @kunovich_systems_2007 also did.
+Similarly, @roex_attitudes_2018 used ISSP indicators for perceived meritocracy and non-meritocracy to build a single score by reverse coding the non-meritocratic items.
+Therefore, the treatment of non-meritocratic items has been rather inconsistent across studies and the assumption that they are the simple opposite of meritocracy certainly requires further conceptual elaboration and empirical assessment.
+For instance, such distinction could open the door to study to what extent different and apparently contradictory views of meritocracy can coexist, to what extent individuals and societies differ to this regard, and of course the associated factors to such differences.
+
+d. _Accounting for measurement error_
+
+Finally, most studies in meritocracy so far have not properly considered the issue of latent structures and measurement error [@brown_confirmatory_2015; @bollen_structural_1989; @ansolabehere_strength_2008], as they mostly use single indicators and/or simple average indexes for measuring meritocracy.
+Such strategy assumes that the latent construct is measured perfectly (i.e.
+no error or residual variance) by the selected indicators, going as far as to propose that "...
+In choosing this strategy of index construction, we argue that support for meritocracy is not a latent variable" [@kunovich_systems_2007 p.653-654].
+Although some advances were made by Reynolds and Xian [-@reynolds_perceptions_2014] through conducting a principal component analysis of meritocratic and non-meritocratic dimensions, they finally chose to build a sum index despite proving a multidimensional latent structure.
+
+There is a relevant trade-off to consider when it comes to measurement error estimation in survey scales.
+Including multiple items per conceptual dimension opens the possibility to validity studies (at least in terms of construct) by using factor analysis techniques, particularly confirmatory ones.
+However, increasing the amount of items reduces the probability of usability of the scale in survey research given space limitations.
+This validity-usability conundrum is not easy to solve and, as we will show later in the methods section, we opted by a minimalistic approach oriented to the usability of the scale but that still gives the possibility of confirmatory factor analysis, although with some limitations.
+
+## A conceptual proposal for studying meritocracy {-#instrumentprop}
+
+Based on the previous assumptions and limitations identified in the empirical study of  meritocracy, we propose a conceptual and measurement framework with the following characteristics:
+
+- _Multidimensionality_, incorporating previous distinctions between preferences and perceptions, as well as between meritocratic and non-meritocratic aspects.
+
+- Multiple indicators for each dimension, in order to _account for measurement error_ in a confirmatory factor analysis context.
+
+- Based on _previous indicators_ as much as possible, for the sake of keeping comparability between studies.
+As most of the survey research on meritocracy so far has been made with the inequality module of the ISSP survey, we attempted to rescue as much as possible the original wording of some of its items in the new scale proposal.
+
+- _Brief_, as to be used in regular public opinion surveys [@davidov_measurement_2009].
+To this regard, it differs from the proposal of  "Preference for the Merit Principle Scale" [@davey_preference_1999], as they use 15 items for just one dimension (aside from the problem of reverse-coding non-meritocratic items).
+
+The proposed conceptual and measurement framework is depicted in Figure \@ref(fig:merit-model):
+
+```{r merit-model, echo=FALSE, fig.cap = "Conceptual Model of perception and preferences for meritocracy and non-meritocracy", fig.align='center'}
+ knitr::include_graphics("input/images/Fig1.generalf.png")
+```
+
+The columns "Perceptions" and "Preferences" represent the distinction between these two concepts, usually confused under the label "beliefs" [@castillo_meritocracia_2019].
+Perceptions refer to the extent to which people observe that meritocracy functions or apply in their society, which relates to items such as "Hard work is important to get ahead in society".
+Preferences refer to normative expectations that are usually linked to a "should" expression (e.g.
+whether hard work should be related to payment).
+The rows in Figure \@ref(fig:merit-model) consider the distinction between meritocratic and non-meritocratic dimensions [@reynolds_perceptions_2014], usually treated as different ends of the same continuum in previous research.
+Non-meritocratic elements refer to the use of resources such as personal contacts and/or family advantages to getting ahead in life.
+
+# Methodology
+
+```{r message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
+  #RUN THIS BEFORE THE ANALYSIS
+  pacman::p_load(knitr)
+  table_format = if(is_html_output()) {
+    "html"
+  } else if(is_latex_output()) {
+    "latex"
+  }
+
+  table_format2 = if(is_html_output()) {
+    T
+  } else if(is_latex_output()) {
+    F
+  }
+```
+
+The analysis is organized into three studies.
+Study 1 serves as a background to our instrument proposal by attempting to operationalize the meritocracy conceptual model with available secondary data from the inequality module of the International Social Survey Programme (ISSP).
+Although we are aware that the available indicators in this dataset do not allow a precise operationalization of the conceptual model, we know that it is the most widely used dataset in comparative meritocracy studies (as well as overall subjective inequality) and therefore it is certainly useful to look at its potentials and limitations to this regard.
+Study 2 is the core of this paper and consists of the proposal of a new scale for measuring meritocratic perceptions and preferences.
+Finally, in Study 3 we perform an additional validity analysis of the meritocracy scale.
+
+# Study 1: Measuring meritocracy with international secondary data
+
+## Data
+
+The data corresponds to the last available wave of the social inequality module from the International Social Survey Programme (ISSP), which is the most specialized international comparative survey in perceptions, attitudes, and beliefs about inequality-related issues.
+This wave corresponds to the year 2009 and covers attitudes towards a series of topics dealing with social inequality across 41 countries and 56,021 individuals.
+Although there is data available from this module for previous waves (1987, 1992, and 1999), unfortunately, there are several variables that are important for this research which were not included, particularly in 1987 and 1999, reason why we only use the 2009 wave (during the development of this research the 2019 wave became available, but so far with fewer countries than the previous one).
+Details about the sample of countries can be seen on Table \@ref(tab:countries).
+
+## Variables
+
+There are a series of indicators in the ISSP survey that in the following we attempt to classify in our meritocracy conceptual scheme, as presented below in Table \@ref(tab:table-issp):
+
+```{r table-issp, echo=FALSE}
+  if (!require("pacman")) install.packages("pacman")
+  # remotes::install_github('haozhu233/kableExtra@a6af5c0')  # for problems with collapse_rows()
+  pacman::p_load(knitr, kableExtra, dplyr)
+  table_format = if(knitr::is_html_output()) {
+    "html"
+  } else if(knitr::is_latex_output()) {
+    "latex"
+  }
+
+  tableissp <- read.csv(file = "input/tables/table-issp.csv",header = 1,sep = ";",encoding = "UTF-8") # call file generated externally for the table
+  cnames <- c("Component", "Dimensions","Item")
+  cap <- "Items of the ISSP meritocratic perceptions and preferences measures"
+
+  knitr::kable(tableissp, table_format, booktabs = T, linesep = "",col.names = cnames, caption = cap) %>%
+    kableExtra::kable_styling(
+      full_width = F,
+      latex_options = c("hold_position"),
+      position = "center",
+      font_size = 10,
+      bootstrap_options=c("striped", "bordered")) %>%
+    kableExtra::column_spec(column = 1, width = "2cm", ) %>%
+    kableExtra::column_spec(column = 2,width = "5cm") %>%
+    kableExtra::column_spec(column = 3,width = "8cm") %>%
+    kableExtra::collapse_rows(columns = 1:2,valign = "middle")
+```
+
+- _Perception of meritocracy/non-meritocracy_: for operationalizing perceptions the closest set of ISSP's indicators comes from the question asking about perceptions for opportunities to get ahead, which are usually considered as "meritocratic beliefs" in previous studies.
+The general heading of the battery is: _"To begin we have some questions about opportunities for getting ahead.
+Please tick one box for each of these to show how important you think it is for getting ahead in life."_ This is followed by a list of statements to be rated from 1 to 5: essential, very important, fairly important, not very important, not important at all.
+
+\break
+
+The classification of the items is based on criteria of internal motivation (meritocratic) and structural constraints (non-meritocratic).
+There were two items from the battery that were excluded from the analysis as they would not fit into the classification.
+The first one was "having good education yourself", since it was not clear whether this is could be due to individual motivation or system opportunities, and the second was "giving bribes", as introduced elements of criminality that were beyond a non-meritocratic perception.
+
+- _Preferences for meritocracy-non meritocracy_: for the operationalization of normative preferences we used a list of items related to reasons for pay battery.
+The ISSP question was: _In deciding how much people ought to earn, how important should each of these things be, in your opinion_, rated in the same _essential-non important at all_ scale (1 to 5) as the questions for meritocratic perceptions.
+
+## Methods
+
+The estimation of the measurement model was performed using Confirmatory Factor Analysis (CFA).
+CFA was conducted using the `lavaan` R package (version 0.6-3; @rosseel_lavaan_2012a), with diagonally weighted least squares (DWLS) estimation due to the items' ordinal level of measurement [@kline_principles_2016; @rosseel_lavaan_2012a].
+As recommended by @brown_confirmatory_2015, we assessed model fit by jointly considering the comparative fit index and Tucker-Lewis Index (CFI and TLI; acceptable fit > 0.95), Root of the average squared residual approximation (RMSEA; acceptable fit < 0.08),  Chi-square: (p-value; acceptable fit > 0.05, and Chi-square ratio > 3).
+
+## Results
+
+### Descriptive analyses
+
+Figure \@ref(fig:likert-issp) shows the distribution of responses across the selected items in their corresponding dimensions.
+On the one side, we see that there is a high degree of importance attributed to factors such as hard work and ambition in the process of getting ahead, concentrating 94.9% and 92.8% in the fairly important to essential categories.
+By contrast, the perception of non-meritocratic aspects is rated lower than the meritocratic ones, particularly for gender, race, and political connections.
+Regarding normative preferences, we observe that the meritocratic ones are deemed as important (from fairly to essential) for nearly the whole sample, support that decreases slightly for the non-meritocratic ones.
+Still, we have to consider that the preferences for non-meritocratic aspects, in this case, refer to the distributive principle of need rather than personal background (as in the perceptions of non-meritocracy), which certainly is one of the limitations of this set of items for the operationalization of the proposed conceptual framework with this dataset.
+
+```{r build-likert}
+  load(file = "input/data/proc/data_s1.RData")
+  if (!require("pacman")) install.packages("pacman")
+  pacman::p_load(dplyr,sjPlot, ggplot2, sjlabelled,sjmisc, ggpubr)
+  likert_issp<- data_s1 %>%
+    dplyr::select(hwork,ambition,wealthy,pareduc,
+                  race,gender,people,polcone,
+                  welljob,hardjob,family,child) %>% na.omit() #Subset variables
+  # Set labels for each variable
+  likert_issp <- sjlabelled::set_label(likert_issp,
+                                       label = c("Hard work",
+                                                 "Having ambition",
+                                                 "Wealthy family",
+                                                 "Educated parents",
+                                                 "Race",
+                                                 "Gender",
+                                                 "Knowing people",
+                                                 "Political connections",
+                                                 "Well job",
+                                                 "Hard job",
+                                                 "Support family",
+                                                 "Has children"))
+  #Reverse coding (ONLY) for plot
+  likert_issp <- sjmisc::rec(likert_issp, rec="rev", append=FALSE)
+  #Set value labels for plot
+  likert_issp <-
+    sjlabelled::set_labels(likert_issp,
+                           labels = c("Essential","Very important","Fairly important",
+                                      "Not very important","Not important at all"))
+
+  # Declare ggplot2 features
+  ggplot2::theme_set(ggplot2::theme(panel.background = ggplot2::element_rect(fill = "gray85",
+                                                                             colour = "gray85"),
+                                    panel.border = ggplot2::element_blank(),
+                                    axis.text.y = ggplot2::element_text(size = 13,
+                                                                        hjust = 1),
+                                    title = ggplot2::element_text(size = 13,
+                                                                  face = "bold"),
+                                    legend.text = ggplot2::element_text(size = 12),
+                                    plot.caption = ggplot2::element_text(size = 10,
+                                                                         face = "plain",
+                                                                         hjust = 1)))
+  #Plot 1: Perception meritocratic
+  p1<- likert_issp %>%
+    dplyr::select(hwork_r:ambition_r) %>%
+    sjPlot::plot_likert(geom.colors = "PuBu",
+                        grid.range  =  c (1.2 , 1.2),
+                        title = "a. Perception meritocratic",
+                          geom.size = c(0.62),
+                          catcount = 5,
+                          expand.grid = T,
+                          values  =  "sum.outside",
+                          reverse.colors = T,
+                          reverse.scale = F,
+                          show.n = FALSE) +
+    ggplot2::theme(legend.position = "none")
+  #Plot 2: Perception Non-meritocratic
+  p2<-likert_issp %>%
+    dplyr::select(wealthy_r:polcone_r) %>%
+    sjPlot::plot_likert(geom.colors = "PuBu",
+                        grid.range  =  c (1.2 , 1.2),
+                        title = "b. Perception Non-meritocratic",
+                          geom.size = c(0.9),
+                          catcount = 5,
+                          expand.grid = T,
+                          values  =  "sum.outside",
+                          reverse.colors = T,
+                          reverse.scale = F,
+                          show.n = FALSE)  +
+    ggplot2::theme(legend.position = "none")
+  #Plot 3: Preference meritocratic
+  p3<-likert_issp %>%
+    dplyr::select(welljob_r:hardjob_r) %>%
+    sjPlot::plot_likert(geom.colors = "PuBu",
+                        grid.range  =  c (1.2 , 1.2),
+                        title = "c. Preference meritocratic",
+                          geom.size = c(0.62),
+                          catcount = 5,
+                          expand.grid = T,
+                          values  =  "sum.outside",
+                          reverse.colors = T,
+                          reverse.scale = F,
+                          show.n = FALSE) +
+    ggplot2::theme(legend.position = "none")
+  #Plot 4: Preference Non-meritocratic
+  p4<-likert_issp %>%
+    dplyr::select(family_r:child_r) %>%
+    sjPlot::plot_likert(geom.colors = "PuBu",
+                        grid.range  =  c (1.2 , 1.2),
+                        title = "d. Preference Non-meritocratic",
+                          geom.size = c(0.62),
+                          catcount = 5,
+                          expand.grid = T,
+                          values  =  "sum.outside",
+                          reverse.colors = T,
+                          reverse.scale = F,
+                          show.n = FALSE) +
+    ggplot2::theme(legend.position = "none")
+
+  #Legend: Extract legend for the final plot using ggpubr
+  leg<- ggpubr::as_ggplot(
+    ggpubr::get_legend(
+      dplyr::select(likert_issp, family_r:child_r) %>%
+        sjPlot::plot_likert(geom.colors = "PuBu", catcount = 5) +
+        ggplot2::scale_fill_brewer(
+          breaks = c(5:1),
+          direction = -1,
+          labels = c("Not important at all","Not very important","Fairly important",
+                     "Very important","Essential")) +
+        ggplot2::theme(legend.position = "bottom")
+      )
+    )
+```
+
+```{r likert-issp,fig.height=11,fig.width=11,fig.cap="Distribution of responses in the ISSP meritocracy items", out.width='80%'}
+  # dev.copy(png,'output/images/Fig2.plotlikert-study1.png', width = 800, height = 600) # save image
+  if (!require("pacman")) install.packages("pacman")
+  pacman::p_load(cowplot, dplyr)
+  #Final plot
+  cowplot::add_sub(
+    cowplot::plot_grid(p1,p2,p3,p4,leg, #plot 1,2,3,4 and legend (bottom)
+      align = "hv",
+      axis = "a",
+      nrow = 5, #5 plots
+      ncol = 1, #1 column
+      rel_heights = c(1,2,1,1,0.25) #fig size within plot
+      ),
+    label = paste0(
+      "Source: Authors calculation based on International Social Survey Programme - Social Inequality 2009",
+      " (n=",dim(likert_issp)[1],")"
+      ),        # Caption: source and sample size
+    size = 10,  # Caption font size
+    hjust = 0.25# Caption align
+    ) %>%
+    cowplot::ggdraw()
+    # dev.off()
+```
+
+In terms of the association between the indicators, Figure \@ref(fig:corr-issp) shows the polychoric correlation matrix.
+Firstly, we observe in general that the moderate to high associations are between the pairs of items representing each of the dimensions as described in \@ref(tab:table-issp).
+The exception here is the dimension of non-meritocratic perception, in which there are six items that appear mostly associated by pairs according to their specific topics (family background, personal background, and connections).
+Still, between this same set of items, there are moderate correlations (around 0.3) which could anticipate a single latent factor underlying non-meritocratic perception and that is tested next through confirmatory factor analysis.
+
+```{r corr-issp,fig.cap = "Perceptions and preferences for ISSP  meritocracy items' polychoric correlations", fig.align='center',fig.height=6,fig.width=8}
+  # dev.copy(png,'output/images/Fig3.plotcor-study1.png', width = 800, height = 600) # save image
+  load(file = "input/data/proc/data_s1.RData")
+  if (!require("pacman")) install.packages("pacman")
+  pacman::p_load(dplyr, corrplot, lavaan)
+  # create polychoric correlation matrix
+  cor_issp<-
+    data_s1 %>%
+    dplyr::select(hwork,ambition,wealthy,pareduc,
+                  race,gender,people,polcone,
+                  welljob,hardjob,family,child) %>%
+    lavaan::lavCor(., ordered=names(.))
+  diag(cor_issp) = NA #set diagonal values to NA
+  # Set Row names of the matrix
+  rownames(cor_issp) <- c("A. Hard work",
+                         "B. Having ambition",
+                         "C. Wealthy family",
+                         "D. Educated parents",
+                         "E. Race",
+                         "F. Gender",
+                         "G. Knowing people",
+                         "H. Political connections",
+                         "I. Well job",
+                         "J. Hard job",
+                         "K. Support family",
+                         "L. Has children")
+  #set Column names of the matrix
+  colnames(cor_issp) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)",
+                         "(H)", "(I)","(J)","(K)","(L)")
+
+  #Plot the matrix using corrplot
+  corrplot::corrplot(cor_issp,
+    method = "color",
+    addCoef.col = "#000390",
+    type = "upper",
+    tl.col = "black",
+    col=colorRampPalette(c("white","#0068DC"))(12),
+    bg = "white",
+    na.label = "-")
+    # dev.off()
+```
+
+### Confirmatory Factor Analysis
+
+Table \@ref(tab:sum-fit-issp) shows the results of the estimation of two confirmatory models.
+The first one (First Order) corresponds to a model that estimate four factors, each for one of the dimensions presented in \@ref(tab:table-issp), with only regular fit indicators (CFI=0.959, TLI=0.944, RMSEA=0.098, $\chi^2$(df=48)= 21308.535).
+Further analysis showed that the sources for poor fit were mostly related to the items of the non-meritocratic perceptions, which as mentioned above displayed correlations among them not taken into account in this model specification.
+
+```{r sum-fit-issp}
+  load(file = "input/data/proc/data_s1.RData")
+  if (!require("pacman")) install.packages("pacman")
+  pacman::p_load(dplyr, kableExtra, knitr,lavaan)
+  table_format = if(knitr::is_html_output()) { #conditional instructions for kable
+    "html" #if html set "html" in format
+  } else if(knitr::is_latex_output()) {
+    "latex"#if latex set "latex" in format
+  }
+
+  # meassurement model: First order
+  cfa_perpref1 <- '
+  	  # latent variables
+  	  percmerit =~ hwork + ambition
+  	  percnmerit=~ wealthy + pareduc +race + gender +people + polcone
+  	  prefmerit =~ welljob + hardjob
+  	  prefnmerit=~ family + child
+  	  '
+  # Model fit: First order
+  fit_perpref1 <- lavaan::cfa(cfa_perpref1, data = data_s1,ordered = T)
+
+  # meassurement model: Second order
+  cfa_perpref2 <- '
+  	  # latent variables
+  	  percmerit =~ hwork + ambition
+  	  parents =~ wealthy + pareduc
+  	  backgrd =~ race + gender
+  	  networks=~ people + polcone
+  	  percnmerit=~ parents + backgrd + networks
+  	  prefmerit =~ welljob + hardjob
+  	  prefnmerit=~ family + child
+  	  '
+  # Model fit: second order
+  fit_perpref2 <- lavaan::cfa(cfa_perpref2, data = data_s1,ordered = T)
+
+  # extract fit indices from models and add to table
+  sum_fit_issp<- dplyr::bind_rows(
+    lavaan::fitmeasures(fit_perpref1)[c("chisq","df","cfi","tli","rmsea")],
+    lavaan::fitmeasures(fit_perpref2)[c("chisq","df","cfi","tli","rmsea")]
+    )
+
+  # Customize object
+  sum_fit_issp$mod <- c("First order","Second order")
+  sum_fit_issp$nobs <- c(lavaan::nobs(fit_perpref2),lavaan::nobs(fit_perpref2))
+  sum_fit_issp$est <- c("DWLS","DWLS")
+  sum_fit_issp <- dplyr::select(sum_fit_issp,mod,nobs,est,dplyr::everything())
+  colnames <- c("Model","$N$","Estimator","$\\chi^2$","df","CFI","TLI","RMSEA")
+
+  # Create table
+  table_issp_fits <-
+    knitr::kable(sum_fit_issp, format=table_format, digits=3, booktabs=T,
+                 col.names=colnames,
+                 caption = "Summary fit indices according to model",
+                 escape = FALSE) %>%
+    kableExtra::kable_styling(font_size = 10,
+                              full_width = F,
+                              latex_options = "HOLD_position",
+                              bootstrap_options=c("striped", "bordered"))
+  table_issp_fits
+```
+
+Attending to the sources of poor fit from the first estimated confirmatory model, we specified a second model that keeps the basic four-factor structure but generates an additional model for the non-meritocratic perceptions as depicted in Figure \@ref(fig:meas02-issp).
+In this model, the pairs of items in this dimension form three latent factors which at the same time generate a second-order factor of non-meritocratic perceptions, improving the fit of the model.
+
+```{r meas02-issp, echo=FALSE, fig.cap = "Descriptive plot", fig.align='center',fig.cap="Second-order confirmatory factor analysis model using ISSP indicators of Perceptions and Preferences for Meritocracy"}
+ knitr::include_graphics('output/images/Fig4.cfa_issp_study1.PNG')
+```
+
+Regarding the correlations between the factors in the Figure \@ref(fig:meas02-issp), we observe that perceptions are correlated with preferences, but more strongly for the meritocratic ($r=0.46$) than for the non-meritocratic ($r=0.27$) dimensions.
+Secondly, both perceptual dimensions (meritocratic and non-meritocratic) depict a moderate correlation ($r=0.30$), suggesting that they are not the opposite poles of the same continuum.
+Something similar occurs for the correlation between the two dimensions of preferences ($r=0.25$).
+
+Based on the measurement validation, we asses if the proposed measurement strategy can be used in cross-national comparisons.
+In this line, Figure \@ref(fig:scat1) depicts the correlation of factor scores for Meritocratic Perception and Meritocratic Preferences at the country level.
+For instance, Bulgaria (BGR) and the United States (USA) appropriately represent the positive association between perception and preferences.
+In other words, countries that describe merit in their society as an essential factor to get ahead in life also consider that effort and talent should be determinant factors in how much an individual earns.
+On the other hand, in countries with lower perceived meritocracy, preferences are also low, where Denmark (DNK) and Venezuela (VEN) are representative societies of this association.
+
+Another finding depicted by Figure \@ref(fig:scat1) is the positive association of Non-meritocratic perception with Meritocratic preferences, showing that countries like Bulgaria (BGR) and Philipines (PHL) represent cases with higher Non-meritocratic perceptions and higher preferences for meritocracy.
+In other words, in societies where people perceive higher importance of non-meritocratic aspects to get ahead in life, people also prefer that merit should be a determinant factor in wages.
+On the other side, countries like Sweden (SWE) or Denmark (DNK), both Scandinavian societies with strong welfare systems, represent contexts in which non-meritocratic aspects are not described as important, and merit is not considered as a factor that should be determinant in wages.
+
+```{r scat1, fig.cap="Cross-national comparison of non-meritocratic and meritocratic perceptions with meritocratic preference",fig.height=6, fig.width=12}
+  load(file = "input/data/proc/data_s1.RData")
+  if (!require("pacman")) install.packages("pacman")
+  pacman::p_load(dplyr, kableExtra, knitr,lavaan,countrycode,ggplot2,haven,
+                 gridExtra)
+  table_format = if(knitr::is_html_output()) { #conditional instructions for kable
+    "html" #if html set "html" in format
+  } else if(knitr::is_latex_output()) {
+    "latex"#if latex set "latex" in format
+  }
+
+  # meassurement model: Second order
+  cfa_perpref2 <- '
+  	  # latent variables
+  	  percmerit =~ hwork + ambition
+  	  parents =~ wealthy + pareduc
+  	  backgrd =~ race + gender
+  	  networks=~ people + polcone
+  	  percnmerit=~ parents + backgrd + networks
+  	  prefmerit =~ welljob + hardjob
+  	  prefnmerit=~ family + child
+  	  '
+  # Model fit: second order
+
+  fit_perpref2 <- lavaan::cfa(cfa_perpref2, data = data_s1,ordered = F)
+
+  df_proc<- data_s1 %>% select(id,v5,wealthy:hardjob)  
+  df_proc$ccode <- as.numeric(data_s1$v5)
+
+  predicted2 <- lavPredict(fit_perpref2,newdata = data_s1)
+
+  idx <- lavInspect(fit_perpref2, "case.idx")
+  fscores <- lavPredict(fit_perpref2)
+  ## loop over factors
+  for (fs in colnames(fscores)) {
+    df_proc[idx, fs] <- fscores[ , fs]
+  }
+
+  df_proc <-  df_proc %>%  select(id,ccode,percmerit,prefmerit,prefnmerit) %>% na.omit()
+
+  df_proc$clab <- countrycode(sourcevar = df_proc$ccode,
+                            origin = "iso3n",
+                            destination = "iso3c")
+
+avg_count <-  
+  df_proc %>% group_by(clab) %>% summarise(prefmerit=mean(prefmerit),
+                                            percmerit=mean(percmerit),
+                                            prefnmerit=mean(prefnmerit))
+
+# create figure 1: perception merit by preferences merit
+figure1 <- 
+ggplot(avg_count, aes(x=percmerit, y=prefmerit)) +
+  # geom_point() + 
+  geom_text(label=as.character(avg_count$clab)) +
+  geom_smooth(method = "lm") +
+  xlab("Meritocratic perception")+
+  ylab("Meritocratic preference")  
+# create figure 2: perception non-merit by preferences merit
+figure2 <- 
+ggplot(avg_count, aes(x=prefnmerit, y=prefmerit)) +
+  # geom_point() + 
+  geom_text(label=as.character(avg_count$clab)) +
+  geom_smooth(method = "lm") +
+  xlab("Non-meritocratic perception")+
+  ylab("Meritocratic preference")
+
+grid.arrange(figure1, figure2, ncol = 2,
+             bottom = grid::textGrob(label = paste0("Source: Authors calculation based on International Social Survey Programme - Social Inequality 2009",
+      " (n=",dim(df_proc)[1],")"),
+                                     hjust = 0.1,
+                                     gp = grid::gpar(fontface = 1, fontsize = 9)))
+```
+
+## Discussion study 1
+
+Although it was possible to get an adequate fit of the model to the pooled data from ISSP, there are several aspects to take into account when using these items for measuring meritocracy.
+First of all, these items were not specifically designed based on a conceptual approach to meritocracy, but still is possible to find a certain fit of the model to the data after minor adjustments.
+Secondly, some of the items in similar dimensions refer to different concepts, as for instance meritocratic perceptions are related to ambition and hard work, whereas meritocratic preferences deal with how well a job is done.
+A better operationalization would require more alignment between the different sides of the same concept.
+And thirdly, the scale is rather unbalanced given he larger amount of non-meritocratic perception items, producing correlated errors.
+All these limitations are taken into account next in the design of Study 2.
+
+With regard to the meritocratic / non-meritocratic distinction in previous studies, we can preliminary conclude that the wide use of reverse-coding for non-meritocratic items attempting to measure meritocracy is not an adequate operationalization of the construct.
+Meritocracy and non-meritocracy are related but different concepts, meaning that it is theoretically possible to find different combinations of both at individual and contextual levels.
+This finding opens a series of avenues for future research that are considered in the conclusions.
+
+# Study 2: The Perceptions and Preferences for Meritocracy Scale
+
+## Data
+
+The data was obtained through an online survey which was part of a larger study on meritocracy and preferences developed in Chile in 2019, funded by the Chilean National Scientific Agency (ANID).
+The questionnaire was programmed in Qualtrics and the fieldwork was conducted by an external online survey agency ([netquest.com](www.netquest.com)) between December 2019 and January 2020.
+The sample was selected from a non-probabilistic quota design in three large cities in Chile (Santiago, Concepción & Antofagasta).
+The quotas for gender, age, and educational levels were generated based on a survey by the Public Studies Center [@cep_encuesta_2019], which is a well-regarded counterpart agency of the ISSP (International Social Survey Programme) in Chile.
+A total sample of 2,141 people was collected, excluding those who did not sign the informed consent as well as those not answering the meritocracy response battery.
+There were no significant differences between our sample and the wider population for most socio-demographic characteristics, with the exception of educational level (see Table \@ref(tab:rep-samp) in Appendix).
+As it is often the case with online surveys, there were some limitations in achieving the quotas for lower educational levels [@zhang_quota_2018; @boas_recruiting_2020].
+This survey was designed as a three-wave panel and the data used in this analysis corresponds to the first wave.
+
+## Study design
+
+### Instrument
+
+The proposed scale of meritocratic perceptions and preferences consisted of eight indicators, two for each of the four dimensions listed earlier: Perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic).
+This design is oriented by two main criteria.
+The first one is usability in social surveys research, for which the length of the scale is restricted due to space reasons in such studies.
+Nevertheless, at the same time the scale should offer possibilities of accounting for measurement error through factor analysis techniques.
+Although a rule of thumb in latent  measurement suggests at least three indicators per dimension, we still opted for two in order to count with a brief scale.
+Although such decision posits some limitation for modeling measurement error, still allows for factor analysis and is a common approach in previous social survey studies with limited items (for instance see @davidov_measurement_2009).
+
+The second criteria is to expand the possibilities of comparability with previous studies on meritocracy, by considering some of the items previously used in this research area.
+To this regard, some of the items were adapted from the items battery "reasons to get ahead" (ISSP/GSS), which has been widely used for operationalizing meritocracy in international research [@mijs_paradox_2019; @duru-bellat_who_2012; @reynolds_perceptions_2014].
+The scale items ordered according to their respective dimensions are presented in Table \@ref(tab:table-indicadores).
+These eight likert-type items have five response alternatives, ranging from "Completely disagree"(1) to "Completely agree" (5).
+
+```{r table-indicadores, echo=FALSE}
+ if (!require("pacman")) install.packages("pacman")
+ # remotes::install_github('haozhu233/kableExtra@a6af5c0')  # for problems with collapse_rows()
+ pacman::p_load(knitr, kableExtra, dplyr)
+ table_format = if (knitr::is_html_output()) {
+   #conditional instructions for kable
+   "html" #if html set "html" in format
+ } else if (knitr::is_latex_output()) {
+   "latex"#if latex set "latex" in format
+ }
+ tabitems <- read.csv(file = "input/tables/table01.csv",header = 1,sep = ",",encoding = "UTF-8") # call file  generated externally for the table
+ cnames <- c("Component", "Dimensions","Item (English)" , "Item original (Spanish)")
+ cap <- "Items of the perceptions and preferences for meritocracy scale."
+
+ knitr::kable(tabitems, table_format, booktabs = T, linesep = "",col.names = cnames, caption = cap) %>%
+   kableExtra::kable_styling(
+     full_width = T,
+     latex_options = c("hold_position"),
+     position = "center",
+     font_size = 10,
+     bootstrap_options=c("striped", "bordered")) %>%
+   kableExtra::column_spec(column = 1, width = "1.5cm", ) %>%
+   kableExtra::column_spec(column = 2,width = "2 cm") %>%
+   kableExtra::collapse_rows(columns = 1:2,valign = "middle")
+```
+
+\pagebreak
+
+### Administration sets
+
+With the objective of evaluating the effect of indicator ordering in the responses, three different versions of items' order were designed and randomly assigned to the respondents, as depicted in Figure \@ref(fig:appmod).
+The scale was presented to the first group (_n = 712_) in the order that appears in Table \@ref(tab:table-indicadores) according to perceptions and preferences.
+For the second group (_n = 717_), the order was reorganized according to perceptions and preferences over the same topic, e.g.
+for the topic of hard work, the item about perception was followed by the item about preference and the same for the rest of the topics.
+Finally, for the third group (_n = 712_), the items were presented as completely randomized.
+
+```{r appmod, echo=FALSE, fig.cap = "Survey flow", fig.align='center'}
+ knitr::include_graphics('output/images/Fig5.app_mod.png')
+```
+
+## Methods
+
+To test the scale's underlying constructs, we employed confirmatory factor analysis models (CFA).
+The models estimated one factor for each of the four proposed dimensions presented in Table \@ref(tab:table-indicadores).
+As in Study 1, CFA was conducted using the `lavaan` R package (version 0.6-3; @rosseel_lavaan_2012a), with diagonally weighted least squares (DWLS) estimation due to the items' ordinal level of measurement [@kline_principles_2016; @rosseel_lavaan_2012a].
+The fit indexes and cut-off criteria were the same as the ones used in Study 1.
+
+A pre-registration was made in the OSF platform, available at the following link: [https://osf.io/z45y2](https://osf.io/z45y2).
+Included in this pre-registration are the hypotheses regarding the four-dimensional conceptual model underlying the scale, the variable measurement levels, the statistical tests to be performed with their respective evaluation parameters, and other important aspects of the research design.
+
+## Results
+
+### Descriptive analyses
+
+The graphs presented in Figure \@ref(fig:plotlikert) display disaggregated and comparable information of the different response categories for each item.
+As it can be observed, in general, there is more agreement in the perception of non-meritocratic items than in meritocratic ones, while in the case of preferences the opposite occurs.
+As far as preferences are concerned, the preponderant role of effort over talent as a criterion of meritocratic preference is noteworthy.
+All in all the descriptive results tend to show a sort of critical view of meritocracy, perceiving the operation of non-meritocratic aspects over meritocratic ones, whereas in the preferences the opposite occurs.
+
+```{r produce-likert_s2, include=FALSE}
+  if (!require("pacman")) install.packages("pacman")
+  pacman::p_load(dplyr,sjPlot, ggplot2, sjlabelled,sjmisc, ggpubr)
+  load(file = "input/data/proc/data_s2.RData")
+  names(data_s2)
+  dat_likert_s2 <- data_s2 %>% dplyr::select(!group) # exclude group variable
+  dat_likert_s2 <- sjmisc::rec(dat_likert_s2, rec="rev", append=FALSE)
+  names(dat_likert_s2) # reverse coding for likert graph
+
+  table(data_s2$perc_effort)
+  table(dat_likert_s2$perc_effort_r) # check ok.
+  # update.packages("sjPlot") # please install version 2.8.9 or above
+  ggplot2::theme_set(ggplot2::theme(panel.background = ggplot2::element_rect(fill = "gray85",
+                                                                             colour = "gray85"),
+                                    panel.border = ggplot2::element_blank(),
+                                    axis.text.y = ggplot2::element_text(size = 13,
+                                                                        hjust = 1),
+                                    title = ggplot2::element_text(size = 13,
+                                                                  face = "bold"),
+                                    legend.text = ggplot2::element_text(size = 12),
+                                    plot.caption = ggplot2::element_text(size = 10,
+                                                                         face = "plain",
+                                                                         hjust = 1)))
+
+  #Plot 1: Perception
+  p1<-
+    dplyr::select(dat_likert_s2,perc_effort_r:perc_netw_r) %>%
+    sjPlot::plot_likert(geom.colors = "PuBu",
+                        title = c("a. Perceptions"),
+                        geom.size = 0.8,
+                        axis.labels = c("Effort", "Talent", "Rich Parents", "Contacts"),
+                        catcount = 4,
+                        cat.neutral = 3,
+                        grid.range  =  c (1.2 , 1.2),
+                        values  =  "sum.outside",
+                        reverse.colors = T,
+                        reverse.scale = F,
+                        show.n = FALSE) +
+    ggplot2::theme(legend.position = "none")
+
+  #Plot 2: Preferences
+  p2<-
+    dplyr::select(dat_likert_s2,pref_effort_r:pref_netw_r) %>%
+    sjPlot::plot_likert(geom.colors = "PuBu",
+                        title = c("b. Preferences"),
+                        geom.size = 0.8,
+                        axis.labels = c("Effort", "Talent", "Rich Parents", "Contacts"),
+                        catcount = 4,
+                        cat.neutral = 3,
+                        grid.range  =  c (1.2 , 1.2),
+                        values  =  "sum.outside",
+                        reverse.colors = T,
+                        reverse.scale = F,
+                        show.n = FALSE)  +
+    ggplot2::theme(legend.position = "none")
+
+  #Legend: Extract legend for the final plot using ggpubr
+  leg<- ggpubr::as_ggplot(
+    ggpubr::get_legend(
+    dplyr::select(dat_likert_s2,pref_effort_r:pref_netw_r) %>%
+        sjPlot::plot_likert(geom.colors = "PuBu",
+                            catcount = 4,
+                            cat.neutral = 3) +
+      ggplot2::theme(legend.position = "bottom") +
+      ggplot2::scale_fill_manual(
+        values = c("#b3b3b3ff", "#f1eef6ff", "#bdc9e1ff","#74a9cfff","#0570b0ff"),
+        breaks=c("neutral",4,3,2,1),
+        labels=c("Neither agree nor disagree","Strongly disagree","Disagree",
+                 "Agree","Totally agree")
+        )
+      )
+    )
+```
+
+```{r plotlikert, echo=FALSE, fig.align='center', fig.cap="Distribution of responses in the Merit Scale items", fig.height=8, fig.width=11}
+  # dev.copy(png,'output/images/Fig6.plotlikert-study2.png', width = 820, height = 600)
+  if (!require("pacman")) install.packages("pacman")
+  pacman::p_load(cowplot)
+  #Final plot
+  cowplot::add_sub(
+    cowplot::plot_grid(p1,p2,leg, #plot 1,2 and legend (bottom)
+      align = "hv",
+      axis = "a",
+      nrow = 3, #2 plots
+      ncol = 1, #1 column
+      rel_heights = c(1,1,0.25) #fig size within plot
+      ),
+    label = paste0("Source: Authors calculations based on Study 2 sample"," (n=",dim(na.omit(dat_likert_s2))[1],")"
+      ),        # Caption: source and sample size
+    size = 12,  # Caption font size
+    hjust = -0.15# Caption align
+    ) %>%
+    cowplot::ggdraw()
+  # dev.off()
+```
+
+Attending now to the association among the scale items, Figure \@ref(fig:corpoly) shows the items' polychoric correlations.
+There are three main aspects to highlight from this correlation matrix.
+Firstly, and as expected,  the largest correlations are between indicators that correspond to the same factors behind the conceptual model (e.g., perception of meritocracy by effort and by talent, $r=0.52$).
+Secondly, among these correlations the highest are those between the non-meritocratic dimension but in perceptions ($r=0.73$) and preferences ($r=0.61$).
+Thirdly, both items for meritocratic preferences (E and F) are the ones that mostly correlate with the rest of the perceptual items, showing medium to high correlations.
+This is noteworthy because it would indicate that the perception of non-meritocracy would be related to larger meritocratic preferences.
+Finally, and similar to what was found in Study 1 with ISSP data, we observe that there are no considerable negative correlations between meritocratic and non-meritocratic aspects, undermining the assumptions of previous studies that suggested that these dimensions would be the opposite poles of one same continuum [@reynolds_perceptions_2014].
+
+```{r corpoly, echo=FALSE, fig.cap = "Perceptions and preferences for meritcoracy items' polychoric correlations", fig.align='center'}
+  # dev.copy(png,'output/images/Fig7.plotcor-study2.png',width = 800, height = 600) # save image
+  if (!require("pacman")) install.packages("pacman")
+  load(file = "input/data/proc/data_s2.RData")
+  pacman::p_load(dplyr, corrplot, lavaan)
+  # generate polychoric matrix
+  cor_s2<- data_s2 %>%
+    dplyr::select(perc_effort,perc_talent,perc_wpart,perc_netw,pref_effort,pref_talent,pref_wpart,pref_netw) %>%
+    lavaan::lavCor(., ordered=names(.))
+
+  diag(cor_s2) = NA
+
+  rownames(cor_s2) <-c(
+      "A. Perception Effort",
+      "B. Perception Talent",
+      "C. Perception Rich parents",
+      "D. Perception Contacts",
+      "E. Preferences Effort",
+      "F. Preferences Talent",
+      "G. Preferences Rich parents",
+      "H. Preferences Contacts")
+  colnames(cor_s2) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)", "(H)")
+
+  corrplot::corrplot(cor_s2,
+    method = "color",
+    addCoef.col = "#000390",
+    type = "upper",
+    tl.col = "black",
+    col=colorRampPalette(c("white","#0068DC"))(8),
+    bg = "white",
+    na.label = "-")
+  # dev.off()
+```
+
+### Confirmatory Factor Analysis
+
+The present section describes the results of the confirmatory factor analysis estimation.
+The model estimates four latent factors: perception meritocratic, perception non-meritocratic, preferences meritocratic, and preferences non-meritocratic.
+Each factor is estimated based on two items of the scale as detailed in Table \@ref(tab:table-indicadores).
+
+The first step in the analysis consists of comparing the model fit indicators for the three versions of the scale that were randomly assigned to the participants: order according to perceptions/preferences, order according to topics, and random order (see Figure \@ref(fig:appmod)).
+
+```{r model-generation-cfa}
+  # Libraries & data
+  if (!require("pacman")) install.packages("pacman")
+  pacman::p_load(lavaan)
+  load(file = "input/data/proc/data_s2.RData")
+  # names(data_s2)
+
+  # model
+  model_cfa <- '
+  perc_merit =~ perc_effort+perc_talent
+  perc_nmerit=~perc_wpart +perc_netw
+  pref_merit =~ pref_effort+pref_talent
+  pref_nmerit=~pref_wpart +pref_netw
+  '
+
+  # estimation for each order set
+  fit_cfa1 <- lavaan::cfa(model_cfa, data = data_s2[data_s2$group == 1,], ordered = TRUE, std.lv=FALSE)
+  fit_cfa2 <- lavaan::cfa(model_cfa, data = data_s2[data_s2$group == 2,], ordered = TRUE, std.lv=FALSE)
+  fit_cfa3 <- lavaan::cfa(model_cfa, data = data_s2[data_s2$group == 3,], ordered = TRUE, std.lv=FALSE)
+```
+
+```{r table-cfafits, echo=FALSE}
+  if (!require("pacman")) install.packages("pacman")
+  pacman::p_load(dplyr, kableExtra, knitr)
+  table_format = if(knitr::is_html_output()) { #conditional instructions for kable
+     "html" #if html set "html" in format
+   } else if(knitr::is_latex_output()) {
+     "latex"#if latex set "latex" in format
+   }
+  # extract fit indices from models and add to table
+  sum_fit<- dplyr::bind_rows(lavaan::fitmeasures(fit_cfa1)[c("chisq","df","cfi","tli","rmsea")],
+                             lavaan::fitmeasures(fit_cfa2)[c("chisq","df","cfi","tli","rmsea")],
+                             lavaan::fitmeasures(fit_cfa3)[c("chisq","df","cfi","tli","rmsea")])
+
+  # Customize object
+  sum_fit$mod <- c("Version 1","Version 2","Version 3")
+  sum_fit$nobs <- c(nobs(fit_cfa1),nobs(fit_cfa2),nobs(fit_cfa3))
+  sum_fit$est <- c("DWLS","DWLS","DWLS")
+  sum_fit <- dplyr::select(sum_fit,mod,nobs,est,everything())
+  colnames <- c("Model","$N$","Estimator","$\\chi^2$","df","CFI","TLI","RMSEA")
+
+  # Create table
+  table_cfa_fits <-kable(sum_fit, format=table_format,
+                         digits=3,
+                         booktabs=T,
+                         col.names=colnames,
+                         caption = "Summary fit indices according to order versions",
+                         escape = FALSE) %>%
+    kable_styling(full_width = F,
+                  font_size = 10,
+                  latex_options = "HOLD_position",
+                  bootstrap_options=c("striped", "bordered"))
+  table_cfa_fits
+```
+
+Table \@ref(tab:table-cfafits) shows the fit indicators of the models estimated for each of the three versions of the items' order described in the methodology.
+Regardless of the version, all models obtained adequate fit indicators, with CFI's above 0.95 and RMSEA's below 0.08.
+However, none of the models achieved a non-significant chi-square, something expected in large samples as the one used here.
+The first version order (perceptions-preferences) was the one obtaining best fit (CFI=0.993, TLI=0.995, RMSEA=0.034, $\chi2$(df=14)=42.276), whereas version 2 with the fixed order according to merit/non-merit items shows the comparatively worst indicators.
+The CFA fit indices for the completely randomized items' order (Model 3) it keeps all the indicators within the acceptable cut-off criteria and besides it controls for possible order effects in the administration of the instrument.
+The model and parameter estimates for this version are depicted in Figure \@ref(fig:meas02):
+
+```{r meas02, echo=FALSE, fig.cap = "Confirmatory factor analysis of the Perceptions and Preferences for Meritocracy Scale", fig.align='center'}
+ # estimation for diagram: summary(fit_cfa3, standardized = TRUE, fit.measures=TRUE)
+ knitr::include_graphics('output/images/Fig8.cfa_randomized_study2.png')
+```
+
+Attending to the correlations between the latent variables as depicted in Figure \@ref(fig:meas02), meritocratic preferences are moderate to highly correlated with perceptions, both meritocratic ($r=0.457$) and non-meritocratic ($r=0.500$).
+The correlation between both types of perceptions ($r=-0.044$) and both types preferences ($r=0.185$) are low, as well as between non-meritocratic preferences and perceptions ($r=-0.059$).
+This last finding gives further evidence regarding the lack of unidimensionality of meritocratic and non-meritocratic aspects as assumed by previous studies.
+
+## Discussion study 2
+
+The estimation performed with different administration sets for the meritocracy scale gives evidence in line with the conceptual model.
+The four dimensions appear as independent but related, although with some differences when compared with the estimation with ISSP data in Study 1.
+The main general difference is that with ISSP data the four latent dimensions were positively correlated, whereas now some correlations are statistically not different from zero.
+For instance, perceptions and preferences for meritocracy keep being related, which is not the case for the non-meritocratic pair.
+In principle this makes more sense in theoretical terms, as if someones perceives that meritocracy functions well, probably will support it more.
+On the other side, if someones perceives non-meritocracy, this should lead to a larger preference for meritocracy as it would constitute a demand for lager meritocracy when this one is not perceived to work.
+
+Despite the good general working of the scale, there are some limitations in the interpretation of the results given the few items (two) per dimension.
+As it was mentioned before, the option of a minimalistic approach in terms of items has advantages and disadvantages: advantages in terms of usability in social surveys with limited space, and disadvantages in terms of measurement validity in the estimation of the latent constructs.
+Being aware of this trade-off, in the following section we offer additional validity evidence as a way of complementing some of the weaknesses identified.
+
+# Study 3: Additional validity analyses
+
+The modeling for invariance testing has been gaining more relevance in sociological survey studies, mainly due to the interest in the validity of measurement scales in comparative studies in various social and cultural contexts [@davidov_measurement_2014].
+In this section we present the results of an analysis of invariance between groups, testing whether the proposed factorial structure for the meritocracy scale is applicable in a different dataset.
+
+The invariance measurement procedure consists of a series of nested models to which restrictions are progressively incorporated on the parameters of the measurement model.
+The literature generally suggests that this modeling should be done through four levels or types of progressive restriction [@milfont_testing_2010; @millsap_statistical_2011; @vandeschoot_checklist_2012]:
+
+1. **Configural:** the model is estimated only indicating the factorial structure used in the CFA.
+
+2. **Weak:** an equality restriction is applied to the factor loadings in the different groups, that is, the loadings are forced to be identical in both measurements.
+
+3. **Strong:** equality restrictions are added to the intercepts of each indicator.
+
+4. **Strict:** equality restrictions are added to the error variances of each indicator.
+
+### Data
+
+For testing the invariance we compared the data described previously in Study 2 to a different  dataset coming from an online survey carried out during the first half of 2020.
+The characteristics of this sample in terms of application and coverage are equivalent to those of the previous study.
+The final sample obtained contains 1,242 cases, where 605 (48.71%) are women and 637 (51.28%) are men.
+48.84% of them are 45 years of age or older, and 33.97% have tertiary education or higher.
+
+### Variables
+
+The items of the meritocracy scale were identical to the original and they  were administrated in a randomized order (which corresponds to the application modality of Group 3 in Study 2, as showed in Figure \@ref(fig:appmod)).
+
+### Results
+
+The first step for invariance testing is the estimation of the configural model, which serves as the baseline for further comparison and is expected to adequately meet the global fit criteria of a measurement model.
+Although the chi-square statistic is used as a global measure of fit, it is usually complemented with other indexes given its high sensitivity to sample size: the Comparative Fit Index (CFI), which should have a value greater than 0.95; the Root Mean Square Error Approximation (RMSEA), which must be lower than 0.06; and the Standardized Root Mean Square Residuals (SRMR) which must be less than 0.08 [@vandeschoot_checklist_2012].
+Besides, the literature on measurement invariance suggests some complementary approaches for the evaluation of the fit, from which we will consider the incremental adjustment of the fit indexes [@cheung_evaluating_2002; @milfont_testing_2010; @dimitrov_testing_2010] and the ANOVA test for means comparison in nested models [@newsom_longitudinal_2015].
+
+```{r invargroup, echo=FALSE}
+  if (!require("pacman")) install.packages("pacman")
+  #---Load packages and data -------------------------------------#
+  pacman::p_load(dplyr,lavaan,semTools,knitr,kableExtra,stats,gtools)
+  options(knitr.kable.NA = '')
+  load(file = "input/data/proc/data_s3_inv.RData")
+  pov01 <- data_s3_inv %>% dplyr::filter(dataset=="pvw01")
+  fs01  <- data_s3_inv %>% dplyr::filter(dataset=="fsw02")
+
+  #-----Meassurement model--------------------------------------#
+  model01 <- '
+  perc_merit  =~ perc_effort + perc_talent
+  perc_nmerit =~ perc_wpart  + perc_netw
+  pref_merit  =~ pref_effort + pref_talent
+  pref_nmerit =~ pref_wpart  + pref_netw'
+
+  #-----Invariance---------------------------------------------#
+  inv01<- semTools::measurementInvariance(model=model01,data=data_s3_inv,group="dataset",estimator = "ML",strict=TRUE,quiet = T)
+
+  conf  <- inv01$fit.configural
+  weak  <- inv01$fit.loadings
+  strong<- inv01$fit.intercepts
+  strict<- inv01$fit.residuals
+
+  tab01<- lavaan::anova(conf,weak,strong,strict,SB.classic=TRUE) %>%
+    dplyr::as_tibble() %>%
+    dplyr::select("Chisq","Df","chisq_diff"=`Chisq diff`,"df_diff"=`Df diff`,"pvalue"=`Pr(>Chisq)`) %>%
+    dplyr::mutate(stars=gtools::stars.pval(pvalue),
+           chisqt=paste0(round(Chisq,2)," (",Df,") "),
+           decision=ifelse(pvalue>0.05,yes = "Accept",no = "Reject"),
+           model=c("Configural","Weak","Strong","Strict"))
+
+  fit.meas<- dplyr::bind_rows(lavaan::fitmeasures(inv01$fit.configural,output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
+                              lavaan::fitmeasures(inv01$fit.loadings,  output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
+                              lavaan::fitmeasures(inv01$fit.intercepts,output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
+                              lavaan::fitmeasures(inv01$fit.residuals, output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),])
+
+  # compute differences in chisq, df, cfi and rmsea (90%, lower.ci - upper.ci )
+  fit.meas<- fit.meas %>%
+    dplyr::mutate(diff.chi2 = chisq    - lag(chisq,default = dplyr::first(chisq)),
+           diff.df   = df       - lag(df,   default = dplyr::first(df)),
+           diff.cfi  = cfi      - lag(cfi,  default = dplyr::first(cfi)),
+           diff.rmsea   = rmsea - lag(rmsea,default = dplyr::first(rmsea))) %>%
+    round(3) %>%
+    dplyr::mutate(rmsea.ci=paste0(rmsea," \n ", "(",rmsea.ci.lower,"-",rmsea.ci.upper,")"))
+
+  tab.inv<- dplyr::bind_cols(tab01,fit.meas) %>%
+    dplyr::select(model,chisqt,cfi,rmsea.ci,diff.chi2,diff.df,diff.cfi,diff.rmsea,stars,decision) %>%
+    dplyr::mutate(diff.chi2=paste0(diff.chi2," (",diff.df,") ",stars)) %>%
+    dplyr::select(model,chisqt,cfi,rmsea.ci,diff.chi2,diff.cfi,diff.rmsea,decision)
+
+  #clean values
+  tab.inv[tab.inv == c("0 (0) ")] <- NA
+  tab.inv[tab.inv == c(0)] <- NA
+
+  ### Check an alternative method
+  configural        <- lavaan::cfa(model01, data=data_s3_inv, group = "dataset", estimator="ML")
+  weak.invariance   <- lavaan::cfa(model01, data=data_s3_inv, group = "dataset",group.equal = "loadings", estimator="ML")
+  strong.invariance <- lavaan::cfa(model01, data=data_s3_inv, group = "dataset",group.equal = c( "loadings", "intercepts"), estimator="ML")
+  strict.invariance <- lavaan::cfa(model01, data=data_s3_inv, group = "dataset",group.equal = c( "loadings", "intercepts", "residuals"), estimator="ML")
+
+  table_format = if(is_html_output()) {
+    "html"
+  } else if(is_latex_output()) {
+    "latex"
+  }
+
+  col.nam <- c("Model","$\\chi^2 (\\text{df})$","CFI","RMSEA (90 CI)",
+               "$\\Delta \\chi^2 (\\Delta \\text{df}$)","$\\Delta \\text{CFI}$","$\\Delta \\text{RMSEA}$","Decision")
+  footnote <- paste0("N = ","; Group 1, n = ",conf@Data@nobs[[1]],"; Group 2, n = ",conf@Data@nobs[[2]], ", ***p < 0.001")
+
+  knitr::kable(tab.inv, col.names = col.nam, align = "l",
+        booktabs=TRUE,format = table_format,escape = FALSE,
+        caption = "Multiple Group meassurement invariance for Perceptions and Preferences for Meritocracy") %>%
+    kableExtra::kable_styling(full_width = FALSE,
+                              latex_options = "HOLD_position",
+                              bootstrap_options=c("striped", "bordered"),
+                              font_size = 10) %>%
+    kableExtra::footnote(general = footnote, footnote_as_chunk = T)
+```
+
+Table \@ref(tab:invargroup) shows the results of the measurement invariance estimation.
+When attending to the traditional invariance test of $\Delta \chi^2 (\Delta \text{df})$, the results support the invariance at the strong level meaning that the fit of the factor model of the merit scale is equivalent across samples when constraining factor loadings and intercepts to being equal.
+Such result is considered in general as evidence of invariance [@fischer_are_2011], as strict forms of measurement invariance rarely hold [@vandeschoot_checklist_2012].
+Still, the comparability of latent means requires strict invariance which in this case does not hold when considering just $\Delta \chi^2 (\Delta \text{df})$.
+Nevertheless, the criteria of $\Delta \text{CFI}$ used for comparing models is close to the rejection criteria of >.01, whereas the $\Delta \text{RMSEA}$ fulfills the requirements of being below the cut-off criteria as suggested by [@chen_sensitivity_2007].
+Therefore, using this last standard, the level for strict invariance would hold for the meritocracy scale.
+
+## Discusion Study 3
+
+Based on the measurement invariance analysis for two datasets, we found evidence supporting the equivalence of the scale when applied in different samples.
+This gives a more robust ground when using this measure for comparing groups, as for instance country comparison.
+Although the results are promising, still the analysis was performed in two samples in one country (Chile) and it requires further examination with social surveys in other societies.
+
+# Conclusions
+
+Studies that attempt to characterize and compare societies by their support for meritocratic beliefs have used different approaches.
+As most studies use secondary survey data, they tend to assume that the available indicators represent an underlying meritocratic construct.
+A review of these studies reveals several non-tested assumptions, as well as the use of similar indicators to represent different constructs and dimensions of meritocracy.
+As the existence of heterogeneous approaches certainly has consequences for the advancement of the study of meritocracy, this paper presented a comprehensive conceptual framework for the empirical study of meritocracy, building upon previous research.
+This framework was operacionalized in a survey scale and tested against available and new survey data.
+
+We identified four critical aspects regarding the measurement of meritocracy in previous studies: unidimensionality, the ambiguous use of the term "beliefs", the use of non-meritocratic indicators as opposed to meritocratic indicators, and the consideration of measurement error.
+The proposed 8-item scale, "Perceptions and Preferences for Meritocracy", was designed and tested in order to deal with these four issues.
+Our results indicate that perceptions and preferences seem to be two related but different dimensions, often confounded in previous research under the label of "beliefs".
+At the same time, meritocratic and non-meritocratic dimensions do not appear to constitute poles of the same continuum, as some previous studies have assumed.
+Regarding the possible effects of the items' order in the estimation of the latent variable, we tested three different order versions and found evidence that suggests the use of the randomized version of the scale.
+Furthermore, we found evidence of measurement invariance that approaches well the requirement of comparability between samples.
+
+One of the criteria in designing the merit scale was to develop a short instrument that could be used in social surveys, assuming that using only two items per dimension posits limitations for the confirmatory analysis estimated here.
+We preferred this confirmatory approach at it offers several advantages in terms of information about the quality of the measurement, and we offered additional validity evidence with measurement invariance analysis.
+Despite the advantages, a more simple and straightforward perspective is always available, as for instance using only the average of the items per dimension.
+Whereas a short scale it is always appreciated in social survey studies, for other audiences could not reach measurement standards.
+This may require further expansions and/or consider using a different instrument as the "Preference for the Merit Principle Scale" [@davey_preference_1999], which measures only one dimension of meritocracy with fifteen items.
+
+The four-dimensional conceptual framework and its operationalization in the Perceptions and Preferences for Meritocracy scale opens several avenues for future research.
+For instance, distinguishing perceptions from preferences will allow us to evaluate the extent to which different societies are accustomed to, or satisfied with, the perceived level of meritocracy, in terms of differences between what is perceived and what is preferred.
+Additionally, given that non-meritocratic factors are not necessarily related to meritocratic ones, our framework makes it possible to assess the perceived legitimacy of practices such as the use of personal contacts and their interference (or not) with meritocratic ideals in different societies.
+The consideration of structural factors, both individual and societal, will allow us to advance future hypotheses relating to meritocratic legitimacy to a more comprehensive level than the current one, which is based solely on meritocratic perceptions.
+Furthermore, the impact of different configurations of the four-dimensional framework on practices and behaviors such as corruption, civic involvement, and political alignment, is an area that requires additional research.
+Such future agendas could be especially relevant in times of economic crisis and growing inequalities, that could entail changes in the legitimation of the current distributive structure based on meritocratic ideals.
+
+# Transparency statement
+
+This research follows a series of open science guidelines as the pre-registration of Study 2 (scale development) as well as the availability of data and codes (R) in an public repository in Github.
+This document was generated with the R package Rmarkdown and contains the code for all the tables and figures in order to make it reproducible.
+
+# References
+
+<div id="refs"> </div>  <!-- Although <div> is an HTML tag, this method also works for other output formats such as PDF. -->
+
+\pagebreak
+
+<!-- # Appendix {-} -->
+
+`r if (knitr::is_latex_output()){ '\\appendix'}`
+
+`r if (knitr::is_latex_output()){ '\\section{Appendix}'}`
+
+**Study 1**
+
+```{r countries, echo=FALSE}
+  load(file = "input/data/proc/data_s1.RData")
+  if (!require("pacman")) install.packages("pacman")
+  pacman::p_load(dplyr, kableExtra, knitr,lavaan,countrycode,ggplot2,haven,
+                 gridExtra)
+  table_format = if(knitr::is_html_output()) { #conditional instructions for kable
+    "html" #if html set "html" in format
+  } else if(knitr::is_latex_output()) {
+    "latex"#if latex set "latex" in format
+  }
+  load(file = "input/data/proc/data_s1.RData")
+  if (!require("pacman")) install.packages("pacman")
+  pacman::p_load(dplyr, kableExtra, knitr,lavaan,countrycode,ggplot2,haven,
+                 gridExtra)
+  table_format = if(knitr::is_html_output()) { #conditional instructions for kable
+    "html" #if html set "html" in format
+  } else if(knitr::is_latex_output()) {
+    "latex"#if latex set "latex" in format
+  }
+ # meassurement model: Second order
+  cfa_perpref2 <- '
+  	  # latent variables
+  	  percmerit =~ hwork + ambition
+  	  parents =~ wealthy + pareduc
+  	  backgrd =~ race + gender
+  	  networks=~ people + polcone
+  	  percnmerit=~ parents + backgrd + networks
+  	  prefmerit =~ welljob + hardjob
+  	  prefnmerit=~ family + child
+  	  '
+  # Model fit: second order
+
+  fit_perpref2 <- lavaan::cfa(cfa_perpref2, data = data_s1,ordered = F)
+
+  df_proc<- data_s1 %>% select(id,v5,wealthy:hardjob)  
+  df_proc$ccode <- as.numeric(data_s1$v5)
+
+  predicted2 <- lavPredict(fit_perpref2,newdata = data_s1)
+
+  idx <- lavInspect(fit_perpref2, "case.idx")
+  fscores <- lavPredict(fit_perpref2)
+  ## loop over factors
+  for (fs in colnames(fscores)) {
+    df_proc[idx, fs] <- fscores[ , fs]
+  }
+
+  df_proc <-  df_proc %>%  select(id,ccode,percmerit,prefmerit,prefnmerit) %>% na.omit()
+
+  df_proc$clab <- countrycode(sourcevar = df_proc$ccode,
+                            origin = "iso3n",
+                            destination = "country.name")
+avg_count <-  
+  df_proc %>% group_by(clab) %>% summarise(N=n())
+avg_count <- arrange(avg_count)
+
+# sum(avg_count[, 'N'])
+  col.nam <- c("Country","N", "Country", "N")
+  footnote <- paste0("N = ", sum(avg_count[, 'N']),". Source: International Social Survey Programme (2009).")
+
+  knitr::kable(bind_cols(avg_count[1:20,],avg_count[21:40,]),
+               col.names = col.nam, align = c("l","l","l","l"),linesep = "",
+        booktabs=TRUE,format = table_format,escape = FALSE,
+        caption = "Sample of Study 1 ") %>%
+    kableExtra::kable_styling(full_width = FALSE,
+                              latex_options = "HOLD_position",
+                              bootstrap_options=c("striped", "bordered"),
+                              font_size = 10) %>%
+    kableExtra::column_spec(column = c(2,4),width = "1cm") %>% 
+    kableExtra::column_spec(column = c(1,3),width = "2.5cm") %>% 
+    column_spec(3, border_left = T) %>% 
+    kableExtra::footnote(general = footnote, footnote_as_chunk = T,threeparttable = T)
+```
+
+\pagebreak
+
+**Study 2**
+
+|                     | Sample  |   CEP  |
+|---------------------|:-------:|:------:|
+| __Gender__          |         |        |
+| Men                 | 49,82%  | 50,52% |
+| Women               | 50.18%  | 49,47% |
+| __Age__             |         |        |
+| 18 - 24             | 18,55%  | 18,17% |
+| 25 - 34             | 18,86%  | 17,48% |
+| 35 - 44             | 19.09%  | 19,98% |
+| 45 - 54             | 17,96%  | 19,23% |
+| 55 - or more        | 25,54%  | 25.11% |
+| __Education__       |         |        |
+| Primary  or less    | 2,93%   | 15,88% |
+| High school         | 43,23%  | 37,04% |
+| Non university      | 32,63%  | 28,93% |
+| university or more  | 21,21%  | 18,13% |
+Table: `r as.character(paste("(\\#tab:rep-samp)", "Representativeness of the study 2 sample"))`.
+
+```{r tb-perpref1, eval=FALSE, include=FALSE}
+ if (!require("pacman")) install.packages("pacman")
+ load(file = "input/data/proc/data_s1.RData")
+ pacman::p_load(dplyr,knitr,kableExtra)
+ options(knitr.kable.NA = '')
+ table_format = if(knitr::is_html_output()) { #conditional instructions for kable
+   "html" #if html set "html" in format
+ } else if(knitr::is_latex_output()) {
+   "latex"#if latex set "latex" in format
+ }
+
+ # First order model for Appendix
+ cfa_perpref1 <- '
+ 	  # latent variables
+ 	  percmerit =~ hwork + ambition
+ 	  percnmerit=~ wealthy + pareduc +race + gender +people + polcone
+ 	  prefmerit =~ welljob + hardjob
+ 	  prefnmerit=~ family + child
+ 	  '
+ # Fit First order model for Appendix
+ fit_perpref1 <- lavaan::cfa(cfa_perpref1, data = data_s1,
+                             ordered = c("wealthy","pareduc","ambition","hwork",
+                                         "people","polcone","race","gender",
+                                         "respons","yeduc",
+                                         "family","child",
+                                         "welljob","hardjob"))
+ # Labels of First order model for Appendix
+ labs <- c(
+   "Importance: hard work",
+   "Importance: having ambition",
+   "Importance: wealthy family",
+   "Importance: educated parents",
+   "Importance: race",
+   "Importance: gender",
+   "Importance: knowing people",
+   "Importance: political connections",
+   "Reasons for pay: well job",
+   "Reasons for pay: hard job",
+   "Reasons for pay: support family",
+   "Reasons for pay: has children"
+ )
+ #Extract factor loadings
+ tb.load<- data.frame(round(cbind(lavaan::inspect(fit_perpref1,
+                                                         what="std")$lambda),
+                            digits = 3))
+ tb.load[tb.load==c(0.00)] <- NA #set to NA factor loadings of 0
+
+ for (i in names(tb.load)) {
+   tb.load[,i] <- sprintf(tb.load[,i], fmt = '%#.3f') #set values as character
+ }
+ tb.load[tb.load=="NA"] <- "" #set NA values to "" (empty space)
+ #---extract fit measures-------#
+ fm01<- data.frame(t(data.frame(lavaan::fitmeasures(fit_perpref1, output ="matrix")[c("chisq","df","cfi","tli","rmsea"),])));  row.names(fm01) ="percnmerit"
+
+ #------chi2, df------#
+ fm04<- round(rbind(fm01),3) #select and round to 4 digits
+ fm04.1 <- fm04 %>% dplyr::select(chisq,df) #select chi2 and degrees of freedom
+ fm04.1$chisq <- round(x = fm04.1$chisq,digits = 1) # round Chi2 to 1 digit
+ fm04.1$df <- round(x = fm04.1$df,digits = 0) #round Df to 0 digits
+ fm04.1$chi2df <- paste0(fm04.1$chisq,"(",fm04.1$df,")") #Paste chi2(df)
+ fm04.1 <- dplyr::select(fm04.1,"chi2df") #select final column
+ for (i in names(fm04.1)) {
+   fm04.1[,i] <- as.character(fm04.1[,i]) # transform all to character
+ }
+
+ #------CFI, RMSEA------#
+ fm04.2 <- fm04 %>% dplyr::select(cfi,tli,rmsea) # select CFI, TLI and RMSEA
+ for (i in names(fm04.2)) {                      # transform all to character
+   fm04.2[,i] <- sprintf(fm04.2[,i], fmt = '%#.3f')
+ }
+
+ fm.df      <- dplyr::bind_cols(fm04.1,fm04.2) # factor loading and labels
+ fm.df$nobs <- c(lavaan::nobs(fit_perpref1))   # add number of observations
+ fm.df <- data.frame(t(fm.df)); colnames(fm.df) <- c("percmerit") #transpose and set colnames
+
+ #------ merge ------#
+ tb.fm<- dplyr::bind_rows(tb.load,fm.df) #merge factor loadings and fit indices
+ tb.fm<- tb.fm %>%
+   dplyr::mutate(vars=c(labs,"$\\chi^2\\text{(df)}$","$\\text{CFI}$",
+                        "$\\text{TLI}$","$\\text{RMSEA}$","$N$")) %>%
+   dplyr::select(vars,everything()) #Change names of Fit meassures to latex symbols
+ tb.perpref1 <- tb.fm #final matrix for Table
+
+ # Table --------------------------------------------------------------------.
+ tb.foot <- paste0("Standardized factor loadings using DWLS estimator ; CFI = Comparative fit index ;TLI = Tucker-Lewis index; RMSEA =  Root mean square error of approximation")
+ tb.col <- c("","Meritocratic","Non-meritocratic",
+             "Meritocratic","Non-meritocratic")
+ tb.caption <- c(" Factor loadings and fit measures of the four first-order factor model in study 1")
+ tb.issp<- kable(tb.perpref1,escape = FALSE,digits = 3,
+                 align = "lcccc",
+                 col.names = tb.col,
+                 caption = tb.caption,
+                 booktabs = T,
+                 linesep = "",
+                 format = table_format,
+                 row.names=F) %>%
+           kable_styling(latex_options = c("hold_position"),
+                         bootstrap_options = c("striped", "bordered"),
+                         position = "center",
+                         font_size = 10) %>%
+           add_header_above(header = c(" "=1,"Perception"= 2,"Preference"= 2)) %>%
+           add_header_above(header = c(" "=1,"Factor loadings"= 4)) %>%
+           row_spec(row = 12,hline_after = TRUE) %>%
+           add_indent(c(13:17)) %>%
+           footnote(general =tb.foot ,footnote_as_chunk = T,threeparttable = T);tb.issp
+```
+
+<!-- **Study 3** -->
+
+```{r fcloads-inv, eval=FALSE, include=FALSE}
+ if (!require("pacman")) install.packages("pacman")
+ pacman::p_load(dplyr, knitr, kableExtra,lavaan)
+ options(knitr.kable.NA = '')
+ table_format = if(knitr::is_html_output()) { #conditional instructions for kable
+   "html" #if html set "html" in format
+ } else if(knitr::is_latex_output()) {
+   "latex"#if latex set "latex" in format
+ }
+ table_format2 = if(knitr::is_html_output()) {
+   T
+ } else if(knitr::is_latex_output()) {
+   F
+ }
+ # Note:
+ # This CFA corresponds to the sample from the Factorial Survey Study
+ load(file = "input/data/proc/data_s3_inv.RData") #load data study 3
+ fs01  <- data_s3_inv %>% dplyr::filter(dataset=="fsw02") #subset sample
+
+ # Set model
+ model01 <- '
+ perc_merit  =~ perc_effort + perc_talent
+ perc_nmerit =~ perc_wpart  + perc_netw
+ pref_merit  =~ pref_effort + pref_talent
+ pref_nmerit =~ pref_wpart  + pref_netw'
+
+ # Model fit
+ fs1_o <- lavaan::cfa(model = model01,
+              data = fs01,
+              ordered = c("perc_effort","perc_talent",
+                          "perc_wpart","perc_netw",
+                          "pref_effort","pref_talent",
+                          "pref_wpart","pref_netw"),std.lv=FALSE)
+
+ # Extract information for table
+ fm01<- data.frame(t(data.frame(lavaan::fitmeasures(fs1_o, output ="matrix")[c("chisq","df","cfi","tli","rmsea"),])));  row.names(fm01) ="M1.1"
+
+ #------chi2, df------#
+ fm04.1 <- fm01 %>% dplyr::select(chisq,df)
+ fm04.1$chisq <- round(x = fm04.1$chisq,digits = 1)
+ fm04.1$df <- round(x = fm04.1$df,digits = 0)
+ fm04.1$chi2df <- paste0(fm04.1$chisq,"(",fm04.1$df,")")
+ fm04.1 <- dplyr::select(fm04.1,"chi2df")
+
+ #------CFI, RMSEA------#
+ fm04.2 <- fm01 %>% dplyr::select(cfi,tli,rmsea)
+ for (i in names(fm04.2)) {
+   fm04.2[,i] <- sprintf(fm04.2[,i], fmt = '%#.3f')
+ }
+
+ #------ merge fit indices ------#
+ fm.df      <- dplyr::bind_cols(fm04.1,fm04.2)
+ fm.df$nobs <- c(lavaan::nobs(fs1_o))
+ fm.df      <- data.frame(t(fm.df))
+
+ #------ extract factor loadings ------#
+ loads01<- data.frame(round(inspect(fs1_o,what="std")$lambda,digits = 3))
+ loads01[loads01==c(0.00)] <- NA; names(loads01) <- c(paste0("M1.",1:4))
+
+ for (i in names(loads01)) {
+   loads01[,i] <- sprintf(loads01[,i], fmt = '%#.3f')
+ }
+ loads01[loads01=="NA"] <- ""
+
+ #------ labels for table ------#
+ labs <-c(
+     "A. Perception Effort",
+     "B. Perception Talent",
+     "C. Perception Rich parents",
+     "D. Perception Contacts",
+     "E. Preferences Effort",
+     "F. Preferences Talent",
+     "G. Preferences Rich parents",
+     "H. Preferences Contacts")
+
+ #------ merge ------#
+ tb.fm2<- dplyr::bind_rows(loads01,fm.df)
+ tb.fm2<- tb.fm2 %>%
+   dplyr::mutate(vars=c(labs,"$\\chi^2\\text{(df)}$","$\\text{CFI}$","$\\text{TLI}$","$\\text{RMSEA}$","$N$")) %>%
+   dplyr::select(vars,everything())
+
+ tb.col  <- c(" ",
+              "Meritocratic","Non-meritocratic",
+              "Meritocratic","Non-meritocratic")
+ tb.caption <- "Factor loadings and fit measures for the second sample for study 3"
+ tb.foot <- paste0("Standardized factor loadings using DWLS estimator ; CFI = Comparative fit index ;TLI = Tucker-Lewis index; RMSEA =  Root mean square error of approximation")
+ tb.fm02<- knitr::kable(tb.fm2,escape = FALSE,
+                        align = "lcccc",col.names = tb.col,
+                        caption = tb.caption,
+                        booktabs = T,linesep = "",
+                        format = table_format,
+                        row.names = F) %>%
+           kableExtra::kable_styling(full_width = table_format2,
+                         latex_options = c("hold_position"),
+                         position = "center",
+                         font_size = 10,
+                         bootstrap_options = c("striped", "bordered")) %>%
+           kableExtra::add_header_above(header = c(" "=1,"Perception"= 2,"Preference"= 2)) %>%
+           kableExtra::add_header_above(header = c(" "=1,"Factor loadings"= 4)) %>%
+           kableExtra::row_spec(row = 8,hline_after = TRUE) %>%
+           kableExtra::add_indent(c(9:13)) %>%
+           kableExtra::footnote(general =tb.foot ,footnote_as_chunk = T,threeparttable = T);tb.fm02
+```

### 2023-02-01 paper-blinded.Rmd
grammarly

@@
 ---
+title: "Measur---
 title: "Measuring Perceptions and Preferences for Meritocracy"
@@
 Despite its wide use as a concept, empirical research on meritocracy is relatively novel and characterized by diverse conceptualizations and measures that make the findings and their interpretation rather inconsistent.
-Most of the studies in the area have relied upon secondary data from different surveys to operationalize meritocracy, with a large variation in the use and interpretation of the same measuring items.
+Most of the studies in the area have relied upon secondary data from different surveys to operationalize meritocracy, with a wide variation in the use and interpretation of the same measuring items.
 Based on a critical analysis of concepts and operationalizations of meritocracy in survey research, we propose a conceptual framework for measuring perceptions and preferences for meritocracy and non-meritocracy in a minimalistic version (few items as possible), in order to be considered in comparative survey research studies.
@@
 ```
-
-ing Perceptions and Preferences for Meritocracy"
-css: "custom.css"
-linestretch: '1.5'
-link-citations: yes
-# author: "Juan Carlos Castillo, Julio Iturra, Luis Maldonado, Francisco Meneses & Jorge Atria <br><br>"
-abstract: "Economic and social inequalities have generated growing concerns and crises across contemporary societies.
-The belief in meritocracy has been suggested as one of the factors that explain the persistence of inequality, as it would legitimize economic disparities based on differences in effort and talent.
-Despite its wide use as a concept, empirical research on meritocracy is relatively novel and characterized by diverse conceptualizations and measures that make the findings and their interpretation rather inconsistent.
-Most of the studies in the area have relied upon secondary data from different surveys to operationalize meritocracy, with a wide variation in the use and interpretation of the same measuring items.
-Based on a critical analysis of concepts and operationalizations of meritocracy in survey research, we propose a conceptual framework for measuring perceptions and preferences for meritocracy and non-meritocracy in a minimalistic version (few items as possible), in order to be considered in comparative survey research studies.
-The proposal is first tested through confirmatory factor analysis (CFA) using available ISSP (International Social Survey Programme) inequality module items.
-Based on these results, in a second stage we developed a improved scale that better suits the perceptions/preferences distinction, which was applied in a Chilean sample (N=2,141).
-The results of the CFA estimation give evidence of a better fit for the new scale.
-Our discussion highlights the importance of considering different dimensions in order to advance in the study of meritocracy."
-output:
-  bookdown::pdf_document2:
-     template: null
-     toc: false
-     number_sections: false
-  bookdown::word_document2:
-    number_sections: false
-  bookdown::html_document2:
-    number_sections: false
-linkcolor: blue
-bibliography:
-  - input/bib/meritocracy.bib
-csl: input/bib/apa7-singlespace.csl
-editor_options:
-  chunk_output_type: console
-geometry: margin=0.78in
-header-includes:
-  - \usepackage{times}           # Times New Roman
-  - \usepackage{caption}
-  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
-  - \usepackage[figurename=Fig.]{caption}
-  - \usepackage{graphicx}
-  - \usepackage{float}
-  - \usepackage{booktabs}
-  - \usepackage{longtable}
-  - \usepackage{array}
-  - \usepackage{multirow}
-  - \usepackage{wrapfig}
-  - \usepackage{float}
-  - \usepackage{colortbl}
-  - \usepackage{pdflscape}
-  - \usepackage{tabu}
-  - \usepackage{threeparttable}
-always_allow_html: true # for word output
-fontsize: 10pt
----
-
-```{r eval=FALSE, include=FALSE}
- # for render in pdf run rmarkdown::render_site("docs/paper.Rmd", output_format = "all")
- # clean #in the yml
- rmarkdown::render("paper-blinded.Rmd", output_format = "bookdown::pdf_document2")
- rmarkdown::render("paper-blinded.Rmd", output_format = "bookdown::html_document2")
- # rmarkdown::render("paper.Rmd", output_format = "bookdown::word_document2")
-```
-
-```{r setup, include=FALSE}
- knitr::opts_chunk$set(echo=FALSE, warning = FALSE,message = FALSE, cache = FALSE,out.width = '85%',fig.pos= "H", fig.align = 'center')
- # knitr::opts_knit$set(base.url = "../") #relative path for .html output file
- # knitr::opts_knit$set(root.dir = "../") #relative path for chunks within .rmd files
- options(scipen=999)
- options(kableExtra.auto_format = FALSE)
- rm(list=ls())
- options(knitr.kable.NA = '')
- options(knitr.graphics.error = FALSE)
- Sys.setlocale("LC_ALL", "ES_ES.UTF-8")
-```
-
-<!---
-\pagebreak
-
-# Acknowledgments
-
-The authors recognize the support of ANID/FONDECYT grant 1160921 and of  the Center for Conflict and Social Cohesion Studies (COES)  grant ANID/FONDAP 1513009.
-
-\pagebreak
-
---->
-
-# Introduction
-
-Economic inequality and income concentration have become topics of growing concern over the last years, leading to a series of social movements in different societies as well as diverse critical analyses regarding the development of capitalism and its consequences [@streeck_politics_2014; @piketty_capital_2014].
-In such a context, the study of views, preferences, and perceptions of inequality has acquired relevance in the social sciences, in topics such as redistributive preferences [@alesina_fairness_2005; @dimick_models_2018], the legitimization of economic inequality [@schroder_income_2017] and the functioning of meritocracy [@duru-bellat_who_2012; @mijs_paradox_2019; @reynolds_perceptions_2014; @atria_economic_2020].
-Within this research area, and particularly in times of economic and health crises, the ideal of meritocracy has been strongly challenged as an unfulfilled promise of modern societies that allows the perpetuation of social inequalities [@sandel_tyranny_2020; @goldthorpe_myth_2003].
-
-In general, meritocracy has been defined as a system of distribution of resources and rewards based on individual merit, which in its original conception is a combination of talent and effort [@young_rise_1962].
-This traditional conception of merit places in a secondary position the possible interference of structural or non-meritocratic factors, such as inheritance, personal contacts, and luck [@breen_class_1999; @saunders_might_1995; @yair_meritocracy_2007; @land_we_2006; @young_rise_1962].
-Social psychology and sociology have studied the characteristics and consequences of beliefs in meritocracy, under the general hypothesis that a greater belief in meritocracy emphasizes the role of the individual over structural factors in personal achievements, leading to greater legitimization of inequalities [@preminger_meritocracy_2020; @trump_when_2020; @hadjar_meritokratie_2008; @madeira_primes_2019].
-Such research has raised criticism of meritocracy as a moral standard of distribution given the preponderant weight of non-meritocratic elements upon the individual status and social mobility [@sandel_tyranny_2020; @witteveen_reconsidering_2020; @arrow_meritocracy_2000; @goldthorpe_myth_2003; @markovits_meritocracy_2019; @khan_privilege_2013].
-
-Due to the role that meritocratic beliefs play in the justification of individual achievement (or failure) in contemporary societies [@hadjar_meritokratie_2008; @markovits_meritocracy_2019; @sandel_tyranny_2020], multiple studies have evaluated the relationship between meritocratic beliefs and personal and/or contextual characteristics.
-For example, some studies have linked meritocracy to the reinforcement of socio-economic, gender, and ethnic stereotypes [@madeira_primes_2019; @girerd_neoliberalism_2020; @girerd_neoliberalism_2020], as well as the effects of meritocratic beliefs in educational [@generett_stories_2020; @owens_engines_2020] and organizational contexts [@perez_advancing_2020; @aiello_new_2019].
-Most of such studies so far have used indicators from existing standard social surveys, such as the International Social Survey Programme (ISSP), to measure meritocracy.
-However, as we will show later, the concepts used as well as the instruments for measuring meritocracy vary extensively among studies.
-In many cases, similar phenomena are associated with different indicators, and conversely, often different phenomena are measured with similar indicators.
-Such trends not only limit the comparability of studies, but also the ability to understanding the effects of meritocratic beliefs across different disciplines and lines of research.
-
-Based on a critical analysis of different approaches to the measurement of meritocracy to date, this article proposes a conceptual and measurement framework that considers two of the main aspects in the study of meritocracy: perceptions and preferences.
-Furthermore, it distinguishes between meritocratic and non-meritocratic dimensions, as they are not necessarily two poles of the same continuum as previous studies seem to suggest.
-This conceptual framework is oriented to develop a measurement instrument as simple and brief as possible, being suitable to be used in public opinion surveys and thus allowing to integrate meritocratic beliefs in the study of different social phenomena.
-
-## The black-box of meritocratic beliefs {-#blackbox}
-
-In the following, we discuss four critical aspects in studies conceptualizing and measuring meritocracy, based upon which we develop a conceptual and measurement proposal.
-
-a. _Conceptual components_: Is merit only effort?
-
-One recent definition of meritocracy by @mijs_paradox_2019 is the following: "When I discuss meritocracy beliefs, I am referring to citizens' belief in the importance of hard work relative to structural factors." [@mijs_paradox_2019, pg.9].
-In the subsequent operationalization, this is associated with the following question and indicator: "how important you think it is for getting ahead in life: (a) hard work", scored from 1 to 5 on a likert scale.
-The assumptions behind such definition are worth discussing in light of the conceptual meaning of meritocracy and its possibilities of operationalization.
-
-The item used by @mijs_paradox_2019 is part of an items' battery present in several international surveys, usually called "reasons to get ahead".
-This battery displays a series of indicators related to what people consider important to get ahead in life: hard work, education, ambition, a wealthy family, the right connections, religion, race, and gender.
-Therefore, when considering only one of the items, it means that other aspects that could be associated with talent, such as education, would not be deemed as meritocratic.
-As he points out: "Hard work is arguably the most meritocratic part of Michael Young's equation: 'Merit = Intelligence + Effort', for the simple fact that intelligence itself is influenced by a non-meritocratic factor: who your parents happen to be" (p.5).
-
-In this conceptual and measurement approach of meritocratic beliefs, we can observe a couple of strong conceptual assumptions: a) effort would not depend on parental influence, and b) talent (as innate ability) is not meritocratic (contrary to Michael Youngs' original conceptualization).
-This conceptual and measurement-based assumption is found in other studies, which also assumes effort to be the main and only aspect of meritocracy [@girerd_neoliberalism_2020; @bubak_perceptions_2019], raising the question: Is effort the only dimension behind the concept of meritocracy?
-The question of whether talent (as intelligence and ability) is or is not considered meritocratic is certainly an interesting topic to discuss from a philosophical point of view, but for those working on empirical studies, it should face empirical scrutiny.
-Besides, the consideration of talent as part of meritocracy certainly opens some interesting avenues of research, as for instance some studies suggest that for the elites meritocracy is actually related to talent, whereas effort is something more characteristic of the meritocracy of the middle and lower classes [@atria_economic_2020].
-Therefore, we propose that a concept (and measure) of meritocracy should include both talent and effort as in its original conception.
-
-b. _Beliefs?_
-
-Several approaches to the empirical study of meritocracy based on public opinion surveys refer to the concept of _beliefs_, but with wide differences in meanings and operationalizations.
-To illustrate this point let us consider the commonly used "reasons to get ahead" battery, mentioned above.
-Some versions of this battery ask "how important you think it is for getting ahead in life" and then lists several factors, whereas another version of this same battery - sometimes presented along the previous one - asks about "how important you think it _should_ be ...", and then listing the same concepts.
-Therefore, the question raised here is: Which one of both is a "belief": what _is_ (the first one) or what it _should_ be (the second one)?
-
-The term belief has an ambiguous character in the literature, conceived as "idea-elements" by @converse_nature_1964 or "considerations" by @zaller_nature_1992.
-As @kluegel_beliefs_1986 pointed out about the scope of beliefs: "This usage encompasses such more specific social-psychological concepts as values, perceptions, and attitudes" (p.30).
-Therefore, beliefs use to cover almost anything related to subjective factors.
-To this regard, a relevant distinction in the field of inequality beliefs was made by @janmaat_subjective_2013: "Perceptions refer to subjective estimates of existing inequality (i.e.
-thoughts about what is).
-Beliefs are here defined as normative ideas about just inequality (i.e.
-thoughts about what should be)"(p.359).
-Several papers dealing with meritocracy use the term beliefs (i.e.
-what should be), while actually referring to perceptions (i.e.
-what is).
-This occurs for instance in @reynolds_perceptions_2014, in which the term belief is used to talk about what @janmaat_subjective_2013 refers to as perceptions, whereas other authors use general terms such as attitudes [@kunovich_systems_2007].
-The first attempt to shed light on this issue in meritocracy research was made by @duru-bellat_who_2012, who used the question "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" as a proxy for "desired" meritocracy (normative beliefs).
-They then determined "perceived" meritocracy, using the questions: "Would you say that in your country, people are rewarded for their efforts?" and "...
-people are rewarded for their skills?".
-
-Is the belief in meritocracy a perception or a preference with normative meaning?
-In order to expand the analytical conceptual framework, we believe that both dimensions should be included as proposed by @duru-bellat_who_2012.
-This opens up the possibility of analyzing whether perceptions and preferences are actually related (i.e.
-have a high correlation), or whether they are independent aspects of the same phenomenon.
-As @sonhing_merit_2011 has pointed out, "People can believe that outcomes ought to be distributed on the basis of merit and yet vary in their perceptions of whether this is how society currently operates" (p.
-435).
-In other words, normative beliefs should be considered while taking perceptions into account: a strong normative belief in meritocracy may mean something totally different to someone perceiving high meritocracy than to someone perceiving low meritocracy.
-To avoid the confusion generated by the term "belief", we propose the terms meritocratic preferences ("what should be"), and meritocratic perceptions ("what is"), as they better reflect the two facets of meritocracy under scrutiny [@castillo_meritocracia_2019].
-
-c. _Non-meritocratic aspects_
-
-Some research in meritocracy considers aspects usually opposed to effort and talent for personal achievement, as for instance family status and the use of personal advantages (as contacts or having a wealthy family) to get ahead in life.
-This distinction refers to the classical achieved and ascribed status dimension from @linton_study_1936.
-For instance, @kunovich_systems_2007 used an items' battery listing a number of factors in relation to "How important each should be in deciding pay..." (as @duru-bellat_who_2012 for desired meritocracy).
-They consider factors such as education and responsibility as meritocratic, giving them a value of 1 is considered "essential" in the scale response, whereas factors such as having a family or children are valued 1 when rated as "not important at all" (i.e.
-reverse coded).
-The assumption behind this approach is that rejecting a supposed ascribed or non-meritocratic^[Although the use of the concept "non-meritocratic" is very broad, we prefer to use it instead of something like ascription, given that this last one leaves out aspects such as the use of personal relationships.] aspect (as having family and/or children) implies a stronger belief in meritocracy.
-A similar approach of reverse-coding non-meritocratic items was taken by @newman_false_2015, using the same principles applied in the "Preference for the Merit Principle Scale" [@davey_preference_1999].
-
-The assumption that meritocratic and non-meritocratic elements are poles of the same continuum was tested by @reynolds_perceptions_2014 using the "get ahead" perceptions' battery items mentioned above.
-They considered education, ambition, and hard work as meritocratic, whereas factors such as family wealth and connections where classified as non-meritocratic.
-Despite making this distinction, the authors end up subtracting one dimension from the other, thus coming back to the assumption that they are two poles of the same continuum, as @kunovich_systems_2007 also did.
-Similarly, @roex_attitudes_2018 used ISSP indicators for perceived meritocracy and non-meritocracy to build a single score by reverse coding the non-meritocratic items.
-Therefore, the treatment of non-meritocratic items has been rather inconsistent across studies and the assumption that they are the simple opposite of meritocracy certainly requires further conceptual elaboration and empirical assessment.
-For instance, such distinction could open the door to study to what extent different and apparently contradictory views of meritocracy can coexist, to what extent individuals and societies differ to this regard, and of course the associated factors to such differences.
-
-d. _Accounting for measurement error_
-
-Finally, most studies in meritocracy so far have not properly considered the issue of latent structures and measurement error [@brown_confirmatory_2015; @bollen_structural_1989; @ansolabehere_strength_2008], as they mostly use single indicators and/or simple average indexes for measuring meritocracy.
-Such strategy assumes that the latent construct is measured perfectly (i.e.
-no error or residual variance) by the selected indicators, going as far as to propose that "...
-In choosing this strategy of index construction, we argue that support for meritocracy is not a latent variable" [@kunovich_systems_2007 p.653-654].
-Although some advances were made by Reynolds and Xian [-@reynolds_perceptions_2014] through conducting a principal component analysis of meritocratic and non-meritocratic dimensions, they finally chose to build a sum index despite proving a multidimensional latent structure.
-
-There is a relevant trade-off to consider when it comes to measurement error estimation in survey scales.
-Including multiple items per conceptual dimension opens the possibility to validity studies (at least in terms of construct) by using factor analysis techniques, particularly confirmatory ones.
-However, increasing the amount of items reduces the probability of usability of the scale in survey research given space limitations.
-This validity-usability conundrum is not easy to solve and, as we will show later in the methods section, we opted by a minimalistic approach oriented to the usability of the scale but that still gives the possibility of confirmatory factor analysis, although with some limitations.
-
-## A conceptual proposal for studying meritocracy {-#instrumentprop}
-
-Based on the previous assumptions and limitations identified in the empirical study of  meritocracy, we propose a conceptual and measurement framework with the following characteristics:
-
-- _Multidimensionality_, incorporating previous distinctions between preferences and perceptions, as well as between meritocratic and non-meritocratic aspects.
-
-- Multiple indicators for each dimension, in order to _account for measurement error_ in a confirmatory factor analysis context.
-
-- Based on _previous indicators_ as much as possible, for the sake of keeping comparability between studies.
-As most of the survey research on meritocracy so far has been made with the inequality module of the ISSP survey, we attempted to rescue as much as possible the original wording of some of its items in the new scale proposal.
-
-- _Brief_, as to be used in regular public opinion surveys [@davidov_measurement_2009].
-To this regard, it differs from the proposal of  "Preference for the Merit Principle Scale" [@davey_preference_1999], as they use 15 items for just one dimension (aside from the problem of reverse-coding non-meritocratic items).
-
-The proposed conceptual and measurement framework is depicted in Figure \@ref(fig:merit-model):
-
-```{r merit-model, echo=FALSE, fig.cap = "Conceptual Model of perception and preferences for meritocracy and non-meritocracy", fig.align='center'}
- knitr::include_graphics("input/images/Fig1.generalf.png")
-```
-
-The columns "Perceptions" and "Preferences" represent the distinction between these two concepts, usually confused under the label "beliefs" [@castillo_meritocracia_2019].
-Perceptions refer to the extent to which people observe that meritocracy functions or apply in their society, which relates to items such as "Hard work is important to get ahead in society".
-Preferences refer to normative expectations that are usually linked to a "should" expression (e.g.
-whether hard work should be related to payment).
-The rows in Figure \@ref(fig:merit-model) consider the distinction between meritocratic and non-meritocratic dimensions [@reynolds_perceptions_2014], usually treated as different ends of the same continuum in previous research.
-Non-meritocratic elements refer to the use of resources such as personal contacts and/or family advantages to getting ahead in life.
-
-# Methodology
-
-```{r message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
-  #RUN THIS BEFORE THE ANALYSIS
-  pacman::p_load(knitr)
-  table_format = if(is_html_output()) {
-    "html"
-  } else if(is_latex_output()) {
-    "latex"
-  }
-
-  table_format2 = if(is_html_output()) {
-    T
-  } else if(is_latex_output()) {
-    F
-  }
-```
-
-The analysis is organized into three studies.
-Study 1 serves as a background to our instrument proposal by attempting to operationalize the meritocracy conceptual model with available secondary data from the inequality module of the International Social Survey Programme (ISSP).
-Although we are aware that the available indicators in this dataset do not allow a precise operationalization of the conceptual model, we know that it is the most widely used dataset in comparative meritocracy studies (as well as overall subjective inequality) and therefore it is certainly useful to look at its potentials and limitations to this regard.
-Study 2 is the core of this paper and consists of the proposal of a new scale for measuring meritocratic perceptions and preferences.
-Finally, in Study 3 we perform an additional validity analysis of the meritocracy scale.
-
-# Study 1: Measuring meritocracy with international secondary data
-
-## Data
-
-The data corresponds to the last available wave of the social inequality module from the International Social Survey Programme (ISSP), which is the most specialized international comparative survey in perceptions, attitudes, and beliefs about inequality-related issues.
-This wave corresponds to the year 2009 and covers attitudes towards a series of topics dealing with social inequality across 41 countries and 56,021 individuals.
-Although there is data available from this module for previous waves (1987, 1992, and 1999), unfortunately, there are several variables that are important for this research which were not included, particularly in 1987 and 1999, reason why we only use the 2009 wave (during the development of this research the 2019 wave became available, but so far with fewer countries than the previous one).
-Details about the sample of countries can be seen on Table \@ref(tab:countries).
-
-## Variables
-
-There are a series of indicators in the ISSP survey that in the following we attempt to classify in our meritocracy conceptual scheme, as presented below in Table \@ref(tab:table-issp):
-
-```{r table-issp, echo=FALSE}
-  if (!require("pacman")) install.packages("pacman")
-  # remotes::install_github('haozhu233/kableExtra@a6af5c0')  # for problems with collapse_rows()
-  pacman::p_load(knitr, kableExtra, dplyr)
-  table_format = if(knitr::is_html_output()) {
-    "html"
-  } else if(knitr::is_latex_output()) {
-    "latex"
-  }
-
-  tableissp <- read.csv(file = "input/tables/table-issp.csv",header = 1,sep = ";",encoding = "UTF-8") # call file generated externally for the table
-  cnames <- c("Component", "Dimensions","Item")
-  cap <- "Items of the ISSP meritocratic perceptions and preferences measures"
-
-  knitr::kable(tableissp, table_format, booktabs = T, linesep = "",col.names = cnames, caption = cap) %>%
-    kableExtra::kable_styling(
-      full_width = F,
-      latex_options = c("hold_position"),
-      position = "center",
-      font_size = 10,
-      bootstrap_options=c("striped", "bordered")) %>%
-    kableExtra::column_spec(column = 1, width = "2cm", ) %>%
-    kableExtra::column_spec(column = 2,width = "5cm") %>%
-    kableExtra::column_spec(column = 3,width = "8cm") %>%
-    kableExtra::collapse_rows(columns = 1:2,valign = "middle")
-```
-
-- _Perception of meritocracy/non-meritocracy_: for operationalizing perceptions the closest set of ISSP's indicators comes from the question asking about perceptions for opportunities to get ahead, which are usually considered as "meritocratic beliefs" in previous studies.
-The general heading of the battery is: _"To begin we have some questions about opportunities for getting ahead.
-Please tick one box for each of these to show how important you think it is for getting ahead in life."_ This is followed by a list of statements to be rated from 1 to 5: essential, very important, fairly important, not very important, not important at all.
-
-\break
-
-The classification of the items is based on criteria of internal motivation (meritocratic) and structural constraints (non-meritocratic).
-There were two items from the battery that were excluded from the analysis as they would not fit into the classification.
-The first one was "having good education yourself", since it was not clear whether this is could be due to individual motivation or system opportunities, and the second was "giving bribes", as introduced elements of criminality that were beyond a non-meritocratic perception.
-
-- _Preferences for meritocracy-non meritocracy_: for the operationalization of normative preferences we used a list of items related to reasons for pay battery.
-The ISSP question was: _In deciding how much people ought to earn, how important should each of these things be, in your opinion_, rated in the same _essential-non important at all_ scale (1 to 5) as the questions for meritocratic perceptions.
-
-## Methods
-
-The estimation of the measurement model was performed using Confirmatory Factor Analysis (CFA).
-CFA was conducted using the `lavaan` R package (version 0.6-3; @rosseel_lavaan_2012a), with diagonally weighted least squares (DWLS) estimation due to the items' ordinal level of measurement [@kline_principles_2016; @rosseel_lavaan_2012a].
-As recommended by @brown_confirmatory_2015, we assessed model fit by jointly considering the comparative fit index and Tucker-Lewis Index (CFI and TLI; acceptable fit > 0.95), Root of the average squared residual approximation (RMSEA; acceptable fit < 0.08),  Chi-square: (p-value; acceptable fit > 0.05, and Chi-square ratio > 3).
-
-## Results
-
-### Descriptive analyses
-
-Figure \@ref(fig:likert-issp) shows the distribution of responses across the selected items in their corresponding dimensions.
-On the one side, we see that there is a high degree of importance attributed to factors such as hard work and ambition in the process of getting ahead, concentrating 94.9% and 92.8% in the fairly important to essential categories.
-By contrast, the perception of non-meritocratic aspects is rated lower than the meritocratic ones, particularly for gender, race, and political connections.
-Regarding normative preferences, we observe that the meritocratic ones are deemed as important (from fairly to essential) for nearly the whole sample, support that decreases slightly for the non-meritocratic ones.
-Still, we have to consider that the preferences for non-meritocratic aspects, in this case, refer to the distributive principle of need rather than personal background (as in the perceptions of non-meritocracy), which certainly is one of the limitations of this set of items for the operationalization of the proposed conceptual framework with this dataset.
-
-```{r build-likert}
-  load(file = "input/data/proc/data_s1.RData")
-  if (!require("pacman")) install.packages("pacman")
-  pacman::p_load(dplyr,sjPlot, ggplot2, sjlabelled,sjmisc, ggpubr)
-  likert_issp<- data_s1 %>%
-    dplyr::select(hwork,ambition,wealthy,pareduc,
-                  race,gender,people,polcone,
-                  welljob,hardjob,family,child) %>% na.omit() #Subset variables
-  # Set labels for each variable
-  likert_issp <- sjlabelled::set_label(likert_issp,
-                                       label = c("Hard work",
-                                                 "Having ambition",
-                                                 "Wealthy family",
-                                                 "Educated parents",
-                                                 "Race",
-                                                 "Gender",
-                                                 "Knowing people",
-                                                 "Political connections",
-                                                 "Well job",
-                                                 "Hard job",
-                                                 "Support family",
-                                                 "Has children"))
-  #Reverse coding (ONLY) for plot
-  likert_issp <- sjmisc::rec(likert_issp, rec="rev", append=FALSE)
-  #Set value labels for plot
-  likert_issp <-
-    sjlabelled::set_labels(likert_issp,
-                           labels = c("Essential","Very important","Fairly important",
-                                      "Not very important","Not important at all"))
-
-  # Declare ggplot2 features
-  ggplot2::theme_set(ggplot2::theme(panel.background = ggplot2::element_rect(fill = "gray85",
-                                                                             colour = "gray85"),
-                                    panel.border = ggplot2::element_blank(),
-                                    axis.text.y = ggplot2::element_text(size = 13,
-                                                                        hjust = 1),
-                                    title = ggplot2::element_text(size = 13,
-                                                                  face = "bold"),
-                                    legend.text = ggplot2::element_text(size = 12),
-                                    plot.caption = ggplot2::element_text(size = 10,
-                                                                         face = "plain",
-                                                                         hjust = 1)))
-  #Plot 1: Perception meritocratic
-  p1<- likert_issp %>%
-    dplyr::select(hwork_r:ambition_r) %>%
-    sjPlot::plot_likert(geom.colors = "PuBu",
-                        grid.range  =  c (1.2 , 1.2),
-                        title = "a. Perception meritocratic",
-                          geom.size = c(0.62),
-                          catcount = 5,
-                          expand.grid = T,
-                          values  =  "sum.outside",
-                          reverse.colors = T,
-                          reverse.scale = F,
-                          show.n = FALSE) +
-    ggplot2::theme(legend.position = "none")
-  #Plot 2: Perception Non-meritocratic
-  p2<-likert_issp %>%
-    dplyr::select(wealthy_r:polcone_r) %>%
-    sjPlot::plot_likert(geom.colors = "PuBu",
-                        grid.range  =  c (1.2 , 1.2),
-                        title = "b. Perception Non-meritocratic",
-                          geom.size = c(0.9),
-                          catcount = 5,
-                          expand.grid = T,
-                          values  =  "sum.outside",
-                          reverse.colors = T,
-                          reverse.scale = F,
-                          show.n = FALSE)  +
-    ggplot2::theme(legend.position = "none")
-  #Plot 3: Preference meritocratic
-  p3<-likert_issp %>%
-    dplyr::select(welljob_r:hardjob_r) %>%
-    sjPlot::plot_likert(geom.colors = "PuBu",
-                        grid.range  =  c (1.2 , 1.2),
-                        title = "c. Preference meritocratic",
-                          geom.size = c(0.62),
-                          catcount = 5,
-                          expand.grid = T,
-                          values  =  "sum.outside",
-                          reverse.colors = T,
-                          reverse.scale = F,
-                          show.n = FALSE) +
-    ggplot2::theme(legend.position = "none")
-  #Plot 4: Preference Non-meritocratic
-  p4<-likert_issp %>%
-    dplyr::select(family_r:child_r) %>%
-    sjPlot::plot_likert(geom.colors = "PuBu",
-                        grid.range  =  c (1.2 , 1.2),
-                        title = "d. Preference Non-meritocratic",
-                          geom.size = c(0.62),
-                          catcount = 5,
-                          expand.grid = T,
-                          values  =  "sum.outside",
-                          reverse.colors = T,
-                          reverse.scale = F,
-                          show.n = FALSE) +
-    ggplot2::theme(legend.position = "none")
-
-  #Legend: Extract legend for the final plot using ggpubr
-  leg<- ggpubr::as_ggplot(
-    ggpubr::get_legend(
-      dplyr::select(likert_issp, family_r:child_r) %>%
-        sjPlot::plot_likert(geom.colors = "PuBu", catcount = 5) +
-        ggplot2::scale_fill_brewer(
-          breaks = c(5:1),
-          direction = -1,
-          labels = c("Not important at all","Not very important","Fairly important",
-                     "Very important","Essential")) +
-        ggplot2::theme(legend.position = "bottom")
-      )
-    )
-```
-
-```{r likert-issp,fig.height=11,fig.width=11,fig.cap="Distribution of responses in the ISSP meritocracy items", out.width='80%'}
-  # dev.copy(png,'output/images/Fig2.plotlikert-study1.png', width = 800, height = 600) # save image
-  if (!require("pacman")) install.packages("pacman")
-  pacman::p_load(cowplot, dplyr)
-  #Final plot
-  cowplot::add_sub(
-    cowplot::plot_grid(p1,p2,p3,p4,leg, #plot 1,2,3,4 and legend (bottom)
-      align = "hv",
-      axis = "a",
-      nrow = 5, #5 plots
-      ncol = 1, #1 column
-      rel_heights = c(1,2,1,1,0.25) #fig size within plot
-      ),
-    label = paste0(
-      "Source: Authors calculation based on International Social Survey Programme - Social Inequality 2009",
-      " (n=",dim(likert_issp)[1],")"
-      ),        # Caption: source and sample size
-    size = 10,  # Caption font size
-    hjust = 0.25# Caption align
-    ) %>%
-    cowplot::ggdraw()
-    # dev.off()
-```
-
-In terms of the association between the indicators, Figure \@ref(fig:corr-issp) shows the polychoric correlation matrix.
-Firstly, we observe in general that the moderate to high associations are between the pairs of items representing each of the dimensions as described in \@ref(tab:table-issp).
-The exception here is the dimension of non-meritocratic perception, in which there are six items that appear mostly associated by pairs according to their specific topics (family background, personal background, and connections).
-Still, between this same set of items, there are moderate correlations (around 0.3) which could anticipate a single latent factor underlying non-meritocratic perception and that is tested next through confirmatory factor analysis.
-
-```{r corr-issp,fig.cap = "Perceptions and preferences for ISSP  meritocracy items' polychoric correlations", fig.align='center',fig.height=6,fig.width=8}
-  # dev.copy(png,'output/images/Fig3.plotcor-study1.png', width = 800, height = 600) # save image
-  load(file = "input/data/proc/data_s1.RData")
-  if (!require("pacman")) install.packages("pacman")
-  pacman::p_load(dplyr, corrplot, lavaan)
-  # create polychoric correlation matrix
-  cor_issp<-
-    data_s1 %>%
-    dplyr::select(hwork,ambition,wealthy,pareduc,
-                  race,gender,people,polcone,
-                  welljob,hardjob,family,child) %>%
-    lavaan::lavCor(., ordered=names(.))
-  diag(cor_issp) = NA #set diagonal values to NA
-  # Set Row names of the matrix
-  rownames(cor_issp) <- c("A. Hard work",
-                         "B. Having ambition",
-                         "C. Wealthy family",
-                         "D. Educated parents",
-                         "E. Race",
-                         "F. Gender",
-                         "G. Knowing people",
-                         "H. Political connections",
-                         "I. Well job",
-                         "J. Hard job",
-                         "K. Support family",
-                         "L. Has children")
-  #set Column names of the matrix
-  colnames(cor_issp) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)",
-                         "(H)", "(I)","(J)","(K)","(L)")
-
-  #Plot the matrix using corrplot
-  corrplot::corrplot(cor_issp,
-    method = "color",
-    addCoef.col = "#000390",
-    type = "upper",
-    tl.col = "black",
-    col=colorRampPalette(c("white","#0068DC"))(12),
-    bg = "white",
-    na.label = "-")
-    # dev.off()
-```
-
-### Confirmatory Factor Analysis
-
-Table \@ref(tab:sum-fit-issp) shows the results of the estimation of two confirmatory models.
-The first one (First Order) corresponds to a model that estimate four factors, each for one of the dimensions presented in \@ref(tab:table-issp), with only regular fit indicators (CFI=0.959, TLI=0.944, RMSEA=0.098, $\chi^2$(df=48)= 21308.535).
-Further analysis showed that the sources for poor fit were mostly related to the items of the non-meritocratic perceptions, which as mentioned above displayed correlations among them not taken into account in this model specification.
-
-```{r sum-fit-issp}
-  load(file = "input/data/proc/data_s1.RData")
-  if (!require("pacman")) install.packages("pacman")
-  pacman::p_load(dplyr, kableExtra, knitr,lavaan)
-  table_format = if(knitr::is_html_output()) { #conditional instructions for kable
-    "html" #if html set "html" in format
-  } else if(knitr::is_latex_output()) {
-    "latex"#if latex set "latex" in format
-  }
-
-  # meassurement model: First order
-  cfa_perpref1 <- '
-  	  # latent variables
-  	  percmerit =~ hwork + ambition
-  	  percnmerit=~ wealthy + pareduc +race + gender +people + polcone
-  	  prefmerit =~ welljob + hardjob
-  	  prefnmerit=~ family + child
-  	  '
-  # Model fit: First order
-  fit_perpref1 <- lavaan::cfa(cfa_perpref1, data = data_s1,ordered = T)
-
-  # meassurement model: Second order
-  cfa_perpref2 <- '
-  	  # latent variables
-  	  percmerit =~ hwork + ambition
-  	  parents =~ wealthy + pareduc
-  	  backgrd =~ race + gender
-  	  networks=~ people + polcone
-  	  percnmerit=~ parents + backgrd + networks
-  	  prefmerit =~ welljob + hardjob
-  	  prefnmerit=~ family + child
-  	  '
-  # Model fit: second order
-  fit_perpref2 <- lavaan::cfa(cfa_perpref2, data = data_s1,ordered = T)
-
-  # extract fit indices from models and add to table
-  sum_fit_issp<- dplyr::bind_rows(
-    lavaan::fitmeasures(fit_perpref1)[c("chisq","df","cfi","tli","rmsea")],
-    lavaan::fitmeasures(fit_perpref2)[c("chisq","df","cfi","tli","rmsea")]
-    )
-
-  # Customize object
-  sum_fit_issp$mod <- c("First order","Second order")
-  sum_fit_issp$nobs <- c(lavaan::nobs(fit_perpref2),lavaan::nobs(fit_perpref2))
-  sum_fit_issp$est <- c("DWLS","DWLS")
-  sum_fit_issp <- dplyr::select(sum_fit_issp,mod,nobs,est,dplyr::everything())
-  colnames <- c("Model","$N$","Estimator","$\\chi^2$","df","CFI","TLI","RMSEA")
-
-  # Create table
-  table_issp_fits <-
-    knitr::kable(sum_fit_issp, format=table_format, digits=3, booktabs=T,
-                 col.names=colnames,
-                 caption = "Summary fit indices according to model",
-                 escape = FALSE) %>%
-    kableExtra::kable_styling(font_size = 10,
-                              full_width = F,
-                              latex_options = "HOLD_position",
-                              bootstrap_options=c("striped", "bordered"))
-  table_issp_fits
-```
-
-Attending to the sources of poor fit from the first estimated confirmatory model, we specified a second model that keeps the basic four-factor structure but generates an additional model for the non-meritocratic perceptions as depicted in Figure \@ref(fig:meas02-issp).
-In this model, the pairs of items in this dimension form three latent factors which at the same time generate a second-order factor of non-meritocratic perceptions, improving the fit of the model.
-
-```{r meas02-issp, echo=FALSE, fig.cap = "Descriptive plot", fig.align='center',fig.cap="Second-order confirmatory factor analysis model using ISSP indicators of Perceptions and Preferences for Meritocracy"}
- knitr::include_graphics('output/images/Fig4.cfa_issp_study1.PNG')
-```
-
-Regarding the correlations between the factors in the Figure \@ref(fig:meas02-issp), we observe that perceptions are correlated with preferences, but more strongly for the meritocratic ($r=0.46$) than for the non-meritocratic ($r=0.27$) dimensions.
-Secondly, both perceptual dimensions (meritocratic and non-meritocratic) depict a moderate correlation ($r=0.30$), suggesting that they are not the opposite poles of the same continuum.
-Something similar occurs for the correlation between the two dimensions of preferences ($r=0.25$).
-
-Based on the measurement validation, we asses if the proposed measurement strategy can be used in cross-national comparisons.
-In this line, Figure \@ref(fig:scat1) depicts the correlation of factor scores for Meritocratic Perception and Meritocratic Preferences at the country level.
-For instance, Bulgaria (BGR) and the United States (USA) appropriately represent the positive association between perception and preferences.
-In other words, countries that describe merit in their society as an essential factor to get ahead in life also consider that effort and talent should be determinant factors in how much an individual earns.
-On the other hand, in countries with lower perceived meritocracy, preferences are also low, where Denmark (DNK) and Venezuela (VEN) are representative societies of this association.
-
-Another finding depicted by Figure \@ref(fig:scat1) is the positive association of Non-meritocratic perception with Meritocratic preferences, showing that countries like Bulgaria (BGR) and Philipines (PHL) represent cases with higher Non-meritocratic perceptions and higher preferences for meritocracy.
-In other words, in societies where people perceive higher importance of non-meritocratic aspects to get ahead in life, people also prefer that merit should be a determinant factor in wages.
-On the other side, countries like Sweden (SWE) or Denmark (DNK), both Scandinavian societies with strong welfare systems, represent contexts in which non-meritocratic aspects are not described as important, and merit is not considered as a factor that should be determinant in wages.
-
-```{r scat1, fig.cap="Cross-national comparison of non-meritocratic and meritocratic perceptions with meritocratic preference",fig.height=6, fig.width=12}
-  load(file = "input/data/proc/data_s1.RData")
-  if (!require("pacman")) install.packages("pacman")
-  pacman::p_load(dplyr, kableExtra, knitr,lavaan,countrycode,ggplot2,haven,
-                 gridExtra)
-  table_format = if(knitr::is_html_output()) { #conditional instructions for kable
-    "html" #if html set "html" in format
-  } else if(knitr::is_latex_output()) {
-    "latex"#if latex set "latex" in format
-  }
-
-  # meassurement model: Second order
-  cfa_perpref2 <- '
-  	  # latent variables
-  	  percmerit =~ hwork + ambition
-  	  parents =~ wealthy + pareduc
-  	  backgrd =~ race + gender
-  	  networks=~ people + polcone
-  	  percnmerit=~ parents + backgrd + networks
-  	  prefmerit =~ welljob + hardjob
-  	  prefnmerit=~ family + child
-  	  '
-  # Model fit: second order
-
-  fit_perpref2 <- lavaan::cfa(cfa_perpref2, data = data_s1,ordered = F)
-
-  df_proc<- data_s1 %>% select(id,v5,wealthy:hardjob)  
-  df_proc$ccode <- as.numeric(data_s1$v5)
-
-  predicted2 <- lavPredict(fit_perpref2,newdata = data_s1)
-
-  idx <- lavInspect(fit_perpref2, "case.idx")
-  fscores <- lavPredict(fit_perpref2)
-  ## loop over factors
-  for (fs in colnames(fscores)) {
-    df_proc[idx, fs] <- fscores[ , fs]
-  }
-
-  df_proc <-  df_proc %>%  select(id,ccode,percmerit,prefmerit,prefnmerit) %>% na.omit()
-
-  df_proc$clab <- countrycode(sourcevar = df_proc$ccode,
-                            origin = "iso3n",
-                            destination = "iso3c")
-
-avg_count <-  
-  df_proc %>% group_by(clab) %>% summarise(prefmerit=mean(prefmerit),
-                                            percmerit=mean(percmerit),
-                                            prefnmerit=mean(prefnmerit))
-
-# create figure 1: perception merit by preferences merit
-figure1 <- 
-ggplot(avg_count, aes(x=percmerit, y=prefmerit)) +
-  # geom_point() + 
-  geom_text(label=as.character(avg_count$clab)) +
-  geom_smooth(method = "lm") +
-  xlab("Meritocratic perception")+
-  ylab("Meritocratic preference")  
-# create figure 2: perception non-merit by preferences merit
-figure2 <- 
-ggplot(avg_count, aes(x=prefnmerit, y=prefmerit)) +
-  # geom_point() + 
-  geom_text(label=as.character(avg_count$clab)) +
-  geom_smooth(method = "lm") +
-  xlab("Non-meritocratic perception")+
-  ylab("Meritocratic preference")
-
-grid.arrange(figure1, figure2, ncol = 2,
-             bottom = grid::textGrob(label = paste0("Source: Authors calculation based on International Social Survey Programme - Social Inequality 2009",
-      " (n=",dim(df_proc)[1],")"),
-                                     hjust = 0.1,
-                                     gp = grid::gpar(fontface = 1, fontsize = 9)))
-```
-
-## Discussion study 1
-
-Although it was possible to get an adequate fit of the model to the pooled data from ISSP, there are several aspects to take into account when using these items for measuring meritocracy.
-First of all, these items were not specifically designed based on a conceptual approach to meritocracy, but still is possible to find a certain fit of the model to the data after minor adjustments.
-Secondly, some of the items in similar dimensions refer to different concepts, as for instance meritocratic perceptions are related to ambition and hard work, whereas meritocratic preferences deal with how well a job is done.
-A better operationalization would require more alignment between the different sides of the same concept.
-And thirdly, the scale is rather unbalanced given he larger amount of non-meritocratic perception items, producing correlated errors.
-All these limitations are taken into account next in the design of Study 2.
-
-With regard to the meritocratic / non-meritocratic distinction in previous studies, we can preliminary conclude that the wide use of reverse-coding for non-meritocratic items attempting to measure meritocracy is not an adequate operationalization of the construct.
-Meritocracy and non-meritocracy are related but different concepts, meaning that it is theoretically possible to find different combinations of both at individual and contextual levels.
-This finding opens a series of avenues for future research that are considered in the conclusions.
-
-# Study 2: The Perceptions and Preferences for Meritocracy Scale
-
-## Data
-
-The data was obtained through an online survey which was part of a larger study on meritocracy and preferences developed in Chile in 2019, funded by the Chilean National Scientific Agency (ANID).
-The questionnaire was programmed in Qualtrics and the fieldwork was conducted by an external online survey agency ([netquest.com](www.netquest.com)) between December 2019 and January 2020.
-The sample was selected from a non-probabilistic quota design in three large cities in Chile (Santiago, Concepción & Antofagasta).
-The quotas for gender, age, and educational levels were generated based on a survey by the Public Studies Center [@cep_encuesta_2019], which is a well-regarded counterpart agency of the ISSP (International Social Survey Programme) in Chile.
-A total sample of 2,141 people was collected, excluding those who did not sign the informed consent as well as those not answering the meritocracy response battery.
-There were no significant differences between our sample and the wider population for most socio-demographic characteristics, with the exception of educational level (see Table \@ref(tab:rep-samp) in Appendix).
-As it is often the case with online surveys, there were some limitations in achieving the quotas for lower educational levels [@zhang_quota_2018; @boas_recruiting_2020].
-This survey was designed as a three-wave panel and the data used in this analysis corresponds to the first wave.
-
-## Study design
-
-### Instrument
-
-The proposed scale of meritocratic perceptions and preferences consisted of eight indicators, two for each of the four dimensions listed earlier: Perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic).
-This design is oriented by two main criteria.
-The first one is usability in social surveys research, for which the length of the scale is restricted due to space reasons in such studies.
-Nevertheless, at the same time the scale should offer possibilities of accounting for measurement error through factor analysis techniques.
-Although a rule of thumb in latent  measurement suggests at least three indicators per dimension, we still opted for two in order to count with a brief scale.
-Although such decision posits some limitation for modeling measurement error, still allows for factor analysis and is a common approach in previous social survey studies with limited items (for instance see @davidov_measurement_2009).
-
-The second criteria is to expand the possibilities of comparability with previous studies on meritocracy, by considering some of the items previously used in this research area.
-To this regard, some of the items were adapted from the items battery "reasons to get ahead" (ISSP/GSS), which has been widely used for operationalizing meritocracy in international research [@mijs_paradox_2019; @duru-bellat_who_2012; @reynolds_perceptions_2014].
-The scale items ordered according to their respective dimensions are presented in Table \@ref(tab:table-indicadores).
-These eight likert-type items have five response alternatives, ranging from "Completely disagree"(1) to "Completely agree" (5).
-
-```{r table-indicadores, echo=FALSE}
- if (!require("pacman")) install.packages("pacman")
- # remotes::install_github('haozhu233/kableExtra@a6af5c0')  # for problems with collapse_rows()
- pacman::p_load(knitr, kableExtra, dplyr)
- table_format = if (knitr::is_html_output()) {
-   #conditional instructions for kable
-   "html" #if html set "html" in format
- } else if (knitr::is_latex_output()) {
-   "latex"#if latex set "latex" in format
- }
- tabitems <- read.csv(file = "input/tables/table01.csv",header = 1,sep = ",",encoding = "UTF-8") # call file  generated externally for the table
- cnames <- c("Component", "Dimensions","Item (English)" , "Item original (Spanish)")
- cap <- "Items of the perceptions and preferences for meritocracy scale."
-
- knitr::kable(tabitems, table_format, booktabs = T, linesep = "",col.names = cnames, caption = cap) %>%
-   kableExtra::kable_styling(
-     full_width = T,
-     latex_options = c("hold_position"),
-     position = "center",
-     font_size = 10,
-     bootstrap_options=c("striped", "bordered")) %>%
-   kableExtra::column_spec(column = 1, width = "1.5cm", ) %>%
-   kableExtra::column_spec(column = 2,width = "2 cm") %>%
-   kableExtra::collapse_rows(columns = 1:2,valign = "middle")
-```
-
-\pagebreak
-
-### Administration sets
-
-With the objective of evaluating the effect of indicator ordering in the responses, three different versions of items' order were designed and randomly assigned to the respondents, as depicted in Figure \@ref(fig:appmod).
-The scale was presented to the first group (_n = 712_) in the order that appears in Table \@ref(tab:table-indicadores) according to perceptions and preferences.
-For the second group (_n = 717_), the order was reorganized according to perceptions and preferences over the same topic, e.g.
-for the topic of hard work, the item about perception was followed by the item about preference and the same for the rest of the topics.
-Finally, for the third group (_n = 712_), the items were presented as completely randomized.
-
-```{r appmod, echo=FALSE, fig.cap = "Survey flow", fig.align='center'}
- knitr::include_graphics('output/images/Fig5.app_mod.png')
-```
-
-## Methods
-
-To test the scale's underlying constructs, we employed confirmatory factor analysis models (CFA).
-The models estimated one factor for each of the four proposed dimensions presented in Table \@ref(tab:table-indicadores).
-As in Study 1, CFA was conducted using the `lavaan` R package (version 0.6-3; @rosseel_lavaan_2012a), with diagonally weighted least squares (DWLS) estimation due to the items' ordinal level of measurement [@kline_principles_2016; @rosseel_lavaan_2012a].
-The fit indexes and cut-off criteria were the same as the ones used in Study 1.
-
-A pre-registration was made in the OSF platform, available at the following link: [https://osf.io/z45y2](https://osf.io/z45y2).
-Included in this pre-registration are the hypotheses regarding the four-dimensional conceptual model underlying the scale, the variable measurement levels, the statistical tests to be performed with their respective evaluation parameters, and other important aspects of the research design.
-
-## Results
-
-### Descriptive analyses
-
-The graphs presented in Figure \@ref(fig:plotlikert) display disaggregated and comparable information of the different response categories for each item.
-As it can be observed, in general, there is more agreement in the perception of non-meritocratic items than in meritocratic ones, while in the case of preferences the opposite occurs.
-As far as preferences are concerned, the preponderant role of effort over talent as a criterion of meritocratic preference is noteworthy.
-All in all the descriptive results tend to show a sort of critical view of meritocracy, perceiving the operation of non-meritocratic aspects over meritocratic ones, whereas in the preferences the opposite occurs.
-
-```{r produce-likert_s2, include=FALSE}
-  if (!require("pacman")) install.packages("pacman")
-  pacman::p_load(dplyr,sjPlot, ggplot2, sjlabelled,sjmisc, ggpubr)
-  load(file = "input/data/proc/data_s2.RData")
-  names(data_s2)
-  dat_likert_s2 <- data_s2 %>% dplyr::select(!group) # exclude group variable
-  dat_likert_s2 <- sjmisc::rec(dat_likert_s2, rec="rev", append=FALSE)
-  names(dat_likert_s2) # reverse coding for likert graph
-
-  table(data_s2$perc_effort)
-  table(dat_likert_s2$perc_effort_r) # check ok.
-  # update.packages("sjPlot") # please install version 2.8.9 or above
-  ggplot2::theme_set(ggplot2::theme(panel.background = ggplot2::element_rect(fill = "gray85",
-                                                                             colour = "gray85"),
-                                    panel.border = ggplot2::element_blank(),
-                                    axis.text.y = ggplot2::element_text(size = 13,
-                                                                        hjust = 1),
-                                    title = ggplot2::element_text(size = 13,
-                                                                  face = "bold"),
-                                    legend.text = ggplot2::element_text(size = 12),
-                                    plot.caption = ggplot2::element_text(size = 10,
-                                                                         face = "plain",
-                                                                         hjust = 1)))
-
-  #Plot 1: Perception
-  p1<-
-    dplyr::select(dat_likert_s2,perc_effort_r:perc_netw_r) %>%
-    sjPlot::plot_likert(geom.colors = "PuBu",
-                        title = c("a. Perceptions"),
-                        geom.size = 0.8,
-                        axis.labels = c("Effort", "Talent", "Rich Parents", "Contacts"),
-                        catcount = 4,
-                        cat.neutral = 3,
-                        grid.range  =  c (1.2 , 1.2),
-                        values  =  "sum.outside",
-                        reverse.colors = T,
-                        reverse.scale = F,
-                        show.n = FALSE) +
-    ggplot2::theme(legend.position = "none")
-
-  #Plot 2: Preferences
-  p2<-
-    dplyr::select(dat_likert_s2,pref_effort_r:pref_netw_r) %>%
-    sjPlot::plot_likert(geom.colors = "PuBu",
-                        title = c("b. Preferences"),
-                        geom.size = 0.8,
-                        axis.labels = c("Effort", "Talent", "Rich Parents", "Contacts"),
-                        catcount = 4,
-                        cat.neutral = 3,
-                        grid.range  =  c (1.2 , 1.2),
-                        values  =  "sum.outside",
-                        reverse.colors = T,
-                        reverse.scale = F,
-                        show.n = FALSE)  +
-    ggplot2::theme(legend.position = "none")
-
-  #Legend: Extract legend for the final plot using ggpubr
-  leg<- ggpubr::as_ggplot(
-    ggpubr::get_legend(
-    dplyr::select(dat_likert_s2,pref_effort_r:pref_netw_r) %>%
-        sjPlot::plot_likert(geom.colors = "PuBu",
-                            catcount = 4,
-                            cat.neutral = 3) +
-      ggplot2::theme(legend.position = "bottom") +
-      ggplot2::scale_fill_manual(
-        values = c("#b3b3b3ff", "#f1eef6ff", "#bdc9e1ff","#74a9cfff","#0570b0ff"),
-        breaks=c("neutral",4,3,2,1),
-        labels=c("Neither agree nor disagree","Strongly disagree","Disagree",
-                 "Agree","Totally agree")
-        )
-      )
-    )
-```
-
-```{r plotlikert, echo=FALSE, fig.align='center', fig.cap="Distribution of responses in the Merit Scale items", fig.height=8, fig.width=11}
-  # dev.copy(png,'output/images/Fig6.plotlikert-study2.png', width = 820, height = 600)
-  if (!require("pacman")) install.packages("pacman")
-  pacman::p_load(cowplot)
-  #Final plot
-  cowplot::add_sub(
-    cowplot::plot_grid(p1,p2,leg, #plot 1,2 and legend (bottom)
-      align = "hv",
-      axis = "a",
-      nrow = 3, #2 plots
-      ncol = 1, #1 column
-      rel_heights = c(1,1,0.25) #fig size within plot
-      ),
-    label = paste0("Source: Authors calculations based on Study 2 sample"," (n=",dim(na.omit(dat_likert_s2))[1],")"
-      ),        # Caption: source and sample size
-    size = 12,  # Caption font size
-    hjust = -0.15# Caption align
-    ) %>%
-    cowplot::ggdraw()
-  # dev.off()
-```
-
-Attending now to the association among the scale items, Figure \@ref(fig:corpoly) shows the items' polychoric correlations.
-There are three main aspects to highlight from this correlation matrix.
-Firstly, and as expected,  the largest correlations are between indicators that correspond to the same factors behind the conceptual model (e.g., perception of meritocracy by effort and by talent, $r=0.52$).
-Secondly, among these correlations the highest are those between the non-meritocratic dimension but in perceptions ($r=0.73$) and preferences ($r=0.61$).
-Thirdly, both items for meritocratic preferences (E and F) are the ones that mostly correlate with the rest of the perceptual items, showing medium to high correlations.
-This is noteworthy because it would indicate that the perception of non-meritocracy would be related to larger meritocratic preferences.
-Finally, and similar to what was found in Study 1 with ISSP data, we observe that there are no considerable negative correlations between meritocratic and non-meritocratic aspects, undermining the assumptions of previous studies that suggested that these dimensions would be the opposite poles of one same continuum [@reynolds_perceptions_2014].
-
-```{r corpoly, echo=FALSE, fig.cap = "Perceptions and preferences for meritcoracy items' polychoric correlations", fig.align='center'}
-  # dev.copy(png,'output/images/Fig7.plotcor-study2.png',width = 800, height = 600) # save image
-  if (!require("pacman")) install.packages("pacman")
-  load(file = "input/data/proc/data_s2.RData")
-  pacman::p_load(dplyr, corrplot, lavaan)
-  # generate polychoric matrix
-  cor_s2<- data_s2 %>%
-    dplyr::select(perc_effort,perc_talent,perc_wpart,perc_netw,pref_effort,pref_talent,pref_wpart,pref_netw) %>%
-    lavaan::lavCor(., ordered=names(.))
-
-  diag(cor_s2) = NA
-
-  rownames(cor_s2) <-c(
-      "A. Perception Effort",
-      "B. Perception Talent",
-      "C. Perception Rich parents",
-      "D. Perception Contacts",
-      "E. Preferences Effort",
-      "F. Preferences Talent",
-      "G. Preferences Rich parents",
-      "H. Preferences Contacts")
-  colnames(cor_s2) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)", "(H)")
-
-  corrplot::corrplot(cor_s2,
-    method = "color",
-    addCoef.col = "#000390",
-    type = "upper",
-    tl.col = "black",
-    col=colorRampPalette(c("white","#0068DC"))(8),
-    bg = "white",
-    na.label = "-")
-  # dev.off()
-```
-
-### Confirmatory Factor Analysis
-
-The present section describes the results of the confirmatory factor analysis estimation.
-The model estimates four latent factors: perception meritocratic, perception non-meritocratic, preferences meritocratic, and preferences non-meritocratic.
-Each factor is estimated based on two items of the scale as detailed in Table \@ref(tab:table-indicadores).
-
-The first step in the analysis consists of comparing the model fit indicators for the three versions of the scale that were randomly assigned to the participants: order according to perceptions/preferences, order according to topics, and random order (see Figure \@ref(fig:appmod)).
-
-```{r model-generation-cfa}
-  # Libraries & data
-  if (!require("pacman")) install.packages("pacman")
-  pacman::p_load(lavaan)
-  load(file = "input/data/proc/data_s2.RData")
-  # names(data_s2)
-
-  # model
-  model_cfa <- '
-  perc_merit =~ perc_effort+perc_talent
-  perc_nmerit=~perc_wpart +perc_netw
-  pref_merit =~ pref_effort+pref_talent
-  pref_nmerit=~pref_wpart +pref_netw
-  '
-
-  # estimation for each order set
-  fit_cfa1 <- lavaan::cfa(model_cfa, data = data_s2[data_s2$group == 1,], ordered = TRUE, std.lv=FALSE)
-  fit_cfa2 <- lavaan::cfa(model_cfa, data = data_s2[data_s2$group == 2,], ordered = TRUE, std.lv=FALSE)
-  fit_cfa3 <- lavaan::cfa(model_cfa, data = data_s2[data_s2$group == 3,], ordered = TRUE, std.lv=FALSE)
-```
-
-```{r table-cfafits, echo=FALSE}
-  if (!require("pacman")) install.packages("pacman")
-  pacman::p_load(dplyr, kableExtra, knitr)
-  table_format = if(knitr::is_html_output()) { #conditional instructions for kable
-     "html" #if html set "html" in format
-   } else if(knitr::is_latex_output()) {
-     "latex"#if latex set "latex" in format
-   }
-  # extract fit indices from models and add to table
-  sum_fit<- dplyr::bind_rows(lavaan::fitmeasures(fit_cfa1)[c("chisq","df","cfi","tli","rmsea")],
-                             lavaan::fitmeasures(fit_cfa2)[c("chisq","df","cfi","tli","rmsea")],
-                             lavaan::fitmeasures(fit_cfa3)[c("chisq","df","cfi","tli","rmsea")])
-
-  # Customize object
-  sum_fit$mod <- c("Version 1","Version 2","Version 3")
-  sum_fit$nobs <- c(nobs(fit_cfa1),nobs(fit_cfa2),nobs(fit_cfa3))
-  sum_fit$est <- c("DWLS","DWLS","DWLS")
-  sum_fit <- dplyr::select(sum_fit,mod,nobs,est,everything())
-  colnames <- c("Model","$N$","Estimator","$\\chi^2$","df","CFI","TLI","RMSEA")
-
-  # Create table
-  table_cfa_fits <-kable(sum_fit, format=table_format,
-                         digits=3,
-                         booktabs=T,
-                         col.names=colnames,
-                         caption = "Summary fit indices according to order versions",
-                         escape = FALSE) %>%
-    kable_styling(full_width = F,
-                  font_size = 10,
-                  latex_options = "HOLD_position",
-                  bootstrap_options=c("striped", "bordered"))
-  table_cfa_fits
-```
-
-Table \@ref(tab:table-cfafits) shows the fit indicators of the models estimated for each of the three versions of the items' order described in the methodology.
-Regardless of the version, all models obtained adequate fit indicators, with CFI's above 0.95 and RMSEA's below 0.08.
-However, none of the models achieved a non-significant chi-square, something expected in large samples as the one used here.
-The first version order (perceptions-preferences) was the one obtaining best fit (CFI=0.993, TLI=0.995, RMSEA=0.034, $\chi2$(df=14)=42.276), whereas version 2 with the fixed order according to merit/non-merit items shows the comparatively worst indicators.
-The CFA fit indices for the completely randomized items' order (Model 3) it keeps all the indicators within the acceptable cut-off criteria and besides it controls for possible order effects in the administration of the instrument.
-The model and parameter estimates for this version are depicted in Figure \@ref(fig:meas02):
-
-```{r meas02, echo=FALSE, fig.cap = "Confirmatory factor analysis of the Perceptions and Preferences for Meritocracy Scale", fig.align='center'}
- # estimation for diagram: summary(fit_cfa3, standardized = TRUE, fit.measures=TRUE)
- knitr::include_graphics('output/images/Fig8.cfa_randomized_study2.png')
-```
-
-Attending to the correlations between the latent variables as depicted in Figure \@ref(fig:meas02), meritocratic preferences are moderate to highly correlated with perceptions, both meritocratic ($r=0.457$) and non-meritocratic ($r=0.500$).
-The correlation between both types of perceptions ($r=-0.044$) and both types preferences ($r=0.185$) are low, as well as between non-meritocratic preferences and perceptions ($r=-0.059$).
-This last finding gives further evidence regarding the lack of unidimensionality of meritocratic and non-meritocratic aspects as assumed by previous studies.
-
-## Discussion study 2
-
-The estimation performed with different administration sets for the meritocracy scale gives evidence in line with the conceptual model.
-The four dimensions appear as independent but related, although with some differences when compared with the estimation with ISSP data in Study 1.
-The main general difference is that with ISSP data the four latent dimensions were positively correlated, whereas now some correlations are statistically not different from zero.
-For instance, perceptions and preferences for meritocracy keep being related, which is not the case for the non-meritocratic pair.
-In principle this makes more sense in theoretical terms, as if someones perceives that meritocracy functions well, probably will support it more.
-On the other side, if someones perceives non-meritocracy, this should lead to a larger preference for meritocracy as it would constitute a demand for lager meritocracy when this one is not perceived to work.
-
-Despite the good general working of the scale, there are some limitations in the interpretation of the results given the few items (two) per dimension.
-As it was mentioned before, the option of a minimalistic approach in terms of items has advantages and disadvantages: advantages in terms of usability in social surveys with limited space, and disadvantages in terms of measurement validity in the estimation of the latent constructs.
-Being aware of this trade-off, in the following section we offer additional validity evidence as a way of complementing some of the weaknesses identified.
-
-# Study 3: Additional validity analyses
-
-The modeling for invariance testing has been gaining more relevance in sociological survey studies, mainly due to the interest in the validity of measurement scales in comparative studies in various social and cultural contexts [@davidov_measurement_2014].
-In this section we present the results of an analysis of invariance between groups, testing whether the proposed factorial structure for the meritocracy scale is applicable in a different dataset.
-
-The invariance measurement procedure consists of a series of nested models to which restrictions are progressively incorporated on the parameters of the measurement model.
-The literature generally suggests that this modeling should be done through four levels or types of progressive restriction [@milfont_testing_2010; @millsap_statistical_2011; @vandeschoot_checklist_2012]:
-
-1. **Configural:** the model is estimated only indicating the factorial structure used in the CFA.
-
-2. **Weak:** an equality restriction is applied to the factor loadings in the different groups, that is, the loadings are forced to be identical in both measurements.
-
-3. **Strong:** equality restrictions are added to the intercepts of each indicator.
-
-4. **Strict:** equality restrictions are added to the error variances of each indicator.
-
-### Data
-
-For testing the invariance we compared the data described previously in Study 2 to a different  dataset coming from an online survey carried out during the first half of 2020.
-The characteristics of this sample in terms of application and coverage are equivalent to those of the previous study.
-The final sample obtained contains 1,242 cases, where 605 (48.71%) are women and 637 (51.28%) are men.
-48.84% of them are 45 years of age or older, and 33.97% have tertiary education or higher.
-
-### Variables
-
-The items of the meritocracy scale were identical to the original and they  were administrated in a randomized order (which corresponds to the application modality of Group 3 in Study 2, as showed in Figure \@ref(fig:appmod)).
-
-### Results
-
-The first step for invariance testing is the estimation of the configural model, which serves as the baseline for further comparison and is expected to adequately meet the global fit criteria of a measurement model.
-Although the chi-square statistic is used as a global measure of fit, it is usually complemented with other indexes given its high sensitivity to sample size: the Comparative Fit Index (CFI), which should have a value greater than 0.95; the Root Mean Square Error Approximation (RMSEA), which must be lower than 0.06; and the Standardized Root Mean Square Residuals (SRMR) which must be less than 0.08 [@vandeschoot_checklist_2012].
-Besides, the literature on measurement invariance suggests some complementary approaches for the evaluation of the fit, from which we will consider the incremental adjustment of the fit indexes [@cheung_evaluating_2002; @milfont_testing_2010; @dimitrov_testing_2010] and the ANOVA test for means comparison in nested models [@newsom_longitudinal_2015].
-
-```{r invargroup, echo=FALSE}
-  if (!require("pacman")) install.packages("pacman")
-  #---Load packages and data -------------------------------------#
-  pacman::p_load(dplyr,lavaan,semTools,knitr,kableExtra,stats,gtools)
-  options(knitr.kable.NA = '')
-  load(file = "input/data/proc/data_s3_inv.RData")
-  pov01 <- data_s3_inv %>% dplyr::filter(dataset=="pvw01")
-  fs01  <- data_s3_inv %>% dplyr::filter(dataset=="fsw02")
-
-  #-----Meassurement model--------------------------------------#
-  model01 <- '
-  perc_merit  =~ perc_effort + perc_talent
-  perc_nmerit =~ perc_wpart  + perc_netw
-  pref_merit  =~ pref_effort + pref_talent
-  pref_nmerit =~ pref_wpart  + pref_netw'
-
-  #-----Invariance---------------------------------------------#
-  inv01<- semTools::measurementInvariance(model=model01,data=data_s3_inv,group="dataset",estimator = "ML",strict=TRUE,quiet = T)
-
-  conf  <- inv01$fit.configural
-  weak  <- inv01$fit.loadings
-  strong<- inv01$fit.intercepts
-  strict<- inv01$fit.residuals
-
-  tab01<- lavaan::anova(conf,weak,strong,strict,SB.classic=TRUE) %>%
-    dplyr::as_tibble() %>%
-    dplyr::select("Chisq","Df","chisq_diff"=`Chisq diff`,"df_diff"=`Df diff`,"pvalue"=`Pr(>Chisq)`) %>%
-    dplyr::mutate(stars=gtools::stars.pval(pvalue),
-           chisqt=paste0(round(Chisq,2)," (",Df,") "),
-           decision=ifelse(pvalue>0.05,yes = "Accept",no = "Reject"),
-           model=c("Configural","Weak","Strong","Strict"))
-
-  fit.meas<- dplyr::bind_rows(lavaan::fitmeasures(inv01$fit.configural,output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
-                              lavaan::fitmeasures(inv01$fit.loadings,  output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
-                              lavaan::fitmeasures(inv01$fit.intercepts,output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
-                              lavaan::fitmeasures(inv01$fit.residuals, output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),])
-
-  # compute differences in chisq, df, cfi and rmsea (90%, lower.ci - upper.ci )
-  fit.meas<- fit.meas %>%
-    dplyr::mutate(diff.chi2 = chisq    - lag(chisq,default = dplyr::first(chisq)),
-           diff.df   = df       - lag(df,   default = dplyr::first(df)),
-           diff.cfi  = cfi      - lag(cfi,  default = dplyr::first(cfi)),
-           diff.rmsea   = rmsea - lag(rmsea,default = dplyr::first(rmsea))) %>%
-    round(3) %>%
-    dplyr::mutate(rmsea.ci=paste0(rmsea," \n ", "(",rmsea.ci.lower,"-",rmsea.ci.upper,")"))
-
-  tab.inv<- dplyr::bind_cols(tab01,fit.meas) %>%
-    dplyr::select(model,chisqt,cfi,rmsea.ci,diff.chi2,diff.df,diff.cfi,diff.rmsea,stars,decision) %>%
-    dplyr::mutate(diff.chi2=paste0(diff.chi2," (",diff.df,") ",stars)) %>%
-    dplyr::select(model,chisqt,cfi,rmsea.ci,diff.chi2,diff.cfi,diff.rmsea,decision)
-
-  #clean values
-  tab.inv[tab.inv == c("0 (0) ")] <- NA
-  tab.inv[tab.inv == c(0)] <- NA
-
-  ### Check an alternative method
-  configural        <- lavaan::cfa(model01, data=data_s3_inv, group = "dataset", estimator="ML")
-  weak.invariance   <- lavaan::cfa(model01, data=data_s3_inv, group = "dataset",group.equal = "loadings", estimator="ML")
-  strong.invariance <- lavaan::cfa(model01, data=data_s3_inv, group = "dataset",group.equal = c( "loadings", "intercepts"), estimator="ML")
-  strict.invariance <- lavaan::cfa(model01, data=data_s3_inv, group = "dataset",group.equal = c( "loadings", "intercepts", "residuals"), estimator="ML")
-
-  table_format = if(is_html_output()) {
-    "html"
-  } else if(is_latex_output()) {
-    "latex"
-  }
-
-  col.nam <- c("Model","$\\chi^2 (\\text{df})$","CFI","RMSEA (90 CI)",
-               "$\\Delta \\chi^2 (\\Delta \\text{df}$)","$\\Delta \\text{CFI}$","$\\Delta \\text{RMSEA}$","Decision")
-  footnote <- paste0("N = ","; Group 1, n = ",conf@Data@nobs[[1]],"; Group 2, n = ",conf@Data@nobs[[2]], ", ***p < 0.001")
-
-  knitr::kable(tab.inv, col.names = col.nam, align = "l",
-        booktabs=TRUE,format = table_format,escape = FALSE,
-        caption = "Multiple Group meassurement invariance for Perceptions and Preferences for Meritocracy") %>%
-    kableExtra::kable_styling(full_width = FALSE,
-                              latex_options = "HOLD_position",
-                              bootstrap_options=c("striped", "bordered"),
-                              font_size = 10) %>%
-    kableExtra::footnote(general = footnote, footnote_as_chunk = T)
-```
-
-Table \@ref(tab:invargroup) shows the results of the measurement invariance estimation.
-When attending to the traditional invariance test of $\Delta \chi^2 (\Delta \text{df})$, the results support the invariance at the strong level meaning that the fit of the factor model of the merit scale is equivalent across samples when constraining factor loadings and intercepts to being equal.
-Such result is considered in general as evidence of invariance [@fischer_are_2011], as strict forms of measurement invariance rarely hold [@vandeschoot_checklist_2012].
-Still, the comparability of latent means requires strict invariance which in this case does not hold when considering just $\Delta \chi^2 (\Delta \text{df})$.
-Nevertheless, the criteria of $\Delta \text{CFI}$ used for comparing models is close to the rejection criteria of >.01, whereas the $\Delta \text{RMSEA}$ fulfills the requirements of being below the cut-off criteria as suggested by [@chen_sensitivity_2007].
-Therefore, using this last standard, the level for strict invariance would hold for the meritocracy scale.
-
-## Discusion Study 3
-
-Based on the measurement invariance analysis for two datasets, we found evidence supporting the equivalence of the scale when applied in different samples.
-This gives a more robust ground when using this measure for comparing groups, as for instance country comparison.
-Although the results are promising, still the analysis was performed in two samples in one country (Chile) and it requires further examination with social surveys in other societies.
-
-# Conclusions
-
-Studies that attempt to characterize and compare societies by their support for meritocratic beliefs have used different approaches.
-As most studies use secondary survey data, they tend to assume that the available indicators represent an underlying meritocratic construct.
-A review of these studies reveals several non-tested assumptions, as well as the use of similar indicators to represent different constructs and dimensions of meritocracy.
-As the existence of heterogeneous approaches certainly has consequences for the advancement of the study of meritocracy, this paper presented a comprehensive conceptual framework for the empirical study of meritocracy, building upon previous research.
-This framework was operacionalized in a survey scale and tested against available and new survey data.
-
-We identified four critical aspects regarding the measurement of meritocracy in previous studies: unidimensionality, the ambiguous use of the term "beliefs", the use of non-meritocratic indicators as opposed to meritocratic indicators, and the consideration of measurement error.
-The proposed 8-item scale, "Perceptions and Preferences for Meritocracy", was designed and tested in order to deal with these four issues.
-Our results indicate that perceptions and preferences seem to be two related but different dimensions, often confounded in previous research under the label of "beliefs".
-At the same time, meritocratic and non-meritocratic dimensions do not appear to constitute poles of the same continuum, as some previous studies have assumed.
-Regarding the possible effects of the items' order in the estimation of the latent variable, we tested three different order versions and found evidence that suggests the use of the randomized version of the scale.
-Furthermore, we found evidence of measurement invariance that approaches well the requirement of comparability between samples.
-
-One of the criteria in designing the merit scale was to develop a short instrument that could be used in social surveys, assuming that using only two items per dimension posits limitations for the confirmatory analysis estimated here.
-We preferred this confirmatory approach at it offers several advantages in terms of information about the quality of the measurement, and we offered additional validity evidence with measurement invariance analysis.
-Despite the advantages, a more simple and straightforward perspective is always available, as for instance using only the average of the items per dimension.
-Whereas a short scale it is always appreciated in social survey studies, for other audiences could not reach measurement standards.
-This may require further expansions and/or consider using a different instrument as the "Preference for the Merit Principle Scale" [@davey_preference_1999], which measures only one dimension of meritocracy with fifteen items.
-
-The four-dimensional conceptual framework and its operationalization in the Perceptions and Preferences for Meritocracy scale opens several avenues for future research.
-For instance, distinguishing perceptions from preferences will allow us to evaluate the extent to which different societies are accustomed to, or satisfied with, the perceived level of meritocracy, in terms of differences between what is perceived and what is preferred.
-Additionally, given that non-meritocratic factors are not necessarily related to meritocratic ones, our framework makes it possible to assess the perceived legitimacy of practices such as the use of personal contacts and their interference (or not) with meritocratic ideals in different societies.
-The consideration of structural factors, both individual and societal, will allow us to advance future hypotheses relating to meritocratic legitimacy to a more comprehensive level than the current one, which is based solely on meritocratic perceptions.
-Furthermore, the impact of different configurations of the four-dimensional framework on practices and behaviors such as corruption, civic involvement, and political alignment, is an area that requires additional research.
-Such future agendas could be especially relevant in times of economic crisis and growing inequalities, that could entail changes in the legitimation of the current distributive structure based on meritocratic ideals.
-
-# Transparency statement
-
-This research follows a series of open science guidelines as the pre-registration of Study 2 (scale development) as well as the availability of data and codes (R) in an public repository in Github.
-This document was generated with the R package Rmarkdown and contains the code for all the tables and figures in order to make it reproducible.
-
-# References
-
-<div id="refs"> </div>  <!-- Although <div> is an HTML tag, this method also works for other output formats such as PDF. -->
-
-\pagebreak
-
-<!-- # Appendix {-} -->
-
-`r if (knitr::is_latex_output()){ '\\appendix'}`
-
-`r if (knitr::is_latex_output()){ '\\section{Appendix}'}`
-
-**Study 1**
-
-```{r countries, echo=FALSE}
-  load(file = "input/data/proc/data_s1.RData")
-  if (!require("pacman")) install.packages("pacman")
-  pacman::p_load(dplyr, kableExtra, knitr,lavaan,countrycode,ggplot2,haven,
-                 gridExtra)
-  table_format = if(knitr::is_html_output()) { #conditional instructions for kable
-    "html" #if html set "html" in format
-  } else if(knitr::is_latex_output()) {
-    "latex"#if latex set "latex" in format
-  }
-  load(file = "input/data/proc/data_s1.RData")
-  if (!require("pacman")) install.packages("pacman")
-  pacman::p_load(dplyr, kableExtra, knitr,lavaan,countrycode,ggplot2,haven,
-                 gridExtra)
-  table_format = if(knitr::is_html_output()) { #conditional instructions for kable
-    "html" #if html set "html" in format
-  } else if(knitr::is_latex_output()) {
-    "latex"#if latex set "latex" in format
-  }
- # meassurement model: Second order
-  cfa_perpref2 <- '
-  	  # latent variables
-  	  percmerit =~ hwork + ambition
-  	  parents =~ wealthy + pareduc
-  	  backgrd =~ race + gender
-  	  networks=~ people + polcone
-  	  percnmerit=~ parents + backgrd + networks
-  	  prefmerit =~ welljob + hardjob
-  	  prefnmerit=~ family + child
-  	  '
-  # Model fit: second order
-
-  fit_perpref2 <- lavaan::cfa(cfa_perpref2, data = data_s1,ordered = F)
-
-  df_proc<- data_s1 %>% select(id,v5,wealthy:hardjob)  
-  df_proc$ccode <- as.numeric(data_s1$v5)
-
-  predicted2 <- lavPredict(fit_perpref2,newdata = data_s1)
-
-  idx <- lavInspect(fit_perpref2, "case.idx")
-  fscores <- lavPredict(fit_perpref2)
-  ## loop over factors
-  for (fs in colnames(fscores)) {
-    df_proc[idx, fs] <- fscores[ , fs]
-  }
-
-  df_proc <-  df_proc %>%  select(id,ccode,percmerit,prefmerit,prefnmerit) %>% na.omit()
-
-  df_proc$clab <- countrycode(sourcevar = df_proc$ccode,
-                            origin = "iso3n",
-                            destination = "country.name")
-avg_count <-  
-  df_proc %>% group_by(clab) %>% summarise(N=n())
-avg_count <- arrange(avg_count)
-
-# sum(avg_count[, 'N'])
-  col.nam <- c("Country","N", "Country", "N")
-  footnote <- paste0("N = ", sum(avg_count[, 'N']),". Source: International Social Survey Programme (2009).")
-
-  knitr::kable(bind_cols(avg_count[1:20,],avg_count[21:40,]),
-               col.names = col.nam, align = c("l","l","l","l"),linesep = "",
-        booktabs=TRUE,format = table_format,escape = FALSE,
-        caption = "Sample of Study 1 ") %>%
-    kableExtra::kable_styling(full_width = FALSE,
-                              latex_options = "HOLD_position",
-                              bootstrap_options=c("striped", "bordered"),
-                              font_size = 10) %>%
-    kableExtra::column_spec(column = c(2,4),width = "1cm") %>% 
-    kableExtra::column_spec(column = c(1,3),width = "2.5cm") %>% 
-    column_spec(3, border_left = T) %>% 
-    kableExtra::footnote(general = footnote, footnote_as_chunk = T,threeparttable = T)
-```
-
-\pagebreak
-
-**Study 2**
-
-|                     | Sample  |   CEP  |
-|---------------------|:-------:|:------:|
-| __Gender__          |         |        |
-| Men                 | 49,82%  | 50,52% |
-| Women               | 50.18%  | 49,47% |
-| __Age__             |         |        |
-| 18 - 24             | 18,55%  | 18,17% |
-| 25 - 34             | 18,86%  | 17,48% |
-| 35 - 44             | 19.09%  | 19,98% |
-| 45 - 54             | 17,96%  | 19,23% |
-| 55 - or more        | 25,54%  | 25.11% |
-| __Education__       |         |        |
-| Primary  or less    | 2,93%   | 15,88% |
-| High school         | 43,23%  | 37,04% |
-| Non university      | 32,63%  | 28,93% |
-| university or more  | 21,21%  | 18,13% |
-Table: `r as.character(paste("(\\#tab:rep-samp)", "Representativeness of the study 2 sample"))`.
-
-```{r tb-perpref1, eval=FALSE, include=FALSE}
- if (!require("pacman")) install.packages("pacman")
- load(file = "input/data/proc/data_s1.RData")
- pacman::p_load(dplyr,knitr,kableExtra)
- options(knitr.kable.NA = '')
- table_format = if(knitr::is_html_output()) { #conditional instructions for kable
-   "html" #if html set "html" in format
- } else if(knitr::is_latex_output()) {
-   "latex"#if latex set "latex" in format
- }
-
- # First order model for Appendix
- cfa_perpref1 <- '
- 	  # latent variables
- 	  percmerit =~ hwork + ambition
- 	  percnmerit=~ wealthy + pareduc +race + gender +people + polcone
- 	  prefmerit =~ welljob + hardjob
- 	  prefnmerit=~ family + child
- 	  '
- # Fit First order model for Appendix
- fit_perpref1 <- lavaan::cfa(cfa_perpref1, data = data_s1,
-                             ordered = c("wealthy","pareduc","ambition","hwork",
-                                         "people","polcone","race","gender",
-                                         "respons","yeduc",
-                                         "family","child",
-                                         "welljob","hardjob"))
- # Labels of First order model for Appendix
- labs <- c(
-   "Importance: hard work",
-   "Importance: having ambition",
-   "Importance: wealthy family",
-   "Importance: educated parents",
-   "Importance: race",
-   "Importance: gender",
-   "Importance: knowing people",
-   "Importance: political connections",
-   "Reasons for pay: well job",
-   "Reasons for pay: hard job",
-   "Reasons for pay: support family",
-   "Reasons for pay: has children"
- )
- #Extract factor loadings
- tb.load<- data.frame(round(cbind(lavaan::inspect(fit_perpref1,
-                                                         what="std")$lambda),
-                            digits = 3))
- tb.load[tb.load==c(0.00)] <- NA #set to NA factor loadings of 0
-
- for (i in names(tb.load)) {
-   tb.load[,i] <- sprintf(tb.load[,i], fmt = '%#.3f') #set values as character
- }
- tb.load[tb.load=="NA"] <- "" #set NA values to "" (empty space)
- #---extract fit measures-------#
- fm01<- data.frame(t(data.frame(lavaan::fitmeasures(fit_perpref1, output ="matrix")[c("chisq","df","cfi","tli","rmsea"),])));  row.names(fm01) ="percnmerit"
-
- #------chi2, df------#
- fm04<- round(rbind(fm01),3) #select and round to 4 digits
- fm04.1 <- fm04 %>% dplyr::select(chisq,df) #select chi2 and degrees of freedom
- fm04.1$chisq <- round(x = fm04.1$chisq,digits = 1) # round Chi2 to 1 digit
- fm04.1$df <- round(x = fm04.1$df,digits = 0) #round Df to 0 digits
- fm04.1$chi2df <- paste0(fm04.1$chisq,"(",fm04.1$df,")") #Paste chi2(df)
- fm04.1 <- dplyr::select(fm04.1,"chi2df") #select final column
- for (i in names(fm04.1)) {
-   fm04.1[,i] <- as.character(fm04.1[,i]) # transform all to character
- }
-
- #------CFI, RMSEA------#
- fm04.2 <- fm04 %>% dplyr::select(cfi,tli,rmsea) # select CFI, TLI and RMSEA
- for (i in names(fm04.2)) {                      # transform all to character
-   fm04.2[,i] <- sprintf(fm04.2[,i], fmt = '%#.3f')
- }
-
- fm.df      <- dplyr::bind_cols(fm04.1,fm04.2) # factor loading and labels
- fm.df$nobs <- c(lavaan::nobs(fit_perpref1))   # add number of observations
- fm.df <- data.frame(t(fm.df)); colnames(fm.df) <- c("percmerit") #transpose and set colnames
-
- #------ merge ------#
- tb.fm<- dplyr::bind_rows(tb.load,fm.df) #merge factor loadings and fit indices
- tb.fm<- tb.fm %>%
-   dplyr::mutate(vars=c(labs,"$\\chi^2\\text{(df)}$","$\\text{CFI}$",
-                        "$\\text{TLI}$","$\\text{RMSEA}$","$N$")) %>%
-   dplyr::select(vars,everything()) #Change names of Fit meassures to latex symbols
- tb.perpref1 <- tb.fm #final matrix for Table
-
- # Table --------------------------------------------------------------------.
- tb.foot <- paste0("Standardized factor loadings using DWLS estimator ; CFI = Comparative fit index ;TLI = Tucker-Lewis index; RMSEA =  Root mean square error of approximation")
- tb.col <- c("","Meritocratic","Non-meritocratic",
-             "Meritocratic","Non-meritocratic")
- tb.caption <- c(" Factor loadings and fit measures of the four first-order factor model in study 1")
- tb.issp<- kable(tb.perpref1,escape = FALSE,digits = 3,
-                 align = "lcccc",
-                 col.names = tb.col,
-                 caption = tb.caption,
-                 booktabs = T,
-                 linesep = "",
-                 format = table_format,
-                 row.names=F) %>%
-           kable_styling(latex_options = c("hold_position"),
-                         bootstrap_options = c("striped", "bordered"),
-                         position = "center",
-                         font_size = 10) %>%
-           add_header_above(header = c(" "=1,"Perception"= 2,"Preference"= 2)) %>%
-           add_header_above(header = c(" "=1,"Factor loadings"= 4)) %>%
-           row_spec(row = 12,hline_after = TRUE) %>%
-           add_indent(c(13:17)) %>%
-           footnote(general =tb.foot ,footnote_as_chunk = T,threeparttable = T);tb.issp
-```
-
-<!-- **Study 3** -->
-
-```{r fcloads-inv, eval=FALSE, include=FALSE}
- if (!require("pacman")) install.packages("pacman")
- pacman::p_load(dplyr, knitr, kableExtra,lavaan)
- options(knitr.kable.NA = '')
- table_format = if(knitr::is_html_output()) { #conditional instructions for kable
-   "html" #if html set "html" in format
- } else if(knitr::is_latex_output()) {
-   "latex"#if latex set "latex" in format
- }
- table_format2 = if(knitr::is_html_output()) {
-   T
- } else if(knitr::is_latex_output()) {
-   F
- }
- # Note:
- # This CFA corresponds to the sample from the Factorial Survey Study
- load(file = "input/data/proc/data_s3_inv.RData") #load data study 3
- fs01  <- data_s3_inv %>% dplyr::filter(dataset=="fsw02") #subset sample
-
- # Set model
- model01 <- '
- perc_merit  =~ perc_effort + perc_talent
- perc_nmerit =~ perc_wpart  + perc_netw
- pref_merit  =~ pref_effort + pref_talent
- pref_nmerit =~ pref_wpart  + pref_netw'
-
- # Model fit
- fs1_o <- lavaan::cfa(model = model01,
-              data = fs01,
-              ordered = c("perc_effort","perc_talent",
-                          "perc_wpart","perc_netw",
-                          "pref_effort","pref_talent",
-                          "pref_wpart","pref_netw"),std.lv=FALSE)
-
- # Extract information for table
- fm01<- data.frame(t(data.frame(lavaan::fitmeasures(fs1_o, output ="matrix")[c("chisq","df","cfi","tli","rmsea"),])));  row.names(fm01) ="M1.1"
-
- #------chi2, df------#
- fm04.1 <- fm01 %>% dplyr::select(chisq,df)
- fm04.1$chisq <- round(x = fm04.1$chisq,digits = 1)
- fm04.1$df <- round(x = fm04.1$df,digits = 0)
- fm04.1$chi2df <- paste0(fm04.1$chisq,"(",fm04.1$df,")")
- fm04.1 <- dplyr::select(fm04.1,"chi2df")
-
- #------CFI, RMSEA------#
- fm04.2 <- fm01 %>% dplyr::select(cfi,tli,rmsea)
- for (i in names(fm04.2)) {
-   fm04.2[,i] <- sprintf(fm04.2[,i], fmt = '%#.3f')
- }
-
- #------ merge fit indices ------#
- fm.df      <- dplyr::bind_cols(fm04.1,fm04.2)
- fm.df$nobs <- c(lavaan::nobs(fs1_o))
- fm.df      <- data.frame(t(fm.df))
-
- #------ extract factor loadings ------#
- loads01<- data.frame(round(inspect(fs1_o,what="std")$lambda,digits = 3))
- loads01[loads01==c(0.00)] <- NA; names(loads01) <- c(paste0("M1.",1:4))
-
- for (i in names(loads01)) {
-   loads01[,i] <- sprintf(loads01[,i], fmt = '%#.3f')
- }
- loads01[loads01=="NA"] <- ""
-
- #------ labels for table ------#
- labs <-c(
-     "A. Perception Effort",
-     "B. Perception Talent",
-     "C. Perception Rich parents",
-     "D. Perception Contacts",
-     "E. Preferences Effort",
-     "F. Preferences Talent",
-     "G. Preferences Rich parents",
-     "H. Preferences Contacts")
-
- #------ merge ------#
- tb.fm2<- dplyr::bind_rows(loads01,fm.df)
- tb.fm2<- tb.fm2 %>%
-   dplyr::mutate(vars=c(labs,"$\\chi^2\\text{(df)}$","$\\text{CFI}$","$\\text{TLI}$","$\\text{RMSEA}$","$N$")) %>%
-   dplyr::select(vars,everything())
-
- tb.col  <- c(" ",
-              "Meritocratic","Non-meritocratic",
-              "Meritocratic","Non-meritocratic")
- tb.caption <- "Factor loadings and fit measures for the second sample for study 3"
- tb.foot <- paste0("Standardized factor loadings using DWLS estimator ; CFI = Comparative fit index ;TLI = Tucker-Lewis index; RMSEA =  Root mean square error of approximation")
- tb.fm02<- knitr::kable(tb.fm2,escape = FALSE,
-                        align = "lcccc",col.names = tb.col,
-                        caption = tb.caption,
-                        booktabs = T,linesep = "",
-                        format = table_format,
-                        row.names = F) %>%
-           kableExtra::kable_styling(full_width = table_format2,
-                         latex_options = c("hold_position"),
-                         position = "center",
-                         font_size = 10,
-                         bootstrap_options = c("striped", "bordered")) %>%
-           kableExtra::add_header_above(header = c(" "=1,"Perception"= 2,"Preference"= 2)) %>%
-           kableExtra::add_header_above(header = c(" "=1,"Factor loadings"= 4)) %>%
-           kableExtra::row_spec(row = 8,hline_after = TRUE) %>%
-           kableExtra::add_indent(c(9:13)) %>%
-           kableExtra::footnote(general =tb.foot ,footnote_as_chunk = T,threeparttable = T);tb.fm02
-```

### 2023-02-02 paper-blinded.Rmd
grammarly vscode check

@@
 ---
-title: "Measur---
 title: "Measuring Perceptions and Preferences for Meritocracy"
@@
 Most of the studies in the area have relied upon secondary data from different surveys to operationalize meritocracy, with a wide variation in the use and interpretation of the same measuring items.
-Based on a critical analysis of concepts and operationalizations of meritocracy in survey research, we propose a conceptual framework for measuring perceptions and preferences for meritocracy and non-meritocracy in a minimalistic version (few items as possible), in order to be considered in comparative survey research studies.
+Based on a critical analysis of concepts and operationalizations of meritocracy in survey research, we propose a conceptual and measurement framework for perceptions and preferences for meritocracy and non-meritocracy in a minimalistic version (few items as possible), in order to be considered in comparative survey research studies.
 The proposal is first tested through confirmatory factor analysis (CFA) using available ISSP (International Social Survey Programme) inequality module items.
-Based on these results, in a second stage we propose  an  improved scale that better suits the perceptions/preferences distinction, which was applied in a Chilean sample (N=2,141).
+Based on these results, in a second study we propose an  improved scale that better suits the perceptions/preferences distinction and that is tested  in a Chilean sample (N=2,141).
 The results of the CFA estimation give evidence of a better fit for the new scale.
-Our discussion highlights the importance of considering different dimensions in order to advance in the study of meritocracy."
+Our discussion highlights the importance of considering different dimensions in the study of meritocracy, both conceptually and empirically."
 output:
@@
 In many cases, similar phenomena are associated with different indicators, and conversely, different phenomena are measured with similar indicators.
-Such trends limit the comparability of studies and the ability to understand the effects of meritocratic beliefs across different disciplines and lines of research.
+Such approaches limit the comparability of studies and the capacity to understand the effects of meritocratic beliefs across different disciplines and lines of research.
 
 Based on a critical analysis of different approaches to the measurement of meritocracy to date, this article proposes a conceptual and measurement framework that considers two of the main aspects in the study of meritocracy: perceptions and preferences.
-Furthermore, it distinguishes between meritocratic and non-meritocratic dimensions, as they are not necessarily two poles of the same continuum as previous studies suggest.
+Furthermore, it distinguishes between meritocratic and non-meritocratic dimensions, as they would not be two poles of the same continuum as previous studies suggest.
 This conceptual framework is oriented to develop a measurement instrument as simple and brief as possible, suitable for public opinion surveys and thus allowing to integrate meritocratic beliefs in studying different social phenomena.
@@
 
-In this conceptual and measurement approach of meritocratic beliefs, we can observe a couple of strong conceptual assumptions: a) effort would not depend on parental influence, and b) talent (as innate ability) is not meritocratic (contrary to Michael Youngs' original conceptualization).
-This conceptual and measurement-based assumption is found in other studies, which also assume effort to be the main and only aspect of meritocracy [@girerd_neoliberalism_2020; @bubak_perceptions_2019], raising the question: Is effort the only dimension behind the concept of meritocracy?
-Whether talent (as intelligence and ability) is or is not considered meritocratic is an interesting topic to discuss from a philosophical point of view, but it also requires empirical evidence.
-Furthermore, considering talent as part of meritocracy certainly opens some interesting avenues of research, as for instance some studies suggest that for the elites meritocracy is related to talent, whereas effort is more characteristic of the meritocracy of the middle and lower classes [@atria_economic_2020].
-Therefore, we propose that a concept (and measure) of meritocracy should include both talent and effort as in its original conception.
+In this conceptual and measurement approach of meritocratic beliefs, we can observe a couple of strong conceptual assumptions: a) effort would not depend on parental influence (at least not as strongly as intelligence), and b) talent (as innate ability) is not meritocratic (contrary to Michael Youngs' original conceptualization).
+This conceptual and measurement-based assumption is found in other studies, which also assume effort to be the main and only aspect of meritocracy [@girerd_neoliberalism_2020; @bubak_perceptions_2019], raising the question: Is effort the only (or main) dimension behind the concept of meritocracy?
+We argue that whether talent (as intelligence and ability) is or is not considered meritocratic is an interesting topic to discuss from a philosophical point of view, but it also requires empirical evidence in studies focused on the subjective aspects of inequality.
+Furthermore, considering talent as part of meritocracy certainly opens some interesting avenues of research, for instance, some studies suggest that for the elites meritocracy is related to talent, whereas effort is more characteristic of the meritocracy of the middle and lower classes [@atria_economic_2020].
+Therefore, we propose that a concept (and measure) of meritocracy benefits from including both talent and effort as in the original Young's conception.
 
@@
 To illustrate this point let us consider the commonly used "reasons to get ahead" battery, mentioned above.
-Some versions of this battery ask "how important you think it is for getting ahead in life" and then lists several factors, whereas another version of this same battery - sometimes presented along the previous one - asks about "how important you think it _should_ be ...", and then listing the same concepts.
+Some versions of this battery ask "how important you think it is for getting ahead in life" and then list several factors, whereas another version of this same battery - sometimes presented along the previous one - asks about "how important you think it _should_ be ...", and then listing the same concepts.
 Therefore, the question raised here is: Which one of both is a "belief": what _is_ (the first one) or what it _should_ be (the second one)?
@@
 Therefore, beliefs use to cover almost anything related to subjective factors.
-To this regard, a relevant distinction in the field of inequality beliefs was made by @janmaat_subjective_2013: "Perceptions refer to subjective estimates of existing inequality (i.e.
-thoughts about what is).
-Beliefs are here defined as normative ideas about just inequality (i.e.
-thoughts about what should be)"(p.359).
+In this regard, a relevant distinction in the field of inequality beliefs was made by @janmaat_subjective_2013: "Perceptions refer to subjective estimates of existing inequality (i.e., thoughts about what is).
+Beliefs are here defined as normative ideas about just inequality (i.e., thoughts about what should be)"(p.359).
 Several papers dealing with meritocracy use the term beliefs (i.e.
-what should be), while referring to perceptions (i.e.
-what is).
+what should be), while referring to perceptions (i.e., what is).
 For instance,  in @reynolds_perceptions_2014, the term belief is used to talk about what @janmaat_subjective_2013 refers to as perceptions, whereas other authors use general terms such as attitudes [@kunovich_systems_2007].
@@
 reverse coded).
-The assumption behind this approach is that rejecting a supposed ascribed or non-meritocratic aspect (as having family and/or children) implies a stronger belief in meritocracy.
+The assumption behind this approach is that rejecting a supposed ascribed or non-meritocratic aspect (such as having family and/or children) implies a stronger belief in meritocracy.
 A similar approach of reverse-coding non-meritocratic items was taken by @newman_false_2015, using the same principles applied in the "Preference for the Merit Principle Scale" [@davey_preference_1999].
@@
 They considered education, ambition, and hard work as meritocratic, whereas factors such as family wealth and connections were classified as non-meritocratic.
-Despite making this distinction, the authors end up subtracting one dimension from the other, thus coming back to the assumption that they are two poles of the same continuum, as @kunovich_systems_2007 also did.
+Despite making this distinction, in the operationalization the authors end up subtracting one dimension from the other, thus coming back to the assumption that they are two poles of the same continuum, as @kunovich_systems_2007 also did.
 Similarly, @roex_attitudes_2018 used ISSP indicators for perceived meritocracy and non-meritocracy to build a single score by reverse coding the non-meritocratic items.
 Therefore, the treatment of non-meritocratic items has been rather inconsistent across studies and the assumption that they are the simple opposite of meritocracy certainly requires further conceptual elaboration and empirical assessment.
-For instance, such distinction could open the door to study to what extent different and apparently contradictory views of meritocracy can coexist, to what extent individuals and societies differ to this regard, and of course the associated factors to such differences.
-Although the use of the concept "non-meritocratic" is very broad, we prefer to use it instead of something like ascription, given that this last one leaves out aspects such as the use of personal relationships.
+For instance, a such distinction could open the door to studying to what extent different and contradictory views of meritocracy can coexist, to what extent individuals and societies differ in this regard, and of course the associated factors to such differences.
+Although the use of the concept "non-meritocratic" is very broad, we prefer to use it instead of other terms like ascription, given that this last one leaves out aspects such as the use of personal relationships as a way to achieve success.
 
@@
 There is a relevant trade-off regarding measurement error estimation in survey scales.
-Including multiple items per conceptual dimension opens the possibility to validity studies (at least in terms of construct) by using factor analysis techniques, particularly confirmatory ones.
-However, increasing the amount of items reduces the probability of usability of the scale in survey research given space limitations.
-This validity-usability conundrum is not easy to solve and, as we will show later in the methods section, we opted by a minimalistic approach oriented to the usability of the scale but that still gives the possibility of confirmatory factor analysis, although with some limitations.
+Including multiple items per conceptual dimension opens the possibility of validity studies (at least in terms of the construct) by using factor analysis techniques, particularly confirmatory ones.
+However, increasing the number of items reduces the probability of usability of the scale in survey research given space limitations.
+This validity-usability conundrum is not easy to solve and, as we will show later in the methods section, we opted for a minimalistic approach oriented to the usability of the scale but that still gives the possibility of confirmatory factor analysis, although with some limitations.
 
@@
 The rows in Figure \@ref(fig:merit-model) consider the distinction between meritocratic and non-meritocratic dimensions [@reynolds_perceptions_2014], usually treated as different ends of the same continuum in previous research.
-Non-meritocratic elements refer to the use of resources such as personal contacts and/or family advantages to getting ahead in life.
+Non-meritocratic elements refer to the use of resources such as personal contacts and/or family advantages to get ahead in life.
 
@@
 There were two items from the battery that were excluded from the analysis as they would not fit into the classification.
-The first one was "having good education yourself", since it was not clear whether this is could be due to individual motivation or system opportunities, and the second was "giving bribes", as introduced elements of criminality that were beyond a non-meritocratic perception.
+The first one was "having good education yourself" since it was not clear whether this is could be due to individual motivation or system opportunities, and the second was "giving bribes", as introduced elements of criminality that were beyond a non-meritocratic perception.
 
@@
 Firstly, we observe in general that the moderate to high associations are between the pairs of items representing each of the dimensions as described in \@ref(tab:table-issp).
-The exception here is the dimension of non-meritocratic perception, in which there are six items that appear mostly associated by pairs according to their specific topics (family background, personal background, and connections).
+The exception here is the dimension of non-meritocratic perception, in which six items appear mostly associated by pairs according to their specific topics (family background, personal background, and connections).
 Still, between this same set of items, moderate correlations (around 0.3) could anticipate a single latent factor underlying non-meritocratic perception, which is tested next through confirmatory factor analysis.
@@
 Table \@ref(tab:sum-fit-issp) shows the results of the estimation of two confirmatory models.
-The first one (First Order) corresponds to a model that estimate four factors, each for one of the dimensions presented in \@ref(tab:table-issp), with only regular fit indicators (CFI=0.959, TLI=0.944, RMSEA=0.098, $\chi^2$(df=48)= 21308.535).
+The first one (First Order) corresponds to a model that estimates four factors, each for one of the dimensions presented in \@ref(tab:table-issp), with only regular fit indicators (CFI=0.959, TLI=0.944, RMSEA=0.098, $\chi^2$(df=48)= 21308.535).
 Further analysis showed that the sources for poor fit were mostly related to the items of the non-meritocratic perceptions, which as mentioned above displayed correlations among them not taken into account in this model specification.
@@
 
-Regarding the correlations between the factors in the Figure \@ref(fig:meas02-issp), we observe that perceptions are correlated with preferences, but more strongly for the meritocratic ($r=0.46$) than for the non-meritocratic ($r=0.27$) dimensions.
+Regarding the correlations between the factors in Figure ref(fig:meas02-issp), we observe that perceptions are correlated with preferences, but more strongly for the meritocratic ($r=0.46$) than for the non-meritocratic ($r=0.27$) dimensions.
 Secondly, both perceptual dimensions (meritocratic and non-meritocratic) depict a moderate correlation ($r=0.30$), suggesting that they are not the opposite poles of the same continuum.
@@
 
-## Discussion study 1
+## Discussion Study 1
 
@@
 First of all, these items were not specifically designed based on a conceptual approach to meritocracy, but it still is possible to find a certain fit of the model to the data after minor adjustments.
-Secondly, some of the items in similar dimensions refer to different concepts, as for instance meritocratic perceptions are related to ambition and hard work, whereas meritocratic preferences deal with how well a job is done.
+Secondly, some of the items in similar dimensions refer to different concepts, for instance, meritocratic perceptions are related to ambition and hard work, whereas meritocratic preferences deal with how well a job is done.
 A better operationalization would require more alignment between the different sides of the same concept.
@@
 
-The data comes from an online survey which was part of a larger study on meritocracy and preferences developed in Chile in 2019, funded by the Chilean National Scientific Agency (ANID).
+The data comes from an online survey that was part of a larger study on meritocracy and preferences developed in Chile in 2019, funded by the Chilean National Scientific Agency (ANID).
 The questionnaire was programmed in Qualtrics and the fieldwork was conducted by an external online survey agency ([netquest.com](www.netquest.com)) between December 2019 and January 2020.
@@
 There were no significant differences between our sample and the wider population for most socio-demographic characteristics, except for the educational level (see Table \@ref(tab:rep-samp) in Appendix).
-As it is often the case with online surveys, there were some limitations in achieving the quotas for lower educational levels [@zhang_quota_2018; @boas_recruiting_2020].
+As is often the case with online surveys, there were some limitations in achieving the quotas for lower educational levels [@zhang_quota_2018; @boas_recruiting_2020].
 This survey was designed as a three-wave panel and the data used in this analysis corresponds to the first wave.
@@
 The first one was usability in social survey research, for which the scale length must attend spacer restrictions in such studies.
-Nevertheless, at the same time the scale should offer possibilities of accounting for measurement error through factor analysis techniques.
+Nevertheless, at the same time, the scale should offer possibilities of accounting for measurement error through factor analysis techniques.
 Although a rule of thumb in latent measurement suggests at least three indicators per dimension, we still opted for two to count with a brief scale.
-Although such decision posits some limitation for modeling measurement error, still allows for factor analysis and is a common approach in previous social survey studies with limited items (for instance see @davidov_measurement_2009).
-
-The second criteria is to expand the possibilities of comparability with previous studies on meritocracy, by considering some of the items previously used in this research area.
+Although such a decision posits some limitations for modeling measurement error, still allows for factor analysis and is a common approach in previous social survey studies with limited items (for instance see @davidov_measurement_2009).
+The second criterion is to expand the possibilities of comparability with previous studies on meritocracy, by considering some of the items previously used in this research area.
 To this regard, some of the items were adapted from the items battery "reasons to get ahead" (ISSP/GSS), which has been widely used for operationalizing meritocracy in international research [@mijs_paradox_2019; @duru-bellat_who_2012; @reynolds_perceptions_2014].
@@
 
-With the objective of evaluating the effect of indicator ordering in the responses, three different versions of items' order were designed and randomly assigned to the respondents, as depicted in Figure \@ref(fig:appmod).
+To evaluate the effect of indicator ordering in the responses, three different versions of items' order were designed and randomly assigned to the respondents, as depicted in Figure \@ref(fig:appmod).
 The scale was presented to the first group (_n = 712_) in the order shown in Table \@ref(tab:table-indicadores) according to perceptions and preferences.
-For the second group (_n = 717_), the order was reorganized according to perceptions and preferences over the same topic, e.g., for the topic of hard work, the item about perception was followed by the item about preference, and the same for the rest of the topics.
+For the second group (_n = 717_), the order was reorganized according to perceptions and preferences over the same topic, e.g., for the topic of hard work, the item about perception was followed by the item about preference and the same for the rest of the topics.
 Finally, for the third group (_n = 712_), the items were presented as completely randomized.
@@
 
-## Discussion study 2
+## Discussion Study 2
 
 The estimation performed with different administration sets for the meritocracy scale gives evidence in line with the conceptual model.
-The four dimensions appear independent but related, with some differences when compared with the estimation with ISSP data in Study 1.
+The four dimensions appear independent but related, with some differences when compared to the estimation with ISSP data in Study 1.
 The main general difference is that with ISSP data, the four latent dimensions were positively correlated, whereas now some correlations are statistically not different from zero.
@@
 Although the chi-square statistic is used as a global measure of fit, it is usually complemented with other indexes given its high sensitivity to sample size: the Comparative Fit Index (CFI), which should have a value greater than 0.95; the Root Mean Square Error Approximation (RMSEA), which must be lower than 0.06; and the Standardized Root Mean Square Residuals (SRMR) which must be less than 0.08 [@vandeschoot_checklist_2012].
-Besides, the literature on measurement invariance suggests some complementary approaches for evaluamerit scale the fit, from which we will consider the incremental adjustment of the fit indexes [@cheung_evaluating_2002; @milfont_testing_2010; @dimitrov_testing_2010] and the ANOVA test for means comparison in nested models [@newsom_longitudinal_2015].
+Besides, the literature on measurement invariance suggests some complementary approaches for evaluating the fit, from which we will consider the incremental adjustment of the fit indexes [@cheung_evaluating_2002; @milfont_testing_2010; @dimitrov_testing_2010] and the ANOVA test for means comparison in nested models [@newsom_longitudinal_2015].
 
@@
 Nevertheless, the criteria of $\Delta \text{CFI}$ used for comparing models is close to the rejection criteria of >.01, whereas the $\Delta \text{RMSEA}$ fulfills the requirements of being below the cut-off criteria as suggested by [@chen_sensitivity_2007].
-Therefore, using this last standard, the level for strict invariance would hold for the meritocracy scale.
+Therefore, using this last standard, the level of strict invariance would hold for the meritocracy scale.
 
@@
 Based on the measurement invariance analysis for two datasets, we found evidence supporting the equivalence of the scale when applied in different samples.
-This gives a more robust ground when using this measure for comparing groups, as for instance in country comparisons.
+This gives a more robust ground when using this measure for comparing groups, for instance in country comparisons.
 Although the results are promising, still the analysis was performed in two samples in a single country (Chile) and it requires further examination with social surveys in other societies.
@@
 As the existence of heterogeneous approaches certainly has consequences for the advancement of the study of meritocracy, this paper presented a comprehensive conceptual framework for the empirical study of meritocracy, building upon previous research.
-This framework was operacionalized in a survey scale and tested against available and new survey data.
+This framework was operationalized in a survey scale and tested against available and new survey data.
 
 We identified four critical aspects regarding the measurement of meritocracy in previous studies: unidimensionality, the ambiguous use of the term "beliefs", the use of non-meritocratic indicators as opposed to meritocratic indicators, and the consideration of measurement error.
-The proposed 8-item scale, "Perceptions and Preferences for Meritocracy", was designed and tested in order to deal with these four issues.
+The proposed 8-item scale, "Perceptions and Preferences for Meritocracy", was designed and tested to deal with these four issues.
 Our results indicate that perceptions and preferences are two related but different dimensions, often confounded in the previous research under the label of "beliefs".
@@
 One of the criteria in designing the merit scale was to develop a short instrument that could be used in social surveys, assuming that using only two items per dimension posits limitations for the confirmatory analysis estimated here.
-We preferred this confirmatory-minimalistic approach at it offers several advantages in terms of information about the quality of the measurement, and we offered additional validity evidence with measurement invariance analysis.
-Despite the advantages, a more simple and straightforward perspective is always available, as for instance using only the average of the items per dimension.
-Whereas a short scale it is always appreciated in social survey studies, for other audiences could not reach measurement standards.
+We preferred this confirmatory-minimalistic approach as it offers several advantages in terms of information about the quality of the measurement, and we offered additional validity evidence with measurement invariance analysis.
+Despite the advantages, a more straightforward perspective is always available, for instance using only the average of the items per dimension.
+Whereas a short scale is always appreciated in social survey studies, for other audiences could not reach measurement standards.
 This may require further expansions and/or consider using a different instrument as the "Preference for the Merit Principle Scale" [@davey_preference_1999] (which measures only one dimension of meritocracy with fifteen items).
@@
 Furthermore, the impact of different configurations of the four-dimensional framework on practices and behaviors, such as corruption, civic involvement, and political alignment, are areas open to further research.
-Such future agendas could be especially relevant in times of economic crisis and growing inequalities, that could entail changes in the legitimation of the current distributive structure based on meritocratic ideals.
+Such future agendas could be especially relevant in times of economic crisis and growing inequalities, which could entail changes in the legitimation of the current distributive structure based on meritocratic ideals.
 
@@
 This research follows a series of open science guidelines as the pre-registration of Study 2 (scale development) as well as the availability of data and codes (R) in a public repository in Github.
-This document was generated with the R package Rmarkdown and contains the code for all the tables and figures in order to make it reproducible.
+This document was generated with the R package Rmarkdown and contains the code for all the tables and figures to make it reproducible.
 

### 2023-02-02 Title Page.Rmd
ajustes layout

---
title: "Measuring Perceptions and Preferences for Meritocracy"
linestretch: '1.5'
output:
  bookdown::pdf_document2:
    template: null
    toc: false
    number_sections: false
geometry: margin=0.78in
header-includes:
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage[figurename=Fig.]{caption}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
fontsize: 12pt

---

Juan Carlos Castillo, Department of Sociology, Universidad de Chile (
juancastillov@uchile.cl - [jc-castillo.com](https://juancarloscastillo.github.io/jc-castillo/) -  Ignacio Carrera Pinto 1045 Ñuñoa, Santiago de Chile )

Luis Maldonado, Institute of Sociology, Pontificia Universidad Católica de Chile

Jorge Atria, Department of Sociology, Universidad Diego Portales

Julio Iturra, Bremen International Graduate School of Social Sciences

Francisco Meneses, Department of Sociology, Universidad de Chile

### 2023-02-02 paper-blinded.Rmd
ajustes layout

@@
 css: "custom.css"
-linestretch: '1.5'
+linestretch: '2.0'
 link-citations: yes
 # author: "Juan Carlos Castillo, Julio Iturra, Luis Maldonado, Francisco Meneses & Jorge Atria <br><br>"
-abstract: "Economic and social inequalities have generated growing concerns and crises across contemporary societies.
+# abstract: "Economic and social inequalities have generated growing concerns and crises across contemporary societies.
 The belief in meritocracy has been suggested as one of the factors that explain the persistence of inequality, as it would legitimize economic disparities based on differences in effort and talent.
@@
      number_sections: false
-  bookdown::word_document2:
-    number_sections: false
-  bookdown::html_document2:
-    number_sections: false
+     keep_text: true
+  # bookdown::word_document2:
+  #   number_sections: false
+  # bookdown::html_document2:
+  #   number_sections: false
 linkcolor: blue
@@
   - input/bib/meritocracy.bib
-csl: input/bib/apa7-singlespace.csl
+csl: input/bib/asa6.csl
 editor_options:
   chunk_output_type: console
-geometry: margin=0.78in
+geometry: margin=1in
 header-includes:
@@
   - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
-  - \usepackage[figurename=Fig.]{caption}
+#  - \usepackage[figurename=Fig.]{caption}
   - \usepackage{graphicx}
@@
 always_allow_html: true # for word output
-fontsize: 10pt
+fontsize: 12pt
+papersize: a4
 ---
@@
 ```{r meas02-issp, echo=FALSE, fig.cap = "Descriptive plot", fig.align='center',fig.cap="Second-order confirmatory factor analysis model using ISSP indicators of Perceptions and Preferences for Meritocracy"}
- knitr::include_graphics('output/images/Fig4.cfa_issp_study1.PNG')
+ knitr::include_graphics('output/images/figure4.PNG')
 ```
@@
 ```{r appmod, echo=FALSE, fig.cap = "Survey flow", fig.align='center'}
- knitr::include_graphics('output/images/Fig5.app_mod.png')
+ knitr::include_graphics('output/images/figure6.png')
 ```
@@
 
-A pre-registration was made in the OSF platform, available at the following link: [https://osf.io/z45y2](https://osf.io/z45y2).
+A pre-registration was made in the OSF platform, available at the following link: [https://osf.io/z45y2?](https://osf.io/z45y2/?view_only=adcd496887f2471597b1d6ffb46e492d).
 This pre-registration includes the hypotheses regarding the four-dimensional conceptual model underlying the scale, the variable measurement levels, the statistical tests to be performed with their respective evaluation parameters, and other important aspects of the research design.
@@
 
-```{r corpoly, echo=FALSE, fig.cap = "Perceptions and preferences for meritcoracy items' polychoric correlations", fig.align='center'}
+```{r corpoly, echo=FALSE, fig.cap = "Perceptions and preferences for meritocracy items' polychoric correlations", fig.align='center'}
   # dev.copy(png,'output/images/Fig7.plotcor-study2.png',width = 800, height = 600) # save image
@@
  # estimation for diagram: summary(fit_cfa3, standardized = TRUE, fit.measures=TRUE)
- knitr::include_graphics('output/images/Fig8.cfa_randomized_study2.png')
+ knitr::include_graphics('output/images/figure9.png')
 ```

### 2023-02-02 Title Page.Rmd
ajustes submit

@@
 ---
-title: "Measuring Perceptions and Preferences for Meritocracy"
+title: "Measuring Perceptions and Preferences for Meritocracy [^*]"
 linestretch: '1.5'
@@
 
-Juan Carlos Castillo, Department of Sociology, Universidad de Chile (
-juancastillov@uchile.cl - [jc-castillo.com](https://juancarloscastillo.github.io/jc-castillo/) -  Ignacio Carrera Pinto 1045 Ñuñoa, Santiago de Chile )
+Juan Carlos Castillo, Department of Sociology, Universidad de Chile *
 
@@
 Francisco Meneses, Department of Sociology, Universidad de Chile
+
+Word count: 8998
+
+[^*]: Corresponding author (juancastillov@uchile.cl - [jc-castillo.com](https://juancarloscastillo.github.io/jc-castillo/) -  Ignacio Carrera Pinto 1045 Ñuñoa, Santiago de Chile).
+The authors recognize the support of ANID/FONDECYT grant 1160921 and of  the Center for Conflict and Social Cohesion Studies (COES)  grant ANID/FONDAP 1513009.

### 2023-02-02 paper-blinded.Rmd
ajustes submit

@@
      keep_text: true
-  # bookdown::word_document2:
-  #   number_sections: false
-  # bookdown::html_document2:
-  #   number_sections: false
+  bookdown::word_document2:
+    number_sections: false
+  bookdown::html_document2:
+    number_sections: false
 linkcolor: blue
@@
  rmarkdown::render("paper-blinded.Rmd", output_format = "bookdown::html_document2")
- # rmarkdown::render("paper.Rmd", output_format = "bookdown::word_document2")
+ rmarkdown::render("paper-blinded.Rmd", output_format = "bookdown::word_document2")
 ```

### 2023-05-12 paper-blinded.Rmd
update

@@
 css: "custom.css"
-linestretch: '2.0'
+linestretch: '1.5'
 link-citations: yes
@@
 Our discussion highlights the importance of considering different dimensions in the study of meritocracy, both conceptually and empirically."
+# keywords: meritocracy, survey methodology, factor analysis, social inequality
 output:
+  bookdown::word_document2:
+    number_sections: false
+    reference_docx: input/template.docx
+    toc: false
   bookdown::pdf_document2:
     template: null
+    number_sections: false    
     toc: false
-     number_sections: false
-     keep_text: true
-  bookdown::word_document2:
-    number_sections: false
+    keep_tex: false
+    pandoc_args:
+      - --template=input/mytemplate.tex #custom template para usar autores con afiliacion    
   bookdown::html_document2:
@@
   - input/bib/meritocracy.bib
-csl: input/bib/asa6.csl
+csl: input/bib/apa7-singlespace.csl
 editor_options:
@@
   - \usepackage{times}           # Times New Roman
+  - \hypersetup {colorlinks = true, linkcolor = blue, urlcolor = blue}    
   - \usepackage{caption}
   - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
-#  - \usepackage[figurename=Fig.]{caption}
   - \usepackage{graphicx}
@@
 always_allow_html: true # for word output
-fontsize: 12pt
+fontsize: 10pt
 papersize: a4
@@
 Please tick one box for each of these to show how important you think it is for getting ahead in life."_ This is followed by a list of statements to be rated from 1 to 5: essential, very important, fairly important, not very important, not important at all.
-
-\break
 
@@
         booktabs=TRUE,format = table_format,escape = FALSE,
-        caption = "Multiple Group meassurement invariance for Perceptions and Preferences for Meritocracy") %>%
+        caption = "Multiple Group measurement invariance for Perceptions and Preferences for Meritocracy") %>%
     kableExtra::kable_styling(full_width = FALSE,
@@
 | Non university      | 32,63%  | 28,93% |
-| university or more  | 21,21%  | 18,13% |
+| University or more  | 21,21%  | 18,13% |
 Table: `r as.character(paste("(\\#tab:rep-samp)", "Representativeness of the study 2 sample"))`.

### 2023-05-16 Title Page.Rmd

@@
 
-Juan Carlos Castillo, Department of Sociology, Universidad de Chile *
+Juan Carlos Castillo, Department of Sociology, Universidad de Chile, Santiago de Chile. *
 
-Luis Maldonado, Institute of Sociology, Pontificia Universidad Católica de Chile
+ORCID: 0000-0003-1265-7854
 
-Jorge Atria, Department of Sociology, Universidad Diego Portales
+Julio Iturra, Bremen International Graduate School of Social Sciences, Bremen, Germany.
 
-Julio Iturra, Bremen International Graduate School of Social Sciences
+ORCID: 0000-0003-1993-3479
 
-Francisco Meneses, Department of Sociology, Universidad de Chile
+Luis Maldonado, Institute of Sociology, Pontificia Universidad Católica de Chile, Santiago de Chile.
+
+ORCID: 0000-0002-0028-4766
+
+Jorge Atria, Department of Sociology, Universidad Diego Portales, Santiago de Chile.
+
+ORCID: 0000-0001-7145-3048
+
+Francisco Meneses, Department of Sociology, Universidad de Chile, Santiago de Chile.
 
@@
 [^*]: Corresponding author (juancastillov@uchile.cl - [jc-castillo.com](https://juancarloscastillo.github.io/jc-castillo/) -  Ignacio Carrera Pinto 1045 Ñuñoa, Santiago de Chile).
-The authors recognize the support of ANID/FONDECYT grant 1160921 and of  the Center for Conflict and Social Cohesion Studies (COES)  grant ANID/FONDAP 1513009.
+The authors recognize the support of ANID/FONDECYT grant 1210847 and of the Center for Conflict and Social Cohesion Studies (COES)  grant ANID/FONDAP 1513009.

### 2023-05-18 paper-blinded.Rmd
submission IJS

@@
 ---
-title: "Measuring Perceptions and Preferences for Meritocracy"
-css: "custom.css"
+title: "Perceptions and Preferences for Meritocracy in Survey Research"
+#css: "custom.css"
 linestretch: '1.5'
@@
 # author: "Juan Carlos Castillo, Julio Iturra, Luis Maldonado, Francisco Meneses & Jorge Atria <br><br>"
-# abstract: "Economic and social inequalities have generated growing concerns and crises across contemporary societies.
-The belief in meritocracy has been suggested as one of the factors that explain the persistence of inequality, as it would legitimize economic disparities based on differences in effort and talent.
-Despite its wide use as a concept, empirical research on meritocracy is relatively novel and characterized by diverse conceptualizations and measures that make the findings and their interpretation rather inconsistent.
-Most of the studies in the area have relied upon secondary data from different surveys to operationalize meritocracy, with a wide variation in the use and interpretation of the same measuring items.
-Based on a critical analysis of concepts and operationalizations of meritocracy in survey research, we propose a conceptual and measurement framework for perceptions and preferences for meritocracy and non-meritocracy in a minimalistic version (few items as possible), in order to be considered in comparative survey research studies.
-The proposal is first tested through confirmatory factor analysis (CFA) using available ISSP (International Social Survey Programme) inequality module items.
-Based on these results, in a second study we propose an  improved scale that better suits the perceptions/preferences distinction and that is tested  in a Chilean sample (N=2,141).
-The results of the CFA estimation give evidence of a better fit for the new scale.
-Our discussion highlights the importance of considering different dimensions in the study of meritocracy, both conceptually and empirically."
-# keywords: meritocracy, survey methodology, factor analysis, social inequality
+abstract: |
+  Economic and social inequalities are causing growing concerns and crises in contemporary societies.
+One factor suggested to explain the persistence of inequality is the belief in meritocracy, which justifies economic disparities based on differences in effort and talent.
+However, empirical research on meritocracy is relatively new and characterized by diverse conceptualizations and measures, leading to inconsistent findings and interpretations.
+To address this issue, the present paper proposes a conceptual and measurement framework for perceptions and preferences regarding meritocracy and non-meritocracy, focusing on comparative survey research methodologies.
+The proposal is first tested through confirmatory factor analysis (CFA) using data from the International Social Survey Programme (ISSP) inequality module, commonly used in comparative research on meritocracy.
+Based on these findings,  a second study focuses on developing a novel scale that better captures the distinction between perceptions and preferences related to meritocracy.
+This scale is tested using a sample of 2,141 individuals from Chile.
+The results of this study contribute to refining the measurement of meritocracy and non-meritocracy and shed light on the importance of considering different dimensions when studying meritocracy in survey research, both conceptually and empirically.
+  \newline
+  \newline
+
+   **Keywords**: meritocracy, survey methodology, social inequality, construct measurement 
+# keywords: "meritocracy, survey methodology, social inequality, construct measurement"
 output:
@@
     toc: false
-    keep_tex: false
+    keep_tex: true
     pandoc_args:
-      - --template=input/mytemplate.tex #custom template para usar autores con afiliacion    
+#      - --template=input/mytemplate.tex #custom template para usar autores con afiliacion    
   bookdown::html_document2:
@@
   - \usepackage{threeparttable}
+  - \usepackage[british,UKenglish,USenglish,american]{babel}
 always_allow_html: true # for word output
@@
 
+\selectlanguage{USenglish}
+
 # Introduction
 
-Economic inequality and income concentration have become topics of growing concern over the last years, leading to a series of social movements in different societies as well as diverse critical analyses regarding the development of capitalism and its consequences [@streeck_politics_2014; @piketty_capital_2014].
+Economic inequality and income concentration have become topics of growing concern over the last years, leading to a series of social upheavals in different societies as well as diverse critical analyses regarding the development of capitalism and its consequences [@streeck_politics_2014; @piketty_capital_2014].
 In such a context, the study of views, preferences, and perceptions of inequality has acquired relevance in the social sciences, in topics such as redistributive preferences [@alesina_fairness_2005; @dimick_models_2018], the legitimization of economic inequality [@schroder_income_2017] and the functioning of meritocracy [@duru-bellat_who_2012; @mijs_paradox_2019; @reynolds_perceptions_2014; @atria_economic_2020].
@@
 One recent definition of meritocracy by @mijs_paradox_2019 is the following: "When I discuss meritocracy beliefs, I am referring to citizens' belief in the importance of hard work relative to structural factors." [@mijs_paradox_2019, pg.9].
-In the subsequent operationalization, this is associated with the following question and indicator: "how important you think it is for getting ahead in life: (a) hard work", scored from 1 to 5 on a likert scale.
+In the subsequent operationalization, this is associated with the following question and indicator: "how important you think it is for getting ahead in life: (a) hard work", scored on a likert scale.
 The assumptions behind such a definition are worth discussing in light of the conceptual meaning of meritocracy and its possibilities of operationalization.
@@
 
-Some research in meritocracy considers aspects usually opposed to effort and talent for personal achievement, for instance, family status and the use of personal advantages (as contacts or having a wealthy family) to get ahead in life.
+Some research in meritocracy considers aspects usually opposed to effort and talent for personal achievement, for instance, family status and the use of personal advantages (such as contacts or having a wealthy family) to get ahead in life.
 This distinction refers to the classical achieved and ascribed status dimension from @linton_study_1936.
@@
 - Based on _previous indicators_ as much as possible to keep comparability between studies.
-As most of the survey research on meritocracy so far has been made with the inequality module of the ISSP survey, we attempted to rescue as much as possible the original wording of some of its items in the new scale proposal.
+As most of the survey research on meritocracy so far has been made with the inequality module of the ISSP survey, we propose an operationalization suitable for this dataset, as well as a novel scale that attempts to rescue as much as possible the original wording of ISSP items.
 
@@
 Study 1 serves as a background to our instrument proposal by attempting to operationalize the meritocracy conceptual model with available secondary data from the inequality module of the International Social Survey Programme (ISSP).
-Although we are aware that the available indicators in this dataset do not allow a precise operationalization of the conceptual model, we know that it is the most widely used dataset in comparative meritocracy studies (as well as overall subjective inequality) and therefore it is certainly useful to look at its potentials and limitations to this regard.
-Study 2 is the core of this paper and consists of the proposal of a new scale for measuring meritocratic perceptions and preferences.
+Study 2 consists of the proposal of a new scale for measuring meritocratic perceptions and preferences.
 Finally, in Study 3 we perform an additional validity analysis of the meritocracy scale.
 
-# Study 1: Measuring meritocracy with international secondary data
+# Study 1: Analyzing meritocracy with ISSP data
 
@@
 
-The data corresponds to the last available wave of the social inequality module from the International Social Survey Programme (ISSP), the most specialized international comparative survey in perceptions, attitudes, and beliefs about inequality-related issues.
+The data corresponds to the social inequality module from the International Social Survey Programme (ISSP), the most specialized international comparative survey in perceptions, attitudes, and beliefs about inequality-related issues [  @issp-research-groupInternationalSocialSurvey2017].
 This wave corresponds to 2009 and covers attitudes towards a series of topics dealing with social inequality across 41 countries and 56,021 individuals.
-Although there is data available from this module for previous waves (1987, 1992, and 1999), unfortunately, there are several variables that are important for this research that were not included, particularly in 1987 and 1999; the reason why we only use the 2009 wave (during the development of this research the 2019 wave became available, but so far with fewer countries than the previous one).
+Although there is data available from this module for a more recent wave (2019), the 2009 wave contains relevant items for studying meritocracy that unfortunately are not present in the last wave.
 Table \@ref(tab:countries) shows details about the sample of countries.
@@
 ```{r table-issp, echo=FALSE}
-  if (!require("pacman")) install.packages("pacman")
+ # if (!require("pacman")) install.packages("pacman")
  # remotes::install_github('haozhu233/kableExtra@a6af5c0')  # for problems with collapse_rows()
-  pacman::p_load(knitr, kableExtra, dplyr)
+library(knitr)
+library(kableExtra)
+library(dplyr)
+#  pacman::p_load(knitr, kableExtra, dplyr)
   table_format = if(knitr::is_html_output()) {
@@
 There were two items from the battery that were excluded from the analysis as they would not fit into the classification.
-The first one was "having good education yourself" since it was not clear whether this is could be due to individual motivation or system opportunities, and the second was "giving bribes", as introduced elements of criminality that were beyond a non-meritocratic perception.
+The first one was "having good education yourself" since it was not clear whether this could be due to individual motivation or system opportunities, and the second was "giving bribes", as introduced elements of criminality that were beyond a non-meritocratic perception.
 
@@
 Figure \@ref(fig:likert-issp) shows the distribution of responses across the selected items in their corresponding dimensions.
-On the one side, we see a high degree of importance attributed to factors such as hard work and ambition for getting ahead, concentrating 94.9% and 92.8% in the fairly important to essential categories.
+On the one side, we observe a high degree of importance attributed to factors such as hard work and ambition for getting ahead, concentrating 94.9% and 92.8% in the _fairly important_ to _essential_ categories, respectively.
 By contrast, the perception of non-meritocratic aspects is rated lower than the meritocratic ones, particularly for gender, race, and political connections.
@@
 In terms of the association between the indicators, Figure \@ref(fig:corr-issp) shows the polychoric correlation matrix.
-Firstly, we observe in general that the moderate to high associations are between the pairs of items representing each of the dimensions as described in \@ref(tab:table-issp).
+Firstly, we observe in general that the moderate to high associations are between the pairs of items representing each of the dimensions as described in Table \@ref(tab:table-issp).
 The exception here is the dimension of non-meritocratic perception, in which six items appear mostly associated by pairs according to their specific topics (family background, personal background, and connections).
@@
 Table \@ref(tab:sum-fit-issp) shows the results of the estimation of two confirmatory models.
-The first one (First Order) corresponds to a model that estimates four factors, each for one of the dimensions presented in \@ref(tab:table-issp), with only regular fit indicators (CFI=0.959, TLI=0.944, RMSEA=0.098, $\chi^2$(df=48)= 21308.535).
+The first one (First Order) corresponds to a model that estimates four factors, each for one of the dimensions presented in  Table \@ref(tab:table-issp), with only regular fit indicators (CFI=0.959, TLI=0.944, RMSEA=0.098, $\chi^2$(df=48)= 21308.535).
 Further analysis showed that the sources for poor fit were mostly related to the items of the non-meritocratic perceptions, which as mentioned above displayed correlations among them not taken into account in this model specification.
@@
 
-Regarding the correlations between the factors in Figure ref(fig:meas02-issp), we observe that perceptions are correlated with preferences, but more strongly for the meritocratic ($r=0.46$) than for the non-meritocratic ($r=0.27$) dimensions.
+Regarding the correlations between the factors in Figure \@ref(fig:meas02-issp), we observe that perceptions are correlated with preferences, but more strongly for the meritocratic ($r=0.46$) than for the non-meritocratic ($r=0.27$) dimensions.
 Secondly, both perceptual dimensions (meritocratic and non-meritocratic) depict a moderate correlation ($r=0.30$), suggesting that they are not the opposite poles of the same continuum.
@@
 
+All participants signed the required informed consent and the survey implementation was approved by the ethical committee from Pontificia Universidad Católica de Chile, protocol ID 150811008, April 6th 2016.
+
 ## Study design
@@
 
-A pre-registration was made in the OSF platform, available at the following link: [https://osf.io/z45y2?](https://osf.io/z45y2/?view_only=adcd496887f2471597b1d6ffb46e492d).
+A pre-registration was made in the OSF platform, available at the following link: [https://osf.io/z45y2](https://osf.io/z45y2/?view_only=adcd496887f2471597b1d6ffb46e492d).
 This pre-registration includes the hypotheses regarding the four-dimensional conceptual model underlying the scale, the variable measurement levels, the statistical tests to be performed with their respective evaluation parameters, and other important aspects of the research design.

### 2023-05-18 paper.Rmd
submission IJS

--- 2021-09-23 paper.Rmd
@@
 ---
-title: "Measuring Perceptions and Preferences for Meritocracy"
+title: "Perceptions and Preferences for Meritocracy in Survey Research"
 css: "custom.css"
@@
 link-citations: yes
-author: "Juan Carlos Castillo, Julio Iturra, Luis Maldonado, Francisco Meneses & Jorge Atria <br><br>"
-abstract: "Economic and social inequalities have generated growing concerns and crises across contemporary societies.
-One of the mechanisms proposed by social sciences to explain the persistence of inequality is the belief in meritocracy, which would legitimize economic disparities based on differences in effort and talent.
-Despite its wide use as a concept, empirical research on meritocracy is relatively novel and characterized by diverse conceptualization and measures that make the findings and their interpretation rather inconsistent.
-Most of the studies in the area have relied upon secondary data to operationalize meritocracy, with a wide variation in the use and interpretation of the same survey items.
-Taking into account the extant literature that uses measures of meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding conceptualization and operationalization.
-Based on this critical analysis, we propose a conceptual framework for studying perceptions and preferences for meritocracy and non-meritocracy, which is then tested through confirmatory analysis using ISSP (International Social Survey Programme) data as well as a novel survey designed specifically with this purpose (N=2,141).
-Our results support the conceptual framework and its operationalizarion, although with a better fit for the proposed scale than for the ISSP survey.
-Our discussion highlights the importance of considering different dimensions in order to advance in the study of meritocracy."
+params:
+  lang: en
+author: 
+- name: Juan Carlos Castillo
+  affiliation: Department of Sociology, Universidad de Chile
+  email: juancastillov@uchile.cl
+  number: 1
+- name: Julio Iturra
+  affiliation: Bremen International Graduate School of Social Sciences, Universität Bremen
+  number: 2
+- name: Luis Maldonado
+  affiliation: Institute of Sociology, Universidad Católica de Chile
+  number: 3
+- name: Francisco Meneses
+  affiliation: 
+  number: 1
+- name: Jorge Atria
+  affiliation: Department of Sociology, Universidad Diego Portales, Chile
+  number: 5
+abstract: |
+  Economic and social inequalities are causing growing concerns and crises in contemporary societies.
+One factor suggested to explain the persistence of inequality is the belief in meritocracy, which justifies economic disparities based on differences in effort and talent.
+However, empirical research on meritocracy is relatively new and characterized by diverse conceptualizations and measures, leading to inconsistent findings and interpretations.
+To address this issue, the present paper proposes a conceptual and measurement framework for perceptions and preferences regarding meritocracy and non-meritocracy, focusing on comparative survey research methodologies.
+The proposal is first tested through confirmatory factor analysis (CFA) using data from the International Social Survey Programme (ISSP) inequality module, commonly used in comparative research on meritocracy.
+Based on these findings,  a second study focuses on developing a novel scale that better captures the distinction between perceptions and preferences related to meritocracy.
+This scale is tested using a sample of 2,141 individuals from Chile.
+The results of this study contribute to refining the measurement of meritocracy and non-meritocracy and shed light on the importance of considering different dimensions when studying meritocracy in survey research, both conceptually and empirically.
+  \newline
+  \newline
+
+   **Keywords**: meritocracy, survey methodology, social inequality, construct measurement 
 output:
@@
     number_sections: false
-  bookdown::html_document2:
-    number_sections: false
+    reference_docx: input/template.docx
+    toc: false
   bookdown::pdf_document2:
     template: null
+    number_sections: false    
     toc: false
+    keep_tex: false
+    pandoc_args:
+      - --template=input/mytemplate.tex #custom template para usar autores con afiliacion    
+  bookdown::html_document2:
     number_sections: false
@@
   chunk_output_type: console
-geometry: margin=0.78in
+geometry: margin=1in
 header-includes:
   - \usepackage{times}           # Times New Roman
+  - \hypersetup {colorlinks = true, linkcolor = blue, urlcolor = blue}    
   - \usepackage{caption}
   - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
-  - \usepackage[figurename=Fig.]{caption}
   - \usepackage{graphicx}
@@
   - \usepackage{threeparttable}
+  - \usepackage[british,UKenglish,USenglish,american]{babel}
 always_allow_html: true # for word output
 fontsize: 10pt
+papersize: a4
 ---
@@
  # clean #in the yml
- rmarkdown::render("paper.Rmd", output_format = "bookdown::pdf_document2")
- rmarkdown::render("paper.Rmd", output_format = "bookdown::html_document2")
- # rmarkdown::render("paper.Rmd", output_format = "bookdown::word_document2")
+ rmarkdown::render("paper-blinded.Rmd", output_format = "bookdown::pdf_document2")
+ rmarkdown::render("paper-blinded.Rmd", output_format = "bookdown::html_document2")
+ rmarkdown::render("paper-blinded.Rmd", output_format = "bookdown::word_document2")
 ```
@@
 
+--->
+
+\selectlanguage{USenglish}
+
 # Introduction
 
-Economic inequality and income concentration have become topics of growing concern over the last years, acquiring even more salience in times of pandemic and economic crises.
-This has led to a series of social movements in different societies as well as diverse critical analyses regarding the development of capitalism and its consequences [@streeck_politics_2014; @piketty_capital_2014].
-In this context, the study of views, preferences, and perceptions of inequality has acquired relevance in the social sciences, in topics such as redistributive preferences [@alesina_fairness_2005; @dimick_models_2018], the legitimization of economic inequality [@schroder_income_2017] and the functioning of meritocracy [@duru-bellat_who_2012; @mijs_paradox_2019; @reynolds_perceptions_2014; @atria_economic_2020].
-Within this context, the ideal of meritocracy has been strongly challenged as an unfulfilled promise of modern societies that allows the perpetuation of social inequalities [@sandel_tyranny_2020; @goldthorpe_myth_2003].
-
-In general, meritocracy is defined as a system of distribution of resources and rewards based on individual merit, which in its original conception is a combination of talent and effort [@young_rise_1962].
+Economic inequality and income concentration have become topics of growing concern over the last years, leading to a series of social upheavals in different societies as well as diverse critical analyses regarding the development of capitalism and its consequences [@streeck_politics_2014; @piketty_capital_2014].
+In such a context, the study of views, preferences, and perceptions of inequality has acquired relevance in the social sciences, in topics such as redistributive preferences [@alesina_fairness_2005; @dimick_models_2018], the legitimization of economic inequality [@schroder_income_2017] and the functioning of meritocracy [@duru-bellat_who_2012; @mijs_paradox_2019; @reynolds_perceptions_2014; @atria_economic_2020].
+Within this research area, and particularly in times of economic and health crises, the ideal of meritocracy has been strongly challenged as an unfulfilled promise of modern societies that allows the perpetuation of social inequalities [@sandel_tyranny_2020; @goldthorpe_myth_2003].
+
+In general, meritocracy has been defined as a system of distribution of resources and rewards based on individual merit, which in its original conception is a combination of talent and effort [@young_rise_1962].
 This traditional conception of merit places in a secondary position the possible interference of structural or non-meritocratic factors, such as inheritance, personal contacts, and luck [@breen_class_1999; @saunders_might_1995; @yair_meritocracy_2007; @land_we_2006; @young_rise_1962].
@@
 
-Due to the role that meritocratic beliefs play in the justification of individual achievement (or failure) in contemporary societies [@hadjar_meritokratie_2008; @markovits_meritocracy_2019; @sandel_tyranny_2020], multiple studies have evaluated the relationship between meritocratic beliefs and personal and/or contextual characteristics.
+Due to the role that meritocratic beliefs play in the justification of individual achievement (or failure) in contemporary societies [@hadjar_meritokratie_2008; @markovits_meritocracy_2019; @sandel_tyranny_2020], multiple studies have evaluated the relationship between meritocratic beliefs and personal and (or) contextual characteristics.
 For example, some studies have linked meritocracy to the reinforcement of socio-economic, gender, and ethnic stereotypes [@madeira_primes_2019; @girerd_neoliberalism_2020; @girerd_neoliberalism_2020], as well as the effects of meritocratic beliefs in educational [@generett_stories_2020; @owens_engines_2020] and organizational contexts [@perez_advancing_2020; @aiello_new_2019].
 Most of such studies so far have used indicators from existing standard social surveys, such as the International Social Survey Programme (ISSP), to measure meritocracy.
-However, as we will show later, the concepts used as well as the instruments for measuring meritocracy vary extensively among studies.
-In many cases, similar phenomena are associated with different indicators, and conversely, often different phenomena are measured with similar indicators.
-Such trends not only limit the comparability of studies, but also the ability to understanding the effects of meritocratic beliefs across different disciplines and lines of research.
-
-Based on a critical analysis of different approaches to the measurement of meritocracy to date, this article proposes an instrument to both measure and relate two main aspects in the study of meritocracy: perceptions and preferences.
-Furthermore, it distinguishes between meritocratic and non-meritocratic dimensions, as they are not necessarily two poles of the same continuum as previous studies seem to suggest.
-This conceptual framework is oriented to develop a measurement instrument as simple and brief as possible, being suitable to be used in public opinion surveys and allowing to integrate meritocratic beliefs in the study of different social phenomena.
-
-## The black-box of meritocratic beliefs {-#blackbox}
+However, as we will show later, the concepts and instruments used for measuring meritocracy vary extensively among studies.
+In many cases, similar phenomena are associated with different indicators, and conversely, different phenomena are measured with similar indicators.
+Such approaches limit the comparability of studies and the capacity to understand the effects of meritocratic beliefs across different disciplines and lines of research.
+
+Based on a critical analysis of different approaches to the measurement of meritocracy to date, this article proposes a conceptual and measurement framework that considers two of the main aspects in the study of meritocracy: perceptions and preferences.
+Furthermore, it distinguishes between meritocratic and non-meritocratic dimensions, as they would not be two poles of the same continuum as previous studies suggest.
+This conceptual framework is oriented to develop a measurement instrument as simple and brief as possible, suitable for public opinion surveys and thus allowing to integrate meritocratic beliefs in studying different social phenomena.
+
+## The black box of meritocratic beliefs {-#blackbox}
 
@@
 One recent definition of meritocracy by @mijs_paradox_2019 is the following: "When I discuss meritocracy beliefs, I am referring to citizens' belief in the importance of hard work relative to structural factors." [@mijs_paradox_2019, pg.9].
-In the subsequent operationalization, this is associated with the following question and indicator: "how important you think it is for getting ahead in life: (a) hard work", scored from 1 to 5 on a likert scale.
-The assumptions behind such definition are worth discussing in light of the conceptual meaning of meritocracy and its possibilities of operationalization.
-
-The item used by @mijs_paradox_2019 is part of an items' battery present in several international surveys, usually called "reasons to get ahead".
-This battery displays a series of indicators related to what people consider important to get ahead in life: hard work, education, ambition, a wealthy family, the right connections, religion, race, and gender.
-Therefore, when considering only one of the items, it means that other aspects that could be associated with talent, such as education, would not be deemed as meritocratic.
+In the subsequent operationalization, this is associated with the following question and indicator: "how important you think it is for getting ahead in life: (a) hard work", scored on a likert scale.
+The assumptions behind such a definition are worth discussing in light of the conceptual meaning of meritocracy and its possibilities of operationalization.
+
+The item used by @mijs_paradox_2019 is part of an item's battery which is available in several international surveys, usually called "reasons to get ahead." This battery displays a series of indicators related to what people consider important to get ahead: hard work, education, ambition, a wealthy family, the right connections, religion, race, and gender.
+Therefore, considering only one of the items means that other aspects associated with talent, such as education, would not be deemed meritocratic.
 As he points out: "Hard work is arguably the most meritocratic part of Michael Young's equation: 'Merit = Intelligence + Effort', for the simple fact that intelligence itself is influenced by a non-meritocratic factor: who your parents happen to be" (p.5).
 
-In this conceptual and measurement approach of meritocratic beliefs, we can observe a couple of strong conceptual assumptions: a) effort would not depend on parental influence, and b) talent (as innate ability) is not meritocratic (contrary to Michael Youngs' original conceptualization).
-This conceptual and measurement-based assumption is found in other studies, which also assumes effort to be the main and only aspect of meritocracy [@girerd_neoliberalism_2020; @bubak_perceptions_2019], raising the question: Is effort the only dimension behind the concept of meritocracy?
-The question of whether talent (as intelligence and ability) is or is not considered meritocratic is certainly an interesting topic to discuss from a philosophical point of view, but for those working on empirical studies, it should face empirical scrutiny.
-Besides, the consideration of talent as part of meritocracy certainly opens some interesting avenues of research, for instance in studies showing that for the elites meritocracy is more related to talent, whereas effort is something more characteristic of the meritocracy of the middle and low classes [@atria_economic_2020].
+In this conceptual and measurement approach of meritocratic beliefs, we can observe a couple of strong conceptual assumptions: a) effort would not depend on parental influence (at least not as strongly as intelligence), and b) talent (as innate ability) is not meritocratic (contrary to Michael Youngs' original conceptualization).
+This conceptual and measurement-based assumption is found in other studies, which also assume effort to be the main and only aspect of meritocracy [@girerd_neoliberalism_2020; @bubak_perceptions_2019], raising the question: Is effort the only (or main) dimension behind the concept of meritocracy?
+We argue that whether talent (as intelligence and ability) is or is not considered meritocratic is an interesting topic to discuss from a philosophical point of view, but it also requires empirical evidence in studies focused on the subjective aspects of inequality.
+Furthermore, considering talent as part of meritocracy certainly opens some interesting avenues of research, for instance, some studies suggest that for the elites meritocracy is related to talent, whereas effort is more characteristic of the meritocracy of the middle and lower classes [@atria_economic_2020].
+Therefore, we propose that a concept (and measure) of meritocracy benefits from including both talent and effort as in the original Young's conception.
 
@@
 Several approaches to the empirical study of meritocracy based on public opinion surveys refer to the concept of _beliefs_, but with wide differences in meanings and operationalizations.
-To illustrate this point let us consider the most commonly used items for operationalizing meritocratic beliefs in survey research (as ISSP), which is the "reasons to get ahead" battery.
-It consists of a series of items asking "how important you think it is for getting ahead in life" and then lists several factors, such as effort, education, parental wealth, and contacts.
-Another version of this same battery used in other surveys - sometimes together with the previous one - asking about "how important you think it _should_ be ...", and then listing the same concepts.
-Therefore, the question raised here is: Which one of both is a "belief": what _is_ or what it _should_ be?
+To illustrate this point let us consider the commonly used "reasons to get ahead" battery, mentioned above.
+Some versions of this battery ask "how important you think it is for getting ahead in life" and then list several factors, whereas another version of this same battery - sometimes presented along the previous one - asks about "how important you think it _should_ be ...", and then listing the same concepts.
+Therefore, the question raised here is: Which one of both is a "belief": what _is_ (the first one) or what it _should_ be (the second one)?
 
@@
 As @kluegel_beliefs_1986 pointed out about the scope of beliefs: "This usage encompasses such more specific social-psychological concepts as values, perceptions, and attitudes" (p.30).
-Therefore, beliefs cover almost anything related to subjective factors.
-To this regard, a relevant distinction in the field of inequality beliefs was made by @janmaat_subjective_2013: "Perceptions refer to subjective estimates of existing inequality (i.e.
-thoughts about what is).
-Beliefs are here defined as normative ideas about just inequality (i.e.
-thoughts about what should be)"(p.359).
+Therefore, beliefs use to cover almost anything related to subjective factors.
+In this regard, a relevant distinction in the field of inequality beliefs was made by @janmaat_subjective_2013: "Perceptions refer to subjective estimates of existing inequality (i.e., thoughts about what is).
+Beliefs are here defined as normative ideas about just inequality (i.e., thoughts about what should be)"(p.359).
 Several papers dealing with meritocracy use the term beliefs (i.e.
-what should be), while actually referring to perceptions (i.e.
-what is).
-This occurs for instance in @reynolds_perceptions_2014, in which the term belief is used to talk about what @janmaat_subjective_2013 refers to as perceptions, whereas other authors use general terms such as attitudes [@kunovich_systems_2007].
+what should be), while referring to perceptions (i.e., what is).
+For instance,  in @reynolds_perceptions_2014, the term belief is used to talk about what @janmaat_subjective_2013 refers to as perceptions, whereas other authors use general terms such as attitudes [@kunovich_systems_2007].
 The first attempt to shed light on this issue in meritocracy research was made by @duru-bellat_who_2012, who used the question "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" as a proxy for "desired" meritocracy (normative beliefs).
@@
 
-Some research in meritocracy considers aspects usually opposed to effort and talent for personal achievement, as for instance the use of personal advantages (as contacts or having a wealthy family) to get ahead in life.
-For instance, @kunovich_systems_2007 used an items' battery listing a number of factors in relation to "How important each should be in deciding pay..." (as @duru-bellat_who_2012 for desired meritocracy).
-They consider factors such as education and responsibility as meritocratic, giving them a value of 1 is considered "essential" in the scale response, whereas factors such as having a family or children are valued 1 when rated as "not important at all" (i.e.
+Some research in meritocracy considers aspects usually opposed to effort and talent for personal achievement, for instance, family status and the use of personal advantages (such as contacts or having a wealthy family) to get ahead in life.
+This distinction refers to the classical achieved and ascribed status dimension from @linton_study_1936.
+For instance, @kunovich_systems_2007 used an items' battery following the question “How important each should be in deciding pay..." (as @duru-bellat_who_2012 for desired meritocracy).
+They consider factors such as education and responsibility as meritocratic, giving them a value of 1 is considered "essential" in the scale response, whereas factors such as having a family or children are valued as 1 when rated as "not important at all" (i.e.
 reverse coded).
-The assumption behind this approach is that rejecting a supposed non-meritocratic aspect (as having family and/or children) implies a stronger belief in meritocracy.
+The assumption behind this approach is that rejecting a supposed ascribed or non-meritocratic aspect (such as having family and/or children) implies a stronger belief in meritocracy.
 A similar approach of reverse-coding non-meritocratic items was taken by @newman_false_2015, using the same principles applied in the "Preference for the Merit Principle Scale" [@davey_preference_1999].
@@
 The assumption that meritocratic and non-meritocratic elements are poles of the same continuum was tested by @reynolds_perceptions_2014 using the "get ahead" perceptions' battery items mentioned above.
-They considered education, ambition, and hard work as meritocratic, and other factors such as family wealth and connections as non-meritocratic.
-Despite making and proving this distinction, however, the authors end up subtracting one dimension from the other, thus coming back to the assumption that they are two poles of the same continuum as @kunovich_systems_2007 did.
+They considered education, ambition, and hard work as meritocratic, whereas factors such as family wealth and connections were classified as non-meritocratic.
+Despite making this distinction, in the operationalization the authors end up subtracting one dimension from the other, thus coming back to the assumption that they are two poles of the same continuum, as @kunovich_systems_2007 also did.
 Similarly, @roex_attitudes_2018 used ISSP indicators for perceived meritocracy and non-meritocracy to build a single score by reverse coding the non-meritocratic items.
-Therefore, the treatment of non-meritocratic items has been rather inconsistent across studies and the assumption that they are the simple opposite of meritocracy certainly requires further empirical assessment.
+Therefore, the treatment of non-meritocratic items has been rather inconsistent across studies and the assumption that they are the simple opposite of meritocracy certainly requires further conceptual elaboration and empirical assessment.
+For instance, a such distinction could open the door to studying to what extent different and contradictory views of meritocracy can coexist, to what extent individuals and societies differ in this regard, and of course the associated factors to such differences.
+Although the use of the concept "non-meritocratic" is very broad, we prefer to use it instead of other terms like ascription, given that this last one leaves out aspects such as the use of personal relationships as a way to achieve success.
 
@@
 
-Finally, most studies in meritocracy so far have not properly considered the issue of latent structures and measurement error [@brown_confirmatory_2015; @bollen_structural_1989; @ansolabehere_strength_2008], as they mostly use single indicators and/or simple average indexes for measuring meritocracy.
-Such strategy assumes that the latent construct is measured perfectly (i.e.
-no error or residual variance) by the selected indicators, going as far as to propose that "...
+Finally, most meritocracy studies have not adequately considered the issue of latent structures and measurement error [@brown_confirmatory_2015; @bollen_structural_1989; @ansolabehere_strength_2008], as they mostly use single indicators and/or simple average indexes for measuring meritocracy.
+Such a strategy assumes that the latent construct is measured perfectly (i.e., no error or residual variance) by the selected indicators, going as far as to propose that "...
 In choosing this strategy of index construction, we argue that support for meritocracy is not a latent variable" [@kunovich_systems_2007 p.653-654].
-Although some advances were made by @reynolds_perceptions_2014 through conducting a principal component analysis of meritocratic and non-meritocratic dimensions, somewhat surprisingly they finally chose to build a sum index despite proving a multidimensional latent structure.
-
-## Conceptual proposal for studying meritocracy {-#instrumentprop}
+Although Reynolds and Xian [-@reynolds_perceptions_2014] made advances by conducting a principal component analysis of meritocratic and non-meritocratic dimensions, they finally chose to build a sum index despite proving a multidimensional latent structure.
+
+There is a relevant trade-off regarding measurement error estimation in survey scales.
+Including multiple items per conceptual dimension opens the possibility of validity studies (at least in terms of the construct) by using factor analysis techniques, particularly confirmatory ones.
+However, increasing the number of items reduces the probability of usability of the scale in survey research given space limitations.
+This validity-usability conundrum is not easy to solve and, as we will show later in the methods section, we opted for a minimalistic approach oriented to the usability of the scale but that still gives the possibility of confirmatory factor analysis, although with some limitations.
+
+## A conceptual proposal for studying meritocracy {-#instrumentprop}
 
@@
 
-- Multiple indicators for each dimension, in order to _account for measurement error_ in a confirmatory factor analysis context.
-
-- Based on _previous indicators_ as much as possible, for the sake of keeping comparability between studies.
-
-- _Brief_, as to be used in regular public opinion surveys.
-In this respect, it differs from the proposal of  "Preference for the Merit Principle Scale" [@davey_preference_1999], as they use 15 items for just one dimension (aside from the problem of reverse-coding non-meritocratic items).
+- Multiple indicators for each dimension to _account for measurement error_ in a confirmatory factor analysis context.
+
+- Based on _previous indicators_ as much as possible to keep comparability between studies.
+As most of the survey research on meritocracy so far has been made with the inequality module of the ISSP survey, we propose an operationalization suitable for this dataset, as well as a novel scale that attempts to rescue as much as possible the original wording of ISSP items.
+
+- _Brief_, as to be used in regular public opinion surveys [@davidov_measurement_2009].
+In this regard, it differs from the proposal of  "Preference for the Merit Principle Scale" [@davey_preference_1999], as they use 15 items for just one dimension (aside from the problem of reverse-coding non-meritocratic items).
 
@@
 The rows in Figure \@ref(fig:merit-model) consider the distinction between meritocratic and non-meritocratic dimensions [@reynolds_perceptions_2014], usually treated as different ends of the same continuum in previous research.
-Non-meritocratic elements refer to the use of resources such as personal contacts or family advantages to getting ahead in life.
+Non-meritocratic elements refer to the use of resources such as personal contacts and/or family advantages to get ahead in life.
 
@@
 The analysis is organized into three studies.
-Study 1 serves as a background to our instrument proposal by attempting to operationalize the meritocracy dimensions with available secondary data from the inequality module of the International Social Survey Programme (ISSP).
-Although we are aware that the available indicators in this dataset do not allow a precise operationalization of the conceptual model, we know that it is the most widely used data in comparative meritocracy studies (as well as overall subjective inequality) and therefore it is certainly useful to look at its potentials and limitations to this regard.
-Study 2 is the core of this paper and consists of the proposal of a new scale for measuring meritocratic perceptions and preferences.
+Study 1 serves as a background to our instrument proposal by attempting to operationalize the meritocracy conceptual model with available secondary data from the inequality module of the International Social Survey Programme (ISSP).
+Study 2 consists of the proposal of a new scale for measuring meritocratic perceptions and preferences.
 Finally, in Study 3 we perform an additional validity analysis of the meritocracy scale.
 
-# Study 1: Measuring meritocracy with international secondary data
+# Study 1: Analyzing meritocracy with ISSP data
 
@@
 
-The data corresponds to the last available wave of the social inequality module from the International Social Survey Programme (ISSP), which is the most specialized international comparative survey in perceptions, attitudes, and beliefs about inequality-related issues.
-This wave corresponds to the year 2009 and covers attitudes towards a series of topics dealing with social inequality across 41 countries and 56021 individuals.
-Although there is data available from this module for previous waves (1987, 1992, and 1999), unfortunately, there are several variables that are important for this research which were not included, particularly in 1987 and 1999, the reason why we only use the 2009 wave (during the development of this research the 2019 wave became available, but so far only with a restricted set of countries).
+The data corresponds to the social inequality module from the International Social Survey Programme (ISSP), the most specialized international comparative survey in perceptions, attitudes, and beliefs about inequality-related issues [  @issp-research-groupInternationalSocialSurvey2017].
+This wave corresponds to 2009 and covers attitudes towards a series of topics dealing with social inequality across 41 countries and 56,021 individuals.
+Although there is data available from this module for a more recent wave (2019), the 2009 wave contains relevant items for studying meritocracy that unfortunately are not present in the last wave.
+Table \@ref(tab:countries) shows details about the sample of countries.
 
@@
 
-There are a series of indicators in the ISSP survey that in the following we attempt to classify in our meritocracy measurement scheme.
-However, it is important to mention that the items were not originally designed as a specific measure of meritocracy.
-The variables are presented below organized in perceptions and preferences:
+There are a series of indicators in the ISSP survey that in the following we attempt to classify in our meritocracy conceptual scheme, as presented below in Table \@ref(tab:table-issp):
 
 ```{r table-issp, echo=FALSE}
-  if (!require("pacman")) install.packages("pacman")
+ # if (!require("pacman")) install.packages("pacman")
  # remotes::install_github('haozhu233/kableExtra@a6af5c0')  # for problems with collapse_rows()
-  pacman::p_load(knitr, kableExtra, dplyr)
+library(knitr)
+library(kableExtra)
+library(dplyr)
+#  pacman::p_load(knitr, kableExtra, dplyr)
   table_format = if(knitr::is_html_output()) {
@@
 There were two items from the battery that were excluded from the analysis as they would not fit into the classification.
-The first one was "having good education yourself", since it was not clear whether this is could be due to individual motivation or system opportunities, and the second was "giving bribes", as introduced elements of criminality that were beyond a non-meritocratic perception.
+The first one was "having good education yourself" since it was not clear whether this could be due to individual motivation or system opportunities, and the second was "giving bribes", as introduced elements of criminality that were beyond a non-meritocratic perception.
 
@@
 
-The estimation was performed using Confirmatory Factor Analysis (CFA).
-CFA was conducted using the `lavaan` R package (version 0.6-3; @rosseel_lavaan_2012), with diagonally weighted least squares (DWLS) estimation due to the items' ordinal level of measurement [@kline_principles_2016; @rosseel_lavaan_2012].
+The estimation of the measurement model was performed using Confirmatory Factor Analysis (CFA).
+CFA was conducted using the `lavaan` R package (version 0.6-3; @rosseel_lavaan_2012a), with diagonally weighted least squares (DWLS) estimation due to the items' ordinal level of measurement [@kline_principles_2016; @rosseel_lavaan_2012a].
 As recommended by @brown_confirmatory_2015, we assessed model fit by jointly considering the comparative fit index and Tucker-Lewis Index (CFI and TLI; acceptable fit > 0.95), Root of the average squared residual approximation (RMSEA; acceptable fit < 0.08),  Chi-square: (p-value; acceptable fit > 0.05, and Chi-square ratio > 3).
@@
 Figure \@ref(fig:likert-issp) shows the distribution of responses across the selected items in their corresponding dimensions.
-On the one side, we see that there is a high degree of importance attributed to factors such as hard work and ambition in the process of getting ahead, concentrating 94.9% and 92.8% in the fairly important to essential categories.
+On the one side, we observe a high degree of importance attributed to factors such as hard work and ambition for getting ahead, concentrating 94.9% and 92.8% in the _fairly important_ to _essential_ categories, respectively.
 By contrast, the perception of non-meritocratic aspects is rated lower than the meritocratic ones, particularly for gender, race, and political connections.
-Regarding normative preferences, we observe that the meritocratic ones are deemed as important (from fairly to essential) for nearly the whole sample, support that decreases slightly for the non-meritocratic ones.
+Regarding normative preferences, we observe that the meritocratic ones are deemed as important (from fairly to essential) for nearly the whole sample, decreasing slightly for the non-meritocratic ones.
 Still, we have to consider that the preferences for non-meritocratic aspects, in this case, refer to the distributive principle of need rather than personal background (as in the perceptions of non-meritocracy), which certainly is one of the limitations of this set of items for the operationalization of the proposed conceptual framework with this dataset.
@@
 In terms of the association between the indicators, Figure \@ref(fig:corr-issp) shows the polychoric correlation matrix.
-Firstly, we observe in general that the moderate to high associations are between the pairs of items representing each of the dimensions as described in \@ref(tab:table-issp).
-The exception here is the dimension of non-meritocratic perception, in which there are six items that appear mostly associated by pairs according to their specific topics (family background, personal background, and connections).
-Still, between this same set of items, there are moderate correlations (around 0.3) which could anticipate a single latent factor underlying non-meritocratic perception and that is tested next through confirmatory factor analysis.
+Firstly, we observe in general that the moderate to high associations are between the pairs of items representing each of the dimensions as described in Table \@ref(tab:table-issp).
+The exception here is the dimension of non-meritocratic perception, in which six items appear mostly associated by pairs according to their specific topics (family background, personal background, and connections).
+Still, between this same set of items, moderate correlations (around 0.3) could anticipate a single latent factor underlying non-meritocratic perception, which is tested next through confirmatory factor analysis.
 
@@
 
-First, a four-dimensional model for meritocratic and non-meritocratic perception and preferences is estimated, considering all indicators of the non-meritocratic perception dimension as a single latent variable.
-Then, we estimate a model that follows the proposed four latent variable measurement approach, however, the non-meritocratic perception dimension is assumed as a second-order factor for the first-order latent variables parental attributes, racial and gender background, and networks.
+Table \@ref(tab:sum-fit-issp) shows the results of the estimation of two confirmatory models.
+The first one (First Order) corresponds to a model that estimates four factors, each for one of the dimensions presented in  Table \@ref(tab:table-issp), with only regular fit indicators (CFI=0.959, TLI=0.944, RMSEA=0.098, $\chi^2$(df=48)= 21308.535).
+Further analysis showed that the sources for poor fit were mostly related to the items of the non-meritocratic perceptions, which as mentioned above displayed correlations among them not taken into account in this model specification.
 
@@
 
-Table \@ref(tab:sum-fit-issp) shows the results of the estimation of two confirmatory models.
-The first one (First Order) corresponds to a model that estimate four factors, each for one of the dimensions presented in \@ref(tab:table-issp), showing only regular fit indicators (CFI=0.959, TLI=0.944, RMSEA=0.098, $\chi^2$(df=48)= 21308.535; more detail in Table \@ref(tab:tb-perpref1)).
-Further analysis showed that the sources for poor fit were mostly related to the items of the non-meritocratic perceptions, which as mentioned above displayed correlations among them not taken into account in this model specification.
-Attending to these constraints, we specified a second model that keeps the basic four-factor structure but generates an additional model for the non-meritocratic perceptions as depicted in \@ref(fig:meas02-issp).
-In this model, the pairs of items in this dimension form three latent factors which at the same time generates a second-order factor of non-meritocratic perceptions, improving the fit of the model.
-
-```{r meas02-issp, echo=FALSE, fig.cap = "Descriptive plot", fig.align='center',fig.cap="Confirmatory factor analysis of the ISSP indicators of Perceptions and Preferences for Meritocracy"}
- knitr::include_graphics('output/images/Fig4.cfa_issp_study1.PNG')
-```
-
-Regarding the correlations between the factors in the Figure \@ref(fig:meas02-issp), we observe that perceptions are correlated with preferences, but more strongly for the meritocratic ($r=0.46$) than for the non-meritocratic ($r=0.27$) dimensions.
+Attending to the sources of poor fit from the first estimated confirmatory model, we specified a second model that keeps the basic four-factor structure but generates an additional model for the non-meritocratic perceptions as depicted in Figure \@ref(fig:meas02-issp).
+In this model, the pairs of items in this dimension form three latent factors which at the same time generate a second-order factor of non-meritocratic perceptions, improving the fit of the model.
+
+```{r meas02-issp, echo=FALSE, fig.cap = "Descriptive plot", fig.align='center',fig.cap="Second-order confirmatory factor analysis model using ISSP indicators of Perceptions and Preferences for Meritocracy"}
+ knitr::include_graphics('output/images/figure4.PNG')
+```
+
+Regarding the correlations between the factors in Figure \@ref(fig:meas02-issp), we observe that perceptions are correlated with preferences, but more strongly for the meritocratic ($r=0.46$) than for the non-meritocratic ($r=0.27$) dimensions.
 Secondly, both perceptual dimensions (meritocratic and non-meritocratic) depict a moderate correlation ($r=0.30$), suggesting that they are not the opposite poles of the same continuum.
 Something similar occurs for the correlation between the two dimensions of preferences ($r=0.25$).
-From this information, we can preliminary conclude that the wide use of reverse-coding for non-meritocratic items attempting to measure meritocracy is not an adequate operationalization of the construct.
-All in all, despite the good fit of the second-order model there are still several limitations regarding the content of the items which we attempt to overcome next by developing a new instrument.
+
+Based on the measurement validation, we asses if the proposed measurement strategy can be used in cross-national comparisons.
+In this line, Figure \@ref(fig:scat1) depicts the correlation of factor scores for Meritocratic Perception and Meritocratic Preferences at the country level.
+For instance, Bulgaria (BGR) and the United States (USA) appropriately represent the positive association between perception and preferences.
+In other words, countries that describe merit in their society as an essential factor to get ahead in life also consider that effort and talent should be determinant factors in how much an individual earns.
+On the other hand, in countries with lower perceived meritocracy, preferences are also low, where Denmark (DNK) and Venezuela (VEN) are representative societies of this association.
+
+Another finding depicted by Figure \@ref(fig:scat1) is the positive association of Non-meritocratic perception with Meritocratic preferences, showing that countries like Bulgaria (BGR) and Philipines (PHL) represent cases with higher Non-meritocratic perceptions and higher preferences for meritocracy.
+In other words, in societies where people perceive higher importance of non-meritocratic aspects to get ahead in life, people also prefer that merit should be a determinant factor in wages.
+On the other side, countries like Sweden (SWE) or Denmark (DNK), both Scandinavian societies with strong welfare systems, represent contexts in which non-meritocratic aspects are not described as important, and merit is not considered as a factor that should be determinant in wages.
+
+```{r scat1, fig.cap="Cross-national comparison of non-meritocratic and meritocratic perceptions with meritocratic preference",fig.height=6, fig.width=12}
+  load(file = "input/data/proc/data_s1.RData")
+  if (!require("pacman")) install.packages("pacman")
+  pacman::p_load(dplyr, kableExtra, knitr,lavaan,countrycode,ggplot2,haven,
+                 gridExtra)
+  table_format = if(knitr::is_html_output()) { #conditional instructions for kable
+    "html" #if html set "html" in format
+  } else if(knitr::is_latex_output()) {
+    "latex"#if latex set "latex" in format
+  }
+
+  # meassurement model: Second order
+  cfa_perpref2 <- '
+  	  # latent variables
+  	  percmerit =~ hwork + ambition
+  	  parents =~ wealthy + pareduc
+  	  backgrd =~ race + gender
+  	  networks=~ people + polcone
+  	  percnmerit=~ parents + backgrd + networks
+  	  prefmerit =~ welljob + hardjob
+  	  prefnmerit=~ family + child
+  	  '
+  # Model fit: second order
+
+  fit_perpref2 <- lavaan::cfa(cfa_perpref2, data = data_s1,ordered = F)
+
+  df_proc<- data_s1 %>% select(id,v5,wealthy:hardjob)  
+  df_proc$ccode <- as.numeric(data_s1$v5)
+
+  predicted2 <- lavPredict(fit_perpref2,newdata = data_s1)
+
+  idx <- lavInspect(fit_perpref2, "case.idx")
+  fscores <- lavPredict(fit_perpref2)
+  ## loop over factors
+  for (fs in colnames(fscores)) {
+    df_proc[idx, fs] <- fscores[ , fs]
+  }
+
+  df_proc <-  df_proc %>%  select(id,ccode,percmerit,prefmerit,prefnmerit) %>% na.omit()
+
+  df_proc$clab <- countrycode(sourcevar = df_proc$ccode,
+                            origin = "iso3n",
+                            destination = "iso3c")
+
+avg_count <-  
+  df_proc %>% group_by(clab) %>% summarise(prefmerit=mean(prefmerit),
+                                            percmerit=mean(percmerit),
+                                            prefnmerit=mean(prefnmerit))
+
+# create figure 1: perception merit by preferences merit
+figure1 <- 
+ggplot(avg_count, aes(x=percmerit, y=prefmerit)) +
+  # geom_point() + 
+  geom_text(label=as.character(avg_count$clab)) +
+  geom_smooth(method = "lm") +
+  xlab("Meritocratic perception")+
+  ylab("Meritocratic preference")  
+# create figure 2: perception non-merit by preferences merit
+figure2 <- 
+ggplot(avg_count, aes(x=prefnmerit, y=prefmerit)) +
+  # geom_point() + 
+  geom_text(label=as.character(avg_count$clab)) +
+  geom_smooth(method = "lm") +
+  xlab("Non-meritocratic perception")+
+  ylab("Meritocratic preference")
+
+grid.arrange(figure1, figure2, ncol = 2,
+             bottom = grid::textGrob(label = paste0("Source: Authors calculation based on International Social Survey Programme - Social Inequality 2009",
+      " (n=",dim(df_proc)[1],")"),
+                                     hjust = 0.1,
+                                     gp = grid::gpar(fontface = 1, fontsize = 9)))
+```
+
+## Discussion Study 1
+
+Although it was possible to get an adequate fit of the model to the pooled data from ISSP, there are several aspects to take into account when using these items for measuring meritocracy.
+First of all, these items were not specifically designed based on a conceptual approach to meritocracy, but it still is possible to find a certain fit of the model to the data after minor adjustments.
+Secondly, some of the items in similar dimensions refer to different concepts, for instance, meritocratic perceptions are related to ambition and hard work, whereas meritocratic preferences deal with how well a job is done.
+A better operationalization would require more alignment between the different sides of the same concept.
+And thirdly, the scale is rather unbalanced given the larger amount of non-meritocratic perception items, producing correlated errors.
+All these limitations are taken into account next in the design of Study 2.
+
+Concerning the meritocratic / non-meritocratic distinction in previous studies, we can preliminary conclude that the wide use of reverse-coding for non-meritocratic items attempting to measure meritocracy is not an adequate operationalization of the construct.
+Meritocracy and non-meritocracy are related but different concepts, meaning that it is theoretically possible to find different combinations at individual and contextual levels.
+This finding opens a series of avenues for future research that are considered in the conclusions.
 
@@
 
-The data was obtained through an online survey which was part of a larger study on meritocracy and preferences developed in Chile in 2019, funded by the national scientific agency ANID.
+The data comes from an online survey that was part of a larger study on meritocracy and preferences developed in Chile in 2019, funded by the Chilean National Scientific Agency (ANID).
 The questionnaire was programmed in Qualtrics and the fieldwork was conducted by an external online survey agency ([netquest.com](www.netquest.com)) between December 2019 and January 2020.
@@
 The quotas for gender, age, and educational levels were generated based on a survey by the Public Studies Center [@cep_encuesta_2019], which is a well-regarded counterpart agency of the ISSP (International Social Survey Programme) in Chile.
-A total sample of 2,141 people was collected, excluding those who did not sign the informed consent as well as those not answering the meritocracy response battery.
-There were no significant differences between our sample and the wider population for most socio-demographic characteristics, with the exception of educational level (see Table \@ref(tab:rep-samp) in Appendix).
-As it is often the case with online surveys, there were some limitations in achieving the quotas for lower educational levels [@zhang_quota_2018; @boas_recruiting_2020].
+A total sample of 2,141 people was collected, excluding those who did not sign the informed consent and those not answering the meritocracy response battery.
+There were no significant differences between our sample and the wider population for most socio-demographic characteristics, except for the educational level (see Table \@ref(tab:rep-samp) in Appendix).
+As is often the case with online surveys, there were some limitations in achieving the quotas for lower educational levels [@zhang_quota_2018; @boas_recruiting_2020].
+This survey was designed as a three-wave panel and the data used in this analysis corresponds to the first wave.
+
+All participants signed the required informed consent and the survey implementation was approved by the ethical committee from Pontificia Universidad Católica de Chile, protocol ID 150811008, April 6th 2016.
 
@@
 
-The proposed scale of perceptions and preferences relating to meritocracy consisted of eight indicators that were grouped into the four dimensions listed earlier: Perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic).
-In order to achieve at least some comparability with previous studies, the questions were adapted from the items battery "reasons to get ahead" (ISSP/GSS), which has been widely used for operationalizing meritocracy in previous studies [@mijs_paradox_2019; @duru-bellat_who_2012; @reynolds_perceptions_2014].
-The aforementioned eight items ordered according to dimensions are presented in Table \@ref(tab:table-indicadores).
+The proposed scale of meritocratic perceptions and preferences consisted of eight indicators, two for each of the four dimensions listed earlier: Perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic).
+Two main criteria oriented this design.
+The first one was usability in social survey research, for which the scale length must attend spacer restrictions in such studies.
+Nevertheless, at the same time, the scale should offer possibilities of accounting for measurement error through factor analysis techniques.
+Although a rule of thumb in latent measurement suggests at least three indicators per dimension, we still opted for two to count with a brief scale.
+Although such a decision posits some limitations for modeling measurement error, still allows for factor analysis and is a common approach in previous social survey studies with limited items (for instance see @davidov_measurement_2009).
+The second criterion is to expand the possibilities of comparability with previous studies on meritocracy, by considering some of the items previously used in this research area.
+To this regard, some of the items were adapted from the items battery "reasons to get ahead" (ISSP/GSS), which has been widely used for operationalizing meritocracy in international research [@mijs_paradox_2019; @duru-bellat_who_2012; @reynolds_perceptions_2014].
+The scale items ordered according to their respective dimensions are presented in Table \@ref(tab:table-indicadores).
 These eight likert-type items have five response alternatives, ranging from "Completely disagree"(1) to "Completely agree" (5).
@@
 
-With the objective of evaluating the effect of indicator ordering in the responses, three different versions of items' order were designed and randomly assigned to the respondents, as depicted in Figure \@ref(fig:appmod).
-The scale was presented to the first group (_n = 712_) in the order that appears in Table \@ref(tab:table-indicadores) according to perceptions and preferences.
-For the second group (_n = 717_), the order was reorganized according to perceptions and preferences over the same topic, e.g.
-for the topic of hard work, the item about perception was followed by the item about preference and the same for the rest of the topics.
+To evaluate the effect of indicator ordering in the responses, three different versions of items' order were designed and randomly assigned to the respondents, as depicted in Figure \@ref(fig:appmod).
+The scale was presented to the first group (_n = 712_) in the order shown in Table \@ref(tab:table-indicadores) according to perceptions and preferences.
+For the second group (_n = 717_), the order was reorganized according to perceptions and preferences over the same topic, e.g., for the topic of hard work, the item about perception was followed by the item about preference and the same for the rest of the topics.
 Finally, for the third group (_n = 712_), the items were presented as completely randomized.
@@
 The models estimated one factor for each of the four proposed dimensions presented in Table \@ref(tab:table-indicadores).
-As in Study 1, CFA was conducted using the `lavaan` R package (version 0.6-3; @rosseel_lavaan_2012), with diagonally weighted least squares (DWLS) estimation due to the items' ordinal level of measurement [@kline_principles_2016; @rosseel_lavaan_2012].
+As in Study 1, CFA was performed with the `lavaan` R package (version 0.6-3; @rosseel_lavaan_2012a), with diagonally weighted least squares (DWLS) estimation due to the items' ordinal level of measurement [@kline_principles_2016; @rosseel_lavaan_2012a].
 The fit indexes and cut-off criteria were the same as the ones used in Study 1.
 
-A pre-registration was made in the OSF platform, available at the following link: [https://osf.io/z45y2](https://osf.io/z45y2).
-Included in this pre-registration are the hypotheses regarding the four-dimensional conceptual model underlying the scale, the variable measurement levels, the statistical tests to be performed with their respective evaluation parameters, and other important aspects of the research design.
+A pre-registration was made in the OSF platform, available at the following link: [https://osf.io/z45y2](https://osf.io/z45y2/?view_only=adcd496887f2471597b1d6ffb46e492d).
+This pre-registration includes the hypotheses regarding the four-dimensional conceptual model underlying the scale, the variable measurement levels, the statistical tests to be performed with their respective evaluation parameters, and other important aspects of the research design.
 
@@
 There are three main aspects to highlight from this correlation matrix.
-Firstly, as expected the largest correlations are between indicators that correspond to the same factors behind the conceptual model (e.g., perception of meritocracy by effort and by talent, $r=0.52$).
-Secondly, among this correlations the highest are those between the non-meritocratic dimension but in perceptions ($r=0.73$) and preferences ($r=0.61$).
-Thirdly, both items for meritocratic preferences (E and F) are the ones that mostly correlate with the rest of the perceptual items, showing medium to high correlations.
-This is noteworthy because it indicates that the perception of non-meritocracy would be related to larger meritocratic preferences.
-Finally, and similar to what was found in Study 1 with ISSP data, we observe that there are no considerable negative correlations between meritocratic and non-meritocratic aspects, undermining the assumptions of previous studies that suggested that these dimensions would be the opposite poles of one same continuum [@reynolds_perceptions_2014].
-
-```{r corpoly, echo=FALSE, fig.cap = "Perceptions and preferences for meritcoracy items' polychoric correlations", fig.align='center'}
+Firstly, and as expected, the largest correlations are between indicators that correspond to the same factors behind the conceptual model (e.g., perception of meritocracy by effort and by talent, $r=0.52$).
+Secondly, among these correlations, the highest are those between the non-meritocratic dimension but in perceptions ($r=0.73$) and preferences ($r=0.61$).
+Thirdly, both items for meritocratic preferences (E and F) are the ones that mostly correlate with the rest of the perceptual items, showing medium to high correlations and indicating that the perception of non-meritocracy would be related to larger meritocratic preferences.
+Finally, similar to the results of Study 1 with ISSP data, we observe no considerable negative correlations between meritocratic and non-meritocratic aspects, undermining previous studies' assumptions that these dimensions would be the opposite poles of one same continuum [@reynolds_perceptions_2014].
+
+```{r corpoly, echo=FALSE, fig.cap = "Perceptions and preferences for meritocracy items' polychoric correlations", fig.align='center'}
   # dev.copy(png,'output/images/Fig7.plotcor-study2.png',width = 800, height = 600) # save image
@@
 Attending to the correlations between the latent variables as depicted in Figure \@ref(fig:meas02), meritocratic preferences are moderate to highly correlated with perceptions, both meritocratic ($r=0.457$) and non-meritocratic ($r=0.500$).
-The correlation between both types of perceptions ($r=-0.044$) and both types preferences ($r=0.185$) are low, as well as between non-meritocratic preferences and perceptions ($r=-0.059$).
+The correlation between both types of perceptions ($r=-0.044$) and both types of preferences ($r=0.185$) was low, as well as between non-meritocratic preferences and perceptions ($r=-0.059$).
 This last finding gives further evidence regarding the lack of unidimensionality of meritocratic and non-meritocratic aspects as assumed by previous studies.
 
+## Discussion Study 2
+
+The estimation performed with different administration sets for the meritocracy scale gives evidence in line with the conceptual model.
+The four dimensions appear independent but related, with some differences when compared to the estimation with ISSP data in Study 1.
+The main general difference is that with ISSP data, the four latent dimensions were positively correlated, whereas now some correlations are statistically not different from zero.
+For instance, perceptions and preferences for meritocracy keep being related, which is not the case for the non-meritocratic pair.
+Theoretically, this makes more sense, as if someone perceives that meritocracy functions well, they probably will support it more.
+On the other side, if someone perceives non-meritocracy, this should lead to a more prominent preference for meritocracy as it would constitute a demand for a larger meritocracy when it is not perceived to work.
+
+Despite the good general working of the scale, there are some limitations in interpreting the results given the few items (two) per dimension.
+As mentioned before, the option of a minimalistic approach in terms of items has advantages and disadvantages: advantages in terms of usability in social surveys with limited space and disadvantages in terms of measurement validity in the estimation of the latent constructs.
+Aware of this trade-off, next we present additional evidence of validity to complement some of the weaknesses identified.
+
 # Study 3: Additional validity analyses
 
-We performed two further analyses in order to add evidence regarding the quality of the scale.
-Firstly we analyzed the convergent validity by exploring the association of the merit scale with related concepts and its correspondent measurements.
-Secondly, we test the internal consistency of the measurement model behind the scale comparing it with a different sample through measurement invariance procedures.
-
-## Convergent validity
+The modeling for invariance testing has been gaining more relevance in sociological survey studies, mainly due to the interest in the validity of measurement scales in comparative studies in various social and cultural contexts [@davidov_measurement_2014].
+In this section, we present the results of an analysis of invariance between groups, testing whether the proposed factorial structure for the meritocracy scale is applicable in a different dataset.
+
+The invariance measurement procedure consists of a series of nested models in which restrictions are progressively incorporated into the parameters of the measurement model.
+The literature generally suggests that this modeling approach should consider four levels or types of progressive restriction [@milfont_testing_2010; @millsap_statistical_2011; @vandeschoot_checklist_2012]:
+
+1. **Configural:** the model is estimated only indicating the factorial structure used in the CFA.
+
+2. **Weak:** an equality restriction applied to the factor loadings in the different groups; that is, the loadings are forced to be identical in both measurements.
+
+3. **Strong:** equality restrictions added to the intercepts of each indicator.
+
+4. **Strict:** equality restrictions added to the error variances of each indicator.
 
@@
 
-For this analysis, we examined data from the last of the three waves of the online panel survey used in Study 2.
-This wave included additional measures that allow testing the correlations of the merit scale with some related constructs as opportunity beliefs and personal wherewithal.
-After a listwise deletion of missing cases in socio-demographic information, a total of 1422 individuals took part in the third wave: 668 (46.97%) women and 754 (53.03%) men.
-The ages of 50.57% of them are 45 years or older, and 36.28% held a tertiary degree.
+For testing the invariance, we compared the data described previously in Study 2 to a different dataset from an online survey carried out during the first half of 2020.
+The characteristics of this sample in terms of application and coverage are equivalent to those of the previous study.
+The final sample contains 1,242 cases, where 605 (48.71%) are women, and 637 (51.28%) are men.
+48.84% are 45 or older, and 33.97% have tertiary education or higher.
 
@@
 
-The following scales were included for testing their correlations with the 8-item meritocracy scale:
-
-- Opportunity beliefs: We used two items of the social inequality module of the International Social Survey Programme.
-Studies use generally both indicators to measure meritocratic and non-meritocratic beliefs [@mccall_exposure_2017; @mijs_paradox_2019].
-One item measures the importance of individual factors in determining life outcomes by asking for the importance of hard work for getting ahead in life (M=3.76, SD=0.93).
-The second item asks for the importance of coming from a wealthy family and captures the importance of structural factors in getting ahead (M=2.75, SD=1.27).
-Participants responded to each item using a 1 (not important at all) to 5 (essential) response scale.
-We call the first item _hard work_ and the second one _social origin_.
-
-- Personal wherewithal: the questionnaire included The Neoliberal Beliefs Inventory (NBI), which considers four factors: Government Interference preferences, Competition preferences, System Inequality perception, and Personal Wherewithal [@bay-cheng_tracking_2015].
-We used the factor _personal wherewithal_ that reflects meritocracy beliefs in terms of the importance of personal attributes as strength and skills to yield success.
-The construct consists of 8 items (e.g., "Any goal can be achieved with enough hard work and talent", "I've benefited from working hard, so there's no reason others can't", "Anyone who is willing to work hard can be successful in Chile").
-Participants answered the items using a 1 (totally disagree) to 6 (totally agree) scale.
-We simply computed mean scores, with higher scores indicating stronger support for meritocracy (M=3.32, SD=0.88, $\alpha$ = 0.89).
-
-Both instruments cover mainly the perceptual side of the concept.
-Therefore, in general, we expect a larger correlation with meritocratic perceptions than with preferences.
-In the case of opportunity beliefs, the hard work item should correlate positively with meritocratic perception whereas the importance of coming from a wealthy family is expected to correlate positively with non-meritocratic perceptions.
-Regarding the second instrument of personal wherewithal, we also anticipate a positive correlation with meritocratic perception and low to null correlations with the other dimensions of the merit scale.
+The items of the meritocracy scale were identical to the original scale of Study 2 administrated in a randomized order (which corresponds to the application modality of Group 3 in Study 2, as shown in Figure \@ref(fig:appmod)).
 
@@
 
-We used polychoric correlations to evaluate relationships of meritocratic and non-meritocratic preferences and perceptions with the items of the opportunity beliefs battery.
-For personal wherewithal, we examined Pearson correlations.
-
-Table \@ref(tab:val-conv) shows the polychoric and Pearson correlations between our meritocracy scale and common measures of meritocracy.
-As we expected, preferences show very weak correlations.
-Findings also indicate that correlations of perceptions with opportunity beliefs are consistent with our expectations.
-The item for social origin shows a positive and moderate association with unmeritocratic perception ($r=0.361$), and a negative and low correlation with meritocratic perception ($r=-0.186$).
-Regarding the item of hard work, the correlation with meritocratic perception is positive but weak ($r=0.201$).
-One explanation for this result is that the support for hard work is very high among respondents and therefore with low variability, as only 8.6 percent indicates _not very important_ or _not important at all_.
-
-```{r val-conv}
-  if (!require("pacman")) install.packages("pacman")
-  pacman::p_load(kableExtra, polycor)
-  table_format = if(knitr::is_html_output()) { #conditional instructions for kable
-     "html" #if html set "html" in format
-   } else if(knitr::is_latex_output()) {
-     "latex"#if latex set "latex" in format
-   }
-  load(file="input/data/proc/data_s3_conv.RData")
-
-  pl1 <- polycor::polyserial(data_s3_conv$perc_merit,data_s3_conv$get_ah_1)
-  pl2 <- polycor::polyserial(data_s3_conv$pref_merit,data_s3_conv$get_ah_1)
-  pl3 <- polycor::polyserial(data_s3_conv$perc_nmerit,data_s3_conv$get_ah_1)
-  pl4 <- polycor::polyserial(data_s3_conv$pref_nmerit,data_s3_conv$get_ah_1)
-  valores1 <- round(c(pl1,pl2,pl3,pl4),4)
-
-  pl1b <- polycor::polyserial(data_s3_conv$perc_merit,data_s3_conv$get_ah_5)
-  pl2b <- polycor::polyserial(data_s3_conv$pref_merit,data_s3_conv$get_ah_5)
-  pl3b <- polycor::polyserial(data_s3_conv$perc_nmerit,data_s3_conv$get_ah_5)
-  pl4b <- polycor::polyserial(data_s3_conv$pref_nmerit,data_s3_conv$get_ah_5)
-  valores2 <- round(c(pl1b,pl2b,pl3b,pl4b),4)
-
-  pl1c <- cor(data_s3_conv$perc_merit,data_s3_conv$nbmerit)
-  pl2c <- cor(data_s3_conv$pref_merit,data_s3_conv$nbmerit)
-  pl3c <- cor(data_s3_conv$perc_nmerit,data_s3_conv$nbmerit)
-  pl4c <- cor(data_s3_conv$pref_nmerit,data_s3_conv$nbmerit)
-  valores3 <- round(c(pl1c,pl2c,pl3c,pl4c),4)
-
-  variables= c("Meritocratic Perception", "Meritocratic Preferences","Non-meritocratic Perception","Non-meritocratic Preferences")
-
-  table01 <- cbind(variables, valores1, valores2, valores3)
-
-  knitr::kable(table01,
-        col.names = c("Merit-Scale","Social Origin ", "Hard Work " , "Wherewithal"),
-        linesep = "",
-        format = table_format,
-        booktabs=T,
-        digits = 3,
-        caption = c("Polyserial and Pearson correlation whit other merit scales")) %>%
-    kableExtra::kable_styling(
-      full_width = F,
-      position = "center",
-      font_size = 10,
-      bootstrap_options = c("striped", "bordered")) %>%
-    kableExtra::column_spec(column = 1, width = "4.5 cm")
-```
-
-The measure of NBI for _personal wherewithal_ also provides evidence for the convergent validity of the merit scale.
-Results indicate a positive correlation of this measure with meritocratic perception ($r=0.406$), suggesting that both variables capture similar concepts.
-The NBI's factor depicts a positive association with unmeritocratic preference ($r=0.225$).
-This finding is consistent with the literature, in as much as the belief for meritocracy is associated with the justification of inequality [@mccall_exposure_2017; @madeira_primes_2019].
-Furthermore, results of our confirmatory factor analysis indicate that meritocratic perceptions and unmeritocratic preferences show a positive and weak correlation.
-
-## Measurement invariance
-
-The modeling for invariance testing has been gaining more relevance in sociological survey studies, mainly due to the interest in the validity of measurement scales in comparative studies in various social and cultural contexts [@davidov_measurement_2014].
-In this section we will use an analysis of invariance between groups, where the proposed factorial structure is expected to be independent of belonging to the sample corresponding to the original study, comparing it with another more recent study available and where the meritocracy scale was also incorporated.
-
-The invariance measurement procedure consists of a series of nested models to which restrictions are progressively incorporated on the parameters of the measurement model.
-The literature generally suggests that this modeling should be done through four levels or types of progressive restriction [@milfont_testing_2010; @millsap_statistical_2011; @vandeschoot_checklist_2012]:
-
-1. **Configural:** the model is estimated only indicating the factorial structure used in the CFA.
-
-2. **Weak:** an equality restriction is applied to the factor loadings in the different groups, that is, the loadings are forced to be identical in both measurements.
-
-3. **Strong:** equality restrictions are added to the intercepts of each indicator.
-
-4. **Strict:** equality restrictions are added to the error variances of each indicator.
-
-### Data
-
-For testing the invariance we compared the data described previously with a new data source coming from an online survey carried out during the first half of 2020.
-The characteristics of this sample in terms of application and coverage are equivalent to those of the previous study.
-The final sample obtained contains 1,242 cases, where 605 (48.71%) are women and 637 (51.28%) are men.
-48.84% of them are 45 years of age or older, and 33.97% have tertiary education or higher.
-
-### Variables
-
-The items of the meritocracy scale were identical to the original.
-The items were administrated in a randomized order (which corresponds to the application modality of Group 3 present in Figure \@ref(fig:appmod)).
-
-### Results
-
-Table \@ref(tab:fcloads-inv) shows the results of the CFA estimation of the meritocracy scale in this new sample.
-As observed, the loadings and the fit indices show similar results as the ones presented for the original study above, which gives a first base from which to start the invariance analysis.
-
-The first step for invariance testing is the estimation of the configural model, which serves as the baseline for further comparison and is expected to adequately meet the global fit criteria of a measurement model.
-Although the chi-square statistic is used as a global measure of fit, since its sensitivity to sample size it is recommended to use three additional fit indices: Comparative Fit Index (CFI), which should have a value greater than 0.95; Root Mean Square Error Approximation (RMSEA), which must be lower than 0.06; and the Standardized Root Mean Square Residuals (SRMR) which must be less than 0.08 [@vandeschoot_checklist_2012].
-Besides, the literature on measurement invariance suggests some complementary approaches for the evaluation of the fit, from which we will consider the incremental adjustment of the fit indexes [@cheung_evaluating_2002; @milfont_testing_2010; @dimitrov_testing_2010] and the ANOVA test for means comparison in nested models [@newsom_longitudinal_2015].
+The first step for invariance testing is estimating the configural model, which serves as the baseline for further comparison and is expected to adequately meet the global fit criteria of a measurement model.
+Although the chi-square statistic is used as a global measure of fit, it is usually complemented with other indexes given its high sensitivity to sample size: the Comparative Fit Index (CFI), which should have a value greater than 0.95; the Root Mean Square Error Approximation (RMSEA), which must be lower than 0.06; and the Standardized Root Mean Square Residuals (SRMR) which must be less than 0.08 [@vandeschoot_checklist_2012].
+Besides, the literature on measurement invariance suggests some complementary approaches for evaluating the fit, from which we will consider the incremental adjustment of the fit indexes [@cheung_evaluating_2002; @milfont_testing_2010; @dimitrov_testing_2010] and the ANOVA test for means comparison in nested models [@newsom_longitudinal_2015].
 
@@
 Such result is considered in general as evidence of invariance [@fischer_are_2011], as strict forms of measurement invariance rarely hold [@vandeschoot_checklist_2012].
-Still, the comparability of latent means requires strict invariance which in this case does not hold when considering $\Delta \chi^2 (\Delta \text{df})$.
-Nevertheless, the criteria of $\Delta \text{CFI}$ used for comparing models is close to the rejection criteria of >.01, whereas the $\Delta \text{RMSEA}$ fulfils the requirements of being below the cut-off criteria as suggested by [@chen_sensitivity_2007].
-Therefore, using this last standard the level for strict invariance holds for the meritocracy scale.
+Still, the comparability of latent means requires strict invariance which in this case does not hold when considering just $\Delta \chi^2 (\Delta \text{df})$.
+Nevertheless, the criteria of $\Delta \text{CFI}$ used for comparing models is close to the rejection criteria of >.01, whereas the $\Delta \text{RMSEA}$ fulfills the requirements of being below the cut-off criteria as suggested by [@chen_sensitivity_2007].
+Therefore, using this last standard, the level of strict invariance would hold for the meritocracy scale.
+
+## Discusion Study 3
+
+Based on the measurement invariance analysis for two datasets, we found evidence supporting the equivalence of the scale when applied in different samples.
+This gives a more robust ground when using this measure for comparing groups, for instance in country comparisons.
+Although the results are promising, still the analysis was performed in two samples in a single country (Chile) and it requires further examination with social surveys in other societies.
 
@@
 As most studies use secondary survey data, they tend to assume that the available indicators represent an underlying meritocratic construct.
-A review of these studies reveals several non-tested assumptions, as well as the use of similar indicators to represent different constructs and dimensions of meritocracy.
+A review of these studies reveals several non-tested assumptions and the use of similar indicators to represent different constructs and dimensions of meritocracy.
 As the existence of heterogeneous approaches certainly has consequences for the advancement of the study of meritocracy, this paper presented a comprehensive conceptual framework for the empirical study of meritocracy, building upon previous research.
-This framework was then tested against available and new survey data.
+This framework was operationalized in a survey scale and tested against available and new survey data.
 
 We identified four critical aspects regarding the measurement of meritocracy in previous studies: unidimensionality, the ambiguous use of the term "beliefs", the use of non-meritocratic indicators as opposed to meritocratic indicators, and the consideration of measurement error.
-The proposed 8-item scale, "Perceptions and Preferences for Meritocracy", was designed and tested in order to deal with these four issues.
-Our results indicate that perceptions and preferences seem to be two related but different dimensions, often confounded in previous research under the label of "beliefs".
-Meritocratic and non-meritocratic dimensions do not appear to constitute poles of the same continuum, as some previous studies have assumed.
+The proposed 8-item scale, "Perceptions and Preferences for Meritocracy", was designed and tested to deal with these four issues.
+Our results indicate that perceptions and preferences are two related but different dimensions, often confounded in the previous research under the label of "beliefs".
+At the same time, meritocratic and non-meritocratic dimensions do not appear to constitute poles of the same continuum, as some previous studies have assumed.
 Regarding the possible effects of the items' order in the estimation of the latent variable, we tested three different order versions and found evidence that suggests the use of the randomized version of the scale.
-Furthermore, we found evidence of convergent validity as well as measurement invariance that approach well the requirement of comparability between samples.
-
-The four-dimensional conceptual framework and its operationalization in the Perceptions and Preferences for Meritocracy scale open several avenues for future research.
+Furthermore, we found evidence of measurement invariance that approaches well the requirement of comparability between samples.
+
+One of the criteria in designing the merit scale was to develop a short instrument that could be used in social surveys, assuming that using only two items per dimension posits limitations for the confirmatory analysis estimated here.
+We preferred this confirmatory-minimalistic approach as it offers several advantages in terms of information about the quality of the measurement, and we offered additional validity evidence with measurement invariance analysis.
+Despite the advantages, a more straightforward perspective is always available, for instance using only the average of the items per dimension.
+Whereas a short scale is always appreciated in social survey studies, for other audiences could not reach measurement standards.
+This may require further expansions and/or consider using a different instrument as the "Preference for the Merit Principle Scale" [@davey_preference_1999] (which measures only one dimension of meritocracy with fifteen items).
+
+The four-dimensional conceptual framework and its operationalization in the Perceptions and Preferences for Meritocracy scale opens several avenues for future research.
 For instance, distinguishing perceptions from preferences will allow us to evaluate the extent to which different societies are accustomed to, or satisfied with, the perceived level of meritocracy, in terms of differences between what is perceived and what is preferred.
 Additionally, given that non-meritocratic factors are not necessarily related to meritocratic ones, our framework makes it possible to assess the perceived legitimacy of practices such as the use of personal contacts and their interference (or not) with meritocratic ideals in different societies.
-The consideration of structural factors, both individual and societal, will allow us to advance future hypotheses relating to meritocratic legitimacy to a more comprehensive level than the current one, which is based solely on meritocratic perceptions.
-Furthermore, the impact of different configurations of the four-dimensional framework on practices and behaviors such as corruption, civic involvement, and political alignment, is an area that requires additional research.
-Such future agendas could be especially relevant in times of economic crisis and growing inequalities, that could entail changes in the legitimation of the current distributive structure based on meritocratic ideals.
+Considering individual and societal structural factors affecting meritocracy will allow us to advance future hypotheses relating to meritocratic legitimacy to a more comprehensive level than the current one, which is based mostly on meritocratic perceptions.
+Furthermore, the impact of different configurations of the four-dimensional framework on practices and behaviors, such as corruption, civic involvement, and political alignment, are areas open to further research.
+Such future agendas could be especially relevant in times of economic crisis and growing inequalities, which could entail changes in the legitimation of the current distributive structure based on meritocratic ideals.
+
+# Transparency statement
+
+This research follows a series of open science guidelines as the pre-registration of Study 2 (scale development) as well as the availability of data and codes (R) in a public repository in Github.
+This document was generated with the R package Rmarkdown and contains the code for all the tables and figures to make it reproducible.
 
@@
 
-# (APPENDIX) Appendix {-}
+<!-- # Appendix {-} -->
 
@@
 
-## Appendix study 1
+**Study 1**
+
+```{r countries, echo=FALSE}
+  load(file = "input/data/proc/data_s1.RData")
+  if (!require("pacman")) install.packages("pacman")
+  pacman::p_load(dplyr, kableExtra, knitr,lavaan,countrycode,ggplot2,haven,
+                 gridExtra)
+  table_format = if(knitr::is_html_output()) { #conditional instructions for kable
+    "html" #if html set "html" in format
+  } else if(knitr::is_latex_output()) {
+    "latex"#if latex set "latex" in format
+  }
+  load(file = "input/data/proc/data_s1.RData")
+  if (!require("pacman")) install.packages("pacman")
+  pacman::p_load(dplyr, kableExtra, knitr,lavaan,countrycode,ggplot2,haven,
+                 gridExtra)
+  table_format = if(knitr::is_html_output()) { #conditional instructions for kable
+    "html" #if html set "html" in format
+  } else if(knitr::is_latex_output()) {
+    "latex"#if latex set "latex" in format
+  }
+ # meassurement model: Second order
+  cfa_perpref2 <- '
+  	  # latent variables
+  	  percmerit =~ hwork + ambition
+  	  parents =~ wealthy + pareduc
+  	  backgrd =~ race + gender
+  	  networks=~ people + polcone
+  	  percnmerit=~ parents + backgrd + networks
+  	  prefmerit =~ welljob + hardjob
+  	  prefnmerit=~ family + child
+  	  '
+  # Model fit: second order
+
+  fit_perpref2 <- lavaan::cfa(cfa_perpref2, data = data_s1,ordered = F)
+
+  df_proc<- data_s1 %>% select(id,v5,wealthy:hardjob)  
+  df_proc$ccode <- as.numeric(data_s1$v5)
+
+  predicted2 <- lavPredict(fit_perpref2,newdata = data_s1)
+
+  idx <- lavInspect(fit_perpref2, "case.idx")
+  fscores <- lavPredict(fit_perpref2)
+  ## loop over factors
+  for (fs in colnames(fscores)) {
+    df_proc[idx, fs] <- fscores[ , fs]
+  }
+
+  df_proc <-  df_proc %>%  select(id,ccode,percmerit,prefmerit,prefnmerit) %>% na.omit()
+
+  df_proc$clab <- countrycode(sourcevar = df_proc$ccode,
+                            origin = "iso3n",
+                            destination = "country.name")
+avg_count <-  
+  df_proc %>% group_by(clab) %>% summarise(N=n())
+avg_count <- arrange(avg_count)
+
+# sum(avg_count[, 'N'])
+  col.nam <- c("Country","N", "Country", "N")
+  footnote <- paste0("N = ", sum(avg_count[, 'N']),". Source: International Social Survey Programme (2009).")
+
+  knitr::kable(bind_cols(avg_count[1:20,],avg_count[21:40,]),
+               col.names = col.nam, align = c("l","l","l","l"),linesep = "",
+        booktabs=TRUE,format = table_format,escape = FALSE,
+        caption = "Sample of Study 1 ") %>%
+    kableExtra::kable_styling(full_width = FALSE,
+                              latex_options = "HOLD_position",
+                              bootstrap_options=c("striped", "bordered"),
+                              font_size = 10) %>%
+    kableExtra::column_spec(column = c(2,4),width = "1cm") %>% 
+    kableExtra::column_spec(column = c(1,3),width = "2.5cm") %>% 
+    column_spec(3, border_left = T) %>% 
+    kableExtra::footnote(general = footnote, footnote_as_chunk = T,threeparttable = T)
+```
+
+\pagebreak
+
+**Study 2**
 
@@
 | Non university      | 32,63%  | 28,93% |
-| university or more  | 21,21%  | 18,13% |
-Table: `r as.character(paste("(\\#tab:rep-samp)", "Representativeness of the study 1 sample"))`.
-
-\pagebreak
-
-```{r tb-perpref1, echo=FALSE}
+| University or more  | 21,21%  | 18,13% |
+Table: `r as.character(paste("(\\#tab:rep-samp)", "Representativeness of the study 2 sample"))`.
+
+```{r tb-perpref1, eval=FALSE, include=FALSE}
  if (!require("pacman")) install.packages("pacman")
@@
 
-## Appendix study 3
-
-```{r fcloads-inv, echo=FALSE}
+<!-- **Study 3** -->
+
+```{r fcloads-inv, eval=FALSE, include=FALSE}
  if (!require("pacman")) install.packages("pacman")

### 2023-05-18 processing/preparation_study1_2019.Rmd
submission IJS

--- 2021-07-08 processing/preparation_study1.Rmd
@@
 ---
-title: "Data preparation: Study 1"
+title: "Data preparation: Study 1_2019"
 date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
@@
 ---
+
+## Note: this is an update  of the original study_1 (which used 2009 data) Here we use the last available file from wave 2019.
 
@@
 ```
-
-This is the preparation code of the data for Study 1: Measuring meritocracy with international secondary data. The prepared data is `data_s1.RData`
+This is the preparation code of the data for Study 1_2019: Measuring meritocracy with international secondary data. The prepared data is `data_s1_2019.RData`
 
@@
 ```{r}
-issp2009 <- read_dta("input/data/original/ZA5400_v4-0-0.dta")
+issp2019 <- read_dta("input/data/original/ZA7600_v3-0-0.dta")
 ```
@@
 ```{r }
-names(issp2009) <- tolower(names(issp2009))
-# Check & generate IDs (v4: respondent, v5: country)
-tapply(issp2009$v4, issp2009$v5, mean, na.rm = TRUE)
+names(issp2019) <- tolower(names(issp2019))
+find_var(data = issp2019,"country") # country and c_sample
+find_var(data = issp2019,"id") #caseid
+dplyr::n_distinct(issp2019$caseid) # ok, id única por caso
+
+# Check & generate IDs (caseid: respondent, country: country)
+tapply(issp2019$country, issp2019$caseid, mean, na.rm = TRUE)
 # Algunos paises (ej:826) tienen ids raros ...
 # Generar nueva id secuencial individuos por país para merges
-issp2009$id <- sequence(tabulate(issp2009$v5))
-tapply(issp2009$id, issp2009$v5, mean, na.rm = TRUE) # ok
-summary(issp2009$id)
-issp2009$idtem = issp2009$v5 * 10000
-issp2009$id = issp2009$idtem + issp2009$id
+issp2019$id <- sequence(tabulate(issp2019$country))
+tapply(issp2019$id, issp2019$country, mean, na.rm = TRUE) # ok
+summary(issp2019$id)
+issp2019$idtem = issp2019$country * 10000
+issp2019$id = issp2019$idtem + issp2019$id
 ```
@@
 
-*  v6: wealthy family
-*  v7: educated parents
-*  v9: ambition 
-* v10: hard work 
-* v11: right people
-* v12: political connections
-* v14: race 
-* v16: gender 
+*  v1: wealthy family
+*  v2: educated parents
+*  XX: ambition 
+*  v4: hard work 
+*  v5: right people
+*  v6: political connections
+*  v8: race 
+* v10: gender 
 * IDs: v5 (country) & id (individual id)

### 2023-05-18 paper.Rmd
ajustes template.tex

@@
 link-citations: yes
-params:
-  lang: en
+# params:
+# lang: en-GB
 author: 
@@
   - \usepackage{threeparttable}
-  - \usepackage[british,UKenglish,USenglish,american]{babel}
+  # - \usepackage[british,UKenglish,USenglish,american]{babel}
 always_allow_html: true # for word output

### 2023-09-07 paper-blinded-RR-IJS.Rmd
- new file paper-blinded-RR-IJS for resubmission to IJS

--- 2023-05-18 paper-blinded.Rmd
@@
 ---
-title: "Perceptions and Preferences for Meritocracy in Survey Research"
+title: "Measuring meritocratic beliefs with the ISSP social inequality survey: Possibilities, limitations and alternatives"
 #css: "custom.css"
@@
 One factor suggested to explain the persistence of inequality is the belief in meritocracy, which justifies economic disparities based on differences in effort and talent.
-However, empirical research on meritocracy is relatively new and characterized by diverse conceptualizations and measures, leading to inconsistent findings and interpretations.
-To address this issue, the present paper proposes a conceptual and measurement framework for perceptions and preferences regarding meritocracy and non-meritocracy, focusing on comparative survey research methodologies.
-The proposal is first tested through confirmatory factor analysis (CFA) using data from the International Social Survey Programme (ISSP) inequality module, commonly used in comparative research on meritocracy.
-Based on these findings,  a second study focuses on developing a novel scale that better captures the distinction between perceptions and preferences related to meritocracy.
-This scale is tested using a sample of 2,141 individuals from Chile.
-The results of this study contribute to refining the measurement of meritocracy and non-meritocracy and shed light on the importance of considering different dimensions when studying meritocracy in survey research, both conceptually and empirically.
+A great part of the international research on meritocratic beliefs has used the social inequality module of the ISSP (International Social Survey Programme), which for the first time has offered the possibility of comparing different societies in this regard.
+However, this research line is still a novel enterprise characterized by diverse conceptualizations and measures, often leading to inconsistent findings and interpretations.
+Based on a series of studies using ISSP data, the present paper proposes a multidimensional measurement framework for meritocratic beliefs that allows us to classify previous research contributions as well as to identify some vacuums and limitations.
+Firstly, we test the multidimensional framework for meritocratic beliefs using the ISSP 2009 inequality module through confirmatory factor analysis (CFA) techniques.
+After identifying the advantages and some limitations of ISSP for a multidiemsional operationalization of meritocratic belifs, in a second study we propose an expanded set of items that better taps the different dimensions of meritocratic beliefs.
+This measurement instrument was applied to a Chilean sample (N=2,141), offering new evidence regarding the multidimensional character of meritocratic beliefs.
   \newline

### 2023-09-10 paper-blinded-RR-IJS.Rmd
ajustes RR -IJS

@@
   Economic and social inequalities are causing growing concerns and crises in contemporary societies.
-One factor suggested to explain the persistence of inequality is the belief in meritocracy, which justifies economic disparities based on differences in effort and talent.
-A great part of the international research on meritocratic beliefs has used the social inequality module of the ISSP (International Social Survey Programme), which for the first time has offered the possibility of comparing different societies in this regard.
-However, this research line is still a novel enterprise characterized by diverse conceptualizations and measures, often leading to inconsistent findings and interpretations.
-Based on a series of studies using ISSP data, the present paper proposes a multidimensional measurement framework for meritocratic beliefs that allows us to classify previous research contributions as well as to identify some vacuums and limitations.
+One factor suggested to explain the persistence of inequality is the belief in meritocracy, which would justify economic disparities based on differences in effort and talent.
+A great part of the international research that has attempted to measure meritocratic beliefs has used the social inequality module of the ISSP (International Social Survey Programme), which offers an unprecedented opportunity to compare meritocratic views in different societies.
+However, this research agenda is still a novel enterprise characterized by diverse conceptualizations and measures of meritocracy, often leading to inconsistent findings and interpretations.
+Based on a series of studies using ISSP data, the present paper proposes a multidimensional measurement framework for meritocratic beliefs that allows us to classify previous research contributions as well as to identify some vacuums and limitations in the extant literature.
 Firstly, we test the multidimensional framework for meritocratic beliefs using the ISSP 2009 inequality module through confirmatory factor analysis (CFA) techniques.
-After identifying the advantages and some limitations of ISSP for a multidiemsional operationalization of meritocratic belifs, in a second study we propose an expanded set of items that better taps the different dimensions of meritocratic beliefs.
-This measurement instrument was applied to a Chilean sample (N=2,141), offering new evidence regarding the multidimensional character of meritocratic beliefs.
+After identifying the advantages and some limitations of ISSP for a multidimensional operationalization of meritocratic beliefs, in a second study, we propose an expanded set of items that better taps the different dimensions of meritocracy and that is suitable for survey research.
+We examined the measuring properties of this new instrument using a sample of Chilean adults (N=2,141).
+While the instrument was tested in the limited context of a single country, the findings offer  new evidence  about the multidimensional character of meritocratic beliefs.
+Based on these results, we suggest improvements in measuring meritocratic beliefs in cross-national studies.
   \newline
@@
   - \usepackage{times}           # Times New Roman
-  - \hypersetup {colorlinks = true, linkcolor = blue, urlcolor = blue}    
+#  - \hypersetup {colorlinks = true, linkcolor = blue, urlcolor = blue}    
   - \usepackage{caption}
@@
  # clean #in the yml
- rmarkdown::render("paper-blinded.Rmd", output_format = "bookdown::pdf_document2")
- rmarkdown::render("paper-blinded.Rmd", output_format = "bookdown::html_document2")
- rmarkdown::render("paper-blinded.Rmd", output_format = "bookdown::word_document2")
+ rmarkdown::render("paper-blinded-RR-IJS.Rmd", output_format = "bookdown::pdf_document2")
+ rmarkdown::render("paper-blinded-RR-IJS.Rmd", output_format = "bookdown::html_document2")
+ rmarkdown::render("paper-blinded-RR-IJS.Rmd", output_format = "bookdown::word_document2")
 ```
@@
 In terms of the association between the indicators, Figure \@ref(fig:corr-issp) shows the polychoric correlation matrix.
-Firstly, we observe in general that the moderate to high associations are between the pairs of items representing each of the dimensions as described in Table \@ref(tab:table-issp).
+Firstly, according to the criteria suggested by @cohen_statistical_1988 for effect size, we observe in general that the moderate to high associations are between the pairs of items representing each of the dimensions as described in Table \@ref(tab:table-issp).
 The exception here is the dimension of non-meritocratic perception, in which six items appear mostly associated by pairs according to their specific topics (family background, personal background, and connections).
-Still, between this same set of items, moderate correlations (around 0.3) could anticipate a single latent factor underlying non-meritocratic perception, which is tested next through confirmatory factor analysis.
+Still, between this same set of items, moderate correlations could anticipate a single latent factor underlying non-meritocratic perception, which is tested next through confirmatory factor analysis.
 
@@
 
-Regarding the correlations between the factors in Figure \@ref(fig:meas02-issp), we observe that perceptions are correlated with preferences, but more strongly for the meritocratic ($r=0.46$) than for the non-meritocratic ($r=0.27$) dimensions.
-Secondly, both perceptual dimensions (meritocratic and non-meritocratic) depict a moderate correlation ($r=0.30$), suggesting that they are not the opposite poles of the same continuum.
-Something similar occurs for the correlation between the two dimensions of preferences ($r=0.25$).
+Regarding the correlations between the factors in Figure \@ref(fig:meas02-issp), we observe that perceptions are correlated with preferences, but more strongly for the meritocratic ($r=0.46, p<0.01$) than for the non-meritocratic ($r=0.27,p<0.01$) dimensions.
+Secondly, both perceptual dimensions (meritocratic and non-meritocratic) depict a moderate correlation ($r=0.30, p<0.01$), suggesting that they are not the opposite poles of the same continuum.
+Something similar occurs for the correlation between the two dimensions of preferences ($r=0.25,p<0.01$).
 
 Based on the measurement validation, we asses if the proposed measurement strategy can be used in cross-national comparisons.
-In this line, Figure \@ref(fig:scat1) depicts the correlation of factor scores for Meritocratic Perception and Meritocratic Preferences at the country level.
+In this line, Figure \@ref(fig:scat1) depicts the correlation of factor scores for Meritocratic Perception and Meritocratic Preferences at the country level ($r=0.65,p<0.01$).
 For instance, Bulgaria (BGR) and the United States (USA) appropriately represent the positive association between perception and preferences.
@@
 
-Another finding depicted by Figure \@ref(fig:scat1) is the positive association of Non-meritocratic perception with Meritocratic preferences, showing that countries like Bulgaria (BGR) and Philipines (PHL) represent cases with higher Non-meritocratic perceptions and higher preferences for meritocracy.
+Another finding depicted by Figure \@ref(fig:scat1) is the positive association of Non-meritocratic perception with Meritocratic preferences ($r=0.55,p<0.01$), showing that countries like Bulgaria (BGR) and Philipines (PHL) represent cases with higher Non-meritocratic perceptions and higher preferences for meritocracy.
 In other words, in societies where people perceive higher importance of non-meritocratic aspects to get ahead in life, people also prefer that merit should be a determinant factor in wages.
@@
 
-  df_proc <-  df_proc %>%  select(id,ccode,percmerit,prefmerit,prefnmerit) %>% na.omit()
+  df_proc <-  df_proc %>%  select(id,ccode,percmerit,prefmerit,prefnmerit,percnmerit) %>% na.omit()
 
@@
                                             percmerit=mean(percmerit),
-                                            prefnmerit=mean(prefnmerit))
+                                            prefnmerit=mean(prefnmerit),
+                                           percnmerit=mean(percnmerit),
+                                           pop=n()
+                                           )
+
+# idea: por continentes o regiones del mundo
+# avg_count %>% filter(clab %in% c("ARG","VEN","CHL"))
+
+avg_count$continent <- countrycode(sourcevar = avg_count$clab,
+                            origin = "iso3c",
+                            destination = "continent")
+
+avg_count$region <- countrycode(sourcevar = avg_count$clab,
+                            origin = "iso3c",
+                            destination = "region23")
+
+avg_count$country <- countrycode(sourcevar = avg_count$clab,
+                            origin = "iso3c",
+                            destination = "country.name")
 
@@
   geom_text(label=as.character(avg_count$clab)) +
-  geom_smooth(method = "lm") +
+  scale_x_continuous(limits = c(-0.6,0.6))+
+  scale_y_continuous(limits = c(-0.6,0.6))+
+  geom_smooth(method = "lm",fullrange=TRUE) +
   xlab("Meritocratic perception")+
@@
 figure2 <- 
-ggplot(avg_count, aes(x=prefnmerit, y=prefmerit)) +
+ggplot(avg_count, aes(x=percnmerit, y=prefmerit)) +
   # geom_point() + 
   geom_text(label=as.character(avg_count$clab)) +
+  scale_x_continuous(limits = c(-0.6,0.6))+
+  scale_y_continuous(limits = c(-0.6,0.6))+ 
+  geom_smooth(method = "lm",fullrange=TRUE) +  
   geom_smooth(method = "lm") +
@@
                                      hjust = 0.1,
+                                     gp = grid::gpar(fontface = 1, fontsize = 9)))
+```
+
+On one hand, when comparing societies based on their level of industrialization [@ishida_trends_2005], Figure \@ref(fig:scat2) illustrates the correlations within the sample.
+It shows that early-industrialized countries exhibit a similar association between Meritocratic Perception and Meritocratic Preference ($r=0.67, p<0.05$), a trend consistent within the group of late-industrialized nations ($r=0.68, p<0.01$).
+This suggests that citizens' perceptions regarding the significance of merit in both industrialized and late-industrialized nations tend to have a similar impact on their beliefs about the role of merit in determining wages.
+On the other hand, the association between Non-meritocratic Perception and Meritocratic Preference displays significant differences when compared to the complete sample.
+Industrialized nations exhibit a strong positive correlation ($r=0.75, p<0.01$).
+In contrast, late-industrialized countries show a weak and non-significant correlation ($r=0.29, p>0.05$).
+In other words, the non-meritocratic aspect holds greater salience in industrialized societies when it comes to citizens' opinions regarding the determinants of wages.
+However, in late-industrialized societies, the importance of non-meritocratic factors diminishes in influencing meritocratic preferences.
+This evidence suggests that in societies with a long history of transformations in their social structure, the perceived importance of both meritocratic and non-meritocratic dimensions in achieving success reinforces people's preferences for how status positions should be distributed.
+In contrast, in countries that experienced changes in their occupational structure during the early decades of the 20th century, meritocratic perception holds greater prominence than non-meritocratic factors such as personal contacts or family background in shaping citizens' opinions regarding the determinants of wages.
+
+```{r scat2, fig.cap="Cross-national comparison between industrialized and late-industrialized",fig.height=10, fig.width=10}
+  load(file = "input/data/proc/data_s1.RData")
+  if (!require("pacman")) install.packages("pacman")
+  pacman::p_load(dplyr, kableExtra, knitr,lavaan,countrycode,ggplot2,haven,
+                 gridExtra)
+  table_format = if(knitr::is_html_output()) { #conditional instructions for kable
+    "html" #if html set "html" in format
+  } else if(knitr::is_latex_output()) {
+    "latex"#if latex set "latex" in format
+  }
+
+  # meassurement model: Second order
+  cfa_perpref2 <- '
+  	  # latent variables
+  	  percmerit =~ hwork + ambition
+  	  parents =~ wealthy + pareduc
+  	  backgrd =~ race + gender
+  	  networks=~ people + polcone
+  	  percnmerit=~ parents + backgrd + networks
+  	  prefmerit =~ welljob + hardjob
+  	  prefnmerit=~ family + child
+  	  '
+  # Model fit: second order
+
+  fit_perpref2 <- lavaan::cfa(cfa_perpref2, data = data_s1,ordered = F)
+
+  df_proc<- data_s1 %>% select(id,v5,wealthy:hardjob)  
+  df_proc$ccode <- as.numeric(data_s1$v5)
+
+  predicted2 <- lavPredict(fit_perpref2,newdata = data_s1)
+
+  idx <- lavInspect(fit_perpref2, "case.idx")
+  fscores <- lavPredict(fit_perpref2)
+  ## loop over factors
+  for (fs in colnames(fscores)) {
+    df_proc[idx, fs] <- fscores[ , fs]
+  }
+
+  df_proc <-  df_proc %>%  select(id,ccode,percmerit,prefmerit,prefnmerit,percnmerit) %>% na.omit()
+
+  df_proc$clab <- countrycode(sourcevar = df_proc$ccode,
+                            origin = "iso3n",
+                            destination = "iso3c")
+
+avg_count <-  
+  df_proc %>% group_by(clab) %>% summarise(prefmerit=mean(prefmerit),
+                                            percmerit=mean(percmerit),
+                                            prefnmerit=mean(prefnmerit),
+                                           percnmerit=mean(percnmerit),
+                                           pop=n()
+                                           )
+
+# idea: por continentes o regiones del mundo
+# avg_count %>% filter(clab %in% c("ARG","VEN","CHL"))
+
+avg_count$continent <- countrycode(sourcevar = avg_count$clab,
+                            origin = "iso3c",
+                            destination = "continent")
+
+avg_count$region <- countrycode(sourcevar = avg_count$clab,
+                            origin = "iso3c",
+                            destination = "region23")
+
+avg_count$country <- countrycode(sourcevar = avg_count$clab,
+                            origin = "iso3c",
+                            destination = "country.name")
+
+western <- 
+avg_count %>%
+  filter(continent=="Europe"| clab %in% c("USA","AUS","NZL"))
+
+nonwestern <- 
+avg_count %>%
+  filter(continent=="Asia" | clab %in% c("VEN","CHL","ARG","ZAF"))
+
+indus <-  avg_count %>% 
+  filter(region %in% c("Northern America","Western Europe"))
+
+lateindus <-  avg_count %>% 
+  filter(!region %in% c("Northern America","Western Europe"))
+
+indus <-  avg_count %>% 
+  filter(country %in% c("Australia","Austria","Belgium","Switzerland","Germany","Denmark","Finland","France","United Kingdom","Iceland","Norway", "Sweden","United States"))
+
+lateindus <-  avg_count %>% 
+  filter(!country %in% c("Australia","Austria","Belgium","Switzerland","Germany","Denmark","Finland","France","United Kingdom","Iceland","Norway", "Sweden","United States"))
+
+# Western countries-------------------------------------------------------------
+wesmerit <- 
+indus %>%  
+ggplot(aes(x=percmerit, y=prefmerit)) +
+  # geom_point() + 
+  geom_text(label=as.character(indus$clab)) +
+  scale_x_continuous(limits = c(-0.6,0.6))+
+  scale_y_continuous(limits = c(-0.6,0.6))+
+  geom_smooth(method = "lm",fullrange=TRUE) +
+  xlab("Meritocratic perception")+
+  ylab("Meritocratic preference") +
+  ggtitle("Industrialized countries")
+wesnonmer <- 
+indus %>%  
+ggplot(aes(x=percnmerit, y=prefmerit)) +
+  # geom_point() + 
+  geom_text(label=as.character(indus$clab)) +
+  scale_x_continuous(limits = c(-0.6,0.6))+
+  scale_y_continuous(limits = c(-0.6,0.6))+ 
+  geom_smooth(method = "lm",fullrange=TRUE) +
+  xlab("Non-meritocratic perception")+
+  ylab("Meritocratic preference") +
+  ggtitle("Industrialized countries")
+
+# Non western countries --------------------------------------------------------
+nonwesmer <- 
+lateindus %>%  
+ggplot(aes(x=percmerit, y=prefmerit)) +
+  # geom_point() + 
+  geom_text(label=as.character(lateindus$clab)) +
+  scale_x_continuous(limits = c(-0.6,0.6))+
+  scale_y_continuous(limits = c(-0.6,0.6))+
+  geom_smooth(method = "lm",fullrange=TRUE) +
+  xlab("Meritocratic perception")+
+  ylab("Meritocratic preference") +
+  ggtitle("Late industralized countries")
+nonwesnonmer <- 
+lateindus %>%  
+ggplot(aes(x=percnmerit, y=prefmerit)) +
+  # geom_point() + 
+  geom_text(label=as.character(lateindus$clab)) +
+  scale_x_continuous(limits = c(-0.6,0.6))+
+  scale_y_continuous(limits = c(-0.6,0.6))+
+  geom_smooth(method = "lm",fullrange=TRUE) +
+  xlab("Non-meritocratic perception")+
+  ylab("Meritocratic preference") +
+  ggtitle("Late industralized countries")
+
+grid.arrange(wesmerit,wesnonmer,
+             nonwesmer,nonwesnonmer, ncol = 2,
+             bottom = grid::textGrob(label = paste0("Source: Authors calculation based on International Social Survey Programme - Social Inequality 2009"),
+                                     hjust = 0.2,
                                      gp = grid::gpar(fontface = 1, fontsize = 9)))
@@
 There are three main aspects to highlight from this correlation matrix.
-Firstly, and as expected, the largest correlations are between indicators that correspond to the same factors behind the conceptual model (e.g., perception of meritocracy by effort and by talent, $r=0.52$).
-Secondly, among these correlations, the highest are those between the non-meritocratic dimension but in perceptions ($r=0.73$) and preferences ($r=0.61$).
+Firstly, and as expected, the largest correlations are between indicators that correspond to the same factors behind the conceptual model (e.g., perception of meritocracy by effort and by talent, $r=0.52,p<0.01$).
+Secondly, among these correlations, the highest are those between the non-meritocratic dimension but in perceptions ($r=0.73,p<0.01$) and preferences ($r=0.61,p<0.01$).
 Thirdly, both items for meritocratic preferences (E and F) are the ones that mostly correlate with the rest of the perceptual items, showing medium to high correlations and indicating that the perception of non-meritocracy would be related to larger meritocratic preferences.
@@
 
-Attending to the correlations between the latent variables as depicted in Figure \@ref(fig:meas02), meritocratic preferences are moderate to highly correlated with perceptions, both meritocratic ($r=0.457$) and non-meritocratic ($r=0.500$).
-The correlation between both types of perceptions ($r=-0.044$) and both types of preferences ($r=0.185$) was low, as well as between non-meritocratic preferences and perceptions ($r=-0.059$).
+Attending to the correlations between the latent variables as depicted in Figure \@ref(fig:meas02), meritocratic preferences are moderate to highly correlated with perceptions, both meritocratic ($r=0.457,p<0.01$) and non-meritocratic ($r=0.500,p<0.01$).
+The correlation between both types of perceptions ($r=-0.044,p>0.05$) and both types of preferences ($r=0.185,p<0.01$) was low, as well as between non-meritocratic preferences and perceptions ($r=-0.059p>0.05$).
 This last finding gives further evidence regarding the lack of unidimensionality of meritocratic and non-meritocratic aspects as assumed by previous studies.
@@
 Aware of this trade-off, next we present additional evidence of validity to complement some of the weaknesses identified.
-
-# Study 3: Additional validity analyses
-
-The modeling for invariance testing has been gaining more relevance in sociological survey studies, mainly due to the interest in the validity of measurement scales in comparative studies in various social and cultural contexts [@davidov_measurement_2014].
-In this section, we present the results of an analysis of invariance between groups, testing whether the proposed factorial structure for the meritocracy scale is applicable in a different dataset.
-
-The invariance measurement procedure consists of a series of nested models in which restrictions are progressively incorporated into the parameters of the measurement model.
-The literature generally suggests that this modeling approach should consider four levels or types of progressive restriction [@milfont_testing_2010; @millsap_statistical_2011; @vandeschoot_checklist_2012]:
-
-1. **Configural:** the model is estimated only indicating the factorial structure used in the CFA.
-
-2. **Weak:** an equality restriction applied to the factor loadings in the different groups; that is, the loadings are forced to be identical in both measurements.
-
-3. **Strong:** equality restrictions added to the intercepts of each indicator.
-
-4. **Strict:** equality restrictions added to the error variances of each indicator.
-
-### Data
-
-For testing the invariance, we compared the data described previously in Study 2 to a different dataset from an online survey carried out during the first half of 2020.
-The characteristics of this sample in terms of application and coverage are equivalent to those of the previous study.
-The final sample contains 1,242 cases, where 605 (48.71%) are women, and 637 (51.28%) are men.
-48.84% are 45 or older, and 33.97% have tertiary education or higher.
-
-### Variables
-
-The items of the meritocracy scale were identical to the original scale of Study 2 administrated in a randomized order (which corresponds to the application modality of Group 3 in Study 2, as shown in Figure \@ref(fig:appmod)).
-
-### Results
-
-The first step for invariance testing is estimating the configural model, which serves as the baseline for further comparison and is expected to adequately meet the global fit criteria of a measurement model.
-Although the chi-square statistic is used as a global measure of fit, it is usually complemented with other indexes given its high sensitivity to sample size: the Comparative Fit Index (CFI), which should have a value greater than 0.95; the Root Mean Square Error Approximation (RMSEA), which must be lower than 0.06; and the Standardized Root Mean Square Residuals (SRMR) which must be less than 0.08 [@vandeschoot_checklist_2012].
-Besides, the literature on measurement invariance suggests some complementary approaches for evaluating the fit, from which we will consider the incremental adjustment of the fit indexes [@cheung_evaluating_2002; @milfont_testing_2010; @dimitrov_testing_2010] and the ANOVA test for means comparison in nested models [@newsom_longitudinal_2015].
-
-```{r invargroup, echo=FALSE}
-  if (!require("pacman")) install.packages("pacman")
-  #---Load packages and data -------------------------------------#
-  pacman::p_load(dplyr,lavaan,semTools,knitr,kableExtra,stats,gtools)
-  options(knitr.kable.NA = '')
-  load(file = "input/data/proc/data_s3_inv.RData")
-  pov01 <- data_s3_inv %>% dplyr::filter(dataset=="pvw01")
-  fs01  <- data_s3_inv %>% dplyr::filter(dataset=="fsw02")
-
-  #-----Meassurement model--------------------------------------#
-  model01 <- '
-  perc_merit  =~ perc_effort + perc_talent
-  perc_nmerit =~ perc_wpart  + perc_netw
-  pref_merit  =~ pref_effort + pref_talent
-  pref_nmerit =~ pref_wpart  + pref_netw'
-
-  #-----Invariance---------------------------------------------#
-  inv01<- semTools::measurementInvariance(model=model01,data=data_s3_inv,group="dataset",estimator = "ML",strict=TRUE,quiet = T)
-
-  conf  <- inv01$fit.configural
-  weak  <- inv01$fit.loadings
-  strong<- inv01$fit.intercepts
-  strict<- inv01$fit.residuals
-
-  tab01<- lavaan::anova(conf,weak,strong,strict,SB.classic=TRUE) %>%
-    dplyr::as_tibble() %>%
-    dplyr::select("Chisq","Df","chisq_diff"=`Chisq diff`,"df_diff"=`Df diff`,"pvalue"=`Pr(>Chisq)`) %>%
-    dplyr::mutate(stars=gtools::stars.pval(pvalue),
-           chisqt=paste0(round(Chisq,2)," (",Df,") "),
-           decision=ifelse(pvalue>0.05,yes = "Accept",no = "Reject"),
-           model=c("Configural","Weak","Strong","Strict"))
-
-  fit.meas<- dplyr::bind_rows(lavaan::fitmeasures(inv01$fit.configural,output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
-                              lavaan::fitmeasures(inv01$fit.loadings,  output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
-                              lavaan::fitmeasures(inv01$fit.intercepts,output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
-                              lavaan::fitmeasures(inv01$fit.residuals, output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),])
-
-  # compute differences in chisq, df, cfi and rmsea (90%, lower.ci - upper.ci )
-  fit.meas<- fit.meas %>%
-    dplyr::mutate(diff.chi2 = chisq    - lag(chisq,default = dplyr::first(chisq)),
-           diff.df   = df       - lag(df,   default = dplyr::first(df)),
-           diff.cfi  = cfi      - lag(cfi,  default = dplyr::first(cfi)),
-           diff.rmsea   = rmsea - lag(rmsea,default = dplyr::first(rmsea))) %>%
-    round(3) %>%
-    dplyr::mutate(rmsea.ci=paste0(rmsea," \n ", "(",rmsea.ci.lower,"-",rmsea.ci.upper,")"))
-
-  tab.inv<- dplyr::bind_cols(tab01,fit.meas) %>%
-    dplyr::select(model,chisqt,cfi,rmsea.ci,diff.chi2,diff.df,diff.cfi,diff.rmsea,stars,decision) %>%
-    dplyr::mutate(diff.chi2=paste0(diff.chi2," (",diff.df,") ",stars)) %>%
-    dplyr::select(model,chisqt,cfi,rmsea.ci,diff.chi2,diff.cfi,diff.rmsea,decision)
-
-  #clean values
-  tab.inv[tab.inv == c("0 (0) ")] <- NA
-  tab.inv[tab.inv == c(0)] <- NA
-
-  ### Check an alternative method
-  configural        <- lavaan::cfa(model01, data=data_s3_inv, group = "dataset", estimator="ML")
-  weak.invariance   <- lavaan::cfa(model01, data=data_s3_inv, group = "dataset",group.equal = "loadings", estimator="ML")
-  strong.invariance <- lavaan::cfa(model01, data=data_s3_inv, group = "dataset",group.equal = c( "loadings", "intercepts"), estimator="ML")
-  strict.invariance <- lavaan::cfa(model01, data=data_s3_inv, group = "dataset",group.equal = c( "loadings", "intercepts", "residuals"), estimator="ML")
-
-  table_format = if(is_html_output()) {
-    "html"
-  } else if(is_latex_output()) {
-    "latex"
-  }
-
-  col.nam <- c("Model","$\\chi^2 (\\text{df})$","CFI","RMSEA (90 CI)",
-               "$\\Delta \\chi^2 (\\Delta \\text{df}$)","$\\Delta \\text{CFI}$","$\\Delta \\text{RMSEA}$","Decision")
-  footnote <- paste0("N = ","; Group 1, n = ",conf@Data@nobs[[1]],"; Group 2, n = ",conf@Data@nobs[[2]], ", ***p < 0.001")
-
-  knitr::kable(tab.inv, col.names = col.nam, align = "l",
-        booktabs=TRUE,format = table_format,escape = FALSE,
-        caption = "Multiple Group measurement invariance for Perceptions and Preferences for Meritocracy") %>%
-    kableExtra::kable_styling(full_width = FALSE,
-                              latex_options = "HOLD_position",
-                              bootstrap_options=c("striped", "bordered"),
-                              font_size = 10) %>%
-    kableExtra::footnote(general = footnote, footnote_as_chunk = T)
-```
-
-Table \@ref(tab:invargroup) shows the results of the measurement invariance estimation.
-When attending to the traditional invariance test of $\Delta \chi^2 (\Delta \text{df})$, the results support the invariance at the strong level meaning that the fit of the factor model of the merit scale is equivalent across samples when constraining factor loadings and intercepts to being equal.
-Such result is considered in general as evidence of invariance [@fischer_are_2011], as strict forms of measurement invariance rarely hold [@vandeschoot_checklist_2012].
-Still, the comparability of latent means requires strict invariance which in this case does not hold when considering just $\Delta \chi^2 (\Delta \text{df})$.
-Nevertheless, the criteria of $\Delta \text{CFI}$ used for comparing models is close to the rejection criteria of >.01, whereas the $\Delta \text{RMSEA}$ fulfills the requirements of being below the cut-off criteria as suggested by [@chen_sensitivity_2007].
-Therefore, using this last standard, the level of strict invariance would hold for the meritocracy scale.
-
-## Discusion Study 3
-
-Based on the measurement invariance analysis for two datasets, we found evidence supporting the equivalence of the scale when applied in different samples.
-This gives a more robust ground when using this measure for comparing groups, for instance in country comparisons.
-Although the results are promising, still the analysis was performed in two samples in a single country (Chile) and it requires further examination with social surveys in other societies.
 
@@
 Table: `r as.character(paste("(\\#tab:rep-samp)", "Representativeness of the study 2 sample"))`.
+
+**Study 3: Additional validity analyses**
+
+The modeling for invariance testing has been gaining more relevance in sociological survey studies, mainly due to the interest in the validity of measurement scales in comparative studies in various social and cultural contexts [@davidov_measurement_2014].
+In this section, we present the results of an analysis of invariance between groups, testing whether the proposed factorial structure for the meritocracy scale is applicable in a different dataset.
+
+The invariance measurement procedure consists of a series of nested models in which restrictions are progressively incorporated into the parameters of the measurement model.
+The literature generally suggests that this modeling approach should consider four levels or types of progressive restriction [@milfont_testing_2010; @millsap_statistical_2011; @vandeschoot_checklist_2012]:
+
+1. **Configural:** the model is estimated only indicating the factorial structure used in the CFA.
+
+2. **Weak:** an equality restriction applied to the factor loadings in the different groups; that is, the loadings are forced to be identical in both measurements.
+
+3. **Strong:** equality restrictions added to the intercepts of each indicator.
+
+4. **Strict:** equality restrictions added to the error variances of each indicator.
+
+**Data**
+
+For testing the invariance, we compared the data described previously in Study 2 to a different dataset from an online survey carried out during the first half of 2020.
+The characteristics of this sample in terms of application and coverage are equivalent to those of the previous study.
+The final sample contains 1,242 cases, where 605 (48.71%) are women, and 637 (51.28%) are men.
+48.84% are 45 or older, and 33.97% have tertiary education or higher.
+
+**Variables**
+
+The items of the meritocracy scale were identical to the original scale of Study 2 administrated in a randomized order (which corresponds to the application modality of Group 3 in Study 2, as shown in Figure \@ref(fig:appmod)).
+
+**Results**
+
+The first step for invariance testing is estimating the configural model, which serves as the baseline for further comparison and is expected to adequately meet the global fit criteria of a measurement model.
+Although the chi-square statistic is used as a global measure of fit, it is usually complemented with other indexes given its high sensitivity to sample size: the Comparative Fit Index (CFI), which should have a value greater than 0.95; the Root Mean Square Error Approximation (RMSEA), which must be lower than 0.06; and the Standardized Root Mean Square Residuals (SRMR) which must be less than 0.08 [@vandeschoot_checklist_2012].
+Besides, the literature on measurement invariance suggests some complementary approaches for evaluating the fit, from which we will consider the incremental adjustment of the fit indexes [@cheung_evaluating_2002; @milfont_testing_2010; @dimitrov_testing_2010] and the ANOVA test for means comparison in nested models [@newsom_longitudinal_2015].
+
+```{r invargroup, echo=FALSE}
+  if (!require("pacman")) install.packages("pacman")
+  #---Load packages and data -------------------------------------#
+  pacman::p_load(dplyr,lavaan,semTools,knitr,kableExtra,stats,gtools)
+  options(knitr.kable.NA = '')
+  load(file = "input/data/proc/data_s3_inv.RData")
+  pov01 <- data_s3_inv %>% dplyr::filter(dataset=="pvw01")
+  fs01  <- data_s3_inv %>% dplyr::filter(dataset=="fsw02")
+
+  #-----Meassurement model--------------------------------------#
+  model01 <- '
+  perc_merit  =~ perc_effort + perc_talent
+  perc_nmerit =~ perc_wpart  + perc_netw
+  pref_merit  =~ pref_effort + pref_talent
+  pref_nmerit =~ pref_wpart  + pref_netw'
+
+  #-----Invariance---------------------------------------------#
+  inv01<- semTools::measurementInvariance(model=model01,data=data_s3_inv,group="dataset",estimator = "ML",strict=TRUE,quiet = T)
+
+  conf  <- inv01$fit.configural
+  weak  <- inv01$fit.loadings
+  strong<- inv01$fit.intercepts
+  strict<- inv01$fit.residuals
+
+  tab01<- lavaan::anova(conf,weak,strong,strict,SB.classic=TRUE) %>%
+    dplyr::as_tibble() %>%
+    dplyr::select("Chisq","Df","chisq_diff"=`Chisq diff`,"df_diff"=`Df diff`,"pvalue"=`Pr(>Chisq)`) %>%
+    dplyr::mutate(stars=gtools::stars.pval(pvalue),
+           chisqt=paste0(round(Chisq,2)," (",Df,") "),
+           decision=ifelse(pvalue>0.05,yes = "Accept",no = "Reject"),
+           model=c("Configural","Weak","Strong","Strict"))
+
+  fit.meas<- dplyr::bind_rows(lavaan::fitmeasures(inv01$fit.configural,output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
+                              lavaan::fitmeasures(inv01$fit.loadings,  output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
+                              lavaan::fitmeasures(inv01$fit.intercepts,output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
+                              lavaan::fitmeasures(inv01$fit.residuals, output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),])
+
+  # compute differences in chisq, df, cfi and rmsea (90%, lower.ci - upper.ci )
+  fit.meas<- fit.meas %>%
+    dplyr::mutate(diff.chi2 = chisq    - lag(chisq,default = dplyr::first(chisq)),
+           diff.df   = df       - lag(df,   default = dplyr::first(df)),
+           diff.cfi  = cfi      - lag(cfi,  default = dplyr::first(cfi)),
+           diff.rmsea   = rmsea - lag(rmsea,default = dplyr::first(rmsea))) %>%
+    round(3) %>%
+    dplyr::mutate(rmsea.ci=paste0(rmsea," \n ", "(",rmsea.ci.lower,"-",rmsea.ci.upper,")"))
+
+  tab.inv<- dplyr::bind_cols(tab01,fit.meas) %>%
+    dplyr::select(model,chisqt,cfi,rmsea.ci,diff.chi2,diff.df,diff.cfi,diff.rmsea,stars,decision) %>%
+    dplyr::mutate(diff.chi2=paste0(diff.chi2," (",diff.df,") ",stars)) %>%
+    dplyr::select(model,chisqt,cfi,rmsea.ci,diff.chi2,diff.cfi,diff.rmsea,decision)
+
+  #clean values
+  tab.inv[tab.inv == c("0 (0) ")] <- NA
+  tab.inv[tab.inv == c(0)] <- NA
+
+  ### Check an alternative method
+  configural        <- lavaan::cfa(model01, data=data_s3_inv, group = "dataset", estimator="ML")
+  weak.invariance   <- lavaan::cfa(model01, data=data_s3_inv, group = "dataset",group.equal = "loadings", estimator="ML")
+  strong.invariance <- lavaan::cfa(model01, data=data_s3_inv, group = "dataset",group.equal = c( "loadings", "intercepts"), estimator="ML")
+  strict.invariance <- lavaan::cfa(model01, data=data_s3_inv, group = "dataset",group.equal = c( "loadings", "intercepts", "residuals"), estimator="ML")
+
+  table_format = if(is_html_output()) {
+    "html"
+  } else if(is_latex_output()) {
+    "latex"
+  }
+
+  col.nam <- c("Model","$\\chi^2 (\\text{df})$","CFI","RMSEA (90 CI)",
+               "$\\Delta \\chi^2 (\\Delta \\text{df}$)","$\\Delta \\text{CFI}$","$\\Delta \\text{RMSEA}$","Decision")
+  footnote <- paste0("N = ","; Group 1, n = ",conf@Data@nobs[[1]],"; Group 2, n = ",conf@Data@nobs[[2]], ", ***p < 0.001")
+
+  knitr::kable(tab.inv, col.names = col.nam, align = "l",
+        booktabs=TRUE,format = table_format,escape = FALSE,
+        caption = "Multiple Group measurement invariance for Perceptions and Preferences for Meritocracy") %>%
+    kableExtra::kable_styling(full_width = FALSE,
+                              latex_options = "HOLD_position",
+                              bootstrap_options=c("striped", "bordered"),
+                              font_size = 10) %>%
+    kableExtra::footnote(general = footnote, footnote_as_chunk = T)
+```
+
+Table \@ref(tab:invargroup) shows the results of the measurement invariance estimation.
+When attending to the traditional invariance test of $\Delta \chi^2 (\Delta \text{df})$, the results support the invariance at the strong level meaning that the fit of the factor model of the merit scale is equivalent across samples when constraining factor loadings and intercepts to being equal.
+Such result is considered in general as evidence of invariance [@fischer_are_2011], as strict forms of measurement invariance rarely hold [@vandeschoot_checklist_2012].
+Still, the comparability of latent means requires strict invariance which in this case does not hold when considering just $\Delta \chi^2 (\Delta \text{df})$.
+Nevertheless, the criteria of $\Delta \text{CFI}$ used for comparing models is close to the rejection criteria of >.01, whereas the $\Delta \text{RMSEA}$ fulfills the requirements of being below the cut-off criteria as suggested by [@chen_sensitivity_2007].
+Therefore, using this last standard, the level of strict invariance would hold for the meritocracy scale.
+
+*Discusion*
+
+Based on the measurement invariance analysis for two datasets, we found evidence supporting the equivalence of the scale when applied in different samples.
+This gives a more robust ground when using this measure for comparing groups, for instance in country comparisons.
+Although the results are promising, still the analysis was performed in two samples in a single country (Chile) and it requires further examination with social surveys in other societies.
 

### 2023-09-13 paper-blinded-RR-IJS.Rmd
update RR-IJS

@@
 ---
-title: "Measuring meritocratic beliefs with the ISSP social inequality survey: Possibilities, limitations and alternatives"
+title: "A multidimensional approach for measuring meritocratic beliefs with the ISSP social inequality survey: Possibilities, limitations and alternatives"
 #css: "custom.css"
@@
   Economic and social inequalities are causing growing concerns and crises in contemporary societies.
-One factor suggested to explain the persistence of inequality is the belief in meritocracy, which would justify economic disparities based on differences in effort and talent.
+One factor suggested explaining the persistence of inequality is the belief in meritocracy, which would justify economic disparities based on differences in effort and talent.
 A great part of the international research that has attempted to measure meritocratic beliefs has used the social inequality module of the ISSP (International Social Survey Programme), which offers an unprecedented opportunity to compare meritocratic views in different societies.
 However, this research agenda is still a novel enterprise characterized by diverse conceptualizations and measures of meritocracy, often leading to inconsistent findings and interpretations.
-Based on a series of studies using ISSP data, the present paper proposes a multidimensional measurement framework for meritocratic beliefs that allows us to classify previous research contributions as well as to identify some vacuums and limitations in the extant literature.
-Firstly, we test the multidimensional framework for meritocratic beliefs using the ISSP 2009 inequality module through confirmatory factor analysis (CFA) techniques.
-After identifying the advantages and some limitations of ISSP for a multidimensional operationalization of meritocratic beliefs, in a second study, we propose an expanded set of items that better taps the different dimensions of meritocracy and that is suitable for survey research.
+Based on a series of studies using ISSP data, the present paper proposes a multidimensional measurement framework for meritocratic beliefs.
+This framework distinguishes, on the one side, between perceptions and preferences, and on the other side, between meritocratic and not meritocratic aspects, a two-by-two scheme that identifies four dimensions of meritocratic research.
+This framework allows us to classify previous research contributions, as well as to identify some vacuums and limitations in the extant literature.
+In a first study, we test the multidimensional framework for meritocratic beliefs using the ISSP 2009 inequality module through confirmatory factor analysis (CFA) techniques.
+After identifying the advantages and some limitations of ISSP items for a multidimensional operationalization of meritocratic beliefs, in a second study, we suggest an expanded set of items that better taps the different dimensions of meritocracy.
 We examined the measuring properties of this new instrument using a sample of Chilean adults (N=2,141).
-While the instrument was tested in the limited context of a single country, the findings offer  new evidence  about the multidimensional character of meritocratic beliefs.
-Based on these results, we suggest improvements in measuring meritocratic beliefs in cross-national studies.
+While the instrument was tested in the limited context of a single country, the findings offer  aditional evidence about the validity of the multidimensional measurement model of meritocratic beliefs.
+Based on these results, we recommend improvements in measuring meritocratic beliefs in cross-national studies.
   \newline
@@
 In many cases, similar phenomena are associated with different indicators, and conversely, different phenomena are measured with similar indicators.
-Such approaches limit the comparability of studies and the capacity to understand the effects of meritocratic beliefs across different disciplines and lines of research.
-
-Based on a critical analysis of different approaches to the measurement of meritocracy to date, this article proposes a conceptual and measurement framework that considers two of the main aspects in the study of meritocracy: perceptions and preferences.
-Furthermore, it distinguishes between meritocratic and non-meritocratic dimensions, as they would not be two poles of the same continuum as previous studies suggest.
-This conceptual framework is oriented to develop a measurement instrument as simple and brief as possible, suitable for public opinion surveys and thus allowing to integrate meritocratic beliefs in studying different social phenomena.
+Such approaches limit the comparability of studies and the capacity to understand the effects of meritocratic beliefs across different disciplines and lines of research, raising doubts about the validity of the instruments used to measure meritocratic beliefs.
+In this sense, this paper is aligned with recent discussions about the estimand [@lundberg_what_2021] in sociology, this is, the need for a clear definition of what we are estimating as well as the connection of this definition with statistical evidence.
+
+Based on a critical analysis of different approaches to the empirical study of meritocracy to date —most of them using data from the inequality module of ISSP—, this article identifies several conceptual and measurement issues.
+In order to address them, firstly, we propose a multidimensional conceptual framework for measuring meritocratic beliefs.
+The proposal consists in a first basic distinction between the dimensions of perceptions and preferences for meritocracy, often confused in the literature.
+Furthermore, the proposal distinguishes between meritocratic and non-meritocratic dimensions, as they would not be two poles of the same continuum as previous studies suggest.
+Secondly, we test the validity of our conceptual framework with two studies.
+Study 1 offers an operationalization and confirmatory analysis of the proposed multidimensional framework using available items from ISSP (inequality module 2009).
+Based on the advantages and limitations of the results obtained in study 1, and in order to gain additional evidence for the validity of the measurement instruments, study 2 suggests modifications and expansion of the items commonly used to measure meritocratic beliefs.
+This measurement proposal is tested with data from a single country case (Chile).
 
@@
 
-The item used by @mijs_paradox_2019 is part of an item's battery which is available in several international surveys, usually called "reasons to get ahead." This battery displays a series of indicators related to what people consider important to get ahead: hard work, education, ambition, a wealthy family, the right connections, religion, race, and gender.
+The item used by @mijs_paradox_2019 is part of an item's battery which is available in several international surveys (as ISSP), usually called “reasons to get ahead”.
+This battery displays a series of indicators related to what people consider important to get ahead: hard work, education, ambition, a wealthy family, the right connections, religion, race, and gender.
 Therefore, considering only one of the items means that other aspects associated with talent, such as education, would not be deemed meritocratic.
@@
 
-In this conceptual and measurement approach of meritocratic beliefs, we can observe a couple of strong conceptual assumptions: a) effort would not depend on parental influence (at least not as strongly as intelligence), and b) talent (as innate ability) is not meritocratic (contrary to Michael Youngs' original conceptualization).
+In this conceptual and measurement approach of meritocratic beliefs, we can observe a couple of strong conceptual assumptions: a) effort would not depend on parental influence (at least not as strongly as intelligence), and b) talent (as innate ability) is not meritocratic (contrary to Michael Young's' original conceptualization).
 This conceptual and measurement-based assumption is found in other studies, which also assume effort to be the main and only aspect of meritocracy [@girerd_neoliberalism_2020; @bubak_perceptions_2019], raising the question: Is effort the only (or main) dimension behind the concept of meritocracy?
-We argue that whether talent (as intelligence and ability) is or is not considered meritocratic is an interesting topic to discuss from a philosophical point of view, but it also requires empirical evidence in studies focused on the subjective aspects of inequality.
-Furthermore, considering talent as part of meritocracy certainly opens some interesting avenues of research, for instance, some studies suggest that for the elites meritocracy is related to talent, whereas effort is more characteristic of the meritocracy of the middle and lower classes [@atria_economic_2020].
-Therefore, we propose that a concept (and measure) of meritocracy benefits from including both talent and effort as in the original Young's conception.
+We argue that whether talent (as intelligence and ability) is or is not considered meritocratic is an interesting topic to discuss from a philosophical point of view.
+However, this distinction requires empirical evidence based in studies focused on the subjective aspects of inequality.
+Furthermore, considering talent as part of meritocracy certainly opens some appealing avenues of research.
+For instance, some studies suggest that for the elites meritocracy is related to talent, whereas effort is more characteristic of the meritocracy of the middle and lower classes [@atria_economic_2020].
+Therefore, we propose that a concept (and measure) of meritocracy benefits from including both talent and effort, as in Young's original conception.
 
@@
 Several approaches to the empirical study of meritocracy based on public opinion surveys refer to the concept of _beliefs_, but with wide differences in meanings and operationalizations.
-To illustrate this point let us consider the commonly used "reasons to get ahead" battery, mentioned above.
-Some versions of this battery ask "how important you think it is for getting ahead in life" and then list several factors, whereas another version of this same battery - sometimes presented along the previous one - asks about "how important you think it _should_ be ...", and then listing the same concepts.
+To illustrate this point, let us consider the commonly used "reasons to get ahead" items’ battery, mentioned above.
+Some versions of this battery ask "how important you think it is for getting ahead in life" and then list various factors, whereas another version of this same battery - sometimes presented along the previous one - asks about "how important you think it _should_ be ...", and then listing the same concepts.
 Therefore, the question raised here is: Which one of both is a "belief": what _is_ (the first one) or what it _should_ be (the second one)?
@@
 As @kluegel_beliefs_1986 pointed out about the scope of beliefs: "This usage encompasses such more specific social-psychological concepts as values, perceptions, and attitudes" (p.30).
-Therefore, beliefs use to cover almost anything related to subjective factors.
+Therefore, beliefs used to cover almost anything related to subjective factors.
 In this regard, a relevant distinction in the field of inequality beliefs was made by @janmaat_subjective_2013: "Perceptions refer to subjective estimates of existing inequality (i.e., thoughts about what is).
@@
 In choosing this strategy of index construction, we argue that support for meritocracy is not a latent variable" [@kunovich_systems_2007 p.653-654].
-Although Reynolds and Xian [-@reynolds_perceptions_2014] made advances by conducting a principal component analysis of meritocratic and non-meritocratic dimensions, they finally chose to build a sum index despite proving a multidimensional latent structure.
-
-There is a relevant trade-off regarding measurement error estimation in survey scales.
+Although Reynolds and Xian [-@reynolds_perceptions_2014] made advances by conducting a principal component analysis of meritocratic and non-meritocratic dimensions, they finally decided to build a sum index despite proving a multidimensional latent structure.
+
+There is a relevant trade-off regarding measurement error estimation in survey instruments.
 Including multiple items per conceptual dimension opens the possibility of validity studies (at least in terms of the construct) by using factor analysis techniques, particularly confirmatory ones.
 However, increasing the number of items reduces the probability of usability of the scale in survey research given space limitations.
-This validity-usability conundrum is not easy to solve and, as we will show later in the methods section, we opted for a minimalistic approach oriented to the usability of the scale but that still gives the possibility of confirmatory factor analysis, although with some limitations.
+This validity-usability conundrum is not easy to solve and, as we will show later in the methods section, we opted for a minimalistic approach .
+Such approach is oriented to the usability of the measurement instrument, but still giving the possibility of confirmatory factor analysis.
 
@@
 
-Based on the previous assumptions and limitations identified in the empirical study of meritocracy, we propose a conceptual and measurement framework with the following characteristics:
+Based on the previous assumptions and limitations identified in the empirical study of meritocracy, mostly using the inequality module of ISSP, we propose a measurement framework for the study of meritocratic beliefs with the following characteristics:
 
@@
 - Based on _previous indicators_ as much as possible to keep comparability between studies.
-As most of the survey research on meritocracy so far has been made with the inequality module of the ISSP survey, we propose an operationalization suitable for this dataset, as well as a novel scale that attempts to rescue as much as possible the original wording of ISSP items.
+As most of the survey research on meritocracy so far has been made with the inequality module of the ISSP survey, we propose an operationalization suitable for this dataset, as well as a novel instrument that attempts to rescue as much as possible the original ISSP items.
 
@@
 Perceptions refer to the extent to which people observe that meritocracy functions or apply in their society, which relates to items such as "Hard work is important to get ahead in society".
-Preferences refer to normative expectations that are usually linked to a "should" expression (e.g.
+Preferences refer to normative expectations that are frequently linked to a "should" expression (e.g.
 whether hard work should be related to payment).
-The rows in Figure \@ref(fig:merit-model) consider the distinction between meritocratic and non-meritocratic dimensions [@reynolds_perceptions_2014], usually treated as different ends of the same continuum in previous research.
+The rows in Figure \@ref(fig:merit-model) consider the distinction between meritocratic and non-meritocratic dimensions [@reynolds_perceptions_2014], often treated as different ends of the same continuum in previous research.
 Non-meritocratic elements refer to the use of resources such as personal contacts and/or family advantages to get ahead in life.
+
+The ISSP survey (inequality module) has some items that researchers have used to measure each of the four different dimensions of meritocracy depicted here, but in a rather inconsistent manner.
+In the next section, we propose a classification of the ISSP meritocracy items into the different dimensions suggested in our multidimensional model depicted in Figure \@ref(fig:merit-model) as well as an empirical test of this model.
 
@@
 
-The analysis is organized into three studies.
-Study 1 serves as a background to our instrument proposal by attempting to operationalize the meritocracy conceptual model with available secondary data from the inequality module of the International Social Survey Programme (ISSP).
-Study 2 consists of the proposal of a new scale for measuring meritocratic perceptions and preferences.
-Finally, in Study 3 we perform an additional validity analysis of the meritocracy scale.
-
-# Study 1: Analyzing meritocracy with ISSP data
+The analysis is organized into two studies.
+The first study is an analysis of the multidimensional model for measuring meritocratic beliefs using ISSP data.
+Based on the results of this first study, the second study recommends some modifications and expansions of the ISSP items in order to tap better each of the dimensions identified in the study of meritocratic beliefs.
+
+# Study 1: Analyzing meritocratic beliefs with ISSP data
 
@@
 This wave corresponds to 2009 and covers attitudes towards a series of topics dealing with social inequality across 41 countries and 56,021 individuals.
-Although there is data available from this module for a more recent wave (2019), the 2009 wave contains relevant items for studying meritocracy that unfortunately are not present in the last wave.
-Table \@ref(tab:countries) shows details about the sample of countries.
+Although there is data available from this module for a more recent wave (2019), the 2009 wave contains relevant items for studying meritocracy that unfortunately are not present in the last wave (details about which items are missing in the last 2019 wave are presented in study 2, Table \@ref(tab:table-indicadores)).
+Table \@ref(tab:countries) shows details about the sample of countries in the ISSP 2009 data.
 
@@
 
-There are a series of indicators in the ISSP survey that in the following we attempt to classify in our meritocracy conceptual scheme, as presented below in Table \@ref(tab:table-issp):
+There are a series of indicators in the ISSP survey that we classified in our meritocracy conceptual scheme, as presented below in Table \@ref(tab:table-issp):
 
@@
 Regarding normative preferences, we observe that the meritocratic ones are deemed as important (from fairly to essential) for nearly the whole sample, decreasing slightly for the non-meritocratic ones.
-Still, we have to consider that the preferences for non-meritocratic aspects, in this case, refer to the distributive principle of need rather than personal background (as in the perceptions of non-meritocracy), which certainly is one of the limitations of this set of items for the operationalization of the proposed conceptual framework with this dataset.
+Still, we have to consider that the preferences for non-meritocratic aspects, in this case, refer to the distributive principle of need rather than personal background (as in the perceptions of non-meritocracy).
 
@@
 The exception here is the dimension of non-meritocratic perception, in which six items appear mostly associated by pairs according to their specific topics (family background, personal background, and connections).
-Still, between this same set of items, moderate correlations could anticipate a single latent factor underlying non-meritocratic perception, which is tested next through confirmatory factor analysis.
+Still, between this same set of items, moderate correlations could anticipate a single latent factor underlying non-meritocratic perceptions, which is tested next through confirmatory factor analysis.
 
@@
 Table \@ref(tab:sum-fit-issp) shows the results of the estimation of two confirmatory models.
-The first one (First Order) corresponds to a model that estimates four factors, each for one of the dimensions presented in  Table \@ref(tab:table-issp), with only regular fit indicators (CFI=0.959, TLI=0.944, RMSEA=0.098, $\chi^2$(df=48)= 21308.535).
+The first one (First Order) corresponds to a model that estimates four factors, each for one of the dimensions of the multidimensional framework for meritocratic beliefs presented in  Table \@ref(tab:table-issp), with only regular fit indicators (CFI=0.959, TLI=0.944, RMSEA=0.098, $\chi^2$(df=48)= 21308.535).
 Further analysis showed that the sources for poor fit were mostly related to the items of the non-meritocratic perceptions, which as mentioned above displayed correlations among them not taken into account in this model specification.
@@
 
-Attending to the sources of poor fit from the first estimated confirmatory model, we specified a second model that keeps the basic four-factor structure but generates an additional model for the non-meritocratic perceptions as depicted in Figure \@ref(fig:meas02-issp).
-In this model, the pairs of items in this dimension form three latent factors which at the same time generate a second-order factor of non-meritocratic perceptions, improving the fit of the model.
+Attending to the sources of poor fit from the first estimated confirmatory model, we specified a second model that keeps the basic four-dimensional structure but generates an additional model for the non-meritocratic perceptions as depicted in Figure \@ref(fig:meas02-issp).
+In this model, the pairs of items in this dimension form three latent factors which at the same time generate a second-order factor of non-meritocratic perceptions, improving the fit of the model significantly.
 
@@
 Regarding the correlations between the factors in Figure \@ref(fig:meas02-issp), we observe that perceptions are correlated with preferences, but more strongly for the meritocratic ($r=0.46, p<0.01$) than for the non-meritocratic ($r=0.27,p<0.01$) dimensions.
-Secondly, both perceptual dimensions (meritocratic and non-meritocratic) depict a moderate correlation ($r=0.30, p<0.01$), suggesting that they are not the opposite poles of the same continuum.
+Secondly, both perceptual dimensions (meritocratic and non-meritocratic) depict a moderate positive correlation ($r=0.30, p<0.01$), suggesting that they are not the opposite poles of the same continuum as some previous studies assume.
 Something similar occurs for the correlation between the two dimensions of preferences ($r=0.25,p<0.01$).
 
-Based on the measurement validation, we asses if the proposed measurement strategy can be used in cross-national comparisons.
+Based on the previous measurement validation, next we assess some of the potential of the proposed measurement strategy for cross-national comparisons of meritocratic beliefs.
 In this line, Figure \@ref(fig:scat1) depicts the correlation of factor scores for Meritocratic Perception and Meritocratic Preferences at the country level ($r=0.65,p<0.01$).
@@
 
-Although it was possible to get an adequate fit of the model to the pooled data from ISSP, there are several aspects to take into account when using these items for measuring meritocracy.
-First of all, these items were not specifically designed based on a conceptual approach to meritocracy, but it still is possible to find a certain fit of the model to the data after minor adjustments.
-Secondly, some of the items in similar dimensions refer to different concepts, for instance, meritocratic perceptions are related to ambition and hard work, whereas meritocratic preferences deal with how well a job is done.
-A better operationalization would require more alignment between the different sides of the same concept.
-And thirdly, the scale is rather unbalanced given the larger amount of non-meritocratic perception items, producing correlated errors.
-All these limitations are taken into account next in the design of Study 2.
-
-Concerning the meritocratic / non-meritocratic distinction in previous studies, we can preliminary conclude that the wide use of reverse-coding for non-meritocratic items attempting to measure meritocracy is not an adequate operationalization of the construct.
+ISSP is - and probably it will continue to be - the most widely used dataset for comparative studies of meritocratic beliefs.
+Although several ISSP items were not conceived originally with this end, they have been extensively used to operationalize different aspects of meritocracy.
+In this study, we ordered and classified ISSP items in four dimensions: perceptions and preferences, and meritocracy and non-meritocracy.
+The confirmatory analysis gave evidence of an adequate fit of the multidimensional model to the ISSP data, this is, it is possible to identify perceptions and preferences as two related but different constructs, and within perceptions and preferences there is evidence of meritocratic and non-meritocratic dimensions.
+Regarding this last distinction, we can preliminary conclude that the wide use of reverse-coding for non-meritocratic items attempting to measure meritocracy is not an adequate operationalization of the construct.
 Meritocracy and non-meritocracy are related but different concepts, meaning that it is theoretically possible to find different combinations at individual and contextual levels.
@@
 
-# Study 2: The Perceptions and Preferences for Meritocracy Scale
+Despite the advantages of using ISSP data for studying meritocratic beliefs, there are some aspects to be aware of regarding the measurement quality offered by the available items.
+Some of the items are not really coherent across dimensions, as for instance, meritocratic perceptions are related to ambition and hard work, whereas meritocratic preferences deal with how well a job is done.
+A better operationalization would require more alignment between the different sides of the same concept, this is, that perceptions and preferences point to the same object of analysis.
+Furthermore, the instrument is rather unbalanced given the larger amount of non-meritocratic perception items, producing correlated errors.
+These limitations are taken into account next in the design of Study 2.
+
+# Study 2: The Multidimensional Meritocracy Measurement Instrument
+
+Taking the analysis of meritocratic beliefs with ISSP as a departing point, the following study offers a measurement instrument that attempts to overcome some limitations identified in the previous study.
+In this sense, the analysis with ISSP data can be considered as an _inductive_ exercise, taking the items and data available and attempting to adjust this information to the proposed multidimensional conceptual model.
+In the following study we take a complementary _deductive_ approach, departing from the conceptual model for the design of a measurement instrument.
+Still, this second approach does not start from scratch, as it is mostly based on ISSP items, but incorporates a balanced representation of items for each dimension, as well as attempts to maintain the coherence in topics across dimensions (for instance, hard work in both perceptions and preferences).
 
@@
 The quotas for gender, age, and educational levels were generated based on a survey by the Public Studies Center [@cep_encuesta_2019], which is a well-regarded counterpart agency of the ISSP (International Social Survey Programme) in Chile.
-A total sample of 2,141 people was collected, excluding those who did not sign the informed consent and those not answering the meritocracy response battery.
+A total sample of 2,141 individuals was collected, excluding those who did not sign the informed consent and those not answering the meritocracy instrument.
 There were no significant differences between our sample and the wider population for most socio-demographic characteristics, except for the educational level (see Table \@ref(tab:rep-samp) in Appendix).
@@
 
-The proposed scale of meritocratic perceptions and preferences consisted of eight indicators, two for each of the four dimensions listed earlier: Perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic).
+The proposed items' battery of meritocratic perceptions and preferences consisted of eight indicators, two for each of the four dimensions listed earlier: Perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic).
 Two main criteria oriented this design.
@@
 To this regard, some of the items were adapted from the items battery "reasons to get ahead" (ISSP/GSS), which has been widely used for operationalizing meritocracy in international research [@mijs_paradox_2019; @duru-bellat_who_2012; @reynolds_perceptions_2014].
-The scale items ordered according to their respective dimensions are presented in Table \@ref(tab:table-indicadores).
+
+The items, organized according to their respective dimensions, are presented in Table \@ref(tab:table-indicadores).
 These eight likert-type items have five response alternatives, ranging from "Completely disagree"(1) to "Completely agree" (5).
+The table presents an additional column with the respective equivalent items from ISSP 2009/2019.
+With this information it is possible to have a more clear picture of the areas in which the proposed measurement instrument is similar to ISSP and in which ones it differs.
 
@@
 
+The measurement of meritocratic beliefs with this new instrument shows, in general, better fit - and therefore better measurement quality - than study 1.
+However, the comparison between the two studies is unbalanced as we are comparing one particular country (study 2) with the whole set of ISSP countries (study 1).
+To clear this point, we performed an additional analysis of ISSP data only for the case of Chile.
+Overall, the results for the chilean ISSP data show a similar fit to the ones of with the complete ISSP dataset (CFI=0.987, TLI=0.980, RMSEA=0.055, $\chi^2$(df=45)= 234), and therefore the instrument of study 2 outperforms their fit indicators.
+
 ## Discussion Study 2
 
-The estimation performed with different administration sets for the meritocracy scale gives evidence in line with the conceptual model.
+The estimation performed with different administration sets for the meritocracy eight-items scale gives evidence in line with the multi dimensional conceptual model proposed in this paper.
 The four dimensions appear independent but related, with some differences when compared to the estimation with ISSP data in Study 1.
@@
 Theoretically, this makes more sense, as if someone perceives that meritocracy functions well, they probably will support it more.
-On the other side, if someone perceives non-meritocracy, this should lead to a more prominent preference for meritocracy as it would constitute a demand for a larger meritocracy when it is not perceived to work.
+On the other side, if someone perceives non-meritocracy, this should lead to a more prominent preference for meritocracy as it would constitute a demand for a larger meritocracy when it is not perceived to function.
 
@@
 As mentioned before, the option of a minimalistic approach in terms of items has advantages and disadvantages: advantages in terms of usability in social surveys with limited space and disadvantages in terms of measurement validity in the estimation of the latent constructs.
-Aware of this trade-off, next we present additional evidence of validity to complement some of the weaknesses identified.
+Regarding the practical use of the measurement instruments, the primary lesson from our study 2 is the importance of distinguishing between the four dimensions.
+We recommend using the eight-items battery when scholars intend to measure both perceptions and preferences.
+Alternatively, scholars can use the individual components separately.
+For instance, researchers interested in the dimension of perception can use only  the four items that assess both the meritocratic and non-meritocratic perception.
+Furthermore, this measurement proposal can be used for relating the dimensions among themselves (for instance, to what extent the perception of meritocracy is related to the preference of it).
+It can also be used to study how each of these dimensions is associated with or moderated by individual and country level variables.
 
@@
 Studies that attempt to characterize and compare societies by their support for meritocratic beliefs have used different approaches.
-As most studies use secondary survey data, they tend to assume that the available indicators represent an underlying meritocratic construct.
+As most studies use secondary survey data, and by and large items from the social inequality module from ISSP, they tend to assume that the available indicators represent an underlying meritocratic construct.
 A review of these studies reveals several non-tested assumptions and the use of similar indicators to represent different constructs and dimensions of meritocracy.
-As the existence of heterogeneous approaches certainly has consequences for the advancement of the study of meritocracy, this paper presented a comprehensive conceptual framework for the empirical study of meritocracy, building upon previous research.
-This framework was operationalized in a survey scale and tested against available and new survey data.
-
-We identified four critical aspects regarding the measurement of meritocracy in previous studies: unidimensionality, the ambiguous use of the term "beliefs", the use of non-meritocratic indicators as opposed to meritocratic indicators, and the consideration of measurement error.
-The proposed 8-item scale, "Perceptions and Preferences for Meritocracy", was designed and tested to deal with these four issues.
-Our results indicate that perceptions and preferences are two related but different dimensions, often confounded in the previous research under the label of "beliefs".
-At the same time, meritocratic and non-meritocratic dimensions do not appear to constitute poles of the same continuum, as some previous studies have assumed.
-Regarding the possible effects of the items' order in the estimation of the latent variable, we tested three different order versions and found evidence that suggests the use of the randomized version of the scale.
-Furthermore, we found evidence of measurement invariance that approaches well the requirement of comparability between samples.
-
-One of the criteria in designing the merit scale was to develop a short instrument that could be used in social surveys, assuming that using only two items per dimension posits limitations for the confirmatory analysis estimated here.
-We preferred this confirmatory-minimalistic approach as it offers several advantages in terms of information about the quality of the measurement, and we offered additional validity evidence with measurement invariance analysis.
-Despite the advantages, a more straightforward perspective is always available, for instance using only the average of the items per dimension.
-Whereas a short scale is always appreciated in social survey studies, for other audiences could not reach measurement standards.
-This may require further expansions and/or consider using a different instrument as the "Preference for the Merit Principle Scale" [@davey_preference_1999] (which measures only one dimension of meritocracy with fifteen items).
-
-The four-dimensional conceptual framework and its operationalization in the Perceptions and Preferences for Meritocracy scale opens several avenues for future research.
+As the existence of heterogeneous approaches certainly has consequences for the advancement of the study of meritocracy, this paper presented a comprehensive multidimensional conceptual framework for the empirical study of meritocracy, building upon previous research.
+We tested the validity of our conceptual framework with two studies.
+In a first study, this framework was operationalized and tested with ISSP international data.
+Based on the results with ISSP, a second study proposed some improvements to the measure of meritocracy  with ISSP, which was tested with novel data from a single country case (Chile).
+
+The results of our empirical studies show evidence in favor of a four dimensional model of meritocracy, opening interesting avenues for future comparative research.
 For instance, distinguishing perceptions from preferences will allow us to evaluate the extent to which different societies are accustomed to, or satisfied with, the perceived level of meritocracy, in terms of differences between what is perceived and what is preferred.
-Additionally, given that non-meritocratic factors are not necessarily related to meritocratic ones, our framework makes it possible to assess the perceived legitimacy of practices such as the use of personal contacts and their interference (or not) with meritocratic ideals in different societies.
-Considering individual and societal structural factors affecting meritocracy will allow us to advance future hypotheses relating to meritocratic legitimacy to a more comprehensive level than the current one, which is based mostly on meritocratic perceptions.
-Furthermore, the impact of different configurations of the four-dimensional framework on practices and behaviors, such as corruption, civic involvement, and political alignment, are areas open to further research.
+Additionally, given that non-meritocratic factors are not necessarily related to meritocratic ones, the multidimensional measurement of meritocratic beliefs makes it possible to assess the perceived legitimacy of practices such as the use of personal contacts and their interference (or not) with meritocratic ideals in different societies.
+Considering how individual and societal structural factors are related to meritocracy will allow us to gain knowledge about the legitimacy of distribution based on meritocratic criteria.
+Furthermore, the impact of different configurations of the four-dimensional framework on practices and behaviors, such as corruption, civic involvement, and political alignment, are areas that become open for development.
 Such future agendas could be especially relevant in times of economic crisis and growing inequalities, which could entail changes in the legitimation of the current distributive structure based on meritocratic ideals.
+
+Besides the areas of research with comparative survey studies as ISSP, single case studies as well as future comparative studies (ISSP included) might benefit from taking into account the measurement instrument for meritocracy proposed here, tested with novel Chilean data.
+The instrument is to some extent comparable with ISSP - as it shares some of its items - but improves the balanced measurement of the four dimensions in a minimalistic eight-items’ battery, suitable for survey research.
+Still, further evidence is needed in order to assess the external validity of the results.
+Upcoming research in the area of meritocratic beliefs, a growing research agenda, would certainly help to keep improving the quality of the measurement instruments in this area, which is a key enterprise when it comes to assess the support and legitimacy of the meritocratic distribution, particularly in unequal societies.
 

### 2023-09-13 paper-blinded-RR-IJS.Rmd
ajustes tablas

@@
  tabitems <- read.csv(file = "input/tables/table01.csv",header = 1,sep = ",",encoding = "UTF-8") # call file  generated externally for the table
- cnames <- c("Component", "Dimensions","Item (English)" , "Item original (Spanish)")
- cap <- "Items of the perceptions and preferences for meritocracy scale."
-
+ tabitems$itemsspa <- NULL
+ tabitems$issp <- c("How important is hard work?","-",
+                    "How important is coming from a wealthy family?", "How important is knowing the right people",
+                    "How hard he or she works at the job?*","How well he or she does the job?*",
+                    "-", "-")
+ cnames <- c("Component", "Dimensions","Items" , "Items ISSP (2009,2019)")
+ cap <- "Items' battery for the multidimensional measurement of meritocratic beliefs and their equivalents in ISSP"
  knitr::kable(tabitems, table_format, booktabs = T, linesep = "",col.names = cnames, caption = cap) %>%
@@
    kableExtra::column_spec(column = 2,width = "2 cm") %>%
-   kableExtra::collapse_rows(columns = 1:2,valign = "middle")
-```
-
+   kableExtra::collapse_rows(columns = 1:2,valign = "middle") %>%
+   kableExtra::footnote(general = "Those with an asterisk (*) are only present in 2009",footnote_as_chunk = T)
+```
 \pagebreak

### 2023-09-14 paper-blinded-RR-IJS.Rmd
last review for render manuscripts

@@
 ---
-title: "A multidimensional approach for measuring meritocratic beliefs with the ISSP social inequality survey: Possibilities, limitations and alternatives"
+title: "A multidimensional approach for measuring meritocratic beliefs: Advantages, limitations and alternatives to the ISSP social inequality survey"
 #css: "custom.css"
@@
 abstract: |
-  Economic and social inequalities are causing growing concerns and crises in contemporary societies.
-One factor suggested explaining the persistence of inequality is the belief in meritocracy, which would justify economic disparities based on differences in effort and talent.
-A great part of the international research that has attempted to measure meritocratic beliefs has used the social inequality module of the ISSP (International Social Survey Programme), which offers an unprecedented opportunity to compare meritocratic views in different societies.
-However, this research agenda is still a novel enterprise characterized by diverse conceptualizations and measures of meritocracy, often leading to inconsistent findings and interpretations.
+  A great part of the comparative international research that has attempted to measure meritocratic beliefs has used the social inequality module of the ISSP (International Social Survey Programme), which offers an unprecedented opportunity to compare meritocratic views in different societies.
 Based on a series of studies using ISSP data, the present paper proposes a multidimensional measurement framework for meritocratic beliefs.
-This framework distinguishes, on the one side, between perceptions and preferences, and on the other side, between meritocratic and not meritocratic aspects, a two-by-two scheme that identifies four dimensions of meritocratic research.
-This framework allows us to classify previous research contributions, as well as to identify some vacuums and limitations in the extant literature.
+This framework distinguishes, on the one side, between perceptions and preferences, and on the other side, between meritocratic and not meritocratic aspects.
 In a first study, we test the multidimensional framework for meritocratic beliefs using the ISSP 2009 inequality module through confirmatory factor analysis (CFA) techniques.
-After identifying the advantages and some limitations of ISSP items for a multidimensional operationalization of meritocratic beliefs, in a second study, we suggest an expanded set of items that better taps the different dimensions of meritocracy.
+After identifying the advantages and some limitations of ISSP items for a multidimensional operationalization of meritocratic beliefs, in a second study, we suggest an modified  set of items that better taps the different dimensions of meritocracy.
 We examined the measuring properties of this new instrument using a sample of Chilean adults (N=2,141).
-While the instrument was tested in the limited context of a single country, the findings offer  aditional evidence about the validity of the multidimensional measurement model of meritocratic beliefs.
 Based on these results, we recommend improvements in measuring meritocratic beliefs in cross-national studies.
@@
 linkcolor: blue
-bibliography:
-  - input/bib/meritocracy.bib
+bibliography: input/bib/meritocracy.bib
 csl: input/bib/apa7-singlespace.csl
@@
   - \usepackage{times}           # Times New Roman
-#  - \hypersetup {colorlinks = true, linkcolor = blue, urlcolor = blue}    
+  - \hypersetup {colorlinks = true, linkcolor = blue, urlcolor = blue}    
   - \usepackage{caption}
@@
 \pagebreak
-
+\selectlanguage{USenglish} 
 --->
-
-\selectlanguage{USenglish}
 
@@
 Secondly, we test the validity of our conceptual framework with two studies.
-Study 1 offers an operationalization and confirmatory analysis of the proposed multidimensional framework using available items from ISSP (inequality module 2009).
-Based on the advantages and limitations of the results obtained in study 1, and in order to gain additional evidence for the validity of the measurement instruments, study 2 suggests modifications and expansion of the items commonly used to measure meritocratic beliefs.
-This measurement proposal is tested with data from a single country case (Chile).
+Study 1 offers an operationalization and confirmatory analysis of the proposed multidimensional framework using available items from ISSP, inequality module 2009 (the reason to use this survey wave instead of the last one on social inequality from 2019 is that the 2009 version includes some key items used in meritocratic reserach that were removed in 2019).
+Taking into account on the advantages and limitations of the results obtained in study 1, and in order to gain additional evidence for the validity of the measurement instruments, study 2 suggests modifications and expansion of the items commonly used to measure meritocratic beliefs.
+This measurement proposal is tested with data available from a Chilean survey on meritocracy and inequality (2020).
 
@@
 As he points out: "Hard work is arguably the most meritocratic part of Michael Young's equation: 'Merit = Intelligence + Effort', for the simple fact that intelligence itself is influenced by a non-meritocratic factor: who your parents happen to be" (p.5).
-
-In this conceptual and measurement approach of meritocratic beliefs, we can observe a couple of strong conceptual assumptions: a) effort would not depend on parental influence (at least not as strongly as intelligence), and b) talent (as innate ability) is not meritocratic (contrary to Michael Young's' original conceptualization).
+In such conceptual and measurement approach of meritocratic beliefs, we can observe a couple of strong conceptual assumptions: a) effort would not depend on parental influence (at least not as strongly as intelligence), and b) talent (as innate ability) is not meritocratic (contrary to Michael Young's' original conceptualization).
 This conceptual and measurement-based assumption is found in other studies, which also assume effort to be the main and only aspect of meritocracy [@girerd_neoliberalism_2020; @bubak_perceptions_2019], raising the question: Is effort the only (or main) dimension behind the concept of meritocracy?
+
 We argue that whether talent (as intelligence and ability) is or is not considered meritocratic is an interesting topic to discuss from a philosophical point of view.
-However, this distinction requires empirical evidence based in studies focused on the subjective aspects of inequality.
+However, when it comes to survey research on meritocratic beliefs, the weight of effort or talent in what it is deemed  meritocratic should be left to public scrutiny.
 Furthermore, considering talent as part of meritocracy certainly opens some appealing avenues of research.
-For instance, some studies suggest that for the elites meritocracy is related to talent, whereas effort is more characteristic of the meritocracy of the middle and lower classes [@atria_economic_2020].
-Therefore, we propose that a concept (and measure) of meritocracy benefits from including both talent and effort, as in Young's original conception.
+For instance, some studies suggest that, for the elites, meritocracy is related to talent, whereas effort is more characteristic of the meritocracy of the middle and lower classes [@atria_economic_2020].
+Therefore, we suggest that a concept (and measure) of meritocracy benefits from including both talent and effort, as in Young's original conception.
 
@@
 As @kluegel_beliefs_1986 pointed out about the scope of beliefs: "This usage encompasses such more specific social-psychological concepts as values, perceptions, and attitudes" (p.30).
-Therefore, beliefs used to cover almost anything related to subjective factors.
+Therefore, the term _beliefs_ used to cover almost anything related to subjective factors.
 In this regard, a relevant distinction in the field of inequality beliefs was made by @janmaat_subjective_2013: "Perceptions refer to subjective estimates of existing inequality (i.e., thoughts about what is).
@@
 - Based on _previous indicators_ as much as possible to keep comparability between studies.
-As most of the survey research on meritocracy so far has been made with the inequality module of the ISSP survey, we propose an operationalization suitable for this dataset, as well as a novel instrument that attempts to rescue as much as possible the original ISSP items.
+As most of the survey research on meritocracy so far has been made with the inequality module of the ISSP survey, we propose a comprehensive operationalization suitable for this dataset, as well as a novel instrument that attempts to rescue as much as possible the original ISSP items.
 
@@
 Although there is data available from this module for a more recent wave (2019), the 2009 wave contains relevant items for studying meritocracy that unfortunately are not present in the last wave (details about which items are missing in the last 2019 wave are presented in study 2, Table \@ref(tab:table-indicadores)).
-Table \@ref(tab:countries) shows details about the sample of countries in the ISSP 2009 data.
+Table \@ref(tab:countries) (appendix) shows details about the sample of countries in the ISSP 2009 data.
 
@@
  # remotes::install_github('haozhu233/kableExtra@a6af5c0')  # for problems with collapse_rows()
-library(knitr)
-library(kableExtra)
-library(dplyr)
-#  pacman::p_load(knitr, kableExtra, dplyr)
+pacman::p_load(knitr, kableExtra, dplyr)
   table_format = if(knitr::is_html_output()) {
@@
 Still, we have to consider that the preferences for non-meritocratic aspects, in this case, refer to the distributive principle of need rather than personal background (as in the perceptions of non-meritocracy).
+Such inconsistency between the target of the items for perceptions an preferences is one of the issues to deal with in study 2.
 
@@
 In terms of the association between the indicators, Figure \@ref(fig:corr-issp) shows the polychoric correlation matrix.
-Firstly, according to the criteria suggested by @cohen_statistical_1988 for effect size, we observe in general that the moderate to high associations are between the pairs of items representing each of the dimensions as described in Table \@ref(tab:table-issp).
+Firstly, according to the criteria suggested by @cohen_statistical_1988 for effect size, we observe in general that the moderate to high associations are between the pairs of items representing each of the four dimensions described in Table \@ref(tab:table-issp).
 The exception here is the dimension of non-meritocratic perception, in which six items appear mostly associated by pairs according to their specific topics (family background, personal background, and connections).
-Still, between this same set of items, moderate correlations could anticipate a single latent factor underlying non-meritocratic perceptions, which is tested next through confirmatory factor analysis.
+Still, between this same set of items, moderate correlations could anticipate a single latent factor underlying non-meritocratic perceptions, which is tested next in the confirmatory factor analysis model.
 
@@
   if (!require("pacman")) install.packages("pacman")
-  pacman::p_load(dplyr, kableExtra, knitr,lavaan)
+ # pacman::p_load(dplyr, knitr,lavaan)
+  library(kableExtra)
   table_format = if(knitr::is_html_output()) { #conditional instructions for kable
@@
 Based on the previous measurement validation, next we assess some of the potential of the proposed measurement strategy for cross-national comparisons of meritocratic beliefs.
-In this line, Figure \@ref(fig:scat1) depicts the correlation of factor scores for Meritocratic Perception and Meritocratic Preferences at the country level ($r=0.65,p<0.01$).
+In this line, Figure \@ref(fig:scat1) depicts the correlation of factor scores for meritocratic perceptions and meritocratic preferences at the country level ($r=0.65,p<0.01$).
 For instance, Bulgaria (BGR) and the United States (USA) appropriately represent the positive association between perception and preferences.
-In other words, countries that describe merit in their society as an essential factor to get ahead in life also consider that effort and talent should be determinant factors in how much an individual earns.
-On the other hand, in countries with lower perceived meritocracy, preferences are also low, where Denmark (DNK) and Venezuela (VEN) are representative societies of this association.
-
-Another finding depicted by Figure \@ref(fig:scat1) is the positive association of Non-meritocratic perception with Meritocratic preferences ($r=0.55,p<0.01$), showing that countries like Bulgaria (BGR) and Philipines (PHL) represent cases with higher Non-meritocratic perceptions and higher preferences for meritocracy.
+In other words, countries that perceived merit in their society as an essential factor to get ahead in life also consider that effort and talent should be determinant factors in how much an individual earns.
+On the other hand, in countries with lower perceived meritocracy, preferences are also low, where Denmark (DNK) and Venezuela (VEN) are representative cases of this association.
+
+Another finding depicted by Figure \@ref(fig:scat1) is the positive association of non-meritocratic perceptions with meritocratic preferences ($r=0.55,p<0.01$), showing that countries like Bulgaria (BGR) and Philipines (PHL) represent cases with higher non-meritocratic perceptions and higher preferences for meritocracy.
 In other words, in societies where people perceive higher importance of non-meritocratic aspects to get ahead in life, people also prefer that merit should be a determinant factor in wages.
-On the other side, countries like Sweden (SWE) or Denmark (DNK), both Scandinavian societies with strong welfare systems, represent contexts in which non-meritocratic aspects are not described as important, and merit is not considered as a factor that should be determinant in wages.
+On the other side, countries like Sweden (SWE) or Denmark (DNK), both Scandinavian societies with strong welfare systems, represent contexts in which non-meritocratic aspects are not described as important, and merit is not considered as a factor that should be relevant for wages.
+This finding, particularly the positive association between non-meritocratic perception and meritocratic preferences, opens the question of to what extent meritocratic preferences are capable to hold even under evidence of a large amount of perceived non-meritocracy .
 
@@
   if (!require("pacman")) install.packages("pacman")
-  pacman::p_load(dplyr, kableExtra, knitr,lavaan,countrycode,ggplot2,haven,
+#  pacman::p_load(dplyr, kableExtra, knitr,lavaan,countrycode,ggplot2,haven,
                  gridExtra)
@@
 
-On one hand, when comparing societies based on their level of industrialization [@ishida_trends_2005], Figure \@ref(fig:scat2) illustrates the correlations within the sample.
-It shows that early-industrialized countries exhibit a similar association between Meritocratic Perception and Meritocratic Preference ($r=0.67, p<0.05$), a trend consistent within the group of late-industrialized nations ($r=0.68, p<0.01$).
-This suggests that citizens' perceptions regarding the significance of merit in both industrialized and late-industrialized nations tend to have a similar impact on their beliefs about the role of merit in determining wages.
-On the other hand, the association between Non-meritocratic Perception and Meritocratic Preference displays significant differences when compared to the complete sample.
-Industrialized nations exhibit a strong positive correlation ($r=0.75, p<0.01$).
-In contrast, late-industrialized countries show a weak and non-significant correlation ($r=0.29, p>0.05$).
-In other words, the non-meritocratic aspect holds greater salience in industrialized societies when it comes to citizens' opinions regarding the determinants of wages.
-However, in late-industrialized societies, the importance of non-meritocratic factors diminishes in influencing meritocratic preferences.
-This evidence suggests that in societies with a long history of transformations in their social structure, the perceived importance of both meritocratic and non-meritocratic dimensions in achieving success reinforces people's preferences for how status positions should be distributed.
-In contrast, in countries that experienced changes in their occupational structure during the early decades of the 20th century, meritocratic perception holds greater prominence than non-meritocratic factors such as personal contacts or family background in shaping citizens' opinions regarding the determinants of wages.
+Another possibility to explore within this multidimensional framework of meritocratic beliefs is the differences between countries.
+For instance, when comparing societies based on their level of industrialization [@ishida_trends_2005], Figure \@ref(fig:scat2) shows that early and late-industrialized countries exhibit a similar association between meritocratic perceptions and preferences ($r=0.67, p<0.05$; $r=0.68, p<0.01$, respectively).
+However, when analyzing the association between non-meritocratic perception and meritocratic preferences, industrialized nations exhibit a strong positive correlation ($r=0.75, p<0.01$), which is not the case for late-industrialized countries ($r=0.29, p>0.05$).
+A possible interpretation is that the perception of non-meritocracy in less industrialized societies threatens the meritocratic ideal, which is not the case for the industrialized ones.
+Still, this preliminary analysis only attempts to show some possiblities that open up when introducing a multidimensional perspective for studying meritocratic beliefs
 
@@
 Although several ISSP items were not conceived originally with this end, they have been extensively used to operationalize different aspects of meritocracy.
-In this study, we ordered and classified ISSP items in four dimensions: perceptions and preferences, and meritocracy and non-meritocracy.
-The confirmatory analysis gave evidence of an adequate fit of the multidimensional model to the ISSP data, this is, it is possible to identify perceptions and preferences as two related but different constructs, and within perceptions and preferences there is evidence of meritocratic and non-meritocratic dimensions.
-Regarding this last distinction, we can preliminary conclude that the wide use of reverse-coding for non-meritocratic items attempting to measure meritocracy is not an adequate operationalization of the construct.
+In this study, we ordered and classified ISSP items, which have been previously used for measuring meritocratic beliefs, into four different dimensions: perceptions and preferences, and meritocracy and non-meritocracy.
+The confirmatory analysis gave evidence of an adequate fit of the multidimensional model to the ISSP data, this is, it is possible to identify perceptions and preferences for meritocracy and non-meritocracy as four related but different constructs.
+Regarding the non-meritocratic dimensions, we can preliminary conclude that the wide use of reverse-coding for non-meritocratic items attempting to measure meritocratic perceptions or preferences, is not an adequate operationalization.
 Meritocracy and non-meritocracy are related but different concepts, meaning that it is theoretically possible to find different combinations at individual and contextual levels.
@@
 Some of the items are not really coherent across dimensions, as for instance, meritocratic perceptions are related to ambition and hard work, whereas meritocratic preferences deal with how well a job is done.
+Such differences impose limitations for a comparative analysis between pereptions and preferences.
 A better operationalization would require more alignment between the different sides of the same concept, this is, that perceptions and preferences point to the same object of analysis.
-Furthermore, the instrument is rather unbalanced given the larger amount of non-meritocratic perception items, producing correlated errors.
-These limitations are taken into account next in the design of Study 2.
-
-# Study 2: The Multidimensional Meritocracy Measurement Instrument
-
-Taking the analysis of meritocratic beliefs with ISSP as a departing point, the following study offers a measurement instrument that attempts to overcome some limitations identified in the previous study.
+Furthermore, the instrument is rather unbalanced given the larger amount of non-meritocratic perception items (six) when compared to the number items related to the other dimensions (two).
+These limitations are taken into account next in the design of study 2.
+
+# Study 2: A proposal for a multidimensional measure of meritocratic beliefs
+
+Taking the analysis of meritocratic beliefs with ISSP as a departing point, the following study offers and test a new measurement instrument that attempts to overcome some limitations identified in the previous study.
 In this sense, the analysis with ISSP data can be considered as an _inductive_ exercise, taking the items and data available and attempting to adjust this information to the proposed multidimensional conceptual model.
 In the following study we take a complementary _deductive_ approach, departing from the conceptual model for the design of a measurement instrument.
-Still, this second approach does not start from scratch, as it is mostly based on ISSP items, but incorporates a balanced representation of items for each dimension, as well as attempts to maintain the coherence in topics across dimensions (for instance, hard work in both perceptions and preferences).
+However, this second approach does not start from scratch, as it is based on ISSP items as much as possible for the sake of comparability with previous studies.
+The two major differences are: a) it incorporates a balanced and minimalistic representation of items for each of the four dimensions,and b) attempts to maintain the coherence in the topics asked across dimensions (for instance, "effort" in items for both perceptions and preferences).
 
@@
 
-The data comes from an online survey that was part of a larger study on meritocracy and preferences developed in Chile in 2019, funded by the Chilean National Scientific Agency (ANID).
+The data comes from an online survey that was part of a larger study on meritocracy and distributive preferences developed in Chile in 2020, funded by the Chilean National Scientific Agency (ANID).
 The questionnaire was programmed in Qualtrics and the fieldwork was conducted by an external online survey agency ([netquest.com](www.netquest.com)) between December 2019 and January 2020.
@@
 
-The proposed items' battery of meritocratic perceptions and preferences consisted of eight indicators, two for each of the four dimensions listed earlier: Perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic).
+The proposed items' battery of meritocratic perceptions and preferences consisted of eight indicators, two for each of the four dimensions listed earlier: perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic).
 Two main criteria oriented this design.
@@
 The models estimated one factor for each of the four proposed dimensions presented in Table \@ref(tab:table-indicadores).
-As in Study 1, CFA was performed with the `lavaan` R package (version 0.6-3; @rosseel_lavaan_2012a), with diagonally weighted least squares (DWLS) estimation due to the items' ordinal level of measurement [@kline_principles_2016; @rosseel_lavaan_2012a].
+As in study 1, CFA was performed with the `lavaan` R package (version 0.6-3; @rosseel_lavaan_2012a), with diagonally weighted least squares (DWLS) estimation due to the items' ordinal level of measurement [@kline_principles_2016; @rosseel_lavaan_2012a].
 The fit indexes and cut-off criteria were the same as the ones used in Study 1.
@@
 There are three main aspects to highlight from this correlation matrix.
-Firstly, and as expected, the largest correlations are between indicators that correspond to the same factors behind the conceptual model (e.g., perception of meritocracy by effort and by talent, $r=0.52,p<0.01$).
-Secondly, among these correlations, the highest are those between the non-meritocratic dimension but in perceptions ($r=0.73,p<0.01$) and preferences ($r=0.61,p<0.01$).
+Firstly, and as expected, the largest correlations are between indicators that correspond to the same factors (dimensions) behind the conceptual model (e.g., perception of meritocracy by effort and by talent, $r=0.52,p<0.01$).
+Secondly, among these correlations, the highest are those between the non-meritocratic dimensions, both in perceptions ($r=0.73,p<0.01$) and preferences ($r=0.61,p<0.01$).
 Thirdly, both items for meritocratic preferences (E and F) are the ones that mostly correlate with the rest of the perceptual items, showing medium to high correlations and indicating that the perception  of non-meritocracy would be related to larger meritocratic preferences.
-Finally, similar to the results of Study 1 with ISSP data, we observe no considerable negative correlations between meritocratic and non-meritocratic aspects, undermining previous studies' assumptions that these dimensions would be the opposite poles of one same continuum [@reynolds_perceptions_2014].
+Finally, similar to the results of study 1 with ISSP data, we observe no considerable negative correlations between meritocratic and non-meritocratic aspects, undermining previous studies' assumptions that these dimensions would be the opposite poles of one same continuum [@reynolds_perceptions_2014].
 
@@
 
-The first step in the analysis consists of comparing the model fit indicators for the three versions of the scale that were randomly assigned to the participants: order according to perceptions/preferences, order according to topics, and a complete random items’ order (see Figure \@ref(fig:appmod)).
+The first step in the analysis consists of comparing the model fit indicators for the three versions of the instrument that were randomly assigned to the participants: order according to perceptions/preferences, order according to topics, and a complete random items’ order (see Figure \@ref(fig:appmod)).
 
@@
 
-The estimation performed with different administration sets for the meritocracy eight-items scale gives evidence in line with the multi dimensional conceptual model proposed in this paper.
-The four dimensions appear independent but related, with some differences when compared to the estimation with ISSP data in Study 1.
-The main general difference is that with ISSP data, the four latent dimensions were positively correlated, whereas now some correlations are statistically not different from zero.
-For instance, perceptions and preferences for meritocracy keep being related, which is not the case for the non-meritocratic pair.
-Theoretically, this makes more sense, as if someone perceives that meritocracy functions well, they probably will support it more.
-On the other side, if someone perceives non-meritocracy, this should lead to a more prominent preference for meritocracy as it would constitute a demand for a larger meritocracy when it is not perceived to function.
-
+The eight-items' battery presented here and the estimation of the confirmatory model with chilean data gives additional evidence for a multidimensional concept of meritocratic beliefs.
 Despite the good general working of the scale, there are some limitations in interpreting the results given the few items (two) per dimension.
@@
 Furthermore, this measurement proposal can be used for relating the dimensions among themselves (for instance, to what extent the perception of meritocracy is related to the preference of it).
-It can also be used to study how each of these dimensions is associated with or moderated by individual and country level variables.
+It can also serve to studies about how each of these dimensions is associated with or moderated by individual and country level variables.
 
@@
 Considering how individual and societal structural factors are related to meritocracy will allow us to gain knowledge about the legitimacy of distribution based on meritocratic criteria.
-Furthermore, the impact of different configurations of the four-dimensional framework on practices and behaviors, such as corruption, civic involvement, and political alignment, are areas that become open for development.
-Such future agendas could be especially relevant in times of economic crisis and growing inequalities, which could entail changes in the legitimation of the current distributive structure based on meritocratic ideals.
-
-Besides the areas of research with comparative survey studies as ISSP, single case studies as well as future comparative studies (ISSP included) might benefit from taking into account the measurement instrument for meritocracy proposed here, tested with novel Chilean data.
-The instrument is to some extent comparable with ISSP - as it shares some of its items - but improves the balanced measurement of the four dimensions in a minimalistic eight-items’ battery, suitable for survey research.
+Furthermore, the impact of different configurations of the four-dimensional framework on practices and behaviors, such as corruption, civic involvement, and political alignment, are areas that open up for future development.
+Such agendas could be especially relevant in times of economic crisis and growing inequalities, which could entail changes in the legitimation of the current distributive structure based on meritocratic ideals.
+
+Besides the areas of research with comparative survey studies as ISSP, single case studies as well as future comparative studies (ISSP included) might benefit from taking into account the measurement instrument for meritocracy proposed here, tested so far with novel Chilean data.
+The instrument is to some extent comparable to ISSP - as some of its items are equivalent - but improves the balanced measurement of the four dimensions in a minimalistic eight-items’ battery, suitable for survey research.
 Still, further evidence is needed in order to assess the external validity of the results.

### 2023-09-14 paper-blinded-RR-IJS.tex
last review for render manuscripts

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  10pt,
  a4paper,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{times}
\hypersetup {colorlinks = true, linkcolor = blue, urlcolor = blue}
\usepackage{caption}
\captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
\usepackage{graphicx}
\usepackage{float}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage[british,UKenglish,USenglish,american]{babel}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={A multidimensional approach for measuring meritocratic beliefs: Advantages, limitations and alternatives to the ISSP social inequality survey},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{A multidimensional approach for measuring meritocratic beliefs: Advantages, limitations and alternatives to the ISSP social inequality survey}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle
\begin{abstract}
A great part of the comparative international research that has attempted to measure meritocratic beliefs has used the social inequality module of the ISSP (International Social Survey Programme), which offers an unprecedented opportunity to compare meritocratic views in different societies. Based on a series of studies using ISSP data, the present paper proposes a multidimensional measurement framework for meritocratic beliefs. This framework distinguishes, on the one side, between perceptions and preferences, and on the other side, between meritocratic and not meritocratic aspects. In a first study, we test the multidimensional framework for meritocratic beliefs using the ISSP 2009 inequality module through confirmatory factor analysis (CFA) techniques. After identifying the advantages and some limitations of ISSP items for a multidimensional operationalization of meritocratic beliefs, in a second study, we suggest an modified set of items that better taps the different dimensions of meritocracy. We examined the measuring properties of this new instrument using a sample of Chilean adults (N=2,141). Based on these results, we recommend improvements in measuring meritocratic beliefs in cross-national studies.
\newline
\newline

\textbf{Keywords}: meritocracy, survey methodology, social inequality, construct measurement
\end{abstract}

\setstretch{1.5}
\section{Introduction}\label{introduction}

Economic inequality and income concentration have become topics of growing concern over the last years, leading to a series of social upheavals in different societies as well as diverse critical analyses regarding the development of capitalism and its consequences (\citeproc{ref-piketty_capital_2014}{Piketty, 2014}; \citeproc{ref-streeck_politics_2014}{Streeck, 2014}). In such a context, the study of views, preferences, and perceptions of inequality has acquired relevance in the social sciences, in topics such as redistributive preferences (\citeproc{ref-alesina_fairness_2005}{Alesina \& Angeletos, 2005}; \citeproc{ref-dimick_models_2018}{Dimick et al., 2018}), the legitimization of economic inequality (\citeproc{ref-schroder_income_2017}{Schröder, 2017}) and the functioning of meritocracy (\citeproc{ref-atria_economic_2020}{Atria et al., 2020}; \citeproc{ref-duru-bellat_who_2012}{Duru-Bellat \& Tenret, 2012}; \citeproc{ref-mijs_paradox_2019}{Mijs, 2019}; \citeproc{ref-reynolds_perceptions_2014}{Reynolds \& Xian, 2014}). Within this research area, and particularly in times of economic and health crises, the ideal of meritocracy has been strongly challenged as an unfulfilled promise of modern societies that allows the perpetuation of social inequalities (\citeproc{ref-goldthorpe_myth_2003}{Goldthorpe, 2003}; \citeproc{ref-sandel_tyranny_2020}{Sandel, 2020}).

In general, meritocracy has been defined as a system of distribution of resources and rewards based on individual merit, which in its original conception is a combination of talent and effort (\citeproc{ref-young_rise_1962}{Young, 1962}). This traditional conception of merit places in a secondary position the possible interference of structural or non-meritocratic factors, such as inheritance, personal contacts, and luck (\citeproc{ref-breen_class_1999}{Breen \& Goldthorpe, 1999}; \citeproc{ref-land_we_2006}{Land, 2006}; \citeproc{ref-saunders_might_1995}{Saunders, 1995}; \citeproc{ref-yair_meritocracy_2007}{Yair, 2007}; \citeproc{ref-young_rise_1962}{Young, 1962}). Social psychology and sociology have studied the characteristics and consequences of beliefs in meritocracy, under the general hypothesis that a greater belief in meritocracy emphasizes the role of the individual over structural factors in personal achievements, leading to greater legitimization of inequalities (\citeproc{ref-hadjar_meritokratie_2008}{Hadjar, 2008}; \citeproc{ref-madeira_primes_2019}{Madeira et al., 2019}; \citeproc{ref-preminger_meritocracy_2020}{Preminger, 2020}; \citeproc{ref-trump_when_2020}{Trump, 2020}). Such research has raised criticism of meritocracy as a moral standard of distribution given the preponderant weight of non-meritocratic elements upon the individual status and social mobility (\citeproc{ref-arrow_meritocracy_2000}{Arrow et al., 2000}; \citeproc{ref-goldthorpe_myth_2003}{Goldthorpe, 2003}; \citeproc{ref-khan_privilege_2013}{Khan, 2013}; \citeproc{ref-markovits_meritocracy_2019}{Markovits, 2019}; \citeproc{ref-sandel_tyranny_2020}{Sandel, 2020}; \citeproc{ref-witteveen_reconsidering_2020}{Witteveen \& Attewell, 2020}).

Due to the role that meritocratic beliefs play in the justification of individual achievement (or failure) in contemporary societies (\citeproc{ref-hadjar_meritokratie_2008}{Hadjar, 2008}; \citeproc{ref-markovits_meritocracy_2019}{Markovits, 2019}; \citeproc{ref-sandel_tyranny_2020}{Sandel, 2020}), multiple studies have evaluated the relationship between meritocratic beliefs and personal and (or) contextual characteristics. For example, some studies have linked meritocracy to the reinforcement of socio-economic, gender, and ethnic stereotypes (\citeproc{ref-girerd_neoliberalism_2020}{Girerd \& Bonnot, 2020}, \citeproc{ref-girerd_neoliberalism_2020}{2020}; \citeproc{ref-madeira_primes_2019}{Madeira et al., 2019}), as well as the effects of meritocratic beliefs in educational (\citeproc{ref-generett_stories_2020}{Generett \& Olson, 2020}; \citeproc{ref-owens_engines_2020}{Owens \& de St Croix, 2020}) and organizational contexts (\citeproc{ref-aiello_new_2019}{Aiello et al., 2019}; \citeproc{ref-perez_advancing_2020}{Pérez \& Sabelis, 2020}). Most of such studies so far have used indicators from existing standard social surveys, such as the International Social Survey Programme (ISSP), to measure meritocracy. However, as we will show later, the concepts and instruments used for measuring meritocracy vary extensively among studies. In many cases, similar phenomena are associated with different indicators, and conversely, different phenomena are measured with similar indicators. Such approaches limit the comparability of studies and the capacity to understand the effects of meritocratic beliefs across different disciplines and lines of research, raising doubts about the validity of the instruments used to measure meritocratic beliefs. In this sense, this paper is aligned with recent discussions about the estimand (\citeproc{ref-lundberg_what_2021}{Lundberg et al., 2021}) in sociology, this is, the need for a clear definition of what we are estimating as well as the connection of this definition with statistical evidence.

Based on a critical analysis of different approaches to the empirical study of meritocracy to date ---most of them using data from the inequality module of ISSP---, this article identifies several conceptual and measurement issues. In order to address them, firstly, we propose a multidimensional conceptual framework for measuring meritocratic beliefs. The proposal consists in a first basic distinction between the dimensions of perceptions and preferences for meritocracy, often confused in the literature. Furthermore, the proposal distinguishes between meritocratic and non-meritocratic dimensions, as they would not be two poles of the same continuum as previous studies suggest. Secondly, we test the validity of our conceptual framework with two studies. Study 1 offers an operationalization and confirmatory analysis of the proposed multidimensional framework using available items from ISSP, inequality module 2009 (the reason to use this survey wave instead of the last one on social inequality from 2019 is that the 2009 version includes some key items used in meritocratic reserach that were removed in 2019). Taking into account on the advantages and limitations of the results obtained in study 1, and in order to gain additional evidence for the validity of the measurement instruments, study 2 suggests modifications and expansion of the items commonly used to measure meritocratic beliefs. This measurement proposal is tested with data available from a Chilean survey on meritocracy and inequality (2020).

\subsection*{The black box of meritocratic beliefs}\label{blackbox}
\addcontentsline{toc}{subsection}{The black box of meritocratic beliefs}

In the following, we discuss four critical aspects in studies conceptualizing and measuring meritocracy, based upon which we develop a conceptual and measurement proposal.

\begin{enumerate}
\def\labelenumi{\alph{enumi}.}
\tightlist
\item
  \emph{Conceptual components}: Is merit only effort?
\end{enumerate}

One recent definition of meritocracy by Mijs (\citeproc{ref-mijs_paradox_2019}{2019}) is the following: ``When I discuss meritocracy beliefs, I am referring to citizens' belief in the importance of hard work relative to structural factors.'' (\citeproc{ref-mijs_paradox_2019}{Mijs, 2019, p. pg.9}). In the subsequent operationalization, this is associated with the following question and indicator: ``how important you think it is for getting ahead in life: (a) hard work'', scored on a likert scale. The assumptions behind such a definition are worth discussing in light of the conceptual meaning of meritocracy and its possibilities of operationalization.

The item used by Mijs (\citeproc{ref-mijs_paradox_2019}{2019}) is part of an item's battery which is available in several international surveys (as ISSP), usually called ``reasons to get ahead''. This battery displays a series of indicators related to what people consider important to get ahead: hard work, education, ambition, a wealthy family, the right connections, religion, race, and gender. Therefore, considering only one of the items means that other aspects associated with talent, such as education, would not be deemed meritocratic. As he points out: ``Hard work is arguably the most meritocratic part of Michael Young's equation: `Merit = Intelligence + Effort', for the simple fact that intelligence itself is influenced by a non-meritocratic factor: who your parents happen to be'' (p.5). In such conceptual and measurement approach of meritocratic beliefs, we can observe a couple of strong conceptual assumptions: a) effort would not depend on parental influence (at least not as strongly as intelligence), and b) talent (as innate ability) is not meritocratic (contrary to Michael Young's' original conceptualization). This conceptual and measurement-based assumption is found in other studies, which also assume effort to be the main and only aspect of meritocracy (\citeproc{ref-bubak_perceptions_2019}{Bubak, 2019}; \citeproc{ref-girerd_neoliberalism_2020}{Girerd \& Bonnot, 2020}), raising the question: Is effort the only (or main) dimension behind the concept of meritocracy?

We argue that whether talent (as intelligence and ability) is or is not considered meritocratic is an interesting topic to discuss from a philosophical point of view. However, when it comes to survey research on meritocratic beliefs, the weight of effort or talent in what it is deemed meritocratic should be left to public scrutiny. Furthermore, considering talent as part of meritocracy certainly opens some appealing avenues of research. For instance, some studies suggest that, for the elites, meritocracy is related to talent, whereas effort is more characteristic of the meritocracy of the middle and lower classes (\citeproc{ref-atria_economic_2020}{Atria et al., 2020}). Therefore, we suggest that a concept (and measure) of meritocracy benefits from including both talent and effort, as in Young's original conception.

\begin{enumerate}
\def\labelenumi{\alph{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  \emph{Beliefs?}
\end{enumerate}

Several approaches to the empirical study of meritocracy based on public opinion surveys refer to the concept of \emph{beliefs}, but with wide differences in meanings and operationalizations. To illustrate this point, let us consider the commonly used ``reasons to get ahead'' items' battery, mentioned above. Some versions of this battery ask ``how important you think it is for getting ahead in life'' and then list various factors, whereas another version of this same battery - sometimes presented along the previous one - asks about ``how important you think it \emph{should} be \ldots{}'', and then listing the same concepts. Therefore, the question raised here is: Which one of both is a ``belief'': what \emph{is} (the first one) or what it \emph{should} be (the second one)?

The term belief has an ambiguous character in the literature, conceived as ``idea-elements'' by Converse (\citeproc{ref-converse_nature_1964}{1964}) or ``considerations'' by Zaller (\citeproc{ref-zaller_nature_1992}{1992}). As Kluegel \& Smith (\citeproc{ref-kluegel_beliefs_1986}{1986}) pointed out about the scope of beliefs: ``This usage encompasses such more specific social-psychological concepts as values, perceptions, and attitudes'' (p.30). Therefore, the term \emph{beliefs} used to cover almost anything related to subjective factors. In this regard, a relevant distinction in the field of inequality beliefs was made by Janmaat (\citeproc{ref-janmaat_subjective_2013}{2013}): ``Perceptions refer to subjective estimates of existing inequality (i.e., thoughts about what is). Beliefs are here defined as normative ideas about just inequality (i.e., thoughts about what should be)''(p.359). Several papers dealing with meritocracy use the term beliefs (i.e.~what should be), while referring to perceptions (i.e., what is). For instance, in Reynolds \& Xian (\citeproc{ref-reynolds_perceptions_2014}{2014}), the term belief is used to talk about what Janmaat (\citeproc{ref-janmaat_subjective_2013}{2013}) refers to as perceptions, whereas other authors use general terms such as attitudes (\citeproc{ref-kunovich_systems_2007}{Kunovich \& Slomczynski, 2007}). The first attempt to shed light on this issue in meritocracy research was made by Duru-Bellat \& Tenret (\citeproc{ref-duru-bellat_who_2012}{2012}), who used the question ``how important should the number of years spent in education and training be in deciding how much money people ought to earn?'' as a proxy for ``desired'' meritocracy (normative beliefs). They then determined ``perceived'' meritocracy, using the questions: ``Would you say that in your country, people are rewarded for their efforts?'' and ``\ldots{} people are rewarded for their skills?''.

Is the belief in meritocracy a perception or a preference with normative meaning? To expand the analytical conceptual framework, we believe that both dimensions should be considered, as proposed by Duru-Bellat \& Tenret (\citeproc{ref-duru-bellat_who_2012}{2012}). Such an approach opens up the possibility of analyzing whether perceptions and preferences are related (i.e., have a high correlation), or constitute independent aspects of the same phenomenon. As Son Hing et al. (\citeproc{ref-sonhing_merit_2011}{2011}) has pointed out, ``People can believe that outcomes ought to be distributed based on merit and yet vary in their perceptions of whether this is how society currently operates'' (p.~435). In other words, normative beliefs should be considered while taking perceptions into account: a strong normative belief in meritocracy may mean something different to someone perceiving high meritocracy than to someone perceiving low meritocracy. To avoid the confusion generated by the term ``belief'', we propose the terms meritocratic preferences (``what should be''), and meritocratic perceptions (``what is''), as they better reflect the two facets of meritocracy under scrutiny (\citeproc{ref-castillo_meritocracia_2019}{Castillo et al., 2019}).

\begin{enumerate}
\def\labelenumi{\alph{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  \emph{Non-meritocratic aspects}
\end{enumerate}

Some research in meritocracy considers aspects usually opposed to effort and talent for personal achievement, for instance, family status and the use of personal advantages (such as contacts or having a wealthy family) to get ahead in life. This distinction refers to the classical achieved and ascribed status dimension from Linton (\citeproc{ref-linton_study_1936}{1936}). For instance, Kunovich \& Slomczynski (\citeproc{ref-kunovich_systems_2007}{2007}) used an items' battery following the question ``How important each should be in deciding pay\ldots{}'' (as Duru-Bellat \& Tenret (\citeproc{ref-duru-bellat_who_2012}{2012}) for desired meritocracy). They consider factors such as education and responsibility as meritocratic, giving them a value of 1 is considered ``essential'' in the scale response, whereas factors such as having a family or children are valued as 1 when rated as ``not important at all'' (i.e.~reverse coded). The assumption behind this approach is that rejecting a supposed ascribed or non-meritocratic aspect (such as having family and/or children) implies a stronger belief in meritocracy. A similar approach of reverse-coding non-meritocratic items was taken by Newman et al. (\citeproc{ref-newman_false_2015}{2015}), using the same principles applied in the ``Preference for the Merit Principle Scale'' (\citeproc{ref-davey_preference_1999}{Davey et al., 1999}).

The assumption that meritocratic and non-meritocratic elements are poles of the same continuum was tested by Reynolds \& Xian (\citeproc{ref-reynolds_perceptions_2014}{2014}) using the ``get ahead'' perceptions' battery items mentioned above. They considered education, ambition, and hard work as meritocratic, whereas factors such as family wealth and connections were classified as non-meritocratic. Despite making this distinction, in the operationalization the authors end up subtracting one dimension from the other, thus coming back to the assumption that they are two poles of the same continuum, as Kunovich \& Slomczynski (\citeproc{ref-kunovich_systems_2007}{2007}) also did. Similarly, Roex et al. (\citeproc{ref-roex_attitudes_2018}{2018}) used ISSP indicators for perceived meritocracy and non-meritocracy to build a single score by reverse coding the non-meritocratic items. Therefore, the treatment of non-meritocratic items has been rather inconsistent across studies and the assumption that they are the simple opposite of meritocracy certainly requires further conceptual elaboration and empirical assessment. For instance, a such distinction could open the door to studying to what extent different and contradictory views of meritocracy can coexist, to what extent individuals and societies differ in this regard, and of course the associated factors to such differences. Although the use of the concept ``non-meritocratic'' is very broad, we prefer to use it instead of other terms like ascription, given that this last one leaves out aspects such as the use of personal relationships as a way to achieve success.

\begin{enumerate}
\def\labelenumi{\alph{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
  \emph{Accounting for measurement error}
\end{enumerate}

Finally, most meritocracy studies have not adequately considered the issue of latent structures and measurement error (\citeproc{ref-ansolabehere_strength_2008}{Ansolabehere et al., 2008}; \citeproc{ref-bollen_structural_1989}{Bollen, 1989}; \citeproc{ref-brown_confirmatory_2015}{Brown, 2015}), as they mostly use single indicators and/or simple average indexes for measuring meritocracy. Such a strategy assumes that the latent construct is measured perfectly (i.e., no error or residual variance) by the selected indicators, going as far as to propose that ``\ldots{} In choosing this strategy of index construction, we argue that support for meritocracy is not a latent variable'' (\citeproc{ref-kunovich_systems_2007}{Kunovich \& Slomczynski, 2007, pp. 653--654}). Although Reynolds and Xian (\citeproc{ref-reynolds_perceptions_2014}{2014}) made advances by conducting a principal component analysis of meritocratic and non-meritocratic dimensions, they finally decided to build a sum index despite proving a multidimensional latent structure.

There is a relevant trade-off regarding measurement error estimation in survey instruments. Including multiple items per conceptual dimension opens the possibility of validity studies (at least in terms of the construct) by using factor analysis techniques, particularly confirmatory ones. However, increasing the number of items reduces the probability of usability of the scale in survey research given space limitations. This validity-usability conundrum is not easy to solve and, as we will show later in the methods section, we opted for a minimalistic approach . Such approach is oriented to the usability of the measurement instrument, but still giving the possibility of confirmatory factor analysis.

\subsection*{A conceptual proposal for studying meritocracy}\label{instrumentprop}
\addcontentsline{toc}{subsection}{A conceptual proposal for studying meritocracy}

Based on the previous assumptions and limitations identified in the empirical study of meritocracy, mostly using the inequality module of ISSP, we propose a measurement framework for the study of meritocratic beliefs with the following characteristics:

\begin{itemize}
\item
  \emph{Multidimensionality}, incorporating previous distinctions between preferences and perceptions, as well as between meritocratic and non-meritocratic aspects.
\item
  Multiple indicators for each dimension to \emph{account for measurement error} in a confirmatory factor analysis context.
\item
  Based on \emph{previous indicators} as much as possible to keep comparability between studies. As most of the survey research on meritocracy so far has been made with the inequality module of the ISSP survey, we propose a comprehensive operationalization suitable for this dataset, as well as a novel instrument that attempts to rescue as much as possible the original ISSP items.
\item
  \emph{Brief}, as to be used in regular public opinion surveys (\citeproc{ref-davidov_measurement_2009}{Davidov, 2009}). In this regard, it differs from the proposal of ``Preference for the Merit Principle Scale'' (\citeproc{ref-davey_preference_1999}{Davey et al., 1999}), as they use 15 items for just one dimension (aside from the problem of reverse-coding non-meritocratic items).
\end{itemize}

The proposed conceptual and measurement framework is depicted in Figure \ref{fig:merit-model}:

\begin{figure}[H]

{\centering \includegraphics[width=0.85\linewidth]{input/images/Fig1.generalf}

}

\caption{Conceptual Model of perception and preferences for meritocracy and non-meritocracy}\label{fig:merit-model}
\end{figure}

The columns ``Perceptions'' and ``Preferences'' represent the distinction between these two concepts, usually confused under the label ``beliefs'' (\citeproc{ref-castillo_meritocracia_2019}{Castillo et al., 2019}). Perceptions refer to the extent to which people observe that meritocracy functions or apply in their society, which relates to items such as ``Hard work is important to get ahead in society''. Preferences refer to normative expectations that are frequently linked to a ``should'' expression (e.g.~whether hard work should be related to payment). The rows in Figure \ref{fig:merit-model} consider the distinction between meritocratic and non-meritocratic dimensions (\citeproc{ref-reynolds_perceptions_2014}{Reynolds \& Xian, 2014}), often treated as different ends of the same continuum in previous research. Non-meritocratic elements refer to the use of resources such as personal contacts and/or family advantages to get ahead in life.

The ISSP survey (inequality module) has some items that researchers have used to measure each of the four different dimensions of meritocracy depicted here, but in a rather inconsistent manner. In the next section, we propose a classification of the ISSP meritocracy items into the different dimensions suggested in our multidimensional model depicted in Figure \ref{fig:merit-model} as well as an empirical test of this model.

\section{Methodology}\label{methodology}

The analysis is organized into two studies. The first study is an analysis of the multidimensional model for measuring meritocratic beliefs using ISSP data. Based on the results of this first study, the second study recommends some modifications and expansions of the ISSP items in order to tap better each of the dimensions identified in the study of meritocratic beliefs.

\section{Study 1: Analyzing meritocratic beliefs with ISSP data}\label{study-1-analyzing-meritocratic-beliefs-with-issp-data}

\subsection{Data}\label{data}

The data corresponds to the social inequality module from the International Social Survey Programme (ISSP), the most specialized international comparative survey in perceptions, attitudes, and beliefs about inequality-related issues (\citeproc{ref-issp-research-groupInternationalSocialSurvey2017}{ISSP-Research-Group, 2017}). This wave corresponds to 2009 and covers attitudes towards a series of topics dealing with social inequality across 41 countries and 56,021 individuals. Although there is data available from this module for a more recent wave (2019), the 2009 wave contains relevant items for studying meritocracy that unfortunately are not present in the last wave (details about which items are missing in the last 2019 wave are presented in study 2, Table \ref{tab:table-indicadores}). Table \ref{tab:countries} (appendix) shows details about the sample of countries in the ISSP 2009 data.

\subsection{Variables}\label{variables}

There are a series of indicators in the ISSP survey that we classified in our meritocracy conceptual scheme, as presented below in Table \ref{tab:table-issp}:

\begin{table}[!h]

\caption{\label{tab:table-issp}Items of the ISSP meritocratic perceptions and preferences measures}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2cm}>{\raggedright\arraybackslash}p{5cm}>{\raggedright\arraybackslash}p{8cm}}
\toprule
Component & Dimensions & Item\\
\midrule
 &  & How important is having ambition?\\
\cmidrule{3-3}
 & \multirow{-2}{5cm}{\raggedright\arraybackslash Meritocratic} & How important is hard work?\\
\cmidrule{2-3}
 &  & How important is coming from a wealthy family?\\
\cmidrule{3-3}
 &  & How important is having well-educated parents?\\
\cmidrule{3-3}
 &  & How important is a person's race?\\
\cmidrule{3-3}
 &  & How important is being born a man or a woman?\\
\cmidrule{3-3}
 &  & How important is knowing the right people\\
\cmidrule{3-3}
\multirow{-8}{2cm}{\raggedright\arraybackslash Perception} & \multirow{-6}{5cm}{\raggedright\arraybackslash Non-meritocratic} & How important is having political connections\\
\cmidrule{1-3}
 &  & How well he or she does the job?\\
\cmidrule{3-3}
 & \multirow{-2}{5cm}{\raggedright\arraybackslash Meritocratic} & How hard he or she works at the job?\\
\cmidrule{2-3}
 &  & What is needed to support a family?\\
\cmidrule{3-3}
\multirow{-4}{2cm}{\raggedright\arraybackslash Preference} & \multirow{-2}{5cm}{\raggedright\arraybackslash Non-meritocratic} & Whether the person has children to support?\\
\bottomrule
\end{tabular}
\end{table}

\begin{itemize}
\tightlist
\item
  \emph{Perception of meritocracy/non-meritocracy}: for operationalizing perceptions the closest set of ISSP's indicators comes from the question asking about perceptions for opportunities to get ahead, which are usually considered as ``meritocratic beliefs'' in previous studies. The general heading of the battery is: \emph{``To begin we have some questions about opportunities for getting ahead. Please tick one box for each of these to show how important you think it is for getting ahead in life.''} This is followed by a list of statements to be rated from 1 to 5: essential, very important, fairly important, not very important, not important at all.
\end{itemize}

The classification of the items is based on criteria of internal motivation (meritocratic) and structural constraints (non-meritocratic). There were two items from the battery that were excluded from the analysis as they would not fit into the classification. The first one was ``having good education yourself'' since it was not clear whether this could be due to individual motivation or system opportunities, and the second was ``giving bribes'', as introduced elements of criminality that were beyond a non-meritocratic perception.

\begin{itemize}
\tightlist
\item
  \emph{Preferences for meritocracy-non meritocracy}: for the operationalization of normative preferences we used a list of items related to reasons for pay battery. The ISSP question was: \emph{In deciding how much people ought to earn, how important should each of these things be, in your opinion}, rated in the same \emph{essential-non important at all} scale (1 to 5) as the questions for meritocratic perceptions.
\end{itemize}

\subsection{Methods}\label{methods}

The estimation of the measurement model was performed using Confirmatory Factor Analysis (CFA). CFA was conducted using the \texttt{lavaan} R package (version 0.6-3; Rosseel (\citeproc{ref-rosseel_lavaan_2012a}{2012})), with diagonally weighted least squares (DWLS) estimation due to the items' ordinal level of measurement (\citeproc{ref-kline_principles_2016}{Kline, 2016}; \citeproc{ref-rosseel_lavaan_2012a}{Rosseel, 2012}). As recommended by Brown (\citeproc{ref-brown_confirmatory_2015}{2015}), we assessed model fit by jointly considering the comparative fit index and Tucker-Lewis Index (CFI and TLI; acceptable fit \textgreater{} 0.95), Root of the average squared residual approximation (RMSEA; acceptable fit \textless{} 0.08), Chi-square: (p-value; acceptable fit \textgreater{} 0.05, and Chi-square ratio \textgreater{} 3).

\subsection{Results}\label{results}

\subsubsection{Descriptive analyses}\label{descriptive-analyses}

Figure \ref{fig:likert-issp} shows the distribution of responses across the selected items in their corresponding dimensions. On the one side, we observe a high degree of importance attributed to factors such as hard work and ambition for getting ahead, concentrating 94.9\% and 92.8\% in the \emph{fairly important} to \emph{essential} categories, respectively. By contrast, the perception of non-meritocratic aspects is rated lower than the meritocratic ones, particularly for gender, race, and political connections. Regarding normative preferences, we observe that the meritocratic ones are deemed as important (from fairly to essential) for nearly the whole sample, decreasing slightly for the non-meritocratic ones. Still, we have to consider that the preferences for non-meritocratic aspects, in this case, refer to the distributive principle of need rather than personal background (as in the perceptions of non-meritocracy). Such inconsistency between the target of the items for perceptions an preferences is one of the issues to deal with in study 2.

\begin{figure}[H]

{\centering \includegraphics[width=0.8\linewidth]{paper-blinded-RR-IJS_files/figure-latex/likert-issp-1}

}

\caption{Distribution of responses in the ISSP meritocracy items}\label{fig:likert-issp}
\end{figure}

In terms of the association between the indicators, Figure \ref{fig:corr-issp} shows the polychoric correlation matrix. Firstly, according to the criteria suggested by Cohen (\citeproc{ref-cohen_statistical_1988}{1988}) for effect size, we observe in general that the moderate to high associations are between the pairs of items representing each of the four dimensions described in Table \ref{tab:table-issp}. The exception here is the dimension of non-meritocratic perception, in which six items appear mostly associated by pairs according to their specific topics (family background, personal background, and connections). Still, between this same set of items, moderate correlations could anticipate a single latent factor underlying non-meritocratic perceptions, which is tested next in the confirmatory factor analysis model.

\begin{figure}[H]

{\centering \includegraphics[width=0.85\linewidth]{paper-blinded-RR-IJS_files/figure-latex/corr-issp-1}

}

\caption{Perceptions and preferences for ISSP  meritocracy items' polychoric correlations}\label{fig:corr-issp}
\end{figure}

\subsubsection{Confirmatory Factor Analysis}\label{confirmatory-factor-analysis}

Table \ref{tab:sum-fit-issp} shows the results of the estimation of two confirmatory models. The first one (First Order) corresponds to a model that estimates four factors, each for one of the dimensions of the multidimensional framework for meritocratic beliefs presented in Table \ref{tab:table-issp}, with only regular fit indicators (CFI=0.959, TLI=0.944, RMSEA=0.098, \(\chi^2\)(df=48)= 21308.535). Further analysis showed that the sources for poor fit were mostly related to the items of the non-meritocratic perceptions, which as mentioned above displayed correlations among them not taken into account in this model specification.

\begin{table}[H]

\caption{\label{tab:sum-fit-issp}Summary fit indices according to model}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lrlrrrrr}
\toprule
Model & $N$ & Estimator & $\chi^2$ & df & CFI & TLI & RMSEA\\
\midrule
First order & 46594 & DWLS & 21308.535 & 48 & 0.959 & 0.944 & 0.098\\
Second order & 46594 & DWLS & 4472.369 & 45 & 0.992 & 0.988 & 0.046\\
\bottomrule
\end{tabular}
\end{table}

Attending to the sources of poor fit from the first estimated confirmatory model, we specified a second model that keeps the basic four-dimensional structure but generates an additional model for the non-meritocratic perceptions as depicted in Figure \ref{fig:meas02-issp}. In this model, the pairs of items in this dimension form three latent factors which at the same time generate a second-order factor of non-meritocratic perceptions, improving the fit of the model significantly.

\begin{figure}[H]

{\centering \includegraphics[width=0.85\linewidth]{output/images/figure4}

}

\caption{Second-order confirmatory factor analysis model using ISSP indicators of Perceptions and Preferences for Meritocracy}\label{fig:meas02-issp}
\end{figure}

Regarding the correlations between the factors in Figure \ref{fig:meas02-issp}, we observe that perceptions are correlated with preferences, but more strongly for the meritocratic (\(r=0.46, p<0.01\)) than for the non-meritocratic (\(r=0.27,p<0.01\)) dimensions. Secondly, both perceptual dimensions (meritocratic and non-meritocratic) depict a moderate positive correlation (\(r=0.30, p<0.01\)), suggesting that they are not the opposite poles of the same continuum as some previous studies assume. Something similar occurs for the correlation between the two dimensions of preferences (\(r=0.25,p<0.01\)).

Based on the previous measurement validation, next we assess some of the potential of the proposed measurement strategy for cross-national comparisons of meritocratic beliefs. In this line, Figure \ref{fig:scat1} depicts the correlation of factor scores for meritocratic perceptions and meritocratic preferences at the country level (\(r=0.65,p<0.01\)). For instance, Bulgaria (BGR) and the United States (USA) appropriately represent the positive association between perception and preferences. In other words, countries that perceived merit in their society as an essential factor to get ahead in life also consider that effort and talent should be determinant factors in how much an individual earns. On the other hand, in countries with lower perceived meritocracy, preferences are also low, where Denmark (DNK) and Venezuela (VEN) are representative cases of this association.

Another finding depicted by Figure \ref{fig:scat1} is the positive association of non-meritocratic perceptions with meritocratic preferences (\(r=0.55,p<0.01\)), showing that countries like Bulgaria (BGR) and Philipines (PHL) represent cases with higher non-meritocratic perceptions and higher preferences for meritocracy. In other words, in societies where people perceive higher importance of non-meritocratic aspects to get ahead in life, people also prefer that merit should be a determinant factor in wages. On the other side, countries like Sweden (SWE) or Denmark (DNK), both Scandinavian societies with strong welfare systems, represent contexts in which non-meritocratic aspects are not described as important, and merit is not considered as a factor that should be relevant for wages. This finding, particularly the positive association between non-meritocratic perception and meritocratic preferences, opens the question of to what extent meritocratic preferences are capable to hold even under evidence of a large amount of perceived non-meritocracy .

\begin{figure}[H]

{\centering \includegraphics[width=0.85\linewidth]{paper-blinded-RR-IJS_files/figure-latex/scat1-1}

}

\caption{Cross-national comparison of non-meritocratic and meritocratic perceptions with meritocratic preference}\label{fig:scat1}
\end{figure}

Another possibility to explore within this multidimensional framework of meritocratic beliefs is the differences between countries. For instance, when comparing societies based on their level of industrialization (\citeproc{ref-ishida_trends_2005}{Ishida \& Miwa, 2005}), Figure \ref{fig:scat2} shows that early and late-industrialized countries exhibit a similar association between meritocratic perceptions and preferences (\(r=0.67, p<0.05\); \(r=0.68, p<0.01\), respectively). However, when analyzing the association between non-meritocratic perception and meritocratic preferences, industrialized nations exhibit a strong positive correlation (\(r=0.75, p<0.01\)), which is not the case for late-industrialized countries (\(r=0.29, p>0.05\)). A possible interpretation is that the perception of non-meritocracy in less industrialized societies threatens the meritocratic ideal, which is not the case for the industrialized ones. Still, this preliminary analysis only attempts to show some possiblities that open up when introducing a multidimensional perspective for studying meritocratic beliefs

\begin{figure}[H]

{\centering \includegraphics[width=0.85\linewidth]{paper-blinded-RR-IJS_files/figure-latex/scat2-1}

}

\caption{Cross-national comparison between industrialized and late-industrialized}\label{fig:scat2}
\end{figure}

\subsection{Discussion Study 1}\label{discussion-study-1}

ISSP is - and probably it will continue to be - the most widely used dataset for comparative studies of meritocratic beliefs. Although several ISSP items were not conceived originally with this end, they have been extensively used to operationalize different aspects of meritocracy. In this study, we ordered and classified ISSP items, which have been previously used for measuring meritocratic beliefs, into four different dimensions: perceptions and preferences, and meritocracy and non-meritocracy. The confirmatory analysis gave evidence of an adequate fit of the multidimensional model to the ISSP data, this is, it is possible to identify perceptions and preferences for meritocracy and non-meritocracy as four related but different constructs. Regarding the non-meritocratic dimensions, we can preliminary conclude that the wide use of reverse-coding for non-meritocratic items attempting to measure meritocratic perceptions or preferences, is not an adequate operationalization. Meritocracy and non-meritocracy are related but different concepts, meaning that it is theoretically possible to find different combinations at individual and contextual levels. This finding opens a series of avenues for future research that are considered in the conclusions.

Despite the advantages of using ISSP data for studying meritocratic beliefs, there are some aspects to be aware of regarding the measurement quality offered by the available items. Some of the items are not really coherent across dimensions, as for instance, meritocratic perceptions are related to ambition and hard work, whereas meritocratic preferences deal with how well a job is done. Such differences impose limitations for a comparative analysis between pereptions and preferences. A better operationalization would require more alignment between the different sides of the same concept, this is, that perceptions and preferences point to the same object of analysis. Furthermore, the instrument is rather unbalanced given the larger amount of non-meritocratic perception items (six) when compared to the number items related to the other dimensions (two). These limitations are taken into account next in the design of study 2.

\section{Study 2: A proposal for a multidimensional measure of meritocratic beliefs}\label{study-2-a-proposal-for-a-multidimensional-measure-of-meritocratic-beliefs}

Taking the analysis of meritocratic beliefs with ISSP as a departing point, the following study offers and test a new measurement instrument that attempts to overcome some limitations identified in the previous study. In this sense, the analysis with ISSP data can be considered as an \emph{inductive} exercise, taking the items and data available and attempting to adjust this information to the proposed multidimensional conceptual model. In the following study we take a complementary \emph{deductive} approach, departing from the conceptual model for the design of a measurement instrument. However, this second approach does not start from scratch, as it is based on ISSP items as much as possible for the sake of comparability with previous studies. The two major differences are: a) it incorporates a balanced and minimalistic representation of items for each of the four dimensions,and b) attempts to maintain the coherence in the topics asked across dimensions (for instance, ``effort'' in items for both perceptions and preferences).

\subsection{Data}\label{data-1}

The data comes from an online survey that was part of a larger study on meritocracy and distributive preferences developed in Chile in 2020, funded by the Chilean National Scientific Agency (ANID). The questionnaire was programmed in Qualtrics and the fieldwork was conducted by an external online survey agency (\href{www.netquest.com}{netquest.com}) between December 2019 and January 2020. The sample was selected from a non-probabilistic quota design in three large cities in Chile (Santiago, Concepción \& Antofagasta). The quotas for gender, age, and educational levels were generated based on a survey by the Public Studies Center (\citeproc{ref-cep_encuesta_2019}{CEP, 2019}), which is a well-regarded counterpart agency of the ISSP (International Social Survey Programme) in Chile. A total sample of 2,141 individuals was collected, excluding those who did not sign the informed consent and those not answering the meritocracy instrument. There were no significant differences between our sample and the wider population for most socio-demographic characteristics, except for the educational level (see Table \ref{tab:rep-samp} in Appendix). As is often the case with online surveys, there were some limitations in achieving the quotas for lower educational levels (\citeproc{ref-boas_recruiting_2020}{Boas et al., 2020}; \citeproc{ref-zhang_quota_2018}{Zhang et al., 2018}). This survey was designed as a three-wave panel and the data used in this analysis corresponds to the first wave.

All participants signed the required informed consent and the survey implementation was approved by the ethical committee from Pontificia Universidad Católica de Chile, protocol ID 150811008, April 6th 2016.

\subsection{Study design}\label{study-design}

\subsubsection{Instrument}\label{instrument}

The proposed items' battery of meritocratic perceptions and preferences consisted of eight indicators, two for each of the four dimensions listed earlier: perceptions (meritocratic/non-meritocratic) and preferences (meritocratic/non-meritocratic). Two main criteria oriented this design. The first one was usability in social survey research, for which the scale length must attend spacer restrictions in such studies. Nevertheless, at the same time, the scale should offer possibilities of accounting for measurement error through factor analysis techniques. Although a rule of thumb in latent measurement suggests at least three indicators per dimension, we still opted for two to count with a brief scale. Although such a decision posits some limitations for modeling measurement error, still allows for factor analysis and is a common approach in previous social survey studies with limited items (for instance see Davidov (\citeproc{ref-davidov_measurement_2009}{2009})). The second criterion is to expand the possibilities of comparability with previous studies on meritocracy, by considering some of the items previously used in this research area. To this regard, some of the items were adapted from the items battery ``reasons to get ahead'' (ISSP/GSS), which has been widely used for operationalizing meritocracy in international research (\citeproc{ref-duru-bellat_who_2012}{Duru-Bellat \& Tenret, 2012}; \citeproc{ref-mijs_paradox_2019}{Mijs, 2019}; \citeproc{ref-reynolds_perceptions_2014}{Reynolds \& Xian, 2014}).

The items, organized according to their respective dimensions, are presented in Table \ref{tab:table-indicadores}. These eight likert-type items have five response alternatives, ranging from ``Completely disagree''(1) to ``Completely agree'' (5). The table presents an additional column with the respective equivalent items from ISSP 2009/2019. With this information it is possible to have a more clear picture of the areas in which the proposed measurement instrument is similar to ISSP and in which ones it differs.

\begin{table}[!h]

\caption{\label{tab:table-indicadores}Items' battery for the multidimensional measurement of meritocratic beliefs and their equivalents in ISSP}
\centering
\fontsize{10}{12}\selectfont
\begin{tabu} to \linewidth {>{\raggedright\arraybackslash}p{1.5cm}>{\raggedright\arraybackslash}p{2 cm}>{\raggedright}X>{\raggedright}X}
\toprule
Component & Dimensions & Items & Items ISSP (2009,2019)\\
\midrule
 &  & Those who make more effort get greater rewards than those who make less effort. & How important is hard work?\\
\cmidrule{3-4}
 & \multirow{-2}{2 cm}{\raggedright\arraybackslash Meritocratic} & Those with more talent get greater rewards than those who have less talent. & -\\
\cmidrule{2-4}
 &  & Those who have rich parents manage to get ahead. & How important is coming from a wealthy family?\\
\cmidrule{3-4}
\multirow{-4}{1.5cm}{\raggedright\arraybackslash Perception} & \multirow{-2}{2 cm}{\raggedright\arraybackslash Non meritocratic} & Those who have good contacts manage to get ahead. & How important is knowing the right people\\
\cmidrule{1-4}
 &  & Those who make more effort should get greater rewards than those who make less effort. & How hard he or she works at the job?*\\
\cmidrule{3-4}
 & \multirow{-2}{2 cm}{\raggedright\arraybackslash Meritocratic} & Those who have more talent should get greater rewards than those who have less talent. & How well he or she does the job?*\\
\cmidrule{2-4}
 &  & It is fine if those with rich parents get ahead. & -\\
\cmidrule{3-4}
\multirow{-4}{1.5cm}{\raggedright\arraybackslash Preference} & \multirow{-2}{2 cm}{\raggedright\arraybackslash Non meritocratic} & It is fine if those with good contacts get ahead. & -\\
\bottomrule
\multicolumn{4}{l}{\rule{0pt}{1em}\textit{Note: } Those with an asterisk (*) are only present in 2009}\\
\end{tabu}
\end{table}
\pagebreak

\subsubsection{Administration sets}\label{administration-sets}

To evaluate the effect of indicator ordering in the responses, three different versions of items' order were designed and randomly assigned to the respondents, as depicted in Figure \ref{fig:appmod}. The scale was presented to the first group (\emph{n = 712}) in the order shown in Table \ref{tab:table-indicadores} according to perceptions and preferences. For the second group (\emph{n = 717}), the order was reorganized according to perceptions and preferences over the same topic, e.g., for the topic of hard work, the item about perception was followed by the item about preference and the same for the rest of the topics. Finally, for the third group (\emph{n = 712}), the items were presented as completely randomized.

\begin{figure}[H]

{\centering \includegraphics[width=0.85\linewidth]{output/images/figure6}

}

\caption{Survey flow}\label{fig:appmod}
\end{figure}

\subsection{Methods}\label{methods-1}

To test the scale's underlying constructs, we employed confirmatory factor analysis models (CFA). The models estimated one factor for each of the four proposed dimensions presented in Table \ref{tab:table-indicadores}. As in study 1, CFA was performed with the \texttt{lavaan} R package (version 0.6-3; Rosseel (\citeproc{ref-rosseel_lavaan_2012a}{2012})), with diagonally weighted least squares (DWLS) estimation due to the items' ordinal level of measurement (\citeproc{ref-kline_principles_2016}{Kline, 2016}; \citeproc{ref-rosseel_lavaan_2012a}{Rosseel, 2012}). The fit indexes and cut-off criteria were the same as the ones used in Study 1.

A pre-registration was made in the OSF platform, available at the following link: \href{https://osf.io/z45y2/?view_only=adcd496887f2471597b1d6ffb46e492d}{https://osf.io/z45y2}. This pre-registration includes the hypotheses regarding the four-dimensional conceptual model underlying the scale, the variable measurement levels, the statistical tests to be performed with their respective evaluation parameters, and other important aspects of the research design.

\subsection{Results}\label{results-1}

\subsubsection{Descriptive analyses}\label{descriptive-analyses-1}

The graphs presented in Figure \ref{fig:plotlikert} display disaggregated and comparable information of the different response categories for each item. Generally, there is more agreement in the perception of non-meritocratic items than in meritocratic ones, while in the preferences, the opposite occurs. Regarding preferences, the predominant role of effort over talent as a criterion of meritocratic preference is noteworthy. All in all, the descriptive results show a critical view of meritocracy, perceiving the operation of non-meritocratic aspects over meritocratic ones, whereas in the preferences the opposite occurs.

\begin{figure}[H]

{\centering \includegraphics[width=0.85\linewidth]{paper-blinded-RR-IJS_files/figure-latex/plotlikert-1}

}

\caption{Distribution of responses in the Merit Scale items}\label{fig:plotlikert}
\end{figure}

Attending now to the association among the scale items, Figure \ref{fig:corpoly} shows the items' polychoric correlations. There are three main aspects to highlight from this correlation matrix. Firstly, and as expected, the largest correlations are between indicators that correspond to the same factors (dimensions) behind the conceptual model (e.g., perception of meritocracy by effort and by talent, \(r=0.52,p<0.01\)). Secondly, among these correlations, the highest are those between the non-meritocratic dimensions, both in perceptions (\(r=0.73,p<0.01\)) and preferences (\(r=0.61,p<0.01\)). Thirdly, both items for meritocratic preferences (E and F) are the ones that mostly correlate with the rest of the perceptual items, showing medium to high correlations and indicating that the perception of non-meritocracy would be related to larger meritocratic preferences. Finally, similar to the results of study 1 with ISSP data, we observe no considerable negative correlations between meritocratic and non-meritocratic aspects, undermining previous studies' assumptions that these dimensions would be the opposite poles of one same continuum (\citeproc{ref-reynolds_perceptions_2014}{Reynolds \& Xian, 2014}).

\begin{figure}[H]

{\centering \includegraphics[width=0.85\linewidth]{paper-blinded-RR-IJS_files/figure-latex/corpoly-1}

}

\caption{Perceptions and preferences for meritocracy items' polychoric correlations}\label{fig:corpoly}
\end{figure}

\subsubsection{Confirmatory Factor Analysis}\label{confirmatory-factor-analysis-1}

The present section describes the results of the confirmatory factor analysis estimation. The model estimates four latent factors: perception meritocratic, perception non-meritocratic, preferences meritocratic, and preferences non-meritocratic. Each factor is estimated based on two scale items as detailed in Table \ref{tab:table-indicadores}.

The first step in the analysis consists of comparing the model fit indicators for the three versions of the instrument that were randomly assigned to the participants: order according to perceptions/preferences, order according to topics, and a complete random items' order (see Figure \ref{fig:appmod}).

\begin{table}[H]

\caption{\label{tab:table-cfafits}Summary fit indices according to order versions}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lrlrrrrr}
\toprule
Model & $N$ & Estimator & $\chi^2$ & df & CFI & TLI & RMSEA\\
\midrule
Version 1 & 712 & DWLS & 25.631 & 14 & 0.998 & 0.996 & 0.034\\
Version 2 & 717 & DWLS & 67.652 & 14 & 0.984 & 0.967 & 0.073\\
Version 3 & 712 & DWLS & 41.633 & 14 & 0.991 & 0.981 & 0.053\\
\bottomrule
\end{tabular}
\end{table}

Table \ref{tab:table-cfafits} shows the fit indicators of the models estimated for each of the three versions of the items' order described in the methodology. Regardless of the version, all models obtained adequate fit indicators, with CFI's above 0.95 and RMSEA's below 0.08. However, none of the models achieved a non-significant chi-square, as expected in large samples like the one used here. The first version order (perceptions-preferences) was the one obtaining the best fit (CFI=0.993, TLI=0.995, RMSEA=0.034, \(\chi2\)(df=14)=42.276), whereas version 2 with the fixed order according to merit/non-merit items shows the comparatively worst indicators. The CFA fit indices for the completely randomized items' order (Model 3) keep all the indicators within the acceptable cut-off criteria and controls for possible order effects in the administration of the instrument. The model and parameter estimates for this version are depicted in Figure \ref{fig:meas02}:

\begin{figure}[H]

{\centering \includegraphics[width=0.85\linewidth]{output/images/figure9}

}

\caption{Confirmatory factor analysis of the Perceptions and Preferences for Meritocracy Scale}\label{fig:meas02}
\end{figure}

Attending to the correlations between the latent variables as depicted in Figure \ref{fig:meas02}, meritocratic preferences are moderate to highly correlated with perceptions, both meritocratic (\(r=0.457,p<0.01\)) and non-meritocratic (\(r=0.500,p<0.01\)). The correlation between both types of perceptions (\(r=-0.044,p>0.05\)) and both types of preferences (\(r=0.185,p<0.01\)) was low, as well as between non-meritocratic preferences and perceptions (\(r=-0.059p>0.05\)). This last finding gives further evidence regarding the lack of unidimensionality of meritocratic and non-meritocratic aspects as assumed by previous studies.

The measurement of meritocratic beliefs with this new instrument shows, in general, better fit - and therefore better measurement quality - than study 1. However, the comparison between the two studies is unbalanced as we are comparing one particular country (study 2) with the whole set of ISSP countries (study 1). To clear this point, we performed an additional analysis of ISSP data only for the case of Chile. Overall, the results for the chilean ISSP data show a similar fit to the ones of with the complete ISSP dataset (CFI=0.987, TLI=0.980, RMSEA=0.055, \(\chi^2\)(df=45)= 234), and therefore the instrument of study 2 outperforms their fit indicators.

\subsection{Discussion Study 2}\label{discussion-study-2}

The eight-items' battery presented here and the estimation of the confirmatory model with chilean data gives additional evidence for a multidimensional concept of meritocratic beliefs. Despite the good general working of the scale, there are some limitations in interpreting the results given the few items (two) per dimension. As mentioned before, the option of a minimalistic approach in terms of items has advantages and disadvantages: advantages in terms of usability in social surveys with limited space and disadvantages in terms of measurement validity in the estimation of the latent constructs. Regarding the practical use of the measurement instruments, the primary lesson from our study 2 is the importance of distinguishing between the four dimensions. We recommend using the eight-items battery when scholars intend to measure both perceptions and preferences. Alternatively, scholars can use the individual components separately. For instance, researchers interested in the dimension of perception can use only the four items that assess both the meritocratic and non-meritocratic perception. Furthermore, this measurement proposal can be used for relating the dimensions among themselves (for instance, to what extent the perception of meritocracy is related to the preference of it). It can also serve to studies about how each of these dimensions is associated with or moderated by individual and country level variables.

\section{Conclusions}\label{conclusions}

Studies that attempt to characterize and compare societies by their support for meritocratic beliefs have used different approaches. As most studies use secondary survey data, and by and large items from the social inequality module from ISSP, they tend to assume that the available indicators represent an underlying meritocratic construct. A review of these studies reveals several non-tested assumptions and the use of similar indicators to represent different constructs and dimensions of meritocracy. As the existence of heterogeneous approaches certainly has consequences for the advancement of the study of meritocracy, this paper presented a comprehensive multidimensional conceptual framework for the empirical study of meritocracy, building upon previous research. We tested the validity of our conceptual framework with two studies. In a first study, this framework was operationalized and tested with ISSP international data. Based on the results with ISSP, a second study proposed some improvements to the measure of meritocracy with ISSP, which was tested with novel data from a single country case (Chile).

The results of our empirical studies show evidence in favor of a four dimensional model of meritocracy, opening interesting avenues for future comparative research. For instance, distinguishing perceptions from preferences will allow us to evaluate the extent to which different societies are accustomed to, or satisfied with, the perceived level of meritocracy, in terms of differences between what is perceived and what is preferred. Additionally, given that non-meritocratic factors are not necessarily related to meritocratic ones, the multidimensional measurement of meritocratic beliefs makes it possible to assess the perceived legitimacy of practices such as the use of personal contacts and their interference (or not) with meritocratic ideals in different societies. Considering how individual and societal structural factors are related to meritocracy will allow us to gain knowledge about the legitimacy of distribution based on meritocratic criteria. Furthermore, the impact of different configurations of the four-dimensional framework on practices and behaviors, such as corruption, civic involvement, and political alignment, are areas that open up for future development. Such agendas could be especially relevant in times of economic crisis and growing inequalities, which could entail changes in the legitimation of the current distributive structure based on meritocratic ideals.

Besides the areas of research with comparative survey studies as ISSP, single case studies as well as future comparative studies (ISSP included) might benefit from taking into account the measurement instrument for meritocracy proposed here, tested so far with novel Chilean data. The instrument is to some extent comparable to ISSP - as some of its items are equivalent - but improves the balanced measurement of the four dimensions in a minimalistic eight-items' battery, suitable for survey research. Still, further evidence is needed in order to assess the external validity of the results. Upcoming research in the area of meritocratic beliefs, a growing research agenda, would certainly help to keep improving the quality of the measurement instruments in this area, which is a key enterprise when it comes to assess the support and legitimacy of the meritocratic distribution, particularly in unequal societies.

\section{Transparency statement}\label{transparency-statement}

This research follows a series of open science guidelines as the pre-registration of Study 2 (scale development) as well as the availability of data and codes (R) in a public repository in Github. This document was generated with the R package Rmarkdown and contains the code for all the tables and figures to make it reproducible.

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-aiello_new_2019}
Aiello, F., Cardamone, P., \& Pupo, V. (2019). New evidence on the firm-university linkages in {Europe}. {The} role of meritocratic management practices. \emph{International Review of Applied Economics}, \emph{33}(6), 813--828. \url{https://doi.org/10.1080/02692171.2019.1608917}

\bibitem[\citeproctext]{ref-alesina_fairness_2005}
Alesina, A., \& Angeletos, G. (2005). Fairness and redistribution. \emph{American Economic Review}, 960--980.

\bibitem[\citeproctext]{ref-ansolabehere_strength_2008}
Ansolabehere, S., Rodden, J., \& Snyder, J. M. (2008). The {Strength} of {Issues}: {Using Multiple Measures} to {Gauge Preference Stability}, {Ideological Constraint}, and {Issue Voting}. \emph{American Political Science Review}, \emph{102}(2), 215--232. \url{https://doi.org/10.1017/S0003055408080210}

\bibitem[\citeproctext]{ref-arrow_meritocracy_2000}
Arrow, K. J., Bowles, S., \& Durlauf, S. N. (Eds.). (2000). \emph{Meritocracy and economic inequality}. {Princeton University Press}.

\bibitem[\citeproctext]{ref-atria_economic_2020}
Atria, J., Castillo, J., Maldonado, L., \& Ramirez, S. (2020). Economic {Elites}' {Attitudes Toward Meritocracy} in {Chile}: {A Moral Economy Perspective}. \emph{American Behavioral Scientist}, \emph{64}(9), 1219--1241. \url{https://doi.org/10.1177/0002764220941214}

\bibitem[\citeproctext]{ref-boas_recruiting_2020}
Boas, T. C., Christenson, D. P., \& Glick, D. M. (2020). Recruiting large online samples in the {United States} and {India}: {Facebook}, {Mechanical Turk}, and {Qualtrics}. \emph{Political Science Research and Methods}, \emph{8}(2), 232--250. \url{https://doi.org/10.1017/psrm.2018.28}

\bibitem[\citeproctext]{ref-bollen_structural_1989}
Bollen, K. A. (1989). \emph{Structural equations with latent variables}. {Wiley}.

\bibitem[\citeproctext]{ref-breen_class_1999}
Breen, R., \& Goldthorpe, J. H. (1999). Class inequality and meritocracy: A critique of {Saunders} and an alternative analysis. \emph{The British Journal of Sociology}, \emph{50}(1), 1--27. \url{https://doi.org/10.1111/j.1468-4446.1999.00001.x}

\bibitem[\citeproctext]{ref-brown_confirmatory_2015}
Brown, T. (2015). \emph{Confirmatory factor analysis for applied research} (Second edition). {The Guilford Press}.

\bibitem[\citeproctext]{ref-bubak_perceptions_2019}
Bubak, O. (2019). Perceptions of meritocracy: {A} note on {China}. \emph{Asian Journal of Comparative Politics}, \emph{4}(2), 192--209. \url{https://doi.org/10.1177/2057891118806065}

\bibitem[\citeproctext]{ref-castillo_meritocracia_2019}
Castillo, J. C., Torres, A., Atria, J., \& Maldonado, L. (2019). Meritocracia y desigualdad económica: {Percepciones}, preferencias e implicancias. \emph{Revista Internacional de Sociología}, \emph{77}(1), 117. \url{https://doi.org/10.3989/ris.2019.77.1.17.114}

\bibitem[\citeproctext]{ref-cep_encuesta_2019}
CEP. (2019). \emph{Encuesta {CEP Mayo} 2019}.

\bibitem[\citeproctext]{ref-chen_sensitivity_2007}
Chen, F. F. (2007). Sensitivity of {Goodness} of {Fit Indexes} to {Lack} of {Measurement Invariance}. \emph{Structural Equation Modeling: A Multidisciplinary Journal}, \emph{14}(3), 464--504. \url{https://doi.org/10.1080/10705510701301834}

\bibitem[\citeproctext]{ref-cheung_evaluating_2002}
Cheung, G. W., \& Rensvold, R. B. (2002). Evaluating {Goodness-of-Fit Indexes} for {Testing Measurement Invariance}. \emph{Structural Equation Modeling: A Multidisciplinary Journal}, \emph{9}(2), 233--255. \url{https://doi.org/10.1207/S15328007SEM0902_5}

\bibitem[\citeproctext]{ref-cohen_statistical_1988}
Cohen, J. (1988). \emph{Statistical power analysis for the behavioral sciences} (2nd ed). {L. Erlbaum Associates}.

\bibitem[\citeproctext]{ref-converse_nature_1964}
Converse, P. E. (1964). The nature of belief systems in mass publics (1964). \emph{Critical Review}, \emph{18}(1-3), 1--74. \url{https://doi.org/10.1080/08913810608443650}

\bibitem[\citeproctext]{ref-davey_preference_1999}
Davey, L. M., Bobocel, D. R., Son Hing, L. S., \& Zanna, M. P. (1999). Preference for the {Merit Principle Scale}: {An} individual difference measure of distributive justice preferences. \emph{Social Justice Research}, \emph{12}(3), 223--240.

\bibitem[\citeproctext]{ref-davidov_measurement_2009}
Davidov, E. (2009). Measurement {Equivalence} of {Nationalism} and {Constructive Patriotism} in the {ISSP}: 34 {Countries} in a {Comparative Perspective}. \emph{Political Analysis}, \emph{17}(1), 64--82. \url{https://doi.org/10.1093/pan/mpn014}

\bibitem[\citeproctext]{ref-davidov_measurement_2014}
Davidov, E., Meuleman, B., Cieciuch, J., Schmidt, P., \& Billiet, J. (2014). Measurement {Equivalence} in {Cross-National Research}. \emph{Annual Review of Sociology}, \emph{40}(1), 55--75. \url{https://doi.org/10.1146/annurev-soc-071913-043137}

\bibitem[\citeproctext]{ref-dimick_models_2018}
Dimick, M., Rueda, D., \& Stegmueller, D. (2018). Models of other-regarding preferences, inequality, and redistribution. \emph{Annual Review of Political Science}, \emph{21}(1), 441--460. \url{https://doi.org/10.1146/annurev-polisci-091515-030034}

\bibitem[\citeproctext]{ref-dimitrov_testing_2010}
Dimitrov, D. M. (2010). Testing for {Factorial Invariance} in the {Context} of {Construct Validation}. \emph{Measurement and Evaluation in Counseling and Development}, \emph{43}(2), 121--149. \url{https://doi.org/10.1177/0748175610373459}

\bibitem[\citeproctext]{ref-duru-bellat_who_2012}
Duru-Bellat, M., \& Tenret, E. (2012). Who's for meritocracy? {Individual} and contextual variations in the faith. \emph{Comparative Education Review}, \emph{56}(2), 223--247. \url{https://doi.org/10.1086/661290}

\bibitem[\citeproctext]{ref-fischer_are_2011}
Fischer, R., Ferreira, M. C., Jiang, D.-Y., Cheng, B.-S., Achoui, M. M., Wong, C. C., Baris, G., Mendoza, S., van Meurs, N., Achmadi, D., Hassan, A., Zeytinoglu, G., Dalyan, F., Harb, C., Darwish, D. D., \& Assmar, E. M. (2011). Are {Perceptions} of {Organizational Justice Universal}? {An Exploration} of {Measurement Invariance Across Thirteen Cultures}. \emph{Social Justice Research}, \emph{24}(4), 297--313. \url{https://doi.org/10.1007/s11211-011-0142-7}

\bibitem[\citeproctext]{ref-generett_stories_2020}
Generett, G. G., \& Olson, A. M. (2020). The stories we tell: {How} merit narratives undermine success for urban youth. \emph{Urban Education}, \emph{55}(3), 394--423. \url{https://doi.org/10.1177/0042085918817342}

\bibitem[\citeproctext]{ref-girerd_neoliberalism_2020}
Girerd, L., \& Bonnot, V. (2020). Neoliberalism: {An} ideological barrier to feminist identification and collective action. \emph{Social Justice Research}, \emph{33}(1), 81--109. \url{https://doi.org/10.1007/s11211-020-00347-8}

\bibitem[\citeproctext]{ref-goldthorpe_myth_2003}
Goldthorpe, J. (2003). The myth of education-based meritocracy. \emph{New Economy}, \emph{10}(4), 234--239. \url{https://doi.org/10.1046/j.1468-0041.2003.00324.x}

\bibitem[\citeproctext]{ref-hadjar_meritokratie_2008}
Hadjar, A. (2008). \emph{Meritokratie als legitimationsprinzip}. {VS Verlag}.

\bibitem[\citeproctext]{ref-ishida_trends_2005}
Ishida, H., \& Miwa, S. (2005). Trends in intergenerational class mobility and education in japan. \emph{Social Stratification and Social Mobility in Late-Industrializing Countries}.

\bibitem[\citeproctext]{ref-issp-research-groupInternationalSocialSurvey2017}
ISSP-Research-Group. (2017). \emph{International {Social Survey Programme}: {Social Inequality Dataset IV} - {ISSP} 2009}. {GESIS Data Archive}. \url{https://doi.org/10.4232/1.12777}

\bibitem[\citeproctext]{ref-janmaat_subjective_2013}
Janmaat, J. G. (2013). Subjective inequality: A review of international comparative studies on people's views about inequality. \emph{European Journal of Sociology}, \emph{54}(3), 357--389. \url{https://doi.org/10.1017/S0003975613000209}

\bibitem[\citeproctext]{ref-khan_privilege_2013}
Khan, S. (2013). \emph{Privilege: {The} making of an adolescent elite at st. {Paul}'s school} (first paperback printing). {Princeton Univ. Press}.

\bibitem[\citeproctext]{ref-kline_principles_2016}
Kline, R. B. (2016). \emph{Principles and practice of structural equation modeling, 4th ed.} (pp. xvii, 534). {Guilford Press}.

\bibitem[\citeproctext]{ref-kluegel_beliefs_1986}
Kluegel, J. R., \& Smith, E. R. (1986). \emph{Beliefs about {Inequality}: {Americans}' {Views} of {What Is} and {What Ought} to {Be}} (First). {Routledge}. \url{https://doi.org/10.4324/9781351329002}

\bibitem[\citeproctext]{ref-kunovich_systems_2007}
Kunovich, S., \& Slomczynski, K. M. (2007). Systems of distribution and a sense of equity: {A} multilevel analysis of meritocratic attitudes in post-industrial societies. \emph{European Sociological Review}, \emph{23}(5), 649--663. \url{https://doi.org/10.1093/esr/jcm026}

\bibitem[\citeproctext]{ref-land_we_2006}
Land, H. (2006). We {Sat Down} at the {Table} of {Privilege} and {Complained} about the {Food} {\textsuperscript{1}}. \emph{The Political Quarterly}, \emph{77}(s1), 45--60. \url{https://doi.org/10.1111/j.1467-923X.2006.00780.x}

\bibitem[\citeproctext]{ref-linton_study_1936}
Linton, R. (1936). \emph{The {Study} of {Man}}. {D. Appleton-Century}.

\bibitem[\citeproctext]{ref-lundberg_what_2021}
Lundberg, I., Johnson, R., \& Stewart, B. M. (2021). What {Is Your Estimand}? {Defining} the {Target Quantity Connects Statistical Evidence} to {Theory}. \emph{American Sociological Review}, \emph{86}(3), 532--565. \url{https://doi.org/10.1177/00031224211004187}

\bibitem[\citeproctext]{ref-madeira_primes_2019}
Madeira, A. F., Costa-Lopes, R., Dovidio, J. F., Freitas, G., \& Mascarenhas, M. F. (2019). Primes and {Consequences}: {A Systematic Review} of {Meritocracy} in {Intergroup Relations}. \emph{Frontiers in Psychology}, \emph{10}. \url{https://doi.org/10.3389/fpsyg.2019.02007}

\bibitem[\citeproctext]{ref-markovits_meritocracy_2019}
Markovits, D. (2019). \emph{The {Meritocracy} trap: {How America}'s foundational myth feeds inequality, dismantles the middle class, and devours the elite}. {Penguin Press}.

\bibitem[\citeproctext]{ref-mijs_paradox_2019}
Mijs, J. J. B. (2019). The paradox of inequality: {Income} inequality and belief in meritocracy go hand in hand. \emph{Socio-Economic Review}. \url{https://doi.org/10.1093/ser/mwy051}

\bibitem[\citeproctext]{ref-milfont_testing_2010}
Milfont, T. L., \& Fischer, R. (2010). Testing measurement invariance across groups : {Applications} in cross-. \emph{International Journal of Psychological Research}, \emph{3}(1), 111--121. \url{https://doi.org/10.1007/s11135-007-9143-x}

\bibitem[\citeproctext]{ref-millsap_statistical_2011}
Millsap, R. E. (2011). \emph{Statistical approaches to measurement invariance.} (pp. xii, 355). {Routledge/Taylor \& Francis Group}.

\bibitem[\citeproctext]{ref-newman_false_2015}
Newman, B. J., Johnston, C. D., \& Lown, P. L. (2015). False consciousness or class awareness? {Local} income inequality, personal economic position, and belief in american meritocracy. \emph{American Journal of Political Science}, \emph{59}(2), 326--340. \url{https://doi.org/10.1111/ajps.12153}

\bibitem[\citeproctext]{ref-newsom_longitudinal_2015}
Newsom, J. T. (2015). \emph{Longitudinal structural equation modeling: {A} comprehensive introduction}. {Routledge}.

\bibitem[\citeproctext]{ref-owens_engines_2020}
Owens, J., \& de St Croix, T. (2020). {ENGINES OF SOCIAL MOBILITY}? {NAVIGATING MERITOCRATIC EDUCATION DISCOURSE IN AN UNEQUAL SOCIETY}. \emph{British Journal of Educational Studies}, 1--21. \url{https://doi.org/10.1080/00071005.2019.1708863}

\bibitem[\citeproctext]{ref-perez_advancing_2020}
Pérez, A., \& Sabelis, I. (2020). Advancing careers through {``merit''}: A rationalized-sensemaking narrative in hierarchical organizations. \emph{Culture and Organization}, \emph{26}(4), 315--332. \url{https://doi.org/10.1080/14759551.2019.1601723}

\bibitem[\citeproctext]{ref-piketty_capital_2014}
Piketty, T. (2014). \emph{Capital in the twenty-first century} (A. Goldhammer, Trans.). {The Belknap Press of Harvard University Press}.

\bibitem[\citeproctext]{ref-preminger_meritocracy_2020}
Preminger, J. (2020). Meritocracy in the service of ethnocracy. \emph{Citizenship Studies}, \emph{24}(2), 247--263. \url{https://doi.org/10.1080/13621025.2020.1720604}

\bibitem[\citeproctext]{ref-reynolds_perceptions_2014}
Reynolds, J., \& Xian, H. (2014). Perceptions of meritocracy in the land of opportunity. \emph{Research in Social Stratification and Mobility}, \emph{36}, 121--137. \url{https://doi.org/10.1016/j.rssm.2014.03.001}

\bibitem[\citeproctext]{ref-roex_attitudes_2018}
Roex, K. L., Huijts, T., \& Sieben, I. (2018). Attitudes towards income inequality: {``{Winners}''} versus {``{Losers}''} of the perceived meritocracy. \emph{Acta Sociologica}, 0001699317748340. \url{https://doi.org/10.1177/0001699317748340}

\bibitem[\citeproctext]{ref-rosseel_lavaan_2012a}
Rosseel, Y. (2012). Lavaan : {An R Package} for {Structural Equation Modeling}. \emph{Journal of Statistical Software}, \emph{48}(2). \url{https://doi.org/10.18637/jss.v048.i02}

\bibitem[\citeproctext]{ref-sandel_tyranny_2020}
Sandel, M. J. (2020). \emph{The tyranny of merit: {What}'s become of the common good?} (First edition). {Farrar, Straus and Giroux}.

\bibitem[\citeproctext]{ref-saunders_might_1995}
Saunders, P. (1995). Might {Britain} be a {Meritocracy}? \emph{Sociology}, \emph{29}(1), 23--41. \url{https://doi.org/10.1177/0038038595029001003}

\bibitem[\citeproctext]{ref-schroder_income_2017}
Schröder, M. (2017). Is income inequality related to tolerance for inequality? \emph{Social Justice Research}, \emph{30}(1), 23--47. \url{https://doi.org/10.1007/s11211-016-0276-8}

\bibitem[\citeproctext]{ref-sonhing_merit_2011}
Son Hing, L. S., Ramona, D., Zanna, M. P., Garcia, D. M., Gee, S. S., \& Orazietti, K. (2011). The merit of meritocracy. \emph{Journal of Personality and Social Psychology}, \emph{101}(3), 433--450. \url{https://doi.org/10.1037/a0024618}

\bibitem[\citeproctext]{ref-streeck_politics_2014}
Streeck, W. (2014). The {Politics} of {Public Debt}: {Neoliberalism}, {Capitalist Development} and the {Restructuring} of the {State}. \emph{German Economic Review}, \emph{15}(1), 143--165. \url{https://doi.org/10.1111/geer.12032}

\bibitem[\citeproctext]{ref-trump_when_2020}
Trump, K.-S. (2020). When and why is economic inequality seen as fair. \emph{Current Opinion in Behavioral Sciences}, \emph{34}, 46--51. \url{https://doi.org/10.1016/j.cobeha.2019.12.001}

\bibitem[\citeproctext]{ref-vandeschoot_checklist_2012}
van de Schoot, R., Lugtig, P., \& Hox, J. (2012). A checklist for testing measurement invariance. \emph{European Journal of Developmental Psychology}, \emph{9}(4), 486--492. \url{https://doi.org/10.1080/17405629.2012.686740}

\bibitem[\citeproctext]{ref-witteveen_reconsidering_2020}
Witteveen, D., \& Attewell, P. (2020). Reconsidering the {``meritocratic power of a college degree.''} \emph{Research in Social Stratification and Mobility}, \emph{66}, 100479. \url{https://doi.org/10.1016/j.rssm.2020.100479}

\bibitem[\citeproctext]{ref-yair_meritocracy_2007}
Yair, G. (2007). Meritocracy. In G. Ritzer (Ed.), \emph{The blackwell encyclopedia of sociology}. {Blackwell Publishing Ltd}.

\bibitem[\citeproctext]{ref-young_rise_1962}
Young, M. (1962). \emph{The rise of the meritocracy}. {Penguin Books}.

\bibitem[\citeproctext]{ref-zaller_nature_1992}
Zaller, J. R. (1992). \emph{The {Nature} and {Origins} of {Mass Opinion}} (First). {Cambridge University Press}. \url{https://doi.org/10.1017/CBO9780511818691}

\bibitem[\citeproctext]{ref-zhang_quota_2018}
Zhang, B., Mildenberger, M., Howe, P. D., Marlon, J., Rosenthal, S. A., \& Leiserowitz, A. (2018). Quota sampling using {Facebook} advertisements. \emph{Political Science Research and Methods}, 1--7. \url{https://doi.org/10.1017/psrm.2018.49}

\end{CSLReferences}

\pagebreak

\appendix

\section{Appendix}

\textbf{Study 1}

\begin{table}[H]

\caption{\label{tab:countries}Sample of Study 1 }
\centering
\fontsize{10}{12}\selectfont
\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2.5cm}>{\raggedright\arraybackslash}p{1cm}|>{}l>{\raggedright\arraybackslash}p{1cm}}
\toprule
Country & N & Country & N\\
\midrule
Argentina & 997 & Latvia & 972\\
Australia & 1259 & Lithuania & 781\\
Austria & 900 & New Zealand & 834\\
Belgium & 895 & Norway & 997\\
Bulgaria & 709 & Philippines & 1161\\
Chile & 1373 & Poland & 995\\
China & 2687 & Russia & 1201\\
Croatia & 995 & Slovakia & 1064\\
Cyprus & 889 & Slovenia & 934\\
Czechia & 1098 & South Africa & 3011\\
Denmark & 1160 & South Korea & 1546\\
Estonia & 822 & Spain & 1011\\
Finland & 668 & Sweden & 865\\
France & 2104 & Switzerland & 1145\\
Germany & 1174 & Taiwan & 1919\\
Hungary & 947 & Turkey & 1375\\
Iceland & 860 & Ukraine & 1387\\
Israel & 1109 & United Kingdom & 764\\
Italy & 859 & United States & 1335\\
Japan & 916 & Venezuela & 876\\
\bottomrule
\end{tabular}
\begin{tablenotes}[para]
\item \textit{Note: } 
\item N = 46594. Source: International Social Survey Programme (2009).
\end{tablenotes}
\end{threeparttable}
\end{table}

\pagebreak

\textbf{Study 2}

\begin{longtable}[]{@{}lcc@{}}
\caption{\label{tab:rep-samp} Representativeness of the study 2 sample.}\tabularnewline
\toprule\noalign{}
& Sample & CEP \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
& Sample & CEP \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Gender} & & \\
Men & 49,82\% & 50,52\% \\
Women & 50.18\% & 49,47\% \\
\textbf{Age} & & \\
18 - 24 & 18,55\% & 18,17\% \\
25 - 34 & 18,86\% & 17,48\% \\
35 - 44 & 19.09\% & 19,98\% \\
45 - 54 & 17,96\% & 19,23\% \\
55 - or more & 25,54\% & 25.11\% \\
\textbf{Education} & & \\
Primary or less & 2,93\% & 15,88\% \\
High school & 43,23\% & 37,04\% \\
Non university & 32,63\% & 28,93\% \\
University or more & 21,21\% & 18,13\% \\
\end{longtable}

\textbf{Study 3: Additional validity analyses}

The modeling for invariance testing has been gaining more relevance in sociological survey studies, mainly due to the interest in the validity of measurement scales in comparative studies in various social and cultural contexts (\citeproc{ref-davidov_measurement_2014}{Davidov et al., 2014}). In this section, we present the results of an analysis of invariance between groups, testing whether the proposed factorial structure for the meritocracy scale is applicable in a different dataset.

The invariance measurement procedure consists of a series of nested models in which restrictions are progressively incorporated into the parameters of the measurement model. The literature generally suggests that this modeling approach should consider four levels or types of progressive restriction (\citeproc{ref-milfont_testing_2010}{Milfont \& Fischer, 2010}; \citeproc{ref-millsap_statistical_2011}{Millsap, 2011}; \citeproc{ref-vandeschoot_checklist_2012}{van de Schoot et al., 2012}):

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Configural:} the model is estimated only indicating the factorial structure used in the CFA.
\item
  \textbf{Weak:} an equality restriction applied to the factor loadings in the different groups; that is, the loadings are forced to be identical in both measurements.
\item
  \textbf{Strong:} equality restrictions added to the intercepts of each indicator.
\item
  \textbf{Strict:} equality restrictions added to the error variances of each indicator.
\end{enumerate}

\textbf{Data}

For testing the invariance, we compared the data described previously in Study 2 to a different dataset from an online survey carried out during the first half of 2020. The characteristics of this sample in terms of application and coverage are equivalent to those of the previous study. The final sample contains 1,242 cases, where 605 (48.71\%) are women, and 637 (51.28\%) are men. 48.84\% are 45 or older, and 33.97\% have tertiary education or higher.

\textbf{Variables}

The items of the meritocracy scale were identical to the original scale of Study 2 administrated in a randomized order (which corresponds to the application modality of Group 3 in Study 2, as shown in Figure \ref{fig:appmod}).

\textbf{Results}

The first step for invariance testing is estimating the configural model, which serves as the baseline for further comparison and is expected to adequately meet the global fit criteria of a measurement model. Although the chi-square statistic is used as a global measure of fit, it is usually complemented with other indexes given its high sensitivity to sample size: the Comparative Fit Index (CFI), which should have a value greater than 0.95; the Root Mean Square Error Approximation (RMSEA), which must be lower than 0.06; and the Standardized Root Mean Square Residuals (SRMR) which must be less than 0.08 (\citeproc{ref-vandeschoot_checklist_2012}{van de Schoot et al., 2012}). Besides, the literature on measurement invariance suggests some complementary approaches for evaluating the fit, from which we will consider the incremental adjustment of the fit indexes (\citeproc{ref-cheung_evaluating_2002}{Cheung \& Rensvold, 2002}; \citeproc{ref-dimitrov_testing_2010}{Dimitrov, 2010}; \citeproc{ref-milfont_testing_2010}{Milfont \& Fischer, 2010}) and the ANOVA test for means comparison in nested models (\citeproc{ref-newsom_longitudinal_2015}{Newsom, 2015}).

\begin{table}[H]

\caption{\label{tab:invargroup}Multiple Group measurement invariance for Perceptions and Preferences for Meritocracy}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{llllllll}
\toprule
Model & $\chi^2 (\text{df})$ & CFI & RMSEA (90 CI) & $\Delta \chi^2 (\Delta \text{df}$) & $\Delta \text{CFI}$ & $\Delta \text{RMSEA}$ & Decision\\
\midrule
Configural & 316.27 (28) & 0.944 & 0.078 
 (0.07-0.086) &  &  &  & \\
Weak & 323.62 (32) & 0.943 & 0.073 
 (0.066-0.081) & 7.347 (4) & -0.001 & -0.005 & Accept\\
Strong & 330.66 (36) & 0.942 & 0.07 
 (0.063-0.077) & 7.045 (4) & -0.001 & -0.004 & Accept\\
Strict & 426.74 (44) & 0.925 & 0.072 
 (0.066-0.078) & 96.08 (8) *** & -0.017 & 0.002 & Reject\\
\bottomrule
\multicolumn{8}{l}{\rule{0pt}{1em}\textit{Note: } N = ; Group 1, n = 1242; Group 2, n = 2141, ***p < 0.001}\\
\end{tabular}
\end{table}

Table \ref{tab:invargroup} shows the results of the measurement invariance estimation. When attending to the traditional invariance test of \(\Delta \chi^2 (\Delta \text{df})\), the results support the invariance at the strong level meaning that the fit of the factor model of the merit scale is equivalent across samples when constraining factor loadings and intercepts to being equal. Such result is considered in general as evidence of invariance (\citeproc{ref-fischer_are_2011}{Fischer et al., 2011}), as strict forms of measurement invariance rarely hold (\citeproc{ref-vandeschoot_checklist_2012}{van de Schoot et al., 2012}). Still, the comparability of latent means requires strict invariance which in this case does not hold when considering just \(\Delta \chi^2 (\Delta \text{df})\). Nevertheless, the criteria of \(\Delta \text{CFI}\) used for comparing models is close to the rejection criteria of \textgreater.01, whereas the \(\Delta \text{RMSEA}\) fulfills the requirements of being below the cut-off criteria as suggested by (\citeproc{ref-chen_sensitivity_2007}{Chen, 2007}). Therefore, using this last standard, the level of strict invariance would hold for the meritocracy scale.

\emph{Discusion}

Based on the measurement invariance analysis for two datasets, we found evidence supporting the equivalence of the scale when applied in different samples. This gives a more robust ground when using this measure for comparing groups, for instance in country comparisons. Although the results are promising, still the analysis was performed in two samples in a single country (Chile) and it requires further examination with social surveys in other societies.

\end{document}

### 2023-09-14 paper-blinded-RR-IJS.Rmd
ajustes autores

@@
 link-citations: yes
-# author: "Juan Carlos Castillo, Julio Iturra, Luis Maldonado, Francisco Meneses & Jorge Atria <br><br>"
+# author: 
+# - name: Juan Carlos Castillo
+#   affiliation: Department of Sociology, Universidad de Chile
+#   email: juancastillov@uchile.cl
+#   number: 1
+# - name: Julio Iturra
+#   affiliation: Bremen International Graduate School of Social Sciences, Universität Bremen
+#   number: 2
+# - name: Luis Maldonado
+#   affiliation: Institute of Sociology, Universidad Católica de Chile
+#   number: 3
+# - name: Francisco Meneses
+#   affiliation: 
+#   number: 1
+# - name: Jorge Atria
+#   affiliation: Department of Sociology, Universidad Diego Portales, Chile
+#   number: 5
 abstract: |
@@
     pandoc_args:
-#      - --template=input/mytemplate.tex #custom template para usar autores con afiliacion    
+     - --template=input/mytemplate.tex #custom template para usar autores con afiliacion
   bookdown::html_document2:
@@
 #  pacman::p_load(dplyr, kableExtra, knitr,lavaan,countrycode,ggplot2,haven,
-                 gridExtra)
   table_format = if(knitr::is_html_output()) { #conditional instructions for kable
@@
  tabitems$itemsspa <- NULL
+ tabitems[9,] <- tabitems[8,]
+ tabitems[9,3] <- "-"
+ tabitems[10,] <- tabitems[8,]
+ tabitems[10,3] <- "-"
  tabitems$issp <- c("How important is hard work?","-",
-                    "How important is coming from a wealthy family?", "How important is knowing the right people",
-                    "How hard he or she works at the job?*","How well he or she does the job?*",
-                    "-", "-")
+                    "How important is coming from a wealthy family?", 
+                    "How important is knowing the right people",
+                    "In deciding how much people ought to earn, how important each of these things be (...) How hard he or she works at the job?¹",
+                    " (...) How well he or she does the job?¹",
+                    "-", "-",
+                    "(...) What is needed to support a family?¹²","(...) Whether the person has children to support²")
+
  cnames <- c("Component", "Dimensions","Items" , "Items ISSP (2009,2019)")
  cap <- "Items' battery for the multidimensional measurement of meritocratic beliefs and their equivalents in ISSP"
- knitr::kable(tabitems, table_format, booktabs = T, linesep = "",col.names = cnames, caption = cap) %>%
+ knitr::kable(tabitems, table_format, booktabs = T, linesep = "",col.names = cnames, caption = cap,
+              longtable = F, escape = T) %>%
    kableExtra::kable_styling(
@@
      bootstrap_options=c("striped", "bordered")) %>%
-   kableExtra::column_spec(column = 1, width = "1.5cm", ) %>%
+   kableExtra::column_spec(column = 1, width = "1.5cm") %>%
    kableExtra::column_spec(column = 2,width = "2 cm") %>%
    kableExtra::collapse_rows(columns = 1:2,valign = "middle") %>%
-   kableExtra::footnote(general = "Those with an asterisk (*) are only present in 2009",footnote_as_chunk = T)
+   kableExtra::footnote(number =  c("Items present only in ISSP 2009",
+                                    "ISSP items that are not equivalent with the new measurement proposal"))
 ```
@@
 
-**Study 1**
-
 ```{r countries, echo=FALSE}
@@
 \pagebreak
-
-**Study 2**
 
@@
 
-**Study 3: Additional validity analyses**
+**Additional validity analyses of study 2**
 
@@
 
-*Discusion*
+**Discusion**
 

### 2023-10-06 paper-blinded-RR-IJS.Rmd
ajustes pre render

@@
 The item used by @mijs_paradox_2019 is part of an item's battery which is available in several international surveys (as ISSP), usually called “reasons to get ahead”.
+In the ISSP, this battery belongs to the social inequality module---see @roberts_issp_2023 for a recent overview of this module.
 This battery displays a series of indicators related to what people consider important to get ahead: hard work, education, ambition, a wealthy family, the right connections, religion, race, and gender.
@@
   if (!require("pacman")) install.packages("pacman")
-#  pacman::p_load(dplyr, kableExtra, knitr,lavaan,countrycode,ggplot2,haven,
+ pacman::p_load(dplyr, kableExtra, knitr,lavaan,countrycode,ggplot2,haven,gridExtra)
   table_format = if(knitr::is_html_output()) { #conditional instructions for kable
@@
 ```
+
+```{r fit-country-s2, eval=FALSE, include=FALSE}
+if (!require("pacman")) install.packages("pacman")
+  #---Load packages and data -------------------------------------#
+  pacman::p_load(dplyr,lavaan,semTools,knitr,kableExtra,stats,gtools,haven)
+  options(knitr.kable.NA = '')
+  load(file = "input/data/proc/data_s1.RData")
+
+# meassurement model: Second order
+  cfa_perpref2 <- '
+  	  # latent variables
+  	  percmerit =~ hwork + ambition
+  	  parents =~ wealthy + pareduc
+  	  backgrd =~ race + gender
+  	  networks=~ people + polcone
+  	  percnmerit=~ parents + backgrd + networks
+  	  prefmerit =~ welljob + hardjob
+  	  prefnmerit=~ family + child
+  	  '
+
+  data_s1$country <- countrycode::countrycode(sourcevar = data_s1$v5,origin = "iso3n",destination = "country.name")
+# estimar modelo por paises
+df1<-data_s1 %>% select(country,wealthy:hardjob)
+
+# sjmisc::frq(df1$country)
+# 
+# Argentina      <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Argentina",],ordered = T)
+# Australia      <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Australia",],ordered = T)
+# Austria        <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Austria",],ordered = T)
+# Belgium        <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Belgium",],ordered = T)
+# # Bulgaria       <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Bulgaria",],ordered = T)
+# Chile          <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Chile",],ordered = T)
+# China          <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="China",],ordered = T)
+# Croatia        <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Croatia",],ordered = T)
+# Cyprus         <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Cyprus",],ordered = T)
+# Czechia        <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Czechia",],ordered = T)
+# Denmark        <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Denmark",],ordered = T)
+# Estonia        <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Estonia",],ordered = T)
+# Finland        <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Finland",],ordered = T)
+# France         <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="France",],ordered = T)
+# Germany        <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Germany",],ordered = T)
+# Hungary        <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Hungary",],ordered = T)
+# # Iceland        <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Iceland",],ordered = T)
+# Israel         <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Israel",],ordered = T)
+# # Italy          <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Italy",],ordered = T)
+# # Japan          <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Japan",],ordered = T)
+# Latvia         <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Latvia",],ordered = T)
+# # Lithuania      <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Lithuania",],ordered = T)
+# NewZealand    <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="New Zealand",],ordered = T)
+# Norway         <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Norway",],ordered = T)
+# Philippines    <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Philippines",],ordered = T)
+# Poland         <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Poland",],ordered = T)
+# # Portugal       <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Portugal",],ordered = T)
+# Russia         <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Russia",],ordered = T)
+# Slovakia       <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Slovakia",],ordered = T)
+# Slovenia       <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Slovenia",],ordered = T)
+# SouthAfrica   <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="South Africa",],ordered = T)
+# SouthKorea    <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="South Korea",],ordered = T)
+# Spain          <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Spain",],ordered = T)
+# # Sweden         <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Sweden",],ordered = T)
+# Switzerland    <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Switzerland",],ordered = T)
+# Taiwan         <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Taiwan",],ordered = T)
+# Turkey         <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Turkey",],ordered = T)
+# Ukraine        <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Ukraine",],ordered = T)
+# UnitedKingdom <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="United Kingdom",],ordered = T)
+# # UnitedStates  <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="United States",],ordered = T)
+# # Venezuela      <- lavaan::cfa(cfa_perpref2,data = df1[df1$country=="Venezuela",],ordered = T)
+# # 
+# # 
+# # 
+# lavaan::fitmeasures(Argentina     )
+# lavaan::fitmeasures(Australia     )
+# lavaan::fitmeasures(Austria       )
+# lavaan::fitmeasures(Belgium       )
+# # lavaan::fitmeasures(# Bulgaria    )
+# lavaan::fitmeasures(Chile         )
+# lavaan::fitmeasures(China         )
+# lavaan::fitmeasures(Croatia       )
+# lavaan::fitmeasures(Cyprus        )
+# lavaan::fitmeasures(Czechia       )
+# lavaan::fitmeasures(Denmark       )
+# lavaan::fitmeasures(Estonia       )
+# lavaan::fitmeasures(Finland       )
+# lavaan::fitmeasures(France        )
+# lavaan::fitmeasures(Germany       )
+# lavaan::fitmeasures(Hungary       )
+# # lavaan::fitmeasures(# Iceland     )
+# # lavaan::fitmeasures(Israel        )
+# # lavaan::fitmeasures(Italy         )
+# # lavaan::fitmeasures(# Japan       )
+# lavaan::fitmeasures(Latvia        )
+# # lavaan::fitmeasures(# Lithuania   )
+# lavaan::fitmeasures(NewZealand    )
+# lavaan::fitmeasures(Norway        )
+# lavaan::fitmeasures(Philippines   )
+# lavaan::fitmeasures(Poland        )
+# # lavaan::fitmeasures(# Portugal    )
+# lavaan::fitmeasures(Russia        )
+# lavaan::fitmeasures(Slovakia      )
+# lavaan::fitmeasures(Slovenia      )
+# lavaan::fitmeasures(SouthAfrica   )
+# lavaan::fitmeasures(SouthKorea    )
+# lavaan::fitmeasures(Spain         )
+# # lavaan::fitmeasures(# Sweden      )
+# lavaan::fitmeasures(Switzerland   )
+# lavaan::fitmeasures(Taiwan        )
+# lavaan::fitmeasures(Turkey        )
+# lavaan::fitmeasures(Ukraine       )
+# lavaan::fitmeasures(UnitedKingdom )
+# # lavaan::fitmeasures(UnitedStates)
+# # lavaan::fitmeasures(Venezuela     )
+
+df2<-data_s1 %>% select(country,wealthy:hardjob) %>% 
+  filter(!country %in% c("Bulgaria","Iceland","Israel","Italy","Japan",
+                         "Lithuania","Portugal","Sweden","United States","Venezuela"))
+
+df_paises <- split(df2, df2$country)
+results = lapply(df_paises, function(df2) lavaan::cfa(cfa_perpref2,data = df2,ordered = T))
+
+#extraer fitmeasures por paises
+results_list<- lapply(results, function(results) data.frame(t(lavaan::fitmeasures(results)[c("chisq","df","cfi","tli","rmsea")])))
+#extraer nobs por paises
+nobs_list <- lapply(results, function(results) data.frame(obs=t(lavaan::nobs(results))))
+nobs_n <- data.table::rbindlist(nobs_list) %>% 
+  mutate(country=names(nobs_list)) %>%
+  select(country,everything())
+
+  # extract fit indices from models and add to table
+  sum_fit <- 
+  data.table::rbindlist(results_list) %>% 
+  mutate(country=names(results_list)) %>% 
+  select(country,everything())
+  # Customize object
+  sum_fit$nobs <- nobs_n$obs
+  sum_fit <- dplyr::select(sum_fit,country,"chisq","df", nobs,"cfi","tli","rmsea")
+  sum_fit <- sum_fit %>% dplyr::arrange(chisq)
+  colnames <- c("País","$\\chi^2$","gl","$N$","CFI","TLI","RMSEA")
+  # Create table
+  tab_cfa_count_1d <-
+    kable(
+      sum_fit,
+      # format=table_format,
+      digits = 3,
+      booktabs = T,
+      col.names = colnames,
+      caption = "Summary of country-specific fit for the ISSP",
+      escape = FALSE
+    ) %>% 
+    kable_styling(full_width = F,
+                  font_size = 10,
+                  latex_options = "HOLD_position",
+                  bootstrap_options=c("striped", "bordered"))
+  tab_cfa_count_1d
+```
+
+```{r invariance-s2, eval=FALSE, include=FALSE}
+df2<-data_s1 %>% select(country,wealthy:hardjob) %>% 
+  filter(!country %in% c("Bulgaria","Iceland","Israel","Italy","Japan",
+                         "Lithuania","Portugal","Sweden","United States","Venezuela","Finland"))
+
+  inv01 <- semTools::measurementInvariance(model=cfa_perpref2,
+                                           data=df2,
+                                           group="country",estimator = "ML",strict=TRUE,quiet = T)
+
+  var_or <- c("hwork","ambition","wealthy", "pareduc","race",
+              "gender","people","polcone","welljob" ,"hardjob","family","child")
+
+  inv02 <- semTools::measurementInvarianceCat(model=cfa_perpref2,
+                                              data=df2,
+                                           group="country",estimator = "DWLS",
+                                           strict=TRUE,quiet = T,ordered=var_or)
+
+conf  <- inv01$fit.configural
+  weak  <- inv01$fit.loadings
+  strong<- inv01$fit.intercepts
+  strict<- inv01$fit.residuals
+
+  ### Check an alternative method
+  configural        <- lavaan::cfa(cfa_perpref2, data=df2, group = "country", estimator="ML")
+  weak.invariance   <- lavaan::cfa(cfa_perpref2, data=df2, group = "country",group.equal = "loadings", estimator="ML")
+  strong.invariance <- lavaan::cfa(cfa_perpref2, data=df2, group = "country",group.equal = c( "loadings", "intercepts"), estimator="ML")
+  strict.invariance <- lavaan::cfa(cfa_perpref2, data=df2, group = "country",group.equal = c( "loadings", "intercepts", "residuals"), estimator="ML")
+
+  tab01<- lavaan::anova(conf,weak,strong,strict,SB.classic=TRUE) %>%
+    dplyr::as_tibble() %>%
+    dplyr::select("Chisq","Df","chisq_diff"=`Chisq diff`,"df_diff"=`Df diff`,"pvalue"=`Pr(>Chisq)`) %>%
+    dplyr::mutate(stars=gtools::stars.pval(pvalue),
+           chisqt=paste0(round(Chisq,2)," (",Df,") "),
+           decision=ifelse(pvalue>0.05,yes = "Accept",no = "Reject"),
+           model=c("Configural","Weak","Strong","Strict"))
+
+  fit.meas<- dplyr::bind_rows(lavaan::fitmeasures(inv01$fit.configural,output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
+                              lavaan::fitmeasures(inv01$fit.loadings,  output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
+                              lavaan::fitmeasures(inv01$fit.intercepts,output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
+                              lavaan::fitmeasures(inv01$fit.residuals, output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),])
+
+  # compute differences in chisq, df, cfi and rmsea (90%, lower.ci - upper.ci )
+  fit.meas<- fit.meas %>%
+    dplyr::mutate(diff.chi2 = chisq    - lag(chisq,default = dplyr::first(chisq)),
+           diff.df   = df       - lag(df,   default = dplyr::first(df)),
+           diff.cfi  = cfi      - lag(cfi,  default = dplyr::first(cfi)),
+           diff.rmsea   = rmsea - lag(rmsea,default = dplyr::first(rmsea))) %>%
+    round(3) %>%
+    dplyr::mutate(rmsea.ci=paste0(rmsea," \n ", "(",rmsea.ci.lower,"-",rmsea.ci.upper,")"))
+
+  tab.inv<- dplyr::bind_cols(tab01,fit.meas) %>%
+    dplyr::select(model,chisqt,cfi,rmsea.ci,diff.chi2,diff.df,diff.cfi,diff.rmsea,stars,decision) %>%
+    dplyr::mutate(diff.chi2=paste0(diff.chi2," (",diff.df,") ",stars)) %>%
+    dplyr::select(model,chisqt,cfi,rmsea.ci,diff.chi2,diff.cfi,diff.rmsea,decision)
+
+  #clean values
+  tab.inv[tab.inv == c("0 (0) ")] <- NA
+  tab.inv[tab.inv == c(0)] <- NA
+
+  ### Check an alternative method
+  configural        <- lavaan::cfa(model01, data=data_s3_inv, group = "dataset", estimator="ML")
+  weak.invariance   <- lavaan::cfa(model01, data=data_s3_inv, group = "dataset",group.equal = "loadings", estimator="ML")
+  strong.invariance <- lavaan::cfa(model01, data=data_s3_inv, group = "dataset",group.equal = c( "loadings", "intercepts"), estimator="ML")
+  strict.invariance <- lavaan::cfa(model01, data=data_s3_inv, group = "dataset",group.equal = c( "loadings", "intercepts", "residuals"), estimator="ML")
+
+  table_format = if(is_html_output()) {
+    "html"
+  } else if(is_latex_output()) {
+    "latex"
+  }
+
+  col.nam <- c("Model","$\\chi^2 (\\text{df})$","CFI","RMSEA (90 CI)",
+               "$\\Delta \\chi^2 (\\Delta \\text{df}$)","$\\Delta \\text{CFI}$","$\\Delta \\text{RMSEA}$","Decision")
+  footnote <- paste0("N = ","; Group 1, n = ",conf@Data@nobs[[1]],"; Group 2, n = ",conf@Data@nobs[[2]], ", ***p < 0.001")
+
+  knitr::kable(tab.inv, col.names = col.nam, align = "l",
+        booktabs=TRUE,format = table_format,escape = FALSE,
+        caption = "Multiple Group measurement invariance for Perceptions and Preferences for Meritocracy") %>%
+    kableExtra::kable_styling(full_width = FALSE,
+                              latex_options = "HOLD_position",
+                              bootstrap_options=c("striped", "bordered"),
+                              font_size = 10) %>%
+    kableExtra::footnote(general = footnote, footnote_as_chunk = T)
+```

### 2023-10-06 paper-blinded-RR-IJS.Rmd
render 1

@@
 ```{r setup, include=FALSE}
- knitr::opts_chunk$set(echo=FALSE, warning = FALSE,message = FALSE, cache = TRUE,out.width = '85%',fig.pos= "H", fig.align = 'center')
+ knitr::opts_chunk$set(echo=FALSE, warning = FALSE,message = FALSE, cache = FALSE,out.width = '85%',fig.pos= "H", fig.align = 'center')
  # knitr::opts_knit$set(base.url = "../") #relative path for .html output file

### 2023-10-26 paper-blinded-RR-IJS.Rmd

--- 2023-09-14 paper-blinded-RR-IJS.Rmd
@@
 output:
-  bookdown::word_document2:
+  bookdown::html_document2:
     number_sections: false
-    reference_docx: input/template.docx
-    toc: false
   bookdown::pdf_document2:
@@
      - --template=input/mytemplate.tex #custom template para usar autores con afiliacion
-  bookdown::html_document2:
+  bookdown::word_document2:
     number_sections: false
+    reference_docx: input/template.docx
+    toc: false
 linkcolor: blue
@@
   - \usepackage{times}           # Times New Roman
-#  - \hypersetup {colorlinks = true, linkcolor = blue, urlcolor = blue}    
   - \usepackage{caption}
@@
     dplyr::select(hwork_r:ambition_r) %>%
-    sjPlot::plot_likert(geom.colors = "PuBu",
+    sjPlot::plot_likert(geom.colors = "Greys",
                         grid.range  =  c (1.2 , 1.2),
@@
     dplyr::select(wealthy_r:polcone_r) %>%
-    sjPlot::plot_likert(geom.colors = "PuBu",
+    sjPlot::plot_likert(geom.colors = "Greys",
                         grid.range  =  c (1.2 , 1.2),
@@
     dplyr::select(welljob_r:hardjob_r) %>%
-    sjPlot::plot_likert(geom.colors = "PuBu",
+    sjPlot::plot_likert(geom.colors = "Greys",
                         grid.range  =  c (1.2 , 1.2),
@@
     dplyr::select(family_r:child_r) %>%
-    sjPlot::plot_likert(geom.colors = "PuBu",
+    sjPlot::plot_likert(geom.colors = "Greys",
                         grid.range  =  c (1.2 , 1.2),
@@
       dplyr::select(likert_issp, family_r:child_r) %>%
-        sjPlot::plot_likert(geom.colors = "PuBu", catcount = 5) +
+        sjPlot::plot_likert(geom.colors = "Greys", catcount = 5) +
         ggplot2::scale_fill_brewer(
@@
           labels = c("Not important at all","Not very important","Fairly important",
-                     "Very important","Essential")) +
+                     "Very important","Essential"),palette = "Greys") +
         ggplot2::theme(legend.position = "bottom")
@@
     method = "color",
-    addCoef.col = "#000390",
+    # addCoef.col = "#000390",
+    addCoef.col = "Gray",
     type = "upper",
     tl.col = "black",
-    col=colorRampPalette(c("white","#0068DC"))(12),
+    # col=colorRampPalette(c("white","#0068DC"))(12),
+    col=colorRampPalette(c("white","gray9"))(12),
     bg = "white",
@@
   scale_y_continuous(limits = c(-0.6,0.6))+
-  geom_smooth(method = "lm",fullrange=TRUE) +
+  geom_smooth(method = "lm",fullrange=TRUE,color="black") +
   xlab("Meritocratic perception")+
@@
   scale_y_continuous(limits = c(-0.6,0.6))+ 
-  geom_smooth(method = "lm",fullrange=TRUE) +  
-  geom_smooth(method = "lm") +
+  geom_smooth(method = "lm",fullrange=TRUE,color="black") +  
   xlab("Non-meritocratic perception")+
@@
   scale_y_continuous(limits = c(-0.6,0.6))+
-  geom_smooth(method = "lm",fullrange=TRUE) +
+  # geom_smooth(method = "lm",fullrange=TRUE) +
+geom_smooth(method = "lm",fullrange=TRUE,color="black") +  
   xlab("Meritocratic perception")+
@@
   scale_y_continuous(limits = c(-0.6,0.6))+ 
-  geom_smooth(method = "lm",fullrange=TRUE) +
+  # geom_smooth(method = "lm",fullrange=TRUE) +
+geom_smooth(method = "lm",fullrange=TRUE,color="black") +    
   xlab("Non-meritocratic perception")+
@@
   scale_y_continuous(limits = c(-0.6,0.6))+
-  geom_smooth(method = "lm",fullrange=TRUE) +
+  # geom_smooth(method = "lm",fullrange=TRUE) +
+  geom_smooth(method = "lm",fullrange=TRUE,color="black") +  
   xlab("Meritocratic perception")+
@@
   geom_smooth(method = "lm",fullrange=TRUE) +
+  geom_smooth(method = "lm",fullrange=TRUE,color="black") +    
   xlab("Non-meritocratic perception")+
@@
     dplyr::select(dat_likert_s2,perc_effort_r:perc_netw_r) %>%
-    sjPlot::plot_likert(geom.colors = "PuBu",
+    sjPlot::plot_likert(geom.colors = "Greys",
                         title = c("a. Perceptions"),
@@
     dplyr::select(dat_likert_s2,pref_effort_r:pref_netw_r) %>%
-    sjPlot::plot_likert(geom.colors = "PuBu",
+    sjPlot::plot_likert(geom.colors = "Greys",
                         title = c("b. Preferences"),
@@
     dplyr::select(dat_likert_s2,pref_effort_r:pref_netw_r) %>%
-        sjPlot::plot_likert(geom.colors = "PuBu",
+        sjPlot::plot_likert(geom.colors = "Greys",
                             catcount = 4,
@@
       ggplot2::scale_fill_manual(
-        values = c("#b3b3b3ff", "#f1eef6ff", "#bdc9e1ff","#74a9cfff","#0570b0ff"),
+        values = c("#B3B3B3","#F7F7F7","#CCCCCC","#969696","#525252"),
+        # values = c("#b3b3b3ff", "#f1eef6ff", "#bdc9e1ff","#74a9cfff","#0570b0ff"),
         breaks=c("neutral",4,3,2,1),
@@
     method = "color",
-    addCoef.col = "#000390",
+    # addCoef.col = "#000390",
+    addCoef.col = "Gray",
     type = "upper",
     tl.col = "black",
-    col=colorRampPalette(c("white","#0068DC"))(8),
+    # col=colorRampPalette(c("white","#0068DC"))(8),
+    col=colorRampPalette(c("white","gray9"))(12),
     bg = "white",
